### 2018-03-28 MSLT/MSLTdoc.Rmd
Cloned MSLT from https://github.com/ITHIM/mslt

---
title: "Proportional multi-state multiple-cohort life table model"
author: "Belen Zapata-Diomedi and Ali Abbas"
date: "26 March 2018"
header-includes:
    - \usepackage{setspace}\onehalfspacing
output:
  pdf_document: 
    toc: true
    number_sections: true
    fig_caption: yes
  fontsize: 12 pt
  geometry: margin=1in
  bibliography: MSLT\Bibliography.txt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

The proportional multi-state multiple-cohort life table model (PMSLT) is a population level model (macro) approach to simulate health (and economic) implications of changes in exposure to health risk factors (e.g. physical inactivity, air pollution and diet). The PMSLT has been widely used to simulate outcomes for population level interventions for the reduction of chronic diseases.

The model was developed by Jan Barendregt and colleagues (REF ACE prevention and core papers) and has been widely used in Australia and New Zealand (Ref ACE-report and BODE website of key publications).

The basic infrastructure of the model consist of three components: (1) Effect size for the intervention of interest (e.g. intervention to urban design that modifies population levels of physical activity); (2) Calculation of the potential impact fraction (PIF) to derive the change in occurrence of disease (incidence rate/mortality rate) attributable to a change in the distribution of the risk factor (e.g. physical activity); and (3) Use of the PMSLT to simulate health (and economic) outcomes attributable to a change in the distribution of health risk factor/s in the population of interest. Figure 1 summaries the basic infrastructure of the model. ITHIM is included in Figure 1 to show that both approaches share in common steps one and two and differ in the mechanisms of calculation of change in health burden.

![Figure 1. Basic model infrastructure.](structure\Figure1.png)

**HALYs, QALYs and DALYs**

In this model we use the term ‘health-adjusted life year’ (HALY). As ‘summary measure of population health’ it measures both quantity and quality of life, where one HALY represent the equivalent of one year in full health (which could be two years with a quality of life of 0.5, for example). Specific types of HALY are the quality-adjusted life year (QALY) and the disability-adjusted life year (DALY). The QALY derives from economics and was first used in the 1960s as a measure of health gain (ref Gold MR, Stevenson D, Fryback DG. HALYs and QALYs and DALYs, oh my: similarities and differences in summary measures of population health. Annual review of public health. 2002;23(1):115-34.). The disability-adjusted life-year (DALY) was developed for use in burden of disease studies as a measure of health loss due to disease (Gold et al.). Our calculated HALYs are neither QALYs not DALYs, but something in between. They are similar to QALYs in that they represent health gains. However, the main difference is in the calculation of the health-related quality of life component. QALYs use measures of utility weights that traditionally represent individual experiences of health, whereas our estimated HALYs use disability weights linked to specific diseases, which were developed for the Global Burden of Disease study (GOld et al). As discussed in past research (REF Roux and Cobiac 2009) the main advantage of using disability weights over utility weights is that disability weights refer to specific diseases rather than health states. We opted to use the more general terms HALYs given that the use of the DALYs terminology may lead to think that our calculations are similar to those in burden of diseases studies (REF Murray 1997). In our study, our model does not explicitly separate years of life lost (YLL) and years lived with disability (YLD) components, but instead calculates the total number of life years lived, adjusted for the average health-related quality of life in those years (by age and sex). In burden of disease studies, DALYs are defined as the sum Years of Life Lost (YLL) and Years Lived with Disability (YLD).

## Contribution to ITHIMR

The PMSLT similar to ITHIM is a comparative risk assessment approach (REF BRIGGS 2016) that consist of calculating the change in the health burden for a population of interest from a change in exposure to health risks factors (e.g. physical inactivity, air pollution and road trauma).As depicted in Figure 1, both methods need estimates of the potential impact fraction (PIF), which indicates the proportion of the disease burden attributable to a risk factor of interest (e.g. physical inactivity). A step further back, is the development of scenarios that bring about change in the distribution of the risk factor of interest. For now, we only focus on calculations from the PIF onward, and provide a hypothetical example of change in population levels of physical activity. Incorporation of additional  health risk factor (air pollution, road trauma, NO2 and noise) will be discussed in the relevant code sections.

### Difference between ITHIM and PMSLT

* **Time component**
    The *PMSLT* follows a population of interest over time. For example, as set up         here, we simulate sex and age (5 years starting at 18) cohorts over time until they     die or reach 100 years of age. This implies that we can include trends for             diseases, time lags between change in exposure to risk factors and change in health     and demographic changes (e.g. population growth). In addition, we can estimate         yearly changes in the burden of diseases over the life course or for a specified       number of years. The *ITHIM* approach is a snapshot of change in burden for one        year.

* **Interaction between multiple diseases**
    The *PMSLT* accounts for the interaction between multiple diseases, with               proportions of the population being able to be in more than one health state (REF      BRIGGS 2016). This avoids  overestimation of outcomes as a result of summing health     outcomes attributable to each disease individually as done in *ITHIM*. It is           important to note that the *PMSLT* assumes that diseases are independent of each       other. That is to say, developing a disease is unrelated to a concurrent               diagnoses of another disease).

* **Mortality rate** 
    The *PMSLT* calculations for changes in life years (and health-adjusted life years)     and mortality outcomes is based on observed mortality rates for the population of      interest. In the *ITHIM* model, if burden of disease estimates from the Global B       Burden of Disease (GBD) study are used, then, the mortality component is based on      the highest attained life expectancy observed in the world.

* **Impact of disability in increased life expectancy**
    In GBD studies, YLLs are not adjusted for disability; hence, their use in              estimating intervention effects results in over-estimation, which the *PMSLT*          approach avoids. Another way of seeing this is that estimated changes in morbidity     using the *ITHIM* do not allow for how implicit increases in life expectancy impact     on morbidity. While the changes in deaths and prevalence using the *PMSLT* are in      some ways more accurate than those from the *ITHIM* approach it should be noted        that that the average age of death and incident disease will change and thus the       disease burden will be on average be shifted later in life (which is a realistic       approach).

# R development

The model is set up as a long script to perform the required mathematical calculations. Where possible, we wrote functions and loops to avoid repetition. We set up the model  with Australian data, for Melbourne. Figure 2 is depicts the PMSLT model framework, which was followed in the code development.

![Figure 2. Proportional multi-state life-table simplified framework. *The simplied PMST shows the interaction between the life table, disease life table and potential impact fraction (PIF). The PIF calculations by age and sex group are the same as those generated for ITHIM. The PIF (or 1-PIF) modifies incidence of disease, which changes prevalence and mortality (disease specific life table). Changes in prevalence and mortality rates from the disease specific life tables feed into the life table by changing all-cause mortality, which in turn changes life years. Change in prevalence of diseases changes total years lived with disability, which in turn modifies health-adjusted life years*](structure\Figure2.png)

In what follows, first, we specify input parameters. Second, we present the code with explaining notes. Third, we present examples of outcomes and lastly we comment on topics related to implementation. Here we only included the physical activity health pathway. In the comments section, implementation of exposure to air pollution and road trauma is discussed. Note that in the presentation of input parameters, those needed to calculate PIFs are excluded, as these are common to the ITHIM, expect if trends are included (refer to comments section)

## Inputs (ADD A SUMMARY TABLE with inputs and potential sources of data)

We specify data requirements for the life table and disease life tables (Figure 2) and potential sources.

### **Life table**:

Inputs of the life table are: population numbers by sex (per 1-year or age grouping of interest), mortality rates or probability of all cause mortality by single age group and sex and total prevalent years lived with disability rate per one year by sex.

####	Population numbers

These data will be provided by the synthetic population. In the code presented here, we created 5-year age and sex cohorts from one-year age groups data. I left potential data sources below as a reference.

Data source: (1) National census; (2) Worldwide population and mortality data:  http://www.mortality.org/ (mostly high income countries; and (3) Calculate from GBD data (rates and numbers).

#### Mortality rates

These data will be provided by the synthetic population. Other sources are the same sources as above for population numbers. Sometimes mortality rates are in age groups (1-4, 5-9, etc). Interpolation can be used to derive in between ages rates (cubic spline). Here, mortality rates are needed per single year and sex.

Note that we need data for population numbers and all cause mortality rates for: (1) PMSLT and (2) Dismod II collection (more in Dismod II section).

####	Total prevalent years lived with disability rates per single year and sex.

These data is available from the GBD (http://ghdx.healthdata.org/gbd-results-tool) per 5-year age groups. We can use interpolation to derive between ages rates.

### **Disease life tables**

#### Incidence and case fatality

For each of the modeled diseases the PMSLT needs incidence and case fatality rates per sex and one-year intervals. Data from the GBD studies with Dismod II (free at https://www.epigear.com/index_files/dismod_ii.html) can be used to derive internally consistent data and generate missing data. For example, the GBD studies provide data for incidence, prevalence and disease mortality, however, not case fatality. Other national level sources may also be explored/used, and compare with estimates produce from GBD data and Dismod II.

**Dismod II** inputs are: (1) population numbers and mortality rates and (2) disease specific inputs.

*Population and mortality*

Within Dismod II, each setting (e.g. country) has a collection that consists of population numbers (preferably the same as used in GBD studies, due to the mortality envelop) and all- cause mortality rates (numbers and calculate rates). The GBD provides 5-year age groups that are acceptable input parameters for Dismod II.

*Disease inputs by age group and sex*

Each setting collection has a given number of diseases. Dismod II works with at least three of: case fatality, prevalence, incidence, mortality (disease), case fatality, remission, duration and the relative risk for mortality. So far, we have been assuming that remission is zero for chronic diseases, that is to say, when people become diseased, they do not recover. Special care should be taken with this assumption, as the GBD data assumes remission for some diseases, for example cancers, where after 10 years cases recover, except for long term sequelae. Since GBD now provides prevalence, incidence and mortality, it may be best to use all three as Dismod II input parameters to compare the effect of the remission assumption by the GBD for some diseases.

#### Disability weights (quality of life weights)

Disability weights (DW) can be derived from disease specific years lived with disability (YLD) and disease specific prevalence by age group (5 years) and sex. Data for YLDs prevalence can be obtained from the online GBD data tool (http://ghdx.healthdata.org/gbd-results-tool). Our calculations of DW in the example here based on the GBD methods for estimating YLDs as the sum of sequelae prevalence multiplied by sequelae disability weights (REF GBD). The GBD has publicly available data at the cause level (e.g. ischemic heart disease) instead of squelae level (e.g. myocardial infarction, angina and heart failure). However, the GBD disability weights are for health states associated with sequelae, hence, we need to calculate DWs. An age and sex specific-correction was introduced to couteract the effects of accumulating comorbid illnesses in the older age groups (Equation 1).

\begin{equation}
\label{DW adjusted for total YLDs-YLDdPd1YLDt}
(YLDd/Pd)/(1-YLDt) = DW adjusted for total YLDs
\end{equation}

Where YLDd is the YLD mean number per age and sex for a given disease, Pd is the prevalence (as reported in GBD )  for a given disease by age and sex and YLDt is total YLD rate per age and sex.

```{r}
inputs <- matrix(c("Life table", "Synthetic population", "", "Life table", "Synthetic population", "", "Life table", "GBD Results tool (http://ghdx.healthdata.org/gbd-results-tool)","", "Disease life table", "GBD Results tool (http://ghdx.healthdata.org/gbd-results-tool) and Dismod II", "", "Disease life table", "GBD Results tool (http://ghdx.healthdata.org/gbd-results-tool) and Dismod II", "", "", "Disease life table", "GBD Results tool (http://ghdx.healthdata.org/gbd-results-tool)"),ncol= 3,byrow=TRUE)
colnames(inputs) <- c("Life table/Disease life table","","")
rownames(inputs) <- c("Population","All-cause mortality", "Total YLD rates", "Incidence", "Case fatality", "Disability weighs")
inputs <- as.table(inputs)
inputs
```

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-03-28 MSLT/code/functions.R

@@ -1,0 +1,410 @@
+#####Packages
+
+require(dplyr)
+require(tidyverse)
+
+#############################Explanation method###################################
+
+#####Briefly, the proportional multi-state multi cohort life table consists of a 
+#####general life table and a life table for each of the modelled diseases.
+#####The diseases are those associated to the studied risk factor/s. 
+#####The link between the general life table and disease life tables is via the 
+#####potential impact fraction (pif), also called paf (population attributable fraction)
+#####The pif combines exposure to the risk factor and relative risks. The pif is 
+#####appleid to modify incidence in the individual disease life tables, which in turn
+#####modify prevalence and mortality. Changes in mortality and prevalence rates
+#####feed bacak into the general life table to modify total mortality and disability. 
+####Changes in total mortality impact on life years and changes in disability impact 
+####the disability adjusment of life years. 
+
+#### Method reference: 1.	Barendregt JJ, Oortmarssen vGJ, Murray CJ, Vos T. A generic model for the assessment of disease epidemiology: the computational basis of DisMod II. Popul Health Metr. 2003;1(1):4-.
+
+
+####The functions: run_life_table, run_disease, and run_pif serve to generate outputs
+####per age and sex (see model.R). The function run_output to facilitate the generation
+####of results. 
+
+
+#############################Function for general life table###################################
+#####Function to generate age and sex life tables. The function is then use in the model.R script
+#####to calculate variables for the baseline and scenario life tables. 
+
+run_life_table <- function(in_idata, in_sex, in_mid_age)
+{
+  
+  # Create a life table data frame
+  
+  # Filter in_idata by age and select columns for lifetable calculations
[…]

### 2018-03-28 MSLT/code/model.R

@@ -1,0 +1,589 @@
+##########################Prepare data and environment to run code###########################################################
+
+##### Clean the Global Environment
+
+rm (list = ls())
+
+##### Prevent scientific notation in data frame
+
+options(scipen=999)
+
+##### Load all functions/packages
+
+source("code/functions.R")
+
+
+##### Read data. idata: life table and disease life tables (including trends). edata: exposure data (e.g. physical activity).
+#### irr.data: relative risks data and ee: energy expenditure
+#### For exposure (edata) and input (idata) data I am using already generated categories, but a function should be developed to work from raw data. 
+
+idata <- read.csv("data/idata.csv", stringsAsFactors = F)
+edata <- read.csv("data/edata.csv", stringsAsFactors = F)
+irr <- read.csv("data/irr.csv", stringsAsFactors = F)
+ee <- read.csv("data/ee.csv", stringsAsFactors = F)
+
+#####To do
+
+## Note that disease trends are excluded for now. We need to generate a trends function for diseases, mortality (all cause), 
+## change in disease due to change in risk factor exposure (discuss with Niel M), road injuries effect over time. 
+## Eventually, we want to add (1) CRA ITHIM approach to compare results and (2) use all cause mortaltiy instead of 
+## disease specific dose responses for incidence to understand the difference in results for all cause mortality. 
+
+#################################Model parameters#########################################################################
+
+p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
+
+p_sex <- c("males", "females")
+
+p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
+
+###As an expample, an increase in 1000 MET-mins per week
+
+p_intervention_effect <- 100
+
+##############################Prepare general life table (age, death rate, population #)########
+
+##### Make lower cases values for sex column
+
+idata$sex <- tolower(idata$sex)
+
+##### Rename mortality_rate to mx
+
+idata$mx <- idata$mortality_rate
+
+##### Remove mortality_rate column
+
+idata$mortality_rate <- NULL
+
+##### Rename male for males and female for females in edata to match idata. 
+
+edata$sex[edata$sex=="male"] <- "males"
[…]

### 2018-03-29 MSLT/MSLTdoc.Rmd
Creating inputs table

@@
   geometry: margin=1in
-  bibliography: MSLT\Bibliography.txt
+
 ---
@@
 
-```{r}
-inputs <- matrix(c("Life table", "Synthetic population", "", "Life table", "Synthetic population", "", "Life table", "GBD Results tool (http://ghdx.healthdata.org/gbd-results-tool)","", "Disease life table", "GBD Results tool (http://ghdx.healthdata.org/gbd-results-tool) and Dismod II", "", "Disease life table", "GBD Results tool (http://ghdx.healthdata.org/gbd-results-tool) and Dismod II", "", "", "Disease life table", "GBD Results tool (http://ghdx.healthdata.org/gbd-results-tool)"),ncol= 3,byrow=TRUE)
-colnames(inputs) <- c("Life table/Disease life table","","")
-rownames(inputs) <- c("Population","All-cause mortality", "Total YLD rates", "Incidence", "Case fatality", "Disability weighs")
+```{r, echo=FALSE}
+inputs <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18)
+colnames(inputs) <- c("Life table","Source","Comments")
+rownames(inputs) <- c("Population", "All-cause mortality", "Total YLD rate", "Incidence", "Case fatality", "Disability weight")
 inputs <- as.table(inputs)
 inputs
-```
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
 ```

### 2018-04-03 MSLT/MSLTdoc.Rmd
Adding table with PMSLT inputs

@@
 ```{r, echo=FALSE}
-inputs <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18)
-colnames(inputs) <- c("Life table","Source","Comments")
-rownames(inputs) <- c("Population", "All-cause mortality", "Total YLD rate", "Incidence", "Case fatality", "Disability weight")
-inputs <- as.table(inputs)
-inputs
+pmslt_inputs <- matrix(c(1:15), ncol = 3, byrow = TRUE)
+colnames(pmslt_inputs) <- c("Input", "Source", "Comments")
+rownames (pmslt_inputs) <- c("Population numbers", "Death rates", "Years lived with disability rates", "Disease epidemilogy", "Disability weights")
+print(pmslt_inputs)
 ```

### 2018-04-03 MSLT/MSLTdoc.Rmd
Added inputs table

@@
     - \usepackage{setspace}\onehalfspacing
+
 output:
@@
 ```{r setup, include=FALSE}
+
+#Global options
 knitr::opts_chunk$set(echo = TRUE)
+
+#Packages to run code
+require(dplyr)
+require(tidyverse)
+require(knitr)
+require(kableExtra)
+options(knitr.table.format = 'latex')
 ```
@@
 
-## Inputs (ADD A SUMMARY TABLE with inputs and potential sources of data)
+## Inputs
 
@@
 
-####	Total prevalent years lived with disability rates per single year and sex.
+####	Total years lived with disability rates per single year and sex.
 
@@
 ```{r, echo=FALSE}
-pmslt_inputs <- matrix(c(1:15), ncol = 3, byrow = TRUE)
-colnames(pmslt_inputs) <- c("Input", "Source", "Comments")
-rownames (pmslt_inputs) <- c("Population numbers", "Death rates", "Years lived with disability rates", "Disease epidemilogy", "Disability weights")
-print(pmslt_inputs)
+pmslt_inputs <- tribble(
+~Input, ~Source, ~Comments, 
+"Life table", "Synthetic population per sex and age group", "Age grouping in life table to match synthetic population", 
+"Life table", "Synthetic population per sex and one-year age group", "If one year age group is not avabilable it can be derive using interpolation from age groups data", 
+"Life table", "Global Burden of Disease (GBD) study per one-year age group and sex", "GBD data is in five-year age groups, interpolation to derive one-year age groups", 
+"Disease life table", "GBD data for prevalence, incidence and mortality and DISMOD II", "Two step process.
+First obtain disease and population data from GBD.
+Second, use Dismod II to derive internally consistent estimates for incidence and case fatality (PMSLT disease life table iputs",
+"Disease life table", "Derive from disease prevalence and years lived with disability from GBD", "Adjustments for comorbidities in later years of life to be applied"
+)
+
+kable(pmslt_inputs,"latex", booktabs = T, caption = "PMSLT inputs") %>%
+ kable_styling(latex_options = c("striped", "hold_position"),
+               full_width = F)%>%
+  column_spec(2, width = "20em") %>%
+  column_spec(3, width = "20em")
 ```

### 2018-04-03 MSLT/code/functions.R

@@ -2,6 +2,8 @@
 
 require(dplyr)
 require(tidyverse)
+require(knitr)
+require(kableExtra)
 
 #############################Explanation method###################################
 

### 2018-04-03 MSLT/MSLTdoc.Rmd
Added packages to include citations

@@
   geometry: margin=1in
+  citation_package: natib
+  bibliography: Bibliography.bib
 
@@
 
-The model was developed by Jan Barendregt and colleagues (REF ACE prevention and core papers) and has been widely used in Australia and New Zealand (Ref ACE-report and BODE website of key publications).
+The model was developed by Jan Barendregt and colleagues  and has been widely used in Australia and New Zealand ().
 
@@
 Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+
+#References

### 2018-04-03 MSLT/code/functions.R

@@ -4,6 +4,7 @@ require(dplyr)
 require(tidyverse)
 require(knitr)
 require(kableExtra)
+require(citr)
 
 #############################Explanation method###################################
 

### 2018-04-03 MSLT/MSLTdoc.Rmd
Added bibliography file

@@
     - \usepackage{setspace}\onehalfspacing
-
 output:
@@
    geometry: margin=1in
-  citation_package: natib
 bibliography: Bibliography.bib
-
 ---
@@
 require(kableExtra)
+require(citr)
 options(knitr.table.format = 'latex')
@@
 
-The model was developed by Jan Barendregt and colleagues  and has been widely used in Australia and New Zealand ().
+The model was developed by Jan Barendregt and colleagues  and has been widely used in Australia and New Zealand [@RN38].
 

### 2018-04-03 MSLT/MSLTdoc.Rmd
Added missing refs

@@
 
-The model was developed by Jan Barendregt and colleagues  and has been widely used in Australia and New Zealand [@RN38].
+The model was developed by Jan Barendregt and colleagues and has been widely used in Australia and New Zealand [@RN38; @RN8299].
 
@@
 Specific types of HALY are the quality-adjusted life year (QALY) and the disability-adjusted life year (DALY).
-The QALY derives from economics and was first used in the 1960s as a measure of health gain (ref Gold MR, Stevenson D, Fryback DG.
-HALYs and QALYs and DALYs, oh my: similarities and differences in summary measures of population health.
-Annual review of public health.
-2002;23(1):115-34.).
-The disability-adjusted life-year (DALY) was developed for use in burden of disease studies as a measure of health loss due to disease (Gold et al.).
+The QALY derives from economics and was first used in the 1960s as a measure of health gain [@RN8158].
+The disability-adjusted life-year (DALY) was developed for use in burden of disease studies as a measure of health loss due to disease [@RN8158].
 Our calculated HALYs are neither QALYs not DALYs, but something in between.
@@
 However, the main difference is in the calculation of the health-related quality of life component.
-QALYs use measures of utility weights that traditionally represent individual experiences of health, whereas our estimated HALYs use disability weights linked to specific diseases, which were developed for the Global Burden of Disease study (GOld et al).
-As discussed in past research (REF Roux and Cobiac 2009) the main advantage of using disability weights over utility weights is that disability weights refer to specific diseases rather than health states.
-We opted to use the more general terms HALYs given that the use of the DALYs terminology may lead to think that our calculations are similar to those in burden of diseases studies (REF Murray 1997).
+QALYs use measures of utility weights that traditionally represent individual experiences of health, whereas our estimated HALYs use disability weights linked to specific diseases, which were developed for the Global Burden of Disease study [@RN8158].
+As discussed in past research [@RN21; @RN8262] the main advantage of using disability weights over utility weights is that disability weights refer to specific diseases rather than health states.
+We opted to use the more general terms HALYs given that the use of the DALYs terminology may lead to think that our calculations are similar to those in burden of diseases studies [@RN8153].
 In our study, our model does not explicitly separate years of life lost (YLL) and years lived with disability (YLD) components, but instead calculates the total number of life years lived, adjusted for the average health-related quality of life in those years (by age and sex).
@@
 
-The PMSLT similar to ITHIM is a comparative risk assessment approach (REF BRIGGS 2016) that consist of calculating the change in the health burden for a population of interest from a change in exposure to health risks factors (e.g.
+The PMSLT similar to ITHIM is a comparative risk assessment approach [@RN8957] that consist of calculating the change in the health burden for a population of interest from a change in exposure to health risks factors (e.g.
 physical inactivity, air pollution and road trauma).As depicted in Figure 1, both methods need estimates of the potential impact fraction (PIF), which indicates the proportion of the disease burden attributable to a risk factor of interest (e.g.
-physical inactivity).
+physical inactivity) [@RN42].
 A step further back, is the development of scenarios that bring about change in the distribution of the risk factor of interest.
@@
 * **Interaction between multiple diseases**
-    The *PMSLT* accounts for the interaction between multiple diseases, with               proportions of the population being able to be in more than one health state (REF      BRIGGS 2016).
+    The *PMSLT* accounts for the interaction between multiple diseases, with proportions of the population being     able to be in more than one health state [@RN8957].
 This avoids  overestimation of outcomes as a result of       summing health outcomes attributable to each disease individually as done in *ITHIM*.
@@
     The *PMSLT* calculations for changes in life years (and health-adjusted life years) and mortality outcomes is     based on observed mortality rates for the population of interest.
-In the *ITHIM* model, if burden of disease estimates from the Global B       Burden of Disease (GBD) study are used, then, the mortality component is based on      the highest attained life expectancy observed in the world.
+In the *ITHIM* model, if burden of disease     estimates from the Global Burden of Disease (GBD) study are used, then, the mortality component is based on      the highest attained life expectancy observed in the world.
 
@@
 In the comments section, implementation of exposure to air pollution and road trauma is discussed.
-Note that in the presentation of input parameters, those needed to calculate PIFs are excluded, as these are common to the ITHIM, expect if trends are included (refer to comments section)
+Note that in the presentation of input parameters, those needed to calculate PIFs are excluded, as these are common to the ITHIM, expect if trends are included (refer to comments section).
 

### 2018-04-03 MSLT/MSLTdoc.Rmd
Start code explanation

@@
 
-#Packages to run code
+#The work in this document requires the following packages (code, rmarkdown and citing packages):
 require(dplyr)
@@
 
+##Code
+
 ## Including Plots

### 2018-04-03 MSLT/MSLTdoc.Rmd
Road trauma comments methods

@@
 However, the GBD disability weights are for health states associated with sequelae, hence, we need to calculate DWs.
-An age and sex specific-correction was introduced to couteract the effects of accumulating comorbid illnesses in the older age groups (Equation 1).
+An age and sex specific-correction was introduced to counteract the effects of accumulating comorbid illnesses in the older age groups (Equation 1).
 
@@
 
+#Comments
+
+##Road injuries in the PMsLT
+
+The disease model used in each of the disease life table is not directly applicable to road injuries, however, similar concept can be follow.
+Firstly, changes in road fatalities impact on the overall mortality rate, hence, by knowing the road fatality rates for baseline and scenarios, we will be able to incorporate changes to mortality attributable to road fatalities.
+For road injuries, methods developed by Kavi Bhalla and Marko Tanio (REFS) that derive the average YLD attributable to life long and short term injuries can be applied to derive the change in total YLDs (CHECK THAT THESE WERE DEVELOPED AS INCIDENCE YLDs).MT's methods assumes that injuries do not reduce the life expectancy of the injured person.
+
 #References

### 2018-04-04 MSLT/MSLTdoc.Rmd
Add code to Rmarkdown doc

@@
     - \usepackage{setspace}\onehalfspacing
+    - \usepackage{float}
 output:
@@
      fig_caption: yes
+
    fontsize: 12 pt
@@
 require(citr)
+require(png)
 options(knitr.table.format = 'latex')
+
 ```
@@
 kable(pmslt_inputs,"latex", booktabs = T, caption = "PMSLT inputs") %>%
- kable_styling(latex_options = c("striped", "hold_position"),
+ kable_styling(latex_options = c("striped", "hold position"),
                full_width = F)%>%
-  column_spec(2, width = "20em") %>%
-  column_spec(3, width = "20em")
+  column_spec(2, width = "15em") %>%
+  column_spec(3, width = "15em") %>%
+  row_spec(0, bold = T)
+```
+\pagebreak
+
+##Code
+
+Following the structure of Figure 2, we developed functions to perform sex and age cohorts calculations for the life table, disease life tables and potential impact fractions: run_life_table, run_disease and and run_pif.
+We also generated two functions for outputs: plot_outputs and gen_aggregate.
+The function plot_outputs creates age-group and sex linear plots for specified outcomes (e.g.
+health-adjusted life years, incidence of diabetes) and gen_aggregate adds up each cohort results.
+Functions were then used in a code script.
+In what follows, we explain each step in the development of the script.
+Here we also include code chunks, however, we also kept them separately in the MSLT folder, in the code file.
+
+In what follows, we start with the **model** script file, and explain the **functions** script file as these are used to perform calculations.
+The **functions** will be explained in detailed to provide clarity for required inputs.
+
+###Set up
+
+We start by *cleaning the global environment* to keep track of our works and ensure that the code is generating our outcomes.
+Then, we set up an option to avoind the use of scientific notation and lastly we load the functions.
+The code chunks are shown in the rmarkdown ouput.
+
+```{r, echo=TRUE}
+rm (list = ls())
+
+##### Prevent scientific notation in data frame
+
+options(scipen=999)
+
+##### Load all functions/packages
+
+source("MSLT/code/functions.R")
 ```
 
-##Code
+The data used here is for an example for Australia, however, this will be updated to link with the ITHIM-R modules (synthetic population, PIF calculations).
+
+```{r, echo=TRUE}
+##### Read data. idata: life table and disease life tables (including trends). edata: exposure data (e.g. physical activity).
+#### irr.data: relative risks data and ee: energy expenditure
+#### For exposure (edata) and input (idata) data I am using already generated categories, but a function should be developed to work from raw data.
+
+idata <- read.csv("MSLT/data/idata.csv", stringsAsFactors = F)
+edata <- read.csv("MSLT/data/edata.csv", stringsAsFactors = F)
+irr <- read.csv("MSLT/data/irr.csv", stringsAsFactors = F)
+ee <- read.csv("MSLT/data/ee.csv", stringsAsFactors = F)
+```
 

### 2018-04-04 MSLT/MSLTdoc.tex
Add code to Rmarkdown doc

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Proportional multi-state multiple-cohort life table model},
            pdfauthor={Belen Zapata-Diomedi and Ali Abbas},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Proportional multi-state multiple-cohort life table model}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Belen Zapata-Diomedi and Ali Abbas}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{26 March 2018}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage[normalem]{ulem}

\usepackage{setspace}\onehalfspacing
\usepackage{float}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\section{Introduction}\label{introduction}

The proportional multi-state multiple-cohort life table model (PMSLT) is
a population level model (macro) approach to simulate health (and
economic) implications of changes in exposure to health risk factors
(e.g.~physical inactivity, air pollution and diet). The PMSLT has been
widely used to simulate outcomes for population level interventions for
the reduction of chronic diseases.

The model was developed by Jan Barendregt and colleagues and has been
widely used in Australia and New Zealand (T. Vos et al. 2010; Blakely et
al. 2015).

The basic infrastructure of the model consist of three components: (1)
Effect size for the intervention of interest (e.g.~intervention to urban
design that modifies population levels of physical activity); (2)
Calculation of the potential impact fraction (PIF) to derive the change
in occurrence of disease (incidence rate/mortality rate) attributable to
a change in the distribution of the risk factor (e.g.~physical
activity); and (3) Use of the PMSLT to simulate health (and economic)
outcomes attributable to a change in the distribution of health risk
factor/s in the population of interest. Figure 1 summaries the basic
infrastructure of the model. ITHIM is included in Figure 1 to show that
both approaches share in common steps one and two and differ in the
mechanisms of calculation of change in health burden.
\includegraphics{structure/Figure1.png}

\textbf{HALYs, QALYs and DALYs}

In this model we use the term `health-adjusted life year' (HALY). As
`summary measure of population health' it measures both quantity and
quality of life, where one HALY represent the equivalent of one year in
full health (which could be two years with a quality of life of 0.5, for
example). Specific types of HALY are the quality-adjusted life year
(QALY) and the disability-adjusted life year (DALY). The QALY derives
from economics and was first used in the 1960s as a measure of health
gain (Gold, Stevenson, and Fryback 2002). The disability-adjusted
life-year (DALY) was developed for use in burden of disease studies as a
measure of health loss due to disease (Gold, Stevenson, and Fryback
2002). Our calculated HALYs are neither QALYs not DALYs, but something
in between. They are similar to QALYs in that they represent health
gains. However, the main difference is in the calculation of the
health-related quality of life component. QALYs use measures of utility
weights that traditionally represent individual experiences of health,
whereas our estimated HALYs use disability weights linked to specific
diseases, which were developed for the Global Burden of Disease study
(Gold, Stevenson, and Fryback 2002). As discussed in past research (L.
Cobiac, Vos, and Barendregt 2009; Roux, Pratt, and Tengs 2008) the main
advantage of using disability weights over utility weights is that
disability weights refer to specific diseases rather than health states.
We opted to use the more general terms HALYs given that the use of the
DALYs terminology may lead to think that our calculations are similar to
those in burden of diseases studies (Murray et al. 2012). In our study,
our model does not explicitly separate years of life lost (YLL) and
years lived with disability (YLD) components, but instead calculates the
total number of life years lived, adjusted for the average
health-related quality of life in those years (by age and sex). In
burden of disease studies, DALYs are defined as the sum Years of Life
Lost (YLL) and Years Lived with Disability (YLD).

\subsection{Contribution to ITHIMR}\label{contribution-to-ithimr}

The PMSLT similar to ITHIM is a comparative risk assessment approach
(Briggs, Scarborough, and Smith 2016) that consist of calculating the
change in the health burden for a population of interest from a change
in exposure to health risks factors (e.g.~physical inactivity, air
pollution and road trauma).As depicted in Figure 1, both methods need
estimates of the potential impact fraction (PIF), which indicates the
proportion of the disease burden attributable to a risk factor of
interest (e.g.~physical inactivity) (Barendregt and Veerman 2010). A
step further back, is the development of scenarios that bring about
change in the distribution of the risk factor of interest. For now, we
only focus on calculations from the PIF onward, and provide a
hypothetical example of change in population levels of physical
activity. Incorporation of additional health risk factor (air pollution,
road trauma, NO2 and noise) will be discussed in the relevant code
sections.

\subsubsection{Difference between ITHIM and
PMSLT}\label{difference-between-ithim-and-pmslt}

\begin{itemize}
\item
  \textbf{Time component} The \emph{PMSLT} follows a population of
  interest over time. For example, as set up here, we simulate sex and
  age (5 years starting at 18) cohorts over time until they die or reach
  100 years of age. This implies that we can include trends for
  diseases, time lags between change in exposure to risk factors and
  change in health and demographic changes (e.g.~population growth). In
  addition, we can estimate yearly changes in the burden of diseases
  over the life course or for a specified number of years. The
  \emph{ITHIM} approach is a snapshot of change in burden for one year.
\item
  \textbf{Interaction between multiple diseases} The \emph{PMSLT}
  accounts for the interaction between multiple diseases, with
  proportions of the population being able to be in more than one health
  state (Briggs, Scarborough, and Smith 2016). This avoids
  overestimation of outcomes as a result of summing health outcomes
  attributable to each disease individually as done in \emph{ITHIM}. It
  is important to note that the \emph{PMSLT} assumes that diseases are
  independent of each other. That is to say, developing a disease is
  unrelated to a concurrent diagnoses of another disease).
\item
  \textbf{Mortality rate} The \emph{PMSLT} calculations for changes in
  life years (and health-adjusted life years) and mortality outcomes is
  based on observed mortality rates for the population of interest. In
  the \emph{ITHIM} model, if burden of disease estimates from the Global
  Burden of Disease (GBD) study are used, then, the mortality component
  is based on the highest attained life expectancy observed in the
  world.
\item
  \textbf{Impact of disability in increased life expectancy} In GBD
  studies, YLLs are not adjusted for disability; hence, their use in
  estimating intervention effects results in over-estimation, which the
  \emph{PMSLT} approach avoids. Another way of seeing this is that
  estimated changes in morbidity using the \emph{ITHIM} do not allow for
  how implicit increases in life expectancy impact on morbidity. While
  the changes in deaths and prevalence using the \emph{PMSLT} are in
  some ways more accurate than those from the \emph{ITHIM} approach it
  should be noted that that the average age of death and incident
  disease will change and thus the disease burden will be on average be
  shifted later in life (which is a realistic approach).
\end{itemize}

\section{R development}\label{r-development}

The model is set up as a long script to perform the required
mathematical calculations. Where possible, we wrote functions and loops
to avoid repetition. We set up the model with Australian data, for
Melbourne. Figure 2 is depicts the PMSLT model framework, which was
followed in the code development.

\begin{figure}
\centering
\includegraphics{structure/Figure2.png}
\caption{Figure 2. Proportional multi-state life-table simplified
framework. \emph{The simplied PMST shows the interaction between the
life table, disease life table and potential impact fraction (PIF). The
PIF calculations by age and sex group are the same as those generated
for ITHIM. The PIF (or 1-PIF) modifies incidence of disease, which
changes prevalence and mortality (disease specific life table). Changes
in prevalence and mortality rates from the disease specific life tables
feed into the life table by changing all-cause mortality, which in turn
changes life years. Change in prevalence of diseases changes total years
lived with disability, which in turn modifies health-adjusted life
years}}
\end{figure}

In what follows, first, we specify input parameters. Second, we present
the code with explaining notes. Third, we present examples of outcomes
and lastly we comment on topics related to implementation. Here we only
included the physical activity health pathway. In the comments section,
implementation of exposure to air pollution and road trauma is
discussed. Note that in the presentation of input parameters, those
needed to calculate PIFs are excluded, as these are common to the ITHIM,
expect if trends are included (refer to comments section).

\subsection{Inputs}\label{inputs}

We specify data requirements for the life table and disease life tables
(Figure 2) and potential sources.

\subsubsection{\texorpdfstring{\textbf{Life
table}:}{Life table:}}\label{life-table}

Inputs of the life table are: population numbers by sex (per 1-year or
age grouping of interest), mortality rates or probability of all cause
mortality by single age group and sex and total prevalent years lived
with disability rate per one year by sex.

\paragraph{Population numbers}\label{population-numbers}

These data will be provided by the synthetic population. In the code
presented here, we created 5-year age and sex cohorts from one-year age
groups data. I left potential data sources below as a reference.

Data source: (1) National census; (2) Worldwide population and mortality
data: \url{http://www.mortality.org/} (mostly high income countries; and
(3) Calculate from GBD data (rates and numbers).

\paragraph{Mortality rates}\label{mortality-rates}

These data will be provided by the synthetic population. Other sources
are the same sources as above for population numbers. Sometimes
mortality rates are in age groups (1-4, 5-9, etc). Interpolation can be
used to derive in between ages rates (cubic spline). Here, mortality
rates are needed per single year and sex.

Note that we need data for population numbers and all cause mortality
rates for: (1) PMSLT and (2) Dismod II collection (more in Dismod II
section).

\paragraph{Total years lived with disability rates per single year and
sex.}\label{total-years-lived-with-disability-rates-per-single-year-and-sex.}

These data is available from the GBD
(\url{http://ghdx.healthdata.org/gbd-results-tool}) per 5-year age
groups. We can use interpolation to derive between ages rates.

\subsubsection{\texorpdfstring{\textbf{Disease life
tables}}{Disease life tables}}\label{disease-life-tables}

\paragraph{Incidence and case
fatality}\label{incidence-and-case-fatality}

For each of the modeled diseases the PMSLT needs incidence and case
fatality rates per sex and one-year intervals. Data from the GBD studies
with Dismod II (free at
\url{https://www.epigear.com/index_files/dismod_ii.html}) can be used to
derive internally consistent data and generate missing data. For
example, the GBD studies provide data for incidence, prevalence and
disease mortality, however, not case fatality. Other national level
sources may also be explored/used, and compare with estimates produce
from GBD data and Dismod II.

\textbf{Dismod II} inputs are: (1) population numbers and mortality
rates and (2) disease specific inputs.

\emph{Population and mortality}

Within Dismod II, each setting (e.g.~country) has a collection that
consists of population numbers (preferably the same as used in GBD
studies, due to the mortality envelop) and all- cause mortality rates
(numbers and calculate rates). The GBD provides 5-year age groups that
are acceptable input parameters for Dismod II.

\emph{Disease inputs by age group and sex}

Each setting collection has a given number of diseases. Dismod II works
with at least three of: case fatality, prevalence, incidence, mortality
(disease), case fatality, remission, duration and the relative risk for
mortality. So far, we have been assuming that remission is zero for
chronic diseases, that is to say, when people become diseased, they do
not recover. Special care should be taken with this assumption, as the
GBD data assumes remission for some diseases, for example cancers, where
after 10 years cases recover, except for long term sequelae. Since GBD
now provides prevalence, incidence and mortality, it may be best to use
all three as Dismod II input parameters to compare the effect of the
remission assumption by the GBD for some diseases.

\paragraph{Disability weights (quality of life
weights)}\label{disability-weights-quality-of-life-weights}

Disability weights (DW) can be derived from disease specific years lived
with disability (YLD) and disease specific prevalence by age group (5
years) and sex. Data for YLDs prevalence can be obtained from the online
GBD data tool (\url{http://ghdx.healthdata.org/gbd-results-tool}). Our
calculations of DW in the example here based on the GBD methods for
estimating YLDs as the sum of sequelae prevalence multiplied by sequelae
disability weights (REF GBD). The GBD has publicly available data at the
cause level (e.g.~ischemic heart disease) instead of squelae level
(e.g.~myocardial infarction, angina and heart failure). However, the GBD
disability weights are for health states associated with sequelae,
hence, we need to calculate DWs. An age and sex specific-correction was
introduced to counteract the effects of accumulating comorbid illnesses
in the older age groups (Equation 1).

\begin{equation}
\label{DW adjusted for total YLDs-YLDdPd1YLDt}
(YLDd/Pd)/(1-YLDt) = DW adjusted for total YLDs
\end{equation}

Where YLDd is the YLD mean number per age and sex for a given disease,
Pd is the prevalence (as reported in GBD ) for a given disease by age
and sex and YLDt is total YLD rate per age and sex.

\rowcolors{2}{gray!6}{white}

\begin{table}

\caption{\label{tab:unnamed-chunk-1}PMSLT inputs}
\centering
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{15em}>{\raggedright\arraybackslash}p{15em}}
\hiderowcolors
\toprule
\textbf{Input} & \textbf{Source} & \textbf{Comments}\\
\midrule
\showrowcolors
Life table & Synthetic population per sex and age group & Age grouping in life table to match synthetic population\\
Life table & Synthetic population per sex and one-year age group & If one year age group is not avabilable it can be derive using interpolation from age groups data\\
Life table & Global Burden of Disease (GBD) study per one-year age group and sex & GBD data is in five-year age groups, interpolation to derive one-year age groups\\
Disease life table & GBD data for prevalence, incidence and mortality and DISMOD II & Two step process. First obtain disease and population data from GBD. Second, use Dismod II to derive internally consistent estimates for incidence and case fatality (PMSLT disease life table iputs\\
Disease life table & Derive from disease prevalence and years lived with disability from GBD & Adjustments for comorbidities in later years of life to be applied\\
\bottomrule
\end{tabular}
\end{table}

\rowcolors{2}{white}{white} \pagebreak

\subsection{Code}\label{code}

Following the structure of Figure 2, we developed functions to perform
sex and age cohorts calculations for the life table, disease life tables
and potential impact fractions: run\_life\_table, run\_disease and and
run\_pif. We also generated two functions for outputs: plot\_outputs and
gen\_aggregate. The function plot\_outputs creates age-group and sex
linear plots for specified outcomes (e.g.~health-adjusted life years,
incidence of diabetes) and gen\_aggregate adds up each cohort results.
Functions were then used in a code script. In what follows, we explain
each step in the development of the script. Here we also include code
chunks, however, we also kept them separately in the MSLT folder, in the
code file.

In what follows, we start with the \textbf{mode} script file, and
explain the \textbf{functions} script file as these are used to perform
calculations. The \textbf{functions} will be explained in detailed to
provide clarity for required inputs.

\subsection{Including Plots}\label{including-plots}

You can also embed plots, for example:

\includegraphics{MSLTdoc_files/figure-latex/pressure-1.pdf}

Note that the \texttt{echo\ =\ FALSE} parameter was added to the code
chunk to prevent printing of the R code that generated the plot.

\section{Comments}\label{comments}

\subsection{Road injuries in the
PMsLT}\label{road-injuries-in-the-pmslt}

The disease model used in each of the disease life table is not directly
applicable to road injuries, however, similar concept can be follow.
Firstly, changes in road fatalities impact on the overall mortality
rate, hence, by knowing the road fatality rates for baseline and
scenarios, we will be able to incorporate changes to mortality
attributable to road fatalities. For road injuries, methods developed by
Kavi Bhalla and Marko Tanio (REFS) that derive the average YLD
attributable to life long and short term injuries can be applied to
derive the change in total YLDs (CHECK THAT THESE WERE DEVELOPED AS
INCIDENCE YLDs).MT's methods assumes that injuries do not reduce the
life expectancy of the injured person.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-RN42}{}
Barendregt, J.J., and J.L. Veerman. 2010. ``Categorical Versus
Continuous Risk Factors and the Calculation of Potential Impact
Fractions.'' Journal Article. \emph{J Epidemiol Community Health} 64
(3): 209--12.
doi:\href{https://doi.org/10.1136/jech.2009.090274}{10.1136/jech.2009.090274}.

\hypertarget{ref-RN8299}{}
Blakely, T., L. J. Cobiac, C. L. Cleghorn, A. L. Pearson, F. S. Deen, G.
Kvizhinadze, N. Nghiem, M. McLeod, and N. Wilson. 2015. ``Health, Health
Inequality, and Cost Impacts of Annual Increases in Tobacco Tax:
Multistate Life Table Modeling in New Zealand.'' Journal Article.
\emph{PLoS Med} 12.
doi:\href{https://doi.org/10.1371/journal.pmed.1001856}{10.1371/journal.pmed.1001856}.

\hypertarget{ref-RN8957}{}
Briggs, Adam, Peter Scarborough, and Adrian Smith. 2016. ``Modelling in
Public Health.'' Book Section. In \emph{Public Health Intelligence:
Issues of Measure and Method}, edited by Krishna Regmi and Ivan Gee,
67--90. Cham: Springer International Publishing.
doi:\href{https://doi.org/10.1007/978-3-319-28326-5_4}{10.1007/978-3-319-28326-5\_4}.

\hypertarget{ref-RN21}{}
Cobiac, L.J., T. Vos, and J.J. Barendregt. 2009. ``Cost-Effectiveness of
Interventions to Promote Physical Activity: A Modelling Study.'' Journal
Article. \emph{Plos Med} 6 (7): e1000110--e1000110.
doi:\href{https://doi.org/10.1371/journal.pmed.1000110}{10.1371/journal.pmed.1000110}.

\hypertarget{ref-RN8158}{}
Gold, Marthe R., David Stevenson, and Dennis G. Fryback. 2002. ``HALYs
and Qalys and Dalys, Oh My: Similarities and Differences in Summary
Measures of Population Health.'' Journal Article. \emph{Annu Rev Public
Health} 23 (1): 115--34.
doi:\href{https://doi.org/doi:10.1146/annurev.publhealth.23.100901.140513}{doi:10.1146/annurev.publhealth.23.100901.140513}.

\hypertarget{ref-RN8153}{}
Murray, Christopher J. L., Majid Ezzati, Abraham D. Flaxman, Stephen
Lim, Rafael Lozano, Catherine Michaud, Mohsen Naghavi, et al. 2012.
``GBD 2010: Design, Definitions, and Metrics.'' Journal Article.
\emph{The Lancet} 380 (9859): 2063--6.
doi:\href{https://doi.org/10.1016/S0140-6736(12)61899-6}{10.1016/S0140-6736(12)61899-6}.

\hypertarget{ref-RN8262}{}
Roux, L., M. Pratt, and T. O. Tengs. 2008. ``Cost Effectiveness of
Community-Based Physical Activity Interventions.'' Journal Article.
\emph{Am J Prev Med} 35.
doi:\href{https://doi.org/10.1016/j.amepre.2008.06.040}{10.1016/j.amepre.2008.06.040}.

\hypertarget{ref-RN38}{}
Vos, T., R. Carter, J. J Barendregt, Mihalopoulos C., JL. Veerman, A.
Magnus, L. Cobiac, Bertram MY., and AL. Wallace. 2010. ``Assessing
Cost-Effectiveness in Prevention (Ace-Prevention): Final Report.''
Report.

\end{document}

### 2018-04-04 MSLT/MSLTdoc.Rmd
R markdown not running code chunks
Fixed path for files

@@
 
-##### Prevent scientific notation in data frame
+# Prevent scientific notation in data frame
 
@@
 
-##### Load all functions/packages
+#Load all functions/packages
 
-source("MSLT/code/functions.R")
+source("code/functions.R")
 ```
@@
 
-idata <- read.csv("MSLT/data/idata.csv", stringsAsFactors = F)
-edata <- read.csv("MSLT/data/edata.csv", stringsAsFactors = F)
-irr <- read.csv("MSLT/data/irr.csv", stringsAsFactors = F)
-ee <- read.csv("MSLT/data/ee.csv", stringsAsFactors = F)
+idata <- read.csv("data/idata.csv", stringsAsFactors = F)
+edata <- read.csv("data/edata.csv", stringsAsFactors = F)
+irr <- read.csv("data/irr.csv", stringsAsFactors = F)
+ee <- read.csv("data/ee.csv", stringsAsFactors = F)
 ```

### 2018-04-04 MSLT/MSLTdoc.Rmd
Working on inputs description and link to framwork

@@
 
-The data used here is for an example for Australia, however, this will be updated to link with the ITHIM-R modules (synthetic population, PIF calculations).
+###Inputs
+
+Table 1 describes data needs for the PMSLT, the data used here is for an example for Australia, however, this will be updated to link with the ITHIM-R modules and additional work required to generate data inputs.
+Here, we expand on Table 1 with the aim of providing clear guidance for the PMSLT data needs.
 

### 2018-04-05 MSLT/MSLTdoc.Rmd
Inputs summary table

--- 2018-04-04 MSLT/MSLTdoc.Rmd
@@
      toc: true
+     toc_depth: 3
      number_sections: true
      fig_caption: yes
-
+     fig_width: 7
+     fig_height: 6
+     df_print: kable
    fontsize: 12 pt
@@
 
-![Figure 1. Basic model infrastructure.](structure\Figure1.png)
+![Basic ITHIMR infrastructure](structure\Figure1.png)
 
@@
 
-![Figure 2.
-Proportional multi-state life-table simplified framework.
+![Proportional multi-state life-table simplified framework.
 *The simplied PMST shows the interaction between the life table, disease life table and potential impact fraction (PIF).
@@
 
-### **Life table**:
+### Life table
 
 Inputs of the life table are: population numbers by sex (per 1-year or age grouping of interest), mortality rates or probability of all cause mortality by single age group and sex and total prevalent years lived with disability rate per one year by sex.
+Disease specific disability weights are presented as inputs here as these adjust the total years lived with disability, hence, the health-adjusted life years.
 
@@
 
-Data source: (1) National census; (2) Worldwide population and mortality data:  http://www.mortality.org/ (mostly high income countries; and (3) Calculate from GBD data (rates and numbers).
+Data source: (1) National census; (2) Worldwide population and mortality data:  http://www.mortality.org/ (mostly high income countries; and (3) Calculate from the Global Buden of Disease by the Institute of Health Metrics and Evaluation (GBD IHME) data (rates and numbers available from (http://ghdx.healthdata.org/gbd-results-tool).
 
@@
 
-These data will be provided by the synthetic population.
-Other sources are the same sources as above for population numbers.
-Sometimes mortality rates are in age groups (1-4, 5-9, etc).
+Mortality rates are needed per single year and sex.
+These data are available from GBD IHME, however, in age groups (1-4, 5-9, etc).
 Interpolation can be used to derive in between ages rates (cubic spline).
-Here, mortality rates are needed per single year and sex.
 
 Note that we need data for population numbers and all cause mortality rates for: (1) PMSLT and (2) Dismod II collection (more in Dismod II section).
+Population data from the synthetic population is used for the PMSLT.
+For Dismod II, popuation and mortality data should be from the same source (GBD IHME)
 
@@
 
-### **Disease life tables**
-
-#### Incidence and case fatality
-
-For each of the modeled diseases the PMSLT needs incidence and case fatality rates per sex and one-year intervals.
-Data from the GBD studies with Dismod II (free at https://www.epigear.com/index_files/dismod_ii.html) can be used to derive internally consistent data and generate missing data.
-For example, the GBD studies provide data for incidence, prevalence and disease mortality, however, not case fatality.
-Other national level sources may also be explored/used, and compare with estimates produce from GBD data and Dismod II.
-
-**Dismod II** inputs are: (1) population numbers and mortality rates and (2) disease specific inputs.
-
-*Population and mortality*
-
-Within Dismod II, each setting (e.g.
-country) has a collection that consists of population numbers (preferably the same as used in GBD studies, due to the mortality envelop) and all- cause mortality rates (numbers and calculate rates).
-The GBD provides 5-year age groups that are acceptable input parameters for Dismod II.
-
-*Disease inputs by age group and sex*
-
-Each setting collection has a given number of diseases.
-Dismod II works with at least three of: case fatality, prevalence, incidence, mortality (disease), case fatality, remission, duration and the relative risk for mortality.
-So far, we have been assuming that remission is zero for chronic diseases, that is to say, when people become diseased, they do not recover.
-Special care should be taken with this assumption, as the GBD data assumes remission for some diseases, for example cancers, where after 10 years cases recover, except for long term sequelae.
-Since GBD now provides prevalence, incidence and mortality, it may be best to use all three as Dismod II input parameters to compare the effect of the remission assumption by the GBD for some diseases.
-
 #### Disability weights (quality of life weights)
@@
 Disability weights (DW) can be derived from disease specific years lived with disability (YLD) and disease specific prevalence by age group (5 years) and sex.
-Data for YLDs prevalence can be obtained from the online GBD data tool (http://ghdx.healthdata.org/gbd-results-tool).
+Data for YLDs prevalence can be obtained from the online GBD IHME data tool (http://ghdx.healthdata.org/gbd-results-tool).
 Our calculations of DW in the example here based on the GBD methods for estimating YLDs as the sum of sequelae prevalence multiplied by sequelae disability weights (REF GBD).
@@
 Where YLDd is the YLD mean number per age and sex for a given disease, Pd is the prevalence (as reported in GBD )  for a given disease by age and sex and YLDt is total YLD rate per age and sex.
+
+### Disease life tables
+
+#### Incidence and case fatality
+
+For each of the modeled diseases the PMSLT needs incidence and case fatality rates per sex and one-year intervals.
+Data from the GBD IHME studies with Dismod II (free at https://www.epigear.com/index_files/dismod_ii.html) can be used to derive internally consistent data and generate missing data.
+For example, the GBD studies provide data for incidence, prevalence and disease mortality, however, not case fatality.
+Other national level sources may also be explored/used, and compare with estimates produce from GBD data and Dismod II.
+
+**Dismod II** inputs are: (1) population numbers and mortality rates and (2) disease specific inputs.
+
+*Population and mortality*
+
+Within Dismod II, each setting (e.g.
+country) has a collection that consists of population numbers (preferably the same as used in GBD IHME studies, due to the mortality envelop) and all- cause mortality rates (numbers and calculate rates).
+The GBD provides 5-year age groups that are acceptable input parameters for Dismod II.
+
+*Disease inputs by age group and sex*
+
+Each setting collection has a given number of diseases.
+Dismod II works with at least three of: case fatality, prevalence, incidence, mortality (disease), case fatality, remission, duration and the relative risk for mortality.
+So far, we have been assuming that remission is zero for chronic diseases, that is to say, when people become diseased, they do not recover.
+Special care should be taken with this assumption, as the GBD data assumes remission for some diseases, for example cancers, where after 10 years cases recover, except for long term sequelae.
+Since GBD now provides prevalence, incidence and mortality, it may be best to use all three as Dismod II input parameters to compare the effect of the remission assumption by the GBD for some diseases.
 
@@
 ```
-\pagebreak
 
@@
 
-We start by *cleaning the global environment* to keep track of our works and ensure that the code is generating our outcomes.
-Then, we set up an option to avoind the use of scientific notation and lastly we load the functions.
+We start by cleaning the global environment (#1) to keep track of our works and ensure that the code is generating our outcomes.
+Then, we set up an option to avoid the use of scientific notation (#2) and lastly we load the functions (#3).
 The code chunks are shown in the rmarkdown ouput.
@@
 ```{r, echo=TRUE}
+#1
 rm (list = ls())
-
-# Prevent scientific notation in data frame
-
+#2
 options(scipen=999)
-
-#Load all functions/packages
-
+#3
 source("code/functions.R")
@@
 
-The data used here is for an example for Australia, however, this will be updated to link with the ITHIM-R modules (synthetic population, PIF calculations).
+###Inputs
+
+Table 1 describes data needs for the PMSLT, here we expand on the data needs and mechanisms (Figure 3) to use the PMSLT approach in ITHIMR (Figure 1).
+Figure 3 is used as a reference in the development of the code.
+
+![Proportional multi-state life-table model. *Three sections are presented in Figure 3: **Data input sources**, **Inputs PMSLT** and **PMSLT mechanisms**.* ](structure\Figure3.png)
+
+####Data for example
+
+The code below is to load the data used in the example provided in this code for Australia (#4).
+*idata* includes data by one-year age groups and sex for: population numbers, mortality rates, total years lived with disability rates, incidence, case fatality, disability weights and costs for included diseases (ischemic heart disease, ischemic stroke, diabates, breast cancer (women) and colon cancer), cost for all other diseases (excludes those for physical inactivity) and trend for incidence and case fatality.
+Here we do not use cost and trends data.
+*edata* is for physical activity prevalence per categories (health risk factor), *irr* are relative risks and **ee** is energy expenditure per physical activity categories.
 
 ```{r, echo=TRUE}
-##### Read data. idata: life table and disease life tables (including trends). edata: exposure data (e.g. physical activity).
-#### irr.data: relative risks data and ee: energy expenditure
-#### For exposure (edata) and input (idata) data I am using already generated categories, but a function should be developed to work from raw data.
-
+#4
 idata <- read.csv("data/idata.csv", stringsAsFactors = F)
@@
 ee <- read.csv("data/ee.csv", stringsAsFactors = F)
+```
+
+####Model parameters
+
+We define parameters for our model that are used multiple times.
+*p_age_cohort* defines the starting age for the cohorts.
+*p_sex* is for the cohort sex.
+*p_disease* specifies the modelled diseases.
+*p_intervention* is a hypothetical increase in MET minutes per week.
+
+```{r, echo=TRUE}
+#5
+p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
+p_sex <- c("males", "females")
+p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
+p_intervention_effect <- 100
 ```

### 2018-04-05 MSLT/MSLTdoc.Rmd
Code

@@
 
-####Data for example
+####Data for example (CHANGE TO ENGLISH DATA, MAY NOT BE AN EXAMPLE, BUT HAVE THE DATA FOR THE CASE STUDIES (NEED TO PREPARE: LONDON, SAO PABLO, DELHI, ACCRA, LOS ANGELES, EDINBURGH)).
+DATA AVAILABILITY FROM GBD AT WHAT LEVEL?
 
@@
 
+####Data preparation
+
+The following code applies to the data used as an exmaple here.
+#6 transforms upper cases to lower cases in the sex variable.
+#7 and #8 is used to rename mortality_rate to mx (used in life table language).
+#9 is used to replace male and female in the exposure data to males and females to facilitate combination of data sets.
+
+```{r, echo=TRUE}
+#6
+idata$sex <- tolower(idata$sex)
+#7
+idata$mx <- idata$mortality_rate
+#8
+idata$mortality_rate <- NULL
+#9
+edata$sex[edata$sex=="male"] <- "males"
+edata$sex[edata$sex=="female"] <- "females"
+```
+
+```{r, echo=TRUE}
+##### Create new variable for 5_year population (age cohorts). Depends on the age-cohorts of interest (5-yrs here)
+
+idata$five_year_population <- NA
+start_index <- 3
+index <- 1
+end_index <- 4
+for (i in 1:20){
+  if (i == 1)
+    index <- 1
+  else 
+    index <- ((i - 1)  * 5) + 1
+  if (index == 96){
+    end_index <- 5
+  }
+  cat("population sum ", start_index, "start index ", index, " and end index ", index + end_index, "\n")
+  idata$five_year_population[start_index] <- sum(idata$population[index:(index + end_index)])
+  start_index <- start_index + 5
+}
+
+end_index <- 4
+start_index <- 104
+
+for (i in 1:20){
+  if (i == 1)
+    index <- 102
+  else
+    index <- (((i - 1)  * 5) + 1) + 101
+  if (index == 197){
+    end_index <- 5
+  }
+  cat("population sum ", start_index, "start index ", index, " and end index ", index + end_index, "\n")
+  idata$five_year_population[start_index] <- sum(idata$population[index:(index + end_index)])
+  start_index <- start_index + 5
+}
+
+```
+
 ## Including Plots

### 2018-04-05 MSLT/MSLTdoc.Rmd
Comments on data for case studies

@@
 ####Data for example (CHANGE TO ENGLISH DATA, MAY NOT BE AN EXAMPLE, BUT HAVE THE DATA FOR THE CASE STUDIES (NEED TO PREPARE: LONDON, SAO PABLO, DELHI, ACCRA, LOS ANGELES, EDINBURGH)).
-DATA AVAILABILITY FROM GBD AT WHAT LEVEL?
+DATA AVAILABILITY FROM ARE IN 5 YEARS UP TO 95+.
+gbd DATA IS NOT AVAILABLE AT THE CITY LEVEL, HOWEVER, SOME LEVEL OF DISSAGREAGATION IS AVAILABLE.
+ENGLAND: EAST MIDLANDS, EAST OF ENGLAND, GRETER LONDON, ETC.
+wE COULD USE GREATER LONDON.
+FOR EDINBURGH WE HAVE DATA FOR SCOTLAND.
+FOR SAO PABLO, DATA IS FOR BRAZIL.
+MEXICO IS NOT LISTED, BUT GBD HAS DATA FOR DISTRITO FEDERAL.
+FOR DELHI THE DATA IS FOR INDIA.
+FOR ACCRA THE DATA IS FOR GHANA.
+FOR LOS ANGELES DATA IS FOR UNITED STATES, BUT WE CAN HAVE THIS FROM NEIL?
 

### 2018-04-06 MSLT/MSLTdoc.Rmd
Figure 3 description

--- 2018-04-05 MSLT/MSLTdoc.Rmd
@@
 Table 1 describes data needs for the PMSLT, here we expand on the data needs and mechanisms (Figure 3) to use the PMSLT approach in ITHIMR (Figure 1).
-Figure 3 is used as a reference in the development of the code.
 
-![Proportional multi-state life-table model. *Three sections are presented in Figure 3: **Data input sources**, **Inputs PMSLT** and **PMSLT mechanisms**.* ](structure\Figure3.png)
+(NEED TO PREPARE: LONDON, SAO PABLO, DELHI, BAY AREA).
+DATA AVAILABILITY FROM ARE IN 5 YEARS UP TO 95+.
+gbd DATA IS NOT AVAILABLE AT THE CITY LEVEL, HOWEVER, SOME LEVEL OF DISSAGREAGATION IS AVAILABLE.
+ENGLAND: EAST MIDLANDS, EAST OF ENGLAND, GRETER LONDON, ETC.
+wE COULD USE GREATER LONDON.
+FOR EDINBURGH WE HAVE DATA FOR SCOTLAND.
+FOR SAO PABLO, DATA IS FOR BRAZIL.
+MEXICO IS NOT LISTED, BUT GBD HAS DATA FOR DISTRITO FEDERAL.
+FOR DELHI THE DATA IS FOR INDIA.
+FOR ACCRA THE DATA IS FOR GHANA.
+FOR LOS ANGELES DATA IS FOR UNITED STATES, BUT WE CAN HAVE THIS FROM NEIL?
 
-####Data for example
+![Proportional multi-state life-table model.
+*Three sections are presented in Figure 3: **Data input sources**, **Inputs PMSLT** and **PMSLT mechanisms**.
+The color coding from Data inputs sources to Inputs PMSLT link sources with inputs for the PMSLT.
+Solid arrows represent final inputs and dashed-arrows represent intermediate inputs that need further processing.
+Purple coding means a process and green coding represent change in mortality and disability prevalence rates to modify the life table parameters.
+Black color coding with white color font represent final model outcomes.
+For both, the life table and disease life table, two sets of each are simulated, one for the baseline and the other for the scenario.
+**Data inputs** are from four sources: Sythetic population, GBD data (from the Institute of Health Metrics and Evaluations (IHME)), ITHIM calculations and Injuries model.
+The synthetic population provides population numbers by age-groups and sex which are used in the life table (**PMSLT Mechanisms**) with each age cohort simulated until they die or reach 100 years of age.
+Population numbers are also used to derive road death rates and road injuries rates from numbers for the baseline and scenario generated by the Injuries model.
+The difference in road deaths rates modifies all cause-mortatliy in the scenario life table and the change in injuries rates modifies total YLDs rates (hence, healthy life years).
+Data from GBD IHME includes: total YLD rates per one, disease specific YLDs (numbers) and prevalence to derive disability weights for the included diseases, all cause mortatliy rates per one, population numbers and diseases parameters (incidence, mortality and prevalence).
+Dismod II is used to derive internally consitent estimates incidence and case fatality rates from GBD IHME data, which are inputs of the disease specific life tables.
+Case fatality derive from Dismod II is the same for the baseline and scenario (however, could be changed) and incidence is different.
+For the baseline life tables incidence is as derived from Dismod II and for the scanrio is modified by the PIF (1-PIF).
+A change in incidence modifies prevalence and mortality in later years.
+The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.* ](structure\Figure3.png)
 
-The code below is to load the data used in the example provided in this code for Australia (#4).
-*idata* includes data by one-year age groups and sex for: population numbers, mortality rates, total years lived with disability rates, incidence, case fatality, disability weights and costs for included diseases (ischemic heart disease, ischemic stroke, diabates, breast cancer (women) and colon cancer), cost for all other diseases (excludes those for physical inactivity) and trend for incidence and case fatality.
-Here we do not use cost and trends data.
-*edata* is for physical activity prevalence per categories (health risk factor), *irr* are relative risks and **ee** is energy expenditure per physical activity categories.
+(NEED TO PREPARE: LONDON, SAO PABLO, DELHI, ACCRA, LOS ANGELES, EDINBURGH)).
+DATA AVAILABILITY FROM ARE IN 5 YEARS UP TO 95+.
+gbd DATA IS NOT AVAILABLE AT THE CITY LEVEL, HOWEVER, SOME LEVEL OF DISSAGREAGATION IS AVAILABLE.
+ENGLAND: EAST MIDLANDS, EAST OF ENGLAND, GRETER LONDON, ETC.
+wE COULD USE GREATER LONDON.
+FOR EDINBURGH WE HAVE DATA FOR SCOTLAND.
+FOR SAO PABLO, DATA IS FOR BRAZIL.
+MEXICO IS NOT LISTED, BUT GBD HAS DATA FOR DISTRITO FEDERAL.
+FOR DELHI THE DATA IS FOR INDIA.
+FOR ACCRA THE DATA IS FOR GHANA.
+FOR LOS ANGELES DATA IS FOR UNITED STATES, BUT WE CAN HAVE THIS FROM NEIL?
 
-```{r, echo=TRUE}
-#4
-idata <- read.csv("data/idata.csv", stringsAsFactors = F)
-edata <- read.csv("data/edata.csv", stringsAsFactors = F)
-irr <- read.csv("data/irr.csv", stringsAsFactors = F)
-ee <- read.csv("data/ee.csv", stringsAsFactors = F)
-```
+The code below is to load the data used will be updated to the study cases.
+
+WRITE CODE HERE TO GET DATA.
 
@@
 
-```{r, echo=TRUE}
-#5
-p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
-p_sex <- c("males", "females")
-p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
-p_intervention_effect <- 100
-```
+# ```{r, echo=TRUE}
+# #5
+# p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
+# p_sex <- c("males", "females")
+# p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
+# p_intervention_effect <- 100
+# ```
+
+####Data preparation
+
+The following code applies to the data used as an exmaple here.
+#6 transforms upper cases to lower cases in the sex variable.
+#7 and #8 is used to rename mortality_rate to mx (used in life table language).
+#9 is used to replace male and female in the exposure data to males and females to facilitate combination of data sets.
+
+# ```{r, echo=TRUE}
+# #6
+# idata$sex <- tolower(idata$sex)
+# #7
+# idata$mx <- idata$mortality_rate
+# #8
+# idata$mortality_rate <- NULL
+# #9
+# edata$sex[edata$sex=="male"] <- "males"
+# edata$sex[edata$sex=="female"] <- "females"
+# ```
+
+# ```{r, echo=TRUE}
+# ##### Create new variable for 5_year population (age cohorts). Depends on the age-cohorts of interest (5-yrs here)
+# 
+# idata$five_year_population <- NA
+# start_index <- 3
+# index <- 1
+# end_index <- 4
+# for (i in 1:20){
+#   if (i == 1)
+#     index <- 1
+#   else 
+#     index <- ((i - 1)  * 5) + 1
+#   if (index == 96){
+#     end_index <- 5
+#   }
+#   cat("population sum ", start_index, "start index ", index, " and end index ", index + end_index, "\n")
+#   idata$five_year_population[start_index] <- sum(idata$population[index:(index + end_index)])
+#   start_index <- start_index + 5
+# }
+# 
+# end_index <- 4
+# start_index <- 104
+# 
+# for (i in 1:20){
+#   if (i == 1)
+#     index <- 102
+#   else
+#     index <- (((i - 1)  * 5) + 1) + 101
+#   if (index == 197){
+#     end_index <- 5
+#   }
+#   cat("population sum ", start_index, "start index ", index, " and end index ", index + end_index, "\n")
+#   idata$five_year_population[start_index] <- sum(idata$population[index:(index + end_index)])
+#   start_index <- start_index + 5
+# }
+# 
+# ```
 

### 2018-04-09 MSLT/MSLTdoc.Rmd
Pdf figures

@@
     - \usepackage{float}
+    - \usepackage{pdflscape}
+    - \newcommand{\blandscape}{\begin{landscape}}
+    - \newcommand{\elandscape}{\end{landscape}}
 output:
@@
      number_sections: true
-     fig_caption: yes
-     fig_width: 7
-     fig_height: 6
+
      df_print: kable
@@
 knitr::opts_chunk$set(echo = TRUE)
+options(knitr.table.format = 'latex')
+knitr::opts_chunk$set(fig.cap = TRUE, fig.align = "center", fig.pos = "H", fig.scap = TRUE)
 
@@
 require(png)
-options(knitr.table.format = 'latex')
 
@@
 
-![Basic ITHIMR infrastructure](structure\Figure1.png)
+![Basic ITHIMR infrastructure](structure\Figure1.png){width=70%, fig.pos="H"}
 
@@
 
+\newpage
+\blandscape
+
 ![Proportional multi-state life-table simplified framework.
@@
 Changes in prevalence and mortality rates from the disease specific life tables feed into the life table by changing all-cause mortality, which in turn changes life years.
-Change in prevalence of diseases changes total years lived with disability, which in turn modifies health-adjusted life years*](structure\Figure2.png)
+Change in prevalence of diseases changes total years lived with disability, which in turn modifies health-adjusted life years*](structure\Figure2.png){width=100%, fig.pos="H"}
+
+\elandscape
 
@@
 
+\newpage
+\blandscape
+
 ![Proportional multi-state life-table model.
@@
 A change in incidence modifies prevalence and mortality in later years.
-The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.* ](structure\Figure3.png)
+The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.* ](structure\Figure3.png){width=70%, hight=50%, fig.pos="H"}
+
+\elandscape
 

### 2018-04-09 MSLT/MSLTdoc.Rmd
Figure labels too far from figures and not showing

@@
 require(citr)
-require(png)
-
 ```

### 2018-04-09 MSLT/MSLTdoc.Rmd
GBD data inputs

@@
 ```
+
+\newpage
+
 # Introduction
@@
 ITHIM is included in Figure 1 to show that both approaches share in common steps one and two and differ in the mechanisms of calculation of change in health burden.
-
-![Basic ITHIMR infrastructure](structure\Figure1.png){width=70%, fig.pos="H"}
 
@@
 In burden of disease studies, DALYs are defined as the sum Years of Life Lost (YLL) and Years Lived with Disability (YLD).
+
+\newpage
+
+![Basic ITHIMR infrastructure](structure\Figure1.png){fig.pos="H"}
 
@@
 
-* **Interaction between multiple diseases**
+***Interaction between multiple diseases**
 The *PMSLT* accounts for the interaction between multiple diseases, with proportions of the population    being able to be in more than one health state [@RN8957].
@@
 
-* **Mortality rate** 
+***Mortality rate** 
 The *PMSLT* calculations for changes in life years (and health-adjusted life years) and mortality outcomes is based on observed mortality rates for the population of interest.
@@
 
-* **Impact of disability in increased life expectancy**
+***Impact of disability in increased life expectancy**
 In GBD studies, YLLs are not adjusted for disability; hence, their use in estimating intervention effects results in over-estimation, which the *PMSLT* approach avoids.
@@
 
-(NEED TO PREPARE: LONDON, SAO PABLO, DELHI, BAY AREA).
-DATA AVAILABILITY FROM ARE IN 5 YEARS UP TO 95+.
-gbd DATA IS NOT AVAILABLE AT THE CITY LEVEL, HOWEVER, SOME LEVEL OF DISSAGREAGATION IS AVAILABLE.
-ENGLAND: EAST MIDLANDS, EAST OF ENGLAND, GRETER LONDON, ETC.
-wE COULD USE GREATER LONDON.
-FOR EDINBURGH WE HAVE DATA FOR SCOTLAND.
-FOR SAO PABLO, DATA IS FOR BRAZIL.
-MEXICO IS NOT LISTED, BUT GBD HAS DATA FOR DISTRITO FEDERAL.
-FOR DELHI THE DATA IS FOR INDIA.
-FOR ACCRA THE DATA IS FOR GHANA.
-FOR LOS ANGELES DATA IS FOR UNITED STATES, BUT WE CAN HAVE THIS FROM NEIL?
+Initial case studies for the ITHIMR are: London, Sao PAblo, Delhi, Accra, Los Angeles and Edinburgh.
+Here, we will start with **greater London** given the availability of dieases epidemiology data from the GBD IHME study.
+For the rest of the case study cities data is available at the country level, hence, a scaling method is needed.
 
@@
 A change in incidence modifies prevalence and mortality in later years.
-The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.* ](structure\Figure3.png){width=70%, hight=50%, fig.pos="H"}
+The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.*](structure\Figure3.png){width=50%,hight=30%,fig.pos="H"}
 
 \elandscape
+
+####Global Burden of Disease data
+
+First, we explain how to obtain the data, second additional processing to derive data not reported (population) and one-year age groups (original data is in five-year age groups) and last procedure to use Dismod II.
+Data from the *Global Burden of Disease data* in Figure 3 can be download from here: http://ghdx.healthdata.org/gbd-results-tool.
+Figure 4 is a screenshot of the GHDx.
+
+![Global Burden of Disease data results tool.](data/Figure4.png){width=70%,hight=50%,fig.pos="H"}
+
+Table 2 specifies the selections to do for each of the tabs in Figure 2.
+
+```{r, echo=FALSE}
+pmslt_inputs <- tribble(
+~Tab, ~Selection,
+"Base", "Single", "Location", "Case study city", "Year", "Latest available", "Context", "Cause", "Age", "Under 5, 5 to 9, 10 to 14, 15 to 19, 20 to 24, 25 to 29, 30 to 34, 35 to 39, 40 to 49, 50 to 54, 55 to 59, 60 to 64, 65 to 69, 70 to 74, 75 to 79, 80 to 84, 89 to 89, 90 to 95, 95 plus", "Metric", "Number, Rate", "Measure", "Deaths, YLDs, Prevalence, Incidence", "Sex", "Male, Female", "Cause", "Total All causes, ischemic heart disease, etc")
+
+kable(pmslt_inputs,"latex", booktabs = T, caption = "Global burden of disease data") %>%
+ kable_styling(latex_options = c("striped", "hold position"),
+               full_width = F)%>%
+  column_spec(2, width = "15em") %>%
+  column_spec(3, width = "15em") %>%
+  row_spec(0, bold = T)
+```
+
+Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.
 

### 2018-04-09 MSLT/MSLTdoc.Rmd
Inputs GBD table

@@
                full_width = F)%>%
-  column_spec(2, width = "15em") %>%
-  column_spec(3, width = "15em") %>%
+  column_spec(2, width = "30em") %>%
+  column_spec(3, width = "30em") %>%
   row_spec(0, bold = T)

### 2018-04-09 MSLT/MSLTdoc.Rmd
Spell check

@@
 
-Data source: (1) National census; (2) Worldwide population and mortality data:  http://www.mortality.org/ (mostly high income countries; and (3) Calculate from the Global Buden of Disease by the Institute of Health Metrics and Evaluation (GBD IHME) data (rates and numbers available from (http://ghdx.healthdata.org/gbd-results-tool).
+Data source: (1) National census; (2) Worldwide population and mortality data:  http://www.mortality.org/ (mostly high income countries; and (3) Calculate from the Global Burden of Disease by the Institute of Health Metrics and Evaluation (GBD IHME) data (rates and numbers available from (http://ghdx.healthdata.org/gbd-results-tool).
 
@@
 Population data from the synthetic population is used for the PMSLT.
-For Dismod II, popuation and mortality data should be from the same source (GBD IHME)
+For Dismod II, population and mortality data should be from the same source (GBD IHME)
 
@@
 Then, we set up an option to avoid the use of scientific notation (#2) and lastly we load the functions (#3).
-The code chunks are shown in the rmarkdown ouput.
+The code chunks are shown in the rmarkdown output.
 
@@
 
-Initial case studies for the ITHIMR are: London, Sao PAblo, Delhi, Accra, Los Angeles and Edinburgh.
-Here, we will start with **greater London** given the availability of dieases epidemiology data from the GBD IHME study.
+Initial case studies for the ITHIMR are: London, Sao Pablo, Delhi, Accra, Los Angeles and Edinburgh.
+Here, we will start with **greater London** given the availability of diseases epidemiology data from the GBD IHME study.
 For the rest of the case study cities data is available at the country level, hence, a scaling method is needed.
@@
 A change in incidence modifies prevalence and mortality in later years.
-The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.*](structure\Figure3.png){width=50%,hight=30%,fig.pos="H"}
+The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.*](structure\Figure3.png){width=50%,height=30%,fig.pos="H"}
 
@@
 
-![Global Burden of Disease data results tool.](data/Figure4.png){width=70%,hight=50%,fig.pos="H"}
+![Global Burden of Disease data results tool.](data/Figure4.png){width=70%,height=50%,fig.pos="H"}
 
@@
 
+Once the selections described in Table 2 are made, the option *Download CVS** in the GHDx website is selected.
+A prompt comes up asking for an email address.
+The data is sent to the designated email address (within minutes) in ZIP format, unzip and use the code below to read the data.
+Here, we selected data for Greater London and England.
+The aim is to compare and derive scaling factors as for most cities the data is not available from the GBD and country level data may be used and scaled to the city level.
+Note that the data input requirement for the PMSLT, except population numbers, is in rates.
+Therefore the scaling is to better reflect the burden of an area, this is a different issue than working with numbers (e.g.
+total mortality numbers, total YLDs numbers) as in the ITHIM approach.
+
+```{r, echo=TRUE}
+options(scipen=999)
+GBDdata <- read.csv("data/IHME-GBD_2016_DATA-8fca12f6-1.csv", stringsAsFactors = F)
+```
+
 Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.
-
-(NEED TO PREPARE: LONDON, SAO PABLO, DELHI, ACCRA, LOS ANGELES, EDINBURGH)).
-DATA AVAILABILITY FROM ARE IN 5 YEARS UP TO 95+.
-gbd DATA IS NOT AVAILABLE AT THE CITY LEVEL, HOWEVER, SOME LEVEL OF DISSAGREAGATION IS AVAILABLE.
-ENGLAND: EAST MIDLANDS, EAST OF ENGLAND, GRETER LONDON, ETC.
-wE COULD USE GREATER LONDON.
-FOR EDINBURGH WE HAVE DATA FOR SCOTLAND.
-FOR SAO PABLO, DATA IS FOR BRAZIL.
-MEXICO IS NOT LISTED, BUT GBD HAS DATA FOR DISTRITO FEDERAL.
-FOR DELHI THE DATA IS FOR INDIA.
-FOR ACCRA THE DATA IS FOR GHANA.
-FOR LOS ANGELES DATA IS FOR UNITED STATES, BUT WE CAN HAVE THIS FROM NEIL?
-
-The code below is to load the data used will be updated to the study cases.
-
-WRITE CODE HERE TO GET DATA.
 
@@
 *p_sex* is for the cohort sex.
-*p_disease* specifies the modelled diseases.
+*p_disease* specifies the modeled diseases.
 *p_intervention* is a hypothetical increase in MET minutes per week.
@@
 
-The following code applies to the data used as an exmaple here.
+The following code applies to the data used as an example here.
 #6 transforms upper cases to lower cases in the sex variable.

### 2018-04-10 MSLT/MSLTdoc.Rmd
Preparing data for England and Greater London

@@
 
+The following code is to compare mortality rates per one, by age groups of 5 and sex for England and England.
+
+ALSO PART OF THE BELOW CODE WILL BE NEEDED FOR THE MSLT AND DISMOD COLLECTION. GENERATE DATA INPUTS FOR DISMOD II FROM HERE.
+
+```{r, echo=TRUE}
+#Data preparation (generate age groups, change capitals to lower case)
+##All upper cases to lower cases
+
+GBDdata <- mutate_all(GBDdata, funs(tolower))
+
+##Create age categories index in GBDdata (rounding mid age)
+
+GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
+GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
+GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
+GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
+GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
+GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
+GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
+GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
+GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
+GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
+GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
+GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
+GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
+GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
+GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
+GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
+GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
+GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
+GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
+GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97
+
+##Convert variable val to numberic
+
+GBDdata$val <- as.numeric(as.character(GBDdata$val))
+
+##Filter data for England and Greater London adn generate per person rates (per 100,000 in original data)
+#England
+GBDEngland <- filter(GBDdata, location == "england" & metric == "rate") %>% select(measure, location, sex, age, metric, val, age_cat)
+
+GBDEngland$one_rate <- GBDEngland$val/100000
+
+#Greater London (GL)
+GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% select(measure, location, sex, age, metric, val, age_cat)
+GBDGL$one_rate <- GBDGL$val/100000
+
+```
+
 Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.

### 2018-04-10 MSLT/code/functions.R

@@ -78,7 +78,8 @@ run_life_table <- function(in_idata, in_sex, in_mid_age)
     lf_df$ex[i] <- sum(lf_df$Lx[i:nrow(lf_df)]) / lf_df$lx[i]
   }
   
-  # create health adjusted life years variable
+  # create health adjusted life years variable 
+  
   lf_df$Lwx <- lf_df$Lx * (1 - lf_df$pyld_rate)
   
   # create health adjusted life expectancy variable

### 2018-04-10 MSLT/MSLTdoc.Rmd
Derived male population numbers from GBD data

@@
 
-ALSO PART OF THE BELOW CODE WILL BE NEEDED FOR THE MSLT AND DISMOD COLLECTION. GENERATE DATA INPUTS FOR DISMOD II FROM HERE.
+ALSO PART OF THE BELOW CODE WILL BE NEEDED FOR THE MSLT AND DISMOD COLLECTION.
+GENERATE DATA INPUTS FOR DISMOD II FROM HERE.
+These data should be used to generate the general life table and disease life tables
 
@@
 
-##Create age categories index in GBDdata (rounding mid age)
+##Create age categories index in GBDdata (rounding mid age), total of 20 age groups
 
@@
 
-##Convert variable val to numberic
+##Create age and sex categories to be able to derive population numbers
+
+GBDdata$sex_age_cat <- paste(GBDdata$age_cat,GBDdata$sex, sep = "_"  )
+
+##Convert variable val to numeric
 
@@
 
-##Filter data for England and Greater London adn generate per person rates (per 100,000 in original data)
+##Generate a data frame with population numbers for England and Greater London (for life table and Dismod II collection). 
+##These are derived from total number of cases and rates per 100,000. To derive values we need to select one cause (E.g. all cause)-does not matter what is chosen.
+
+GBD_England_population_male <- filter(GBDdata, location == "england", measure == "deaths", cause == "all causes", sex == "male", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, val, sex_age_cat)
+
+###Generate population numbers frame males England
+ for (i in 1:nrow(GBD_England_population_male)) {
+    if (GBD_England_population_male$metric[i] == "number") {
+      GBD_England_population_male$val_pop[i] <- GBD_England_population_male$val[i] * 10000/ GBD_England_population_male$val[i + 1]}
+   else {GBD_England_population_male$val_pop[i] <- NA}
+ }
+##Remove rows with zero
+GBD_England_population_male <- GBD_England_population_male[!is.na(GBD_England_population_male$val_pop),]
+
+##Last only keep relevant variables
+
+GBD_England_population_male <- filter(GBD_England_population_male) %>% select(age_cat, val_pop)
+
+##Filter data for England and Greater London and generate per person rates (per 100,000 in original data)
 #England
@@
 GBDGL$one_rate <- GBDGL$val/100000
-
 ```

### 2018-04-10 MSLT/MSLTdoc.Rmd
Population numbers from GBD 2
Modified the previous formula (only men in England) to apply to all the data set and generate population numbers  for male and females in England and Greater London

@@
 ##Generate a data frame with population numbers for England and Greater London (for life table and Dismod II collection). 
-##These are derived from total number of cases and rates per 100,000. To derive values we need to select one cause (E.g. all cause)-does not matter what is chosen.
+##These are derived from total number of cases and rates per 100,000. To derive values we need to select one cause (E.g. all cause)-does not matter what we choose.
 
-GBD_England_population_male <- filter(GBDdata, location == "england", measure == "deaths", cause == "all causes", sex == "male", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, val, sex_age_cat)
+###First generate population numbers for England and Greater London and then separate in dataframes
+GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, val, sex_age_cat, location)
 
-###Generate population numbers frame males England
- for (i in 1:nrow(GBD_England_population_male)) {
-    if (GBD_England_population_male$metric[i] == "number") {
-      GBD_England_population_male$val_pop[i] <- GBD_England_population_male$val[i] * 10000/ GBD_England_population_male$val[i + 1]}
-   else {GBD_England_population_male$val_pop[i] <- NA}
+###Generate population numbers from GBD data
+ for (i in 1:nrow(GBD_population)) {
+    if (GBD_population$metric[i] == "number") {
+      GBD_population$val_pop[i] <- GBD_population$val[i] * 100000/ GBD_population$val[i + 2]}
+   else {GBD_population$val_pop[i] <- NA}
  }
 ##Remove rows with zero
-GBD_England_population_male <- GBD_England_population_male[!is.na(GBD_England_population_male$val_pop),]
+GBD_population <- GBD_population[!is.na(GBD_population$val_pop),]
 
-##Last only keep relevant variables
+##Keep relevant variables
 
-GBD_England_population_male <- filter(GBD_England_population_male) %>% select(age_cat, val_pop)
+GBD_population <- filter(GBD_population) %>% select(sex_age_cat, val_pop, location)
+
+##Create data frames for Greater London and England population
+
+GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, val_pop, location)
+
+GBD_population_England <-  filter(GBD_population, location == "england") %>% select(sex_age_cat, val_pop, location)
+
+##Check population total numbers
+
+GreaterLondon <- sum(GBD_population_GL$val_pop)
+England <- sum(GBD_population_England$val_pop)
 

### 2018-04-10 MSLT/MSLTdoc.Rmd
Added comment to code
Discuss with Ali code from line 234 as this will apply to multiple geographies, may be a way to do it more efficiently.

--- 2018-04-10 MSLT/MSLTdoc.Rmd
@@
 
-The following code is to compare mortality rates per one, by age groups of 5 and sex for England and England.
+The following code serves to sort out the GBD data to the inputs required for the life table, disease life table and Dismod II
 
-ALSO PART OF THE BELOW CODE WILL BE NEEDED FOR THE MSLT AND DISMOD COLLECTION. GENERATE DATA INPUTS FOR DISMOD II FROM HERE.
+ALSO PART OF THE BELOW CODE WILL BE NEEDED FOR THE MSLT AND DISMOD COLLECTION.
+GENERATE DATA INPUTS FOR DISMOD II FROM HERE.
+These data should be used to generate the general life table and disease life tables (Figure 3).
 
@@
 
-##Create age categories index in GBDdata (rounding mid age)
+##Create age categories index in GBDdata (rounding mid age), total of 20 age groups. These are the age cohorts to simulate.
 
@@
 
-##Convert variable val to numberic
+##Create age and sex categories to obtain population numbers. Population numbers from GBD are used in 
+##Dismod II. For the Life table (Figure 3), the numbers may be from the synthetic population.
+
+GBDdata$sex_age_cat <- paste(GBDdata$age_cat,GBDdata$sex, sep = "_"  )
+
+##Convert variable "val" to numeric as its latter used in calculations.
 
@@
 
-##Filter data for England and Greater London adn generate per person rates (per 100,000 in original data)
+##Generate population numbers for England and Greater London in a new data frame ("GBD_population") and then separate in dataframes for England and Greater London (this avoids repeating the same code).
+
+GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, val, sex_age_cat, location)
+
+###Generate population numbers from given number of cases and rates per 100,000 people. 
+ for (i in 1:nrow(GBD_population)) {
+    if (GBD_population$metric[i] == "number") {
+      GBD_population$val_pop[i] <- GBD_population$val[i] * 100000/ GBD_population$val[i + 2]}
+   else {GBD_population$val_pop[i] <- NA}
+ }
+##Remove rows with zero
+GBD_population <- GBD_population[!is.na(GBD_population$val_pop),]
+
+##Keep relevant variables
+
+GBD_population <- filter(GBD_population) %>% select(sex_age_cat, val_pop, location)
+
+##Create data frames for Greater London and England population
+
+GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, val_pop, location)
+
+GBD_population_England <-  filter(GBD_population, location == "england") %>% select(sex_age_cat, val_pop, location)
+
+##Check population total numbers
+
+GreaterLondon <- sum(GBD_population_GL$val_pop)
+England <- sum(GBD_population_England$val_pop)
+
+##Generate data frames for England and Grater London with per person rates (per 100,000 in original data).
+
 #England

### 2018-04-10 MSLT/MSLTdoc.Rmd
Plot England all_cause mortality rates

@@
 #England
-GBDEngland <- filter(GBDdata, location == "england" & metric == "rate") %>% select(measure, location, sex, age, metric, val, age_cat)
+GBDEngland <- filter(GBDdata, location == "england" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
 
@@
 #Greater London (GL)
-GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% select(measure, location, sex, age, metric, val, age_cat)
+GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
 GBDGL$one_rate <- GBDGL$val/100000
+```
+
+These graphs are to compare GBD data for England and Greater London.
+
+```{r, echo=FALSE}
+
+mx_plot_compare <- ggplot(data = filter(GBDEngland, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat), aes(x = age_cat)) +
+    geom_line(mapping = aes(y = one_rate, colour = "one_rate")) +
+  theme_classic()
+
+#Print to view
+print(mx_plot_compare)
+
 ```

### 2018-04-10 MSLT/MSLTdoc.Rmd
Plot mortality rates males

@@
 
-mx_plot_compare <- ggplot(data = filter(GBDEngland, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat), aes(x = age_cat)) +
-    geom_line(mapping = aes(y = one_rate, colour = "one_rate")) +
-  theme_classic()
+##Select data from data frames for all cause mortality males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause mortality rates males 
+mx_plot_compare_male <- ggplot() + 
+geom_line(data=ggplot_data_England, aes(x= age_cat, y= one_rate), color='green') +
+geom_line(data=ggplot_data_GL, aes(x= age_cat, y= one_rate), color='red') +
+xlab ("Age") + ylab ("Rate per one") + labs (title = "Mortality rates comparison") +
+theme_classic() +
+scale_alpha_manual("Location", values = c("England", "Greater London")) +
 
 #Print to view
-print(mx_plot_compare)
+print(mx_plot_compare_male)
 

### 2018-04-10 MSLT/MSLTdoc.Rmd
Change plot type for mortality rates comparison

@@
 ##Plot all cause mortality rates males 
-mx_plot_compare_male <- ggplot() + 
-geom_line(data=ggplot_data_England, aes(x= age_cat, y= one_rate), color='green') +
-geom_line(data=ggplot_data_GL, aes(x= age_cat, y= one_rate), color='red') +
-xlab ("Age") + ylab ("Rate per one") + labs (title = "Mortality rates comparison") +
-theme_classic() +
-scale_alpha_manual("Location", values = c("England", "Greater London")) +
+mx_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+    geom_line(aes(color = "England"))+
+    geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+    labs(colour="Locations",x="Age",y="Mortality rate")+
+    theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+    scale_color_manual(values = c("blue","red"))+
+    theme_classic()
 
-#Print to view
 print(mx_plot_compare_male)

### 2018-04-10 MSLT/MSLTdoc.Rmd
Added females graphs
Next, do a loop for all measures (deaths, prevalenc,e, incidence, prevalence, ylds)

@@
     theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-    scale_color_manual(values = c("blue","red"))+
+    scale_color_manual(values = c("blue","red")) +
+    labs (title = "Mortality rates difference males") +
     theme_classic()
@@
 print(mx_plot_compare_male)
+
+##Select data from data frames for all cause mortality females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause mortality rates males 
+mx_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+    geom_line(aes(color = "England"))+
+    geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+    labs(colour="Locations",x="Age",y="Mortality rate")+
+    theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+    scale_color_manual(values = c("blue","red")) +
+    labs (title = "Mortality rates difference females") +
+    theme_classic()
+
+print(mx_plot_compare_female)
 

### 2018-04-12 MSLT/MSLTdoc.Rmd
Added comparison graphs
Error when knit to pdf

@@
 #Data preparation (generate age groups, change capitals to lower case)
+
 ##All upper cases to lower cases
@@
 
-##Create data frames for Greater London and England population
+##Create data frames for Greater London and England population to be later used as inputs for PMSLT calculations
 
@@
 ```{r, echo=FALSE}
+###TO DO: loops for sex, cause and measure and interpolation of rates, care with interpolating rates, need to convert to log for interpolation using cubic spline. 
+#FOR NOW THE CODE FOR GRAPH IS repetitive
+####################MORTALITY RATES ALL CAUSE PER ONE
+
+###############################Mortality rates###################################################################
 
@@
 
+#############YLD RATES ALL CAUSE PER ONE
+
+##################################yld rates all cause############################################################
+
+##Select data from data frames for all cause yld rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause yld rates males 
+totalyld_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Total YLDs rate")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Total YLDs rates difference males") +
+  theme_classic()
+
+print(totalyld_plot_compare_male)
+
+##Select data from data frames for all cause yld females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause yld rates females 
+totalyld_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Total YLDs rate")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Total YLDs rates difference females") +
+  theme_classic()
+
+print(totalyld_plot_compare_female)
+
+################IHD
+##############################Mortality rate ihd #############################################################
+
+##Select data from data frames for ihd yld rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause yld rates males 
+mx_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality IHD")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "IHD mortality rates difference males") +
+  theme_classic()
+
+print(mx_ihd_plot_compare_male)
+
+##Select data from data frames for ihd mortality rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause yld rates females 
+mx_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality IHD")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "IHD mortality rates difference females") +
+  theme_classic()
+
+print(mx_ihd_plot_compare_female)
+
+##############################Prevalence rate ihd #############################################################
+
+##Select data from data frames for ihd prevalence rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates males 
+px_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence IHD")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "IHD prevalence rates difference males") +
+  theme_classic()
+
+print(px_ihd_plot_compare_male)
+
+##Select data from data frames for ihd prevalence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates females 
+px_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence IHD")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "IHD prevalence rates difference females") +
+  theme_classic()
+
+print(px_ihd_plot_compare_female)
+
+##############################Incidence rate ihd #############################################################
+
+##Select data from data frames for ihd prevalence rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates males 
+ix_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Incidence IHD")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "IHD incidence rates difference males") +
+  theme_classic()
+
+print(ix_ihd_plot_compare_male)
+
+##Select data from data frames for ihd prevalence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates females 
+ix_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Incidence IHD")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "IHD incidence rates difference females") +
+  theme_classic()
+
+print(ix_ihd_plot_compare_female)
+
+################ISCHEMIC STROKE
+##############################Mortality rate ischemic stroke #############################################################
+
+##Select data from data frames for ischemic stroke mortality rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot ischemic stroke mortality rates males
+
+mx_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality IHD")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Ischemic stroke mortality rates difference males") +
+  theme_classic()
+
+print(mx_is_plot_compare_male)
+
+##Select data from data frames for ischemic stroke mortality rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot ischemic stroke mortality rates females 
+mx_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality ischemic stroke")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Ischemic stroke mortality rates difference females") +
+  theme_classic()
+
+print(mx_is_plot_compare_female)
+
+##############################Prevalence rate ischemic stroke #############################################################
+
+##Select data from data frames for ihd prevalence rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates males 
+px_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence IHD")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Ischemic stroke prevalence rates difference males") +
+  theme_classic()
+
+print(px_is_plot_compare_male)
+
+##Select data from data frames for ischemic stroke prevalence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates females 
+px_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence ischemic stroke")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Ischemic stroke prevalence rates difference females") +
+  theme_classic()
+
+print(px_is_plot_compare_female)
+
+##############################Incidence rate ischemic stroke #############################################################
+
+##Select data from data frames for ishcemic stroke incidence rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates males 
+ix_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Incidence ischemic stroke")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Ischemic stroke incidence rates difference males") +
+  theme_classic()
+
+print(ix_is_plot_compare_male)
+
+##Select data from data frames for ischemic stroke incidence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot ischemic stroke incidence rates females 
+ix_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Incidence ischemic stroke")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Ischemic stroke incidence rates difference females") +
+  theme_classic()
+
+print(ix_is_plot_compare_female)
+
+################DIABETES MELLITUS
+##############################Mortality rate diabetes mellitus #############################################################
+
+##Select data from data frames for diabetes mellitus mortality rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot ischemic stroke mortality rates males
+
+mx_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality diabetes mellitus")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Diabetes mellitus mortality rates difference males") +
+  theme_classic()
+
+print(mx_DM_plot_compare_male)
+
+##Select data from data frames for diabetes mellitus mortality rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot diabetes mellitus mortality rates females 
+mx_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality diabetes mellitus")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Diabetes mellitus mortality rates difference females") +
+  theme_classic()
+
+print(mx_DM_plot_compare_female)
+
+##############################Prevalence rate diabetes mellitus #############################################################
+
+##Select data from data frames for diabetes mellitus prevalence rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates males 
+px_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence diabetes mellitus")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Diabetes mellitus prevalence rates difference males") +
+  theme_classic()
+
+print(px_DM_plot_compare_male)
+
+##Select data from data frames for diabetes mellitus prevalence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates females 
+px_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence diabetes mellitus")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Ischemic stroke prevalence rates difference females") +
+  theme_classic()
+
+print(px_DM_plot_compare_female)
+
+##############################Incidence rate diabetes mellitus #############################################################
+
+##Select data from data frames for diabetes mellitus incidence rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot incidence rate diabetes mellitus males 
+ix_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Diabetes mellitus")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Diabetes mellitus incidence rates difference males") +
+  theme_classic()
+
+print(ix_DM_plot_compare_male)
+
+##Select data from data frames for diabetes mellitus incidence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot ischemic stroke incidence rates females 
+ix_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Incidence diabetes mellitus")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Diabetes mellitus incidence rates difference females") +
+  theme_classic()
+
+print(ix_DM_plot_compare_female)
+
+#####Breast cancer
+##############################Mortality rate breast cancer #############################################################
+
+##Select data from data frames for breast cancer mortality rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot breast cancer mortality rates females 
+mx_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality breast cancer")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Breast cancer mortality rates difference females") +
+  theme_classic()
+
+print(mx_BC_plot_compare_female)
+
+##############################Prevalence rate breast cancer #############################################################
+
+##Select data from data frames for breast cancer prevalence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot all cause prevalence rates females 
+px_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence breast cancer")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Breast cancer prevalence rates difference females") +
+  theme_classic()
+
+print(px_BC_plot_compare_female)
+
+##############################Incidence rate breast cancer #############################################################
+
+##Select data from data frames for breast cancer incidence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot breast cancer incidence rates females 
+ix_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Incidence breast cancer")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Breast cancer incidence rates difference females") +
+  theme_classic()
+
+print(ix_BC_plot_compare_female)
+
+################Colon and rectum cancer (DOSE RESPONSE FOR COLON CANCER?)
+##############################Mortality rate colon and rectum cancer #############################################################
+
+##Select data from data frames for colon and rectum cancer mortality rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot colon and rectum cancer mortality rates males
+
+mx_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality colon and rectum cancer")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "colon and rectum cancer mortality rates difference males") +
+  theme_classic()
+
+print(mx_CC_plot_compare_male)
+
+##Select data from data frames for colon and rectum cancer mortality rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot colon and rectum cancer mortality rates females 
+mx_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Mortality colon and rectum cancer")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Colon and rectum cancers mortality rates difference females") +
+  theme_classic()
+
+print(mx_CC_plot_compare_female)
+
+##############################Prevalence rate colon and rectum cancer #############################################################
+
+##Select data from data frames for diabetes mellitus prevalence rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot colon and rectum cancer prevalence rates males 
+px_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence colon and rectum cancer")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "colon and rectum cancer prevalence rates difference males") +
+  theme_classic()
+
+print(px_CC_plot_compare_male)
+
+##Select data from data frames for colon and rectum cancer prevalence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot colon and rectum cancer prevalence rates females 
+px_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Prevalence colon and rectum cancer")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Colon and rectum cancer prevalence rates difference females") +
+  theme_classic()
+
+print(px_CC_plot_compare_female)
+
+##############################Incidence rate colon and rectum cancer #############################################################
+
+##Select data from data frames for colon and rectum cancer incidence rates males
+
+ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot incidence rate colon and rectum cancer males 
+ix_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Colon and rectum cancer incidence rate")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "colon and rectum cancer incidence rates difference males") +
+  theme_classic()
+
+print(ix_CC_plot_compare_male)
+
+##Select data from data frames for colon and rectum cancer incidence rates females
+
+ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+
+##Plot colon and rectum cancer incidence rates females 
+ix_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
+  geom_line(aes(color = "England"))+
+  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
+  labs(colour="Locations",x="Age",y="Incidence colon and rectum cancer")+
+  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+  scale_color_manual(values = c("blue","red")) +
+  labs (title = "Colon and rectum cancer incidence rates difference females") +
+  theme_classic()
+
+print(ix_CC_plot_compare_female)
+```
+
+```{r, echo=FALSE}
+
+plot(mx_plot_compare_male)
+plot(mx_plot_compare_female)
+plot(totalyld_plot_compare_male)
+plot(totalyld_plot_compare_female)
+plot(mx_ihd_plot_compare_male)
+plot(mx_ihd_plot_compare_female)
+plot(px_ihd_plot_compare_male)
+plot(px_ihd_plot_compare_female)
+plot(ix_ihd_plot_compare_male)
+plot(ix_ihd_plot_compare_female)
+plot(mx_is_plot_compare_male)
+plot(mx_is_plot_compare_female)
+plot(px_is_plot_compare_male)
+plot(px_is_plot_compare_female)
+plot(ix_is_plot_compare_male)
+plot(ix_is_plot_compare_female)
+plot(mx_DM_plot_compare_male)
+plot(mx_DM_plot_compare_female)
+plot(px_DM_plot_compare_male)
+plot(px_DM_plot_compare_female)
+plot(ix_DM_plot_compare_male)
+plot(ix_DM_plot_compare_female)
+plot(mx_CC_plot_compare_male)
+plot(mx_CC_plot_compare_female)
+plot(px_CC_plot_compare_male)
+plot(px_CC_plot_compare_female)
+plot(ix_CC_plot_compare_male)
+plot(ix_CC_plot_compare_female)
+plot(mx_BC_plot_compare_female)
+plot(px_BC_plot_compare_female)
+plot(ix_BC_plot_compare_female)
 ```
@@
 ####Model parameters
-
-We define parameters for our model that are used multiple times.
-*p_age_cohort* defines the starting age for the cohorts.
-*p_sex* is for the cohort sex.
-*p_disease* specifies the modeled diseases.
-*p_intervention* is a hypothetical increase in MET minutes per week.
-
-# ```{r, echo=TRUE}
-# #5
-# p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
-# p_sex <- c("males", "females")
-# p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
-# p_intervention_effect <- 100
-# ```
-
-####Data preparation
-
-The following code applies to the data used as an example here.
-#6 transforms upper cases to lower cases in the sex variable.
-#7 and #8 is used to rename mortality_rate to mx (used in life table language).
-#9 is used to replace male and female in the exposure data to males and females to facilitate combination of data sets.
-
-# ```{r, echo=TRUE}
-# #6
-# idata$sex <- tolower(idata$sex)
-# #7
-# idata$mx <- idata$mortality_rate
-# #8
-# idata$mortality_rate <- NULL
-# #9
-# edata$sex[edata$sex=="male"] <- "males"
-# edata$sex[edata$sex=="female"] <- "females"
-# ```
-
-# ```{r, echo=TRUE}
-# ##### Create new variable for 5_year population (age cohorts). Depends on the age-cohorts of interest (5-yrs here)
-# 
-# idata$five_year_population <- NA
-# start_index <- 3
-# index <- 1
-# end_index <- 4
-# for (i in 1:20){
-#   if (i == 1)
-#     index <- 1
-#   else 
-#     index <- ((i - 1)  * 5) + 1
-#   if (index == 96){
-#     end_index <- 5
-#   }
-#   cat("population sum ", start_index, "start index ", index, " and end index ", index + end_index, "\n")
-#   idata$five_year_population[start_index] <- sum(idata$population[index:(index + end_index)])
-#   start_index <- start_index + 5
-# }
-# 
-# end_index <- 4
-# start_index <- 104
-# 
-# for (i in 1:20){
-#   if (i == 1)
-#     index <- 102
-#   else
-#     index <- (((i - 1)  * 5) + 1) + 101
-#   if (index == 197){
-#     end_index <- 5
-#   }
-#   cat("population sum ", start_index, "start index ", index, " and end index ", index + end_index, "\n")
-#   idata$five_year_population[start_index] <- sum(idata$population[index:(index + end_index)])
-#   start_index <- start_index + 5
-# }
-# 
-# ```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
 

### 2018-04-13 MSLT/MSLTdoc.Rmd
Graphs GBD data for greater London and England

@@
      number_sections: true
-
      df_print: kable

### 2018-04-13 MSLT/MSLTdoc.Rmd
Added function to run comparison plots GBD data

@@
 #FOR NOW THE CODE FOR GRAPH IS repetitive
-####################MORTALITY RATES ALL CAUSE PER ONE
+
+#in_data1: GBDEngland
+#in_data2: GBDGL
+#in_sex: male, female
+#in_cause: all causes, ischemic heart disease, ischemic stroke, diabetes mellitus, colon and rectum cancers, breast cancer, tracheal, bronchus, and lung cancer
+#in_measure: deaths, incidence, prevalence, ylds (years lived with disability)
+
+plot_GBD_compare_death_all_cause_males <- function(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = "male", in_cause = "all causes", in_measure = "deaths")
 

### 2018-04-13 MSLT/code/functions.R

@@ -412,3 +412,35 @@ gen_aggregate <- function(in_data, in_cohorts, in_population, in_outcomes){
   
   aggr
 }
+
+
+#######################################Function to generate GBD graphs to compare data national to local############################
+
+
+plot_GBD <- function(in_data1, in_data2, in_sex, in_cause, in_measure) {
+  
+ in_data1 <- GBDEngland
+ in_data2 <- GBDGL
+ in_sex <- "male"
+ in_cause <- "all causes"
+ in_measure <- "deaths"
+
+
+data1 <- filter(in_data1, sex == in_sex, cause == in_cause & measure == in_measure) %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     
+                     
+data2 <- filter(in_data2, sex == in_sex, cause == in_cause & measure == in_measure) %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     
+
+  
+                      p <- ggplot(data1, aes(age_cat,one_rate))+
+                       geom_line(aes(color = "England"))+
+                       geom_line(data=data2,aes(color="Greater London"))+
+                       labs(colour="Locations",x="Age",y= paste(in_cause, in_measure, sep = " "))+
+                       theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+                       scale_color_manual(values = c("blue","red")) +
+                       labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " ")) +
+                       theme_classic()
+
+                     print(p)
+}
+
+plot_GBD_compare_death_all_cause_males <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = "male", in_cause = "all causes", in_measure = "deaths") 

### 2018-04-13 MSLT/MSLTdoc.Rmd
Added loop to generate graphs

--- 2018-04-13 MSLT/MSLTdoc.Rmd
@@
 ```{r, echo=FALSE}
-###TO DO: loops for sex, cause and measure and interpolation of rates, care with interpolating rates, need to convert to log for interpolation using cubic spline. 
-#FOR NOW THE CODE FOR GRAPH IS repetitive
-####################MORTALITY RATES ALL CAUSE PER ONE
+###TO DO: cubic splines for input rates
 
-###############################Mortality rates###################################################################
+#in_data1: GBDEngland
+#in_data2: GBDGL
 
-##Select data from data frames for all cause mortality males
+sex <- c("male", "female")
+cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancers", "breast cancer", "tracheal, bronchus, and lung cancer")
+measure <- c("all causes", "ylds (years lived with disability)", "prevalence", "incidence")
 
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
+for (i in sex){
+  for (j in cause) {
+    for (k in measure) {
+      p <- paste("Plot", "compare", sex, cause, measure, sep = "_")
+      assign(p, plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = sex, in_cause = cause, in_measure = measure))
+      print(p)
+    }
+  }
+}
 
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause mortality rates males 
-mx_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality rate")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Mortality rates difference males") +
-  theme_classic()
-
-print(mx_plot_compare_male)
-
-##Select data from data frames for all cause mortality females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause mortality rates males 
-mx_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality rate")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Mortality rates difference females") +
-  theme_classic()
-
-print(mx_plot_compare_female)
-
-#############YLD RATES ALL CAUSE PER ONE
-
-##################################yld rates all cause############################################################
-
-##Select data from data frames for all cause yld rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause yld rates males 
-totalyld_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Total YLDs rate")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Total YLDs rates difference males") +
-  theme_classic()
-
-print(totalyld_plot_compare_male)
-
-##Select data from data frames for all cause yld females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "all causes" & measure == "ylds (years lived with disability)") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause yld rates females 
-totalyld_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Total YLDs rate")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Total YLDs rates difference females") +
-  theme_classic()
-
-print(totalyld_plot_compare_female)
-
-################IHD
-##############################Mortality rate ihd #############################################################
-
-##Select data from data frames for ihd yld rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause yld rates males 
-mx_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality IHD")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "IHD mortality rates difference males") +
-  theme_classic()
-
-print(mx_ihd_plot_compare_male)
-
-##Select data from data frames for ihd mortality rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause yld rates females 
-mx_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality IHD")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "IHD mortality rates difference females") +
-  theme_classic()
-
-print(mx_ihd_plot_compare_female)
-
-##############################Prevalence rate ihd #############################################################
-
-##Select data from data frames for ihd prevalence rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates males 
-px_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence IHD")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "IHD prevalence rates difference males") +
-  theme_classic()
-
-print(px_ihd_plot_compare_male)
-
-##Select data from data frames for ihd prevalence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates females 
-px_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence IHD")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "IHD prevalence rates difference females") +
-  theme_classic()
-
-print(px_ihd_plot_compare_female)
-
-##############################Incidence rate ihd #############################################################
-
-##Select data from data frames for ihd prevalence rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates males 
-ix_ihd_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Incidence IHD")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "IHD incidence rates difference males") +
-  theme_classic()
-
-print(ix_ihd_plot_compare_male)
-
-##Select data from data frames for ihd prevalence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic heart disease" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates females 
-ix_ihd_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Incidence IHD")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "IHD incidence rates difference females") +
-  theme_classic()
-
-print(ix_ihd_plot_compare_female)
-
-################ISCHEMIC STROKE
-##############################Mortality rate ischemic stroke #############################################################
-
-##Select data from data frames for ischemic stroke mortality rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot ischemic stroke mortality rates males
-
-mx_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality IHD")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Ischemic stroke mortality rates difference males") +
-  theme_classic()
-
-print(mx_is_plot_compare_male)
-
-##Select data from data frames for ischemic stroke mortality rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot ischemic stroke mortality rates females 
-mx_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality ischemic stroke")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Ischemic stroke mortality rates difference females") +
-  theme_classic()
-
-print(mx_is_plot_compare_female)
-
-##############################Prevalence rate ischemic stroke #############################################################
-
-##Select data from data frames for ihd prevalence rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates males 
-px_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence IHD")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Ischemic stroke prevalence rates difference males") +
-  theme_classic()
-
-print(px_is_plot_compare_male)
-
-##Select data from data frames for ischemic stroke prevalence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates females 
-px_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence ischemic stroke")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Ischemic stroke prevalence rates difference females") +
-  theme_classic()
-
-print(px_is_plot_compare_female)
-
-##############################Incidence rate ischemic stroke #############################################################
-
-##Select data from data frames for ishcemic stroke incidence rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates males 
-ix_is_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Incidence ischemic stroke")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Ischemic stroke incidence rates difference males") +
-  theme_classic()
-
-print(ix_is_plot_compare_male)
-
-##Select data from data frames for ischemic stroke incidence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "ischemic stroke" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot ischemic stroke incidence rates females 
-ix_is_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Incidence ischemic stroke")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Ischemic stroke incidence rates difference females") +
-  theme_classic()
-
-print(ix_is_plot_compare_female)
-
-################DIABETES MELLITUS
-##############################Mortality rate diabetes mellitus #############################################################
-
-##Select data from data frames for diabetes mellitus mortality rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot ischemic stroke mortality rates males
-
-mx_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality diabetes mellitus")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Diabetes mellitus mortality rates difference males") +
-  theme_classic()
-
-print(mx_DM_plot_compare_male)
-
-##Select data from data frames for diabetes mellitus mortality rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot diabetes mellitus mortality rates females 
-mx_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality diabetes mellitus")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Diabetes mellitus mortality rates difference females") +
-  theme_classic()
-
-print(mx_DM_plot_compare_female)
-
-##############################Prevalence rate diabetes mellitus #############################################################
-
-##Select data from data frames for diabetes mellitus prevalence rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates males 
-px_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence diabetes mellitus")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Diabetes mellitus prevalence rates difference males") +
-  theme_classic()
-
-print(px_DM_plot_compare_male)
-
-##Select data from data frames for diabetes mellitus prevalence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates females 
-px_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence diabetes mellitus")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Ischemic stroke prevalence rates difference females") +
-  theme_classic()
-
-print(px_DM_plot_compare_female)
-
-##############################Incidence rate diabetes mellitus #############################################################
-
-##Select data from data frames for diabetes mellitus incidence rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot incidence rate diabetes mellitus males 
-ix_DM_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Diabetes mellitus")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Diabetes mellitus incidence rates difference males") +
-  theme_classic()
-
-print(ix_DM_plot_compare_male)
-
-##Select data from data frames for diabetes mellitus incidence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "diabetes mellitus" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot ischemic stroke incidence rates females 
-ix_DM_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Incidence diabetes mellitus")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Diabetes mellitus incidence rates difference females") +
-  theme_classic()
-
-print(ix_DM_plot_compare_female)
-
-#####Breast cancer
-##############################Mortality rate breast cancer #############################################################
-
-##Select data from data frames for breast cancer mortality rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot breast cancer mortality rates females 
-mx_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality breast cancer")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Breast cancer mortality rates difference females") +
-  theme_classic()
-
-print(mx_BC_plot_compare_female)
-
-##############################Prevalence rate breast cancer #############################################################
-
-##Select data from data frames for breast cancer prevalence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot all cause prevalence rates females 
-px_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence breast cancer")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Breast cancer prevalence rates difference females") +
-  theme_classic()
-
-print(px_BC_plot_compare_female)
-
-##############################Incidence rate breast cancer #############################################################
-
-##Select data from data frames for breast cancer incidence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "breast cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "breast cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot breast cancer incidence rates females 
-ix_BC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Incidence breast cancer")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Breast cancer incidence rates difference females") +
-  theme_classic()
-
-print(ix_BC_plot_compare_female)
-
-################Colon and rectum cancer (DOSE RESPONSE FOR COLON CANCER?)
-##############################Mortality rate colon and rectum cancer #############################################################
-
-##Select data from data frames for colon and rectum cancer mortality rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot colon and rectum cancer mortality rates males
-
-mx_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality colon and rectum cancer")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "colon and rectum cancer mortality rates difference males") +
-  theme_classic()
-
-print(mx_CC_plot_compare_male)
-
-##Select data from data frames for colon and rectum cancer mortality rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot colon and rectum cancer mortality rates females 
-mx_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Mortality colon and rectum cancer")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Colon and rectum cancers mortality rates difference females") +
-  theme_classic()
-
-print(mx_CC_plot_compare_female)
-
-##############################Prevalence rate colon and rectum cancer #############################################################
-
-##Select data from data frames for diabetes mellitus prevalence rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot colon and rectum cancer prevalence rates males 
-px_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence colon and rectum cancer")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "colon and rectum cancer prevalence rates difference males") +
-  theme_classic()
-
-print(px_CC_plot_compare_male)
-
-##Select data from data frames for colon and rectum cancer prevalence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "prevalence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot colon and rectum cancer prevalence rates females 
-px_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Prevalence colon and rectum cancer")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Colon and rectum cancer prevalence rates difference females") +
-  theme_classic()
-
-print(px_CC_plot_compare_female)
-
-##############################Incidence rate colon and rectum cancer #############################################################
-
-##Select data from data frames for colon and rectum cancer incidence rates males
-
-ggplot_data_England <- filter(GBDEngland, sex == "male", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "male", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot incidence rate colon and rectum cancer males 
-ix_CC_plot_compare_male <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Colon and rectum cancer incidence rate")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "colon and rectum cancer incidence rates difference males") +
-  theme_classic()
-
-print(ix_CC_plot_compare_male)
-
-##Select data from data frames for colon and rectum cancer incidence rates females
-
-ggplot_data_England <- filter(GBDEngland, sex == "female", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-ggplot_data_GL <- filter(GBDGL, sex == "female", cause == "colon and rectum cancer" & measure == "incidence") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)
-
-##Plot colon and rectum cancer incidence rates females 
-ix_CC_plot_compare_female <- ggplot(ggplot_data_England, aes(age_cat,one_rate))+
-  geom_line(aes(color = "England"))+
-  geom_line(data=ggplot_data_GL,aes(color="Greater London"))+
-  labs(colour="Locations",x="Age",y="Incidence colon and rectum cancer")+
-  theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-  scale_color_manual(values = c("blue","red")) +
-  labs (title = "Colon and rectum cancer incidence rates difference females") +
-  theme_classic()
-
-print(ix_CC_plot_compare_female)
 ```
@@
 ```{r, echo=FALSE}
+#Call plots here, add titles to figures
 
-plot(mx_plot_compare_male)
-plot(mx_plot_compare_female)
-plot(totalyld_plot_compare_male)
-plot(totalyld_plot_compare_female)
-plot(mx_ihd_plot_compare_male)
-plot(mx_ihd_plot_compare_female)
-plot(px_ihd_plot_compare_male)
-plot(px_ihd_plot_compare_female)
-plot(ix_ihd_plot_compare_male)
-plot(ix_ihd_plot_compare_female)
-plot(mx_is_plot_compare_male)
-plot(mx_is_plot_compare_female)
-plot(px_is_plot_compare_male)
-plot(px_is_plot_compare_female)
-plot(ix_is_plot_compare_male)
-plot(ix_is_plot_compare_female)
-plot(mx_DM_plot_compare_male)
-plot(mx_DM_plot_compare_female)
-plot(px_DM_plot_compare_male)
-plot(px_DM_plot_compare_female)
-plot(ix_DM_plot_compare_male)
-plot(ix_DM_plot_compare_female)
-plot(mx_CC_plot_compare_male)
-plot(mx_CC_plot_compare_female)
-plot(px_CC_plot_compare_male)
-plot(px_CC_plot_compare_female)
-plot(ix_CC_plot_compare_male)
-plot(ix_CC_plot_compare_female)
-plot(mx_BC_plot_compare_female)
-plot(px_BC_plot_compare_female)
-plot(ix_BC_plot_compare_female)
 ```

### 2018-04-13 MSLT/code/functions.R

@@ -443,4 +443,4 @@ data2 <- filter(in_data2, sex == in_sex, cause == in_cause & measure == in_measu
                      print(p)
 }
 
-plot_GBD_compare_death_all_cause_males <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = "male", in_cause = "all causes", in_measure = "deaths") 
+ 

### 2018-04-13 MSLT/MSLTdoc.Rmd
Loop graphs

@@
 
-sex <- c("male", "female")
-cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancers", "breast cancer", "tracheal, bronchus, and lung cancer")
-measure <- c("all causes", "ylds (years lived with disability)", "prevalence", "incidence")
+p_sex <- c("male", "female")
+p_cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancers", "breast cancer", "tracheal, bronchus, and lung cancer")
+p_measure <- c("all causes", "ylds (years lived with disability)", "prevalence", "incidence")
 
-for (i in sex){
-  for (j in cause) {
-    for (k in measure) {
-      p <- paste("Plot", "compare", sex, cause, measure, sep = "_")
-      assign(p, plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = sex, in_cause = cause, in_measure = measure))
-      print(p)
+for (sex in p_sex){
+  for (cause in p_cause) {
+    for (measure in p_measure) {
+      p1 <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = sex, in_cause = cause, in_measure = measure)
+      dev.new()
+      print(p1)
+
     }

### 2018-04-13 MSLT/code/functions.R

@@ -419,11 +419,11 @@ gen_aggregate <- function(in_data, in_cohorts, in_population, in_outcomes){
 
 plot_GBD <- function(in_data1, in_data2, in_sex, in_cause, in_measure) {
   
- in_data1 <- GBDEngland
- in_data2 <- GBDGL
- in_sex <- "male"
- in_cause <- "all causes"
- in_measure <- "deaths"
+ # in_data1 <- GBDEngland
+ # in_data2 <- GBDGL
+ # in_sex <- "male"
+ # in_cause <- "all causes"
+ # in_measure <- "deaths"
 
 
 data1 <- filter(in_data1, sex == in_sex, cause == in_cause & measure == in_measure) %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     

### 2018-04-16 MSLT/MSLTdoc.Rmd
Issues with loop for GBD_plot

@@
 
-ALSO PART OF THE BELOW CODE WILL BE NEEDED FOR THE MSLT AND DISMOD COLLECTION.
-GENERATE DATA INPUTS FOR DISMOD II FROM HERE.
 These data should be used to generate the general life table and disease life tables (Figure 3).
@@
 p_cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancers", "breast cancer", "tracheal, bronchus, and lung cancer")
-p_measure <- c("all causes", "ylds (years lived with disability)", "prevalence", "incidence")
+p_measure <- c("deaths", "ylds (years lived with disability)", "prevalence", "incidence")
 

### 2018-04-16 MSLT/code/functions.R

@@ -418,11 +418,11 @@ gen_aggregate <- function(in_data, in_cohorts, in_population, in_outcomes){
 
 
 plot_GBD <- function(in_data1, in_data2, in_sex, in_cause, in_measure) {
-  
+   
  # in_data1 <- GBDEngland
  # in_data2 <- GBDGL
- # in_sex <- "male"
- # in_cause <- "all causes"
+ # in_sex <- "female"
+ # in_cause <- "ischemic stroke"
  # in_measure <- "deaths"
 
 
@@ -431,16 +431,17 @@ data1 <- filter(in_data1, sex == in_sex, cause == in_cause & measure == in_measu
 data2 <- filter(in_data2, sex == in_sex, cause == in_cause & measure == in_measure) %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     
 
   
-                      p <- ggplot(data1, aes(age_cat,one_rate))+
-                       geom_line(aes(color = "England"))+
-                       geom_line(data=data2,aes(color="Greater London"))+
-                       labs(colour="Locations",x="Age",y= paste(in_cause, in_measure, sep = " "))+
-                       theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
-                       scale_color_manual(values = c("blue","red")) +
-                       labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " ")) +
-                       theme_classic()
+                    p <- ggplot(data = in_data1, aes(age_cat,one_rate))+
+                      geom_line(aes(color = "England"))+
+                      geom_line(data = in_data2, aes(age_cat,one_rate))+
+                      geom_line(aes(color = "Greater London"))+
+                      labs(colour="Locations",x="Age",y= paste(in_cause, in_measure, sep = " "))+
+                      theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
+                      scale_color_manual(values = c("blue","red")) +
+                      labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " ")) +
+                      theme_classic()
 
                      print(p)
 }
 
- 
+plot_test <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = "male", in_cause = "all causes", in_measure = "deaths")

### 2018-04-16 MSLT/MSLTdoc.Rmd
Loop plots, not solved

@@
 
+#USE demography package to interpolate moratlity rates.
+
 ##All upper cases to lower cases
@@
 p_sex <- c("male", "female")
-p_cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancers", "breast cancer", "tracheal, bronchus, and lung cancer")
+p_cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancer", "breast cancer", "tracheal, bronchus, and lung cancer")
 p_measure <- c("deaths", "ylds (years lived with disability)", "prevalence", "incidence")
@@
 
+##This works, check what is the issue with the loop. Seems that the problem was with colon and rectum cancer mispelling.
+
+p_test <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = "male", in_cause = "colon and rectum cancer", in_measure = "deaths")
 ```

### 2018-04-16 MSLT/MSLTdoc.Rmd
Plots compare finish
Improve labels and add graphs names

@@
 
-We start by cleaning the global environment (#1) to keep track of our works and ensure that the code is generating our outcomes.
-Then, we set up an option to avoid the use of scientific notation (#2) and lastly we load the functions (#3).
+We start by cleaning the global environment (1) to keep track of our works and ensure that the code is generating our outcomes.
+Then, we set up an option to avoid the use of scientific notation (2) and lastly we load the functions (3).
 The code chunks are shown in the rmarkdown output.
 
+1) Clearn Global Environment
 ```{r, echo=TRUE}
-#1
 rm (list = ls())
-#2
+```
+2) Avoid scientific notation
+```{r, echo=TRUE}
 options(scipen=999)
-#3
+```
+3) Load functions 
+```{r, echo=TRUE}
 source("code/functions.R")
@@
 Initial case studies for the ITHIMR are: London, Sao Pablo, Delhi, Accra, Los Angeles and Edinburgh.
-Here, we will start with **greater London** given the availability of diseases epidemiology data from the GBD IHME study.
-For the rest of the case study cities data is available at the country level, hence, a scaling method is needed.
+Here, we will start with **Greater London** given the availability of disease epidemiology data from the GBD IHME study.
+For the rest of the case study cities data is available at the country level, hence, a scaling method is needed to reflect the local burden of disease.
 
@@
 Black color coding with white color font represent final model outcomes.
-For both, the life table and disease life table, two sets of each are simulated, one for the baseline and the other for the scenario.
-**Data inputs** are from four sources: Sythetic population, GBD data (from the Institute of Health Metrics and Evaluations (IHME)), ITHIM calculations and Injuries model.
-The synthetic population provides population numbers by age-groups and sex which are used in the life table (**PMSLT Mechanisms**) with each age cohort simulated until they die or reach 100 years of age.
-Population numbers are also used to derive road death rates and road injuries rates from numbers for the baseline and scenario generated by the Injuries model.
-The difference in road deaths rates modifies all cause-mortatliy in the scenario life table and the change in injuries rates modifies total YLDs rates (hence, healthy life years).
-Data from GBD IHME includes: total YLD rates per one, disease specific YLDs (numbers) and prevalence to derive disability weights for the included diseases, all cause mortatliy rates per one, population numbers and diseases parameters (incidence, mortality and prevalence).
-Dismod II is used to derive internally consitent estimates incidence and case fatality rates from GBD IHME data, which are inputs of the disease specific life tables.
-Case fatality derive from Dismod II is the same for the baseline and scenario (however, could be changed) and incidence is different.
-For the baseline life tables incidence is as derived from Dismod II and for the scanrio is modified by the PIF (1-PIF).
-A change in incidence modifies prevalence and mortality in later years.
-The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.*](structure\Figure3.png){width=50%,height=30%,fig.pos="H"}
+For both, the life table and disease life table, two sets of each are simulated, one for the baseline and the other for the scenario.*](structure\Figure3.png){width=50%,height=30%,fig.pos="H"}
 
@@
 A prompt comes up asking for an email address.
-The data is sent to the designated email address (within minutes) in ZIP format, unzip and use the code below to read the data.
+The data is sent to the designated email address (within minutes) in ZIP format, unzip and use the code below to read the data (4).
 Here, we selected data for Greater London and England.
@@
 
+4) Read GBD data
 ```{r, echo=TRUE}
-options(scipen=999)
 GBDdata <- read.csv("data/IHME-GBD_2016_DATA-8fca12f6-1.csv", stringsAsFactors = F)
@@
 
-The following code serves to sort out the GBD data to the inputs required for the life table, disease life table and Dismod II
+The following codes serves to sort out the GBD data to the inputs required for the life table, disease life table and Dismod II
 
@@
 
+5) Change all uper cases to lower case
 ```{r, echo=TRUE}
-#Data preparation (generate age groups, change capitals to lower case)
+GBDdata <- mutate_all(GBDdata, funs(tolower))
+```
 
-#USE demography package to interpolate moratlity rates.
-
-##All upper cases to lower cases
-
-GBDdata <- mutate_all(GBDdata, funs(tolower))
-
-##Create age categories index in GBDdata (rounding mid age), total of 20 age groups. These are the age cohorts to simulate.
-
+6) Create age categories index in GBDdata (rounding mid age), total of 20 age groups. These are the age cohorts to simulate. 
+```{r, echo=TRUE}
 GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
@@
 GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97
+```
 
-##Create age and sex categories to obtain population numbers. Population numbers from GBD are used in 
-##Dismod II. For the Life table (Figure 3), the numbers may be from the synthetic population.
-
+7) Create age and sex categories to obtain population numbers. Population numbers from GBD are used in Dismod II. For the Life table (Figure 3), the numbers may be from the synthetic population. 
+```{r, echo=TRUE}
 GBDdata$sex_age_cat <- paste(GBDdata$age_cat,GBDdata$sex, sep = "_"  )
-
-##Convert variable "val" to numeric as its latter used in calculations.
-
+```
+8) Conver string variables to numberic to do calculations.
+```{r, echo=TRUE}
 GBDdata$val <- as.numeric(as.character(GBDdata$val))
-
-##Generate population numbers for England and Greater London in a new data frame ("GBD_population") and then separate in dataframes for England and Greater London (this avoids repeating the same code).
-
+```
+9) Generate population numbers for England and Greater London in a new data frame ("GBD_population") and then separate in dataframes for England and Greater London (this avoids repeating the same code).
+Population numbers are derived from rates per 100,000 and total numbers of cases.
+```{r, echo=TRUE}
 GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, val, sex_age_cat, location)
-
-###Generate population numbers from given number of cases and rates per 100,000 people. 
+```
+10) Generate population numbers from given number of cases and rates per 100,000 people. 
+```{r, echo=TRUE}
  for (i in 1:nrow(GBD_population)) {
@@
  }
-##Remove rows with zero
+```
+11) Remove rows with zero
+```{r, echo=TRUE}
 GBD_population <- GBD_population[!is.na(GBD_population$val_pop),]
-
-##Keep relevant variables
-
+```
+12) Keep relevant variables
+```{r, echo=TRUE}
 GBD_population <- filter(GBD_population) %>% select(sex_age_cat, val_pop, location)
-
-##Create data frames for Greater London and England population to be later used as inputs for PMSLT calculations
-
+```
+13) Create data frames for Greater London and England population to be later used as inputs for PMSLT calculations
+```{r, echo=TRUE}
 GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, val_pop, location)
@@
 GBD_population_England <-  filter(GBD_population, location == "england") %>% select(sex_age_cat, val_pop, location)
-
-##Check population total numbers
-
+```
+12) Check population total numbers
+```{r, echo=TRUE}
 GreaterLondon <- sum(GBD_population_GL$val_pop)
 England <- sum(GBD_population_England$val_pop)
+```
 
-##Generate data frames for England and Grater London with per person rates (per 100,000 in original data).
-
-#England
+13) Generate data frames for England and Grater London with per person rates (per 100,000 in original data).  
+```{r, echo=TRUE}
 GBDEngland <- filter(GBDdata, location == "england" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
-
 GBDEngland$one_rate <- GBDEngland$val/100000
-
-#Greater London (GL)
 GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
@@
 
-These graphs are to compare GBD data for England and Greater London.
+14) Generate graphs are to compare GBD data for England and Greater London.
 
@@
   for (cause in p_cause) {
-    for (measure in p_measure) {
+    for (measure in p_measure){
+
+# Exclude breast_cancer for Males
+      if (sex == "male" && cause == "breast cancer" || cause == "all causes" && measure == "prevalence" || cause == "all causes" && measure == "incidence"){
+        p1 <- NULL
+
+      }
+      else {
       p1 <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = sex, in_cause = cause, in_measure = measure)
-      dev.new()
-      print(p1)
 
@@
   }
+}
 
-##This works, check what is the issue with the loop. Seems that the problem was with colon and rectum cancer mispelling.
-
-p_test <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = "male", in_cause = "colon and rectum cancer", in_measure = "deaths")
 ```

### 2018-04-16 MSLT/code/functions.R
Loop graphs  names

@@ -436,12 +436,10 @@ data2 <- filter(in_data2, sex == in_sex, cause == in_cause & measure == in_measu
                       geom_line(data = data2, aes(age_cat,one_rate))+
                       geom_line(aes(color = "Greater London"))+
                       labs(colour="Locations",x="Age",y= paste(in_cause, in_measure, sep = " "))+
-                      theme(legend.position = c(0, 1),legend.justification = c(0, 1))+
                       scale_color_manual(values = c("blue","red")) +
-                      labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " ")) +
+                      labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " "), size=14) +
                       theme_classic()
 
                      print(p)
 }
 
-# plot_test <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = "male", in_cause = "all causes", in_measure = "deaths")

### 2018-04-16 MSLT/code/functions.R
Plots

@@ -421,8 +421,8 @@ plot_GBD <- function(in_data1, in_data2, in_sex, in_cause, in_measure) {
    
 # in_data1 <- GBDEngland
 # in_data2 <- GBDGL
-# in_sex <- "female"
-# in_cause <- "ischemic stroke"
+# in_sex <- "male"
+# in_cause <- "all causes"
 # in_measure <- "deaths"
 
 
@@ -436,10 +436,13 @@ data2 <- filter(in_data2, sex == in_sex, cause == in_cause & measure == in_measu
                       geom_line(data = data2, aes(age_cat,one_rate))+
                       geom_line(aes(color = "Greater London"))+
                       labs(colour="Locations",x="Age",y= paste(in_cause, in_measure, sep = " "))+
-                      scale_color_manual(values = c("blue","red")) +
-                      labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " "), size=14) +
+                      scale_color_manual(values=c("red", "blue"))+labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " "), size=14) +
                       theme_classic()
 
                      print(p)
 }
 
+# data1 <- filter(GBDEngland, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     
+# 
+# data2 <- filter(GBDGL, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     
+# 

### 2018-04-17 MSLT/MSLTdoc.Rmd
Loop graphs  names
Plots
Sort data

@@
 ```{r, echo=TRUE}
-GBDdata <- read.csv("data/IHME-GBD_2016_DATA-8fca12f6-1.csv", stringsAsFactors = F)
+GBDdata <- read.csv("MSLT/data/UK/englandandgreaterlondon.csv", stringsAsFactors = F)
 ```
@@
 
-14) Generate graphs are to compare GBD data for England and Greater London.
+14) Generate graphs to compare GBD data for England and Greater London.
 
@@
 
-```{r, echo=FALSE}
-#Call plots here, add titles to figures
-
-```
+FIX GRAPHS TITLES.
 

### 2018-04-17 MSLT/code/functions.R
FInalised compare plots UK

@@ -436,10 +436,9 @@ data2 <- filter(in_data2, sex == in_sex, cause == in_cause & measure == in_measu
                       geom_line(data = data2, aes(age_cat,one_rate))+
                       geom_line(aes(color = "Greater London"))+
                       labs(colour="Locations",x="Age",y= paste(in_cause, in_measure, sep = " "))+
-                      scale_color_manual(values=c("red", "blue"))+labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " "), size=14) +
+                      scale_color_manual(values=c("black", "blue")) +labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " "), size=14) + 
                       theme_classic()
-
-                     print(p)
+                    print(p)
 }
 
 # data1 <- filter(GBDEngland, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     

### 2018-04-17 MSLT/GBDCompare.Rmd
Separate compare plots to another doc

---
title: "GBDCompare"
author: "Belen Zapata Diomedi"
date: "17 April 2018"
output: pdf_document
---

The following graphs compare rates per one by age (5-year groups) and sex for England and Greater London.

```{r, echo = FALSE}
###TO DO: cubic splines for input rates

#Global options

#in_data1: GBDEngland
#in_data2: GBDGL
require(dplyr)
require(tidyverse)
require(knitr)
require(kableExtra)
require(citr)
source("code/functions.R")

GBDdata <- read.csv("data/UK/englandandgreaterlondon.csv", stringsAsFactors = F)
GBDdata <- mutate_all(GBDdata, funs(tolower))

GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97

GBDdata$sex_age_cat <- paste(GBDdata$age_cat,GBDdata$sex, sep = "_"  )

GBDdata$val <- as.numeric(as.character(GBDdata$val))

GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, val, sex_age_cat, location)

GBDEngland <- filter(GBDdata, location == "england" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
GBDEngland$one_rate <- GBDEngland$val/100000
GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
GBDGL$one_rate <- GBDGL$val/100000

p_sex <- c("male", "female")
p_cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancer", "breast cancer", "tracheal, bronchus, and lung cancer")
p_measure <- c("deaths", "ylds (years lived with disability)", "prevalence", "incidence")

for (sex in p_sex){
  for (cause in p_cause) {
    for (measure in p_measure){

# Exclude breast_cancer for Males
      if (sex == "male" && cause == "breast cancer" || cause == "all causes" && measure == "prevalence" || cause == "all causes" && measure == "incidence"){
        p1 <- NULL

      }
      else {
      p1 <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = sex, in_cause = cause, in_measure = measure)

      }
    }
  }
}

```

### 2018-04-17 MSLT/MSLTdoc.Rmd
FInalised compare plots UK
Separate compare plots to another doc

@@
 ```{r, echo=TRUE}
-GBDdata <- read.csv("MSLT/data/UK/englandandgreaterlondon.csv", stringsAsFactors = F)
+GBDdata <- read.csv("data/UK/englandandgreaterlondon.csv", stringsAsFactors = F)
 ```
@@
 
-14) Generate graphs to compare GBD data for England and Greater London.
-
-```{r, echo=FALSE}
-###TO DO: cubic splines for input rates
-
-#in_data1: GBDEngland
-#in_data2: GBDGL
-
-p_sex <- c("male", "female")
-p_cause <- c("all causes", "ischemic heart disease", "ischemic stroke", "diabetes mellitus", "colon and rectum cancer", "breast cancer", "tracheal, bronchus, and lung cancer")
-p_measure <- c("deaths", "ylds (years lived with disability)", "prevalence", "incidence")
-
-for (sex in p_sex){
-  for (cause in p_cause) {
-    for (measure in p_measure){
-
-# Exclude breast_cancer for Males
-      if (sex == "male" && cause == "breast cancer" || cause == "all causes" && measure == "prevalence" || cause == "all causes" && measure == "incidence"){
-        p1 <- NULL
-
-      }
-      else {
-      p1 <- plot_GBD(in_data1 = GBDEngland, in_data2 = GBDGL, in_sex = sex, in_cause = cause, in_measure = measure)
-
-      }
-    }
-  }
-}
-
-```
-
-FIX GRAPHS TITLES.
-
 Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.

### 2018-04-17 MSLT/code/functions.R

@@ -431,12 +431,12 @@ data1 <- filter(in_data1, sex == in_sex, cause == in_cause & measure == in_measu
 data2 <- filter(in_data2, sex == in_sex, cause == in_cause & measure == in_measure) %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     
 
   
-                    p <- ggplot(data = data1, aes(age_cat,one_rate))+
+                    p <- ggplot(data = data1, aes(age_cat,one_rate), color = "blue")+
                       geom_line(aes(color = "England"))+
-                      geom_line(data = data2, aes(age_cat,one_rate))+
+                      geom_line(data = data2, aes(age_cat,one_rate), color = "red")+
                       geom_line(aes(color = "Greater London"))+
                       labs(colour="Locations",x="Age",y= paste(in_cause, in_measure, sep = " "))+
-                      scale_color_manual(values=c("black", "blue")) +labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " "), size=14) + 
+                      labs (title = paste(in_cause, in_measure, in_sex, "compare", sep = " "), size=14) + 
                       theme_classic()
                     print(p)
 }

### 2018-04-17 MSLT/GBDCompare.Rmd
GBD COmpare to html
Add options to set messages and warnings to false

@@
 date: "17 April 2018"
-output: pdf_document
+output: html_document
 ---
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, message=FALSE, warning=FALSE}
 ###TO DO: cubic splines for input rates

### 2018-04-17 MSLT/MSLTdoc.Rmd
Added mortality spline to derive between ages rates

@@
 require(citr)
+require(demography)
 ```
@@
 
+14) For the life table, we need to mortatlity and total yld rates in one year age intervals.
+The original data is in five years.
+Thus, the following code is used to interpolate a single-year age distribution form a five-yearly distribution.
+
+```{r, echo=TRUE}
+##Example for Greater London. First, extract rates of interest and add to GBD_population frame (change the name of the data frame).
+##COULD USE A FUNCTION AND ALOOP AS WITH THE GRAPHS TO GENERATE DATA FOR EACH OF THE DISEASE FOR DISMOD.
+##Create dataframe with mortality data males for Greater London (test spline, then need to better organise the row data from GBD)
+
+Mortality_males <- filter(GBDGL, measure == "deaths", sex == "male", cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate)
+
+x=Mortality_males$age_cat
+y=log(Mortality_males$one_rate)
+plot(x,y)
+Mortality_males_interpolated <- as.data.frame(spline(x, y, n = 5 *length (x), xmin = min(x-2), xmax = max(x+3)))
+Mortality_males_interpolated$y <- exp(Mortality_males_interpolated$y)
+
+```
 Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.

### 2018-04-17 MSLT/MSLTdoc.Rmd
Interpolation with stat function splines

--- 2018-04-17 MSLT/MSLTdoc.Rmd
@@
 
+14) For the life table, we need to mortatlity and total yld rates in one year age intervals.
+The original data is in five years.
+Thus, the following code is used to interpolate a single-year age distribution form a five-yearly distribution.
+
+```{r, echo=TRUE}
+##Example for Greater London. First, extract rates of interest and add to GBD_population frame (change the name of the data frame).
+##COULD USE A FUNCTION AND ALOOP AS WITH THE GRAPHS TO GENERATE DATA FOR EACH OF THE DISEASE FOR DISMOD.
+##Create dataframe with mortality data males for Greater London (test spline, then need to better organise the row data from GBD)
+##I used spline function as the rates are not always monotonic.
+
+Mortality_males <- filter(GBDGL, measure == "deaths", sex == "male", cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate)
+
+x=Mortality_males$age_cat
+y=log(Mortality_males$one_rate)
+plot(x,y)
+Mortality_males_interpolated <- as.data.frame(spline(x, y, n = 5 *length (x), xmin = min(x-2), xmax = max(x + 2)))
+Mortality_males_interpolated$y <- exp(Mortality_males_interpolated$y)
+plot(Mortality_males_interpolated$x,Mortality_males_interpolated$y)
+```
 Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.

### 2018-04-18 MSLT/MSLTdoc.Rmd
fix knitt issue

@@
 
-In this model we use the term ‘health-adjusted life year’ (HALY).
-As ‘summary measure of population health’ it measures both quantity and quality of life, where one HALY represent the equivalent of one year in full health (which could be two years with a quality of life of 0.5, for example).
+In this model we use the term *health-adjusted life year* (HALY).
+As *summary measure of population health* it measures both quantity and quality of life, where one HALY represent the equivalent of one year in full health (which could be two years with a quality of life of 0.5, for example).
 Specific types of HALY are the quality-adjusted life year (QALY) and the disability-adjusted life year (DALY).
@@
 
-1) Clearn Global Environment
+1) Clean Global Environment
 ```{r, echo=TRUE}

### 2018-04-18 MSLT/MSLTdoc.Rmd
Text and code for MSLT.doc for Greater London

@@
 The basic infrastructure of the model consist of three components: (1) Effect size for the intervention of interest (e.g.
-intervention to urban design that modifies population levels of physical activity); (2) Calculation of the potential impact fraction (PIF) to derive the change in occurrence of disease (incidence rate/mortality rate) attributable to a change in the distribution of the risk factor (e.g.
+intervention to urban design that modifies population levels of physical activity); (2) Calculation of the potential impact fraction (PIF) to derive the change in occurrence of disease (incidence rate/case fatality rate) attributable to a change in the distribution of the risk factor (e.g.
 physical activity); and (3) Use of the PMSLT to simulate health (and economic) outcomes attributable to a change in the distribution of health risk factor/s in the population of interest.
@@
 QALYs use measures of utility weights that traditionally represent individual experiences of health, whereas our estimated HALYs use disability weights linked to specific diseases, which were developed for the Global Burden of Disease study [@RN8158].
-As discussed in past research [@RN21; @RN8262] the main advantage of using disability weights over utility weights is that disability weights refer to specific diseases rather than health states.
+As discussed in past research [@RN21; @RN8262] the main advantage of using disability weights over utility weights is that disability weights refer to specific diseases rather than health states (which are difficult to link to risk factors-e.g.
+physical inactivity).
 We opted to use the more general terms HALYs given that the use of the DALYs terminology may lead to think that our calculations are similar to those in burden of diseases studies [@RN8153].
@@
 A step further back, is the development of scenarios that bring about change in the distribution of the risk factor of interest.
-For now, we only focus on calculations from the PIF onward, and provide a hypothetical example of change in population levels of physical activity.
+For now, we only focus on calculations from the PIF onward, and provide a hypothetical example of change in the population distribution of physical activity.
 Incorporation of additional  health risk factor (air pollution, road trauma, NO2 and noise) will be discussed in the relevant code sections.
@@
 For example, as set up here, we simulate sex and age (5 years starting at 18) cohorts over time until they die or reach 100 years of age.
-This implies that we can include trends for diseases, time lags between change in exposure to risk factors and change in health and demographic changes (e.g.
+This implies that we can include trends for diseases, time lags for change in exposure to risk factors and change in health and demographic changes (e.g.
 population growth).
@@
 Where possible, we wrote functions and loops to avoid repetition.
-We set up the model  with Australian data, for Melbourne.
+We set up the model  with data for Great London.
 Figure 2 is depicts the PMSLT model framework, which was followed in the code development.
@@
 Here we only included the physical activity health pathway.
-In the comments section, implementation of exposure to air pollution and road trauma is discussed.
+In the comments section, we discuss the implementation of exposure to air pollution and road trauma.
 Note that in the presentation of input parameters, those needed to calculate PIFs are excluded, as these are common to the ITHIM, expect if trends are included (refer to comments section).
@@
 
-Inputs of the life table are: population numbers by sex (per 1-year or age grouping of interest), mortality rates or probability of all cause mortality by single age group and sex and total prevalent years lived with disability rate per one year by sex.
+Inputs of the life table are: population numbers by sex (per 1-year or age grouping of interest), mortality rates or probability of all cause mortality by single age group and sex and total prevalent years lived with disability rate per single year by sex.
 Disease specific disability weights are presented as inputs here as these adjust the total years lived with disability, hence, the health-adjusted life years.
@@
 
-These data will be provided by the synthetic population.
-In the code presented here, we created 5-year age and sex cohorts from one-year age groups data.
+These data will be provided by the synthetic population or derived from other data.
+In the code presented here, we simulate 5-year age and sex cohorts.
+Data for population may be in five-year age groups or one-year.
+For the example for Greater London, we derive 5-year age groups from GBD IHME data, however, we also provide a script if five-year age groups are to be derived from one-year age groups data.
 I left potential data sources below as a reference.
@@
 These data are available from GBD IHME, however, in age groups (1-4, 5-9, etc).
-Interpolation can be used to derive in between ages rates (cubic spline).
+We provide an interpolation script to derive in between ages rates (cubic spline).
 
 Note that we need data for population numbers and all cause mortality rates for: (1) PMSLT and (2) Dismod II collection (more in Dismod II section).
-Population data from the synthetic population is used for the PMSLT.
-For Dismod II, population and mortality data should be from the same source (GBD IHME)
+Population data from the synthetic population is used for the PMSLT (if available).
+For Dismod II, population and mortality data should be from the same source (GBD IHME).
 
@@
 
-These data is available from the GBD (http://ghdx.healthdata.org/gbd-results-tool) per 5-year age groups. We can use interpolation to derive between ages rates.
+These data is available from the GBD (http://ghdx.healthdata.org/gbd-results-tool) per 5-year age groups. We can use interpolation to derive between ages rates (script provided).
 
@@
 
-Disability weights (DW) can be derived from disease specific years lived with disability (YLD) and disease specific prevalence by age group (5 years) and sex.
-Data for YLDs prevalence can be obtained from the online GBD IHME data tool (http://ghdx.healthdata.org/gbd-results-tool).
+Disability weights (DW) are derived from disease specific years lived with disability (YLD) and disease specific prevalence by age group (5 years) and sex.
+Data for YLDs prevalence is obtained from the online GBD IHME data tool (http://ghdx.healthdata.org/gbd-results-tool).
 Our calculations of DW in the example here based on the GBD methods for estimating YLDs as the sum of sequelae prevalence multiplied by sequelae disability weights (REF GBD).
@@
 For each of the modeled diseases the PMSLT needs incidence and case fatality rates per sex and one-year intervals.
-Data from the GBD IHME studies with Dismod II (free at https://www.epigear.com/index_files/dismod_ii.html) can be used to derive internally consistent data and generate missing data.
+Data from the GBD IHME studies with Dismod II (free at https://www.epigear.com/index_files/dismod_ii.html) is used to derive internally consistent data and generate missing data.
 For example, the GBD studies provide data for incidence, prevalence and disease mortality, however, not case fatality.
@@
 
-In what follows, we start with the **model** script file, and explain the **functions** script file as these are used to perform calculations.
-The **functions** will be explained in detailed to provide clarity for required inputs.
+In what follows, we start explain the script step by step.
 
@@
 
-We start by cleaning the global environment (1) to keep track of our works and ensure that the code is generating our outcomes.
+We start by cleaning the global environment (1) to keep track of our works and ensure that the code is generating our desired outcomes.
 Then, we set up an option to avoid the use of scientific notation (2) and lastly we load the functions (3).
@@
 
-7) Create age and sex categories to obtain population numbers. Population numbers from GBD are used in Dismod II. For the Life table (Figure 3), the numbers may be from the synthetic population. 
+7) Create age and sex categories to obtain population numbers.
+Population numbers from GBD are used in Dismod II.
+For the Life table (Figure 3), the numbers may be from the synthetic population.
+For now, the Life table is set up with population numbers derived from the GBD data.
 ```{r, echo=TRUE}
@@
 ```
-9) Generate population numbers for England and Greater London in a new data frame ("GBD_population") and then separate in dataframes for England and Greater London (this avoids repeating the same code).
+9) Generate population numbers for Greater London in a new data frame ("GBD_population").
+Note that there is data for England as well, which is used in a separate rmarkdown dowcument (GBDCompare).
 Population numbers are derived from rates per 100,000 and total numbers of cases.
@@
 ```
-13) Create data frames for Greater London and England population to be later used as inputs for PMSLT calculations
+13) Create data frames for Greater London to be later used for: a) interpolation of rates and b) PMLT cohorts. 
 ```{r, echo=TRUE}
 GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, val_pop, location)
-
-GBD_population_England <-  filter(GBD_population, location == "england") %>% select(sex_age_cat, val_pop, location)
 ```
@@
 GreaterLondon <- sum(GBD_population_GL$val_pop)
-England <- sum(GBD_population_England$val_pop)
 ```
 
-13) Generate data frames for England and Grater London with per person rates (per 100,000 in original data).  
+13) Generate data frames for Greater London with per person rates (per 100,000 in original data).  
 ```{r, echo=TRUE}
-GBDEngland <- filter(GBDdata, location == "england" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
-GBDEngland$one_rate <- GBDEngland$val/100000
 GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
@@
 
-14) For the life table, we need to mortatlity and total yld rates in one year age intervals.
+14) For the life table, we need to mortality and total yld rates in one year age intervals.
 The original data is in five years.
@@
 ##I used spline function as the rates are not always monotonic.
+
+##Mortality rates
 

### 2018-04-18 MSLT/code/functions.R

@@ -413,8 +413,11 @@ gen_aggregate <- function(in_data, in_cohorts, in_population, in_outcomes){
   aggr
 }
 
+#######################################Funtion to interpolate rates##############################
 
-#######################################Function to generate GBD graphs to compare data national to local############################
+
+
+#######################################Function to generate GBD graphs to compare data national to local (USED in GBD COMPARE############################
 
 
 plot_GBD <- function(in_data1, in_data2, in_sex, in_cause, in_measure) {
@@ -440,7 +443,3 @@ data2 <- filter(in_data2, sex == in_sex, cause == in_cause & measure == in_measu
           print(p)
 }
 
-# data1 <- filter(GBDEngland, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     
-# 
-# data2 <- filter(GBDGL, sex == "male", cause == "all causes" & measure == "deaths") %>% select(measure, location, sex, age, metric, cause, one_rate, age_cat)     
-# 

### 2018-04-18 MSLT/MSLTdoc.Rmd
loop spline

@@
 ```{r, echo=TRUE}
-##Example for Greater London. First, extract rates of interest and add to GBD_population frame (change the name of the data frame).
-##COULD USE A FUNCTION AND ALOOP AS WITH THE GRAPHS TO GENERATE DATA FOR EACH OF THE DISEASE FOR DISMOD.
-##Create dataframe with mortality data males for Greater London (test spline, then need to better organise the row data from GBD)
 ##I used spline function as the rates are not always monotonic.
 
-##Mortality rates
+##Loop to generate interpolated values for all cause mortality and all-cause years lived with disability.
 
-Mortality_males <- filter(GBDGL, measure == "deaths", sex == "male", cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate)
+i_sex <- c("male", "female")
+i_measure <- c("deaths", "ylds (years lived with disability)")
 
-x=Mortality_males$age_cat
-y=log(Mortality_males$one_rate)
+for(sex in i_sex) {
+  for (measure in i_measure) {
+
+data <- filter(GBDGL, measure == "ylds (years lived with disability)", sex == "female", cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate)
+
+x=data$age_cat
+y=log(data$one_rate)
+
+interpolated<- as.data.frame(spline(x, y, n = 5 *length (x), xmin = min(x-2), xmax = max(x + 2)))
+interpolated$y <- exp(y)
 plot(x,y)
-Mortality_males_interpolated <- as.data.frame(spline(x, y, n = 5 *length (x), xmin = min(x-2), xmax = max(x + 2)))
-Mortality_males_interpolated$y <- exp(Mortality_males_interpolated$y)
-plot(Mortality_males_interpolated$x,Mortality_males_interpolated$y)
+
+  }
+}
+
+##Test here to generate all 4 data frames simultaneously.
+
+i_sex <- c("male", "female")
+i_measure <- c("deaths", "ylds (years lived with disability)")
+
+for(sex in i_sex) {
+  for (measure in i_measure) {
+
+    data <- filter(GBDGL, measure == "ylds (years lived with disability)", sex == "female", cause == "all       causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate)
+    assign(paste(i_sex, i_measure, "interpolated_data", sep = " "), data)
+x=data$age_cat
+y=log(data$one_rate)
+
+interpolated <- as.data.frame(spline(x, y, n = 5 *length (x), xmin = min(x-2), xmax = max(x + 2)))
+interpolated$y <- exp(y)
+assign(paste(i_sex, i_measure, "interpolated", sep = " "), interpolated)
+plot(x,y)
+
+  }
+}
+
 ```

### 2018-04-18 MSLT/code/functions.R

@@ -413,9 +413,6 @@ gen_aggregate <- function(in_data, in_cohorts, in_population, in_outcomes){
   aggr
 }
 
-#######################################Funtion to interpolate rates##############################
-
-
 
 #######################################Function to generate GBD graphs to compare data national to local (USED in GBD COMPARE############################
 

### 2018-04-18 MSLT/MSLTdoc.Rmd
loop spline

@@
 
-##Loop to generate interpolated values for all cause mortality and all-cause years lived with disability.
-
-i_sex <- c("male", "female")
-i_measure <- c("deaths", "ylds (years lived with disability)")
-
-for(sex in i_sex) {
-  for (measure in i_measure) {
-
 data <- filter(GBDGL, measure == "ylds (years lived with disability)", sex == "female", cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate)
@@
 
-  }
-}
-
 ##Test here to generate all 4 data frames simultaneously.
@@
 
-    data <- filter(GBDGL, measure == "ylds (years lived with disability)", sex == "female", cause == "all       causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate)
+    data <- filter(GBDGL, measure == i_measure, sex == i_sex, cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate) 
     assign(paste(i_sex, i_measure, "interpolated_data", sep = " "), data)

### 2018-04-18 MSLT/MSLTdoc.Rmd
Remove brackets from var names and fix var names

@@
 i_sex <- c("male", "female")
-i_measure <- c("deaths", "ylds (years lived with disability)")
+i_measure <- c("deaths", "ylds") #" (years lived with disability)")
 
@@
 
-    data <- filter(GBDGL, measure == i_measure, sex == i_sex, cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate) 
-    assign(paste(i_sex, i_measure, "interpolated_data", sep = " "), data)
+    #print(paste(sex, measure, "interpolated_data", sep = "_"))
+    #print(paste(sex, measure, "interpolated", sep = "_"))
+
+    data <- filter(GBDGL, measure == i_measure, sex == sex, cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate) 
+    assign(paste(sex, measure, "interpolated_data", sep = "_"), data)
 x=data$age_cat
@@
 interpolated$y <- exp(y)
-assign(paste(i_sex, i_measure, "interpolated", sep = " "), interpolated)
+assign(paste(sex, measure, "interpolated", sep = "_"), interpolated)
 plot(x,y)

### 2018-04-18 MSLT/MSLTdoc.Rmd
Fixed spline
the argument n within the spline was 5 * 5, should only be x

@@
 
-6) Create age categories index in GBDdata (rounding mid age), total of 20 age groups. These are the age cohorts to simulate. 
+6) Create age categories index in GBDdata (upper limit age group to match injury data), total of 20 age groups. These are the age cohorts to simulate. 
 ```{r, echo=TRUE}
@@
 ```{r, echo=TRUE}
-##I used spline function as the rates are not always monotonic.
 
@@
 
-##Test here to generate all 4 data frames simultaneously.
+##Loop to generate interpolated rates for all cause mortality and ylds for males and females
 
@@
 
+plot(male_deaths_interpolated_data$age_cat, male_deaths_interpolated_data$one_rate)
+
+plot(female_deaths_interpolated$x, female_deaths_interpolated$y)
+
 ```

### 2018-04-18 MSLT/MSLTdoc.Rmd
Replaces spline for splinefunc
The spline function is best to use here, as before I was not interpolating within ages.

@@
 ```{r, echo=TRUE}
-GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
-GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
-GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
-GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
-GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
-GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
-GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
-GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
-GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
-GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
-GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
-GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
-GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
-GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
-GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
-GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
-GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
-GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
-GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
-GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97
+GBDdata$age_cat [GBDdata$age =="under 5"] <- 5
+GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 10
+GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 15
+GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 20
+GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 25
+GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 30
+GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 35
+GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 40
+GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 45
+GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 50
+GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 55
+GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 60
+GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 65
+GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 70
+GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 75
+GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 80
+GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 85
+GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 90
+GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 95
+GBDdata$age_cat [GBDdata$age =="95 plus"] <- 100
 ```
@@
 
-data <- filter(GBDGL, measure == "ylds (years lived with disability)", sex == "female", cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate)
-
-x=data$age_cat
-y=log(data$one_rate)
-
-interpolated<- as.data.frame(spline(x, y, n = 5 *length (x), xmin = min(x-2), xmax = max(x + 2)))
-interpolated$y <- exp(y)
-plot(x,y)
-
 ##Loop to generate interpolated rates for all cause mortality and ylds for males and females
@@
 
-    #print(paste(sex, measure, "interpolated_data", sep = "_"))
-    #print(paste(sex, measure, "interpolated", sep = "_"))
-
 data <- filter(GBDGL, measure == i_measure, sex == sex, cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate) 
@@
 
-interpolated <- as.data.frame(spline(x, y, n = 5 *length (x), xmin = min(x-2), xmax = max(x + 2)))
-interpolated$y <- exp(y)
+interpolation_func <- splinefun(x, y, method = "natural", ties = mean)
+
+interpolated <- as.data.frame(interpolation_func(seq(0, 99, 1)))
+ages <- seq(0,99,by=1)
+interpolated <- cbind(interpolated, ages)
+interpolated[,1] <- exp(interpolated[,1])
+colnames(interpolated)[1] <- "rate"
+
 assign(paste(sex, measure, "interpolated", sep = "_"), interpolated)
-plot(x,y)
+plot(interpolated$age,interpolated$rate)
 
@@
 }
-
-plot(male_deaths_interpolated_data$age_cat, male_deaths_interpolated_data$one_rate)
-
-plot(female_deaths_interpolated$x, female_deaths_interpolated$y)
 

### 2018-04-18 MSLT/MSLTdoc.Rmd
Checking different spline methods

@@
 plot(interpolated$age,interpolated$rate)
-
+ p <- ggplot(data = interpolated, aes(age,rate)) +
+            geom_line(aes(color = "Interpolated")) 
+            theme_classic()
+            print(p)
   }

### 2018-04-18 MSLT/MSLTdoc.Rmd
Generate inputs life table
From produced population numbers, mortality rates and yld rates. TO DO

--- 2018-04-18 MSLT/MSLTdoc.Rmd
@@
 
+###Do graph with another layer for original rates in age groups for comparison purposes. 
+ # p <- ggplot(data = interpolated, aes(age,rate)) +
+ #            geom_line(aes(color = "Interpolated")) 
+ #            theme_classic()
+ #            print(p)
   }
@@
 ```
+
+15) The following code brings together all the required data to create the inputs life table (Figure 3)
+
 Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.

### 2018-04-19 MSLT/MSLTdoc.Rmd
Deleted () from ylds variable

@@
 
-5) Change all uper cases to lower case
+5) Change all uper cases to lower case and delete () from variables. 
 ```{r, echo=TRUE}
 GBDdata <- mutate_all(GBDdata, funs(tolower))
+GBDdata$measure[GBDdata$measure=="ylds (years lived with disability)"] <- "ylds"
 ```

### 2018-04-19 MSLT/MSLTdoc.Rmd
Fix variable names

@@
 
-for(sex in i_sex) {
-  for (measure in i_measure) {
+for(sex_index in i_sex) {
+  for (measure_index in i_measure) {
 
-data <- filter(GBDGL, measure == i_measure, sex == sex, cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate) 
-assign(paste(sex, measure, "interpolated_data", sep = "_"), data)
+data <- filter(GBDGL, measure == measure_index, sex == sex_index, cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate) 
+assign(paste(sex_index, measure_index, "interpolated_data", sep = "_"), data)
 x=data$age_cat
@@
 
-assign(paste(sex, measure, "interpolated", sep = "_"), interpolated)
+assign(paste(sex_index, measure_index, "interpolated", sep = "_"), interpolated)
 plot(interpolated$age,interpolated$rate)

### 2018-04-19 MSLT/MSLTdoc.Rmd
Added original data to interpolated data graphs

@@
 assign(paste(sex_index, measure_index, "interpolated", sep = "_"), interpolated)
-plot(interpolated$age,interpolated$rate)
+# plot(interpolated$age,interpolated$rate)
 
-###Do graph with another layer for original rates in age groups for comparison purposes. 
- # p <- ggplot(data = interpolated, aes(age,rate)) +
- #            geom_line(aes(color = "Interpolated")) 
- #            theme_classic()
- #            print(p)
+##Do graph with another layer for original rates in age groups for comparison purposes.
+p <- ggplot() +
+            geom_line (
+              data = interpolated,
+              mapping = aes(ages,rate) 
+              ) +
+              geom_point(
+                data = data, 
+                mapping = aes(age_cat, one_rate)
+                ) +
+           theme_classic()
+           print(p)
   }

### 2018-04-19 MSLT/MSLTdoc.Rmd
Graphs aes

@@
 ##Do graph with another layer for original rates in age groups for comparison purposes.
-p <- ggplot() +
-            geom_line (
-              data = interpolated,
-              mapping = aes(ages,rate) 
-              ) +
+p <- ggplot(data = interpolated,mapping = aes(ages,rate)) +
+              geom_line(aes(color = "Interpolated")) +
               geom_point(
                 data = data, 
-                mapping = aes(age_cat, one_rate)
-                ) +
-           theme_classic()
+                mapping = aes(age_cat, one_rate, color = "Original")) +
+              labs(colour="",x="Age",y= paste(i_sex, i_measure, "rates", sep = " "))+
+            labs (title = paste("Rates", i_sex, i_measure, sep = " "), size=14) +
+            theme_classic() +
+            theme(plot.title = element_text(hjust = 0.5))
            print(p)

### 2018-04-19 MSLT/code/model.R

@@ -582,7 +582,8 @@ total_plot <- ggplot(total_aggr[1:20,], aes(x = sim_year)) +
   geom_line(mapping = aes(y = total_inc_num_diff_ihd, colour = "total_mx_num_diff_ihd")) +
   xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
   theme(plot.title = element_text(hjust = 0.5, size = 12)) +
-  scale_color_discrete(name = paste(""), labels = c("Baseline", "Scenario", "Difference"))
+  scale_color_discrete(name = paste(""), labels = c("Baseline", "Scenario", "Difference")) +
+  theme(plot.title = element_text(hjust = 0.5))
   
 
 #Print to view

### 2018-04-19 MSLT/MSLTdoc.Rmd
life table input

@@
 ```{r, echo=TRUE}
-GBDdata$sex_age_cat <- paste(GBDdata$age_cat,GBDdata$sex, sep = "_"  )
+GBDdata$sex_age_cat <- paste(GBDdata$sex,GBDdata$age_cat, sep = "_"  )
 ```
@@
 ```{r, echo=TRUE}
-GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, val, sex_age_cat, location)
+GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, sex, val, sex_age_cat, location)
 ```
@@
     if (GBD_population$metric[i] == "number") {
-      GBD_population$val_pop[i] <- GBD_population$val[i] * 100000/ GBD_population$val[i + 2]}
-   else {GBD_population$val_pop[i] <- NA}
+      GBD_population$pop_five_year[i] <- GBD_population$val[i] * 100000/ GBD_population$val[i + 2]}
+   else {GBD_population$pop_five_year[i] <- NA}
  }
@@
 ```{r, echo=TRUE}
-GBD_population <- GBD_population[!is.na(GBD_population$val_pop),]
+GBD_population <- GBD_population[!is.na(GBD_population$pop_five_year),]
 ```
@@
 ```{r, echo=TRUE}
-GBD_population <- filter(GBD_population) %>% select(sex_age_cat, val_pop, location)
+GBD_population <- filter(GBD_population) %>% select(sex_age_cat, sex, age_cat, pop_five_year, location)
 ```
@@
 ```{r, echo=TRUE}
-GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, val_pop, location)
+GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, age_cat, sex, pop_five_year, location)
 ```
@@
 ```{r, echo=TRUE}
-GreaterLondon <- sum(GBD_population_GL$val_pop)
+GreaterLondon <- sum(GBD_population_GL$pop_five_year)
 ```
@@
 
+```{r, echo=TRUE}
+##Create life table input data frame (population numbers, all-cause moratlity rates ##(per one) and all cause yld rates)
+
+##Create empty data frame
+input_life_table <- as.data.frame(matrix(0, ncol=5, nrow = 202))
+##Give variables names (mx=total mortality rate, yldsx = total years lived with disability rates )
+names(input_life_table) <- c("age", "sex", "pop_five_year", "mx", "yldsx")
+##Populate life table: Age, 0 to 100
+input_life_table[1:101, 1] <- c(0:100)
+input_life_table[102:202, 1] <- c(0:100)
+##Populate life table: sex, male, female
+input_life_table[1:101, 2] <- "male"
+input_life_table[102:202, 2] <- "female"
+##Create variable age_sex to match with population data. 
+input_life_table$sex_age_cat <- paste(input_life_table$sex,input_life_table$age, sep = "_"  )
+
+##Populate life table: five_year_pop. SEE LIFE TABLE LOOPS THAT USES 5 YEAR POP. 
+input_life_table$pop_five_year <- if(
+  input_life_table$sex_age_cat == GBD_population_GL$pop_five_year, GBD_population_GL$pop_five_year, NA)
+
+```
 Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.

### 2018-04-19 MSLT/code/model.R

@@ -37,7 +37,7 @@ p_sex <- c("males", "females")
 
 p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
 
-###As an expample, an increase in 1000 MET-mins per week
+###As an expample, an increase in 100 MET-mins per week
 
 p_intervention_effect <- 100
 

### 2018-04-20 MSLT/MSLTdoc.Rmd
Remove pop_five_year from input_life_table dataframe and then join it GBD_population_gl to get pop_five_year

@@
 ##Create empty data frame
-input_life_table <- as.data.frame(matrix(0, ncol=5, nrow = 202))
+input_life_table <- as.data.frame(matrix(0, ncol=4, nrow = 202))
 ##Give variables names (mx=total mortality rate, yldsx = total years lived with disability rates )
-names(input_life_table) <- c("age", "sex", "pop_five_year", "mx", "yldsx")
+names(input_life_table) <- c("age", "sex", "mx", "yldsx")
 ##Populate life table: Age, 0 to 100
@@
 ##Populate life table: five_year_pop. SEE LIFE TABLE LOOPS THAT USES 5 YEAR POP.
-input_life_table$pop_five_year <- if(
-  input_life_table$sex_age_cat == GBD_population_GL$pop_five_year, GBD_population_GL$pop_five_year, NA)
+
+input_life_table <- merge(input_life_table, select(GBD_population_GL
+                                                     ,c(sex_age_cat, pop_five_year))
+                                                     , by = 'sex_age_cat', all.x = TRUE)
+
+#input_life_table$pop_five_year <- ifelse(
+#  input_life_table$sex_age_cat == GBD_population_GL$pop_five_year, #GBD_population_GL$pop_five_year, NA)
 

### 2018-04-20 MSLT/MSLTdoc.Rmd
Order input life dataset

@@
 
-#input_life_table$pop_five_year <- ifelse(
-#  input_life_table$sex_age_cat == GBD_population_GL$pop_five_year, #GBD_population_GL$pop_five_year, NA)
+##Cross check population numbers
+sum(input_life_table$pop_five_year, na.rm = TRUE)
+#Sort data by sex and age to use in life table function. 
+input_life_table<-input_life_table[order(input_life_table$sex, input_life_table$age),]
 

### 2018-04-20 MSLT/MSLTdoc.Rmd
Modify age groups for cohort simulation

@@
 
-6) Create age categories index in GBDdata (upper limit age group to match injury data), total of 20 age groups. These are the age cohorts to simulate. 
+6) Create age categories index in GBDdata (mid age, to match cohort running), total of 20 age groups. These are the age cohorts to simulate. 
 ```{r, echo=TRUE}
-GBDdata$age_cat [GBDdata$age =="under 5"] <- 5
-GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 10
-GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 15
-GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 20
-GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 25
-GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 30
-GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 35
-GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 40
-GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 45
-GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 50
-GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 55
-GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 60
-GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 65
-GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 70
-GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 75
-GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 80
-GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 85
-GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 90
-GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 95
-GBDdata$age_cat [GBDdata$age =="95 plus"] <- 100
+# 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97
+GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
+GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
+GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
+GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
+GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
+GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
+GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
+GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
+GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
+GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
+GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
+GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
+GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
+GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
+GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
+GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
+GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
+GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
+GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
+GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97
 ```

### 2018-04-23 MSLT/MSLTdoc.Rmd
Generated general life tables

@@
     if (GBD_population$metric[i] == "number") {
-      GBD_population$pop_five_year[i] <- GBD_population$val[i] * 100000/ GBD_population$val[i + 2]}
-   else {GBD_population$pop_five_year[i] <- NA}
+      GBD_population$five_year_population[i] <- GBD_population$val[i] * 100000/ GBD_population$val[i + 2]}
+   else {GBD_population$five_year_population[i] <- NA}
  }
@@
 ```{r, echo=TRUE}
-GBD_population <- GBD_population[!is.na(GBD_population$pop_five_year),]
+GBD_population <- GBD_population[!is.na(GBD_population$five_year_population),]
 ```
@@
 ```{r, echo=TRUE}
-GBD_population <- filter(GBD_population) %>% select(sex_age_cat, sex, age_cat, pop_five_year, location)
+GBD_population <- filter(GBD_population) %>% select(sex_age_cat, sex, age_cat, five_year_population, location)
 ```
@@
 ```{r, echo=TRUE}
-GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, age_cat, sex, pop_five_year, location)
+GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, age_cat, sex, five_year_population, location)
 ```
@@
 ```{r, echo=TRUE}
-GreaterLondon <- sum(GBD_population_GL$pop_five_year)
+GreaterLondon <- sum(GBD_population_GL$five_year_population)
 ```
@@
 
-interpolated <- as.data.frame(interpolation_func(seq(0, 99, 1)))
-ages <- seq(0,99,by=1)
-interpolated <- cbind(interpolated, ages)
+interpolated <- as.data.frame(interpolation_func(seq(0, 100, 1)))
+age <- seq(0,100,by=1)
+interpolated <- cbind(interpolated, age)
 interpolated[,1] <- exp(interpolated[,1])
-colnames(interpolated)[1] <- "rate"
+colnames(interpolated)[1] <- paste(measure_index)
+# Add column with sex to create age_sex category to then merge with input_life table
+interpolated$sex <- paste(sex_index)
+interpolated$sex_age_cat <- paste(interpolated$sex, interpolated$age, sep = "_")
+# Change name of column death to mx and ylds to pyld_rate to then merge with input_life table
 
+if (colnames(interpolated)[1] == "deaths")
+    colnames(interpolated)[1] <- paste("mx")
+  else
+    colnames(interpolated)[1] <- paste("pyld_rate")
+
+# Name data frame
 assign(paste(sex_index, measure_index, "interpolated", sep = "_"), interpolated)
@@
 ##Do graph with another layer for original rates in age groups for comparison purposes.
-p <- ggplot(data = interpolated,mapping = aes(ages,rate)) +
+p <- ggplot(data = interpolated,mapping = aes(age, interpolated[,1])) +
               geom_line(aes(color = "Interpolated")) +
@@
             theme_classic() +
-            theme(plot.title = element_text(hjust = 0.5))
+            theme (plot.title = element_text(hjust = 0.5))
             print(p)
@@
 ```{r, echo=TRUE}
-##Create life table input data frame (population numbers, all-cause moratlity rates ##(per one) and all cause yld rates)
+##Create life table input data frame (population numbers, all-cause moratlity rates ##(per one), all cause yld rates) and 5-year age groups 
+## to run the cohorts (from population life table).
 
 ##Create empty data frame
-input_life_table <- as.data.frame(matrix(0, ncol=4, nrow = 202))
+input_life_table <- as.data.frame(matrix(0, ncol=2, nrow = 202))
 ##Give variables names (mx=total mortality rate, yldsx = total years lived with disability rates )
-names(input_life_table) <- c("age", "sex", "mx", "yldsx")
+names(input_life_table) <- c("age", "sex")
 ##Populate life table: Age, 0 to 100
@@
 
-##Populate life table: five_year_pop. SEE LIFE TABLE LOOPS THAT USES 5 YEAR POP.
+##Populate life table: five_year_pop.
 
 input_life_table <- merge(input_life_table, select(GBD_population_GL
-                                                     ,c(sex_age_cat, pop_five_year))
+                                                     ,c(sex_age_cat, five_year_population))
+                                                     , by = 'sex_age_cat', all = TRUE) 
+##Populate life table: mortality rates (from interpolated rates)
+## Males mortality rates. NEED MATCHING NAMES
+input_life_table <- merge(input_life_table, select(male_deaths_interpolated
+                                                     ,c(sex_age_cat, mx))
+                                                     , by = 'sex_age_cat', all.x = TRUE)
+input_life_table <- merge(input_life_table, select(female_deaths_interpolated
+                                                     ,c(sex_age_cat, mx))
+                                                     , by = 'sex_age_cat', all.x = TRUE)
+input_life_table <- merge(input_life_table, select(male_ylds_interpolated
+                                                     ,c(sex_age_cat, pyld_rate))
+                                                     , by = 'sex_age_cat', all.x = TRUE) 
+input_life_table <- merge(input_life_table, select(female_ylds_interpolated
+                                                     ,c(sex_age_cat, pyld_rate))
                                                      , by = 'sex_age_cat', all.x = TRUE)
 
+## I created mx.x and mx.y and pyld.x and pyld.y and I only need mx and pyld_rate in one colument. I wil do a quick fix here. 
+## DISCUSS WITH ALI TO IMPROVE
+
+input_life_table$mx <-  ifelse(input_life_table$sex == "male", input_life_table$mx.x, input_life_table$mx.y)
+input_life_table$pyld_rate <-  ifelse(input_life_table$sex == "male", input_life_table$pyld_rate.x, input_life_table$pyld_rate.y)
+
 ##Cross check population numbers
-sum(input_life_table$pop_five_year, na.rm = TRUE)
+sum(input_life_table$five_year_population, na.rm = TRUE)
 #Sort data by sex and age to use in life table function. 
@@
 ```
+16) Generate baseline life tables from input_data_frame. Life tables are generated for each age (5-years) and sex cohort. 
+It is assumed that current observed rates of mortality and disability are going to be observed in the future.
+
+``` {r, echo=TRUE}
+
+##Generate list of life tables. 
+## The function run_life_table was created to run the basline and scenario life tables (see functions.R script). 
+## The input_life_table data frame was generated to work with the function.
+
+## Generate looping variables (sex and age_group. Sex already generated in code for interpolation (i_sex))
+
+i_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
+
+general_life_table_list_bl <- list()
+
+index <- 1
+
+for (age in i_age_cohort){
+  for (sex in i_sex){
+    cat("age ", age, " and sex ", sex, "\n")
+    general_life_table_list_bl[[index]] <- run_life_table(in_idata = input_life_table, in_sex = sex, in_mid_age = age)
+    index <- index + 1
+  }
+}
+
+##### Uncommnet to check life table list
+View(general_life_table_list_bl[[2]])
+
+```
+
+17) Generate disease life tables. 
+Generate two sets for Greater London: 1) GBD data and 2) English survey data.
+
 Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.

### 2018-04-23 MSLT/MSLTdoc.Rmd
Preparing data for DISMOD

@@
 Data for YLDs prevalence is obtained from the online GBD IHME data tool (http://ghdx.healthdata.org/gbd-results-tool).
-Our calculations of DW in the example here based on the GBD methods for estimating YLDs as the sum of sequelae prevalence multiplied by sequelae disability weights (REF GBD).
-The GBD has publicly available data at the cause level (e.g.
-ischemic heart disease) instead of squelae level (e.g.
-myocardial infarction, angina and heart failure).
-However, the GBD disability weights are for health states associated with sequelae, hence, we need to calculate DWs.
 An age and sex specific-correction was introduced to counteract the effects of accumulating comorbid illnesses in the older age groups (Equation 1).
@@
 First obtain disease and population data from GBD.
-Second, use Dismod II to derive internally consistent estimates for incidence and case fatality (PMSLT disease life table iputs",
+Second, use Dismod II to derive internally consistent estimates for incidence and case fatality (PMSLT disease life table iputs)",
 "Disease life table", "Derive from disease prevalence and years lived with disability from GBD", "Adjustments for comorbidities in later years of life to be applied"
@@
 ```{r, echo=TRUE}
-GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", metric == "rate" | metric == "number" ) %>% select(metric, age_cat, sex, val, sex_age_cat, location)
+GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", 
+metric == "rate" | metric == "number" ) %>% select(metric, age_cat, sex, val, 
+sex_age_cat, location)
 ```
@@
     if (GBD_population$metric[i] == "number") {
-      GBD_population$five_year_population[i] <- GBD_population$val[i] * 100000/ GBD_population$val[i + 2]}
+      GBD_population$five_year_population[i] <- GBD_population$val[i] * 100000/
+      GBD_population$val[i + 2]}     
     else {GBD_population$five_year_population[i] <- NA}
@@
 ```{r, echo=TRUE}
-GBD_population <- filter(GBD_population) %>% select(sex_age_cat, sex, age_cat, five_year_population, location)
+GBD_population <- filter(GBD_population) %>% 
+select(sex_age_cat, sex, age_cat, five_year_population, location)
 ```
@@
 ```{r, echo=TRUE}
-GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% select(sex_age_cat, age_cat, sex, five_year_population, location)
+GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% 
+select(sex_age_cat, age_cat, sex, five_year_population, location)
 ```
@@
 ```{r, echo=TRUE}
-GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% select(measure, location, sex, age, metric, cause, val, age_cat)
+GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% 
+select(measure, location, sex, age, metric, cause, val, age_cat)
 GBDGL$one_rate <- GBDGL$val/100000
@@
 
-##Loop to generate interpolated rates for all cause mortality and ylds for males and females
+## Loop to generate interpolated rates for all cause mortality and ylds for males and females
 
@@
 
-data <- filter(GBDGL, measure == measure_index, sex == sex_index, cause == "all causes") %>% select(measure, location, sex, age, metric, cause, val, age_cat, one_rate) 
+data <- filter(GBDGL, measure == measure_index, sex == sex_index, 
+cause == "all causes") %>% select(measure, location, sex, age, metric, 
+cause, val, age_cat, one_rate) 
 assign(paste(sex_index, measure_index, "interpolated_data", sep = "_"), data)
@@
 colnames(interpolated)[1] <- paste(measure_index)
-# Add column with sex to create age_sex category to then merge with input_life table
+## Add column with sex to create age_sex category to then merge with input_life table
 interpolated$sex <- paste(sex_index)
 interpolated$sex_age_cat <- paste(interpolated$sex, interpolated$age, sep = "_")
-# Change name of column death to mx and ylds to pyld_rate to then merge with input_life table
+## Change name of column death to mx and ylds to pyld_rate to then merge 
+## with input_life table
 
@@
               mapping = aes(age_cat, one_rate, color = "Original")) +
-              labs(colour="",x="Age",y= paste(i_sex, i_measure, "rates", sep = " "))+
-            labs (title = paste("Rates", i_sex, i_measure, sep = " "), size=14) +
+              labs(colour="",x="Age",y="Rates", sep = " ") +
+              labs (title = paste("Rates", sex_index, "all cause", 
+              measure_index, sep = " "), size=14) +
               theme_classic() +
@@
 ```{r, echo=TRUE}
-##Create life table input data frame (population numbers, all-cause moratlity rates ##(per one), all cause yld rates) and 5-year age groups 
-## to run the cohorts (from population life table).
 
-##Create empty data frame
+## Create empty data frame
 input_life_table <- as.data.frame(matrix(0, ncol=2, nrow = 202))
-##Give variables names (mx=total mortality rate, yldsx = total years lived with disability rates )
+## Give variables names (mx=total mortality rate, 
+## yldsx = total years lived with disability rates )
 names(input_life_table) <- c("age", "sex")
-##Populate life table: Age, 0 to 100
+## Populate life table: Age, 0 to 100
 input_life_table[1:101, 1] <- c(0:100)
 input_life_table[102:202, 1] <- c(0:100)
-##Populate life table: sex, male, female
+## Populate life table: sex, male, female
 input_life_table[1:101, 2] <- "male"
 input_life_table[102:202, 2] <- "female"
-##Create variable age_sex to match with population data. 
-input_life_table$sex_age_cat <- paste(input_life_table$sex,input_life_table$age, sep = "_"  )
+## Create variable age_sex to match with population data. 
+input_life_table$sex_age_cat <- paste(input_life_table$sex,input_life_table$age,
+                                sep = "_"  )
 
@@
 
-## I created mx.x and mx.y and pyld.x and pyld.y and I only need mx and pyld_rate in one colument. I wil do a quick fix here. 
 ## DISCUSS WITH ALI TO IMPROVE
 
-input_life_table$mx <-  ifelse(input_life_table$sex == "male", input_life_table$mx.x, input_life_table$mx.y)
-input_life_table$pyld_rate <-  ifelse(input_life_table$sex == "male", input_life_table$pyld_rate.x, input_life_table$pyld_rate.y)
-
-##Cross check population numbers
+input_life_table$mx <-  ifelse(input_life_table$sex == "male",
+input_life_table$mx.x, input_life_table$mx.y)
+input_life_table$pyld_rate <-  ifelse(input_life_table$sex == "male",
+input_life_table$pyld_rate.x, input_life_table$pyld_rate.y)
+## Cross check population numbers
 sum(input_life_table$five_year_population, na.rm = TRUE)
-#Sort data by sex and age to use in life table function. 
+## Sort data by sex and age to use in life table function. 
 input_life_table<-input_life_table[order(input_life_table$sex, input_life_table$age),]
@@
 ```
-16) Generate baseline life tables from input_data_frame. Life tables are generated for each age (5-years) and sex cohort. 
+16) Generate **baseline life tables** from input_data_frame. Life tables are generated for each age (5-years) and sex cohort. 
 It is assumed that current observed rates of mortality and disability are going to be observed in the future.
 
-``` {r, echo=TRUE}
+```{r, echo=TRUE}
 
 ##Generate list of life tables. 
-## The function run_life_table was created to run the basline and scenario life tables (see functions.R script). 
+## The function run_life_table was created to run the basline and scenario life tables 
+## (see functions.R script). 
 ## The input_life_table data frame was generated to work with the function.
 
-## Generate looping variables (sex and age_group. Sex already generated in code for interpolation (i_sex))
+## Generate looping variables (sex and age_group. Sex already generated 
+## in code for interpolation (i_sex))
 
@@
     cat("age ", age, " and sex ", sex, "\n")
-    general_life_table_list_bl[[index]] <- run_life_table(in_idata = input_life_table, in_sex = sex, in_mid_age = age)
+    general_life_table_list_bl[[index]] <- run_life_table(in_idata = input_life_table,
+    in_sex = sex, in_mid_age = age)
     index <- index + 1
@@
 
-##### Uncommnet to check life table list
-View(general_life_table_list_bl[[2]])
+## Uncommnet to check life table list
+# View(general_life_table_list_bl[[2]])
 
@@
 
-17) Generate disease life tables. 
-Generate two sets for Greater London: 1) GBD data and 2) English survey data.
+17) Generate **baseline disease life tables**.
+Inputs of the disease life tables are: incidence,   case fatality (Figure 3, purple shaded area) and disability weights.
+Internally consistent estimates of incidence and case fatality were derived from GBD data (by five-year age groups and sex) and Dismod II.
+Disability weights were derived from GBD disease specific YLDs and prevalence (by five-year age groups and sex) adjusted for all-cause YLDs (see section 2.1.1.4).
+Note that Dismod II is an external software.
+The code below is used to generate the inputs for Dismod II using GBD data (see Figure 3, purple shaded area).
+Dismod II requires data for the *collection* (population numbers and all cause mortality rates) and *data set* (disease specific paramenters).
+
+```{r, echo=TRUE}
+## Collection input for Greater London: population numbers (sex and 5-year age groups). 
+## Already processed in past code, here, we sort it out and save in cvs format in a Dismod II folder.
+
+## Get population numbers
+
+collection_dismodii <- GBD_population_GL[order(GBD_population_GL$sex, GBD_population_GL$age_cat),]
+
+## Add mortality rates (these were generated for the life table, so we use the same code)
+
+collection_dismodii  <- merge(collection_dismodii, select(male_deaths_interpolated
+                                                     ,c(sex_age_cat, mx))
+                                                     , by = 'sex_age_cat', all.x = TRUE)
+collection_dismodii  <- merge(collection_dismodii, select(female_deaths_interpolated
+                                                     ,c(sex_age_cat, mx))
+                                                     , by = 'sex_age_cat', all.x = TRUE)
+
+collection_dismodii$mx <-  ifelse(collection_dismodii$sex == "male",
+collection_dismodii$mx.x, collection_dismodii$mx.y)
+
+## Save as CVS (Dismod II format)
+
+```
 

### 2018-04-23 MSLT/MSLTdoc.Rmd
Disease life table data input

@@
 
-15) The following code brings together all the required data to create the inputs life table (Figure 3)
+15) The following code brings together all the required data to create the inputs life table (Figure 3).
+ADD DIDMOD GENERATED DATA ONCE THESE ARE READY TO FOLLOW THE FORMAT OF THE IDATA USED IN THE FUNCTIONS.
 
@@
 input_life_table$pyld_rate.x, input_life_table$pyld_rate.y)
+## Drop redundant variables
+input_life_table <- subset(input_life_table, select = -c(mx.y, mx.x, pyld_rate.x, pyld_rate.y))
+
 ## Cross check population numbers
@@
 Inputs of the disease life tables are: incidence,   case fatality (Figure 3, purple shaded area) and disability weights.
-Internally consistent estimates of incidence and case fatality were derived from GBD data (by five-year age groups and sex) and Dismod II.
-Disability weights were derived from GBD disease specific YLDs and prevalence (by five-year age groups and sex) adjusted for all-cause YLDs (see section 2.1.1.4).
+Internally consistent estimates of incidence and case fatality are derived from GBD data (by five-year age groups and sex) and Dismod II.
+Disability weights are derived from GBD disease specific YLDs and prevalence (by five-year age groups and sex) adjusted for all-cause YLDs (see section 2.1.1.4).
 Note that Dismod II is an external software.
@@
 Dismod II requires data for the *collection* (population numbers and all cause mortality rates) and *data set* (disease specific paramenters).
+We generate a data set with all inputs for dismod by 5 year-age (represente at mid age) groups and sex.
 
@@
 
-collection_dismodii <- GBD_population_GL[order(GBD_population_GL$sex, GBD_population_GL$age_cat),]
+data_dismodii <- GBD_population_GL[order(GBD_population_GL$sex, GBD_population_GL$age_cat),]
 
@@
 
-collection_dismodii  <- merge(collection_dismodii, select(male_deaths_interpolated
+data_dismodii  <- merge(data_dismodii, select(male_deaths_interpolated
                                                      ,c(sex_age_cat, mx))
                                                      , by = 'sex_age_cat', all.x = TRUE)
-collection_dismodii  <- merge(collection_dismodii, select(female_deaths_interpolated
+data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
                                                      ,c(sex_age_cat, mx))
@@
 
-collection_dismodii$mx <-  ifelse(collection_dismodii$sex == "male",
-collection_dismodii$mx.x, collection_dismodii$mx.y)
+data_dismodii$mx <-  ifelse(data_dismodii$sex == "male",
+data_dismodii$mx.x, data_dismodii$mx.y)
+## Drop redundant variables 
+data_dismodii <- subset(data_dismodii, select = -c(mx.y, mx.x))
+##Sort data
+data_dismodii<-data_dismodii[order(data_dismodii$sex, data_dismodii$age_cat),] 
+## Add disease rates per one(incidence, prevalence and mortality)
 
-## Save as CVS (Dismod II format)
+## Generate variable to do a loop that creates dataframes for each of the disease inputs.
+
+i_disease <- c("ischemic heart disease", "ischemic stroke", "diabetes mellitus", "tracheal, bronchus, and lung cancer", "breast cancer", "colon and rectum cancer")
+i_sex <- c("male", "female")
+i_measure <- c("deaths", "ylds", "incidence", "prevalence")
+
+for(sex_index in i_sex) {
+  for (measure_index in i_measure) {
+    for (disease_index in i_disease) {
+
+data <- filter(GBDGL, measure == measure_index, sex == sex_index, 
+cause == disease_index) %>% select(measure, location, sex, age, metric, 
+cause, val, age_cat, one_rate) 
+assign(paste(sex_index, measure_index, disease_index, sep = "_"), data)
+## Give one rate a name to match merge in data_dismodii input
+
+    }
+  }
+}
+
+## Generate empy variables in data_dismodii for death, incidence and prevalence rates per disease
+
+ for (measure_index in i_measure) {
+    for (disease_index in i_disease) {
+
+  data_dismodii[,paste(measure_index, disease_index, sep =  "_")] <- 0
+
+  }
+ }
 

### 2018-04-23 MSLT/MSLTdoc.Rmd
Disease life table inputs

@@
 GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% 
-select(sex_age_cat, age_cat, sex, five_year_population, location)
+select(sex_age_cat, age_cat, sex, five_year_population, location, sex_age_cat)
 ```
@@
 GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% 
-select(measure, location, sex, age, metric, cause, val, age_cat)
+select(measure, location, sex, age, metric, cause, val, age_cat, sex_age_cat)
 GBDGL$one_rate <- GBDGL$val/100000
@@
 cause == disease_index) %>% select(measure, location, sex, age, metric, 
-cause, val, age_cat, one_rate) 
+cause, val, age_cat, one_rate, sex_age_cat) 
+colnames(data)[9] <- paste(measure_index, disease_index, sep =  "_")
 assign(paste(sex_index, measure_index, disease_index, sep = "_"), data)
 ## Give one rate a name to match merge in data_dismodii input
+    }
+  }
+}
+
+## Generate empy variables in data_dismodii for death, incidence and prevalence rates per disease (MAY NOT NEED TO USE, WITH MERGING VARIABLES)
+
+ for (measure_index in i_measure) {
+    for (disease_index in i_disease) {
+
+  # data_dismodii[,paste(measure_index, disease_index, sep =  "_")] <- 0
+data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
+                                                     ,c(sex_age_cat, mx))
+                                                     , by = 'sex_age_cat', all.x = TRUE)
+  }
+ }
+
+for (measure_index in i_measure) {
+    for (disease_index in i_disease) {
+      for (sex_index in i_sex){
+
+data_dismodii  <- merge(data_dismodii, select(paste(sex_index, measure_index, disease_index, sep = "_"),c(sex_age_cat, paste(measure_index, disease_index, sep = "_")), by = 'sex_age_cat', all.x = TRUE))
 
@@
 
-## Generate empy variables in data_dismodii for death, incidence and prevalence rates per disease
-
- for (measure_index in i_measure) {
-    for (disease_index in i_disease) {
-
-  data_dismodii[,paste(measure_index, disease_index, sep =  "_")] <- 0
-
-  }
- }
+# data_dismodii  <- merge(data_dismodii, select(male_deaths_interpolated
+#                                                      ,c(sex_age_cat, mx))
+#                                                      , by = 'sex_age_cat', all.x = TRUE)
 

### 2018-04-23 MSLT/MSLTdoc.Rmd
Data inputs dismod
Incomplete

@@
 
+####THIS CODE IS INCOMPLETE, I WILL PREPARE THE DATA FOR DISMOD II IN EXCEL AND AFTER PRESENTATION THURSDAY 26 APRIL I WILL DEVELOP THE SCRIPT.
+
 ## Get population numbers
@@
 
+## Loop to generate dataframes
 for(sex_index in i_sex) {
@@
 assign(paste(sex_index, measure_index, disease_index, sep = "_"), data)
-## Give one rate a name to match merge in data_dismodii input
+
+## The code below is not generating the multiple datasets saved as csv.
+
+# data_name <- file.path(tempdir(), paste0(sex_index, measure_index, disease_index, ".csv"))
+# write.csv(data, file = data_name)
     }
@@
 
+## Generate TO DO, ADD disease parameters.
+
 ## Generate empy variables in data_dismodii for death, incidence and prevalence rates per disease (MAY NOT NEED TO USE, WITH MERGING VARIABLES)
 
- for (measure_index in i_measure) {
-    for (disease_index in i_disease) {
-
-  # data_dismodii[,paste(measure_index, disease_index, sep =  "_")] <- 0
-data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
-                                                     ,c(sex_age_cat, mx))
-                                                     , by = 'sex_age_cat', all.x = TRUE)
-  }
- }
-
-for (measure_index in i_measure) {
-    for (disease_index in i_disease) {
-      for (sex_index in i_sex){
-
-data_dismodii  <- merge(data_dismodii, select(paste(sex_index, measure_index, disease_index, sep = "_"),c(sex_age_cat, paste(measure_index, disease_index, sep = "_")), by = 'sex_age_cat', all.x = TRUE))
-
-    }
-  }
-}
-
-# data_dismodii  <- merge(data_dismodii, select(male_deaths_interpolated
+#  for (measure_index in i_measure) {
+#     for (disease_index in i_disease) {
+# 
+#   # data_dismodii[,paste(measure_index, disease_index, sep =  "_")] <- 0
+# data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
 #                                                      ,c(sex_age_cat, mx))
 #                                                      , by = 'sex_age_cat', all.x = TRUE)
+#   }
+#  }
 

### 2018-04-24 MSLT/MSLTdoc.Rmd
Greater London example MSLT
To do: adjust dw with dismod, add loop to plot function and inspect outcomes.

@@
 ```{r, echo=TRUE}
-GBDdata <- read.csv("data/UK/englandandgreaterlondon.csv", stringsAsFactors = F)
+GBDdata <- read.csv("data/UK/gbd2016.csv", stringsAsFactors = F)
 ```
@@
 ## Loop to generate interpolated rates for all cause mortality and ylds for males and females
+## UPDATE WITH YLDS RATES ALL CAUSES ADJUSTED FOR ALL OTHER MODELLED DISEASES.
 
@@
 Dismod II requires data for the *collection* (population numbers and all cause mortality rates) and *data set* (disease specific paramenters).
-We generate a data set with all inputs for dismod by 5 year-age (represente at mid age) groups and sex.
+We generate a data set with all inputs for dismod by 5 year-age (represente at mid age) groups and sex for five diseases (ischemic heart disease, ischemic stroke, diabetes mellitus, colon and rectum cancer and breast cancer-women).
+WE WILL NEED ADJUSTMENT FOR TO MATCH RRS TO DISEASES (COLON AND RECTUM CANCER AND DIABETES MELLITUS)
 
@@
 
-## Get population numbers
-
-data_dismodii <- GBD_population_GL[order(GBD_population_GL$sex, GBD_population_GL$age_cat),]
-
-## Add mortality rates (these were generated for the life table, so we use the same code)
-
-data_dismodii  <- merge(data_dismodii, select(male_deaths_interpolated
-                                                     ,c(sex_age_cat, mx))
-                                                     , by = 'sex_age_cat', all.x = TRUE)
-data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
-                                                     ,c(sex_age_cat, mx))
-                                                     , by = 'sex_age_cat', all.x = TRUE)
-
-data_dismodii$mx <-  ifelse(data_dismodii$sex == "male",
-data_dismodii$mx.x, data_dismodii$mx.y)
-## Drop redundant variables 
-data_dismodii <- subset(data_dismodii, select = -c(mx.y, mx.x))
-##Sort data
-data_dismodii<-data_dismodii[order(data_dismodii$sex, data_dismodii$age_cat),] 
-## Add disease rates per one(incidence, prevalence and mortality)
-
-## Generate variable to do a loop that creates dataframes for each of the disease inputs.
-
-i_disease <- c("ischemic heart disease", "ischemic stroke", "diabetes mellitus", "tracheal, bronchus, and lung cancer", "breast cancer", "colon and rectum cancer")
-i_sex <- c("male", "female")
-i_measure <- c("deaths", "ylds", "incidence", "prevalence")
-
-## Loop to generate dataframes
-for(sex_index in i_sex) {
-  for (measure_index in i_measure) {
-    for (disease_index in i_disease) {
-
-data <- filter(GBDGL, measure == measure_index, sex == sex_index, 
-cause == disease_index) %>% select(measure, location, sex, age, metric, 
-cause, val, age_cat, one_rate, sex_age_cat) 
-colnames(data)[9] <- paste(measure_index, disease_index, sep =  "_")
-assign(paste(sex_index, measure_index, disease_index, sep = "_"), data)
+# ## Get population numbers
+# 
+# data_dismodii <- GBD_population_GL[order(GBD_population_GL$sex, GBD_population_GL$age_cat),] 
+# 
+# ## Add mortality rates (these were generated for the life table, so we use the same code)
+# 
+# data_dismodii  <- merge(data_dismodii, select(male_deaths_interpolated
+#                                                      ,c(sex_age_cat, mx))
+#                                                      , by = 'sex_age_cat', all.x = TRUE)
+# data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
+#                                                      ,c(sex_age_cat, mx))
+#                                                      , by = 'sex_age_cat', all.x = TRUE)
+# 
+# data_dismodii$mx <-  ifelse(data_dismodii$sex == "male",
+# data_dismodii$mx.x, data_dismodii$mx.y)
+# ## Drop redundant variables 
+# data_dismodii <- subset(data_dismodii, select = -c(mx.y, mx.x))
+# ##Sort data
+# data_dismodii<-data_dismodii[order(data_dismodii$sex, data_dismodii$age_cat),] 
+# ## Add disease rates per one(incidence, prevalence and mortality)
+# 
+# ## Generate variable to do a loop that creates dataframes for each of the disease inputs.
+# 
+# i_disease <- c("ischemic heart disease", "ischemic stroke", "diabetes mellitus", "tracheal, bronchus, and lung cancer", "breast cancer", "colon and rectum cancer")
+# i_sex <- c("male", "female")
+# i_measure <- c("deaths", "ylds", "incidence", "prevalence")
+# 
+# ## Loop to generate dataframes
+# for(sex_index in i_sex) {
+#   for (measure_index in i_measure) {
+#     for (disease_index in i_disease) {
+#    
+# data <- filter(GBDGL, measure == measure_index, sex == sex_index, 
+# cause == disease_index) %>% select(measure, location, sex, age, metric, 
+# cause, val, age_cat, one_rate, sex_age_cat) 
+# colnames(data)[9] <- paste(measure_index, disease_index, sep =  "_")
+# assign(paste(sex_index, measure_index, disease_index, sep = "_"), data)
 
@@
 # write.csv(data, file = data_name)
-    }
-  }
-}
+#     }
+#   }
+# }
 
@@
 
-Health outcomes included in leisure time meta-analysis: all-cause mortality, cardiovascular disease, stroke, coronary heart disease, total cancer, colon, lung, endometrial, and breast cancer.
+```{r, echo=true}
+## Use externaly generated inputs for for disease life table.
+The previous code will serve to to some of the data preparation for DISMOD.
+Also need to devlop code to calculated DWs, for now, all generated in excel.
 
-####Model parameters
+idata <- read.csv("data/UK/idata.csv", stringsAsFactors = F)
+## Use run_disease
 
+i_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
+i_sex <- c("male", "female")
+i_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
+
+disease_life_table_list_bl <- list()
+index <- 1
+
+for (age in i_age_cohort){
+  for (sex in i_sex){
+    for (disease in i_disease) {
+      # Exclude breast_cancer for Males
+      if (sex == "male" && disease == "breast_cancer"){
+        cat("\n")
+      }
+      else {
+        cat("age ", age, " sex ", sex, "and disease", disease, "\n")
+        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, in_disease = disease)
+        index <- index + 1
+      }
+    }
+  }
+}
+##### Uncommnet to check disease life table list
+# View(disease_life_table_list_bl[[8]])
+
+```
+
+18) Generate mock change in incidence of disease to generate sceanrio life tables. In the final version, this will come from the calculated PIFs by disease. 
+PIFs are applied here to incidence, however, may also be applied to case fatality, depending on the RRs.
+Also, we can include time delays from change in exposure to change in health outcomes via the PIF.
+
+```{r, echo=TRUE}
+## Create value to use as factor changing incidence rates.
+
+incidence_change <- 0.95
+
+#####Generate scenario incidence (for each disease)
+
+incidence_sc <- list()
+index <- 1
+
+for (age in i_age_cohort){
+  for (sex in i_sex){
+    for (disease in i_disease) {
+
+      # Exclude breast_cancer for Males
+      if (sex == "male" && disease == "breast_cancer"){
+        cat("\n")
+      }
+      else {
+
+        incidence_sc[[index]] <- disease_life_table_list_bl[[index]]$incidence_disease * incidence_change
+        index <- index + 1
+
+      }
+    }
+  }
+}
+
+##### Uncommnet to check scenario incidence
+# View(incidence_sc[[1]])
+```
+
+19) Use scenario incidence to calculate scenario life tables.
+
+```{r, echo=TRUE}
+
+disease_life_table_list_sc <- list()
+index <- 1
+
+for (age in i_age_cohort){
+  for (sex in i_sex){
+    for (disease in i_disease) {
+      # Exclude breast_cancer for Males
+      if (sex == "male" && disease == "breast_cancer"){
+        cat("\n")
+      }
+      else {
+
+        cat("age ", age, " sex ", sex, "and disease", disease, "\n")
+        # modify idata's incidence for the said scenario
+        td1 <- idata
+        td1[td1$age >= age & td1$sex == sex,][[paste("incidence", disease, sep = "_")]] <- incidence_sc[[index]]
+
+        # Instead of idata, feed td to run scenarios
+        disease_life_table_list_sc[[index]] <- run_disease(in_idata = td1, in_sex = sex, in_mid_age = age, in_disease = disease)
+        disease_life_table_list_sc[[index]]$diff_inc_disease <- disease_life_table_list_sc[[index]]$incidence_disease - disease_life_table_list_bl[[index]]$incidence_disease
+        disease_life_table_list_sc[[index]]$diff_prev_disease <- disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px
+        disease_life_table_list_sc[[index]]$diff_mort_disease <- disease_life_table_list_sc[[index]]$mx - disease_life_table_list_bl[[index]]$mx
+        disease_life_table_list_sc[[index]]$diff_pylds_disease <- (disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px) * disease_life_table_list_bl[[index]]$dw_disease
+
+        index <- index + 1
+      }
+    }
+  }
+}
+##### Uncommnet to check scenario life tables
+# View(disease_life_table_list_sc[[1]])
+
+```
+
+20) Calculate life tables scearnio.
+
+```{r, echo=TRUE}
+#####Generate total change in mortality rate
+##### Sum mortality rate scenarios (mx_sc_total)
+
+mx_sc_total <- list()
+l_index <- 1
+index <- 1
+for (age in i_age_cohort){
+  for (sex in i_sex){
+    mortality_sum <- NULL
+    create_new <- T
+
+    for (disease in i_disease) {
+      if (sex == "male" && disease == "breast_cancer"){
+        cat("\n")
+      }else{
+
+        if (create_new){
+          mortality_sum <- select(disease_life_table_list_sc[[index]], c('age', 'sex'))
+          mortality_sum$total <- 0
+          create_new <- F
+          mortality_sum$total <- mortality_sum$total + (disease_life_table_list_sc[[index]]$diff_mort_disease)
+        }else{
+          mortality_sum$total <- mortality_sum$total + (disease_life_table_list_sc[[index]]$diff_mort_disease)
+        }
+
+        cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
+        index <- index + 1
+      }
+
+    }
+    mx_sc_total[[l_index]] <- mortality_sum 
+    l_index <- l_index + 1
+  }
+}
+
+##### Uncommnet to check sceanrio mortality and changes 
+# View(mx_sc_total[[1]])
+
+#####Generate total change in prevalent yld rates
+#####total ylds rate= sum (change prevalence disease*dw)
+
+pylds_sc_total <- list()
+l_index <- 1
+index <- 1
+for (age in i_age_cohort){
+  for (sex in i_sex){
+    pylds_sum <- NULL
+    create_new <- T
+
+    for (disease in i_disease) {
+      if (sex == "male" && disease == "breast_cancer"){
+        cat("\n")
+      }else{
+
+        if (create_new){
+          pylds_sum <- select(disease_life_table_list_sc[[index]], c('age', 'sex'))
+          pylds_sum$total <- 0
+          create_new <- F
+          pylds_sum$total <- pylds_sum$total + (disease_life_table_list_sc[[index]]$diff_pylds_disease)
+        }else{
+          pylds_sum$total <- pylds_sum$total + (disease_life_table_list_sc[[index]]$diff_pylds_disease)
+        }
+
+        cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
+        index <- index + 1
+      }
+
+    }
+    pylds_sc_total[[l_index]] <- pylds_sum
+    l_index <- l_index + 1
+  }
+}
+
+##### Uncommnet to check scenario pyld change
+# View(pylds_sc_total[[2]])
+
+##### Calculate general life tables with modified mortality and pylds total#############################
+###Original mortality rate is modified by the mx_sc_total (total change in mortality from diseases)
+###Original pyld rate is modified by the change in each disease pylds
+
+general_life_table_list_sc <- list()
+index <- 1
+
+for (age in i_age_cohort){
+  for (sex in i_sex){
+
+    cat("age ", age, " and sex ", sex, "\n")
+    # modify idata's mortality and pyld total for the said scenario
+    td2 <- input_life_table
+    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx + mx_sc_total[[index]]$total
+    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate + pylds_sc_total[[index]]$total
+
+    # Instead of idata, feed td to run scenarios
+    general_life_table_list_sc[[index]] <- run_life_table(in_idata = td2, in_sex = sex, in_mid_age = age)
+
+    index <- index + 1
+  }
+}
+##### Uncommnet to check scenario life tables
+# View(general_life_table_list_sc[[1]])
+# View(general_life_table_list_bl[[2]])
+```
+
+21) Generate list of outputs by age and sex
+
+```{r, echo=TRUE}
+#####In the following list "output_life_table", 32 data frames are nested per age and sex cohort
+
+output_burden <- list()
+l_index <- 1
+index <- 1
+for (age in i_age_cohort){
+  for (sex in i_sex){
+
+    #Males do not have breast cancer, that is why we need the if/else. 
+    #We create a TRUE/FALSE variable for the loop to move into the next disease
+
+    create_new <- T
+    for (disease in i_disease) {
+      if (sex == "male" && disease == "breast_cancer"){
+        cat("\n")
+      }else{
+
+        if (create_new){
+          output_burden_sc <- select(disease_life_table_list_sc[[index]], c('age', 'sex', 'incidence_disease', 'mx', 'px'))
+          names(output_burden_sc)[names(output_burden_sc) == 'incidence_disease'] <- paste('incidence_disease', disease, "sc", sep = "_")
+          names(output_burden_sc)[names(output_burden_sc) == 'mx'] <- paste('mx', disease, "sc", sep = "_")
+          names(output_burden_sc)[names(output_burden_sc) == 'px'] <- paste('px', disease, "sc", sep = "_")
+
+          output_burden_bl <- select(disease_life_table_list_bl[[index]], c('incidence_disease', 'mx', 'px'))
+          names(output_burden_bl)[names(output_burden_bl) == 'incidence_disease'] <- paste('incidence_disease', disease, "bl", sep = "_")
+          names(output_burden_bl)[names(output_burden_bl) == 'mx'] <- paste('mx', disease, "bl", sep = "_")
+          names(output_burden_bl)[names(output_burden_bl) == 'px'] <- paste('px', disease, "bl", sep = "_")
+
+          ####New list to add calculations for changes in burden of disease (incidence and mortality numbers)
+
+          output_burden_change <- list()
+
+          output_burden_change$inc_num_bl <- disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) *       general_life_table_list_bl[[l_index]]$Lx
+          output_burden_change$inc_num_sc <- disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx
+          output_burden_change$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx)
+
+          output_burden_change$mx_num_bl <- disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx
+          output_burden_change$mx_num_sc <- disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx
+          output_burden_change$mx_num_diff <- (disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx)
+
+          names(output_burden_change)[names(output_burden_change) == 'inc_num_bl'] <- paste('inc_num_bl', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'inc_num_sc'] <- paste('inc_num_sc', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'inc_num_diff'] <- paste('inc_num_diff', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'mx_num_bl'] <- paste('mx_num_bl', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'mx_num_sc'] <- paste('mx_num_sc', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'mx_num_diff'] <- paste('mx_num_diff', disease, sep = "_")
+
+          ###Bind all lists
+
+          output_burden_sc <- cbind(output_burden_sc, output_burden_bl)
+          output_burden_sc <- cbind(output_burden_sc, output_burden_change)
+
+          create_new <- F
+
+          #Here the calculations above are repeated, here is where the F is telling to move into the next disease
+
+        }else{
+
+          td3 <- select(disease_life_table_list_sc[[index]], c('incidence_disease', 'mx', 'px'))
+          names(td3)[names(td3) == 'incidence_disease'] <- paste('incidence_disease', disease, "sc", sep = "_")
+          names(td3)[names(td3) == 'mx'] <- paste('mx', disease, "sc", sep = "_")
+          names(td3)[names(td3) == 'px'] <- paste('px', disease, "sc", sep = "_")
+
+          td4 <- select(disease_life_table_list_bl[[index]], c('incidence_disease', 'mx', 'px'))
+          names(td4)[names(td4) == 'incidence_disease'] <- paste('incidence_disease', disease, "bl", sep = "_")
+          names(td4)[names(td4) == 'mx'] <- paste('mx', disease, "bl", sep = "_")
+          names(td4)[names(td4) == 'px'] <- paste('px', disease, "bl", sep = "_")
+
+          output_burden_change2 <- list()
+
+          output_burden_change2$inc_num_bl <- disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx
+          output_burden_change2$inc_num_sc <- disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx
+          output_burden_change2$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx)
+
+          output_burden_change2$mx_num_bl <- disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx
+          output_burden_change2$mx_num_sc <- disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx
+          output_burden_change2$mx_num_diff <- (disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx)
+
+          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_bl'] <- paste('inc_num_bl', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_sc'] <- paste('inc_num_sc', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_diff'] <- paste('inc_num_diff', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_bl'] <- paste('mx_num_bl', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_sc'] <- paste('mx_num_sc', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_diff'] <- paste('mx_num_diff', disease, sep = "_")
+
+          ###Bind all lists
+
+          output_burden_sc <- cbind(output_burden_sc, td3)
+          output_burden_sc <- cbind(output_burden_sc, td4)
+          output_burden_sc$age_cohort <- age 
+          output_burden_sc <- cbind(output_burden_sc, output_burden_change2)
+
+        }
+
+        cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
+        index <- index + 1
+      }
+
+    }
+
+    # general_life_table_list_sc and general_life_table_list_bl (Lx)
+    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], c('Lx', 'Lwx'))
+    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- paste('Lx', "sc", sep = "_")
+    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- paste('Lwx', "sc", sep = "_")
+
+    output_burden_lf_bl <- select(general_life_table_list_bl[[l_index]], c('Lx', 'Lwx'))
+    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lx'] <- paste('Lx', "bl", sep = "_")
+    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lwx'] <- paste('Lwx', "bl", sep = "_")
+
+    output_burden_lf_sc$Lx_diff <- general_life_table_list_bl[[l_index]]$Lx - general_life_table_list_sc[[l_index]]$Lx
+    output_burden_lf_sc$Lwx_diff <- general_life_table_list_bl[[l_index]]$Lwx - general_life_table_list_sc[[l_index]]$Lwx
+
+    output_burden_sc <- cbind(output_burden_sc, output_burden_lf_sc)
+    output_burden_sc <- cbind(output_burden_sc, output_burden_lf_bl)
+
+    output_burden[[l_index]] <- output_burden_sc
+    l_index <- l_index + 1
+
+  }
+}
+
+#Uncomment to check
+# View(output_burden[[32]])
+
+```
+
+22) Combine all lists of outputs in a data frame to facilitate extraction of outcomes of interest and plotting.
+
+```{r, echo=TRUE}
+#####Generate a data frame for all results and create function to get outcomes.
+
+output_df <- plyr::ldply(output_burden, rbind)
+
+#Remove variables that are not used in the generation of outputs. CHANGE THIS NAMES, TOO LONG
+
+output_df <- subset(output_df, select = -c(incidence_disease_ihd_bl, incidence_disease_ihd_sc, 
+                                           incidence_disease_istroke_bl, incidence_disease_istroke_sc, 
+                                           incidence_disease_diabetes_bl, incidence_disease_diabetes_sc, 
+                                           incidence_disease_breast_cancer_bl, incidence_disease_breast_cancer_sc, 
+                                           incidence_disease_colon_cancer_bl, incidence_disease_colon_cancer_sc, 
+                                           mx_ihd_bl, mx_ihd_sc, mx_istroke_bl, mx_istroke_sc, mx_diabetes_bl, mx_diabetes_sc, 
+                                           mx_breast_cancer_bl, mx_breast_cancer_sc, mx_colon_cancer_bl, mx_colon_cancer_sc, 
+                                           px_ihd_bl, px_ihd_sc, px_istroke_bl, px_istroke_sc, px_diabetes_bl, px_diabetes_sc, 
+                                           px_breast_cancer_bl, px_breast_cancer_sc, px_colon_cancer_bl, px_colon_cancer_sc))
+
+```
+
+23) Plot outcomes for each age and sex cohort using created function. ADD A LOOP TO GENERATE ALL OUTCOMES FOR IN_AGE, IN_POPULATION, IN_OUTCOMES AND IN_LEGEND. FIX THE COLORS IN THE FUNCTION.
+
+```{r, echo=true}
+plot_eg <- plot_output(in_data = output_df, in_age = 22, in_population = "male", in_outcomes = c("age", "mx_num_bl_ihd", "mx_num_sc_ihd", "mx_num_diff_ihd"), in_legend = "Ischemic Heart Disease Deaths")
+```
+
+24) Add up all outcomes per year of simulation with generated function.
+
+```{r, echo=true}
+####Generate data frame with all outputs to graph total change in burden by simulation year. 
+###first, need to run function for males and females separetly, in_cohorts indicates the number of age cohorts to 
+####include. To show all select 16. what is specified in in_outcomes will be graphed or can also be presented as a total. 
+aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('inc_num_bl_ihd','inc_num_sc_ihd', "inc_num_diff_ihd"))
+
+aggregate_frame_females <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c('inc_num_bl_ihd','inc_num_sc_ihd', "inc_num_diff_ihd"))
+
+#####The following adds up both males and females
+# Remove non-numeric columns starting with age and sex
+aggregate_frame_males <- aggregate_frame_males %>% select(-starts_with("age"), -starts_with("sex"))
+
+aggregate_frame_females <- aggregate_frame_females %>% select(-starts_with("age"), -starts_with("sex"))
+
+# Create a copy of aggregate_frame_females
+total_aggr <- aggregate_frame_females
+# Add aggregate_frame_males values to it
+for (i in 1:ncol(aggregate_frame_females)){
+  total_aggr[i] <- total_aggr[i] + aggregate_frame_males[i]
+}
+total_aggr$sim_year <- seq.int(nrow(total_aggr))
+```
+
+25) Plots using total aggregated change in burden.
+
+```{r, echo=true}
+
+####This plot has to be customised to in_outcomes, here, only totals shown, but specifications are up to the user
+
+####[] is used here to indicate the number of simulation years into the future. 
+####Disease outcomes has to be changed to the outcome of interest
+
+disease_outcome <- "Deaths ischemic heart disease"
+
+total_plot <- ggplot(total_aggr[1:20,], aes(x = sim_year)) +
+
+  geom_line(mapping = aes(y = total_inc_num_bl_ihd, colour = "total_mx_num_bl_ihd")) +
+  theme_classic() +
+  geom_hline(yintercept=0, linetype="dashed", color = "black") +
+  geom_line(mapping = aes(y = total_inc_num_sc_ihd, colour = "total_mx_num_sc_ihd")) +
+  geom_line(mapping = aes(y = total_inc_num_diff_ihd, colour = "total_mx_num_diff_ihd")) +
+  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
+  theme(plot.title = element_text(hjust = 0.5, size = 12)) +
+  scale_color_discrete(name = paste(""), labels = c("Baseline", "Scenario", "Difference")) +
+  theme(plot.title = element_text(hjust = 0.5))
+
+#Print to view
+print(total_plot)
+```
 #Comments

### 2018-04-24 MSLT/MSLTdoc.Rmd
Updated DWs

@@
 
-output_df <- subset(output_df, select = -c(incidence_disease_ihd_bl, incidence_disease_ihd_sc, 
-                                           incidence_disease_istroke_bl, incidence_disease_istroke_sc, 
-                                           incidence_disease_diabetes_bl, incidence_disease_diabetes_sc, 
-                                           incidence_disease_breast_cancer_bl, incidence_disease_breast_cancer_sc, 
-                                           incidence_disease_colon_cancer_bl, incidence_disease_colon_cancer_sc, 
-                                           mx_ihd_bl, mx_ihd_sc, mx_istroke_bl, mx_istroke_sc, mx_diabetes_bl, mx_diabetes_sc, 
-                                           mx_breast_cancer_bl, mx_breast_cancer_sc, mx_colon_cancer_bl, mx_colon_cancer_sc, 
-                                           px_ihd_bl, px_ihd_sc, px_istroke_bl, px_istroke_sc, px_diabetes_bl, px_diabetes_sc, 
-                                           px_breast_cancer_bl, px_breast_cancer_sc, px_colon_cancer_bl, px_colon_cancer_sc))
+output_df <- subset(output_df, select = -c(incidence_disease_ihd_bl, incidence_disease_ihd_sc, incidence_disease_istroke_bl, incidence_disease_istroke_sc, incidence_disease_diabetes_bl, incidence_disease_diabetes_sc, incidence_disease_breast_cancer_bl, incidence_disease_breast_cancer_sc,
+incidence_disease_colon_cancer_bl, incidence_disease_colon_cancer_sc,  mx_ihd_bl, mx_ihd_sc, mx_istroke_bl, mx_istroke_sc, mx_diabetes_bl, mx_diabetes_sc, mx_breast_cancer_bl, mx_breast_cancer_sc, mx_colon_cancer_bl, mx_colon_cancer_sc, px_ihd_bl, px_ihd_sc, px_istroke_bl, px_istroke_sc, px_diabetes_bl, px_diabetes_sc, px_breast_cancer_bl, px_breast_cancer_sc, px_colon_cancer_bl, px_colon_cancer_sc))
 
@@
   theme(plot.title = element_text(hjust = 0.5, size = 12)) +
-  scale_color_discrete(name = paste(""), labels = c("Baseline", "Scenario", "Difference")) +
+  scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
   theme(plot.title = element_text(hjust = 0.5))

### 2018-04-24 MSLT/MSLTdoc.Rmd
Ammended dws

@@
 
-5) Change all uper cases to lower case and delete () from variables. 
+5) Change all upper cases to lower case and delete () from variables. 
 ```{r, echo=TRUE}
@@
 ```
-8) Conver string variables to numberic to do calculations.
+8) Convert string variables to numeric to do calculations.
 ```{r, echo=TRUE}
@@
 9) Generate population numbers for Greater London in a new data frame ("GBD_population").
-Note that there is data for England as well, which is used in a separate rmarkdown dowcument (GBDCompare).
+Note that there is data for England as well, which is used in a separate rmarkdown document (GBDCompare).
 Population numbers are derived from rates per 100,000 and total numbers of cases.
@@
 15) The following code brings together all the required data to create the inputs life table (Figure 3).
-ADD DIDMOD GENERATED DATA ONCE THESE ARE READY TO FOLLOW THE FORMAT OF THE IDATA USED IN THE FUNCTIONS.
-
-```{r, echo=TRUE}
-
+
+```{r, echo=TRUE}
 ## Create empty data frame
@@
                                 sep = "_"  )
-
 ##Populate life table: five_year_pop.
 
-input_life_table <- merge(input_life_table, select(GBD_population_GL
-                                                     ,c(sex_age_cat, five_year_population))
-                                                     , by = 'sex_age_cat', all = TRUE) 
+input_life_table <- merge(input_life_table, select(GBD_population_GL, c(sex_age_cat, five_year_population)), by = 'sex_age_cat', all = TRUE) 
 ##Populate life table: mortality rates (from interpolated rates)
 ## Males mortality rates. NEED MATCHING NAMES
-input_life_table <- merge(input_life_table, select(male_deaths_interpolated
-                                                     ,c(sex_age_cat, mx))
-                                                     , by = 'sex_age_cat', all.x = TRUE)
-input_life_table <- merge(input_life_table, select(female_deaths_interpolated
-                                                     ,c(sex_age_cat, mx))
-                                                     , by = 'sex_age_cat', all.x = TRUE)
-input_life_table <- merge(input_life_table, select(male_ylds_interpolated
-                                                     ,c(sex_age_cat, pyld_rate))
-                                                     , by = 'sex_age_cat', all.x = TRUE) 
-input_life_table <- merge(input_life_table, select(female_ylds_interpolated
-                                                     ,c(sex_age_cat, pyld_rate))
-                                                     , by = 'sex_age_cat', all.x = TRUE)
+input_life_table <- merge(input_life_table, select(male_deaths_interpolated , c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
+input_life_table <- merge(input_life_table, select(female_deaths_interpolated ,c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
+input_life_table <- merge(input_life_table, select(male_ylds_interpolated ,c(sex_age_cat, pyld_rate)) , by = 'sex_age_cat', all.x = TRUE) 
+input_life_table <- merge(input_life_table, select(female_ylds_interpolated ,c(sex_age_cat, pyld_rate)), by = 'sex_age_cat', all.x = TRUE)
 
@@
 ```
-16) Generate **baseline life tables** from input_data_frame. Life tables are generated for each age (5-years) and sex cohort. 
+16) Generate **baseline life tables** from input_data_frame using run_life_table function. Life tables are generated for each age (5-years) and sex cohort. 
 It is assumed that current observed rates of mortality and disability are going to be observed in the future.
@@
 ```{r, echo=TRUE}
-
-##Generate list of life tables. 
-## The function run_life_table was created to run the basline and scenario life tables 
-## (see functions.R script). 
-## The input_life_table data frame was generated to work with the function.
-
-## Generate looping variables (sex and age_group. Sex already generated 
-## in code for interpolation (i_sex))
 
@@
   for (sex in i_sex){
-    cat("age ", age, " and sex ", sex, "\n")
+    # cat("age ", age, " and sex ", sex, "\n") #Uncomment to see index
     general_life_table_list_bl[[index]] <- run_life_table(in_idata = input_life_table,
@@
 
-17) Generate **baseline disease life tables**.
+17) Generate **baseline disease life tables** using run_disase function.
 Inputs of the disease life tables are: incidence, case fatality (Figure 3, purple shaded area) and disability weights.
@@
 The code below is used to generate the inputs for Dismod II using GBD data (see Figure 3, purple shaded area).
-Dismod II requires data for the *collection* (population numbers and all cause mortality rates) and *data set* (disease specific paramenters).
-We generate a data set with all inputs for dismod by 5 year-age (represente at mid age) groups and sex for five diseases (ischemic heart disease, ischemic stroke, diabetes mellitus, colon and rectum cancer and breast cancer-women).
-WE WILL NEED ADJUSTMENT FOR TO MATCH RRS TO DISEASES (COLON AND RECTUM CANCER AND DIABETES MELLITUS)
-
-```{r, echo=TRUE}
+Dismod II requires data for the *collection* (population numbers and all cause mortality rates) and *data set* (disease specific parameters).
+We will generate a data set with all inputs for dismod by 5 year-age (represent at mid age) groups and sex for five diseases (ischemic heart disease, ischemic stroke, diabetes mellitus, colon and rectum cancer and breast cancer-women).
+WE WILL NEED ADJUSTMENT FOR TO MATCH RRS TO DISEASES (COLON AND RECTUM CANCER AND DIABETES MELLITUS).
+*Code here is work in progress**, for now the disease input data is organised externally with excel.
+
+```{r, echo=FALSE}
 ## Collection input for Greater London: population numbers (sex and 5-year age groups). 
@@
 
-```{r, echo=true}
+```{r, echo=TRUE}
 ## Use externaly generated inputs for for disease life table.
@@
       if (sex == "male" && disease == "breast_cancer"){
-        cat("\n")
+       # cat("\n") #Uncomment to see list
       }
       else {
-        cat("age ", age, " sex ", sex, "and disease", disease, "\n")
+        # cat("age ", age, " sex ", sex, "and disease", disease, "\n") #Uncomment to see list
         disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, in_disease = disease)
@@
 
-18) Generate mock change in incidence of disease to generate sceanrio life tables. In the final version, this will come from the calculated PIFs by disease. 
+18) Generate mock change in incidence of disease to generate scenario life tables. In the final version, this will come from the calculated PIFs by disease. 
 PIFs are applied here to incidence, however, may also be applied to case fatality, depending on the RRs.
@@
       if (sex == "male" && disease == "breast_cancer"){
-        cat("\n")
+        # cat("\n") # Uncomment to see list
       }
@@
 
-19) Use scenario incidence to calculate scenario life tables.
+19) Use scenario incidence to calculate scenario disease life tables.
 
@@
 index <- 1
-
 for (age in i_age_cohort){
@@
       if (sex == "male" && disease == "breast_cancer"){
-        cat("\n")
+        # cat("\n")
       }
       else {
-
-        cat("age ", age, " sex ", sex, "and disease", disease, "\n")
+        # cat("age ", age, " sex ", sex, "and disease", disease, "\n")
         # modify idata's incidence for the said scenario
@@
         disease_life_table_list_sc[[index]]$diff_pylds_disease <- (disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px) * disease_life_table_list_bl[[index]]$dw_disease
-
         index <- index + 1
@@
 
-20) Calculate life tables scearnio.
+20) Calculate life tables scenario.
 
@@
       if (sex == "male" && disease == "breast_cancer"){
-        cat("\n")
+        # cat("\n")
       }else{
@@
 
-        cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
+        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
         index <- index + 1
       }
-
     }
@@
 
-        cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
+        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
         index <- index + 1
@@
   for (sex in i_sex){
-
-    cat("age ", age, " and sex ", sex, "\n")
+    # cat("age ", age, " and sex ", sex, "\n")
     # modify idata's mortality and pyld total for the said scenario
@@
           names(output_burden_sc)[names(output_burden_sc) == 'px'] <- paste('px', disease, "sc", sep = "_")
-
           output_burden_bl <- select(disease_life_table_list_bl[[index]], c('incidence_disease', 'mx', 'px'))
@@
 
-```{r, echo=true}
+```{r, echo=TRUE}
 plot_eg <- plot_output(in_data = output_df, in_age = 22, in_population = "male", in_outcomes = c("age", "mx_num_bl_ihd", "mx_num_sc_ihd", "mx_num_diff_ihd"), in_legend = "Ischemic Heart Disease Deaths")
@@
 
-```{r, echo=true}
+```{r, echo=TRUE}
 ####Generate data frame with all outputs to graph total change in burden by simulation year. 
@@
 
-```{r, echo=true}
+```{r, echo=TRUE}
 

### 2018-04-24 MSLT/MSLTdoc.Rmd
Added loop to age and sex cohort change in burden

@@
 ```{r, echo=TRUE}
-plot_eg <- plot_output(in_data = output_df, in_age = 22, in_population = "male", in_outcomes = c("age", "mx_num_bl_ihd", "mx_num_sc_ihd", "mx_num_diff_ihd"), in_legend = "Ischemic Heart Disease Deaths")
+## Do a loop, the first in_outcomes is age.
+
+i_outcome <- c("mx", "inc")
+
+for (age in i_age_cohort){
+  for (sex in i_sex) {
+    for (disease in i_disease){
+      for (outcome in i_outcome) {
+
+plot_eg <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, age, sex, sep = "_"))
+      }
+    }
+  }
+}
 ```

### 2018-04-24 MSLT/MSLTdoc.Rmd
Loop to age and sex results plot

--- 2018-04-24 MSLT/MSLTdoc.Rmd
@@
 ```{r, echo=TRUE}
-plot_eg <- plot_output(in_data = output_df, in_age = 22, in_population = "male", in_outcomes = c("age", "mx_num_bl_ihd", "mx_num_sc_ihd", "mx_num_diff_ihd"), in_legend = "Ischemic Heart Disease Deaths")
+## Do a loop, the first in_outcomes is age.
+
+i_outcome <- c("mx", "inc")
+
+for (age in i_age_cohort){
+  for (sex in i_sex) {
+    for (disease in i_disease){
+      for (outcome in i_outcome) {
+        {
+      if (sex == "male" && disease == "breast_cancer"){
+        # cat("\n")
+      }else{
+
+plot_eg <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, age, sex, sep = "_"))
+          }
+        }
+      }
+    }
+  }
+}
 ```

### 2018-04-25 MSLT/MSLTdoc.Rmd
Plots

@@
 require(citr)
+require(gridExtra)
 ```
@@
 
-plot_eg <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, age, sex, sep = "_"))
-          }
-        }
-      }
-    }
-  }
-}
-```
+ p <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, sep = " "))
+           }
+         }
+       }
+     }
+   }
+ }
+
+```
+
+paste(age, sex, disease, outcome, sep = "_")
 

### 2018-04-25 MSLT/MSLTdoc.Rmd
WHO data availability

@@
 The *PMSLT* follows a population of interest over time.
-For example, as set up here, we simulate sex and age (5 years starting at 18) cohorts over time until they die or reach 100 years of age.
+For example, as set up here, we simulate sex and age (5 years starting at 20) cohorts over time until they die or reach 100 years of age.
 This implies that we can include trends for diseases, time lags for change in exposure to risk factors and change in health and demographic changes (e.g.
@@
 
-***Interaction between multiple diseases**
+* **Interaction between multiple diseases**
 The *PMSLT* accounts for the interaction between multiple diseases, with proportions of the population being able to be in more than one health state [@RN8957].
@@
 
-***Mortality rate** 
+* **Mortality rate** 
 The *PMSLT* calculations for changes in life years (and health-adjusted life years) and mortality outcomes is based on observed mortality rates for the population of interest.
@@
 
-***Impact of disability in increased life expectancy**
+* **Impact of disability in increased life expectancy**
 In GBD studies, YLLs are not adjusted for disability; hence, their use in estimating intervention effects results in over-estimation, which the *PMSLT* approach avoids.
@@
 Where possible, we wrote functions and loops to avoid repetition.
-We set up the model  with data for Great London.
-Figure 2 is depicts the PMSLT model framework, which was followed in the code development.
+We set up the model  with data for Greater London.
+Figure 2 depicts the PMSLT model framework, which was followed in the code development.
 
@@
 ####include. To show all select 16. what is specified in in_outcomes will be graphed or can also be presented as a total. 
-aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('inc_num_bl_ihd','inc_num_sc_ihd', "inc_num_diff_ihd"))
-
-aggregate_frame_females <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c('inc_num_bl_ihd','inc_num_sc_ihd', "inc_num_diff_ihd"))
+aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))
+
+aggregate_frame_females <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))
 
@@
 
-disease_outcome <- "Deaths ischemic heart disease"
-
-total_plot <- ggplot(total_aggr[1:20,], aes(x = sim_year)) +
-
-  geom_line(mapping = aes(y = total_inc_num_bl_ihd, colour = "total_mx_num_bl_ihd")) +
+disease_outcome <- "Health adjusted life years"
+
+total_plot <- ggplot(total_aggr[1:79,], aes(x = sim_year)) +
+
+  geom_line(mapping = aes(y = total_Lx_bl, colour = "total_Lx_bl")) +
   theme_classic() +
   geom_hline(yintercept=0, linetype="dashed", color = "black") +
-  geom_line(mapping = aes(y = total_inc_num_sc_ihd, colour = "total_mx_num_sc_ihd")) +
-  geom_line(mapping = aes(y = total_inc_num_diff_ihd, colour = "total_mx_num_diff_ihd")) +
+  geom_line(mapping = aes(y = total_Lx_sc, colour = "total_total_Lx_sc")) +
+  geom_line(mapping = aes(y = total_Lx_diff, colour = "total_Lx_diff")) +
   xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +

### 2018-04-27 MSLT/MSLTdoc.Rmd
Multiple outputs in one page

@@
 require(gridExtra)
+require(cowplot)
 ```
@@
 i_outcome <- c("mx", "inc")
-
+p_list <- list()
+
+index <- 1
 for (age in i_age_cohort){
@@
        }else{
-
- p <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, sep = " "))
-           }
-         }
-       }
-     }
-   }
- }
-
-```
-
-paste(age, sex, disease, outcome, sep = "_")
+# names(p_list) <- unique(age, sex, disease, outcome)
+ p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, sep = " "))
+p_list[[index]] <- p_index
+index <- index + 1
+
+            }
+         }
+       }
+     }
+   }
+}
+
+# m<-plot_grid(p_index, ncol=2, nrow=2)
+# library(cowplot)
+# save_plot("output.tiff",m ,ncol = 2, nrow = 2) # make room for figure legend)
+
+# png(file="superplot.png")
+# marrangeGrob(grobs=p_list, nrow=1, ncol=1)
+# # do.call("grid.arrange", p_index)
+
+ml <- do.call(marrangeGrob, list(grobs=p_list, nrow = 2, ncol = 2))
+ml 
+```
 

### 2018-04-27 MSLT/code/functions.R

@@ -328,6 +328,7 @@ plot_output <- function(in_data, in_age, in_population, in_outcomes, in_legend){
   # in_population <- "males"
   # in_age <- 22
   # in_outcomes <- c('age', 'inc_num_bl_ihd', 'inc_num_sc_ihd')
+  # in_legend <- "none"
   # in_cols <- c('alpha', 'beta')
   
   data <- in_data
@@ -346,6 +347,7 @@ plot_output <- function(in_data, in_age, in_population, in_outcomes, in_legend){
   for (i in 2:length(in_outcomes)) {
     # use aes_string with names of the data.frame
     p <- p + geom_line(aes_string(y = td[[in_outcomes[i]]], color = as.factor(in_outcomes[i])), size = 0.8) +
+      
       theme_classic() 
     
     
@@ -357,6 +359,10 @@ plot_output <- function(in_data, in_age, in_population, in_outcomes, in_legend){
     theme(plot.title = element_text(hjust = 0.5, size = 12)) + 
     xlim(in_age, 100) +
     geom_hline(yintercept=0, linetype="dashed", color = "black")
+  
+  grob_all <- arrangeGrob(p)
+  return(grob_all)
+  
   
   
   # print the result

### 2018-04-27 MSLT/MSLTdoc.Rmd
Multiple plots

@@
 require(citr)
-require(gridExtra)
-require(cowplot)
+# require(gridExtra)
+# require(cowplot)
+# install.packages("ggpubr")
+require(ggpubr)
 ```
@@
 
-```{r, echo=TRUE}
+```{r echo=TRUE}
 ## Do a loop, the first in_outcomes is age.
@@
        }else{
-# names(p_list) <- unique(age, sex, disease, outcome)
+
  p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, sep = " "))
@@
 
-ml <- do.call(marrangeGrob, list(grobs=p_list, nrow = 2, ncol = 2))
+ml <- do.call(marrangeGrob, list(grobs=p_list, nrow = 3, ncol = 2))
 ml
+
+#Check ggarrange in ggpubr
+
+#http://www.sthda.com/english/wiki/print.php?id=177
 ```

### 2018-04-27 MSLT/code/functions.R

@@ -353,10 +353,13 @@ plot_output <- function(in_data, in_age, in_population, in_outcomes, in_legend){
     
   }
   
-  p <- p + scale_color_discrete(name = paste(in_legend), labels = c("Baseline", "Difference", "Scenario"))
+  p <- p + scale_color_discrete(name = paste(in_legend), labels = c("Baseline", "Difference", "Scenario")) +
+  theme(legend.title = element_text(size = 7))
   
   p <- p + xlab ('Age') + ylab ('Cases') + labs (title = paste('Cohort', in_age, "years old", in_population, sep = " ")) +
-    theme(plot.title = element_text(hjust = 0.5, size = 12)) + 
+    theme(plot.title = element_text(hjust = 0.5, size = 9)) +
+    theme(legend.text = element_text(size = 7)) +
+    theme(axis.title.x = element_text(size = 7)) +
     xlim(in_age, 100) +
     geom_hline(yintercept=0, linetype="dashed", color = "black")
   

### 2018-04-27 MSLT/MSLTdoc.Rmd
Updated PDF

@@
 require(citr)
-# require(gridExtra)
-# require(cowplot)
-# install.packages("ggpubr")
+require(gridExtra)
+require(cowplot)
 require(ggpubr)

### 2018-04-27 MSLT/code/functions.R

@@ -359,7 +359,7 @@ plot_output <- function(in_data, in_age, in_population, in_outcomes, in_legend){
   p <- p + xlab ('Age') + ylab ('Cases') + labs (title = paste('Cohort', in_age, "years old", in_population, sep = " ")) +
     theme(plot.title = element_text(hjust = 0.5, size = 9)) +
     theme(legend.text = element_text(size = 7)) +
-    theme(axis.title.x = element_text(size = 7)) +
+    # theme(axis.title.x = element_text(size = 7)) +
     xlim(in_age, 100) +
     geom_hline(yintercept=0, linetype="dashed", color = "black")
   

### 2018-05-01 MSLT/MSLTdoc.Rmd
Tidy up code

@@
 
-#Global options
+## Global options
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-#The work in this document requires the following packages (code, rmarkdown and citing packages):
+## The work in this document requires the following packages (code, rmarkdown and citing packages):
+
 require(dplyr)
@@
 require(ggpubr)
+
 ```
@@
 ```{r, echo=FALSE}
+
 pmslt_inputs <- tribble(
@@
   row_spec(0, bold = T)
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBDdata <- read.csv("data/UK/gbd2016.csv", stringsAsFactors = F)
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBDdata <- mutate_all(GBDdata, funs(tolower))
 GBDdata$measure[GBDdata$measure=="ylds (years lived with disability)"] <- "ylds"
+
 ```
@@
 ```{r, echo=TRUE}
+
 # 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97
@@
 GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBDdata$sex_age_cat <- paste(GBDdata$sex,GBDdata$age_cat, sep = "_"  )
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBDdata$val <- as.numeric(as.character(GBDdata$val))
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", 
@@
 sex_age_cat, location)
+
 ```
@@
 ```{r, echo=TRUE}
+
  for (i in 1:nrow(GBD_population)) {
@@
  }
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBD_population <- GBD_population[!is.na(GBD_population$five_year_population),]
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBD_population <- filter(GBD_population) %>% 
 select(sex_age_cat, sex, age_cat, five_year_population, location)
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% 
 select(sex_age_cat, age_cat, sex, five_year_population, location, sex_age_cat)
+
 ```
@@
 ```{r, echo=TRUE}
+
 GreaterLondon <- sum(GBD_population_GL$five_year_population)
+
 ```
@@
 ```{r, echo=TRUE}
+
 GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% 
@@
 GBDGL$one_rate <- GBDGL$val/100000
+
 ```
@@
 i_measure <- c("deaths", "ylds") #" (years lived with disability)")
+
+index <- 1
 
@@
 assign(paste(sex_index, measure_index, "interpolated", sep = "_"), interpolated)
-# plot(interpolated$age,interpolated$rate)
-
-##Do graph with another layer for original rates in age groups for comparison purposes.
-p <- ggplot(data = interpolated,mapping = aes(age, interpolated[,1])) +
+
+  }
+}
+
+## Do graph with another layer for original rates in age groups for comparison purposes.
+
+p_interpolation_list <- list()
+
+interpolation_index <- 1
+
+for(sex_index in i_sex) {
+  for (measure_index in i_measure) {
+
+p_interpolation_index <- ggplot(data = interpolated,mapping = aes(age, interpolated[,1])) +
               geom_line(aes(color = "Interpolated")) +
@@
               theme (plot.title = element_text(hjust = 0.5))
-              print(p)
-  }
-}
+
+              p_interpolation_list[[interpolation_index]] <- p_interpolation_index
+              interpolation_index <- interpolation_index + 1
+
+  }
+}
+
+## Save plots to tiff. Makes a separate file for each plot.
+
+interpolation_index <- 1
+for(sex_index in i_sex) {
+  for (measure_index in i_measure) {
+    file_name = paste("graphs/", "Interpolation rates", sex_index, measure_index, ".tiff", sep=" ")
+    tiff(file_name)
+    print(p_interpolation_list[[interpolation_index]])
+    interpolation_index <- interpolation_index + 1
+    dev.off()
+  }
+}
+
+p_interpolated <- do.call(marrangeGrob, list(grobs=p_interpolation_list, nrow = 2, ncol = 2))
+p_interpolated
 
@@
 ```{r, echo=TRUE}
+
 ## Create empty data frame
+
 input_life_table <- as.data.frame(matrix(0, ncol=2, nrow = 202))
+
 ## Give variables names (mx=total mortality rate, 
 ## yldsx = total years lived with disability rates )
+
 names(input_life_table) <- c("age", "sex")
+
 ## Populate life table: Age, 0 to 100
+
 input_life_table[1:101, 1] <- c(0:100)
 input_life_table[102:202, 1] <- c(0:100)
+
 ## Populate life table: sex, male, female
+
 input_life_table[1:101, 2] <- "male"
 input_life_table[102:202, 2] <- "female"
-## Create variable age_sex to match with population data. 
+
+## Create variable age_sex to match with population data
+
 input_life_table$sex_age_cat <- paste(input_life_table$sex,input_life_table$age,
                                 sep = "_"  )
-##Populate life table: five_year_pop.
-
-input_life_table <- merge(input_life_table, select(GBD_population_GL, c(sex_age_cat, five_year_population)), by = 'sex_age_cat', all = TRUE) 
-##Populate life table: mortality rates (from interpolated rates)
+## Populate life table: five_year_pop.
+
+input_life_table <- merge(input_life_table, select(GBD_population_GL, 
+                    c(sex_age_cat, five_year_population)), by = 'sex_age_cat', all = TRUE)
+
+## Populate life table: mortality rates (from interpolated rates)
 ## Males mortality rates. NEED MATCHING NAMES
-input_life_table <- merge(input_life_table, select(male_deaths_interpolated , c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
-input_life_table <- merge(input_life_table, select(female_deaths_interpolated ,c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
-input_life_table <- merge(input_life_table, select(male_ylds_interpolated ,c(sex_age_cat, pyld_rate)) , by = 'sex_age_cat', all.x = TRUE) 
-input_life_table <- merge(input_life_table, select(female_ylds_interpolated ,c(sex_age_cat, pyld_rate)), by = 'sex_age_cat', all.x = TRUE)
+
+input_life_table <- merge(input_life_table, select(male_deaths_interpolated, 
+                    c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
+input_life_table <- merge(input_life_table, select(female_deaths_interpolated,
+                    c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
+input_life_table <- merge(input_life_table, select(male_ylds_interpolated, 
+                    c(sex_age_cat, pyld_rate)) , by = 'sex_age_cat', all.x = TRUE) 
+input_life_table <- merge(input_life_table, select(female_ylds_interpolated, 
+                    c(sex_age_cat, pyld_rate)), by = 'sex_age_cat', all.x = TRUE)
 
@@
                         input_life_table$mx.x, input_life_table$mx.y)
+
 input_life_table$pyld_rate <-  ifelse(input_life_table$sex == "male",
@@
 ## Cross check population numbers
+
 sum(input_life_table$five_year_population, na.rm = TRUE)
-## Sort data by sex and age to use in life table function. 
+
+## Sort data by sex and age to use in life table function
+
 input_life_table<-input_life_table[order(input_life_table$sex, input_life_table$age),]
@@
 ```{r, echo=TRUE}
+
 ## Use externaly generated inputs for for disease life table.
@@
         # cat("age ", age, " sex ", sex, "and disease", disease, "\n") #Uncomment to see list
-        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, in_disease = disease)
+        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, 
+                                               in_disease = disease)
         index <- index + 1
@@
 }
-##### Uncommnet to check disease life table list
+## Uncommnet to check disease life table list
 # View(disease_life_table_list_bl[[8]])
@@
 
-#####Generate scenario incidence (for each disease)
+## Generate scenario incidence (for each disease)
 
@@
 
-        incidence_sc[[index]] <- disease_life_table_list_bl[[index]]$incidence_disease * incidence_change
+        incidence_sc[[index]] <- disease_life_table_list_bl[[index]]$incidence_disease * 
+        incidence_change
         index <- index + 1
@@
 
-##### Uncommnet to check scenario incidence
+## Uncommnet to check scenario incidence
 # View(incidence_sc[[1]])
@@
         # Instead of idata, feed td to run scenarios
-        disease_life_table_list_sc[[index]] <- run_disease(in_idata = td1, in_sex = sex, in_mid_age = age, in_disease = disease)
-        disease_life_table_list_sc[[index]]$diff_inc_disease <- disease_life_table_list_sc[[index]]$incidence_disease - disease_life_table_list_bl[[index]]$incidence_disease
-        disease_life_table_list_sc[[index]]$diff_prev_disease <- disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px
-        disease_life_table_list_sc[[index]]$diff_mort_disease <- disease_life_table_list_sc[[index]]$mx - disease_life_table_list_bl[[index]]$mx
-        disease_life_table_list_sc[[index]]$diff_pylds_disease <- (disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px) * disease_life_table_list_bl[[index]]$dw_disease
+        disease_life_table_list_sc[[index]] <- run_disease(in_idata = td1, in_sex = sex, 
+                                               in_mid_age = age, in_disease = disease)
+        disease_life_table_list_sc[[index]]$diff_inc_disease <- 
+                                              disease_life_table_list_sc[[index]]$incidence_disease 
+                                               -   disease_life_table_list_bl[[index]]$incidence_disease
+        disease_life_table_list_sc[[index]]$diff_prev_disease <- 
+                                              disease_life_table_list_sc[[index]]$px 
+                                              - disease_life_table_list_bl[[index]]$px
+        disease_life_table_list_sc[[index]]$diff_mort_disease <- 
+                                              disease_life_table_list_sc[[index]]$mx 
+                                              - disease_life_table_list_bl[[index]]$mx
+        disease_life_table_list_sc[[index]]$diff_pylds_disease <- 
+                                              (disease_life_table_list_sc[[index]]$px  
+                                              - disease_life_table_list_bl[[index]]$px) * 
+                                              disease_life_table_list_bl[[index]]$dw_disease
         index <- index + 1
@@
 }
-##### Uncommnet to check scenario life tables
+## Uncommnet to check scenario life tables
 # View(disease_life_table_list_sc[[1]])
@@
 ```{r, echo=TRUE}
-#####Generate total change in mortality rate
-##### Sum mortality rate scenarios (mx_sc_total)
+
+## Generate total change in mortality rate
+## Sum mortality rate scenarios (mx_sc_total)
 
@@
         if (create_new){
-          mortality_sum <- select(disease_life_table_list_sc[[index]], c('age', 'sex'))
+          mortality_sum <- select(disease_life_table_list_sc[[index]], 
+                            c('age', 'sex'))
                             mortality_sum$total <- 0
           create_new <- F
-          mortality_sum$total <- mortality_sum$total + (disease_life_table_list_sc[[index]]$diff_mort_disease)
+          mortality_sum$total <- mortality_sum$total + 
+                                (disease_life_table_list_sc[[index]]$diff_mort_disease)
         }else{
-          mortality_sum$total <- mortality_sum$total + (disease_life_table_list_sc[[index]]$diff_mort_disease)
+          mortality_sum$total <- mortality_sum$total +
+                                (disease_life_table_list_sc[[index]]$diff_mort_disease)
         }
@@
 
-##### Uncommnet to check sceanrio mortality and changes 
+## Uncommnet to check sceanrio mortality and changes 
 # View(mx_sc_total[[1]])
 
-#####Generate total change in prevalent yld rates
-#####total ylds rate= sum (change prevalence disease*dw)
+## Generate total change in prevalent yld rates
+## Total ylds rate= sum (change prevalence disease*dw)
 
@@
           create_new <- F
-          pylds_sum$total <- pylds_sum$total +   (disease_life_table_list_sc[[index]]$diff_pylds_disease)
+          pylds_sum$total <- pylds_sum$total +   
+                            (disease_life_table_list_sc[[index]]$diff_pylds_disease)
         }else{
-          pylds_sum$total <- pylds_sum$total + (disease_life_table_list_sc[[index]]$diff_pylds_disease)
+          pylds_sum$total <- pylds_sum$total + 
+                            (disease_life_table_list_sc[[index]]$diff_pylds_disease)
         }
@@
 
-##### Uncommnet to check scenario pyld change
+## Uncommnet to check scenario pyld change
 # View(pylds_sc_total[[2]])
 
-##### Calculate general life tables with modified mortality and pylds total#############################
-###Original mortality rate is modified by the mx_sc_total (total change in mortality from diseases)
-###Original pyld rate is modified by the change in each disease pylds
+## Calculate general life tables with modified mortality and pylds total
+## Original mortality rate is modified by the mx_sc_total (total change in mortality from diseases)
+## Original pyld rate is modified by the change in each disease pylds
 
@@
     td2 <- input_life_table
-    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx + mx_sc_total[[index]]$total
-    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate + pylds_sc_total[[index]]$total
+    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx 
+                                                            + mx_sc_total[[index]]$total
+    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate 
+                                                                    + pylds_sc_total[[index]]$total
 
@@
 }
-##### Uncommnet to check scenario life tables
+## Uncommnet to check scenario life tables
 # View(general_life_table_list_sc[[1]])
 # View(general_life_table_list_bl[[2]])
-```
-
-21) Generate list of outputs by age and sex
-
-```{r, echo=TRUE}
-#####In the following list "output_life_table", 32 data frames are nested per age and sex cohort
+
+```
+
+21) Generate list of outputs by age and sex.
+
+```{r, echo=TRUE}
+
+## In the following list "output_life_table", 32 data frames are nested per age and sex cohort
 
@@
 
-    #Males do not have breast cancer, that is why we need the if/else. 
-    #We create a TRUE/FALSE variable for the loop to move into the next disease
+    # Males do not have breast cancer, that is why we need the if/else. 
+    # We create a TRUE/FALSE variable for the loop to move into the next disease
 
@@
         if (create_new){
-          output_burden_sc <- select(disease_life_table_list_sc[[index]], c('age', 'sex', 'incidence_disease', 'mx', 'px'))
-          names(output_burden_sc)[names(output_burden_sc) == 'incidence_disease'] <- paste('incidence_disease', disease, "sc", sep = "_")
-          names(output_burden_sc)[names(output_burden_sc) == 'mx'] <- paste('mx', disease, "sc", sep = "_")
-          names(output_burden_sc)[names(output_burden_sc) == 'px'] <- paste('px', disease, "sc", sep = "_")
-          output_burden_bl <- select(disease_life_table_list_bl[[index]], c('incidence_disease', 'mx', 'px'))
-          names(output_burden_bl)[names(output_burden_bl) == 'incidence_disease'] <- paste('incidence_disease', disease, "bl", sep = "_")
-          names(output_burden_bl)[names(output_burden_bl) == 'mx'] <- paste('mx', disease, "bl", sep = "_")
-          names(output_burden_bl)[names(output_burden_bl) == 'px'] <- paste('px', disease, "bl", sep = "_")
-
-          ####New list to add calculations for changes in burden of disease (incidence and mortality numbers)
+          output_burden_sc <- select(disease_life_table_list_sc[[index]], 
+                              c('age', 'sex', 'incidence_disease', 'mx', 'px'))
+          names(output_burden_sc)[names(output_burden_sc) == 'incidence_disease'] <- 
+                              paste('incidence_disease', disease, "sc", sep = "_")
+          names(output_burden_sc)[names(output_burden_sc) == 'mx'] <- 
+                              paste('mx', disease, "sc", sep = "_")
+          names(output_burden_sc)[names(output_burden_sc) == 'px'] <- 
+                              paste('px', disease, "sc", sep = "_")
+          output_burden_bl <- select(disease_life_table_list_bl[[index]], 
+                              c('incidence_disease', 'mx', 'px'))
+          names(output_burden_bl)[names(output_burden_bl) == 'incidence_disease'] <- 
+                              paste('incidence_disease', disease, "bl", sep = "_")
+          names(output_burden_bl)[names(output_burden_bl) == 'mx'] <- 
+                              paste('mx', disease, "bl", sep = "_")
+          names(output_burden_bl)[names(output_burden_bl) == 'px'] <- 
+                              paste('px', disease, "bl", sep = "_")
+
+          ## New list to add calculations for changes in burden of disease (incidence and mortality numbers)
 
@@
 
-          output_burden_change$inc_num_bl <- disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx
-          output_burden_change$inc_num_sc <- disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx
-          output_burden_change$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx)
+          output_burden_change$inc_num_bl <- disease_life_table_list_bl[[index]]$incidence_disease * 
+          (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx
+          output_burden_change$inc_num_sc <- disease_life_table_list_sc[[index]]$incidence_disease * 
+          (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx
+          output_burden_change$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * 
+          (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) 
+          - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) 
+          * general_life_table_list_bl[[l_index]]$Lx)
 
@@
 
-          names(output_burden_change)[names(output_burden_change) == 'inc_num_bl'] <- paste('inc_num_bl', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'inc_num_sc'] <- paste('inc_num_sc', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'inc_num_diff'] <- paste('inc_num_diff', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'mx_num_bl'] <- paste('mx_num_bl', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'mx_num_sc'] <- paste('mx_num_sc', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'mx_num_diff'] <- paste('mx_num_diff', disease, sep = "_")
-
-          ###Bind all lists
+          names(output_burden_change)[names(output_burden_change) == 'inc_num_bl'] <- 
+          paste('inc_num_bl', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'inc_num_sc'] <- 
+          paste('inc_num_sc', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'inc_num_diff'] <- 
+          paste('inc_num_diff', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'mx_num_bl'] <- 
+          paste('mx_num_bl', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'mx_num_sc'] <- 
+          paste('mx_num_sc', disease, sep = "_")
+          names(output_burden_change)[names(output_burden_change) == 'mx_num_diff'] <- 
+          paste('mx_num_diff', disease, sep = "_")
+
+          ## Bind all lists
 
@@
 
-          #Here the calculations above are repeated, here is where the F is telling to move into the next disease
+          ## Here the calculations above are repeated, here is where the F is telling to move into the next disease
 
@@
 
-          td3 <- select(disease_life_table_list_sc[[index]], c('incidence_disease', 'mx', 'px'))
-          names(td3)[names(td3) == 'incidence_disease'] <- paste('incidence_disease', disease, "sc", sep = "_")
-          names(td3)[names(td3) == 'mx'] <- paste('mx', disease, "sc", sep = "_")
-          names(td3)[names(td3) == 'px'] <- paste('px', disease, "sc", sep = "_")
-
-          td4 <- select(disease_life_table_list_bl[[index]], c('incidence_disease', 'mx', 'px'))
-          names(td4)[names(td4) == 'incidence_disease'] <- paste('incidence_disease', disease, "bl", sep = "_")
-          names(td4)[names(td4) == 'mx'] <- paste('mx', disease, "bl", sep = "_")
-          names(td4)[names(td4) == 'px'] <- paste('px', disease, "bl", sep = "_")
+          td3 <- select(disease_life_table_list_sc[[index]], 
+                c('incidence_disease', 'mx', 'px'))
+          names(td3)[names(td3) == 'incidence_disease'] <-
+          paste('incidence_disease', disease, "sc", sep = "_")
+          names(td3)[names(td3) == 'mx'] <- 
+          paste('mx', disease, "sc", sep = "_")
+          names(td3)[names(td3) == 'px'] <- 
+          paste('px', disease, "sc", sep = "_")
+
+          td4 <- select(disease_life_table_list_bl[[index]], 
+                c('incidence_disease', 'mx', 'px'))
+          names(td4)[names(td4) == 'incidence_disease'] <- 
+          paste('incidence_disease', disease, "bl", sep = "_")
+          names(td4)[names(td4) == 'mx'] <- 
+          paste('mx', disease, "bl", sep = "_")
+          names(td4)[names(td4) == 'px'] <- 
+          paste('px', disease, "bl", sep = "_")
 
@@
 
-          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_bl'] <- paste('inc_num_bl', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_sc'] <- paste('inc_num_sc', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_diff'] <- paste('inc_num_diff', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_bl'] <- paste('mx_num_bl', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_sc'] <- paste('mx_num_sc', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_diff'] <- paste('mx_num_diff', disease, sep = "_")
-
-          ###Bind all lists
+          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_bl'] <- 
+          paste('inc_num_bl', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_sc'] <- 
+          paste('inc_num_sc', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_diff'] <- 
+          paste('inc_num_diff', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_bl'] <- 
+          paste('mx_num_bl', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_sc'] <- 
+          paste('mx_num_sc', disease, sep = "_")
+          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_diff'] <- 
+          paste('mx_num_diff', disease, sep = "_")
+
+          ## Bind all lists
 
@@
 
-    # general_life_table_list_sc and general_life_table_list_bl (Lx)
-    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], c('Lx', 'Lwx'))
-    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- paste('Lx', "sc", sep = "_")
-    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- paste('Lwx', "sc", sep = "_")
-
-    output_burden_lf_bl <- select(general_life_table_list_bl[[l_index]], c('Lx', 'Lwx'))
-    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lx'] <- paste('Lx', "bl", sep = "_")
-    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lwx'] <- paste('Lwx', "bl", sep = "_")
-
-    output_burden_lf_sc$Lx_diff <- general_life_table_list_bl[[l_index]]$Lx - general_life_table_list_sc[[l_index]]$Lx
-    output_burden_lf_sc$Lwx_diff <- general_life_table_list_bl[[l_index]]$Lwx - general_life_table_list_sc[[l_index]]$Lwx
+    ## general_life_table_list_sc and general_life_table_list_bl (Lx)
+    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], 
+                          c('Lx', 'Lwx'))
+    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- 
+                          paste('Lx', "sc", sep = "_")
+    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- 
+                          paste('Lwx', "sc", sep = "_")
+
+    output_burden_lf_bl <- select(general_life_table_list_bl[[l_index]], 
+                          c('Lx', 'Lwx'))
+    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lx'] <- 
+                          paste('Lx', "bl", sep = "_")
+    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lwx'] <- 
+                          paste('Lwx', "bl", sep = "_")
+
+    output_burden_lf_sc$Lx_diff <- general_life_table_list_bl[[l_index]]$Lx 
+                                - general_life_table_list_sc[[l_index]]$Lx
+    output_burden_lf_sc$Lwx_diff <- general_life_table_list_bl[[l_index]]$Lwx 
+                                - general_life_table_list_sc[[l_index]]$Lwx
 
@@
 
-#Uncomment to check
+## Uncomment to check
 # View(output_burden[[32]])
@@
 p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, sep = " "))
+
 p_list[[index]] <- p_index
@@
 
-# m<-plot_grid(p_index, ncol=2, nrow=2)
-# library(cowplot)
-# save_plot("output.tiff",m ,ncol = 2, nrow = 2) # make room for figure legend)
-
-# png(file="superplot.png")
-# marrangeGrob(grobs=p_list, nrow=1, ncol=1)
-# # do.call("grid.arrange", p_index)
-
-ml <- do.call(marrangeGrob, list(grobs=p_list, nrow = 3, ncol = 2))
-ml
+## Use plot_grid in a loo/grid.arrange
+
+mpage <- do.call(marrangeGrob, list(grobs=p_list, nrow = 3, ncol = 2))
+mpage
+
+## Trying to use ggarrange
+# mpage2 <- ggarrange(plotlist = p_list, ncol = 3, nrow = 3)
+
+# ggsave("graphs/multipage.jpeg", mpage)
 

### 2018-05-01 MSLT/code/functions.R

@@ -363,6 +363,7 @@ plot_output <- function(in_data, in_age, in_population, in_outcomes, in_legend){
     xlim(in_age, 100) +
     geom_hline(yintercept=0, linetype="dashed", color = "black")
   
+  
   grob_all <- arrangeGrob(p)
   return(grob_all)
   

### 2018-05-02 MSLT/MSLTdoc.Rmd
Generate age, sex and outcome specific group plots

@@
           output_burden_change$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * 
-          (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) 
-          - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) 
+          (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px)
           * general_life_table_list_bl[[l_index]]$Lx)
@@
 
+require(ggplot2)
+require(gridExtra)
+require(grid)
+
+grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, position = c("bottom", "right"), mainTitle = "", mainLeft = "", mainBottom = "") {
+
+  plots <- list(...)
+  position <- match.arg(position)
+  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
+  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
+  lheight <- sum(legend$height)
+  lwidth <- sum(legend$width)
+  gl <- lapply(plots, function(x) x + theme(legend.position="none"))
+  gl <- c(gl, ncol = ncol, nrow = nrow, top = mainTitle, left = mainLeft, bottom = mainBottom)
+
+  combined <- switch(position,
+                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
+                                            legend,
+                                            ncol = 1,
+                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
+                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
+                                           legend,
+                                           ncol = 2,
+                                           widths = unit.c(unit(1, "npc") - lwidth, lwidth)))
+
+  #grid.newpage()
+  grid.draw(combined)
+
+  # return gtable invisibly
+  invisible(combined)
+
+}
+
+g_legend <- function(a.gplot){
+  tmp <- ggplot_gtable(ggplot_build(a.gplot))
+  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
+  legend <- tmp$grobs[[leg]]
+  return(legend)
+}
+
+output_dir = "graphs"
+
 i_outcome <- c("mx", "inc")
-p_list <- list()
-
-index <- 1
+p_list_male <- list()
+p_list_female <- list()
+male_index <- 1
+female_index <- 1
 for (age in i_age_cohort){
       for (sex in i_sex) {
+        for (outcome in i_outcome) {
           for (disease in i_disease){
-       for (outcome in i_outcome) {
-         {
+
        if (sex == "male" && disease == "breast_cancer"){
@@
 
-p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_legend = paste(outcome, disease, sep = " "))
-
-p_list[[index]] <- p_index
-index <- index + 1
-
-            }
+            p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_disease = disease)
+
+            if (sex == "male"){
+
+              p_list_male[[male_index]] <- p_index
+
+              if (male_index %% 4 == 0 && male_index > 0){
+
+                p1 <- p_list_male[[male_index - 3]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+                p2 <- p_list_male[[male_index - 2]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+                p3 <- p_list_male[[male_index - 1]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+                p4 <- p_index + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+
+                pdf(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".pdf"))
+                grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(age, sex, outcome),
+                                          mainLeft = 'Cases', mainBottom = 'Age')
+                dev.off()
+
+              }
+
+              male_index <- male_index + 1
+
+            }
+
+            if (sex == "female" && female_index > 0){
+              p_list_female[[female_index]] <- p_index
+
+              if (female_index %% 5 == 0){
+                p1 <- p_list_female[[female_index - 4]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+                p2 <- p_list_female[[female_index - 3]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+                p3 <- p_list_female[[female_index - 2]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+                p4 <- p_list_female[[female_index - 1]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+                p5 <- p_index + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
+
+                pdf(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".pdf"))
+                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(age, sex, outcome),
+                                            mainLeft = 'Cases', mainBottom = 'Age')
+                dev.off()
+
+              }
+              female_index <- female_index + 1
+            }
+
             }
@@
 
-mpage <- do.call(marrangeGrob, list(grobs=p_list, nrow = 3, ncol = 2))
-mpage
+# mpage_male <- do.call(marrangeGrob, list(grobs=p_list_male, nrow = 2, ncol = 2))
+# mpage_male
+# 
+# 
+# lay <- rbind(c(1, 2),
+#              c(3, 4),
+#              c(5, NA))
+# 
+# 
+# mpage_female <- do.call(marrangeGrob, list(grobs=p_list_female, layout_matrix = lay))
+# mpage_female
 

### 2018-05-02 MSLT/MSLTdoc.Rmd
Move legend function to function scr
ipt

@@
 
-23) Plot outcomes for each age and sex cohort using created function. ADD A LOOP TO GENERATE ALL OUTCOMES FOR IN_AGE, IN_POPULATION, IN_OUTCOMES AND IN_LEGEND. FIX THE COLORS IN THE FUNCTION.
+23) Plot outcomes for each age and sex cohort using created functions. Two functions are used: plot_output (plots outputs) and grid_arrange_shared_legend (places one label per page).
 
@@
 ## Do a loop, the first in_outcomes is age.
-
-require(ggplot2)
-require(gridExtra)
-require(grid)
-
-grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, position = c("bottom", "right"), mainTitle = "", mainLeft = "", mainBottom = "") {
-
-  plots <- list(...)
-  position <- match.arg(position)
-  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
-  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
-  lheight <- sum(legend$height)
-  lwidth <- sum(legend$width)
-  gl <- lapply(plots, function(x) x + theme(legend.position="none"))
-  gl <- c(gl, ncol = ncol, nrow = nrow, top = mainTitle, left = mainLeft, bottom = mainBottom)
-
-  combined <- switch(position,
-                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
-                                            legend,
-                                            ncol = 1,
-                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
-                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
-                                           legend,
-                                           ncol = 2,
-                                           widths = unit.c(unit(1, "npc") - lwidth, lwidth)))
-
-  #grid.newpage()
-  grid.draw(combined)
-
-  # return gtable invisibly
-  invisible(combined)
-
-}
-
-g_legend <- function(a.gplot){
-  tmp <- ggplot_gtable(ggplot_build(a.gplot))
-  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
-  legend <- tmp$grobs[[leg]]
-  return(legend)
-}
 
@@
 
+            # There are 4 diseases for males, this should be modified according to the modelled 
+            # diseases.
+
             if (sex == "male"){
@@
                 pdf(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".pdf"))
-                grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(age, sex, outcome),
-                                          mainLeft = 'Cases', mainBottom = 'Age')
+                grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(age, sex, outcome),mainLeft = 'Cases', mainBottom = 'Age')
                 dev.off()
@@
             }
+
+            # There are 5 diseases for females, this should be modified according to the modelled 
+            # diseases.
 
@@
                 pdf(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".pdf"))
-                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(age, sex, outcome),
-                                            mainLeft = 'Cases', mainBottom = 'Age')
+                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(age, sex, outcome), mainLeft = 'Cases', mainBottom = 'Age')
                 dev.off()
@@
 
-## Use plot_grid in a loo/grid.arrange
-
-# mpage_male <- do.call(marrangeGrob, list(grobs=p_list_male, nrow = 2, ncol = 2))
-# mpage_male
-# 
-# 
-# lay <- rbind(c(1, 2),
-#              c(3, 4),
-#              c(5, NA))
-# 
-# 
-# mpage_female <- do.call(marrangeGrob, list(grobs=p_list_female, layout_matrix = lay))
-# mpage_female
-
-## Trying to use ggarrange
-# mpage2 <- ggarrange(plotlist = p_list, ncol = 3, nrow = 3)
-
-# ggsave("graphs/multipage.jpeg", mpage)
-
-#Check ggarrange in ggpubr
-
-#http://www.sthda.com/english/wiki/print.php?id=177
 ```

### 2018-05-02 MSLT/code/functions.R

@@ -425,6 +425,38 @@ gen_aggregate <- function(in_data, in_cohorts, in_population, in_outcomes){
 }
 
 
+#####################################Function to generate combined labels for multiple plots in a page####
+
+grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, position = c("bottom", "right"), mainTitle = "", mainLeft = "", mainBottom = "") {
+  
+  plots <- list(...)
+  position <- match.arg(position)
+  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
+  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
+  lheight <- sum(legend$height)
+  lwidth <- sum(legend$width)
+  gl <- lapply(plots, function(x) x + theme(legend.position="none"))
+  gl <- c(gl, ncol = ncol, nrow = nrow, top = mainTitle, left = mainLeft, bottom = mainBottom)
+  
+  combined <- switch(position,
+                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
+                                            legend,
+                                            ncol = 1,
+                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
+                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
+                                           legend,
+                                           ncol = 2,
+                                           widths = unit.c(unit(1, "npc") - lwidth, lwidth)))
+  
+  
+  #grid.newpage()
+  grid.draw(combined)
+  
+  # return gtable invisibly
+  invisible(combined)
+  
+}
+
 #######################################Function to generate GBD graphs to compare data national to local (USED in GBD COMPARE############################
 
 

### 2018-05-02 MSLT/MSLTdoc.Rmd
Tidy graph labels

@@
 require(ggpubr)
+require(grid)
+require(ggplot2)
 
@@
 
-23) Plot outcomes for each age and sex cohort using created functions. Two functions are used: plot_output (plots outputs) and grid_arrange_shared_legend (places one label per page).
+23) Plot outcomes for each age and sex cohort using created functions.
+Founr functions are used: plot_output (plots outputs), grid_arrange_shared_legend and g_legend (places one label per page) and get_qualified_disease_name (replaces full name of disease)
 
 ```{r echo=TRUE}
-## Do a loop, the first in_outcomes is age.
 
@@
 
-            p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_disease = disease)
-
-            # There are 4 diseases for males, this should be modified according to the modelled 
-            # diseases.
+            p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_disease = get_qualified_disease_name(disease))
+
+                      #                 paste0(if(disease == "ihd") {
+                      # disease == "Ischemic heart disease"
+                      # } else {
+                      # if (disease == "istroke") {
+                      # disease == "Ischemic stroke"
+                      # } else {
+                      # if (disease == "colon_cancer") {
+                      # disease == "Colon cancer"
+                      # } else {
+                      # if (disease == "diabetes") {
+                      # disease == "Diabetes"
+                      # } else {
+                      # if (disease == "breast_cancer") {
+                      # disease == "Breast Cancer"}}}}}))
 
@@
                 pdf(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".pdf"))
-                grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(age, sex, outcome),mainLeft = 'Cases', mainBottom = 'Age')
+                grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age),
+                                          mainLeft = 'Cases', mainBottom = 'Age')
                 dev.off()
@@
             }
-
-            # There are 5 diseases for females, this should be modified according to the modelled 
-            # diseases.
 
@@
                 pdf(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".pdf"))
-                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(age, sex, outcome), mainLeft = 'Cases', mainBottom = 'Age')
+                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age), mainLeft = 'Cases', mainBottom = 'Age')
                 dev.off()
@@
 }
+
+## Insert the plots in the rmarkdown doc
 

### 2018-05-02 MSLT/MSLTdoc.Rmd
Included age and sex cohort graphs in pdf

@@
 
-                      #                 paste0(if(disease == "ihd") {
-                      # disease == "Ischemic heart disease"
-                      # } else {
-                      # if (disease == "istroke") {
-                      # disease == "Ischemic stroke"
-                      # } else {
-                      # if (disease == "colon_cancer") {
-                      # disease == "Colon cancer"
-                      # } else {
-                      # if (disease == "diabetes") {
-                      # disease == "Diabetes"
-                      # } else {
-                      # if (disease == "breast_cancer") {
-                      # disease == "Breast Cancer"}}}}}))
-
             if (sex == "male"){
@@
 
-                pdf(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".pdf"))
+                jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
                 grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age),
@@
 
-                pdf(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".pdf"))
+               jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
                 grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age), mainLeft = 'Cases', mainBottom = 'Age')
@@
 
-## Insert the plots in the rmarkdown doc
-
+```
+
+```{r echo=FALSE, out.width='100%'}
+## Loop to include graphs in the document
+graphs_doc <- list()
+index <- 1
+for (age in i_age_cohort) {
+      for (sex in i_sex)  {
+        for (outcome in i_outcome) {
+
+graphs_doc [[index]] <- c(paste("graphs/",age,"_",sex,"_", outcome,".jpeg", sep = ""))
+          knitr::include_graphics(graphs_doc [[index]])
+          index <- index + 1
+        }
+    }
+}
+ knitr::include_graphics(graphs_doc[c(1:64)])
 ```

### 2018-05-08 MSLT/MSLTdoc.Rmd
Checked total Lx and LXw calcs

@@
 ## Uncommnet to check life table list
-# View(general_life_table_list_bl[[2]])
+# View(general_life_table_list_bl[[32]])
 
@@
 i_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
-i_sex <- c("male", "female")
-i_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
 
@@
         # cat("age ", age, " sex ", sex, "and disease", disease, "\n") #Uncomment to see list
-        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, 
-                                               in_disease = disease)
+        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, in_disease = disease)
         index <- index + 1
@@
 # View(incidence_sc[[1]])
+
 ```
@@
         disease_life_table_list_sc[[index]]$diff_inc_disease <- 
-                                              disease_life_table_list_sc[[index]]$incidence_disease 
-                                               -   disease_life_table_list_bl[[index]]$incidence_disease
+                                              disease_life_table_list_sc[[index]]$incidence_disease -   disease_life_table_list_bl[[index]]$incidence_disease
         disease_life_table_list_sc[[index]]$diff_prev_disease <- 
-                                              disease_life_table_list_sc[[index]]$px 
-                                              - disease_life_table_list_bl[[index]]$px
+                                              disease_life_table_list_sc[[index]]$px  - disease_life_table_list_bl[[index]]$px
         disease_life_table_list_sc[[index]]$diff_mort_disease <- 
-                                              disease_life_table_list_sc[[index]]$mx 
-                                              - disease_life_table_list_bl[[index]]$mx
+                                              disease_life_table_list_sc[[index]]$mx - disease_life_table_list_bl[[index]]$mx
         disease_life_table_list_sc[[index]]$diff_pylds_disease <- 
-                                              (disease_life_table_list_sc[[index]]$px  
-                                              - disease_life_table_list_bl[[index]]$px) * 
-                                              disease_life_table_list_bl[[index]]$dw_disease
+                                              (disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px) * disease_life_table_list_bl[[index]]$dw_disease
         index <- index + 1
@@
 ## Generate total change in mortality rate
-## Sum mortality rate scenarios (mx_sc_total)
+
+## Sum mortality rate change scenarios (mx_sc_total)
 
@@
     mx_sc_total[[l_index]] <- mortality_sum
+
     l_index <- l_index + 1
@@
 ## Uncommnet to check sceanrio mortality and changes 
-# View(mx_sc_total[[1]])
+# View(mx_sc_total[[3]])
 
@@
   for (sex in i_sex){
-    # cat("age ", age, " and sex ", sex, "\n")
+
+    cat("age ", age, " and sex ", sex, "\n")
     # modify idata's mortality and pyld total for the said scenario
     td2 <- input_life_table
-    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx 
-                                                            + mx_sc_total[[index]]$total
-    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate 
-                                                                    + pylds_sc_total[[index]]$total
+    # td2 <- subset(td2, select = -c(mx, pyld_rate))
+    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx + mx_sc_total[[index]]$total
+    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate + pylds_sc_total[[index]]$total
 
@@
     general_life_table_list_sc[[index]] <- run_life_table(in_idata = td2, in_sex = sex, in_mid_age = age)
+    #
 
@@
 ## Uncommnet to check scenario life tables
-# View(general_life_table_list_sc[[1]])
-# View(general_life_table_list_bl[[2]])
+# View(general_life_table_list_sc[[32]])
+# View(general_life_table_list_bl[[1]])
+
+## Check difference life table baseline and scenario
+general_life_table_list_bl[[1]]$Lx - general_life_table_list_sc[[1]]$Lx
+general_life_table_list_bl[[1]]$Lwx - general_life_table_list_sc[[1]]$Lwx
 
@@
     ## general_life_table_list_sc and general_life_table_list_bl (Lx)
-    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], 
-                          c('Lx', 'Lwx'))
-    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- 
-                          paste('Lx', "sc", sep = "_")
-    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- 
-                          paste('Lwx', "sc", sep = "_")
-
-    output_burden_lf_bl <- select(general_life_table_list_bl[[l_index]], 
-                          c('Lx', 'Lwx'))
-    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lx'] <- 
-                          paste('Lx', "bl", sep = "_")
-    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lwx'] <- 
-                          paste('Lwx', "bl", sep = "_")
-
-    output_burden_lf_sc$Lx_diff <- general_life_table_list_bl[[l_index]]$Lx 
-                                - general_life_table_list_sc[[l_index]]$Lx
-    output_burden_lf_sc$Lwx_diff <- general_life_table_list_bl[[l_index]]$Lwx 
-                                - general_life_table_list_sc[[l_index]]$Lwx
+    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], c('Lx', 'Lwx'))
+    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- paste('Lx', "sc", sep = "_")
+    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- paste('Lwx', "sc", sep = "_")
+
+    output_burden_lf_bl <- select(general_life_table_list_bl[[l_index]], c('Lx', 'Lwx'))
+    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lx'] <- paste('Lx', "bl", sep = "_")
+    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lwx'] <- paste('Lwx', "bl", sep = "_")
+
+    output_burden_lf_sc$Lx_diff <- general_life_table_list_bl[[l_index]]$Lx - general_life_table_list_sc[[l_index]]$Lx
+        output_burden_lf_sc$Lwx_diff <- general_life_table_list_bl[[l_index]]$Lwx - general_life_table_list_sc[[l_index]]$Lwx
 
@@
 ## Uncomment to check
-# View(output_burden[[32]])
+# View(output_burden[[1]])
 
@@
 
-                jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
+  tiff(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".tiff"))
   grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age),
@@
 
-               jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
+               tiff(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".tiff"))
                 grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age), mainLeft = 'Cases', mainBottom = 'Age')
@@
 
-graphs_doc [[index]] <- c(paste("graphs/",age,"_",sex,"_", outcome,".jpeg", sep = ""))
+graphs_doc [[index]] <- c(paste("graphs/",age,"_",sex,"_", outcome,".tiff", sep = ""))
           knitr::include_graphics(graphs_doc [[index]]) 
+          # %>%
+          #   set_names(paste("Change in", ifelse(outcome == "mx", "Deaths", "Incidence"), "over time", sep = ""))
+
           index <- index + 1
@@
 }
+```
+
+```{r, echo=FALSE}
+## fig.cap=paste(names(graphs_doc)), message=FALSE, 
  knitr::include_graphics(graphs_doc[c(1:64)])
@@
 ```{r, echo=TRUE}
-####Generate data frame with all outputs to graph total change in burden by simulation year. 
-###first, need to run function for males and females separetly, in_cohorts indicates the number of age cohorts to 
-####include. To show all select 16. what is specified in in_outcomes will be graphed or can also be presented as a total. 
+## Generate data frame with all outputs to graph total change in burden by simulation year. 
+## First, need to run function for males and females separetly, in_cohorts indicates the number of age cohorts to include.
+To show all select 16.
+what is specified in in_outcomes will be graphed or can also be presented as a total.
+
+## DO A LIST TO GENERATE ALL OUTCOMES ALL. CHECK SUMMARION OF COHORTS ORDER.
+
 aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))
@@
 
-#####The following adds up both males and females
-# Remove non-numeric columns starting with age and sex
+## The following adds up both males and females
+## Remove non-numeric columns starting with age and sex
 aggregate_frame_males <- aggregate_frame_males %>% select(-starts_with("age"), -starts_with("sex"))
@@
 
-# Create a copy of aggregate_frame_females
+## Create a copy of aggregate_frame_females
 total_aggr <- aggregate_frame_females
-# Add aggregate_frame_males values to it
+## Add aggregate_frame_males values to it
 for (i in 1:ncol(aggregate_frame_females)){
@@
 total_aggr$sim_year <- seq.int(nrow(total_aggr))
+
+## Uncomment to view aggregated data
+sum(total_aggr$total_Lx_diff)
 ```
@@
 25) Plots using total aggregated change in burden.
+This graphs are not very good for life years and health adjusted life years as the difference between baseline and scenario is very small relative to the total, hence the graphs show almost no change.
 
@@
 
-disease_outcome <- "Health adjusted life years"
+disease_outcome <- "Life years"
 
@@
   scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
-  theme(plot.title = element_text(hjust = 0.5))
+  theme(plot.title = element_text(hjust = 0.5)) +
+  scale_y_continuous(breaks=seq(0,7000000,by=1000000))
 

### 2018-05-08 MSLT/code/model.R

@@ -8,9 +8,8 @@ rm (list = ls())
 
 options(scipen=999)
 
-##### Load all functions/packages
+##### Set WD
 
-source("MSLT/code/functions.R")
 
 
 ##### Read data. idata: life table and disease life tables (including trends). edata: exposure data (e.g. physical activity).

### 2018-05-08 MSLT/MSLTdoc.Rmd
Graphs in pdf

@@
 
-## Save plots to tiff. Makes a separate file for each plot.
+## Save plots to jpeg. Makes a separate file for each plot.
 
@@
   for (measure_index in i_measure) {
-    file_name = paste("graphs/", "Interpolation rates", sex_index, measure_index, ".tiff", sep=" ")
-    tiff(file_name)
+    file_name = paste("graphs1/", "Interpolation rates", sex_index, measure_index, ".jpeg", sep=" ")
+    jpeg(file_name)
     print(p_interpolation_list[[interpolation_index]])
@@
 
-17) Generate **baseline disease life tables** using run_disase function.
+17) Generate **baseline disease life tables** using run_disease function.
 Inputs of the disease life tables are: incidence, case fatality (Figure 3, purple shaded area) and disability weights.
@@
 23) Plot outcomes for each age and sex cohort using created functions.
-Founr functions are used: plot_output (plots outputs), grid_arrange_shared_legend and g_legend (places one label per page) and get_qualified_disease_name (replaces full name of disease)
+Four functions are used: plot_output (plots outputs), grid_arrange_shared_legend and g_legend (places one label per page) and get_qualified_disease_name (replaces full name of disease)
 
@@
 
-output_dir = "graphs"
+output_dir = "graphs2"
 
@@
 
-  tiff(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".tiff"))
+  jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
   grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age),
@@
 
-               tiff(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".tiff"))
+               jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
                 grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age), mainLeft = 'Cases', mainBottom = 'Age')
@@
 ## Loop to include graphs in the document
+
+## MAY NOT NEED THIS ONE
 graphs_doc <- list()
@@
 
-graphs_doc [[index]] <- c(paste("graphs/",age,"_",sex,"_", outcome,".tiff", sep = ""))
+graphs_doc [[index]] <- c(paste(output_dir, "/",age,"_",sex,"_", outcome,".jpeg", sep = ""))
 knitr::include_graphics(graphs_doc [[index]])
-          # %>%
-          #   set_names(paste("Change in", ifelse(outcome == "mx", "Deaths", "Incidence"), "over time", sep = ""))
 
@@
         }
-    }
-}
+
+      }
+}
+
+ knitr::include_graphics(graphs_doc[c(1:64)])
 ```
@@
 ```{r, echo=FALSE}
-## fig.cap=paste(names(graphs_doc)), message=FALSE, 
- knitr::include_graphics(graphs_doc[c(1:64)])
+## THIS IS AN ERROR WHEN TRYING TO KNIT THE DOC, PROBLEM WITH THE FILE PATH.
+
+# knitr::include_graphics(graphs_doc[[c(1:64)]])
+
+# image <- graphs_doc[[1]]
+# knitr::include_graphics(image)
+# cat("\n\n")
+# for (i in 2:64){
+#   cat("![](",graphs_doc[[i]],")")
+# }
+
+# 
+# myimages <- list.files("graphs2/", pattern = ".jpeg", full.names = TRUE)
+# include_graphics(myimages)
+
+#for (i in 1:64){
+#  knitr::include_graphics(graphs_doc[[i]]) 
+#}
 ```
@@
 ## DO A LIST TO GENERATE ALL OUTCOMES ALL. CHECK SUMMARION OF COHORTS ORDER. 
-
-aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))
-
-aggregate_frame_females <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))
-
-## The following adds up both males and females
-## Remove non-numeric columns starting with age and sex
-aggregate_frame_males <- aggregate_frame_males %>% select(-starts_with("age"), -starts_with("sex"))
-
-aggregate_frame_females <- aggregate_frame_females %>% select(-starts_with("age"), -starts_with("sex"))
-
-## Create a copy of aggregate_frame_females
-total_aggr <- aggregate_frame_females
-## Add aggregate_frame_males values to it
-for (i in 1:ncol(aggregate_frame_females)){
-  total_aggr[i] <- total_aggr[i] + aggregate_frame_males[i]
-}
-total_aggr$sim_year <- seq.int(nrow(total_aggr))
-
-## Uncomment to view aggregated data
-sum(total_aggr$total_Lx_diff)
+## Generate vectors with names of outcomes of interest: baseline, scenario and difference (from output dataframe)
+
+i_output_bl <- c("inc_num_bl_ihd", "inc_num_bl_istroke", "inc_num_bl_diabetes", "inc_num_bl_colon_cancer", "inc_num_bl_breast_cancer", "mx_num_bl_ihd", "mx_num_bl_istroke", "mx_num_bl_diabetes", "mx_num_bl_colon_cancer", "mx_num_bl_breast_cancer", "Lx_bl", "Lwx_bl")
+
+i_output_sc <- c("inc_num_sc_ihd", "inc_num_sc_istroke", "inc_num_sc_diabetes", "inc_num_sc_colon_cancer", "inc_num_sc_breast_cancer", "mx_num_sc_ihd", "mx_num_sc_istroke", "mx_num_sc_diabetes", "mx_num_sc_colon_cancer", "mx_num_sc_breast_cancer", "Lx_sc", "Lwx_sc")
+
+i_output_diff <- c("inc_num_diff_ihd", "inc_num_diff_istroke", "inc_num_diff_diabetes", "inc_num_diff_colon_cancer", "inc_num_diff_breast_cancer", "mx_num_diff_ihd", "mx_num_diff_istroke", "mx_num_diff_diabetes", "mx_num_diff_colon_cancer", "mx_num_diff_breast_cancer", "Lx_diff", "Lwx_diff")
+
+## SEPARATE LOOPS FOR MALES AND FEMALES
+aggregate_frame_males <- list()
+aggregate_frame_females <- list()
+index <- 1
+
+for (output_bl in i_output_bl) {
+  for (output_sc in i_output_sc) {
+    for (output_diff in i_output_diff) {
+
+aggregate_frame_males[[index]]<- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c(paste(output_bl), paste(output_sc), paste(output_diff)))
+
+# aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c(output_bl, output_sc, output_diff))
+
+aggregate_frame_females[[index]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c(paste(output_bl), paste(output_sc), paste(output_diff)))
+
+# aggregate_frame_females <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c(output_bl, output_sc, output_diff))
+
+index <- index + 1
+    }
+  }
+}
+
+# ## The following adds up both males and females
+# ## Remove non-numeric columns starting with age and sex
+# aggregate_frame_males <- aggregate_frame_males %>% select(-starts_with("age"), -starts_with("sex"))
+# 
+# aggregate_frame_females <- aggregate_frame_females %>% select(-starts_with("age"), -starts_with("sex"))
+# 
+# ## Create a copy of aggregate_frame_females
+# total_aggr <- aggregate_frame_females
+# ## Add aggregate_frame_males values to it
+# for (i in 1:ncol(aggregate_frame_females)){
+#   total_aggr[i] <- total_aggr[i] + aggregate_frame_males[i]
+# }
+# total_aggr$sim_year <- seq.int(nrow(total_aggr))
+# 
+# ## Uncomment to view aggregated data
+# sum(total_aggr$total_Lx_diff)
 ```
@@
 
-```{r, echo=TRUE}
-
-####This plot has to be customised to in_outcomes, here, only totals shown, but specifications are up to the user
-
-####[] is used here to indicate the number of simulation years into the future. 
-####Disease outcomes has to be changed to the outcome of interest
-
-disease_outcome <- "Life years"
-
-total_plot <- ggplot(total_aggr[1:79,], aes(x = sim_year)) +
-
-  geom_line(mapping = aes(y = total_Lx_bl, colour = "total_Lx_bl")) +
-  theme_classic() +
-  geom_hline(yintercept=0, linetype="dashed", color = "black") +
-  geom_line(mapping = aes(y = total_Lx_sc, colour = "total_total_Lx_sc")) +
-  geom_line(mapping = aes(y = total_Lx_diff, colour = "total_Lx_diff")) +
-  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
-  theme(plot.title = element_text(hjust = 0.5, size = 12)) +
-  scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  scale_y_continuous(breaks=seq(0,7000000,by=1000000))
-
-#Print to view
-print(total_plot)
+# ```{r, echo=TRUE}
+# 
+# ####This plot has to be customised to in_outcomes, here, only totals shown, but specifications are up to the user
+# 
+# ####[] is used here to indicate the number of simulation years into the future. 
+# ####Disease outcomes has to be changed to the outcome of interest
+# 
+# disease_outcome <- "Life years"
+# 
+# total_plot <- ggplot(total_aggr[1:79,], aes(x = sim_year)) +
+#   
+#   geom_line(mapping = aes(y = total_Lx_bl, colour = "total_Lx_bl")) +
+#   theme_classic() +
+#   geom_hline(yintercept=0, linetype="dashed", color = "black") +
+#   geom_line(mapping = aes(y = total_Lx_sc, colour = "total_total_Lx_sc")) +
+#   geom_line(mapping = aes(y = total_Lx_diff, colour = "total_Lx_diff")) +
+#   xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
+#   theme(plot.title = element_text(hjust = 0.5, size = 12)) +
+#   scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
+#   theme(plot.title = element_text(hjust = 0.5)) +
+#   scale_y_continuous(breaks=seq(0,7000000,by=1000000))
+#   
+#   
+# 
+# #Print to view
+# print(total_plot)
 ```

### 2018-05-08 MSLT/MSLTdoc.Rmd
tried a loop of aggregated outcomes, did not work

--- 2018-05-08 MSLT/MSLTdoc.Rmd
@@
         }
-    }
-}
-```
-
-```{r, echo=FALSE}
-## fig.cap=paste(names(graphs_doc)), message=FALSE, 
+
+      }
+}
+
  knitr::include_graphics(graphs_doc[c(1:64)])
@@
 24) Add up all outcomes per year of simulation with generated function.
+BETTER TO HAVE A LOOP, I TRIED BY LOOPING IN THE FUNCITON ARGUMENT I_OUTCOMES, BUT THIS DID NOT WORK.
+THEN, IF LOOP HERE ALSO IN NEXT CODE.
 
@@
 
-aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))
-
-aggregate_frame_females <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c('Lx_bl','Lx_sc', "Lx_diff"))
+aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('inc_num_bl_ihd','inc_num_sc_ihd', "inc_num_diff_ihd"))
+
+aggregate_frame_females <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c('inc_num_bl_ihd','inc_num_sc_ihd', "inc_num_diff_ihd"))
 
@@
 total_aggr$sim_year <- seq.int(nrow(total_aggr))
-
-## Uncomment to view aggregated data
-sum(total_aggr$total_Lx_diff)
 ```
@@
 
-disease_outcome <- "Life years"
+disease_outcome <- "Incidence of Ischemic Heart Disease"
 
@@
 
-  geom_line(mapping = aes(y = total_Lx_bl, colour = "total_Lx_bl")) +
+  geom_line(mapping = aes(y = total_inc_num_bl_ihd, colour = "total_inc_num_bl_ihd")) +
   theme_classic() +
   geom_hline(yintercept=0, linetype="dashed", color = "black") +
-  geom_line(mapping = aes(y = total_Lx_sc, colour = "total_total_Lx_sc")) +
-  geom_line(mapping = aes(y = total_Lx_diff, colour = "total_Lx_diff")) +
+  geom_line(mapping = aes(y = total_inc_num_sc_ihd, colour = "total_inc_num_sc_ihd")) +
+  geom_line(mapping = aes(y = total_inc_num_diff_ihd, colour = "total_inc_num_diff_ihd")) +
   xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
@@
   scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  scale_y_continuous(breaks=seq(0,7000000,by=1000000))
+  theme(plot.title = element_text(hjust = 0.5))
 
@@
 print(total_plot)
-```
+
+```
+
 #Comments

### 2018-05-09 MSLT/MSLTdoc.Rmd
Added loop to generate dataframe with total outcomes

@@
 
+# View(output_df)
+
 #Remove variables that are not used in the generation of outputs. CHANGE THIS NAMES, TOO LONG
@@
 
-## MAY NOT NEED THIS ONE
 graphs_doc <- list()
@@
 ```{r, echo=TRUE}
-## Generate data frame with all outputs to graph total change in burden by simulation year. 
-## First, need to run function for males and females separetly, in_cohorts indicates the number of age cohorts to include.
-To show all select 16.
-what is specified in in_outcomes will be graphed or can also be presented as a total.
-
-## DO A LIST TO GENERATE ALL OUTCOMES ALL. CHECK SUMMARION OF COHORTS ORDER.
-
-aggregate_frame_males <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c('inc_num_bl_ihd','inc_num_sc_ihd', "inc_num_diff_ihd"))
-
-aggregate_frame_females <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c('inc_num_bl_ihd','inc_num_sc_ihd', "inc_num_diff_ihd"))
-
-## The following adds up both males and females
-## Remove non-numeric columns starting with age and sex
-aggregate_frame_males <- aggregate_frame_males %>% select(-starts_with("age"), -starts_with("sex"))
-
-aggregate_frame_females <- aggregate_frame_females %>% select(-starts_with("age"), -starts_with("sex"))
-
-## Create a copy of aggregate_frame_females
-total_aggr <- aggregate_frame_females
+
+## Try binding function
+
+aggregate_frame_males <- list()
+aggregate_frame_females <- list()
+
+index <- 1
+
+for (outcome in i_outcome) {
+  for (disease in i_disease) {
+
+aggregate_frame_males[[index]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c(paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")))
+
+aggregate_frame_females[[index]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c(paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")))
+
+# Remove non-numeric columns starting with age and sex
+aggregate_frame_males[[index]] <- aggregate_frame_males[[index]] %>% select(-starts_with("age"), -starts_with("sex"))
+
+aggregate_frame_females[[index]] <- aggregate_frame_females[[index]] %>% select(-starts_with("age"), -starts_with("sex"))
+
+index <- index + 1
+  }
+}
+
+## Loop for life years (Lx) and health adjusted life years (Lwx) to then add to total aggregated data.
+
+## RUN ANOTHER LOOP FOR LX AND LWX
+
+i_outcome2 <- c("Lx", "Lwx")
+
+aggregate_frame_males2 <- list()
+aggregate_frame_females2 <- list()
+
+for (i in i_outcome2){
+
+aggregate_frame_males2[[i]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c(paste(i, "bl", sep = "_"), paste(i, "sc", sep = "_"), paste(i, "diff",sep = "_")))
+
+aggregate_frame_females2[[i]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c(paste(i, "bl", sep = "_"), paste(i,  "sc", sep = "_"), paste(i, "diff",sep = "_")))
+
+aggregate_frame_males2[[i]] <- aggregate_frame_males2[[i]] %>% select(-starts_with("age"), -starts_with("sex"))
+
+aggregate_frame_females2[[i]] <- aggregate_frame_females2[[i]] %>% select(-starts_with("age"), -starts_with("sex"))
+
+}
+
+## Uncomment to check data frames (list correspond to a disease and outcome combination, for example,
+## incidence breast cancer)
+
+# View(aggregate_frame_females[[2]])
+
+## Transform list to data frame (this is not working after the first disease)
+
+aggregate_frame_females <- do.call(cbind, aggregate_frame_females)
+aggregate_frame_males <- do.call(cbind, aggregate_frame_males)
+aggregate_frame_females2 <- do.call(cbind, aggregate_frame_females2)
+aggregate_frame_males2 <- do.call(cbind, aggregate_frame_males2)
+
+View(aggregate_frame_females2)
+
+## Drop ending of variables names _males/_females to add up all outcomes in next step.
+
+names(aggregate_frame_females) = gsub(pattern = "_female", replacement = "", x = names(aggregate_frame_females))
+
+names(aggregate_frame_males) = gsub(pattern = "_male", replacement = "", x = names(aggregate_frame_males))
+
+names(aggregate_frame_females2) = gsub(pattern = "_female", replacement = "", x = names(aggregate_frame_females2))
+
+names(aggregate_frame_males2) = gsub(pattern = "_male", replacement = "", x = names(aggregate_frame_males2))
+
+## Uncomment to check that the code is dropping the male/female ending. 
+# View(aggregate_frame_males)
+
+## Create a copy of aggregate_frame_females.  
+total_aggr1 <- aggregate_frame_females
 ## Add aggregate_frame_males values to it
 for (i in 1:ncol(aggregate_frame_females)){
-  total_aggr[i] <- total_aggr[i] + aggregate_frame_males[i]
-}
+  total_aggr1[i] <- total_aggr1[i] + aggregate_frame_males[i]
+}
+
+## Add data frames 2 with life years (check adds Lx and Lwx at the beginning of the variable name)
+total_aggr2 <- aggregate_frame_females2
+## Add aggregate_frame_males values to it
+for (i in 1:ncol(aggregate_frame_females2)){
+  total_aggr2[i] <- total_aggr2[i] + aggregate_frame_males2[i]
+}
+
+## Combine data frames
+
+total_aggr <- cbind.data.frame(total_aggr1, total_aggr2)
+
 total_aggr$sim_year <- seq.int(nrow(total_aggr))
+
+## Uncomment to see total data frame
+View(total_aggr)
+
+## Test that total_aggr is adding males and females dataframes
+# 
+# (aggregate_frame_females2$Lx.Lx_bl_22 + aggregate_frame_males2$Lx.Lx_bl_22) -total_aggr$Lx.Lx_bl_22
+
 ```
@@
 
-####This plot has to be customised to in_outcomes, here, only totals shown, but specifications are up to the user
+####This plot has to be customised to in_outcomes, here, only totals shown, but specifications are up to the user. ADD LOOP for all outcomes over time and total TABLE.
 
@@
 
+## ADD LOOP
 disease_outcome <- "Incidence of Ischemic Heart Disease"

### 2018-05-09 MSLT/MSLTdoc.Rmd
Loop graphs

@@
     }
-
   }
@@
 
-## ADD LOOP
+#### Test code with loops for aggregated outcomes diseases burden. NOT WORKING.
+
+### Compare with loops for age and sex cohort outcomes.
+
+p_aggr_list <- list()
+index <- 1
+
+for (outcome in i_outcome) {
+  for (disease in i_disease) {
+
+p_aggr_list_index <- ggplot(total_aggr[1:79,], aes(x = sim_year)) +
+
+  geom_line(mapping = aes(y = paste("total", outcome, "num_bl", disease, sep = "_"), colour = paste("total", outcome, "num_bl", disease, sep = "_"))) +
+  theme_classic() +
+  geom_hline(yintercept=0, linetype="dashed", color = "black") +
+  geom_line(mapping = aes(y = paste("total", outcome, "num_sc", disease, sep = "_"), colour = paste("total", outcome, "num_sc", disease, sep = "_"))) +
+  geom_line(mapping = aes(y = paste("total", outcome, "num_dif", disease, sep = "_"), colour = paste("total", outcome, "num_diff", disease, sep = "_"))) +
+  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease)) +
+  theme(plot.title = element_text(hjust = 0.5, size = 12)) +
+  scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
+  theme(plot.title = element_text(hjust = 0.5))
+p_aggr_list[[index]] <- p_aggr_list_index  
+index <- index + 1
+
+  }
+}
+
+#####
+## original code with single outcomes
 disease_outcome <- "Incidence of Ischemic Heart Disease"

### 2018-05-22 MSLT/MSLTdoc.Rmd
fixed path names in rmarkdown

--- 2018-05-09 MSLT/MSLTdoc.Rmd
@@
 author: "Belen Zapata-Diomedi and Ali Abbas"
-date: "26 March 2018"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 header-includes:
@@
    geometry: margin=1in
-bibliography: Bibliography.bib
+bibliography: documentation/Bibliography.bib
 ---
@@
 ## Global options
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, width = 75)
 options(knitr.table.format = 'latex')
-knitr::opts_chunk$set(fig.cap = TRUE, fig.align = "center", fig.pos = "H", fig.scap = TRUE)
+knitr::opts_chunk$set(fig.cap = FALSE, fig.align = "center", fig.pos = "H", fig.scap = TRUE)
 
@@
 
-![Basic ITHIMR infrastructure](structure\Figure1.png){fig.pos="H"}
+![Basic ITHIMR infrastructure](documentation\structure\Figure1.png){fig.pos="H"}
 
@@
 Changes in prevalence and mortality rates from the disease specific life tables feed into the life table by changing all-cause mortality, which in turn changes life years.
-Change in prevalence of diseases changes total years lived with disability, which in turn modifies health-adjusted life years*](structure\Figure2.png){width=100%, fig.pos="H"}
+Change in prevalence of diseases changes total years lived with disability, which in turn modifies health-adjusted life years*](documentation\structure\Figure2.png){width=100%, fig.pos="H"}
 
@@
 Black color coding with white color font represent final model outcomes.
-For both, the life table and disease life table, two sets of each are simulated, one for the baseline and the other for the scenario.*](structure\Figure3.png){width=50%,height=30%,fig.pos="H"}
+For both, the life table and disease life table, two sets of each are simulated, one for the baseline and the other for the scenario.*](documentation\structure\Figure3.png){width=50%,height=30%,fig.pos="H"}
 
@@
 
-![Global Burden of Disease data results tool.](data/Figure4.png){width=70%,height=50%,fig.pos="H"}
+![Global Burden of Disease data results tool.](documentation\data\Figure4.png){width=70%,height=50%,fig.pos="H"}
 
@@
 
-GBDdata <- read.csv("data/UK/gbd2016.csv", stringsAsFactors = F)
+GBDdata <- read.csv("data/legacy/UK/gbd2016.csv", stringsAsFactors = F)
 
@@
   for (measure_index in i_measure) {
-    file_name = paste("graphs1/", "Interpolation rates", sex_index, measure_index, ".jpeg", sep=" ")
+    file_name = paste("output/graphs1/", "Interpolation rates", sex_index, measure_index, ".jpeg", sep=" ")
     jpeg(file_name)
@@
 Note that Dismod II is an external software.
-The code below is used to generate the inputs for Dismod II using GBD data (see Figure 3, purple shaded area).
-Dismod II requires data for the *collection* (population numbers and all cause mortality rates) and *data set* (disease specific parameters).
-We will generate a data set with all inputs for dismod by 5 year-age (represent at mid age) groups and sex for five diseases (ischemic heart disease, ischemic stroke, diabetes mellitus, colon and rectum cancer and breast cancer-women).
-WE WILL NEED ADJUSTMENT FOR TO MATCH RRS TO DISEASES (COLON AND RECTUM CANCER AND DIABETES MELLITUS).
-*Code here is work in progress**, for now the disease input data is organised externally with excel.
+The data generation for Dismod II is now done externally and the end product, for the example of Greater London is idata.
+For now, we plan to get the data processing for all data for the MSLT and Dismod using code (space blank left below).
+We now have all the data needs in the main data folder in data.
 
 ```{r, echo=FALSE}
-## Collection input for Greater London: population numbers (sex and 5-year age groups). 
-## Already processed in past code, here, we sort it out and save in cvs format in a Dismod II folder.
-
-####THIS CODE IS INCOMPLETE, I WILL PREPARE THE DATA FOR DISMOD II IN EXCEL AND AFTER PRESENTATION THURSDAY 26 APRIL I WILL DEVELOP THE SCRIPT.
-
-# ## Get population numbers
-# 
-# data_dismodii <- GBD_population_GL[order(GBD_population_GL$sex, GBD_population_GL$age_cat),] 
-# 
-# ## Add mortality rates (these were generated for the life table, so we use the same code)
-# 
-# data_dismodii  <- merge(data_dismodii, select(male_deaths_interpolated
-#                                                      ,c(sex_age_cat, mx))
-#                                                      , by = 'sex_age_cat', all.x = TRUE)
-# data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
-#                                                      ,c(sex_age_cat, mx))
-#                                                      , by = 'sex_age_cat', all.x = TRUE)
-# 
-# data_dismodii$mx <-  ifelse(data_dismodii$sex == "male",
-# data_dismodii$mx.x, data_dismodii$mx.y)
-# ## Drop redundant variables 
-# data_dismodii <- subset(data_dismodii, select = -c(mx.y, mx.x))
-# ##Sort data
-# data_dismodii<-data_dismodii[order(data_dismodii$sex, data_dismodii$age_cat),] 
-# ## Add disease rates per one(incidence, prevalence and mortality)
-# 
-# ## Generate variable to do a loop that creates dataframes for each of the disease inputs.
-# 
-# i_disease <- c("ischemic heart disease", "ischemic stroke", "diabetes mellitus", "tracheal, bronchus, and lung cancer", "breast cancer", "colon and rectum cancer")
-# i_sex <- c("male", "female")
-# i_measure <- c("deaths", "ylds", "incidence", "prevalence")
-# 
-# ## Loop to generate dataframes
-# for(sex_index in i_sex) {
-#   for (measure_index in i_measure) {
-#     for (disease_index in i_disease) {
-#    
-# data <- filter(GBDGL, measure == measure_index, sex == sex_index, 
-# cause == disease_index) %>% select(measure, location, sex, age, metric, 
-# cause, val, age_cat, one_rate, sex_age_cat) 
-# colnames(data)[9] <- paste(measure_index, disease_index, sep =  "_")
-# assign(paste(sex_index, measure_index, disease_index, sep = "_"), data)
-
-## The code below is not generating the multiple datasets saved as csv.
-
-# data_name <- file.path(tempdir(), paste0(sex_index, measure_index, disease_index, ".csv"))
-# write.csv(data, file = data_name)
-#     }
-#   }
-# }
-
-## Generate TO DO, ADD disease parameters.
-
-## Generate empy variables in data_dismodii for death, incidence and prevalence rates per disease (MAY NOT NEED TO USE, WITH MERGING VARIABLES)
-
-#  for (measure_index in i_measure) {
-#     for (disease_index in i_disease) {
-# 
-#   # data_dismodii[,paste(measure_index, disease_index, sep =  "_")] <- 0
-# data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
-#                                                      ,c(sex_age_cat, mx))
-#                                                      , by = 'sex_age_cat', all.x = TRUE)
-#   }
-#  }
 
@@
 
-idata <- read.csv("data/UK/idata.csv", stringsAsFactors = F)
+idata <- read.csv("data/legacy/UK/idata.csv", stringsAsFactors = F)
 ## Use run_disease
@@
 
-output_dir = "graphs2"
+output_dir = "output/graphs2"
 
@@
 
-## ADD LOOP
-disease_outcome <- "Incidence of Ischemic Heart Disease"
-
-total_plot <- ggplot(total_aggr[1:79,], aes(x = sim_year)) +
-
-  geom_line(mapping = aes(y = total_inc_num_bl_ihd, colour = "total_inc_num_bl_ihd")) +
+#### Test code with loops for aggregated outcomes diseases burden. NOT WORKING.
+
+### Compare with loops for age and sex cohort outcomes.
+
+p_aggr_list <- list()
+index <- 1
+
+for (outcome in i_outcome) {
+  for (disease in i_disease) {
+    # outcome <- i_outcome[1]
+    # disease <- i_disease[1]
+
+p_aggr_list_index <- ggplot(total_aggr[1:79,], aes(x = total_aggr[["sim_year"]])) +
+
+  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_bl", disease, sep = "_")]], colour = paste("total", outcome, "num_bl", disease, sep = "_"))) +
   theme_classic() +
   geom_hline(yintercept=0, linetype="dashed", color = "black") +
-  geom_line(mapping = aes(y = total_inc_num_sc_ihd, colour = "total_inc_num_sc_ihd")) +
-  geom_line(mapping = aes(y = total_inc_num_diff_ihd, colour = "total_inc_num_diff_ihd")) +
-  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
+  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_sc", disease, sep = "_")]], colour = paste("total", outcome, "num_sc", disease, sep = "_"))) +
+  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_diff", disease, sep = "_")]], colour = paste("total", outcome, "num_diff", disease, sep = "_"))) +
+  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease, outcome)) +
   theme(plot.title = element_text(hjust = 0.5, size = 12)) +
@@
   theme(plot.title = element_text(hjust = 0.5))
-
-#Print to view
-print(total_plot)
+p_aggr_list[[index]] <- p_aggr_list_index  
+index <- index + 1
+
+  }
+}
+
+index <- 1
+
+interpolation_index <- 1
+for (outcome in i_outcome) {
+  for (disease in i_disease) {
+    file_name = paste("output/graphs2/", "Aggregated Outcomes", outcome, disease, ".jpeg", sep=" ")
+    jpeg(file_name)
+    print(p_aggr_list[[index]])
+    index <- index + 1
+    dev.off()
+  }
+}
+
+p_aggregated <- do.call(marrangeGrob, list(grobs=p_aggr_list, nrow = 2, ncol = 2))
+p_aggregated
 

### 2018-05-25 MSLT/MSLTdoc.Rmd
Separate code and markdown script. Also fix chunk numbers

@@
 
+```{r cache=FALSE}
+knitr::read_chunk('mslt_code.r')
+# source("code/functions.R")
+```
+
 1) Clean Global Environment
-```{r, echo=TRUE}
-rm (list = ls())
+```{r chunk-1}
 ```
 2) Avoid scientific notation
-```{r, echo=TRUE}
-options(scipen=999)
+```{r chunk-2}
 ```
 3) Load functions 
-```{r, echo=TRUE}
-source("code/functions.R")
+```{r chunk-3}
 ```
@@
 4) Read GBD data
-```{r, echo=TRUE}
-
-GBDdata <- read.csv("data/legacy/UK/gbd2016.csv", stringsAsFactors = F)
-
+```{r chunk-4}
 ```
@@
 5) Change all upper cases to lower case and delete () from variables.
-```{r, echo=TRUE}
-
-GBDdata <- mutate_all(GBDdata, funs(tolower))
-GBDdata$measure[GBDdata$measure=="ylds (years lived with disability)"] <- "ylds"
-
+```{r chunk-5}
 ```
@@
 6) Create age categories index in GBDdata (mid age, to match cohort running), total of 20 age groups. These are the age cohorts to simulate.
-```{r, echo=TRUE}
-
-# 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97
-GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
-GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
-GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
-GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
-GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
-GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
-GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
-GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
-GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
-GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
-GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
-GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
-GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
-GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
-GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
-GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
-GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
-GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
-GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
-GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97
-
+```{r chunk-6}
 ```
@@
 For now, the Life table is set up with population numbers derived from the GBD data.
-```{r, echo=TRUE}
+```{r chunk-7}
+```
 
-GBDdata$sex_age_cat <- paste(GBDdata$sex,GBDdata$age_cat, sep = "_"  )
+8) Convert string variables to numeric to do calculations.
+```{r chunk-8}
+```
 
-```
-8) Convert string variables to numeric to do calculations.
-```{r, echo=TRUE}
-
-GBDdata$val <- as.numeric(as.character(GBDdata$val))
-
-```
 9) Generate population numbers for Greater London in a new data frame ("GBD_population").
@@
 Population numbers are derived from rates per 100,000 and total numbers of cases.
-```{r, echo=TRUE}
-
-GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", 
-metric == "rate" | metric == "number" ) %>% select(metric, age_cat, sex, val, 
-sex_age_cat, location)
-
+```{r chunk-9}
 ```
 10) Generate population numbers from given number of cases and rates per 100,000 people. 
-```{r, echo=TRUE}
-
- for (i in 1:nrow(GBD_population)) {
-    if (GBD_population$metric[i] == "number") {
-      GBD_population$five_year_population[i] <- GBD_population$val[i] * 100000/
-      GBD_population$val[i + 2]}     
-    else {GBD_population$five_year_population[i] <- NA}
- }
-
+```{r chunk-10}
 ```
 11) Remove rows with zero
-```{r, echo=TRUE}
-
-GBD_population <- GBD_population[!is.na(GBD_population$five_year_population),]
-
+```{r chunk-11}
 ```
 12) Keep relevant variables
-```{r, echo=TRUE}
-
-GBD_population <- filter(GBD_population) %>% 
-select(sex_age_cat, sex, age_cat, five_year_population, location)
-
+```{r chunk-12}
 ```
 13) Create data frames for Greater London to be later used for: a) interpolation of rates and b) PMLT cohorts. 
-```{r, echo=TRUE}
-
-GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% 
-select(sex_age_cat, age_cat, sex, five_year_population, location, sex_age_cat)
-
-```
-12) Check population total numbers
-```{r, echo=TRUE}
-
-GreaterLondon <- sum(GBD_population_GL$five_year_population)
-
+```{r chunk-13}
 ```
 
-13) Generate data frames for Greater London with per person rates (per 100,000 in original data).  
-```{r, echo=TRUE}
-
-GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% 
-select(measure, location, sex, age, metric, cause, val, age_cat, sex_age_cat)
-GBDGL$one_rate <- GBDGL$val/100000
-
+14) Check population total numbers
+```{r chunk-14}
 ```
 
-14) For the life table, we need to mortality and total yld rates in one year age intervals.
+15) Generate data frames for Greater London with per person rates (per 100,000 in original data).  
+```{r chunk-15}
+```
+
+16) For the life table, we need to mortality and total yld rates in one year age intervals.
 The original data is in five years.
@@
 
-```{r, echo=TRUE}
-
-## Loop to generate interpolated rates for all cause mortality and ylds for males and females
-## UPDATE WITH YLDS RATES ALL CAUSES ADJUSTED FOR ALL OTHER MODELLED DISEASES.
-
-i_sex <- c("male", "female")
-i_measure <- c("deaths", "ylds") #" (years lived with disability)")
-
-index <- 1
-
-for(sex_index in i_sex) {
-  for (measure_index in i_measure) {
-
-data <- filter(GBDGL, measure == measure_index, sex == sex_index, 
-cause == "all causes") %>% select(measure, location, sex, age, metric, 
-cause, val, age_cat, one_rate) 
-assign(paste(sex_index, measure_index, "interpolated_data", sep = "_"), data)
-x=data$age_cat
-y=log(data$one_rate)
-
-interpolation_func <- splinefun(x, y, method = "natural", ties = mean)
-
-interpolated <- as.data.frame(interpolation_func(seq(0, 100, 1)))
-age <- seq(0,100,by=1)
-interpolated <- cbind(interpolated, age)
-interpolated[,1] <- exp(interpolated[,1])
-colnames(interpolated)[1] <- paste(measure_index)
-## Add column with sex to create age_sex category to then merge with input_life table
-interpolated$sex <- paste(sex_index)
-interpolated$sex_age_cat <- paste(interpolated$sex, interpolated$age, sep = "_")
-## Change name of column death to mx and ylds to pyld_rate to then merge 
-## with input_life table
-
-if (colnames(interpolated)[1] == "deaths")
-    colnames(interpolated)[1] <- paste("mx")
-  else
-    colnames(interpolated)[1] <- paste("pyld_rate")
-
-# Name data frame
-assign(paste(sex_index, measure_index, "interpolated", sep = "_"), interpolated)
-
-  }
-}
-
-## Do graph with another layer for original rates in age groups for comparison purposes.
-
-p_interpolation_list <- list()
-
-interpolation_index <- 1
-
-for(sex_index in i_sex) {
-  for (measure_index in i_measure) {
-
-p_interpolation_index <- ggplot(data = interpolated,mapping = aes(age, interpolated[,1])) +
-              geom_line(aes(color = "Interpolated")) +
-              geom_point(
-              data = data, 
-              mapping = aes(age_cat, one_rate, color = "Original")) +
-              labs(colour="",x="Age",y="Rates", sep = " ") +
-              labs (title = paste("Rates", sex_index, "all cause", 
-              measure_index, sep = " "), size=14) +
-              theme_classic() +
-              theme (plot.title = element_text(hjust = 0.5))
-
-              p_interpolation_list[[interpolation_index]] <- p_interpolation_index
-              interpolation_index <- interpolation_index + 1
-
-  }
-}
-
-## Save plots to jpeg. Makes a separate file for each plot.
-
-interpolation_index <- 1
-for(sex_index in i_sex) {
-  for (measure_index in i_measure) {
-    file_name = paste("output/graphs1/", "Interpolation rates", sex_index, measure_index, ".jpeg", sep=" ")
-    jpeg(file_name)
-    print(p_interpolation_list[[interpolation_index]])
-    interpolation_index <- interpolation_index + 1
-    dev.off()
-  }
-}
-
-p_interpolated <- do.call(marrangeGrob, list(grobs=p_interpolation_list, nrow = 2, ncol = 2))
-p_interpolated
-
+```{r chunk-16}
 ```
 
-15) The following code brings together all the required data to create the inputs life table (Figure 3).
+17) The following code brings together all the required data to create the inputs life table (Figure 3).
 
-```{r, echo=TRUE}
+```{r chunk-17}
+```
 
-## Create empty data frame
-
-input_life_table <- as.data.frame(matrix(0, ncol=2, nrow = 202))
-
-## Give variables names (mx=total mortality rate, 
-## yldsx = total years lived with disability rates )
-
-names(input_life_table) <- c("age", "sex")
-
-## Populate life table: Age, 0 to 100
-
-input_life_table[1:101, 1] <- c(0:100)
-input_life_table[102:202, 1] <- c(0:100)
-
-## Populate life table: sex, male, female
-
-input_life_table[1:101, 2] <- "male"
-input_life_table[102:202, 2] <- "female"
-
-## Create variable age_sex to match with population data
-
-input_life_table$sex_age_cat <- paste(input_life_table$sex,input_life_table$age,
-                                sep = "_"  )
-## Populate life table: five_year_pop.
-
-input_life_table <- merge(input_life_table, select(GBD_population_GL, 
-                    c(sex_age_cat, five_year_population)), by = 'sex_age_cat', all = TRUE)
-
-## Populate life table: mortality rates (from interpolated rates)
-## Males mortality rates. NEED MATCHING NAMES
-
-input_life_table <- merge(input_life_table, select(male_deaths_interpolated, 
-                    c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
-input_life_table <- merge(input_life_table, select(female_deaths_interpolated,
-                    c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
-input_life_table <- merge(input_life_table, select(male_ylds_interpolated, 
-                    c(sex_age_cat, pyld_rate)) , by = 'sex_age_cat', all.x = TRUE) 
-input_life_table <- merge(input_life_table, select(female_ylds_interpolated, 
-                    c(sex_age_cat, pyld_rate)), by = 'sex_age_cat', all.x = TRUE)
-
-## DISCUSS WITH ALI TO IMPROVE
-
-input_life_table$mx <-  ifelse(input_life_table$sex == "male",
-                        input_life_table$mx.x, input_life_table$mx.y)
-
-input_life_table$pyld_rate <-  ifelse(input_life_table$sex == "male",
-                              input_life_table$pyld_rate.x, input_life_table$pyld_rate.y)
-## Drop redundant variables
-input_life_table <- subset(input_life_table, select = -c(mx.y, mx.x, pyld_rate.x, pyld_rate.y))
-
-## Cross check population numbers
-
-sum(input_life_table$five_year_population, na.rm = TRUE)
-
-## Sort data by sex and age to use in life table function
-
-input_life_table<-input_life_table[order(input_life_table$sex, input_life_table$age),]
-
-```
-16) Generate **baseline life tables** from input_data_frame using run_life_table function. Life tables are generated for each age (5-years) and sex cohort. 
+18) Generate **baseline life tables** from input_data_frame using run_life_table function. Life tables are generated for each age (5-years) and sex cohort. 
 It is assumed that current observed rates of mortality and disability are going to be observed in the future.
 
-```{r, echo=TRUE}
-
-i_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
-
-general_life_table_list_bl <- list()
-
-index <- 1
-
-for (age in i_age_cohort){
-  for (sex in i_sex){
-    # cat("age ", age, " and sex ", sex, "\n") #Uncomment to see index
-    general_life_table_list_bl[[index]] <- run_life_table(in_idata = input_life_table,
-    in_sex = sex, in_mid_age = age)
-    index <- index + 1
-  }
-}
-
-## Uncommnet to check life table list
-# View(general_life_table_list_bl[[32]])
-
+```{r chunk-18}
 ```
 
-17) Generate **baseline disease life tables** using run_disease function.
+19) Generate **baseline disease life tables** using run_disease function.
 Inputs of the disease life tables are: incidence, case fatality (Figure 3, purple shaded area) and disability weights.
@@
 
-```{r, echo=FALSE}
+```{r chunk-19}
+```
+
+20) Generate mock change in incidence of disease to generate scenario life tables. In the final version, this will come from the calculated PIFs by disease. 
+PIFs are applied here to incidence, however, may also be applied to case fatality, depending on the RRs.
+Also, we can include time delays from change in exposure to change in health outcomes via the PIF.
+
+```{r chunk-20}
+```
+
+21) Use scenario incidence to calculate scenario disease life tables.
+
+```{r chunk-21}
+```
+
+22) Calculate life tables scenario.
+
+```{r chunk-22}
+```
+
+23) Generate list of outputs by age and sex.
+
+```{r chunk-23}
+```
+
+24) Combine all lists of outputs in a data frame to facilitate extraction of outcomes of interest and plotting.
+
+```{r chunk-24}
+```
+
+25) Plot outcomes for each age and sex cohort using created functions.
+Four functions are used: plot_output (plots outputs), grid_arrange_shared_legend and g_legend (places one label per page) and get_qualified_disease_name (replaces full name of disease)
+
+```{r chunk-25}
+
+knitr::include_graphics(graphs_doc[c(1:64)])
 
@@
 
-```{r, echo=TRUE}
-
-## Use externaly generated inputs for for disease life table.
-The previous code will serve to to some of the data preparation for DISMOD.
-Also need to devlop code to calculated DWs, for now, all generated in excel.
-
-idata <- read.csv("data/legacy/UK/idata.csv", stringsAsFactors = F)
-## Use run_disease
-
-i_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
-
-disease_life_table_list_bl <- list()
-index <- 1
-
-for (age in i_age_cohort){
-  for (sex in i_sex){
-    for (disease in i_disease) {
-      # Exclude breast_cancer for Males
-      if (sex == "male" && disease == "breast_cancer"){
-       # cat("\n") #Uncomment to see list
-      }
-      else {
-        # cat("age ", age, " sex ", sex, "and disease", disease, "\n") #Uncomment to see list
-        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, in_disease = disease)
-        index <- index + 1
-      }
-    }
-  }
-}
-## Uncommnet to check disease life table list
-# View(disease_life_table_list_bl[[8]])
-
-```
-
-18) Generate mock change in incidence of disease to generate scenario life tables. In the final version, this will come from the calculated PIFs by disease. 
-PIFs are applied here to incidence, however, may also be applied to case fatality, depending on the RRs.
-Also, we can include time delays from change in exposure to change in health outcomes via the PIF.
-
-```{r, echo=TRUE}
-## Create value to use as factor changing incidence rates.
-
-incidence_change <- 0.95
-
-## Generate scenario incidence (for each disease)
-
-incidence_sc <- list()
-index <- 1
-
-for (age in i_age_cohort){
-  for (sex in i_sex){
-    for (disease in i_disease) {
-
-      # Exclude breast_cancer for Males
-      if (sex == "male" && disease == "breast_cancer"){
-        # cat("\n") # Uncomment to see list
-      }
-      else {
-
-        incidence_sc[[index]] <- disease_life_table_list_bl[[index]]$incidence_disease * 
-        incidence_change
-        index <- index + 1
-
-      }
-    }
-  }
-}
-
-## Uncommnet to check scenario incidence
-# View(incidence_sc[[1]])
-
-```
-
-19) Use scenario incidence to calculate scenario disease life tables.
-
-```{r, echo=TRUE}
-
-disease_life_table_list_sc <- list()
-index <- 1
-for (age in i_age_cohort){
-  for (sex in i_sex){
-    for (disease in i_disease) {
-      # Exclude breast_cancer for Males
-      if (sex == "male" && disease == "breast_cancer"){
-        # cat("\n")
-      }
-      else {
-        # cat("age ", age, " sex ", sex, "and disease", disease, "\n")
-        # modify idata's incidence for the said scenario
-        td1 <- idata
-        td1[td1$age >= age & td1$sex == sex,][[paste("incidence", disease, sep = "_")]] <- incidence_sc[[index]]
-
-        # Instead of idata, feed td to run scenarios
-        disease_life_table_list_sc[[index]] <- run_disease(in_idata = td1, in_sex = sex,
-                                                           in_mid_age = age, in_disease = disease)
-        disease_life_table_list_sc[[index]]$diff_inc_disease <- 
-                                              disease_life_table_list_sc[[index]]$incidence_disease -   disease_life_table_list_bl[[index]]$incidence_disease
-        disease_life_table_list_sc[[index]]$diff_prev_disease <- 
-                                              disease_life_table_list_sc[[index]]$px  - disease_life_table_list_bl[[index]]$px
-        disease_life_table_list_sc[[index]]$diff_mort_disease <- 
-                                              disease_life_table_list_sc[[index]]$mx - disease_life_table_list_bl[[index]]$mx
-        disease_life_table_list_sc[[index]]$diff_pylds_disease <- 
-                                              (disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px) * disease_life_table_list_bl[[index]]$dw_disease
-        index <- index + 1
-      }
-    }
-  }
-}
-## Uncommnet to check scenario life tables
-# View(disease_life_table_list_sc[[1]])
-
-```
-
-20) Calculate life tables scenario.
-
-```{r, echo=TRUE}
-
-## Generate total change in mortality rate
-
-## Sum mortality rate change scenarios (mx_sc_total)
-
-mx_sc_total <- list()
-l_index <- 1
-index <- 1
-for (age in i_age_cohort){
-  for (sex in i_sex){
-    mortality_sum <- NULL
-    create_new <- T
-
-    for (disease in i_disease) {
-      if (sex == "male" && disease == "breast_cancer"){
-        # cat("\n")
-      }else{
-
-        if (create_new){
-          mortality_sum <- select(disease_life_table_list_sc[[index]], 
-                            c('age', 'sex'))
-                            mortality_sum$total <- 0
-          create_new <- F
-          mortality_sum$total <- mortality_sum$total + 
-                                (disease_life_table_list_sc[[index]]$diff_mort_disease)
-        }else{
-          mortality_sum$total <- mortality_sum$total +
-                                (disease_life_table_list_sc[[index]]$diff_mort_disease)
-        }
-
-        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
-        index <- index + 1
-      }
-    }
-    mx_sc_total[[l_index]] <- mortality_sum
-
-    l_index <- l_index + 1
-  }
-}
-
-## Uncommnet to check sceanrio mortality and changes 
-# View(mx_sc_total[[3]])
-
-## Generate total change in prevalent yld rates
-## Total ylds rate= sum (change prevalence disease*dw)
-
-pylds_sc_total <- list()
-l_index <- 1
-index <- 1
-for (age in i_age_cohort){
-  for (sex in i_sex){
-    pylds_sum <- NULL
-    create_new <- T
-
-    for (disease in i_disease) {
-      if (sex == "male" && disease == "breast_cancer"){
-        # cat("\n")
-      }else{
-
-        if (create_new){
-          pylds_sum <- select(disease_life_table_list_sc[[index]], c('age', 'sex'))
-          pylds_sum$total <- 0
-          create_new <- F
-          pylds_sum$total <- pylds_sum$total +   
-                            (disease_life_table_list_sc[[index]]$diff_pylds_disease)
-        }else{
-          pylds_sum$total <- pylds_sum$total + 
-                            (disease_life_table_list_sc[[index]]$diff_pylds_disease)
-        }
-
-        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
-        index <- index + 1
-      }
-
-    }
-    pylds_sc_total[[l_index]] <- pylds_sum
-    l_index <- l_index + 1
-  }
-}
-
-## Uncommnet to check scenario pyld change
-# View(pylds_sc_total[[2]])
-
-## Calculate general life tables with modified mortality and pylds total
-## Original mortality rate is modified by the mx_sc_total (total change in mortality from diseases)
-## Original pyld rate is modified by the change in each disease pylds
-
-general_life_table_list_sc <- list()
-index <- 1
-
-for (age in i_age_cohort){
-  for (sex in i_sex){
-
-    cat("age ", age, " and sex ", sex, "\n")
-    # modify idata's mortality and pyld total for the said scenario
-    td2 <- input_life_table
-    # td2 <- subset(td2, select = -c(mx, pyld_rate))
-    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx + mx_sc_total[[index]]$total
-    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate + pylds_sc_total[[index]]$total
-
-    # Instead of idata, feed td to run scenarios
-    general_life_table_list_sc[[index]] <- run_life_table(in_idata = td2, in_sex = sex, in_mid_age = age)
-    #
-
-    index <- index + 1
-  }
-}
-## Uncommnet to check scenario life tables
-# View(general_life_table_list_sc[[32]])
-# View(general_life_table_list_bl[[1]])
-
-## Check difference life table baseline and scenario
-general_life_table_list_bl[[1]]$Lx - general_life_table_list_sc[[1]]$Lx
-general_life_table_list_bl[[1]]$Lwx - general_life_table_list_sc[[1]]$Lwx
-
-```
-
-21) Generate list of outputs by age and sex.
-
-```{r, echo=TRUE}
-
-## In the following list "output_life_table", 32 data frames are nested per age and sex cohort
-
-output_burden <- list()
-l_index <- 1
-index <- 1
-for (age in i_age_cohort){
-  for (sex in i_sex){
-
-    # Males do not have breast cancer, that is why we need the if/else. 
-    # We create a TRUE/FALSE variable for the loop to move into the next disease
-
-    create_new <- T
-    for (disease in i_disease) {
-      if (sex == "male" && disease == "breast_cancer"){
-        # cat("\n")
-      }else{
-
-        if (create_new){
-          output_burden_sc <- select(disease_life_table_list_sc[[index]], 
-                              c('age', 'sex', 'incidence_disease', 'mx', 'px'))
-          names(output_burden_sc)[names(output_burden_sc) == 'incidence_disease'] <- 
-                              paste('incidence_disease', disease, "sc", sep = "_")
-          names(output_burden_sc)[names(output_burden_sc) == 'mx'] <- 
-                              paste('mx', disease, "sc", sep = "_")
-          names(output_burden_sc)[names(output_burden_sc) == 'px'] <- 
-                              paste('px', disease, "sc", sep = "_")
-          output_burden_bl <- select(disease_life_table_list_bl[[index]], 
-                              c('incidence_disease', 'mx', 'px'))
-          names(output_burden_bl)[names(output_burden_bl) == 'incidence_disease'] <- 
-                              paste('incidence_disease', disease, "bl", sep = "_")
-          names(output_burden_bl)[names(output_burden_bl) == 'mx'] <- 
-                              paste('mx', disease, "bl", sep = "_")
-          names(output_burden_bl)[names(output_burden_bl) == 'px'] <- 
-                              paste('px', disease, "bl", sep = "_")
-
-          ## New list to add calculations for changes in burden of disease (incidence and mortality numbers)
-
-          output_burden_change <- list()
-
-          output_burden_change$inc_num_bl <- disease_life_table_list_bl[[index]]$incidence_disease * 
-          (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx
-          output_burden_change$inc_num_sc <- disease_life_table_list_sc[[index]]$incidence_disease * 
-          (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx
-          output_burden_change$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * 
-          (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px)
-          * general_life_table_list_bl[[l_index]]$Lx)
-
-          output_burden_change$mx_num_bl <- disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx
-          output_burden_change$mx_num_sc <- disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx
-          output_burden_change$mx_num_diff <- (disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx)
-
-          names(output_burden_change)[names(output_burden_change) == 'inc_num_bl'] <- 
-          paste('inc_num_bl', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'inc_num_sc'] <- 
-          paste('inc_num_sc', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'inc_num_diff'] <- 
-          paste('inc_num_diff', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'mx_num_bl'] <- 
-          paste('mx_num_bl', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'mx_num_sc'] <- 
-          paste('mx_num_sc', disease, sep = "_")
-          names(output_burden_change)[names(output_burden_change) == 'mx_num_diff'] <- 
-          paste('mx_num_diff', disease, sep = "_")
-
-          ## Bind all lists
-
-          output_burden_sc <- cbind(output_burden_sc, output_burden_bl)
-          output_burden_sc <- cbind(output_burden_sc, output_burden_change)
-
-          create_new <- F
-
-          ## Here the calculations above are repeated, here is where the F is telling to move into the next disease
-
-          }else{
-
-          td3 <- select(disease_life_table_list_sc[[index]], 
-                c('incidence_disease', 'mx', 'px'))
-          names(td3)[names(td3) == 'incidence_disease'] <-
-          paste('incidence_disease', disease, "sc", sep = "_")
-          names(td3)[names(td3) == 'mx'] <- 
-          paste('mx', disease, "sc", sep = "_")
-          names(td3)[names(td3) == 'px'] <- 
-          paste('px', disease, "sc", sep = "_")
-
-          td4 <- select(disease_life_table_list_bl[[index]], 
-                c('incidence_disease', 'mx', 'px'))
-          names(td4)[names(td4) == 'incidence_disease'] <- 
-          paste('incidence_disease', disease, "bl", sep = "_")
-          names(td4)[names(td4) == 'mx'] <- 
-          paste('mx', disease, "bl", sep = "_")
-          names(td4)[names(td4) == 'px'] <- 
-          paste('px', disease, "bl", sep = "_")
-
-          output_burden_change2 <- list()
-
-          output_burden_change2$inc_num_bl <- disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx
-          output_burden_change2$inc_num_sc <- disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx
-          output_burden_change2$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx)
-
-          output_burden_change2$mx_num_bl <- disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx
-          output_burden_change2$mx_num_sc <- disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx
-          output_burden_change2$mx_num_diff <- (disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx)
-
-          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_bl'] <- 
-          paste('inc_num_bl', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_sc'] <- 
-          paste('inc_num_sc', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_diff'] <- 
-          paste('inc_num_diff', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_bl'] <- 
-          paste('mx_num_bl', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_sc'] <- 
-          paste('mx_num_sc', disease, sep = "_")
-          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_diff'] <- 
-          paste('mx_num_diff', disease, sep = "_")
-
-          ## Bind all lists
-
-          output_burden_sc <- cbind(output_burden_sc, td3)
-          output_burden_sc <- cbind(output_burden_sc, td4)
-          output_burden_sc$age_cohort <- age 
-          output_burden_sc <- cbind(output_burden_sc, output_burden_change2)
-
-        }
-
-        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
-        index <- index + 1
-      }
-
-    }
-
-    ## general_life_table_list_sc and general_life_table_list_bl (Lx)
-    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], c('Lx', 'Lwx'))
-    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- paste('Lx', "sc", sep = "_")
-    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- paste('Lwx', "sc", sep = "_")
-
-    output_burden_lf_bl <- select(general_life_table_list_bl[[l_index]], c('Lx', 'Lwx'))
-    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lx'] <- paste('Lx', "bl", sep = "_")
-    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lwx'] <- paste('Lwx', "bl", sep = "_")
-
-    output_burden_lf_sc$Lx_diff <- general_life_table_list_bl[[l_index]]$Lx - general_life_table_list_sc[[l_index]]$Lx
-        output_burden_lf_sc$Lwx_diff <- general_life_table_list_bl[[l_index]]$Lwx - general_life_table_list_sc[[l_index]]$Lwx
-
-    output_burden_sc <- cbind(output_burden_sc, output_burden_lf_sc)
-    output_burden_sc <- cbind(output_burden_sc, output_burden_lf_bl)
-
-    output_burden[[l_index]] <- output_burden_sc
-    l_index <- l_index + 1
-
-  }
-}
-
-## Uncomment to check
-# View(output_burden[[1]])
-
-```
-
-22) Combine all lists of outputs in a data frame to facilitate extraction of outcomes of interest and plotting.
-
-```{r, echo=TRUE}
-#####Generate a data frame for all results and create function to get outcomes.
-
-output_df <- plyr::ldply(output_burden, rbind)
-
-# View(output_df)
-
-#Remove variables that are not used in the generation of outputs. CHANGE THIS NAMES, TOO LONG
-
-output_df <- subset(output_df, select = -c(incidence_disease_ihd_bl, incidence_disease_ihd_sc, incidence_disease_istroke_bl, incidence_disease_istroke_sc, incidence_disease_diabetes_bl, incidence_disease_diabetes_sc, incidence_disease_breast_cancer_bl, incidence_disease_breast_cancer_sc,
-incidence_disease_colon_cancer_bl, incidence_disease_colon_cancer_sc,  mx_ihd_bl, mx_ihd_sc, mx_istroke_bl, mx_istroke_sc, mx_diabetes_bl, mx_diabetes_sc, mx_breast_cancer_bl, mx_breast_cancer_sc, mx_colon_cancer_bl, mx_colon_cancer_sc, px_ihd_bl, px_ihd_sc, px_istroke_bl, px_istroke_sc, px_diabetes_bl, px_diabetes_sc, px_breast_cancer_bl, px_breast_cancer_sc, px_colon_cancer_bl, px_colon_cancer_sc))
-
-```
-
-23) Plot outcomes for each age and sex cohort using created functions.
-Four functions are used: plot_output (plots outputs), grid_arrange_shared_legend and g_legend (places one label per page) and get_qualified_disease_name (replaces full name of disease)
-
-```{r echo=TRUE}
-
-output_dir = "output/graphs2"
-
-i_outcome <- c("mx", "inc")
-p_list_male <- list()
-p_list_female <- list()
-male_index <- 1
-female_index <- 1
-for (age in i_age_cohort){
-      for (sex in i_sex) {
-        for (outcome in i_outcome) {
-          for (disease in i_disease){
-
-       if (sex == "male" && disease == "breast_cancer"){
-         # cat("\n")
-       }else{
-
-p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_disease = get_qualified_disease_name(disease))
-
-            if (sex == "male"){
-
-              p_list_male[[male_index]] <- p_index
-
-              if (male_index %% 4 == 0 && male_index > 0){
-
-                p1 <- p_list_male[[male_index - 3]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-                p2 <- p_list_male[[male_index - 2]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-                p3 <- p_list_male[[male_index - 1]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-                p4 <- p_index + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-
-  jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
-  grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age),
-                                          mainLeft = 'Cases', mainBottom = 'Age')
-                dev.off()
-
-              }
-
-              male_index <- male_index + 1
-
-            }
-
-            if (sex == "female" && female_index > 0){
-              p_list_female[[female_index]] <- p_index
-
-              if (female_index %% 5 == 0){
-                p1 <- p_list_female[[female_index - 4]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-                p2 <- p_list_female[[female_index - 3]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-                p3 <- p_list_female[[female_index - 2]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-                p4 <- p_list_female[[female_index - 1]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-                p5 <- p_index + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
-
-               jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
-                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age), mainLeft = 'Cases', mainBottom = 'Age')
-                dev.off()
-
-                }
-              female_index <- female_index + 1
-              }
-
-            }
-         }
-       }
-   }
-}
-
-```
-
-```{r echo=FALSE, out.width='100%'}
-## Loop to include graphs in the document
-
-graphs_doc <- list()
-index <- 1
-for (age in i_age_cohort) {
-      for (sex in i_sex)  {
-        for (outcome in i_outcome) {
-
-graphs_doc [[index]] <- c(paste(output_dir, "/",age,"_",sex,"_", outcome,".jpeg", sep = ""))
-knitr::include_graphics(graphs_doc [[index]])
-
-          index <- index + 1
-    }
-  }
-}
-
- knitr::include_graphics(graphs_doc[c(1:64)])
-```
-
-24) Add up all outcomes per year of simulation with generated function.
+26) Add up all outcomes per year of simulation with generated function.
 BETTER TO HAVE A LOOP, I TRIED BY LOOPING IN THE FUNCITON ARGUMENT I_OUTCOMES, BUT THIS DID NOT WORK.
@@
 
-```{r, echo=TRUE}
-
-## Try binding function
-
-aggregate_frame_males <- list()
-aggregate_frame_females <- list()
-
-index <- 1
-
-for (outcome in i_outcome) {
-  for (disease in i_disease) {
-
-aggregate_frame_males[[index]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c(paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")))
-
-aggregate_frame_females[[index]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c(paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")))
-
-# Remove non-numeric columns starting with age and sex
-aggregate_frame_males[[index]] <- aggregate_frame_males[[index]] %>% select(-starts_with("age"), -starts_with("sex"))
-
-aggregate_frame_females[[index]] <- aggregate_frame_females[[index]] %>% select(-starts_with("age"), -starts_with("sex"))
-
-index <- index + 1
-  }
-}
-
-## Loop for life years (Lx) and health adjusted life years (Lwx) to then add to total aggregated data.
-
-## RUN ANOTHER LOOP FOR LX AND LWX
-
-i_outcome2 <- c("Lx", "Lwx")
-
-aggregate_frame_males2 <- list()
-aggregate_frame_females2 <- list()
-
-for (i in i_outcome2){
-
-aggregate_frame_males2[[i]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c(paste(i, "bl", sep = "_"), paste(i, "sc", sep = "_"), paste(i, "diff",sep = "_")))
-
-aggregate_frame_females2[[i]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c(paste(i, "bl", sep = "_"), paste(i,  "sc", sep = "_"), paste(i, "diff",sep = "_")))
-
-aggregate_frame_males2[[i]] <- aggregate_frame_males2[[i]] %>% select(-starts_with("age"), -starts_with("sex"))
-
-aggregate_frame_females2[[i]] <- aggregate_frame_females2[[i]] %>% select(-starts_with("age"), -starts_with("sex"))
-
-}
-
-## Uncomment to check data frames (list correspond to a disease and outcome combination, for example,
-## incidence breast cancer)
-
-# View(aggregate_frame_females[[2]])
-
-## Transform list to data frame (this is not working after the first disease)
-
-aggregate_frame_females <- do.call(cbind, aggregate_frame_females)
-aggregate_frame_males <- do.call(cbind, aggregate_frame_males)
-aggregate_frame_females2 <- do.call(cbind, aggregate_frame_females2)
-aggregate_frame_males2 <- do.call(cbind, aggregate_frame_males2)
-
-View(aggregate_frame_females2)
-
-## Drop ending of variables names _males/_females to add up all outcomes in next step.
-
-names(aggregate_frame_females) = gsub(pattern = "_female", replacement = "", x = names(aggregate_frame_females))
-
-names(aggregate_frame_males) = gsub(pattern = "_male", replacement = "", x = names(aggregate_frame_males))
-
-names(aggregate_frame_females2) = gsub(pattern = "_female", replacement = "", x = names(aggregate_frame_females2))
-
-names(aggregate_frame_males2) = gsub(pattern = "_male", replacement = "", x = names(aggregate_frame_males2))
-
-## Uncomment to check that the code is dropping the male/female ending. 
-# View(aggregate_frame_males)
-
-## Create a copy of aggregate_frame_females.  
-total_aggr1 <- aggregate_frame_females
-## Add aggregate_frame_males values to it
-for (i in 1:ncol(aggregate_frame_females)){
-  total_aggr1[i] <- total_aggr1[i] + aggregate_frame_males[i]
-}
-
-## Add data frames 2 with life years (check adds Lx and Lwx at the beginning of the variable name)
-total_aggr2 <- aggregate_frame_females2
-## Add aggregate_frame_males values to it
-for (i in 1:ncol(aggregate_frame_females2)){
-  total_aggr2[i] <- total_aggr2[i] + aggregate_frame_males2[i]
-}
-
-## Combine data frames
-
-total_aggr <- cbind.data.frame(total_aggr1, total_aggr2)
-
-total_aggr$sim_year <- seq.int(nrow(total_aggr))
-
-## Uncomment to see total data frame
-View(total_aggr)
-
-## Test that total_aggr is adding males and females dataframes
-# 
-# (aggregate_frame_females2$Lx.Lx_bl_22 + aggregate_frame_males2$Lx.Lx_bl_22) -total_aggr$Lx.Lx_bl_22
-
+```{r chunk-26}
 ```
 
-25) Plots using total aggregated change in burden.
+27) Plots using total aggregated change in burden.
 This graphs are not very good for life years and health adjusted life years as the difference between baseline and scenario is very small relative to the total, hence the graphs show almost no change.
 
-```{r, echo=TRUE}
-
-####This plot has to be customised to in_outcomes, here, only totals shown, but specifications are up to the user. ADD LOOP for all outcomes over time and total TABLE.
-
-####[] is used here to indicate the number of simulation years into the future. 
-####Disease outcomes has to be changed to the outcome of interest
-
-#### Test code with loops for aggregated outcomes diseases burden. NOT WORKING.
-
-### Compare with loops for age and sex cohort outcomes.
-
-p_aggr_list <- list()
-index <- 1
-
-for (outcome in i_outcome) {
-  for (disease in i_disease) {
-    # outcome <- i_outcome[1]
-    # disease <- i_disease[1]
-
-p_aggr_list_index <- ggplot(total_aggr[1:79,], aes(x = total_aggr[["sim_year"]])) +
-
-  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_bl", disease, sep = "_")]], colour = paste("total", outcome, "num_bl", disease, sep = "_"))) +
-  theme_classic() +
-  geom_hline(yintercept=0, linetype="dashed", color = "black") +
-  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_sc", disease, sep = "_")]], colour = paste("total", outcome, "num_sc", disease, sep = "_"))) +
-  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_diff", disease, sep = "_")]], colour = paste("total", outcome, "num_diff", disease, sep = "_"))) +
-  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease, outcome)) +
-  theme(plot.title = element_text(hjust = 0.5, size = 12)) +
-  scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
-  theme(plot.title = element_text(hjust = 0.5))
-p_aggr_list[[index]] <- p_aggr_list_index  
-index <- index + 1
-
-  }
-}
-
-index <- 1
-
-interpolation_index <- 1
-for (outcome in i_outcome) {
-  for (disease in i_disease) {
-    file_name = paste("output/graphs2/", "Aggregated Outcomes", outcome, disease, ".jpeg", sep=" ")
-    jpeg(file_name)
-    print(p_aggr_list[[index]])
-    index <- index + 1
-    dev.off()
-  }
-}
-
-p_aggregated <- do.call(marrangeGrob, list(grobs=p_aggr_list, nrow = 2, ncol = 2))
-p_aggregated
-
+```{r chunk-27}
 ```

### 2018-06-07 R/MSLT/MSLTdoc.Rmd
Move PA, injury and MSLT inside R folder

--- 2018-05-25 MSLT/MSLTdoc.Rmd
@@
     - \newcommand{\elandscape}{\end{landscape}}
+
 output:
@@
 ```
+
 12) Keep relevant variables

### 2018-06-26 R/MSLT/documentation/MSLTdoc.Rmd
merging

---
title: "Proportional multi-state multiple-cohort life table model"
author: "Belen Zapata-Diomedi and Ali Abbas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
    - \usepackage{setspace}\onehalfspacing
    - \usepackage{float}
    - \usepackage{pdflscape}
    - \newcommand{\blandscape}{\begin{landscape}}
    - \newcommand{\elandscape}{\end{landscape}}
output:
   pdf_document: 
     toc: true
     toc_depth: 3
     number_sections: true
     df_print: kable
   fontsize: 12 pt
   geometry: margin=1in
bibliography: Bibliography.bib
---

```{r setup, include=FALSE}

## Global options
knitr::opts_chunk$set(echo = TRUE, width = 75)
options(knitr.table.format = 'latex')
knitr::opts_chunk$set(fig.cap = FALSE, fig.align = "center", fig.pos = "H", fig.scap = TRUE)

## The work in this document requires the following packages (code, rmarkdown and citing packages):

require(dplyr)
require(tidyverse)
require(knitr)
require(kableExtra)
require(citr)
require(gridExtra)
require(cowplot)
require(ggpubr)
require(grid)
require(ggplot2)

```

\newpage

# Introduction

The proportional multi-state multiple-cohort life table model (PMSLT) is a population level model (macro) approach to simulate health (and economic) implications of changes in exposure to health risk factors (e.g. physical inactivity, air pollution and diet). The PMSLT has been widely used to simulate outcomes for population level interventions for the reduction of chronic diseases.

The model was developed by Jan Barendregt and colleagues and has been widely used in Australia and New Zealand [@RN38; @RN8299].

The basic infrastructure of the model consist of three components: (1) Effect size for the intervention of interest (e.g. intervention to urban design that modifies population levels of physical activity); (2) Calculation of the potential impact fraction (PIF) to derive the change in occurrence of disease (incidence rate/case fatality rate) attributable to a change in the distribution of the risk factor (e.g. physical activity); and (3) Use of the PMSLT to simulate health (and economic) outcomes attributable to a change in the distribution of health risk factor/s in the population of interest. Figure 1 summaries the basic infrastructure of the model. ITHIM is included in Figure 1 to show that both approaches share in common steps one and two and differ in the mechanisms of calculation of change in health burden.

**HALYs, QALYs and DALYs**

In this model we use the term *health-adjusted life year* (HALY). As *summary measure of population health* it measures both quantity and quality of life, where one HALY represent the equivalent of one year in full health (which could be two years with a quality of life of 0.5, for example). Specific types of HALY are the quality-adjusted life year (QALY) and the disability-adjusted life year (DALY). The QALY derives from economics and was first used in the 1960s as a measure of health gain [@RN8158]. The disability-adjusted life-year (DALY) was developed for use in burden of disease studies as a measure of health loss due to disease [@RN8158]. Our calculated HALYs are neither QALYs not DALYs, but something in between. They are similar to QALYs in that they represent health gains. However, the main difference is in the calculation of the health-related quality of life component. QALYs use measures of utility weights that traditionally represent individual experiences of health, whereas our estimated HALYs use disability weights linked to specific diseases, which were developed for the Global Burden of Disease study [@RN8158]. As discussed in past research [@RN21; @RN8262] the main advantage of using disability weights over utility weights is that disability weights refer to specific diseases rather than health states (which are difficult to link to risk factors-e.g. physical inactivity). We opted to use the more general terms HALYs given that the use of the DALYs terminology may lead to think that our calculations are similar to those in burden of diseases studies [@RN8153]. In our study, our model does not explicitly separate years of life lost (YLL) and years lived with disability (YLD) components, but instead calculates the total number of life years lived, adjusted for the average health-related quality of life in those years (by age and sex). In burden of disease studies, DALYs are defined as the sum Years of Life Lost (YLL) and Years Lived with Disability (YLD).

\newpage

![Basic ITHIMR infrastructure](structure\Figure1.png){fig.pos="H"}

## Contribution to ITHIMR

The PMSLT similar to ITHIM is a comparative risk assessment approach [@RN8957] that consist of calculating the change in the health burden for a population of interest from a change in exposure to health risks factors (e.g. physical inactivity, air pollution and road trauma).As depicted in Figure 1, both methods need estimates of the potential impact fraction (PIF), which indicates the proportion of the disease burden attributable to a risk factor of interest (e.g. physical inactivity) [@RN42]. A step further back, is the development of scenarios that bring about change in the distribution of the risk factor of interest. For now, we only focus on calculations from the PIF onward, and provide a hypothetical example of change in the population distribution of physical activity. Incorporation of additional  health risk factor (air pollution, road trauma, NO2 and noise) will be discussed in the relevant code sections.

### Difference between ITHIM and PMSLT

* **Time component**
The *PMSLT* follows a population of interest over time. For example, as set up here, we simulate sex and age (5 years starting at 20) cohorts over time until they die or reach 100 years of age. This implies that we can include trends for diseases, time lags for change in exposure to risk factors and change in health and demographic changes (e.g. population growth). In addition, we can estimate yearly changes in the burden of diseases over the life course or for a specified number of years. The *ITHIM* approach is a snapshot of change in burden for one year.

* **Interaction between multiple diseases**
The *PMSLT* accounts for the interaction between multiple diseases, with proportions of the population being able to be in more than one health state [@RN8957]. This avoids  overestimation of outcomes as a result of summing health outcomes attributable to each disease individually as done in *ITHIM*. It is important to note that the *PMSLT* assumes that diseases are independent of each other. That is to say, developing a disease is unrelated to a concurrent diagnoses of another disease).

* **Mortality rate** 
The *PMSLT* calculations for changes in life years (and health-adjusted life years) and mortality outcomes is based on observed mortality rates for the population of interest. In the *ITHIM* model, if burden of  disease estimates from the Global Burden of Disease (GBD) study are used, then, the mortality component is based on the highest attained life expectancy observed in the world.

* **Impact of disability in increased life expectancy**
In GBD studies, YLLs are not adjusted for disability; hence, their use in estimating intervention effects results in over-estimation, which the *PMSLT* approach avoids. Another way of seeing this is that estimated changes in morbidity using the *ITHIM* do not allow for how implicit increases in life expectancy impact on morbidity. While the changes in deaths and prevalence using the *PMSLT* are in some  ways more accurate than those from the *ITHIM* approach it should be noted that that the average age of death and incident disease will change and thus the disease burden will be on average be shifted later in life (which is a realistic approach).

# R development

The model is set up as a long script to perform the required mathematical calculations. Where possible, we wrote functions and loops to avoid repetition. We set up the model  with data for Greater London. Figure 2 depicts the PMSLT model framework, which was followed in the code development.

\newpage
\blandscape

![Proportional multi-state life-table simplified framework. *The simplied PMST shows the interaction between the life table, disease life table and potential impact fraction (PIF). The PIF calculations by age and sex group are the same as those generated for ITHIM. The PIF (or 1-PIF) modifies incidence of disease, which changes prevalence and mortality (disease specific life table). Changes in prevalence and mortality rates from the disease specific life tables feed into the life table by changing all-cause mortality, which in turn changes life years. Change in prevalence of diseases changes total years lived with disability, which in turn modifies health-adjusted life years*](structure\Figure2.png){width=100%, fig.pos="H"}

\elandscape

In what follows, first, we specify input parameters. Second, we present the code with explaining notes. Third, we present examples of outcomes and lastly we comment on topics related to implementation. Here we only included the physical activity health pathway. In the comments section, we discuss the implementation of exposure to air pollution and road trauma. Note that in the presentation of input parameters, those needed to calculate PIFs are excluded, as these are common to the ITHIM, expect if trends are included (refer to comments section).

## Inputs

We specify data requirements for the life table and disease life tables (Figure 2) and potential sources.

### Life table

Inputs of the life table are: population numbers by sex (per 1-year or age grouping of interest), mortality rates or probability of all cause mortality by single age group and sex and total prevalent years lived with disability rate per single year by sex. Disease specific disability weights are presented as inputs here as these adjust the total years lived with disability, hence, the health-adjusted life years.

####	Population numbers

These data will be provided by the synthetic population or derived from other data. In the code presented here, we simulate 5-year age and sex cohorts. Data for population may be in five-year age groups or one-year. For the example for Greater London, we derive 5-year age groups from GBD IHME data, however, we also provide a script if five-year age groups are to be derived from one-year age groups data. I left potential data sources below as a reference.

Data source: (1) National census; (2) Worldwide population and mortality data:  http://www.mortality.org/ (mostly high income countries; and (3) Calculate from the Global Burden of Disease by the Institute of Health Metrics and Evaluation (GBD IHME) data (rates and numbers available from (http://ghdx.healthdata.org/gbd-results-tool).

#### Mortality rates

Mortality rates are needed per single year and sex. These data are available from GBD IHME, however, in age groups (1-4, 5-9, etc). We provide an interpolation script to derive in between ages rates (cubic spline).

Note that we need data for population numbers and all cause mortality rates for: (1) PMSLT and (2) Dismod II collection (more in Dismod II section). Population data from the synthetic population is used for the PMSLT (if available). For Dismod II, population and mortality data should be from the same source (GBD IHME).

####	Total years lived with disability rates per single year and sex.

These data is available from the GBD (http://ghdx.healthdata.org/gbd-results-tool) per 5-year age groups. We can use interpolation to derive between ages rates (script provided).

#### Disability weights (quality of life weights)

Disability weights (DW) are derived from disease specific years lived with disability (YLD) and disease specific prevalence by age group (5 years) and sex. Data for YLDs prevalence is obtained from the online GBD IHME data tool (http://ghdx.healthdata.org/gbd-results-tool). An age and sex specific-correction was introduced to counteract the effects of accumulating comorbid illnesses in the older age groups (Equation 1).

\begin{equation}
\label{DW adjusted for total YLDs-YLDdPd1YLDt}
(YLDd/Pd)/(1-YLDt) = DW adjusted for total YLDs
\end{equation}

Where YLDd is the YLD mean number per age and sex for a given disease, Pd is the prevalence (as reported in GBD )  for a given disease by age and sex and YLDt is total YLD rate per age and sex.

### Disease life tables

#### Incidence and case fatality

For each of the modeled diseases the PMSLT needs incidence and case fatality rates per sex and one-year intervals. Data from the GBD IHME studies with Dismod II (free at https://www.epigear.com/index_files/dismod_ii.html) is used to derive internally consistent data and generate missing data. For example, the GBD studies provide data for incidence, prevalence and disease mortality, however, not case fatality. Other national level sources may also be explored/used, and compare with estimates produce from GBD data and Dismod II.

**Dismod II** inputs are: (1) population numbers and mortality rates and (2) disease specific inputs.

*Population and mortality*

Within Dismod II, each setting (e.g. country) has a collection that consists of population numbers (preferably the same as used in GBD IHME studies, due to the mortality envelop) and all- cause mortality rates (numbers and calculate rates). The GBD provides 5-year age groups that are acceptable input parameters for Dismod II.

*Disease inputs by age group and sex*

Each setting collection has a given number of diseases. Dismod II works with at least three of: case fatality, prevalence, incidence, mortality (disease), case fatality, remission, duration and the relative risk for mortality. So far, we have been assuming that remission is zero for chronic diseases, that is to say, when people become diseased, they do not recover. Special care should be taken with this assumption, as the GBD data assumes remission for some diseases, for example cancers, where after 10 years cases recover, except for long term sequelae. Since GBD now provides prevalence, incidence and mortality, it may be best to use all three as Dismod II input parameters to compare the effect of the remission assumption by the GBD for some diseases.

```{r, echo=FALSE}

pmslt_inputs <- tribble(
~Input, ~Source, ~Comments, 
"Life table", "Synthetic population per sex and age group", "Age grouping in life table to match synthetic population", 
"Life table", "Synthetic population per sex and one-year age group", "If one year age group is not avabilable it can be derive using interpolation from age groups data", 
"Life table", "Global Burden of Disease (GBD) study per one-year age group and sex", "GBD data is in five-year age groups, interpolation to derive one-year age groups", 
"Disease life table", "GBD data for prevalence, incidence and mortality and DISMOD II", "Two step process. First obtain disease and population data from GBD. Second, use Dismod II to derive internally consistent estimates for incidence and case fatality (PMSLT disease life table iputs)", 
"Disease life table", "Derive from disease prevalence and years lived with disability from GBD", "Adjustments for comorbidities in later years of life to be applied"
)

kable(pmslt_inputs,"latex", booktabs = T, caption = "PMSLT inputs") %>%
 kable_styling(latex_options = c("striped", "hold position"),
               full_width = F)%>%
  column_spec(2, width = "15em") %>%
  column_spec(3, width = "15em") %>%
  row_spec(0, bold = T)
```

##Code

Following the structure of Figure 2, we developed functions to perform sex and age cohorts calculations for the life table, disease life tables and potential impact fractions: run_life_table, run_disease and and run_pif. We also generated two functions for outputs: plot_outputs and gen_aggregate. The function plot_outputs creates age-group and sex linear plots for specified outcomes (e.g. health-adjusted life years, incidence of diabetes) and gen_aggregate adds up each cohort results. Functions were then used in a code script. In what follows, we explain each step in the development of the script. Here we also include code chunks, however, we also kept them separately in the MSLT folder, in the code file.

In what follows, we start explain the script step by step.

###Set up

We start by cleaning the global environment (1) to keep track of our works and ensure that the code is generating our desired outcomes. Then, we set up an option to avoid the use of scientific notation (2) and lastly we load the functions (3). The code chunks are shown in the rmarkdown output.

1) Clean Global Environment
```{r, echo=TRUE}
rm (list = ls())
```
2) Avoid scientific notation
```{r, echo=TRUE}
options(scipen=999)
```
3) Load functions 
```{r, echo=TRUE}
source("code/functions.R")
```

###Inputs

Table 1 describes data needs for the PMSLT, here we expand on the data needs and mechanisms (Figure 3) to use the PMSLT approach in ITHIMR (Figure 1).

Initial case studies for the ITHIMR are: London, Sao Pablo, Delhi, Accra, Los Angeles and Edinburgh. Here, we will start with **Greater London** given the availability of disease epidemiology data from the GBD IHME study. For the rest of the case study cities data is available at the country level, hence, a scaling method is needed to reflect the local burden of disease.

\newpage
\blandscape

![Proportional multi-state life-table model. *Three sections are presented in Figure 3: **Data input sources**, **Inputs PMSLT** and **PMSLT mechanisms**. The color coding from Data inputs sources to Inputs PMSLT link sources with inputs for the PMSLT. Solid arrows represent final inputs and dashed-arrows represent intermediate inputs that need further processing. Purple coding means a process and green coding represent change in mortality and disability prevalence rates to modify the life table parameters. Black color coding with white color font represent final model outcomes. For both, the life table and disease life table, two sets of each are simulated, one for the baseline and the other for the scenario.*](structure\Figure3.png){width=50%,height=30%,fig.pos="H"}

\elandscape

####Global Burden of Disease data

First, we explain how to obtain the data, second additional processing to derive data not reported (population) and one-year age groups (original data is in five-year age groups) and last procedure to use Dismod II. Data from the *Global Burden of Disease data* in Figure 3 can be download from here: http://ghdx.healthdata.org/gbd-results-tool. Figure 4 is a screenshot of the GHDx.

![Global Burden of Disease data results tool.](data/Figure4.png){width=70%,height=50%,fig.pos="H"}

Table 2 specifies the selections to do for each of the tabs in Figure 2.

```{r, echo=FALSE}

pmslt_inputs <- tribble(
~Tab, ~Selection,
"Base", "Single", "Location", "Case study city", "Year", "Latest available", "Context", "Cause", "Age", "Under 5, 5 to 9, 10 to 14, 15 to 19, 20 to 24, 25 to 29, 30 to 34, 35 to 39, 40 to 49, 50 to 54, 55 to 59, 60 to 64, 65 to 69, 70 to 74, 75 to 79, 80 to 84, 89 to 89, 90 to 95, 95 plus", "Metric", "Number, Rate", "Measure", "Deaths, YLDs, Prevalence, Incidence", "Sex", "Male, Female", "Cause", "Total All causes, ischemic heart disease, etc")

kable(pmslt_inputs,"latex", booktabs = T, caption = "Global burden of disease data") %>%
 kable_styling(latex_options = c("striped", "hold position"),
               full_width = F)%>%
  column_spec(2, width = "30em") %>%
  column_spec(3, width = "30em") %>%
  row_spec(0, bold = T)

```

Once the selections described in Table 2 are made, the option *Download CVS** in the GHDx website is selected. A prompt comes up asking for an email address. The data is sent to the designated email address (within minutes) in ZIP format, unzip and use the code below to read the data (4). Here, we selected data for Greater London and England. The aim is to compare and derive scaling factors as for most cities the data is not available from the GBD and country level data may be used and scaled to the city level. Note that the data input requirement for the PMSLT, except population numbers, is in rates. Therefore the scaling is to better reflect the burden of an area, this is a different issue than working with numbers (e.g. total mortality numbers, total YLDs numbers) as in the ITHIM approach.

4) Read GBD data
```{r, echo=TRUE}

GBDdata <- read.csv("data/UK/gbd2016.csv", stringsAsFactors = F)

```

The following codes serves to sort out the GBD data to the inputs required for the life table, disease life table and Dismod II

These data should be used to generate the general life table and disease life tables (Figure 3).

5) Change all upper cases to lower case and delete () from variables. 
```{r, echo=TRUE}

GBDdata <- mutate_all(GBDdata, funs(tolower))
GBDdata$measure[GBDdata$measure=="ylds (years lived with disability)"] <- "ylds"

```

6) Create age categories index in GBDdata (mid age, to match cohort running), total of 20 age groups. These are the age cohorts to simulate. 
```{r, echo=TRUE}

# 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97
GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
GBDdata$age_cat [GBDdata$age =="80 to 84"] <- 82
GBDdata$age_cat [GBDdata$age =="85 to 89"] <- 87
GBDdata$age_cat [GBDdata$age =="90 to 94"] <- 92
GBDdata$age_cat [GBDdata$age =="95 plus"] <- 97

```

7) Create age and sex categories to obtain population numbers. Population numbers from GBD are used in Dismod II. For the Life table (Figure 3), the numbers may be from the synthetic population. For now, the Life table is set up with population numbers derived from the GBD data. 
```{r, echo=TRUE}

GBDdata$sex_age_cat <- paste(GBDdata$sex,GBDdata$age_cat, sep = "_"  )

```
8) Convert string variables to numeric to do calculations.
```{r, echo=TRUE}

GBDdata$val <- as.numeric(as.character(GBDdata$val))

```
9) Generate population numbers for Greater London in a new data frame ("GBD_population"). Note that there is data for England as well, which is used in a separate rmarkdown document (GBDCompare). Population numbers are derived from rates per 100,000 and total numbers of cases. 
```{r, echo=TRUE}

GBD_population <- filter(GBDdata, measure == "deaths", cause == "all causes", 
metric == "rate" | metric == "number" ) %>% select(metric, age_cat, sex, val, 
sex_age_cat, location)

```
10) Generate population numbers from given number of cases and rates per 100,000 people. 
```{r, echo=TRUE}

 for (i in 1:nrow(GBD_population)) {
    if (GBD_population$metric[i] == "number") {
      GBD_population$five_year_population[i] <- GBD_population$val[i] * 100000/
      GBD_population$val[i + 2]}     
    else {GBD_population$five_year_population[i] <- NA}
 }

```
11) Remove rows with zero
```{r, echo=TRUE}

GBD_population <- GBD_population[!is.na(GBD_population$five_year_population),]

```
12) Keep relevant variables
```{r, echo=TRUE}

GBD_population <- filter(GBD_population) %>% 
select(sex_age_cat, sex, age_cat, five_year_population, location)

```
13) Create data frames for Greater London to be later used for: a) interpolation of rates and b) PMLT cohorts. 
```{r, echo=TRUE}

GBD_population_GL <-  filter(GBD_population, location == "greater london") %>% 
select(sex_age_cat, age_cat, sex, five_year_population, location, sex_age_cat)

```
12) Check population total numbers
```{r, echo=TRUE}

GreaterLondon <- sum(GBD_population_GL$five_year_population)

```

13) Generate data frames for Greater London with per person rates (per 100,000 in original data).  
```{r, echo=TRUE}

GBDGL <- filter(GBDdata, location == "greater london" & metric == "rate") %>% 
select(measure, location, sex, age, metric, cause, val, age_cat, sex_age_cat)
GBDGL$one_rate <- GBDGL$val/100000

```

14) For the life table, we need to mortality and total yld rates in one year age intervals. The original data is in five years. Thus, the following code is used to interpolate a single-year age distribution form a five-yearly distribution.

```{r, echo=TRUE}

## Loop to generate interpolated rates for all cause mortality and ylds for males and females
## UPDATE WITH YLDS RATES ALL CAUSES ADJUSTED FOR ALL OTHER MODELLED DISEASES.

i_sex <- c("male", "female")
i_measure <- c("deaths", "ylds") #" (years lived with disability)")

index <- 1

for(sex_index in i_sex) {
  for (measure_index in i_measure) {

data <- filter(GBDGL, measure == measure_index, sex == sex_index, 
cause == "all causes") %>% select(measure, location, sex, age, metric, 
cause, val, age_cat, one_rate) 
assign(paste(sex_index, measure_index, "interpolated_data", sep = "_"), data)
x=data$age_cat
y=log(data$one_rate)

interpolation_func <- splinefun(x, y, method = "natural", ties = mean)

interpolated <- as.data.frame(interpolation_func(seq(0, 100, 1)))
age <- seq(0,100,by=1)
interpolated <- cbind(interpolated, age)
interpolated[,1] <- exp(interpolated[,1])
colnames(interpolated)[1] <- paste(measure_index)
## Add column with sex to create age_sex category to then merge with input_life table
interpolated$sex <- paste(sex_index)
interpolated$sex_age_cat <- paste(interpolated$sex, interpolated$age, sep = "_")
## Change name of column death to mx and ylds to pyld_rate to then merge 
## with input_life table

if (colnames(interpolated)[1] == "deaths")
    colnames(interpolated)[1] <- paste("mx")
  else
    colnames(interpolated)[1] <- paste("pyld_rate")

# Name data frame
assign(paste(sex_index, measure_index, "interpolated", sep = "_"), interpolated)

  }
}

## Do graph with another layer for original rates in age groups for comparison purposes.

p_interpolation_list <- list()

interpolation_index <- 1

for(sex_index in i_sex) {
  for (measure_index in i_measure) {

p_interpolation_index <- ggplot(data = interpolated,mapping = aes(age, interpolated[,1])) +
              geom_line(aes(color = "Interpolated")) +
              geom_point(
              data = data, 
              mapping = aes(age_cat, one_rate, color = "Original")) +
              labs(colour="",x="Age",y="Rates", sep = " ") +
              labs (title = paste("Rates", sex_index, "all cause", 
              measure_index, sep = " "), size=14) +
              theme_classic() +
              theme (plot.title = element_text(hjust = 0.5))

              p_interpolation_list[[interpolation_index]] <- p_interpolation_index
              interpolation_index <- interpolation_index + 1

  }
}

## Save plots to jpeg. Makes a separate file for each plot.

interpolation_index <- 1
for(sex_index in i_sex) {
  for (measure_index in i_measure) {
    file_name = paste("graphs1/", "Interpolation rates", sex_index, measure_index, ".jpeg", sep=" ")
    jpeg(file_name)
    print(p_interpolation_list[[interpolation_index]])
    interpolation_index <- interpolation_index + 1
    dev.off()
  }
}

p_interpolated <- do.call(marrangeGrob, list(grobs=p_interpolation_list, nrow = 2, ncol = 2))
p_interpolated

```

15) The following code brings together all the required data to create the inputs life table (Figure 3).

```{r, echo=TRUE}

## Create empty data frame

input_life_table <- as.data.frame(matrix(0, ncol=2, nrow = 202))

## Give variables names (mx=total mortality rate, 
## yldsx = total years lived with disability rates )

names(input_life_table) <- c("age", "sex")

## Populate life table: Age, 0 to 100

input_life_table[1:101, 1] <- c(0:100)
input_life_table[102:202, 1] <- c(0:100)

## Populate life table: sex, male, female

input_life_table[1:101, 2] <- "male"
input_life_table[102:202, 2] <- "female"

## Create variable age_sex to match with population data

input_life_table$sex_age_cat <- paste(input_life_table$sex,input_life_table$age,
                                sep = "_"  )
## Populate life table: five_year_pop.

input_life_table <- merge(input_life_table, select(GBD_population_GL, 
                    c(sex_age_cat, five_year_population)), by = 'sex_age_cat', all = TRUE)

## Populate life table: mortality rates (from interpolated rates)
## Males mortality rates. NEED MATCHING NAMES

input_life_table <- merge(input_life_table, select(male_deaths_interpolated, 
                    c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
input_life_table <- merge(input_life_table, select(female_deaths_interpolated,
                    c(sex_age_cat, mx)), by = 'sex_age_cat', all.x = TRUE)
input_life_table <- merge(input_life_table, select(male_ylds_interpolated, 
                    c(sex_age_cat, pyld_rate)) , by = 'sex_age_cat', all.x = TRUE) 
input_life_table <- merge(input_life_table, select(female_ylds_interpolated, 
                    c(sex_age_cat, pyld_rate)), by = 'sex_age_cat', all.x = TRUE)

## DISCUSS WITH ALI TO IMPROVE

input_life_table$mx <-  ifelse(input_life_table$sex == "male",
                        input_life_table$mx.x, input_life_table$mx.y)

input_life_table$pyld_rate <-  ifelse(input_life_table$sex == "male",
                              input_life_table$pyld_rate.x, input_life_table$pyld_rate.y)
## Drop redundant variables
input_life_table <- subset(input_life_table, select = -c(mx.y, mx.x, pyld_rate.x, pyld_rate.y))

## Cross check population numbers

sum(input_life_table$five_year_population, na.rm = TRUE)

## Sort data by sex and age to use in life table function

input_life_table<-input_life_table[order(input_life_table$sex, input_life_table$age),]

```
16) Generate **baseline life tables** from input_data_frame using run_life_table function. Life tables are generated for each age (5-years) and sex cohort. 
It is assumed that current observed rates of mortality and disability are going to be observed in the future.

```{r, echo=TRUE}

i_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)

general_life_table_list_bl <- list()

index <- 1

for (age in i_age_cohort){
  for (sex in i_sex){
    # cat("age ", age, " and sex ", sex, "\n") #Uncomment to see index
    general_life_table_list_bl[[index]] <- run_life_table(in_idata = input_life_table,
    in_sex = sex, in_mid_age = age)
    index <- index + 1
  }
}

## Uncommnet to check life table list
# View(general_life_table_list_bl[[32]])

```

17) Generate **baseline disease life tables** using run_disease function. Inputs of the disease life tables are: incidence, case fatality (Figure 3, purple shaded area) and disability weights. Internally consistent estimates of incidence and case fatality are derived from GBD data (by five-year age groups and sex) and Dismod II. Disability weights are derived from GBD disease specific YLDs and prevalence (by five-year age groups and sex) adjusted for all-cause YLDs (see section 2.1.1.4).  
Note that Dismod II is an external software. The code below is used to generate the inputs for Dismod II using GBD data (see Figure 3, purple shaded area). Dismod II requires data for the *collection* (population numbers and all cause mortality rates) and *data set* (disease specific parameters). We will generate a data set with all inputs for dismod by 5 year-age (represent at mid age) groups and sex for five diseases (ischemic heart disease, ischemic stroke, diabetes mellitus, colon and rectum cancer and breast cancer-women). WE WILL NEED ADJUSTMENT FOR TO MATCH RRS TO DISEASES (COLON AND RECTUM CANCER AND DIABETES MELLITUS). *Code here is work in progress**, for now the disease input data is organised externally with excel.

```{r, echo=FALSE}
## Collection input for Greater London: population numbers (sex and 5-year age groups). 
## Already processed in past code, here, we sort it out and save in cvs format in a Dismod II folder.

####THIS CODE IS INCOMPLETE, I WILL PREPARE THE DATA FOR DISMOD II IN EXCEL AND AFTER PRESENTATION THURSDAY 26 APRIL I WILL DEVELOP THE SCRIPT.

# ## Get population numbers
# 
# data_dismodii <- GBD_population_GL[order(GBD_population_GL$sex, GBD_population_GL$age_cat),] 
# 
# ## Add mortality rates (these were generated for the life table, so we use the same code)
# 
# data_dismodii  <- merge(data_dismodii, select(male_deaths_interpolated
#                                                      ,c(sex_age_cat, mx))
#                                                      , by = 'sex_age_cat', all.x = TRUE)
# data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
#                                                      ,c(sex_age_cat, mx))
#                                                      , by = 'sex_age_cat', all.x = TRUE)
# 
# data_dismodii$mx <-  ifelse(data_dismodii$sex == "male",
# data_dismodii$mx.x, data_dismodii$mx.y)
# ## Drop redundant variables 
# data_dismodii <- subset(data_dismodii, select = -c(mx.y, mx.x))
# ##Sort data
# data_dismodii<-data_dismodii[order(data_dismodii$sex, data_dismodii$age_cat),] 
# ## Add disease rates per one(incidence, prevalence and mortality)
# 
# ## Generate variable to do a loop that creates dataframes for each of the disease inputs.
# 
# i_disease <- c("ischemic heart disease", "ischemic stroke", "diabetes mellitus", "tracheal, bronchus, and lung cancer", "breast cancer", "colon and rectum cancer")
# i_sex <- c("male", "female")
# i_measure <- c("deaths", "ylds", "incidence", "prevalence")
# 
# ## Loop to generate dataframes
# for(sex_index in i_sex) {
#   for (measure_index in i_measure) {
#     for (disease_index in i_disease) {
#    
# data <- filter(GBDGL, measure == measure_index, sex == sex_index, 
# cause == disease_index) %>% select(measure, location, sex, age, metric, 
# cause, val, age_cat, one_rate, sex_age_cat) 
# colnames(data)[9] <- paste(measure_index, disease_index, sep =  "_")
# assign(paste(sex_index, measure_index, disease_index, sep = "_"), data)

## The code below is not generating the multiple datasets saved as csv.

# data_name <- file.path(tempdir(), paste0(sex_index, measure_index, disease_index, ".csv"))
# write.csv(data, file = data_name)
#     }
#   }
# }

## Generate TO DO, ADD disease parameters.

## Generate empy variables in data_dismodii for death, incidence and prevalence rates per disease (MAY NOT NEED TO USE, WITH MERGING VARIABLES)

#  for (measure_index in i_measure) {
#     for (disease_index in i_disease) {
# 
#   # data_dismodii[,paste(measure_index, disease_index, sep =  "_")] <- 0
# data_dismodii  <- merge(data_dismodii, select(female_deaths_interpolated
#                                                      ,c(sex_age_cat, mx))
#                                                      , by = 'sex_age_cat', all.x = TRUE)
#   }
#  }

```

```{r, echo=TRUE}

## Use externaly generated inputs for for disease life table. The previous code will serve to to some of the data preparation for DISMOD. Also need to devlop code to calculated DWs, for now, all generated in excel.

idata <- read.csv("data/UK/idata.csv", stringsAsFactors = F)
## Use run_disease

i_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")

disease_life_table_list_bl <- list()
index <- 1

for (age in i_age_cohort){
  for (sex in i_sex){
    for (disease in i_disease) {
      # Exclude breast_cancer for Males
      if (sex == "male" && disease == "breast_cancer"){
       # cat("\n") #Uncomment to see list
      }
      else {
        # cat("age ", age, " sex ", sex, "and disease", disease, "\n") #Uncomment to see list
        disease_life_table_list_bl[[index]] <- run_disease(in_idata = idata, in_sex = sex, in_mid_age = age, in_disease = disease)
        index <- index + 1
      }
    }
  }
}
## Uncommnet to check disease life table list
# View(disease_life_table_list_bl[[8]])

```

18) Generate mock change in incidence of disease to generate scenario life tables. In the final version, this will come from the calculated PIFs by disease. 
PIFs are applied here to incidence, however, may also be applied to case fatality, depending on the RRs. Also, we can include time delays from change in exposure to change in health outcomes via the PIF.

```{r, echo=TRUE}
## Create value to use as factor changing incidence rates.

incidence_change <- 0.95

## Generate scenario incidence (for each disease)

incidence_sc <- list()
index <- 1

for (age in i_age_cohort){
  for (sex in i_sex){
    for (disease in i_disease) {

      # Exclude breast_cancer for Males
      if (sex == "male" && disease == "breast_cancer"){
        # cat("\n") # Uncomment to see list
      }
      else {

        incidence_sc[[index]] <- disease_life_table_list_bl[[index]]$incidence_disease * 
        incidence_change
        index <- index + 1

      }
    }
  }
}

## Uncommnet to check scenario incidence
# View(incidence_sc[[1]])

```

19) Use scenario incidence to calculate scenario disease life tables.

```{r, echo=TRUE}

disease_life_table_list_sc <- list()
index <- 1
for (age in i_age_cohort){
  for (sex in i_sex){
    for (disease in i_disease) {
      # Exclude breast_cancer for Males
      if (sex == "male" && disease == "breast_cancer"){
        # cat("\n")
      }
      else {
        # cat("age ", age, " sex ", sex, "and disease", disease, "\n")
        # modify idata's incidence for the said scenario
        td1 <- idata
        td1[td1$age >= age & td1$sex == sex,][[paste("incidence", disease, sep = "_")]] <- incidence_sc[[index]]

        # Instead of idata, feed td to run scenarios
        disease_life_table_list_sc[[index]] <- run_disease(in_idata = td1, in_sex = sex,
                                                           in_mid_age = age, in_disease = disease)
        disease_life_table_list_sc[[index]]$diff_inc_disease <- 
                                              disease_life_table_list_sc[[index]]$incidence_disease -   disease_life_table_list_bl[[index]]$incidence_disease
        disease_life_table_list_sc[[index]]$diff_prev_disease <- 
                                              disease_life_table_list_sc[[index]]$px  - disease_life_table_list_bl[[index]]$px
        disease_life_table_list_sc[[index]]$diff_mort_disease <- 
                                              disease_life_table_list_sc[[index]]$mx - disease_life_table_list_bl[[index]]$mx
        disease_life_table_list_sc[[index]]$diff_pylds_disease <- 
                                              (disease_life_table_list_sc[[index]]$px - disease_life_table_list_bl[[index]]$px) * disease_life_table_list_bl[[index]]$dw_disease
        index <- index + 1
      }
    }
  }
}
## Uncommnet to check scenario life tables
# View(disease_life_table_list_sc[[1]])

```

20) Calculate life tables scenario.

```{r, echo=TRUE}

## Generate total change in mortality rate

## Sum mortality rate change scenarios (mx_sc_total)

mx_sc_total <- list()
l_index <- 1
index <- 1
for (age in i_age_cohort){
  for (sex in i_sex){
    mortality_sum <- NULL
    create_new <- T

    for (disease in i_disease) {
      if (sex == "male" && disease == "breast_cancer"){
        # cat("\n")
      }else{

        if (create_new){
          mortality_sum <- select(disease_life_table_list_sc[[index]], 
                            c('age', 'sex'))
                            mortality_sum$total <- 0
          create_new <- F
          mortality_sum$total <- mortality_sum$total + 
                                (disease_life_table_list_sc[[index]]$diff_mort_disease)
        }else{
          mortality_sum$total <- mortality_sum$total +
                                (disease_life_table_list_sc[[index]]$diff_mort_disease)
        }

        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
        index <- index + 1
      }
    }
    mx_sc_total[[l_index]] <- mortality_sum

    l_index <- l_index + 1
  }
}

## Uncommnet to check sceanrio mortality and changes 
# View(mx_sc_total[[3]])

## Generate total change in prevalent yld rates
## Total ylds rate= sum (change prevalence disease*dw)

pylds_sc_total <- list()
l_index <- 1
index <- 1
for (age in i_age_cohort){
  for (sex in i_sex){
    pylds_sum <- NULL
    create_new <- T

    for (disease in i_disease) {
      if (sex == "male" && disease == "breast_cancer"){
        # cat("\n")
      }else{

        if (create_new){
          pylds_sum <- select(disease_life_table_list_sc[[index]], c('age', 'sex'))
          pylds_sum$total <- 0
          create_new <- F
          pylds_sum$total <- pylds_sum$total +   
                            (disease_life_table_list_sc[[index]]$diff_pylds_disease)
        }else{
          pylds_sum$total <- pylds_sum$total + 
                            (disease_life_table_list_sc[[index]]$diff_pylds_disease)
        }

        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
        index <- index + 1
      }

    }
    pylds_sc_total[[l_index]] <- pylds_sum
    l_index <- l_index + 1
  }
}

## Uncommnet to check scenario pyld change
# View(pylds_sc_total[[2]])

## Calculate general life tables with modified mortality and pylds total
## Original mortality rate is modified by the mx_sc_total (total change in mortality from diseases)
## Original pyld rate is modified by the change in each disease pylds

general_life_table_list_sc <- list()
index <- 1

for (age in i_age_cohort){
  for (sex in i_sex){

    cat("age ", age, " and sex ", sex, "\n")
    # modify idata's mortality and pyld total for the said scenario
    td2 <- input_life_table
    # td2 <- subset(td2, select = -c(mx, pyld_rate))
    td2[td2$age >= age & td2$sex == sex,][[paste("mx")]] <- general_life_table_list_bl[[index]]$mx + mx_sc_total[[index]]$total
    td2[td2$age >= age & td2$sex == sex,][[paste("pyld_rate")]] <- general_life_table_list_bl[[index]]$pyld_rate + pylds_sc_total[[index]]$total

    # Instead of idata, feed td to run scenarios
    general_life_table_list_sc[[index]] <- run_life_table(in_idata = td2, in_sex = sex, in_mid_age = age)
    #

    index <- index + 1
  }
}
## Uncommnet to check scenario life tables
# View(general_life_table_list_sc[[32]])
# View(general_life_table_list_bl[[1]])

## Check difference life table baseline and scenario
general_life_table_list_bl[[1]]$Lx - general_life_table_list_sc[[1]]$Lx
general_life_table_list_bl[[1]]$Lwx - general_life_table_list_sc[[1]]$Lwx

```

21) Generate list of outputs by age and sex.

```{r, echo=TRUE}

## In the following list "output_life_table", 32 data frames are nested per age and sex cohort

output_burden <- list()
l_index <- 1
index <- 1
for (age in i_age_cohort){
  for (sex in i_sex){

    # Males do not have breast cancer, that is why we need the if/else. 
    # We create a TRUE/FALSE variable for the loop to move into the next disease

    create_new <- T
    for (disease in i_disease) {
      if (sex == "male" && disease == "breast_cancer"){
        # cat("\n")
      }else{

        if (create_new){
          output_burden_sc <- select(disease_life_table_list_sc[[index]], 
                              c('age', 'sex', 'incidence_disease', 'mx', 'px'))
          names(output_burden_sc)[names(output_burden_sc) == 'incidence_disease'] <- 
                              paste('incidence_disease', disease, "sc", sep = "_")
          names(output_burden_sc)[names(output_burden_sc) == 'mx'] <- 
                              paste('mx', disease, "sc", sep = "_")
          names(output_burden_sc)[names(output_burden_sc) == 'px'] <- 
                              paste('px', disease, "sc", sep = "_")
          output_burden_bl <- select(disease_life_table_list_bl[[index]], 
                              c('incidence_disease', 'mx', 'px'))
          names(output_burden_bl)[names(output_burden_bl) == 'incidence_disease'] <- 
                              paste('incidence_disease', disease, "bl", sep = "_")
          names(output_burden_bl)[names(output_burden_bl) == 'mx'] <- 
                              paste('mx', disease, "bl", sep = "_")
          names(output_burden_bl)[names(output_burden_bl) == 'px'] <- 
                              paste('px', disease, "bl", sep = "_")

          ## New list to add calculations for changes in burden of disease (incidence and mortality numbers)

          output_burden_change <- list()

          output_burden_change$inc_num_bl <- disease_life_table_list_bl[[index]]$incidence_disease * 
          (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx
          output_burden_change$inc_num_sc <- disease_life_table_list_sc[[index]]$incidence_disease * 
          (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx
          output_burden_change$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * 
          (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) 
          * general_life_table_list_bl[[l_index]]$Lx)

          output_burden_change$mx_num_bl <- disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx
          output_burden_change$mx_num_sc <- disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx
          output_burden_change$mx_num_diff <- (disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx)

          names(output_burden_change)[names(output_burden_change) == 'inc_num_bl'] <- 
          paste('inc_num_bl', disease, sep = "_")
          names(output_burden_change)[names(output_burden_change) == 'inc_num_sc'] <- 
          paste('inc_num_sc', disease, sep = "_")
          names(output_burden_change)[names(output_burden_change) == 'inc_num_diff'] <- 
          paste('inc_num_diff', disease, sep = "_")
          names(output_burden_change)[names(output_burden_change) == 'mx_num_bl'] <- 
          paste('mx_num_bl', disease, sep = "_")
          names(output_burden_change)[names(output_burden_change) == 'mx_num_sc'] <- 
          paste('mx_num_sc', disease, sep = "_")
          names(output_burden_change)[names(output_burden_change) == 'mx_num_diff'] <- 
          paste('mx_num_diff', disease, sep = "_")

          ## Bind all lists

          output_burden_sc <- cbind(output_burden_sc, output_burden_bl)
          output_burden_sc <- cbind(output_burden_sc, output_burden_change)

          create_new <- F

          ## Here the calculations above are repeated, here is where the F is telling to move into the next disease

          }else{

          td3 <- select(disease_life_table_list_sc[[index]], 
                c('incidence_disease', 'mx', 'px'))
          names(td3)[names(td3) == 'incidence_disease'] <-
          paste('incidence_disease', disease, "sc", sep = "_")
          names(td3)[names(td3) == 'mx'] <- 
          paste('mx', disease, "sc", sep = "_")
          names(td3)[names(td3) == 'px'] <- 
          paste('px', disease, "sc", sep = "_")

          td4 <- select(disease_life_table_list_bl[[index]], 
                c('incidence_disease', 'mx', 'px'))
          names(td4)[names(td4) == 'incidence_disease'] <- 
          paste('incidence_disease', disease, "bl", sep = "_")
          names(td4)[names(td4) == 'mx'] <- 
          paste('mx', disease, "bl", sep = "_")
          names(td4)[names(td4) == 'px'] <- 
          paste('px', disease, "bl", sep = "_")

          output_burden_change2 <- list()

          output_burden_change2$inc_num_bl <- disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx
          output_burden_change2$inc_num_sc <- disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx
          output_burden_change2$inc_num_diff <- (disease_life_table_list_sc[[index]]$incidence_disease * (1 - disease_life_table_list_sc[[index]]$px) * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$incidence_disease * (1 - disease_life_table_list_bl[[index]]$px) * general_life_table_list_bl[[l_index]]$Lx)

          output_burden_change2$mx_num_bl <- disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx
          output_burden_change2$mx_num_sc <- disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx
          output_burden_change2$mx_num_diff <- (disease_life_table_list_sc[[index]]$mx * general_life_table_list_sc[[l_index]]$Lx) - (disease_life_table_list_bl[[index]]$mx * general_life_table_list_bl[[l_index]]$Lx)

          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_bl'] <- 
          paste('inc_num_bl', disease, sep = "_")
          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_sc'] <- 
          paste('inc_num_sc', disease, sep = "_")
          names(output_burden_change2)[names(output_burden_change2) == 'inc_num_diff'] <- 
          paste('inc_num_diff', disease, sep = "_")
          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_bl'] <- 
          paste('mx_num_bl', disease, sep = "_")
          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_sc'] <- 
          paste('mx_num_sc', disease, sep = "_")
          names(output_burden_change2)[names(output_burden_change2) == 'mx_num_diff'] <- 
          paste('mx_num_diff', disease, sep = "_")

          ## Bind all lists

          output_burden_sc <- cbind(output_burden_sc, td3)
          output_burden_sc <- cbind(output_burden_sc, td4)
          output_burden_sc$age_cohort <- age 
          output_burden_sc <- cbind(output_burden_sc, output_burden_change2)

        }

        # cat(age, " - ", sex," - ",  disease," - ",  index, " - ", l_index,  "\n")
        index <- index + 1
      }

    }

    ## general_life_table_list_sc and general_life_table_list_bl (Lx)
    output_burden_lf_sc <- select(general_life_table_list_sc[[l_index]], c('Lx', 'Lwx'))
    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lx'] <- paste('Lx', "sc", sep = "_")
    names(output_burden_lf_sc)[names(output_burden_lf_sc) == 'Lwx'] <- paste('Lwx', "sc", sep = "_")

    output_burden_lf_bl <- select(general_life_table_list_bl[[l_index]], c('Lx', 'Lwx'))
    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lx'] <- paste('Lx', "bl", sep = "_")
    names(output_burden_lf_bl)[names(output_burden_lf_bl) == 'Lwx'] <- paste('Lwx', "bl", sep = "_")

    output_burden_lf_sc$Lx_diff <- general_life_table_list_bl[[l_index]]$Lx - general_life_table_list_sc[[l_index]]$Lx
        output_burden_lf_sc$Lwx_diff <- general_life_table_list_bl[[l_index]]$Lwx - general_life_table_list_sc[[l_index]]$Lwx

    output_burden_sc <- cbind(output_burden_sc, output_burden_lf_sc)
    output_burden_sc <- cbind(output_burden_sc, output_burden_lf_bl)

    output_burden[[l_index]] <- output_burden_sc
    l_index <- l_index + 1

  }
}

## Uncomment to check
# View(output_burden[[1]])

```

22) Combine all lists of outputs in a data frame to facilitate extraction of outcomes of interest and plotting.

```{r, echo=TRUE}
#####Generate a data frame for all results and create function to get outcomes.

output_df <- plyr::ldply(output_burden, rbind)

# View(output_df)

#Remove variables that are not used in the generation of outputs. CHANGE THIS NAMES, TOO LONG

output_df <- subset(output_df, select = -c(incidence_disease_ihd_bl, incidence_disease_ihd_sc, incidence_disease_istroke_bl, incidence_disease_istroke_sc, incidence_disease_diabetes_bl, incidence_disease_diabetes_sc, incidence_disease_breast_cancer_bl, incidence_disease_breast_cancer_sc, 
incidence_disease_colon_cancer_bl, incidence_disease_colon_cancer_sc,  mx_ihd_bl, mx_ihd_sc, mx_istroke_bl, mx_istroke_sc, mx_diabetes_bl, mx_diabetes_sc, mx_breast_cancer_bl, mx_breast_cancer_sc, mx_colon_cancer_bl, mx_colon_cancer_sc, px_ihd_bl, px_ihd_sc, px_istroke_bl, px_istroke_sc, px_diabetes_bl, px_diabetes_sc, px_breast_cancer_bl, px_breast_cancer_sc, px_colon_cancer_bl, px_colon_cancer_sc))

```

23) Plot outcomes for each age and sex cohort using created functions. Four functions are used: plot_output (plots outputs), grid_arrange_shared_legend and g_legend (places one label per page) and get_qualified_disease_name (replaces full name of disease)

```{r echo=TRUE}

output_dir = "graphs2"

i_outcome <- c("mx", "inc")
p_list_male <- list()
p_list_female <- list()
male_index <- 1
female_index <- 1
for (age in i_age_cohort){
      for (sex in i_sex) {
        for (outcome in i_outcome) {
          for (disease in i_disease){

       if (sex == "male" && disease == "breast_cancer"){
         # cat("\n")
       }else{

p_index  <- plot_output(in_data = output_df, in_age = age, in_population = sex, in_outcomes = c("age", paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")), in_disease = get_qualified_disease_name(disease))

            if (sex == "male"){

              p_list_male[[male_index]] <- p_index

              if (male_index %% 4 == 0 && male_index > 0){

                p1 <- p_list_male[[male_index - 3]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
                p2 <- p_list_male[[male_index - 2]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
                p3 <- p_list_male[[male_index - 1]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
                p4 <- p_index + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())

  jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
  grid_arrange_shared_legend (p1, p2, p3, p4, ncol = 2, nrow = 2, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age),
                                          mainLeft = 'Cases', mainBottom = 'Age')
                dev.off()

              }

              male_index <- male_index + 1

            }

            if (sex == "female" && female_index > 0){
              p_list_female[[female_index]] <- p_index

              if (female_index %% 5 == 0){
                p1 <- p_list_female[[female_index - 4]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
                p2 <- p_list_female[[female_index - 3]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
                p3 <- p_list_female[[female_index - 2]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
                p4 <- p_list_female[[female_index - 1]] + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())
                p5 <- p_index + theme(legend.position="none", axis.title.x = element_blank(),  axis.title.y = element_blank())

               jpeg(paste0(output_dir, "/", paste(age, sex, outcome, sep="_"), ".jpeg"))
                grid_arrange_shared_legend (p1, p2, p3, p4, p5, ncol = 2, nrow = 3, mainTitle = paste(ifelse(outcome == "mx", "Deaths", "Incidence"), sex, "cohort mid age", age), mainLeft = 'Cases', mainBottom = 'Age')
                dev.off()

                }
              female_index <- female_index + 1
              }

            }
         }
       }
   }
}

```

```{r echo=FALSE, out.width='100%'}
## Loop to include graphs in the document

graphs_doc <- list()
index <- 1
for (age in i_age_cohort) {
      for (sex in i_sex)  {
        for (outcome in i_outcome) {

graphs_doc [[index]] <- c(paste(output_dir, "/",age,"_",sex,"_", outcome,".jpeg", sep = ""))
knitr::include_graphics(graphs_doc [[index]])

          index <- index + 1
    }
  }
}

 knitr::include_graphics(graphs_doc[c(1:64)])
```

24) Add up all outcomes per year of simulation with generated function. BETTER TO HAVE A LOOP, I TRIED BY LOOPING IN THE FUNCITON ARGUMENT I_OUTCOMES, BUT THIS DID NOT WORK. THEN, IF LOOP HERE ALSO IN NEXT CODE.

```{r, echo=TRUE}

## Try binding function

aggregate_frame_males <- list()
aggregate_frame_females <- list()

index <- 1

for (outcome in i_outcome) {
  for (disease in i_disease) {

aggregate_frame_males[[index]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c(paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")))

aggregate_frame_females[[index]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c(paste(outcome, "num", "bl", disease, sep = "_"), paste(outcome, "num", "sc", disease, sep = "_"), paste(outcome, "num", "diff", disease, sep = "_")))

# Remove non-numeric columns starting with age and sex
aggregate_frame_males[[index]] <- aggregate_frame_males[[index]] %>% select(-starts_with("age"), -starts_with("sex"))

aggregate_frame_females[[index]] <- aggregate_frame_females[[index]] %>% select(-starts_with("age"), -starts_with("sex"))

index <- index + 1
  }
}

## Loop for life years (Lx) and health adjusted life years (Lwx) to then add to total aggregated data.

## RUN ANOTHER LOOP FOR LX AND LWX

i_outcome2 <- c("Lx", "Lwx")

aggregate_frame_males2 <- list()
aggregate_frame_females2 <- list()

for (i in i_outcome2){

aggregate_frame_males2[[i]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "male", in_outcomes = c(paste(i, "bl", sep = "_"), paste(i, "sc", sep = "_"), paste(i, "diff",sep = "_")))

aggregate_frame_females2[[i]] <- gen_aggregate(in_data = output_df, in_cohorts = 16, in_population = "female", in_outcomes = c(paste(i, "bl", sep = "_"), paste(i,  "sc", sep = "_"), paste(i, "diff",sep = "_")))

aggregate_frame_males2[[i]] <- aggregate_frame_males2[[i]] %>% select(-starts_with("age"), -starts_with("sex"))

aggregate_frame_females2[[i]] <- aggregate_frame_females2[[i]] %>% select(-starts_with("age"), -starts_with("sex"))

}

## Uncomment to check data frames (list correspond to a disease and outcome combination, for example,
## incidence breast cancer)

# View(aggregate_frame_females[[2]])

## Transform list to data frame (this is not working after the first disease)

aggregate_frame_females <- do.call(cbind, aggregate_frame_females)
aggregate_frame_males <- do.call(cbind, aggregate_frame_males)
aggregate_frame_females2 <- do.call(cbind, aggregate_frame_females2)
aggregate_frame_males2 <- do.call(cbind, aggregate_frame_males2)

View(aggregate_frame_females2)

## Drop ending of variables names _males/_females to add up all outcomes in next step.

names(aggregate_frame_females) = gsub(pattern = "_female", replacement = "", x = names(aggregate_frame_females))

names(aggregate_frame_males) = gsub(pattern = "_male", replacement = "", x = names(aggregate_frame_males))

names(aggregate_frame_females2) = gsub(pattern = "_female", replacement = "", x = names(aggregate_frame_females2))

names(aggregate_frame_males2) = gsub(pattern = "_male", replacement = "", x = names(aggregate_frame_males2))

## Uncomment to check that the code is dropping the male/female ending. 
# View(aggregate_frame_males)

## Create a copy of aggregate_frame_females.  
total_aggr1 <- aggregate_frame_females
## Add aggregate_frame_males values to it
for (i in 1:ncol(aggregate_frame_females)){
  total_aggr1[i] <- total_aggr1[i] + aggregate_frame_males[i]
}

## Add data frames 2 with life years (check adds Lx and Lwx at the beginning of the variable name)
total_aggr2 <- aggregate_frame_females2
## Add aggregate_frame_males values to it
for (i in 1:ncol(aggregate_frame_females2)){
  total_aggr2[i] <- total_aggr2[i] + aggregate_frame_males2[i]
}

## Combine data frames

total_aggr <- cbind.data.frame(total_aggr1, total_aggr2)

total_aggr$sim_year <- seq.int(nrow(total_aggr))

## Uncomment to see total data frame
View(total_aggr)

## Test that total_aggr is adding males and females dataframes
# 
# (aggregate_frame_females2$Lx.Lx_bl_22 + aggregate_frame_males2$Lx.Lx_bl_22) -total_aggr$Lx.Lx_bl_22

```

25) Plots using total aggregated change in burden. This graphs are not very good for life years and health adjusted life years as the difference between baseline and scenario is very small relative to the total, hence the graphs show almost no change.

```{r, echo=TRUE}

####This plot has to be customised to in_outcomes, here, only totals shown, but specifications are up to the user. ADD LOOP for all outcomes over time and total TABLE.

####[] is used here to indicate the number of simulation years into the future. 
####Disease outcomes has to be changed to the outcome of interest

#### Test code with loops for aggregated outcomes diseases burden. NOT WORKING.

### Compare with loops for age and sex cohort outcomes.

p_aggr_list <- list()
index <- 1

for (outcome in i_outcome) {
  for (disease in i_disease) {
    # outcome <- i_outcome[1]
    # disease <- i_disease[1]

p_aggr_list_index <- ggplot(total_aggr[1:79,], aes(x = total_aggr[["sim_year"]])) +

  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_bl", disease, sep = "_")]], colour = paste("total", outcome, "num_bl", disease, sep = "_"))) +
  theme_classic() +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_sc", disease, sep = "_")]], colour = paste("total", outcome, "num_sc", disease, sep = "_"))) +
  geom_line(mapping = aes(y = total_aggr[[paste("total", outcome, "num_diff", disease, sep = "_")]], colour = paste("total", outcome, "num_diff", disease, sep = "_"))) +
  xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease, outcome)) +
  theme(plot.title = element_text(hjust = 0.5, size = 12)) +
  scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
  theme(plot.title = element_text(hjust = 0.5))
p_aggr_list[[index]] <- p_aggr_list_index  
index <- index + 1

  }
}

index <- 1

interpolation_index <- 1
for (outcome in i_outcome) {
  for (disease in i_disease) {
    file_name = paste("graphs2/", "Aggregated Outcomes", outcome, disease, ".jpeg", sep=" ")
    jpeg(file_name)
    print(p_aggr_list[[index]])
    index <- index + 1
    dev.off()
  }
}

p_aggregated <- do.call(marrangeGrob, list(grobs=p_aggr_list, nrow = 2, ncol = 2))
p_aggregated

#####
## original code with single outcomes
# disease_outcome <- "Incidence of Ischemic Heart Disease"
# 
# total_plot <- ggplot(total_aggr[1:79,], aes(x = sim_year)) +
#   
#   geom_line(mapping = aes(y = total_aggr[["total_inc_num_bl_ihd"]], colour = "total_inc_num_bl_ihd")) +
#   theme_classic() +
#   geom_hline(yintercept=0, linetype="dashed", color = "black") +
#   geom_line(mapping = aes(y = total_aggr[["total_inc_num_sc_ihd"]], colour = "total_inc_num_sc_ihd")) +
#   geom_line(mapping = aes(y = total_aggr[["total_inc_num_diff_ihd"]], colour = "total_inc_num_diff_ihd")) +
#   xlab ("Simulation years") + ylab ("Cases") + labs (title = paste(disease_outcome)) +
#   theme(plot.title = element_text(hjust = 0.5, size = 12)) +
#   scale_color_discrete(name = paste(""), labels = c("Baseline", "Difference", "Scenario")) +
#   theme(plot.title = element_text(hjust = 0.5))
#   
# 
# #Print to view
# print(total_plot)

```

#Comments

##Road injuries in the PMsLT

The disease model used in each of the disease life table is not directly applicable to road injuries, however, similar concept can be follow. Firstly, changes in road fatalities impact on the overall mortality rate, hence, by knowing the road fatality rates for baseline and scenarios, we will be able to incorporate changes to mortality attributable to road fatalities. For road injuries, methods developed by Kavi Bhalla and Marko Tanio (REFS) that derive the average YLD attributable to life long and short term injuries can be applied to derive the change in total YLDs (CHECK THAT THESE WERE DEVELOPED AS INCIDENCE YLDs).MT's methods assumes that injuries do not reduce the life expectancy of the injured person.

#References

### 2018-06-26 MSLT/code/functions.R

@@ -1,504 +1,0 @@
-#####Packages
-
-require(dplyr)
-require(tidyverse)
-require(knitr)
-require(kableExtra)
-require(citr)
-
-#############################Explanation method###################################
-
-#####Briefly, the proportional multi-state multi cohort life table consists of a 
-#####general life table and a life table for each of the modelled diseases.
-#####The diseases are those associated to the studied risk factor/s. 
-#####The link between the general life table and disease life tables is via the 
-#####potential impact fraction (pif), also called paf (population attributable fraction)
-#####The pif combines exposure to the risk factor and relative risks. The pif is 
-#####appleid to modify incidence in the individual disease life tables, which in turn
-#####modify prevalence and mortality. Changes in mortality and prevalence rates
-#####feed bacak into the general life table to modify total mortality and disability. 
-####Changes in total mortality impact on life years and changes in disability impact 
-####the disability adjusment of life years. 
-
-#### Method reference: 1.	Barendregt JJ, Oortmarssen vGJ, Murray CJ, Vos T. A generic model for the assessment of disease epidemiology: the computational basis of DisMod II. Popul Health Metr. 2003;1(1):4-.
-
-
-####The functions: run_life_table, run_disease, and run_pif serve to generate outputs
-####per age and sex (see model.R). The function run_output to facilitate the generation
-####of results. 
-
-
-#############################Function for general life table###################################
-#####Function to generate age and sex life tables. The function is then use in the model.R script
-#####to calculate variables for the baseline and scenario life tables. 
-
-run_life_table <- function(in_idata, in_sex, in_mid_age)
-{
-  
-  # Create a life table data frame
-  
-  # Filter in_idata by age and select columns for lifetable calculations
[…]

### 2018-06-26 MSLT/code/model.R

@@ -1,589 +1,0 @@
-##########################Prepare data and environment to run code###########################################################
-
-##### Clean the Global Environment
-
-rm (list = ls())
-
-##### Prevent scientific notation in data frame
-
-options(scipen=999)
-
-##### Set WD
-
-
-
-##### Read data. idata: life table and disease life tables (including trends). edata: exposure data (e.g. physical activity).
-#### irr.data: relative risks data and ee: energy expenditure
-#### For exposure (edata) and input (idata) data I am using already generated categories, but a function should be developed to work from raw data. 
-
-idata <- read.csv("MSLT/data/Australia/idata.csv", stringsAsFactors = F)
-edata <- read.csv("MSLT/data/Australia/edata.csv", stringsAsFactors = F)
-irr <- read.csv("MSLT/data/Australia/irr.csv", stringsAsFactors = F)
-ee <- read.csv("MSLT/data/Australia/ee.csv", stringsAsFactors = F)
-
-#####To do
-
-## Note that disease trends are excluded for now. We need to generate a trends function for diseases, mortality (all cause), 
-## change in disease due to change in risk factor exposure (discuss with Niel M), road injuries effect over time. 
-## Eventually, we want to add (1) CRA ITHIM approach to compare results and (2) use all cause mortaltiy instead of 
-## disease specific dose responses for incidence to understand the difference in results for all cause mortality. 
-
-#################################Model parameters#########################################################################
-
-p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
-
-p_sex <- c("males", "females")
-
-p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
-
-###As an expample, an increase in 100 MET-mins per week
-
-p_intervention_effect <- 100
-
-##############################Prepare general life table (age, death rate, population #)########
-
-##### Make lower cases values for sex column
-
-idata$sex <- tolower(idata$sex)
-
-##### Rename mortality_rate to mx
-
-idata$mx <- idata$mortality_rate
-
-##### Remove mortality_rate column
-
-idata$mortality_rate <- NULL
-
-##### Rename male for males and female for females in edata to match idata. 
-
-edata$sex[edata$sex=="male"] <- "males"
[…]

### 2018-06-26 PA/code/functions.R

@@ -1,130 +1,0 @@
-# Function to clear columns with labelled class
-# taken from: https://stackoverflow.com/a/24070958
-clear.labels <- function(x) {
-  if(is.list(x)) {
-    for(i in 1 : length(x)) class(x[[i]]) <- setdiff(class(x[[i]]), 'labelled') 
-    for(i in 1 : length(x)) attr(x[[i]],"label") <- NULL
-  }
-  else {
-    class(x) <- setdiff(class(x), "labelled")
-    attr(x, "label") <- NULL
-  }
-  return(x)
-}
-
-
-create.lookups <- function(df, cols){
-  lt <- data.frame("name" = as.character(), "id" = as.integer(), "val" = as.character())
-  for (i in 1:length(cols)){
-    val <- as.character(unique(as_factor(df[[cols[i]]], labels = "values")))
-    id <- unique(df[[cols[i]]])
-    m <- matrix(c(rep(cols[i], length(id)), id, val), nrow = length(id), ncol = 3)
-    lt <- rbind(lt, as.data.frame(m, stringsAsFactors = F))
-  }
-  colnames(lt) <- c("names", "id", "val")
-  lt$id <- as.numeric(lt$id)
-  lt
-}
-
-
-mmet2RR <- function(m, cn){
-    for (i in 1:length(cn)){
-      lcn <- as.character(cn[i])
-      m[[lcn]]  <- apply(data.frame(m[[lcn]]), 1, function(x) mmet2RRVal(x[1]))
-    }
-    m
-  }
-
-
-mmet2RRVal <-function(val) {
-  if ((!is.null( val) && !is.na(val))){
-    # Cap values at 35 mmet
-    mmet2RR_mat[which.min(abs(ifelse(mmet2RR_mat$dose < 35, mmet2RR_mat$dose, 35)  - val)), 2]
-  }
-  else
-    0
-}
-
-
-PAF <- function(pop, attr, cn){
-  # pop = rr
-  # attr = c('female', 'agecat')
-  # cn = c('total_mmet', 'total_mmet_sc')
-  
-  unique_gender <- unique(pop[[attr[1]]])
-  unique_age_group <- unique(pop[[attr[2]]])
-  combinations <- length(unique_age_group)*length(unique_gender)
-  
-  m = matrix(nrow=combinations,ncol=(2 + length(cn)))
-  
-  colnames(m) <- append(c("age band", "gender"), cn)
-  mi <- 1
-  
[…]

### 2018-06-26 PA/code/script.R

@@ -1,129 +1,0 @@
-# Remove everything
-rm (list = ls())
-# Load packages
-library(tidyverse)
-library(haven)
-
-#Set seed
-set.seed(1)
-
-# Load all functions
-source("PA/code/functions.R")
-
-# Constants
-# Initialize  energy expenditure constants - taken from ICT
-METCycling <- 5.63
-METWalking <- 3.53
-METEbikes <- 4.50
-
-# Read combined individual travel survey and Physical Activity data
-raw_data <- haven::read_dta("PA/data/SPtrip_CensusNTSAPS_E06000001.dta")
-
-# Create id columns for all trips (and people without any trips)
-raw_data$id <- 1:nrow(raw_data)
-
-# Create a lookup table for selected columns - labelled columns from stata
-lt <- create.lookups(raw_data, c("female", "agecat", "agecat_det", "trip_mainmode"))
-
-# Sort it
-lt <- arrange(lt, names, id)
-
-
-# Sample 10k unique IDs
-# Select trips for the 10k people
-baseline <- raw_data %>% filter(census_id %in% sample(unique(census_id), 1000)) 
-
-# Remove labels
-baseline <- clear.labels(baseline)
-# Replace NAs with 0
-baseline$trip_cycletime_min <- ifelse(is.na(baseline$trip_cycletime_min), 0, baseline$trip_cycletime_min)
-# Replace NAs with 0
-baseline$trip_walktime_min <- ifelse(is.na(baseline$trip_walktime_min), 0, baseline$trip_walktime_min)
-# Calculate trip_cycletime_hr by dividing trip_cycletime_min by 60
-baseline$trip_cycletime_hr <- baseline$trip_cycletime_min / 60
-# Calculate trip_walktime_hr by dividing trip_walktime_min by 60
-baseline$trip_walktime_hr <- baseline$trip_walktime_min / 60
-
-# Get total individual level walking and cycling and sport mmets 
-individual_mmet <- baseline %>% group_by(census_id) %>% summarise (female = first(female), 
-                                                                   agecat_det = first(agecat_det), 
-                                                                   cycleNTS_wkhr = sum(trip_cycletime_hr),
-                                                                   walkNTS_wkhr = sum(trip_walktime_hr),
-                                                                   sport_wkmmets = first(sport_wkmmets))
-
-
[…]

### 2018-06-26 R/MSLT/Package/PMSLT.R

@@ -1,0 +1,8 @@
+###Packages
+
+require(c("devtools", "roxygen2", "testthat", "knitr"))
+
+###Use the following code to access new devtools functions as I develop them. This is particularly important during the development of the book.
+
+
+devtools::install_github("hadley/devtools")

### 2018-06-26 R/MSLT/code/functions.R

@@ -1,0 +1,504 @@
+#####Packages
+
+require(dplyr)
+require(tidyverse)
+require(knitr)
+require(kableExtra)
+require(citr)
+
+#############################Explanation method###################################
+
+#####Briefly, the proportional multi-state multi cohort life table consists of a 
+#####general life table and a life table for each of the modelled diseases.
+#####The diseases are those associated to the studied risk factor/s. 
+#####The link between the general life table and disease life tables is via the 
+#####potential impact fraction (pif), also called paf (population attributable fraction)
+#####The pif combines exposure to the risk factor and relative risks. The pif is 
+#####appleid to modify incidence in the individual disease life tables, which in turn
+#####modify prevalence and mortality. Changes in mortality and prevalence rates
+#####feed bacak into the general life table to modify total mortality and disability. 
+####Changes in total mortality impact on life years and changes in disability impact 
+####the disability adjusment of life years. 
+
+#### Method reference: 1.	Barendregt JJ, Oortmarssen vGJ, Murray CJ, Vos T. A generic model for the assessment of disease epidemiology: the computational basis of DisMod II. Popul Health Metr. 2003;1(1):4-.
+
+
+####The functions: run_life_table, run_disease, and run_pif serve to generate outputs
+####per age and sex (see model.R). The function run_output to facilitate the generation
+####of results. 
+
+
+#############################Function for general life table###################################
+#####Function to generate age and sex life tables. The function is then use in the model.R script
+#####to calculate variables for the baseline and scenario life tables. 
+
+run_life_table <- function(in_idata, in_sex, in_mid_age)
+{
+  
+  # Create a life table data frame
+  
+  # Filter in_idata by age and select columns for lifetable calculations
[…]

### 2018-06-26 R/MSLT/code/model.R

@@ -1,0 +1,589 @@
+##########################Prepare data and environment to run code###########################################################
+
+##### Clean the Global Environment
+
+rm (list = ls())
+
+##### Prevent scientific notation in data frame
+
+options(scipen=999)
+
+##### Set WD
+
+
+
+##### Read data. idata: life table and disease life tables (including trends). edata: exposure data (e.g. physical activity).
+#### irr.data: relative risks data and ee: energy expenditure
+#### For exposure (edata) and input (idata) data I am using already generated categories, but a function should be developed to work from raw data. 
+
+idata <- read.csv("MSLT/data/Australia/idata.csv", stringsAsFactors = F)
+edata <- read.csv("MSLT/data/Australia/edata.csv", stringsAsFactors = F)
+irr <- read.csv("MSLT/data/Australia/irr.csv", stringsAsFactors = F)
+ee <- read.csv("MSLT/data/Australia/ee.csv", stringsAsFactors = F)
+
+#####To do
+
+## Note that disease trends are excluded for now. We need to generate a trends function for diseases, mortality (all cause), 
+## change in disease due to change in risk factor exposure (discuss with Niel M), road injuries effect over time. 
+## Eventually, we want to add (1) CRA ITHIM approach to compare results and (2) use all cause mortaltiy instead of 
+## disease specific dose responses for incidence to understand the difference in results for all cause mortality. 
+
+#################################Model parameters#########################################################################
+
+p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
+
+p_sex <- c("males", "females")
+
+p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
+
+###As an expample, an increase in 100 MET-mins per week
+
+p_intervention_effect <- 100
+
+##############################Prepare general life table (age, death rate, population #)########
+
+##### Make lower cases values for sex column
+
+idata$sex <- tolower(idata$sex)
+
+##### Rename mortality_rate to mx
+
+idata$mx <- idata$mortality_rate
+
+##### Remove mortality_rate column
+
+idata$mortality_rate <- NULL
+
+##### Rename male for males and female for females in edata to match idata. 
+
+edata$sex[edata$sex=="male"] <- "males"
[…]

### 2018-06-26 R/MSLT/mslt_code.r

@@ -1,0 +1,965 @@
+
+# ---- chunk-intro ----
+require(dplyr)
+require(tidyverse)
+require(knitr)
+require(kableExtra)
+require(citr)
+require(gridExtra)
+require(cowplot)
+require(ggpubr)
+require(grid)
+require(ggplot2)
+#source("code/functions.R")
+
+
+# ---- chunk-1 ----
+rm (list = ls())
+
+# ---- chunk-2 ----
+options(scipen=999)
+
+# ---- chunk-3 ----
+source("code/functions.R")
+
+# ---- chunk-4 ----
+GBDdata <- read.csv("data/legacy/UK/gbd2016.csv", stringsAsFactors = F)
+
+
+# ---- chunk-5 ----
+GBDdata <- mutate_all(GBDdata, funs(tolower))
+
+GBDdata <- GBDdata %>% mutate(measure = ifelse(measure == "ylds (years lived with disability)", "ylds", measure))
+
+# ---- chunk-6 ----
+# 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97
+GBDdata$age_cat [GBDdata$age =="under 5"] <- 2
+GBDdata$age_cat [GBDdata$age =="5 to 9"] <- 7
+GBDdata$age_cat [GBDdata$age =="10 to 14"] <- 12
+GBDdata$age_cat [GBDdata$age =="15 to 19"] <- 17
+GBDdata$age_cat [GBDdata$age =="20 to 24"] <- 22
+GBDdata$age_cat [GBDdata$age =="25 to 29"] <- 27
+GBDdata$age_cat [GBDdata$age =="30 to 34"] <- 32
+GBDdata$age_cat [GBDdata$age =="35 to 39"] <- 37
+GBDdata$age_cat [GBDdata$age =="40 to 44"] <- 42
+GBDdata$age_cat [GBDdata$age =="45 to 49"] <- 47
+GBDdata$age_cat [GBDdata$age =="50 to 54"] <- 52
+GBDdata$age_cat [GBDdata$age =="55 to 59"] <- 57
+GBDdata$age_cat [GBDdata$age =="60 to 64"] <- 62
+GBDdata$age_cat [GBDdata$age =="65 to 69"] <- 67
+GBDdata$age_cat [GBDdata$age =="70 to 74"] <- 72
+GBDdata$age_cat [GBDdata$age =="75 to 79"] <- 77
[…]

### 2018-06-26 R/PA/code/functions.R

@@ -1,0 +1,136 @@
+# Function to clear columns with labelled class
+# taken from: https://stackoverflow.com/a/24070958
+clear.labels <- function(x) {
+  if(is.list(x)) {
+    for(i in 1 : length(x)) class(x[[i]]) <- setdiff(class(x[[i]]), 'labelled') 
+    for(i in 1 : length(x)) attr(x[[i]],"label") <- NULL
+  }
+  else {
+    class(x) <- setdiff(class(x), "labelled")
+    attr(x, "label") <- NULL
+  }
+  return(x)
+}
+
+
+create.lookups <- function(df, cols){
+  lt <- data.frame("name" = as.character(), "id" = as.integer(), "val" = as.character())
+  for (i in 1:length(cols)){
+    val <- as.character(unique(as_factor(df[[cols[i]]], labels = "values")))
+    id <- unique(df[[cols[i]]])
+    m <- matrix(c(rep(cols[i], length(id)), id, val), nrow = length(id), ncol = 3)
+    lt <- rbind(lt, as.data.frame(m, stringsAsFactors = F))
+  }
+  colnames(lt) <- c("names", "id", "val")
+  lt$id <- as.numeric(lt$id)
+  lt
+}
+
+
+mmet2RR <- function(m, cn){
+    for (i in 1:length(cn)){
+      lcn <- as.character(cn[i])
+      m[[lcn]]  <- apply(data.frame(m[[lcn]]), 1, function(x) mmet2RRVal(x[1]))
+    }
+    m
+  }
+
+
+mmet2RRVal <-function(val) {
+  if ((!is.null( val) && !is.na(val))){
+    # Cap values at 35 mmet
+    mmet2RR_mat[which.min(abs(ifelse(mmet2RR_mat$dose < 35, mmet2RR_mat$dose, 35)  - val)), 2]
+  }
+  else
+    0
+}
+
+
+PAF <- function(pop, attr, cn){
+  # pop = rr
+  # attr = c('female', 'agecat_det')
+  # cn = c('total_mmet', 'total_mmet_sc')
+  
+  unique_gender <- unique(pop[[attr[1]]])
+  unique_age_group <- unique(pop[[attr[2]]])
+  combinations <- length(unique_age_group)*length(unique_gender)
+  
+  m = matrix(nrow=combinations,ncol=(2 + length(cn)))
+  
+  colnames(m) <- append(c("age band", "gender"), cn)
+  mi <- 1
+  
+  for (i in 1:length(unique_age_group)){
+    
[…]

### 2018-06-26 R/PA/code/script.R

@@ -1,0 +1,170 @@
+# Remove everything
+rm (list = ls())
+# Load packages
+library(tidyverse)
+library(haven)
+library(plotly)
+
+
+#Set seed
+set.seed(1)
+
+# Load all functions
+source("R/PA/code/functions.R")
+
+# Sample size for scenario
+sample_size <- 110
+
+# Constants
+# Initialize  energy expenditure constants - taken from ICT
+METCycling <- 5.63
+METWalking <- 3.53
+METEbikes <- 4.50
+
+# Read combined individual travel survey and Physical Activity data
+raw_data <- haven::read_dta("data/synth_pop_data/england/SPtrip_CensusNTSAPS_E06000001.dta")
+
+# Create id columns for all trips (and people without any trips)
+raw_data$id <- 1:nrow(raw_data)
+
+# Create a lookup table for selected columns - labelled columns from stata
+lt <- create.lookups(raw_data, c("female", "agecat", "agecat_det", "trip_mainmode"))
+
+# Sort it
+lt <- arrange(lt, names, id)
+
+
+# Sample 10k unique IDs
+# Select trips for the 10k people
+baseline <- raw_data %>% filter(census_id %in% sample(unique(census_id), 1000)) 
+
+lbls <- as.character(unique(as_factor(baseline$trip_mainmode, labels = "values")))
+
+ggplot(baseline, aes (x = as_factor(trip_mainmode, labels = "values"))) + geom_bar() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+# Remove labels
+baseline <- clear.labels(baseline)
+# Replace NAs with 0
+baseline$trip_cycletime_min <- ifelse(is.na(baseline$trip_cycletime_min), 0, baseline$trip_cycletime_min)
+# Replace NAs with 0
+baseline$trip_walktime_min <- ifelse(is.na(baseline$trip_walktime_min), 0, baseline$trip_walktime_min)
+# Calculate trip_cycletime_hr by dividing trip_cycletime_min by 60
+baseline$trip_cycletime_hr <- baseline$trip_cycletime_min / 60
+# Calculate trip_walktime_hr by dividing trip_walktime_min by 60
+baseline$trip_walktime_hr <- baseline$trip_walktime_min / 60
+
+# Get total individual level walking and cycling and sport mmets 
+individual_mmet <- baseline %>% group_by(census_id) %>% summarise (female = first(female), 
[…]

### 2018-06-26 R/injuries/functions/injury_calculation_functions.R

@@ -1,0 +1,965 @@
+setwd('/home/rob/overflow_dropbox/ITHIM-R/injuries/')
+library(shiny)
+library(shinyjs)
+library(shinydashboard)
+library(shinyFiles)
+library(shinyWidgets)
+library(xlsx)
+library(readODS)
+library(MASS)
+library(splines)
+library(gtools)
+library(tools)
+library(tcltk2)
+library(dplyr)
+library(mvtnorm)
+library(distr)
+
+if(exists('useShiny')) rm(useShiny)
+
+## choose whether to compute or recompute
+getModelFits <- function(object_store,input){
+  ##TODO replicate all input objects as object_store objects. Check them against each other when computing/plotting.
+  ## recompute if SIN, data, model changed
+  ## could keep same formula?
+  ## resummarise if lq, uq, SIN uncertainty, reporting rate changed
+  
+  
+  ## process data if they haven't been processed yet
+  if(is.null(object_store$scenario_tabs)){
+    object_store <- processData(object_store)
+  }
+  
+  ## fit models if they haven't been fit yet
+  if(is.null(object_store$fit_whw)&&is.null(object_store$fit_nov)){
+    object_store <- fitModel(object_store,input)
+  }else{ ## re-fit models if model type or sin have changed
+    if((as.numeric(object_store$model=='poisson')!=as.numeric(object_store$fit_whw[[1]]$family$family=='poisson'))||
+        input$sin!=object_store$sin ){
+      object_store <- fitModel(object_store,input)
+    #}else{
+    #  object_store$scenario_tabs <- get_scenarios(object_store)
+    }##TODO}else{ if 
+      #(quantiles in the model different from quantile inputs and sin T/F in the model different from sin T/F in the inputs)
+      #calculateModels
+    
+  }
+  
+  object_store$summary_table <- assemble_output(object_store,input)
+  object_store
+}
+
+## function to get injury and distance dataset
+processData <- function(object_store){
+  injuries <- object_store$injuries
+  rownames(injuries) <- injuries[,1]
+  injuries <- injuries[,-1]
+  
+  # get distance data
[…]

### 2018-06-26 R/injuries/functions/parse_mc_travel_data.R

@@ -1,0 +1,111 @@
+setwd('~/overflow_dropbox/ITHIM-R/injuries/')
+library(xlsx)
+
+## read trip dataset
+trips0 <- read.csv('tviaje.csv')
+trips <- trips0[,c(1,2,26*2+6,26*3+3:4)]
+names(trips) <- c('trip_id','person_id','passenger','gen','age')
+
+## read stage dataset
+stages0 <- read.csv('ttransporte.csv')
+stages0 <- subset(stages0,p5_16_1_1<24)
+stages <- stages0[,c(1,2,3,5,7,8,13)]
+names(stages) <- c('stage_id','trip_id','weekend','mode','hours','minutes','factor')
+stages$time <- stages$minutes + 60*stages$hours
+stages$weekend <- stages$weekend - 1
+
+stages_temp <- (subset(stages0,edad>79&p5_14==14&(p5_16_1_2>0|p5_16_1_1>0)))
+x11(); plot(stages_temp$factor,stages_temp$p5_16_1_1*60+stages_temp$p5_16_1_2)
+points(stages_temp$factor[stages_temp$edad>90],stages_temp$p5_16_1_1[stages_temp$edad>90]*60+stages_temp$p5_16_1_2[stages_temp$edad>90],col='red')
+
+## list modes
+modes <- c('Car','Minibus','Taxi (Internet app)','Taxi (street)','Subway',
+  'Bus','Bycicle','Bus','Motorcycle','Tram',
+  'Bus Rapid Transit (BRT)','Light Rail Transit (LRT)','Train','Walk','Cable car',
+  'Bycicle taxi','Motorcycle taxi','School bus','Charter bus','Other')
+## define whether modes are by default passenger or not
+passenger <- c(0,1,1,1,1,
+  1,0,1,0,1,
+  1,1,1,0,1,
+  1,1,1,1,1)
+## list modes to keep (exclude train, tram, subway, cable car, other)
+keep_mode <- c(1,2,3,4,6,7,8,9,11,14,16,17,18,19)
+
+stages <- subset(stages,mode%in%keep_mode)
+
+trips$passenger <- trips$passenger - 1
+
+stages$passenger <- stages$passenger_trip <- trips$passenger[match(stages$trip_id,trips$trip_id)]
+stages$passenger[is.na(stages$passenger)] <- passenger[stages$mode[is.na(stages$passenger)]]
+stages$passenger[!stages$mode%in%c(1,9)] <- passenger[stages$mode[!stages$mode%in%c(1,9)]]
[…]

### 2018-06-26 R/injuries/get_london_injuries.R

@@ -1,0 +1,97 @@
+setwd('~/overflow_dropbox/ITHIM-R/injuries')
+library(matrixStats)
+library(latex2exp)
+library(readstata13)
+library(xlsx)
+library(readODS)
+library(tensorA)
+library(dplyr)
+library(tidyr)
+library(MASS); 
+library(splines)
+loc <- readRDS('~/overflow_dropbox/HPC/locality.RData')
+
+#########################################################################################################################
+## code name map
+lacodes <- read.ods('~/Dropbox/ITHIM/InjuryModel/spatial/UKLAcodes.ods')[[1]][-1,]
+lacodes10 <- read.ods('~/Dropbox/ITHIM/InjuryModel/spatial/UKLAcodes10.ods')[[1]]
+#########################################################################################################################
+## road distance name
+raw_road <- read.xlsx('~/overflow_dropbox/injury_predictions/VehicleTypeRoadTypeRegionalLevel.xlsx',sheetIndex=1,rowIndex=6:402,colClasses=c('numeric',rep('character',2),rep('numeric',7)))
+raw_la <- read.xlsx('~/overflow_dropbox/injury_predictions/VehicleType_LALevel.xlsx',sheetIndex=1,rowIndex=6:1670,colClasses=c('numeric',rep('character',2),rep('numeric',7)))
+regions <- levels(raw_road[,2])
+las <- unique(raw_la[,3])
+
+#########################################################################################################################
+## injury code
+load('~/overflow_dropbox/ITHIM/InjuryModel/Stats19.rda')
+ss19<-tbl_df(ss19)
+ss19nov <- ss19[ss19$strike_mode=='No other vehicle',c(1:7)]
+ss19nov<-ss19nov[complete.cases(ss19nov),]
+ss19<-ss19[complete.cases(ss19),]
+ss19nov$district <- loc$local_authority_.highway.[match(ss19nov$accident_index,loc$accident_index)]
+ss19$district <- loc$local_authority_.highway.[match(ss19$accident_index,loc$accident_index)]
+injury_codes <- unique(ss19$district)
+#########################################################################################################################
+## combine
+lacodes <- lacodes[lacodes$A%in%injury_codes,]
+lacodes <- lacodes[lacodes$C%in%las,]
+lacodes10 <- lacodes10[lacodes10$B%in%injury_codes,]
[…]

### 2018-06-26 R/injuries/linear_injury.R

@@ -1,0 +1,81 @@
+setwd('/home/rob/overflow_dropbox/ITHIM-R/injuries/')
+source('shinyinjury.R')
+
+input <- list()
+input$injuryfile <- 'input_injuries.xlsx'
+input$distancefile <- 'mexico_city_travel_data.csv' #'synthetic_population.ods' #
+input$file2 <- 'saved_mexico_model.Rdata'
+input$file3 <- 'saved_mexico_model_NB.Rdata'
+input$group <- 'casualty mode'
+input$over <- 'strike mode'
+input$subgroup <- 'pedestrian'
+input$SE <- F
+input$lq <- 0.05
+input$uq <- 0.95
+input$sin <- T
+input$sinuncertainty <- T
+inFile <- list()
+object_store <- list(fit_whw=NULL,fit_nov=NULL,scenario_tabs=NULL,injuries=NULL,plotButton=NULL,covariates=NULL,
+  mexicoButton=NULL,distance=NULL,model='poisson',nScenarios=1,lq=0.25,uq=0.75,sinfile='default_sin_exponents.Rdata')
+
+model <- 'poisson'
+if (model=='NB') object_store$model <- 'NB'
+
+useShiny <- F
+useMexicoModel <- 0
+
+if(useMexicoModel==0){
+  ## GET INJURY INPUT
+  inFile$datapath <- input$injuryfile
+  if (is.null(inFile)) return(NULL)
+  ##TODO allow many file types
+  injuries <- read.xlsx(inFile$datapath,sheetIndex=1,stringsAsFactors=FALSE,header=F)
+  ##TODO many checks here
+  # are the ages interpretable? How are they coded?
+  # are the ages the same as the distance ages?
+  # are there travel data for hgvs?
+  # what are the covariates? are they all in the distance data?
+  # V codes comprehensible, numbers as numbers not strings...
+  object_store$injuries <- injuries
+  
+  inFile$datapath <- input$distancefile
+  if (is.null(inFile)) return(NULL)
+  if(file_ext(inFile$datapath)=='ods'){
+    distance <- read.ods(inFile$datapath)[[1]]
+    names(distance) <- distance[1,]
+    distance <- distance[-1,]
+    ##TODO choose columns to change based on column names
+    distance[,c(2,4:dim(distance)[2])]<-sapply(distance[,c(2,4:dim(distance)[2])],as.numeric)
+  }else if(file_ext(inFile$datapath)=='csv'){
[…]

### 2018-06-26 R/injuries/server.R

@@ -1,0 +1,281 @@
+server <- function(input, output){
+  source('functions/injury_calculation_functions.R')
+  # object storing all values
+  object_store <- reactiveValues( ## models
+    fit_whw=NULL,
+    fit_nov=NULL,
+    ## data
+    nScenarios=1,
+    scenario_tabs=NULL,
+    injuries=NULL,
+    distance=NULL,
+    covariates=NULL,
+    ## switches
+    plotButton=NULL,
+    mexicoButton=NULL,
+    model='poisson',
+    ## quantiles
+    lq=0.25,
+    uq=0.75,
+    ## SIN
+    sin=F,
+    sinuncertainty=F,
+    sinfile='data/default_sin_exponents.Rdata',
+    ## reporting rate
+    rr=F,
+    temp_rrdistribution='Beta',
+    temp_rrp=c(5,1),
+    temp_rrdist=Beta(5,1),
+    rrdistribution='Beta',
+    rrp=c(5,1),
+    rrdist=Beta(5,1))
+  # response to uploading injury. Saves injury; reveals distance file button
+  ##TODO remove distance data file response when we have true synthetic population
+  output$ui.distance <- renderUI({
+    inFile <- input$injuryfile
+    if (is.null(inFile)) return(NULL)
+    ##TODO allow many file types
+    injuries <- read.xlsx(inFile$datapath,sheetIndex=1,stringsAsFactors=FALSE,header=F)
+    ##TODO many checks here
+    # are the ages interpretable? How are they coded?
+    # are the ages the same as the distance ages?
+    # are there travel data for hgvs?
+    # what are the covariates? are they all in the distance data?
+    # V codes comprehensible, numbers as numbers not strings...
+    object_store$injuries <- injuries
+    ##TODO this will eventually be read directly from synthetic population, not uploaded by user.
+    fileInput(inputId='distancefile', label='Upload distance data',
+      accept = c(
+        "text/csv",
+        "text/comma-separated-values,text/plain",
+        ".xlsx",
+        ".csv")
+    )
+  })
+  # reads distance file
+  observeEvent(input$distancefile, {
+    inFile <- input$distancefile
+    if(file_ext(inFile$datapath)=='ods'){
+      distance <- read.ods(inFile$datapath)[[1]]
+      names(distance) <- distance[1,]
+      distance <- distance[-1,]
+      ##TODO choose columns to change based on column names
[…]

### 2018-06-26 injuries/functions/injury_calculation_functions.R

@@ -1,965 +1,0 @@
-setwd('/home/rob/overflow_dropbox/ITHIM-R/injuries/')
-library(shiny)
-library(shinyjs)
-library(shinydashboard)
-library(shinyFiles)
-library(shinyWidgets)
-library(xlsx)
-library(readODS)
-library(MASS)
-library(splines)
-library(gtools)
-library(tools)
-library(tcltk2)
-library(dplyr)
-library(mvtnorm)
-library(distr)
-
-if(exists('useShiny')) rm(useShiny)
-
-## choose whether to compute or recompute
-getModelFits <- function(object_store,input){
-  ##TODO replicate all input objects as object_store objects. Check them against each other when computing/plotting.
-  ## recompute if SIN, data, model changed
-  ## could keep same formula?
-  ## resummarise if lq, uq, SIN uncertainty, reporting rate changed
-  
-  
-  ## process data if they haven't been processed yet
-  if(is.null(object_store$scenario_tabs)){
-    object_store <- processData(object_store)
-  }
-  
-  ## fit models if they haven't been fit yet
-  if(is.null(object_store$fit_whw)&&is.null(object_store$fit_nov)){
-    object_store <- fitModel(object_store,input)
-  }else{ ## re-fit models if model type or sin have changed
-    if((as.numeric(object_store$model=='poisson')!=as.numeric(object_store$fit_whw[[1]]$family$family=='poisson'))||
-        input$sin!=object_store$sin ){
-      object_store <- fitModel(object_store,input)
-    #}else{
-    #  object_store$scenario_tabs <- get_scenarios(object_store)
-    }##TODO}else{ if 
-      #(quantiles in the model different from quantile inputs and sin T/F in the model different from sin T/F in the inputs)
-      #calculateModels
-    
-  }
-  
-  object_store$summary_table <- assemble_output(object_store,input)
-  object_store
-}
-
-## function to get injury and distance dataset
-processData <- function(object_store){
-  injuries <- object_store$injuries
-  rownames(injuries) <- injuries[,1]
-  injuries <- injuries[,-1]
-  
-  # get distance data
[…]

### 2018-06-26 injuries/functions/predict.lm.hack.R

@@ -1,250 +1,0 @@
-
-#qr.solve(qr.R(stats:::qr.lm(object))[p1, p1])
-
-
-# qr.solve <- function (a, b, tol = .Machine$double.eps) 
-# {
-#   if (!inherits(a, "qr")) 
-#     a <- qr(a, tol = tol)
-#   nc <- ncol(a$qr)
-#   nr <- nrow(a$qr)
-#   print(a$rank)
-#   if (a$rank != min(nc, nr)) 
-#     stop("singular matrix 'a' in solve")
-#   if (missing(b)) {
-#     if (nc != nr) 
-#       stop("only square matrices can be inverted")
-#     b <- diag(1, nc)
-#   }
-#   res <- qr.coef(a, b)
-#   res[is.na(res)] <- 0
-#   res
-# }
-
-predict.lm <- function (object, newdata, se.fit = FALSE, scale = NULL, df = Inf, 
-  interval = c("none", "confidence", "prediction"), level = 0.95, 
-  type = c("response", "terms"), terms = NULL, na.action = na.pass, 
-  pred.var = res.var/weights, weights = 1, tol = 1e-7, ...) 
-{
-  tt <- terms(object)
-  if (!inherits(object, "lm")) 
-    warning("calling predict.lm(<fake-lm-object>) ...")
-  if (missing(newdata) || is.null(newdata)) {
-    mm <- X <- model.matrix(object)
-    mmDone <- TRUE
-    offset <- object$offset
-  }
-  else {
-    Terms <- delete.response(tt)
-    m <- model.frame(Terms, newdata, na.action = na.action, 
-      xlev = object$xlevels)
-    if (!is.null(cl <- attr(Terms, "dataClasses"))) 
-      .checkMFClasses(cl, m)
-    X <- model.matrix(Terms, m, contrasts.arg = object$contrasts)
-    offset <- rep(0, nrow(X))
-    if (!is.null(off.num <- attr(tt, "offset"))) 
-      for (i in off.num) offset <- offset + eval(attr(tt, 
-        "variables")[[i + 1]], newdata)
-    if (!is.null(object$call$offset)) 
-      offset <- offset + eval(object$call$offset, newdata)
-    mmDone <- FALSE
-  }
-  n <- length(object$residuals)
-  p <- object$rank
-  p1 <- seq_len(p)
-  piv <- if (p) 
-    stats:::qr.lm(object)$pivot[p1]
[…]

### 2018-06-26 injuries/server.R

@@ -1,281 +1,0 @@
-server <- function(input, output){
-  source('functions/injury_calculation_functions.R')
-  # object storing all values
-  object_store <- reactiveValues( ## models
-    fit_whw=NULL,
-    fit_nov=NULL,
-    ## data
-    nScenarios=1,
-    scenario_tabs=NULL,
-    injuries=NULL,
-    distance=NULL,
-    covariates=NULL,
-    ## switches
-    plotButton=NULL,
-    mexicoButton=NULL,
-    model='poisson',
-    ## quantiles
-    lq=0.25,
-    uq=0.75,
-    ## SIN
-    sin=F,
-    sinuncertainty=F,
-    sinfile='data/default_sin_exponents.Rdata',
-    ## reporting rate
-    rr=F,
-    temp_rrdistribution='Beta',
-    temp_rrp=c(5,1),
-    temp_rrdist=Beta(5,1),
-    rrdistribution='Beta',
-    rrp=c(5,1),
-    rrdist=Beta(5,1))
-  # response to uploading injury. Saves injury; reveals distance file button
-  ##TODO remove distance data file response when we have true synthetic population
-  output$ui.distance <- renderUI({
-    inFile <- input$injuryfile
-    if (is.null(inFile)) return(NULL)
-    ##TODO allow many file types
-    injuries <- read.xlsx(inFile$datapath,sheetIndex=1,stringsAsFactors=FALSE,header=F)
-    ##TODO many checks here
-    # are the ages interpretable? How are they coded?
-    # are the ages the same as the distance ages?
-    # are there travel data for hgvs?
-    # what are the covariates? are they all in the distance data?
-    # V codes comprehensible, numbers as numbers not strings...
-    object_store$injuries <- injuries
-    ##TODO this will eventually be read directly from synthetic population, not uploaded by user.
-    fileInput(inputId='distancefile', label='Upload distance data',
-      accept = c(
-        "text/csv",
-        "text/comma-separated-values,text/plain",
-        ".xlsx",
-        ".csv")
-    )
-  })
-  # reads distance file
-  observeEvent(input$distancefile, {
-    inFile <- input$distancefile
-    if(file_ext(inFile$datapath)=='ods'){
-      distance <- read.ods(inFile$datapath)[[1]]
-      names(distance) <- distance[1,]
-      distance <- distance[-1,]
-      ##TODO choose columns to change based on column names
[…]

### 2018-06-26 injuries/ui.R

@@ -1,63 +1,0 @@
-ui <- fluidPage(
-  tags$head(
-    tags$style(HTML("hr {border-top: 1px solid #000000;}"))
-  ),
-  sidebarLayout(position='left',
-    # Sidebar panel for inputs
-    sidebarPanel('Inputs',
-      # Injury file input
-      fileInput(inputId='injuryfile', label='Upload injury data',
-        accept = c(
-          "text/csv",
-          "text/comma-separated-values,text/plain",
-          ".xlsx",
-          ".csv")
-      ),
-      # Distance file input
-      ##TODO this will go when we have a true synthetic population from which to obtain distance data
-      uiOutput('ui.distance'),
-      ## To use SIN values
-      uiOutput("ui.sin"),
-      # Use SIN uncertainty
-      uiOutput('ui.sinuncertainty'),
-      ## To use reporting rate
-      uiOutput('ui.rrbox'),
-      # Toggle switch to choose between poisson and NB models
-      uiOutput("ui.rr.distribution"),
-      # Toggle switch to choose between poisson and NB models
-      uiOutput('ui.modeltoggle'),
-      # Choose quantiles
-      uiOutput('ui.lq'),
-      uiOutput('ui.uq'),
-      # Button to compute model
-      uiOutput('ui.compute'),
-      # Button to save model
-      uiOutput('ui.save'),
-      hr(),
-      # Button to import saved model
-      fileInput(inputId='file2', label='Import saved model',
-        accept = c(
-          ".Rdata",
-          ".RData")
-      ),
-      # Rob's shortcut to get Mexico model
-      ##TODO remove
-      actionButton('mexico','Use saved Mexico model'),
-      hr(),
-      # Choose which covariate to plot
-      uiOutput('ui.group'),
-      # Choose which realisation of covariate to plot
-      uiOutput('ui.subgroup'),
-      # Choose which other covariate to plot over
-      uiOutput('ui.over'),
-      # Radio button: plot SE or not
-      uiOutput('ui.se'),
-      # Button to plot/renew plot
-      uiOutput('ui.plot')
-    ),
-    # Main panel shows the results
-    mainPanel(
-      plotOutput(outputId = "plot", width = "500px", height = "800px")
-    )
-  )
-)

### 2018-07-09 doc/accra/process_flow.rmd
Add rmd doc for accra process flow

---
title: "Accra - process flow"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scenario Creation - AA

Read leandro’s created processed raw data (which already has domain specific duration and days) as:

rd <- read.csv("data/synth_pop_data/accra/travel_survey/synthetic_population_with_trips.csv")

Leandro has already removed >70 population from it
This has people without trips (trip_mode 99 and trip_id 0) as well

### Walk trips to Baseline

* Add walk trips out of bus trips to baseline

### Create scenarios

* Scenario 1: Shift modes - 50% of all walking walking trips to Private Car

* Scenario 1: Shift modes - 50% of all car trips less than 7km to cycle

* Scenario 3: Shift and create new trips - 50% of all trips longer than 10km to Bus (as each bus trip might create a new short walking trip)

* Save scenarios as: data/scenarios/accra/baseline_and_three_scenarios.csv
(contains mode and duration of each trips for baseline and all three scenarios)

## Physical Activity – AA

Calculate total PA per individual
1.	Multiply trips with AT by 6 (5 * 1.2) to get weekly trips. 
2.	Sum them for each individual to get total PA

## Air Pollution – RG
Fill in the info

## Health – RG

Identify PA and AP related diseases and then calculate their RR.

Multiply PA’s and AP’s RR to get combined effect

Normalize them – see a message from James below:

*When adjusting AP for PA I think we should renormalize the values if we are using non-linear RRs. The epid studies didn’t allow for PA levels so they will be coming up with lower equivalent levels than we are. Thus I would suggest that for an average individual (how to define that is a challenge) we estimate the ratio of our modelled inhaled exposure to background and apply this to everyone. Does that make sense?*
*The same would apply for diurnal patterns etc.*

### Calculate disease specific burden 
* Calculate PIFs – CRA for baseline and Scenario
* Calculate disease burden (using GBD’s data) for deaths and YLLs

### 2018-07-09 doc/accra/process_flow.rmd
Formatting
Remove a repeated word

@@
 Leandro has already removed >70 population from it
-This has people without trips (trip_mode 99 and trip_id 0) as well
+
+(this has people without trips (trip_mode 99 and trip_id 0) as well)
 
@@
 
-* Scenario 1: Shift modes - 50% of all walking walking trips to Private Car
+* Scenario 1: Shift modes - 50% of all walking trips to Private Car
 

### 2018-08-14 doc/accra/process_flow.rmd

@@
 ## Air Pollution – RG
-Fill in the info
+* Ideally, air pollution calculations will account for four activity domains totalling to 24 hours: these include transport, work, leisure, and sleep.
+Each domain will have specific air inhalation rate which will be based on intensity of the activity.
+* In case work and leisure are not available, these columns will be left as NA and air pollution calculation will assume work and leisure time concentration same as background concentration-- sleep can still be calculated assuming X number of average sleeping hours
+* The background concentration for scenario is a function of the changing transport emission inventory.
+For this, the input sheet will include emission inventory classified by pre-specified vehicle categories.
+The emission estimates for scenarios will be revised based on scenario distances, and scenario background concentration will be calculated.
+* Emission calculation will be a part of the input sheet outside the modelling framework
 

### 2018-10-11 code/MSLT/MSLTdoc.Rmd
Updating RMarkdown code

--- 2018-06-07 R/MSLT/MSLTdoc.Rmd
@@
 The data is sent to the designated email address (within minutes) in ZIP format, unzip and use the code below to read the data (4).
-Here, we selected data for Greater London and England.
+Here, we selected data for Greater London.
 The aim is to compare and derive scaling factors as for most cities the data is not available from the GBD and country level data may be used and scaled to the city level.
@@
 
-4) Read GBD data
+4) Generate Dismod input dataset
 ```{r chunk-4}
@@
 
-5) Change all upper cases to lower case and delete () from variables.
+5) Generate data for the proportional-multi state life table by single age-year (mslt_df).
 ```{r chunk-5}
@@
 
-6) Create age categories index in GBDdata (mid age, to match cohort running), total of 20 age groups. These are the age cohorts to simulate.
+6) Add Dismod output to mslt_df.
 ```{r chunk-6}
@@
 
-7) Create age and sex categories to obtain population numbers.
-Population numbers from GBD are used in Dismod II.
-For the Life table (Figure 3), the numbers may be from the synthetic population.
-For now, the Life table is set up with population numbers derived from the GBD data.
+7) Generate baseline general life tables.
 ```{r chunk-7}
@@
 
-8) Convert string variables to numeric to do calculations.
+8) Generate baseline disease life table.
 ```{r chunk-8}
@@
 
-9) Generate population numbers for Greater London in a new data frame ("GBD_population").
-Note that there is data for England as well, which is used in a separate rmarkdown document (GBDCompare).
-Population numbers are derived from rates per 100,000 and total numbers of cases.
+9) Factor to multiply by incidence, will be replaced by (1-PIF) and generate scenario incidence.
 ```{r chunk-9}
 ```
-10) Generate population numbers from given number of cases and rates per 100,000 people. 
+10) Generate scenario disease life tables. 
 ```{r chunk-10}
 ```
-11) Remove rows with zero
+11) Generate all-cause mortality and all-cause years with disability rate for scenario and use in scenario general life table. 
 ```{r chunk-11}
@@
 
-12) Keep relevant variables
+12) Generate a list of disease outputs by age and sex.  
 ```{r chunk-12}
 ```
-13) Create data frames for Greater London to be later used for: a) interpolation of rates and b) PMLT cohorts. 
+13)  Generate a general dataframe for all age-sex cohort disease specific outcomes.
 ```{r chunk-13}
@@
 
-14) Check population total numbers
+14) Generate graphs for changes in burden of disease by age and sex (incidence and mortality)
 ```{r chunk-14}
@@
 
-15) Generate data frames for Greater London with per person rates (per 100,000 in original data).  
+15) Generate aggregated (all cohorts) outcomes by sex.
+
 ```{r chunk-15}

### 2018-10-11 code/MSLT/mslt_code.r

@@ -323,6 +323,21 @@ for (i in 2:nrow(disease_short_names)){
   }
 }
   
+# ---- chunk-6 ----
+
+## Use dismod output and add to mslt_df
+
+idata <- read.csv("data/legacy/UK/idata.csv", stringsAsFactors = F)
+
+# Add age_sex category to match with mslt_df
+
+idata$sex_age_cat <- paste(idata$sex,idata$age, sep = "_"  )
+
+idata <- select(idata, -c(age, sex))
+
+mslt_df <- left_join(mslt_df, idata, by = "sex_age_cat")
+
+
 
 # ---- chunk-7 ----
 
@@ -347,18 +362,6 @@ for (age in i_age_cohort){
 ## View(general_life_table_list_bl[[2]])
 
 # ---- chunk-8 ----
-
-## Use dismod output and add to mslt_df
-
-idata <- read.csv("data/legacy/UK/idata.csv", stringsAsFactors = F)
-
-# Add age_sex category to match with mslt_df
-
-idata$sex_age_cat <- paste(idata$sex,idata$age, sep = "_"  )
-
-idata <- select(idata, -c(age, sex))
-
-mslt_df <- left_join(mslt_df, idata, by = "sex_age_cat")
 
 ## Use run_disease
 
@@ -575,9 +578,9 @@ for (age in i_age_cohort){
 # View(general_life_table_list_sc[[32]])
 # View(general_life_table_list_bl[[1]])
 
-## Check difference life table baseline and scenario
-general_life_table_list_bl[[1]]$Lx - general_life_table_list_sc[[1]]$Lx
-general_life_table_list_bl[[1]]$Lwx - general_life_table_list_sc[[1]]$Lwx
+## Check difference life table baseline and scenario (uncomment to check)
+# general_life_table_list_bl[[1]]$Lx - general_life_table_list_sc[[1]]$Lx
+# general_life_table_list_bl[[1]]$Lwx - general_life_table_list_sc[[1]]$Lwx
 
 
 
@@ -853,10 +856,7 @@ for (age in i_age_cohort){
 
 
 
-# ---- chunk-26 ----
-
-
-## Try binding function
+# ---- chunk-15 ----
 
 aggregate_frame_males <- list()
 aggregate_frame_females <- list()

### 2018-10-11 code/MSLT/mslt_code.r
Updating rmarkdown

@@ -961,7 +961,7 @@ View(total_aggr)
 
 
 
-# ---- chunk-27 ----
+# ---- chunk-16 ----
 
 
 ####This plot has to be customised to in_outcomes, here, only totals shown, but specifications are up to the user. ADD LOOP for all outcomes over time and total TABLE.

### 2018-10-22 ithim_r_function_descriptions.Rmd
function descriptions

---
title: "ITHIM Functions"
author: "Daniel Fuller"
date: '2018-10-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ITHIM R Functions

CAPITAL LETTERS: Background constant that is fixed
snake_case: all variable names and functions

### Common Abbreviations
- PA: Physical Activity
- AP: Air Pollution
- PM: Particulate Matter? 
- GDB: Global Burden of Disease
- HC: ??

### add_walk_time

Description  
* A function to add walking time to bus trips assuming XYZ

Generic: TRUE or FALSE

Input  
*

Output  
*

### combine_health_and_pif
* A function to reorder HC? values and combine those values

Generic: TRUE or FALSE

Input
*

Output
*

### combined_rr_pa_ap
* A function to join and calculate disease burden for pa and ap?

Generic: TRUE or FALSE

Input
*

Output
*

### create_all_scenarios
* A function to create scenarios for ITHIM model. This functions requires manual input from the user with assumptions.

Generic: TRUE or FALSE

Input
* Mode shift assumptions (expressed as %) for at least one mode.

Output: 
* Scenario data appended to an updated dataframe with Scenario column identifying each scenario.

### create_scenario
* A function to create a scenario for ITHIM model. This functions requires manual input from the user with assumptions.

Generic: TRUE or FALSE

Input
* Mode shift assumptions (expressed as %) for at least one mode.

Output: 
* Scenario data appended to an updated dataframe with Scenario column identifying each scenario.

### dist_dur_tbls
* A function to create trip distance for the baseline and scenarios

Generic: TRUE or FALSE

Input
* Data frame from household travel survey data with baseline data and scenarios

Output
*

### distances_for_injury_function
* A function to calculate distances travelled by mode for each scenario using the baseline data

Generic: TRUE or FALSE

Input
* Data frame from household travel survey data with baseline data and scenarios

Output
* 1 new column with distances travelled by mode for each scenario

### gen_ap_rr
* A function to generate health risk ratios for air pollution combining the PA data from the function (total_mmet.R) and the PM2.5 concentration data (scenario_pm_calculations.R) at the individual level

Generic: TRUE or FALSE

Input
*

Output
*

### gen_pa_rr
* A function to generate health risk ratios for physical activity

Generic: TRUE or FALSE

Input
*

Output
*

### health_burden
* A function to create a reference scenario and to calculate the overall heath burden of the scenarios compared to the reference (!! I think)

Generic: TRUE or FALSE

Input
*

Output
*

### injuries_function
* _This is the script for distance-based injury model for Accra using safety-in-numbers_

Generic: FALSE

Input: 
*

Output:
*

### ithim_load_data
* A function to load in the necessary data for running an ITHIM model

* There is also a simple function to calculate the ratio of YLL to deaths for each age and sex group. This should probably be pulled out.

* There is also a long comment about data structure in there that I think should be adopted.

Generic: TRUE or FALSE

Input
*

Output
*

### ithim_setup_baseline_scenario
* A function to create a baseline scenario? or setup the baseline data frame as it should be. Not sure. Baseline dataframe should come from user... ?

Generic: TRUE or FALSE

Input
*

Output
*

### ithim_setup_parameters
* A function to setup baseline parameters for the model. Parameters are assigned to the environment and so are set for every function. They are over-written when `sample_parameters` is called.

Defaults are  
* MEAN_BUS_WALK_TIME= 5  
* MMET_CYCLING = 4.63  
* MMET_WALKING = 2.53  
* PM_CONC_BASE = 50  
* PM_TRANS_SHARE = 0.225  
* PA_DOSE_RESPONSE_QUANTILE = F  
* AP_DOSE_RESPONSE_QUANTILE = F  
* BACKGROUND_PA_SCALAR = 1  
* SAFETY_SCALAR = 1  
* CHRONIC_DISEASE_SCALAR = 1

Generic: TRUE or FALSE

Input
*

Output
*

### ithim_uncertainty
* A function to with uncertainty, but sure what precisely it does.

Generic: TRUE or FALSE

Input
*

Output
*

### PA_dose_response
* A function to define the dose reponse for physical activity and all of the diseases of interest

* all-cause-mortality  
* breast-cancer  
* cardiovascular-disease  
* colon-cancer  
* coronary-heart-disease   
* endometrial-cancer  
* heart-failure  
* lung-cancer  
* stroke  
* total-cancer

Generic: TRUE or FALSE

Input
*

Output
*

### PAF
* A function to hard code the indices to something??

Generic: TRUE or FALSE

Input
*

Output
*

### parrellel_evppi_for_AP
* A function that ???

Generic: TRUE or FALSE

Input
*

Output
*

### run_ithim
* A function that runs the ITHIM model. This function requires all of the data to be in the correct format and for the scenario functions to have been input by the user.

Generic: TRUE or FALSE

Input
*

Output
*

### run_ithim_setup
* A function that setups up the assumptions for ITHIM and creates the necessary objects. The fixed paramters the user can specify are as follows:

* NSAMPLES = 1
* modes = c("Bus", "Private Car", "Taxi", "Walking","Short Walking", "Bicycle", "Motorcycle")
* speeds = c(15, 21, 21, 4.8, 4.8, 14.5, 25)
* DIST_CAT = c("0-6 km", "7-9 km", "10+ km")
* AGE_CATEGORY = c("15-49", "50-69", "70+")
* MEAN_BUS_WALK_TIME= 5
* MMET_CYCLING = 4.63
* MMET_WALKING = 2.53
* PM_CONC_BASE = 50
* PM_TRANS_SHARE = 0.225
* PA_DOSE_RESPONSE_QUANTILE = F
* AP_DOSE_RESPONSE_QUANTILE = F
* BACKGROUND_PA_SCALAR = 1
* SAFETY_SCALAR = 1
* CHRONIC_DISEASE_SCALAR = 1

Generic: TRUE or FALSE

Input
*

Output
*

### scenario_pm_calculations
* A function that calculates the particular matter concentrations for trips and for each scenario

Generic: TRUE or FALSE

Input
*

Output
*

### set_scenario_specific_variable
* A function that???

Generic: TRUE or FALSE

Input
*

Output
*

### total_mmet
* A function that calculates the marginal MET values for transport related physical activity for the baseline and the scenarios

Generic: TRUE or FALSE

Input
*

Output
*

### 2019-03-13 code/MSLT/legacy/GBDCompare.Rmd
Re-add MSLT folder. Was deleted in: 183398fdc8d691c44eae468a5165c0450be41d22

--- 2018-04-17 MSLT/GBDCompare.Rmd
@@
 
-The following graphs compare rates per one by age (5-year groups) and sex for England and Greater London.
+The following graphs compare rates per one by age (5-year groups) and sex for England and Greater London for all cause mortality, all cause years lived with disability and diseases specific mortality, ylds, incidence and prevalence.
 
-```{r, echo = FALSE}
+Rates are in five year age groups, as per the GBD age grouping.
+
+```{r, echo = FALSE, message=FALSE, warning=FALSE}
 ###TO DO: cubic splines for input rates

### 2019-03-13 code/MSLT/pif_calculations.R

@@ -1,0 +1,184 @@
+rm (list = ls())
+
+require(dplyr)
+require(tidyverse)
+
+# load dr for pa objects
+source("code/drpa/dose_response.R")
+
+# Load all health related functions
+source("code/PA/code/functions.R")
+
+
+
+mmets <- read_csv("code/MSLT/data/england/mmets_london.csv")
+
+## Remove all scenario but one for simplicity
+
+MS_cols <- names(mmets %>% dplyr::select(starts_with("MS")))
+
+mmets <- select(mmets, -c(MS_cols[-1]))
+
+disease_short_names <- data.frame(disease = c("all-cause-mortality", 
+                                              "diabetes",
+                                              "lung-cancer",
+                                              "breast-cancer", 
+                                              "colon-cancer",
+                                              "coronary-heart-disease",
+                                              "stroke"),
+                                  sname = c("ac", "dm", "tblc", "bc",
+                                            "cc", "ihd", "is"))
+
+# 
+# 'all-cause-mortality', 'breast-cancer', 'cardiovascular-disease',
+# 'colon-cancer', 'coronary-heart-disease', 'diabetes', 'endometrial-cancer',
+# 'heart-failure', 'lung-cancer', 'stroke', 'total-cancer'
+
+pa_certainty <- T
+
+### iterating over all all disease outcomes
+for ( j in 1:nrow(disease_short_names)){
+  ## checking whether to calculate this health outcome for PA
+  pa_dn <- disease_short_names$sname[j] %>% as.character()
+  local_cause <- disease_short_names$disease[j] %>% as.character()
+  
+  for (scen in append('baseline', names(mmets %>% dplyr::select(starts_with("MS"))))){
+    
+    mmets[[paste('RR_pa', scen, pa_dn, sep = '_')]] <- -1
+    
+    
+    for (i in 1:nrow(mmets)){
+      
+      
+      if (local_cause == 'all-cause-mortality'){
+        
+        mmets[[paste('RR_pa', scen, pa_dn, sep = '_')]][i] <- dose_response(cause = local_cause, outcome_type = 'mortality', certainty = pa_certainty, 
+                                                                            dose = mmets[[scen]][i], use_75_pert = T)$rr %>% as.numeric()
+        
+      }else if(local_cause == 'coronary-heart-disease'){
[…]

### 2019-03-13 code/MSLT/MSLTdoc.Rmd
Updating rmarkdown
Revert "Remove MSLT folder"
This reverts commit 183398fdc8d691c44eae468a5165c0450be41d22.

@@
 
-15) Generate aggregated (all cohorts) outcomes by sex.
+15) Generate aggregated (all cohorts) outcomes by sex for disease and life years and health-adjusted life years.
 
@@
 
-16) For the life table, we need to mortality and total yld rates in one year age intervals.
-The original data is in five years.
-Thus, the following code is used to interpolate a single-year age distribution form a five-yearly distribution.
+16) UPDATED TO HERE.
 
 ```{r chunk-16}
-```
-
-17) The following code brings together all the required data to create the inputs life table (Figure 3).
-
-```{r chunk-17}
-```
-
-18) Generate **baseline life tables** from input_data_frame using run_life_table function. Life tables are generated for each age (5-years) and sex cohort. 
-It is assumed that current observed rates of mortality and disability are going to be observed in the future.
-
-```{r chunk-18}
-```
-
-19) Generate **baseline disease life tables** using run_disease function.
-Inputs of the disease life tables are: incidence, case fatality (Figure 3, purple shaded area) and disability weights.
-Internally consistent estimates of incidence and case fatality are derived from GBD data (by five-year age groups and sex) and Dismod II.
-Disability weights are derived from GBD disease specific YLDs and prevalence (by five-year age groups and sex) adjusted for all-cause YLDs (see section 2.1.1.4).
-Note that Dismod II is an external software.
-The data generation for Dismod II is now done externally and the end product, for the example of Greater London is idata.
-For now, we plan to get the data processing for all data for the MSLT and Dismod using code (space blank left below).
-We now have all the data needs in the main data folder in data.
-
-```{r chunk-19}
-```
-
-20) Generate mock change in incidence of disease to generate scenario life tables. In the final version, this will come from the calculated PIFs by disease. 
-PIFs are applied here to incidence, however, may also be applied to case fatality, depending on the RRs.
-Also, we can include time delays from change in exposure to change in health outcomes via the PIF.
-
-```{r chunk-20}
-```
-
-21) Use scenario incidence to calculate scenario disease life tables.
-
-```{r chunk-21}
-```
-
-22) Calculate life tables scenario.
-
-```{r chunk-22}
-```
-
-23) Generate list of outputs by age and sex.
-
-```{r chunk-23}
-```
-
-24) Combine all lists of outputs in a data frame to facilitate extraction of outcomes of interest and plotting.
-
-```{r chunk-24}
-```
-
-25) Plot outcomes for each age and sex cohort using created functions.
-Four functions are used: plot_output (plots outputs), grid_arrange_shared_legend and g_legend (places one label per page) and get_qualified_disease_name (replaces full name of disease)
-
-```{r chunk-25}
-
-knitr::include_graphics(graphs_doc[c(1:64)])
-
-```
-
-26) Add up all outcomes per year of simulation with generated function.
-BETTER TO HAVE A LOOP, I TRIED BY LOOPING IN THE FUNCITON ARGUMENT I_OUTCOMES, BUT THIS DID NOT WORK.
-THEN, IF LOOP HERE ALSO IN NEXT CODE.
-
-```{r chunk-26}
-```
-
-27) Plots using total aggregated change in burden.
-This graphs are not very good for life years and health adjusted life years as the difference between baseline and scenario is very small relative to the total, hence the graphs show almost no change.
-
-```{r chunk-27}
 ```

### 2019-03-13 code/MSLT/code/functions.R

@@ -1,0 +1,505 @@
+#####Packages
+
+require(dplyr)
+require(tidyverse)
+require(knitr)
+require(kableExtra)
+require(citr)
+
+#############################Explanation method###################################
+
+#####Briefly, the proportional multi-state multi cohort life table consists of a 
+#####general life table and a life table for each of the modelled diseases.
+#####The diseases are those associated to the studied risk factor/s. 
+#####The link between the general life table and disease life tables is via the 
+#####potential impact fraction (pif), also called paf (population attributable fraction)
+#####The pif combines exposure to the risk factor and relative risks. The pif is 
+#####appleid to modify incidence in the individual disease life tables, which in turn
+#####modify prevalence and mortality. Changes in mortality and prevalence rates
+#####feed bacak into the general life table to modify total mortality and disability. 
+####Changes in total mortality impact on life years and changes in disability impact 
+####the disability adjusment of life years. 
+
+#### Method reference: 1.	Barendregt JJ, Oortmarssen vGJ, Murray CJ, Vos T. A generic model for the assessment of disease epidemiology: the computational basis of DisMod II. Popul Health Metr. 2003;1(1):4-.
+
+
+####The functions: run_life_table, run_disease, and run_pif serve to generate outputs
+####per age and sex (see model.R). The function run_output to facilitate the generation
+####of results. 
+
+
+#############################Function for general life table###################################
+#####Function to generate age and sex life tables. The function is then use in the model.R script
+#####to calculate variables for the baseline and scenario life tables. 
+
+run_life_table <- function(in_idata, in_sex, in_mid_age)
+{
+  
+  # Create a life table data frame
+  
+  # Filter in_idata by age and select columns for lifetable calculations
[…]

### 2019-03-13 code/MSLT/code/model.R

@@ -1,0 +1,588 @@
+##########################Prepare data and environment to run code###########################################################
+
+##### Clean the Global Environment
+
+rm (list = ls())
+
+##### Prevent scientific notation in data frame
+
+options(scipen=999)
+
+##### Set WD
+
+
+##### Read data. idata: life table and disease life tables (including trends). edata: exposure data (e.g. physical activity).
+#### irr.data: relative risks data and ee: energy expenditure
+#### For exposure (edata) and input (idata) data I am using already generated categories, but a function should be developed to work from raw data. 
+
+idata <- read.csv("code/MSLT/data/legacy/Australia/idata.csv", stringsAsFactors = F)
+edata <- read.csv("code/MSLT/data/legacy/Australia/edata.csv", stringsAsFactors = F)
+irr <- read.csv("code/MSLT/data/legacy/Australia/irr.csv", stringsAsFactors = F)
+ee <- read.csv("code/MSLT/data/legacy/Australia/ee.csv", stringsAsFactors = F)
+
+#####To do
+
+## Note that disease trends are excluded for now. We need to generate a trends function for diseases, mortality (all cause), 
+## change in disease due to change in risk factor exposure (discuss with Niel M), road injuries effect over time. 
+## Eventually, we want to add (1) CRA ITHIM approach to compare results and (2) use all cause mortaltiy instead of 
+## disease specific dose responses for incidence to understand the difference in results for all cause mortality. 
+
+#################################Model parameters#########################################################################
+
+p_age_cohort <- c(22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 87, 92, 97)
+
+p_sex <- c("males", "females")
+
+p_disease <- c("ihd", "istroke", "diabetes", "colon_cancer", "breast_cancer")
+
+###As an expample, an increase in 100 MET-mins per week
+
+p_intervention_effect <- 100
+
+##############################Prepare general life table (age, death rate, population #)########
+
+##### Make lower cases values for sex column
+
+idata$sex <- tolower(idata$sex)
+
+##### Rename mortality_rate to mx
+
+idata$mx <- idata$mortality_rate
+
+##### Remove mortality_rate column
+
+idata$mortality_rate <- NULL
+
+##### Rename male for males and female for females in edata to match idata. 
+
[…]

### 2019-03-13 code/MSLT/mslt_code.r

@@ -1,0 +1,1021 @@
+setwd("code/MSLT")
+# Change to own wd
+
+# ---- chunk-intro ----
+require(dplyr)
+require(tidyverse)
+require(knitr)
+require(kableExtra)
+require(citr)
+require(gridExtra)
+require(cowplot)
+require(ggpubr)
+require(grid)
+require(ggplot2)
+require(gridExtra)
+
+# ---- chunk-1 ----
+rm (list = ls())
+
+# ---- chunk-2 ----
+options(scipen=999)
+
+# ---- chunk-3 ----
+source("code/functions.R")
+
+# ---- chunk-4 ----
+
+# ------------------- Dismod input data set ---------------------------#
+
+
+gbd <- readxl::read_excel("data/england/gbd2016.xlsx")
+
+disease_short_names <- data.frame(disease = c("All causes", 
+                                              "Diabetes mellitus",
+                                              "Tracheal, bronchus, and lung cancer",
+                                              "Breast cancer", 
+                                              "Colon and rectum cancer",
+                                              "Ischemic heart disease",
+                                              "Ischemic stroke"),
+                                  sname = c("ac", "dm", "tblc", "bc",
+                                            "cc", "ihd", "is"))
+
+disease_short_names$disease <- as.character(disease_short_names$disease)
+disease_short_names$sname <- as.character(disease_short_names$sname)
+
+disease_measures <- list("Prevalence", "Incidence", "Deaths", "YLDs (Years Lived with Disability)")
+
+gbd_df <- NULL
+
+for (ag in 1:length(unique(gbd$age))){
+  for (gender in c("Male", "Female")){
+    age_sex_df <- NULL
+    for (dm in 1:length(disease_measures)){
+      for (d in 1:nrow(disease_short_names)){
+        dn <- disease_short_names$disease[d]
+        dmeasure <- disease_measures[dm] %>% as.character()
+        # gender <- "Male"
+        agroup <- unique(gbd$age)[ag]
+        
[…]

### 2019-04-08 summary_tables.Rmd
Add summary table rmd

---
title: "Summary Tables"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadLibraries, echo = F}
library(summarytools)
library(knitr)
library(summarytools)
library(tidyverse)
```

```{r load_objects = "asis"}
io <- readRDS("results/multi_city/io.rds")
# Assumes that multi_city_script.R has been run till 
cities <- c('accra','sao_paulo','delhi','bangalore')

```

```{r trip_tables = "asis"}

for(city in cities){
    td <- io[[city]]$dist
    print(kable(td, caption = paste("Distance table for ", city)))
    cat("\n")
}

for(city in cities){
    td <- io[[city]]$dur
    print(kable(td, caption = paste("Duration table for ", city)))
    cat("\n")
}
```

```{r pa="asis"}

for (city in cities){
  td <- io[[city]]$outcomes$mmets %>% dplyr::select(-c(participant_id))
  print(summarytools::dfSummary(td))
  cat("\n")

}

```

### 2019-04-09 summary_tables.Rmd
Add opts chunk param

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=TRUE, results='asis')
+
 ```
@@
 library(summarytools)
-library(tidyverse)
+#library(tidyverse)
+
+st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
+           plain.ascii       = FALSE,       # One of the essential settings
+           style             = "rmarkdown", # Idem.
+           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
+           footnote          = NA,          # Keeping the results minimalistic
+           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
+                                            # much better results. Your mileage may vary
+
 ```
@@
 }
+
 ```
 
-```{r pa="asis"}
+```{r pa = "asis"}
 
 for (city in cities){
-  td <- io[[city]]$outcomes$mmets %>% dplyr::select(-c(participant_id))
-  print(summarytools::dfSummary(td))
+  td <- io[[city]]$outcomes$mmets
+  print(dfSummary(td))
   cat("\n")

### 2019-04-09 summary_tables.Rmd
Add summarytools as examples, and add some desc

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=TRUE, results='asis')
+knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
 
@@
 
-```{r loadLibraries, echo = F}
+```{r loadLibraries, echo = F, message = F}
+suppressWarnings({
+
 library(summarytools)
@@
 library(summarytools)
-#library(tidyverse)
+library(dplyr)
+})
 
@@
 
+# Set plot_mmets to F
+plot_mmets <- F
+
 ```
 
-```{r trip_tables = "asis"}
+## Scenario definition
+Table displays highest propensity for each distance category, for a given mode.
+
+```{r scen_prop = "asis"}
+
+scen_prop <- round(io$scen_prop, 2)
+
+rownames(scen_prop) <- paste(rownames(io$scen_prop), "scen", sep = "_")
+
+kable(scen_prop, headings = "Scenario Proportions")
+
+```
+
+## Distance tables
+Case study specific distance tables for baseline and five scenarios
+
+```{r trip_dist = "asis"}
 
 for(city in cities){
-    td <- io[[city]]$dist
+    td <- io[[city]]$dist %>% mutate_if(is.numeric, round, digits = 2)
+    colnames(td)[3:7] <- rownames(scen_prop)
     print(kable(td, caption = paste("Distance table for ", city)))
@@
 
+```
+
+## Distance tables
+Case study specific duration tables for baseline and five scenarios
+
+```{r trip_dur = "asis"}
+
 for(city in cities){
-    td <- io[[city]]$dur
+    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = 2)
+    colnames(td)[3:7] <- rownames(scen_prop)
     print(kable(td, caption = paste("Duration table for ", city)))
@@
 
-```{r pa = "asis"}
+```{r pa, echo = F, message = F}
 
-for (city in cities){
-  td <- io[[city]]$outcomes$mmets
-  print(dfSummary(td))
-  cat("\n")
+if (plot_mmets){
 
+  dfSummary(accra_mmets <- io[["accra"]]$outcomes$mmets, style = 'grid', graph.magnif = 0.75, tmp.img.dir = "/tmp")
+
+  dfSummary(sao_paulo_mmets <- io[["sao_paulo"]]$outcomes$mmets, style = 'grid', graph.magnif = 0.75, tmp.img.dir = "/tmp")
+
+  dfSummary(delhi_mmets <- io[["delhi"]]$outcomes$mmets, style = 'grid', graph.magnif = 0.75, tmp.img.dir = "/tmp")
+
+  dfSummary(bangalore_mmets <- io[["bangalore"]]$outcomes$mmets, style = 'grid', graph.magnif = 0.75, tmp.img.dir = "/tmp")
 }

### 2019-04-09 summary_tables.Rmd
Add trip prop for each case study

@@
 
+## Trip proportion
+Case study specific trip proportions by mode, for baseline and five scenarios
+
+```{r trip_mode_dist = "asis", echo = F}
+suppressWarnings({
+  require(tidyverse)  
+})
+
+for(city in cities){
+    td <- io[[city]]$trip_scen_sets %>% group_by(trip_mode, scenario) %>% summarise(p = round(n() / nrow(.) * 100, 2)) %>% spread(key = trip_mode, value = p)
+    td[2:6, 1] <- rownames(scen_prop)
+    print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
+    cat("\n")
+}
+
+suppressWarnings({
+  detach("package:tidyverse", character.only = T)
+})
+
+```
+
 ## Distance tables

### 2019-04-09 summary_tables.Rmd
changing n() to dplyr::n() to avoid conflict with distr function n()

--- 2019-04-09 summary_tables.Rmd
@@
 
+## Trip proportion
+Case study specific trip proportions by mode, for baseline and five scenarios
+
+```{r trip_mode_dist = "asis", echo = F}
+suppressWarnings({
+  require(tidyverse)  
+})
+
+for(city in cities){
+    td <- io[[city]]$trip_scen_sets %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() / nrow(.) * 100, 2)) %>% spread(key = trip_mode, value = p)
+    td[2:6, 1] <- rownames(scen_prop)
+    print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
+    cat("\n")
+}
+
+suppressWarnings({
+  detach("package:tidyverse", character.only = T)
+})
+
+```
+
 ## Distance tables

### 2019-04-09 summary_tables.Rmd
Fix a conflict and also fix a bug to calculate trip count

--- 2019-04-09 summary_tables.Rmd
@@
 
+## Trip proportion
+Case study specific trip proportions by mode, for baseline and five scenarios
+
+```{r load_tidyverse, echo = F, message = F}
+suppressWarnings({
+  require(tidyverse)  
+})
+```
+
+```{r trip_mode_dist = "asis", echo = F}
+
+for(city in cities){
+  df <- io[[city]]$trip_scen_sets
+  df1 <- df %>% filter(scenario == "Baseline") %>% group_by(trip_id, trip_mode, scenario)
+  u_trips <- nrow(df1)
+  td <- df %>% group_by(trip_id, trip_mode, scenario) %>% group_by(trip_mode, scenario) %>% summarise(p = round(n() /  u_trips * 100, 2)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
+  td[2:6, 1] <- rownames(scen_prop)
+  print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
+  cat("\n")
+}
+
+```
+
+```{r unload_tidyverse, echo = F, message = F}
+suppressWarnings({
+  detach("package:tidyverse", character.only = T)
+})
+
+```
+
 ## Distance tables
@@
 
-## Distance tables
+## Duration tables
 Case study specific duration tables for baseline and five scenarios

### 2019-04-09 summary_tables.Rmd
Use dplyr's n()

@@
   u_trips <- nrow(df1)
-  td <- df %>% group_by(trip_id, trip_mode, scenario) %>% group_by(trip_mode, scenario) %>% summarise(p = round(n() /  u_trips * 100, 2)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
+  td <- df %>% group_by(trip_id, trip_mode, scenario) %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() /  u_trips * 100, 2)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
   td[2:6, 1] <- rownames(scen_prop)

### 2019-04-09 summary_tables.Rmd
Use a global var to round to a specific value

--- 2019-04-09 summary_tables.Rmd
@@
 
+# round to
+round_to <- 1
+
 # Set plot_mmets to F
@@
 
-scen_prop <- round(io$scen_prop, 2)
+scen_prop <- round(io$scen_prop, round_to)
 
@@
   u_trips <- nrow(df1)
-  td <- df %>% group_by(trip_id, trip_mode, scenario) %>% group_by(trip_mode, scenario) %>% summarise(p = round(n() /  u_trips * 100, 2)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
+  td <- df %>% group_by(trip_id, trip_mode, scenario) %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() /  u_trips * 100, round_to)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
   td[2:6, 1] <- rownames(scen_prop)
@@
 for(city in cities){
-    td <- io[[city]]$dist %>% mutate_if(is.numeric, round, digits = 2)
+    td <- io[[city]]$dist %>% mutate_if(is.numeric, round, digits = round_to)
     colnames(td)[3:7] <- rownames(scen_prop)
@@
 for(city in cities){
-    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = 2)
+    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to)
     colnames(td)[3:7] <- rownames(scen_prop)

### 2019-04-11 summary_tables.Rmd
Normalize to per person for distance and duration tables

@@
 for(city in cities){
-    td <- io[[city]]$dist %>% mutate_if(is.numeric, round, digits = round_to)
+  count_people <- nrow(io[[city]]$synth_pop)
+  td <- io[[city]]$dist %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)
-    print(kable(td, caption = paste("Distance table for ", city)))
+    print(kable(td, caption = paste("Distance table for ", city, "( ", count_people, " )")))
     cat("\n")
+
 }
@@
 for(city in cities){
-    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to)
+  count_people <- nrow(io[[city]]$synth_pop)
+    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)
-    print(kable(td, caption = paste("Duration table for ", city)))
+    print(kable(td, caption = paste("Duration table for ", city, "( ", count_people, " )")))
     cat("\n")

### 2019-04-11 summary_tables.Rmd
Update the titles of the tables @rahuliitd the delhi duration table does not look right. Looking into it.

--- 2019-04-09 summary_tables.Rmd
@@
 for(city in cities){
-    td <- io[[city]]$dist %>% mutate_if(is.numeric, round, digits = round_to)
+  count_people <- nrow(io[[city]]$synth_pop)
+  td <- io[[city]]$dist %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)
-    print(kable(td, caption = paste("Distance table for ", city)))
+    print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person")))
     cat("\n")
+
 }
@@
 for(city in cities){
-    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to)
+  count_people <- nrow(io[[city]]$synth_pop)
+    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)
-    print(kable(td, caption = paste("Duration table for ", city)))
+    print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person")))
     cat("\n")

### 2019-04-11 summary_tables.Rmd
Regen summary tables

@@
   df <- io[[city]]$trip_scen_sets
-  df1 <- df %>% filter(scenario == "Baseline") %>% group_by(trip_id, trip_mode, scenario)
+  df1 <- df %>% dplyr::filter(scenario == "Baseline") %>% group_by(trip_id, trip_mode, scenario)
   u_trips <- nrow(df1)

### 2019-04-25 summary_tables_PA_AP.Rmd
Create summary_tables_PA_AP.Rmd
Rmarkdown file for summary tables of AP and PA for baseline and scenarios

---
title: "Summary Tables ITHIMR"
author: "Rahul Goel, MRC Epidemiology Unit, University of Cambridge, UK"
date: "25 April 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Introduction
These are the summary tables of the following items
1) Individual-level PM2.5 concentrations for baseline and scenarios
2) Individual-level physical activity for baseline and scenarios
3) Distance mode shares for baseline and scenarios

```{r, echo=FALSE}
#setwd('C:/Users/rg574/Dropbox/spatiotemporal analysis fatalities inida/Rajasthan tourism road deaths') #to create pretty tables
```

```{r, message=FALSE, warning=FALSE}
#library(INLA)     #loading the INLA package
library(ggplot2)  #loading ggplot package for plotting graphs
library(knitr)     
library(tidyr)
library(dplyr)

```

# Boxplots of Individual-level PM2.5 Concentrations

```{r, message=FALSE, warning=FALSE, echo=FALSE}
io <- readRDS("V:/Group/RG_PHM/ITHIM-R/results/multi_city/io.rds")

cities<- names(io)
cities<- cities[-2]

for (x in 1: length(cities))
{
  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, baseline:bus_scen, factor_key=TRUE)
  x<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot()+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")
  print(x)

}
```

# Descriptive tables of Individual-level PM2.5 Concentrations

```{r, message=FALSE, warning=FALSE, echo=FALSE}

for (x in 1: length(cities))
{
  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, baseline:bus_scen, factor_key=TRUE)
  summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(pm_conc),'5th'=quantile(pm_conc, 0.05),'50th'=quantile(pm_conc, 0.5),'95th'=quantile(pm_conc, 0.9)))
  summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1))
  names(summary)[1]<-"Scenario"
  print (as.character(cities[[x]]))
  print(summary)

}
```

# Boxplots of Individual-level Physical activity (MMETs)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

for (x in 1: length(cities))
{
  names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE)
  x<-ggplot(data_long, aes(x=scenario, y=mmet, fill=scenario)) + geom_boxplot()+ ggtitle(cities[x])+ labs(y="Weekly MMET", x = "Scenarios")
  print(x)

}
```

# Descriptive tables of Individual-level Physical activity (MMETs)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

for (x in 1: length(cities))
{
  names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE)
  summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(mmet),'5th'=quantile(mmet, 0.05),'50th'=quantile(mmet, 0.5),'95th'=quantile(mmet, 0.9)))
  summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1))
  names(summary)[1]<-"Scenario"
  print (as.character(cities[[x]]))
  print(summary)

}
```

# Distance share by scenarios in percentages (excludes city-specific minority modes-- auto rickshaws, subway, etc.)

```{r, message=FALSE, warning=FALSE, echo=FALSE}
for (x in 1: length(cities))
{

select<- c("car", "motorcycle", "walking", "bicycle", "bus")
io[[cities[x]]]$trip_scen_sets <-io[[cities[x]]]$trip_scen_sets[io[[cities[x]]]$trip_scen_sets$trip_mode %in% select,]

dist_scen<- io[[cities[x]]]$trip_scen_sets %>% group_by(trip_mode, scenario) %>% summarise(sum(trip_distance))
dist_scen<- as.data.frame (dist_scen)
names(dist_scen)[3]<-"distance"
dist_scen<-spread(dist_scen, trip_mode, distance)
dist_scen$sum<-rowSums (dist_scen[2:(ncol(dist_scen)-1)], na.rm = FALSE, dims = 1)
dist_scen_prop<- round(dist_scen[,2:(ncol(dist_scen)-1)]*100/dist_scen$sum, digits=1)
dist_scen_prop<- cbind(dist_scen[,1], dist_scen_prop)

names(dist_scen_prop)[1]<-"Scenario"
dist_scen_prop$Scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
print (as.character(cities[[x]]))
print(dist_scen_prop)
#library(knitr)
#library(kableExtra)
#kable(dist_scen_prop)
}
```

### 2019-04-25 summary_tables_PA_AP.Rmd
injury outcomes included

@@
 output:
-  pdf_document: default
   html_document: default
   word_document: default
+  pdf_document: default
 ---
@@
   names(summary)[1]<-"Scenario"
-  print (as.character(cities[[x]]))
+  print (as.character(cities[x]))
   print(summary)
@@
   names(summary)[1]<-"Scenario"
-  print (as.character(cities[[x]]))
+  print (as.character(cities[x]))
+  print(summary)
+
+}
+```
+
+# Descriptive tables of injury outcomes
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
+for (x in 1: length(cities))
+{
+
+  summary<- as.data.frame(io[[cities[x]]]$outcomes$injuries %>% group_by(scenario) %>% summarise('Ped'=sum(pedestrian,na.rm=T),'Bike'= sum(bicycle), 'Car'=sum(car), "Motorcycle"=sum(motorcycle,na.rm=T), "All modes"=sum(Deaths,na.rm=T)))
+  summary$scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
+  summary<- cbind(summary$scenario ,round(summary[,2:6], digits=0))
+  names(summary)[1]<-"Scenario"
+  print (as.character(cities[x]))
   print(summary)
@@
 dist_scen_prop$Scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-print (as.character(cities[[x]]))
+print (as.character(cities[x]))
 print(dist_scen_prop)

### 2019-04-25 summary_tables.Rmd
Added PA summary tables

@@
     colnames(td)[3:7] <- rownames(scen_prop)
-    print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person")))
+
+    count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
+    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
+    colnames(td)[3:7] <- rownames(scen_prop)
+    print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))
+    cat("\n")
+}
+
+```
+
+## Physical activity tables
+Case study specific physical activity tables for backgound (non-travel) PA, and baseline and five scenarios (non-travel + travel)
+
+```{r mmets = "asis"}
+
+for(city in cities){
+  df1 <- as.data.frame(io[[city]]$synth_pop$work_ltpa_marg_met)
+  colnames(df1) <- "mmet"
+
+  td1 <- df1 %>% filter(mmet == 0) %>% 
+    summarise(perct_zeros = round(n() / nrow(df1) * 100, 1))
+
+  td2 <- df1 %>%
+    summarise(min = min(mmet),
+              q_25th = round(quantile(mmet, 0.25), 1),
+              median = round(quantile(mmet, 0.50), 1),
+              mean = round(mean(mmet), 1),
+              q_75th = round(quantile(mmet, 0.75), 1),
+              q_90th = round(quantile(mmet, 0.95), 1),
+              max = round(max(mmet), 1))
+
+  td3 <- cbind(td1, td2)
+  td3 <- as.data.frame(t(td3))
+
+  df2 <- select(io[[city]]$outcomes$mmets, base_mmet : scen5_mmet)
+
+  td4 <- df2 %>% gather(key = "scenario", value = "mmet") %>% 
+    filter(mmet == 0) %>% 
+    group_by(scenario) %>% 
+    summarise(perct_zeros = round(n() / nrow(df2) * 100, 1))
+
+  td5 <- df2 %>% gather(key = "scenario", value = "mmet") %>% 
+    group_by(scenario) %>% 
+    summarise(min = min(mmet),
+              q_25th = round(quantile(mmet, 0.25), 1),
+              median = round(quantile(mmet, 0.50), 1),
+              mean = round(mean(mmet), 1),
+              q_75th = round(quantile(mmet, 0.75), 1),
+              q_90th = round(quantile(mmet, 0.95), 1),
+              max = round(max(mmet), 1))
+
+  td6 <- left_join(td4, td5, "scenario")
+  td6 <- as.data.frame(t(td6))
+  td6 <- td6[-1,]
+  td6 <- mutate_all(td6, function(x) as.numeric(as.character(x)))
+
+  td <- cbind(td3, td6)
+
+  colnames(td)[1] <- "Background (non-travel)"
+  colnames(td)[2] <- "Baseline"
+  colnames(td)[3:7] <- rownames(scen_prop)
+
+  print(kable(td, caption = paste("Marginal MET-h/wk for ", city)))
   cat("\n")

### 2019-04-28 summary_tables.Rmd

--- 2019-04-11 summary_tables.Rmd
@@
     colnames(td)[3:7] <- rownames(scen_prop)
-    print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person")))
+    print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (everyone)")))
+
+    count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
+    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
+    colnames(td)[3:7] <- rownames(scen_prop)
+    print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))
     cat("\n")

### 2019-04-29 summary_tables_PA_AP.Rmd
changes made to summary code

@@
 library(dplyr)
+library(kableExtra)
 
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
-io <- readRDS("V:/Group/RG_PHM/ITHIM-R/results/multi_city/io.rds")
+io <- readRDS("results/multi_city/io.rds")
 
@@
   print(summary)
+
+  summary %>%
+  kable() %>%
+  kable_styling()
 

### 2019-04-29 summary_tables.Rmd

--- 2019-04-25 summary_tables.Rmd
@@
     colnames(td)[3:7] <- rownames(scen_prop)
+    print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (everyone)")))
 

### 2019-04-29 summary_tables.Rmd

@@
 ## Scenario definition
-Table displays highest propensity for each distance category, for a given mode.
+Table displays highest propensity (%) for each distance category, for a given mode.
+
+In scenario generation, trips are sampled without replacement to increase the share of the requested mode up to the total shown.
+
+Distance categories are:
+
+0-1 km = \{trip distance < 2\}
+
+2-5 km = \{2 <= trip distance < 6\}
+
+6+ km = \{6 <= trip distance\}
 

### 2019-04-30 summary_tables_outcomes.Rmd

---
title: "Summary Tables"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(pracma)
library(summarytools)
library(dplyr)
knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')

```

```{r loadLibraries, echo = F, message = F}

st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
           plain.ascii       = FALSE,       # One of the essential settings
           style             = "rmarkdown", # Idem.
           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
           footnote          = NA,          # Keeping the results minimalistic
           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
                                            # much better results. Your mileage may vary

```

```{r load_objects = "asis"}
io <- readRDS("results/multi_city/io.rds")
# Assumes that multi_city_script.R has been run till 
cities <- c('accra','sao_paulo','delhi','bangalore')

# round to
round_to <- 1

# Set plot_mmets to F
plot_mmets <- F

```

```{r preprocessing}
scen_prop <- io$scen_prop
io <- io[-2]
for(i in 1:length(io)) io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
for(i in 1:length(io)) {
  io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
  io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
}
for(i in 1:length(io)) io[[i]]$demographic$age[io[[i]]$demographic$age=='5-9'] <- '05-9'
pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population)))
pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
for(i in 1:length(io)) io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
injury_col <- which(colnames(io$accra$outcomes$hb$deaths)=='scen1_deaths_inj')
scen_names <- rownames(scen_prop)

```

## TOTAL

Change in deaths total (for the city based on real population size) by age group by scenario

```{r scen_prop = "asis"}
death_totals <- lapply(io,function(x){
  temp <- sapply(1:5,function(y){
    xx <- x$outcomes$hb$deaths
    xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)])
    sapply(sort(unique(xx$age_cat)),function(z)
      sum(xxx[xx$age_cat==z]))
  })
  #rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
  colnames(temp) <- scen_names
  temp
  })

death_totals_rounded <- lapply(death_totals, function(x)round(x,round_to))

for(city in cities) {
  print(kable(death_totals_rounded[[city]], caption = paste("Change in deaths total in ", city)))
  cat("\n")
}
```

Change in deaths per 100,000 people by age group by scenario

```{r scen_prop = "asis"}
death_rates <- lapply(cities,function(x) death_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
names(death_rates) <- cities
death_rates_rounded <- lapply(death_rates, function(x)round(x,round_to))

for(city in cities) {
  print(kable(death_rates_rounded[[city]], caption = paste("Change in deaths per 100,000 in ", city)))
  cat("\n")
}
```

YLLs total (for the city based on real population size) by age group  by scenario

```{r scen_prop = "asis"}
yll_totals <- lapply(io,function(x){
  temp <- sapply(1:5,function(y){
    xx <- x$outcomes$hb$ylls
    xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)])
    sapply(sort(unique(xx$age_cat)),function(z)
      sum(xxx[xx$age_cat==z]))
  })
  colnames(temp) <- scen_names
  temp
  })

yll_totals_rounded <- lapply(yll_totals, function(x)round(x,round_to))

for(city in cities) {
  print(kable(yll_totals_rounded[[city]], caption = paste("Change in YLL total in ", city)))
  cat("\n")
}
```

YLLs per 100,000 people by age group by scenario

```{r scen_prop = "asis"}
yll_rates <- lapply(cities,function(x) yll_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
names(yll_rates) <- cities
yll_rates_rounded <- lapply(yll_rates, function(x)round(x,round_to))

for(city in cities) {
  print(kable(yll_rates_rounded[[city]], caption = paste("Change in YLLs per 100,000 in ", city)))
  cat("\n")
}
```

YLLs per 100,000 people by gender by age group scenario

```{r scen_prop = "asis"}
yll_totals <- lapply(io,function(x){
  temp <- sapply(1:5,function(y){
    xx <- x$outcomes$hb$ylls
    rowSums(xx[,seq(2+y,ncol(xx),by=5)])
  })
  rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
  colnames(temp) <- scen_names
  temp
  })

yll_rates <- lapply(cities,function(x) 
  yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[1],'_',z[2])),rownames(yll_totals[[x]])),]/
    t(repmat(io[[x]]$demographic$population,5,1))*100000)
names(yll_rates) <- cities
yll_rates_rounded <- lapply(yll_rates, function(x)round(x,round_to))

for(city in cities) {
  print(kable(yll_rates_rounded[[city]], caption = paste("Change in YLLs per 100,000 in ", city)))
  cat("\n")
}
```

## BY DISEASE

By disease (with non- injury separate & summed)

Change in deaths total (for the city based on real population size) by age group by scenario

```{r scen_prop = "asis"}
disease_totals <- lapply(io,function(x){
  temp <- sapply(1:5,function(y){
    xx <- x$outcomes$hb$deaths
    xxx <- rowSums(xx[,seq(2+y,injury_col-1,by=5)])
    sapply(sort(unique(xx$age_cat)),function(z)
      sum(xxx[xx$age_cat==z]))
  })
  #rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
  colnames(temp) <- scen_names
  temp
  })

disease_totals_rounded <- lapply(disease_totals, function(x)round(x,round_to))

for(city in cities) {
  print(kable(disease_totals_rounded[[city]], caption = paste("Change in deaths due to disease in ", city)))
  cat("\n")
}

```

Change in deaths per 100,000 people by age group by scenario

```{r scen_prop = "asis"}
disease_rates <- lapply(cities,function(x) disease_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
names(disease_rates) <- cities
disease_rates_rounded <- lapply(disease_rates, function(x)round(x,round_to))

for(city in cities) {
  print(kable(disease_rates_rounded[[city]], caption = paste("Change in deaths due to disease per 100,000 in ", city)))
  cat("\n")
}
```

YLLs total (for the city based on real population size) by age group by scenario

```{r scen_prop = "asis"}
disease_totals <- lapply(io,function(x){
  temp <- sapply(1:5,function(y){
    xx <- x$outcomes$hb$ylls
    xxx <- rowSums(xx[,seq(2+y,injury_col-1,by=5)])
    sapply(sort(unique(xx$age_cat)),function(z)
      sum(xxx[xx$age_cat==z]))
  })
  #rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
  colnames(temp) <- scen_names
  temp
  })

disease_totals_rounded <- lapply(disease_totals, function(x)round(x,round_to))

for(city in cities) {
  print(kable(disease_totals_rounded[[city]], caption = paste("Change in YLL due to disease in ", city)))
  cat("\n")
}

#injury_totals <- lapply(io,function(x){
#  xx <- x$outcomes$hb$deaths
#  sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y])))
#  })

#injury_totals_rounded <- lapply(injury_totals, function(x)round(x,round_to))

#for(city in cities) {
#  print(kable(injury_totals_rounded[[city]], caption = paste("Change in deaths due to injury in ", city)))
#  cat("\n")
#}
```

YLLs per 100,000 people by age group by scenario

```{r scen_prop = "asis"}
disease_rates <- lapply(cities,function(x) disease_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
names(disease_rates) <- cities
disease_rates_rounded <- lapply(disease_rates, function(x)round(x,round_to))

for(city in cities) {
  print(kable(disease_rates_rounded[[city]], caption = paste("Change in YLL due to disease per 100,000 in ", city)))
  cat("\n")
}
```

YLLs per 100,000 people by gender by scenario

```{r scen_prop = "asis"}

yll_totals <- lapply(io,function(x){
  temp <- sapply(1:5,function(y){
    xx <- x$outcomes$hb$ylls
    xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)])
  sapply(unique(xx$sex), function(z) sum(xxx[xx$sex==z],na.rm=T))
  })
  #rownames(temp) <- apply(x$outcomes$hb$ylls[,1:2],1,function(z)paste0(z[1],'_',z[2]))
  colnames(temp) <- rownames(scen_prop)
  temp
  })

yll_rates <- lapply(cities,function(x) 
  yll_totals[[x]][match(names(pop_by_gender[[x]]),rownames(yll_totals[[x]])),]/
    t(repmat(pop_by_gender[[x]],5,1))*100000)
names(yll_rates) <- cities
yll_rates_rounded <- lapply(yll_rates, function(x)round(x,round_to))

for(city in cities) {
  print(kable(yll_rates_rounded[[city]], caption = paste("Change in YLLs by gender per 100,000 in ", city)))
  cat("\n")
}
```

## BY PATHWAY

By pathway (with non- injury separate & summed)

Change in deaths total (for the city based on real population size) by age group by scenario

```{r}
injury_totals <- lapply(io,function(x){
  xx <- x$outcomes$hb$deaths
  xxx <- sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y])))
  colnames(xxx) <- scen_names
  xxx
  })
injury_totals_rounded <- lapply(injury_totals, function(x)round(x,round_to))

for(city in cities) {
  print(kable(injury_totals_rounded[[city]], caption = paste("Change in deaths due to injury in ", city)))
  cat("\n")
}
```

Change in deaths per 100,000 people by age group by scenario

```{r scen_prop = "asis"}
injury_rates <- lapply(cities,function(x) injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
names(injury_rates) <- cities
injury_rates_rounded <- lapply(injury_rates, function(x)round(x,round_to))

for(city in cities) {
  print(kable(injury_rates_rounded[[city]], caption = paste("Change in deaths due to injury per 100,000 in ", city)))
  cat("\n")
}
```

YLLs total (for the city based on real population size) by age group by scenario

```{r}
injury_totals <- lapply(io,function(x){
  xx <- x$outcomes$hb$ylls
  xxx <- sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y])))
  colnames(xxx) <- scen_names
  xxx
  })
injury_totals_rounded <- lapply(injury_totals, function(x)round(x,round_to))

for(city in cities) {
  print(kable(injury_totals_rounded[[city]], caption = paste("Change in YLLs due to injury in ", city)))
  cat("\n")
}
```

YLLs per 100,000 people by age group by scenario

```{r scen_prop = "asis"}
injury_rates <- lapply(cities,function(x) injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
names(injury_rates) <- cities
injury_rates_rounded <- lapply(injury_rates, function(x)round(x,round_to))

for(city in cities) {
  print(kable(injury_rates_rounded[[city]], caption = paste("Change in YLLs due to injury per 100,000 in ", city)))
  cat("\n")
}
```

### 2019-04-30 summary_tables_outcomes.Rmd

@@
 io <- io[-2]
-for(i in 1:length(io)) io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
 for(i in 1:length(io)) {
+  io[[i]]$demographic$age[io[[i]]$demographic$age=='5-9'] <- '05-9'
+  io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
+  io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
+  io[[i]]$outcomes$pathway_hb$deaths$age_cat[io[[i]]$outcomes$pathway_hb$deaths$age_cat=='5-9'] <- '05-9'
+  io[[i]]$outcomes$pathway_hb$ylls$age_cat[io[[i]]$outcomes$pathway_hb$ylls$age_cat=='5-9'] <- '05-9'
   io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
   io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
+  io[[i]]$outcomes$pathway_hb$deaths <- io[[i]]$outcomes$pathway_hb$deaths[,!sapply(names(io[[i]]$outcomes$pathway_hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
+  io[[i]]$outcomes$pathway_hb$ylls <- io[[i]]$outcomes$pathway_hb$ylls[,!sapply(names(io[[i]]$outcomes$pathway_hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
 }
-for(i in 1:length(io)) io[[i]]$demographic$age[io[[i]]$demographic$age=='5-9'] <- '05-9'
 pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population)))
 pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
-for(i in 1:length(io)) io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
 injury_col <- which(colnames(io$accra$outcomes$hb$deaths)=='scen1_deaths_inj')
+ap_cols <- which(sapply(colnames(io$accra$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
+pa_cols <- which(sapply(colnames(io$accra$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
 scen_names <- rownames(scen_prop)
@@
 }
+
+pa_totals <- lapply(io,function(x){
+  xx <- x$outcomes$pathway_hb$deaths
+  xxx <- sapply(1:5,function(y){
+    xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]])
+    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
+  })
+  colnames(xxx) <- scen_names
+  xxx
+  })
+pa_totals_rounded <- lapply(pa_totals, function(x)round(x,round_to))
+
+for(city in cities) {
+  print(kable(pa_totals_rounded[[city]], caption = paste("Change in deaths due to PA in ", city)))
+  cat("\n")
+}
+
+ap_totals <- lapply(io,function(x){
+  xx <- x$outcomes$pathway_hb$deaths
+  xxx <- sapply(1:5,function(y){
+    xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]])
+    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
+  })
+  colnames(xxx) <- scen_names
+  xxx
+  })
+ap_totals_rounded <- lapply(ap_totals, function(x)round(x,round_to))
+
+for(city in cities) {
+  print(kable(ap_totals_rounded[[city]], caption = paste("Change in deaths due to AP in ", city)))
+  cat("\n")
+}
 ```

### 2019-05-01 summary_tables_PA_AP.Rmd

@@
 dist_scen<-spread(dist_scen, trip_mode, distance)
+
 dist_scen$sum<-rowSums (dist_scen[2:(ncol(dist_scen)-1)], na.rm = FALSE, dims = 1)
@@
 print(dist_scen_prop)
+print(dist_scen)
 #library(knitr)

### 2019-05-02 summary_tables.Rmd
more refixes
Remove bus_driver from dist and dur dfs

@@
 ## Scenario definition
+Table displays highest propensity for each distance category, for a given mode.
+
 Table displays highest propensity (%) for each distance category, for a given mode.
@@
   count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dist %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
+  td <- io[[city]]$dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)
@@
   count_people <- nrow(io[[city]]$synth_pop)
-    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
+    td <- io[[city]]$dur %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)
@@
     count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-    td <- io[[city]]$dur %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
+    td <- io[[city]]$dur %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)

### 2019-05-02 summary_tables.Rmd
Transpose trip data frames

@@
 for(city in cities){
+
   df <- io[[city]]$trip_scen_sets
@@
   td[2:6, 1] <- rownames(scen_prop)
-  print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
+  td <- as.data.frame(t(td))
+  print(kable(t(td), caption = paste("Trip proportion (%) by mode for ", city)))
   cat("\n")

### 2019-05-03 summary_tables_PA_AP.Rmd
changed summary PA_AP - distance and source apportionment

@@
 
+# Descriptive tables of emission inventory
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
+city<-cities
+trans_share<-c(22.5, 40.0, 22.5, 28.1) 
+pm_conc<- c(50, 18, 122, 47 )
+shares<- as.data.frame(cbind(city,as.numeric(trans_share), pm_conc))
+names(shares)[2]<- "trans_share"
+for (x in 1:length(cities))
+{
+  modes<-names(unlist(io[[cities[x]]]$emission_inventory))
+  emissions<-as.data.frame(unlist(io[[cities[x]]]$emission_inventory))
+  city_emissions<-cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$emission_inventory)`))
+  names(city_emissions)[2]<- "emissions"
+  select<- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
+  city_emissions$modes<- as.character(city_emissions$modes)
+  city_emissions$modes[!(city_emissions$modes %in% select)]<- "other"
+  summary<-city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
+  names(summary)[2]<- "emissions"
+  summary$emissions <- round(summary$emissions*100/sum(summary$emissions), digits=1)
+  print (as.character(cities[x]))
+  summary(as.data.frame(summary))
+  summary[nrow(summary)+1,1]<-"Transport share"
+  summary[nrow(summary),2]<-as.character(shares$trans_share[x])
+  summary[nrow(summary)+1,1]<-"PM2.5 Conc"
+  summary[nrow(summary),2]<-as.character(shares$pm_conc[x])
+  print (summary)
+
+}
+
+```
+
 # Boxplots of Individual-level Physical activity (MMETs)
@@
   select<- c("car", "motorcycle", "walking", "bicycle", "bus")
-io[[cities[x]]]$trip_scen_sets <-io[[cities[x]]]$trip_scen_sets[io[[cities[x]]]$trip_scen_sets$trip_mode %in% select,]
+
+  io[[cities[x]]]$trip_scen_sets$trip_mode[!io[[cities[x]]]$trip_scen_sets$trip_mode %in% select]<- "other"
 
@@
   names(dist_scen)[3]<-"distance"
+  dist_scen$distance<- round(dist_scen$distance, digits=0)
     dist_scen<-spread(dist_scen, trip_mode, distance)
-
-dist_scen$sum<-rowSums (dist_scen[2:(ncol(dist_scen)-1)], na.rm = FALSE, dims = 1)
-dist_scen_prop<- round(dist_scen[,2:(ncol(dist_scen)-1)]*100/dist_scen$sum, digits=1)
+  dist_scen$sum<-rowSums (dist_scen[2:(ncol(dist_scen))], na.rm = FALSE, dims = 1)
+  dist_scen_prop<- round(dist_scen[,2:(ncol(dist_scen))]*100/dist_scen$sum, digits=1)
   dist_scen_prop<- cbind(dist_scen[,1], dist_scen_prop)

### 2019-05-03 summary_tables_PA_AP.Rmd
added striking vehicles

@@
 cities<- names(io)
-cities<- cities[-2]
+cities<- cities[-c(1,2)]
 
@@
 
+# Striking vehicles in scenarios
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
+for (x in 1: length(cities))
+{
+
+scenarios<- c('Baseline', 'Scenario 1', 'Scenario 2', 'Scenario 3', 'Scenario 4', 'Scenario 5')
+for (i in 1: length(scenarios))
+{
+whw<-as.data.frame(rowSums(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$whw))
+noov<-as.data.frame(rowSums(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$noov))
+names(whw)[1]<- as.character(scenarios[i])
+names(noov)[1]<-as.character(scenarios[i])
+
+if (i ==1 )
+{
+ summary<-round(rbind(whw,noov), digits = 1)
+} 
+else
+{
+summary<- cbind(summary, round(rbind(whw, noov), digits=1) ) 
+}
+
+}
+
+print (as.character(cities[x]))
+print(summary)
+
+}
+
+```
+
 # Distance share by scenarios in percentages (excludes city-specific minority modes-- auto rickshaws, subway, etc.)

### 2019-05-03 summary_tables.Rmd

--- 2019-05-02 summary_tables.Rmd
@@
   df1 <- df %>% dplyr::filter(scenario == "Baseline") %>% group_by(trip_id, trip_mode, scenario)
-  u_trips <- nrow(df1)
+  u_trips <- df1 %>% distinct(trip_id) %>% nrow()
   td <- df %>% group_by(trip_id, trip_mode, scenario) %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() /  u_trips * 100, round_to)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
-  td[2:6, 1] <- rownames(scen_prop)
+  #td[1:6, 1] <- append('baseline', rownames(scen_prop))
+  td <- as.data.frame(t(td))
+  names(td) <- lapply(append('baseline', rownames(scen_prop)), as.character)
+  td <- td[-1, ]
   print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))

### 2019-05-03 summary_tables_PA_AP.Rmd
made few changes in summary tables

@@
 
-city<-cities
+cities<- names(io)
+cities<- cities[-2]
 trans_share<-c(22.5, 40.0, 22.5, 28.1) 
 pm_conc<- c(50, 18, 122, 47 )
+city<-cities
 shares<- as.data.frame(cbind(city,as.numeric(trans_share), pm_conc))
@@
   print (as.character(cities[x]))
-  summary(as.data.frame(summary))
+  summary<-as.data.frame(summary)
   summary[nrow(summary)+1,1]<-"Transport share"
@@
 
+cities<- names(io)
+cities<- cities[-2]
 for (x in 1: length(cities))
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
-
+cities<- names(io)
+cities<- cities[-2]
 for (x in 1: length(cities))
@@
 print (as.character(cities[x]))
+names(summary)<-c("base", "walk", "bike", "car", "MC", "bus")
 print(summary)
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
+cities<- names(io)
+cities<- cities[-2]
 for (x in 1: length(cities))

### 2019-05-03 summary_tables.Rmd
Fix denominator for trips prop

--- 2019-05-02 summary_tables.Rmd
@@
   df1 <- df %>% dplyr::filter(scenario == "Baseline") %>% group_by(trip_id, trip_mode, scenario)
-  u_trips <- nrow(df1)
-  td <- df %>% group_by(trip_id, trip_mode, scenario) %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() /  u_trips * 100, round_to)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
-  td[2:6, 1] <- rownames(scen_prop)
+  u_trips <- df1 %>% distinct(trip_id) %>% nrow()
+  td <- df %>% distinct(trip_id, trip_mode, scenario, .keep_all = T) %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() /  u_trips * 100, round_to)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
+  #td[1:6, 1] <- append('baseline', rownames(scen_prop))
+  td <- as.data.frame(t(td))
+  names(td) <- lapply(append('baseline', rownames(scen_prop)), as.character)
+  td <- td[-1, ]
+  # a[-nrow(a),] 
   print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))

### 2019-05-03 summary_tables.Rmd
Fix trip prop by identifying unique trips and then using them as denominator

--- 2019-05-02 summary_tables.Rmd
@@
 for(city in cities){
-
   df <- io[[city]]$trip_scen_sets
-  df1 <- df %>% dplyr::filter(scenario == "Baseline") %>% group_by(trip_id, trip_mode, scenario)
-  u_trips <- nrow(df1)
-  td <- df %>% group_by(trip_id, trip_mode, scenario) %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() /  u_trips * 100, round_to)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
-  td[2:6, 1] <- rownames(scen_prop)
+  u_trips <- df %>% dplyr::filter(scenario == "Baseline") %>% summarise(uid <- n_distinct(trip_id)) %>% as.numeric()
+  td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() / u_trips * 100, 1)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
   td <- as.data.frame(t(td))
-  print(kable(t(td), caption = paste("Trip proportion (%) by mode for ", city)))
+  names(td) <- lapply(append('baseline', rownames(scen_prop)), as.character)
+  td <- td[-1, ]
+  print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
   cat("\n")

### 2019-05-08 summary_tables_outcomes.Rmd
new results tables

--- 2019-04-30 summary_tables_outcomes.Rmd
@@
 ```{r preprocessing}
+collapse_ages <- function(data,ages=c('15-49','50-69'),age_lab='age_cat'){
+  target_min_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][1]))
+  target_max_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][2]))
+  min_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][1]))
+  max_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][2]))
+  genders <- unique(data$sex)
+  if(ncol(data)>3) {
+    reformatted <- do.call(rbind,lapply(1:length(ages),
+    function(x) t(sapply(genders,
+          function(y) 
+            if(ncol(data)>3) colSums(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
+          else sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
+         ))
+  ))
+  data.frame(age_cat=rep(ages,each=2),sex=rep(genders,2),reformatted,stringsAsFactors=F)
+  }else{
+  reformatted <- do.call(c,lapply(1:length(ages),
+    function(x) sapply(genders,
+          function(y) 
+            sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
+         )
+  ))
+  data.frame(age=rep(ages,each=2),sex=rep(genders,2),population=as.numeric(reformatted),stringsAsFactors=F)
+  }
+}
+
 scen_prop <- io$scen_prop
 io <- io[-2]
-for(i in 1:length(io)) io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
 for(i in 1:length(io)) {
+  io[[i]]$demographic$age[io[[i]]$demographic$age=='5-9'] <- '05-9'
+  io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
+  io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
+  io[[i]]$outcomes$pathway_hb$deaths$age_cat[io[[i]]$outcomes$pathway_hb$deaths$age_cat=='5-9'] <- '05-9'
+  io[[i]]$outcomes$pathway_hb$ylls$age_cat[io[[i]]$outcomes$pathway_hb$ylls$age_cat=='5-9'] <- '05-9'
   io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
   io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
+  io[[i]]$outcomes$pathway_hb$deaths <- io[[i]]$outcomes$pathway_hb$deaths[,!sapply(names(io[[i]]$outcomes$pathway_hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
+  io[[i]]$outcomes$pathway_hb$ylls <- io[[i]]$outcomes$pathway_hb$ylls[,!sapply(names(io[[i]]$outcomes$pathway_hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
 }
-for(i in 1:length(io)) io[[i]]$demographic$age[io[[i]]$demographic$age=='5-9'] <- '05-9'
+for(city in cities)
+  for(type in c('hb','pathway_hb'))
+    for(out in c('deaths','ylls'))
+      io[[city]]$outcomes[[type]][[out]] <- collapse_ages(io[[city]]$outcomes[[type]][[out]])
+for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')
 pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population)))
 pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
-for(i in 1:length(io)) io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
 injury_col <- which(colnames(io$accra$outcomes$hb$deaths)=='scen1_deaths_inj')
+ap_cols <- which(sapply(colnames(io$accra$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
+pa_cols <- which(sapply(colnames(io$accra$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
 scen_names <- rownames(scen_prop)
@@
 
+death_totals_rounded <- lapply(death_totals_rounded,function(x)rbind(x,Total=colSums(x)))
+
 for(city in cities) {
@@
 names(death_rates) <- cities
+
 death_rates_rounded <- lapply(death_rates, function(x)round(x,round_to))
+
+death_rates_rounded <- lapply(death_rates_rounded,function(x)rbind(x,Total=colSums(x)))
 
@@
   })
-
-yll_totals_rounded <- lapply(yll_totals, function(x)round(x,round_to))
+yll_totals_rounded <- lapply(yll_totals,function(x)rbind(x,Total=colSums(x)))
+yll_totals_rounded <- lapply(yll_totals_rounded, function(x)round(x,round_to))
 
@@
 names(yll_rates) <- cities
-yll_rates_rounded <- lapply(yll_rates, function(x)round(x,round_to))
+yll_rates_rounded <- lapply(yll_rates,function(x)rbind(x,Total=colSums(x)))
+yll_rates_rounded <- lapply(yll_rates_rounded, function(x)round(x,round_to))
 
@@
 yll_rates <- lapply(cities,function(x) 
-  yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[1],'_',z[2])),rownames(yll_totals[[x]])),]/
+  yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[2],'_',z[1])),rownames(yll_totals[[x]])),]/
     t(repmat(io[[x]]$demographic$population,5,1))*100000)
 names(yll_rates) <- cities
-yll_rates_rounded <- lapply(yll_rates, function(x)round(x,round_to))
+yll_rates_rounded <- lapply(yll_rates,function(x)rbind(x,Total=colSums(x)))
+yll_rates_rounded <- lapply(yll_rates_rounded, function(x)round(x,round_to))
 
@@
   print(kable(yll_rates_rounded[[city]], caption = paste("Change in YLLs per 100,000 in ", city)))
-  cat("\n")
-}
-```
-
-## BY DISEASE
-
-By disease (with non- injury separate & summed)
-
-Change in deaths total (for the city based on real population size) by age group by scenario
-
-```{r scen_prop = "asis"}
-disease_totals <- lapply(io,function(x){
-  temp <- sapply(1:5,function(y){
-    xx <- x$outcomes$hb$deaths
-    xxx <- rowSums(xx[,seq(2+y,injury_col-1,by=5)])
-    sapply(sort(unique(xx$age_cat)),function(z)
-      sum(xxx[xx$age_cat==z]))
-  })
-  #rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
-  colnames(temp) <- scen_names
-  temp
-  })
-
-disease_totals_rounded <- lapply(disease_totals, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(disease_totals_rounded[[city]], caption = paste("Change in deaths due to disease in ", city)))
-  cat("\n")
-}
-
-```
-
-Change in deaths per 100,000 people by age group by scenario
-
-```{r scen_prop = "asis"}
-disease_rates <- lapply(cities,function(x) disease_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
-names(disease_rates) <- cities
-disease_rates_rounded <- lapply(disease_rates, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(disease_rates_rounded[[city]], caption = paste("Change in deaths due to disease per 100,000 in ", city)))
-  cat("\n")
-}
-```
-
-YLLs total (for the city based on real population size) by age group by scenario
-
-```{r scen_prop = "asis"}
-disease_totals <- lapply(io,function(x){
-  temp <- sapply(1:5,function(y){
-    xx <- x$outcomes$hb$ylls
-    xxx <- rowSums(xx[,seq(2+y,injury_col-1,by=5)])
-    sapply(sort(unique(xx$age_cat)),function(z)
-      sum(xxx[xx$age_cat==z]))
-  })
-  #rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
-  colnames(temp) <- scen_names
-  temp
-  })
-
-disease_totals_rounded <- lapply(disease_totals, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(disease_totals_rounded[[city]], caption = paste("Change in YLL due to disease in ", city)))
-  cat("\n")
-}
-
-#injury_totals <- lapply(io,function(x){
-#  xx <- x$outcomes$hb$deaths
-#  sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y])))
-#  })
-
-#injury_totals_rounded <- lapply(injury_totals, function(x)round(x,round_to))
-
-#for(city in cities) {
-#  print(kable(injury_totals_rounded[[city]], caption = paste("Change in deaths due to injury in ", city)))
-#  cat("\n")
-#}
-```
-
-YLLs per 100,000 people by age group by scenario
-
-```{r scen_prop = "asis"}
-disease_rates <- lapply(cities,function(x) disease_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
-names(disease_rates) <- cities
-disease_rates_rounded <- lapply(disease_rates, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(disease_rates_rounded[[city]], caption = paste("Change in YLL due to disease per 100,000 in ", city)))
-  cat("\n")
-}
-```
-
-YLLs per 100,000 people by gender by scenario
-
-```{r scen_prop = "asis"}
-
-yll_totals <- lapply(io,function(x){
-  temp <- sapply(1:5,function(y){
-    xx <- x$outcomes$hb$ylls
-    xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)])
-  sapply(unique(xx$sex), function(z) sum(xxx[xx$sex==z],na.rm=T))
-  })
-  #rownames(temp) <- apply(x$outcomes$hb$ylls[,1:2],1,function(z)paste0(z[1],'_',z[2]))
-  colnames(temp) <- rownames(scen_prop)
-  temp
-  })
-
-yll_rates <- lapply(cities,function(x) 
-  yll_totals[[x]][match(names(pop_by_gender[[x]]),rownames(yll_totals[[x]])),]/
-    t(repmat(pop_by_gender[[x]],5,1))*100000)
-names(yll_rates) <- cities
-yll_rates_rounded <- lapply(yll_rates, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(yll_rates_rounded[[city]], caption = paste("Change in YLLs by gender per 100,000 in ", city)))
   cat("\n")
@@
   })
-injury_totals_rounded <- lapply(injury_totals, function(x)round(x,round_to))
+injury_totals_rounded <- lapply(injury_totals,function(x)rbind(x,Total=colSums(x)))
+injury_totals_rounded <- lapply(injury_totals_rounded, function(x)round(x,round_to))
 
@@
   print(kable(injury_totals_rounded[[city]], caption = paste("Change in deaths due to injury in ", city)))
+  cat("\n")
+}
+
+pa_totals <- lapply(io,function(x){
+  xx <- x$outcomes$pathway_hb$deaths
+  xxx <- sapply(1:5,function(y){
+    xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]])
+    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
+  })
+  colnames(xxx) <- scen_names
+  xxx
+  })
+pa_totals_rounded <- lapply(pa_totals,function(x)rbind(x,Total=colSums(x)))
+pa_totals_rounded <- lapply(pa_totals_rounded, function(x)round(x,round_to))
+
+for(city in cities) {
+  print(kable(pa_totals_rounded[[city]], caption = paste("Change in deaths due to PA in ", city)))
+  cat("\n")
+}
+
+ap_totals <- lapply(io,function(x){
+  xx <- x$outcomes$pathway_hb$deaths
+  xxx <- sapply(1:5,function(y){
+    xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]])
+    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
+  })
+  colnames(xxx) <- scen_names
+  xxx
+  })
+ap_totals_rounded <- lapply(ap_totals,function(x)rbind(x,Total=colSums(x)))
+ap_totals_rounded <- lapply(ap_totals_rounded, function(x)round(x,round_to))
+
+for(city in cities) {
+  print(kable(ap_totals_rounded[[city]], caption = paste("Change in deaths due to AP in ", city)))
   cat("\n")
@@
 names(injury_rates) <- cities
-injury_rates_rounded <- lapply(injury_rates, function(x)round(x,round_to))
+injury_rates_rounded <- lapply(injury_rates,function(x)rbind(x,Total=colSums(x)))
+injury_rates_rounded <- lapply(injury_rates_rounded, function(x)round(x,round_to))
 
@@
 ```
+
+## BY DISEASE
+
+Change in deaths total (for the city based on real population size) by scenario
+
+```{r}
+
+disease_totals <- lapply(io,function(x){
+  xx <- x$outcomes$hb$deaths
+  xxx <- sapply(1:5,function(y){
+    colSums(xx[,seq(y+2,ncol(xx),by=5)])
+  })
+  colnames(xxx) <- scen_names
+  rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y))
+  xxx
+  })
+disease_totals_rounded <- lapply(disease_totals,function(x)rbind(x,Total=colSums(x)))
+disease_totals_rounded <- lapply(disease_totals_rounded, function(x)round(x,round_to))
+
+for(city in cities) {
+  print(kable(disease_totals_rounded[[city]], caption = paste("Change in deaths due to disease in ", city)))
+  cat("\n")
+}
+
+```
+
+Change in deaths per 100,000 people by scenario
+
+```{r scen_prop = "asis"}
+disease_rates <- lapply(cities,function(x) disease_totals[[x]]/sum(pop_by_age[[x]])*100000)
+names(disease_rates) <- cities
+disease_rates_rounded <- lapply(disease_rates,function(x)rbind(x,Total=colSums(x)))
+disease_rates_rounded <- lapply(disease_rates_rounded, function(x)round(x,round_to))
+
+for(city in cities) {
+  print(kable(disease_rates_rounded[[city]], caption = paste("Change in deaths due to disease per 100,000 in ", city)))
+  cat("\n")
+}
+```
+
+Change in YLL total (for the city based on real population size) by scenario
+
+```{r}
+
+disease_totals <- lapply(io,function(x){
+  xx <- x$outcomes$hb$ylls
+  xxx <- sapply(1:5,function(y){
+    colSums(xx[,seq(y+2,ncol(xx),by=5)])
+  })
+  colnames(xxx) <- scen_names
+  rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y))
+  xxx
+  })
+disease_totals_rounded <- lapply(disease_totals,function(x)rbind(x,Total=colSums(x)))
+disease_totals_rounded <- lapply(disease_totals_rounded, function(x)round(x,round_to))
+
+for(city in cities) {
+  print(kable(disease_totals_rounded[[city]], caption = paste("Change in YLL due to disease in ", city)))
+  cat("\n")
+}
+
+```
+
+Change in YLL per 100,000 people by scenario
+
+```{r scen_prop = "asis"}
+disease_rates <- lapply(cities,function(x) disease_totals[[x]]/sum(pop_by_age[[x]])*100000)
+names(disease_rates) <- cities
+disease_rates_rounded <- lapply(disease_rates,function(x)rbind(x,Total=colSums(x)))
+disease_rates_rounded <- lapply(disease_rates_rounded, function(x)round(x,round_to))
+
+for(city in cities) {
+  print(kable(disease_rates_rounded[[city]], caption = paste("Change in YLL due to disease per 100,000 in ", city)))
+  cat("\n")
+}
+```

### 2019-05-08 summary_tables.Rmd
Add date

@@
 output: html_document
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 ---
@@
   count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
+  td <- io[[city]]$dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)
@@
   count_people <- nrow(io[[city]]$synth_pop)
-    td <- io[[city]]$dur %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
+    td <- io[[city]]$dur %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)
@@
     count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-    td <- io[[city]]$dur %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, funs(round((.) / count_people, round_to)))
+    td <- io[[city]]$dur %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
     colnames(td)[3:7] <- rownames(scen_prop)

### 2019-05-08 summary_tables_outcomes.Rmd
new results tables

@@
 plot_mmets <- F
+
+sum_and_round_and_print <- function(data,text=''){
+  data <- lapply(data, function(x)round(x,round_to))
+  data <- lapply(data,function(x)rbind(x,Total=colSums(x)))
+  for(city in cities) {
+    print(kable(data[[city]], caption = paste(text, city)))
+    cat("\n")
+  }
+}
 
@@
 
-death_totals_rounded <- lapply(death_totals, function(x)round(x,round_to))
-
-death_totals_rounded <- lapply(death_totals_rounded,function(x)rbind(x,Total=colSums(x)))
-
-for(city in cities) {
-  print(kable(death_totals_rounded[[city]], caption = paste("Change in deaths total in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(death_totals,"Change in deaths total in ")
 ```
@@
 
-death_rates_rounded <- lapply(death_rates, function(x)round(x,round_to))
-
-death_rates_rounded <- lapply(death_rates_rounded,function(x)rbind(x,Total=colSums(x)))
-
-for(city in cities) {
-  print(kable(death_rates_rounded[[city]], caption = paste("Change in deaths per 100,000 in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(death_rates,"Change in deaths per 100,000 in ")
 ```
@@
   })
-yll_totals_rounded <- lapply(yll_totals,function(x)rbind(x,Total=colSums(x)))
-yll_totals_rounded <- lapply(yll_totals_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(yll_totals_rounded[[city]], caption = paste("Change in YLL total in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(yll_totals,"Change in YLL total in ")
 ```
@@
 names(yll_rates) <- cities
-yll_rates_rounded <- lapply(yll_rates,function(x)rbind(x,Total=colSums(x)))
-yll_rates_rounded <- lapply(yll_rates_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(yll_rates_rounded[[city]], caption = paste("Change in YLLs per 100,000 in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
 ```
@@
   })
-injury_totals_rounded <- lapply(injury_totals,function(x)rbind(x,Total=colSums(x)))
-injury_totals_rounded <- lapply(injury_totals_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(injury_totals_rounded[[city]], caption = paste("Change in deaths due to injury in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(injury_totals,"Change in deaths due to injury in ")
 
@@
   })
-pa_totals_rounded <- lapply(pa_totals,function(x)rbind(x,Total=colSums(x)))
-pa_totals_rounded <- lapply(pa_totals_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(pa_totals_rounded[[city]], caption = paste("Change in deaths due to PA in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(pa_totals,"Change in deaths due to PA in ")
 
@@
   })
-ap_totals_rounded <- lapply(ap_totals,function(x)rbind(x,Total=colSums(x)))
-ap_totals_rounded <- lapply(ap_totals_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(ap_totals_rounded[[city]], caption = paste("Change in deaths due to AP in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(ap_totals,"Change in deaths due to AP in ")
 ```
@@
 names(injury_rates) <- cities
-injury_rates_rounded <- lapply(injury_rates,function(x)rbind(x,Total=colSums(x)))
-injury_rates_rounded <- lapply(injury_rates_rounded, function(x)round(x,round_to))
+sum_and_round_and_print(injury_rates,"Change in deaths due to injury per 100,000 in ")
 
-for(city in cities) {
-  print(kable(injury_rates_rounded[[city]], caption = paste("Change in deaths due to injury per 100,000 in ", city)))
-  cat("\n")
-}
+pa_rates <- lapply(cities,function(x) pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+names(pa_rates) <- cities
+sum_and_round_and_print(pa_rates,"Change in deaths due to PA per 100,000 in ")
+
+ap_rates <- lapply(cities,function(x) ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+names(ap_rates) <- cities
+sum_and_round_and_print(ap_rates,"Change in deaths due to AP per 100,000 in ")
 ```
@@
   })
-injury_totals_rounded <- lapply(injury_totals,function(x)rbind(x,Total=colSums(x)))
-injury_totals_rounded <- lapply(injury_totals_rounded, function(x)round(x,round_to))
+sum_and_round_and_print(injury_totals,"Change in YLLs due to injury in ")
 
-for(city in cities) {
-  print(kable(injury_totals_rounded[[city]], caption = paste("Change in YLLs due to injury in ", city)))
-  cat("\n")
-}
+pa_totals <- lapply(io,function(x){
+  xx <- x$outcomes$pathway_hb$ylls
+  xxx <- sapply(1:5,function(y){
+    xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]])
+    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
+  })
+  colnames(xxx) <- scen_names
+  xxx
+  })
+sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
+
+ap_totals <- lapply(io,function(x){
+  xx <- x$outcomes$pathway_hb$ylls
+  xxx <- sapply(1:5,function(y){
+    xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]])
+    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
+  })
+  colnames(xxx) <- scen_names
+  xxx
+  })
+sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ")
 ```
@@
 names(injury_rates) <- cities
-injury_rates_rounded <- lapply(injury_rates,function(x)rbind(x,Total=colSums(x)))
-injury_rates_rounded <- lapply(injury_rates_rounded, function(x)round(x,round_to))
+sum_and_round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ")
 
-for(city in cities) {
-  print(kable(injury_rates_rounded[[city]], caption = paste("Change in YLLs due to injury per 100,000 in ", city)))
-  cat("\n")
-}
+pa_rates <- lapply(cities,function(x) pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+names(pa_rates) <- cities
+sum_and_round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ")
+
+ap_rates <- lapply(cities,function(x) ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+names(ap_rates) <- cities
+sum_and_round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ")
 ```
@@
   })
-disease_totals_rounded <- lapply(disease_totals,function(x)rbind(x,Total=colSums(x)))
-disease_totals_rounded <- lapply(disease_totals_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(disease_totals_rounded[[city]], caption = paste("Change in deaths due to disease in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(disease_totals,"Change in deaths due to disease in ")
 
@@
 names(disease_rates) <- cities
-disease_rates_rounded <- lapply(disease_rates,function(x)rbind(x,Total=colSums(x)))
-disease_rates_rounded <- lapply(disease_rates_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(disease_rates_rounded[[city]], caption = paste("Change in deaths due to disease per 100,000 in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ")
 ```
@@
   })
-disease_totals_rounded <- lapply(disease_totals,function(x)rbind(x,Total=colSums(x)))
-disease_totals_rounded <- lapply(disease_totals_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(disease_totals_rounded[[city]], caption = paste("Change in YLL due to disease in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(disease_totals,"Change in YLL due to disease in ")
 
@@
 names(disease_rates) <- cities
-disease_rates_rounded <- lapply(disease_rates,function(x)rbind(x,Total=colSums(x)))
-disease_rates_rounded <- lapply(disease_rates_rounded, function(x)round(x,round_to))
-
-for(city in cities) {
-  print(kable(disease_rates_rounded[[city]], caption = paste("Change in YLL due to disease per 100,000 in ", city)))
-  cat("\n")
-}
+sum_and_round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ")
 ```

### 2019-05-08 summary_tables_PA_AP.Rmd

--- 2019-05-03 summary_tables_PA_AP.Rmd
@@
 
-cities<- names(io)
-cities<- cities[-c(1,2)]
+cities <- c("accra", "sao_paulo", "delhi", "bangalore")
 
@@
 {
+  print(cities[x])
   names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
@@
 
-city<-cities
 trans_share<-c(22.5, 40.0, 22.5, 28.1) 
 pm_conc<- c(50, 18, 122, 47 )
+city<-cities
 shares<- as.data.frame(cbind(city,as.numeric(trans_share), pm_conc))
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
-
 for (x in 1: length(cities))

### 2019-05-09 summary_tables.Rmd
Add case study specific propensities - as outputted from multi-city script

@@
 cities <- c('accra','sao_paulo','delhi','bangalore')
+# cities <- c('sao_paulo')
 
@@
 ```
+
+## Case study specific propensities
+
+|accra|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 96.7|66.2|2.0 |
+| bicycle | 0.6|0.3|0.4 |
+| car | 1.0|10.4|16.9 |
+| motorcycle | 0.2|0.9|2.7 |
+| bus | 0.0|18.6|69.4 |
+
+|sao_paulo|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 65.4|6.9|2.9 |
+| bicycle | 0.7|0.9|0.3 |
+| car | 22.2|41.9|27.2 |
+| motorcycle | 1.3|2.7|3.5 |
+| bus | 9.7|43.0|35.8 |
+
+|delhi|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 83.5|28.8|1.8 |
+| bicycle | 2.0|5.8|4.7 |
+| car | 2.0|10.7|16.8 |
+| motorcycle | 5.6|23.3|20.7 |
+| bus | 1.2|18.2|39.1 |
+
+|bangalore|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 89.7|28.6|1.2 |
+| bicycle | 2.6|10.9|0.9 |
+| car | 0.8|5.0|4.9 |
+| motorcycle | 4.8|37.3|48.2 |
+| bus | 0.6|8.3|29.3 |
 

### 2019-05-13 summary_tables.Rmd

@@
 
+    count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
+    td <- io[[city]]$dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+    colnames(td)[3:7] <- rownames(scen_prop)
+    print(kable(td, caption = paste("Distance table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))
+
 }

### 2019-05-13 diagnostic/dr_curves.Rmd
dose--response curves added and appended below parameter_table_all

---
title: "Dose--response curves"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(ithimr)
```

```{r pa, include=FALSE}

global_path <- file.path(find.package('ithimr',lib.loc=.libPaths()), 'extdata/global/')
global_path <- paste0(global_path, "/")

DISEASE_INVENTORY <- read.csv(paste0(global_path,"dose_response/disease_outcomes_lookup.csv"))

list_of_files <- list.files(path = paste0(global_path,"dose_response/drpa/extdata/"), recursive = TRUE, pattern = "\\.csv$", full.names = TRUE)
for (i in 1:length(list_of_files)){
  assign(stringr::str_sub(basename(list_of_files[[i]]), end = -5),
         readr::read_csv(list_of_files[[i]],col_types = cols()),
         pos = 1)
}

########## dose response

doses_vector <- as.numeric(0:50)
cols <- rainbow(sum(DISEASE_INVENTORY$physical_activity == 1))
cause_indices <- which(DISEASE_INVENTORY$physical_activity == 1)

############# median dose-response curves
# PA_DOSE_RESPONSE_QUANTILE <<- 0.5
# #pdf('results/dose_response/PA_dose_response.pdf',width=11,height=11);par(mar=c(5,5,2,1))
# for ( j in 1:length(cause_indices)){
#   cause <- as.character(DISEASE_INVENTORY$pa_acronym[cause_indices[j]])
#   # get name of disease
#   print(cause)
#   return_vector <- PA_dose_response(cause = cause,dose = doses_vector)
#   
#   if(j==1){
#     plot(doses_vector,return_vector$rr,type='l',ylim=0:1,main='',ylab='Relative risk',xlab='Dose',
#          frame=F,col=cols[j],xlim=range(doses_vector),cex.axis=1.5,cex.lab=1.5,cex.main=1.5,lwd=2) 
#   }else{
#     lines(doses_vector,return_vector$rr,col=cols[j],lwd=2)
#   }
# }
# legend(legend=DISEASE_INVENTORY$pa_acronym[DISEASE_INVENTORY$physical_activity==1],col=cols,bty='n',x=0,y=0.5,lwd=2)
# #dev.off()
# 
# ############# sample dose-response curves
# nsamples <- 10
# parameters <- ithim_setup_parameters(NSAMPLES=nsamples,
#                                      PA_DOSE_RESPONSE_QUANTILE=T)
# #pdf('results/dose_response/PA_dose_response_sample.pdf',width=11,height=11);
# par(mar=c(5,5,2,1),mfrow=c(3,3))
# for ( j in 1:length(cause_indices)){
#   cause <- as.character(DISEASE_INVENTORY$pa_acronym[cause_indices[j]])
#   # get name of disease
#   print(cause)
#   for(seed in 1:nsamples){
#     for(i in 1:length(parameters))
#       assign(names(parameters)[i],parameters[[i]][[seed]],pos=1)
#     return_vector <- PA_dose_response(cause = cause,dose = doses_vector)
#     
#     if(seed==1){
#       plot(doses_vector,return_vector$rr,type='l',ylim=0:1,main=cause,ylab='Relative risk',xlab='Dose',
#            frame=F,col=cols[j],xlim=range(doses_vector),cex.axis=1.5,cex.lab=1.5,cex.main=1.5,lwd=1) 
#     }else{
#       lines(doses_vector,return_vector$rr,col=cols[j],lwd=1)
#     }
#   }
# }
# #dev.off()
# 
# ################ explanatory graph
# library(diagram)
# j <- 3
# ylow <- 0.6
# doses_clean <- doses_vector
# {x11(); par(mfrow=c(2,2),mar=c(2,5,1,1))
#   pa_dn <- as.character(DISEASE_INVENTORY$pa_acronym[j])
#   pa_n <- as.character(DISEASE_INVENTORY$acronym[j])
#   outcome_type <- ifelse(pa_dn%in%c('lung_cancer','stroke'), 'incidence' , 'mortality')
#   dose <- doses_clean
#   if(pa_dn %in% c('total_cancer','coronary_heart_disease')) dose[dose>35] <- 35
#   cause<-pa_dn
#   fname <- paste(cause, outcome_type, sep = "_")
#   lookup_table <- get(paste0(fname))
#   lookup_df <- as.data.frame(lookup_table)
#   rr <- approx(x=lookup_df$dose,y=lookup_df$RR,xout=dose,yleft=1,yright=min(lookup_df$RR))$y
#   lb <- approx(x = lookup_df$dose,y = lookup_df$lb,xout = dose,yleft = 1,yright = min(lookup_df$lb))$y
#   ub <-approx(x = lookup_df$dose,y = lookup_df$ub,xout = dose,yleft = 1,yright = min(lookup_df$ub))$y
#   plot(doses_clean,rr,lwd=2,type='l',col=rgb(0.6,0.2,0.2),
#        xlim=range(doses_clean),ylim=c(ylow,1),ylab='Response',frame=F,cex.lab=1.5,cex.axis=1.5,xlab='')
#   lines(doses_clean,lb,lwd=2,type='l',lty=2,col=rgb(0.6,0.2,0.2,1))
#   lines(doses_clean,ub,lwd=2,type='l',lty=2,col=rgb(0.6,0.2,0.2,1))
#   text('Mean',x=40,y=0.8,col=rgb(0.6,0.2,0.2),cex=1.5)
#   text('Lower bound',x=35,y=0.66,col=rgb(0.6,0.2,0.2),cex=1.5)
#   text('Upper bound',x=35,y=0.93,col=rgb(0.6,0.2,0.2),cex=1.5)
#   par(mar=c(2,2,1,1))
#   for(i in 1:500){
#     x<-runif(1)
#     rr0 <- truncnorm::qtruncnorm(p=x, mean=rr, sd=(rr-lb)/1.96,a=0)#a=lb,b=ub)#
#     if(i==1){plot(doses_clean,rr0,type='l',col=rgb(0.6,0.2,0.2,0.1),
#                   xlim=range(doses_clean),ylim=c(ylow,1),ylab='',frame=F,cex.lab=1.5,cex.axis=1.5,xlab='') 
#       text('Samples',x=35,y=0.97,col=rgb(0.6,0.2,0.2),cex=1.5)
#     }else lines(doses_clean,rr0,col=rgb(0.6,0.2,0.2,0.1))
#   }
#   quant <- 0.5
#   par(mar=c(5,5,1,1))
#   rr0 <- truncnorm::qtruncnorm(p=quant, mean=rr, sd=(rr-lb)/1.96,a=0)
#   plot(doses_clean,rr0,lwd=2,type='l',col='grey',
#        xlim=range(doses_clean),ylim=c(ylow,1),ylab='Response',frame=F,cex.lab=1.5,cex.axis=1.5,xlab='Dose')
#   doses <- c(5,25,45)
#   text('Median',x=10,y=0.97,col='grey',cex=1.5)
#   abline(h=ylow,lwd=2,lty=2,col='grey')
#   points(doses,rep(ylow,length(doses)),col=rgb(0.6,0.2,0.2),pch=18,cex=1.5)
#   points(doses,rr0[which(doses_clean%in%doses)],col=rgb(0.6,0.2,0.2),pch=15,cex=1.5)
#   for(dose in doses) arrows(x0=dose,y0=ylow,y1=rr0[which(doses_clean==dose)]-0.02,col=rgb(0.6,0.2,0.2),length=0.1,lty=1,lwd=2)
#   par(mar=c(5,2,1,1))
#   cols <- c('hotpink','navyblue','darkorange2')
#   quants <- c(0.1,0.6,0.9)
#   for(i in 1:length(quants)){
#     quant <- quants[i]
#     rr0 <- truncnorm::qtruncnorm(p=quant, mean=rr, sd=(rr-lb)/1.96,a=0)
#     if(i==1) {plot(doses_clean,rr0,lwd=2,type='l',col=cols[i],
#                    xlim=range(doses_clean),ylim=c(ylow,1),ylab='',frame=F,cex.lab=1.5,cex.axis=1.5,xlab='Dose')
#       for(dose in doses) arrows(x0=dose,y0=ylow,y1=rr0[which(doses_clean==dose)]-0.02,col=cols[i],length=0.1,lty=1,lwd=2)
#     }  else{
#       lines(doses_clean,rr0,lwd=2,type='l',col=cols[i])
#       for(dose in doses) 
#         curvedarrow(from=c(dose,ylow),to=c(dose,rr0[which(doses_clean==dose)]-0.01),lty=1,lwd=2,lcol=cols[i],
#                     arr.col=cols[i],arr.pos=0.65+0.05*i,curve=-70+30*i,arr.type='curved',arr.adj=0,endhead=T)
#     }
#     abline(h=ylow,lwd=2,lty=2,col='grey')
#     points(doses,rr0[which(doses_clean%in%doses)],col=cols[i],pch=15,cex=1.5)
#     
#     points(doses,rep(ylow,length(doses)),col=rgb(0.6,0.2,0.2),pch=18,cex=1.5)
#   }
# }

#### plot to html

ylow <- 0.2
doses_clean <- as.numeric(0:60)
```

## PA dose--response curves (4 sd)

```{r plot pa, include=T}
par(mar=c(5,5,2,1),mfrow=c(3,3))
for ( j in 1:length(cause_indices)){
  cause <- as.character(DISEASE_INVENTORY$pa_acronym[cause_indices[j]])
  # get name of disease
  # decide whether to use "all" or "mortality"
  outcome_type <- ifelse(cause%in%c('lung_cancer','breast_cancer','endometrial_cancer','colon_cancer'), 'all' , 'mortality')

  dose <- doses_clean
  # apply disease-specific thresholds
  if(cause %in% c('total_cancer','coronary_heart_disease','breast_cancer','endometrial_cancer','colon_cancer')) dose[dose>35] <- 35
  else if(cause == 'lung_cancer') dose[dose>10] <- 10
  else if(cause == 'stroke') dose[dose>32] <- 32
  else if(cause == 'all_cause') dose[dose>16.08] <- 16.08
  fname <- paste(cause, outcome_type, sep = "_")
  lookup_table <- get(paste0(fname))
  lookup_df <- as.data.frame(lookup_table)
  rr <- approx(x=lookup_df$dose,y=lookup_df$RR,xout=dose,yleft=1,yright=min(lookup_df$RR))$y
  lb <- approx(x = lookup_df$dose,y = lookup_df$lb,xout = dose,yleft = 1,yright = min(lookup_df$lb))$y
  ub <-approx(x = lookup_df$dose,y = lookup_df$ub,xout = dose,yleft = 1,yright = min(lookup_df$ub))$y
  plot(doses_clean,rr,lwd=2,type='l',col='navyblue',main=cause,xlab='MMET per week',
       xlim=range(doses_clean),ylim=c(ylow,1),ylab='Response',frame=F,cex.lab=1.5,cex.axis=1.5)
  lines(doses_clean,lb,lwd=2,type='l',lty=1,col='grey')
  lines(doses_clean,ub,lwd=2,type='l',lty=1,col='grey')
  #text('Mean',x=40,y=0.8,col=rgb(0.6,0.2,0.2),cex=1.5)
  #text('Lower bound',x=35,y=0.66,col=rgb(0.6,0.2,0.2),cex=1.5)
  #text('Upper bound',x=35,y=0.93,col=rgb(0.6,0.2,0.2),cex=1.5)

}

```

```{r get ap, include=F}
nsamples <- 100
DR_AP <- read.csv(paste0(global_path,"dose_response/drap/dose_response.csv"))
parameters <- ithim_setup_parameters(NSAMPLES=nsamples,
                                     AP_DOSE_RESPONSE_QUANTILE=T)
for(i in 1:length(parameters))
  assign(names(parameters)[i],parameters[[i]],pos=1)
causes <- unique(DR_AP$cause_code)

############## plot parameter samples
dr_ap_list <- list()
for ( j in c(1:nrow(DISEASE_INVENTORY))[DISEASE_INVENTORY$air_pollution == 1]){ 
  cause <- as.character(DISEASE_INVENTORY$ap_acronym[j])
  dr_ap <- subset(DR_AP,cause_code==cause)
  ages <- unique(dr_ap$age_code)
  dr_ap_list[[cause]] <- list()
  #print(cause)
  #print(ages)
  for(age in ages){
    dr_ap_list[[cause]][[age]] <- list()
    dr_ap_age <- subset(dr_ap,age_code==age)
    alpha <- sapply(DR_AP_LIST,function(x)x[[cause]][[as.character(age)]]$alpha)
    beta <- sapply(DR_AP_LIST,function(x)x[[cause]][[as.character(age)]]$beta)
    gamma <- sapply(DR_AP_LIST,function(x)x[[cause]][[as.character(age)]]$gamma)
    tmrel <- sapply(DR_AP_LIST,function(x)x[[cause]][[as.character(age)]]$tmrel)
    dr_ap_list[[cause]][[age]]$alpha <- alpha
    dr_ap_list[[cause]][[age]]$beta <- beta
    dr_ap_list[[cause]][[age]]$gamma <- gamma
    dr_ap_list[[cause]][[age]]$tmrel <- tmrel
    #pdf(paste0('results/dose_response/AP/',cause,age,'.pdf'))
    #x11(width=8); 
    #par(mfrow=c(2,3),mar=c(4,4,1,1))
    #plot(log(dr_ap_age$alpha),log(dr_ap_age$beta),main=paste0(cause,age),col='grey'); points(log(alpha),log(beta),col='hotpink',pch=18)
    #plot(log(dr_ap_age$alpha),(log(dr_ap_age$gamma)),main=paste0(cause,age),col='grey'); points(log(alpha),log(gamma),col='hotpink',pch=18)
    #plot(log(dr_ap_age$alpha),log(dr_ap_age$tmrel),main=paste0(cause,age),col='grey'); points(log(alpha),log(tmrel),col='hotpink',pch=18)
    #plot(log(dr_ap_age$beta),log(dr_ap_age$gamma),main=paste0(cause,age),col='grey'); points(log(beta),log(gamma),col='hotpink',pch=18)
    #plot(log(dr_ap_age$beta),log(dr_ap_age$tmrel),main=paste0(cause,age),col='grey'); points(log(beta),log(tmrel),col='hotpink',pch=18)
    #plot(log(dr_ap_age$gamma),log(dr_ap_age$tmrel),main=paste0(cause,age),col='grey'); points(log(gamma),log(tmrel),col='hotpink',pch=18)
    #dev.off()
  }
}

```

## AP dose&ndash;response curves (95% intervals)

```{r plot ap}
pm <- 1:300
for ( j in 1:nrow(DISEASE_INVENTORY)) if (DISEASE_INVENTORY$air_pollution[j] == 1){ 
  cause <- as.character(DISEASE_INVENTORY$ap_acronym[j])
  dr_ap <- subset(DR_AP,cause_code==cause)
  ages <- unique(dr_ap$age_code)
  #print(cause)
  #print(ages)
  if(length(ages)>1){ 
    #pdf(paste0('results/dose_response/AP/',cause,'DR-sim.pdf'),width=6,height=8); 
    par(mfrow=c(1,1))
    layout.matrix <- matrix(1:length(ages),nrow=3,ncol=5,byrow=T)
    graphics::layout(mat=layout.matrix,heights=c(4,3,4),widths=c(3.8,3,3,3,3))
  }else{
    #pdf(paste0('results/dose_response/AP/',cause,'DR-sim.pdf'))
    par(mfrow=c(1,1))
  }
  for(age in ages){
    par(mar=c(ifelse(age>70,5,1),ifelse(age%in%c(99,25,50,75),5,1),ifelse(age%in%c(25:45,99),5,1),1))
    dr_ap_age <- subset(dr_ap,age_code==age)
    alpha <- dr_ap_list[[cause]][[age]]$alpha
    beta <- dr_ap_list[[cause]][[age]]$beta
    gamma <- dr_ap_list[[cause]][[age]]$gamma
    tmrel <- dr_ap_list[[cause]][[age]]$tmrel
    pmcurve <- matrix(0,nrow=nsamples,ncol=length(pm))
    for(i in 1:nsamples){#length(alpha)){
      pmcurve[i,] <- ap_dose_response_curve(pm,alpha[i],beta[i],gamma[i],tmrel[i])
      pmcurve[i,is.na(pmcurve[i,])] <- 1
    }
    pmcurvemean <- colMeans(pmcurve)
    pmcurveupper <- apply(pmcurve,2,quantile,0.975)
    pmcurvelower <- apply(pmcurve,2,quantile,0.025)
    plot(pm,pmcurvemean,typ='l',col='navyblue',ylim=c(1,4),main='',cex.lab=1.5,lwd=2,frame=F,xlab=ifelse(age>70,'PM dose',''),ylab=ifelse(age%in%c(99,25,50,75),'RR',''),xaxt='n',yaxt='n',cex.main=1.5)
    axis(1,labels=ifelse(age>70,T,F),cex.axis=1.5)
    axis(2,labels=ifelse(age%in%c(99,25,50,75),T,F),cex.axis=1.5)
    if(length(ages)>1) mtext(side=3,text=paste0('Age ',age),line=0)
    if(age%in%c(35,99)) mtext(side=3,text=as.character(DISEASE_INVENTORY$GBD_name)[j],line=1)
    lines(pm,pmcurveupper,col='grey',lwd=2,cex.axis=1.5)
    lines(pm,pmcurvelower,col='grey',lwd=2,cex.axis=1.5)
  }
  #dev.off()

}

```

### 2019-05-13 diagnostic/parameter_tables_script.R

@@ -74,3 +74,9 @@ for(i in 1:length(cities)){
 }
 
 setwd('..')
+
+
+
+
+## to append:
+## iconv -s -t utf-8 parameter_table_all.html dr_curves.html | pandoc -s -f html -t html -o parameter_table_all_plus_dr.html

### 2019-05-14 summary_tables_PA_AP.Rmd
newest results

@@
 whw<-as.data.frame(rowSums(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$whw))
-noov<-as.data.frame(rowSums(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$noov))
+noov<-as.data.frame(sum(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$nov))
 names(whw)[1]<- as.character(scenarios[i])
@@
 {
- summary<-round(rbind(whw,noov), digits = 1)
+ summary<-round(rbind(whw,nov=noov), digits = 1)
 } 
@@
 {
-summary<- cbind(summary, round(rbind(whw, noov), digits=1) ) 
+summary<- cbind(summary, round(rbind(whw, nov=noov), digits=1) ) 
 }

### 2019-05-15 summary_tables_PA_AP.Rmd
modified PA_AP summary

@@
 cities <- c("accra", "sao_paulo", "delhi", "bangalore")
-
+high<- c(70, 25, 150,60)
+low<- c(40,15,100,40)
 for (x in 1: length(cities))
@@
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, baseline:bus_scen, factor_key=TRUE)
-  x<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot()+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")
+  x<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")+ylim(low[x],high[x])
+
   print(x)
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
-
+limit=100
 for (x in 1: length(cities))
@@
   data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE)
-  x<-ggplot(data_long, aes(x=scenario, y=mmet, fill=scenario)) + geom_boxplot()+ ggtitle(cities[x])+ labs(y="Weekly MMET", x = "Scenarios")
+  x<-ggplot(data_long, aes(x=scenario, y=mmet, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Weekly MMET", x = "Scenarios")+  ylim(0, limit)
   print(x)

### 2019-05-15 summary_tables_PA_AP.Rmd

--- 2019-05-14 summary_tables_PA_AP.Rmd
@@
 ---
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
+
+```
 
@@
 library(dplyr)
-library(kableExtra)
 
@@
 cities <- c("accra", "sao_paulo", "delhi", "bangalore")
-
+high<- c(70, 25, 150,60)
+low<- c(40,15,100,40)
 for (x in 1: length(cities))
 {
-  print(cities[x])
+
   names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, baseline:bus_scen, factor_key=TRUE)
-  x<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot()+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")
-  print(x)
+  y<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")+ylim(low[x],high[x])
+  print(y)
 
@@
   summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1))
+  summary$change_PM<- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits=2)
   names(summary)[1]<-"Scenario"
-  print (as.character(cities[x]))
-  print(summary)
-
-  summary %>%
-  kable() %>%
-  kable_styling()
+  print(kable(summary, caption= cities[x]))
 
@@
   summary$emissions <- round(summary$emissions*100/sum(summary$emissions), digits=1)
-  print (as.character(cities[x]))
   summary<-as.data.frame(summary)
@@
   summary[nrow(summary),2]<-as.character(shares$pm_conc[x])
-  print (summary)
+  print(kable(summary, caption= cities[x]))
 
@@
   data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE)
-  x<-ggplot(data_long, aes(x=scenario, y=mmet, fill=scenario)) + geom_boxplot()+ ggtitle(cities[x])+ labs(y="Weekly MMET", x = "Scenarios")
-  print(x)
+  y<-ggplot(data_long, aes(x=scenario, y=mmet, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Weekly MMET", x = "Scenarios")+  ylim(0, limit)
+  print(y)
 
@@
   names(summary)[1]<-"Scenario"
-  print (as.character(cities[x]))
-  print(summary)
+  print(kable(summary, caption= cities[x]))
 
@@
 
-print (as.character(cities[x]))
 names(summary)<-c("base", "walk", "bike", "car", "MC", "bus")
-print(summary)
+print(kable(summary, caption= cities[x]))
 
@@
   dist_scen_prop$Scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  print (as.character(cities[x]))
-  print(dist_scen_prop)
-  print(dist_scen)
-  #library(knitr)
-  #library(kableExtra)
-  #kable(dist_scen_prop)
+  print(kable(dist_scen_prop, caption= cities[x]))
+  print(kable(dist_scen, caption= cities[x]))
+
 }

### 2019-05-16 summary_tables.Rmd
Add mode order

--- 2019-05-09 summary_tables.Rmd
@@
 library(summarytools)
-library(dplyr)
+library(tidyverse)
 })
@@
 # cities <- c('sao_paulo')
+
+# Read trip_order
+trip_order <- read_csv("data/global/trips/mode_order.csv")
 
@@
   td <- td[-1, ]
+  x <- match(row.names(td), trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
   print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
@@
 for(city in cities){
+  city <- 'bangalore'
   count_people <- nrow(io[[city]]$synth_pop)
@@
   colnames(td)[3:7] <- rownames(scen_prop)
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
   print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person")))
   cat("\n")
+
+  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
+  td <- io[[city]]$dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+
+  print(kable(td, caption = paste("Distance table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))
 
@@
   colnames(td)[3:7] <- rownames(scen_prop)
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
   print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (everyone)")))
@@
   colnames(td)[3:7] <- rownames(scen_prop)
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
   print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))

### 2019-05-16 summary_tables.Rmd
Remove fixed city

@@
 for(city in cities){
-  city <- 'bangalore'
   count_people <- nrow(io[[city]]$synth_pop)

### 2019-05-16 summary_tables_PA_AP.Rmd
emission inventory in one table

@@
 
+<!-- # Descriptive tables of emission inventory -->
+
+<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
+
+<!-- trans_share<-c(22.5, 40.0, 22.5, 28.1)  -->
+<!-- pm_conc<- c(50, 18, 122, 47 ) -->
+<!-- city<-cities -->
+<!-- shares<- as.data.frame(cbind(city,as.numeric(trans_share), pm_conc)) -->
+<!-- names(shares)[2]<- "trans_share" -->
+<!-- for (x in 1:length(cities)) -->
+<!-- { -->
+<!--   modes<-names(unlist(io[[cities[x]]]$emission_inventory)) -->
+<!--   emissions<-as.data.frame(unlist(io[[cities[x]]]$emission_inventory)) -->
+<!--   city_emissions<-cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$emission_inventory)`)) -->
+<!--   names(city_emissions)[2]<- "emissions" -->
+<!--   select<- c("car", "motorcycle", "bus_driver", "truck", "big_truck") -->
+<!--   city_emissions$modes<- as.character(city_emissions$modes) -->
+<!--   city_emissions$modes[!(city_emissions$modes %in% select)]<- "other" -->
+<!--   summary<-city_emissions %>% group_by(modes) %>% summarise(sum(emissions)) -->
+<!--   names(summary)[2]<- "emissions" -->
+<!--   summary$emissions <- round(summary$emissions*100/sum(summary$emissions), digits=1) -->
+<!--   summary<-as.data.frame(summary) -->
+<!--   summary[nrow(summary)+1,1]<-"Transport share" -->
+<!--   summary[nrow(summary),2]<-as.character(shares$trans_share[x]) -->
+<!--   summary[nrow(summary)+1,1]<-"PM2.5 Conc" -->
+<!--   summary[nrow(summary),2]<-as.character(shares$pm_conc[x]) -->
+<!--   print(kable(summary, caption= cities[x])) -->
+
+<!-- } -->
+
+<!-- ``` -->
+
 # Descriptive tables of emission inventory
@@
   summary<-city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
-  names(summary)[2]<- "emissions"
-  summary$emissions <- round(summary$emissions*100/sum(summary$emissions), digits=1)
+  names(summary)[2]<- cities[x]
+  summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits=1)
   summary<-as.data.frame(summary)
@@
   summary[nrow(summary),2]<-as.character(shares$pm_conc[x])
-  print(kable(summary, caption= cities[x]))
+  io[[cities[x]]]$summary_emission<- summary
 
@@
 
+summary_all<- cbind(io[[cities[1]]]$summary_emission,io[[cities[2]]]$summary_emission,io[[cities[3]]]$summary_emission,io[[cities[4]]]$summary_emission)
+summary_all<- summary_all[-c(3,5,7)]
+summary_all$modes[2]<- "bus"
+
+print(kable(summary_all))
 ```

### 2019-05-17 summary_tables.Rmd

@@
   td <- td[order(x),]
+
   print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
@@
   colnames(td)[3:7] <- rownames(scen_prop)
+
   x <- match(td$stage_mode, trip_order$mode)
@@
   td <- td[order(x),]
+  row.names(td) <- NULL
+
   print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person")))
@@
   td <- td[order(x),]
+  row.names(td) <- NULL
 
@@
   td <- td[order(x),]
+  row.names(td) <- NULL
+
   print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (everyone)")))
@@
   td <- td[order(x),]
+  row.names(td) <- NULL
+
   print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))

### 2019-05-20 summary_tables.Rmd
add yll per hundred thousand by age outcome

@@
   td1 <- df1 %>% filter(mmet == 0) %>% 
-    summarise(perct_zeros = round(n() / nrow(df1) * 100, 1))
+    summarise(perct_zeros = round(dplyr::n() / nrow(df1) * 100, 1))
 
@@
     group_by(scenario) %>% 
-    summarise(perct_zeros = round(n() / nrow(df2) * 100, 1))
+    summarise(perct_zeros = round(dplyr::n() / nrow(df2) * 100, 1))
 

### 2019-05-20 multi_city_voi.R

@@ -192,7 +192,7 @@ save(cities,setting_parameters,injury_reporting_rate,chronic_disease_scalar,pm_c
 
 
 parameters_only <- F
-multi_city_ithim <- outcome <- outcome_pp <- list()
+multi_city_ithim <- outcome <- outcome_pp <- yll_per_hundred_thousand <- list()
 numcores <- 16
 nsamples <- 1024
 print(system.time(
@@ -271,10 +271,22 @@ print(system.time(
       max_pop_ages <- sapply(DEMOGRAPHIC$age,function(x)as.numeric(strsplit(x,'-')[[1]][2]))
       outcome_pp[[city]] <- outcome_pp[[city]]/sum(subset(DEMOGRAPHIC,min_pop_ages>=min_age&max_pop_ages<=max_age)$population)
       colnames(outcome_pp[[city]]) <- paste0(colnames(outcome_pp[[city]]),'_',city)
-      
+      ## get yll per 100,000 by age
+      outcome_age_min <- c(15,50)
+      outcome_age_max <- c(49,69)
+      outcome_age_groups <- c('15-49','50-69')
+      yll_per_hundred_thousand[[city]] <- list()
+      for(aa in 1:length(outcome_age_groups)){
+        age <- outcome_age_groups[aa]
+        keep_rows2 <- which(min_ages>=outcome_age_min[aa]&max_ages<=outcome_age_max[aa])
+        tmp <- t(sapply(multi_city_ithim[[ci]]$outcomes, function(x) colSums(x$hb$ylls[keep_rows2,keep_cols],na.rm=T)))
+        tmp <- tmp/sum(subset(DEMOGRAPHIC,min_pop_ages>=outcome_age_min[aa]&max_pop_ages<=outcome_age_max[aa])$population)*100000
+        yll_per_hundred_thousand[[city]][[age]] <- tmp
+      }
       ## omit ac (all cause) and neoplasms (neo) and age and gender columns
       outcome[[city]] <- t(sapply(multi_city_ithim[[ci]]$outcomes, function(x) colSums(x$hb$ylls[keep_rows,keep_cols],na.rm=T)))
       colnames(outcome[[city]]) <- paste0(colnames(outcome[[city]]),'_',city)
+      
     }
     
     ## rename city-specific parameters according to city

### 2019-06-21 summary_tables_outcomes.Rmd

@@
   data <- lapply(data,function(x)rbind(x,Total=colSums(x)))
+  for(city in cities) {
+    print(kable(data[[city]], caption = paste(text, city)))
+    cat("\n")
+  }
+}
+round_and_print <- function(data,text=''){
+  data <- lapply(data, function(x)round(x,round_to))
   for(city in cities) {
@@
 ```{r scen_prop = "asis"}
-death_rates <- lapply(cities,function(x) death_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+death_rates <- lapply(cities,function(x) rbind(death_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(death_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(death_rates) <- cities
 
-sum_and_round_and_print(death_rates,"Change in deaths per 100,000 in ")
+round_and_print(death_rates,"Change in deaths per 100,000 in ")
 ```
@@
 ```{r scen_prop = "asis"}
-yll_rates <- lapply(cities,function(x) yll_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+yll_rates <- lapply(cities,function(x) rbind(yll_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(yll_rates) <- cities
-sum_and_round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
+round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
 ```
@@
 yll_rates <- lapply(cities,function(x) 
-  yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[2],'_',z[1])),rownames(yll_totals[[x]])),]/
-    t(repmat(io[[x]]$demographic$population,5,1))*100000)
+  rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[2],'_',z[1])),rownames(yll_totals[[x]])),]/
+    t(repmat(io[[x]]$demographic$population,5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(yll_rates) <- cities
-sum_and_round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
+round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
 ```
@@
 ```{r scen_prop = "asis"}
-injury_rates <- lapply(cities,function(x) injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+injury_rates <- lapply(cities,function(x) rbind(injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(injury_rates) <- cities
-sum_and_round_and_print(injury_rates,"Change in deaths due to injury per 100,000 in ")
+round_and_print(injury_rates,"Change in deaths due to injury per 100,000 in ")
 
-pa_rates <- lapply(cities,function(x) pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(pa_rates) <- cities
-sum_and_round_and_print(pa_rates,"Change in deaths due to PA per 100,000 in ")
+round_and_print(pa_rates,"Change in deaths due to PA per 100,000 in ")
 
-ap_rates <- lapply(cities,function(x) ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(ap_rates) <- cities
-sum_and_round_and_print(ap_rates,"Change in deaths due to AP per 100,000 in ")
+round_and_print(ap_rates,"Change in deaths due to AP per 100,000 in ")
 ```
@@
 ```{r scen_prop = "asis"}
-injury_rates <- lapply(cities,function(x) injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+injury_rates <- lapply(cities,function(x) rbind(injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(injury_rates) <- cities
-sum_and_round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ")
+round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ")
 
-pa_rates <- lapply(cities,function(x) pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(pa_rates) <- cities
-sum_and_round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ")
+round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ")
 
-ap_rates <- lapply(cities,function(x) ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000)
+ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(ap_rates) <- cities
-sum_and_round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ")
+round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ")
 ```
@@
 ```{r scen_prop = "asis"}
-disease_rates <- lapply(cities,function(x) disease_totals[[x]]/sum(pop_by_age[[x]])*100000)
+disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(disease_rates) <- cities
-sum_and_round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ")
+round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ")
 ```
@@
 ```{r scen_prop = "asis"}
-disease_rates <- lapply(cities,function(x) disease_totals[[x]]/sum(pop_by_age[[x]])*100000)
+disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
 names(disease_rates) <- cities
-sum_and_round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ")
+round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ")
 ```

### 2019-08-05 code/trips_analysis/trip_distance.Rmd
Markdown document for trip distance analysis

---
title: "Distance distribution by modes"
output: html_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
options(warn = -1)

```

```{r loadLibraries, echo = F, message = F}
suppressWarnings({

  library(tidyverse)
  library(ggthemes)
  library(plotly)
  library(knitr)

})

```

```{r loadData, message = F, echo = F, warning = F}

  io <- read_rds("../../results/multi_city/io.rds")
```

```{r genPlots, message = T, echo = F, warning = F}

suppressWarnings({
  require(knitr)

  l <- htmltools::tagList()
  tl <- list()
  index <- 1

  for (cities in c('accra','sao_paulo','delhi')){

  rd <- io[[cities]]$trip_scen_sets %>% filter(scenario == "Baseline")

  rd <- rd %>% filter(trip_mode %in% c("bus", "walking", "car", "motorcycle", "bicycle"))

  # ggplot(rd) +
  #   aes(x = trip_distance) +
  #   geom_density(adjust = 1L) +
  #   scale_fill_hue() +
  #   theme_tufte() +
  #   facet_wrap(vars(trip_mode), scales = "free") +
  #   labs(title = "Distribution of trips by distance - Accra")

  p <- ggplot(rd) +
  aes(x = trip_mode, y = trip_distance) +
  geom_boxplot(adjust = 1L, scale = "area", fill = "#39486b") +
  geom_hline(yintercept = c(2, 6), linetype="dashed", color = "red", size=0.1) +
  labs(title = paste("Distribution of trips by distance - ", cities), x = "Mode", y = "Distance (km)")

  if (cities == 'delhi'){
    p <- p + coord_cartesian(ylim = c(0, 100)) 
  }

  tl[[index]] <-  filter(p$data, scenario == 'Baseline') %>% select(trip_mode, trip_distance) %>% rename(var = trip_distance) %>% group_by(trip_mode) %>%  summarise(mean = mean(var), median =median(var), Q1 = quantile (var, probs=0.25), Q2 = quantile (var, probs=0.50), Q3 = quantile(var, probs=0.75)) %>% mutate_if(is.numeric, round, digits = 1)

  pp <- plotly_build(p)

  pp$x$data <- lapply(pp$x$data, FUN = function(x){
    x$marker = list(opacity = 0)
    return(x)
  })

  l[[index]] <- as_widget(ggplotly(p))
  index <- index + 1

  }

})

```

## Plots

```{r printPlots, echo = F, message = F}

suppressWarnings({
 l
})
```

## Tables

```{r printTables, echo = F, results = 'asis'}

suppressWarnings({
  require(knitr)
  local_cities <- c('Accra','Sao Paulo','Delhi')
   for (i in 1:3){
      print(kable(tl[[i]], caption = paste("Distribution of trips by distance - ", local_cities[i])))
      cat('\n')
   }
})
```

### 2019-08-07 ithim_r_function_descriptions.Rmd
Update ithim_r_function_descriptions.Rmd
A test push for trying pull request on a single file
documentation

@@
 ---
-title: "ITHIM Functions"
+title: "ITHIM Functions (BETA)"
 author: "Daniel Fuller"
@@
 - AP: Air Pollution
-- PM: Particulate Matter? 
-- GDB: Global Burden of Disease
-- HC: ??
+- PM: Particulate Matter (PM2.5)
+- GBD: Global Burden of Disease
+- RR: Relative Risk
+- DR: Dose Response
+- PT: Public Transport
+- mMET: marginal Metabolic Equivalent Task
+- YLL: Years of Life Lost
 

### 2019-08-07 R/run_ithim_setup.R

@@ -2,6 +2,51 @@
 #' 
 #' Sets up the basic ITHIM object for onward calculation. Data loading, processing and harmonisation. Setting global values.
 #' 
+#' Parameters have two options: to be set to a constant, and to be sampled from a prespecified distribution.
+#' Each parameter is given as an argument of length 1 or 2. 
+#' If length 1, it's constant, and set to the global environment. 
+#' If length 2, a distribution is defined and sampled from NSAMPLE times.
+#' There are some exceptions, listed above
+#' 
+#' @param seed
+#' @param CITY name of the city, and name of the directory containing city data files
+#' @param speeds named list of mode speeds
+#' @param emission_inventory named list of mode emissions
+#' @param setup_call_summary_filename name to write setup call summary to
+#' @param DIST_CAT vector string of distance categories in the form '0-6'. (The unit is assumed to be the same as in the trip set.)
+#' @param AGE_RANGE vector of minimum and maximum ages to include
+#' @param ADD_WALK_TO_BUS_TRIPS logic: whether or not to add short walks to all PT trips
+#' @param ADD_BUS_DRIVERS logic: whether or not to add bus drivers
+#' @param ADD_TRUCK_DRIVERS logic: whether or not to add truck drivers
+#' @param TEST_WALK_SCENARIO logic: whether or not to run the walk scenario
+#' @param TEST_CYCLE_SCENARIO logic: whether or not to run the cycle scenario
+#' @param MAX_MODE_SHARE_SCENARIO logic: whether or not to run the max mode share scenario
+#' @param REFERENCE_SCENARIO which scenario forms the reference for the health comparison
+#' @param PATH_TO_LOCAL_DATA path to CITY directory, if not using package
+#' @param NSAMPLES constant integer: number of samples to take
+#' @param BUS_WALK_TIME lognormal parameter: duration of walk to PT
+#' @param MMET_CYCLING lognormal parameter: mMETs when cycling
+#' @param MMET_WALKING lognormal parameter: mMETs when walking
+#' @param PM_CONC_BASE lognormal parameter: background PM2.5 concentration
+#' @param PM_TRANS_SHARE beta parameter: fraction of background PM2.5 attributable to transport
+#' @param PA_DOSE_RESPONSE_QUANTILE logic: whether or not to sample from PA RR DR functions
[…]

### 2019-08-07 code/trips_analysis/trip_distance.Rmd

@@
   io <- read_rds("../../results/multi_city/io.rds")
+```
+
+## Trip distance - histogram
+
+```{r genTotalHistPlots, message = T, echo = F, warning = F}
+
+rd1 <- io$accra$trip_scen_sets %>% mutate(cs = 'Accra') %>% select(cs, trip_distance, trip_distance_cat)
+rd2 <- io$sao_paulo$trip_scen_sets %>% mutate(cs = 'Sao Paulo') %>% select(cs, trip_distance, trip_distance_cat)
+rd3 <- io$delhi$trip_scen_sets %>% mutate(cs = 'Delhi') %>% select(cs, trip_distance, trip_distance_cat)
+
+rd <- plyr::rbind.fill(rd1, rd2, rd3)
+
+op <- ggplot(rd %>% filter(trip_distance < 51)) +
+  aes(x = trip_distance, fill = cs) +
+  geom_histogram(breaks = c(0:10,seq(11, 51, by = 5)), alpha=0.8, position="identity", aes(y = ..density..)) +
+  labs(x = '') + 
+  #geom_density(alpha = 0.4, color = 'lightgrey') +
+  facet_wrap(vars(cs))
+
+ggplotly(op)
+
+```
+
+## Trip distance - density
+
+```{r genTotalDensityPlots, message = T, echo = F, warning = F}
+
+op <- ggplot(rd %>% filter(trip_distance < 50)) +
+  aes(x = trip_distance, fill = cs) +
+  labs(x = '') + 
+  #geom_histogram(breaks = c(0:10,seq(11, 61, by = 5)), alpha=0.8, position="identity", aes(y = ..density..)) +
+  geom_density(alpha = 0.8) +
+  facet_wrap(vars(cs))
+
+ggplotly(op)
+
 ```

### 2019-08-07 code/trips_analysis/trip_distance.Rmd
select -> dplyr::select

@@
 
-rd1 <- io$accra$trip_scen_sets %>% mutate(cs = 'Accra') %>% select(cs, trip_distance, trip_distance_cat)
-rd2 <- io$sao_paulo$trip_scen_sets %>% mutate(cs = 'Sao Paulo') %>% select(cs, trip_distance, trip_distance_cat)
-rd3 <- io$delhi$trip_scen_sets %>% mutate(cs = 'Delhi') %>% select(cs, trip_distance, trip_distance_cat)
+rd1 <- io$accra$trip_scen_sets %>% mutate(cs = 'Accra') %>% dplyr::select(cs, trip_distance, trip_distance_cat)
+rd2 <- io$sao_paulo$trip_scen_sets %>% mutate(cs = 'Sao Paulo') %>% dplyr::select(cs, trip_distance, trip_distance_cat)
+rd3 <- io$delhi$trip_scen_sets %>% mutate(cs = 'Delhi') %>% dplyr::select(cs, trip_distance, trip_distance_cat)
 
@@
 
-  tl[[index]] <-  filter(p$data, scenario == 'Baseline') %>% select(trip_mode, trip_distance) %>% rename(var = trip_distance) %>% group_by(trip_mode) %>%  summarise(mean = mean(var), median =median(var), Q1 = quantile (var, probs=0.25), Q2 = quantile (var, probs=0.50), Q3 = quantile(var, probs=0.75)) %>% mutate_if(is.numeric, round, digits = 1)
+  tl[[index]] <-  filter(p$data, scenario == 'Baseline') %>% dplyr::select(trip_mode, trip_distance) %>% rename(var = trip_distance) %>% group_by(trip_mode) %>%  summarise(mean = mean(var), median =median(var), Q1 = quantile (var, probs=0.25), Q2 = quantile (var, probs=0.50), Q3 = quantile(var, probs=0.75)) %>% mutate_if(is.numeric, round, digits = 1)
 

### 2019-08-12 code/trips_analysis/trip_distance.Rmd
Use janitor to create grouped summarise

@@
   library(tidyverse)
+  library(janitor)
   library(ggthemes)
@@
 
-```{r genTotalHistPlots, message = T, echo = F, warning = F}
+```{r genTotalHistPlots, message = F, echo = F, warning = F}
 
@@
 
-op <- ggplot(rd %>% filter(trip_distance < 51)) +
-  aes(x = trip_distance, fill = cs) +
-  geom_histogram(breaks = c(0:10,seq(11, 51, by = 5)), alpha=0.8, position="identity", aes(y = ..density..)) +
-  labs(x = '') + 
-  #geom_density(alpha = 0.4, color = 'lightgrey') +
-  facet_wrap(vars(cs))
+rd <- rd %>% filter(trip_distance < 51) %>% mutate(category = cut(trip_distance, breaks = c(0:10,seq(11, 51, by = 5))))
 
+# td <- rd %>% filter(!is.na(category)) %>% group_by(category, cs) %>% summarise(percentage = round(n() / nrow(.) * 100, 1))
+
+td <- rd %>% filter(!is.na(category)) %>% tabyl(cs, category) %>% adorn_percentages("row")
+td[, 2:ncol(td)] <- round(100 * td[, 2:ncol(td)], 1)
+td <- reshape2::melt(td, by = cs)
+
+op <- ggplot(td) +
+     aes(x = variable, fill = cs, weight = value) +
+     geom_bar() +
+     scale_fill_hue() +
+     theme_minimal() +
+     facet_wrap(vars(cs)) + 
+     labs(x = '', y = 'percentage (%)') +
+     coord_flip()
+
+# op <- ggplot(rd) +
+#   aes(x = category, fill = cs) +
+#   geom_bar() +
+#   labs(x = '') + 
+#   #geom_density(alpha = 0.4, color = 'lightgrey') +
+#   facet_wrap(vars(cs), scales="free")  
 ggplotly(op)
@@
 
-## Trip distance - density
+<!-- ## Trip distance - density -->
 
-```{r genTotalDensityPlots, message = T, echo = F, warning = F}
+<!-- ```{r genTotalDensityPlots, message = T, echo = F, warning = F} -->
 
-op <- ggplot(rd %>% filter(trip_distance < 50)) +
-  aes(x = trip_distance, fill = cs) +
-  labs(x = '') + 
-  #geom_histogram(breaks = c(0:10,seq(11, 61, by = 5)), alpha=0.8, position="identity", aes(y = ..density..)) +
-  geom_density(alpha = 0.8) +
-  facet_wrap(vars(cs))
+<!-- op <- ggplot(rd %>% filter(trip_distance < 50)) + -->
+<!--   aes(x = trip_distance, fill = cs) + -->
+<!--   labs(x = '') +  -->
+<!--   #geom_histogram(breaks = c(0:10,seq(11, 61, by = 5)), alpha=0.8, position="identity", aes(y = ..density..)) + -->
+<!--   geom_density(alpha = 0.8) + -->
+<!--   facet_wrap(vars(cs)) -->
 
-ggplotly(op)
+<!-- ggplotly(op) -->
 
-```
+<!-- ``` -->
 
-```{r genPlots, message = T, echo = F, warning = F}
+```{r genPlots, message = F, echo = F, warning = F}
 

### 2019-10-04 code/producing_trips_rd/trip-summary.Rmd
addin rmd for trip reults

---
title: "Summary Tables ITHIMR"
author: "Lambed Tatah, MRC Epidemiology Unit, University of Cambridge, UK"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
---

```{r, setup, comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F, include=FALSE, results='asis'}

source("J:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/buenos_aires/buenos_aires_trip.csv")
quality_check(trip)
trip_summary$buenos_aires <- value

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/sao_paulo/sao_paulo_trip.csv")
quality_check(trip)
trip_summary$sao_paulo <- value

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/belo_horizonte/belo_horizonte_trip.csv")
quality_check(trip)
trip_summary$belo_horizonte <- value

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/santiago/santiago_trip.csv")
quality_check(trip)
trip_summary$santiago <- value

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/bogota/bogota_trip.csv")
quality_check(trip)
trip_summary$bogota <- value

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/accra/accra_trip.csv")
quality_check(trip)
trip_summary$accra <- value

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/delhi/delhi_trip.csv")
quality_check(trip)
trip_summary$delhi <- value

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/bangalore/bangalore_trip.csv")
quality_check(trip)
trip_summary$bangalore <- value

trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/mexico/mexico_city_trip.csv")
quality_check(trip)
trip_summary$mexico_city <- value

print(kable_styling(kable(trip_summary)))

```

### 2019-10-04 code/producing_trips_rd/travel_survey.R

@@ -9,8 +9,6 @@
 rm(list =ls())
 
 source("J:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
-package()
-
 
 setwd('J:/Studies/MOVED/HealthImpact/Data/TIGTHAT/Argentina/WP1-TS/Buenos Aires/')
 
@@ -75,7 +73,13 @@ trip <- person %>%
            trip_mode, trip_duration, trip_distance, stage_id,
            stage_mode, stage_duration, stage_distance)
 
-quality_check(trip)
+
+trip$year <- "2012"
+
+buenos_aires_trip <- trip
+#quality_check(Buenos_Aires)
+write.csv(buenos_aires_trip, "J:/Group/lambed/ITHIM-R/data/local/buenas_aires/buenos_aires_trip.csv")
+
 #write.csv(trip, "trips_buenas_aires.csv")
 
 #####Argentina Cordoba###########
@@ -948,7 +952,6 @@ rm("person","trips")
 rm(list =ls())
 
 source("J:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
-package()
 
 
 setwd('J:/Studies/MOVED/HealthImpact/Data/TIGTHAT/Brazil/Sao Paulo/Pesquisa Origem Destino 2012')
@@ -993,15 +996,16 @@ trip <- setdiff(trip_0, no_trip) %>%
            trip_duration = DURACAO) %>% 
     select(cluster_id, household_id, participant_id, participant_wt, age, sex, trip_id, trip_purpose, trip_mode, trip_duration,trip_distance, stage_id, stage_mode)
 
-quality_check(trip)
+trip$year <- "2012"
 
+#quality_check(trip)
+write.csv(trip, "J:/Group/lambed/ITHIM-R/data/local/sao_paulo/sao_paulo.csv")
 
 #####Brazil Belo Horizonte######
 
 rm(list =ls())
 
 source("J:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
-package()
 
 setwd('J:/Studies/MOVED/HealthImpact/Data/TIGTHAT/Brazil/Belo Horizonte/Travel survey')
 
@@ -1041,8 +1045,11 @@ trip <- person %>%
     left_join(trip) %>%
     rename(age= IDADE, trip_id = Viagem)
 trip[129,9] <- "train"
+#quality_check(trip)
 
-quality_check(trip)
+trip$year <- "2012"
+
[…]

### 2019-10-04 code/producing_trips_rd/used_functions.R

@@ -1,38 +1,109 @@
-package <- function(){
-    #required packages
-    library(tidyverse)
-    library(readxl)
-    library(haven)
-    library(nnet)
-    library(foreign)
-    library(gsubfn)
+## @knitr package
+
+#required packages
+library(tidyverse)
+library(readxl)
+library(haven)
+library(nnet)
+library(foreign)
+library(gsubfn)
+library(knitr)
+library(kableExtra)
+
     
-    #mode speeds --> pls add other modes and their average mode speeds.
-    mode_speed <<- data.frame(mode = c("bicycle","bus","car","metro", "motorcycle",
-                                             "other", "rickshaw", "taxi", "train", "walk" ),
-                              mode_speed = c(15, 15, 25, 25, 25,21 ,25,25, 30, 5 ))
-    
-    #function for evaluating data quality
-    quality_check <<- function(trip){
+#mode speeds --> pls add other modes and their average mode speeds.
+mode_speed <- data.frame(mode = c("bicycle","bus","car","metro", "motorcycle",
+                                "other", "rickshaw", "taxi","train","truck", "van","walk" ),
+                              mode_speed = c(15, 15, 25, 25, 25,21 ,25,25, 30,25,25, 5 ))
+
+
+## @knitr trip_summary
+trip_summary <- data.frame(row.names = paste(seq(1:49),
+                                             c("Year of survey",
+                                               "Number of households",
+                                               "Number of individuals",
+                                               "Adults (% > 17 years)",
+                                               "Household size",
+                                               "Adults per household",
+                                               "Male to female ratio",
+                                               "People with trips (%)",
+                                               "Trip distribution by sex (%)",
+                                               "Trip rates by sex (%)",
+                                               "People with 1 trip (%)",
+                                               "People with 2 trips (%)",
+                                               "People with 3 trips (%)",
+                                               "People with 4 trips (%)",
+                                               "Trip per capita (overall)",
+                                               "Trip per capita (adults)",
+                                               "Trip duration (mins)",
+                                               "travel time per person",
+                                               "Mean trip duration",
+                                               "Bicycle",
+                                               "Bus",
+                                               "Car",
+                                               "Metro",
[…]

### 2019-10-04 code/producing_trips_rd/trip-summary.Rmd
adding trip summary table

@@
 ---
-title: "Summary Tables ITHIMR"
+title: "Trip Summary ITHIM-R"
 author: "Lambed Tatah, MRC Epidemiology Unit, University of Cambridge, UK"
@@
 output:
-  html_document: default
+    html_document:
+        theme: null
 ---
 
-```{r, setup, comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F, include=FALSE, results='asis'}
+```{r, include = FALSE}
+source("J:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
 
-source("J:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
+```
+
+```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
 
@@
 quality_check(trip)
-trip_summary$santiago <- value
+trip_summary$santiago_city <- value
 
@@
 quality_check(trip)
-trip_summary$bogota <- value
+trip_summary$bogota_city <- value
 
@@
 quality_check(trip)
-trip_summary$accra <- value
+trip_summary$accra_city <- value
 
@@
 quality_check(trip)
-trip_summary$delhi <- value
+trip_summary$delhi_city <- value
 
@@
 quality_check(trip)
-trip_summary$bangalore <- value
+trip_summary$bangalore_city <- value
 
@@
 
-print(kable_styling(kable(trip_summary)))
+```
+
+```{r, include = TRUE,  echo= FALSE, results='asis'}
+kable(trip_summary) %>%
+kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14, fixed_thead = T)
 

### 2019-10-04 code/producing_trips_rd/used_functions.R

@@ -1,6 +1,5 @@
 ## @knitr package
 
-#required packages
 library(tidyverse)
 library(readxl)
 library(haven)
@@ -11,63 +10,68 @@ library(knitr)
 library(kableExtra)
 
     
-#mode speeds --> pls add other modes and their average mode speeds.
+## @knitr mode_speed
+#pls add other modes and their average mode speeds.
 mode_speed <- data.frame(mode = c("bicycle","bus","car","metro", "motorcycle",
                                 "other", "rickshaw", "taxi","train","truck", "van","walk" ),
                               mode_speed = c(15, 15, 25, 25, 25,21 ,25,25, 30,25,25, 5 ))
 
 
 ## @knitr trip_summary
-trip_summary <- data.frame(row.names = paste(seq(1:49),
-                                             c("Year of survey",
+trip_summary <- data.frame(row.names =       c("Year of survey",
+                                               "1",
                                                "Number of households",
                                                "Number of individuals",
                                                "Adults (% > 17 years)",
                                                "Household size",
                                                "Adults per household",
-                                               "Male to female ratio",
+                                               "Male to female ratio (M:F)",
+                                               "2",
                                                "People with trips (%)",
-                                               "Trip distribution by sex (%)",
-                                               "Trip rates by sex (%)",
+                                               "Prop. of sex with trips (M:F,%)",
+                                               "Trip Distribution by sex (M:F,%)",
                                                "People with 1 trip (%)",
                                                "People with 2 trips (%)",
                                                "People with 3 trips (%)",
                                                "People with 4 trips (%)",
                                                "Trip per capita (overall)",
                                                "Trip per capita (adults)",
+                                               "3",
                                                "Trip duration (mins)",
                                                "travel time per person",
                                                "Mean trip duration",
-                                               "Bicycle",
-                                               "Bus",
-                                               "Car",
-                                               "Metro",
-                                               "Motocycle",
-                                               "Other",
-                                               "Rickshaw",
-                                               "Taxi",
-                                               "Train",
-                                               "Truck",
-                                               "Van",
-                                               "Walk",
+                                               "Bicycle_duration",
+                                               "Bus_duration",
+                                               "Car_duration",
+                                               "Metro_duration",
+                                               "Motocycle_duration",
+                                               "Other_duration",
+                                               "Rickshaw_duration",
+                                               "Taxi_duration",
[…]

### 2019-10-05 code/producing_trips_rd/trip-summary.Rmd
modify trip summary tables

@@
 title: "Trip Summary ITHIM-R"
-author: "Lambed Tatah, MRC Epidemiology Unit, University of Cambridge, UK"
+author: ""
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:
-    html_document:
-        theme: null
+output: html_document
 ---
@@
 ```{r, include = FALSE}
-source("J:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
+source("V:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
+knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
+```
+
+```{r loadLibraries, echo = F, message = F}
+
+st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
+           plain.ascii       = FALSE,       # One of the essential settings
+           style             = "rmarkdown", # Idem.
+           dfSummary.silent  = TRUE,        # Suppresses messages about temporary files
+           footnote          = NA,          # Keeping the results minimalistic
+           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
+                                            # much better results. Your mileage may vary
 
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/buenos_aires/buenos_aires_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/buenos_aires/buenos_aires_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/sao_paulo/sao_paulo_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/sao_paulo/sao_paulo_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/belo_horizonte/belo_horizonte_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/belo_horizonte/belo_horizonte_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/santiago/santiago_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/santiago/santiago_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/bogota/bogota_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/bogota/bogota_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/accra/accra_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/accra/accra_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/delhi/delhi_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/delhi/delhi_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/bangalore/bangalore_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/bangalore/bangalore_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("J:/Group/lambed/ITHIM-R/data/local/mexico/mexico_city_trip.csv")
+trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/mexico/mexico_city_trip.csv")
 quality_check(trip)

### 2019-10-05 code/producing_trips_rd/used_functions.R

@@ -8,6 +8,7 @@ library(foreign)
 library(gsubfn)
 library(knitr)
 library(kableExtra)
+library(summarytools)
 
     
 ## @knitr mode_speed
@@ -19,14 +20,14 @@ mode_speed <- data.frame(mode = c("bicycle","bus","car","metro", "motorcycle",
 
 ## @knitr trip_summary
 trip_summary <- data.frame(row.names =       c("Year of survey",
-                                               "1",
+                                               "",
                                                "Number of households",
                                                "Number of individuals",
                                                "Adults (% > 17 years)",
                                                "Household size",
                                                "Adults per household",
                                                "Male to female ratio (M:F)",
-                                               "2",
+                                               "Trip Frequency",
                                                "People with trips (%)",
                                                "Prop. of sex with trips (M:F,%)",
                                                "Trip Distribution by sex (M:F,%)",
@@ -36,7 +37,6 @@ trip_summary <- data.frame(row.names =       c("Year of survey",
                                                "People with 4 trips (%)",
                                                "Trip per capita (overall)",
                                                "Trip per capita (adults)",
-                                               "3",
                                                "Trip duration (mins)",
                                                "travel time per person",
                                                "Mean trip duration",
@@ -52,7 +52,6 @@ trip_summary <- data.frame(row.names =       c("Year of survey",
                                                "Truck_duration",
                                                "Van_duration",
                                                "Walk_duration",
-                                               "4",
                                                "Trip mode Shares (%)",
                                                "Bicycle_share",
                                                "Bus_share",
@@ -66,7 +65,6 @@ trip_summary <- data.frame(row.names =       c("Year of survey",
                                                "Truck_share",
                                                "Van_share",
                                                "Walk_share",
-                                               "5",
                                                "Trip Purpose (%)",
                                                "Work related",
                                                "School related",
@@ -275,7 +273,6 @@ quality_check <- function(trip){
                    round(trip_per_capita,1),
                    round(trip_per_capita_adult,1),
                    "",
-                   "",
                    round(travel_time_per_person),
                    round(average_trip_duration_trip$x),
                    avg_mode_time$avg_mode_time[1],
@@ -291,7 +288,6 @@ quality_check <- function(trip){
                    avg_mode_time$avg_mode_time[11],
                    avg_mode_time$avg_mode_time[12],
                    "",
-                   "",
                    mode_share$mode_share[1],
[…]

### 2019-10-11 code/producing_trips_rd/trip-summary.Rmd

@@
 ```{r, include = FALSE}
-source("V:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
+source("used_functions.R")
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
@@
 ```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
-
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/buenos_aires/buenos_aires_trip.csv")
+trip <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/sao_paulo/sao_paulo_trip.csv")
+trip <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/belo_horizonte/belo_horizonte_trip.csv")
+trip <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/santiago/santiago_trip.csv")
+trip <- read_csv("../../data/local/santiago/santiago_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/bogota/bogota_trip.csv")
+trip <- read_csv("../../data/local/bogota/bogota_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/accra/accra_trip.csv")
+trip <- read_csv("../../data/local/accra/accra_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/delhi/delhi_trip.csv")
+trip <- read_csv("../../data/local/delhi/delhi_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/bangalore/bangalore_trip.csv")
+trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("V:/Group/lambed/ITHIM-R/data/local/mexico/mexico_city_trip.csv")
+trip <- read_csv("../../data/local/mexico/mexico_city_trip.csv")
 quality_check(trip)

### 2019-11-04 project_mgmt/data_source_summary.Rmd
summarise data sources

---
title: "Data source summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a summary of data sources as saved in project_mgmt/case_cities_data.xlsx

```{r start, echo=FALSE}
library(xlsx)
#setwd('project_mgmt/')
x <- xlsx::read.xlsx2('case_cities_data.xlsx',sheetIndex=1,colClasses = 'character',stringsAsFactors=FALSE)

start_index <- which(x[,1]=='Country') + 1
stop_index <- which(x[,1]=='Extras') - 1

countries <- x[start_index:stop_index,1]
cities <- x[start_index:stop_index,2]

```

## Travel

```{r travel, echo=FALSE}

travel_section <- which(colnames(x)=='Background.travel.data')
travel_year <- which(x[1,travel_section:ncol(x)]=='Year of data collection')[1] + travel_section - 1
travel_source <- which(x[1,travel_section:ncol(x)]=='Source of data')[1] + travel_section - 1

for(i in start_index:stop_index){
  cat('Travel data for ')
  cat(paste0(cities[i-start_index+1],' (',countries[i-start_index+1],') '))
  cat('for year ')
  cat(x[i,travel_year])
  cat(' were obtained from ')
  cat(x[i,travel_source])
  cat('.\n')
}
```

### 2019-12-10 summary_tables.Rmd
Add 9 cities, and then respective dist_cat tables

@@
 # Assumes that multi_city_script.R has been run till 
-cities <- c('accra','sao_paulo','delhi','bangalore')
+cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota')
+
 # cities <- c('sao_paulo')
@@
 
+|santiago|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 92.4|29.7|0.3 |
+| bicycle | 1.5|10.5|2.5 |
+| car | 5.4|42.3|27.4 |
+| motorcycle | 0.0|0.6|0.7 |
+| bus | 0.0|3.1|36.0 |
+
+|belo_horizonte|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 91.7|36.0|0.6 |
+| bicycle | 0.4|3.6|1.0 |
+| car | 7.2|39.9|38.9 |
+| motorcycle | 0.2|6.6|7.3 |
+| bus | 0.0|0.9|39.4 |
+
+|buenos_aires|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 95.9|32.9|0.2 |
+| bicycle | 3.0|10.1|1.5 |
+| car | 0.5|24.1|25.4 |
+| motorcycle | 0.0|2.2|1.7 |
+| bus | 0.6|27.4|49.9 |
+
+|mexico_city|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 95.1|38.1|0.3 |
+| bicycle | 0.8|7.7|1.3 |
+| car | 2.3|23.9|28.7 |
+| motorcycle | 0.0|3.3|1.9 |
+| bus | 0.3|19.0|48.3 |
+
+|bogota|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 94.9|54.5|4.2 |
+| bicycle | 2.1|14.3|6.2 |
+| car | 1.1|10.3|18.9 |
+| motorcycle | 0.0|0.0|0.1 |
+| bus | 1.4|17.6|65.7 |
+
 ## Trip proportion
@@
 for(city in cities){
+  print(city)
   df <- io[[city]]$trip_scen_sets

### 2019-12-18 summary_tables.Rmd

@@
 
-```{r load_objects = "asis"}
+```{r load_objects = "asis", echo = F, message = F}
 io <- readRDS("results/multi_city/io.rds")

### 2019-12-18 summary_tables.Rmd
Add capetown and vizag

--- 2019-12-10 summary_tables.Rmd
@@
 
-```{r load_objects = "asis"}
+```{r load_objects = "asis", echo = F, message = F}
 io <- readRDS("results/multi_city/io.rds")
 # Assumes that multi_city_script.R has been run till 
-cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota')
+cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota', 'cape_town', 'vizag')
 
@@
 | bus | 1.4|17.6|65.7 |
+
+|cape_town|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 70.8|23.9|2.3 |
+| bicycle | 0.0|0.5|0.4 |
+| car | 7.5|47.2|54.6 |
+| motorcycle | 0.0|0.3|0.7 |
+| bus | 1.6|28.1|42.1 |
+
+|vizag|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| walking | 81.4|19.6|0.1 |
+| bicycle | 3.9|7.6|1.9 |
+| car | 0.6|2.5|2.2 |
+| motorcycle | 9.6|25.4|22.9 |
+| bus | 0.1|21.7|56.1 |
 

### 2019-12-19 summary_tables.Rmd
Regen with new props - fixed for capetown

@@
 |---|---|---|---|
-| walking | 83.5|28.8|1.8 |
-| bicycle | 2.0|5.8|4.7 |
-| car | 2.0|10.7|16.8 |
-| motorcycle | 5.6|23.3|20.7 |
-| bus | 1.2|18.2|39.1 |
+| walking | 84.3|11.1|0.8 |
+| bicycle | 1.9|7.9|3.6 |
+| car | 1.9|14.7|19.6 |
+| motorcycle | 5.6|31.9|25.3 |
+| bus | 0.7|19.7|35.5 |
 
@@
 |---|---|---|---|
-| walking | 92.4|29.7|0.3 |
+| walking | 92.5|29.7|0.3 |
 | bicycle | 1.5|10.5|2.5 |
@@
 |---|---|---|---|
-| walking | 95.1|38.1|0.3 |
+| walking | 95.4|38.1|0.3 |
 | bicycle | 0.8|7.7|1.3 |
-| car | 2.3|23.9|28.7 |
+| car | 2.3|23.9|28.6 |
 | motorcycle | 0.0|3.3|1.9 |
-| bus | 0.3|19.0|48.3 |
+| bus | 0.3|19.0|48.2 |
 
@@
 |---|---|---|---|
-| walking | 94.9|54.5|4.2 |
-| bicycle | 2.1|14.3|6.2 |
-| car | 1.1|10.3|18.9 |
-| motorcycle | 0.0|0.0|0.1 |
-| bus | 1.4|17.6|65.7 |
+| walking | 90.0|27.3|2.4 |
+| bicycle | 4.5|15.1|4.2 |
+| car | 1.4|13.3|18.7 |
+| motorcycle | 0.4|5.6|7.2 |
+| bus | 3.2|34.1|62.9 |
 
@@
 |---|---|---|---|
-| walking | 70.8|23.9|2.3 |
-| bicycle | 0.0|0.5|0.4 |
-| car | 7.5|47.2|54.6 |
-| motorcycle | 0.0|0.3|0.7 |
-| bus | 1.6|28.1|42.1 |
+| walking | 88.5|23.3|2.0 |
+| bicycle | 0.0|0.4|0.3 |
+| car | 9.4|46.1|47.3 |
+| motorcycle | 0.0|0.3|0.6 |
+| bus | 2.0|27.5|36.5 |
 

### 2020-01-09 summary_tables.Rmd
Add prop by mode and distance

@@
 
+## Distance by mode by distance category
+
+```{r trip_dist_mode_figs = "asis"}
+
+for (city in cities){
+
+  df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline' & trip_mode %in% c('walking', 'bicycle', 'car', 'motorcycle', 'bus')) %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_mode, trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n()) %>%
+    group_by(trip_mode) %>% mutate(n_vals_1 = sum(n_vals)) %>% 
+    mutate(prop = round(n_vals/n_vals_1 * 100, 1))
+  print(kable(df, caption = paste("Distance by mode by for  ", city)))
+}
+
+```
+
+# Figures
+
+```{r trip_dist_mode_figs = "asis"}
+
+for (city in cities){
+
+  df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline' & trip_mode %in% c('walking', 'bicycle', 'car', 'motorcycle', 'bus')) %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_mode, trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n()) %>%
+    group_by(trip_mode) %>% mutate(n_vals_1 = sum(n_vals)) %>% 
+    mutate(prop = round(n_vals/n_vals_1 * 100, 1))
+
+  print(ggplot(data=df, aes(x=trip_distance_cat, y=prop)) +
+          geom_bar(stat="identity", fill="steelblue")+
+          facet_wrap(vars(trip_mode)) +
+          geom_text(aes(label=prop), vjust = -0.3, size = 3.5)+
+          theme_minimal() +
+          theme(panel.spacing = unit(2, "lines")) +
+          labs(title = paste(city, ' by dist by mode')))
+}
+
+```
+
 ## Duration tables

### 2020-01-09 summary_tables.Rmd
Distance prop by categories

--- 2019-12-19 summary_tables.Rmd
@@
 
+## Distance by distance category
+### Tables
+
+```{r trip_dist_mode_figs = "asis"}
+
+for (city in cities){
+
+  df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n(), prop = round(n_vals / nrow(.) * 100, 1))
+  print(kable(df, caption = paste("Distance by mode by for  ", city)))
+}
+
+```
+
+<!-- ### Figures -->
+
+<!-- ```{r trip_dist_mode_figs = "asis"} -->
+
+<!-- for (city in cities){ -->
+
+<!--   df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline' & trip_mode %in% c('walking', 'bicycle', 'car', 'motorcycle', 'bus')) %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_mode, trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n()) %>%  -->
+<!--     group_by(trip_mode) %>% mutate(n_vals_1 = sum(n_vals)) %>%  -->
+<!--     mutate(prop = round(n_vals/n_vals_1 * 100, 1)) -->
+
+<!--   print(ggplot(data=df, aes(x=trip_distance_cat, y=prop)) + -->
+<!--           geom_bar(stat="identity", fill="steelblue")+ -->
+<!--           facet_wrap(vars(trip_mode)) + -->
+<!--           geom_text(aes(label=prop), vjust = -0.3, size = 3.5)+ -->
+<!--           theme_minimal() + -->
+<!--           theme(panel.spacing = unit(2, "lines")) + -->
+<!--           labs(title = paste(city, ' by dist by mode'))) -->
+<!-- } -->
+
+<!-- ``` -->
+
 ## Duration tables

### 2020-01-20 code/producing_trips_rd/trip-summary.Rmd
adding trip summary for cape town
updating repo

@@
 
-trip <- read_csv("../../data/local/bogota/bogota_trip.csv")
-quality_check(trip)
-trip_summary$bogota_city <- value
+#trip <- read_csv("../../data/local/bogota/bogota_trip.csv")
+#quality_check(trip)
+#trip_summary$bogota_city <- value
 
@@
 
+trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv")
+quality_check(trip)
+trip_summary$cape_town <- value
+
 ```

### 2020-01-20 summary_tables_outcomes.Rmd

@@
 # Assumes that multi_city_script.R has been run till 
-cities <- c('accra','sao_paulo','delhi','bangalore')
-
+cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota', 'cape_town', 'vizag')
 # round to

### 2020-01-27 summary_tables_outcomes.Rmd
Instead of hardcoded case study, Accra, use the first index of the cities to get column names

--- 2019-06-21 summary_tables_outcomes.Rmd
@@
 # Assumes that multi_city_script.R has been run till 
-cities <- c('accra','sao_paulo','delhi','bangalore')
+cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota','cape_town', 'vizag')
 
-# round to
 round_to <- 1
@@
 pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
-injury_col <- which(colnames(io$accra$outcomes$hb$deaths)=='scen1_deaths_inj')
-ap_cols <- which(sapply(colnames(io$accra$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
-pa_cols <- which(sapply(colnames(io$accra$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
+injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths)=='scen1_deaths_inj')
+ap_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
+pa_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
 scen_names <- rownames(scen_prop)

### 2020-02-11 code/producing_trips_rd/travel_survey.R
updating repo

@@ -2068,6 +2068,34 @@ write.csv(trip, "J:/Group/lambed/ITHIM-R/data/local/bangalore/bangalore_trip.csv
 
 #quality_check(trip)
 
+#####india Visakhapatnam ####
+rm(list =ls())
+
+source("J:/Group/lambed/ITHIM-R/code/producing_trips_rd/used_functions.R")
+
+visakhapatnam_trips <- read_csv("J:/Group/lambed/ITHIM-R/data/local/vizag/visakhapatnam_trips.csv")
+
+trip <- visakhapatnam_trips
+
+trip %<>% 
+  mutate(cluster_id = 1,
+         household_id = 1,
+         participant_id= ind_id,
+         trip_id = ifelse(is.na(travel_time), NA, X1),
+         sex = ifelse(sex=="M", "Male", "Female"),
+         trip_duration = travel_time,
+         trip_distance = distance,
+         trip_purpose = "unknown",
+         trip_mode = ifelse(mode=="autorickshaw"|mode=="shared_autorickshaw"|mode=="cyclerickshaw", "rickshaw",
+                            ifelse(mode=="mtw", "motorcycle",
+                                   ifelse(mode=="company_bus", "bus", mode))),
+         trip_wkly_frequency = wk_freq) %>% 
+  select(cluster_id, household_id, participant_id,age, sex, trip_id,trip_mode, trip_duration, trip_distance,trip_wkly_frequency, trip_purpose) 
+trip$year <- "2011"
+write.csv(trip, "J:/Group/lambed/ITHIM-R/data/local/vizag/visakhapatnam_trip.csv")
+
+#quality_check(trip)
+
 #####Mexico city ####
 
 rm(list =ls())

### 2020-02-12 summary_tables_outcomes.Rmd
Regen
Add anchor and its reference as example

@@
 
+Click here: [change in yll](#change_in_yll_100k)
+
 ```{r preprocessing}
+
 collapse_ages <- function(data,ages=c('15-49','50-69'),age_lab='age_cat'){
@@
 
-Change in YLL per 100,000 people by scenario
+## Change in YLL due to disease per 100,000 {#change_in_yll_100k}
 

### 2020-02-12 summary_tables_outcomes.Rmd
Add whw matrices for baseline
Add caption
Add anchors

@@
 Click here: [change in yll](#change_in_yll_100k)
+
+```{r whw}
+
+for(city in cities){
+  if (!city %in% c('buenos_aires', 'mexico_city')){
+    print(kable(format(round(io[[city]]$outcomes$whw$Baseline$whw)), caption = paste('Who Hit Whom (WHW) baseline matrix for ', city)))
+    cat("\n")
+  }
+}
+
+```
 
@@
 
+### Change in deaths due to injury
+
 ```{r}
@@
 sum_and_round_and_print(injury_totals,"Change in deaths due to injury in ")
+```
+
+### Change in deaths due to PA
+
+```{r}
 
@@
 sum_and_round_and_print(pa_totals,"Change in deaths due to PA in ")
+```
 
+### Change in deaths due to AP
+
+```{r}
 ap_totals <- lapply(io,function(x){
@@
 
-Change in deaths per 100,000 people by age group by scenario
+## Change in deaths per 100,000 people by age group by scenario
+
+### Change in deaths due to injury per 100,000
 
@@
 round_and_print(injury_rates,"Change in deaths due to injury per 100,000 in ")
+```
 
+### Change in deaths due to PA per 100,000
+```{r}
 pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
@@
 round_and_print(pa_rates,"Change in deaths due to PA per 100,000 in ")
+
+```
+
+### Change in deaths due to AP per 100,000
+```{r}
 
@@
 YLLs total (for the city based on real population size) by age group by scenario
+
+### Change in YLLs due to injury
 
@@
 sum_and_round_and_print(injury_totals,"Change in YLLs due to injury in ")
+```
 
+### Change in YLLs due to PA
+```{r}
 pa_totals <- lapply(io,function(x){
@@
 sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
+```
 
+### Change in YLLs due to AP
+```{r}
 ap_totals <- lapply(io,function(x){
@@
 
+### Change in YLLs due to injury per 100,000
+
 ```{r scen_prop = "asis"}
@@
 round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ")
+```
 
+### Change in YLLs due to PA per 100,000
+```{r}
 pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
@@
 round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ")
+```
 
+### Change in YLLs due to injury AP 100,000
+```{r}
 ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
@@
 Change in deaths total (for the city based on real population size) by scenario
+
+### Change in deaths due to disease
 
@@
 
+### Change in deaths due to disease per 100,000
+
 ```{r scen_prop = "asis"}
@@
 Change in YLL total (for the city based on real population size) by scenario
+
+### Change in YLL due to disease
 

### 2020-02-12 summary_tables_outcomes.Rmd
Add anchors throughout the document

@@
 
-Click here: [change in yll](#change_in_yll_100k)
+Click here: [change in yll](#change_yll_100k)
 
@@
 
+### Change in deaths {#change_death}
+
 ```{r scen_prop = "asis"}
@@
 
+### Change in deaths per 100,000 {#change_death_100k}
+
 ```{r scen_prop = "asis"}
@@
 YLLs total (for the city based on real population size) by age group  by scenario
+
+### Change in YLL {#change_YLL}
 
@@
 
+### Change in YLLs per 100,000 {#change_YLL_age_100k}
+
 ```{r scen_prop = "asis"}
@@
 YLLs per 100,000 people by gender by age group scenario
+
+### Change in YLLs per 100,000 {#change_YLL_gender_age_100k}
 
@@
 
-### Change in deaths due to injury
+### Change in deaths due to injury {#change_death_injury}
 
@@
 
-### Change in deaths due to PA
+### Change in deaths due to PA {#change_death_PA}
 
@@
 
-### Change in deaths due to AP
+### Change in deaths due to AP {#change_death_AP}
 
@@
 
-### Change in deaths due to injury per 100,000
+### Change in deaths due to injury per 100,000 {#change_death_injury_100k}
 
@@
 
-### Change in deaths due to PA per 100,000
+### Change in deaths due to PA per 100,000 {#change_death_PA_100k}
 ```{r}
@@
 
-### Change in deaths due to AP per 100,000
+### Change in deaths due to AP per 100,000 {#change_death_AP_100k}
 ```{r}
@@
 
-### Change in YLLs due to injury
+### Change in YLLs due to injury {#change_yll_injury}
 
@@
 
-### Change in YLLs due to PA
+### Change in YLLs due to PA {#change_yll_PA}
 ```{r}
@@
 
-### Change in YLLs due to AP
+### Change in YLLs due to AP {#change_yll_AP}
 ```{r}
@@
 
-### Change in YLLs due to injury per 100,000
+### Change in YLLs due to injury per 100,000 {#change_yll_injury_100k}
 
@@
 
-### Change in YLLs due to PA per 100,000
+### Change in YLLs due to PA per 100,000 {#change_yll_PA_100k}
 ```{r}
@@
 
-### Change in YLLs due to injury AP 100,000
+### Change in YLLs due to injury AP 100,000 {#change_yll_AP_100k}
 ```{r}
@@
 
-### Change in deaths due to disease
+### Change in deaths due to disease {#change_death_disease}
 
@@
 
-### Change in deaths due to disease per 100,000
+### Change in deaths due to disease per 100,000 {#change_death_disease_100k}
 
@@
 
-### Change in YLL due to disease
+### Change in YLL due to disease {#change_yll_disease}
 
@@
 
-## Change in YLL due to disease per 100,000 {#change_in_yll_100k}
+## Change in YLL due to disease per 100,000 {#change_yll_disease_100k}
 

### 2020-02-12 summary_tables_outcomes.Rmd
Add cross references

@@
 ```
-
-Click here: [change in yll](#change_yll_100k)
+* [change in yll](#change_yll_100k)
+* [Change in total deaths](#change_death)
+* [Change in deaths per 100k](#change_death_100k)
+* [Change in total YLLs](#change_YLL)
+* [Change in YLLs by age per 100k](#change_YLL_age_100k)
+* [Change in YLLs by gender by age per 100k](#change_in_YLL_gender_age_100k)
+* [Change in deaths due to injury](#change_death_injury)
+* [Change in deaths due to PA](#change_death_PA)
+* [Change in deaths due to AP](#change_death_AP)
+* [Change in deathe due to injury per 100k](#change_death_injury_100k)
+* [Change in deahts due to PA per 100k](#change_death_PA_100k)
+* [Change in deaths due to AP per 100k](#change_death_AP_100k)
+* [Change in YLLs due to injury](#change_yll_injury)
+* [Change in YLLs due to PA](#change_yll_PA)
+* [Change in YLLs due to Ap](#change_yll_AP)
+* [Change in YLLs due to injury per 100k](#change_yll_injury_100k)
+* [Change in YLLs due to PA per 100k](#change_yll_PA_100k)
+* [Change in YLLs due to AP per 100k](#change_yll_AP_100k)
+* [Change in deaths by disease due per 100k](#change_death_disease_100k)
+* [Change in YLLs by disease](#change_yll_disease)
+* [Change in deaths due per 100k](#change_yll_disease_100k)
 

### 2020-02-12 summary_tables_outcomes.Rmd
Use table of contents
Fix header

--- 2020-02-12 summary_tables_outcomes.Rmd
@@
 title: "Summary Tables"
-output: html_document
+output:
+  html_document:
+    toc: true
+    theme: united
 ---
@@
 ```
-
-Click here: [change in yll](#change_yll_100k)
 
@@
 
-## Change in YLL due to disease per 100,000 {#change_yll_disease_100k}
+### Change in YLL due to disease per 100,000 {#change_yll_disease_100k}
 

### 2020-02-12 summary_tables_outcomes.Rmd
Add whw for all scenarios

@@
     print(kable(data[[city]], caption = paste(text, city)))
-    cat("\n")
-  }
-}
-
-```
-
-```{r whw}
-
-for(city in cities){
-  if (!city %in% c('buenos_aires', 'mexico_city')){
-    print(kable(format(round(io[[city]]$outcomes$whw$Baseline$whw)), caption = paste('Who Hit Whom (WHW) baseline matrix for ', city)))
     cat("\n")
@@
 scen_names <- rownames(scen_prop)
+
+```
+
+## Who Hit Whom matrix
+
+Who hit whom matrix for baseline for each case study
+
+### Who hit whom matrix for baseline {#whw_baseline}
+
+```{r whw}
+
+for(city in cities){
+  if (!city %in% c('buenos_aires', 'mexico_city')){
+    for (cs in names(io$accra$outcomes$whw)){
+      print(kable(format(round(io[[city]]$outcomes$whw[[cs]]$whw)), caption = paste('Who Hit Whom (WHW) ', cs, ' matrix for ', city)))
+      cat("\n")
+    }
+
+  }
+}
 

### 2020-02-12 summary_tables_outcomes.Rmd
Add NOV table

@@
   if (!city %in% c('buenos_aires', 'mexico_city')){
-    for (cs in names(io$accra$outcomes$whw)){
+    for (cs in names(io$accra$outcomes$whw)[1]){
       print(kable(format(round(io[[city]]$outcomes$whw[[cs]]$whw)), caption = paste('Who Hit Whom (WHW) ', cs, ' matrix for ', city)))
+      cat("\n")
+    }
+
+  }
+}
+
+```
+
+No Other Vehicle matrix for baseline for each case study
+
+### No Other Vehicle matrix matrix for baseline {#nov_baseline}
+
+```{r nov}
+
+for(city in cities){
+  if (!city %in% c('bogota', 'vizag')){
+    for (cs in names(io$accra$outcomes$whw)[1]){
+      print(kable(format(round(io[[city]]$outcomes$whw[[cs]]$nov)), caption = paste('NOV ', cs, ' matrix for ', city)))
       cat("\n")

### 2020-02-13 summary_tables_outcomes.Rmd
Fix minor mistake
Add table for injury per 100k

@@
 
-### No Other Vehicle matrix matrix for baseline {#nov_baseline}
+### No Other Vehicle matrix for baseline {#nov_baseline}
 
@@
 ```
+
+### Change in inury per 100K
+
+```{r inj_100k = "asis"}
+require(tibble)
+
+for (city in cities){
+
+  print(city)
+
+  el <- list()
+
+  if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
+
+    for (cs in names(io[[city]]$outcomes$whw)){
+      #cs <- 'Baseline'
+      #city <- 'accra'
+      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+      td4 <- io[[city]]$dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
+
+      if (length(el) == 0){
+        el <- td4 %>% dplyr::select(stage_mode)
+      }
+
+      td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+      td5 <- td4
+
+      td4[, 2] <- round((td4[,3] / td4[,2]) * 100000, 1)
+
+      names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")
+
+      el <- inner_join(el, td4, by = 'stage_mode')
+
+      # for (i in 1:nrow(td4)){
+      #   td4[i, 2:(ncol(td4) - 2)] <- round((td4$total[i] /(td4[i, 2:6])) * 100000, 1)
+      #   
+      #   
+      # }
+
+      # print(td4)
+    }
+    print(kable(el, caption = city))
+    cat("\n")
+    #kable(print(el), caption = city)
+  }
+
+}
+```

### 2020-02-13 summary_tables_outcomes.Rmd
Move to the right section

@@
 
+### Change in deaths due to injury per 100,000K
+
+```{r inj_100k = "asis"}
+require(tibble)
+
+for (city in cities){
+
+  print(city)
+
+  el <- list()
+
+  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
+
+    for (cs in names(io[[city]]$outcomes$whw)){
+      # cs <- 'Baseline'
+      # city <- 'mexico_city'
+      if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+      }else if(names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+        td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+      }else {
+
+        td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+
+      }
+
+      td4 <- io[[city]]$dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
+
+      if (length(el) == 0){
+        el <- td4 %>% dplyr::select(stage_mode)
+      }
+
+      td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+      td5 <- td4
+
+      td4[, 2] <- round((td4[,3] / td4[,2]) * 100000, 1)
+
+      names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")
+
+      el <- inner_join(el, td4, by = 'stage_mode')
+
+      # for (i in 1:nrow(td4)){
+      #   td4[i, 2:(ncol(td4) - 2)] <- round((td4$total[i] /(td4[i, 2:6])) * 100000, 1)
+      #   
+      #   
+      # }
+
+      # print(td4)
+    }
+    print(kable(el, caption = city))
+    cat("\n")
+    #kable(print(el), caption = city)
+  #}
+
+}
+```
+
 ### Change in deaths due to PA {#change_death_PA}
@@
 ```
-
-### Change in inury per 100K
-
-```{r inj_100k = "asis"}
-require(tibble)
-
-for (city in cities){
-
-  print(city)
-
-  el <- list()
-
-  if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
-
-    for (cs in names(io[[city]]$outcomes$whw)){
-      #cs <- 'Baseline'
-      #city <- 'accra'
-      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
-      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-
-      td4 <- io[[city]]$dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
-
-      if (length(el) == 0){
-        el <- td4 %>% dplyr::select(stage_mode)
-      }
-
-      td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
-
-      td5 <- td4
-
-      td4[, 2] <- round((td4[,3] / td4[,2]) * 100000, 1)
-
-      names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")
-
-      el <- inner_join(el, td4, by = 'stage_mode')
-
-      # for (i in 1:nrow(td4)){
-      #   td4[i, 2:(ncol(td4) - 2)] <- round((td4$total[i] /(td4[i, 2:6])) * 100000, 1)
-      #   
-      #   
-      # }
-
-      # print(td4)
-    }
-    print(kable(el, caption = city))
-    cat("\n")
-    #kable(print(el), caption = city)
-  }
-
-}
-```

### 2020-02-20 summary_tables_outcomes.Rmd
Fix typo
Use true_dist instead
Add whw and nov together, and also produce all cities wide injury risk per 100k kms

@@
   if (!city %in% c('buenos_aires', 'mexico_city')){
-    for (cs in names(io$accra$outcomes$whw)[1]){
-      print(kable(format(round(io[[city]]$outcomes$whw[[cs]]$whw)), caption = paste('Who Hit Whom (WHW) ', cs, ' matrix for ', city)))
+    for (cs in names(io$accra$outcomes$whw)){
+      print(city)
+      td1 <- (io[[city]]$outcomes$whw[[cs]]$whw) %>% as.data.frame() %>% tibble::rownames_to_column("mode")
+      td3 <- td1
+      if (!is.null(io[[city]]$outcomes$whw[[cs]]$nov)){
+        td2 <- t(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame()
+        td2$mode <- 'NOV'
+
+        td3 <- plyr::rbind.fill(td1, td2)
+      }
+      td3 <- td3 %>% mutate(rowSum = rowSums(.[2:ncol(td3)], na.rm = T))
+      td3 <- td3 %>% janitor::adorn_totals("row")
+      td3[, 2:ncol(td3)] <- round(td3[, 2:ncol(td3)], 2)
+
+        print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', cs, ' matrix for ', city)))
       cat("\n")
@@
   if (!city %in% c('bogota', 'vizag')){
-    for (cs in names(io$accra$outcomes$whw)[1]){
+    for (cs in names(io$accra$outcomes$whw)){
       print(kable(format(round(io[[city]]$outcomes$whw[[cs]]$nov)), caption = paste('NOV ', cs, ' matrix for ', city)))
@@
 
-### Change in deaths due to injury per 100,000K
+### Change in deaths due to injury per 100,000 km
 
@@
 
+overall_el <- list()
+
 for (city in cities){
+
+  #city <- 'delhi'
 
@@
 
-      td4 <- io[[city]]$dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
+      td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
 
@@
     print(kable(el, caption = city))
+
+    td <- el %>% dplyr::select(-contains('count'))
+    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+
+    overall_el[[city]] <- td
+
     cat("\n")
@@
 }
+```
+
+### Change in deaths due to injury per 100,000 km across cities
+
+```{r inj_100k_all_cities = "asis"}
+require(tibble)
+
+td <- overall_el %>% purrr::reduce(full_join, by = "stage_mode") %>% as.data.frame()
+
+td <- td %>% dplyr::select(stage_mode, sort(names(.)))
+
+print(kable(format(td, scientific = F), caption = paste('Injury per 100k across all cities')))
+
 ```

### 2020-02-20 summary_tables.Rmd
Use qualified name for a function (n), and comment out PA calculations

@@
 
-  df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n(), prop = round(n_vals / nrow(.) * 100, 1))
+  df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), prop = round(n_vals / nrow(.) * 100, 1))
   print(kable(df, caption = paste("Distance by mode by for  ", city)))
@@
 ```
-
-## Physical activity tables
-Case study specific physical activity tables for backgound (non-travel) PA, and baseline and five scenarios (non-travel + travel)
-
-```{r mmets = "asis"}
-
-for(city in cities){
-  df1 <- as.data.frame(io[[city]]$synth_pop$work_ltpa_marg_met)
-  colnames(df1) <- "mmet"
-
-  td1 <- df1 %>% filter(mmet == 0) %>% 
-    summarise(perct_zeros = round(dplyr::n() / nrow(df1) * 100, 1))
-
-  td2 <- df1 %>%
-    summarise(min = min(mmet),
-              q_25th = round(quantile(mmet, 0.25), 1),
-              median = round(quantile(mmet, 0.50), 1),
-              mean = round(mean(mmet), 1),
-              q_75th = round(quantile(mmet, 0.75), 1),
-              q_90th = round(quantile(mmet, 0.95), 1),
-              max = round(max(mmet), 1))
-
-  td3 <- cbind(td1, td2)
-  td3 <- as.data.frame(t(td3))
-
-  df2 <- select(io[[city]]$outcomes$mmets, base_mmet : scen5_mmet)
-
-  td4 <- df2 %>% gather(key = "scenario", value = "mmet") %>% 
-    filter(mmet == 0) %>% 
-    group_by(scenario) %>% 
-    summarise(perct_zeros = round(dplyr::n() / nrow(df2) * 100, 1))
-
-  td5 <- df2 %>% gather(key = "scenario", value = "mmet") %>% 
-    group_by(scenario) %>% 
-    summarise(min = min(mmet),
-              q_25th = round(quantile(mmet, 0.25), 1),
-              median = round(quantile(mmet, 0.50), 1),
-              mean = round(mean(mmet), 1),
-              q_75th = round(quantile(mmet, 0.75), 1),
-              q_90th = round(quantile(mmet, 0.95), 1),
-              max = round(max(mmet), 1))
-
-  td6 <- left_join(td4, td5, "scenario")
-  td6 <- as.data.frame(t(td6))
-  td6 <- td6[-1,]
-  td6 <- mutate_all(td6, function(x) as.numeric(as.character(x)))
-
-  td <- cbind(td3, td6)
-
-  colnames(td)[1] <- "Background (non-travel)"
-  colnames(td)[2] <- "Baseline"
-  colnames(td)[3:7] <- rownames(scen_prop)
-
-  print(kable(td, caption = paste("Marginal MET-h/wk for ", city)))
-  cat("\n")
-}
-
-```
-
-```{r pa, echo = F, message = F}
-
-if (plot_mmets){
-
-  dfSummary(accra_mmets <- io[["accra"]]$outcomes$mmets, style = 'grid', graph.magnif = 0.75, tmp.img.dir = "/tmp")
-
-  dfSummary(sao_paulo_mmets <- io[["sao_paulo"]]$outcomes$mmets, style = 'grid', graph.magnif = 0.75, tmp.img.dir = "/tmp")
-
-  dfSummary(delhi_mmets <- io[["delhi"]]$outcomes$mmets, style = 'grid', graph.magnif = 0.75, tmp.img.dir = "/tmp")
-
-  dfSummary(bangalore_mmets <- io[["bangalore"]]$outcomes$mmets, style = 'grid', graph.magnif = 0.75, tmp.img.dir = "/tmp")
-}
-
-```

### 2020-02-20 summary_tables_outcomes.Rmd
WHW for all cities

@@
 library(dplyr)
+library(purrr)
+library(DT)
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
@@
 
-Who hit whom matrix for baseline for each case study
+Who hit whom matrix for all scenarios for each case study
 
-### Who hit whom matrix for baseline {#whw_baseline}
+### Who hit whom matrix {#whw}
 
 ```{r whw}
+
+all_whw <- list()
 
@@
       cat("\n")
+
+      td <- td3
+      names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+
+      all_whw[[cs]][[city]] <- td
     }
@@
   }
+}
+
+```
+
+### Who hit whom matrix for all case cities {#whw_all}
+
+```{r whw_for_all_cs}
+
+for (cs in names(all_whw)){
+
+  td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = "mode") %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))
+
+  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total'))
+
+   td[is.na(td)] <- 0
+
+  print(kable(format(td, scientific = F), caption = paste('Who Hit Whom (WHW) for all case cities for ', cs)))
+
 }

### 2020-02-20 summary_tables_outcomes.Rmd
Use datatable to select columns

@@
 
+```
+
+### Test DT
+
+```{r dt}
+
+fname <- "whw"
+
+DT::datatable(td, 
+                  extensions = 'Buttons',
+                  escape = FALSE,
+                  rownames = FALSE,
+                  options = list(dom = 'Bfrtip',
+                                 buttons = 
+                                   list('colvis', list(
+                                     extend = 'collection',
+                                     buttons = list(list(extend='csv',
+                                                         filename = fname),
+                                                    list(extend='excel',
+                                                         filename = fname)),
+                                     text = 'Download'
+                                   )),
+                                 scrollX = TRUE,
+                                 pageLength = nrow(td),
+                                 order=list(list(2,'desc'))))
 ```

### 2020-02-21 summary_tables_outcomes.Rmd
Add DT for whw for all cities where users can select columns of interest

--- 2020-02-20 summary_tables_outcomes.Rmd
@@
 
+st <- list()
+
+td <- NULL
+
 for (cs in names(all_whw)){
+  # cs <- 'Baseline'
 
   td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = "mode") %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))
+
+   td[is.na(td)] <- 0
+
+   unk <- td %>% filter(mode %in% c('unknown', 'unk')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'unknown')
+
+   unsp <- td %>% filter(mode %in% c('unspecified', 'listed_na', '?')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'unspecified')
+
+   other <- unsp <- td %>% filter(mode %in% c('other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'other')
+
+   pu <- td %>% filter(mode %in% c('picku', 'pickup/light goods')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'pickup')
+
+   bus <- td %>% filter(mode %in% c('bus', 'bus_driver')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'bus')
+
+   td <- td %>% filter(!mode %in% c('bus', 'bus_driver', 'unknown', 'unk', 'unspecified', 'listed_na', '?', 
+                                  'other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle',
+                                  'picku', 'pickup/light goods'))
+
+   td <- plyr::rbind.fill(td, bus, unk, unsp, other, pu)
+
+   td <- (td[!duplicated(td), ])
 
@@
 
+   td <- td %>% mutate_if(is.numeric, round, 2)
+
+   readr::write_csv(td, paste0('results/multi_city/whw_', cs, '.csv'))
+
+   st[[cs]] <- format(td, scientific = F)
+
   print(kable(format(td, scientific = F), caption = paste('Who Hit Whom (WHW) for all case cities for ', cs)))
@@
 }
+
+```
+
+### DT - Who hit whom matrix for all case cities {#whw_all}
+
+#### Baseline
+
+```{r dt}
+
+#for (cs in names(all_whw)){
+td <- NULL
+
+cs <- 'Baseline'
+
+fname <- paste("whw", cs)
+
+td <- st[[cs]]
+
+DT::datatable(td, 
+              extensions = 'Buttons',
+              escape = FALSE,
+              rownames = FALSE,
+              options = list(dom = 'Bfrtip',
+                             buttons = 
+                               list('colvis', list(
+                                 extend = 'collection',
+                                 buttons = list(list(extend='csv',
+                                                     filename = fname),
+                                                list(extend='excel',
+                                                     filename = fname)),
+                                 text = 'Download'
+                               )),
+                             scrollX = TRUE,
+                             pageLength = nrow(td),
+                             order=list(list(2,'desc')),
+                             columnDefs = list(list(visible=FALSE, targets=c(1:(ncol(td) - 1))))))
+
+# }
 

### 2020-02-21 summary_tables_outcomes.Rmd
Remove NOV as it's already part of WHW

@@
 # }
-
-```
-
-No Other Vehicle matrix for baseline for each case study
-
-### No Other Vehicle matrix for baseline {#nov_baseline}
-
-```{r nov}
-
-for(city in cities){
-  if (!city %in% c('bogota', 'vizag')){
-    for (cs in names(io$accra$outcomes$whw)){
-      print(kable(format(round(io[[city]]$outcomes$whw[[cs]]$nov)), caption = paste('NOV ', cs, ' matrix for ', city)))
-      cat("\n")
-    }
-
-  }
-}
 

### 2020-02-21 summary_tables.Rmd
Add true dist, and also add TOC

@@
 title: "Summary Tables"
-output: html_document
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+output:
+  html_document:
+    toc: true
+    theme: united
 ---
@@
 
+## True Distance by total population
+### Tables
+
+```{r trip_dist_mode_figs = "asis"}
+
+for (city in cities){
+
+  df <- io[[city]]$true_dist
+  print(kable(df, caption = paste("True distance by mode for  ", city)))
+}
+
+```
+
 <!-- ### Figures -->

### 2020-02-21 summary_tables_outcomes.Rmd
Update path
Regen
Add injury risks to a csv file

@@
 
-   readr::write_csv(td, paste0('results/multi_city/whw_', cs, '.csv'))
+   readr::write_csv(td, paste0('results/multi_city/whw_matrices/whw_', cs, '.csv'))
 
@@
 ### DT - Who hit whom matrix for all case cities {#whw_all}
-
-#### Baseline
+Baseline
 
@@
 
+readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks.csv')
+
 print(kable(format(td, scientific = F), caption = paste('Injury per 100k across all cities')))

### 2020-02-21 summary_tables_outcomes.Rmd
Fix walking in whw for mexico city and buenos aires

@@
   if (!city %in% c('buenos_aires', 'mexico_city')){
+     # city <- 'buenos_aires'
     for (cs in names(io$accra$outcomes$whw)){
       print(city)
+
+      if (is.null(io[[city]]$outcomes$whw[[cs]]$whw)){
+        if (!is.null(io[[city]]$outcomes$whw[[cs]]$nov)){
+          td2 <- t(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame()
+          td2$mode <- 'NOV'
+
+          td2 <- td2 %>% dplyr::select(mode, names(.))
+
+          td3 <- td2
+        }
+
+      }else{
         td1 <- (io[[city]]$outcomes$whw[[cs]]$whw) %>% as.data.frame() %>% tibble::rownames_to_column("mode")
@@
           td3 <- plyr::rbind.fill(td1, td2)
+        }
       }
@@
         td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
 

### 2020-02-24 summary_tables_outcomes.Rmd
Use qualified scen names

@@
 
+scen_names_w_baseline <- append('baseline', scen_names)
+scen_lt <- data.frame("scen_name" = names(io$accra$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)
+
 all_whw <- list()
@@
 
-      print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', cs, ' matrix for ', city)))
+      var <- scen_lt %>% filter(scen_name == cs) %>% select(val) %>% as.character()
+
+      print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', var, ' matrix for ', city)))
       cat("\n")
@@
 
-      all_whw[[cs]][[city]] <- td
+      all_whw[[var]][[city]] <- td
     }
@@
 
+   #var <- scen_lt %>% filter(scen_name == cs) %>% select(val) %>% as.character()
+
+   #print(var)
+
    readr::write_csv(td, paste0('results/multi_city/whw_matrices/whw_', cs, '.csv'))
@@
 
-fname <- paste("whw", cs)
+var <- scen_lt %>% filter(scen_name == cs) %>% select(val) %>% as.character()
 
-td <- st[[cs]]
+fname <- paste("whw_", var, ".csv")
+
+td <- st[[var]]
 
@@
       # cs <- 'Baseline'
-      # city <- 'mexico_city'
+      # city <- 'vizag'
       if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
@@
 
-      }else if(names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
         td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
         td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-      }else {
+      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
 
@@
 
-      td4[, 2] <- round((td4[,3] / td4[,2]) * 100000, 1)
+      td4[, 2] <- round((td4[,3] / td4[,2] ) * 100000, 4)
+      # / sum(io[[city]]$demographic$population, na.rm = T)
 

### 2020-02-24 summary_tables.Rmd
Add datasets of true distances
Use qualified scen names

@@
   df <- io[[city]]$true_dist
+
+  names(df)[-1] <- lapply(append('baseline', rownames(scen_prop)), as.character)
+
+  readr::write_csv(df, paste0('results/multi_city/true_distances/true_distance_', city, '.csv'))
+
   print(kable(df, caption = paste("True distance by mode for  ", city)))

### 2020-02-25 code/producing_trips_rd/trip-summary.Rmd
updating repo

--- 2019-10-11 code/producing_trips_rd/trip-summary.Rmd
@@
 quality_check(trip)
-trip_summary$buenos_aires <- value
+trip_summary$Buenos_Aires <- value
 
@@
 quality_check(trip)
-trip_summary$sao_paulo <- value
+trip_summary$Sao_Paulo <- value
 
@@
 quality_check(trip)
-trip_summary$belo_horizonte <- value
+trip_summary$Belo_Horizonte <- value
+
+trip <- read_csv("../../data/local/mexico/mexico_city_trip.csv")
+quality_check(trip)
+trip_summary$Mexico <- value
+
+trip <- read_csv("../../data/local/bogota/bogota_trip.csv")
+quality_check(trip)
+trip_summary$Bogota <- value
 
@@
 quality_check(trip)
-trip_summary$santiago_city <- value
+trip_summary$Santiago <- value
 
-trip <- read_csv("../../data/local/bogota/bogota_trip.csv")
+trip <- read_csv("../../data/local/vizag/visakhapatnam_trip.csv")
 quality_check(trip)
-trip_summary$bogota_city <- value
+trip_summary$Visakhapatnam <- value
+
+trip <- read_csv("../../data/local/delhi/delhi_trip.csv")
+quality_check(trip)
+trip_summary$Delhi <- value
+
+trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv")
+quality_check(trip)
+trip_summary$Bangalore <- value
 
@@
 quality_check(trip)
-trip_summary$accra_city <- value
+trip_summary$Accra <- value
 
-trip <- read_csv("../../data/local/delhi/delhi_trip.csv")
+trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv")
 quality_check(trip)
-trip_summary$delhi_city <- value
-
-trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv")
-quality_check(trip)
-trip_summary$bangalore_city <- value
-
-trip <- read_csv("../../data/local/mexico/mexico_city_trip.csv")
-quality_check(trip)
-trip_summary$mexico_city <- value
+trip_summary$Cape_Town <- value
 

### 2020-02-25 summary_tables_outcomes.Rmd
Use qualified package name with select function
Use more self explanatory scenario names

@@
 
-scen_names_w_baseline <- append('baseline', scen_names)
+scen_names_w_baseline <- c('bl', 'w_sc', 'bi_sc', 'car_sc', 'mc_sc', 'bus_sc')
 scen_lt <- data.frame("scen_name" = names(io$accra$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)
@@
 
-      var <- scen_lt %>% filter(scen_name == cs) %>% select(val) %>% as.character()
+      var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
 
@@
 
-   #print(var)
-
    readr::write_csv(td, paste0('results/multi_city/whw_matrices/whw_', cs, '.csv'))
@@
 
-var <- scen_lt %>% filter(scen_name == cs) %>% select(val) %>% as.character()
+var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
 
@@
 
+      var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+
+      names(td4)[2] <- var
+
       td5 <- td4
 
-      td4[, 2] <- round((td4[,3] / td4[,2] ) * 100000, 4)
+      td4[, 2] <- round((td4[,3] / ( td4[,2] )) * 100000, 4)
       # / sum(io[[city]]$demographic$population, na.rm = T)

### 2020-02-26 summary_tables_outcomes.Rmd
Scale true distance to year (by 52)
Write long form, interactive and static plots of whw

@@
 library(DT)
+library(ggplot2)
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
@@
 
-```{r whw_for_all_cs}
+```{r whw_for_all_cs, fig.width=7, fig.height=4, echo=FALSE}
 
 st <- list()
+
+file_list <- list()
 
@@
 for (cs in names(all_whw)){
-  # cs <- 'Baseline'
+  # cs <- names(all_whw)[1]
 
@@
 
-   #var <- scen_lt %>% filter(scen_name == cs) %>% select(val) %>% as.character()
+   backup_td <- td
 
    readr::write_csv(td, paste0('results/multi_city/whw_matrices/whw_', cs, '.csv'))
+
+   colnames(td) = gsub("auto_rickshaw", "ar", colnames(td))
+
+   whw_lng <- reshape2::melt(td)
+
+   col_split <- stringr::str_split(whw_lng$variable, "_", simplify = TRUE, n = 2)
+
+   whw_lng <- cbind(whw_lng, col_split)
+   names(whw_lng)[4] <- 'strike_mode'
+   names(whw_lng)[5] <- 'city'
+   whw_lng <- whw_lng %>% dplyr::select(-variable)
+   whw_lng <- whw_lng %>% dplyr::filter(strike_mode != "rowSum")
+   whw_lng <- whw_lng %>% rename(str_mode = mode, cas_mode = strike_mode)
+
+   whw_lng$str_mode <- factor(whw_lng$str_mode, levels = unique(whw_lng$str_mode))
+
+   qualified_scen_name <- 'Bus'
+   scen <- 'Scenario'
+
+   if (cs == 'bl'){
+     qualified_scen_name <- 'Baseline'
+     scen <- ''
+   }else if(cs == "w_sc"){
+     qualified_scen_name <- 'Walking'
+   }else if(cs == "bi_sc"){
+     qualified_scen_name <- 'Bicycle'
+   }else if(cs == "car_sc"){
+     qualified_scen_name <- 'Car'
+   }else if(cs == "mc_sc"){
+     qualified_scen_name <- 'Motorcyle'
+   }
+
+   readr::write_csv(whw_lng, paste0('results/multi_city/whw_matrices/whw_lng_', cs, '.csv'))
+
+   f <- ggplot(data = whw_lng) +
+     aes(x = str_mode, fill = cas_mode, weight = value) +
+     geom_bar(position = "stack") +
+     scale_fill_brewer(palette = "Dark2") +
+     labs(x = "Strike Mode", y = "# of incidents", title = paste("WHW - ", qualified_scen_name, scen)) +
+     coord_flip() +
+     facet_wrap(vars(city), scales = "free_y")
+
+   file_list[[cs]] <- f
+
+   ggsave(f, file = paste0("results/multi_city/whw_matrices/plots/whw_", cs,".png"), limitsize = F, dpi = 300, scale = 1.5)
+
+   d <- paste0("results/multi_city/whw_matrices/interactive_plots/whw_", cs,".html")
+
+   htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
+  #ggsave(f, file=paste0("results/multi_city/whw_matrices/whw_", cs,".png"), width = 14, height = 10, units = "cm")
 
@@
 
-      td4[, 2] <- round((td4[,3] / ( td4[,2] )) * 100000, 4)
+      td4[, 2] <- round((td4[,3] / ( td4[,2] * 52)) * 100000, 4)
       # / sum(io[[city]]$demographic$population, na.rm = T)

### 2020-02-28 summary_tables_outcomes.Rmd
Add buenos aires and mexico city's whw matrices

@@
 cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota','cape_town', 'vizag')
+
+#cities <- c('accra','sao_paulo','buenos_aires')
 
@@
 for(city in cities){
-  if (!city %in% c('buenos_aires', 'mexico_city')){
+  # if (!city %in% c('buenos_aires', 'mexico_city'))
+    {
      # city <- 'buenos_aires'
@@
       # cs <- 'Baseline'
-      # city <- 'vizag'
+      # cs <- "Scenario 4"
+      # city <- 'buenos_aires'
       if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
@@
 readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks.csv')
+
+colnames(td) = gsub("bi_sc", "bisc", colnames(td))
+
+colnames(td) = gsub("bus_sc", "bssc", colnames(td))
+
+colnames(td) = gsub("car_sc", "csc", colnames(td))
+
+colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
+
+colnames(td) = gsub("w_sc", "wsc", colnames(td))
+
+injury_risks_lng <- reshape2::melt(td)
+
+col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)
+
+injury_risks_lng <- cbind(injury_risks_lng, col_split)
+names(injury_risks_lng)[4] <- 'scenario'
+names(injury_risks_lng)[5] <- 'city'
+injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
+cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
+
+injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
+
+injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycle Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Bus Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycle Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking Scenario'
+
+d <- ggplot(injury_risks_lng) +
+  aes(x = stage_mode, fill = scenario, weight = value) +
+  geom_bar(position = "dodge", colour = 'black') +
+  scale_color_manual(values = cols) +
+  coord_flip() +
+  theme_minimal() +
+  facet_wrap(vars(city)) +
+  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per 100k km"))
+
+ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_100k_km.png"), limitsize = F, dpi = 300, scale = 1.5)
+
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_100k_km.html")
+
+htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
 

### 2020-02-28 summary_tables_outcomes.Rmd
Add deaths per 100k for injuries

@@
 
+### # of deaths due to injury per 100,000 people
+
+```{r inj_100k = "asis"}
+require(tibble)
+
+overall_el <- list()
+
+for (city in cities){
+
+  print(city)
+
+  el <- list()
+
+  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
+
+    for (cs in names(io[[city]]$outcomes$whw)){
+      # cs <- 'Baseline'
+      # cs <- "Baseline"
+      # city <- 'accra'
+      if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+        td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+
+        td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+      }
+
+      td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
+
+      if (length(el) == 0){
+        el <- td4 %>% dplyr::select(stage_mode)
+      }
+
+      td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+      var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+
+      names(td4)[2] <- var
+
+      td5 <- td4
+
+      td4[, 2] <- round((td4[,3] / ( td4[,2] * 52)) * 100000, 4)
+
+      td4[, 3] <- round(td4[, 3] / sum(io[[city]]$demographic$population) * 100000, 2)
+      # / sum(io[[city]]$demographic$population, na.rm = T)
+
+      names(td4)[2] <- paste(names(td4)[2], names(td4)[3], sep = "_")
+
+      td4$count <- NULL
+
+      el <- inner_join(el, td4, by = 'stage_mode')
+
+      # for (i in 1:nrow(td4)){
+      #   td4[i, 2:(ncol(td4) - 2)] <- round((td4$total[i] /(td4[i, 2:6])) * 100000, 1)
+      #   
+      #   
+      # }
+
+      # print(td4)
+    }
+    print(kable(el, caption = city))
+
+    #td <- el %>% dplyr::select(-contains('count'))
+    td <- el
+    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+
+    overall_el[[city]] <- td
+
+    cat("\n")
+    #kable(print(el), caption = city)
+  #}
+
+}
+```
+
 ### Change in deaths due to PA {#change_death_PA}

### 2020-02-28 summary_tables.Rmd
Add true distances per person as well

@@
 cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota', 'cape_town', 'vizag')
+
+#cities <- c('accra','sao_paulo','buenos_aires')
 
@@
 
+## Distance tables by using true distance
+Case study specific distance tables for baseline and five scenarios
+
+```{r trip_dist = "asis"}
+
+for(city in cities){
+  count_people <- nrow(io[[city]]$synth_pop)
+  td <- io[[city]]$true_dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+  row.names(td) <- NULL
+
+  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person")))
+  cat("\n")
+
+  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
+  td <- io[[city]]$true_dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+  row.names(td) <- NULL
+
+  print(kable(td, caption = paste("Distance table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))
+
+}
+
+```
+
 ## Distance by distance category

### 2020-03-02 summary_tables_outcomes.Rmd
Fix number of deaths per 100k

--- 2020-02-28 summary_tables_outcomes.Rmd
@@
 
+### # of deaths due to injury per 100,000 people
+
+```{r inj_100k = "asis"}
+require(tibble)
+
+overall_el <- list()
+
+for (city in cities){
+
+  print(city)
+
+  el <- list()
+
+  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
+
+    for (cs in names(io[[city]]$outcomes$whw)){
+      # cs <- 'Baseline'
+      # # cs <- "Baseline"
+      # city <- 'accra'
+      if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+        td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+
+        td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+      }
+
+      td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode)
+
+      if (length(el) == 0){
+        el <- td4 %>% dplyr::select(stage_mode)
+      }
+
+      td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+      var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+
+      names(td4)[2] <- var
+
+      td5 <- td4
+
+      td4[, 2] <- round(td4[, 2] / sum(io[[city]]$demographic$population) * 100000, 2)
+
+      el <- inner_join(el, td4, by = 'stage_mode')
+    }
+    print(kable(el, caption = city))
+    td <- el
+    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+
+    overall_el[[city]] <- td
+
+    cat("\n")
+    #kable(print(el), caption = city)
+  #}
+
+}
+```
+
 ### Change in deaths due to PA {#change_death_PA}

### 2020-03-02 summary_tables.Rmd
Create baseline total population distance with suffix for million and billion

--- 2020-02-24 summary_tables.Rmd
@@
 
+## Distance tables scaled by using total population (per year in km)
+Case study specific distance tables for baseline and five scenarios
+
+```{r trip_tot_dist = "asis", message=F}
+
+bl_td <- list()
+
+require(gt)
+
+for(city in cities){
+
+  print(city)
+  count_people <- round(sum(io[[city]]$demographic$population))
+  td <- io[[city]]$true_dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 52, 3)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+
+  #
+
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+  row.names(td) <- NULL
+
+  if (length(bl_td) == 0){
+    bl_td <- td %>% dplyr::select(stage_mode)
+  }
+
+  td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
+
+  td1 <- td %>% dplyr::select(stage_mode, Baseline)
+
+  names(td1)[2] <- city
+
+  bl_td <- inner_join(bl_td, td1, by = 'stage_mode')
+
+  td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)
+
+  print(td1)
+
+  # print(kable(td, caption = paste("Distance table (km) for ", city, "( ", sitools::f2si(count_people), " ) per person")))
+  cat("\n")
+
+}
+
+```
+
+## Baseline total distance tables for all case studies (per year in km)
+
+```{r trip_dist_all_cities = "asis", message=F}
+
+# tab <- bl_td %>% gt() %>% fmt_number(columns = 2:ncol(bl_td), decimals = T, suffixing = T)
+
+# Source: https://stackoverflow.com/questions/28159936/formatting-large-currency-or-dollar-values-to-millions-billions
+comprss <- function(tx) { 
+  div <- findInterval(as.numeric(gsub("\\,", "", tx)), 
+                      c(0, 1e3, 1e6, 1e9, 1e12) )  # modify this if negative numbers are possible
+  paste(round( as.numeric(gsub("\\,","",tx))/10^(3*(div-1)), 2), 
+        c("","K","M","B","T")[div] )
+}
+
+backup <- bl_td
+
+td <- as.data.frame(bl_td)
+
+for (i in 2:ncol(td)){
+  td[, i] <- comprss(td[, i])
+
+}
+
+print(kable(td, caption = paste("Distance table (km)")))
+
+```
+
 ## Distance by distance category

### 2020-03-02 summary_tables_outcomes.Rmd
Reformat

@@
 
-overall_el <- list()
+overall_el_normalized <- list()
 
@@
   }
+
+  el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
+
   print(kable(el, caption = city))
@@
 
-    overall_el[[city]] <- td
+  overall_el_normalized[[city]] <- td
 

### 2020-03-02 summary_tables.Rmd
Keep help functions at top
Add trip prop and trip distance for all case studies, for baseline

@@
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
+
+```
+
+```{r helper_functions, echo = F, message = F}
+
+# Source: https://stackoverflow.com/questions/28159936/formatting-large-currency-or-dollar-values-to-millions-billions
+comprss <- function(tx) { 
+  div <- findInterval(as.numeric(gsub("\\,", "", tx)), 
+                      c(0, 1e3, 1e6, 1e9, 1e12) )  # modify this if negative numbers are possible
+  paste(round( as.numeric(gsub("\\,","",tx))/10^(3*(div-1)), 2), 
+        c("","K","M","B","T")[div] )
+}
 
@@
 
+trip_prop <- list()
+
 for(city in cities){
@@
 
+  td1 <- td %>% rownames_to_column()
+
+  names(td1)[1] <- 'stage_mode'
+
+  if (length(trip_prop) == 0){
+    trip_prop <- td1 %>% dplyr::select(stage_mode)
+  }
+
+  td1 <- td1 %>% dplyr::select(stage_mode, baseline)
+
+  names(td1)[2] <- city
+
+  trip_prop <- inner_join(trip_prop, td1, by = 'stage_mode')
+
   print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
@@
 }
+
+```
+
+### Trip proportion Baseline
+```{r}
+
+kable(trip_prop)
 
@@
 
+trip_dist <- list()
+
 for(city in cities){
+
   count_people <- nrow(io[[city]]$synth_pop)
@@
 
+  td1 <- td
+
+  if (length(trip_dist) == 0){
+    trip_dist <- td1 %>% dplyr::select(stage_mode)
+  }
+
+  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
+
+  names(td1)[2] <- city
+
+  trip_dist <- inner_join(trip_dist, td1, by = 'stage_mode')
+
   count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
@@
 }
+
+```
+
+### Average distance Baseline
+```{r}
+
+kable(trip_dist)
 
@@
 # tab <- bl_td %>% gt() %>% fmt_number(columns = 2:ncol(bl_td), decimals = T, suffixing = T)
-
-# Source: https://stackoverflow.com/questions/28159936/formatting-large-currency-or-dollar-values-to-millions-billions
-comprss <- function(tx) { 
-  div <- findInterval(as.numeric(gsub("\\,", "", tx)), 
-                      c(0, 1e3, 1e6, 1e9, 1e12) )  # modify this if negative numbers are possible
-  paste(round( as.numeric(gsub("\\,","",tx))/10^(3*(div-1)), 2), 
-        c("","K","M","B","T")[div] )
-}
 

### 2020-03-02 summary_tables.Rmd
Add captions

@@
 
-### Trip proportion Baseline
+### Baseline trip proportion (%)
 ```{r}
 
-kable(trip_prop)
+kable(trip_prop, caption = 'Trip prop (%) for all case studies - only selected modes')
 
@@
 
-  print(kable(td, caption = paste("Distance table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))
+  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips)")))
 
@@
 
-### Average distance Baseline
+### Baseline average distance (km) for all case studies
 ```{r}
 
-kable(trip_dist)
+kable(trip_dist, caption = "Avg. distance per person (km) across all case studies")
 

### 2020-03-02 summary_tables.Rmd
Fix title
Change heading number
Add row sums for average and total distance, and trip prop

@@
 
+trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>% mutate_if(is.factor, ~ as.numeric(as.character(.x))) %>% janitor::adorn_totals('row')
+
 kable(trip_prop, caption = 'Trip prop (%) for all case studies - only selected modes')
@@
 
+trip_dist <- trip_dist %>% janitor::adorn_totals('row')
+
 kable(trip_dist, caption = "Avg. distance per person (km) across all case studies")
@@
 
-## Baseline total distance tables for all case studies (per year in km)
+### Baseline total distance for all case studies (per year in km)
 
@@
 
-td <- as.data.frame(bl_td)
+td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')
 

### 2020-03-03 summary_tables_outcomes.Rmd
Make warning, echo and message to false

@@
 library(ggplot2)
-knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
+knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')
 
@@
 for (cs in names(all_whw)){
-  # cs <- names(all_whw)[1]
+  cs <- names(all_whw)[1]
 
@@
 
-```{r scen_prop = "asis"}
+```{r scen_prop = "asis", echo = F, message = F}
 death_rates <- lapply(cities,function(x) rbind(death_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(death_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
@@
 
-```{r inj_100k = "asis"}
+```{r inj_100k = "asis", echo = F, message = F}
 require(tibble)

### 2020-03-03 summary_tables_outcomes.Rmd
Remove deprecated functions

@@
 for (cs in names(all_whw)){
-  cs <- names(all_whw)[1]
+  # cs <- names(all_whw)[1]
 
@@
 
-  unk <- td %>% filter(mode %in% c('unknown', 'unk')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'unknown')
+  unk <- td %>% filter(mode %in% c('unknown', 'unk')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unknown')
 
-  unsp <- td %>% filter(mode %in% c('unspecified', 'listed_na', '?')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'unspecified')
+  unsp <- td %>% filter(mode %in% c('unspecified', 'listed_na', '?')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unspecified')
 
-  other <- unsp <- td %>% filter(mode %in% c('other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'other')
+  other <- unsp <- td %>% filter(mode %in% c('other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'other')
 
-  pu <- td %>% filter(mode %in% c('picku', 'pickup/light goods')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'pickup')
+  pu <- td %>% filter(mode %in% c('picku', 'pickup/light goods')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'pickup')
 
-  bus <- td %>% filter(mode %in% c('bus', 'bus_driver')) %>% summarise_if(is.numeric, funs(sum)) %>% mutate(mode = 'bus')
+  bus <- td %>% filter(mode %in% c('bus', 'bus_driver')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'bus')
 

### 2020-03-03 summary_tables.Rmd
Add unit with distance and duration tables

@@
 
-  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person")))
+  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person")))
   cat("\n")
@@
 
-### Baseline average distance (km) for all case studies
+### Baseline average distance (km) per day for all case studies
 ```{r}
@@
 
-kable(trip_dist, caption = "Avg. distance per person (km) across all case studies")
+kable(trip_dist, caption = "Avg. distance per person (km) per day across all case studies")
 
@@
   count_people <- round(sum(io[[city]]$demographic$population))
-  td <- io[[city]]$true_dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 52, 3)))
+  td <- io[[city]]$true_dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 365, 3)))
   colnames(td)[3:7] <- rownames(scen_prop)
@@
 
-## True Distance by total population
+## True Distance by total population (per day in km)
 ### Tables
@@
 
-  print(kable(df, caption = paste("True distance by mode for  ", city)))
+  print(kable(df, caption = paste("True distance by mode per day for  ", city)))
 }
@@
 ## Duration tables
-Case study specific duration tables for baseline and five scenarios
+Case study specific duration tables for baseline and five scenarios (per day in km)
 
@@
 
-  print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (everyone)")))
+  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (everyone)")))
 
@@
 
-  print(kable(td, caption = paste("Duration table (mins) for ", city, "( ", count_people, " ) per person (people with trips)")))
+  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (people with trips)")))
   cat("\n")

### 2020-03-03 summary_tables_outcomes.Rmd
Fix scaling of daily travel - not weekly, by multiplying it by 365
Add injury risk per billion kms

@@
 
-    td4[, 2] <- round((td4[,3] / ( td4[,2] * 52)) * 100000, 4)
+    td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 100000, 4)
     # / sum(io[[city]]$demographic$population, na.rm = T)
@@
 print(kable(format(td, scientific = F), caption = paste('Injury per 100k across all cities')))
+
+```
+
+### Change in deaths due to injury per Billion km across cities
+
+```{r inj_100k_all_cities = "asis"}
+require(tibble)
+
+injury_risks_b <- td
+
+# Convert it into Billion by multiping the values by 10000
+injury_risks_b[, 2:ncol(injury_risks_b)] <- injury_risks_b[, 2:ncol(injury_risks_b)] * 10000
+
+readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')
+
+colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))
+
+colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))
+
+colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))
+
+colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))
+
+colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))
+
+injury_risks_lng <- reshape2::melt(injury_risks_b)
+
+col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)
+
+injury_risks_lng <- cbind(injury_risks_lng, col_split)
+names(injury_risks_lng)[4] <- 'scenario'
+names(injury_risks_lng)[5] <- 'city'
+injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
+cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
+
+injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
+
+injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycle Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Bus Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycle Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking Scenario'
+
+d <- ggplot(injury_risks_lng) +
+  aes(x = stage_mode, fill = scenario, weight = value) +
+  geom_bar(position = "dodge", colour = 'black') +
+  scale_color_manual(values = cols) +
+  coord_flip() +
+  theme_minimal() +
+  facet_wrap(vars(city)) +
+  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km"))
+
+ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
+
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
+
+htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+
+print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
 

### 2020-03-04 summary_tables_outcomes.Rmd
Fix filename
Make all risks as 0 where unavailable
Fix legend order

@@
 
+td[is.na(td)] <- 0
+
 td <- td %>% dplyr::select(stage_mode, sort(names(.)))
 
-readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks.csv')
+readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks_per_100k_kms.csv')
 
@@
   scale_color_manual(values = cols) +
+  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
   coord_flip() +

### 2020-03-06 summary_tables_outcomes.Rmd
Add plots by mode for injury risk per billion kms

@@
 
+cols_city <- c("#e41a1c",
+"#377eb8",
+"#4daf4a",
+"#984ea3",
+"#ff7f00",
+"#ffff33",
+"#a65628",
+"#f781bf",
+"#999999")
+
+cols_cbrewer <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#c51b7d','#ff7f00','#cab2d6','#6a3d9a','#543005','#b15928')
+
 d <- ggplot(injury_risks_lng) +
@@
 
+temp <- injury_risks_lng %>% group_by(stage_mode, scenario) %>% summarise(mv = mean(value), sd = sd(value))
+# injury_risks_lng <- left_join(injury_risks_lng, td, by = c('stage_mode', 'scenario'))
+
 d <- ggplot(injury_risks_lng) +
@@
 
+df <- injury_risks_lng %>% filter(city != 'cape_town') %>% spread(value = value, key = city) 
+df[df == 0] <- NA
+df <- df %>% mutate(mv = rowMeans(select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
+df1 <- df %>% dplyr::filter(variable == 'mv')
+df <- df %>% dplyr::filter(variable != 'mv')
+
+d1 <- ggplot(df, aes(x = stage_mode, fill = variable, weight = value)) +
+   geom_bar(position = "dodge", colour = 'black') +
+   scale_fill_manual(values = cols_cbrewer) +
+   theme_minimal() + 
+   geom_point(data = df1, aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
+   facet_wrap(vars(scenario)) + 
+   scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 25)) +
+   labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
+   theme(axis.text.x = element_text(face="bold", color="#993333", size=8, angle=90))
+
 ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
@@
 htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+
+ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
+
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
+
+htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
 

### 2020-03-06 summary_tables_outcomes.Rmd
Add missing library tidyr

@@
 library(ggplot2)
+library(tidyr)
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')

### 2020-03-16 summary_tables_outcomes.Rmd
Add injury risks for selected modes

@@
 
+d2 <- ggplot(df %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
+    geom_bar(position = "dodge", colour = 'black') +
+    scale_fill_manual(values = cols_cbrewer) +
+    theme_minimal() + 
+    geom_point(data = df1 %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')) , aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
+    facet_grid(vars(scenario), scales='free_x') + 
+    scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
+    labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
+    theme(axis.text.x = element_text(face="bold", color="#993333", size=10))
+
 ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
@@
 print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
+
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
+
+htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
 

### 2020-03-16 summary_tables_outcomes.Rmd
Increse font size

--- 2020-03-06 summary_tables_outcomes.Rmd
@@
 
+d2 <- ggplot(df %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
+    geom_bar(position = "dodge", colour = 'black') +
+    scale_fill_manual(values = cols_cbrewer) +
+    theme_minimal() + 
+    geom_point(data = df1 %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')) , aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
+    facet_grid(vars(scenario), scales='free_x') + 
+    scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
+    labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
+    theme(axis.text.x = element_text(face="bold", color="#993333", size=14))
+
 ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)

### 2020-03-19 summary_tables_outcomes.Rmd
Comment out YLLs

@@
 
-### Change in YLL {#change_YLL}
+<!-- ### Change in YLL {#change_YLL} -->
 
-```{r scen_prop = "asis"}
-yll_totals <- lapply(io,function(x){
-  temp <- sapply(1:5,function(y){
-    xx <- x$outcomes$hb$ylls
-    xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)])
-    sapply(sort(unique(xx$age_cat)),function(z)
-      sum(xxx[xx$age_cat==z]))
-  })
-  colnames(temp) <- scen_names
-  temp
-})
-sum_and_round_and_print(yll_totals,"Change in YLL total in ")
-```
+<!-- ```{r scen_prop = "asis"} -->
+<!-- yll_totals <- lapply(io,function(x){ -->
+<!--   temp <- sapply(1:5,function(y){ -->
+<!--     xx <- x$outcomes$hb$ylls -->
+<!--     xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)]) -->
+<!--     sapply(sort(unique(xx$age_cat)),function(z) -->
+<!--       sum(xxx[xx$age_cat==z])) -->
+<!--   }) -->
+<!--   colnames(temp) <- scen_names -->
+<!--   temp -->
+<!-- }) -->
+<!-- sum_and_round_and_print(yll_totals,"Change in YLL total in ") -->
+<!-- ``` -->
 
-YLLs per 100,000 people by age group by scenario
+<!-- YLLs per 100,000 people by age group by scenario -->
 
-### Change in YLLs per 100,000 {#change_YLL_age_100k}
+<!-- ### Change in YLLs per 100,000 {#change_YLL_age_100k} -->
 
-```{r scen_prop = "asis"}
-yll_rates <- lapply(cities,function(x) rbind(yll_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(yll_rates) <- cities
-round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
-```
+<!-- ```{r scen_prop = "asis"} -->
+<!-- yll_rates <- lapply(cities,function(x) rbind(yll_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
+<!-- names(yll_rates) <- cities -->
+<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
+<!-- ``` -->
 
-YLLs per 100,000 people by gender by age group scenario
+<!-- YLLs per 100,000 people by gender by age group scenario -->
 
-### Change in YLLs per 100,000 {#change_YLL_gender_age_100k}
+<!-- ### Change in YLLs per 100,000 {#change_YLL_gender_age_100k} -->
 
-```{r scen_prop = "asis"}
-yll_totals <- lapply(io,function(x){
-  temp <- sapply(1:5,function(y){
-    xx <- x$outcomes$hb$ylls
-    rowSums(xx[,seq(2+y,ncol(xx),by=5)])
-  })
-  rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
-  colnames(temp) <- scen_names
-  temp
-})
+<!-- ```{r scen_prop = "asis"} -->
+<!-- yll_totals <- lapply(io,function(x){ -->
+<!--   temp <- sapply(1:5,function(y){ -->
+<!--     xx <- x$outcomes$hb$ylls -->
+<!--     rowSums(xx[,seq(2+y,ncol(xx),by=5)]) -->
+<!--   }) -->
+<!--   rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2])) -->
+<!--   colnames(temp) <- scen_names -->
+<!--   temp -->
+<!-- }) -->
 
-yll_rates <- lapply(cities,function(x) 
-  rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[2],'_',z[1])),rownames(yll_totals[[x]])),]/
-          t(repmat(io[[x]]$demographic$population,5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(yll_rates) <- cities
-round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
-```
+<!-- yll_rates <- lapply(cities,function(x)  -->
+<!--   rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[2],'_',z[1])),rownames(yll_totals[[x]])),]/ -->
+<!--           t(repmat(io[[x]]$demographic$population,5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
+<!-- names(yll_rates) <- cities -->
+<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
+<!-- ```  -->
 
@@
 
-### Change in YLLs due to injury {#change_yll_injury}
+<!-- ### Change in YLLs due to injury {#change_yll_injury} -->
 
-```{r}
-injury_totals <- lapply(io,function(x){
-  xx <- x$outcomes$hb$ylls
-  xxx <- sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y])))
-  colnames(xxx) <- scen_names
-  xxx
-})
-sum_and_round_and_print(injury_totals,"Change in YLLs due to injury in ")
-```
+<!-- ```{r} -->
+<!-- injury_totals <- lapply(io,function(x){ -->
+<!--   xx <- x$outcomes$hb$ylls -->
+<!--   xxx <- sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y]))) -->
+<!--   colnames(xxx) <- scen_names -->
+<!--   xxx -->
+<!-- }) -->
+<!-- sum_and_round_and_print(injury_totals,"Change in YLLs due to injury in ") -->
+<!-- ``` -->
 
-### Change in YLLs due to PA {#change_yll_PA}
-```{r}
-pa_totals <- lapply(io,function(x){
-  xx <- x$outcomes$pathway_hb$ylls
-  xxx <- sapply(1:5,function(y){
-    xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]])
-    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
-  })
-  colnames(xxx) <- scen_names
-  xxx
-})
-sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
-```
+<!-- ### Change in YLLs due to PA {#change_yll_PA} -->
+<!-- ```{r} -->
+<!-- pa_totals <- lapply(io,function(x){ -->
+<!--   xx <- x$outcomes$pathway_hb$ylls -->
+<!--   xxx <- sapply(1:5,function(y){ -->
+<!--     xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]]) -->
+<!--     sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z])) -->
+<!--   }) -->
+<!--   colnames(xxx) <- scen_names -->
+<!--   xxx -->
+<!-- }) -->
+<!-- sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ") -->
+<!-- ``` -->
 
-### Change in YLLs due to AP {#change_yll_AP}
-```{r}
-ap_totals <- lapply(io,function(x){
-  xx <- x$outcomes$pathway_hb$ylls
-  xxx <- sapply(1:5,function(y){
-    xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]])
-    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
-  })
-  colnames(xxx) <- scen_names
-  xxx
-})
-sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ")
-```
+<!-- ### Change in YLLs due to AP {#change_yll_AP} -->
+<!-- ```{r} -->
+<!-- ap_totals <- lapply(io,function(x){ -->
+<!--   xx <- x$outcomes$pathway_hb$ylls -->
+<!--   xxx <- sapply(1:5,function(y){ -->
+<!--     xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]]) -->
+<!--     sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z])) -->
+<!--   }) -->
+<!--   colnames(xxx) <- scen_names -->
+<!--   xxx -->
+<!-- }) -->
+<!-- sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ") -->
+<!-- ``` -->
 
-YLLs per 100,000 people by age group by scenario
+<!-- YLLs per 100,000 people by age group by scenario -->
 
-### Change in YLLs due to injury per 100,000 {#change_yll_injury_100k}
+<!-- ### Change in YLLs due to injury per 100,000 {#change_yll_injury_100k} -->
 
-```{r scen_prop = "asis"}
-injury_rates <- lapply(cities,function(x) rbind(injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(injury_rates) <- cities
-round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ")
-```
+<!-- ```{r scen_prop = "asis"} -->
+<!-- injury_rates <- lapply(cities,function(x) rbind(injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
+<!-- names(injury_rates) <- cities -->
+<!-- round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ") -->
+<!-- ``` -->
 
-### Change in YLLs due to PA per 100,000 {#change_yll_PA_100k}
-```{r}
-pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(pa_rates) <- cities
-round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ")
-```
+<!-- ### Change in YLLs due to PA per 100,000 {#change_yll_PA_100k} -->
+<!-- ```{r} -->
+<!-- pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
+<!-- names(pa_rates) <- cities -->
+<!-- round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ") -->
+<!-- ``` -->
 
-### Change in YLLs due to injury AP 100,000 {#change_yll_AP_100k}
-```{r}
-ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(ap_rates) <- cities
-round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ")
-```
+<!-- ### Change in YLLs due to injury AP 100,000 {#change_yll_AP_100k} -->
+<!-- ```{r} -->
+<!-- ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
+<!-- names(ap_rates) <- cities -->
+<!-- round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ") -->
+<!-- ``` -->
 
-## BY DISEASE
+<!-- ## BY DISEASE -->
 
-Change in deaths total (for the city based on real population size) by scenario
+<!-- Change in deaths total (for the city based on real population size) by scenario -->
 
-### Change in deaths due to disease {#change_death_disease}
+<!-- ### Change in deaths due to disease {#change_death_disease} -->
 
-```{r}
+<!-- ```{r} -->
 
-disease_totals <- lapply(io,function(x){
-  xx <- x$outcomes$hb$deaths
-  xxx <- sapply(1:5,function(y){
-    colSums(xx[,seq(y+2,ncol(xx),by=5)])
-  })
-  colnames(xxx) <- scen_names
-  rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y))
-  xxx
-})
-sum_and_round_and_print(disease_totals,"Change in deaths due to disease in ")
+<!-- disease_totals <- lapply(io,function(x){ -->
+<!--   xx <- x$outcomes$hb$deaths -->
+<!--   xxx <- sapply(1:5,function(y){ -->
+<!--     colSums(xx[,seq(y+2,ncol(xx),by=5)]) -->
+<!--   }) -->
+<!--   colnames(xxx) <- scen_names -->
+<!--   rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y)) -->
+<!--   xxx -->
+<!-- }) -->
+<!-- sum_and_round_and_print(disease_totals,"Change in deaths due to disease in ") -->
 
-```
+<!-- ``` -->
 
-Change in deaths per 100,000 people by scenario
+<!-- Change in deaths per 100,000 people by scenario -->
 
-### Change in deaths due to disease per 100,000 {#change_death_disease_100k}
+<!-- ### Change in deaths due to disease per 100,000 {#change_death_disease_100k} -->
 
-```{r scen_prop = "asis"}
-disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(disease_rates) <- cities
-round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ")
-```
+<!-- ```{r scen_prop = "asis"} -->
+<!-- disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
+<!-- names(disease_rates) <- cities -->
+<!-- round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ") -->
+<!-- ``` -->
 
-Change in YLL total (for the city based on real population size) by scenario
+<!-- Change in YLL total (for the city based on real population size) by scenario -->
 
-### Change in YLL due to disease {#change_yll_disease}
+<!-- ### Change in YLL due to disease {#change_yll_disease} -->
 
-```{r}
+<!-- ```{r} -->
 
-disease_totals <- lapply(io,function(x){
-  xx <- x$outcomes$hb$ylls
-  xxx <- sapply(1:5,function(y){
-    colSums(xx[,seq(y+2,ncol(xx),by=5)])
-  })
-  colnames(xxx) <- scen_names
-  rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y))
-  xxx
-})
-sum_and_round_and_print(disease_totals,"Change in YLL due to disease in ")
+<!-- disease_totals <- lapply(io,function(x){ -->
+<!--   xx <- x$outcomes$hb$ylls -->
+<!--   xxx <- sapply(1:5,function(y){ -->
+<!--     colSums(xx[,seq(y+2,ncol(xx),by=5)]) -->
+<!--   }) -->
+<!--   colnames(xxx) <- scen_names -->
+<!--   rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y)) -->
+<!--   xxx -->
+<!-- }) -->
+<!-- sum_and_round_and_print(disease_totals,"Change in YLL due to disease in ") -->
 
-```
+<!-- ``` -->
 
-### Change in YLL due to disease per 100,000 {#change_yll_disease_100k}
+<!-- ### Change in YLL due to disease per 100,000 {#change_yll_disease_100k} -->
 
-```{r scen_prop = "asis"}
-disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(disease_rates) <- cities
-round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ")
-```
+<!-- ```{r scen_prop = "asis"} -->
+<!-- disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
+<!-- names(disease_rates) <- cities -->
+<!-- round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ") -->
+<!-- ``` -->

### 2020-03-19 summary_tables_outcomes.Rmd
Add unit and comment out DT

@@
 
-## Who Hit Whom matrix
+## Who Hit Whom matrix (normalized to annual figures)
 
@@
 
-### DT - Who hit whom matrix for all case cities {#whw_all}
-Baseline
+<!-- ### DT - Who hit whom matrix for all case cities {#whw_all} -->
+<!-- Baseline -->
 
-```{r dt}
+<!-- ```{r dt} -->
 
-#for (cs in names(all_whw)){
-td <- NULL
+<!-- #for (cs in names(all_whw)){ -->
+<!-- td <- NULL -->
 
-cs <- 'Baseline'
+<!-- cs <- 'Baseline' -->
 
-var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+<!-- var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character() -->
 
-fname <- paste("whw_", var, ".csv")
+<!-- fname <- paste("whw_", var, ".csv") -->
 
-td <- st[[var]]
+<!-- td <- st[[var]] -->
 
-DT::datatable(td, 
-              extensions = 'Buttons',
-              escape = FALSE,
-              rownames = FALSE,
-              options = list(dom = 'Bfrtip',
-                             buttons = 
-                               list('colvis', list(
-                                 extend = 'collection',
-                                 buttons = list(list(extend='csv',
-                                                     filename = fname),
-                                                list(extend='excel',
-                                                     filename = fname)),
-                                 text = 'Download'
-                               )),
-                             scrollX = TRUE,
-                             pageLength = nrow(td),
-                             order=list(list(2,'desc')),
-                             columnDefs = list(list(visible=FALSE, targets=c(1:(ncol(td) - 1))))))
+<!-- DT::datatable(td,  -->
+<!--               extensions = 'Buttons', -->
+<!--               escape = FALSE, -->
+<!--               rownames = FALSE, -->
+<!--               options = list(dom = 'Bfrtip', -->
+<!--                              buttons =  -->
+<!--                                list('colvis', list( -->
+<!--                                  extend = 'collection', -->
+<!--                                  buttons = list(list(extend='csv', -->
+<!--                                                      filename = fname), -->
+<!--                                                 list(extend='excel', -->
+<!--                                                      filename = fname)), -->
+<!--                                  text = 'Download' -->
+<!--                                )), -->
+<!--                              scrollX = TRUE, -->
+<!--                              pageLength = nrow(td), -->
+<!--                              order=list(list(2,'desc')), -->
+<!--                              columnDefs = list(list(visible=FALSE, targets=c(1:(ncol(td) - 1)))))) -->
 
-# }
+<!-- # } -->
 
-```
+<!-- ``` -->
 
@@
 
-YLLs total (for the city based on real population size) by age group by scenario
+<!-- YLLs total (for the city based on real population size) by age group by scenario -->
 

### 2020-05-13 summary_tables_outcomes.Rmd
Add table of contents, with numbered sections

@@
     toc: true
+    toc_depth: 5
+    toc_float: true
+    number_sections: true
     theme: united
@@
 
-## Who Hit Whom matrix (normalized to annual figures)
+# Who Hit Whom matrix (normalized to annual figures)
 
@@
 
-### Who hit whom matrix {#whw}
+## Who hit whom matrix {#whw}
 
@@
 
-```{r whw_for_all_cs, fig.width=7, fig.height=4, echo=FALSE}
+```{r whw_for_all_cs, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
 
@@
 
+  qual_name <- paste(qualified_scen_name, scen)
   st[[cs]] <- format(td, scientific = F)
-
-  print(kable(format(td, scientific = F), caption = paste('Who Hit Whom (WHW) for all case cities for ', cs)))
+  cat('####', qual_name, '\n')
+  print(kable(format(td, scientific = F), caption = paste('Who Hit Whom (WHW) for all case cities for ', qual_name)))
+  cat('\n')
 
@@
 
-## TOTAL
+# TOTAL
 
@@
 
-### Change in deaths {#change_death}
+## Change in deaths {#change_death}
 
@@
 
-### Change in deaths per 100,000 {#change_death_100k}
+## Change in deaths per 100,000 {#change_death_100k}
 
@@
 
-## BY PATHWAY
+# BY PATHWAY
 
@@
 
-## Change in deaths per 100,000 people by age group by scenario
+# Change in deaths per 100,000 people by age group by scenario
 
-### Change in deaths due to injury per 100,000 {#change_death_injury_100k}
+## Change in deaths due to injury per 100,000 {#change_death_injury_100k}
 
@@
 
-### Change in deaths due to PA per 100,000 {#change_death_PA_100k}
+## Change in deaths due to PA per 100,000 {#change_death_PA_100k}
 ```{r}
@@
 
-### Change in deaths due to AP per 100,000 {#change_death_AP_100k}
+## Change in deaths due to AP per 100,000 {#change_death_AP_100k}
 ```{r}

### 2020-05-18 summary_tables_outcomes.Rmd
Qualified scen names

@@
     toc_depth: 5
-    toc_float: true
-    number_sections: true
-    theme: united
+    toc_float: false
 ---
@@
 
-## Who hit whom matrix {#whw}
+## Who hit whom matrix of each case study (by baseline + scenarios) {#whw}
 
@@
 
-      print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', var, ' matrix for ', city)))
+      if (var == 'bl'){
+        qualified_scen_name <- 'Baseline'
+        scen <- ''
+      }else if(var == "w_sc"){
+        qualified_scen_name <- 'Walking'
+      }else if(var == "bi_sc"){
+        qualified_scen_name <- 'Bicycle'
+      }else if(var == "car_sc"){
+        qualified_scen_name <- 'Car'
+      }else if(var == "mc_sc"){
+        qualified_scen_name <- 'Motorcyle'
+      }
+
+      print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', qualified_scen_name, ' scenario matrix for ', city)))
       cat("\n")

### 2020-07-06 summary_tables_outcomes.Rmd
Add time and date, and also update TOC

@@
 ---
-title: "Summary Tables"
+title: "Summary Outcome Tables"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:
@@
     toc_depth: 5
-    toc_float: false
+    toc_float: true
 ---
@@
 # Assumes that multi_city_script.R has been run till 
-cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota','cape_town', 'vizag')
+cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'bogota',
+             'cape_town', 'vizag')
 

### 2020-07-15 summary_tables.Rmd
Floating TOC

@@
     toc: true
-    theme: united
+    toc_depth: 5
+    toc_float: true
 ---
@@
 # Assumes that multi_city_script.R has been run till 
-cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'mexico_city','bogota', 'cape_town', 'vizag')
+cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'bogota',
+             'mexico_city', 'vizag')
 
@@
 |---|---|---|---|
-| walking | 96.7|66.2|2.0 |
-| bicycle | 0.6|0.3|0.4 |
-| car | 1.0|10.4|16.9 |
-| motorcycle | 0.2|0.9|2.7 |
-| bus | 0.0|18.6|69.4 |
+| walking | 98.1|72.3|2.9 |
+| bicycle | 0.5|0.3|0.5 |
+| car | 1.0|9.0|17.1 |
+| motorcycle | 0.0|0.0|0.0 |
+| bus | 0.0|15.0|64.4 |
 
@@
 |---|---|---|---|
-| walking | 65.4|6.9|2.9 |
+| walking | 62.2|7.2|3.0 |
 | bicycle | 0.7|0.9|0.3 |
-| car | 22.2|41.9|27.2 |
-| motorcycle | 1.3|2.7|3.5 |
-| bus | 9.7|43.0|35.8 |
+| car | 21.0|42.9|28.7 |
+| motorcycle | 1.3|2.9|3.8 |
+| bus | 13.3|39.2|32.7 |
 
@@
 |---|---|---|---|
-| walking | 84.3|11.1|0.8 |
-| bicycle | 1.9|7.9|3.6 |
-| car | 1.9|14.7|19.6 |
-| motorcycle | 5.6|31.9|25.3 |
-| bus | 0.7|19.7|35.5 |
+| walking | 84.4|10.1|0.9 |
+| bicycle | 2.0|7.6|3.7 |
+| car | 2.4|17.3|21.4 |
+| motorcycle | 5.6|30.1|26.5 |
+| bus | 0.7|24.7|32.7 |
 
@@
 |---|---|---|---|
-| walking | 89.7|28.6|1.2 |
-| bicycle | 2.6|10.9|0.9 |
-| car | 0.8|5.0|4.9 |
-| motorcycle | 4.8|37.3|48.2 |
-| bus | 0.6|8.3|29.3 |
+| walking | 74.9|23.0|1.0 |
+| bicycle | 2.3|9.8|1.1 |
+| car | 1.0|4.4|6.4 |
+| motorcycle | 5.1|33.9|63.2 |
+| bus | 15.1|24.3|25.0 |
 
@@
 |---|---|---|---|
-| walking | 92.5|29.7|0.3 |
-| bicycle | 1.5|10.5|2.5 |
-| car | 5.4|42.3|27.4 |
-| motorcycle | 0.0|0.6|0.7 |
-| bus | 0.0|3.1|36.0 |
+| walking | 88.4|23.2|0.3 |
+| bicycle | 3.5|8.7|2.3 |
+| car | 7.2|45.5|35.5 |
+| motorcycle | 0.1|1.4|1.4 |
+| bus | 0.3|10.1|29.6 |
 
@@
 |---|---|---|---|
-| walking | 91.7|36.0|0.6 |
-| bicycle | 0.4|3.6|1.0 |
-| car | 7.2|39.9|38.9 |
-| motorcycle | 0.2|6.6|7.3 |
-| bus | 0.0|0.9|39.4 |
+| walking | 91.7|34.7|0.6 |
+| bicycle | 0.4|3.5|1.0 |
+| car | 7.2|38.4|39.5 |
+| motorcycle | 0.2|6.3|7.4 |
+| bus | 0.1|4.5|38.4 |
 
@@
 |---|---|---|---|
-| walking | 95.9|32.9|0.2 |
-| bicycle | 3.0|10.1|1.5 |
-| car | 0.5|24.1|25.4 |
-| motorcycle | 0.0|2.2|1.7 |
-| bus | 0.6|27.4|49.9 |
+| walking | 96.5|33.8|0.2 |
+| bicycle | 2.4|8.3|1.2 |
+| car | 0.5|24.8|25.7 |
+| motorcycle | 0.0|1.8|1.5 |
+| bus | 0.5|27.7|50.1 |
 
@@
 |---|---|---|---|
-| walking | 95.4|38.1|0.3 |
-| bicycle | 0.8|7.7|1.3 |
-| car | 2.3|23.9|28.6 |
-| motorcycle | 0.0|3.3|1.9 |
-| bus | 0.3|19.0|48.2 |
+| walking | 95.3|37.6|0.3 |
+| bicycle | 0.8|7.4|1.2 |
+| car | 2.4|24.9|29.6 |
+| motorcycle | 0.0|3.1|1.9 |
+| bus | 0.3|18.8|47.1 |
 
@@
 |---|---|---|---|
-| walking | 81.4|19.6|0.1 |
-| bicycle | 3.9|7.6|1.9 |
-| car | 0.6|2.5|2.2 |
-| motorcycle | 9.6|25.4|22.9 |
-| bus | 0.1|21.7|56.1 |
+| walking | 80.6|18.3|0.1 |
+| bicycle | 3.9|7.2|2.0 |
+| car | 0.6|2.3|2.4 |
+| motorcycle | 9.5|23.8|25.0 |
+| bus | 1.2|27.5|56.0 |
 
@@
 
-trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>% mutate_if(is.factor, ~ as.numeric(as.character(.x))) %>% janitor::adorn_totals('row')
+trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>% mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>% janitor::adorn_totals('row')
 

### 2020-07-15 summary_tables_outcomes.Rmd
Create injury risks per billion kms for each scenarios
Combine whw matrices in a long format

@@
 cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'bogota',
-             'cape_town', 'vizag')
+             'mexico_city', 'vizag')
 
@@
 
+whw_list <- list()
+
 for (cs in names(all_whw)){
@@
   readr::write_csv(whw_lng, paste0('results/multi_city/whw_matrices/whw_lng_', cs, '.csv'))
+
+  whw_lng$scenario <- qualified_scen_name
+
+  if (length(whw_list) == 0)
+    whw_list <- whw_lng
+  else
+    whw_list <- rbind(whw_list, whw_lng)
 
@@
 df[df == 0] <- NA
-df <- df %>% mutate(mv = rowMeans(select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
+df <- df %>% mutate(mv = rowMeans(dplyr::select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
 df1 <- df %>% dplyr::filter(variable == 'mv')
 df <- df %>% dplyr::filter(variable != 'mv')
+
+df_rr <- df1 %>% dplyr::filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck'))
+df_rr <- tidyr::spread(data = df_rr, key = stage_mode, value = value)
+
+readr::write_csv(df_rr, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms_by_scen.csv')
 
@@
 
-d2 <- ggplot(df %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
+d2 <- ggplot(df %>% dplyr::filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
     geom_bar(position = "dodge", colour = 'black') +

### 2020-07-15 summary_tables_outcomes.Rmd
Remove unused code

@@
 ```
-
-<!-- YLLs total (for the city based on real population size) by age group by scenario -->
-
-<!-- ### Change in YLLs due to injury {#change_yll_injury} -->
-
-<!-- ```{r} -->
-<!-- injury_totals <- lapply(io,function(x){ -->
-<!--   xx <- x$outcomes$hb$ylls -->
-<!--   xxx <- sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y]))) -->
-<!--   colnames(xxx) <- scen_names -->
-<!--   xxx -->
-<!-- }) -->
-<!-- sum_and_round_and_print(injury_totals,"Change in YLLs due to injury in ") -->
-<!-- ``` -->
-
-<!-- ### Change in YLLs due to PA {#change_yll_PA} -->
-<!-- ```{r} -->
-<!-- pa_totals <- lapply(io,function(x){ -->
-<!--   xx <- x$outcomes$pathway_hb$ylls -->
-<!--   xxx <- sapply(1:5,function(y){ -->
-<!--     xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]]) -->
-<!--     sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z])) -->
-<!--   }) -->
-<!--   colnames(xxx) <- scen_names -->
-<!--   xxx -->
-<!-- }) -->
-<!-- sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ") -->
-<!-- ``` -->
-
-<!-- ### Change in YLLs due to AP {#change_yll_AP} -->
-<!-- ```{r} -->
-<!-- ap_totals <- lapply(io,function(x){ -->
-<!--   xx <- x$outcomes$pathway_hb$ylls -->
-<!--   xxx <- sapply(1:5,function(y){ -->
-<!--     xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]]) -->
-<!--     sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z])) -->
-<!--   }) -->
-<!--   colnames(xxx) <- scen_names -->
-<!--   xxx -->
-<!-- }) -->
-<!-- sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ") -->
-<!-- ``` -->
-
-<!-- YLLs per 100,000 people by age group by scenario -->
-
-<!-- ### Change in YLLs due to injury per 100,000 {#change_yll_injury_100k} -->
-
-<!-- ```{r scen_prop = "asis"} -->
-<!-- injury_rates <- lapply(cities,function(x) rbind(injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
-<!-- names(injury_rates) <- cities -->
-<!-- round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ") -->
-<!-- ``` -->
-
-<!-- ### Change in YLLs due to PA per 100,000 {#change_yll_PA_100k} -->
-<!-- ```{r} -->
-<!-- pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
-<!-- names(pa_rates) <- cities -->
-<!-- round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ") -->
-<!-- ``` -->
-
-<!-- ### Change in YLLs due to injury AP 100,000 {#change_yll_AP_100k} -->
-<!-- ```{r} -->
-<!-- ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
-<!-- names(ap_rates) <- cities -->
-<!-- round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ") -->
-<!-- ``` -->
-
-<!-- ## BY DISEASE -->
-
-<!-- Change in deaths total (for the city based on real population size) by scenario -->
-
-<!-- ### Change in deaths due to disease {#change_death_disease} -->
-
-<!-- ```{r} -->
-
-<!-- disease_totals <- lapply(io,function(x){ -->
-<!--   xx <- x$outcomes$hb$deaths -->
-<!--   xxx <- sapply(1:5,function(y){ -->
-<!--     colSums(xx[,seq(y+2,ncol(xx),by=5)]) -->
-<!--   }) -->
-<!--   colnames(xxx) <- scen_names -->
-<!--   rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y)) -->
-<!--   xxx -->
-<!-- }) -->
-<!-- sum_and_round_and_print(disease_totals,"Change in deaths due to disease in ") -->
-
-<!-- ``` -->
-
-<!-- Change in deaths per 100,000 people by scenario -->
-
-<!-- ### Change in deaths due to disease per 100,000 {#change_death_disease_100k} -->
-
-<!-- ```{r scen_prop = "asis"} -->
-<!-- disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
-<!-- names(disease_rates) <- cities -->
-<!-- round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ") -->
-<!-- ``` -->
-
-<!-- Change in YLL total (for the city based on real population size) by scenario -->
-
-<!-- ### Change in YLL due to disease {#change_yll_disease} -->
-
-<!-- ```{r} -->
-
-<!-- disease_totals <- lapply(io,function(x){ -->
-<!--   xx <- x$outcomes$hb$ylls -->
-<!--   xxx <- sapply(1:5,function(y){ -->
-<!--     colSums(xx[,seq(y+2,ncol(xx),by=5)]) -->
-<!--   }) -->
-<!--   colnames(xxx) <- scen_names -->
-<!--   rownames(xxx) <- sapply(rownames(xxx),function(y)gsub('scen1_','',y)) -->
-<!--   xxx -->
-<!-- }) -->
-<!-- sum_and_round_and_print(disease_totals,"Change in YLL due to disease in ") -->
-
-<!-- ``` -->
-
-<!-- ### Change in YLL due to disease per 100,000 {#change_yll_disease_100k} -->
-
-<!-- ```{r scen_prop = "asis"} -->
-<!-- disease_rates <- lapply(cities,function(x) rbind(disease_totals[[x]]/sum(pop_by_age[[x]])*100000, Total=colSums(disease_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
-<!-- names(disease_rates) <- cities -->
-<!-- round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ") -->
-<!-- ``` -->

### 2020-07-16 summary_tables_outcomes.Rmd
Comment out uncessary bits

@@
 
-### Who hit whom matrix for all case cities {#whw_all}
-
 ```{r whw_for_all_cs, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
@@
 
-  readr::write_csv(td, paste0('results/multi_city/whw_matrices/whw_', cs, '.csv'))
+  # readr::write_csv(td, paste0('results/multi_city/whw_matrices/whw_', cs, '.csv'))
 
@@
 
-  readr::write_csv(whw_lng, paste0('results/multi_city/whw_matrices/whw_lng_', cs, '.csv'))
+  # readr::write_csv(whw_lng, paste0('results/multi_city/whw_matrices/whw_lng_', cs, '.csv'))
 
@@
 
-  ggsave(f, file = paste0("results/multi_city/whw_matrices/plots/whw_", cs,".png"), limitsize = F, dpi = 300, scale = 1.5)
+  #ggsave(f, file = paste0("results/multi_city/whw_matrices/plots/whw_", cs,".png"), limitsize = F, dpi = 300, scale = 1.5)
 
@@
 
-  htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
+  #htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
   #ggsave(f, file=paste0("results/multi_city/whw_matrices/whw_", cs,".png"), width = 14, height = 10, units = "cm")
@@
   st[[cs]] <- format(td, scientific = F)
-  cat('####', qual_name, '\n')
-  print(kable(format(td, scientific = F), caption = paste('Who Hit Whom (WHW) for all case cities for ', qual_name)))
-  cat('\n')
+  #cat('####', qual_name, '\n')
+  #print(kable(format(td, scientific = F), caption = paste('Who Hit Whom (WHW) for all case cities for ', qual_name)))
+  #cat('\n')
 
@@
 
-```
-
-<!-- ### DT - Who hit whom matrix for all case cities {#whw_all} -->
-<!-- Baseline -->
-
-<!-- ```{r dt} -->
-
-<!-- #for (cs in names(all_whw)){ -->
-<!-- td <- NULL -->
-
-<!-- cs <- 'Baseline' -->
-
-<!-- var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character() -->
-
-<!-- fname <- paste("whw_", var, ".csv") -->
-
-<!-- td <- st[[var]] -->
-
-<!-- DT::datatable(td, -->
-<!--               extensions = 'Buttons', -->
-<!--               escape = FALSE, -->
-<!--               rownames = FALSE, -->
-<!--               options = list(dom = 'Bfrtip', -->
-<!--                              buttons = -->
-<!--                                list('colvis', list( -->
-<!--                                  extend = 'collection', -->
-<!--                                  buttons = list(list(extend='csv', -->
-<!--                                                      filename = fname), -->
-<!--                                                 list(extend='excel', -->
-<!--                                                      filename = fname)), -->
-<!--                                  text = 'Download' -->
-<!--                                )), -->
-<!--                              scrollX = TRUE, -->
-<!--                              pageLength = nrow(td), -->
-<!--                              order=list(list(2,'desc')), -->
-<!--                              columnDefs = list(list(visible=FALSE, targets=c(1:(ncol(td) - 1)))))) -->
-
-<!-- # } -->
-
-<!-- ``` -->
-
-# TOTAL
-
-Change in deaths total (for the city based on real population size) by age group by scenario
-
-## Change in deaths {#change_death}
-
-```{r scen_prop = "asis"}
-death_totals <- lapply(io,function(x){
-  temp <- sapply(1:5,function(y){
-    xx <- x$outcomes$hb$deaths
-    xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)])
-    sapply(sort(unique(xx$age_cat)),function(z)
-      sum(xxx[xx$age_cat==z]))
-  })
-  #rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2]))
-  colnames(temp) <- scen_names
-  temp
-})
-
-sum_and_round_and_print(death_totals,"Change in deaths total in ")
-```
-
-Change in deaths per 100,000 people by age group by scenario
-
-## Change in deaths per 100,000 {#change_death_100k}
-
-```{r scen_prop = "asis", echo = F, message = F}
-death_rates <- lapply(cities,function(x) rbind(death_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(death_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(death_rates) <- cities
-
-round_and_print(death_rates,"Change in deaths per 100,000 in ")
-```
-
-YLLs total (for the city based on real population size) by age group  by scenario
-
-<!-- ### Change in YLL {#change_YLL} -->
-
-<!-- ```{r scen_prop = "asis"} -->
-<!-- yll_totals <- lapply(io,function(x){ -->
-<!--   temp <- sapply(1:5,function(y){ -->
-<!--     xx <- x$outcomes$hb$ylls -->
-<!--     xxx <- rowSums(xx[,seq(2+y,ncol(xx),by=5)]) -->
-<!--     sapply(sort(unique(xx$age_cat)),function(z) -->
-<!--       sum(xxx[xx$age_cat==z])) -->
-<!--   }) -->
-<!--   colnames(temp) <- scen_names -->
-<!--   temp -->
-<!-- }) -->
-<!-- sum_and_round_and_print(yll_totals,"Change in YLL total in ") -->
-<!-- ``` -->
-
-<!-- YLLs per 100,000 people by age group by scenario -->
-
-<!-- ### Change in YLLs per 100,000 {#change_YLL_age_100k} -->
-
-<!-- ```{r scen_prop = "asis"} -->
-<!-- yll_rates <- lapply(cities,function(x) rbind(yll_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
-<!-- names(yll_rates) <- cities -->
-<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
-<!-- ``` -->
-
-<!-- YLLs per 100,000 people by gender by age group scenario -->
-
-<!-- ### Change in YLLs per 100,000 {#change_YLL_gender_age_100k} -->
-
-<!-- ```{r scen_prop = "asis"} -->
-<!-- yll_totals <- lapply(io,function(x){ -->
-<!--   temp <- sapply(1:5,function(y){ -->
-<!--     xx <- x$outcomes$hb$ylls -->
-<!--     rowSums(xx[,seq(2+y,ncol(xx),by=5)]) -->
-<!--   }) -->
-<!--   rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2],1,function(z)paste0(z[1],'_',z[2])) -->
-<!--   colnames(temp) <- scen_names -->
-<!--   temp -->
-<!-- }) -->
-
-<!-- yll_rates <- lapply(cities,function(x)  -->
-<!--   rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')],1,function(z)paste0(z[2],'_',z[1])),rownames(yll_totals[[x]])),]/ -->
-<!--           t(repmat(io[[x]]$demographic$population,5,1))*100000, Total=colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000)) -->
-<!-- names(yll_rates) <- cities -->
-<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
-<!-- ```  -->
-
-# BY PATHWAY
-
-By pathway (with non- injury separate & summed)
-
-Change in deaths total (for the city based on real population size) by age group by scenario
-
-### Change in deaths due to injury {#change_death_injury}
-
-```{r}
-injury_totals <- lapply(io,function(x){
-  xx <- x$outcomes$hb$deaths
-  xxx <- sapply(1:5,function(y)sapply(sort(unique(xx$age_cat)),function(z)sum(xx[xx$age_cat==z,injury_col-1+y])))
-  colnames(xxx) <- scen_names
-  xxx
-})
-sum_and_round_and_print(injury_totals,"Change in deaths due to injury in ")
-```
-
-### Change in deaths due to injury per 100,000 km
-
-```{r inj_100k = "asis", echo = F, message = F}
-require(tibble)
-
-overall_el <- list()
-
-for (city in cities){
-
-  #city <- 'delhi'
-
-  print(city)
-
-  el <- list()
-
-  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
-
-  for (cs in names(io[[city]]$outcomes$whw)){
-    # cs <- 'Baseline'
-    # cs <- "Scenario 4"
-    # city <- 'buenos_aires'
-    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
-      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
-      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-
-    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
-      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
-
-      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-
-    }
-
-    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
-
-    if (length(el) == 0){
-      el <- td4 %>% dplyr::select(stage_mode)
-    }
-
-    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
-
-    var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
-
-    names(td4)[2] <- var
-
-    td5 <- td4
-
-    td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 100000, 4)
-    # / sum(io[[city]]$demographic$population, na.rm = T)
-
-    names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")
-
-    el <- inner_join(el, td4, by = 'stage_mode')
-
-    # for (i in 1:nrow(td4)){
-    #   td4[i, 2:(ncol(td4) - 2)] <- round((td4$total[i] /(td4[i, 2:6])) * 100000, 1)
-    #   
-    #   
-    # }
-
-    # print(td4)
-  }
-  print(kable(el, caption = city))
-
-  td <- el %>% dplyr::select(-contains('count'))
-  names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
-
-  overall_el[[city]] <- td
-
-  cat("\n")
-  #kable(print(el), caption = city)
-  #}
-
-}
-```
-
-### Change in deaths due to injury per 100,000 km across cities
-
-```{r inj_100k_all_cities = "asis"}
-require(tibble)
-
-td <- overall_el %>% purrr::reduce(full_join, by = "stage_mode") %>% as.data.frame()
-
-td[is.na(td)] <- 0
-
-td <- td %>% dplyr::select(stage_mode, sort(names(.)))
-
-readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks_per_100k_kms.csv')
-
-colnames(td) = gsub("bi_sc", "bisc", colnames(td))
-
-colnames(td) = gsub("bus_sc", "bssc", colnames(td))
-
-colnames(td) = gsub("car_sc", "csc", colnames(td))
-
-colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
-
-colnames(td) = gsub("w_sc", "wsc", colnames(td))
-
-injury_risks_lng <- reshape2::melt(td)
-
-col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)
-
-injury_risks_lng <- cbind(injury_risks_lng, col_split)
-names(injury_risks_lng)[4] <- 'scenario'
-names(injury_risks_lng)[5] <- 'city'
-injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
-cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
-
-injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
-
-injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
-injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycle Scenario'
-injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Bus Scenario'
-injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car Scenario'
-injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycle Scenario'
-injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking Scenario'
-
-cols_city <- c("#e41a1c",
-"#377eb8",
-"#4daf4a",
-"#984ea3",
-"#ff7f00",
-"#ffff33",
-"#a65628",
-"#f781bf",
-"#999999")
-
-cols_cbrewer <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#c51b7d','#ff7f00','#cab2d6','#6a3d9a','#543005','#b15928')
-
-d <- ggplot(injury_risks_lng) +
-  aes(x = stage_mode, fill = scenario, weight = value) +
-  geom_bar(position = "dodge", colour = 'black') +
-  scale_color_manual(values = cols) +
-  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
-  coord_flip() +
-  theme_minimal() +
-  facet_wrap(vars(city)) +
-  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per 100k km"))
-
-ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_100k_km.png"), limitsize = F, dpi = 300, scale = 1.5)
-
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_100k_km.html")
-
-htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
-
-print(kable(format(td, scientific = F), caption = paste('Injury per 100k across all cities')))
+readr::write_csv(whw_list, paste0('results/multi_city/whw_matrices/whw_lng.csv'))
 
 ```
-
-### Change in deaths due to injury per Billion km across cities
-
-```{r inj_100k_all_cities = "asis"}
-require(tibble)
-
-injury_risks_b <- td
-
-# Convert it into Billion by multiping the values by 10000
-injury_risks_b[, 2:ncol(injury_risks_b)] <- injury_risks_b[, 2:ncol(injury_risks_b)] * 10000
-
-readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')
-
-colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))
-
-colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))
-
-colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))
-
-colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))
-
-colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))
-
-injury_risks_lng <- reshape2::melt(injury_risks_b)
-
-col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)
-
-injury_risks_lng <- cbind(injury_risks_lng, col_split)
-names(injury_risks_lng)[4] <- 'scenario'
-names(injury_risks_lng)[5] <- 'city'
-injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
-cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
-
-injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
-
-injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
-injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycle Scenario'
-injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Bus Scenario'
-injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car Scenario'
-injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycle Scenario'
-injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking Scenario'
-
-temp <- injury_risks_lng %>% group_by(stage_mode, scenario) %>% summarise(mv = mean(value), sd = sd(value))
-# injury_risks_lng <- left_join(injury_risks_lng, td, by = c('stage_mode', 'scenario'))
-
-d <- ggplot(injury_risks_lng) +
-  aes(x = stage_mode, fill = scenario, weight = value) +
-  geom_bar(position = "dodge", colour = 'black') +
-  scale_color_manual(values = cols) +
-  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
-  coord_flip() +
-  theme_minimal() +
-  facet_wrap(vars(city)) +
-  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km"))
-
-df <- injury_risks_lng %>% filter(city != 'cape_town') %>% spread(value = value, key = city) 
-df[df == 0] <- NA
-df <- df %>% mutate(mv = rowMeans(dplyr::select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
-df1 <- df %>% dplyr::filter(variable == 'mv')
-df <- df %>% dplyr::filter(variable != 'mv')
-
-df_rr <- df1 %>% dplyr::filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck'))
-df_rr <- tidyr::spread(data = df_rr, key = stage_mode, value = value)
-
-readr::write_csv(df_rr, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms_by_scen.csv')
-
-d1 <- ggplot(df, aes(x = stage_mode, fill = variable, weight = value)) +
-   geom_bar(position = "dodge", colour = 'black') +
-   scale_fill_manual(values = cols_cbrewer) +
-   theme_minimal() + 
-   geom_point(data = df1, aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
-   facet_wrap(vars(scenario)) + 
-   scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 25)) +
-   labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
-   theme(axis.text.x = element_text(face="bold", color="#993333", size=8, angle=90))
-
-d2 <- ggplot(df %>% dplyr::filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
-    geom_bar(position = "dodge", colour = 'black') +
-    scale_fill_manual(values = cols_cbrewer) +
-    theme_minimal() + 
-    geom_point(data = df1 %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')) , aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
-    facet_grid(vars(scenario), scales='free_x') + 
-    scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
-    labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
-    theme(axis.text.x = element_text(face="bold", color="#993333", size=14))
-
-ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
-
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
-
-htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
-
-ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
-
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
-
-htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
-
-print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
-
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
-
-htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
-
-```
-
-### # of deaths due to injury per 100,000 people
-
-```{r inj_100k = "asis"}
-require(tibble)
-
-overall_el_normalized <- list()
-
-for (city in cities){
-
-  print(city)
-
-  el <- list()
-
-  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
-
-  for (cs in names(io[[city]]$outcomes$whw)){
-    # cs <- 'Baseline'
-    # # cs <- "Baseline"
-    # city <- 'accra'
-    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
-      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
-      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-
-    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
-      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
-
-      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-
-    }
-
-    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode)
-
-    if (length(el) == 0){
-      el <- td4 %>% dplyr::select(stage_mode)
-    }
-
-    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
-
-    var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
-
-    names(td4)[2] <- var
-
-    td5 <- td4
-
-    td4[, 2] <- round(td4[, 2] / sum(io[[city]]$demographic$population) * 100000, 2)
-
-    el <- inner_join(el, td4, by = 'stage_mode')
-  }
-
-  el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
-
-  print(kable(el, caption = city))
-  td <- el
-  names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
-
-  overall_el_normalized[[city]] <- td
-
-  cat("\n")
-  #kable(print(el), caption = city)
-  #}
-
-}
-```
-
-### Change in deaths due to PA {#change_death_PA}
-
-```{r}
-
-pa_totals <- lapply(io,function(x){
-  xx <- x$outcomes$pathway_hb$deaths
-  xxx <- sapply(1:5,function(y){
-    xxx <- rowSums(xx[,pa_cols[seq(y,length(pa_cols),by=5)]])
-    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
-  })
-  colnames(xxx) <- scen_names
-  xxx
-})
-sum_and_round_and_print(pa_totals,"Change in deaths due to PA in ")
-```
-
-### Change in deaths due to AP {#change_death_AP}
-
-```{r}
-ap_totals <- lapply(io,function(x){
-  xx <- x$outcomes$pathway_hb$deaths
-  xxx <- sapply(1:5,function(y){
-    xxx <- rowSums(xx[,ap_cols[seq(y,length(ap_cols),by=5)]])
-    sapply(sort(unique(xx$age_cat)),function(z)sum(xxx[xx$age_cat==z]))
-  })
-  colnames(xxx) <- scen_names
-  xxx
-})
-sum_and_round_and_print(ap_totals,"Change in deaths due to AP in ")
-```
-
-# Change in deaths per 100,000 people by age group by scenario
-
-## Change in deaths due to injury per 100,000 {#change_death_injury_100k}
-
-```{r scen_prop = "asis"}
-injury_rates <- lapply(cities,function(x) rbind(injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(injury_rates) <- cities
-round_and_print(injury_rates,"Change in deaths due to injury per 100,000 in ")
-```
-
-## Change in deaths due to PA per 100,000 {#change_death_PA_100k}
-```{r}
-pa_rates <- lapply(cities,function(x) rbind(pa_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(pa_rates) <- cities
-round_and_print(pa_rates,"Change in deaths due to PA per 100,000 in ")
-
-```
-
-## Change in deaths due to AP per 100,000 {#change_death_AP_100k}
-```{r}
-
-ap_rates <- lapply(cities,function(x) rbind(ap_totals[[x]]/t(repmat(pop_by_age[[x]],5,1))*100000, Total=colSums(ap_totals[[x]])/rep(sum(pop_by_age[[x]]),length=5)*100000))
-names(ap_rates) <- cities
-round_and_print(ap_rates,"Change in deaths due to AP per 100,000 in ")
-```

### 2020-07-16 summary_tables_outcomes.Rmd
Add two levels of WHW matrices

@@
 ```
+
+### Who hit whom matrix for all case cities {#whw_all}
+
+```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
+
+list_temp <- list()
+index <- 1
+
+scen_names <- unique(whw_list$scenario)
+for (i in 1:length(scen_names)){
+  sn <- scen_names[i]
+  cat('###', sn, '\n')
+
+  temp <- whw_list %>% filter(scenario == sn) %>% dplyr::select(cas_mode)
+
+  cas_mode <- unique(temp$cas_mode)
+
+  for (j in 1:length(cas_mode)){
+    # print(cat(scen_names[i], ' - ', cas_mode[j], ' - '))
+    cm <- cas_mode[j]
+    temp1 <- whw_list %>% filter(scenario == sn & cas_mode == cm) %>% spread(value = value, key = str_mode)
+
+    temp1 <- temp1 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))
+
+    list_temp[[index]] <- temp1
+
+    index <- index + 1
+    cat('####', cm, '\n')
+    print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))
+
+    cat('\n')
+
+  }
+
+}
+
+```
+
+[View from](https://github.com/ITHIM/ITHIM-R/blob/master/results/multi_city/whw_matrices/whw_lng.csv)
+
+<!-- ### DT - Who hit whom matrix for all case cities {#whw_all} -->
+<!-- Baseline -->
+
+<!-- ```{r dt} -->
+
+<!-- #for (cs in names(all_whw)){ -->
+<!-- td <- NULL -->
+
+<!-- cs <- 'Baseline' -->
+
+<!-- var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character() -->
+
+<!-- fname <- paste("whw_", var, ".csv") -->
+
+<!-- td <- st[[var]] -->
+
+<!-- DT::datatable(td, -->
+<!--               extensions = 'Buttons', -->
+<!--               escape = FALSE, -->
+<!--               rownames = FALSE, -->
+<!--               options = list(dom = 'Bfrtip', -->
+<!--                              buttons = -->
+<!--                                list('colvis', list( -->
+<!--                                  extend = 'collection', -->
+<!--                                  buttons = list(list(extend='csv', -->
+<!--                                                      filename = fname), -->
+<!--                                                 list(extend='excel', -->
+<!--                                                      filename = fname)), -->
+<!--                                  text = 'Download' -->
+<!--                                )), -->
+<!--                              scrollX = TRUE, -->
+<!--                              pageLength = nrow(td), -->
+<!--                              order=list(list(2,'desc')), -->
+<!--                              columnDefs = list(list(visible=FALSE, targets=c(1:(ncol(td) - 1)))))) -->
+
+<!-- # } -->
+
+<!-- ``` -->

### 2020-07-16 summary_tables_outcomes.Rmd
Add flextable for multi-line headers

@@
 library(purrr)
-library(DT)
+library(flextable)
 library(ggplot2)
@@
     index <- index + 1
+    cat('\n')
     cat('####', cm, '\n')
-    print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))
+    #print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))
+
+    ft <- flextable(temp1)
+    ft <- add_header_row(ft, values = c(' ', 'strike mode'), colwidths = c(3, ncol(temp1) - 3))
+    ft <- merge_v(ft, j = 1:3, part = 'body')
+
+    cat('\n')
+
+    cat(knit_print(ft))
 
@@
   }
+
+  cat('\n')
 

### 2020-07-29 summary_tables_outcomes.Rmd
Set city names from the ithim object

@@
              'mexico_city', 'vizag')
+
+cities <- names(io)[!names(io) %in% 'scen_prop']
 
@@
     ft <- merge_v(ft, j = 1:3, part = 'body')
+    #fix_border_issues(ft)
 
@@
 ```
+
+### Who hit whom matrix for all case cities as CSV {#whw_all_csv}
 

### 2020-07-30 summary_tables_outcomes.Rmd

@@
 cities <- names(io)[!names(io) %in% 'scen_prop']
-
-#cities <- c('accra','sao_paulo','buenos_aires')
 
@@
 scen_names <- rownames(scen_prop)
+# scen_names <- c('Walking', 'Bicycling', 'Driving', 'Public Transport', 'Motorcycling')
 
@@
         qualified_scen_name <- 'Motorcyle'
+      }else if(var == "bus_sc"){
+        qualified_scen_name <- 'Bus'
       }
@@
 
-  unk <- td %>% filter(mode %in% c('unknown', 'unk')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unknown')
-
-  unsp <- td %>% filter(mode %in% c('unspecified', 'listed_na', '?')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unspecified')
-
-  other <- unsp <- td %>% filter(mode %in% c('other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'other')
-
-  pu <- td %>% filter(mode %in% c('picku', 'pickup/light goods')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'pickup')
-
-  bus <- td %>% filter(mode %in% c('bus', 'bus_driver')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'bus')
-
-  td <- td %>% filter(!mode %in% c('bus', 'bus_driver', 'unknown', 'unk', 'unspecified', 'listed_na', '?', 
-                                   'other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle',
-                                   'picku', 'pickup/light goods'))
-
-  td <- plyr::rbind.fill(td, bus, unk, unsp, other, pu)
-
+  # unk <- td %>% filter(mode %in% c('unknown', 'unk')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unknown')
+  # 
+  # unsp <- td %>% filter(mode %in% c('unspecified', 'listed_na', '?')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unspecified')
+  # 
+  # other <- unsp <- td %>% filter(mode %in% c('other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'other')
+  # 
+  # pu <- td %>% filter(mode %in% c('picku', 'pickup/light goods')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'pickup')
+  # 
+  # bus <- td %>% filter(mode %in% c('bus', 'bus_driver')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'bus')
+  # 
+  # td <- td %>% filter(!mode %in% c('bus', 'bus_driver', 'unknown', 'unk', 'unspecified', 'listed_na', '?', 
+  #                                  'other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle',
+  #                                  'picku', 'pickup/light goods'))
+  # 
+  # td <- plyr::rbind.fill(td, bus, unk, unsp, other, pu)
+  # 
   td <- (td[!duplicated(td), ])

### 2020-07-30 summary_tables_outcomes.Rmd
Rename scenarios

@@
 
+      # scen_names <- c('Walking', 'Bicycling', 'Driving', 'Public Transport', 'Motorcycling')
+
       if (var == 'bl'){
@@
       }else if(var == "bi_sc"){
-        qualified_scen_name <- 'Bicycle'
+        qualified_scen_name <- 'Bicycling'
       }else if(var == "car_sc"){
-        qualified_scen_name <- 'Car'
+        qualified_scen_name <- 'Driving'
       }else if(var == "mc_sc"){
-        qualified_scen_name <- 'Motorcyle'
+        qualified_scen_name <- 'Motorcyling'
       }else if(var == "bus_sc"){
-        qualified_scen_name <- 'Bus'
+        qualified_scen_name <- 'Public Transport'
       }
@@
 
-  qualified_scen_name <- 'Bus'
+  # scen_names <- c('Walking', 'Bicycling', 'Driving', 'Public Transport', 'Motorcycling')
+
+  qualified_scen_name <- 'Public Transport'
   scen <- 'Scenario'
@@
   }else if(cs == "bi_sc"){
-    qualified_scen_name <- 'Bicycle'
+    qualified_scen_name <- 'Bicycling'
   }else if(cs == "car_sc"){
-    qualified_scen_name <- 'Car'
+    qualified_scen_name <- 'Driving'
   }else if(cs == "mc_sc"){
-    qualified_scen_name <- 'Motorcyle'
+    qualified_scen_name <- 'Motorcycling'
   }
@@
   sn <- scen_names[i]
-  cat('###', sn, '\n')
+  cat('###', sn, ' scenario', '\n')
 

### 2020-07-30 summary_tables_outcomes.Rmd
Convert whw table into flextable with rounding

@@
 
-      print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', qualified_scen_name, ' scenario matrix for ', city)))
+      temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))
+
+      ft <- flextable(temp)
+      cat(knit_print(ft))
+
+      #print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', qualified_scen_name, ' scenario matrix for ', city)))
       cat("\n")

### 2020-07-31 summary_tables_outcomes.Rmd
Add sections for cities and their WHW matrices
Add table headers

@@
 
-```{r whw}
+```{r whw, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
 
@@
 for(city in cities){
-  # if (!city %in% c('buenos_aires', 'mexico_city'))
   {
-    # city <- 'buenos_aires'
+    cat('\n')
+
+    cat('###', city, '\n')
+
     for (cs in names(io$accra$outcomes$whw)){
-      print(city)
 
@@
 
+      cat('\n')
+
+      cat('####', qualified_scen_name, ' WHW', '\n')
+
+      cat('\n')
+
       temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))
@@
       ft <- flextable(temp)
+      ft <- add_header_row(ft, values = c('Strike Mode', 'Casualty Mode'), colwidths = c(1, ncol(temp) - 1))
+
+      #ft <- autofit(ft)
       cat(knit_print(ft))
@@
 
-### Who hit whom matrix for all case cities {#whw_all}
+## Who hit whom matrix for all case cities {#whw_all}
 
@@
   sn <- scen_names[i]
-  cat('###', sn, ' scenario', '\n')
+  cat('##', sn, ' scenario', '\n')
 
@@
     cat('\n')
-    cat('####', cm, '\n')
+    cat('###', cm, '\n')
     #print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))

### 2020-07-31 summary_tables_outcomes.Rmd
Remove unused code

@@
 
-      #ft <- autofit(ft)
       cat(knit_print(ft))
 
-      #print(kable(format(td3, scientific = F), caption = paste('Who Hit Whom (WHW) ', qualified_scen_name, ' scenario matrix for ', city)))
       cat("\n")
@@
 
-  # unk <- td %>% filter(mode %in% c('unknown', 'unk')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unknown')
-  # 
-  # unsp <- td %>% filter(mode %in% c('unspecified', 'listed_na', '?')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'unspecified')
-  # 
-  # other <- unsp <- td %>% filter(mode %in% c('other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'other')
-  # 
-  # pu <- td %>% filter(mode %in% c('picku', 'pickup/light goods')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'pickup')
-  # 
-  # bus <- td %>% filter(mode %in% c('bus', 'bus_driver')) %>% summarise_if(is.numeric, sum) %>% mutate(mode = 'bus')
-  # 
-  # td <- td %>% filter(!mode %in% c('bus', 'bus_driver', 'unknown', 'unk', 'unspecified', 'listed_na', '?', 
-  #                                  'other', 'non.motor.vehicle', 'tract', 'railway.train.railway.vehicle',
-  #                                  'picku', 'pickup/light goods'))
-  # 
-  # td <- plyr::rbind.fill(td, bus, unk, unsp, other, pu)
-  # 
   td <- (td[!duplicated(td), ])
@@
   backup_td <- td
-
-  # readr::write_csv(td, paste0('results/multi_city/whw_matrices/whw_', cs, '.csv'))
 
@@
 
-  # scen_names <- c('Walking', 'Bicycling', 'Driving', 'Public Transport', 'Motorcycling')
-
   qualified_scen_name <- 'Public Transport'
@@
   }
-
-  # readr::write_csv(whw_lng, paste0('results/multi_city/whw_matrices/whw_lng_', cs, '.csv'))
 
@@
   st[[cs]] <- format(td, scientific = F)
-  #cat('####', qual_name, '\n')
-  #print(kable(format(td, scientific = F), caption = paste('Who Hit Whom (WHW) for all case cities for ', qual_name)))
-  #cat('\n')
 
@@
   for (j in 1:length(cas_mode)){
-    # print(cat(scen_names[i], ' - ', cas_mode[j], ' - '))
+
     cm <- cas_mode[j]
@@
     ft <- merge_v(ft, j = 1:3, part = 'body')
-    #fix_border_issues(ft)
 
@@
 [View from](https://github.com/ITHIM/ITHIM-R/blob/master/results/multi_city/whw_matrices/whw_lng.csv)
-
-<!-- ### DT - Who hit whom matrix for all case cities {#whw_all} -->
-<!-- Baseline -->
-
-<!-- ```{r dt} -->
-
-<!-- #for (cs in names(all_whw)){ -->
-<!-- td <- NULL -->
-
-<!-- cs <- 'Baseline' -->
-
-<!-- var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character() -->
-
-<!-- fname <- paste("whw_", var, ".csv") -->
-
-<!-- td <- st[[var]] -->
-
-<!-- DT::datatable(td, -->
-<!--               extensions = 'Buttons', -->
-<!--               escape = FALSE, -->
-<!--               rownames = FALSE, -->
-<!--               options = list(dom = 'Bfrtip', -->
-<!--                              buttons = -->
-<!--                                list('colvis', list( -->
-<!--                                  extend = 'collection', -->
-<!--                                  buttons = list(list(extend='csv', -->
-<!--                                                      filename = fname), -->
-<!--                                                 list(extend='excel', -->
-<!--                                                      filename = fname)), -->
-<!--                                  text = 'Download' -->
-<!--                                )), -->
-<!--                              scrollX = TRUE, -->
-<!--                              pageLength = nrow(td), -->
-<!--                              order=list(list(2,'desc')), -->
-<!--                              columnDefs = list(list(visible=FALSE, targets=c(1:(ncol(td) - 1)))))) -->
-
-<!-- # } -->
-
-<!-- ``` -->

### 2020-07-31 summary_tables_outcomes.Rmd
Fix TOC levels

@@
 ---
-title: "Summary Outcome Tables"
+title: "Injrury Outcomes - WHW matrices"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 # Assumes that multi_city_script.R has been run till 
-cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'bogota',
-            'mexico_city', 'vizag')
-
+# Get names of cities from the io object
 cities <- names(io)[!names(io) %in% 'scen_prop']
@@
   sn <- scen_names[i]
-  cat('##', sn, ' scenario', '\n')
+  cat('###', sn, ' scenario', '\n')
 
@@
     cat('\n')
-    cat('###', cm, '\n')
+    cat('####', cm, '\n')
     #print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))

### 2020-07-31 summary_tables_outcomes.Rmd
Reorder rows and columns according to the standardized modes

@@
 library(ggplot2)
-library(tidyr)
+library(tidyverse)
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')
+
+smodes <- read_csv('data/global/modes/standardized_modes.csv')
 
@@
 
+      td3 <- td3[,c(1, na.omit(match(smodes$exhaustive_list, colnames(td3))), ncol(td3))]
+
+      td3 <- td3[c(na.omit(match(smodes$exhaustive_list, td3$mode)), nrow(td3)),]
+
       temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))
@@
 
+    temp1 <- temp1[,c(1, 2, 3, na.omit(match(smodes$exhaustive_list, colnames(temp1))), ncol(temp1))]
+
     temp1 <- temp1 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))

### 2020-07-31 summary_tables_outcomes.Rmd
Capitalize title

@@
 
-    cat('###', city, '\n')
+    cat('###', stringr::str_to_title(stringr::str_replace(city, '_', ' ')), '\n')
 
@@
     cat('\n')
-    cat('####', cm, '\n')
+    cat('#### Casualty mode: ', stringr::str_to_title(cm), '\n')
     #print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))

### 2020-08-11 injury_tables.Rmd
Use first element of io to populate names of whw

--- 2020-07-31 summary_tables_outcomes.Rmd
@@
 scen_names_w_baseline <- c('bl', 'w_sc', 'bi_sc', 'car_sc', 'mc_sc', 'bus_sc')
-scen_lt <- data.frame("scen_name" = names(io$accra$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)
+scen_lt <- data.frame("scen_name" = names(io[[1]]$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)
 
@@
 
-    for (cs in names(io$accra$outcomes$whw)){
+    for (cs in names(io[[1]]$outcomes$whw)){
 

### 2020-08-12 injury_tables.Rmd
Make link for direct download
Fix section

@@
 
-### Who hit whom matrix for all case cities as CSV {#whw_all_csv}
+## Who hit whom matrix for all case cities as CSV {#whw_all_csv}
 
-[View from](https://github.com/ITHIM/ITHIM-R/blob/master/results/multi_city/whw_matrices/whw_lng.csv)
+[Download from](https://raw.githubusercontent.com/ITHIM/ITHIM-R/master/results/multi_city/whw_matrices/whw_lng.csv)

### 2020-08-17 injury_tables.Rmd
Remove totals from whw long form
Add injury risks

@@
 
-readr::write_csv(whw_list, paste0('results/multi_city/whw_matrices/whw_lng.csv'))
+# Remove total from whw long form, and write as csv
+readr::write_csv(whw_list %>% filter(str_mode != 'Total'), paste0('results/multi_city/whw_matrices/whw_lng.csv'))
 
@@
 [Download from](https://raw.githubusercontent.com/ITHIM/ITHIM-R/master/results/multi_city/whw_matrices/whw_lng.csv)
+
+### Change in deaths due to injury per 100,000 km
+
+```{r inj_100k = "asis", echo = F, message = F}
+require(tibble)
+overall_el <- list()
+for (city in cities){
+
+  #city <- 'delhi'
+
+  print(city)
+
+  el <- list()
+
+  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
+
+  for (cs in names(io[[city]]$outcomes$whw)){
+    # cs <- 'Baseline'
+    # cs <- "Scenario 4"
+    # city <- 'buenos_aires'
+    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+
+      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+    }
+
+    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
+
+    if (length(el) == 0){
+      el <- td4 %>% dplyr::select(stage_mode)
+    }
+
+    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+    var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+
+    names(td4)[2] <- var
+
+    td5 <- td4
+
+    td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 100000, 4)
+    # / sum(io[[city]]$demographic$population, na.rm = T)
+
+    names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")
+
+    el <- inner_join(el, td4, by = 'stage_mode')
+
+    # for (i in 1:nrow(td4)){
+    #   td4[i, 2:(ncol(td4) - 2)] <- round((td4$total[i] /(td4[i, 2:6])) * 100000, 1)
+    #   
+    #   
+    # }
+
+    # print(td4)
+  }
+  print(kable(el, caption = city))
+
+  td <- el %>% dplyr::select(-contains('count'))
+  names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+
+  overall_el[[city]] <- td
+
+  cat("\n")
+  #kable(print(el), caption = city)
+  #}
+
+}
+```
+
+### Change in deaths due to injury per 100,000 km across cities
+
+```{r inj_100k_all_cities = "asis"}
+require(tibble)
+td <- overall_el %>% purrr::reduce(full_join, by = "stage_mode") %>% as.data.frame()
+td[is.na(td)] <- 0
+td <- td %>% dplyr::select(stage_mode, sort(names(.)))
+#readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks_per_100k_kms.csv')
+colnames(td) = gsub("bi_sc", "bisc", colnames(td))
+colnames(td) = gsub("bus_sc", "bssc", colnames(td))
+colnames(td) = gsub("car_sc", "csc", colnames(td))
+colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
+colnames(td) = gsub("w_sc", "wsc", colnames(td))
+injury_risks_lng <- reshape2::melt(td)
+col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)
+injury_risks_lng <- cbind(injury_risks_lng, col_split)
+names(injury_risks_lng)[4] <- 'scenario'
+names(injury_risks_lng)[5] <- 'city'
+injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
+cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
+injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
+injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycle Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Bus Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycle Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking Scenario'
+cols_city <- c("#e41a1c",
+"#377eb8",
+"#4daf4a",
+"#984ea3",
+"#ff7f00",
+"#ffff33",
+"#a65628",
+"#f781bf",
+"#999999")
+cols_cbrewer <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#c51b7d','#ff7f00','#cab2d6','#6a3d9a','#543005','#b15928')
+d <- ggplot(injury_risks_lng) +
+  aes(x = stage_mode, fill = scenario, weight = value) +
+  geom_bar(position = "dodge", colour = 'black') +
+  scale_color_manual(values = cols) +
+  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
+  coord_flip() +
+  theme_minimal() +
+  facet_wrap(vars(city)) +
+  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per 100k km")) 
+# ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_100k_km.png"), limitsize = F, dpi = 300, scale = 1.5)
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_100k_km.html")
+#htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+print(kable(format(td, scientific = F), caption = paste('Injury per 100k across all cities')))
+```
+
+### Change in deaths due to injury per Billion km across cities
+
+```{r inj_100k_all_cities = "asis"}
+require(tibble)
+injury_risks_b <- td
+# Convert it into Billion by multiping the values by 10000
+injury_risks_b[, 2:ncol(injury_risks_b)] <- injury_risks_b[, 2:ncol(injury_risks_b)] * 10000
+readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')
+colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))
+colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))
+colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))
+colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))
+colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))
+injury_risks_lng <- reshape2::melt(injury_risks_b)
+col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)
+injury_risks_lng <- cbind(injury_risks_lng, col_split)
+names(injury_risks_lng)[4] <- 'scenario'
+names(injury_risks_lng)[5] <- 'city'
+injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
+cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
+injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
+injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycle Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Bus Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycle Scenario'
+injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking Scenario'
+temp <- injury_risks_lng %>% group_by(stage_mode, scenario) %>% summarise(mv = mean(value), sd = sd(value))
+# injury_risks_lng <- left_join(injury_risks_lng, td, by = c('stage_mode', 'scenario'))
+d <- ggplot(injury_risks_lng) +
+  aes(x = stage_mode, fill = scenario, weight = value) +
+  geom_bar(position = "dodge", colour = 'black') +
+  scale_color_manual(values = cols) +
+  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
+  coord_flip() +
+  theme_minimal() +
+  facet_wrap(vars(city)) +
+  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) 
+df <- injury_risks_lng %>% filter(city != 'cape_town') %>% spread(value = value, key = city) 
+df[df == 0] <- NA
+df <- df %>% mutate(mv = rowMeans(select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
+df1 <- df %>% dplyr::filter(variable == 'mv')
+df <- df %>% dplyr::filter(variable != 'mv')
+d1 <- ggplot(df, aes(x = stage_mode, fill = variable, weight = value)) +
+   geom_bar(position = "dodge", colour = 'black') +
+   scale_fill_manual(values = cols_cbrewer) +
+   theme_minimal() + 
+   geom_point(data = df1, aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
+   facet_wrap(vars(scenario)) + 
+   scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 25)) +
+   labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
+   theme(axis.text.x = element_text(face="bold", color="#993333", size=8, angle=90))
+d2 <- ggplot(df %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
+    geom_bar(position = "dodge", colour = 'black') +
+    scale_fill_manual(values = cols_cbrewer) +
+    theme_minimal() + 
+    geom_point(data = df1 %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')) , aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
+    facet_grid(vars(scenario), scales='free_x') + 
+    scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
+    labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
+    theme(axis.text.x = element_text(face="bold", color="#993333", size=14))
+ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
+#htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
+#htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
+print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
+#htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
+```
+
+### # of deaths due to injury per 100,000 people
+
+```{r inj_100k = "asis"}
+require(tibble)
+overall_el_normalized <- list()
+for (city in cities){
+
+  print(city)
+
+  el <- list()
+
+  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
+
+  for (cs in names(io[[city]]$outcomes$whw)){
+    # cs <- 'Baseline'
+    # # cs <- "Baseline"
+    # city <- 'accra'
+    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+
+      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+    }
+
+    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode)
+
+    if (length(el) == 0){
+      el <- td4 %>% dplyr::select(stage_mode)
+    }
+
+    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+    var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+
+    names(td4)[2] <- var
+
+    td5 <- td4
+
+    td4[, 2] <- round(td4[, 2] / sum(io[[city]]$demographic$population) * 100000, 2)
+
+    el <- inner_join(el, td4, by = 'stage_mode')
+  }
+
+  el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
+
+  print(kable(el, caption = city))
+  td <- el
+  names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+
+  overall_el_normalized[[city]] <- td
+
+  cat("\n")
+  #kable(print(el), caption = city)
+  #}
+
+}
+```

### 2020-08-18 injury_tables.Rmd

--- 2020-08-12 injury_tables.Rmd
@@
 [Download from](https://raw.githubusercontent.com/ITHIM/ITHIM-R/master/results/multi_city/whw_matrices/whw_lng.csv)
+
+### Change in deaths due to injury per billion km
+
+```{r inj_100k = "asis", echo = F, message = F}
+require(tibble)
+overall_el <- list()
+for (city in cities){
+
+  #city <- 'delhi'
+
+  print(city)
+
+  el <- list()
+
+  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
+
+  for (cs in names(io[[city]]$outcomes$whw)){
+    # cs <- 'Baseline'
+    # cs <- "Scenario 4"
+    # city <- 'buenos_aires'
+    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+
+      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+    }
+
+    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
+
+    if (length(el) == 0){
+      el <- td4 %>% dplyr::select(stage_mode)
+    }
+
+    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+    var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+
+    names(td4)[2] <- var
+
+    td5 <- td4
+
+    td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 
+                        1000000000, 4)
+
+    names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")
+
+    el <- inner_join(el, td4, by = 'stage_mode')
+
+  }
+  print(kable(el, caption = city))
+
+  td <- el %>% dplyr::select(-contains('count'))
+  names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+
+  overall_el[[city]] <- td
+
+  cat("\n")
+
+}
+```
+
+### Change in deaths due to injury per Billion km across cities
+
+```{r inj_100k_all_cities = "asis"}
+require(tibble)
+
+td <- overall_el %>% purrr::reduce(full_join, by = "stage_mode") %>% as.data.frame()
+td[is.na(td)] <- 0
+td <- td %>% dplyr::select(stage_mode, sort(names(.)))
+#readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks_per_100k_kms.csv')
+colnames(td) = gsub("bi_sc", "bisc", colnames(td))
+colnames(td) = gsub("bus_sc", "bssc", colnames(td))
+colnames(td) = gsub("car_sc", "csc", colnames(td))
+colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
+colnames(td) = gsub("w_sc", "wsc", colnames(td))
+
+injury_risks_b <- td
+# # Convert it into Billion by multiping the values by 10000
+# injury_risks_b[, 2:ncol(injury_risks_b)] <- injury_risks_b[, 2:ncol(injury_risks_b)] * 10000
+
+#readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')
+colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))
+colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))
+colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))
+colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))
+colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))
+injury_risks_lng <- reshape2::melt(injury_risks_b)
+col_split <- stringr::str_split(injury_risks_lng$variable, "_", simplify = TRUE, n = 2)
+injury_risks_lng <- cbind(injury_risks_lng, col_split)
+names(injury_risks_lng)[4] <- 'scenario'
+names(injury_risks_lng)[5] <- 'city'
+injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
+cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
+injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
+injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycling'
+injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Public Transport'
+injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Driving'
+injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcyling'
+injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
+
+rd <- rename(injury_risks_lng, mode = stage_mode)
+rd <- rd %>% filter(mode != 'Total')
+
+readr::write_csv(rd, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms_lng.csv')
+
+temp <- injury_risks_lng %>% group_by(stage_mode, scenario) %>% summarise(mv = mean(value), sd = sd(value))
+# injury_risks_lng <- left_join(injury_risks_lng, td, by = c('stage_mode', 'scenario'))
+
+cols_cbrewer <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#c51b7d','#ff7f00','#cab2d6','#6a3d9a','#543005','#b15928')
+
+d <- ggplot(injury_risks_lng) +
+  aes(x = stage_mode, fill = scenario, weight = value) +
+  geom_bar(position = "dodge", colour = 'black') +
+  scale_color_manual(values = cols) +
+  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
+  coord_flip() +
+  theme_minimal() +
+  facet_wrap(vars(city)) +
+  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) 
+df <- injury_risks_lng %>% spread(value = value, key = city) 
+df[df == 0] <- NA
+df <- df %>% mutate(mv = rowMeans(select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
+df1 <- df %>% dplyr::filter(variable == 'mv')
+df <- df %>% dplyr::filter(variable != 'mv')
+d1 <- ggplot(df, aes(x = stage_mode, fill = variable, weight = value)) +
+   geom_bar(position = "dodge", colour = 'black') +
+   scale_fill_manual(values = cols_cbrewer) +
+   theme_minimal() + 
+   geom_point(data = df1, aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
+   facet_wrap(vars(scenario)) + 
+   scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 25)) +
+   labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
+   theme(axis.text.x = element_text(face="bold", color="#993333", size=8, angle=90))
+d2 <- ggplot(df %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
+    geom_bar(position = "dodge", colour = 'black') +
+    scale_fill_manual(values = cols_cbrewer) +
+    theme_minimal() + 
+    geom_point(data = df1 %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')) , aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
+    facet_grid(vars(scenario), scales='free_x') + 
+    scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
+    labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
+    theme(axis.text.x = element_text(face="bold", color="#993333", size=14))
+# ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
+#htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+# ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
+#htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
+print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
+fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
+#htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
+```
+
+### # of deaths due to injury per 100,000 people
+
+```{r inj_100k = "asis"}
+require(tibble)
+overall_el_normalized <- list()
+for (city in cities){
+
+  print(city)
+
+  el <- list()
+
+  for (cs in names(io[[city]]$outcomes$whw)){
+    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+
+    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+    }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+
+      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+
+    }
+
+    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode)
+
+    if (length(el) == 0){
+      el <- td4 %>% dplyr::select(stage_mode)
+    }
+
+    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+    var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+
+    names(td4)[2] <- var
+
+    td5 <- td4
+
+    td4[, 2] <- round(td4[, 2] / sum(io[[city]]$demographic$population) * 100000, 2)
+
+    el <- inner_join(el, td4, by = 'stage_mode')
+  }
+
+  el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
+
+  print(kable(el, caption = city))
+  td <- el
+  names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+
+  overall_el_normalized[[city]] <- td
+
+  cat("\n")
+
+}
+
+td <- overall_el_normalized %>% purrr::reduce(full_join, by = "stage_mode") %>% as.data.frame()
+td[is.na(td)] <- 0
+td <- td %>% dplyr::select(stage_mode, sort(names(.)))
+
+td <- rename(td, mode = stage_mode)
+colnames(td) = gsub("bi_sc", "bisc", colnames(td))
+colnames(td) = gsub("bus_sc", "bssc", colnames(td))
+colnames(td) = gsub("car_sc", "csc", colnames(td))
+colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
+colnames(td) = gsub("w_sc", "wsc", colnames(td))
+injury_risks_per_100k <- reshape2::melt(td)
+col_split <- stringr::str_split(injury_risks_per_100k$variable, "_", simplify = TRUE, n = 2)
+injury_risks_per_100k <- cbind(injury_risks_per_100k, col_split)
+names(injury_risks_per_100k)[4] <- 'scenario'
+names(injury_risks_per_100k)[5] <- 'city'
+injury_risks_per_100k <- injury_risks_per_100k %>% dplyr::select(-variable)
+injury_risks_per_100k$scenario <- as.character(injury_risks_per_100k$scenario)
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'bisc',]$scenario <- 'Bicycling'
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'bssc',]$scenario <- 'Public Transport'
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Driving'
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
+
+readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/whw_matrices/injury_risks_per_100k_pop.csv')
+
+```

### 2020-08-18 injury_tables.Rmd
Use full join to union modes in WHW and NOV matrices

@@
       td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
       #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
@@
       td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td3 <- left_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T))
+      td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
 

### 2020-08-25 injury_tables.Rmd
Make qualified name of rickshaw

@@
 
+  if (nrow(whw_lng[whw_lng$cas_mode == 'ar',]) > 0){
+    whw_lng[whw_lng$cas_mode == 'ar',]$cas_mode <- "auto_rickshaw"
+  }
+
+  if (nrow(whw_lng[whw_lng$str_mode == 'ar',]) > 0){
+    whw_lng[whw_lng$str_mode == 'ar',]$str_mode <- "auto_rickshaw"
+  }
+
   whw_lng$str_mode <- factor(whw_lng$str_mode, levels = unique(whw_lng$str_mode))
@@
 
-  print(city)
+  # print(city)
 
@@
   for (cs in names(io[[city]]$outcomes$whw)){
-    # cs <- 'Baseline'
-    # cs <- "Scenario 4"
-    # city <- 'buenos_aires'
     if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
@@
 
-    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs)
+    td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs) %>% as.data.frame()
 
@@
     td5 <- td4
+
+    td4[, 2] <- as.numeric(td4[,2])
 

### 2020-08-25 injury_tables.Rmd
Fix typo

@@
       }else if(var == "mc_sc"){
-        qualified_scen_name <- 'Motorcyling'
+        qualified_scen_name <- 'Motorcycling'
       }else if(var == "bus_sc"){
@@
 injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Public Transport'
-injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Driving'
-injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcyling'
+injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario2 <- 'Driving'
+injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
 injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'

### 2020-08-26 injury_tables.Rmd
Expand specific section of markdown
Fix typo

@@
 
+```
+
+```{js echo=FALSE}
+window.location.href='#Who_Hit_Whom_matrix_(normalized_to_annual_figures)';
 ```
@@
 injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Public Transport'
-injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario2 <- 'Driving'
+injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Driving'
 injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'

### 2020-09-03 summary_tables_PA_AP.Rmd
Get city names from ithim object, and replace 'bicycle' with 'cycle'

@@
 
-cities <- c("accra", "sao_paulo", "delhi", "bangalore")
+# Assumes that multi_city_script.R has been run till 
+# Get names of cities from the io object
+cities <- names(io)[!names(io) %in% 'scen_prop']
 high<- c(70, 25, 150,60)
@@
 
-  summary<- as.data.frame(io[[cities[x]]]$outcomes$injuries %>% group_by(scenario) %>% summarise('Ped'=sum(pedestrian,na.rm=T),'Bike'= sum(bicycle), 'Car'=sum(car), "Motorcycle"=sum(motorcycle,na.rm=T), "All modes"=sum(Deaths,na.rm=T)))
+  summary<- as.data.frame(io[[cities[x]]]$outcomes$injuries %>% group_by(scenario) %>% summarise('Ped'=sum(pedestrian,na.rm=T),'Bike'= sum(cycle), 'Car'=sum(car), "Motorcycle"=sum(motorcycle,na.rm=T), "All modes"=sum(Deaths,na.rm=T)))
   summary$scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
@@
 
-  select<- c("car", "motorcycle", "walking", "bicycle", "bus")
+  select<- c("car", "motorcycle", "walking", "cycle", "bus")
 

### 2020-09-04 summary_tables_PA_AP.Rmd

@@
 author: "Rahul Goel, MRC Epidemiology Unit, University of Cambridge, UK"
-date: "25 April 2019"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:

### 2020-09-30 summary_tables.Rmd
Read names of case studies from ithim object

@@
 # Assumes that multi_city_script.R has been run till 
-cities <- c('accra','sao_paulo','delhi','bangalore', 'santiago', 'belo_horizonte', 'buenos_aires', 'bogota',
-             'mexico_city', 'vizag')
+# Get names of cities from the io object
+cities <- names(io)[!names(io) %in% 'scen_prop']
 

### 2020-09-30 summary_tables_PA_AP.Rmd
Read params from csv file and add CO2 emissions

--- 2019-05-16 summary_tables_PA_AP.Rmd
@@
 ---
-title: "Summary Tables ITHIMR"
-author: "Rahul Goel, MRC Epidemiology Unit, University of Cambridge, UK"
-date: "25 April 2019"
+title: "AP/PA Tables ITHIM Global"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:
@@
 These are the summary tables of the following items
+
 1) Individual-level PM2.5 concentrations for baseline and scenarios
-2) Individual-level physical activity for baseline and scenarios
-3) Distance mode shares for baseline and scenarios
+2) Individual-level CO2 for baseline and scenarios
+3) Individual-level physical activity for baseline and scenarios
 
@@
 
-cities <- c("accra", "sao_paulo", "delhi", "bangalore")
-high<- c(70, 25, 150,60)
-low<- c(40,15,100,40)
+# Assumes that multi_city_script.R has been run till 
+# Get names of cities from the io object
+cities <- names(io)[!names(io) %in% 'scen_prop']
+
+all_inputs <- read.csv('all_city_parameter_inputs.csv',stringsAsFactors = F)
+
+#all_inputs$cape_town <- all_inputs$accra
+#all_inputs$vizag <- all_inputs$sao_paulo
+
+parameter_names <- all_inputs$parameter
+parameter_starts <- which(parameter_names!='')
+parameter_stops <- c(parameter_starts[-1] - 1, nrow(all_inputs))
+parameter_names <- parameter_names[parameter_names!='']
+parameter_list <- list()
+compute_mode <- 'constant'
+for(i in 1:length(parameter_names)){
+  parameter_list[[parameter_names[i]]] <- list()
+  parameter_index <- which(all_inputs$parameter==parameter_names[i])
+  if(all_inputs[parameter_index,2]=='')  {
+    parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
+      city_index <- which(colnames(all_inputs)==x)
+      val <- all_inputs[parameter_index,city_index]
+      ifelse(val%in%c('T','F'),val,as.numeric(val))
+    })
+    names(parameter_list[[parameter_names[i]]]) <- cities
+  }else if(all_inputs[parameter_index,2]=='constant'){
+    indices <- 0
+    if(compute_mode=='sample') indices <- 1:2
+    parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
+      city_index <- which(colnames(all_inputs)==x)
+      val <- all_inputs[parameter_index+indices,city_index]
+      ifelse(val=='',0,as.numeric(val))
+    })
+    names(parameter_list[[parameter_names[i]]]) <- cities
+  }else{
+    parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
+      city_index <- which(colnames(all_inputs)==x)
+      if(any(all_inputs[parameter_starts[i]:parameter_stops[i],city_index]!='')){
+        sublist_indices <- which(all_inputs[parameter_starts[i]:parameter_stops[i],city_index]!='')
+        thing <- as.list(as.numeric(c(all_inputs[parameter_starts[i]:parameter_stops[i],city_index])[sublist_indices]))
+        names(thing) <- c(all_inputs[parameter_starts[i]:parameter_stops[i],2])[sublist_indices]
+        thing
+      }
+    }
+    )
+    names(parameter_list[[parameter_names[i]]]) <- cities
+  }
+}
+
+for(i in 1:length(parameter_list)) assign(names(parameter_list)[i],parameter_list[[i]])
+
+# high<- c(70, 25, 150,60)
+# low<- c(40,15,100,40)
 for (x in 1: length(cities))
 {
-
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, baseline:bus_scen, factor_key=TRUE)
-  y<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")+ylim(low[x],high[x])
+  # print(cities[x])
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, Baseline:`Public Transport`, factor_key=TRUE)
+  y<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")
+  #+ylim(low[x],high[x])
   print(y)
@@
 {
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, baseline:bus_scen, factor_key=TRUE)
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, Baseline:`Public Transport`, factor_key=TRUE)
   summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(pm_conc),'5th'=quantile(pm_conc, 0.05),'50th'=quantile(pm_conc, 0.5),'95th'=quantile(pm_conc, 0.9)))
@@
 
-# Descriptive tables of emission inventory
+# Descriptive tables of PM 2.5 emission inventory
 
@@
 
-trans_share<-c(22.5, 40.0, 22.5, 28.1) 
-pm_conc<- c(50, 18, 122, 47 )
+#trans_share<-c(22.5, 40.0, 22.5, 28.1) 
+#pm_conc<- c(50, 18, 122, 47 )
 city<-cities
-shares<- as.data.frame(cbind(city,as.numeric(trans_share), pm_conc))
-names(shares)[2]<- "trans_share"
+#shares<- as.data.frame(cbind(city,as.numeric(io), pm_conc))
+#names(shares)[2]<- "trans_share"
+
+sl <- list()
 for (x in 1:length(cities))
 {
-  modes<-names(unlist(io[[cities[x]]]$emission_inventory))
-  emissions<-as.data.frame(unlist(io[[cities[x]]]$emission_inventory))
-  city_emissions<-cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$emission_inventory)`))
+  modes<-names(unlist(io[[cities[x]]]$PM_emission_inventory))
+  emissions<-as.data.frame(unlist(io[[cities[x]]]$PM_emission_inventory))
+  city_emissions<-cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))
   names(city_emissions)[2]<- "emissions"
@@
   summary<-as.data.frame(summary)
-  summary[nrow(summary)+1,1]<-"Transport share"
-  summary[nrow(summary),2]<-as.character(shares$trans_share[x])
-  summary[nrow(summary)+1,1]<-"PM2.5 Conc"
-  summary[nrow(summary),2]<-as.character(shares$pm_conc[x])
+  summary[nrow(summary)+1,1] <- "Transport share"
+  summary[nrow(summary),2]<- as.character(round(pm_trans_share[[cities[x]]] * 100))
+  summary[nrow(summary)+1,1] <- "PM2.5 Conc"
+  summary[nrow(summary),2] <- as.character(pm_conc_base[[cities[x]]])
   io[[cities[x]]]$summary_emission<- summary
+
+  if (length(sl) == 0){
+    sl <- summary
+  }else{
+    sl <- left_join(sl , summary)
+  }
 
@@
 
-summary_all<- cbind(io[[cities[1]]]$summary_emission,io[[cities[2]]]$summary_emission,io[[cities[3]]]$summary_emission,io[[cities[4]]]$summary_emission)
-summary_all<- summary_all[-c(3,5,7)]
-summary_all$modes[2]<- "bus"
+# summary_all<- cbind(io[[cities[1]]]$summary_emission,io[[cities[2]]]$summary_emission,io[[cities[3]]]$summary_emission,io[[cities[4]]]$summary_emission)
+# summary_all<- summary_all[-c(3,5,7)]
+# summary_all$modes[2]<- "bus"
+# 
+# 
+print(kable(sl))
+```
 
-print(kable(summary_all))
+# Descriptive table for CO2 Emissions
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
+cl <- list()
+for (city in cities)
+{
+
+  td <- round(colSums(io[[city]]$outcomes$co2_conc, na.rm = T), 1) %>% as.data.frame() %>% tibble::rownames_to_column()
+  names(td) <- c('Scenario', city)
+
+  if (length(cl) == 0)
+    cl <- td
+  else
+    cl <- left_join(cl, td)
+
+}
+
+cl$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+
+print(kable(cl))
+
 ```
@@
 {
-  names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE)
+  names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Public Transport`, factor_key=TRUE)
   y<-ggplot(data_long, aes(x=scenario, y=mmet, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Weekly MMET", x = "Scenarios")+  ylim(0, limit)
@@
 
-# Descriptive tables of Individual-level Physical activity (MMETs)
+<!-- # Descriptive tables of Individual-level Physical activity (MMETs) -->
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1: length(cities))
-{
-  names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE)
-  summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(mmet),'5th'=quantile(mmet, 0.05),'50th'=quantile(mmet, 0.5),'95th'=quantile(mmet, 0.9)))
-  summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1))
-  names(summary)[1]<-"Scenario"
-  print(kable(summary, caption= cities[x]))
+<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
+<!-- for (x in 1: length(cities)) -->
+<!-- { -->
+<!--   names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen") -->
+<!--   data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE) -->
+<!--   summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(mmet),'5th'=quantile(mmet, 0.05),'50th'=quantile(mmet, 0.5),'95th'=quantile(mmet, 0.9))) -->
+<!--   summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1)) -->
+<!--   names(summary)[1]<-"Scenario" -->
+<!--   print(kable(summary, caption= cities[x])) -->
 
-}
-```
+<!-- } -->
+<!-- ``` -->
 
-# Descriptive tables of injury outcomes
+<!-- # Descriptive tables of injury outcomes -->
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1: length(cities))
-{
+<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
+<!-- for (x in 1: length(cities)) -->
+<!-- { -->
 
-  summary<- as.data.frame(io[[cities[x]]]$outcomes$injuries %>% group_by(scenario) %>% summarise('Ped'=sum(pedestrian,na.rm=T),'Bike'= sum(bicycle), 'Car'=sum(car), "Motorcycle"=sum(motorcycle,na.rm=T), "All modes"=sum(Deaths,na.rm=T)))
-  summary$scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  summary<- cbind(summary$scenario ,round(summary[,2:6], digits=0))
-  names(summary)[1]<-"Scenario"
-  print(kable(summary, caption= cities[x]))
+<!--   summary<- as.data.frame(io[[cities[x]]]$outcomes$injuries %>% group_by(scenario) %>% summarise('Ped'=sum(pedestrian,na.rm=T),'Bike'= sum(cycle), 'Car'=sum(car), "Motorcycle"=sum(motorcycle,na.rm=T), "All modes"=sum(Deaths,na.rm=T))) -->
+<!--   summary$scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen") -->
+<!--   summary<- cbind(summary$scenario ,round(summary[,2:6], digits=0)) -->
+<!--   names(summary)[1]<-"Scenario" -->
+<!--   print(kable(summary, caption= cities[x])) -->
 
-}
-```
+<!-- } -->
+<!-- ``` -->
 
-# Striking vehicles in scenarios
+<!-- # Striking vehicles in scenarios -->
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1: length(cities))
-{
+<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
+<!-- for (x in 1: length(cities)) -->
+<!-- { -->
 
-scenarios<- c('Baseline', 'Scenario 1', 'Scenario 2', 'Scenario 3', 'Scenario 4', 'Scenario 5')
-for (i in 1: length(scenarios))
-{
-whw<-as.data.frame(rowSums(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$whw))
-noov<-as.data.frame(sum(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$nov))
-names(whw)[1]<- as.character(scenarios[i])
-names(noov)[1]<-as.character(scenarios[i])
+<!-- scenarios<- c('Baseline', 'Scenario 1', 'Scenario 2', 'Scenario 3', 'Scenario 4', 'Scenario 5') -->
+<!-- for (i in 1: length(scenarios)) -->
+<!-- { -->
+<!-- whw<-as.data.frame(rowSums(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$whw)) -->
+<!-- noov<-as.data.frame(sum(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$nov)) -->
+<!-- names(whw)[1]<- as.character(scenarios[i]) -->
+<!-- names(noov)[1]<-as.character(scenarios[i]) -->
 
-if (i ==1 )
-{
- summary<-round(rbind(whw,nov=noov), digits = 1)
-} 
-else
-{
-summary<- cbind(summary, round(rbind(whw, nov=noov), digits=1) ) 
-}
+<!-- if (i ==1 ) -->
+<!-- { -->
+<!--  summary<-round(rbind(whw,nov=noov), digits = 1) -->
+<!-- }  -->
+<!-- else -->
+<!-- { -->
+<!-- summary<- cbind(summary, round(rbind(whw, nov=noov), digits=1) )  -->
+<!-- } -->
 
-}
+<!-- } -->
 
-names(summary)<-c("base", "walk", "bike", "car", "MC", "bus")
-print(kable(summary, caption= cities[x]))
+<!-- names(summary)<-c("base", "walk", "bike", "car", "MC", "bus") -->
+<!-- print(kable(summary, caption= cities[x])) -->
 
-}
+<!-- } -->
 
-```
+<!-- ``` -->
 
-# Distance share by scenarios in percentages (excludes city-specific minority modes-- auto rickshaws, subway, etc.)
+<!-- # Distance share by scenarios in percentages (excludes city-specific minority modes-- auto rickshaws, subway, etc.) -->
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1: length(cities))
-{
+<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
+<!-- for (x in 1: length(cities)) -->
+<!-- { -->
 
-  select<- c("car", "motorcycle", "walking", "bicycle", "bus")
+<!--   select<- c("car", "motorcycle", "walking", "cycle", "bus") -->
 
-  io[[cities[x]]]$trip_scen_sets$trip_mode[!io[[cities[x]]]$trip_scen_sets$trip_mode %in% select]<- "other"
+<!--   io[[cities[x]]]$trip_scen_sets$trip_mode[!io[[cities[x]]]$trip_scen_sets$trip_mode %in% select]<- "other" -->
 
-  dist_scen<- io[[cities[x]]]$trip_scen_sets %>% group_by(trip_mode, scenario) %>% summarise(sum(trip_distance))
-  dist_scen<- as.data.frame (dist_scen)
-  names(dist_scen)[3]<-"distance"
-  dist_scen$distance<- round(dist_scen$distance, digits=0)
-    dist_scen<-spread(dist_scen, trip_mode, distance)
-  dist_scen$sum<-rowSums (dist_scen[2:(ncol(dist_scen))], na.rm = FALSE, dims = 1)
-  dist_scen_prop<- round(dist_scen[,2:(ncol(dist_scen))]*100/dist_scen$sum, digits=1)
-  dist_scen_prop<- cbind(dist_scen[,1], dist_scen_prop)
+<!--   dist_scen<- io[[cities[x]]]$trip_scen_sets %>% group_by(trip_mode, scenario) %>% summarise(sum(trip_distance)) -->
+<!--   dist_scen<- as.data.frame (dist_scen) -->
+<!--   names(dist_scen)[3]<-"distance" -->
+<!--   dist_scen$distance<- round(dist_scen$distance, digits=0) -->
+<!--     dist_scen<-spread(dist_scen, trip_mode, distance) -->
+<!--   dist_scen$sum<-rowSums (dist_scen[2:(ncol(dist_scen))], na.rm = FALSE, dims = 1) -->
+<!--   dist_scen_prop<- round(dist_scen[,2:(ncol(dist_scen))]*100/dist_scen$sum, digits=1) -->
+<!--   dist_scen_prop<- cbind(dist_scen[,1], dist_scen_prop) -->
 
-  names(dist_scen_prop)[1]<-"Scenario"
-  dist_scen_prop$Scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  print(kable(dist_scen_prop, caption= cities[x]))
-  print(kable(dist_scen, caption= cities[x]))
+<!--   names(dist_scen_prop)[1]<-"Scenario" -->
+<!--   dist_scen_prop$Scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen") -->
+<!--   print(kable(dist_scen_prop, caption= cities[x])) -->
+<!--   print(kable(dist_scen, caption= cities[x])) -->
 
-}
-```
+<!-- } -->
+<!-- ``` -->

### 2020-09-30 summary_tables_PA_AP.Rmd
Update description of CO2 emissions

@@
 1) Individual-level PM2.5 concentrations for baseline and scenarios
-2) Individual-level CO2 for baseline and scenarios
+2) Overall CO2 emissions for baseline and scenarios
 3) Individual-level physical activity for baseline and scenarios

### 2020-10-14 injury_tables.Rmd
Save plots to local files

@@
 df[df == 0] <- NA
-df <- df %>% mutate(mv = rowMeans(select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
+df <- df %>% mutate(mv = rowMeans(dplyr::select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
 df1 <- df %>% dplyr::filter(variable == 'mv')
@@
    facet_wrap(vars(scenario)) + 
-   scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 25)) +
+   #scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 25)) +
    labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
@@
     facet_grid(vars(scenario), scales='free_x') + 
-    scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
+    # scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
     labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
     theme(axis.text.x = element_text(face="bold", color="#993333", size=14))
-# ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
+ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
 fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
-# ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
+htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
 fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
+htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
 print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
+ggsave(d2, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_specific_mode.png"), limitsize = F, dpi = 300, scale = 1.5)
 fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
+htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
 ```

### 2020-10-27 summary_tables.Rmd

@@
 |---|---|---|---|
-| walking | 98.1|72.3|2.9 |
-| bicycle | 0.5|0.3|0.5 |
-| car | 1.0|9.0|17.1 |
-| motorcycle | 0.0|0.0|0.0 |
-| bus | 0.0|15.0|64.4 |
+| pedestrian | 98.5|0.0|0.0 |
+| cycle | 0.5|1.5|0.4 |
+| car | 0.8|29.8|14.4 |
+| motorcycle | 0.0|0.1|1.0 |
+| bus | 0.0|60.0|74.0 |
 
-|sao_paulo|0-1 km|2-5 km|6+ km|
+|bangalore|0-1 km|2-5 km|6+ km|
 |---|---|---|---|
-| walking | 62.2|7.2|3.0 |
-| bicycle | 0.7|0.9|0.3 |
-| car | 21.0|42.9|28.7 |
-| motorcycle | 1.3|2.9|3.8 |
-| bus | 13.3|39.2|32.7 |
+| pedestrian | 75.0|23.1|1.0 |
+| cycle | 2.3|9.8|1.2 |
+| car | 1.0|4.3|6.4 |
+| motorcycle | 5.1|33.9|63.2 |
+| bus | 15.1|24.3|24.9 |
+
+|belo_horizonte|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 91.7|36.7|0.6 |
+| cycle | 0.4|3.7|1.0 |
+| car | 7.5|47.2|49.9 |
+| motorcycle | 0.2|6.7|7.2 |
+| bus | 0.1|4.8|37.5 |
+
+|bogota|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 90.0|26.8|2.3 |
+| cycle | 4.5|14.8|4.1 |
+| car | 1.5|15.0|21.0 |
+| motorcycle | 0.4|5.5|6.9 |
+| bus | 3.1|33.5|61.0 |
+
+|buenos_aires|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 96.6|34.2|0.1 |
+| cycle | 2.3|8.3|1.0 |
+| car | 0.5|24.4|21.8 |
+| motorcycle | 0.0|1.9|1.2 |
+| bus | 0.6|27.8|48.2 |
+
+|cape_town|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 88.5|23.3|2.0 |
+| cycle | 0.0|0.4|0.3 |
+| car | 9.4|46.1|47.3 |
+| motorcycle | 0.0|0.3|0.6 |
+| bus | 2.0|27.5|36.5 |
 
@@
 |---|---|---|---|
-| walking | 84.4|10.1|0.9 |
-| bicycle | 2.0|7.6|3.7 |
+| pedestrian | 84.4|10.1|0.9 |
+| cycle | 2.0|7.6|3.7 |
 | car | 2.4|17.3|21.4 |
@@
 
-|bangalore|0-1 km|2-5 km|6+ km|
+|mexico_city|0-1 km|2-5 km|6+ km|
 |---|---|---|---|
-| walking | 74.9|23.0|1.0 |
-| bicycle | 2.3|9.8|1.1 |
-| car | 1.0|4.4|6.4 |
-| motorcycle | 5.1|33.9|63.2 |
-| bus | 15.1|24.3|25.0 |
+| pedestrian | 96.0|37.7|0.3 |
+| cycle | 0.8|6.4|0.7 |
+| car | 2.3|21.8|16.7 |
+| motorcycle | 0.0|2.7|1.1 |
+| bus | 0.4|24.3|58.0 |
 
@@
 |---|---|---|---|
-| walking | 88.4|23.2|0.3 |
-| bicycle | 3.5|8.7|2.3 |
-| car | 7.2|45.5|35.5 |
-| motorcycle | 0.1|1.4|1.4 |
-| bus | 0.3|10.1|29.6 |
+| pedestrian | 88.5|23.7|0.2 |
+| cycle | 3.5|8.8|1.9 |
+| car | 7.2|49.5|32.3 |
+| motorcycle | 0.1|1.6|1.0 |
+| bus | 0.3|10.3|29.8 |
 
-|belo_horizonte|0-1 km|2-5 km|6+ km|
+|sao_paulo|0-1 km|2-5 km|6+ km|
 |---|---|---|---|
-| walking | 91.7|34.7|0.6 |
-| bicycle | 0.4|3.5|1.0 |
-| car | 7.2|38.4|39.5 |
-| motorcycle | 0.2|6.3|7.4 |
-| bus | 0.1|4.5|38.4 |
-
-|buenos_aires|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| walking | 96.5|33.8|0.2 |
-| bicycle | 2.4|8.3|1.2 |
-| car | 0.5|24.8|25.7 |
-| motorcycle | 0.0|1.8|1.5 |
-| bus | 0.5|27.7|50.1 |
-
-|mexico_city|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| walking | 95.3|37.6|0.3 |
-| bicycle | 0.8|7.4|1.2 |
-| car | 2.4|24.9|29.6 |
-| motorcycle | 0.0|3.1|1.9 |
-| bus | 0.3|18.8|47.1 |
-
-|bogota|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| walking | 90.0|27.3|2.4 |
-| bicycle | 4.5|15.1|4.2 |
-| car | 1.4|13.3|18.7 |
-| motorcycle | 0.4|5.6|7.2 |
-| bus | 3.2|34.1|62.9 |
-
-|cape_town|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| walking | 88.5|23.3|2.0 |
-| bicycle | 0.0|0.4|0.3 |
-| car | 9.4|46.1|47.3 |
-| motorcycle | 0.0|0.3|0.6 |
-| bus | 2.0|27.5|36.5 |
+| pedestrian | 62.2|7.2|3.0 |
+| cycle | 0.7|0.9|0.3 |
+| car | 22.0|45.1|29.9 |
+| motorcycle | 1.3|2.9|3.8 |
+| bus | 13.3|39.2|32.7 |
 
@@
 |---|---|---|---|
-| walking | 80.6|18.3|0.1 |
-| bicycle | 3.9|7.2|2.0 |
-| car | 0.6|2.3|2.4 |
-| motorcycle | 9.5|23.8|25.0 |
-| bus | 1.2|27.5|56.0 |
+| pedestrian | 84.1|23.1|0.1 |
+| cycle | 4.0|9.0|2.4 |
+| car | 0.6|2.9|2.8 |
+| motorcycle | 9.9|30.1|29.2 |
+| bus | 1.3|34.8|65.4 |
 

### 2020-11-02 code/trips_analysis/mode_speed_summary.Rmd
adding summary tables for distance

---
title: "mode_speed_summary"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(dplyr)
```

Summary table for mode speeds
```{r}
summary <- read.csv("../../data/analysis/mode_speed/speed_df.csv")
print(summary)
```

Boxplots for mode_speeds
```{r}
bangalore <- read.csv("../../data/local/bangalore/bangalore_trip.csv")
delhi <- read.csv("../../data/local/delhi/delhi_trip.csv")
vizag <- read.csv("../../data/local/vizag/vizag_trip.csv")
sao_paulo <- read.csv("../../data/local/sao_paulo/sao_paulo_trip.csv")

bangalore <- bangalore %>% 
  filter(stage_distance > 0 & stage_duration > 0 & !is.na(stage_mode)) %>%
  mutate(speed = stage_distance / ( stage_duration / 60),
         city = "Bangalore",
         id = paste0("Bangalore",stage_id)) %>% 
  select(id, city, speed, stage_mode) %>% 
  rename(mode = stage_mode)

delhi <- delhi %>% 
  filter(stage_distance > 0 & stage_duration > 0 & !is.na(stage_mode)) %>%
  mutate(speed = stage_distance / ( stage_duration / 60),
         city = "Delhi",
         id = paste0("Delhi", stage_id)) %>% 
  select(id, city, speed, stage_mode) %>% 
  rename(mode = stage_mode)

vizag <- vizag%>% 
  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
  mutate(speed = trip_distance / ( trip_duration / 60),
         city = "Vizag",
         id = paste0("Vizag", trip_id)) %>% 
  select(id, city, speed, trip_mode) %>% 
  rename(mode = trip_mode)

sao_paulo <- sao_paulo%>% 
  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
  mutate(speed = trip_distance / ( trip_duration / 60),
         city = "Sao_Paulo",
         id = paste0("Sao_Paulo", trip_id)) %>% 
  select(id, city, speed, trip_mode) %>% 
  rename(mode = trip_mode)

trip <- rbind(bangalore, delhi, vizag, sao_paulo)
trip_walk <- filter(trip, mode=="walk")
trip_bicycle <- trip[trip$mode=="bicycle",]
trip_bus <- trip[trip$mode == "bus",]
trip_car <- trip[trip$mode == "car",]

```

```{r}
boxplot(speed~mode, data = droplevels(trip), ylim=c(0,40))
```

## Walking
```{r}
boxplot(speed~mode+city, data = droplevels(trip_walk), ylim=c(0, 10))
```

## Cycling
```{r}
boxplot(speed~mode+city, data = droplevels(trip_bicycle), ylim=c(0, 20))
```

## Bus
```{r}
boxplot(speed~mode+city, data = droplevels(trip_bus), ylim=c(0, 40))
```

## Car
```{r}
boxplot(speed~mode+city, data = droplevels(trip_car), ylim=c(0, 40))
```

### 2020-11-02 code/producing_trips_rd/travel_survey.R

@@ -2296,9 +2296,9 @@ trip$year <- 2011
 trip$gdppc2014 <- 23162
 trip$population2014 <- 1878000
   
-write.csv(trip, "data/local/vizag/visakhapatnam_trip.csv")
+write.csv(trip, "data/local/vizag/vizag_trip.csv")
 
-trip <- read_csv('data/local/vizag/visakhapatnam_trip.csv')
+trip <- read_csv('data/local/vizag/vizag_trip.csv')
 
 # Standardized travel modes
 trip <- standardize_modes(trip, mode = c('trip'))

### 2020-11-09 code/trips_analysis/mode_speed_summary.Rmd
add motorcycle speeds to speed distancea analysis

@@
 trip_bicycle <- trip[trip$mode=="bicycle",]
+trip_motorcylce <- trip[trip$mode=="motorcycle",]
 trip_bus <- trip[trip$mode == "bus",]
@@
 
+## Motorcycling
+```{r}
+boxplot(speed~mode+city, data = droplevels(trip_bicycle), ylim=c(0, 20))
+```
+
 ## Bus

### 2020-11-09 code/trips_analysis/mode_speed_summary.Rmd

@@
 library(dplyr)
+library(reshape2)
+library(ggplot2)
+library(wesanderson)
 ```
 
-Summary table for mode speeds
+## Notes
+
+Bangalore had a beautiful variable called transfer time = walk time plus wait time, but when closely examined, what was recoreded there seemd to just be the travel time.
+So in our data processing, we had used both the variables on transfer time and start and stop time to come up with trip duration.
+Are there any concerns here?
+
+## Summary table for mode speeds
 ```{r}
@@
          city = "Bangalore",
-         id = paste0("Bangalore",stage_id)) %>% 
-  select(id, city, speed, stage_mode) %>% 
+         id = paste0("Bangalore",stage_id),
+         age_group = ifelse(age<=50, "<50", "50+")) %>% 
+  select(id, city, speed, sex, age_group ,stage_mode) %>% 
   rename(mode = stage_mode)
@@
          city = "Delhi",
-         id = paste0("Delhi", stage_id)) %>% 
-  select(id, city, speed, stage_mode) %>% 
+         id = paste0("Delhi",stage_id),
+         age_group = ifelse(age<=50, "<50", "50+")) %>% 
+  select(id, city, speed, sex, age_group ,stage_mode) %>% 
   rename(mode = stage_mode)
@@
          city = "Vizag",
-         id = paste0("Vizag", trip_id)) %>% 
-  select(id, city, speed, trip_mode) %>% 
+         id = paste0("Vizag", trip_id),
+          age_group = ifelse(age<=50, "<50", "50+")) %>% 
+  select(id, city, sex, age_group, speed, trip_mode) %>% 
   rename(mode = trip_mode)
 
-sao_paulo <- sao_paulo%>% 
+sao_paulo_1 <- sao_paulo%>% 
   filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
   mutate(speed = trip_distance / ( trip_duration / 60),
-         city = "Sao_Paulo",
-         id = paste0("Sao_Paulo", trip_id)) %>% 
-  select(id, city, speed, trip_mode) %>% 
+         city = "Sao_Paulo_1",
+         id = paste0("Sao_Paulo_1", trip_id),
+          age_group = ifelse(age<=50, "<50", "50+")) %>% 
+  select(id, city,sex, age_group, speed, trip_mode) %>% 
   rename(mode = trip_mode)
 
-trip <- rbind(bangalore, delhi, vizag, sao_paulo)
+sao_paulo_2 <- sao_paulo%>% 
+  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
+  mutate(trip_duration = (trip_duration - (walk_to_pt + walk_from_pt)),
+         speed = trip_distance / ( trip_duration / 60),
+         city = "Sao_Paulo_2",
+         id = paste0("Sao_Paulo_2", trip_id),
+         age_group = ifelse(age<=50, "<50", "50+")) %>% 
+  select(id, city,sex, age_group,  speed, trip_mode) %>% 
+  rename(mode = trip_mode)
+
+trip <- rbind(bangalore, delhi, vizag, sao_paulo_1, sao_paulo_2)
 trip_walk <- filter(trip, mode=="walk")
 trip_bicycle <- trip[trip$mode=="bicycle",]
-trip_motorcylce <- trip[trip$mode=="motorcycle",]
+trip_motorcylce <- trip[trip$mode=="motcycle",]
 trip_bus <- trip[trip$mode == "bus",]
@@
 ```
+## Walking by sex and age
+```{r}
+trip_walk %>% 
+  filter(!is.na(sex) & !is.na(age_group)) %>% 
+  ggplot(aes(x = city, y = speed,  fill = sex))+
+  geom_boxplot(aes(col = age_group)) + 
+  ylim(0, 8)+
+  scale_fill_manual(values= c("grey80", "bisque2"))
+
+```
 
@@
 boxplot(speed~mode+city, data = droplevels(trip_bicycle), ylim=c(0, 20))
+```
+
+## Cycling by sex and age
+```{r}
+trip_bicycle %>% 
+  filter(!is.na(sex) & !is.na(age_group)) %>% 
+  ggplot(aes(x = city, y = speed,  fill = sex))+
+  geom_boxplot(aes(col = age_group)) + 
+  ylim(0, 15)+
+  scale_fill_manual(values= c("grey80", "bisque2"))
+
+#trip_bicycle %>% filter(city=="Delhi") %>% View
 ```
@@
 
+## Car
+```{r}
+boxplot(speed~mode+city, data = droplevels(trip_car), ylim=c(0, 40))
+```
+
 ## Bus
@@
 
-## Car
+## Walk to/from PT
 ```{r}
-boxplot(speed~mode+city, data = droplevels(trip_car), ylim=c(0, 40))
+walk_pt <- sao_paulo %>% 
+  select(walk_to_pt, walk_from_pt) %>% 
+  mutate(walk_to_from_pt = (walk_to_pt + walk_from_pt)) %>% 
+  filter(walk_to_from_pt > 0)
+
+walk_pt <- melt(walk_pt)
+
+boxplot(value~variable, data = droplevels(walk_pt), ylim=c(0, 20))
 ```

### 2020-11-15 code/trips_analysis/mode_speed_summary.Rmd
calculating new defualt speeds

@@
 ---
-title: "mode_speed_summary"
+title: "mode speeds used in the ithim-global model"
 output: html_document
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, include = FALSE}
+rm(list = ls())
 library(dplyr)
+library(magrittr)
 library(reshape2)
@@
 library(wesanderson)
+library(kableExtra)
 ```
 
-## Notes
+We previously used the following default mode speeds from the literature to calculate trip distances [ref]:
+ default_speeds <- list(
+    bus=15,
+    bus_driver=15,
+    minibus=15,
+    minibus_driver=15,
+    car=21,
+    taxi=21,
+    pedestrian=4.8,
+    walk_to_pt=4.8,
+    cycle=14.5,
+    motorcycle=25,
+    truck=21,
+    van=15,
+    subway=28,
+    rail=35,
+    auto_rickshaw=22,
+    shared_auto=22,
+    shared_taxi=21,
+    cycle_rickshaw=10
+  )
 
-Bangalore had a beautiful variable called transfer time = walk time plus wait time, but when closely examined, what was recoreded there seemd to just be the travel time.
-So in our data processing, we had used both the variables on transfer time and start and stop time to come up with trip duration.
-Are there any concerns here?
+We later doubted these estimates and wanted to generated default speeds from the datasets we had [please I might not have gotten the exact reason why we needed these new speeds, kinly edit].
 
-## Summary table for mode speeds
-```{r}
-summary <- read.csv("../../data/analysis/mode_speed/speed_df.csv")
-print(summary)
-```
+From the 11 cities we are working on, we selected cities with both trip duration and trip distances (4 in total: Sao Paulo, Vizag, Delhi and Bangalore) to calculate the default speeds.
 
-Boxplots for mode_speeds
-```{r}
+- Delhi had stage distances and stage duration
+- Bangalore had stage distances and stage duration
+- Sao Paulo had only trip level infomation (trip distance and trip duration). In addition, it had walk to PT
+- Vizag had only trip level data (not stages)
+
+NB: Bangalore initialy had variable called transfer time = walk time plus wait time, but when we examined closely it matched the travel time.
+So in our data processing, we had used both variables to come up with trip duration.
+
+## 1 First we explored the walk to PT time in the Sao Paulo data
+
+```{r echo = FALSE, message=FALSE, warning=FALSE }
 bangalore <- read.csv("../../data/local/bangalore/bangalore_trip.csv")
@@
 
-vizag <- vizag%>% 
+vizag_1 <- vizag%>% 
   filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
   mutate(speed = trip_distance / ( trip_duration / 60),
-         city = "Vizag",
-         id = paste0("Vizag", trip_id),
+         city = "Vizag_1",
+         id = paste0("Vizag_1", trip_id),
           age_group = ifelse(age<=50, "<50", "50+")) %>% 
   select(id, city, sex, age_group, speed, trip_mode) %>% 
+  rename(mode = trip_mode)
+
+vizag_2 <- vizag %>% 
+  filter(trip_distance > 0 & trip_duration > 0 & !is.na(trip_mode)) %>%
+  mutate(trip_duration = ifelse(trip_mode == "bus", abs(trip_duration - 10), trip_duration),
+         speed = trip_distance / ( trip_duration / 60),
+         city = "Vizag_2",
+         id = paste0("Vizag_2", trip_id),
+         age_group = ifelse(age<=50, "<50", "50+")) %>% 
+  select(id, city,sex, age_group,  speed, trip_mode) %>% 
   rename(mode = trip_mode)
@@
 
-trip <- rbind(bangalore, delhi, vizag, sao_paulo_1, sao_paulo_2)
+trip <- rbind(bangalore, delhi, vizag_1, vizag_2, sao_paulo_1, sao_paulo_2)
 trip_walk <- filter(trip, mode=="walk")
 trip_bicycle <- trip[trip$mode=="bicycle",]
-trip_motorcylce <- trip[trip$mode=="motcycle",]
+trip_motorcycle <- trip[trip$mode=="motorcycle",]
 trip_bus <- trip[trip$mode == "bus",]
@@
 
+trip_2 <- rbind(bangalore, delhi, vizag_2, sao_paulo_2)
+trip_2_walk <- filter(trip_2, mode=="walk")
+trip_2_bicycle <- trip[trip_2$mode=="bicycle",]
+trip_2_motorcycle <- trip[trip_2$mode=="motorcycle",]
+trip_2_bus <- trip[trip_2$mode == "bus",]
+trip_2_car <- trip[trip_2$mode == "car",]
+
 ```
 
-```{r}
-boxplot(speed~mode, data = droplevels(trip), ylim=c(0,40))
+### 1.1 Walk to/from PT box plots (Sao Paulo)
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+walk_pt <- sao_paulo %>% 
+  select(walk_to_pt, walk_from_pt) %>% 
+  mutate(walk_to_from_pt = (walk_to_pt + walk_from_pt)) %>% 
+  filter(walk_to_from_pt > 0)
+
+walk_pt <- melt(walk_pt)
+
+#View(walk_pt)
+
+boxplot(value~variable, data = droplevels(walk_pt), ylim=c(0, 20))
 ```
 
-## Walking
-```{r}
+### 1.2 Walk to/from PT (Sao Paulo) - histogram shows skewedness
+```{r echo = FALSE, message=FALSE, warning=FALSE}
+walk_pt <- sao_paulo %>% 
+  select(walk_to_pt, walk_from_pt) %>% 
+  mutate(walk_to_from_pt = (walk_to_pt + walk_from_pt)) %>% 
+  filter(walk_to_from_pt > 0)
+
+hist(walk_pt$walk_to_from_pt)
+
+```
+
+### 1.3 Walk to/from PT overall (Sao Paulo)
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+walk_pt <- sao_paulo %>% 
+  select(walk_to_pt, walk_from_pt) %>% 
+  mutate(walk_to_from_pt = (walk_to_pt + walk_from_pt)) %>% 
+  filter(walk_to_from_pt > 0)
+
+summary(walk_pt$walk_to_from_pt) 
+```
+
+### 1.4 Walk to mode by mode (Sao Paulo) since all modes apart from walking had some walking component. 
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+sao_paulo %>% 
+  mutate(walk_to_from_pt = (walk_to_pt + walk_from_pt)) %>% 
+  filter(walk_to_from_pt>0) %>% 
+  group_by(trip_mode) %>% 
+  summarise(median_walk_time = median(walk_to_from_pt)) %>% 
+  kable()
+```
+We applied the median walk time for bus to the Vizag dataset where we had only trip level data.
+It did not seem logical to apply walk to to all other modes because of the small value and because they are not likely to be captures as separate stage in trip survey.
+
+## 2. We looked at the speeds for the different modes.
+
+### 2.1 Walking speed did not vary much by city (Bangalore stood out though). 
+We agreed to set the same default waking speeds for all cities.
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 boxplot(speed~mode+city, data = droplevels(trip_walk), ylim=c(0, 10))
 ```
-## Walking by sex and age
-```{r}
+Walking speeds by sex and age only showed slight variations, and mostly so in Vizag. The variation by age was least pronounced, so our default walking speeds was split by sex.
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 trip_walk %>% 
@@
 
-## Cycling
-```{r}
+### 2.2 Cycling speeds did not seem to vary much across the cities
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 boxplot(speed~mode+city, data = droplevels(trip_bicycle), ylim=c(0, 20))
@@
 
-## Cycling by sex and age
-```{r}
+Cycling speed by sex and age showed variation in sex but not much by age, so we split default cycling speeds by sex.
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 trip_bicycle %>% 
@@
 
-## Motorcycling
-```{r}
-boxplot(speed~mode+city, data = droplevels(trip_bicycle), ylim=c(0, 20))
+### 2.3 Motorcycle speed
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+boxplot(speed~mode+city, data = droplevels(trip_motorcycle), ylim=c(0, 30))
 ```
 
-## Car
-```{r}
+## 2.4 Car speeds
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 boxplot(speed~mode+city, data = droplevels(trip_car), ylim=c(0, 40))
@@
 
-## Bus
-```{r}
+## 2.5 Bus speeds
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 boxplot(speed~mode+city, data = droplevels(trip_bus), ylim=c(0, 40))
@@
 
-## Walk to/from PT
-```{r}
-walk_pt <- sao_paulo %>% 
-  select(walk_to_pt, walk_from_pt) %>% 
-  mutate(walk_to_from_pt = (walk_to_pt + walk_from_pt)) %>% 
-  filter(walk_to_from_pt > 0)
+## 3. Conclusions on default speeds
+- Walking: use median speeds for all cities split by sex 
+- Cycling: Use median speeds split by sex
+- Cycled Rickshaw should be 2/3 the speed of bicycle
+- Motorcycle Rickshaw should be 2/3 the speed of motorcycle
+- Van should be the same as bus
+- Others:  use same median speeds for both sex and age
 
-walk_pt <- melt(walk_pt)
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+a <- trip_2 %>% 
+  filter(!is.na(sex)) %>% 
+  filter(mode == "bicycle" | mode == "walk") %>% 
+  group_by(mode, sex) %>% 
+  summarise(mode_speed = median(speed))
 
-boxplot(value~variable, data = droplevels(walk_pt), ylim=c(0, 20))
+b <-  trip_2 %>% 
+  filter(!(mode == "bicycle" | mode == "walk")) %>% 
+  group_by(mode) %>% 
+  summarise(mode_speed = median(speed), sex = "All")
+
+kable(rbind(a, b))
 ```
+
+### Adjsted speeds
+
+default_speeds <- list(
+    bus=10.8,
+    bus_driver=10.8,
+    minibus=10.8,
+    minibus_driver=10.8,
+    car=11.3,
+    taxi=11.3,
+    pedestrian_male=4,
+    pedestrian_female=3,
+    walk_to_pt=10,
+    cycle_male=8,
+    cycle_female=6,
+    motorcycle=12,
+    truck=10.8,
+    van=10.8,
+    subway=12.9,
+    rail=13.5,
+    auto_rickshaw=8,
+    shared_auto=11.3,
+    shared_taxi=11.3,
+    cycle_rickshaw=5.3
+  )

### 2020-11-15 injury_tables.Rmd

--- 2020-08-26 injury_tables.Rmd
@@
     facet_grid(vars(scenario), scales='free_x') + 
-    scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
+    # scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
     labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
@@
 # ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+# fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
+# htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
 # ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
-print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
+# fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
+# htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
+# print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
+# ggsave(d2, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_specific_mode.png"), limitsize = F, dpi = 300, scale = 1.5)
+# fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
+# htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
 ```

### 2020-11-18 summary_tables.Rmd
Add baseline duration

@@
 
+trip_dur <- list()
+
+l <- list()
+
 for(city in cities){
@@
   row.names(td) <- NULL
+
+  l[[city]] <- td
+
+  td1 <- td
+
+  if (length(trip_dur) == 0){
+    trip_dur <- td1 %>% dplyr::select(stage_mode)
+  }
+
+  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
+
+  names(td1)[2] <- city
+
+  trip_dur <- inner_join(trip_dur, td1, by = 'stage_mode')
 
@@
 ```
+
+### Baseline average duration (mins) per day for all case studies
+```{r}
+
+trip_dur <- trip_dur %>% janitor::adorn_totals('row')
+
+kable(trip_dur, caption = "Avg. duration per person (mins) per day across all case studies")
+
+```

### 2020-11-21 code/trips_analysis/mode_speed_summary.Rmd
change default speeds to mean of median speeds for cities

@@
 - Motorcycle Rickshaw should be 2/3 the speed of motorcycle
-- Van should be the same as bus
+- Van should be the same as car
 - Others:  use same median speeds for both sex and age
@@
     truck=10.8,
-    van=10.8,
+    van=11.3,
     subway=12.9,

### 2020-11-21 code/trips_analysis/mode_speed_summary.Rmd
minor edits to trip distances

@@
   )
+
+### Use mean of median speeds instead of pooled medians to get default speeds.
+Current work stream does not account for males and females, so we can keep that differentiation for later.  
+```{r}
+trip_2 %>% 
+  group_by(city, mode) %>% 
+  summarise(median_speed = median(speed)) %>% 
+  group_by(mode) %>% 
+  summarise(mean_median = mean(median_speed)) %>% 
+  kable()
+
+```
+
+default_speeds <- list(
+    bus=11,
+    bus_driver=11,
+    minibus=11,
+    minibus_driver=11,
+    car=14,
+    taxi=9,
+    pedestrian=3,
+    walk_to_pt=10,
+    cycle=7,
+    motorcycle=14,
+    truck=11,
+    van=14,
+    subway=16,
+    rail=20,
+    auto_rickshaw=9,
+    shared_auto=14,
+    shared_taxi=9,
+    cycle_rickshaw=5
+  )

### 2020-11-23 code/producing_trips_rd/trip-summary.Rmd
Update paths

@@
 
-trip <- read_csv("../../data/local/mexico/mexico_city_trip.csv")
+trip <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv")
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/vizag/visakhapatnam_trip.csv")
+trip <- read_csv("../../data/local/vizag/vizag_trip.csv")
 quality_check(trip)

### 2020-11-24 summary_tables.Rmd
Remove writing true distance
Update propensities

@@
 |---|---|---|---|
-| pedestrian | 98.5|0.0|0.0 |
-| cycle | 0.5|1.5|0.4 |
-| car | 0.8|29.8|14.4 |
-| motorcycle | 0.0|0.1|1.0 |
-| bus | 0.0|60.0|74.0 |
+| pedestrian | 95.6|37.9|2.5 |
+| cycle | 0.6|0.9|0.3 |
+| car | 3.2|18.1|10.0 |
+| motorcycle | 0.0|0.1|1.1 |
+| bus | 0.0|29.7|80.5 |
 
@@
 |---|---|---|---|
-| pedestrian | 91.7|36.7|0.6 |
-| cycle | 0.4|3.7|1.0 |
-| car | 7.5|47.2|49.9 |
-| motorcycle | 0.2|6.7|7.2 |
-| bus | 0.1|4.8|37.5 |
+| pedestrian | 89.2|8.7|0.3 |
+| cycle | 1.6|2.7|0.2 |
+| car | 7.3|65.6|47.1 |
+| motorcycle | 1.3|10.1|5.3 |
+| bus | 0.4|11.1|43.1 |
 
@@
 |---|---|---|---|
-| pedestrian | 90.0|26.8|2.3 |
-| cycle | 4.5|14.8|4.1 |
-| car | 1.5|15.0|21.0 |
-| motorcycle | 0.4|5.5|6.9 |
-| bus | 3.1|33.5|61.0 |
+| pedestrian | 80.8|11.1|1.0 |
+| cycle | 7.9|11.7|2.1 |
+| car | 3.2|18.9|17.7 |
+| motorcycle | 1.4|8.0|6.4 |
+| bus | 4.9|43.2|70.1 |
 
@@
 |---|---|---|---|
-| pedestrian | 96.6|34.2|0.1 |
-| cycle | 2.3|8.3|1.0 |
-| car | 0.5|24.4|21.8 |
-| motorcycle | 0.0|1.9|1.2 |
-| bus | 0.6|27.8|48.2 |
+| pedestrian | 79.7|3.0|0.1 |
+| cycle | 8.0|3.9|0.3 |
+| car | 6.2|35.0|21.8 |
+| motorcycle | 0.6|2.8|1.3 |
+| bus | 3.3|49.2|50.5 |
 
@@
 |---|---|---|---|
-| pedestrian | 88.5|23.3|2.0 |
-| cycle | 0.0|0.4|0.3 |
-| car | 9.4|46.1|47.3 |
-| motorcycle | 0.0|0.3|0.6 |
-| bus | 2.0|27.5|36.5 |
+| pedestrian | 84.1|5.3|1.9 |
+| cycle | 0.2|0.5|0.1 |
+| car | 8.0|55.6|48.3 |
+| motorcycle | 0.0|0.7|0.4 |
+| bus | 6.9|34.8|33.3 |
 
@@
 |---|---|---|---|
-| pedestrian | 96.0|37.7|0.3 |
-| cycle | 0.8|6.4|0.7 |
-| car | 2.3|21.8|16.7 |
-| motorcycle | 0.0|2.7|1.1 |
-| bus | 0.4|24.3|58.0 |
+| pedestrian | 90.2|7.4|0.1 |
+| cycle | 3.3|4.6|0.3 |
+| car | 2.1|33.1|30.2 |
+| motorcycle | 0.6|4.5|1.1 |
+| bus | 1.6|37.0|49.1 |
 
@@
 |---|---|---|---|
-| pedestrian | 88.5|23.7|0.2 |
-| cycle | 3.5|8.8|1.9 |
-| car | 7.2|49.5|32.3 |
-| motorcycle | 0.1|1.6|1.0 |
-| bus | 0.3|10.3|29.8 |
+| pedestrian | 83.9|4.0|0.1 |
+| cycle | 6.4|6.3|0.4 |
+| car | 7.5|54.7|29.8 |
+| motorcycle | 0.2|1.8|0.9 |
+| bus | 0.6|21.2|30.3 |
 
@@
 |---|---|---|---|
-| pedestrian | 84.1|23.1|0.1 |
-| cycle | 4.0|9.0|2.4 |
-| car | 0.6|2.9|2.8 |
-| motorcycle | 9.9|30.1|29.2 |
-| bus | 1.3|34.8|65.4 |
+| pedestrian | 86.2|19.7|0.1 |
+| cycle | 2.6|6.6|2.1 |
+| car | 0.6|2.4|2.8 |
+| motorcycle | 6.9|24.3|24.6 |
+| bus | 0.8|26.9|55.5 |
 
@@
 
-  readr::write_csv(df, paste0('results/multi_city/true_distances/true_distance_', city, '.csv'))
-
   print(kable(df, caption = paste("True distance by mode per day for  ", city)))

### 2020-11-25 summary_tables.Rmd
add numbers to sections

@@
     toc_float: true
+    number_sections: true
 ---
@@
 library(tidyverse)
+library(gt)
 })
@@
   count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
   colnames(td)[3:7] <- rownames(scen_prop)
@@
   count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-  td <- io[[city]]$dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
   colnames(td)[3:7] <- rownames(scen_prop)
@@
 
-require(gt)
-
 for(city in cities){
@@
   count_people <- round(sum(io[[city]]$demographic$population))
-  td <- io[[city]]$true_dist %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 365, 3)))
+  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 365, 3)))
   colnames(td)[3:7] <- rownames(scen_prop)
@@
 
-  df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), prop = round(n_vals / nrow(.) * 100, 1))
+  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), prop = round(n_vals / nrow(.) * 100, 1))
   print(kable(df, caption = paste("Distance by mode by for  ", city)))
@@
   count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dur %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
   colnames(td)[3:7] <- rownames(scen_prop)
@@
   count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-  td <- io[[city]]$dur %>% filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
   colnames(td)[3:7] <- rownames(scen_prop)

### 2021-01-13 summary_tables.Rmd
Update propensities

@@
 |---|---|---|---|
-| pedestrian | 95.6|37.9|2.5 |
-| cycle | 0.6|0.9|0.3 |
-| car | 3.2|18.1|10.0 |
-| motorcycle | 0.0|0.1|1.1 |
-| bus | 0.0|29.7|80.5 |
+| pedestrian | 95.5|27.9|2.6 |
+| cycle | 0.6|0.7|0.4 |
+| car | 0.9|12.4|17.2 |
+| motorcycle | 0.0|0.0|1.3 |
+| bus | 2.4|48.3|71.8 |
 
@@
 |---|---|---|---|
-| pedestrian | 75.0|23.1|1.0 |
+| pedestrian | 74.9|23.0|1.0 |
 | cycle | 2.3|9.8|1.2 |
-| car | 1.0|4.3|6.4 |
+| car | 1.0|4.4|6.4 |
 | motorcycle | 5.1|33.9|63.2 |
-| bus | 15.1|24.3|24.9 |
+| bus | 15.1|24.4|25.0 |
 
@@
 |---|---|---|---|
-| pedestrian | 80.8|11.1|1.0 |
-| cycle | 7.9|11.7|2.1 |
-| car | 3.2|18.9|17.7 |
-| motorcycle | 1.4|8.0|6.4 |
-| bus | 4.9|43.2|70.1 |
+| pedestrian | 80.5|10.7|1.0 |
+| cycle | 7.9|11.3|2.0 |
+| car | 3.5|21.6|19.7 |
+| motorcycle | 1.4|7.7|6.2 |
+| bus | 4.9|41.6|68.3 |
 
@@
 |---|---|---|---|
-| pedestrian | 90.2|7.4|0.1 |
-| cycle | 3.3|4.6|0.3 |
-| car | 2.1|33.1|30.2 |
-| motorcycle | 0.6|4.5|1.1 |
-| bus | 1.6|37.0|49.1 |
+| pedestrian | 89.6|6.5|0.1 |
+| cycle | 3.4|4.5|0.3 |
+| car | 2.2|34.9|29.7 |
+| motorcycle | 0.6|4.7|1.0 |
+| bus | 1.6|36.2|49.9 |
 
@@
 |---|---|---|---|
-| pedestrian | 83.9|4.0|0.1 |
-| cycle | 6.4|6.3|0.4 |
-| car | 7.5|54.7|29.8 |
-| motorcycle | 0.2|1.8|0.9 |
-| bus | 0.6|21.2|30.3 |
+| pedestrian | 83.9|4.1|0.1 |
+| cycle | 6.2|6.0|0.4 |
+| car | 7.7|54.8|30.0 |
+| motorcycle | 0.2|2.0|0.9 |
+| bus | 0.6|21.1|30.1 |
 

### 2021-03-16 code/producing_trips_rd/trip-summary.Rmd
Adding summary table with details of each survey
Also updating results from Bogota, Mexico city and sao paulo

@@
 ```
+```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
+data.frame(
+  Ciudad = c("Buenos Aires", "Sao Paulo", "Belo Horizonte", "Mexico", "Bogota",
+             "Santiago", "Visakhapatnam", "Delhi", "Bangalore", "Accra",
+             "Cape Town"),
+  Study_area = c("BA and 27 municipalities.",
+                 "Metropolitan region of SP.",
+                 "Metropolitan region of BH.",
+                 "Metropolitan area of Mexico city.",
+                 "Bogota and surrounding municipalities.",
+                 "Metropolitan region of Santiago.",
+                 "There's not a report with survey details.",
+                 "There's not a report with survey details.",
+                 "Not sure",
+                 "The whole country but filtered only Greater Accra",
+                 "Cape Town, Malmesbury and Paarl"),
+  # Is there information at stage level?
+  Trip_Stage = c("Stage", "Trip", "Stage", "Stage", "Trip", 
+                 "Trip", "Trip", "Stage", "Stage", "Trip",
+                 "Trip"),
+  # If not, what is missing?
+  Missing = c("N/A", 
+              "There's no stage duration", 
+              "N/A", 
+              "N/A", 
+              "There's no stage duration",
+              "There's no stage duration",
+              "There's no stage mode, duration",
+              "N/A",
+              "N/A",
+              "Not sure, because it asks about modes and duration but don't know if people answered with this level of detail.",
+              "Looks like the survey capture up to 5 stage modes, but duration was not asked."),
+  # Is there information about short walks to pt
+  Walk_to_from_pt = c("Yes, in blocks (not minutes)",
+                      "No, walking trips are only considered when it is a unique mode. It never appears combined with other modes",
+                      "Yes",
+                      "Yes",
+                      "Yes",
+                      "Yes",
+                      "No",
+                      "Not sure",
+                      "Yes",
+                      "Not sure, because it asks about modes and duration but don't know if people answered with this level of detail.",
+                      "No"),
+  # Is the data ok?
+  OK = c("--", "--", "--", "--", "--", 
+         "--", "--", "--", "--", "--", "--"),
+  ToDo = c("Need to check if injuries dataset covers the same area that travel survey does. \n There's information about walking stages (blocks not minutes) but is not used",
+           "Need to check if injuries dataset covers the same area that travel survey does. \n Travel survey from 2012, there is one from 2017.",
+           "Need to check if injuries dataset covers the same area that travel survey does.
+\n According to the documentation, it looks like there's information at stage level (including stage duration).",
+           "Need to check if injuries dataset covers the same area that travel survey does.
+Trip mode is defined in terms of speed and duration, while in other surveys there's a hierarchy (check this).
+There's information on weekdays and weekends, decide what to do with this.",
+           "Need to check if injuries dataset covers the same area that travel survey does (The one that I got it's from only Bogota). Need to update travel survey.",
+           "Need to check if injuries dataset covers the same area that travel survey does.
+There's information about trips on working days in regular season and and working days in summer season, decide what to do with this.",
+           "Need to find a report to check the details of this survey",
+           "Need to find a report to check the details of this survey",
+           "There's a report called Comprehensive Traffic and Transportation Study for Bangalore Metropolitan region but I'm not sure if it corresponds to the data.
+Trip mode is defined in terms of speed and duration, while in other surveys there's a hierarchy (check this)",
+           "Need to check if injuries dataset covers the same area that time use survey does. Need to check if stages were captured.",
+           "Need to check if injuries dataset covers the same area that time use survey does.")
+  ) %>% kbl() %>% kable_classic()
+
+# (If short walk is not explicitly recorded, we have to find a way to remove; if they are missing, what should we do?)
+
+# Sao Paulo:
+# 1) Walking trips: Page 60, Pesquisa de Mobilidade da RMSP –2012 –Manual da Pesquisa Domiciliar
+# Belo horizonte:
+# 1) Study area: page 35, codebook and protocol.pdf
+# Cape town:
+# 1) Main mode Hierarchy: page 45, Final Draft HH Survey Report.pdf
+```
 
@@
 
-trip <- read_csv("../../data/local/bogota/bogota_trip.csv")
+trip <- read_csv("../../data/local/bogota/bogota_trips.csv")
 quality_check(trip)

### 2021-03-17 code/producing_trips_rd/trip-summary.Rmd
Simple changes in script
Adding some survey details

@@
                  "Cape Town, Malmesbury and Paarl"),
+  year = c("2010",
+           "2017",
+           "2012",
+           "2017",
+           "2019",
+           "2012",
+           "",
+           "",
+           "",
+           "",
+           ""),
+  # Definition of a trip
+  Trip_definition = c(
+    # BA: Pag 23- Report.pdf
+    "Moving from a origin to a destiny with a specific reason/motive",
+                      "Moving for a specific reason between two specific points (origin and destination), using one or more modes of transport.
+Walking trips where the reason for the trip is work or school, regardless of the distance travelled; or the distance covered is more than 500 mts for other reasons.",
+    # Belo Horizonte: pag 16, "Codebook and protocol.pdf"
+    "Moving from one part to another with a specific reason/motive, using one or more modes of transport",
+                      "Moving from one part to another with a specific reason/motive, using one or multiples modes of transport",
+                      "Moving from one part to another with a specific reason/motive, a definite hour of start and end, a mode of transport, and a duration greater than 3 minutes.
+Or moving from one part to another with reason/motive work or study of any duration",
+                      "Any movement carried out on public roads with a purpose
+determined, between two places (origin and destination) at a certain time of day; It can be carried out in several modes of transport and consist of one or more stages",
+"",
+"",
+"",
+"",
+""),
+  Reference_period = c(
+    # BA: page 4, report.pdf
+    "Trips made in working days",
+    "Trips made the day before the survey (from 4am yesterday to 3:59am today).",
+    "",
+                       "Trips made during the week (Tuesday, Wednesday, Thursday) and in Saturdays (weekends)",
+                       "Trips made the day of reference, i.e., the day before the survey (from 4am yesterday to 4am today). Surveys made from Monday to Sunday (according to the dataset).",
+                       "Trips made in working days (regular season), in weekends (regular season) and in working days(summer season)",
+                       "",
+                       "",
+                       "",
+                       "",
+                       ""),
   # Is there information at stage level?
@@
   # If not, what is missing?
-  Missing = c("N/A", 
+  Stage_missing = c("N/A", 
               "There's no stage duration", 
@@
   Walk_to_from_pt = c("Yes, in blocks (not minutes)",
-                      "No, walking trips are only considered when it is a unique mode. It never appears combined with other modes",
+                      "Yes",
                       "Yes",
@@
                       "No"),
-  # Is the data ok?
-  OK = c("--", "--", "--", "--", "--", 
-         "--", "--", "--", "--", "--", "--"),
+Walking_in_trip_duration = c("Not sure, yet.",
+                             "",
+                             "",
+                             "",
+                             "",
+                             "",
+                             "",
+                             "",
+                             "",
+                             "",
+                             ""),
   ToDo = c("Need to check if injuries dataset covers the same area that travel survey does. \n There's information about walking stages (blocks not minutes) but is not used",
@@
            "Need to check if injuries dataset covers the same area that time use survey does. Need to check if stages were captured.",
-           "Need to check if injuries dataset covers the same area that time use survey does.")
+           "Need to check if injuries dataset covers the same area that time use survey does."),
+           # Is the data ok?
+  OK = c("--", "--", "--", "--", "--", 
+         "--", "--", "--", "--", "--", "--")
   ) %>% kbl() %>% kable_classic()

### 2021-03-17 code/producing_trips_rd/trip-summary.Rmd
Updating detail table

@@
     "Moving from one part to another with a specific reason/motive, using one or more modes of transport",
+    # Mexico city
     "Moving from one part to another with a specific reason/motive, using one or multiples modes of transport",
+    # Bogota
     "Moving from one part to another with a specific reason/motive, a definite hour of start and end, a mode of transport, and a duration greater than 3 minutes.
@@
     "",
+    # Mexico city
                        "Trips made during the week (Tuesday, Wednesday, Thursday) and in Saturdays (weekends)",
+    # Bogota
                        "Trips made the day of reference, i.e., the day before the survey (from 4am yesterday to 4am today). Surveys made from Monday to Sunday (according to the dataset).",
@@
   # Is there information at stage level?
-  Trip_Stage = c("Stage", "Trip", "Stage", "Stage", "Trip", 
+  Trip_Stage = c("Stage", "Trip", "Trip", "Stage", "Trip", 
                  "Trip", "Trip", "Stage", "Stage", "Trip",
@@
   # If not, what is missing?
-  Stage_missing = c("N/A", 
+  Stage_missing = c(
+    "N/A", 
     "There's no stage duration", 
+    "No stage information", 
+    # Mexico city
     "N/A", 
-              "N/A", 
+    # Bogota
     "There's no stage duration",
@@
     "Not sure, because it asks about modes and duration but don't know if people answered with this level of detail.",
-              "Looks like the survey capture up to 5 stage modes, but duration was not asked."),
+    "Looks like the survey capture up to 5 stage modes, but duration was not asked."
+    ),
   # Is there information about short walks to pt
@@
                       "Yes",
-                      "Yes",
+                      "No",
                       "Yes",
@@
 Walking_in_trip_duration = c("Not sure, yet.",
-                             "",
-                             "",
-                             "",
-                             "",
-                             "",
-                             "",
-                             "",
-                             "",
-                             "",
-                             ""),
+                             "Yes",
+                             "N/A",
+                             "Yes",
+                             "Yes",
+                             "Not included",
+                             "N/A",
+                             "Not sure, yet.",
+                             "Yes",
+                             "Not sure, yet",
+                             "No"),
   ToDo = c("Need to check if injuries dataset covers the same area that travel survey does. \n There's information about walking stages (blocks not minutes) but is not used",
            "Need to check if injuries dataset covers the same area that travel survey does. \n Travel survey from 2012, there is one from 2017.",
+           # Belo horizonte: page 81, "Codebook and protocol.pdf"
            "Need to check if injuries dataset covers the same area that travel survey does.

### 2021-03-24 code/producing_trips_rd/trip-summary.Rmd
Adding guess_max option when reading datasets

@@
 ```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
-trip <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv")
+trip <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv")
+trip <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv")
+trip <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv")
+trip <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/bogota/bogota_trips.csv")
+trip <- read_csv("../../data/local/bogota/bogota_trips.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/santiago/santiago_trip.csv")
+trip <- read_csv("../../data/local/santiago/santiago_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/vizag/vizag_trip.csv")
+trip <- read_csv("../../data/local/vizag/vizag_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/delhi/delhi_trip.csv")
+trip <- read_csv("../../data/local/delhi/delhi_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv")
+trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/accra/accra_trip.csv")
+trip <- read_csv("../../data/local/accra/accra_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)
@@
 
-trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv")
+trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv",
+                 guess_max = 100000)
 quality_check(trip)

### 2021-04-13 code/producing_trips_rd/trip-summary.Rmd
Updating trip summary

@@
 ```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
-data.frame(
-  Ciudad = c("Buenos Aires", "Sao Paulo", "Belo Horizonte", "Mexico", "Bogota",
-             "Santiago", "Visakhapatnam", "Delhi", "Bangalore", "Accra",
-             "Cape Town"),
-  Study_area = c("BA and 27 municipalities.",
-                 "Metropolitan region of SP.",
-                 "Metropolitan region of BH.",
-                 "Metropolitan area of Mexico city.",
-                 "Bogota and surrounding municipalities.",
-                 "Metropolitan region of Santiago.",
-                 "There's not a report with survey details.",
-                 "There's not a report with survey details.",
-                 "Not sure",
-                 "The whole country but filtered only Greater Accra",
-                 "Cape Town, Malmesbury and Paarl"),
-  year = c("2010",
-           "2017",
-           "2012",
-           "2017",
-           "2019",
-           "2012",
-           "",
-           "",
-           "",
-           "",
-           ""),
-  # Definition of a trip
-  Trip_definition = c(
-    # BA: Pag 23- Report.pdf
-    "Moving from a origin to a destiny with a specific reason/motive",
-                      "Moving for a specific reason between two specific points (origin and destination), using one or more modes of transport.
-Walking trips where the reason for the trip is work or school, regardless of the distance travelled; or the distance covered is more than 500 mts for other reasons.",
-    # Belo Horizonte: pag 16, "Codebook and protocol.pdf"
-    "Moving from one part to another with a specific reason/motive, using one or more modes of transport",
-    # Mexico city
-    "Moving from one part to another with a specific reason/motive, using one or multiples modes of transport",
-    # Bogota
-    "Moving from one part to another with a specific reason/motive, a definite hour of start and end, a mode of transport, and a duration greater than 3 minutes.
-Or moving from one part to another with reason/motive work or study of any duration",
-                      "Any movement carried out on public roads with a purpose
-determined, between two places (origin and destination) at a certain time of day; It can be carried out in several modes of transport and consist of one or more stages",
-"",
-"",
-"",
-"",
-""),
-  Reference_period = c(
-    # BA: page 4, report.pdf
-    "Trips made in working days",
-    "Trips made the day before the survey (from 4am yesterday to 3:59am today).",
-    "",
-    # Mexico city
-                       "Trips made during the week (Tuesday, Wednesday, Thursday) and in Saturdays (weekends)",
-    # Bogota
-                       "Trips made the day of reference, i.e., the day before the survey (from 4am yesterday to 4am today). Surveys made from Monday to Sunday (according to the dataset).",
-                       "Trips made in working days (regular season), in weekends (regular season) and in working days(summer season)",
-                       "",
-                       "",
-                       "",
-                       "",
-                       ""),
-  # Is there information at stage level?
-  Trip_Stage = c("Stage", "Trip", "Trip", "Stage", "Trip", 
-                 "Trip", "Trip", "Stage", "Stage", "Trip",
-                 "Trip"),
-  # If not, what is missing?
-  Stage_missing = c(
-    "N/A", 
-    "There's no stage duration", 
-    "No stage information", 
-    # Mexico city
-    "N/A", 
-    # Bogota
-    "There's no stage duration",
-    "There's no stage duration",
-    "There's no stage mode, duration",
-    "N/A",
-    "N/A",
-    "Not sure, because it asks about modes and duration but don't know if people answered with this level of detail.",
-    "Looks like the survey capture up to 5 stage modes, but duration was not asked."
-    ),
-  # Is there information about short walks to pt
-  Walk_to_from_pt = c("Yes, in blocks (not minutes)",
-                      "Yes",
-                      "No",
-                      "Yes",
-                      "Yes",
-                      "Yes",
-                      "No",
-                      "Not sure",
-                      "Yes",
-                      "Not sure, because it asks about modes and duration but don't know if people answered with this level of detail.",
-                      "No"),
-Walking_in_trip_duration = c("Not sure, yet.",
-                             "Yes",
-                             "N/A",
-                             "Yes",
-                             "Yes",
-                             "Not included",
-                             "N/A",
-                             "Not sure, yet.",
-                             "Yes",
-                             "Not sure, yet",
-                             "No"),
-  ToDo = c("Need to check if injuries dataset covers the same area that travel survey does. \n There's information about walking stages (blocks not minutes) but is not used",
-           "Need to check if injuries dataset covers the same area that travel survey does. \n Travel survey from 2012, there is one from 2017.",
-           # Belo horizonte: page 81, "Codebook and protocol.pdf"
-           "Need to check if injuries dataset covers the same area that travel survey does.
-\n According to the documentation, it looks like there's information at stage level (including stage duration).",
-           "Need to check if injuries dataset covers the same area that travel survey does.
-Trip mode is defined in terms of speed and duration, while in other surveys there's a hierarchy (check this).
-There's information on weekdays and weekends, decide what to do with this.",
-           "Need to check if injuries dataset covers the same area that travel survey does (The one that I got it's from only Bogota). Need to update travel survey.",
-           "Need to check if injuries dataset covers the same area that travel survey does.
-There's information about trips on working days in regular season and and working days in summer season, decide what to do with this.",
-           "Need to find a report to check the details of this survey",
-           "Need to find a report to check the details of this survey",
-           "There's a report called Comprehensive Traffic and Transportation Study for Bangalore Metropolitan region but I'm not sure if it corresponds to the data.
-Trip mode is defined in terms of speed and duration, while in other surveys there's a hierarchy (check this)",
-           "Need to check if injuries dataset covers the same area that time use survey does. Need to check if stages were captured.",
-           "Need to check if injuries dataset covers the same area that time use survey does."),
-           # Is the data ok?
-  OK = c("--", "--", "--", "--", "--", 
-         "--", "--", "--", "--", "--", "--")
+(readxl::read_excel("C:/Users/danie/Documents/Daniel_Gil/Consultorias/2020/WorldBank/ITHIM-R/code/producing_trips_rd/Travel_survey_details.xlsx")
   ) %>% kbl() %>% kable_classic()

### 2021-04-14 summary_tables.Rmd
Add space
Updating summary tables before imputation

@@
 |---|---|---|---|
-| pedestrian | 74.9|23.0|1.0 |
-| cycle | 2.3|9.8|1.2 |
+| pedestrian | 75.0|23.0|1.0 |
+| cycle | 2.3|9.8|1.1 |
 | car | 1.0|4.4|6.4 |
 | motorcycle | 5.1|33.9|63.2 |
-| bus | 15.1|24.4|25.0 |
+| bus | 15.1|24.3|25.0 |
 
@@
 |---|---|---|---|
-| pedestrian | 80.5|10.7|1.0 |
-| cycle | 7.9|11.3|2.0 |
-| car | 3.5|21.6|19.7 |
-| motorcycle | 1.4|7.7|6.2 |
-| bus | 4.9|41.6|68.3 |
+| pedestrian | 91.0|20.9|1.2 |
+| cycle | 4.9|12.1|2.0 |
+| car | 1.5|26.7|29.3 |
+| motorcycle | 0.5|5.9|6.5 |
+| bus | 0.8|20.7|56.8 |
 
@@
 |---|---|---|---|
-| pedestrian | 79.7|3.0|0.1 |
-| cycle | 8.0|3.9|0.3 |
-| car | 6.2|35.0|21.8 |
-| motorcycle | 0.6|2.8|1.3 |
-| bus | 3.3|49.2|50.5 |
+| pedestrian | 80.2|2.9|0.1 |
+| cycle | 7.4|3.6|0.3 |
+| car | 6.2|34.5|21.4 |
+| motorcycle | 0.6|2.8|1.2 |
+| bus | 3.5|50.1|51.0 |
 
@@
 |---|---|---|---|
-| pedestrian | 89.6|6.5|0.1 |
-| cycle | 3.4|4.5|0.3 |
-| car | 2.2|34.9|29.7 |
-| motorcycle | 0.6|4.7|1.0 |
-| bus | 1.6|36.2|49.9 |
+| pedestrian | 89.9|3.5|0.0 |
+| cycle | 2.5|3.2|0.2 |
+| car | 2.2|37.9|31.8 |
+| motorcycle | 0.4|3.2|0.9 |
+| bus | 1.7|34.8|33.8 |
 
@@
 |---|---|---|---|
-| pedestrian | 83.9|4.1|0.1 |
-| cycle | 6.2|6.0|0.4 |
-| car | 7.7|54.8|30.0 |
-| motorcycle | 0.2|2.0|0.9 |
-| bus | 0.6|21.1|30.1 |
+| pedestrian | 84.0|3.9|0.0 |
+| cycle | 6.2|6.1|0.5 |
+| car | 8.1|58.7|32.3 |
+| motorcycle | 0.1|2.2|1.1 |
+| bus | 0.5|20.5|28.8 |
 
@@
 |---|---|---|---|
-| pedestrian | 62.2|7.2|3.0 |
-| cycle | 0.7|0.9|0.3 |
-| car | 22.0|45.1|29.9 |
-| motorcycle | 1.3|2.9|3.8 |
-| bus | 13.3|39.2|32.7 |
+| pedestrian | 82.7|1.9|0.0 |
+| cycle | 1.3|1.4|0.1 |
+| car | 12.6|65.4|37.0 |
+| motorcycle | 0.8|3.2|1.9 |
+| bus | 1.0|21.3|28.9 |
 

### 2021-04-22 code/producing_trips_rd/trip-summary.Rmd
Removed first table trip summary

@@
 ---
-title: "Trip Summary ITHIM-R"
+title: "Baseline Travel Survey Summary ITHIM-R"
 author: ""
@@
 ```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
-(readxl::read_excel("C:/Users/danie/Documents/Daniel_Gil/Consultorias/2020/WorldBank/ITHIM-R/code/producing_trips_rd/Travel_survey_details.xlsx")
-  ) %>% kbl() %>% kable_classic()
+# (readxl::read_excel("C:/Users/danie/Documents/Daniel_Gil/Consultorias/2020/WorldBank/ITHIM-R/code/producing_trips_rd/Travel_survey_details.xlsx")
+#   ) %>% kbl() %>% kable_classic()
 
@@
 ```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
-trip <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv",
+trip_buenos_aires <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv",
                  guess_max = 100000)
-quality_check(trip)
+quality_check(trip_buenos_aires)
 trip_summary$Buenos_Aires <- value
 
-trip <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv",
+trip_sao_paulo <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv",
                  guess_max = 100000)
-quality_check(trip)
+quality_check(trip_sao_paulo)
 trip_summary$Sao_Paulo <- value
 
-trip <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Belo_Horizonte <- value
-
-trip <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Mexico <- value
-
-trip <- read_csv("../../data/local/bogota/bogota_trips.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Bogota <- value
-
-trip <- read_csv("../../data/local/santiago/santiago_trip.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Santiago <- value
-
-trip <- read_csv("../../data/local/vizag/vizag_trip.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Visakhapatnam <- value
-
-trip <- read_csv("../../data/local/delhi/delhi_trip.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Delhi <- value
-
-trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Bangalore <- value
-
-trip <- read_csv("../../data/local/accra/accra_trip.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Accra <- value
-
-trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv",
-                 guess_max = 100000)
-quality_check(trip)
-trip_summary$Cape_Town <- value
+# trip_belo_horizonte <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv",
+#                  guess_max = 100000)
+# quality_check(trip_belo_horizonte)
+# trip_summary$Belo_Horizonte <- value
+# 
+# trip_mexico_city <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv",
+#                  guess_max = 100000)
+# quality_check(trip_mexico_city)
+# trip_summary$Mexico <- value
+# 
+# trip_bogota <- read_csv("../../data/local/bogota/bogota_trips.csv",
+#                  guess_max = 100000)
+# quality_check(trip_bogota)
+# trip_summary$Bogota <- value
+# 
+# trip_santiago <- read_csv("../../data/local/santiago/santiago_trip.csv",
+#                  guess_max = 100000)
+# quality_check(trip_santiago)
+# trip_summary$Santiago <- value
+# 
+# trip_vizag <- read_csv("../../data/local/vizag/vizag_trip.csv",
+#                  guess_max = 100000)
+# quality_check(trip_vizag)
+# trip_summary$Visakhapatnam <- value
+# 
+# trip_delhi <- read_csv("../../data/local/delhi/delhi_trip.csv",
+#                  guess_max = 100000)
+# quality_check(trip_delhi)
+# trip_summary$Delhi <- value
+# 
+# trip_bangalore <- read_csv("../../data/local/bangalore/bangalore_trip.csv",
+#                  guess_max = 100000)
+# quality_check(trip_bangalore)
+# trip_summary$Bangalore <- value
+# 
+# trip_accra <- read_csv("../../data/local/accra/accra_trip.csv",
+#                  guess_max = 100000)
+# quality_check(trip_accra)
+# trip_summary$Accra <- value
+# 
+# trip_cape_town <- read_csv("../../data/local/cape_town/cape_town_trip.csv",
+#                  guess_max = 100000)
+# quality_check(trip_cape_town)
+# trip_summary$Cape_Town <- value
+#
 
@@
 ```
+
+```{r, include = TRUE,  echo= FALSE, results='asis'}
+rownames(trip_summary)
+trip_summary[50:61,]
+```

### 2021-04-26 summary_tables_PA_AP.Rmd
Intermediate results

--- 2020-09-04 summary_tables_PA_AP.Rmd
@@
 ---
-title: "Summary Tables ITHIMR"
-author: "Rahul Goel, MRC Epidemiology Unit, University of Cambridge, UK"
+title: "AP/PA Tables ITHIM Global"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:
-  html_document: default
+  html_document: 
+    toc: true
+    toc_depth: 5
+    toc_float: true
   word_document: default
@@
 These are the summary tables of the following items
+
 1) Individual-level PM2.5 concentrations for baseline and scenarios
-2) Individual-level physical activity for baseline and scenarios
-3) Distance mode shares for baseline and scenarios
+2) Overall CO2 emissions for baseline and scenarios
+3) Individual-level physical activity for baseline and scenarios
 
@@
 library(dplyr)
+library(plotly)
 
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
 io <- readRDS("results/multi_city/io.rds")
@@
 cities <- names(io)[!names(io) %in% 'scen_prop']
-high<- c(70, 25, 150,60)
-low<- c(40,15,100,40)
-for (x in 1: length(cities))
-{
 
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, baseline:bus_scen, factor_key=TRUE)
-  y<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")+ylim(low[x],high[x])
-  print(y)
+all_inputs <- read.csv('all_city_parameter_inputs.csv',stringsAsFactors = F)
 
+#all_inputs$cape_town <- all_inputs$accra
+#all_inputs$vizag <- all_inputs$sao_paulo
+
+parameter_names <- all_inputs$parameter
+parameter_starts <- which(parameter_names != '')
+parameter_stops <- c(parameter_starts[-1] - 1, nrow(all_inputs))
+parameter_names <- parameter_names[parameter_names != '']
+parameter_list <- list()
+compute_mode <- 'constant'
+for (i in 1:length(parameter_names)) {
+  parameter_list[[parameter_names[i]]] <- list()
+  parameter_index <- which(all_inputs$parameter == parameter_names[i])
+  if (all_inputs[parameter_index,2] == '') {
+    parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
+      city_index <- which(colnames(all_inputs) == x)
+      val <- all_inputs[parameter_index,city_index]
+      ifelse(val %in% c('T','F'),val,as.numeric(val))
+    })
+    names(parameter_list[[parameter_names[i]]]) <- cities
+  }else if (all_inputs[parameter_index,2] == 'constant') {
+    indices <- 0
+    if (compute_mode == 'sample') indices <- 1:2
+    parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
+      city_index <- which(colnames(all_inputs) == x)
+      val <- all_inputs[parameter_index + indices,city_index]
+      ifelse(val == '',0,as.numeric(val))
+    })
+    names(parameter_list[[parameter_names[i]]]) <- cities
+  }else{
+    parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
+      city_index <- which(colnames(all_inputs) == x)
+      if (any(all_inputs[parameter_starts[i]:parameter_stops[i],
+                         city_index] != '')) {
+        sublist_indices <- which(all_inputs[parameter_starts[i]:parameter_stops[i],city_index] != '')
+        thing <- as.list(as.numeric(c(all_inputs[parameter_starts[i]:parameter_stops[i],city_index])[sublist_indices]))
+        names(thing) <- c(all_inputs[parameter_starts[i]:parameter_stops[i],2])[sublist_indices]
+        thing
+      }
+    }
+    )
+    names(parameter_list[[parameter_names[i]]]) <- cities
+  }
+}
+
+for (i in 1:length(parameter_list))  assign(names(parameter_list)[i],parameter_list[[i]])
+```
+
+```{r}
+# high<- c(70, 25, 150,60)
+# low<- c(40,15,100,40)
+for (x in 1:length(cities)) {
+  # print(cities[x])
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
+                                                        "Bicycling", "Driving",
+                                                        "Motorcycling", 
+                                                        "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
+                      Baseline:`Public Transport`, factor_key = TRUE)
+  y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
+    geom_boxplot(outlier.shape = 8) + ggtitle(cities[x]) +
+    labs(y = "Daily PM2.5 Concentration", x = "Scenarios") #+ 
+    #ylim(min(data_long$pm_conc), max(data_long$pm_conc) + 10)
+    #+ylim(low[x],high[x])
+  yy <- ggplot() +
+    geom_density(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
+    ggtitle(cities[x]) +
+    labs(y = "Daily PM2.5 Concentration", x = "Scenarios")
+  #print(y)
+  #print(yy)
+  print(ggplotly(y))
+  print(ggplotly(yy))
 }
 ```
-
-# Descriptive tables of Individual-level PM2.5 Concentrations
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-
-for (x in 1: length(cities))
-{
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, baseline:bus_scen, factor_key=TRUE)
-  summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(pm_conc),'5th'=quantile(pm_conc, 0.05),'50th'=quantile(pm_conc, 0.5),'95th'=quantile(pm_conc, 0.9)))
-  summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1))
-  summary$change_PM<- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits=2)
-  names(summary)[1]<-"Scenario"
-  print(kable(summary, caption= cities[x]))
-
-}
-```
-
-<!-- # Descriptive tables of emission inventory -->
-
-<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
-
-<!-- trans_share<-c(22.5, 40.0, 22.5, 28.1)  -->
-<!-- pm_conc<- c(50, 18, 122, 47 ) -->
-<!-- city<-cities -->
-<!-- shares<- as.data.frame(cbind(city,as.numeric(trans_share), pm_conc)) -->
-<!-- names(shares)[2]<- "trans_share" -->
-<!-- for (x in 1:length(cities)) -->
-<!-- { -->
-<!--   modes<-names(unlist(io[[cities[x]]]$emission_inventory)) -->
-<!--   emissions<-as.data.frame(unlist(io[[cities[x]]]$emission_inventory)) -->
-<!--   city_emissions<-cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$emission_inventory)`)) -->
-<!--   names(city_emissions)[2]<- "emissions" -->
-<!--   select<- c("car", "motorcycle", "bus_driver", "truck", "big_truck") -->
-<!--   city_emissions$modes<- as.character(city_emissions$modes) -->
-<!--   city_emissions$modes[!(city_emissions$modes %in% select)]<- "other" -->
-<!--   summary<-city_emissions %>% group_by(modes) %>% summarise(sum(emissions)) -->
-<!--   names(summary)[2]<- "emissions" -->
-<!--   summary$emissions <- round(summary$emissions*100/sum(summary$emissions), digits=1) -->
-<!--   summary<-as.data.frame(summary) -->
-<!--   summary[nrow(summary)+1,1]<-"Transport share" -->
-<!--   summary[nrow(summary),2]<-as.character(shares$trans_share[x]) -->
-<!--   summary[nrow(summary)+1,1]<-"PM2.5 Conc" -->
-<!--   summary[nrow(summary),2]<-as.character(shares$pm_conc[x]) -->
-<!--   print(kable(summary, caption= cities[x])) -->
-
-<!-- } -->
-
-<!-- ``` -->
-
-# Descriptive tables of emission inventory
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-
-trans_share<-c(22.5, 40.0, 22.5, 28.1) 
-pm_conc<- c(50, 18, 122, 47 )
-city<-cities
-shares<- as.data.frame(cbind(city,as.numeric(trans_share), pm_conc))
-names(shares)[2]<- "trans_share"
-for (x in 1:length(cities))
-{
-  modes<-names(unlist(io[[cities[x]]]$emission_inventory))
-  emissions<-as.data.frame(unlist(io[[cities[x]]]$emission_inventory))
-  city_emissions<-cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$emission_inventory)`))
-  names(city_emissions)[2]<- "emissions"
-  select<- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
-  city_emissions$modes<- as.character(city_emissions$modes)
-  city_emissions$modes[!(city_emissions$modes %in% select)]<- "other"
-  summary<-city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
-  names(summary)[2]<- cities[x]
-  summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits=1)
-  summary<-as.data.frame(summary)
-  summary[nrow(summary)+1,1]<-"Transport share"
-  summary[nrow(summary),2]<-as.character(shares$trans_share[x])
-  summary[nrow(summary)+1,1]<-"PM2.5 Conc"
-  summary[nrow(summary),2]<-as.character(shares$pm_conc[x])
-  io[[cities[x]]]$summary_emission<- summary
-
-}
-
-summary_all<- cbind(io[[cities[1]]]$summary_emission,io[[cities[2]]]$summary_emission,io[[cities[3]]]$summary_emission,io[[cities[4]]]$summary_emission)
-summary_all<- summary_all[-c(3,5,7)]
-summary_all$modes[2]<- "bus"
-
-print(kable(summary_all))
-```
-
-# Boxplots of Individual-level Physical activity (MMETs)
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-limit=100
-for (x in 1: length(cities))
-{
-  names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE)
-  y<-ggplot(data_long, aes(x=scenario, y=mmet, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Weekly MMET", x = "Scenarios")+  ylim(0, limit)
-  print(y)
-
-}
-```
-
-# Descriptive tables of Individual-level Physical activity (MMETs)
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1: length(cities))
-{
-  names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE)
-  summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(mmet),'5th'=quantile(mmet, 0.05),'50th'=quantile(mmet, 0.5),'95th'=quantile(mmet, 0.9)))
-  summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1))
-  names(summary)[1]<-"Scenario"
-  print(kable(summary, caption= cities[x]))
-
-}
-```
-
-# Descriptive tables of injury outcomes
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1: length(cities))
-{
-
-  summary<- as.data.frame(io[[cities[x]]]$outcomes$injuries %>% group_by(scenario) %>% summarise('Ped'=sum(pedestrian,na.rm=T),'Bike'= sum(cycle), 'Car'=sum(car), "Motorcycle"=sum(motorcycle,na.rm=T), "All modes"=sum(Deaths,na.rm=T)))
-  summary$scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  summary<- cbind(summary$scenario ,round(summary[,2:6], digits=0))
-  names(summary)[1]<-"Scenario"
-  print(kable(summary, caption= cities[x]))
-
-}
-```
-
-# Striking vehicles in scenarios
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1: length(cities))
-{
-
-scenarios<- c('Baseline', 'Scenario 1', 'Scenario 2', 'Scenario 3', 'Scenario 4', 'Scenario 5')
-for (i in 1: length(scenarios))
-{
-whw<-as.data.frame(rowSums(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$whw))
-noov<-as.data.frame(sum(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$nov))
-names(whw)[1]<- as.character(scenarios[i])
-names(noov)[1]<-as.character(scenarios[i])
-
-if (i ==1 )
-{
- summary<-round(rbind(whw,nov=noov), digits = 1)
-} 
-else
-{
-summary<- cbind(summary, round(rbind(whw, nov=noov), digits=1) ) 
-}
-
-}
-
-names(summary)<-c("base", "walk", "bike", "car", "MC", "bus")
-print(kable(summary, caption= cities[x]))
-
-}
-
-```
-
-# Distance share by scenarios in percentages (excludes city-specific minority modes-- auto rickshaws, subway, etc.)
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1: length(cities))
-{
-
-  select<- c("car", "motorcycle", "walking", "cycle", "bus")
-
-  io[[cities[x]]]$trip_scen_sets$trip_mode[!io[[cities[x]]]$trip_scen_sets$trip_mode %in% select]<- "other"
-
-  dist_scen<- io[[cities[x]]]$trip_scen_sets %>% group_by(trip_mode, scenario) %>% summarise(sum(trip_distance))
-  dist_scen<- as.data.frame (dist_scen)
-  names(dist_scen)[3]<-"distance"
-  dist_scen$distance<- round(dist_scen$distance, digits=0)
-    dist_scen<-spread(dist_scen, trip_mode, distance)
-  dist_scen$sum<-rowSums (dist_scen[2:(ncol(dist_scen))], na.rm = FALSE, dims = 1)
-  dist_scen_prop<- round(dist_scen[,2:(ncol(dist_scen))]*100/dist_scen$sum, digits=1)
-  dist_scen_prop<- cbind(dist_scen[,1], dist_scen_prop)
-
-  names(dist_scen_prop)[1]<-"Scenario"
-  dist_scen_prop$Scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  print(kable(dist_scen_prop, caption= cities[x]))
-  print(kable(dist_scen, caption= cities[x]))
-
-}
-```

### 2021-04-26 summary_tables_PA_AP.Rmd
Intermediate code

--- 2020-09-30 summary_tables_PA_AP.Rmd
@@
 output:
-  html_document: default
+  html_document: 
+    toc: true
+    toc_depth: 5
+    toc_float: true
   word_document: default
@@
 library(dplyr)
+library(plotly)
+library(htmltools)
 
@@
 parameter_names <- all_inputs$parameter
-parameter_starts <- which(parameter_names!='')
+parameter_starts <- which(parameter_names != '')
 parameter_stops <- c(parameter_starts[-1] - 1, nrow(all_inputs))
-parameter_names <- parameter_names[parameter_names!='']
+parameter_names <- parameter_names[parameter_names != '']
 parameter_list <- list()
 compute_mode <- 'constant'
-for(i in 1:length(parameter_names)){
+for (i in 1:length(parameter_names)) {
   parameter_list[[parameter_names[i]]] <- list()
-  parameter_index <- which(all_inputs$parameter==parameter_names[i])
-  if(all_inputs[parameter_index,2]=='')  {
+  parameter_index <- which(all_inputs$parameter == parameter_names[i])
+  if (all_inputs[parameter_index,2] == '') {
     parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
-      city_index <- which(colnames(all_inputs)==x)
+      city_index <- which(colnames(all_inputs) == x)
       val <- all_inputs[parameter_index,city_index]
-      ifelse(val%in%c('T','F'),val,as.numeric(val))
+      ifelse(val %in% c('T','F'),val,as.numeric(val))
     })
     names(parameter_list[[parameter_names[i]]]) <- cities
-  }else if(all_inputs[parameter_index,2]=='constant'){
+  }else if (all_inputs[parameter_index,2] == 'constant') {
     indices <- 0
-    if(compute_mode=='sample') indices <- 1:2
+    if (compute_mode == 'sample') indices <- 1:2
     parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
-      city_index <- which(colnames(all_inputs)==x)
-      val <- all_inputs[parameter_index+indices,city_index]
-      ifelse(val=='',0,as.numeric(val))
+      city_index <- which(colnames(all_inputs) == x)
+      val <- all_inputs[parameter_index + indices,city_index]
+      ifelse(val == '',0,as.numeric(val))
     })
@@
     parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
-      city_index <- which(colnames(all_inputs)==x)
-      if(any(all_inputs[parameter_starts[i]:parameter_stops[i],city_index]!='')){
-        sublist_indices <- which(all_inputs[parameter_starts[i]:parameter_stops[i],city_index]!='')
+      city_index <- which(colnames(all_inputs) == x)
+      if (any(all_inputs[parameter_starts[i]:parameter_stops[i],
+                         city_index] != '')) {
+        sublist_indices <- which(all_inputs[parameter_starts[i]:parameter_stops[i],city_index] != '')
         thing <- as.list(as.numeric(c(all_inputs[parameter_starts[i]:parameter_stops[i],city_index])[sublist_indices]))
@@
 
-for(i in 1:length(parameter_list)) assign(names(parameter_list)[i],parameter_list[[i]])
+for (i in 1:length(parameter_list))  assign(names(parameter_list)[i],parameter_list[[i]])
+```
 
+```{r}
 # high<- c(70, 25, 150,60)
 # low<- c(40,15,100,40)
-for (x in 1: length(cities))
-{
+# l <- htmltools::tagList()
+for (x in 1:length(cities)) {
   # print(cities[x])
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, Baseline:`Public Transport`, factor_key=TRUE)
-  y<-ggplot(data_long, aes(x=scenario, y=pm_conc, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Daily PM2.5 Concentration", x = "Scenarios")
-  #+ylim(low[x],high[x])
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
+                                                        "Bicycling", "Driving",
+                                                        "Motorcycling", 
+                                                        "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
+                      Baseline:`Public Transport`, factor_key = TRUE)
+  y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
+    geom_boxplot(outlier.shape = 8) + ggtitle(cities[x]) +
+    labs(y = "Daily PM2.5 Concentration", x = "Scenarios") #+
+  # +ylim(low[x],high[x])
+  # yy <- ggplot() +
+  #   geom_density(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
+  #   ggtitle(cities[x]) +
+  #   labs(y = "Daily PM2.5 Concentration", x = "Scenarios")
   print(y)
-
+  #print(yy)
+  # l[[i]] <- ggplotly(ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
+  #   geom_boxplot(outlier.shape = 8) + ggtitle(cities[x]) +
+  #   labs(y = "Daily PM2.5 Concentration", x = "Scenarios"))
+  # print(ggplotly(yy))
 }

### 2021-04-27 code/producing_trips_rd/trip-summary.Rmd
Updating summary files

--- 2021-04-13 code/producing_trips_rd/trip-summary.Rmd
@@
 ```
+
 ```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
-(readxl::read_excel("C:/Users/danie/Documents/Daniel_Gil/Consultorias/2020/WorldBank/ITHIM-R/code/producing_trips_rd/Travel_survey_details.xlsx")
-  ) %>% kbl() %>% kable_classic()
+# (readxl::read_excel("C:/Users/danie/Documents/Daniel_Gil/Consultorias/2020/WorldBank/ITHIM-R/code/producing_trips_rd/Travel_survey_details.xlsx")
+#   ) %>% kbl() %>% kable_classic()
 
@@
 trip_summary$Cape_Town <- value
-
 ```
 
+## summary table
 ```{r, include = TRUE,  echo= FALSE, results='asis'}

### 2021-04-27 summary_tables.Rmd
Updating summary files

@@
 ```
-
+# **1st option**
 ## Scenario definition 
@@
 ```
+
+# **2nd option**
+
+## Scenario definition 
+Table displays highest propensity for each distance category, for a given mode.
+
+Table displays highest propensity (%) for each distance category, for a given mode.
+
+In scenario generation, trips are sampled without replacement to increase the share of the requested mode up to the total shown.
+
+Distance categories are:
+
+0-1 km = \{trip distance < 2\}
+
+2-5 km = \{2 <= trip distance < 6\}
+
+6+ km = \{6 <= trip distance\}
+
+```{r scen_prop = "asis"}
+
+scen_prop <- round(io$scen_prop, round_to)
+
+rownames(scen_prop) <- paste(rownames(io$scen_prop), "scen", sep = "_")
+
+kable(scen_prop, headings = "Scenario Proportions")
+ggplot(as.data.frame(scen_prop) %>% rownames_to_column("scenario") %>% 
+         pivot_longer(!scenario, names_to = "distance_cat", 
+                      values_to = "values") %>% 
+         mutate(scenario = factor(scenario, 
+                                  levels = c("pedestrian_scen", "cycle_scen",
+                                             "car_scen", "motorcycle_scen",
+                                             "bus_scen"))
+                ), 
+       aes(scenario, values, fill = distance_cat)) + geom_col() + 
+  facet_grid(~ distance_cat) +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
+        legend.position = "none")
+```
+
+## Case study specific propensities
+
+|accra|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 95.5|27.9|2.6 |
+| cycle | 0.6|0.7|0.4 |
+| car | 0.9|12.4|17.2 |
+| motorcycle | 0.0|0.0|1.3 |
+| bus | 2.4|48.3|71.8 |
+
+|bangalore|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 75.0|23.0|1.0 |
+| cycle | 2.3|9.8|1.1 |
+| car | 1.0|4.4|6.4 |
+| motorcycle | 5.1|33.9|63.2 |
+| bus | 15.1|24.3|25.0 |
+
+|belo_horizonte|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 89.2|8.7|0.3 |
+| cycle | 1.6|2.7|0.2 |
+| car | 7.3|65.6|47.1 |
+| motorcycle | 1.3|10.1|5.3 |
+| bus | 0.4|11.1|43.1 |
+
+|bogota|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 91.0|20.9|1.2 |
+| cycle | 4.9|12.1|2.0 |
+| car | 1.5|26.7|29.3 |
+| motorcycle | 0.5|5.9|6.5 |
+| bus | 0.8|20.7|56.8 |
+
+|buenos_aires|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 80.2|2.9|0.1 |
+| cycle | 7.4|3.6|0.3 |
+| car | 6.2|34.5|21.4 |
+| motorcycle | 0.6|2.8|1.2 |
+| bus | 3.5|50.1|51.0 |
+
+|cape_town|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 84.1|5.3|1.9 |
+| cycle | 0.2|0.5|0.1 |
+| car | 8.0|55.6|48.3 |
+| motorcycle | 0.0|0.7|0.4 |
+| bus | 6.9|34.8|33.3 |
+
+|delhi|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 84.4|10.1|0.9 |
+| cycle | 2.0|7.6|3.7 |
+| car | 2.4|17.3|21.4 |
+| motorcycle | 5.6|30.1|26.5 |
+| bus | 0.7|24.7|32.7 |
+
+|mexico_city|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 89.9|3.5|0.0 |
+| cycle | 2.5|3.2|0.2 |
+| car | 2.2|37.9|31.8 |
+| motorcycle | 0.4|3.2|0.9 |
+| bus | 1.7|34.8|33.8 |
+
+|santiago|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 84.0|3.9|0.0 |
+| cycle | 6.2|6.1|0.5 |
+| car | 8.1|58.7|32.3 |
+| motorcycle | 0.1|2.2|1.1 |
+| bus | 0.5|20.5|28.8 |
+
+|sao_paulo|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 82.7|1.9|0.0 |
+| cycle | 1.3|1.4|0.1 |
+| car | 12.6|65.4|37.0 |
+| motorcycle | 0.8|3.2|1.9 |
+| bus | 1.0|21.3|28.9 |
+
+|vizag|0-1 km|2-5 km|6+ km|
+|---|---|---|---|
+| pedestrian | 86.2|19.7|0.1 |
+| cycle | 2.6|6.6|2.1 |
+| car | 0.6|2.4|2.8 |
+| motorcycle | 6.9|24.3|24.6 |
+| bus | 0.8|26.9|55.5 |
+
+## Trip proportion
+Case study specific trip proportions by mode, for baseline and five scenarios
+
+```{r load_tidyverse2, echo = F, message = F}
+suppressWarnings({
+  require(tidyverse)  
+})
+```
+
+```{r trip_mode_dist = "asis", echo = F, message = F}
+
+trip_prop <- list()
+
+for (city in cities) {
+  cat("\n \n")
+  cat( paste0('### **', city, '** \n' )) 
+  df <- io[[city]]$trip_scen_sets
+  u_trips <- df %>% dplyr::filter(scenario == "Baseline") %>% 
+    summarise(uid = n_distinct(trip_id)) %>% as.numeric()
+  td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% 
+    group_by(trip_mode, scenario) %>% 
+    summarise(p = round(dplyr::n() / u_trips * 100, 1)) %>% 
+    spread(key = trip_mode, value = p) %>% 
+    mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
+  td <- as.data.frame(t(td))
+  names(td) <- lapply(append('baseline', rownames(scen_prop)), as.character)
+  td <- td[-1, ]
+  x <- match(row.names(td), trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+
+  td1 <- td %>% rownames_to_column()
+
+  names(td1)[1] <- 'stage_mode'
+
+  if (length(trip_prop) == 0) {
+    trip_prop <- td1 %>% dplyr::select(stage_mode)
+  }
+
+  td1 <- td1 %>% dplyr::select(stage_mode, baseline)
+
+  names(td1)[2] <- city
+
+  trip_prop <- inner_join(trip_prop, td1, by = 'stage_mode')
+
+  y <- ggplot(td %>% rownames_to_column("stage_mode") %>% 
+         pivot_longer(!stage_mode, names_to = "scenario", 
+                      values_to = "proportion") %>% 
+           mutate(scenario = factor(scenario,
+                                    levels = c("baseline", "pedestrian_scen",
+                                               "cycle_scen", "car_scen",
+                                               "motorcycle_scen", "bus_scen")),
+                  stage_mode = factor(stage_mode,
+                                      levels = c("car", "motorcycle", "bus",
+                                                 "pedestrian", "taxi", "rail",
+                                                 "cycle")),
+                  proportion = as.numeric(proportion)) %>% 
+           dplyr::filter(stage_mode != "row_sums"), 
+       aes(scenario, proportion, fill = stage_mode)) + geom_col() + 
+  facet_grid(~ stage_mode) +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
+        legend.position = "none")
+
+  cat(" ")
+  print(y)
+  cat("\n ")
+  print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
+  cat("\n")
+}
+
+```
+
+### Baseline trip proportion (%)
+```{r}
+
+trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
+  mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>%
+  janitor::adorn_totals('row')
+
+ggplot(trip_prop %>% 
+         pivot_longer(!stage_mode, names_to = "city", 
+                      values_to = "proportion") %>% 
+         dplyr::filter(stage_mode != "Total") %>%
+         mutate(stage_mode = factor(stage_mode, 
+                                    levels = c("car", "motorcycle", "bus",
+                                               "pedestrian", "taxi", "rail",
+                                               "cycle"))), 
+       aes(stage_mode, proportion, fill = city)) + geom_col() + 
+  facet_grid(~ city) +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
+        legend.position = "none")
+
+kable(trip_prop, caption = 'Trip prop (%) for all case studies - only selected modes')
+
+```
+
+```{r unload_tidyverse2, echo = F, message = F}
+suppressWarnings({
+  detach("package:tidyverse", character.only = T)
+})
+
+```
+
+## Distance tables
+Case study specific distance tables for baseline and five scenarios
+
+```{r trip_dist = "asis"}
+
+trip_dist <- list()
+
+for (city in cities) {
+  cat("\n \n")
+  cat( paste0('### **', city, '** \n' )) 
+  count_people <- nrow(io[[city]]$synth_pop)
+  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+  row.names(td) <- NULL
+
+  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person")))
+  cat("\n")
+
+  td1 <- td
+
+  if (length(trip_dist) == 0){
+    trip_dist <- td1 %>% dplyr::select(stage_mode)
+  }
+
+  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
+
+  names(td1)[2] <- city
+
+  trip_dist <- inner_join(trip_dist, td1, by = 'stage_mode')
+
+  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
+  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+  row.names(td) <- NULL
+
+  y <- ggplot(td %>% 
+         pivot_longer(!stage_mode, names_to = "scenario", 
+                      values_to = "distance") %>% 
+           mutate(scenario = factor(scenario,
+                                    levels = c("Baseline", "pedestrian_scen",
+                                               "cycle_scen", "car_scen",
+                                               "motorcycle_scen", "bus_scen")),
+                  stage_mode = factor(stage_mode,
+                                      levels = c("car", "motorcycle", "bus",
+                                                 "pedestrian", "taxi", "rail",
+                                                 "cycle"))),
+         aes(scenario, distance, fill = stage_mode)) + geom_col() + 
+    facet_grid(~ stage_mode) +
+    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
+        legend.position = "none")
+
+  cat(" ")
+  print(y)
+  cat("\n ")
+  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips)")))
+  cat("\n") 
+}
+
+```
+
+### Baseline average distance (km) per day for all case studies
+```{r}
+
+trip_dist <- trip_dist %>% janitor::adorn_totals('row')
+
+ggplot(trip_dist %>% 
+         pivot_longer(!stage_mode, names_to = "city", 
+                      values_to = "distance") %>% 
+         mutate(stage_mode = factor(stage_mode, 
+                                    levels = c("car", "motorcycle", "bus",
+                                               "pedestrian", "taxi", "rail",
+                                               "cycle", "Total"))), 
+       aes(stage_mode, distance, fill = city)) + geom_col() + 
+  facet_grid(~ city) +
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
+        legend.position = "none")
+
+kable(trip_dist, caption = "Avg. distance per person (km) per day across all case studies")
+```
+
+## Distance tables scaled by using total population (per year in km)
+Case study specific distance tables for baseline and five scenarios
+
+```{r trip_tot_dist = "asis", message=F}
+
+bl_td <- list()
+
+for(city in cities){
+
+  print(city)
+  count_people <- round(sum(io[[city]]$demographic$population))
+  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 365, 3)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+
+  #
+
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+  row.names(td) <- NULL
+
+  if (length(bl_td) == 0){
+    bl_td <- td %>% dplyr::select(stage_mode)
+  }
+
+  td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
+
+  td1 <- td %>% dplyr::select(stage_mode, Baseline)
+
+  names(td1)[2] <- city
+
+  bl_td <- inner_join(bl_td, td1, by = 'stage_mode')
+
+  td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)
+
+  print(td1)
+
+  # print(kable(td, caption = paste("Distance table (km) for ", city, "( ", sitools::f2si(count_people), " ) per person")))
+  cat("\n")
+
+}
+
+```
+
+### Baseline total distance for all case studies (per year in km)
+
+```{r trip_dist_all_cities = "asis", message=F}
+
+# tab <- bl_td %>% gt() %>% fmt_number(columns = 2:ncol(bl_td), decimals = T, suffixing = T)
+
+backup <- bl_td
+
+td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')
+
+for (i in 2:ncol(td)){
+  td[, i] <- comprss(td[, i])
+
+}
+
+print(kable(td, caption = paste("Distance table (km)")))
+
+```
+
+## Distance by distance category
+### Tables
+
+```{r trip_dist_mode_figs = "asis"}
+
+for (city in cities){
+
+  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), prop = round(n_vals / nrow(.) * 100, 1))
+  print(kable(df, caption = paste("Distance by mode by for  ", city)))
+}
+
+```
+
+## True Distance by total population (per day in km)
+### Tables
+
+```{r trip_dist_mode_figs = "asis"}
+
+for (city in cities){
+
+  df <- io[[city]]$true_dist
+
+  names(df)[-1] <- lapply(append('baseline', rownames(scen_prop)), as.character)
+
+  print(kable(df, caption = paste("True distance by mode per day for  ", city)))
+}
+
+```
+
+<!-- ### Figures -->
+
+<!-- ```{r trip_dist_mode_figs = "asis"} -->
+
+<!-- for (city in cities){ -->
+
+<!--   df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline' & trip_mode %in% c('walking', 'bicycle', 'car', 'motorcycle', 'bus')) %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_mode, trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n()) %>%  -->
+<!--     group_by(trip_mode) %>% mutate(n_vals_1 = sum(n_vals)) %>%  -->
+<!--     mutate(prop = round(n_vals/n_vals_1 * 100, 1)) -->
+
+<!--   print(ggplot(data=df, aes(x=trip_distance_cat, y=prop)) + -->
+<!--           geom_bar(stat="identity", fill="steelblue")+ -->
+<!--           facet_wrap(vars(trip_mode)) + -->
+<!--           geom_text(aes(label=prop), vjust = -0.3, size = 3.5)+ -->
+<!--           theme_minimal() + -->
+<!--           theme(panel.spacing = unit(2, "lines")) + -->
+<!--           labs(title = paste(city, ' by dist by mode'))) -->
+<!-- } -->
+
+<!-- ``` -->
+
+## Duration tables
+Case study specific duration tables for baseline and five scenarios (per day in km)
+
+```{r trip_dur = "asis"}
+
+trip_dur <- list()
+
+l <- list()
+
+for(city in cities){
+  count_people <- nrow(io[[city]]$synth_pop)
+  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+  row.names(td) <- NULL
+
+  l[[city]] <- td
+
+  td1 <- td
+
+  if (length(trip_dur) == 0){
+    trip_dur <- td1 %>% dplyr::select(stage_mode)
+  }
+
+  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
+
+  names(td1)[2] <- city
+
+  trip_dur <- inner_join(trip_dur, td1, by = 'stage_mode')
+
+  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (everyone)")))
+
+  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
+  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  colnames(td)[3:7] <- rownames(scen_prop)
+
+  x <- match(td$stage_mode, trip_order$mode)
+  x[!is.na(x)]
+  td <- td[order(x),]
+  row.names(td) <- NULL
+
+  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (people with trips)")))
+  cat("\n")
+}
+
+```
+
+### Baseline average duration (mins) per day for all case studies
+```{r}
+
+trip_dur <- trip_dur %>% janitor::adorn_totals('row')
+
+kable(trip_dur, caption = "Avg. duration per person (mins) per day across all case studies")
+
+```

### 2021-04-27 summary_tables_PA_AP.Rmd
Updating summary files

@@
 library(dplyr)
-library(plotly)
-library(htmltools)
+#library(plotly)
+#library(htmltools)
 
@@
 <!-- ``` -->
+
+# **2nd option**
+
+# PM 2.5 summary and CO2
+```{r message=FALSE, warning=FALSE, echo=FALSE}
+# I have to keep the first chunk of code to run this part
+# high<- c(70, 25, 150,60)
+# low<- c(40,15,100,40)
+# l <- htmltools::tagList()
+for (x in 1:length(cities)) {
+  # print(cities[x])
+  cat("\n \n")
+  cat( paste0('## **', cities[x], '** \n' )) 
+  # Emission inventory
+  modes <- names(unlist(io[[cities[x]]]$PM_emission_inventory))
+  emissions <- as.data.frame(unlist(io[[cities[x]]]$PM_emission_inventory))
+  city_emissions <- cbind(as.data.frame(modes),                           as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))
+  names(city_emissions)[2] <- "emissions"
+  select <- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
+  city_emissions$modes <- as.character(city_emissions$modes)
+  city_emissions$modes[!(city_emissions$modes %in% select)] <- "other"
+  summary <- city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
+  names(summary)[2] <- cities[x]
+  summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits = 1)
+  summary <- as.data.frame(summary)
+  summary[nrow(summary)+1,1] <- "Transport share"
+  summary[nrow(summary),2] <- as.character(round(pm_trans_share[[cities[x]]] * 100))
+  summary[nrow(summary)+1,1] <- "PM2.5 Conc"
+  summary[nrow(summary),2] <- as.character(pm_conc_base[[cities[x]]])
+  io[[cities[x]]]$summary_emission <- summary
+
+  # PM 2.5 table and plot
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
+                                                        "Bicycling", "Driving",
+                                                        "Motorcycling", 
+                                                        "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
+                      Baseline:`Public Transport`, factor_key = TRUE)
+
+  summary2 <- as.data.frame(data_long %>% group_by(scenario) %>%
+                             summarise('mean' = mean(pm_conc),
+                                       '5th' = quantile(pm_conc, 0.05),
+                                       '50th' = quantile(pm_conc, 0.5),
+                                       '95th' = quantile(pm_conc, 0.9)))
+  summary2 <- cbind(summary2$scenario, round(summary2[,2:5], digits = 1))
+  summary2$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits = 2)
+  names(summary2)[1] <- "Scenario"
+
+  # Plot
+  y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
+    geom_boxplot(outlier.shape = 8) + ggtitle(cities[x]) +
+    labs(y = "Daily PM2.5 Concentration", x = "Scenarios")
+
+  # CO2
+  td <- round(colSums(io[[cities[x]]]$outcomes$co2_conc, na.rm = T), 1) %>%
+    as.data.frame() %>% tibble::rownames_to_column()
+  #names(td) <- c('Scenario', cities[x])
+  names(td) <- c('Scenario', 'CO2')
+  td$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling",
+                   "Public Transport")
+
+  cat("\n \n")
+  cat( paste('#### Boxplot of Individual-level PM2.5 concentrations \n' ))
+  print(y)
+
+  cat("\n \n")
+  cat( paste('#### Emission inventory \n' )) 
+  print(kable(summary))
+
+  cat("\n \n")
+  cat( paste('#### Descriptive table of Individual-level PM2.5 concentrations and CO2 emissions\n' ))
+  print(kable(cbind(summary2, td), caption = cities[x]))
+
+}
+```
+
+# Boxplots of Individual-level Physical activity (MMETs)
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+limit = 100
+for (x in 1:length(cities)) {
+  names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("Baseline","Walking",
+                                                   "Bicycling", "Driving",
+                                                   "Motorcycling", 
+                                                   "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Public Transport`, factor_key = TRUE)
+  y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) + geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 
+    labs(y = "Weekly MMET", x = "Scenarios") + ylim(0, limit)
+  print(y)
+
+}
+```

### 2021-05-04 code/producing_trips_rd/trip-summary.Rmd
Regen travel survey summary

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    toc_depth: 5
+    toc_float: true
+    number_sections: true
 ---
@@
 
-## summary table
+```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
+trip <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Buenos_Aires <- value
+
+trip <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Sao_Paulo <- value
+
+trip <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Belo_Horizonte <- value
+
+trip <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Mexico <- value
+
+trip <- read_csv("../../data/local/bogota/bogota_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Bogota <- value
+
+trip <- read_csv("../../data/local/santiago/santiago_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Santiago <- value
+
+trip <- read_csv("../../data/local/vizag/vizag_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Visakhapatnam <- value
+
+trip <- read_csv("../../data/local/delhi/delhi_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Delhi <- value
+
+trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Bangalore <- value
+
+trip <- read_csv("../../data/local/accra/accra_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Accra <- value
+
+trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Cape_Town <- value
+```
+
+## summary table (all rows)
 ```{r, include = TRUE,  echo= FALSE, results='asis'}
@@
 ```
+
+## summary table (only people between 15 and 69)
+```{r, include = TRUE,  echo= FALSE, results='asis'}
+kable(trip_summary2) %>%
+kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 14, fixed_thead = T)
+
+```

### 2021-05-04 code/producing_trips_rd/used_functions.R

@@ -86,9 +86,14 @@ trip_summary <- data.frame(row.names =       c("Meta Data",
                                                "Return home",
                                                "Other"))
 
+## trip_summary2 for dataset filtered in age range
+trip_summary2 <- trip_summary
 
 ## function_quality_check
-quality_check <- function(trip){
+quality_check <- function(trip, min_age = NA, max_age = NA){
+        if (!is.na(min_age) & !is.na(max_age))
+            trip <- trip %>% filter(age >= min_age & age <= max_age)
+        
         trip$trip_mode <- factor(trip$trip_mode, levels = c("bicycle","bus","car","metro", "motorcycle",
                                                             "other", "rickshaw", "taxi", "train", "truck", "van", "walk" ))
         

### 2021-05-18 injury_tables.Rmd

--- 2020-08-26 injury_tables.Rmd
@@
 
-  #htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
+  htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
   #ggsave(f, file=paste0("results/multi_city/whw_matrices/whw_", cs,".png"), width = 14, height = 10, units = "cm")
@@
 
-    el <- inner_join(el, td4, by = 'stage_mode')
+    el <- full_join(el, td4, by = 'stage_mode')
 
@@
 fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
 # ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
  fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
-print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
+ htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
+# print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
+# ggsave(d2, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_specific_mode.png"), limitsize = F, dpi = 300, scale = 1.5)
 fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
-#htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
+htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
 ```

### 2021-05-18 summary_tables.Rmd

@@
 |---|---|---|---|
-| pedestrian | 75.0|23.0|1.0 |
+| pedestrian | 74.8|23.0|1.0 |
 | cycle | 2.3|9.8|1.1 |
 | car | 1.0|4.4|6.4 |
-| motorcycle | 5.1|33.9|63.2 |
-| bus | 15.1|24.3|25.0 |
+| motorcycle | 5.1|33.9|62.9 |
+| bus | 15.1|24.3|24.9 |
 
@@
 |---|---|---|---|
-| pedestrian | 89.2|8.7|0.3 |
+| pedestrian | 89.1|8.6|0.3 |
 | cycle | 1.6|2.7|0.2 |
-| car | 7.3|65.6|47.1 |
-| motorcycle | 1.3|10.1|5.3 |
-| bus | 0.4|11.1|43.1 |
+| car | 7.3|65.1|43.9 |
+| motorcycle | 1.3|10.1|5.0 |
+| bus | 0.4|11.0|40.1 |
 
@@
 | pedestrian | 91.0|20.9|1.2 |
-| cycle | 4.9|12.1|2.0 |
-| car | 1.5|26.7|29.3 |
+| cycle | 4.8|12.0|2.0 |
+| car | 1.5|26.6|29.2 |
 | motorcycle | 0.5|5.9|6.5 |
-| bus | 0.8|20.7|56.8 |
+| bus | 0.8|20.6|56.7 |
 
@@
 |---|---|---|---|
-| pedestrian | 80.2|2.9|0.1 |
+| pedestrian | 80.1|2.9|0.1 |
 | cycle | 7.4|3.6|0.3 |
-| car | 6.2|34.5|21.4 |
+| car | 6.2|34.4|21.2 |
 | motorcycle | 0.6|2.8|1.2 |
-| bus | 3.5|50.1|51.0 |
+| bus | 3.5|50.0|50.7 |
 
@@
 |---|---|---|---|
-| pedestrian | 84.1|5.3|1.9 |
+| pedestrian | 84.0|5.3|1.9 |
 | cycle | 0.2|0.5|0.1 |
-| car | 8.0|55.6|48.3 |
+| car | 8.0|55.5|48.1 |
 | motorcycle | 0.0|0.7|0.4 |
-| bus | 6.9|34.8|33.3 |
+| bus | 6.9|34.8|33.1 |
 
@@
 | pedestrian | 89.9|3.5|0.0 |
-| cycle | 2.5|3.2|0.2 |
+| cycle | 2.4|3.2|0.2 |
 | car | 2.2|37.9|31.8 |
@@
 | motorcycle | 0.1|2.2|1.1 |
-| bus | 0.5|20.5|28.8 |
+| bus | 0.5|20.5|28.7 |
 
@@
 |---|---|---|---|
-| pedestrian | 82.7|1.9|0.0 |
+| pedestrian | 82.6|1.9|0.0 |
 | cycle | 1.3|1.4|0.1 |
-| car | 12.6|65.4|37.0 |
-| motorcycle | 0.8|3.2|1.9 |
-| bus | 1.0|21.3|28.9 |
+| car | 12.6|65.2|36.8 |
+| motorcycle | 0.8|3.1|1.9 |
+| bus | 1.0|21.3|28.8 |
 
@@
 |---|---|---|---|
-| pedestrian | 86.2|19.7|0.1 |
-| cycle | 2.6|6.6|2.1 |
+| pedestrian | 86.1|19.6|0.1 |
+| cycle | 2.6|6.6|2.0 |
 | car | 0.6|2.4|2.8 |
-| motorcycle | 6.9|24.3|24.6 |
-| bus | 0.8|26.9|55.5 |
+| motorcycle | 6.9|24.1|24.3 |
+| bus | 0.8|26.7|54.8 |
 

### 2021-05-18 summary_tables_PA_AP.Rmd
Adding one PA table

@@
 ```
+# Descriptive tables of Individual-level Physical activity (MMETs)
 
-<!-- # Descriptive tables of Individual-level Physical activity (MMETs) -->
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+for (x in 1:length(cities)) {
+  names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key = TRUE)
+  summary <- as.data.frame(data_long %>% group_by(scenario) %>% 
+                             summarise('mean' = mean(mmet),
+                                       '5th' = quantile(mmet, 0.05),
+                                       '50th' = quantile(mmet, 0.5),
+                                       '95th' = quantile(mmet, 0.9)))
+  summary <- cbind(summary$scenario ,round(summary[,2:5], digits = 1))
+  names(summary)[1] <- "Scenario"
+  print(kable(summary, caption = cities[x]))
 
-<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
-<!-- for (x in 1: length(cities)) -->
-<!-- { -->
-<!--   names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen") -->
-<!--   data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key=TRUE) -->
-<!--   summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(mmet),'5th'=quantile(mmet, 0.05),'50th'=quantile(mmet, 0.5),'95th'=quantile(mmet, 0.9))) -->
-<!--   summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1)) -->
-<!--   names(summary)[1]<-"Scenario" -->
-<!--   print(kable(summary, caption= cities[x])) -->
+}
+```
 
-<!-- } -->
-<!-- ``` -->
+# Descriptive table for Physical Activity (MMETs)
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
+cl <- list()
+for (city in cities)
+{
+
+  td <- round(colMeans(io[[city]]$outcomes$mmets[,5:10], na.rm = T), 1) %>% as.data.frame() %>% tibble::rownames_to_column()
+  names(td) <- c('Scenario', city)
+
+  if (length(cl) == 0)
+    cl <- td
+  else
+    cl <- left_join(cl, td)
+
+}
+
+cl$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+
+print(kable(cl))
+
+```
 

### 2021-05-19 summary_tables.Rmd
regen summary tables

@@
 |---|---|---|---|
-| pedestrian | 95.5|27.9|2.6 |
-| cycle | 0.6|0.7|0.4 |
-| car | 0.9|12.4|17.2 |
-| motorcycle | 0.0|0.0|1.3 |
-| bus | 2.4|48.3|71.8 |
+| pedestrian | 95.8|23.8|1.7 |
+| cycle | 0.6|0.5|0.2 |
+| car | 0.8|16.3|20.6 |
+| motorcycle | 0.0|0.0|3.1 |
+| bus | 2.4|47.1|62.9 |
 
@@
 |---|---|---|---|
-| pedestrian | 74.8|23.0|1.0 |
-| cycle | 2.3|9.8|1.1 |
-| car | 1.0|4.4|6.4 |
-| motorcycle | 5.1|33.9|62.9 |
-| bus | 15.1|24.3|24.9 |
+| pedestrian | 65.5|21.2|1.3 |
+| cycle | 2.0|8.7|1.0 |
+| car | 1.1|4.0|5.8 |
+| motorcycle | 5.8|31.7|54.0 |
+| bus | 23.9|30.6|34.6 |
 

### 2021-06-01 injury_tables.Rmd
Use full join instead

@@
 
-  htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
+  # htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
   #ggsave(f, file=paste0("results/multi_city/whw_matrices/whw_", cs,".png"), width = 14, height = 10, units = "cm")
@@
 
-    td4 <- left_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+    td4 <- full_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
 
@@
 fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
-htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
+# htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
 # ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
  fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
- htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
+ # htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
 # print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
@@
 fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
-htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
+# htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
 ```

### 2021-06-01 summary_tables.Rmd
Update propensities

@@
 | motorcycle | 0.0|0.0|3.1 |
-| bus | 2.4|47.1|62.9 |
+| bus | 2.4|47.2|62.9 |
 

### 2021-06-22 summary_tables.Rmd
Regen

--- 2021-05-18 summary_tables.Rmd
@@
 |---|---|---|---|
-| pedestrian | 95.5|27.9|2.6 |
-| cycle | 0.6|0.7|0.4 |
-| car | 0.9|12.4|17.2 |
-| motorcycle | 0.0|0.0|1.3 |
-| bus | 2.4|48.3|71.8 |
+| pedestrian | 95.8|23.8|1.7 |
+| cycle | 0.6|0.5|0.2 |
+| car | 0.8|16.3|20.6 |
+| motorcycle | 0.0|0.0|3.1 |
+| bus | 2.5|47.4|63.1 |
 
@@
 |---|---|---|---|
-| pedestrian | 74.8|23.0|1.0 |
-| cycle | 2.3|9.8|1.1 |
-| car | 1.0|4.4|6.4 |
-| motorcycle | 5.1|33.9|62.9 |
-| bus | 15.1|24.3|24.9 |
+| pedestrian | 65.5|21.2|1.3 |
+| cycle | 2.0|8.7|1.0 |
+| car | 1.1|4.0|5.8 |
+| motorcycle | 5.8|31.7|54.0 |
+| bus | 24.0|30.6|34.7 |
 
@@
 | motorcycle | 1.3|10.1|5.0 |
-| bus | 0.4|11.0|40.1 |
+| bus | 0.4|11.0|43.5 |
 
@@
 | motorcycle | 0.6|2.8|1.2 |
-| bus | 3.5|50.0|50.7 |
+| bus | 3.5|51.0|75.9 |
 
@@
 | motorcycle | 0.0|0.7|0.4 |
-| bus | 6.9|34.8|33.1 |
+| bus | 7.5|37.9|49.0 |
 
@@
 | motorcycle | 5.6|30.1|26.5 |
-| bus | 0.7|24.7|32.7 |
+| bus | 0.8|25.0|40.9 |
 
@@
 | motorcycle | 0.4|3.2|0.9 |
-| bus | 1.7|34.8|33.8 |
+| bus | 1.8|35.3|64.7 |
 
@@
 | motorcycle | 0.1|2.2|1.1 |
-| bus | 0.5|20.5|28.7 |
+| bus | 0.6|21.8|64.6 |
 
@@
 | motorcycle | 0.8|3.1|1.9 |
-| bus | 1.0|21.3|28.8 |
+| bus | 1.0|22.0|59.7 |
 

### 2021-07-07 injury_tables.Rmd
Regen removing age and sex in the fatalities' model
Accra, Belohorizonte, Bogota, Cape Town and Sao Paulo results should change because these are the cities where age and sex were removed.

@@
 require(tibble)
-overall_el <- list()
+overall_el <- list() # To save all cities results
+overall_dd <- list() # To save all cities results
 for (city in cities){
@@
 
-  el <- list()
+  el <- list() # To save city specific results
 
@@
     if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      # td1: Number of fatalities in nov
       td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      # td2: Number of fatalities in whw
       td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      # td3: Number of fatalities: sum of nov and whw
       td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
@@
     }
-
+    # td4: distance travelled per mode
     td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs) %>% as.data.frame()
@@
       el <- td4 %>% dplyr::select(stage_mode)
+      dd <- el
     }
-
+    # td4: merge total fatalities to distance
     td4 <- full_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
@@
     td4[, 2] <- as.numeric(td4[,2])
-
+    # Compute risk per bn km travelled
     td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 
@@
 
+    # City's population
+    pop <- sum(io[[city]]$demographic$population)
+    # Compute distance travelled per capita
+    td6 <- td5
+    td6[[paste0(var,"_risk")]] <- unlist(td4[2])
+    td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop
+    names(td6)[3] <- paste(names(td6)[2], names(td6)[3], sep = "_")
+    dd <- full_join(dd, td6, by = 'stage_mode')
   }
@@
   overall_el[[city]] <- td
-
+  overall_dd[[city]] <- dd
   cat("\n")
-
 }

### 2021-07-07 summary_tables.Rmd
Regen removing age and sex in the fatalities' model
Accra, Belohorizonte, Bogota, Cape Town and Sao Paulo results should change because these are the cities where age and sex were removed.

--- 2021-06-01 summary_tables.Rmd
@@
 
-|accra|0-1 km|2-5 km|6+ km|
+|accra|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 0.0|0.0|3.1 |
-| bus | 2.4|47.2|62.9 |
+| bus | 2.5|47.4|63.1 |
 
-|bangalore|0-1 km|2-5 km|6+ km|
+|bangalore|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 5.8|31.7|54.0 |
-| bus | 23.9|30.6|34.6 |
+| bus | 24.0|30.6|34.7 |
 
-|belo_horizonte|0-1 km|2-5 km|6+ km|
+|belo_horizonte|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 1.3|10.1|5.0 |
-| bus | 0.4|11.0|40.1 |
+| bus | 0.4|11.0|43.5 |
 
-|bogota|0-1 km|2-5 km|6+ km|
+|bogota|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 
-|buenos_aires|0-1 km|2-5 km|6+ km|
+|buenos_aires|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 0.6|2.8|1.2 |
-| bus | 3.5|50.0|50.7 |
+| bus | 3.5|51.0|75.9 |
 
-|cape_town|0-1 km|2-5 km|6+ km|
+|cape_town|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 0.0|0.7|0.4 |
-| bus | 6.9|34.8|33.1 |
+| bus | 7.5|37.9|49.0 |
 
-|delhi|0-1 km|2-5 km|6+ km|
+|delhi|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 5.6|30.1|26.5 |
-| bus | 0.7|24.7|32.7 |
+| bus | 0.8|25.0|40.9 |
 
-|mexico_city|0-1 km|2-5 km|6+ km|
+|mexico_city|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 0.4|3.2|0.9 |
-| bus | 1.7|34.8|33.8 |
+| bus | 1.8|35.3|64.7 |
 
-|santiago|0-1 km|2-5 km|6+ km|
+|santiago|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 0.1|2.2|1.1 |
-| bus | 0.5|20.5|28.7 |
+| bus | 0.6|21.8|64.6 |
 
-|sao_paulo|0-1 km|2-5 km|6+ km|
+|sao_paulo|0-2 km|2-6 km|6+ km|
 |---|---|---|---|
@@
 | motorcycle | 0.8|3.1|1.9 |
-| bus | 1.0|21.3|28.8 |
+| bus | 1.0|22.0|59.7 |
 
-|vizag|0-1 km|2-5 km|6+ km|
+|vizag|0-2 km|2-6 km|6+ km|
 |---|---|---|---|

### 2021-07-14 summary_tables.Rmd
Regen

@@
 |---|---|---|---|
-| pedestrian | 65.5|21.2|1.3 |
-| cycle | 2.0|8.7|1.0 |
-| car | 1.1|4.0|5.8 |
-| motorcycle | 5.8|31.7|54.0 |
-| bus | 24.0|30.6|34.7 |
+| pedestrian | 89.3|2.6|0.1 |
+| cycle | 3.3|10.1|0.6 |
+| car | 0.2|5.3|6.6 |
+| motorcycle | 1.0|35.4|60.9 |
+| bus | 4.4|40.2|29.7 |
 

### 2021-07-15 summary_tables.Rmd
Get propensities from get_scenario_settings instead

@@
 ## Case study specific propensities
+```{r message=F, error=F, warning=FALSE}
+get_scen_settings <- ithimr::get_scenario_settings()
 
-|accra|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 95.8|23.8|1.7 |
-| cycle | 0.6|0.5|0.2 |
-| car | 0.8|16.3|20.6 |
-| motorcycle | 0.0|0.0|3.1 |
-| bus | 2.5|47.4|63.1 |
-
-|bangalore|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 89.3|2.6|0.1 |
-| cycle | 3.3|10.1|0.6 |
-| car | 0.2|5.3|6.6 |
-| motorcycle | 1.0|35.4|60.9 |
-| bus | 4.4|40.2|29.7 |
-
-|belo_horizonte|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 89.1|8.6|0.3 |
-| cycle | 1.6|2.7|0.2 |
-| car | 7.3|65.1|43.9 |
-| motorcycle | 1.3|10.1|5.0 |
-| bus | 0.4|11.0|43.5 |
-
-|bogota|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 91.0|20.9|1.2 |
-| cycle | 4.8|12.0|2.0 |
-| car | 1.5|26.6|29.2 |
-| motorcycle | 0.5|5.9|6.5 |
-| bus | 0.8|20.6|56.7 |
-
-|buenos_aires|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 80.1|2.9|0.1 |
-| cycle | 7.4|3.6|0.3 |
-| car | 6.2|34.4|21.2 |
-| motorcycle | 0.6|2.8|1.2 |
-| bus | 3.5|51.0|75.9 |
-
-|cape_town|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 84.0|5.3|1.9 |
-| cycle | 0.2|0.5|0.1 |
-| car | 8.0|55.5|48.1 |
-| motorcycle | 0.0|0.7|0.4 |
-| bus | 7.5|37.9|49.0 |
-
-|delhi|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 84.4|10.1|0.9 |
-| cycle | 2.0|7.6|3.7 |
-| car | 2.4|17.3|21.4 |
-| motorcycle | 5.6|30.1|26.5 |
-| bus | 0.8|25.0|40.9 |
-
-|mexico_city|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 89.9|3.5|0.0 |
-| cycle | 2.4|3.2|0.2 |
-| car | 2.2|37.9|31.8 |
-| motorcycle | 0.4|3.2|0.9 |
-| bus | 1.8|35.3|64.7 |
-
-|santiago|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 84.0|3.9|0.0 |
-| cycle | 6.2|6.1|0.5 |
-| car | 8.1|58.7|32.3 |
-| motorcycle | 0.1|2.2|1.1 |
-| bus | 0.6|21.8|64.6 |
-
-|sao_paulo|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 82.6|1.9|0.0 |
-| cycle | 1.3|1.4|0.1 |
-| car | 12.6|65.2|36.8 |
-| motorcycle | 0.8|3.1|1.9 |
-| bus | 1.0|22.0|59.7 |
-
-|vizag|0-2 km|2-6 km|6+ km|
-|---|---|---|---|
-| pedestrian | 86.1|19.6|0.1 |
-| cycle | 2.6|6.6|2.0 |
-| car | 0.6|2.4|2.8 |
-| motorcycle | 6.9|24.1|24.3 |
-| bus | 0.8|26.7|54.8 |
+```
 

### 2021-07-19 summary_tables.Rmd
Update distance categories
Regen

@@
 
-0-1 km = \{trip distance < 2\}
+0-2 km = \{trip distance < 2\}
 
-2-5 km = \{2 <= trip distance < 6\}
+2-6 km = \{2 <= trip distance < 6\}
 
@@
 ## Case study specific propensities
-
-|accra|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 95.5|27.9|2.6 |
-| cycle | 0.6|0.7|0.4 |
-| car | 0.9|12.4|17.2 |
-| motorcycle | 0.0|0.0|1.3 |
-| bus | 2.4|48.3|71.8 |
-
-|bangalore|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 75.0|23.0|1.0 |
-| cycle | 2.3|9.8|1.1 |
-| car | 1.0|4.4|6.4 |
-| motorcycle | 5.1|33.9|63.2 |
-| bus | 15.1|24.3|25.0 |
-
-|belo_horizonte|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 89.2|8.7|0.3 |
-| cycle | 1.6|2.7|0.2 |
-| car | 7.3|65.6|47.1 |
-| motorcycle | 1.3|10.1|5.3 |
-| bus | 0.4|11.1|43.1 |
-
-|bogota|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 91.0|20.9|1.2 |
-| cycle | 4.9|12.1|2.0 |
-| car | 1.5|26.7|29.3 |
-| motorcycle | 0.5|5.9|6.5 |
-| bus | 0.8|20.7|56.8 |
-
-|buenos_aires|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 80.2|2.9|0.1 |
-| cycle | 7.4|3.6|0.3 |
-| car | 6.2|34.5|21.4 |
-| motorcycle | 0.6|2.8|1.2 |
-| bus | 3.5|50.1|51.0 |
-
-|cape_town|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 84.1|5.3|1.9 |
-| cycle | 0.2|0.5|0.1 |
-| car | 8.0|55.6|48.3 |
-| motorcycle | 0.0|0.7|0.4 |
-| bus | 6.9|34.8|33.3 |
-
-|delhi|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 84.4|10.1|0.9 |
-| cycle | 2.0|7.6|3.7 |
-| car | 2.4|17.3|21.4 |
-| motorcycle | 5.6|30.1|26.5 |
-| bus | 0.7|24.7|32.7 |
-
-|mexico_city|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 89.9|3.5|0.0 |
-| cycle | 2.5|3.2|0.2 |
-| car | 2.2|37.9|31.8 |
-| motorcycle | 0.4|3.2|0.9 |
-| bus | 1.7|34.8|33.8 |
-
-|santiago|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 84.0|3.9|0.0 |
-| cycle | 6.2|6.1|0.5 |
-| car | 8.1|58.7|32.3 |
-| motorcycle | 0.1|2.2|1.1 |
-| bus | 0.5|20.5|28.8 |
-
-|sao_paulo|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 82.7|1.9|0.0 |
-| cycle | 1.3|1.4|0.1 |
-| car | 12.6|65.4|37.0 |
-| motorcycle | 0.8|3.2|1.9 |
-| bus | 1.0|21.3|28.9 |
-
-|vizag|0-1 km|2-5 km|6+ km|
-|---|---|---|---|
-| pedestrian | 86.2|19.7|0.1 |
-| cycle | 2.6|6.6|2.1 |
-| car | 0.6|2.4|2.8 |
-| motorcycle | 6.9|24.3|24.6 |
-| bus | 0.8|26.9|55.5 |
+get_scen_settings <- ithimr::get_scenario_settings()
 

### 2021-08-09 summary_tables.Rmd
Adding health outcomes to summary_tables markdown

@@
 library(gt)
+library(pracma)
 })
@@
 ## Duration tables
+
 Case study specific duration tables for baseline and five scenarios (per day in km)
@@
 
+## Health outcomes
+
+### Change in YLL {#change_YLL}
+```{r scen_prop = "asis"}
+injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths) == 'scen1_deaths_inj')
+ap_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
+pa_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
+
+sum_and_round_and_print <- function(data, text = '') {
+  data <- lapply(data, function(x) round(x, round_to))
+  data <- lapply(data, function(x) rbind(x, Total = colSums(x)))
+  for (city in cities) {
+    print(kable(data[[city]], caption = paste(text, city)))
+    cat("\n")
+  }
+}
+round_and_print <- function(data,text=''){
+  data <- lapply(data, function(x)round(x,round_to))
+  for(city in cities) {
+    print(kable(data[[city]], caption = paste(text, city)))
+    cat("\n")
+  }
+}
+
+scen_names <- rownames(scen_prop)
+
+#yll_totals <- lapply(io,function(x){
+yll_totals <- lapply(io[-2],function(x){
+  temp <- sapply(1:5,function(y){
+    xx <- x$outcomes$hb$ylls
+    xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = 5)])
+    sapply(sort(unique(xx$age_cat)),function(z)
+      sum(xxx[xx$age_cat == z]))
+  })
+  colnames(temp) <- scen_names
+  temp
+})
+sum_and_round_and_print(yll_totals,"Change in YLL total in ")
+```
+
+YLLs per 100,000 people by age group by scenario
+
+### Change in YLLs per 100,000 {#change_YLL_age_100k}
+```{r scen_prop = "asis"}
+pop_by_age <- lapply(io[-2], function(x)
+  sapply(unique(x$demographic$age), 
+         function(y)
+           sum(subset(x$demographic, age == y)$population)
+         )
+  )
+
+yll_rates <- lapply(cities, function(x)
+  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],5,1))*100000,
+        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), 
+                                             length = 5)*100000
+        )
+  )
+
+names(yll_rates) <- cities
+round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
+```
+
+YLLs per 100,000 people by gender by age group scenario
+
+### Change in YLLs per 100,000 {#change_YLL_gender_age_100k}
+```{r scen_prop = "asis"}
+#yll_totals <- lapply(io,function(x){
+yll_totals <- lapply(io[-2], function(x) {
+  temp <- sapply(1:5,function(y){
+    xx <- x$outcomes$hb$ylls
+    rowSums(xx[,seq(2 + y, ncol(xx), by = 5)])
+  })
+  rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2], 1,
+                          function(z)paste0(z[1], '_', z[2]))
+  colnames(temp) <- scen_names
+  temp
+})
+
+yll_rates <- lapply(cities, function(x)
+  rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')], 1,
+                                    function(z) paste0(z[2],'_',z[1])), 
+                              rownames(yll_totals[[x]])),] /
+          t(repmat(io[[x]]$demographic$population,5,1)) * 100000,
+        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),
+                                             length = 5) * 100000))
+names(yll_rates) <- cities
+round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
+```
+
+### Change in YLLs due to injury {#change_yll_injury}
+```{r}
+injury_totals <- lapply(io[-2], function(x) {
+  xx <- x$outcomes$hb$ylls
+  injury_col <- grep("inj", names(xx))[1] # Dan: This line is mine (not sure is ok)
+  xxx <- sapply(1:5, function(y)
+    sapply(sort(unique(xx$age_cat)), function(z)
+      sum(xx[xx$age_cat == z, injury_col - 1 + y])))
+  colnames(xxx) <- scen_names
+  xxx
+})
+sum_and_round_and_print(injury_totals, "Change in YLLs due to injury in ")
+```
+
+### Change in YLLs due to PA {#change_yll_PA}
+```{r}
+pa_totals <- lapply(io[-2],function(x){
+  xx <- x$outcomes$pathway_hb$ylls
+  xxx <- sapply(1:5, function(y){
+    xxx <- rowSums(xx[, pa_cols[seq(y, length(pa_cols), by = 5)]])
+    sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
+  })
+  colnames(xxx) <- scen_names
+  xxx
+})
+
+sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
+```
+
+### Change in YLLs due to AP {#change_yll_AP}
+```{r}
+ap_totals <- lapply(io[-2],function(x){
+  xx <- x$outcomes$pathway_hb$ylls
+  xxx <- sapply(1:5,function(y) {
+    xxx <- rowSums(xx[, ap_cols[seq(y, length(ap_cols), by = 5)]])
+    sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
+  })
+  colnames(xxx) <- scen_names
+  xxx
+})
+sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ")
+```
+
+YLLs per 100,000 people by age group by scenario
+
+### Change in YLLs due to injury per 100,000 {#change_yll_injury_100k}
+```{r scen_prop = "asis"}
+injury_rates <- lapply(cities, function(x)
+  rbind(
+    injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1)) * 100000,
+    Total = colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),
+                                            length = 5) * 100000)
+  )
+
+names(injury_rates) <- cities
+round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ")
+```
+
+### Change in YLLs due to PA per 100,000 {#change_yll_PA_100k}
+```{r}
+pa_rates <- lapply(cities, function(x)
+  rbind(
+    pa_totals[[x]] / t(repmat(pop_by_age[[x]], 5, 1)) * 100000,
+    Total = colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),
+                                        length = 5) * 100000))
+
+names(pa_rates) <- cities
+round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ")
+```
+
+### Change in YLLs due to injury AP 100,000 {#change_yll_AP_100k}
+```{r}
+ap_rates <- lapply(cities, function(x)
+  rbind(
+    ap_totals[[x]] / t(repmat(pop_by_age[[x]], 5 ,1)) * 100000,
+    Total = colSums(ap_totals[[x]]) / rep(sum(pop_by_age[[x]]),
+                                          length = 5) * 100000))
+names(ap_rates) <- cities
+round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ")
+```
+
+## BY DISEASE
+Change in deaths total (for the city based on real population size) by scenario
+
+### Change in deaths due to disease {#change_death_disease}
+```{r}
+disease_totals <- lapply(io[-2],function(x){
+  xx <- x$outcomes$hb$deaths
+  xxx <- sapply(1:5,function(y){
+    colSums(xx[, seq(y + 2, ncol(xx), by = 5)])
+  })
+  colnames(xxx) <- scen_names
+  rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_', '', y))
+  xxx
+})
+sum_and_round_and_print(disease_totals,"Change in deaths due to disease in ")
+```
+
+Change in deaths per 100,000 people by scenario
+
+### Change in deaths due to disease per 100,000 {#change_death_disease_100k}
+```{r scen_prop = "asis"}
+disease_rates <- lapply(cities, function(x)
+  rbind(
+    disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000,
+    Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]), 
+                                               length = 5)*100000))
+
+names(disease_rates) <- cities
+round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ")
+```
+
+Change in YLL total (for the city based on real population size) by scenario
+
+### Change in YLL due to disease {#change_yll_disease}
+```{r}
+disease_totals <- lapply(io[-2],function(x){
+  xx <- x$outcomes$hb$ylls
+  xxx <- sapply(1:5,function(y) {
+    colSums(xx[,seq(y + 2, ncol(xx), by = 5)])
+  })
+  colnames(xxx) <- scen_names
+  rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_','',y))
+  xxx
+})
+sum_and_round_and_print(disease_totals,"Change in YLL due to disease in ")
+```
+
+### Change in YLL due to disease per 100,000 {#change_yll_disease_100k}
+```{r scen_prop = "asis"}
+disease_rates <- lapply(cities, function(x)
+  rbind(
+    disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000,
+    Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]),
+                                               length = 5) * 100000))
+names(disease_rates) <- cities
+round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ")
+```
+
+<!-- #------------------------------------------------------------------------------ -->
 # **2nd option**

### 2021-08-10 summary_tables.Rmd
Applying comments to health outcomes
1) Fixing 1.9.3 because it was full of NAs. 2) Changing scenarios names

@@
 
-rownames(scen_prop) <- paste(rownames(io$scen_prop), "scen", sep = "_")
+#rownames(scen_prop) <- paste(rownames(io$scen_prop), "scen", sep = "_")
+rownames(scen_prop) <- paste0(substr(toupper(rownames(io$scen_prop)),1,3),"_SC")
 
@@
   td <- as.data.frame(t(td))
-  names(td) <- lapply(append('baseline', rownames(scen_prop)), as.character)
+  names(td) <- lapply(append('BASELINE', rownames(scen_prop)), as.character)
   td <- td[-1, ]
@@
 
-  td1 <- td1 %>% dplyr::select(stage_mode, baseline)
+  td1 <- td1 %>% dplyr::select(stage_mode, BASELINE)
 
@@
   td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  colnames(td)[3:7] <- rownames(scen_prop)
+  #colnames(td)[3:7] <- rownames(scen_prop)
+  colnames(td)[2:7] <- c("BASELINE", rownames(scen_prop))
 
@@
 
-  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
+  td1 <- td1 %>% dplyr::select(stage_mode, BASELINE)
 
@@
   td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  colnames(td)[3:7] <- rownames(scen_prop)
+  #colnames(td)[3:7] <- rownames(scen_prop)
+  colnames(td)[2:7] <- c("BASELINE", rownames(scen_prop))
   x <- match(td$stage_mode, trip_order$mode)
@@
   td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 365, 3)))
-  colnames(td)[3:7] <- rownames(scen_prop)
+  #colnames(td)[3:7] <- rownames(scen_prop)
+  colnames(td)[2:7] <- c("BASELINE", rownames(scen_prop))
 
@@
 
-  td1 <- td %>% dplyr::select(stage_mode, Baseline)
+  td1 <- td %>% dplyr::select(stage_mode, BASELINE)
 
@@
 
-  names(df)[-1] <- lapply(append('baseline', rownames(scen_prop)), as.character)
+  names(df)[-1] <- lapply(append('BASELINE', rownames(scen_prop)), as.character)
 
@@
   td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  colnames(td)[3:7] <- rownames(scen_prop)
-
+  #colnames(td)[3:7] <- rownames(scen_prop)
+  colnames(td)[2:7] <- c("BASELINE", rownames(scen_prop))
   x <- match(td$stage_mode, trip_order$mode)
@@
   })
-  rownames(temp) <- apply(x$outcomes$hb$deaths[,1:2], 1,
-                          function(z)paste0(z[1], '_', z[2]))
+  rownames(temp) <- apply(x$outcomes$hb$deaths[,c('sex','age_cat')], 1,
+                          function(z)paste0(z[2], '_', z[1]))
   colnames(temp) <- scen_names

### 2021-08-10 summary_tables.Rmd
Removing second option in summary_tables

@@
 ```
-
-<!-- #------------------------------------------------------------------------------ -->
-# **2nd option**
-
-## Scenario definition 
-Table displays highest propensity for each distance category, for a given mode.
-
-Table displays highest propensity (%) for each distance category, for a given mode.
-
-In scenario generation, trips are sampled without replacement to increase the share of the requested mode up to the total shown.
-
-Distance categories are:
-
-0-1 km = \{trip distance < 2\}
-
-2-5 km = \{2 <= trip distance < 6\}
-
-6+ km = \{6 <= trip distance\}
-
-```{r scen_prop = "asis"}
-
-scen_prop <- round(io$scen_prop, round_to)
-
-rownames(scen_prop) <- paste(rownames(io$scen_prop), "scen", sep = "_")
-
-kable(scen_prop, headings = "Scenario Proportions")
-ggplot(as.data.frame(scen_prop) %>% rownames_to_column("scenario") %>% 
-         pivot_longer(!scenario, names_to = "distance_cat", 
-                      values_to = "values") %>% 
-         mutate(scenario = factor(scenario, 
-                                  levels = c("pedestrian_scen", "cycle_scen",
-                                             "car_scen", "motorcycle_scen",
-                                             "bus_scen"))
-                ), 
-       aes(scenario, values, fill = distance_cat)) + geom_col() + 
-  facet_grid(~ distance_cat) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
-        legend.position = "none")
-```
-
-## Case study specific propensities
-get_scen_settings <- ithimr::get_scenario_settings()
-
-## Trip proportion
-Case study specific trip proportions by mode, for baseline and five scenarios
-
-```{r load_tidyverse2, echo = F, message = F}
-suppressWarnings({
-  require(tidyverse)  
-})
-```
-
-```{r trip_mode_dist = "asis", echo = F, message = F}
-
-trip_prop <- list()
-
-for (city in cities) {
-  cat("\n \n")
-  cat( paste0('### **', city, '** \n' )) 
-  df <- io[[city]]$trip_scen_sets
-  u_trips <- df %>% dplyr::filter(scenario == "Baseline") %>% 
-    summarise(uid = n_distinct(trip_id)) %>% as.numeric()
-  td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% 
-    group_by(trip_mode, scenario) %>% 
-    summarise(p = round(dplyr::n() / u_trips * 100, 1)) %>% 
-    spread(key = trip_mode, value = p) %>% 
-    mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
-  td <- as.data.frame(t(td))
-  names(td) <- lapply(append('baseline', rownames(scen_prop)), as.character)
-  td <- td[-1, ]
-  x <- match(row.names(td), trip_order$mode)
-  x[!is.na(x)]
-  td <- td[order(x),]
-
-  td1 <- td %>% rownames_to_column()
-
-  names(td1)[1] <- 'stage_mode'
-
-  if (length(trip_prop) == 0) {
-    trip_prop <- td1 %>% dplyr::select(stage_mode)
-  }
-
-  td1 <- td1 %>% dplyr::select(stage_mode, baseline)
-
-  names(td1)[2] <- city
-
-  trip_prop <- inner_join(trip_prop, td1, by = 'stage_mode')
-
-  y <- ggplot(td %>% rownames_to_column("stage_mode") %>% 
-         pivot_longer(!stage_mode, names_to = "scenario", 
-                      values_to = "proportion") %>% 
-           mutate(scenario = factor(scenario,
-                                    levels = c("baseline", "pedestrian_scen",
-                                               "cycle_scen", "car_scen",
-                                               "motorcycle_scen", "bus_scen")),
-                  stage_mode = factor(stage_mode,
-                                      levels = c("car", "motorcycle", "bus",
-                                                 "pedestrian", "taxi", "rail",
-                                                 "cycle")),
-                  proportion = as.numeric(proportion)) %>% 
-           dplyr::filter(stage_mode != "row_sums"), 
-       aes(scenario, proportion, fill = stage_mode)) + geom_col() + 
-  facet_grid(~ stage_mode) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
-        legend.position = "none")
-
-  cat(" ")
-  print(y)
-  cat("\n ")
-  print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
-  cat("\n")
-}
-
-```
-
-### Baseline trip proportion (%)
-```{r}
-
-trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
-  mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>%
-  janitor::adorn_totals('row')
-
-ggplot(trip_prop %>% 
-         pivot_longer(!stage_mode, names_to = "city", 
-                      values_to = "proportion") %>% 
-         dplyr::filter(stage_mode != "Total") %>%
-         mutate(stage_mode = factor(stage_mode, 
-                                    levels = c("car", "motorcycle", "bus",
-                                               "pedestrian", "taxi", "rail",
-                                               "cycle"))), 
-       aes(stage_mode, proportion, fill = city)) + geom_col() + 
-  facet_grid(~ city) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
-        legend.position = "none")
-
-kable(trip_prop, caption = 'Trip prop (%) for all case studies - only selected modes')
-
-```
-
-```{r unload_tidyverse2, echo = F, message = F}
-suppressWarnings({
-  detach("package:tidyverse", character.only = T)
-})
-
-```
-
-## Distance tables
-Case study specific distance tables for baseline and five scenarios
-
-```{r trip_dist = "asis"}
-
-trip_dist <- list()
-
-for (city in cities) {
-  cat("\n \n")
-  cat( paste0('### **', city, '** \n' )) 
-  count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  colnames(td)[3:7] <- rownames(scen_prop)
-
-  x <- match(td$stage_mode, trip_order$mode)
-  x[!is.na(x)]
-  td <- td[order(x),]
-  row.names(td) <- NULL
-
-  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person")))
-  cat("\n")
-
-  td1 <- td
-
-  if (length(trip_dist) == 0){
-    trip_dist <- td1 %>% dplyr::select(stage_mode)
-  }
-
-  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
-
-  names(td1)[2] <- city
-
-  trip_dist <- inner_join(trip_dist, td1, by = 'stage_mode')
-
-  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  colnames(td)[3:7] <- rownames(scen_prop)
-  x <- match(td$stage_mode, trip_order$mode)
-  x[!is.na(x)]
-  td <- td[order(x),]
-  row.names(td) <- NULL
-
-  y <- ggplot(td %>% 
-         pivot_longer(!stage_mode, names_to = "scenario", 
-                      values_to = "distance") %>% 
-           mutate(scenario = factor(scenario,
-                                    levels = c("Baseline", "pedestrian_scen",
-                                               "cycle_scen", "car_scen",
-                                               "motorcycle_scen", "bus_scen")),
-                  stage_mode = factor(stage_mode,
-                                      levels = c("car", "motorcycle", "bus",
-                                                 "pedestrian", "taxi", "rail",
-                                                 "cycle"))),
-         aes(scenario, distance, fill = stage_mode)) + geom_col() + 
-    facet_grid(~ stage_mode) +
-    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
-        legend.position = "none")
-
-  cat(" ")
-  print(y)
-  cat("\n ")
-  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips)")))
-  cat("\n") 
-}
-
-```
-
-### Baseline average distance (km) per day for all case studies
-```{r}
-
-trip_dist <- trip_dist %>% janitor::adorn_totals('row')
-
-ggplot(trip_dist %>% 
-         pivot_longer(!stage_mode, names_to = "city", 
-                      values_to = "distance") %>% 
-         mutate(stage_mode = factor(stage_mode, 
-                                    levels = c("car", "motorcycle", "bus",
-                                               "pedestrian", "taxi", "rail",
-                                               "cycle", "Total"))), 
-       aes(stage_mode, distance, fill = city)) + geom_col() + 
-  facet_grid(~ city) +
-  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
-        legend.position = "none")
-
-kable(trip_dist, caption = "Avg. distance per person (km) per day across all case studies")
-```
-
-## Distance tables scaled by using total population (per year in km)
-Case study specific distance tables for baseline and five scenarios
-
-```{r trip_tot_dist = "asis", message=F}
-
-bl_td <- list()
-
-for(city in cities){
-
-  print(city)
-  count_people <- round(sum(io[[city]]$demographic$population))
-  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 365, 3)))
-  colnames(td)[3:7] <- rownames(scen_prop)
-
-  #
-
-  x <- match(td$stage_mode, trip_order$mode)
-  x[!is.na(x)]
-  td <- td[order(x),]
-  row.names(td) <- NULL
-
-  if (length(bl_td) == 0){
-    bl_td <- td %>% dplyr::select(stage_mode)
-  }
-
-  td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
-
-  td1 <- td %>% dplyr::select(stage_mode, Baseline)
-
-  names(td1)[2] <- city
-
-  bl_td <- inner_join(bl_td, td1, by = 'stage_mode')
-
-  td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)
-
-  print(td1)
-
-  # print(kable(td, caption = paste("Distance table (km) for ", city, "( ", sitools::f2si(count_people), " ) per person")))
-  cat("\n")
-
-}
-
-```
-
-### Baseline total distance for all case studies (per year in km)
-
-```{r trip_dist_all_cities = "asis", message=F}
-
-# tab <- bl_td %>% gt() %>% fmt_number(columns = 2:ncol(bl_td), decimals = T, suffixing = T)
-
-backup <- bl_td
-
-td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')
-
-for (i in 2:ncol(td)){
-  td[, i] <- comprss(td[, i])
-
-}
-
-print(kable(td, caption = paste("Distance table (km)")))
-
-```
-
-## Distance by distance category
-### Tables
-
-```{r trip_dist_mode_figs = "asis"}
-
-for (city in cities){
-
-  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), prop = round(n_vals / nrow(.) * 100, 1))
-  print(kable(df, caption = paste("Distance by mode by for  ", city)))
-}
-
-```
-
-## True Distance by total population (per day in km)
-### Tables
-
-```{r trip_dist_mode_figs = "asis"}
-
-for (city in cities){
-
-  df <- io[[city]]$true_dist
-
-  names(df)[-1] <- lapply(append('baseline', rownames(scen_prop)), as.character)
-
-  print(kable(df, caption = paste("True distance by mode per day for  ", city)))
-}
-
-```
-
-<!-- ### Figures -->
-
-<!-- ```{r trip_dist_mode_figs = "asis"} -->
-
-<!-- for (city in cities){ -->
-
-<!--   df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline' & trip_mode %in% c('walking', 'bicycle', 'car', 'motorcycle', 'bus')) %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_mode, trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n()) %>%  -->
-<!--     group_by(trip_mode) %>% mutate(n_vals_1 = sum(n_vals)) %>%  -->
-<!--     mutate(prop = round(n_vals/n_vals_1 * 100, 1)) -->
-
-<!--   print(ggplot(data=df, aes(x=trip_distance_cat, y=prop)) + -->
-<!--           geom_bar(stat="identity", fill="steelblue")+ -->
-<!--           facet_wrap(vars(trip_mode)) + -->
-<!--           geom_text(aes(label=prop), vjust = -0.3, size = 3.5)+ -->
-<!--           theme_minimal() + -->
-<!--           theme(panel.spacing = unit(2, "lines")) + -->
-<!--           labs(title = paste(city, ' by dist by mode'))) -->
-<!-- } -->
-
-<!-- ``` -->
-
-## Duration tables
-Case study specific duration tables for baseline and five scenarios (per day in km)
-
-```{r trip_dur = "asis"}
-
-trip_dur <- list()
-
-l <- list()
-
-for(city in cities){
-  count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  colnames(td)[3:7] <- rownames(scen_prop)
-
-  x <- match(td$stage_mode, trip_order$mode)
-  x[!is.na(x)]
-  td <- td[order(x),]
-  row.names(td) <- NULL
-
-  l[[city]] <- td
-
-  td1 <- td
-
-  if (length(trip_dur) == 0){
-    trip_dur <- td1 %>% dplyr::select(stage_mode)
-  }
-
-  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
-
-  names(td1)[2] <- city
-
-  trip_dur <- inner_join(trip_dur, td1, by = 'stage_mode')
-
-  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (everyone)")))
-
-  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  colnames(td)[3:7] <- rownames(scen_prop)
-
-  x <- match(td$stage_mode, trip_order$mode)
-  x[!is.na(x)]
-  td <- td[order(x),]
-  row.names(td) <- NULL
-
-  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (people with trips)")))
-  cat("\n")
-}
-
-```
-
-### Baseline average duration (mins) per day for all case studies
-```{r}
-
-trip_dur <- trip_dur %>% janitor::adorn_totals('row')
-
-kable(trip_dur, caption = "Avg. duration per person (mins) per day across all case studies")
-
-```

### 2021-08-11 summary_tables.Rmd
Changing the label of causes in health outcomes and Regen

@@
   colnames(xxx) <- scen_names
-  rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_', '', y))
+  #rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_', '', y))
+  rownames(xxx) <- c("PA_AllCauses", "PA_AP_IschemicHeartDisease", "PA_Neoplasms",
+                     "PA_AP_LungCancer", "AP_COPD", "PA_AP_Stroke",
+                     "PA_AP_T2Diabetes", "AP_LowerRespiratoryInfections",
+                     "PA_BreastCancer", "PA_ColonCancer", "PA_UterineCancer",
+                     "Injuries")
   xxx
@@
   colnames(xxx) <- scen_names
-  rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_','',y))
+  #rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_','',y))
+  rownames(xxx) <- c("PA_AllCauses", "PA_AP_IschemicHeartDisease", "PA_Neoplasms",
+                     "PA_AP_LungCancer", "AP_COPD", "PA_AP_Stroke",
+                     "PA_AP_T2Diabetes", "AP_LowerRespiratoryInfections",
+                     "PA_BreastCancer", "PA_ColonCancer", "PA_UterineCancer",
+                     "Injuries")
   xxx

### 2021-08-25 latex/main.tex
Add overleaf document - with PDFS, and its converted Rmd file

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{lscape}
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{amsfonts, amsmath, amssymb,array,comment}
\usepackage[round]{natbib}
\usepackage{soul}
\usepackage{float,graphicx}

\usepackage{footnote}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}
\makesavenoteenv{tabular}
\makesavenoteenv{table}
\newcommand*{\id}{{\normalfont\hbox{1\kern-0.15em \vrule width .8pt depth-.5pt}}}

\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{fit}
\usetikzlibrary{chains}
\usetikzlibrary{arrows}
% Latent node
\tikzstyle{latent} = [circle,fill=white,draw=black,inner sep=1pt,
minimum size=20pt, font=\fontsize{10}{10}\selectfont, node distance=1]
% Observed node
\tikzstyle{obs} = [latent,fill=gray!25]
% Marginal node
\tikzstyle{margin} = [latent,dashed,draw=gray!40!black]
% Constant node
\tikzstyle{const} = [rectangle, inner sep=1pt, node distance=1]
\tikzstyle{parameter} = [rectangle, inner sep=0pt, node distance=1]
% Factor node
\tikzstyle{factor} = [rectangle, fill=black,minimum size=5pt, inner
sep=0pt, node distance=0.4]
% Deterministic node
\tikzstyle{det} = [latent, diamond]

% Plate node
\tikzstyle{plate} = [draw, rectangle, rounded corners, fit=#1]
% Invisible wrapper node
\tikzstyle{wrap} = [inner sep=0pt, fit=#1]
% Gate
\tikzstyle{gate} = [draw, rectangle, dashed, fit=#1]

% Caption node
\tikzstyle{caption} = [font=\footnotesize, node distance=0] %
\tikzstyle{plate caption} = [caption, node distance=0, inner sep=0pt,
below left=5pt and 0pt of #1.south east] %
\tikzstyle{factor caption} = [caption] %
\tikzstyle{every label} += [caption] %

\tikzset{>={triangle 45}}

%\pgfdeclarelayer{b}
%\pgfdeclarelayer{f}
%\pgfsetlayers{b,main,f}

% \factoredge [options] {inputs} {factors} {outputs}
\newcommand{\factoredge}[4][]{ %
  % Connect all nodes #2 to all nodes #4 via all factors #3.
  \foreach \f in {#3} { %
    \foreach \x in {#2} { %
      \path (\x) edge[-,#1] (\f) ; %
      %\draw[-,#1] (\x) edge[-] (\f) ; %
    } ;
    \foreach \y in {#4} { %
      \path (\f) edge[->,#1] (\y) ; %
      %\draw[->,#1] (\f) -- (\y) ; %
    } ;
  } ;
}

% \edge [options] {inputs} {outputs}
\newcommand{\edge}[3][]{ %
  % Connect all nodes #2 to all nodes #3.
  \foreach \x in {#2} { %
    \foreach \y in {#3} { %
      \path (\x) edge [->,#1] (\y) ;%
      %\draw[->,#1] (\x) -- (\y) ;%
    } ;
  } ;
}

% \factor [options] {name} {caption} {inputs} {outputs}
\newcommand{\factor}[5][]{ %
  % Draw the factor node. Use alias to allow empty names.
  \node[factor, label={[name=#2-caption]#3}, name=#2, #1,
  alias=#2-alias] {} ; %
  % Connect all inputs to outputs via this factor
  \factoredge {#4} {#2-alias} {#5} ; %
}

% \plate [options] {name} {fitlist} {caption}
\newcommand{\plate}[4][]{ %
  \node[wrap=#3] (#2-wrap) {}; %
  \node[plate caption=#2-wrap] (#2-caption) {#4}; %
  \node[plate=(#2-wrap)(#2-caption), #1] (#2) {}; %
}

% \gate [options] {name} {fitlist} {inputs}
\newcommand{\gate}[4][]{ %
  \node[gate=#3, name=#2, #1, alias=#2-alias] {}; %
  \foreach \x in {#4} { %
    \draw [-*,thick] (\x) -- (#2-alias); %
  } ;%
}

% \vgate {name} {fitlist-left} {caption-left} {fitlist-right}
% {caption-right} {inputs}
\newcommand{\vgate}[6]{ %
  % Wrap the left and right parts
  \node[wrap=#2] (#1-left) {}; %
  \node[wrap=#4] (#1-right) {}; %
  % Draw the gate
  \node[gate=(#1-left)(#1-right)] (#1) {}; %
  % Add captions
  \node[caption, below left=of #1.north ] (#1-left-caption)
  {#3}; %
  \node[caption, below right=of #1.north ] (#1-right-caption)
  {#5}; %
  % Draw middle separation
  \draw [-, dashed] (#1.north) -- (#1.south); %
  % Draw inputs
  \foreach \x in {#6} { %
    \draw [-*,thick] (\x) -- (#1); %
  } ;%
}

% \hgate {name} {fitlist-top} {caption-top} {fitlist-bottom}
% {caption-bottom} {inputs}
\newcommand{\hgate}[6]{ %
  % Wrap the left and right parts
  \node[wrap=#2] (#1-top) {}; %
  \node[wrap=#4] (#1-bottom) {}; %
  % Draw the gate
  \node[gate=(#1-top)(#1-bottom)] (#1) {}; %
  % Add captions
  \node[caption, above right=of #1.west ] (#1-top-caption)
  {#3}; %
  \node[caption, below right=of #1.west ] (#1-bottom-caption)
  {#5}; %
  % Draw middle separation
  \draw [-, dashed] (#1.west) -- (#1.east); %
  % Draw inputs
  \foreach \x in {#6} { %
    \draw [-*,thick] (\x) -- (#1); %
  } ;%
}

\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    linktoc=all,
    filecolor=blue,
    linkcolor=black,
    urlcolor=black
}

%\pagenumbering{gobble} %% suppresses page numbers
\usepackage[left=2cm, right=2cm, top=2cm, bottom=2cm]{geometry}
\title{ITHIM}

\begin{document}
\maketitle
\tableofcontents
\clearpage
\section{Introduction}

\subsection{Summary of literature}

\subsection{Summary of aims}

\subsection{Overview}

The remainder of the document presents a current working version of the ITHIM, and is a comprehensive record of all the calculations of which it is composed. The equations for the ITHIM are described in Section \ref{equations} (tabulated in Section \ref{tableequations}). The method is concretised through the example application to Accra in Section \ref{example} and particular columns in Tables \ref{indices} and \ref{inputs}, with results presented in Section \ref{results}.

The ITHIM as described here is implemented in the ITHIM-R R package (\url{https://github.com/ITHIM/ITHIM-R}). Other cities are also included in that package (S\~{a}o Paulo, Delhi, Bangalore), and reference will be made to those where relevant.
% Some details here are specific to the Accra version of that model, as Accra has been the setting for construction of the prototype.

\section{Method: ITHIM}\label{equations}

In the ITHIM, we combine many information inputs and compute the health burden through three pathways (depicted in Figure \ref{workflow}). Each pathway has its own set of definitions and outputs, and we refer to them as ``modules''.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\textwidth]{flowchart.pdf}
  \caption{\small ITHIM workflow. Inputs to the model have a bold, black outline. The graph depicts parent--child relationships, where a child (at the head of an arrow) depends upon all its parents (at the source(s) of the arrow(s)). To aid visualisation, an arrow connected to the outside of a grey box indicates that the arrow connects to each item within the box. \textit{AT: active travel. MET: metabolic equivalent task. pp: per person. AP: air pollution. PA: physical activity. DR: dose--response relationship. GBD: global burden of disease. BOD: (local) burden of disease.}}
\label{workflow}
\end{figure}

To describe the model, we use lower-case letters to denote indices, or dimensions, of objects. They take one of a set of possible values, detailed in Table \ref{indices}. The set of input items are denoted by capital letters, and variable parameters are denoted by Greek letters. These are presented in Table \ref{inputs}. The equations of the ITHIM are presented below, and summarised in Table \ref{calculationstab}. These use letter modifiers to show how the inputs are modified in the course of output generation. %A description of the dose--response relationship parametrisation is given in Section \ref{dr}.

We calculate the ITHIM by combining the modules' outputs in terms of the health burden ($\hat{U}_{a,g,h,o,s}$). Other outputs of the model include: MMETs per person ($M_{i,s}$), PM2.5 in each scenario ($\bar{P}_s$), PM2.5 per person ($\check{W}_{i,s}$), and injuries burden ($\check{I}_{a,g,m,o,s}$).

To assess uncertainty, we sample from the ITHIM output by sampling uncertain parameters multiple times and calculating the ITHIM. We specify parametric distributions or sampling strategies for all the uncertain parameters. Evaluating the ITHIM with uncertain parameters allows assessment of their impact on the outcome (AKA sensitivity analysis). We use EVPPI to calculate the expected reduction in uncertainty in the outcome were we to learn a parameter perfectly. This means we can implement models that are basic in their parametrisation, and learn at the end for which parameters it would be worthwhile spending dedicated time learning better.

\begin{table}[ht]
\begin{center}
\caption{\small Indices used in the ITHIM variables. The ``Label'' is the subscript used to refer to the index in this document. The ``Name'' is descriptive. The ``Levels'' are the levels used in the Accra model.}
\begin{tabular}{lll}
\label{indices}
Label & Name & Levels  \\
\hline
$a$ & Age group & 15--49 \\
&  & 50--69  \\
$g$ & Gender & Male \\
&  & Female  \\
$h$ & Disease &  \\
$i$ & Individual index &  \\
$j$ & Trip index & \\
$m$ & Transport mode& walk  \\
&& cycling (cyc.)  \\
&& bus \\
&& car  \\
&& motorbike (mot.)  \\
&& goods vehicles (GV)  \\
&& subway (sub.)  \\
$o$ & Outcome & Death  \\
& & YLL  \\
%$p$ & PM2.5 levels \\
$s$ & Scenario & \\
$w$ & AP RR hyperparameter & $1,2,3,4$ \\
$x$ & PA RR hyperparameter & $1,2,3$ \\
$z$ & PA dose & $0,...,300$ \\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[ht]
\begin{center}
\caption{\small Common abbreviations.}
\begin{tabular}{ll}
\label{abbrev}
Abbreviation & Meaning  \\
\hline
AP & Air pollution \\
CDF & Cumulative distribution function \\
COPD & Chronic obstructive pulmonary disease \\
DR & Dose response \\
EVPPI & Expected value of perfect partial information \\
GAM & Generalised additive model \\
GBD & Global burden of disease\\
IHD & Ischemic heart disease\\
ITHIM & Integrated transport and health impact model \\
LRI & Lower respiratory infection \\
MMET & Marginal metabolic equivalent task \\
PA & Physical activity\\
PAF & Population attributable fraction\\
PIF & Potential impact fraction\\
PM2.5 & Particulate matter (diameter $< 2.5 \mu$m)\\
PT & Public transport\\
RR & Relative risk \\
VOI & Value of information \\
YLL & Years of life lost \\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{landscape}

\begin{center}
\begin{longtable}{lp{5cm}|lrr}
\caption{\small All values input to the ITHIM model.}\\
\label{inputs}
&&\multicolumn{3}{r}{\textit{ITHIM-R reference}}\\

Name & Description & Value in Accra model & Label & Model/ Setting\\

\hline
\multicolumn{2}{l|}{\textit{Global values}}&&&\\
\hline
$U_{a,g,h,o}$ & Background burden of disease   & Constant & GBD\_DATA &  Setting \\

$N_{a,g}$ & Population by age and gender in GBD\_DATA & Constant & GBD\_DATA & Setting \\

$\bar{N}_{a,g}$ & Population by age and gender & Constant & DEMOGRAPHIC & Setting \\

$\rho_h$ & Chronic disease scalar & 1 or Lnorm(0,0.18) & CHRONIC\_DISEASE\_SCALAR & Setting \\ 
\hline
\multicolumn{2}{l|}{\textit{Travel values}}\\
\hline

$Z_{m=walk}$ & Walk speed & 4.8 & MODE\_SPEEDS &  Setting \\
$Z_{m=cyc.}$ & Cycle speed & 14.5 & MODE\_SPEEDS  & Setting \\
$Z_{m=bus}$ & Bus speed & 15 & MODE\_SPEEDS & Setting \\
$Z_{m=car}$ & Car speed & 21 & MODE\_SPEEDS &  Setting \\
$Z_{m=mot.}$ & Motorbike speed & 25 & MODE\_SPEEDS  & Setting \\

$\epsilon$ & Walk time to bus & 5 or Lnorm(log(5),0.18) & BUS\_WALK\_TIME &  Setting \\

%$\mu_{m=mot.}$ & Motorcycle distance relative to car & 0.2 or Lnorm(-1.4,0.4) & MOTORCYCLE\_TO\_CAR\_RATIO &  & Setting \\

$\mu_{m=HGV}$ & Truck distance relative to car & 0.21 or Beta(3,10) & TRUCK\_TO\_CAR\_RATIO &  Setting \\

$\mu_{m=bus}$ & Bus driver distance relative to bus passenger distance & 1/45 or Beta(20,600) & BUS\_TO\_PASSENGER\_RATIO & Setting \\

$\mu_{m=mot.}$ & Motorcycle fleet distance relative to total motorcycle distance & 37/68 or Beta(37/3+1,31/3+1) & FLEET\_TO\_MOTORCYCLE\_RATIO & Setting \\

$T_j$ & Trip-level data & Constant & SYNTHETIC\_TRIPS &    Setting\\

%$Q_i$ & Person weights & Constant (1) & SYNTHETIC\_POPULATION\$ weight (TBI) &   & Setting\\

\hline
\multicolumn{2}{l|}{\textit{Injury model values}}\\
\hline

$I_{a,g,m_{\text{vic}},m_{\text{str}}}$ & Fatalities table & Constant & INJURY\_TABLE &   Setting\\

$\sigma$ & Injury reporting rate & 1 or Beta(8,3) & INJURY\_REPORTING\_RATE &  Setting \\

$\omega$ & Sum of distance exponents & 1.9 or Lnorm(log(1.9),log(1.03)) & SIN\_EXPONENT\_SUM & Model \\

$\psi$ & Casualty fraction of $\omega$  & 0.5 or Beta(20,20) & CASUALTY\_EXPONENT\_ FRACTION &  Model \\

\hline
\multicolumn{2}{l|}{\textit{Pollution model values}}\\
\hline

$\eta$ & {Background PM2.5 concentration} & 50 or Lnorm(log(50),0.18) &PM\_CONC\_BASE&  Setting\\
$\zeta$ & Fraction of measured PM2.5 concentration due to road transport & 0.225 or Beta(5,20) & PM\_TRANS\_SHARE &   Setting\\

$\gamma_{m}$ & Vehicle emission inventory  & Constant or Dirichlet & EMISSION\_INVENTORY & Setting \\

$P$ & Vehicle emission confidence  & 1 & EMISSION\_INVENTORY\_CONFIDENCE & Setting \\

\hline
\multicolumn{2}{l|}{\textit{Pollution \& health model values}}\\
\hline

$C_1$ & Base-level inhalation rate & 1 & BASE\_LEVEL\_INHALATION \_RATE &   Model \\
$C_2$ & Exposure ratio window closed & 0.5 & CLOSED\_WINDOW\_PM\_ RATIO &   Model \\
$C_3$ & Proportion of travel with closed windows & 0.5 & CLOSED\_WINDOW\_RATIO   & Setting \\
$C_4$ & Parameter for on-road PM2.5 level & 3.216 & ROAD\_RATIO\_MAX   & Model \\
$C_5$ & Parameter for on-road PM2.5 level & 0.379 & ROAD\_RATIO\_SLOPE  & Model \\
$C_6$ & Exposure ratio for underground & 0.8 & SUBWAY\_PM\_RATIO &  Setting \\

$H_{a,h,w}$ & {AP dose--response--curve parameters} & Constant & DR\_AP  & Model\\

$\xi_{h,w}$ & Quantiles for AP RR functions & 0.5 or Unif(0,1) & AP\_DOSE\_RESPONSE\_ QUANTILE &   Model\\

\hline
\multicolumn{2}{l|}{\textit{Physical activity model values}}\\
\hline
$\chi$ & Day-to-week travel scalar & 7 or 7$\times$Beta(20,3) & DAY\_TO\_WEEK\_TRAVEL\_ SCALAR &  Model \\

$Y_i$ & Individual non-transport MMETs & Constant & PA\_SET\$work\_ltpa\_marg\_met &    Setting\\

$\theta$ & Background PA scalar & 1 or Lnorm(0,0.18) & BACKGROUND\_PA\_SCALAR & Setting \\

$\tilde{Y}$ & Confidence in PA survey & 1 & BACKGROUND\_PA\_CONFIDENCE &    Setting\\

$\tilde{\theta}$ & Background PA zeros quantile & 0.5 or Unif(0,1) & BACKGROUND\_PA\_ZEROS & Setting \\

$\lambda_{m=\text{cyc.}}$ & Cycling MET surplus & 4.63 or Lnorm(log(4.63),1) & MMET\_CYCLING   &  Model \\

$\lambda_{m=\text{walk}}$ & Walking  MET surplus & 2.53 or Lnorm(log(2.53),1) & MMET\_WALKING  &  Model \\
$\lambda_{m\not\in\text{walk,cyc.}}$ & In-vehicle  MET surplus & 0 &   &  Model \\

\hline
\multicolumn{2}{l|}{\textit{Physical activity \& health model values}}\\
\hline
$G_{h,x,z}$  & Look-up table for relative risk truncated normal distribution & Constant &  & Model \\

$\phi_{h}$  & Quantiles for PA RR functions & 0.5 or Unif(0,1) & PA\_DOSE\_RESPONSE\_ QUANTILE & Model \\

\hline
\end{longtable}
\end{center}

\end{landscape}

%\clearpage
%\section{ITHIM workflow}\label{workflow-section}

\subsection{Data module}\label{data}

Input data for the ITHIM are listed as Roman capital letters in Table \ref{inputs}. They include (vehicle) mode speeds, injury/fatality records, disease dose--response relationships, measures of air pollution, surveys of travel and physical activity, population numbers by demographic group, MMETs associated with active travel, and current burden of disease (see Figure \ref{workflow}).

\subsubsection{Synthetic}\label{synth}

We create a ``synthetic population'' by matching individuals surveyed for travel to the individuals surveyed for physical activity, based on age and gender. Then each person in the synthetic population has a set of trips taken and an amount of non-travel physical activity. This population is assumed to be representative of the population of the city or setting under study.

\paragraph{Alternatively/additionally} we could create a synthetic population as described in Appendix \ref{synthpop}, which allows uncertainty in the propensity to travel. % and (b) models weekly (rather than daily) travel.
This method operates with person-level, rather than trip-level, information.

\subsubsection{Scenarios} From the synthetic population we create ``scenarios'' of different pictures of travel by changing certain trips (e.g. swapping the mode of travel for particular trips and/or particular people). For the purposes of illustration, we choose a simple scenario for Accra in which every person gains one one-kilometre walk.

\subsubsection{Distances}\label{distances}

For the ITHIM, we process the travel data into distance data of various formats. First, we augment PT trips with a walk component: we take the set $\{T_j:m(j)=\text{bus}\}$, which has $J$ entries. We add $J$ journeys with mode ``walk'' and duration $\epsilon$ to the set $T$. (This step would not be necessary for a travel survey which already includes all stages for each trip.)

Then we summarise the total distances and durations. The distance set is labelled $\hat{T}_{m,s}$, representing the total distance travelled per mode per scenario, and the duration set is constructed analogously. We make an assumption that the distance for bus drivers scales linearly with bus passengers, where the distance in the baseline is defined based on the bus driver distance relative to car.

Distance and duration summaries include all person travel in the synthetic population and, in addition, some trips that are attributed to no-one in the synthetic population. These trips are used to represent ``fleet'' travel, which is uncaptured in travel surveys, and unchanged in scenarios. It's important to include these trips because for modes such as motorcycle, where a considerable component of travel comes from fleet, attributing all the harms associated with the mode to only those surveyed dramatically overestimates the contribution of the individuals. In scenarios, then, we would be predicting emissions and injury rates that are far too high per person.

%\subsection{The ITHIM programme}\label{programme}

\subsection{AP module}

\subsubsection{Background PM2.5}\label{dir}

We use the total distance by mode ($\hat{T}_{m,s}$) to calculate the total PM2.5 in the scenario ($\bar{P}_s$). First we calculate the emission inventory in terms of fractions. If  the confidence $P$ is 1 then we use $\dot{P}_{m}=\gamma_m$ directly. If $P<1$, we sample the emission inventory fractions as follows:
$$\tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)),$$
$$\dot{P}_m=\tilde{\gamma}_m$$
where we choose some mapping function such as $f(P,\gamma_m)=\frac{\gamma_m}{\sum_m\gamma_m}10^{5P}$ (Figure \ref{emission_dist}).

Then we multiply vehicle distance by vehicle emission factor (where the emission factor is the emission inventory (fraction) divided by distance at baseline).
$$\tilde{P}_{m,s}=\dot{P}_m\frac{\hat{T}_{m,s}}{\hat{T}_{m,s=\text{baseline}}}.$$

%$\tilde{P}_{m,s}$ is augmented with rows corresponding to any vehicle missing from the trip set. For Accra, these modes include ``big trucks'' and ``other''. 
For the baseline, $\tilde{P}_{m,s}$ is equal to the emission inventory fraction. For each scenario, it is scaled. We take the sum over modes to get a scalar for emissions in scenarios.
%For the baseline scenario, this emission total equates to the transport proportion of background PM2.5. Therefore, for each scenario, we use the ratio of transport emissions to scale the transport-contributed PM2.5:
$$\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}.$$
Then the background PM2.5 in each scenario is the sum of the transport component and the non-transport component:
$$\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta).$$

\subsubsection{PM2.5 per person}

Dose--response relationships are based on background concentrations in cities, and not on personal exposures. In our person-level model, we expect that individuals will have different exposures based on their different travel behaviours. While we can neither estimate individual exposures nor map them via any dose--response relationship to a relative risk, we aim to capture interpersonal differences via assumptions about the mechanistic relationships between behaviours and exposures, and anchor those to the background air pollution in their locality, for which dose--response estimates of relative risk exist.

Therefore, when we write ``individual exposure'', it should not be read as an aetiologically relevant exposure, but rather it should be read as a unit- and scale-agnostic metric that maps an individual to a particular dose with reference to city-level measurements of air pollution. New research is required to elicit dose--response relationships between individual health exposures and health outcomes [citation from JPO?]. Our approximations act as assumptions for now and placeholders until more detailed data are available. [Perhaps we should use words other than ``individual exposure''. How about ``individual PM2.5 reference values''?]

\bigskip

Individual exposures to PM2.5 are calculated using the background PM2.5 ($\bar{P}_s$)  and the trip sets. There are three major components to daily exposure: one, a person's total inhalation off road; two, a person's inhalation on road in a vehicle, and, three, a person's inhalation on road while undertaking active transport. Each category has an amount of background PM2.5 and a ventilation rate which together inform overall exposure.

The ratio of exposure off road to that on road is a function of total PM2.5, defined as
$$K_s=C_4-C_5\log(\bar{P_s})$$
as in \citet{Goel2015}. This defines the exposure of a person in an open vehicle (i.e. pedestrian, cyclist or motorist):
$$\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$$
and it is used to calculate in-vehicle exposure, assuming an exposure of $C_2$ with the window closed, and a proportion $C_3$ of vehicles having closed windows:% (\hl{we use this also for motorcycles, which don't have windows...}):
$$\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3).$$
The exposure in a subway is constant and not dependent on the road ratio $K_s$:
$$\tilde{K}_{m=\text{sub.},s}=C_6.$$

Ventilation rates are calculated for each mode, assuming a base-level inhalation rate of $C_1$:
$$V_m=C_1+\frac{1}{2}\lambda_m,$$
where $\lambda_m$ is the MMETs for mode $m$.\footnote{Ainsworth compendium 2011 sites.google.com/site/ compendiumofphysicalactivities for walking and cycling} Then the air inhaled during travel per person is
$$\tilde{V}_{i,m,s} = \smashoperator{\sum_{\substack{j:i(j)=i,\\m(j)=m,\\s(j)=s}}}T_j\cdot V_{m(j)}/Z_{m(j)}$$
and the rate of inhalation during travel is
$$\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}.$$
The air inhaled when not travelling is
$$\hat{V}_{i,s} = C_1\left(24-\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}T_j/Z_{m(j)}\right).$$
Together, the PM2.5 exposure is calculated as the total PM2.5 inhaled per hour %divided by the total volume of air inhaled
as follows:
%$$\check{V}_{i,s} = \bar{P}_s\frac{\bar{V}_{i,s}+\hat{V}_{i,s}}{\sum_m\tilde{V}_{i,m,s}+\hat{V}_{i,s}}.$$
$$\check{W}_{i,s} = \frac{\bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})}{24}.$$

%\bigskip

%Dose--response relationships are based on background concentrations in cities, and not on personal exposures. Therefore, our person-level exposures, which will relate in some way to the city's background level, represent some variation about the city's precise point. We note that the mechanistic details therefore reflect our assumptions about how individual variation in inhalation arises from variation in travel behaviours, and we assume that these aggregate across a city to realise the result seen in studies of dose--response relationships. New research is required to elicit dose--response relationships between individual health exposures and health outcomes [citation from JPO?]. Our approximations act as assumptions for now and placeholders until then.

%Finally, each person's exposure is rescaled following the ratio between the mean exposure in the baseline and the background PM2.5 in the baseline.
%$$\check{P}=\frac{\text{mean}(\check{V}_{i,s=\text{Baseline}})}{\bar{P}_{s=\text{Baseline}}},$$
%$$\check{W}_{i,s} = \check{P}\check{V}_{i,s}.$$

%We normalise so that the mean exposure corresponds to the given background air pollution, using the Baseline to define the normalisation factor.

\subsubsection{AP--disease dose--response relative risk}
We use each person's exposure to PM2.5 ($\check{W}_{i,s}$) to calculate their relative risk of five diseases (IHD, lung cancer, COPD, stroke, LRI), using curves parametrised by four disease-specific variables \citep{Burnett2014}. Of the five diseases, two (IHD and stroke) have parameters specific to age groups starting at age 25.\footnote{For any person of age lower than 25, we set the relative risk to 1.} The other three (lung cancer, LRI and COPD) have one set of parameters for all ages.

The curves are in the form of samples of the set of four parameters. We model the densities of these samples (using a quantile for parameter 3, kernel density estimation for parameter 2, and GAMs for parameters 1 and 4)\footnote{The four parameters refer, in numerical order, to $\alpha, \beta, \gamma$ and tmrel in \citet{Burnett2014}.} in order to draw either the median or random samples via their quantiles ($\xi_{h,w}$) as follows:

%  (RR_AP_calculations <- gen_ap_rr(pm_conc_pp))

\begin{align}\tilde{H}_{a,h,w=3}=&\text{CDF}_{H_{a,h,w=3}}^{-1}(\xi_{h,w=3})\\
\tilde{H}_{a,h,w=2}=&\text{CDF}_{H_{a,h,w=2}|\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=2})\\
\tilde{H}_{a,h,w=1}=&\text{CDF}_{H_{a,h,w=1}|\tilde{H}_{a,h,w=2},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=1})\\
\tilde{H}_{a,h,w=4}=&\text{CDF}_{H_{a,h,w=4}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=2},\tilde{H}_{a,h,w=1}}^{-1}(\xi_{h,w=4}).\end{align}

From these parameters, the relative risk of mortality is defined
$$\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=1}\times
\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=2}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=4}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}.$$
For $h\not\in\{\text{IHD, lung cancer, COPD, stroke, LRI}\}$, $\check{H}_{h,i,s} \equiv 1$.

\subsection{PA module}
\subsubsection{Individual-level MMETs}\label{beta}

Using trip sets and the synthetic population, we calculate total MMETs per person ($M_{i,s}$) as the sum of walking MMETs per day and cycling MMETs per day, which are scaled up to a week via the scalar $\chi$, and work/leisure MMETs per week as follows:
\begin{equation}M_{i,s}=\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}
\chi\lambda_{m(j)}\frac{T_j}{Z_{m(j)}}+\theta Y_i.\end{equation}

Here, $\lambda_m$ is the MMETs for mode $m$, $T_j$ the distance of trip $j$, $Z_m$ the speed of mode $m$, $\theta$ the scalar for background PA, and $\check{Y}_i$ the amount of work and leisure (background) PA of person $i$.

We calculate $\check{Y}_i$ according to the confidence in the PA survey, $\tilde{Y}$. First, we calculate the raw probability that a person in demographic group completes no non-travel PA:
$$\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}.$$
Then, we set the probability to use, $\bar{Y}_{a,g}$, as $\dot{Y}_{a,g}$ if our confidence $\tilde{Y}$ is 1. If $\tilde{Y}<1$, we map the confidence to parametrise a Beta distribution somehow, e.g. (Figure \ref{prob_zero_PA})
\begin{align}
\hat{Y} &= 500^{\tilde{Y}+0.2},\\
\beta & = \hat{Y}\dot{Y}_{a,g}\left(\frac{1}{\dot{Y}_{a,g}} -1\right),\\
\alpha & = \hat{Y} - \beta,\\
\bar{Y}_{a,g} & = \text{CDF}_{\text{Beta}(\alpha,\beta)}^{-1}(\tilde{\theta}).
\end{align}

Finally, for each person in the population, we sample non-travel MMETs as zero with probability $\bar{Y}_{a,g}$ and from the raw non-zero density of their demographic group with probability $1-\bar{Y}_{a,g}$.

\subsubsection{PA--disease dose--response relative risk}

We use each person's MMETs per week (${M}_{i,s}$) to calculate their relative risk of six diseases, using curves found from meta analysis. Each disease (except type 2 diabetes) has a threshold beyond which there is no further change in relative risk. (Total cancer, breast cancer, colon \& rectum cancer, endometrial cancer, \& coronary heart disease: 35; lung cancer: 10; stroke: 32; all cause: 16.08.) For all individual cancers, we estimate the burden in terms of \hl{...}. For all other diseases, we use mortality.

The interpolation of the dose--response curve can be described as follows:
$$\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$$
where we interpolate the mean, the lower bound, and the upper bound. These define the normal from which we sample, truncated at 0:
$$F_S(v_x) = \mathcal{N}\left.\left(v_{x=2},\frac{v_{x=3}-v_{x=1}}{1.96}\right)\right|_{F_S(v_x)>0}. $$

Then the relative risk for each person for each disease is calculated for the quantile $\phi_h$ as
$$W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right).$$
See Figure \ref{cancer} for an example. For $h\not\in$ \{diabetes, total cancer, breast cancer, colon \& rectum cancer, endometrial cancer, IHD, lung cancer, stroke, all cause\}, $W_{h,i,s} \equiv 1$.

\begin{figure}[H]
\centering
\includegraphics[width=0.6\textwidth]{cancer_PA_dose_response.pdf}
\caption{\small Example of the dose--response workflow for PA and total cancer. Top left: results of the metaanalysis: mean, lower bound and upper bound of the relationship between PA dose and relative risk of disease. Top right: examples of 500 samples from the normal distribution defined by the mean of the top-left panel and the standard deviation defined by (upper bound minus lower bound) divided by 1.96 and truncated at 0. Bottom left: relative risks of individuals with PA doses 5, 25 and 45 are found by mapping their doses onto the median curve. This is used in the ``constant'' ITHIM use case. Bottom right: relative risks of individuals with PA doses 5, 25 and 45 for three different samples from the distribution shown in the top-right panel. For each sample, each individual is mapped onto the response curve it defines. }
\label{cancer}
\end{figure}

\subsection{Injury module}

\subsubsection{Processing}

For the injury module, we require distance per travel mode per demographic group per scenario. It requires matching of mode names in the travel survey to mode names in the injury records, for example ``taxi'', ``shared auto'', ``shared taxi'' and ``car'' combine to form ``car''. Some work is required to separate drivers from passengers; we currently assume all travellers are drivers, with the exception of ``bus''.\footnote{This is a problem for the examples of Delhi and Bangalore, whose trip sets have car travellers under the age at which driving is permitted and who therefore should not contribute to striking distance.}

\subsubsection{Modelling}

We model injuries via regression by predicting the number of fatalities of each demographic group on each mode, which we calculate as a sum over all the ways in which they might have been injured (i.e., all the modes with which they might have collided).

The predictive covariates include the distances travelled by the parties and their demographic details. These requirements lead to a natural separation of the (training) dataset into two groups: a set for which we have distance data for the other party, and a set for which we do not have distance data for the other party. The former equates to a ``who hit whom'' (``whw'') matrix (albeit in a higher dimension), and will account for changes in injuries resulting of a change in strike-mode travel. The latter corresponds to causes of injury that will not change across scenarios, including ``no other vehicle'' and modes of transport that we do not consider to change, which might include trucks and buses if they are not somehow explicitly included in the trip set. We label this group ``noov'': no or other vehicle.

We model the number of injuries as a Poisson-distributed variable with an offset depending on the distance(s) and the reporting rate. We write the model as follows:
\begin{align}
\tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s,y} \sim &
\text{Poisson}(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}); \\
% 'count~
\log\left(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}|m_{\text{str}}\in\{m_{\text{whw}}\}\right) = &
% cas_mode*strike_mode-log(injury_reporting_rate)+log(weight)
{\kappa_0} +\sum_{i=1}^n\kappa_iX_i-\log(\sigma)+\nonumber\\
% +log(cas_distance)
 & \log(\bar{T}_{a,g,m_{\text{cas}},s})
 % +(CAS_EXPONENT-1)*log(cas_distance_sum)
 +(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s})
 %+log(strike_distance)+(STR_EXPONENT-1)*log(strike_distance_sum)
 +E_{m_{\text{str}}}\log(\hat{T}_{m_{\text{str}},s});\\
\log\left(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}|m_{\text{str}}\in\{m_{\text{noov}}\}\right) =&
{\kappa'_0}
+\sum_{i=1}^n\kappa'_iX'_i 
-\log(\sigma)
% +log(cas_distance)
+ \log(\bar{T}_{a,g,m_{\text{cas}},s})
 % +(CAS_EXPONENT-1)*log(cas_distance_sum)
 +(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s}).
\end{align}
We choose this form of equation to enable (a) linearity in injuries with respect to total travel combined across modes and (b) linearity in injuries as subdivided by travellers of each mode.\footnote{Note that the assymmetry in the offset term for the whw expression arises due to the absence of demographic information for striking vehicles. Were we to have that, the offset would read $\log(\bar{T}_{a,g,m_{\text{cas}},s})+(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s})+\log(\bar{T}_{a,g,m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})$. For our case, the latter two terms simplify to $\log(\hat{T}_{m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})=E_{m_{\text{str}}}\log(\hat{T}_{m_{\text{str}},s})$.} Note that this methodology is under development. See \url{https://github.com/robj411/safety_in_numbers_power_law}.

In general, one makes the best model that one can with the data available. We use, for Accra, the covariates \texttt{casualty mode*strike mode, casualty age,} and \texttt{casualty gender}, to form the model matrix $X$. We also have another covariate, \texttt{year}, but our travel data are for one year only, so we model instead many observations of the same quantities.

We fit the coefficients $\kappa$ of the model using data for ten years, assuming the same travel each year, which corresponds to the scenario $s=\text{baseline}$. To incorporate the injury reporting rate ($\sigma$), we set it as an offset, whose value is 1 in fitting the model, and as specified in making predictions. %(For settings where there are no demographic data, we fit the model using \texttt{casualty mode} and \texttt{strike mode} alone. Predictions are made, as before, by demographic group. )

\subsubsection{Prediction}

We predict the number of injuries in each scenario based on the training model built from the baseline scenario. For Accra, we predict for the year 2016, using scenario-specific travel data.
The number of fatalities is taken as the sum of fatalities over the strike modes:
$$\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}.$$
We extrapolate injury deaths to injury YLLs via the ratio in the GBD:
$$R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}},$$
$$\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g}. $$

\subsection{Health module}
\subsubsection{PA and AP relative risks combined}

  We combine the relative risks of disease as a function of AP and disease as a function of PA through multiplication:
  $$\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s}. $$
  Not all diseases have a dose--response relationship for both AP and PA. Just stroke, lung cancer, and IHD have both. For the other diseases, only one of $W_{h,i,s}$ and $\check{H}_{h,i,s}$ is different from one.

  \subsubsection{Total disease burden}

We calculate the total health burden relative to the reference scenario (which, for our Accra example, is the baseline) using the injury and health outputs combined with GBD data, via population attributable fractions (PAF). The PAF is defined:
$$\hat{W}_{a,g,h,s}=\smashoperator{\sum_{\substack{i:a(i)=a,\\g(i)=g}}}\tilde{W}_{h,i,s}.$$

We then calculate the PAF relative to the reference scenario (``ref'') as:
$$\bar{W}_{a,g,h,s}= \frac{\hat{W}_{a,g,h,s=\text{ref}}-\hat{W}_{a,g,h,s}}{\hat{W}_{a,g,h,s=\text{ref}}}.$$

We estimate the background burden of disease using GBD data and scaling based on the ratio of populations between country and city. For country populations $N_{a,g}$, city populations $\bar{N}_{a,g}$ and country burden $U_{a,g,h,o}$, we estimate city burden as
$$\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}.$$

If we are scaling the background burden of non-communicable diseases (``Neoplasms'', ``Ischemic heart disease'', ``Tracheal, bronchus, and lung cancer'', ``Breast cancer'', ``Colon and rectum cancer'', ``Uterine cancer''), we do so here. (Otherwise set $\rho\equiv 1$.)
$$\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}.$$

We combine the burden with the PAF:
$$\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}.$$

\subsubsection{Injury burden}

And, finally, for the injuries, we sum over modes to compute the burden,
$$\bar{U}_{a,g,o,s}=\sum_m\check{I}_{a,g,m,o,s},$$
and subtract from the values for the reference scenario:
$$\hat{U}_{a,g,h=\text{road injury},o,s}=\bar{U}_{a,g,o,s=\text{ref}}-\bar{U}_{a,g,o,s}.$$

\section{Example: Accra}\label{example}

Here we show the ITHIM as evaluated for the city of Accra, with a simple scenario of every person in the synthetic population gaining a single, 1 km walk. We define XX uncertain univariate parameters (Section \ref{uni}), two confidences, and two uncertain parameter sets (Section \ref{dr}).

\subsection{Parametric distributions for uncertain variables}\label{uni}

The XX uncertain univariate parameters are shown in Figure \ref{distributions}. Most of the parameters are self explanatory. Cycling and walking MMETs are the number of MMETs per hour when undertaking cycling and walking, and determine also the ventilation rates. Motorcycle distance is the total distance travelled by motorcycles relative to the total distance travelled by cars in the baseline scenario.  Non-travel PA, injury reporting rate and NCD burden all act as scalars for the corresponding datasets. Note that the non-travel PA scalar does not affect the $\approx$40\% of the population whose non-travel PA is 0.

\begin{figure}[h]
\centering
\includegraphics[width=0.95\textwidth]{uncertain_parameters.pdf}
\caption{\small Distributions of univariate uncertain parameters (1024 samples).}
\label{distributions}
\end{figure}

\subsection{Confidences}

For physical activity propensity and emission inventories, we use ``confidences'', values between 0 and 1 that represent how confident we are about the data source. We map these values to Beta (Figure \ref{prob_zero_PA}) and Dirichlet (Figure \ref{emission_dist}) distributions, respectively. The Beta-distributed propensity towards non-travel physical activity is described in Section \ref{beta}. The Dirichlet-distributed emissions inventory is described in Section \ref{dir}.

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{prob_zero_PA.pdf}
\caption{\small Distributions of a Beta-distributed propensity to non-travel PA with varying confidence values in the raw input data. The sampled value gives the fraction of the population that engages in non-travel PA. \hl{[Re-do with Accra parameters.]}}
\label{prob_zero_PA}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.65\textwidth]{emission_dist.pdf}
\caption{\small Distributions of Dirichlet-distributed emission inventories for four modes and varying confidence values. Raw input values were 4, 4, 20, and 60. \hl{[Re-do with Accra parameters and inventory.]}}
\label{emission_dist}
\end{figure}

\subsection{Dose--response relationships}\label{dr}

For the dose--response relationships between physical activity (PA) and disease and air pollution (AP) and disease, we assume that there is uncertainty, but no variability, in the relationship. This means that we sample a relationship from the distribution of relationships, and apply that relationship to all individuals precisely. This means that, given fixed doses, responses between individuals will be perfectly correlated.

We achieve this by use of the probability integral transform: we sample a random variable uniformly distributed on the space (0,1) and map it, via a cumulative distribution function, to the distribution describing the dose--response relationship. See Figure \ref{cdf} for an illustration.

\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{DR_transform.pdf}
\caption{\small We use the same uniform samples to generate random variables from different distributions. It means that the samples are perfectly correlated.}
\label{cdf}
\end{figure}

\subsubsection{Physical activity}

Each disease's PA dose--response relationship is defined by a truncated normal. For each dose, there is a mean value, an upper bound, and a lower bound. For each person's dose, we get the response by mapping the uniform random variable onto the truncated normal defined by the mean and bounds for that dosage.

\subsubsection{Air pollution}

For the AP relationship, there are four parameters per disease. We sample the first from an empirical distribution using the probability integral transform. We sample the second via the same method, conditioned on the value of the first, constructing their joint density with e.g. kde2d. The third parameter is sampled conditioned on the first and second, constructing their joint density using a GAM. The final parameter is sampled conditioned on the first, second, and third, constructing their joint density using a GAM.

As before, there is perfect correlation between individuals, i.e. if person A's dose is greater than person B's, then person A's response is strictly greater than person B's response.

The empirical distributions come from \citet{Burnett2014}. There are four parameters per disease: IHD, lung cancer, COPD, stroke, and LRI. In addition, for stroke and IHD, there is a set of four parameters for each age group from 25 to 95 in five-year increments. In addition to our assumption that there is perfect correlation between individuals for diseases, we assume perfect correlation between ages for diseases. I.e., our four quantiles per disease will be applied to all age groups. 
\clearpage
\section{Results}\label{results}
\subsection{Health burdens in scenarios}
%The scenarios for Accra are:
%\begin{itemize}
%\item Baseline (base) -- the current state of affairs
%\item Scenario 1 (scen1) --  bus + walk to car 
%\item Scenario 2 (scen2) -- scen1 plus long trips from car + taxi to bus 
%\item Scenario 3 (scen3) -- scen1 plus car to motorbike 
%\item Scenario 4 (scen4) -- scen1 plus short trips from car + taxi + motorbike to cycle 
%\item Scenario 5 (scen5) -- scen1 plus short trips from car + taxi to walk
%\end{itemize}

\begin{comment}
\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{combined_yll_pp.pdf}
\caption{\small Results (YLLs per person) for four cities combined, relative to scen1.}
\label{results}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{city_yll.pdf}
\caption{\small Results (YLLs) for four cities, relative to scen1.}
\label{results}
\end{figure}
\end{comment}

\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{accra_walk_yll.pdf}
\caption{\small YLLs in the scenario minus YLLs in the baseline for Accra.}
\label{results}
\end{figure}
\clearpage
\subsection{Value of information}

We calculate value of information following \citet{Jackson2017}. The ``value of information'' we implement is the expected value of perfect partial information (EVPPI). It is the expected amount by which variation in the outcome diminishes should a parameter, or set of parameters, be known perfectly. With reference to information theory, it is the mutual information between the parameter, or parameter set, and the outcome, where the outcome is described only by its second moment (its variance).

The objective of learning the value of information formulated in this way is to identify the parameter or parameters that most drive variability (variance) in the outcome. The results guide future research, identifying which area or areas offer the greatest benefit in terms of prediction accuracy given further research. For example, for our model of Accra, learning better the dose--response relationship between physical activity and stroke (incidence) would likely increase precision of our estimate of total years of life lost (YLL) more so than learning any other parameter (Figure \ref{evppi}).

\begin{figure}[h]
\includegraphics[width=0.7\textwidth]{accra_walk_evppi.pdf}
\caption{\small EVPPI for Accra's ``walking'' scenario for all causes of YLLs excluding ``all cause'' and ``neoplasm''.}
\label{evppi}
\end{figure}
\clearpage

\begin{appendix}

\section{ITHIM with parameters}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{ithim_with_parameters.pdf}
  \caption{\small ITHIM-R workflow, with variable parameters in purple. Fixed inputs to the model have a bold, black outline, which is solid for inputs we consider ``global'' and dashed for inputs we consider ``local''. There are four global inputs, which will be embedded in ITHIM-R. There are seven local inputs, which users should provide. The graph depicts parent--child relationships, where a child (at the head of an arrow) depends upon all its parents (at the source(s) of the arrow(s)). To aid visualisation, an arrow connected to the outside of a grey box indicates that the arrow connects to each item within the box. \textit{AT: active travel. MET: metabolic equivalent task. pp: per person. AP: air pollution. PA: physical activity. DR: dose--response relationship. GBD: global burden of disease.} NB: $\rho$ only impacts the NCD burden of GBD, not all of it. Missing parameter: motorcycle distance. $\lambda$ is a 2D parameter. $\xi$ is 4D.}
\label{ithim_with_parameters}
\end{figure}

\clearpage
\section{Dose--response relationships}

\begin{figure}[H]
\centering
\includegraphics[width=0.9\textwidth]{PA_dose_response.pdf}
\caption{\small 1000 samples from PA dose--response curves}
\label{PADR}
\end{figure}

\begin{figure}[H]
\centering
\begin{tabular}{cc}
\includegraphics[width=0.4\textwidth]{cvd_ihdDR-sim.pdf}&
\includegraphics[width=0.4\textwidth]{cvd_strokeDR-sim.pdf}\\
\end{tabular}
\begin{tabular}{ccc}
\includegraphics[width=0.3\textwidth]{neo_lungDR-sim.pdf}&
\includegraphics[width=0.3\textwidth]{lriDR-sim.pdf}&
\includegraphics[width=0.3\textwidth]{resp_copdDR-sim.pdf}\\
\end{tabular}
\caption{\small 100 samples from AP dose--response curves}
\label{APDR}
\end{figure}

\clearpage

\begin{landscape}
\section{Tabulated ITHIM equations}\label{tableequations}

\begin{center}
\begin{ThreePartTable}
  \begin{TableNotes}  
  \small
  \item[\textdagger] {\footnotesize CDF: cumulative distribution function. We use generative distributions to generate random numbers so that samples are correlated. For $t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$ with CDF CDF$_F$. (I'd like to find an improved notation for this.)}
  \end{TableNotes}
\begin{longtable}{lp{7cm}p{5cm}}
\caption{\small All calculations in the ITHIM.}\label{calculationstab}\\
Name & Description & Label \\
\hline
\multicolumn{3}{l}{\textit{Scenario generation}}\\
\hline
...&&\\[5pt]
\hline
\multicolumn{3}{l}{\textit{Travel calculations}}\\
\hline
$\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk}\}\, \forall\, T_j:m(j)=\text{bus}$ &New walk trips to bus&{walk\_to\_bus\_and\_combine\_scen}\\[5pt]
%$T_j\leftarrow T_j-\epsilon\, \forall\, j:m(j)=\text{bus}\text{ \& }T_j>\epsilon$ &Subtract walk distance from bus trip&{walk\_to\_bus\_and\_combine\_scen}\\[5pt]
$\bar{T}_{a,g,m,s}=\sum_{j:a(j)=a,g(j)=g,m(j)=m,s(j)=s}Q_{i(j)}T_j$ & (Weighted) total travel per age, gender, mode and scenario &  \\[5pt]
$\check{T}_{a,g,m,s}=\frac{\bar{T}_{a,g,m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}$ & Relative travel per mode and scenario by age and gender & inj\_distances[[1]] \\[5pt]
$\hat{T}_{m,s}=\sum_{a,g}\bar{T}_{a,g,m,s}$ & Total travel per mode and scenario & dist \\[5pt]
$\tilde{T}_{m,s}=\frac{\hat{T}_{m,s}}{\hat{T}_{m,s=\text{baseline}}}$ & Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1 &  inj\_distances[[2]] \\[5pt]
\hline
\multicolumn{3}{l}{\textit{Injury calculations}}\\
\hline

$E_{m_{\text{cas}}}=\psi\omega$ & Casualty exponent & CAS\_EXPONENT  \\[5pt]

$E_{m_{\text{str}}}=\omega-E_{m_{\text{cas}}}$ & Striker exponent & STR\_EXPONENT \\[5pt]

$\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}=\mathcal{S}\left(\sigma\cdot {I}_{a,g,m_{\text{vic}},m_{\text{str}}},\bar{T}_{a,g,m,s},E_{m_{\text{cas}}},E_{m_{\text{str}}}\right)$
%\cdot \tilde{T}_{m=\text{vic},s}^{S_{m_{\text{vic}},m_{\text{str}},r=\text{vic}}}\cdot \tilde{T}_{m=\text{str},s}^{S_{m_{\text{vic}},m_{\text{str}},r=\text{str}}}$
& Fatalities in scenarios by strike mode & injuries\\[5pt]
$\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}$ & Fatalities in scenarios  & victim\_deaths\\[5pt]
%$\check{I}_{a,g,m,o=\text{death},s}=\hat{I}_{a,g,m,s}$ & Fatalities distributed to age-and-gender groups & injuries \\[5pt]

$R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}}$ & Road YLL-to-death ratio & GBD\_INJ\_YLL\\[5pt]

$\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g} $ & YLLs scaled from fatalities & deaths\_yll\_injuries \\[5pt]

\hline
\multicolumn{3}{l}{\textit{Pollution calculations}}\\
\hline
$\dot{P}_{m}=\left\{\begin{array}{lr}\gamma_m & P=1 \\ \tilde{\gamma}_m, \tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)) & P<1 \end{array}\right.$ & Emissions by mode &   trans\_emissions \\[5pt]

$\tilde{P}_{m,s}=\dot{P}_m\hat{T}_{m,s}/\hat{T}_{m,s=\text{baseline}}$& Emissions by mode in scenario relative to baseline &   trans\_emissions \\[5pt]

$\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}$ & Emissions scalar for scenarios & baseline\_sum \\[5pt]
%$\check{P}_s = \eta\zeta\hat{P}_s$ & Background PM2.5 concentration due to transport in scenarios &    \\[5pt]
$\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta)$ & Background PM2.5 concentration in scenarios &   scenario\_pm  \\[5pt]
\hline 
\hline 
$K_s=C_4-C_5\log(\bar{P_s})$ & On-road PM2.5 exposure ratio  & on\_road\_off\_road\_ratio \\[5pt]
$\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$ & On-road PM2.5 exposure ratio for open vehicle  & on\_road\_off\_road\_ratio  \\[5pt]
$\tilde{K}_{m\in\{\text{sub.}\},s}=C_6$ & Subway PM2.5 exposure ratio   & subway\_ratio  \\[5pt]
$\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3)$ & In-vehicle PM2.5 exposure ratio  & in\_vehicle\_ratio  \\[5pt]
\hline
\hline 
$V_m=C_1+\frac{1}{2}\lambda_m$ & Ventilation rates &   vent\_rates \\[5pt] 
$\tilde{V}_{i,m,s} = \sum_{j:i(j)=i,m(j)=m,s(j)=s}T_j\cdot V_{m(j)}/Z_{m(j)}$ & In-travel air inhaled &   \\[5pt]
$\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}$ & In-travel PM2.5 inhalation rate &   \\[5pt]
$\hat{V}_{i,s} = C_1\left(24-\sum_{j:i(j)=i,s(j)=s}T_j/Z_{m(j)}\right)$ & Non-travel air inhaled &   \\[5pt]
$\check{W}_{i,s} = \bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})/24
%/(\sum_m\tilde{V}_{i,m,s}+\hat{V}_{i,s})
$ & Total PM2.5 inhaled per hour  &  \\[5pt]

%$\check{P}=\frac{\text{mean}(\check{V}_{i,s=\text{Baseline}})}{\bar{P}_{s=\text{Baseline}}}$ & Normalisation constant   &  \\[5pt]
%$\check{W}_{i,s} = \check{P}\check{V}_{i,s}$ & Normalised total PM2.5 inhaled   & pm\_conc\_pp \\[5pt]

\hline
\hline 
$\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ &{AP dose--response--curve parameter}& DR\_AP\_LIST \\[5pt]
$\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ &{AP dose--response--curve parameter}& DR\_AP\_LIST \\[5pt]
$\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ & {AP dose--response--curve parameter} & DR\_AP\_LIST \\[5pt]
$\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ &{AP dose--response--curve parameter}& DR\_AP\_LIST \\[5pt]
$\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=3}\times$ &&\\[5pt]
$\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=3}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=3}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}$& RR of disease given pollution &  RR\_AP\_calculations \\[5pt]

\hline 
\multicolumn{3}{l}{\textit{Physical activity calculations}}\\
\hline

$\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}$ & Raw probability non-travel MMETs = 0 for demographic group $a,g$ & raw\_zero \\[5pt]

$\bar{Y}_{a,g}=\left\{\begin{array}{lr} \dot{Y}_{a,g} & \tilde{Y}=1 \\ \text{CDF}_{\text{Beta}(f(\tilde{Y},\dot{Y}_{a,g}))}^{-1}(\tilde{\theta}) & \tilde{Y}<1 \end{array}\right.$& Probability non-travel MMETs = 0 for demographic group $a,g$ &  zeros  \\[5pt]

$\check{Y}_i = \left\{\begin{array}{lr} 0 & \bar{Y}_{a(i),g(i)} \\ \text{Sample}(\{Y_{i'}\}|_{a(i')=a(i),g(i')=g(i),Y_{i'}>0}) & 1-\bar{Y}_{a(i),g(i)} \end{array}\right.$& Non-travel MMETs &  SYNTHETIC\_POPULATION\$ work\_ltpa\_marg\_met  \\[5pt]

$M_{i,s}=\sum_{j:i(j)=i,s(j)=s}(\chi T_j\cdot \lambda_{m(j)}/Z_{m(j)})+\theta \check{Y}_i$ & Total MMETs per person & mmets \\[5pt]

$\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$ & Interpolated dose--response parameters & rr, lb, ub \\[5pt]

$F_S(s_x) = \mathcal{N}(s_{x=2},(s_{x=3}-s_{x=1})/1.96)|_{0<F_S(s_x)} $ & Truncated normal function &  \\[5pt]

$W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right)^{\dagger}$ &Relative risk of disease given PA&RR\_PA\_calculations\\[5pt]

\hline 
\multicolumn{3}{l}{\textit{Burden-of-disease calculations}}\\
\hline
$\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s} $ & Combined PA and AP relative risks& RR\_PA\_AP\_calculations \\[5pt]

$\hat{W}_{a,g,h,s}= \sum_{i:a(i)=a,g(i)=g}\tilde{W}_{h,i,s}$ & Population-attributable fractions& pif\_temp \\[5pt]

$\bar{W}_{a,g,h,s}= (\hat{W}_{a,g,h,s=\text{scen1}}-\hat{W}_{a,g,h,s})/\hat{W}_{a,g,h,s=\text{scen1}}$ & Population-attributable fractions relative to scenario 1& pif\_scen \\[5pt]

$\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$ & GBD scaled to local population & DISEASE\_BURDEN \\[5pt]

$\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}$ & Scaled background disease for Neoplasms, IHD, Lung cancer& gbd\_data\_scaled \\[5pt]

$\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}$ & Combined health and PIF & yll\_dfs, death\_dfs \\[5pt]

$\hat{U}_{a,g,h=\text{road injury},o,s}=\sum_m\check{I}_{a,g,m,o,s}$ & Injury health burden & inj \\[5pt]

\hline

%$\bar{W}_{a,g,h,s}=U_{a,g,h,o}\cdot(1-\check{W}_{a,g,h,s})$ & Burden given activity &   PA\_burden  \\[5pt]

\hline
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\end{center}

\end{landscape}

\section{Uncertainty in travel data}

We model uncertainty in travel via the definition of ``travel attributes'' (in the terminology of \citet{Mohammadian2010}). Specifically, we define the propensity to travel in a day, and the distance travelled in a day, given that travelled occurred. Each travel attribute is defined per demographic group, per mode.

In contrast to other travel-related synthetic populations, we have four demographic groups (rather than clusters) which consist of individuals (rather than households). A crucial difference between our objectives and the existing work on simulating  populations in travel-demand models (that I've found so far) is that we don't need trip-level data for our calculations. Simulating trip-level data is computationally expensive and high-dimensional in terms of uncertain parameters. For our downstream computations, we need only summary data: travel per person per mode. Beyond scenario generation, there is no need for us to consider trips.

%Rather than having clusters of households, we have four demographic groups of individuals, and we match travel attributes with probability 1. With these analogies, our approaches are similar. It suggests an approach for future settings with more complex survey structures, or should we choose to model trips rather than total travel. That is, (a) to form clusters from the data using household-level data; (b) to find summary statistics for each cluster; (c) generate a synthetic population where each individual belongs to a cluster with appropriate proportions; (d) sample travel for each individual given their cluster.

E.g. \citet{Saadi2016} focuses on (a) synthetic population via matching covariates and (b) the sequence of trips taken by each person. (a) isn't a big thing for Accra; we're only using two covariates. And for (b), I think we can focus on people as it's a lot simpler. For our purposes, the things in this work that might be useful to us include matching principles for (a), when we revisit it, and distributions (in particular joint distributions) for travel. More useful to us would be distributions for total travel, but I haven't found any in this literature (e.g. zero-inflated compound Poisson distribution).

In contrast, my main questions are:
\begin{itemize}
\item How to account for multi-mode trips.
\item What to do about zero travel for e.g. older cyclists. One possibility is to smooth over all modes and/or demographic groups.
\item In the ITHIM-R programme, are bus and truck drivers participants?
\item How should we include correlations between modes? This could also be aided by a smooth model across modes.
\item Whether there is a better method to get distance travelled per person than by resampling raw data.
\end{itemize}

\subsection{Method}

\subsubsection{Summarise the travel survey}

Our first step is to summarise the demographic information and the travel survey. In Table \ref{population} are the demographic data: the populations of the city, and the number of people in the survey. In Table \ref{calculations} are the summary statistics from the travel survey which we will use to generate synthetic travel. From the raw travel data (actually the adjusted trip set, which is the raw travel plus motorcycle, bus and truck trips), we calculate the probability to travel, $B_{a,g,m}$, as the number of people who travelled by that mode divided by the number of people in the demographic group. It is to these values that we assign uncertainty through the variable PROPENSITY\_TO\_TRAVEL.

\begin{table}[ht]
\begin{center}
\caption{\small Population by demographic group.}
\begin{tabular}{llrr}
\label{population}
Age & Gender & Surveyed & Accra population \\
\hline
15--49  & Male & 279 & 5,847,716 \\
50--69  & Male & 56 & 1,016,476 \\
15--49  & Female & 328 & 6,360,535 \\
50--69  & Female & 69 & 1,110,037 \\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{center}
\begin{ThreePartTable}
  \begin{TableNotes}  
  \small
  %\item[\textdagger] {\footnotesize These values are hypothesised, as no such trips were recorded in the original dataset.}
  \end{TableNotes}
\begin{longtable}{lll|rr}
\caption{\small Travel summary table.  }\label{calculations}\\
Age & Gender & Mode & Probability in raw trip set & Probability in baseline trip set  \\
\hline
15-49 &  Male &  Bus & 0.4014 &  0.3862\\ 
50-69 &  Male &  Bus & 0.4286 &  0.3636   \\ 
15-49 &  Female &  Bus & 0.3506 &  0.3495   \\ 
50-69 &  Female &  Bus & 0.4203 &  0.4203 \\ 
15-49 &  Male &  Taxi & 0.0681 & 0.0655  \\ 
50-69 &  Male &  Taxi & 0.0357 & 0.0303  \\ 
15-49 &  Female &  Taxi & 0.0671 & 0.0669  \\ 
50-69 &  Female &  Taxi & 0.1159 & 0.1159  \\ 
15-49 &  Male &  Walking & 0.6559 & 0.631 \\ 
50-69 &  Male &  Walking & 0.4643 &   0.3939 \\ 
15-49 &  Female &  Walking & 0.5427 & 0.541  \\ 
50-69 &  Female &  Walking & 0.4203 &  0.4203 \\ 
15-49 &  Male &  Private Car & 0.1254 & 0.1207  \\ 
50-69 &  Male &  Private Car & 0.1964 &   0.1667 \\ 
15-49 &  Female &  Private Car & 0.0549 &   0.0547 \\ 
50-69 &  Female &  Private Car & 0.058 &  0.058 \\ 
15-49 &  Male &  Bicycle & 0.0179 &  0.0172 \\ 
50-69 &  Male &  Bicycle & 0 &  0 \\ 
15-49 &  Female &  Bicycle & 0.003 & 0.003\\ 
50-69 &  Female &  Bicycle & 0 & 0 \\ 
15-49 &  Male & Short Walking & 0 & 0.3862 \\ 
50-69 &  Male &  Short Walking & 0 & 0.3636 \\ 
15-49 &  Female &  Short Walking & 0 & 0.3495 \\ 
50-69 &  Female &  Short Walking & 0 & 0.4203 \\ 
15-49 &  Male &  Motorcycle & 0 & 0.0621  \\ 
50-69 &  Male &  Motorcycle & 0 & 0.1515 \\ 
15-49 &  Female &  Motorcycle & 0 & 0.0061 \\ 
50-69 &  Female &  Motorcycle & 0 & 0.0145 \\ 
\hline
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\end{center}

\subsubsection{Adjust for trip weight}

We define a trip-mode weight per mode, $L_m$, so that a single trip by mode $m$ in the baseline scenario represents $L_m$ trips in reality. If in a scenario the mode of the trip is changed ($m'\rightarrow m$), it retains the weight of its original mode $m'$. We use the trip-mode weight to (a) sample trips for scenario generation and (b) scale the probabilities to travel. Note that we don't use it as trip weights are ordinarily used: to scale trip distance/duration. The result is a scaled set of probability parameters:
$$\hat{B}_{a,g,m} = L_m{B}_{a,g,m}.$$
($L_m$ should actually be the average trip-mode weight for all trips in the scenario. For baseline, it will be the same. It can/should differ for scenarios. Need to work out notation.)

\subsubsection{Smooth probabilities}

We estimate smoothed probabilities for any mode that has zero or one probability, using the rest of the probabilities for reference, replacing $\hat{B}_{a,g,m}$. The reference to other modes is slightly problematic if e.g. in the scenario all travel increases. It means a 0 in the baseline is not the same as a 0 in the scenario. This could be fixed with more rigid coding.

\subsubsection{Resample probabilities}

We resample the adjusted probability parameters of Table \ref{calculations}. We describe each parameter with a Beta distribution as follows:
\begin{align}\label{pitilde}
\tilde{B}_{a,g,m} & \sim \text{Beta}(\alpha_{a,g,m},\beta_{a,g,m}); \\
\alpha_{a,g,m}+\beta_{a,g,m} & = 200; \\
\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}.
\end{align}
Here, $\tilde{B}_{a,g,m}$ is the new, variable parameter that we resample for each simulation. We sample via a quantile, $\nu_m\sim\mathcal{U}(0,1)$, which we use to calculate $\tilde{B}_{a,g,m}$.  We set $\alpha_{a,g,m}$ and $\beta_{a,g,m}$ to sum to 200, making the distributions concentrated, and we constrain their relative values to enforce the distribution's mean to equal $\hat{B}_{a,g,m}$ (the right-most column in Table \ref{calculations}). %The resulting distributions are shown in Figure \ref{travel_parameters}.

%Rather than sample many $\tilde{B}_{a,g,m}$ random variables, we sample just one standard uniform distribution and use the same quantile for every demographic group and every mode.

\subsubsection{Inform individual travel}

Each individual is assigned their own propensity to travel per mode, $\delta_{i,m}$, which is a uniformly distributed random variable that is fixed and not resampled. Whether or not individual $i$ travels by mode $m$ depends on $\tilde{B}_{a(i),g(i),m}$, which does vary: specifically, individual $i$ travels by mode $m$ if $\delta_{i,m}<\tilde{B}_{a(i),g(i),m}$. (To mitigate the effects of these fixed random variables, a large population is required, e.g. more than 5000 individuals.)

%\begin{figure}[H]
%\centering
%\includegraphics[width=0.7\textwidth]{propensity_to_travel.pdf}
%\caption{\small Distributions for propensities to travel. The propensity is the probability that a member of a demographic group will travel by a particular mode on a particular day. In orange are the raw probabilities based on the synthetic trip set, here shown for the baseline scenario. Beta distributions are defined via a pointiness parameter, set to 100, such that $\alpha+\beta=100$. }
%\label{propensity_to_travel}
%\end{figure}
\begin{figure}[H]
\centering
\includegraphics[width=0.7\textwidth]{propensity_to_travel_200.pdf}
\caption{\small Distributions for propensities to travel. The propensity is the probability that a member of a demographic group will travel by a particular mode on a particular day. In orange are the raw probabilities based on the synthetic trip set, here shown for the baseline scenario. Beta distributions are defined via a pointiness parameter, set to 200, such that $\alpha+\beta=200$. }
\label{propensity_to_travel}
\end{figure}

%\begin{comment}
\subsection{Extrapolating to one week of travel}\label{Extrapolation}

%Do we want to extrapolate to one week, or one year? We could feasibly sample one week of trips. One year of trips, however, is likely prohibitively expensive.

In order to extrapolate to one week of travel, we consider ``trips'' only at the stage of scenario generation, and proceed with individual-level statistics. Then, in our synthetic population, we have ``duration travelled per mode'' columns. None of the core ITHIM-R functions will use the ``synthetic trip set''.

To get duration per mode per person, we use, for each demographic-group--mode combination, the probability to travel, and the density of the total duration travelled given that travel occurred. To get from one day to one week, consider two cases: one, each individual follows the same pattern of travel every day. That is, we sample one day's travel and apply it to seven days. Two, each individual follows the travel pattern of their demographic group on each day. That is, we sample seven times from the group's density to make up seven days of travel.%We can then sample, for each individual, the likely number of days travelled out of seven (or 365), and sample the likely duration per day. Their total travel will be the sum of these durations.

%To introduce sampling into the calculation, 
In implementation, we blend individual and group-level travel through a ``repetitiveness'' parameter, $\pi$. Each individual has two travel sets: one from their simulated daily travel pattern, and one from the shared, convolved distribution. These travel sets are weighted by $\pi$ and $1-\pi$, respectively. $\delta_{i,m}$ defines the quantile taken for both the individual density, which is scaled up to one week, and the shared distribution, which  is the convolution of the raw density seven times.\footnote{Using the same random variable for both travel sets increases the variability of the resulting dataset. Conversely, using different random variables would result in more homogenous durations, and fewer zeros.} The densities are shown in Figure \ref{raw_densities}.

\begin{figure}[H]
\centering
\includegraphics[width=0.8\textwidth]{raw_densities.pdf}
\caption{\small Raw distributions for duration travelled per day (left) from the travel survey. Convolution to one week is shown on the right. We use ````walking'''' mode for demographic group ``Male'', ``15--49''.}
\label{raw_densities}
\end{figure}

%We use the raw density of travel duration per day per mode per demographic group. We can manipulate this density directly, in order to simulate scenarios, or to vary the probability-to-travel parameter, $B_{a,g,m}$. %Note that we are no longer interested in the other parameters of Table \ref{calculations}, and we consider six random variables (five propensities for the five modes, and repetitiveness, $\pi$).
%\end{comment}

\subsection{Simulating a synthetic population}\label{synthpop}

\begin{enumerate}
\item Choose a population size of individuals ${N}$. Label each individual $i=1,...,N$.
\item Allocate each individual to a demographic group according to the proportions in the Accra population (Table \ref{population}). Then each demographic group has number $\tilde{N}_{a,g}$ and  $\sum_{a,g}\tilde{N}_{a,g}=N$.
\item Generate one standard uniform variable per person per mode, $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual $i$'s propensity to travel by mode $m$. 
\item $\tilde{B}_{a,g,m,s}$ is the propensity of group \{$a,g$\} to travel by mode $m$ in scenario $s$. It is the proportion of raw person--duration values that are zero. It is parametrised by $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation \ref{pitilde}). 
\item Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw person--duration values for demographic group  \{$a,g$\} and mode $m$ in scenario $s$: $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$ i.e. we isolate all individuals $i$ in group $\{a,g\}$ and sum their trip durations by mode $m$ in scenario $s$.% (see Figure \ref{raw_densities}, left).
\item Let $\tilde{\mathcal{F}}_{a,g,m,s}$ be that density corrected to have proportion $\tilde{B}_{a,g,m,s}$ zeros.
%\item Let $\tilde{\mathcal{F}}_{a,g,m,s}^7$ be the week convolution of $\tilde{\mathcal{F}}_{a,g,m,s}$ (see Figure \ref{raw_densities}, right).
\item Define each individual's travel duration as %$D_{i,m,s}=7\pi\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})+(1-\pi)\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}^7}^{-1}(\delta_{i,m})$ (see Figure \ref{distances_beta} for resulting distributions).
$D_{i,m,s}=7\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})$ (see Figure \ref{distances_beta} for resulting distributions).
\end{enumerate}

For sampling, we loop over steps 6--7, where each sample has unique $\tilde{B}_{a,g,m,s}$ values. In total, the number of uncertain variables is one per mode. 
%For sampling, we loop over steps 6--8, where each sample has unique $\tilde{B}_{a,g,m,s}$ and $\pi$ values. In total, the number of uncertain variables is one per mode plus one. %Alternatively, we sample a unique $T_j$ set for each iteration similarly to Section \ref{synthtrip},\footnote{For scenario generation, we might want to vary trip duration.} omit step 7, and loop over steps 6, 8 and 9.

\subsection{Simulating a synthetic population with repetitiveness parameter \& uncertain extrapolation}\label{synthpop2}

\begin{enumerate}
\item Choose a population size of individuals ${N}$. Label each individual $i=1,...,N$.
\item Allocate each individual to a demographic group according to the proportions in the Accra population (Table \ref{population}). Then each demographic group has number $\tilde{N}_{a,g}$ and  $\sum_{a,g}\tilde{N}_{a,g}=N$.
\item Generate one standard uniform variable per person per mode, $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual $i$'s propensity to travel by mode $m$. 
\item $\tilde{B}_{a,g,m,s}$ is the propensity of group \{$a,g$\} to travel by mode $m$ in scenario $s$. It is the proportion of raw person--duration values that are zero. It is parametrised by $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation \ref{pitilde}). 
\item Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw person--duration values for demographic group  \{$a,g$\} and mode $m$ in scenario $s$: $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$ i.e. we isolate all individuals $i$ in group $\{a,g\}$ and sum their trip durations by mode $m$ in scenario $s$.% (see Figure \ref{raw_densities}, left).
%\item Let $\tilde{\mathcal{F}}_{a,g,m,s}$ be that density corrected to have proportion $\tilde{B}_{a,g,m,s}$ zeros.
\item Let $\tilde{\mathcal{F}}_{a,g,m,s}^7$ be the week convolution of $\tilde{\mathcal{F}}_{a,g,m,s}$ (see Figure \ref{raw_densities}, right).
\item Define each individual's travel duration as $D_{i,m,s}=7\pi\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})+(1-\pi)\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}^7}^{-1}(\delta_{i,m})$ (see Figure \ref{distances_beta} for resulting distributions).
%$D_{i,m,s}=7\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})$ (see Figure \ref{distances_beta} for resulting distributions).
\end{enumerate}

%For sampling, we loop over steps 6--7, where each sample has unique $\tilde{B}_{a,g,m,s}$ values. In total, the number of uncertain variables is one per mode. 
For sampling, we loop over steps 6--8, where each sample has unique $\tilde{B}_{a,g,m,s}$ and $\pi$ values. In total, the number of uncertain variables is one per mode plus one. %Alternatively, we sample a unique $T_j$ set for each iteration similarly to Section \ref{synthtrip},\footnote{For scenario generation, we might want to vary trip duration.} omit step 7, and loop over steps 6, 8 and 9.

\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{distances_beta.pdf}
\caption{\small Samples of total distances travelled by mode and scenario.}
\label{distances_beta}
\end{figure}

\clearpage

\subsection{Results}

\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{accra_walk_yll_uncertain_travel.pdf}
\caption{\small YLLs in the scenario minus YLLs in the baseline for Accra, with motorcycle probability scalar 2 in Accra.}
\label{results}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.7\textwidth]{accra_walk_evppi_uncertain_travel.pdf}
\caption{\small EVPPI for Accra's ``walking'' scenario for all causes of YLLs excluding ``all cause'' and ``neoplasm'', with propensity to motorcycle = 2.}
\label{evppi_mode_propensities_200}
\end{figure}

%\subsection{Questions arising}

%\begin{itemize}
%\item How to model multi-mode trips
%\item Programming: logic of which modes to use
%\item What do we do about zero trips for e.g. older cyclists?
%\item Are bus and truck drivers participants?
%\item Including correlations between modes
%\end{itemize}

\clearpage
\section{Glossary by letter}

\begin{table}[h]
\begin{tabular}{|>{$}r<{$}l|>{$}r<{$}l|>{$}r<{$}l|}
\hline
\text{Letter} & Meaning & \text{Letter} & Meaning & \text{Letter} & Meaning \\
\hline
\alpha &  & a & age & A &  \\
\beta &   & b &   & B & raw travel probabilities \\
\gamma & emission uncertainty  & c &   & C & ventilation constants \\
\delta &  propensity to travel & d & demographic group  & D &  \\
\epsilon & walk-to-bus time  & e &   & E & injury distance exponent \\
\zeta & traffic PM2.5 fraction  & f &   & F & PA DR function \\
\eta & PM2.5 concentration  & g & gender  & G & PA DR curves \\
\theta &  PA uncertainty & h &  disease & H & AP DR curves \\
\iota &   & i &  individual & I & injuries \\
 &   & j & trip  & J &  \\
\kappa & injury regression parameter  & k &   & K & AP exposure \\
\lambda &  travel MMETs & l &   & L &  \\
\mu &  mode distances & m & mode  & M & MMETs per person \\
\nu &   & n &   & N & population \\
\xi &  AP DR quantile & o & outcome  & O &  \\
\pi & travel repetitiveness  & p &   & P & vehicle emissions \\
 &   & q &   & Q &  person weights \\
\rho & chronic disease scalar  & r &   & R & death-to-injury ratio \\
\sigma & injury reporting rate  & s & scenario  & S &  \\
\tau &   & t &   & T & trips \\
\upsilon &   & u &   & U & burden of disease \\
\phi & PA DR quantile  & v &   & V & AP per person \\
 &   & w & AP DR parameter   & W & relative risks \\
\chi & day-to-week travel scalar  & x & PA DR parameter  & X & injury model matrix \\
\psi & casualty fraction of $\omega$  & y & year & Y & individual MMETs \\
\omega & sum of distance exponents  & z & PA dose  & Z & speeds \\
\hline
\end{tabular}
\end{table}

\end{appendix}
\clearpage
\bibliographystyle{agsm}
\bibliography{ithim}
\end{document}

\clearpage
\section{Injury model}

\subsection{Basic set up}

We begin with the trip set, $T_j$. There is one value (a distance) per trip $j$. $T_j$ is informed by a set of covariates $z=\{a,g,m\}$. $a$ is the age of the person who took the trip, $g$ their gender, and $m$ the mode. The complete set $M$ is the set of all the unique modes $m$ in the trip set $T_j$.

From $T_j$ we calculate the total distance per set $x\in X$: $$D_x=\sum_{j:z_j= x}T_j.$$ $X$ is the complete set of all unique combinations of the variables $z$.

We augment $D_X$ with distances from modes not included in the trip set. We denote the augmented set of modes $M^+$ and the augmented distance set $D_{X^+}$.

For each mode in $M^+$ we calculate the total distance travelled as $$\hat{D}_{m}=\sum_{a,g}D_{a,g,m}.$$

We define a set $y=\{a',g',m'\}$ analogous to $x$, so that they each represent one party in a collision/encounter. Then the injury set is defined as $I_{x,y}$. From this we calculate rates of injury as $$R_{x,y}=\frac{I_{x,y} \sqrt{\hat{D}_m\hat{D}_{m'}}}{D_xD_y}.$$

\subsection{Multinational}

We can add/subtract covariates to/from $x$ and $y$ in the above framework. For example, we can include the country, $c$, as in Figure \ref{plate}.

\begin{figure}[h]
  \centering
  \tikz{ %
    \node[latent] (sigD) {$\sigma_x^{(D)}$} ; %
    \node[latent, right=of sigD] (D0) {$D_x^{(0)}$} ; %
    \node[latent, right=of D0] (R0) {$R_{x,y}^{(0)}$} ; %
    \node[latent, right=of R0] (sigR) {$\sigma_{x,y}^{(R)}$} ; %
    \node[latent, below=of D0] (mux) {$\mu_{x,c}$} ; %
    \node[latent, below=of mux] (Dx) {$D_{x,c}$} ; %
    \node[latent, left=of mux] (lambda) {$\lambda_{x,c}$} ; %
    \node[latent, left=of lambda,xshift=-1cm] (lambda0) {$\lambda_{x}^{(0)}$} ; %
    \node[latent, below=of R0] (Rxy) {$R_{x,y,c}$} ; %
    \node[obs, below=of lambda] (T) {$T_{j,c}$} ; %
    \node[obs, below=of Rxy] (I) {$I_{x,y,c}$} ; %
    \plate[inner sep=0.25cm, yshift=0.12cm] {plateT} {(T)} {$j=1,...,J$}; %
    \plate[inner sep=0.25cm, yshift=0.12cm] {plate1} {(lambda) (plateT) (I)} {$c=1,...,C$}; %
    \edge {sigD,D0} {mux} ; %
    \edge {mux,lambda} {Dx} ; %
    \edge {sigR,R0} {Rxy} ; %
    \edge {mux,lambda} {T} ; %
    \edge {lambda0} {lambda} ; %
    \edge {Dx,Rxy} {I} ; %
  }
  \caption{\small Plate diagram relating latent distance and rate distributions to observed travel surveys and injuries. This diagram could be a single set $\{x,y\}$ or it could represent the joint distributions for all $\{x,y\}$. $D_x^{(0)}$ is the distance distribution for covariates $x$, and is described with a log-normal distribution. $R_{x,y}^{(0)}$ is the rate distribution for collision between covariates $x$ and $y$ and can also be log-normally distributed. $\sigma_x^{(D)}$ and $\sigma_{x,y}^{(R)}$ are the inter-country variability for distances and rates, respectively. We might want to impose that $\sigma_x^{(D)}>\sigma_{x,y}^{(R)}$? Then $\mu_{x,c}|D_x^{(0)},\sigma_x^{(D)}$ and $R_{x,y,c}|R_{x,y}^{(0)},\sigma_{x,y}^{(R)}$ can be log-normally distributed. $\lambda_{x,c}$ is a zero-inflated Poisson distribution counting the number of trips per person for covariates $x$, so $D_{x,c}|\mu_{x,c},\lambda_{x,c}$ is a compound Poisson log-normal distribution.  $I_{x,y,c}|D_{x,c},R_{x,y,c}$ is Poisson distributed, and $T_{j,c}|D_{x,c}$ would be a non-summed compound Poisson log-normal distribution.}
\label{plate}
\end{figure}
\begin{figure}[h]
  \centering
  \tikz{ %
    \node[latent] (sigD) {$\sigma_x^{(D)}$} ; %
    \node[latent, right=of sigD] (D0) {$D_x^{(0)}$} ; %
    \node[latent, right=of D0] (R0) {$R_{x,y}^{(0)}$} ; %
    \node[latent, right=of R0] (sigR) {$\sigma_{x,y}^{(R)}$} ; %
    \node[latent, below=of D0] (mux) {$\mu_{x,c}$} ; %
    \node[latent, below=of mux] (Dx) {$D_{x,c}$} ; %
    \node[latent, left=of mux] (lambda) {$\lambda_{x,c}$} ; %
    \node[latent, left=of lambda,xshift=-1cm] (lambda0) {$\lambda_{x}^{(0)}$} ; %
    \node[latent, below=of R0] (Rxy) {$R_{x,y,c}$} ; %
    \node[obs, below=of lambda] (T) {$T_{j,c}$} ; %
    \node[obs, below=of Rxy] (I) {$I_{x,y,c}$} ; %
    \plate[inner sep=0.25cm, yshift=0.12cm] {plateT} {(T)} {$j=1,...,J$}; %
    \edge {sigD,D0} {mux} ; %
    \edge {mux,lambda} {Dx} ; %
    \edge {sigR,R0} {Rxy} ; %
    \edge {mux,lambda} {T} ; %
    \edge {lambda0} {lambda} ; %
    \node[latent, below=of Rxy] (I) {$I_{x,y,c}$} ; %
    \node[obs, below=of I] (Ixc) {$\hat{I}_{x,c}$} ; %
    \plate[inner sep=0.25cm, yshift=0.12cm] {plate1} {(mux) (plateT) (Ixc)} {$c=1,...,C$}; %
    \edge {I} {Ixc} ; %
    \edge {Dx,Rxy} {I} ; %
  }
  \caption{\small Question: can we combine the model in Figure \ref{plate} with the case where we are missing some information in the injuries? Suppose, for example, that we don't have the details of the other vehicle ($y$), only the sum $\hat{I}_{x,c}=\sum_yI_{x,y,c}$}
\label{plate2}
\end{figure}

### 2021-08-25 latex/output-md.Rmd
Add overleaf document - with PDFS, and its converted Rmd file

# Introduction

## Summary of literature

## Summary of aims

## Overview

The remainder of the document presents a current working version of the
ITHIM, and is a comprehensive record of all the calculations of which it
is composed. The equations for the ITHIM are described in Section
[2](#equations){reference-type="ref" reference="equations"} (tabulated
in Section [7](#tableequations){reference-type="ref"
reference="tableequations"}). The method is concretised through the
example application to Accra in Section
[3](#example){reference-type="ref" reference="example"} and particular
columns in Tables [1](#indices){reference-type="ref"
reference="indices"} and [3](#inputs){reference-type="ref"
reference="inputs"}, with results presented in Section
[19](#results){reference-type="ref" reference="results"}.

The ITHIM as described here is implemented in the ITHIM-R R package
(<https://github.com/ITHIM/ITHIM-R>). Other cities are also included in
that package (São Paulo, Delhi, Bangalore), and reference will be made
to those where relevant.

# Method: ITHIM {#equations}

In the ITHIM, we combine many information inputs and compute the health
burden through three pathways (depicted in Figure
[1](#workflow){reference-type="ref" reference="workflow"}). Each pathway
has its own set of definitions and outputs, and we refer to them as
"modules".

![ITHIM workflow. Inputs to the model have a bold, black outline. The
graph depicts parent--child relationships, where a child (at the head of
an arrow) depends upon all its parents (at the source(s) of the
arrow(s)). To aid visualisation, an arrow connected to the outside of a
grey box indicates that the arrow connects to each item within the box.
*AT: active travel. MET: metabolic equivalent task. pp: per person. AP:
air pollution. PA: physical activity. DR: dose--response relationship.
GBD: global burden of disease. BOD: (local) burden of
disease.*](flowchart.pdf){#workflow width="90%"}

To describe the model, we use lower-case letters to denote indices, or
dimensions, of objects. They take one of a set of possible values,
detailed in Table [1](#indices){reference-type="ref"
reference="indices"}. The set of input items are denoted by capital
letters, and variable parameters are denoted by Greek letters. These are
presented in Table [3](#inputs){reference-type="ref"
reference="inputs"}. The equations of the ITHIM are presented below, and
summarised in Table [4](#calculationstab){reference-type="ref"
reference="calculationstab"}. These use letter modifiers to show how the
inputs are modified in the course of output generation.

We calculate the ITHIM by combining the modules' outputs in terms of the
health burden ($\hat{U}_{a,g,h,o,s}$). Other outputs of the model
include: MMETs per person ($M_{i,s}$), PM2.5 in each scenario
($\bar{P}_s$), PM2.5 per person ($\check{W}_{i,s}$), and injuries burden
($\check{I}_{a,g,m,o,s}$).

To assess uncertainty, we sample from the ITHIM output by sampling
uncertain parameters multiple times and calculating the ITHIM. We
specify parametric distributions or sampling strategies for all the
uncertain parameters. Evaluating the ITHIM with uncertain parameters
allows assessment of their impact on the outcome (AKA sensitivity
analysis). We use EVPPI to calculate the expected reduction in
uncertainty in the outcome were we to learn a parameter perfectly. This
means we can implement models that are basic in their parametrisation,
and learn at the end for which parameters it would be worthwhile
spending dedicated time learning better.

::: {.center}
::: {#indices}
  Label   Name                   Levels
  ------- ---------------------- ---------------------
  $a$     Age group              15--49
                                 50--69
  $g$     Gender                 Male
                                 Female
  $h$     Disease                
  $i$     Individual index       
  $j$     Trip index             
  $m$     Transport mode         walk
                                 cycling (cyc.)
                                 bus
                                 car
                                 motorbike (mot.)
                                 goods vehicles (GV)
                                 subway (sub.)
  $o$     Outcome                Death
                                 YLL
  $s$     Scenario               
  $w$     AP RR hyperparameter   $1,2,3,4$
  $x$     PA RR hyperparameter   $1,2,3$
  $z$     PA dose                $0,...,300$

  : Indices used in the ITHIM variables. The "Label" is the subscript
  used to refer to the index in this document. The "Name" is
  descriptive. The "Levels" are the levels used in the Accra model.
:::
:::

::: {.center}
::: {#abbrev}
  Abbreviation   Meaning
  -------------- -----------------------------------------------
  AP             Air pollution
  CDF            Cumulative distribution function
  COPD           Chronic obstructive pulmonary disease
  DR             Dose response
  EVPPI          Expected value of perfect partial information
  GAM            Generalised additive model
  GBD            Global burden of disease
  IHD            Ischemic heart disease
  ITHIM          Integrated transport and health impact model
  LRI            Lower respiratory infection
  MMET           Marginal metabolic equivalent task
  PA             Physical activity
  PAF            Population attributable fraction
  PIF            Potential impact fraction
  PM2.5          Particulate matter (diameter $< 2.5 \mu$m)
  PT             Public transport
  RR             Relative risk
  VOI            Value of information
  YLL            Years of life lost

  : Common abbreviations.
:::
:::

::: {.landscape}
::: {.center}
::: {#inputs}
  ------------------------------------------- ----------------------------------------------------------------- ---------------------------------- ------------------------------- ----------------
  [\[inputs\]]{#inputs label="inputs"}                                                                          *ITHIM-R reference*                                                
  Name                                        Description                                                       Value in Accra model                                         Label   Model/ Setting
  *Global values*                                                                                                                                                                  
  $U_{a,g,h,o}$                               Background burden of disease                                      Constant                                                  GBD_DATA          Setting
  $N_{a,g}$                                   Population by age and gender in GBD_DATA                          Constant                                                  GBD_DATA          Setting
  $\bar{N}_{a,g}$                             Population by age and gender                                      Constant                                               DEMOGRAPHIC          Setting
  $\rho_h$                                    Chronic disease scalar                                            1 or Lnorm(0,0.18)                          CHRONIC_DISEASE_SCALAR          Setting
  *Travel values*                                                                                                                                                                  
  $Z_{m=walk}$                                Walk speed                                                        4.8                                                    MODE_SPEEDS          Setting
  $Z_{m=cyc.}$                                Cycle speed                                                       14.5                                                   MODE_SPEEDS          Setting
  $Z_{m=bus}$                                 Bus speed                                                         15                                                     MODE_SPEEDS          Setting
  $Z_{m=car}$                                 Car speed                                                         21                                                     MODE_SPEEDS          Setting
  $Z_{m=mot.}$                                Motorbike speed                                                   25                                                     MODE_SPEEDS          Setting
  $\epsilon$                                  Walk time to bus                                                  5 or Lnorm(log(5),0.18)                              BUS_WALK_TIME          Setting
  $\mu_{m=HGV}$                               Truck distance relative to car                                    0.21 or Beta(3,10)                              TRUCK_TO_CAR_RATIO          Setting
  $\mu_{m=bus}$                               Bus driver distance relative to bus passenger distance            1/45 or Beta(20,600)                        BUS_TO_PASSENGER_RATIO          Setting
  $\mu_{m=mot.}$                              Motorcycle fleet distance relative to total motorcycle distance   37/68 or Beta(37/3+1,31/3+1)             FLEET_TO_MOTORCYCLE_RATIO          Setting
  $T_j$                                       Trip-level data                                                   Constant                                           SYNTHETIC_TRIPS          Setting
  *Injury model values*                                                                                                                                                            
  $I_{a,g,m_{\text{vic}},m_{\text{str}}}$     Fatalities table                                                  Constant                                              INJURY_TABLE          Setting
  $\sigma$                                    Injury reporting rate                                             1 or Beta(8,3)                               INJURY_REPORTING_RATE          Setting
  $\omega$                                    Sum of distance exponents                                         1.9 or Lnorm(log(1.9),log(1.03))                  SIN_EXPONENT_SUM            Model
  $\psi$                                      Casualty fraction of $\omega$                                     0.5 or Beta(20,20)                    CASUALTY_EXPONENT\_ FRACTION            Model
  *Pollution model values*                                                                                                                                                         
  $\eta$                                      Background PM2.5 concentration                                    50 or Lnorm(log(50),0.18)                             PM_CONC_BASE          Setting
  $\zeta$                                     Fraction of measured PM2.5 concentration due to road transport    0.225 or Beta(5,20)                                 PM_TRANS_SHARE          Setting
  $\gamma_{m}$                                Vehicle emission inventory                                        Constant or Dirichlet                           EMISSION_INVENTORY          Setting
  $P$                                         Vehicle emission confidence                                       1                                    EMISSION_INVENTORY_CONFIDENCE          Setting
  *Pollution & health model values*                                                                                                                                                
  $C_1$                                       Base-level inhalation rate                                        1                                     BASE_LEVEL_INHALATION \_RATE            Model
  $C_2$                                       Exposure ratio window closed                                      0.5                                       CLOSED_WINDOW_PM\_ RATIO            Model
  $C_3$                                       Proportion of travel with closed windows                          0.5                                            CLOSED_WINDOW_RATIO          Setting
  $C_4$                                       Parameter for on-road PM2.5 level                                 3.216                                               ROAD_RATIO_MAX            Model
  $C_5$                                       Parameter for on-road PM2.5 level                                 0.379                                             ROAD_RATIO_SLOPE            Model
  $C_6$                                       Exposure ratio for underground                                    0.8                                                SUBWAY_PM_RATIO          Setting
  $H_{a,h,w}$                                 AP dose--response--curve parameters                               Constant                                                     DR_AP            Model
  $\xi_{h,w}$                                 Quantiles for AP RR functions                                     0.5 or Unif(0,1)                       AP_DOSE_RESPONSE\_ QUANTILE            Model
  *Physical activity model values*                                                                                                                                                 
  $\chi$                                      Day-to-week travel scalar                                         7 or 7$\times$Beta(20,3)               DAY_TO_WEEK_TRAVEL\_ SCALAR            Model
  $Y_i$                                       Individual non-transport MMETs                                    Constant                                PA_SET\$work_ltpa_marg_met          Setting
  $\theta$                                    Background PA scalar                                              1 or Lnorm(0,0.18)                            BACKGROUND_PA_SCALAR          Setting
  $\tilde{Y}$                                 Confidence in PA survey                                           1                                         BACKGROUND_PA_CONFIDENCE          Setting
  $\tilde{\theta}$                            Background PA zeros quantile                                      0.5 or Unif(0,1)                               BACKGROUND_PA_ZEROS          Setting
  $\lambda_{m=\text{cyc.}}$                   Cycling MET surplus                                               4.63 or Lnorm(log(4.63),1)                            MMET_CYCLING            Model
  $\lambda_{m=\text{walk}}$                   Walking MET surplus                                               2.53 or Lnorm(log(2.53),1)                            MMET_WALKING            Model
  $\lambda_{m\not\in\text{walk,cyc.}}$        In-vehicle MET surplus                                            0                                                                             Model
  *Physical activity & health model values*                                                                                                                                        
  $G_{h,x,z}$                                 Look-up table for relative risk truncated normal distribution     Constant                                                                      Model
  $\phi_{h}$                                  Quantiles for PA RR functions                                     0.5 or Unif(0,1)                       PA_DOSE_RESPONSE\_ QUANTILE            Model
  ------------------------------------------- ----------------------------------------------------------------- ---------------------------------- ------------------------------- ----------------

  : All values input to the ITHIM model.
:::
:::
:::

## Data module {#data}

Input data for the ITHIM are listed as Roman capital letters in Table
[3](#inputs){reference-type="ref" reference="inputs"}. They include
(vehicle) mode speeds, injury/fatality records, disease dose--response
relationships, measures of air pollution, surveys of travel and physical
activity, population numbers by demographic group, MMETs associated with
active travel, and current burden of disease (see Figure
[1](#workflow){reference-type="ref" reference="workflow"}).

### Synthetic {#synth}

We create a "synthetic population" by matching individuals surveyed for
travel to the individuals surveyed for physical activity, based on age
and gender. Then each person in the synthetic population has a set of
trips taken and an amount of non-travel physical activity. This
population is assumed to be representative of the population of the city
or setting under study.

#### Alternatively/additionally

we could create a synthetic population as described in Appendix
[8.3](#synthpop){reference-type="ref" reference="synthpop"}, which
allows uncertainty in the propensity to travel. This method operates
with person-level, rather than trip-level, information.

### Scenarios

From the synthetic population we create "scenarios" of different
pictures of travel by changing certain trips (e.g. swapping the mode of
travel for particular trips and/or particular people). For the purposes
of illustration, we choose a simple scenario for Accra in which every
person gains one one-kilometre walk.

### Distances

For the ITHIM, we process the travel data into distance data of various
formats. First, we augment PT trips with a walk component: we take the
set $\{T_j:m(j)=\text{bus}\}$, which has $J$ entries. We add $J$
journeys with mode "walk" and duration $\epsilon$ to the set $T$. (This
step would not be necessary for a travel survey which already includes
all stages for each trip.)

Then we summarise the total distances and durations. The distance set is
labelled $\hat{T}_{m,s}$, representing the total distance travelled per
mode per scenario, and the duration set is constructed analogously. We
make an assumption that the distance for bus drivers scales linearly
with bus passengers, where the distance in the baseline is defined based
on the bus driver distance relative to car.

Distance and duration summaries include all person travel in the
synthetic population and, in addition, some trips that are attributed to
no-one in the synthetic population. These trips are used to represent
"fleet" travel, which is uncaptured in travel surveys, and unchanged in
scenarios. It's important to include these trips because for modes such
as motorcycle, where a considerable component of travel comes from
fleet, attributing all the harms associated with the mode to only those
surveyed dramatically overestimates the contribution of the individuals.
In scenarios, then, we would be predicting emissions and injury rates
that are far too high per person.

## AP module

### Background PM2.5 {#dir}

We use the total distance by mode ($\hat{T}_{m,s}$) to calculate the
total PM2.5 in the scenario ($\bar{P}_s$). First we calculate the
emission inventory in terms of fractions. If the confidence $P$ is 1
then we use $\dot{P}_{m}=\gamma_m$ directly. If $P<1$, we sample the
emission inventory fractions as follows:
$$\tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)),$$
$$\dot{P}_m=\tilde{\gamma}_m$$ where we choose some mapping function
such as $f(P,\gamma_m)=\frac{\gamma_m}{\sum_m\gamma_m}10^{5P}$ (Figure
[5](#emission_dist){reference-type="ref" reference="emission_dist"}).

Then we multiply vehicle distance by vehicle emission factor (where the
emission factor is the emission inventory (fraction) divided by distance
at baseline).
$$\tilde{P}_{m,s}=\dot{P}_m\frac{\hat{T}_{m,s}}{\hat{T}_{m,s=\text{baseline}}}.$$

For the baseline, $\tilde{P}_{m,s}$ is equal to the emission inventory
fraction. For each scenario, it is scaled. We take the sum over modes to
get a scalar for emissions in scenarios.
$$\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}.$$ Then the background PM2.5 in
each scenario is the sum of the transport component and the
non-transport component: $$\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta).$$

### PM2.5 per person

Dose--response relationships are based on background concentrations in
cities, and not on personal exposures. In our person-level model, we
expect that individuals will have different exposures based on their
different travel behaviours. While we can neither estimate individual
exposures nor map them via any dose--response relationship to a relative
risk, we aim to capture interpersonal differences via assumptions about
the mechanistic relationships between behaviours and exposures, and
anchor those to the background air pollution in their locality, for
which dose--response estimates of relative risk exist.

Therefore, when we write "individual exposure", it should not be read as
an aetiologically relevant exposure, but rather it should be read as a
unit- and scale-agnostic metric that maps an individual to a particular
dose with reference to city-level measurements of air pollution. New
research is required to elicit dose--response relationships between
individual health exposures and health outcomes \[citation from JPO?\].
Our approximations act as assumptions for now and placeholders until
more detailed data are available. \[Perhaps we should use words other
than "individual exposure". How about "individual PM2.5 reference
values"?\]

Individual exposures to PM2.5 are calculated using the background PM2.5
($\bar{P}_s$) and the trip sets. There are three major components to
daily exposure: one, a person's total inhalation off road; two, a
person's inhalation on road in a vehicle, and, three, a person's
inhalation on road while undertaking active transport. Each category has
an amount of background PM2.5 and a ventilation rate which together
inform overall exposure.

The ratio of exposure off road to that on road is a function of total
PM2.5, defined as $$K_s=C_4-C_5\log(\bar{P_s})$$ as in @Goel2015. This
defines the exposure of a person in an open vehicle (i.e. pedestrian,
cyclist or motorist):
$$\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$$ and it is used to
calculate in-vehicle exposure, assuming an exposure of $C_2$ with the
window closed, and a proportion $C_3$ of vehicles having closed windows:
$$\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3).$$
The exposure in a subway is constant and not dependent on the road ratio
$K_s$: $$\tilde{K}_{m=\text{sub.},s}=C_6.$$

Ventilation rates are calculated for each mode, assuming a base-level
inhalation rate of $C_1$: $$V_m=C_1+\frac{1}{2}\lambda_m,$$ where
$\lambda_m$ is the MMETs for mode $m$.[^1] Then the air inhaled during
travel per person is
$$\tilde{V}_{i,m,s} = \smashoperator{\sum_{\substack{j:i(j)=i,\\m(j)=m,\\s(j)=s}}}T_j\cdot V_{m(j)}/Z_{m(j)}$$
and the rate of inhalation during travel is
$$\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}.$$ The
air inhaled when not travelling is
$$\hat{V}_{i,s} = C_1\left(24-\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}T_j/Z_{m(j)}\right).$$
Together, the PM2.5 exposure is calculated as the total PM2.5 inhaled
per hour as follows:
$$\check{W}_{i,s} = \frac{\bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})}{24}.$$

### AP--disease dose--response relative risk

We use each person's exposure to PM2.5 ($\check{W}_{i,s}$) to calculate
their relative risk of five diseases (IHD, lung cancer, COPD, stroke,
LRI), using curves parametrised by four disease-specific variables
[@Burnett2014]. Of the five diseases, two (IHD and stroke) have
parameters specific to age groups starting at age 25.[^2] The other
three (lung cancer, LRI and COPD) have one set of parameters for all
ages.

The curves are in the form of samples of the set of four parameters. We
model the densities of these samples (using a quantile for parameter 3,
kernel density estimation for parameter 2, and GAMs for parameters 1 and
4)[^3] in order to draw either the median or random samples via their
quantiles ($\xi_{h,w}$) as follows:

$$\begin{aligned}
\tilde{H}_{a,h,w=3}=&\text{CDF}_{H_{a,h,w=3}}^{-1}(\xi_{h,w=3})\\
\tilde{H}_{a,h,w=2}=&\text{CDF}_{H_{a,h,w=2}|\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=2})\\
\tilde{H}_{a,h,w=1}=&\text{CDF}_{H_{a,h,w=1}|\tilde{H}_{a,h,w=2},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=1})\\
\tilde{H}_{a,h,w=4}=&\text{CDF}_{H_{a,h,w=4}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=2},\tilde{H}_{a,h,w=1}}^{-1}(\xi_{h,w=4}).\end{aligned}$$

From these parameters, the relative risk of mortality is defined
$$\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=1}\times
\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=2}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=4}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}.$$
For $h\not\in\{\text{IHD, lung cancer, COPD, stroke, LRI}\}$,
$\check{H}_{h,i,s} \equiv 1$.

## PA module

### Individual-level MMETs {#beta}

Using trip sets and the synthetic population, we calculate total MMETs
per person ($M_{i,s}$) as the sum of walking MMETs per day and cycling
MMETs per day, which are scaled up to a week via the scalar $\chi$, and
work/leisure MMETs per week as follows:
$$M_{i,s}=\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}
\chi\lambda_{m(j)}\frac{T_j}{Z_{m(j)}}+\theta Y_i.$$

Here, $\lambda_m$ is the MMETs for mode $m$, $T_j$ the distance of trip
$j$, $Z_m$ the speed of mode $m$, $\theta$ the scalar for background PA,
and $\check{Y}_i$ the amount of work and leisure (background) PA of
person $i$.

We calculate $\check{Y}_i$ according to the confidence in the PA survey,
$\tilde{Y}$. First, we calculate the raw probability that a person in
demographic group completes no non-travel PA:
$$\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}.$$ Then, we set the
probability to use, $\bar{Y}_{a,g}$, as $\dot{Y}_{a,g}$ if our
confidence $\tilde{Y}$ is 1. If $\tilde{Y}<1$, we map the confidence to
parametrise a Beta distribution somehow, e.g. (Figure
[4](#prob_zero_PA){reference-type="ref" reference="prob_zero_PA"})
$$\begin{aligned}
\hat{Y} &= 500^{\tilde{Y}+0.2},\\
\beta & = \hat{Y}\dot{Y}_{a,g}\left(\frac{1}{\dot{Y}_{a,g}} -1\right),\\
\alpha & = \hat{Y} - \beta,\\
\bar{Y}_{a,g} & = \text{CDF}_{\text{Beta}(\alpha,\beta)}^{-1}(\tilde{\theta}).\end{aligned}$$

Finally, for each person in the population, we sample non-travel MMETs
as zero with probability $\bar{Y}_{a,g}$ and from the raw non-zero
density of their demographic group with probability $1-\bar{Y}_{a,g}$.

### PA--disease dose--response relative risk

We use each person's MMETs per week (${M}_{i,s}$) to calculate their
relative risk of six diseases, using curves found from meta analysis.
Each disease (except type 2 diabetes) has a threshold beyond which there
is no further change in relative risk. (Total cancer, breast cancer,
colon & rectum cancer, endometrial cancer, & coronary heart disease: 35;
lung cancer: 10; stroke: 32; all cause: 16.08.) For all individual
cancers, we estimate the burden in terms of . For all other diseases, we
use mortality.

The interpolation of the dose--response curve can be described as
follows:
$$\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$$
where we interpolate the mean, the lower bound, and the upper bound.
These define the normal from which we sample, truncated at 0:
$$F_S(v_x) = \mathcal{N}\left.\left(v_{x=2},\frac{v_{x=3}-v_{x=1}}{1.96}\right)\right|_{F_S(v_x)>0}.$$

Then the relative risk for each person for each disease is calculated
for the quantile $\phi_h$ as
$$W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right).$$
See Figure [2](#cancer){reference-type="ref" reference="cancer"} for an
example. For $h\not\in$ {diabetes, total cancer, breast cancer, colon &
rectum cancer, endometrial cancer, IHD, lung cancer, stroke, all cause},
$W_{h,i,s} \equiv 1$.

![Example of the dose--response workflow for PA and total cancer. Top
left: results of the metaanalysis: mean, lower bound and upper bound of
the relationship between PA dose and relative risk of disease. Top
right: examples of 500 samples from the normal distribution defined by
the mean of the top-left panel and the standard deviation defined by
(upper bound minus lower bound) divided by 1.96 and truncated at 0.
Bottom left: relative risks of individuals with PA doses 5, 25 and 45
are found by mapping their doses onto the median curve. This is used in
the "constant" ITHIM use case. Bottom right: relative risks of
individuals with PA doses 5, 25 and 45 for three different samples from
the distribution shown in the top-right panel. For each sample, each
individual is mapped onto the response curve it defines.
](cancer_PA_dose_response.pdf){#cancer width="60%"}

## Injury module

### Processing

For the injury module, we require distance per travel mode per
demographic group per scenario. It requires matching of mode names in
the travel survey to mode names in the injury records, for example
"taxi", "shared auto", "shared taxi" and "car" combine to form "car".
Some work is required to separate drivers from passengers; we currently
assume all travellers are drivers, with the exception of "bus".[^4]

### Modelling

We model injuries via regression by predicting the number of fatalities
of each demographic group on each mode, which we calculate as a sum over
all the ways in which they might have been injured (i.e., all the modes
with which they might have collided).

The predictive covariates include the distances travelled by the parties
and their demographic details. These requirements lead to a natural
separation of the (training) dataset into two groups: a set for which we
have distance data for the other party, and a set for which we do not
have distance data for the other party. The former equates to a "who hit
whom" ("whw") matrix (albeit in a higher dimension), and will account
for changes in injuries resulting of a change in strike-mode travel. The
latter corresponds to causes of injury that will not change across
scenarios, including "no other vehicle" and modes of transport that we
do not consider to change, which might include trucks and buses if they
are not somehow explicitly included in the trip set. We label this group
"noov": no or other vehicle.

We model the number of injuries as a Poisson-distributed variable with
an offset depending on the distance(s) and the reporting rate. We write
the model as follows: $$\begin{aligned}
\tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s,y} \sim &
\text{Poisson}(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}); \\
% 'count~
\log\left(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}|m_{\text{str}}\in\{m_{\text{whw}}\}\right) = &
% cas_mode*strike_mode-log(injury_reporting_rate)+log(weight)
{\kappa_0} +\sum_{i=1}^n\kappa_iX_i-\log(\sigma)+\nonumber\\
% +log(cas_distance)
 & \log(\bar{T}_{a,g,m_{\text{cas}},s})
 % +(CAS_EXPONENT-1)*log(cas_distance_sum)
 +(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s})
 %+log(strike_distance)+(STR_EXPONENT-1)*log(strike_distance_sum)
 +E_{m_{\text{str}}}\log(\hat{T}_{m_{\text{str}},s});\\
\log\left(\bar{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}|m_{\text{str}}\in\{m_{\text{noov}}\}\right) =&
{\kappa'_0}
+\sum_{i=1}^n\kappa'_iX'_i 
-\log(\sigma)
% +log(cas_distance)
+ \log(\bar{T}_{a,g,m_{\text{cas}},s})
 % +(CAS_EXPONENT-1)*log(cas_distance_sum)
 +(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s}).\end{aligned}$$
We choose this form of equation to enable (a) linearity in injuries with
respect to total travel combined across modes and (b) linearity in
injuries as subdivided by travellers of each mode.[^5] Note that this
methodology is under development. See
<https://github.com/robj411/safety_in_numbers_power_law>.

In general, one makes the best model that one can with the data
available. We use, for Accra, the covariates
`casualty mode*strike mode, casualty age,` and `casualty gender`, to
form the model matrix $X$. We also have another covariate, `year`, but
our travel data are for one year only, so we model instead many
observations of the same quantities.

We fit the coefficients $\kappa$ of the model using data for ten years,
assuming the same travel each year, which corresponds to the scenario
$s=\text{baseline}$. To incorporate the injury reporting rate
($\sigma$), we set it as an offset, whose value is 1 in fitting the
model, and as specified in making predictions.

### Prediction

We predict the number of injuries in each scenario based on the training
model built from the baseline scenario. For Accra, we predict for the
year 2016, using scenario-specific travel data. The number of fatalities
is taken as the sum of fatalities over the strike modes:
$$\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}.$$
We extrapolate injury deaths to injury YLLs via the ratio in the GBD:
$$R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}},$$
$$\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g}.$$

## Health module

### PA and AP relative risks combined

We combine the relative risks of disease as a function of AP and disease
as a function of PA through multiplication:
$$\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s}.$$ Not all diseases
have a dose--response relationship for both AP and PA. Just stroke, lung
cancer, and IHD have both. For the other diseases, only one of
$W_{h,i,s}$ and $\check{H}_{h,i,s}$ is different from one.

### Total disease burden

We calculate the total health burden relative to the reference scenario
(which, for our Accra example, is the baseline) using the injury and
health outputs combined with GBD data, via population attributable
fractions (PAF). The PAF is defined:
$$\hat{W}_{a,g,h,s}=\smashoperator{\sum_{\substack{i:a(i)=a,\\g(i)=g}}}\tilde{W}_{h,i,s}.$$

We then calculate the PAF relative to the reference scenario ("ref") as:
$$\bar{W}_{a,g,h,s}= \frac{\hat{W}_{a,g,h,s=\text{ref}}-\hat{W}_{a,g,h,s}}{\hat{W}_{a,g,h,s=\text{ref}}}.$$

We estimate the background burden of disease using GBD data and scaling
based on the ratio of populations between country and city. For country
populations $N_{a,g}$, city populations $\bar{N}_{a,g}$ and country
burden $U_{a,g,h,o}$, we estimate city burden as
$$\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}.$$

If we are scaling the background burden of non-communicable diseases
("Neoplasms", "Ischemic heart disease", "Tracheal, bronchus, and lung
cancer", "Breast cancer", "Colon and rectum cancer", "Uterine cancer"),
we do so here. (Otherwise set $\rho\equiv 1$.)
$$\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}.$$

We combine the burden with the PAF:
$$\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}.$$

### Injury burden

And, finally, for the injuries, we sum over modes to compute the burden,
$$\bar{U}_{a,g,o,s}=\sum_m\check{I}_{a,g,m,o,s},$$ and subtract from the
values for the reference scenario:
$$\hat{U}_{a,g,h=\text{road injury},o,s}=\bar{U}_{a,g,o,s=\text{ref}}-\bar{U}_{a,g,o,s}.$$

# Example: Accra {#example}

Here we show the ITHIM as evaluated for the city of Accra, with a simple
scenario of every person in the synthetic population gaining a single, 1
km walk. We define XX uncertain univariate parameters (Section
[3.1](#uni){reference-type="ref" reference="uni"}), two confidences, and
two uncertain parameter sets (Section [3.3](#dr){reference-type="ref"
reference="dr"}).

## Parametric distributions for uncertain variables {#uni}

The XX uncertain univariate parameters are shown in Figure
[3](#distributions){reference-type="ref" reference="distributions"}.
Most of the parameters are self explanatory. Cycling and walking MMETs
are the number of MMETs per hour when undertaking cycling and walking,
and determine also the ventilation rates. Motorcycle distance is the
total distance travelled by motorcycles relative to the total distance
travelled by cars in the baseline scenario. Non-travel PA, injury
reporting rate and NCD burden all act as scalars for the corresponding
datasets. Note that the non-travel PA scalar does not affect the
$\approx$`<!-- -->`{=html}40% of the population whose non-travel PA is
0.

![Distributions of univariate uncertain parameters (1024
samples).](uncertain_parameters.pdf){#distributions width="95%"}

## Confidences

For physical activity propensity and emission inventories, we use
"confidences", values between 0 and 1 that represent how confident we
are about the data source. We map these values to Beta (Figure
[4](#prob_zero_PA){reference-type="ref" reference="prob_zero_PA"}) and
Dirichlet (Figure [5](#emission_dist){reference-type="ref"
reference="emission_dist"}) distributions, respectively. The
Beta-distributed propensity towards non-travel physical activity is
described in Section [2.3.1](#beta){reference-type="ref"
reference="beta"}. The Dirichlet-distributed emissions inventory is
described in Section [2.2.1](#dir){reference-type="ref"
reference="dir"}.

![Distributions of a Beta-distributed propensity to non-travel PA with
varying confidence values in the raw input data. The sampled value gives
the fraction of the population that engages in non-travel PA.
](prob_zero_PA.pdf){#prob_zero_PA width="50%"}

![Distributions of Dirichlet-distributed emission inventories for four
modes and varying confidence values. Raw input values were 4, 4, 20, and
60. ](emission_dist.pdf){#emission_dist width="65%"}

## Dose--response relationships {#dr}

For the dose--response relationships between physical activity (PA) and
disease and air pollution (AP) and disease, we assume that there is
uncertainty, but no variability, in the relationship. This means that we
sample a relationship from the distribution of relationships, and apply
that relationship to all individuals precisely. This means that, given
fixed doses, responses between individuals will be perfectly correlated.

We achieve this by use of the probability integral transform: we sample
a random variable uniformly distributed on the space (0,1) and map it,
via a cumulative distribution function, to the distribution describing
the dose--response relationship. See Figure
[6](#cdf){reference-type="ref" reference="cdf"} for an illustration.

![We use the same uniform samples to generate random variables from
different distributions. It means that the samples are perfectly
correlated.](DR_transform.pdf){#cdf width="80%"}

### Physical activity

Each disease's PA dose--response relationship is defined by a truncated
normal. For each dose, there is a mean value, an upper bound, and a
lower bound. For each person's dose, we get the response by mapping the
uniform random variable onto the truncated normal defined by the mean
and bounds for that dosage.

### Air pollution

For the AP relationship, there are four parameters per disease. We
sample the first from an empirical distribution using the probability
integral transform. We sample the second via the same method,
conditioned on the value of the first, constructing their joint density
with e.g. kde2d. The third parameter is sampled conditioned on the first
and second, constructing their joint density using a GAM. The final
parameter is sampled conditioned on the first, second, and third,
constructing their joint density using a GAM.

As before, there is perfect correlation between individuals, i.e. if
person A's dose is greater than person B's, then person A's response is
strictly greater than person B's response.

The empirical distributions come from @Burnett2014. There are four
parameters per disease: IHD, lung cancer, COPD, stroke, and LRI. In
addition, for stroke and IHD, there is a set of four parameters for each
age group from 25 to 95 in five-year increments. In addition to our
assumption that there is perfect correlation between individuals for
diseases, we assume perfect correlation between ages for diseases. I.e.,
our four quantiles per disease will be applied to all age groups.

# Results

## Health burdens in scenarios

![YLLs in the scenario minus YLLs in the baseline for
Accra.](accra_walk_yll.pdf){#results width="80%"}

## Value of information

We calculate value of information following @Jackson2017. The "value of
information" we implement is the expected value of perfect partial
information (EVPPI). It is the expected amount by which variation in the
outcome diminishes should a parameter, or set of parameters, be known
perfectly. With reference to information theory, it is the mutual
information between the parameter, or parameter set, and the outcome,
where the outcome is described only by its second moment (its variance).

The objective of learning the value of information formulated in this
way is to identify the parameter or parameters that most drive
variability (variance) in the outcome. The results guide future
research, identifying which area or areas offer the greatest benefit in
terms of prediction accuracy given further research. For example, for
our model of Accra, learning better the dose--response relationship
between physical activity and stroke (incidence) would likely increase
precision of our estimate of total years of life lost (YLL) more so than
learning any other parameter (Figure [8](#evppi){reference-type="ref"
reference="evppi"}).

![EVPPI for Accra's "walking" scenario for all causes of YLLs excluding
"all cause" and "neoplasm".](accra_walk_evppi.pdf){#evppi width="70%"}

::: {.appendix}
# ITHIM with parameters

![ITHIM-R workflow, with variable parameters in purple. Fixed inputs to
the model have a bold, black outline, which is solid for inputs we
consider "global" and dashed for inputs we consider "local". There are
four global inputs, which will be embedded in ITHIM-R. There are seven
local inputs, which users should provide. The graph depicts
parent--child relationships, where a child (at the head of an arrow)
depends upon all its parents (at the source(s) of the arrow(s)). To aid
visualisation, an arrow connected to the outside of a grey box indicates
that the arrow connects to each item within the box. *AT: active travel.
MET: metabolic equivalent task. pp: per person. AP: air pollution. PA:
physical activity. DR: dose--response relationship. GBD: global burden
of disease.* NB: $\rho$ only impacts the NCD burden of GBD, not all of
it. Missing parameter: motorcycle distance. $\lambda$ is a 2D parameter.
$\xi$ is 4D.](ithim_with_parameters.pdf){#ithim_with_parameters
width="\\textwidth"}

# Dose--response relationships {#doseresponse-relationships}

![1000 samples from PA dose--response
curves](PA_dose_response.pdf){#PADR width="90%"}

  ------------------------------------------------------------------------------------ ---------------------------------------------------------------------------------------
   ![100 samples from AP dose--response curves](cvd_ihdDR-sim.pdf){#APDR width="40%"}   ![100 samples from AP dose--response curves](cvd_strokeDR-sim.pdf){#APDR width="40%"}
  ------------------------------------------------------------------------------------ ---------------------------------------------------------------------------------------

  ------------------------------------------------------------------------------------- -------------------------------------------------------------------------------- --------------------------------------------------------------------------------------
   ![100 samples from AP dose--response curves](neo_lungDR-sim.pdf){#APDR width="30%"}   ![100 samples from AP dose--response curves](lriDR-sim.pdf){#APDR width="30%"}   ![100 samples from AP dose--response curves](resp_copdDR-sim.pdf){#APDR width="30%"}
  ------------------------------------------------------------------------------------- -------------------------------------------------------------------------------- --------------------------------------------------------------------------------------

::: {.landscape}
# Tabulated ITHIM equations {#tableequations}

::: {.center}
::: {.ThreePartTable}
::: {.TableNotes}
CDF: cumulative distribution function. We use generative distributions
to generate random numbers so that samples are correlated. For
$t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$
with CDF CDF$_F$. (I'd like to find an improved notation for this.)
:::

::: {#calculationstab}
  Name                                                                                                                                                                                    Description                                                                                      Label
  --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- ------------------------------------------------------------------------------------------------ -------------------------------------------
  *Scenario generation*                                                                                                                                                                                                                                                                    
  \...                                                                                                                                                                                                                                                                                     
  *Travel calculations*                                                                                                                                                                                                                                                                    
  $\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk}\}\, \forall\, T_j:m(j)=\text{bus}$                                                                                           New walk trips to bus                                                                            walk_to_bus_and_combine_scen
  $\bar{T}_{a,g,m,s}=\sum_{j:a(j)=a,g(j)=g,m(j)=m,s(j)=s}Q_{i(j)}T_j$                                                                                                                     (Weighted) total travel per age, gender, mode and scenario                                       
  $\check{T}_{a,g,m,s}=\frac{\bar{T}_{a,g,m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}$                                                                                                             Relative travel per mode and scenario by age and gender                                          inj_distances\[\[1\]\]
  $\hat{T}_{m,s}=\sum_{a,g}\bar{T}_{a,g,m,s}$                                                                                                                                             Total travel per mode and scenario                                                               dist
  $\tilde{T}_{m,s}=\frac{\hat{T}_{m,s}}{\hat{T}_{m,s=\text{baseline}}}$                                                                                                                   Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1   inj_distances\[\[2\]\]
  *Injury calculations*                                                                                                                                                                                                                                                                    
  $E_{m_{\text{cas}}}=\psi\omega$                                                                                                                                                         Casualty exponent                                                                                CAS_EXPONENT
  $E_{m_{\text{str}}}=\omega-E_{m_{\text{cas}}}$                                                                                                                                          Striker exponent                                                                                 STR_EXPONENT
  $\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}=\mathcal{S}\left(\sigma\cdot {I}_{a,g,m_{\text{vic}},m_{\text{str}}},\bar{T}_{a,g,m,s},E_{m_{\text{cas}}},E_{m_{\text{str}}}\right)$   Fatalities in scenarios by strike mode                                                           injuries
  $\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}$                                                                               Fatalities in scenarios                                                                          victim_deaths
  $R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}}$                                                                                     Road YLL-to-death ratio                                                                          GBD_INJ_YLL
  $\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g}$                                                                                                      YLLs scaled from fatalities                                                                      deaths_yll_injuries
  *Pollution calculations*                                                                                                                                                                                                                                                                 
  $\dot{P}_{m}=\left\{\begin{array}{lr}\gamma_m & P=1 \\ \tilde{\gamma}_m, \tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)) & P<1 \end{array}\right.$                                      Emissions by mode                                                                                trans_emissions
  $\tilde{P}_{m,s}=\dot{P}_m\hat{T}_{m,s}/\hat{T}_{m,s=\text{baseline}}$                                                                                                                  Emissions by mode in scenario relative to baseline                                               trans_emissions
  $\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}$                                                                                                                                                  Emissions scalar for scenarios                                                                   baseline_sum
  $\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta)$                                                                                                                                              Background PM2.5 concentration in scenarios                                                      scenario_pm
  $K_s=C_4-C_5\log(\bar{P_s})$                                                                                                                                                            On-road PM2.5 exposure ratio                                                                     on_road_off_road_ratio
  $\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$                                                                                                                                       On-road PM2.5 exposure ratio for open vehicle                                                    on_road_off_road_ratio
  $\tilde{K}_{m\in\{\text{sub.}\},s}=C_6$                                                                                                                                                 Subway PM2.5 exposure ratio                                                                      subway_ratio
  $\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3)$                                                                                                                In-vehicle PM2.5 exposure ratio                                                                  in_vehicle_ratio
  $V_m=C_1+\frac{1}{2}\lambda_m$                                                                                                                                                          Ventilation rates                                                                                vent_rates
  $\tilde{V}_{i,m,s} = \sum_{j:i(j)=i,m(j)=m,s(j)=s}T_j\cdot V_{m(j)}/Z_{m(j)}$                                                                                                           In-travel air inhaled                                                                            
  $\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}$                                                                                                                        In-travel PM2.5 inhalation rate                                                                  
  $\hat{V}_{i,s} = C_1\left(24-\sum_{j:i(j)=i,s(j)=s}T_j/Z_{m(j)}\right)$                                                                                                                 Non-travel air inhaled                                                                           
  $\check{W}_{i,s} = \bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})/24                                                                                                                            Total PM2.5 inhaled per hour                                                                     
  %/(\sum_m\tilde{V}_{i,m,s}+\hat{V}_{i,s})$                                                                                                                                                                                                                                               
  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                                                              AP dose--response--curve parameter                                                               DR_AP_LIST
  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                                          AP dose--response--curve parameter                                                               DR_AP_LIST
  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                      AP dose--response--curve parameter                                                               DR_AP_LIST
  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                  AP dose--response--curve parameter                                                               DR_AP_LIST
  $\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=3}\times$                                                                                                                                                                                                                                 
  $\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=3}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=3}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}$                                             RR of disease given pollution                                                                    RR_AP_calculations
  *Physical activity calculations*                                                                                                                                                                                                                                                         
  $\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}$                                                                                                                                     Raw probability non-travel MMETs = 0 for demographic group $a,g$                                 raw_zero
  $\bar{Y}_{a,g}=\left\{\begin{array}{lr} \dot{Y}_{a,g} & \tilde{Y}=1 \\ \text{CDF}_{\text{Beta}(f(\tilde{Y},\dot{Y}_{a,g}))}^{-1}(\tilde{\theta}) & \tilde{Y}<1 \end{array}\right.$      Probability non-travel MMETs = 0 for demographic group $a,g$                                     zeros
  $\check{Y}_i = \left\{\begin{array}{lr} 0 & \bar{Y}_{a(i),g(i)} \\ \text{Sample}(\{Y_{i'}\}|_{a(i')=a(i),g(i')=g(i),Y_{i'}>0}) & 1-\bar{Y}_{a(i),g(i)} \end{array}\right.$              Non-travel MMETs                                                                                 SYNTHETIC_POPULATION\$ work_ltpa_marg_met
  $M_{i,s}=\sum_{j:i(j)=i,s(j)=s}(\chi T_j\cdot \lambda_{m(j)}/Z_{m(j)})+\theta \check{Y}_i$                                                                                              Total MMETs per person                                                                           mmets
  $\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$           Interpolated dose--response parameters                                                           rr, lb, ub
  $F_S(s_x) = \mathcal{N}(s_{x=2},(s_{x=3}-s_{x=1})/1.96)|_{0<F_S(s_x)}$                                                                                                                  Truncated normal function                                                                        
  $W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right)^{\dagger}$                                                                                   Relative risk of disease given PA                                                                RR_PA_calculations
  *Burden-of-disease calculations*                                                                                                                                                                                                                                                         
  $\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s}$                                                                                                                                     Combined PA and AP relative risks                                                                RR_PA_AP_calculations
  $\hat{W}_{a,g,h,s}= \sum_{i:a(i)=a,g(i)=g}\tilde{W}_{h,i,s}$                                                                                                                            Population-attributable fractions                                                                pif_temp
  $\bar{W}_{a,g,h,s}= (\hat{W}_{a,g,h,s=\text{scen1}}-\hat{W}_{a,g,h,s})/\hat{W}_{a,g,h,s=\text{scen1}}$                                                                                  Population-attributable fractions relative to scenario 1                                         pif_scen
  $\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$                                                                                                                            GBD scaled to local population                                                                   DISEASE_BURDEN
  $\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}$                                                                                                                                    Scaled background disease for Neoplasms, IHD, Lung cancer                                        gbd_data_scaled
  $\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}$                                                                                                                         Combined health and PIF                                                                          yll_dfs, death_dfs
  $\hat{U}_{a,g,h=\text{road injury},o,s}=\sum_m\check{I}_{a,g,m,o,s}$                                                                                                                    Injury health burden                                                                             inj

  : All calculations in the ITHIM.
:::
:::
:::
:::

# Uncertainty in travel data

We model uncertainty in travel via the definition of "travel attributes"
(in the terminology of @Mohammadian2010). Specifically, we define the
propensity to travel in a day, and the distance travelled in a day,
given that travelled occurred. Each travel attribute is defined per
demographic group, per mode.

In contrast to other travel-related synthetic populations, we have four
demographic groups (rather than clusters) which consist of individuals
(rather than households). A crucial difference between our objectives
and the existing work on simulating populations in travel-demand models
(that I've found so far) is that we don't need trip-level data for our
calculations. Simulating trip-level data is computationally expensive
and high-dimensional in terms of uncertain parameters. For our
downstream computations, we need only summary data: travel per person
per mode. Beyond scenario generation, there is no need for us to
consider trips.

E.g. @Saadi2016 focuses on (a) synthetic population via matching
covariates and (b) the sequence of trips taken by each person. (a) isn't
a big thing for Accra; we're only using two covariates. And for (b), I
think we can focus on people as it's a lot simpler. For our purposes,
the things in this work that might be useful to us include matching
principles for (a), when we revisit it, and distributions (in particular
joint distributions) for travel. More useful to us would be
distributions for total travel, but I haven't found any in this
literature (e.g. zero-inflated compound Poisson distribution).

In contrast, my main questions are:

-   How to account for multi-mode trips.

-   What to do about zero travel for e.g. older cyclists. One
    possibility is to smooth over all modes and/or demographic groups.

-   In the ITHIM-R programme, are bus and truck drivers participants?

-   How should we include correlations between modes? This could also be
    aided by a smooth model across modes.

-   Whether there is a better method to get distance travelled per
    person than by resampling raw data.

## Method

### Summarise the travel survey

Our first step is to summarise the demographic information and the
travel survey. In Table [5](#population){reference-type="ref"
reference="population"} are the demographic data: the populations of the
city, and the number of people in the survey. In Table
[6](#calculations){reference-type="ref" reference="calculations"} are
the summary statistics from the travel survey which we will use to
generate synthetic travel. From the raw travel data (actually the
adjusted trip set, which is the raw travel plus motorcycle, bus and
truck trips), we calculate the probability to travel, $B_{a,g,m}$, as
the number of people who travelled by that mode divided by the number of
people in the demographic group. It is to these values that we assign
uncertainty through the variable PROPENSITY_TO_TRAVEL.

::: {.center}
::: {#population}
  Age      Gender     Surveyed   Accra population
  -------- -------- ---------- ------------------
  15--49   Male            279          5,847,716
  50--69   Male             56          1,016,476
  15--49   Female          328          6,360,535
  50--69   Female           69          1,110,037

  : Population by demographic group.
:::
:::

::: {.center}
::: {.ThreePartTable}
::: {.TableNotes}
:::

::: {#calculations}
  Age     Gender   Mode              Probability in raw trip set   Probability in baseline trip set
  ------- -------- --------------- ----------------------------- ----------------------------------
  15-49   Male     Bus                                    0.4014                             0.3862
  50-69   Male     Bus                                    0.4286                             0.3636
  15-49   Female   Bus                                    0.3506                             0.3495
  50-69   Female   Bus                                    0.4203                             0.4203
  15-49   Male     Taxi                                   0.0681                             0.0655
  50-69   Male     Taxi                                   0.0357                             0.0303
  15-49   Female   Taxi                                   0.0671                             0.0669
  50-69   Female   Taxi                                   0.1159                             0.1159
  15-49   Male     Walking                                0.6559                              0.631
  50-69   Male     Walking                                0.4643                             0.3939
  15-49   Female   Walking                                0.5427                              0.541
  50-69   Female   Walking                                0.4203                             0.4203
  15-49   Male     Private Car                            0.1254                             0.1207
  50-69   Male     Private Car                            0.1964                             0.1667
  15-49   Female   Private Car                            0.0549                             0.0547
  50-69   Female   Private Car                             0.058                              0.058
  15-49   Male     Bicycle                                0.0179                             0.0172
  50-69   Male     Bicycle                                     0                                  0
  15-49   Female   Bicycle                                 0.003                              0.003
  50-69   Female   Bicycle                                     0                                  0
  15-49   Male     Short Walking                               0                             0.3862
  50-69   Male     Short Walking                               0                             0.3636
  15-49   Female   Short Walking                               0                             0.3495
  50-69   Female   Short Walking                               0                             0.4203
  15-49   Male     Motorcycle                                  0                             0.0621
  50-69   Male     Motorcycle                                  0                             0.1515
  15-49   Female   Motorcycle                                  0                             0.0061
  50-69   Female   Motorcycle                                  0                             0.0145

  : Travel summary table.
:::
:::
:::

### Adjust for trip weight

We define a trip-mode weight per mode, $L_m$, so that a single trip by
mode $m$ in the baseline scenario represents $L_m$ trips in reality. If
in a scenario the mode of the trip is changed ($m'\rightarrow m$), it
retains the weight of its original mode $m'$. We use the trip-mode
weight to (a) sample trips for scenario generation and (b) scale the
probabilities to travel. Note that we don't use it as trip weights are
ordinarily used: to scale trip distance/duration. The result is a scaled
set of probability parameters: $$\hat{B}_{a,g,m} = L_m{B}_{a,g,m}.$$
($L_m$ should actually be the average trip-mode weight for all trips in
the scenario. For baseline, it will be the same. It can/should differ
for scenarios. Need to work out notation.)

### Smooth probabilities

We estimate smoothed probabilities for any mode that has zero or one
probability, using the rest of the probabilities for reference,
replacing $\hat{B}_{a,g,m}$. The reference to other modes is slightly
problematic if e.g. in the scenario all travel increases. It means a 0
in the baseline is not the same as a 0 in the scenario. This could be
fixed with more rigid coding.

### Resample probabilities

We resample the adjusted probability parameters of Table
[6](#calculations){reference-type="ref" reference="calculations"}. We
describe each parameter with a Beta distribution as follows:
$$\begin{aligned}
\label{pitilde}
\tilde{B}_{a,g,m} & \sim \text{Beta}(\alpha_{a,g,m},\beta_{a,g,m}); \\
\alpha_{a,g,m}+\beta_{a,g,m} & = 200; \\
\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}.\end{aligned}$$
Here, $\tilde{B}_{a,g,m}$ is the new, variable parameter that we
resample for each simulation. We sample via a quantile,
$\nu_m\sim\mathcal{U}(0,1)$, which we use to calculate
$\tilde{B}_{a,g,m}$. We set $\alpha_{a,g,m}$ and $\beta_{a,g,m}$ to sum
to 200, making the distributions concentrated, and we constrain their
relative values to enforce the distribution's mean to equal
$\hat{B}_{a,g,m}$ (the right-most column in Table
[6](#calculations){reference-type="ref" reference="calculations"}).

### Inform individual travel

Each individual is assigned their own propensity to travel per mode,
$\delta_{i,m}$, which is a uniformly distributed random variable that is
fixed and not resampled. Whether or not individual $i$ travels by mode
$m$ depends on $\tilde{B}_{a(i),g(i),m}$, which does vary: specifically,
individual $i$ travels by mode $m$ if
$\delta_{i,m}<\tilde{B}_{a(i),g(i),m}$. (To mitigate the effects of
these fixed random variables, a large population is required, e.g. more
than 5000 individuals.)

![Distributions for propensities to travel. The propensity is the
probability that a member of a demographic group will travel by a
particular mode on a particular day. In orange are the raw probabilities
based on the synthetic trip set, here shown for the baseline scenario.
Beta distributions are defined via a pointiness parameter, set to 200,
such that $\alpha+\beta=200$.
](propensity_to_travel_200.pdf){#propensity_to_travel width="70%"}

## Extrapolating to one week of travel {#Extrapolation}

In order to extrapolate to one week of travel, we consider "trips" only
at the stage of scenario generation, and proceed with individual-level
statistics. Then, in our synthetic population, we have "duration
travelled per mode" columns. None of the core ITHIM-R functions will use
the "synthetic trip set".

To get duration per mode per person, we use, for each
demographic-group--mode combination, the probability to travel, and the
density of the total duration travelled given that travel occurred. To
get from one day to one week, consider two cases: one, each individual
follows the same pattern of travel every day. That is, we sample one
day's travel and apply it to seven days. Two, each individual follows
the travel pattern of their demographic group on each day. That is, we
sample seven times from the group's density to make up seven days of
travel.

In implementation, we blend individual and group-level travel through a
"repetitiveness" parameter, $\pi$. Each individual has two travel sets:
one from their simulated daily travel pattern, and one from the shared,
convolved distribution. These travel sets are weighted by $\pi$ and
$1-\pi$, respectively. $\delta_{i,m}$ defines the quantile taken for
both the individual density, which is scaled up to one week, and the
shared distribution, which is the convolution of the raw density seven
times.[^6] The densities are shown in Figure
[17](#raw_densities){reference-type="ref" reference="raw_densities"}.

![Raw distributions for duration travelled per day (left) from the
travel survey. Convolution to one week is shown on the right. We use
""walking"" mode for demographic group "Male",
"15--49".](raw_densities.pdf){#raw_densities width="80%"}

## Simulating a synthetic population {#synthpop}

1.  Choose a population size of individuals ${N}$. Label each individual
    $i=1,...,N$.

2.  Allocate each individual to a demographic group according to the
    proportions in the Accra population (Table
    [5](#population){reference-type="ref" reference="population"}). Then
    each demographic group has number $\tilde{N}_{a,g}$ and
    $\sum_{a,g}\tilde{N}_{a,g}=N$.

3.  Generate one standard uniform variable per person per mode,
    $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual
    $i$'s propensity to travel by mode $m$.

4.  $\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel
    by mode $m$ in scenario $s$. It is the proportion of raw
    person--duration values that are zero. It is parametrised by
    $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation
    [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).

5.  Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw
    person--duration values for demographic group {$a,g$} and mode $m$
    in scenario $s$:
    $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$
    i.e. we isolate all individuals $i$ in group $\{a,g\}$ and sum their
    trip durations by mode $m$ in scenario $s$.

6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}$ be that density corrected to
    have proportion $\tilde{B}_{a,g,m,s}$ zeros.

7.  Define each individual's travel duration as
    $D_{i,m,s}=7\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})$
    (see Figure [18](#distances_beta){reference-type="ref"
    reference="distances_beta"} for resulting distributions).

For sampling, we loop over steps 6--7, where each sample has unique
$\tilde{B}_{a,g,m,s}$ values. In total, the number of uncertain
variables is one per mode.

## Simulating a synthetic population with repetitiveness parameter & uncertain extrapolation {#synthpop2}

1.  Choose a population size of individuals ${N}$. Label each individual
    $i=1,...,N$.

2.  Allocate each individual to a demographic group according to the
    proportions in the Accra population (Table
    [5](#population){reference-type="ref" reference="population"}). Then
    each demographic group has number $\tilde{N}_{a,g}$ and
    $\sum_{a,g}\tilde{N}_{a,g}=N$.

3.  Generate one standard uniform variable per person per mode,
    $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual
    $i$'s propensity to travel by mode $m$.

4.  $\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel
    by mode $m$ in scenario $s$. It is the proportion of raw
    person--duration values that are zero. It is parametrised by
    $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation
    [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).

5.  Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw
    person--duration values for demographic group {$a,g$} and mode $m$
    in scenario $s$:
    $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$
    i.e. we isolate all individuals $i$ in group $\{a,g\}$ and sum their
    trip durations by mode $m$ in scenario $s$.

6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}^7$ be the week convolution of
    $\tilde{\mathcal{F}}_{a,g,m,s}$ (see Figure
    [17](#raw_densities){reference-type="ref"
    reference="raw_densities"}, right).

7.  Define each individual's travel duration as
    $D_{i,m,s}=7\pi\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})+(1-\pi)\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}^7}^{-1}(\delta_{i,m})$
    (see Figure [18](#distances_beta){reference-type="ref"
    reference="distances_beta"} for resulting distributions).

For sampling, we loop over steps 6--8, where each sample has unique
$\tilde{B}_{a,g,m,s}$ and $\pi$ values. In total, the number of
uncertain variables is one per mode plus one.

![Samples of total distances travelled by mode and
scenario.](distances_beta.pdf){#distances_beta width="80%"}

## Results

![YLLs in the scenario minus YLLs in the baseline for Accra, with
motorcycle probability scalar 2 in
Accra.](accra_walk_yll_uncertain_travel.pdf){#results width="80%"}

![EVPPI for Accra's "walking" scenario for all causes of YLLs excluding
"all cause" and "neoplasm", with propensity to motorcycle =
2.](accra_walk_evppi_uncertain_travel.pdf){#evppi_mode_propensities_200
width="70%"}

# Glossary by letter

    $\text{Letter}$ Meaning                           $\text{Letter}$ Meaning               $\text{Letter}$ Meaning
  ----------------- ------------------------------- ----------------- ------------------- ----------------- --------------------------
           $\alpha$                                               $a$ age                               $A$ 
            $\beta$                                               $b$                                   $B$ raw travel probabilities
           $\gamma$ emission uncertainty                          $c$                                   $C$ ventilation constants
           $\delta$ propensity to travel                          $d$ demographic group                 $D$ 
         $\epsilon$ walk-to-bus time                              $e$                                   $E$ injury distance exponent
            $\zeta$ traffic PM2.5 fraction                        $f$                                   $F$ PA DR function
             $\eta$ PM2.5 concentration                           $g$ gender                            $G$ PA DR curves
           $\theta$ PA uncertainty                                $h$ disease                           $H$ AP DR curves
            $\iota$                                               $i$ individual                        $I$ injuries
                 $$                                               $j$ trip                              $J$ 
           $\kappa$ injury regression parameter                   $k$                                   $K$ AP exposure
          $\lambda$ travel MMETs                                  $l$                                   $L$ 
              $\mu$ mode distances                                $m$ mode                              $M$ MMETs per person
              $\nu$                                               $n$                                   $N$ population
              $\xi$ AP DR quantile                                $o$ outcome                           $O$ 
              $\pi$ travel repetitiveness                         $p$                                   $P$ vehicle emissions
                 $$                                               $q$                                   $Q$ person weights
             $\rho$ chronic disease scalar                        $r$                                   $R$ death-to-injury ratio
           $\sigma$ injury reporting rate                         $s$ scenario                          $S$ 
             $\tau$                                               $t$                                   $T$ trips
         $\upsilon$                                               $u$                                   $U$ burden of disease
             $\phi$ PA DR quantile                                $v$                                   $V$ AP per person
                 $$                                               $w$ AP DR parameter                   $W$ relative risks
             $\chi$ day-to-week travel scalar                     $x$ PA DR parameter                   $X$ injury model matrix
             $\psi$ casualty fraction of $\omega$                 $y$ year                              $Y$ individual MMETs
           $\omega$ sum of distance exponents                     $z$ PA dose                           $Z$ speeds
:::

[^1]: Ainsworth compendium 2011 sites.google.com/site/
    compendiumofphysicalactivities for walking and cycling

[^2]: For any person of age lower than 25, we set the relative risk to
    1.

[^3]: The four parameters refer, in numerical order, to
    $\alpha, \beta, \gamma$ and tmrel in @Burnett2014.

[^4]: This is a problem for the examples of Delhi and Bangalore, whose
    trip sets have car travellers under the age at which driving is
    permitted and who therefore should not contribute to striking
    distance.

[^5]: Note that the assymmetry in the offset term for the whw expression
    arises due to the absence of demographic information for striking
    vehicles. Were we to have that, the offset would read
    $\log(\bar{T}_{a,g,m_{\text{cas}},s})+(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s})+\log(\bar{T}_{a,g,m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})$.
    For our case, the latter two terms simplify to
    $\log(\hat{T}_{m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})=E_{m_{\text{str}}}\log(\hat{T}_{m_{\text{str}},s})$.

[^6]: Using the same random variable for both travel sets increases the
    variability of the resulting dataset. Conversely, using different
    random variables would result in more homogenous durations, and
    fewer zeros.

### 2021-09-08 summary_tables.Rmd
Fixing title
Regen markdown files

--- 2021-08-10 summary_tables.Rmd
@@
 
-Case study specific duration tables for baseline and five scenarios (per day in km)
+Case study specific duration tables for baseline and five scenarios (per day in minutes)
 
@@
   colnames(xxx) <- scen_names
-  rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_', '', y))
+  rownames(xxx) <- gsub("scen1_deaths_", "", rownames(xxx))
+  #rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_', '', y))
+  # rownames(xxx) <- c("PA_AllCauses", "PA_AP_IschemicHeartDisease",
+  #                    "PA_Neoplasms",
+  #                    "PA_AP_LungCancer", "AP_COPD", "PA_AP_Stroke",
+  #                    "PA_AP_T2Diabetes", "AP_LowerRespiratoryInfections",
+  #                    "PA_BreastCancer", "PA_ColonCancer", "PA_UterineCancer",
+  #                    "Injuries")
   xxx
@@
   colnames(xxx) <- scen_names
-  rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_','',y))
+  rownames(xxx) <- gsub("scen1_deaths_", "", rownames(xxx))
+  #rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_','',y))
+  # rownames(xxx) <- c("PA_AllCauses", "PA_AP_IschemicHeartDisease",
+  #                    "PA_Neoplasms",
+  #                    "PA_AP_LungCancer", "AP_COPD", "PA_AP_Stroke",
+  #                    "PA_AP_T2Diabetes", "AP_LowerRespiratoryInfections",
+  #                    "PA_BreastCancer", "PA_ColonCancer", "PA_UterineCancer",
+  #                    "Injuries")
   xxx

### 2021-09-08 summary_tables.Rmd
Fixing labels in health results

@@
   colnames(xxx) <- scen_names
-  rownames(xxx) <- gsub("scen1_deaths_", "", rownames(xxx))
+  rownames(xxx) <- gsub("scen1_ylls_", "", rownames(xxx))
   #rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_','',y))

### 2021-09-17 summary_tables.Rmd
Regen including health impacts

@@
 library(pracma)
+library(data.table)
 })
@@
 ```
+
+```{r, echo=FALSE, results='hide', message=FALSE}
+# Export all health results
+health <- list()
+measure <- c("ylls", "deaths")
+pathway <- c("hb", "pathway_hb")
+for (k in measure) { # loop for ylls or deaths
+  for (j in pathway) { # loop for hb or pathway_hb
+    for (i in names(io[-2])) { # loop for cities
+      # Temporal datset
+      temp <- io[[i]]$outcomes[[j]][[k]]
+      # Population by age, and age-sex
+      pop_by_age <- io[[i]]$demographic %>% group_by(age) %>% 
+        summarize(pop_age = sum(population))
+      pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>% 
+        summarize(pop_age_sex = sum(population))
+
+      # Long format and adding new variables
+      health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure", 
+                                      -sex, -age_cat) %>% 
+        mutate(city = i,
+               scenario = case_when(
+                 grepl("scen1", cause) ~ scen_names[1],
+                 grepl("scen2", cause) ~ scen_names[2],
+                 grepl("scen3", cause) ~ scen_names[3],
+                 grepl("scen4", cause) ~ scen_names[4],
+                 grepl("scen5", cause) ~ scen_names[5],
+               ),
+               level1 = case_when(
+                 grepl("all_cause", cause) ~ "L1: All Cause",
+                 grepl("inj", cause) ~ "L1: RTI"
+               ),
+               level2 = case_when(
+                 grepl("total_cancer", cause) ~ "L2: Cancer",
+                 grepl("CVD", cause) ~ "L2: CVD",
+                 grepl("COPD", cause) ~ "L2: Respiratory",
+                 grepl("LRI", cause) ~ "L2: Respiratory",
+                 grepl("T2D", cause) ~ "L2: Other",
+                 grepl("total_dementia", cause) ~ "L2: Other",
+                 grepl("Parkinson", cause) ~ "L2: Other",
+                 grepl("inj", cause) ~ "L2: RTI"
+               ),
+               level3 = case_when(
+                 grepl("IHD", cause) ~ "L3: IHD",
+                 grepl("lung_cancer", cause) ~ "L3: lung_cancer",
+                 grepl("COPD", cause) ~ "L3: COPD",
+                 grepl("stroke", cause) ~ "L3: stroke",
+                 grepl("T2D", cause) ~ "L3: T2D",
+                 grepl("LRI", cause) ~ "L3: LRI",
+                 grepl("breast_cancer", cause) ~ "L3: breast_cancer",
+                 grepl("colon_cancer", cause) ~ "L3: colon_cancer",
+                 grepl("endo_cancer", cause) ~ "L3: endo_cancer",
+                 grepl("liver_cancer", cause) ~ "L3: liver_cancer",
+                 grepl("CVD", cause) ~ "L3: CVD",
+                 grepl("total_dementia", cause) ~ "L3: total_dementia",
+                 grepl("myeloma", cause) ~ "L3: myeloma",
+                 grepl("Parkinson", cause) ~ "L3: Parkinson",
+                 grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer",
+                 grepl("stomach_cancer", cause) ~ "L3: stomach_cancer",
+                 grepl("inj", cause) ~ "L3: RTI"
+               ),
+               dose = case_when(
+                 grepl("_pa_ap_", cause) ~ "PA and AP",
+                 grepl("_pa_", cause) ~ "PA",
+                 grepl("_ap_", cause) ~ "AP",
+                 grepl("_inj", cause) ~ "RTI"
+               )
+        ) %>% # End mutate 
+        left_join(pop_by_age, by = c("age_cat" = "age")) %>% 
+        left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>% 
+        mutate(measure_100k = measure / pop_age * 100000,
+               measeure_100k_sex = measure / pop_age_sex * 100000)
+
+    } # Loop for each city
+  } # Loop for each pathway
+} # Loop for each measure
+
+# Export ylls HB
+write.csv(rbindlist(health$ylls$hb, use.names = T),
+            "results/multi_city/health_impacts/ylls_hb_lng.csv", 
+            row.names = F)
+# Export ylls PATHWAY_HB
+write.csv(rbindlist(health$ylls$pathway_hb, use.names = T),
+            "results/multi_city/health_impacts/ylls_pathway_hb_lng.csv", 
+            row.names = F)
+# Export deaths HB
+write.csv(rbindlist(health$deaths$hb, use.names = T),
+            "results/multi_city/health_impacts/deaths_hb_lng.csv", 
+            row.names = F)
+# Export deaths PATHWAY_HB
+write.csv(rbindlist(health$deaths$pathway_hb, use.names = T),
+            "results/multi_city/health_impacts/deaths_pathway_hb_lng.csv", 
+            row.names = F)
+
+```

### 2021-09-29 summary_tables.Rmd
Regen and removing CVD from level3

--- 2021-09-08 summary_tables.Rmd
@@
 ```
+
+```{r, echo=FALSE, results='hide', message=FALSE}
+# Export all health results
+health <- list()
+measure <- c("ylls", "deaths")
+pathway <- c("hb", "pathway_hb")
+for (k in measure) { # loop for ylls or deaths
+  for (j in pathway) { # loop for hb or pathway_hb
+    for (i in names(io[-2])) { # loop for cities
+      # Temporal datset
+      temp <- io[[i]]$outcomes[[j]][[k]]
+      # Population by age, and age-sex
+      pop_by_age <- io[[i]]$demographic %>% group_by(age) %>% 
+        summarize(pop_age = sum(population))
+      pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>% 
+        summarize(pop_age_sex = sum(population))
+
+      # Long format and adding new variables
+      health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure", 
+                                      -sex, -age_cat) %>% 
+        mutate(city = i,
+               scenario = case_when(
+                 grepl("scen1", cause) ~ scen_names[1],
+                 grepl("scen2", cause) ~ scen_names[2],
+                 grepl("scen3", cause) ~ scen_names[3],
+                 grepl("scen4", cause) ~ scen_names[4],
+                 grepl("scen5", cause) ~ scen_names[5],
+               ),
+               level1 = case_when(
+                 grepl("all_cause", cause) ~ "L1: All Cause",
+                 grepl("inj", cause) ~ "L1: RTI"
+               ),
+               level2 = case_when(
+                 grepl("total_cancer", cause) ~ "L2: Cancer",
+                 grepl("CVD", cause) ~ "L2: CVD",
+                 grepl("COPD", cause) ~ "L2: Respiratory",
+                 grepl("LRI", cause) ~ "L2: Respiratory",
+                 grepl("T2D", cause) ~ "L2: Other",
+                 grepl("total_dementia", cause) ~ "L2: Other",
+                 grepl("Parkinson", cause) ~ "L2: Other",
+                 grepl("inj", cause) ~ "L2: RTI"
+               ),
+               level3 = case_when(
+                 grepl("IHD", cause) ~ "L3: IHD",
+                 grepl("lung_cancer", cause) ~ "L3: lung_cancer",
+                 grepl("COPD", cause) ~ "L3: COPD",
+                 grepl("stroke", cause) ~ "L3: stroke",
+                 grepl("T2D", cause) ~ "L3: T2D",
+                 grepl("LRI", cause) ~ "L3: LRI",
+                 grepl("breast_cancer", cause) ~ "L3: breast_cancer",
+                 grepl("colon_cancer", cause) ~ "L3: colon_cancer",
+                 grepl("endo_cancer", cause) ~ "L3: endo_cancer",
+                 grepl("liver_cancer", cause) ~ "L3: liver_cancer",
+                 grepl("total_dementia", cause) ~ "L3: total_dementia",
+                 grepl("myeloma", cause) ~ "L3: myeloma",
+                 grepl("Parkinson", cause) ~ "L3: Parkinson",
+                 grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer",
+                 grepl("stomach_cancer", cause) ~ "L3: stomach_cancer",
+                 grepl("inj", cause) ~ "L3: RTI"
+               ),
+               dose = case_when(
+                 grepl("_pa_ap_", cause) ~ "PA and AP",
+                 grepl("_pa_", cause) ~ "PA",
+                 grepl("_ap_", cause) ~ "AP",
+                 grepl("_inj", cause) ~ "RTI"
+               )
+        ) %>% # End mutate 
+        left_join(pop_by_age, by = c("age_cat" = "age")) %>% 
+        left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>% 
+        mutate(measure_100k = measure / pop_age * 100000,
+               measeure_100k_sex = measure / pop_age_sex * 100000)
+
+    } # Loop for each city
+  } # Loop for each pathway
+} # Loop for each measure
+
+# Export ylls HB
+write.csv(rbindlist(health$ylls$hb, use.names = T),
+            "results/multi_city/health_impacts/ylls_hb_lng.csv", 
+            row.names = F)
+# Export ylls PATHWAY_HB
+write.csv(rbindlist(health$ylls$pathway_hb, use.names = T),
+            "results/multi_city/health_impacts/ylls_pathway_hb_lng.csv", 
+            row.names = F)
+# Export deaths HB
+write.csv(rbindlist(health$deaths$hb, use.names = T),
+            "results/multi_city/health_impacts/deaths_hb_lng.csv", 
+            row.names = F)
+# Export deaths PATHWAY_HB
+write.csv(rbindlist(health$deaths$pathway_hb, use.names = T),
+            "results/multi_city/health_impacts/deaths_pathway_hb_lng.csv", 
+            row.names = F)
+
+```

### 2021-10-18 latex/output-md.Rmd
Fix layout and typos

@@
+---
+title: "ITHIM"
+author: "Rob Johnson, Ali Abbas"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+
+output:
+  rmdformats::downcute:
+    self_contained: true
+    thumbnails: true
+    lightbox: true
+    gallery: false
+    highlight: tango
+---
+
 # Introduction
@@
 
-The remainder of the document presents a current working version of the
-ITHIM, and is a comprehensive record of all the calculations of which it
-is composed. The equations for the ITHIM are described in Section
-[2](#equations){reference-type="ref" reference="equations"} (tabulated
-in Section [7](#tableequations){reference-type="ref"
-reference="tableequations"}). The method is concretised through the
-example application to Accra in Section
-[3](#example){reference-type="ref" reference="example"} and particular
-columns in Tables [1](#indices){reference-type="ref"
-reference="indices"} and [3](#inputs){reference-type="ref"
-reference="inputs"}, with results presented in Section
-[19](#results){reference-type="ref" reference="results"}.
-
-The ITHIM as described here is implemented in the ITHIM-R R package
-(<https://github.com/ITHIM/ITHIM-R>). Other cities are also included in
-that package (São Paulo, Delhi, Bangalore), and reference will be made
-to those where relevant.
+The remainder of the document presents a current working version of the ITHIM, and is a comprehensive record of all the calculations of which it is composed.
+The equations for the ITHIM are described in Section [2](#equations){reference-type="ref" reference="equations"} (tabulated in Section [7](#tableequations){reference-type="ref" reference="tableequations"}).
+The method is concretised through the example application to Accra in Section [3](#example){reference-type="ref" reference="example"} and particular columns in Tables [1](#indices){reference-type="ref" reference="indices"} and [3](#inputs){reference-type="ref" reference="inputs"}, with results presented in Section [19](#results){reference-type="ref" reference="results"}.
+
+The ITHIM as described here is implemented in the ITHIM-R R package (<https://github.com/ITHIM/ITHIM-R>).
+Other cities are also included in that package (São Paulo, Delhi, Bangalore), and reference will be made to those where relevant.
 
@@
 
-In the ITHIM, we combine many information inputs and compute the health
-burden through three pathways (depicted in Figure
-[1](#workflow){reference-type="ref" reference="workflow"}). Each pathway
-has its own set of definitions and outputs, and we refer to them as
-"modules".
-
-![ITHIM workflow. Inputs to the model have a bold, black outline. The
-graph depicts parent--child relationships, where a child (at the head of
-an arrow) depends upon all its parents (at the source(s) of the
-arrow(s)). To aid visualisation, an arrow connected to the outside of a
-grey box indicates that the arrow connects to each item within the box.
-*AT: active travel. MET: metabolic equivalent task. pp: per person. AP:
-air pollution. PA: physical activity. DR: dose--response relationship.
-GBD: global burden of disease. BOD: (local) burden of
-disease.*](flowchart.pdf){#workflow width="90%"}
-
-To describe the model, we use lower-case letters to denote indices, or
-dimensions, of objects. They take one of a set of possible values,
-detailed in Table [1](#indices){reference-type="ref"
-reference="indices"}. The set of input items are denoted by capital
-letters, and variable parameters are denoted by Greek letters. These are
-presented in Table [3](#inputs){reference-type="ref"
-reference="inputs"}. The equations of the ITHIM are presented below, and
-summarised in Table [4](#calculationstab){reference-type="ref"
-reference="calculationstab"}. These use letter modifiers to show how the
-inputs are modified in the course of output generation.
-
-We calculate the ITHIM by combining the modules' outputs in terms of the
-health burden ($\hat{U}_{a,g,h,o,s}$). Other outputs of the model
-include: MMETs per person ($M_{i,s}$), PM2.5 in each scenario
-($\bar{P}_s$), PM2.5 per person ($\check{W}_{i,s}$), and injuries burden
-($\check{I}_{a,g,m,o,s}$).
-
-To assess uncertainty, we sample from the ITHIM output by sampling
-uncertain parameters multiple times and calculating the ITHIM. We
-specify parametric distributions or sampling strategies for all the
-uncertain parameters. Evaluating the ITHIM with uncertain parameters
-allows assessment of their impact on the outcome (AKA sensitivity
-analysis). We use EVPPI to calculate the expected reduction in
-uncertainty in the outcome were we to learn a parameter perfectly. This
-means we can implement models that are basic in their parametrisation,
-and learn at the end for which parameters it would be worthwhile
-spending dedicated time learning better.
-
-::: {.center}
+In the ITHIM, we combine many information inputs and compute the health burden through three pathways (depicted in Figure [1](#workflow){reference-type="ref" reference="workflow"}).
+Each pathway has its own set of definitions and outputs, and we refer to them as "modules".
+
+!\[ITHIM workflow.
+Inputs to the model have a bold, black outline.
+The graph depicts parent--child relationships, where a child (at the head of an arrow) depends upon all its parents (at the source(s) of the arrow(s)).
+To aid visualisation, an arrow connected to the outside of a grey box indicates that the arrow connects to each item within the box.
+\*AT: active travel.
+MET: metabolic equivalent task.
+pp: per person.
+AP: air pollution.
+PA: physical activity.
+DR: dose--response relationship.
+GBD: global burden of disease.
+BOD:
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("flowchart.pdf")
+```
+
+To describe the model, we use lower-case letters to denote indices, or dimensions, of objects.
+They take one of a set of possible values, detailed in Table [1](#indices){reference-type="ref" reference="indices"}.
+The set of input items are denoted by capital letters, and variable parameters are denoted by Greek letters.
+These are presented in Table [3](#inputs){reference-type="ref" reference="inputs"}.
+The equations of the ITHIM are presented below, and summarised in Table [4](#calculationstab){reference-type="ref" reference="calculationstab"}.
+These use letter modifiers to show how the inputs are modified in the course of output generation.
+
+We calculate the ITHIM by combining the modules' outputs in terms of the health burden ($\hat{U}_{a,g,h,o,s}$).
+Other outputs of the model include: MMETs per person ($M_{i,s}$), PM2.5 in each scenario ($\bar{P}_s$), PM2.5 per person ($\check{W}_{i,s}$), and injuries burden ($\check{I}_{a,g,m,o,s}$).
+
+To assess uncertainty, we sample from the ITHIM output by sampling uncertain parameters multiple times and calculating the ITHIM.
+We specify parametric distributions or sampling strategies for all the uncertain parameters.
+Evaluating the ITHIM with uncertain parameters allows assessment of their impact on the outcome (AKA sensitivity analysis).
+We use EVPPI to calculate the expected reduction in uncertainty in the outcome were we to learn a parameter perfectly.
+This means we can implement models that are basic in their parametrisation, and learn at the end for which parameters it would be worthwhile spending dedicated time learning better.
+
+::: center
 ::: {#indices}
-  Label   Name                   Levels
-  ------- ---------------------- ---------------------
-  $a$     Age group              15--49
-                                 50--69
-  $g$     Gender                 Male
-                                 Female
-  $h$     Disease                
-  $i$     Individual index       
-  $j$     Trip index             
-  $m$     Transport mode         walk
-                                 cycling (cyc.)
-                                 bus
-                                 car
-                                 motorbike (mot.)
-                                 goods vehicles (GV)
-                                 subway (sub.)
-  $o$     Outcome                Death
-                                 YLL
-  $s$     Scenario               
-  $w$     AP RR hyperparameter   $1,2,3,4$
-  $x$     PA RR hyperparameter   $1,2,3$
-  $z$     PA dose                $0,...,300$
-
-  : Indices used in the ITHIM variables. The "Label" is the subscript
-  used to refer to the index in this document. The "Name" is
-  descriptive. The "Levels" are the levels used in the Accra model.
-:::
-:::
-
-::: {.center}
+| Label | Name                 | Levels              |
+|:------|:---------------------|:--------------------|
+| $a$   | Age group            | 15--49              |
+|       |                      | 50--69              |
+| $g$   | Gender               | Male                |
+|       |                      | Female              |
+| $h$   | Disease              |                     |
+| $i$   | Individual index     |                     |
+| $j$   | Trip index           |                     |
+| $m$   | Transport mode       | walk                |
+|       |                      | cycling (cyc.)      |
+|       |                      | bus                 |
+|       |                      | car                 |
+|       |                      | motorbike (mot.)    |
+|       |                      | goods vehicles (GV) |
+|       |                      | subway (sub.)       |
+| $o$   | Outcome              | Death               |
+|       |                      | YLL                 |
+| $s$   | Scenario             |                     |
+| $w$   | AP RR hyperparameter | $1,2,3,4$           |
+| $x$   | PA RR hyperparameter | $1,2,3$             |
+| $z$   | PA dose              | $0,...,300$         |
+
+: Indices used in the ITHIM variables. The "Label" is the subscript used to refer to the index in this document. The "Name" is descriptive. The "Levels" are the levels used in the Accra model.
+:::
+:::
+
+::: center
 ::: {#abbrev}
-  Abbreviation   Meaning
-  -------------- -----------------------------------------------
-  AP             Air pollution
-  CDF            Cumulative distribution function
-  COPD           Chronic obstructive pulmonary disease
-  DR             Dose response
-  EVPPI          Expected value of perfect partial information
-  GAM            Generalised additive model
-  GBD            Global burden of disease
-  IHD            Ischemic heart disease
-  ITHIM          Integrated transport and health impact model
-  LRI            Lower respiratory infection
-  MMET           Marginal metabolic equivalent task
-  PA             Physical activity
-  PAF            Population attributable fraction
-  PIF            Potential impact fraction
-  PM2.5          Particulate matter (diameter $< 2.5 \mu$m)
-  PT             Public transport
-  RR             Relative risk
-  VOI            Value of information
-  YLL            Years of life lost
+| Abbreviation | Meaning                                       |
+|:-------------|:----------------------------------------------|
+| AP           | Air pollution                                 |
+| CDF          | Cumulative distribution function              |
+| COPD         | Chronic obstructive pulmonary disease         |
+| DR           | Dose response                                 |
+| EVPPI        | Expected value of perfect partial information |
+| GAM          | Generalised additive model                    |
+| GBD          | Global burden of disease                      |
+| IHD          | Ischemic heart disease                        |
+| ITHIM        | Integrated transport and health impact model  |
+| LRI          | Lower respiratory infection                   |
+| MMET         | Marginal metabolic equivalent task            |
+| PA           | Physical activity                             |
+| PAF          | Population attributable fraction              |
+| PIF          | Potential impact fraction                     |
+| PM2.5        | Particulate matter (diameter $< 2.5 \mu$m)    |
+| PT           | Public transport                              |
+| RR           | Relative risk                                 |
+| VOI          | Value of information                          |
+| YLL          | Years of life lost                            |
 
@@
 
-::: {.landscape}
-::: {.center}
+::: landscape
+::: center
 ::: {#inputs}
-  ------------------------------------------- ----------------------------------------------------------------- ---------------------------------- ------------------------------- ----------------
-  [\[inputs\]]{#inputs label="inputs"}                                                                          *ITHIM-R reference*                                                
-  Name                                        Description                                                       Value in Accra model                                         Label   Model/ Setting
-  *Global values*                                                                                                                                                                  
-  $U_{a,g,h,o}$                               Background burden of disease                                      Constant                                                  GBD_DATA          Setting
-  $N_{a,g}$                                   Population by age and gender in GBD_DATA                          Constant                                                  GBD_DATA          Setting
-  $\bar{N}_{a,g}$                             Population by age and gender                                      Constant                                               DEMOGRAPHIC          Setting
-  $\rho_h$                                    Chronic disease scalar                                            1 or Lnorm(0,0.18)                          CHRONIC_DISEASE_SCALAR          Setting
-  *Travel values*                                                                                                                                                                  
-  $Z_{m=walk}$                                Walk speed                                                        4.8                                                    MODE_SPEEDS          Setting
-  $Z_{m=cyc.}$                                Cycle speed                                                       14.5                                                   MODE_SPEEDS          Setting
-  $Z_{m=bus}$                                 Bus speed                                                         15                                                     MODE_SPEEDS          Setting
-  $Z_{m=car}$                                 Car speed                                                         21                                                     MODE_SPEEDS          Setting
-  $Z_{m=mot.}$                                Motorbike speed                                                   25                                                     MODE_SPEEDS          Setting
-  $\epsilon$                                  Walk time to bus                                                  5 or Lnorm(log(5),0.18)                              BUS_WALK_TIME          Setting
-  $\mu_{m=HGV}$                               Truck distance relative to car                                    0.21 or Beta(3,10)                              TRUCK_TO_CAR_RATIO          Setting
-  $\mu_{m=bus}$                               Bus driver distance relative to bus passenger distance            1/45 or Beta(20,600)                        BUS_TO_PASSENGER_RATIO          Setting
-  $\mu_{m=mot.}$                              Motorcycle fleet distance relative to total motorcycle distance   37/68 or Beta(37/3+1,31/3+1)             FLEET_TO_MOTORCYCLE_RATIO          Setting
-  $T_j$                                       Trip-level data                                                   Constant                                           SYNTHETIC_TRIPS          Setting
-  *Injury model values*                                                                                                                                                            
-  $I_{a,g,m_{\text{vic}},m_{\text{str}}}$     Fatalities table                                                  Constant                                              INJURY_TABLE          Setting
-  $\sigma$                                    Injury reporting rate                                             1 or Beta(8,3)                               INJURY_REPORTING_RATE          Setting
-  $\omega$                                    Sum of distance exponents                                         1.9 or Lnorm(log(1.9),log(1.03))                  SIN_EXPONENT_SUM            Model
-  $\psi$                                      Casualty fraction of $\omega$                                     0.5 or Beta(20,20)                    CASUALTY_EXPONENT\_ FRACTION            Model
-  *Pollution model values*                                                                                                                                                         
-  $\eta$                                      Background PM2.5 concentration                                    50 or Lnorm(log(50),0.18)                             PM_CONC_BASE          Setting
-  $\zeta$                                     Fraction of measured PM2.5 concentration due to road transport    0.225 or Beta(5,20)                                 PM_TRANS_SHARE          Setting
-  $\gamma_{m}$                                Vehicle emission inventory                                        Constant or Dirichlet                           EMISSION_INVENTORY          Setting
-  $P$                                         Vehicle emission confidence                                       1                                    EMISSION_INVENTORY_CONFIDENCE          Setting
-  *Pollution & health model values*                                                                                                                                                
-  $C_1$                                       Base-level inhalation rate                                        1                                     BASE_LEVEL_INHALATION \_RATE            Model
-  $C_2$                                       Exposure ratio window closed                                      0.5                                       CLOSED_WINDOW_PM\_ RATIO            Model
-  $C_3$                                       Proportion of travel with closed windows                          0.5                                            CLOSED_WINDOW_RATIO          Setting
-  $C_4$                                       Parameter for on-road PM2.5 level                                 3.216                                               ROAD_RATIO_MAX            Model
-  $C_5$                                       Parameter for on-road PM2.5 level                                 0.379                                             ROAD_RATIO_SLOPE            Model
-  $C_6$                                       Exposure ratio for underground                                    0.8                                                SUBWAY_PM_RATIO          Setting
-  $H_{a,h,w}$                                 AP dose--response--curve parameters                               Constant                                                     DR_AP            Model
-  $\xi_{h,w}$                                 Quantiles for AP RR functions                                     0.5 or Unif(0,1)                       AP_DOSE_RESPONSE\_ QUANTILE            Model
-  *Physical activity model values*                                                                                                                                                 
-  $\chi$                                      Day-to-week travel scalar                                         7 or 7$\times$Beta(20,3)               DAY_TO_WEEK_TRAVEL\_ SCALAR            Model
-  $Y_i$                                       Individual non-transport MMETs                                    Constant                                PA_SET\$work_ltpa_marg_met          Setting
-  $\theta$                                    Background PA scalar                                              1 or Lnorm(0,0.18)                            BACKGROUND_PA_SCALAR          Setting
-  $\tilde{Y}$                                 Confidence in PA survey                                           1                                         BACKGROUND_PA_CONFIDENCE          Setting
-  $\tilde{\theta}$                            Background PA zeros quantile                                      0.5 or Unif(0,1)                               BACKGROUND_PA_ZEROS          Setting
-  $\lambda_{m=\text{cyc.}}$                   Cycling MET surplus                                               4.63 or Lnorm(log(4.63),1)                            MMET_CYCLING            Model
-  $\lambda_{m=\text{walk}}$                   Walking MET surplus                                               2.53 or Lnorm(log(2.53),1)                            MMET_WALKING            Model
-  $\lambda_{m\not\in\text{walk,cyc.}}$        In-vehicle MET surplus                                            0                                                                             Model
-  *Physical activity & health model values*                                                                                                                                        
-  $G_{h,x,z}$                                 Look-up table for relative risk truncated normal distribution     Constant                                                                      Model
-  $\phi_{h}$                                  Quantiles for PA RR functions                                     0.5 or Unif(0,1)                       PA_DOSE_RESPONSE\_ QUANTILE            Model
-  ------------------------------------------- ----------------------------------------------------------------- ---------------------------------- ------------------------------- ----------------
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| \[\[inputs\]\]{label= "inputs"}           |                                                                 | *ITHIM-R reference*                |                                |                |
++:==========================================+=================================================================+:===================================+================================+================+
+| Name                                      | Description                                                     | Value in Accra model               | Label                          | Model/ Setting |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| *Global values*                           |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $U_ { a,g,h,o}$                           | Background burden of disease                                    | Constant                           | GBD_DATA                       | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $N_{a,g}$                                 | Population by age and gender in GBD_DATA                        | Constant                           | GBD_DATA                       | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\bar { N}_{a,g}$                         | Population by age and gender                                    | Constant                           | DEMOGRAPHIC                    | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\rho_h$                                  | Chronic disease scalar                                          | 1 or Lnorm(0,0.18)                 | CHRONIC_DISEASE_SCALAR         | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| *Travel values*                           |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $Z _ {m=walk}$                            | Walk speed                                                      | 4.8                                | MODE_SPEEDS                    | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $Z _ {m=cyc.}$                            | Cycle speed                                                     | 14.5                               | MODE_SPEEDS                    | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $Z _ {m=bus}$                             | Bus speed                                                       | 15                                 | MODE_SPEEDS                    | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $Z _ {m=car}$                             | Car speed                                                       | 21                                 | MODE_SPEEDS                    | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $Z _ {m=mot.}$                            | Motorbike speed                                                 | 25                                 | MODE_SPEEDS                    | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| \$\\epsilon \$                            | Walk time to bus                                                | 5 or Lnorm(log (5),0.18)           | BUS\_ WALK_TIME                | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\mu_{m=HGV}$ \|                          | Truck distance relative to car \|                               | 0.21 or Beta(3,10)                 | TRUCK_TO\_ CAR_RATIO \| S      | etting         |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\mu_{m=bus}$ \|                          | Bus driver distance relative to bus passenger distance \|       | 1/45 or Bet a(20,600) \|           | BUS\_ TO_PASSEN GER_RATIO      | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\mu_ {m=mot.}$                           | Motorcycle fleet distance relative to total motorcycle distance | 37/68 or Beta(37/3+ 1,31/3+1) \|   | FLEET_TO_MOTORCYCLE_RATIO      | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $T_j$                                     | Trip-level data                                                 | Constant                           | SYNTHETIC_TRIPS                | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| *Injury model values*                     |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $I _ {a,g,m_{\                            | Fatalities table                                                | Constant                           | INJURY_TABLE                   | Setting        |
+| text{vi c} },m_{\text{str}}}$             |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\sigma$                                  | Injury reporting rate                                           | 1 or Beta(8,3)                     | INJURY_REPORTING_RATE          | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\omega$                                  | Sum of distance exponents                                       | 1.9 or Lnorm(l og(1.9),l og(1.03)) | SIN_EXPONENT_SUM               | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\psi$                                    | Casualty fraction of $\omega$                                   | 0.5 or Beta(20,20) \|              | CASUALTY_EXPONENT_FRACTION     | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| -   Pollution model                       |                                                                 |                                    |                                |                |
+|                                           |                                                                 |                                    |                                |                |
+|     values\*                              |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\eta$                                    | Background PM2.5 concentration                                  | 50 or Lnorm(log( 50),0.18) \|      | PM_CONC_BASE \|                | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\zeta$                                   | Fraction of measured PM2.5 concentration due to road transport  | 0.225 or Beta(5,20) \|             | PM_TRANS_SHARE \|              | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\                                        | Vehicle emission inventory                                      | Constant or Dirichlet              | EMISSION_INVENTORY             | Setting        |
+| g amma_{m}$                               |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $P$                                       | Vehicle emission confidence                                     | 1                                  | EM ISSION_INVENTORY_CONFIDENCE | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| -   Pollution & health model              |                                                                 |                                    |                                |                |
+|                                           |                                                                 |                                    |                                |                |
+|     values\*                              |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $C_1$                                     | Base-level inhalation rate                                      | 1                                  | BAS E_LEVEL_I NHALATION \_RATE | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $C_2$                                     | Exposure ratio window closed                                    | 0.5                                | CLOSED_WI NDOW_PM\_ RATIO      | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $C_3$                                     | Proportion of travel with closed windows                        | 0.5                                | C LOSED_WIN DOW_RATIO          | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $C_4$                                     | Parameter for on-road PM2.5 level                               | 3.216                              | ROAD\_ RATIO_MAX               | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $C_5$                                     | Parameter for on-road PM2.5 level                               | 0.379                              | ROAD_RA TIO_SLOPE              | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $C_6$                                     | Exposure ratio for underground                                  | 0.8                                | SUBWAY \_PM_RATIO              | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| \$ H \_{a,h,w}\$                          | AP dose- -res ponse--curve parameters                           | Constant                           | DR_AP                          | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\                                        | Quantiles for AP RR functions                                   | 0.5 or Unif(0,1)                   | AP_DOSE_R ESPONSE\_ QUANTILE   | Model          |
+| xi_{h,w}$                                 |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| *Physical activity model values*          |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\chi$                                    | Day-to-week travel scalar                                       | 7 or 7 $\times$Beta(20,3)          | DAY_TO_WEEK \_TRAVEL_SCALAR    | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $Y_i$                                     | Individual n on-transport MMETs                                 | Constant                           | PA_SET\$ work_ltpa \_marg_met  | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\theta$                                  | Background PA scalar                                            | 1 or Lnorm(0,0.18)                 | BACKGROUND\_ PA_SCALAR         | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\                                        | Confidence in PA survey                                         | 1                                  | BACKGROUND_PA_C ONFIDENCE      | Setting        |
+| tilde{Y}$                                 |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\tilde {\theta}$                         | Background PA zeros quantile                                    | 0.5 or Unif(0,1)                   | BACKGROUND \_PA_ZEROS          | Setting        |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\lambda_{m=\text{cyc.}}$                 | Cycling MET surplus                                             | 4.63 or Lnorm(log (4.63),1)        | MMET_CYCLING                   | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\lambda_{m=\text{walk}}$                 | Walking MET surplus                                             | 2.53 or Lnorm(log (2.53),1)        | MMET_WALKING                   | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| $\lambda_{m\not\in\text{wal               | In-vehicle MET surplus                                          | 0                                  |                                | Model          |
+| k,cyc.}}$                                 |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| *Physical activity & health model values* |                                                                 |                                    |                                |                |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| \$\\G\_{h,x,z}\$                          | Look-up table for relative risk truncated normal distribution   | Constant                           |                                | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
+| \$\phi\_{h}\$                             | Quantiles for PA RR functions                                   | 0.5 or Unif(0,1)                   | PA_DOSE_RESPONSE_QUANTILE      | Model          |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
 
@@
 
-Input data for the ITHIM are listed as Roman capital letters in Table
-[3](#inputs){reference-type="ref" reference="inputs"}. They include
-(vehicle) mode speeds, injury/fatality records, disease dose--response
-relationships, measures of air pollution, surveys of travel and physical
-activity, population numbers by demographic group, MMETs associated with
-active travel, and current burden of disease (see Figure
-[1](#workflow){reference-type="ref" reference="workflow"}).
+Input data for the ITHIM are listed as Roman capital letters in Table [3](#inputs){reference-type="ref" reference="inputs"}.
+They include (vehicle) mode speeds, injury/fatality records, disease dose--response relationships, measures of air pollution, surveys of travel and physical activity, population numbers by demographic group, MMETs associated with active travel, and current burden of disease (see Figure [1](#workflow){reference-type="ref" reference="workflow"}).
 
@@
 
-We create a "synthetic population" by matching individuals surveyed for
-travel to the individuals surveyed for physical activity, based on age
-and gender. Then each person in the synthetic population has a set of
-trips taken and an amount of non-travel physical activity. This
-population is assumed to be representative of the population of the city
-or setting under study.
+We create a "synthetic population" by matching individuals surveyed for travel to the individuals surveyed for physical activity, based on age and gender.
+Then each person in the synthetic population has a set of trips taken and an amount of non-travel physical activity.
+This population is assumed to be representative of the population of the city or setting under study.
 
@@
 
-we could create a synthetic population as described in Appendix
-[8.3](#synthpop){reference-type="ref" reference="synthpop"}, which
-allows uncertainty in the propensity to travel. This method operates
-with person-level, rather than trip-level, information.
+we could create a synthetic population as described in Appendix [8.3](#synthpop){reference-type="ref" reference="synthpop"}, which allows uncertainty in the propensity to travel.
+This method operates with person-level, rather than trip-level, information.
 
@@
 
-From the synthetic population we create "scenarios" of different
-pictures of travel by changing certain trips (e.g. swapping the mode of
-travel for particular trips and/or particular people). For the purposes
-of illustration, we choose a simple scenario for Accra in which every
-person gains one one-kilometre walk.
+From the synthetic population we create "scenarios" of different pictures of travel by changing certain trips (e.g.
+swapping the mode of travel for particular trips and/or particular people).
+For the purposes of illustration, we choose a simple scenario for Accra in which every person gains one one-kilometre walk.
 
@@
 
-For the ITHIM, we process the travel data into distance data of various
-formats. First, we augment PT trips with a walk component: we take the
-set $\{T_j:m(j)=\text{bus}\}$, which has $J$ entries. We add $J$
-journeys with mode "walk" and duration $\epsilon$ to the set $T$. (This
-step would not be necessary for a travel survey which already includes
-all stages for each trip.)
-
-Then we summarise the total distances and durations. The distance set is
-labelled $\hat{T}_{m,s}$, representing the total distance travelled per
-mode per scenario, and the duration set is constructed analogously. We
-make an assumption that the distance for bus drivers scales linearly
-with bus passengers, where the distance in the baseline is defined based
-on the bus driver distance relative to car.
-
-Distance and duration summaries include all person travel in the
-synthetic population and, in addition, some trips that are attributed to
-no-one in the synthetic population. These trips are used to represent
-"fleet" travel, which is uncaptured in travel surveys, and unchanged in
-scenarios. It's important to include these trips because for modes such
-as motorcycle, where a considerable component of travel comes from
-fleet, attributing all the harms associated with the mode to only those
-surveyed dramatically overestimates the contribution of the individuals.
-In scenarios, then, we would be predicting emissions and injury rates
-that are far too high per person.
+For the ITHIM, we process the travel data into distance data of various formats.
+First, we augment PT trips with a walk component: we take the set $\{T_j:m(j)=\text{bus}\}$, which has $J$ entries.
+We add $J$ journeys with mode "walk" and duration $\epsilon$ to the set $T$.
+(This step would not be necessary for a travel survey which already includes all stages for each trip.)
+
+Then we summarise the total distances and durations.
+The distance set is labelled $\hat{T}_{m,s}$, representing the total distance travelled per mode per scenario, and the duration set is constructed analogously.
+We make an assumption that the distance for bus drivers scales linearly with bus passengers, where the distance in the baseline is defined based on the bus driver distance relative to car.
+
+Distance and duration summaries include all person travel in the synthetic population and, in addition, some trips that are attributed to no-one in the synthetic population.
+These trips are used to represent "fleet" travel, which is uncaptured in travel surveys, and unchanged in scenarios.
+It's important to include these trips because for modes such as motorcycle, where a considerable component of travel comes from fleet, attributing all the harms associated with the mode to only those surveyed dramatically overestimates the contribution of the individuals.
+In scenarios, then, we would be predicting emissions and injury rates that are far too high per person.
 
@@
 
-We use the total distance by mode ($\hat{T}_{m,s}$) to calculate the
-total PM2.5 in the scenario ($\bar{P}_s$). First we calculate the
-emission inventory in terms of fractions. If the confidence $P$ is 1
-then we use $\dot{P}_{m}=\gamma_m$ directly. If $P<1$, we sample the
-emission inventory fractions as follows:
-$$\tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)),$$
-$$\dot{P}_m=\tilde{\gamma}_m$$ where we choose some mapping function
-such as $f(P,\gamma_m)=\frac{\gamma_m}{\sum_m\gamma_m}10^{5P}$ (Figure
-[5](#emission_dist){reference-type="ref" reference="emission_dist"}).
-
-Then we multiply vehicle distance by vehicle emission factor (where the
-emission factor is the emission inventory (fraction) divided by distance
-at baseline).
+We use the total distance by mode ($\hat{T}_{m,s}$) to calculate the total PM2.5 in the scenario ($\bar{P}_s$).
+First we calculate the emission inventory in terms of fractions.
+If the confidence $P$ is 1 then we use $\dot{P}_{m}=\gamma_m$ directly.
+If $P<1$, we sample the emission inventory fractions as follows: $$\tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)),$$ $$\dot{P}_m=\tilde{\gamma}_m$$ where we choose some mapping function such as $f(P,\gamma_m)=\frac{\gamma_m}{\sum_m\gamma_m}10^{5P}$ (Figure [5](#emission_dist){reference-type="ref" reference="emission_dist"}).
+
+Then we multiply vehicle distance by vehicle emission factor (where the emission factor is the emission inventory (fraction) divided by distance at baseline).
 $$\tilde{P}_{m,s}=\dot{P}_m\frac{\hat{T}_{m,s}}{\hat{T}_{m,s=\text{baseline}}}.$$
 
-For the baseline, $\tilde{P}_{m,s}$ is equal to the emission inventory
-fraction. For each scenario, it is scaled. We take the sum over modes to
-get a scalar for emissions in scenarios.
-$$\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}.$$ Then the background PM2.5 in
-each scenario is the sum of the transport component and the
-non-transport component: $$\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta).$$
+For the baseline, $\tilde{P}_{m,s}$ is equal to the emission inventory fraction.
+For each scenario, it is scaled.
+We take the sum over modes to get a scalar for emissions in scenarios.
+$$\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}.$$ Then the background PM2.5 in each scenario is the sum of the transport component and the non-transport component: $$\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta).$$
 
@@
 
-Dose--response relationships are based on background concentrations in
-cities, and not on personal exposures. In our person-level model, we
-expect that individuals will have different exposures based on their
-different travel behaviours. While we can neither estimate individual
-exposures nor map them via any dose--response relationship to a relative
-risk, we aim to capture interpersonal differences via assumptions about
-the mechanistic relationships between behaviours and exposures, and
-anchor those to the background air pollution in their locality, for
-which dose--response estimates of relative risk exist.
-
-Therefore, when we write "individual exposure", it should not be read as
-an aetiologically relevant exposure, but rather it should be read as a
-unit- and scale-agnostic metric that maps an individual to a particular
-dose with reference to city-level measurements of air pollution. New
-research is required to elicit dose--response relationships between
-individual health exposures and health outcomes \[citation from JPO?\].
-Our approximations act as assumptions for now and placeholders until
-more detailed data are available. \[Perhaps we should use words other
-than "individual exposure". How about "individual PM2.5 reference
-values"?\]
-
-Individual exposures to PM2.5 are calculated using the background PM2.5
-($\bar{P}_s$) and the trip sets. There are three major components to
-daily exposure: one, a person's total inhalation off road; two, a
-person's inhalation on road in a vehicle, and, three, a person's
-inhalation on road while undertaking active transport. Each category has
-an amount of background PM2.5 and a ventilation rate which together
-inform overall exposure.
-
-The ratio of exposure off road to that on road is a function of total
-PM2.5, defined as $$K_s=C_4-C_5\log(\bar{P_s})$$ as in @Goel2015. This
-defines the exposure of a person in an open vehicle (i.e. pedestrian,
-cyclist or motorist):
-$$\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$$ and it is used to
-calculate in-vehicle exposure, assuming an exposure of $C_2$ with the
-window closed, and a proportion $C_3$ of vehicles having closed windows:
-$$\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3).$$
-The exposure in a subway is constant and not dependent on the road ratio
-$K_s$: $$\tilde{K}_{m=\text{sub.},s}=C_6.$$
-
-Ventilation rates are calculated for each mode, assuming a base-level
-inhalation rate of $C_1$: $$V_m=C_1+\frac{1}{2}\lambda_m,$$ where
-$\lambda_m$ is the MMETs for mode $m$.[^1] Then the air inhaled during
-travel per person is
-$$\tilde{V}_{i,m,s} = \smashoperator{\sum_{\substack{j:i(j)=i,\\m(j)=m,\\s(j)=s}}}T_j\cdot V_{m(j)}/Z_{m(j)}$$
-and the rate of inhalation during travel is
-$$\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}.$$ The
-air inhaled when not travelling is
-$$\hat{V}_{i,s} = C_1\left(24-\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}T_j/Z_{m(j)}\right).$$
-Together, the PM2.5 exposure is calculated as the total PM2.5 inhaled
-per hour as follows:
-$$\check{W}_{i,s} = \frac{\bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})}{24}.$$
+Dose--response relationships are based on background concentrations in cities, and not on personal exposures.
+In our person-level model, we expect that individuals will have different exposures based on their different travel behaviours.
+While we can neither estimate individual exposures nor map them via any dose--response relationship to a relative risk, we aim to capture interpersonal differences via assumptions about the mechanistic relationships between behaviours and exposures, and anchor those to the background air pollution in their locality, for which dose--response estimates of relative risk exist.
+
+Therefore, when we write "individual exposure", it should not be read as an aetiologically relevant exposure, but rather it should be read as a unit- and scale-agnostic metric that maps an individual to a particular dose with reference to city-level measurements of air pollution.
+New research is required to elicit dose--response relationships between individual health exposures and health outcomes \[citation from JPO?\].
+Our approximations act as assumptions for now and placeholders until more detailed data are available.
+\[Perhaps we should use words other than "individual exposure".
+How about "individual PM2.5 reference values"?\]
+
+Individual exposures to PM2.5 are calculated using the background PM2.5 ($\bar{P}_s$) and the trip sets.
+There are three major components to daily exposure: one, a person's total inhalation off road; two, a person's inhalation on road in a vehicle, and, three, a person's inhalation on road while undertaking active transport.
+Each category has an amount of background PM2.5 and a ventilation rate which together inform overall exposure.
+
+The ratio of exposure off road to that on road is a function of total PM2.5, defined as $$K_s=C_4-C_5\log(\bar{P_s})$$ as in @Goel2015.
+This defines the exposure of a person in an open vehicle (i.e.
+pedestrian, cyclist or motorist): $$\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$$ and it is used to calculate in-vehicle exposure, assuming an exposure of $C_2$ with the window closed, and a proportion $C_3$ of vehicles having closed windows: $$\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3).$$ The exposure in a subway is constant and not dependent on the road ratio $K_s$: $$\tilde{K}_{m=\text{sub.},s}=C_6.$$
+
+Ventilation rates are calculated for each mode, assuming a base-level inhalation rate of $C_1$: $$V_m=C_1+\frac{1}{2}\lambda_m,$$ where $\lambda_m$ is the MMETs for mode $m$.[^1] Then the air inhaled during travel per person is $$\tilde{V}_{i,m,s} = \smashoperator{\sum_{\substack{j:i(j)=i,\\m(j)=m,\\s(j)=s}}}T_j\cdot V_{m(j)}/Z_{m(j)}$$ and the rate of inhalation during travel is $$\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}.$$ The air inhaled when not travelling is $$\hat{V}_{i,s} = C_1\left(24-\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}T_j/Z_{m(j)}\right).$$ Together, the PM2.5 exposure is calculated as the total PM2.5 inhaled per hour as follows: $$\check{W}_{i,s} = \frac{\bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})}{24}.$$
+
+[^1]: Ainsworth compendium 2011 sites.google.com/site/ compendiumofphysicalactivities for walking and cycling
 
@@
 
-We use each person's exposure to PM2.5 ($\check{W}_{i,s}$) to calculate
-their relative risk of five diseases (IHD, lung cancer, COPD, stroke,
-LRI), using curves parametrised by four disease-specific variables
-[@Burnett2014]. Of the five diseases, two (IHD and stroke) have
-parameters specific to age groups starting at age 25.[^2] The other
-three (lung cancer, LRI and COPD) have one set of parameters for all
-ages.
-
-The curves are in the form of samples of the set of four parameters. We
-model the densities of these samples (using a quantile for parameter 3,
-kernel density estimation for parameter 2, and GAMs for parameters 1 and
-4)[^3] in order to draw either the median or random samples via their
-quantiles ($\xi_{h,w}$) as follows:
+We use each person's exposure to PM2.5 ($\check{W}_{i,s}$) to calculate their relative risk of five diseases (IHD, lung cancer, COPD, stroke, LRI), using curves parametrised by four disease-specific variables [@Burnett2014].
+Of the five diseases, two (IHD and stroke) have parameters specific to age groups starting at age 25.[^2] The other three (lung cancer, LRI and COPD) have one set of parameters for all ages.
+
+[^2]: For any person of age lower than 25, we set the relative risk to 1.
+
+The curves are in the form of samples of the set of four parameters.
+We model the densities of these samples (using a quantile for parameter 3, kernel density estimation for parameter 2, and GAMs for parameters 1 and 4)[^3] in order to draw either the median or random samples via their quantiles ($\xi_{h,w}$) as follows:
+
+[^3]: The four parameters refer, in numerical order, to $\alpha, \beta, \gamma$ and tmrel in @Burnett2014.
 
@@
 
-From these parameters, the relative risk of mortality is defined
-$$\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=1}\times
-\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=2}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=4}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}.$$
-For $h\not\in\{\text{IHD, lung cancer, COPD, stroke, LRI}\}$,
-$\check{H}_{h,i,s} \equiv 1$.
+From these parameters, the relative risk of mortality is defined $$\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=1}\times
+\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=2}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=4}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}.$$ For $h\not\in\{\text{IHD, lung cancer, COPD, stroke, LRI}\}$, $\check{H}_{h,i,s} \equiv 1$.
 
@@
 
-Using trip sets and the synthetic population, we calculate total MMETs
-per person ($M_{i,s}$) as the sum of walking MMETs per day and cycling
-MMETs per day, which are scaled up to a week via the scalar $\chi$, and
-work/leisure MMETs per week as follows:
-$$M_{i,s}=\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}
+Using trip sets and the synthetic population, we calculate total MMETs per person ($M_{i,s}$) as the sum of walking MMETs per day and cycling MMETs per day, which are scaled up to a week via the scalar $\chi$, and work/leisure MMETs per week as follows: $$M_{i,s}=\smashoperator{\sum_{\substack{j:i(j)=i,\\s(j)=s}}}
 \chi\lambda_{m(j)}\frac{T_j}{Z_{m(j)}}+\theta Y_i.$$
 
-Here, $\lambda_m$ is the MMETs for mode $m$, $T_j$ the distance of trip
-$j$, $Z_m$ the speed of mode $m$, $\theta$ the scalar for background PA,
-and $\check{Y}_i$ the amount of work and leisure (background) PA of
-person $i$.
-
-We calculate $\check{Y}_i$ according to the confidence in the PA survey,
-$\tilde{Y}$. First, we calculate the raw probability that a person in
-demographic group completes no non-travel PA:
-$$\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}.$$ Then, we set the
-probability to use, $\bar{Y}_{a,g}$, as $\dot{Y}_{a,g}$ if our
-confidence $\tilde{Y}$ is 1. If $\tilde{Y}<1$, we map the confidence to
-parametrise a Beta distribution somehow, e.g. (Figure
-[4](#prob_zero_PA){reference-type="ref" reference="prob_zero_PA"})
-$$\begin{aligned}
+Here, $\lambda_m$ is the MMETs for mode $m$, $T_j$ the distance of trip $j$, $Z_m$ the speed of mode $m$, $\theta$ the scalar for background PA, and $\check{Y}_i$ the amount of work and leisure (background) PA of person $i$.
+
+We calculate $\check{Y}_i$ according to the confidence in the PA survey, $\tilde{Y}$.
+First, we calculate the raw probability that a person in demographic group completes no non-travel PA: $$\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}.$$ Then, we set the probability to use, $\bar{Y}_{a,g}$, as $\dot{Y}_{a,g}$ if our confidence $\tilde{Y}$ is 1.
+If $\tilde{Y}<1$, we map the confidence to parametrise a Beta distribution somehow, e.g.
+(Figure [4](#prob_zero_PA){reference-type="ref" reference="prob_zero_PA"}) $$\begin{aligned}
 \hat{Y} &= 500^{\tilde{Y}+0.2},\\
@@
 
-Finally, for each person in the population, we sample non-travel MMETs
-as zero with probability $\bar{Y}_{a,g}$ and from the raw non-zero
-density of their demographic group with probability $1-\bar{Y}_{a,g}$.
+Finally, for each person in the population, we sample non-travel MMETs as zero with probability $\bar{Y}_{a,g}$ and from the raw non-zero density of their demographic group with probability $1-\bar{Y}_{a,g}$.
 
@@
 
-We use each person's MMETs per week (${M}_{i,s}$) to calculate their
-relative risk of six diseases, using curves found from meta analysis.
-Each disease (except type 2 diabetes) has a threshold beyond which there
-is no further change in relative risk. (Total cancer, breast cancer,
-colon & rectum cancer, endometrial cancer, & coronary heart disease: 35;
-lung cancer: 10; stroke: 32; all cause: 16.08.) For all individual
-cancers, we estimate the burden in terms of . For all other diseases, we
-use mortality.
-
-The interpolation of the dose--response curve can be described as
-follows:
-$$\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$$
-where we interpolate the mean, the lower bound, and the upper bound.
-These define the normal from which we sample, truncated at 0:
-$$F_S(v_x) = \mathcal{N}\left.\left(v_{x=2},\frac{v_{x=3}-v_{x=1}}{1.96}\right)\right|_{F_S(v_x)>0}.$$
-
-Then the relative risk for each person for each disease is calculated
-for the quantile $\phi_h$ as
-$$W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right).$$
-See Figure [2](#cancer){reference-type="ref" reference="cancer"} for an
-example. For $h\not\in$ {diabetes, total cancer, breast cancer, colon &
-rectum cancer, endometrial cancer, IHD, lung cancer, stroke, all cause},
-$W_{h,i,s} \equiv 1$.
-
-![Example of the dose--response workflow for PA and total cancer. Top
-left: results of the metaanalysis: mean, lower bound and upper bound of
-the relationship between PA dose and relative risk of disease. Top
-right: examples of 500 samples from the normal distribution defined by
-the mean of the top-left panel and the standard deviation defined by
-(upper bound minus lower bound) divided by 1.96 and truncated at 0.
-Bottom left: relative risks of individuals with PA doses 5, 25 and 45
-are found by mapping their doses onto the median curve. This is used in
-the "constant" ITHIM use case. Bottom right: relative risks of
-individuals with PA doses 5, 25 and 45 for three different samples from
-the distribution shown in the top-right panel. For each sample, each
-individual is mapped onto the response curve it defines.
-](cancer_PA_dose_response.pdf){#cancer width="60%"}
+We use each person's MMETs per week (${M}_{i,s}$) to calculate their relative risk of six diseases, using curves found from meta analysis.
+Each disease (except type 2 diabetes) has a threshold beyond which there is no further change in relative risk.
+(Total cancer, breast cancer, colon & rectum cancer, endometrial cancer, & coronary heart disease: 35; lung cancer: 10; stroke: 32; all cause: 16.08.) For all individual cancers, we estimate the burden in terms of .
+For all other diseases, we use mortality.
+
+The interpolation of the dose--response curve can be described as follows: $$\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$$ where we interpolate the mean, the lower bound, and the upper bound.
+These define the normal from which we sample, truncated at 0: $$F_S(v_x) = \mathcal{N}\left.\left(v_{x=2},\frac{v_{x=3}-v_{x=1}}{1.96}\right)\right|_{F_S(v_x)>0}.$$
+
+Then the relative risk for each person for each disease is calculated for the quantile $\phi_h$ as $$W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right).$$ See Figure [2](#cancer){reference-type="ref" reference="cancer"} for an example.
+For $h\not\in$ {diabetes, total cancer, breast cancer, colon & rectum cancer, endometrial cancer, IHD, lung cancer, stroke, all cause}, $W_{h,i,s} \equiv 1$.
+
+\[Example of the dose--response workflow for PA and total cancer.
+Top left: results of the metaanalysis: mean, lower bound and upper bound of the relationship between PA dose and relative risk of disease.
+Top right: examples of 500 samples from the normal distribution defined by the mean of the top-left panel and the standard deviation defined by (upper bound minus lower bound) divided by 1.96 and truncated at 0.
+Bottom left: relative risks of individuals with PA doses 5, 25 and 45 are found by mapping their doses onto the median curve.
+This is used in the "constant" ITHIM use case.
+Bottom right: relative risks of individuals with PA doses 5, 25 and 45 for three different samples from the distribution shown in the top-right panel.
+For each sample, each individual is mapped onto the response curve it defines.\]
+
+```{r out.height = "800px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("cancer_PA_dose_response.pdf")
+```
 
@@
 
-For the injury module, we require distance per travel mode per
-demographic group per scenario. It requires matching of mode names in
-the travel survey to mode names in the injury records, for example
-"taxi", "shared auto", "shared taxi" and "car" combine to form "car".
-Some work is required to separate drivers from passengers; we currently
-assume all travellers are drivers, with the exception of "bus".[^4]
+For the injury module, we require distance per travel mode per demographic group per scenario.
+It requires matching of mode names in the travel survey to mode names in the injury records, for example "taxi", "shared auto", "shared taxi" and "car" combine to form "car".
+Some work is required to separate drivers from passengers; we currently assume all travellers are drivers, with the exception of "bus".[^4]
+
+[^4]: This is a problem for the examples of Delhi and Bangalore, whose trip sets have car travellers under the age at which driving is permitted and who therefore should not contribute to striking distance.
 
@@
 
-We model injuries via regression by predicting the number of fatalities
-of each demographic group on each mode, which we calculate as a sum over
-all the ways in which they might have been injured (i.e., all the modes
-with which they might have collided).
-
-The predictive covariates include the distances travelled by the parties
-and their demographic details. These requirements lead to a natural
-separation of the (training) dataset into two groups: a set for which we
-have distance data for the other party, and a set for which we do not
-have distance data for the other party. The former equates to a "who hit
-whom" ("whw") matrix (albeit in a higher dimension), and will account
-for changes in injuries resulting of a change in strike-mode travel. The
-latter corresponds to causes of injury that will not change across
-scenarios, including "no other vehicle" and modes of transport that we
-do not consider to change, which might include trucks and buses if they
-are not somehow explicitly included in the trip set. We label this group
-"noov": no or other vehicle.
-
-We model the number of injuries as a Poisson-distributed variable with
-an offset depending on the distance(s) and the reporting rate. We write
-the model as follows: $$\begin{aligned}
+We model injuries via regression by predicting the number of fatalities of each demographic group on each mode, which we calculate as a sum over all the ways in which they might have been injured (i.e., all the modes with which they might have collided).
+
+The predictive covariates include the distances travelled by the parties and their demographic details.
+These requirements lead to a natural separation of the (training) dataset into two groups: a set for which we have distance data for the other party, and a set for which we do not have distance data for the other party.
+The former equates to a "who hit whom" ("whw") matrix (albeit in a higher dimension), and will account for changes in injuries resulting of a change in strike-mode travel.
+The latter corresponds to causes of injury that will not change across scenarios, including "no other vehicle" and modes of transport that we do not consider to change, which might include trucks and buses if they are not somehow explicitly included in the trip set.
+We label this group "noov": no or other vehicle.
+
+We model the number of injuries as a Poisson-distributed variable with an offset depending on the distance(s) and the reporting rate. We write the model as follows: $$\begin{aligned}
 \tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s,y} \sim &
@@
  % +(CAS_EXPONENT-1)*log(cas_distance_sum)
- +(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s}).\end{aligned}$$
-We choose this form of equation to enable (a) linearity in injuries with
-respect to total travel combined across modes and (b) linearity in
-injuries as subdivided by travellers of each mode.[^5] Note that this
-methodology is under development. See
-<https://github.com/robj411/safety_in_numbers_power_law>.
-
-In general, one makes the best model that one can with the data
-available. We use, for Accra, the covariates
-`casualty mode*strike mode, casualty age,` and `casualty gender`, to
-form the model matrix $X$. We also have another covariate, `year`, but
-our travel data are for one year only, so we model instead many
-observations of the same quantities.
-
-We fit the coefficients $\kappa$ of the model using data for ten years,
-assuming the same travel each year, which corresponds to the scenario
-$s=\text{baseline}$. To incorporate the injury reporting rate
-($\sigma$), we set it as an offset, whose value is 1 in fitting the
-model, and as specified in making predictions.
+ +(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s}).\end{aligned}$$ We choose this form of equation to enable (a) linearity in injuries with respect to total travel combined across modes and (b) linearity in injuries as subdivided by travellers of each mode.[^5] Note that this methodology is under development.
+See <https://github.com/robj411/safety_in_numbers_power_law>.
+
+[^5]: Note that the assymmetry in the offset term for the whw expression arises due to the absence of demographic information for striking vehicles.
+Were we to have that, the offset would read $\log(\bar{T}_{a,g,m_{\text{cas}},s})+(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s})+\log(\bar{T}_{a,g,m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})$.
+For our case, the latter two terms simplify to $\log(\hat{T}_{m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})=E_{m_{\text{str}}}\log(\hat{T}_{m_{\text{str}},s})$.
+
+In general, one makes the best model that one can with the data available.
+We use, for Accra, the covariates `casualty mode*strike mode, casualty age,` and `casualty gender`, to form the model matrix $X$.
+We also have another covariate, `year`, but our travel data are for one year only, so we model instead many observations of the same quantities.
+
+We fit the coefficients $\kappa$ of the model using data for ten years, assuming the same travel each year, which corresponds to the scenario $s=\text{baseline}$.
+To incorporate the injury reporting rate ($\sigma$), we set it as an offset, whose value is 1 in fitting the model, and as specified in making predictions.
 
@@
 
-We predict the number of injuries in each scenario based on the training
-model built from the baseline scenario. For Accra, we predict for the
-year 2016, using scenario-specific travel data. The number of fatalities
-is taken as the sum of fatalities over the strike modes:
-$$\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}.$$
-We extrapolate injury deaths to injury YLLs via the ratio in the GBD:
-$$R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}},$$
-$$\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g}.$$
+We predict the number of injuries in each scenario based on the training model built from the baseline scenario.
+For Accra, we predict for the year 2016, using scenario-specific travel data.
+The number of fatalities is taken as the sum of fatalities over the strike modes: $$\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{cas}},m_{\text{str}},s}.$$ We extrapolate injury deaths to injury YLLs via the ratio in the GBD: $$R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}},$$ $$\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g}.$$
 
@@
 
-We combine the relative risks of disease as a function of AP and disease
-as a function of PA through multiplication:
-$$\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s}.$$ Not all diseases
-have a dose--response relationship for both AP and PA. Just stroke, lung
-cancer, and IHD have both. For the other diseases, only one of
-$W_{h,i,s}$ and $\check{H}_{h,i,s}$ is different from one.
+We combine the relative risks of disease as a function of AP and disease as a function of PA through multiplication: $$\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s}.$$ Not all diseases have a dose--response relationship for both AP and PA.
+Just stroke, lung cancer, and IHD have both.
+For the other diseases, only one of $W_{h,i,s}$ and $\check{H}_{h,i,s}$ is different from one.
 
@@
 
-We calculate the total health burden relative to the reference scenario
-(which, for our Accra example, is the baseline) using the injury and
-health outputs combined with GBD data, via population attributable
-fractions (PAF). The PAF is defined:
-$$\hat{W}_{a,g,h,s}=\smashoperator{\sum_{\substack{i:a(i)=a,\\g(i)=g}}}\tilde{W}_{h,i,s}.$$
-
-We then calculate the PAF relative to the reference scenario ("ref") as:
-$$\bar{W}_{a,g,h,s}= \frac{\hat{W}_{a,g,h,s=\text{ref}}-\hat{W}_{a,g,h,s}}{\hat{W}_{a,g,h,s=\text{ref}}}.$$
-
-We estimate the background burden of disease using GBD data and scaling
-based on the ratio of populations between country and city. For country
-populations $N_{a,g}$, city populations $\bar{N}_{a,g}$ and country
-burden $U_{a,g,h,o}$, we estimate city burden as
-$$\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}.$$
-
-If we are scaling the background burden of non-communicable diseases
-("Neoplasms", "Ischemic heart disease", "Tracheal, bronchus, and lung
-cancer", "Breast cancer", "Colon and rectum cancer", "Uterine cancer"),
-we do so here. (Otherwise set $\rho\equiv 1$.)
-$$\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}.$$
-
-We combine the burden with the PAF:
-$$\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}.$$
+We calculate the total health burden relative to the reference scenario (which, for our Accra example, is the baseline) using the injury and health outputs combined with GBD data, via population attributable fractions (PAF).
+The PAF is defined: $$\hat{W}_{a,g,h,s}=\smashoperator{\sum_{\substack{i:a(i)=a,\\g(i)=g}}}\tilde{W}_{h,i,s}.$$
+
+We then calculate the PAF relative to the reference scenario ("ref") as: $$\bar{W}_{a,g,h,s}= \frac{\hat{W}_{a,g,h,s=\text{ref}}-\hat{W}_{a,g,h,s}}{\hat{W}_{a,g,h,s=\text{ref}}}.$$
+
+We estimate the background burden of disease using GBD data and scaling based on the ratio of populations between country and city.
+For country populations $N_{a,g}$, city populations $\bar{N}_{a,g}$ and country burden $U_{a,g,h,o}$, we estimate city burden as $$\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}.$$
+
+If we are scaling the background burden of non-communicable diseases ("Neoplasms", "Ischemic heart disease", "Tracheal, bronchus, and lung cancer", "Breast cancer", "Colon and rectum cancer", "Uterine cancer"), we do so here.
+(Otherwise set $\rho\equiv 1$.) $$\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}.$$
+
+We combine the burden with the PAF: $$\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}.$$
 
@@
 
-And, finally, for the injuries, we sum over modes to compute the burden,
-$$\bar{U}_{a,g,o,s}=\sum_m\check{I}_{a,g,m,o,s},$$ and subtract from the
-values for the reference scenario:
-$$\hat{U}_{a,g,h=\text{road injury},o,s}=\bar{U}_{a,g,o,s=\text{ref}}-\bar{U}_{a,g,o,s}.$$
+And, finally, for the injuries, we sum over modes to compute the burden, $$\bar{U}_{a,g,o,s}=\sum_m\check{I}_{a,g,m,o,s},$$ and subtract from the values for the reference scenario: $$\hat{U}_{a,g,h=\text{road injury},o,s}=\bar{U}_{a,g,o,s=\text{ref}}-\bar{U}_{a,g,o,s}.$$
 
@@
 
-Here we show the ITHIM as evaluated for the city of Accra, with a simple
-scenario of every person in the synthetic population gaining a single, 1
-km walk. We define XX uncertain univariate parameters (Section
-[3.1](#uni){reference-type="ref" reference="uni"}), two confidences, and
-two uncertain parameter sets (Section [3.3](#dr){reference-type="ref"
-reference="dr"}).
+Here we show the ITHIM as evaluated for the city of Accra, with a simple scenario of every person in the synthetic population gaining a single, 1 km walk.
+We define XX uncertain univariate parameters (Section [3.1](#uni){reference-type="ref" reference="uni"}), two confidences, and two uncertain parameter sets (Section [3.3](#dr){reference-type="ref" reference="dr"}).
 
@@
 
-The XX uncertain univariate parameters are shown in Figure
-[3](#distributions){reference-type="ref" reference="distributions"}.
-Most of the parameters are self explanatory. Cycling and walking MMETs
-are the number of MMETs per hour when undertaking cycling and walking,
-and determine also the ventilation rates. Motorcycle distance is the
-total distance travelled by motorcycles relative to the total distance
-travelled by cars in the baseline scenario. Non-travel PA, injury
-reporting rate and NCD burden all act as scalars for the corresponding
-datasets. Note that the non-travel PA scalar does not affect the
-$\approx$`<!-- -->`{=html}40% of the population whose non-travel PA is
-0.
-
-![Distributions of univariate uncertain parameters (1024
-samples).](uncertain_parameters.pdf){#distributions width="95%"}
+The XX uncertain univariate parameters are shown in Figure [3](#distributions){reference-type="ref" reference="distributions"}.
+Most of the parameters are self explanatory.
+Cycling and walking MMETs are the number of MMETs per hour when undertaking cycling and walking, and determine also the ventilation rates.
+Motorcycle distance is the total distance travelled by motorcycles relative to the total distance travelled by cars in the baseline scenario.
+Non-travel PA, injury reporting rate and NCD burden all act as scalars for the corresponding datasets.
+Note that the non-travel PA scalar does not affect the $\approx$<!-- -->40% of the population whose non-travel PA is 0.
+
+```{r out.height = "460px", out.width='800px', fig.align='center', fg.cap = 'Distributions of univariate uncertain parameters (1024 samples).', echo=F}
+knitr::include_graphics("uncertain_parameters.pdf")
+```
+
+(){#distributions width="95%"}
 
@@
 
-For physical activity propensity and emission inventories, we use
-"confidences", values between 0 and 1 that represent how confident we
-are about the data source. We map these values to Beta (Figure
-[4](#prob_zero_PA){reference-type="ref" reference="prob_zero_PA"}) and
-Dirichlet (Figure [5](#emission_dist){reference-type="ref"
-reference="emission_dist"}) distributions, respectively. The
-Beta-distributed propensity towards non-travel physical activity is
-described in Section [2.3.1](#beta){reference-type="ref"
-reference="beta"}. The Dirichlet-distributed emissions inventory is
-described in Section [2.2.1](#dir){reference-type="ref"
-reference="dir"}.
-
-![Distributions of a Beta-distributed propensity to non-travel PA with
-varying confidence values in the raw input data. The sampled value gives
-the fraction of the population that engages in non-travel PA.
-](prob_zero_PA.pdf){#prob_zero_PA width="50%"}
-
-![Distributions of Dirichlet-distributed emission inventories for four
-modes and varying confidence values. Raw input values were 4, 4, 20, and
-60. ](emission_dist.pdf){#emission_dist width="65%"}
+For physical activity propensity and emission inventories, we use "confidences", values between 0 and 1 that represent how confident we are about the data source.
+We map these values to Beta (Figure [4](#prob_zero_PA){reference-type="ref" reference="prob_zero_PA"}) and Dirichlet (Figure [5](#emission_dist){reference-type="ref" reference="emission_dist"}) distributions, respectively.
+The Beta-distributed propensity towards non-travel physical activity is described in Section [2.3.1](#beta){reference-type="ref" reference="beta"}.
+The Dirichlet-distributed emissions inventory is described in Section [2.2.1](#dir){reference-type="ref" reference="dir"}.
+
+\[Distributions of a Beta-distributed propensity to non-travel PA with varying confidence values in the raw input data.
+The sampled value gives the fraction of the population that engages in non-travel PA.\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("prob_zero_PA.pdf")
+```
+
+(){#prob_zero_PA width="50%"}
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+"Distributions of Dirichlet-distributed emission inventories for four modes and varying confidence values. Raw input values were 4, 4, 20, and 60."
+
+knitr::include_graphics("emission_dist.pdf")
+```
+
+(){#emission_dist width="65%"}
 
@@
 
-For the dose--response relationships between physical activity (PA) and
-disease and air pollution (AP) and disease, we assume that there is
-uncertainty, but no variability, in the relationship. This means that we
-sample a relationship from the distribution of relationships, and apply
-that relationship to all individuals precisely. This means that, given
-fixed doses, responses between individuals will be perfectly correlated.
-
-We achieve this by use of the probability integral transform: we sample
-a random variable uniformly distributed on the space (0,1) and map it,
-via a cumulative distribution function, to the distribution describing
-the dose--response relationship. See Figure
-[6](#cdf){reference-type="ref" reference="cdf"} for an illustration.
-
-![We use the same uniform samples to generate random variables from
-different distributions. It means that the samples are perfectly
-correlated.](DR_transform.pdf){#cdf width="80%"}
+For the dose--response relationships between physical activity (PA) and disease and air pollution (AP) and disease, we assume that there is uncertainty, but no variability, in the relationship.
+This means that we sample a relationship from the distribution of relationships, and apply that relationship to all individuals precisely.
+This means that, given fixed doses, responses between individuals will be perfectly correlated.
+
+We achieve this by use of the probability integral transform: we sample a random variable uniformly distributed on the space (0,1) and map it, via a cumulative distribution function, to the distribution describing the dose--response relationship.
+See Figure [6](#cdf){reference-type="ref" reference="cdf"} for an illustration.
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F, fig.show='hold', fig.cap="We use the same uniform samples to generate random variables from different distributions.
+It means that the samples are perfectly correlated."}
+knitr::include_graphics("DR_transform.pdf")
+```
+
+(){#cdf width="80%"}
 
@@
 
-Each disease's PA dose--response relationship is defined by a truncated
-normal. For each dose, there is a mean value, an upper bound, and a
-lower bound. For each person's dose, we get the response by mapping the
-uniform random variable onto the truncated normal defined by the mean
-and bounds for that dosage.
+Each disease's PA dose--response relationship is defined by a truncated normal.
+For each dose, there is a mean value, an upper bound, and a lower bound.
+For each person's dose, we get the response by mapping the uniform random variable onto the truncated normal defined by the mean and bounds for that dosage.
 
@@
 
-For the AP relationship, there are four parameters per disease. We
-sample the first from an empirical distribution using the probability
-integral transform. We sample the second via the same method,
-conditioned on the value of the first, constructing their joint density
-with e.g. kde2d. The third parameter is sampled conditioned on the first
-and second, constructing their joint density using a GAM. The final
-parameter is sampled conditioned on the first, second, and third,
-constructing their joint density using a GAM.
-
-As before, there is perfect correlation between individuals, i.e. if
-person A's dose is greater than person B's, then person A's response is
-strictly greater than person B's response.
-
-The empirical distributions come from @Burnett2014. There are four
-parameters per disease: IHD, lung cancer, COPD, stroke, and LRI. In
-addition, for stroke and IHD, there is a set of four parameters for each
-age group from 25 to 95 in five-year increments. In addition to our
-assumption that there is perfect correlation between individuals for
-diseases, we assume perfect correlation between ages for diseases. I.e.,
-our four quantiles per disease will be applied to all age groups.
-
-# Results
+For the AP relationship, there are four parameters per disease.
+We sample the first from an empirical distribution using the probability integral transform.
+We sample the second via the same method, conditioned on the value of the first, constructing their joint density with e.g.
+kde2d.
+The third parameter is sampled conditioned on the first and second, constructing their joint density using a GAM.
+The final parameter is sampled conditioned on the first, second, and third, constructing their joint density using a GAM.
+
+As before, there is perfect correlation between individuals, i.e. if person A's dose is greater than person B's, then person A's response is strictly greater than person B's response.
+
+The empirical distributions come from @Burnett2014.
+There are four parameters per disease: IHD, lung cancer, COPD, stroke, and LRI.
+In addition, for stroke and IHD, there is a set of four parameters for each age group from 25 to 95 in five-year increments.
+In addition to our assumption that there is perfect correlation between individuals for diseases, we assume perfect correlation between ages for diseases.
+I.e., our four quantiles per disease will be applied to all age groups.
+
+# Results {#results}
 
@@
 
-![YLLs in the scenario minus YLLs in the baseline for
-Accra.](accra_walk_yll.pdf){#results width="80%"}
+***YLLs in the scenario minus YLLs in the baseline for Accra***
+
+```{r out.height = "600px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("accra_walk_yll.pdf")
+```
 
@@
 
-We calculate value of information following @Jackson2017. The "value of
-information" we implement is the expected value of perfect partial
-information (EVPPI). It is the expected amount by which variation in the
-outcome diminishes should a parameter, or set of parameters, be known
-perfectly. With reference to information theory, it is the mutual
-information between the parameter, or parameter set, and the outcome,
-where the outcome is described only by its second moment (its variance).
-
-The objective of learning the value of information formulated in this
-way is to identify the parameter or parameters that most drive
-variability (variance) in the outcome. The results guide future
-research, identifying which area or areas offer the greatest benefit in
-terms of prediction accuracy given further research. For example, for
-our model of Accra, learning better the dose--response relationship
-between physical activity and stroke (incidence) would likely increase
-precision of our estimate of total years of life lost (YLL) more so than
-learning any other parameter (Figure [8](#evppi){reference-type="ref"
-reference="evppi"}).
-
-![EVPPI for Accra's "walking" scenario for all causes of YLLs excluding
-"all cause" and "neoplasm".](accra_walk_evppi.pdf){#evppi width="70%"}
-
-::: {.appendix}
+We calculate value of information following @Jackson2017.
+The "value of information" we implement is the expected value of perfect partial information (EVPPI).
+It is the expected amount by which variation in the outcome diminishes should a parameter, or set of parameters, be known perfectly.
+With reference to information theory, it is the mutual information between the parameter, or parameter set, and the outcome, where the outcome is described only by its second moment (its variance).
+
+The objective of learning the value of information formulated in this way is to identify the parameter or parameters that most drive variability (variance) in the outcome.
+The results guide future research, identifying which area or areas offer the greatest benefit in terms of prediction accuracy given further research.
+For example, for our model of Accra, learning better the dose--response relationship between physical activity and stroke (incidence) would likely increase precision of our estimate of total years of life lost (YLL) more so than learning any other parameter (Figure [8](#evppi){reference-type="ref" reference="evppi"}).
+
+\[EVPPI for Accra's "walking" scenario for all causes of YLLs excluding "all cause" and "neoplasm".\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', fig.cap="EVPPI for Accra",fig.path="accra_walk_evppi.pdf", echo=F}
+#knitr::include_graphics("accra_walk_evppi.pdf")
+```
+
 # ITHIM with parameters
 
-![ITHIM-R workflow, with variable parameters in purple. Fixed inputs to
-the model have a bold, black outline, which is solid for inputs we
-consider "global" and dashed for inputs we consider "local". There are
-four global inputs, which will be embedded in ITHIM-R. There are seven
-local inputs, which users should provide. The graph depicts
-parent--child relationships, where a child (at the head of an arrow)
-depends upon all its parents (at the source(s) of the arrow(s)). To aid
-visualisation, an arrow connected to the outside of a grey box indicates
-that the arrow connects to each item within the box. *AT: active travel.
-MET: metabolic equivalent task. pp: per person. AP: air pollution. PA:
-physical activity. DR: dose--response relationship. GBD: global burden
-of disease.* NB: $\rho$ only impacts the NCD burden of GBD, not all of
-it. Missing parameter: motorcycle distance. $\lambda$ is a 2D parameter.
-$\xi$ is 4D.](ithim_with_parameters.pdf){#ithim_with_parameters
-width="\\textwidth"}
+::: appendix
+\[ITHIM-R workflow, with variable parameters in purple.
+Fixed inputs to the model have a bold, black outline, which is solid for inputs we consider "global" and dashed for inputs we consider "local".
+There are four global inputs, which will be embedded in ITHIM-R.
+There are seven local inputs, which users should provide.
+The graph depicts parent--child relationships, where a child (at the head of an arrow) depends upon all its parents (at the source(s) of the arrow(s)).
+To aid visualisation, an arrow connected to the outside of a grey box indicates that the arrow connects to each item within the box.
+*AT: active travel.
+MET: metabolic equivalent task.
+pp: per person.
+AP: air pollution.
+PA: physical activity.
+DR: dose--response relationship.
+GBD: global burden of disease.* NB: $\rho$ only impacts the NCD burden of GBD, not all of it.
+Missing parameter: motorcycle distance.
+$\lambda$ is a 2D parameter.
+$\xi$ is 4D.\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("ithim_with_parameters.pdf")
+```
+
+(){#ithim_with_parameters width="\\textwidth"}
 
@@
 
-![1000 samples from PA dose--response
-curves](PA_dose_response.pdf){#PADR width="90%"}
-
-  ------------------------------------------------------------------------------------ ---------------------------------------------------------------------------------------
-   ![100 samples from AP dose--response curves](cvd_ihdDR-sim.pdf){#APDR width="40%"}   ![100 samples from AP dose--response curves](cvd_strokeDR-sim.pdf){#APDR width="40%"}
-  ------------------------------------------------------------------------------------ ---------------------------------------------------------------------------------------
-
-  ------------------------------------------------------------------------------------- -------------------------------------------------------------------------------- --------------------------------------------------------------------------------------
-   ![100 samples from AP dose--response curves](neo_lungDR-sim.pdf){#APDR width="30%"}   ![100 samples from AP dose--response curves](lriDR-sim.pdf){#APDR width="30%"}   ![100 samples from AP dose--response curves](resp_copdDR-sim.pdf){#APDR width="30%"}
-  ------------------------------------------------------------------------------------- -------------------------------------------------------------------------------- --------------------------------------------------------------------------------------
-
-::: {.landscape}
+\[1000 samples from PA dose--response curves\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("PA_dose_response.pdf")
+```
+
+(){#PADR width="90%"}
+
++:----------------------------------------------------------------------------------:+:---------------------------------------------------------------------------------:+
+| ![100 samples from AP dose--response curves](cvd_ihdDR-sim.pdf){#APDR width="40%"} | \[100 samples from AP dose--response curves\] (cvd_strokeDR-sim.pdf){width="40%"} |
++------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
+
++:-----------------------------------------------------------------:+:------------------------------------------------------------:+:------------------------------------------------------------------:+
+| ![100 samples from AP dose--response curv es](neo_lungDR-sim.pdf) | ![100 samples from AP dose--response c urves](lriDR-sim.pdf) | ![100 samples from AP dose--response curve s](resp_copdDR-sim.pdf) |
++-------------------------------------------------------------------+--------------------------------------------------------------+--------------------------------------------------------------------+
+
+::: landscape
 # Tabulated ITHIM equations {#tableequations}
 
-::: {.center}
-::: {.ThreePartTable}
-::: {.TableNotes}
-CDF: cumulative distribution function. We use generative distributions
-to generate random numbers so that samples are correlated. For
-$t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$
-with CDF CDF$_F$. (I'd like to find an improved notation for this.)
+::: center
+::: ThreePartTable
+::: TableNotes
+CDF: cumulative distribution function.
+We use generative distributions to generate random numbers so that samples are correlated.
+For $t=\text{CDF}_{F}^{-1}(q)$, $t$ is the $q$-th quantile of function $F$ with CDF CDF$_F$.
+(I'd like to find an improved notation for this.)
 :::
@@
 ::: {#calculationstab}
-  Name                                                                                                                                                                                    Description                                                                                      Label
-  --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- ------------------------------------------------------------------------------------------------ -------------------------------------------
-  *Scenario generation*
-  \...
-  *Travel calculations*
-  $\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk}\}\, \forall\, T_j:m(j)=\text{bus}$                                                                                           New walk trips to bus                                                                            walk_to_bus_and_combine_scen
-  $\bar{T}_{a,g,m,s}=\sum_{j:a(j)=a,g(j)=g,m(j)=m,s(j)=s}Q_{i(j)}T_j$                                                                                                                     (Weighted) total travel per age, gender, mode and scenario
-  $\check{T}_{a,g,m,s}=\frac{\bar{T}_{a,g,m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}$                                                                                                             Relative travel per mode and scenario by age and gender                                          inj_distances\[\[1\]\]
-  $\hat{T}_{m,s}=\sum_{a,g}\bar{T}_{a,g,m,s}$                                                                                                                                             Total travel per mode and scenario                                                               dist
-  $\tilde{T}_{m,s}=\frac{\hat{T}_{m,s}}{\hat{T}_{m,s=\text{baseline}}}$                                                                                                                   Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1   inj_distances\[\[2\]\]
-  *Injury calculations*
-  $E_{m_{\text{cas}}}=\psi\omega$                                                                                                                                                         Casualty exponent                                                                                CAS_EXPONENT
-  $E_{m_{\text{str}}}=\omega-E_{m_{\text{cas}}}$                                                                                                                                          Striker exponent                                                                                 STR_EXPONENT
-  $\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}=\mathcal{S}\left(\sigma\cdot {I}_{a,g,m_{\text{vic}},m_{\text{str}}},\bar{T}_{a,g,m,s},E_{m_{\text{cas}}},E_{m_{\text{str}}}\right)$   Fatalities in scenarios by strike mode                                                           injuries
-  $\check{I}_{a,g,m,o=\text{death},s}=\sum_{m_{\text{str}}}\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}$                                                                               Fatalities in scenarios                                                                          victim_deaths
-  $R_{a,g}=U_{a,g,h=\text{road injuries},o=\text{YLL}}/U_{a,g,h=\text{road injuries},o=\text{death}}$                                                                                     Road YLL-to-death ratio                                                                          GBD_INJ_YLL
-  $\check{I}_{a,g,m,o=\text{YLL},s}=\check{I}_{a,g,m,o=\text{death},s}\cdot R_{a,g}$                                                                                                      YLLs scaled from fatalities                                                                      deaths_yll_injuries
-  *Pollution calculations*
-  $\dot{P}_{m}=\left\{\begin{array}{lr}\gamma_m & P=1 \\ \tilde{\gamma}_m, \tilde{\gamma}_m \sim \text{Dir}(f(P,\gamma_m)) & P<1 \end{array}\right.$                                      Emissions by mode                                                                                trans_emissions
-  $\tilde{P}_{m,s}=\dot{P}_m\hat{T}_{m,s}/\hat{T}_{m,s=\text{baseline}}$                                                                                                                  Emissions by mode in scenario relative to baseline                                               trans_emissions
-  $\hat{P}_{s}=\sum_{m} \tilde{P}_{m,s}$                                                                                                                                                  Emissions scalar for scenarios                                                                   baseline_sum
-  $\bar{P}_s=\eta(\zeta\hat{P}_s + 1-\zeta)$                                                                                                                                              Background PM2.5 concentration in scenarios                                                      scenario_pm
-  $K_s=C_4-C_5\log(\bar{P_s})$                                                                                                                                                            On-road PM2.5 exposure ratio                                                                     on_road_off_road_ratio
-  $\tilde{K}_{m\in\{\text{walk,cyc.,mot.}\},s}=K_s$                                                                                                                                       On-road PM2.5 exposure ratio for open vehicle                                                    on_road_off_road_ratio
-  $\tilde{K}_{m\in\{\text{sub.}\},s}=C_6$                                                                                                                                                 Subway PM2.5 exposure ratio                                                                      subway_ratio
-  $\tilde{K}_{m\not\in\{\text{walk,cyc.,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3)$                                                                                                                In-vehicle PM2.5 exposure ratio                                                                  in_vehicle_ratio
-  $V_m=C_1+\frac{1}{2}\lambda_m$                                                                                                                                                          Ventilation rates                                                                                vent_rates
-  $\tilde{V}_{i,m,s} = \sum_{j:i(j)=i,m(j)=m,s(j)=s}T_j\cdot V_{m(j)}/Z_{m(j)}$                                                                                                           In-travel air inhaled
-  $\bar{V}_{i,s} = \sum_{m}\tilde{V}_{i,m,s}\cdot \tilde{K}_{m,s}$                                                                                                                        In-travel PM2.5 inhalation rate
-  $\hat{V}_{i,s} = C_1\left(24-\sum_{j:i(j)=i,s(j)=s}T_j/Z_{m(j)}\right)$                                                                                                                 Non-travel air inhaled
-  $\check{W}_{i,s} = \bar{P}_s(\bar{V}_{i,s}+\hat{V}_{i,s})/24                                                                                                                            Total PM2.5 inhaled per hour
-  %/(\sum_m\tilde{V}_{i,m,s}+\hat{V}_{i,s})$
-  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                                                              AP dose--response--curve parameter                                                               DR_AP_LIST
-  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                                          AP dose--response--curve parameter                                                               DR_AP_LIST
-  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                                      AP dose--response--curve parameter                                                               DR_AP_LIST
-  $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a,h,w=3}|\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3},\tilde{H}_{a,h,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                  AP dose--response--curve parameter                                                               DR_AP_LIST
-  $\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=3}\times$
-  $\left\{1 - \exp\left(-\tilde{H}_{a=a(i),h,w=3}  (\check{W}_{i,s} - \tilde{H}_{a=a(i),h,w=3}) ^ {\tilde{H}_{a=a(i),h,w=3} }\right)\right\}$                                             RR of disease given pollution                                                                    RR_AP_calculations
-  *Physical activity calculations*
-  $\dot{Y}_{a,g}=\text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}$                                                                                                                                     Raw probability non-travel MMETs = 0 for demographic group $a,g$                                 raw_zero
-  $\bar{Y}_{a,g}=\left\{\begin{array}{lr} \dot{Y}_{a,g} & \tilde{Y}=1 \\ \text{CDF}_{\text{Beta}(f(\tilde{Y},\dot{Y}_{a,g}))}^{-1}(\tilde{\theta}) & \tilde{Y}<1 \end{array}\right.$      Probability non-travel MMETs = 0 for demographic group $a,g$                                     zeros
-  $\check{Y}_i = \left\{\begin{array}{lr} 0 & \bar{Y}_{a(i),g(i)} \\ \text{Sample}(\{Y_{i'}\}|_{a(i')=a(i),g(i')=g(i),Y_{i'}>0}) & 1-\bar{Y}_{a(i),g(i)} \end{array}\right.$              Non-travel MMETs                                                                                 SYNTHETIC_POPULATION\$ work_ltpa_marg_met
-  $M_{i,s}=\sum_{j:i(j)=i,s(j)=s}(\chi T_j\cdot \lambda_{m(j)}/Z_{m(j)})+\theta \check{Y}_i$                                                                                              Total MMETs per person                                                                           mmets
-  $\tilde{G}_{h,x}(m) = {G}_{h,x,z=\lfloor{m}\rfloor}+({G}_{h,x,z=\lceil{m}\rceil}-{G}_{h,x,z=\lfloor{m}\rfloor})\frac{m-\lfloor{m}\rfloor}{\lceil{m}\rceil-\lfloor{m}\rfloor}$           Interpolated dose--response parameters                                                           rr, lb, ub
-  $F_S(s_x) = \mathcal{N}(s_{x=2},(s_{x=3}-s_{x=1})/1.96)|_{0<F_S(s_x)}$                                                                                                                  Truncated normal function
-  $W_{h,i,s}=\text{CDF}_{F_S\left(\tilde{G}_{h,x}(M_{i,s})\right)}^{-1}\left(\phi_{h}\right)^{\dagger}$                                                                                   Relative risk of disease given PA                                                                RR_PA_calculations
-  *Burden-of-disease calculations*
-  $\tilde{W}_{h,i,s}=W_{h,i,s}\cdot\check{H}_{h,i,s}$                                                                                                                                     Combined PA and AP relative risks                                                                RR_PA_AP_calculations
-  $\hat{W}_{a,g,h,s}= \sum_{i:a(i)=a,g(i)=g}\tilde{W}_{h,i,s}$                                                                                                                            Population-attributable fractions                                                                pif_temp
-  $\bar{W}_{a,g,h,s}= (\hat{W}_{a,g,h,s=\text{scen1}}-\hat{W}_{a,g,h,s})/\hat{W}_{a,g,h,s=\text{scen1}}$                                                                                  Population-attributable fractions relative to scenario 1                                         pif_scen
-  $\bar{U}_{a,g,h,o}=U_{a,g,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$                                                                                                                            GBD scaled to local population                                                                   DISEASE_BURDEN
-  $\tilde{U}_{a,g,h,o}=\rho_h\cdot  \bar{U}_{a,g,h,o}$                                                                                                                                    Scaled background disease for Neoplasms, IHD, Lung cancer                                        gbd_data_scaled
-  $\hat{U}_{a,g,h,o,s}=\bar{W}_{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}$                                                                                                                         Combined health and PIF                                                                          yll_dfs, death_dfs
-  $\hat{U}_{a,g,h=\text{road injury},o,s}=\sum_m\check{I}_{a,g,m,o,s}$                                                                                                                    Injury health burden                                                                             inj
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| Name                                                                                                                                                         | Description                                                                                    | Label                                        |
++:=============================================================================================================================================================+:===============================================================================================+:=============================================+
+| *Scenario generation*                                                                                                                                        |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| ...
+|                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| *Travel calculations*                                                                                                                                        |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk} \}\, \forall\, T_j:m(j)=\text{bus}$                                                               | New walk trips to bus                                                                          | walk *to_bus_and* combine_scen               |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\bar{T}_{a,g,m,s}=\sum_{j:a(j )= a,g(j)=g,m(j)=m,s(j)=s}Q_{i(j)}T_j$                                                                                        | (Weighted) total travel per age, gender, mode and scenario                                     |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| \$ \che ck{T}\_{a,g,m,s}=\frac{\bar{T}_{a,                                                                                                                   | Relative travel per mode and scenario by age and gender                                        | inj_distan ces\[\[1\]\]                      |
+| g, m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}\$                                                                                                                      |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\hat{ T} _{m,s}=\sum_{a,g}\bar{T}_{a,g,m,s}$                                                                                                                | Total travel per mode and scenario                                                             | dist                                         |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{T}_{m,s}=\frac{\hat{T}_{ m, s}}{\hat{T}_{m,s=\text{baseline}}}$                                                                                      | Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1 | inj_distan ces\[\[2\]\]                      |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| *Injury calculations*                                                                                                                                        |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $E_{m_{\text{cas}}}=\psi\omega$                                                                                                                              | Casualty exponent                                                                              | CAS_EXPONENT                                 |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $E_{m_{\t ex t{str}}}=\omega-E_{m_{\text{cas}}}$                                                                                                             | Striker exponent                                                                               | STR_EXPONENT                                 |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}=\mathcal{S}\left(\sigma\c dot {I}_{a,g,m_{\text{vic}},m_{\text{| Fataliti                                   | es in scenarios by strike mode \| injuries                                                     |                                              |
+| str}}},\bar{T}_{a,g,m,s},E_{m_{\t ex t{cas}}},E_{m_{\text{str}}}\right)$                                                                                     |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\check{I}_{a,g,m,o=\text{death                                                                                                                              | Fatalities in scenarios                                                                        | v ictim_deaths                               |
+| },s} =\sum_{m_{\text{str}}}\tilde{I}_{ a, g,m_{\text{vic}},m_{\text{str}},s}$                                                                                |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $R_{a,g}=U_{a,g,h=\text{road                                                                                                                                 | Road YLL-to-death ratio                                                                        | GBD_INJ_YLL                                  |
+| injuries},o=\text{YLL}}/U_{a,g,h= \t ext{road injuries},o=\text{death}}$                                                                                     |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\check{ I}_{ a,g,m,o=\text{YLL},s}=\check{I}_{ a, g,m,o=\text{death},s}\cdot R_{a,g}$                                                                       | YLLs scaled from fatalities                                                                    | deaths\_ yll_injuries                        |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| *Pollution calculations*                                                                                                                                     |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\dot{P}_{m}=\left\{\begin{array}{lr} \gam ma_m & P=1 \\ \tilde{\gamma}_m, \  | tilde{\gamma}_m \sim \text{Dir}(f(P,\  | gamma_m)) & P<1 \end{array}\right.$ | Emissions by mode \| \|                                                                        | tra ns_emissions                             |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{P}_{m,s}=\dot{P}_m\hat{T} _{ m,s}/\hat{T}_{m,s=\text{baseline}}$                                                                                     | Emissions by mode in scenario relative to baseline                                             | tra ns_emissions                             |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| \$ \h at{P}*{s}=*\sum{m} \tilde{P}\_{m,s}\$                                                                                                                  | Emissions scalar for scenarios                                                                 | baseline_sum                                 |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\bar {P}_s=\eta(\zeta\hat{P}_s + 1-\zeta)$ \|                                                                                                               | Background PM2.5 concentration in scenarios \|                                                 | scenario_pm                                  |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $K_s=C_4-C_5\log(\bar{P_s})$                                                                                                                                 | On-road PM2.5 exposure ratio                                                                   | on_road_of f_road_ratio                      |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{K}_{m\                                                                                                                                               | On-road PM2.5 exposure ratio for open vehicle                                                  | on_road_of f_road_ratio                      |
+| in\{\text{walk,cyc.,mot.}\},s}=K_s$                                                                                                                          |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\                                                                                                                                                           | Subway PM2.5 exposure ratio \|                                                                 | subway_ratio                                 |
+| tilde{K}_{m\in\{\text{sub.}\},s}=C_6$ \|                                                                                                                     |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\                                                                                                                                                           | In-vehicle PM2.5 exposure ratio                                                                | in_v ehicle_ratio                            |
+| tilde{K}_{m\not\in\{\text{walk,cyc.
+|                                                                                                |                                              |
+| ,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3)$                                                                                                                          |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $V_m=C_1+\frac{1}{2}\lambda_m$                                                                                                                               | Ventilation rates                                                                              | vent_rates                                   |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde {V}_{i,m,s} = \sum_{j:i(j)=i,m(j) =m ,s(j)=s}T_j\cdot V_{m(j)}/Z_{m(j)}$ \|                                                                          | In-travel air inhaled \|                                                                       |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\bar{V}_{i,s} = \sum_{m}\t il de{V}_{i,m,s}\cdot \tilde{K}_{m,s}$                                                                                           | In-travel PM2.5 inhalation rate                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| \$ \hat{V}*{i,s} = C_1*\left(24-\sum{j                                                                                                                       | Non-travel air inhaled                                                                         |                                              |
+| :i(j)=i,s(j)=s}T_j/Z\_{m(j)}\right)\$                                                                                                                        |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| \$\check{W}*{i,s} =* \bar{ P}*s(*\bar{V}{i,s}+\hat{V}{i,s})/24                                                                                               | Total PM2.5 inhaled per hour                                                                   |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| %/(\sum\*m\                                                                                                                                                  |                                                                                                |                                              |
+| \*\tilde{V}{i,m,s}+\hat{V}\_{i,s})\$                                                                                                                         |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ \|                                                                              | AP dose-- response--curve parameter \|                                                         | DR_AP_LIST                                   |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{H}_{a,h,w=3} =\te xt{CDF}_{H_{a,h,w=3}|\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                           | AP dose-- response--curve parameter                                                            | DR_AP_LIST                                   |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{H }_{a,h,w=3}=\text{CDF}_{H_{a,h, w=3} |\tilde{H}_{a,h,w=3},\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ \|                                   | AP dose-- response--curve parameter \|                                                         | DR_AP_LIST                                   |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{H}_{a,h,w=3}=\text{CD                                                                                                                                | AP dose-- response--curve parameter                                                            | DR_AP_LIST                                   |
+| F}_{H_{a,h,w=3}|\tilde{H}_{a,h, w=3} ,\tilde{H}_{a,h,w=3},\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=3}\times$                                                                                                     |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\left\{1 - \exp\left(-\tilde{H }_ {a=a(i),h,w=3} (\check{W}_{i,s} - \                                                                                       | RR of disease given pollution                                                                  | RR_AP\_ calculations                         |
+| tilde{H}_{a=a(i),h,w=3}) ^ {\tild e{ H}_{a=a(i),h,w=3} }\right)\right\}$                                                                                     |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| *Physical activity calculations*                                                                                                                             |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\dot{Y}_{a,g}=\                                                                                                                                             | Raw probability non-travel MMETs = 0 for demographic group $a,g$                               | raw_zero                                     |
+| text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}$                                                                                                                          |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\b ar{Y}_{a,g}=\left\{\begin{array }{lr } \dot{Y}_{a,g} & \tilde{Y}=1 \\ \t ext{CDF}_{\text{Beta}(f(\tilde{Y},\                                             | Probability non-travel MMETs = 0 for demographic group $a,g$                                   | zeros                                        |
+| dot{Y}_{a,g}))}^{-1}(\tilde{\thet a} ) & \tilde{Y}<1 \end{array}\right.$                                                                                     |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\check{Y}_i = \left\{\begin {a rray}{lr} 0 & \bar{Y}_{a(i),g(i)} \                                                                                          | Non-travel MMETs                                                                               | SYNTHETIC\_ POPULATION\$ work_l tpa_marg_met |
+| \ \text{Sample}(\{Y_{i'}\}|_{a( i')= a(i),g(i')=g(i),Y_{i'}>0}) & 1-\b ar {Y}_{a(i),g(i)} \end{array}\right.$                                                |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $M_{i,s}=\sum_{j :i(j )=i,s(j)=s}(\chi T_j\cdot \lambda _{ m(j)}/Z_{m(j)})+\theta \check{Y}_i$                                                               | Total MMETs per person                                                                         | mmets                                        |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{G}_{h,x}(m) = {G}_{h, x,z= \lfloor{m}\rfloor}+({G}_{h,x,z=\l ce il{m}\rceil}-{G}_{h,x,z=\lfloor{m}\                                                  | Interpolated dose--response parameters                                                         | rr, lb, ub                                   |
+| rfloor})\frac{m-\lfloor{m}\rfloor }{ \lceil{m}\rceil-\lfloor{m}\rfloor}$                                                                                     |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $F_S(s_x) = \mathcal{N}(s_{x=2},( s_ {x=3}-s_{x=1})/1.96)|_{0<F_S(s_x)}$                                                                                     | Truncated normal function                                                                      |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $W_{h,i,s}=\text{CDF}_{F_S\                                                                                                                                  | Relative risk of disease given PA                                                              | RR_PA\_ calculations                         |
+| left (\tilde{G}_{h,x}(M_{i,s})\right)} ^{ -1}\left(\phi_{h}\right)^{\dagger}$                                                                                |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| *Burden-of-disease calculations*                                                                                                                             |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{W}_{h, i, s}=W_{h,i,s}\cdot\check{H}_{h,i,s}$                                                                                                        | Combined PA and AP relative risks                                                              | RR_PA_AP\_ calculations                      |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\hat{W}_{a,g,h,s}= \su m_ {i:a(i)=a,g(i)=g}\tilde{W}_{h,i,s}$                                                                                               | Populati on-attributable fractions                                                             | pif_temp                                     |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\bar{W}_{a,g,h,s}= (\hat{W} _{a, g,h,s=\text{scen1}}-\hat{W}_{a,g, h, s})/\hat{W}_{a,g,h,s=\text{scen1}}$                                                   | Populati on-attributable fractions relative to scenario 1                                      | pif_scen                                     |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\bar{U}_{a,g,h,o}=U_{a ,g ,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$                                                                                               | GBD scaled to local population                                                                 | DI SEASE_BURDEN                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{U}_{a, g, h,o}=\rho_h\cdot \bar{U}_{a,g,h,o}$                                                                                                        | Scaled background disease for Neoplasms, IHD, Lung cancer                                      | gbd \_data_scaled                            |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\hat{U}_{a,g,h,o,s}=\bar{ W} _{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}$                                                                                            | Combined health and PIF                                                                        | yll_dfs, death_dfs                           |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\hat{U}_{a,g,h=\text{road injury                                                                                                                            | Injury health burden                                                                           | inj                                          |
+| },o,s}=\sum_m\check{I}_{a,g,m,o,s}$                                                                                                                          |                                                                                                |                                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
 
@@
 
-We model uncertainty in travel via the definition of "travel attributes"
-(in the terminology of @Mohammadian2010). Specifically, we define the
-propensity to travel in a day, and the distance travelled in a day,
-given that travelled occurred. Each travel attribute is defined per
-demographic group, per mode.
-
-In contrast to other travel-related synthetic populations, we have four
-demographic groups (rather than clusters) which consist of individuals
-(rather than households). A crucial difference between our objectives
-and the existing work on simulating populations in travel-demand models
-(that I've found so far) is that we don't need trip-level data for our
-calculations. Simulating trip-level data is computationally expensive
-and high-dimensional in terms of uncertain parameters. For our
-downstream computations, we need only summary data: travel per person
-per mode. Beyond scenario generation, there is no need for us to
-consider trips.
-
-E.g. @Saadi2016 focuses on (a) synthetic population via matching
-covariates and (b) the sequence of trips taken by each person. (a) isn't
-a big thing for Accra; we're only using two covariates. And for (b), I
-think we can focus on people as it's a lot simpler. For our purposes,
-the things in this work that might be useful to us include matching
-principles for (a), when we revisit it, and distributions (in particular
-joint distributions) for travel. More useful to us would be
-distributions for total travel, but I haven't found any in this
-literature (e.g. zero-inflated compound Poisson distribution).
+We model uncertainty in travel via the definition of "travel attributes" (in the terminology of @Mohammadian2010).
+Specifically, we define the propensity to travel in a day, and the distance travelled in a day, given that travelled occurred.
+Each travel attribute is defined per demographic group, per mode.
+
+In contrast to other travel-related synthetic populations, we have four demographic groups (rather than clusters) which consist of individuals (rather than households).
+A crucial difference between our objectives and the existing work on simulating populations in travel-demand models (that I've found so far) is that we don't need trip-level data for our calculations.
+Simulating trip-level data is computationally expensive and high-dimensional in terms of uncertain parameters.
+For our downstream computations, we need only summary data: travel per person per mode.
+Beyond scenario generation, there is no need for us to consider trips.
+
+E.g.
+@Saadi2016 focuses on (a) synthetic population via matching covariates and (b) the sequence of trips taken by each person.
+(a) isn't a big thing for Accra; we're only using two covariates.
+And for (b), I think we can focus on people as it's a lot simpler.
+For our purposes, the things in this work that might be useful to us include matching principles for (a), when we revisit it, and distributions (in particular joint distributions) for travel.
+More useful to us would be distributions for total travel, but I haven't found any in this literature (e.g.
+zero-inflated compound Poisson distribution).
 
@@
 
--   What to do about zero travel for e.g. older cyclists. One
-    possibility is to smooth over all modes and/or demographic groups.
+-   What to do about zero travel for e.g. older cyclists. One possibility is to smooth over all modes and/or demographic groups.
 
@@
 
--   How should we include correlations between modes? This could also be
-    aided by a smooth model across modes.
-
--   Whether there is a better method to get distance travelled per
-    person than by resampling raw data.
+-   How should we include correlations between modes? This could also be aided by a smooth model across modes.
+
+-   Whether there is a better method to get distance travelled per person than by resampling raw data.
 
@@
 
-Our first step is to summarise the demographic information and the
-travel survey. In Table [5](#population){reference-type="ref"
-reference="population"} are the demographic data: the populations of the
-city, and the number of people in the survey. In Table
-[6](#calculations){reference-type="ref" reference="calculations"} are
-the summary statistics from the travel survey which we will use to
-generate synthetic travel. From the raw travel data (actually the
-adjusted trip set, which is the raw travel plus motorcycle, bus and
-truck trips), we calculate the probability to travel, $B_{a,g,m}$, as
-the number of people who travelled by that mode divided by the number of
-people in the demographic group. It is to these values that we assign
-uncertainty through the variable PROPENSITY_TO_TRAVEL.
-
-::: {.center}
+Our first step is to summarise the demographic information and the travel survey.
+In Table [5](#population){reference-type="ref" reference="population"} are the demographic data: the populations of the city, and the number of people in the survey.
+In Table [6](#calculations){reference-type="ref" reference="calculations"} are the summary statistics from the travel survey which we will use to generate synthetic travel.
+From the raw travel data (actually the adjusted trip set, which is the raw travel plus motorcycle, bus and truck trips), we calculate the probability to travel, $B_{a,g,m}$, as the number of people who travelled by that mode divided by the number of people in the demographic group.
+It is to these values that we assign uncertainty through the variable PROPENSITY_TO_TRAVEL.
+
+::: center
 ::: {#population}
-  Age      Gender     Surveyed   Accra population
-  -------- -------- ---------- ------------------
-  15--49   Male            279          5,847,716
-  50--69   Male             56          1,016,476
-  15--49   Female          328          6,360,535
-  50--69   Female           69          1,110,037
+| Age    | Gender | Surveyed | Accra population |
+|:-------|:-------|---------:|-----------------:|
+| 15--49 | Male   |      279 |        5,847,716 |
+| 50--69 | Male   |       56 |        1,016,476 |
+| 15--49 | Female |      328 |        6,360,535 |
+| 50--69 | Female |       69 |        1,110,037 |
 
@@
 
-::: {.center}
-::: {.ThreePartTable}
-::: {.TableNotes}
+::: center
+::: ThreePartTable
+::: TableNotes
 :::
@@
 ::: {#calculations}
-  Age     Gender   Mode              Probability in raw trip set   Probability in baseline trip set
-  ------- -------- --------------- ----------------------------- ----------------------------------
-  15-49   Male     Bus                                    0.4014                             0.3862
-  50-69   Male     Bus                                    0.4286                             0.3636
-  15-49   Female   Bus                                    0.3506                             0.3495
-  50-69   Female   Bus                                    0.4203                             0.4203
-  15-49   Male     Taxi                                   0.0681                             0.0655
-  50-69   Male     Taxi                                   0.0357                             0.0303
-  15-49   Female   Taxi                                   0.0671                             0.0669
-  50-69   Female   Taxi                                   0.1159                             0.1159
-  15-49   Male     Walking                                0.6559                              0.631
-  50-69   Male     Walking                                0.4643                             0.3939
-  15-49   Female   Walking                                0.5427                              0.541
-  50-69   Female   Walking                                0.4203                             0.4203
-  15-49   Male     Private Car                            0.1254                             0.1207
-  50-69   Male     Private Car                            0.1964                             0.1667
-  15-49   Female   Private Car                            0.0549                             0.0547
-  50-69   Female   Private Car                             0.058                              0.058
-  15-49   Male     Bicycle                                0.0179                             0.0172
-  50-69   Male     Bicycle                                     0                                  0
-  15-49   Female   Bicycle                                 0.003                              0.003
-  50-69   Female   Bicycle                                     0                                  0
-  15-49   Male     Short Walking                               0                             0.3862
-  50-69   Male     Short Walking                               0                             0.3636
-  15-49   Female   Short Walking                               0                             0.3495
-  50-69   Female   Short Walking                               0                             0.4203
-  15-49   Male     Motorcycle                                  0                             0.0621
-  50-69   Male     Motorcycle                                  0                             0.1515
-  15-49   Female   Motorcycle                                  0                             0.0061
-  50-69   Female   Motorcycle                                  0                             0.0145
+| Age   | Gender | Mode          | Probability in raw trip set | Probability in baseline trip set |
+|:------|:-------|:--------------|----------------------------:|---------------------------------:|
+| 15-49 | Male   | Bus           |                      0.4014 |                           0.3862 |
+| 50-69 | Male   | Bus           |                      0.4286 |                           0.3636 |
+| 15-49 | Female | Bus           |                      0.3506 |                           0.3495 |
+| 50-69 | Female | Bus           |                      0.4203 |                           0.4203 |
+| 15-49 | Male   | Taxi          |                      0.0681 |                           0.0655 |
+| 50-69 | Male   | Taxi          |                      0.0357 |                           0.0303 |
+| 15-49 | Female | Taxi          |                      0.0671 |                           0.0669 |
+| 50-69 | Female | Taxi          |                      0.1159 |                           0.1159 |
+| 15-49 | Male   | Walking       |                      0.6559 |                            0.631 |
+| 50-69 | Male   | Walking       |                      0.4643 |                           0.3939 |
+| 15-49 | Female | Walking       |                      0.5427 |                            0.541 |
+| 50-69 | Female | Walking       |                      0.4203 |                           0.4203 |
+| 15-49 | Male   | Private Car   |                      0.1254 |                           0.1207 |
+| 50-69 | Male   | Private Car   |                      0.1964 |                           0.1667 |
+| 15-49 | Female | Private Car   |                      0.0549 |                           0.0547 |
+| 50-69 | Female | Private Car   |                       0.058 |                            0.058 |
+| 15-49 | Male   | Bicycle       |                      0.0179 |                           0.0172 |
+| 50-69 | Male   | Bicycle       |                           0 |                                0 |
+| 15-49 | Female | Bicycle       |                       0.003 |                            0.003 |
+| 50-69 | Female | Bicycle       |                           0 |                                0 |
+| 15-49 | Male   | Short Walking |                           0 |                           0.3862 |
+| 50-69 | Male   | Short Walking |                           0 |                           0.3636 |
+| 15-49 | Female | Short Walking |                           0 |                           0.3495 |
+| 50-69 | Female | Short Walking |                           0 |                           0.4203 |
+| 15-49 | Male   | Motorcycle    |                           0 |                           0.0621 |
+| 50-69 | Male   | Motorcycle    |                           0 |                           0.1515 |
+| 15-49 | Female | Motorcycle    |                           0 |                           0.0061 |
+| 50-69 | Female | Motorcycle    |                           0 |                           0.0145 |
 
@@
 
-We define a trip-mode weight per mode, $L_m$, so that a single trip by
-mode $m$ in the baseline scenario represents $L_m$ trips in reality. If
-in a scenario the mode of the trip is changed ($m'\rightarrow m$), it
-retains the weight of its original mode $m'$. We use the trip-mode
-weight to (a) sample trips for scenario generation and (b) scale the
-probabilities to travel. Note that we don't use it as trip weights are
-ordinarily used: to scale trip distance/duration. The result is a scaled
-set of probability parameters: $$\hat{B}_{a,g,m} = L_m{B}_{a,g,m}.$$
-($L_m$ should actually be the average trip-mode weight for all trips in
-the scenario. For baseline, it will be the same. It can/should differ
-for scenarios. Need to work out notation.)
+We define a trip-mode weight per mode, $L_m$, so that a single trip by mode $m$ in the baseline scenario represents $L_m$ trips in reality.
+If in a scenario the mode of the trip is changed ($m'\rightarrow m$), it retains the weight of its original mode $m'$.
+We use the trip-mode weight to (a) sample trips for scenario generation and (b) scale the probabilities to travel.
+Note that we don't use it as trip weights are ordinarily used: to scale trip distance/duration.
+The result is a scaled set of probability parameters: $$\hat{B}_{a,g,m} = L_m{B}_{a,g,m}.$$ ($L_m$ should actually be the average trip-mode weight for all trips in the scenario.
+For baseline, it will be the same.
+It can/should differ for scenarios.
+Need to work out notation.)
 
@@
 
-We estimate smoothed probabilities for any mode that has zero or one
-probability, using the rest of the probabilities for reference,
-replacing $\hat{B}_{a,g,m}$. The reference to other modes is slightly
-problematic if e.g. in the scenario all travel increases. It means a 0
-in the baseline is not the same as a 0 in the scenario. This could be
-fixed with more rigid coding.
+We estimate smoothed probabilities for any mode that has zero or one probability, using the rest of the probabilities for reference, replacing $\hat{B}_{a,g,m}$.
+The reference to other modes is slightly problematic if e.g.
+in the scenario all travel increases.
+It means a 0 in the baseline is not the same as a 0 in the scenario.
+This could be fixed with more rigid coding.
 
@@
 
-We resample the adjusted probability parameters of Table
-[6](#calculations){reference-type="ref" reference="calculations"}. We
-describe each parameter with a Beta distribution as follows:
-$$\begin{aligned}
+We resample the adjusted probability parameters of Table [6](#calculations){reference-type="ref" reference="calculations"}.
+We describe each parameter with a Beta distribution as follows: $$\begin{aligned}
 \label{pitilde}
@@
 \alpha_{a,g,m}+\beta_{a,g,m} & = 200; \\
-\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}.\end{aligned}$$
-Here, $\tilde{B}_{a,g,m}$ is the new, variable parameter that we
-resample for each simulation. We sample via a quantile,
-$\nu_m\sim\mathcal{U}(0,1)$, which we use to calculate
-$\tilde{B}_{a,g,m}$. We set $\alpha_{a,g,m}$ and $\beta_{a,g,m}$ to sum
-to 200, making the distributions concentrated, and we constrain their
-relative values to enforce the distribution's mean to equal
-$\hat{B}_{a,g,m}$ (the right-most column in Table
-[6](#calculations){reference-type="ref" reference="calculations"}).
+\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}.\end{aligned}$$ Here, $\tilde{B}_{a,g,m}$ is the new, variable parameter that we resample for each simulation.
+We sample via a quantile, $\nu_m\sim\mathcal{U}(0,1)$, which we use to calculate $\tilde{B}_{a,g,m}$.
+We set $\alpha_{a,g,m}$ and $\beta_{a,g,m}$ to sum to 200, making the distributions concentrated, and we constrain their relative values to enforce the distribution's mean to equal $\hat{B}_{a,g,m}$ (the right-most column in Table [6](#calculations){reference-type="ref" reference="calculations"}).
 
@@
 
-Each individual is assigned their own propensity to travel per mode,
-$\delta_{i,m}$, which is a uniformly distributed random variable that is
-fixed and not resampled. Whether or not individual $i$ travels by mode
-$m$ depends on $\tilde{B}_{a(i),g(i),m}$, which does vary: specifically,
-individual $i$ travels by mode $m$ if
-$\delta_{i,m}<\tilde{B}_{a(i),g(i),m}$. (To mitigate the effects of
-these fixed random variables, a large population is required, e.g. more
-than 5000 individuals.)
-
-![Distributions for propensities to travel. The propensity is the
-probability that a member of a demographic group will travel by a
-particular mode on a particular day. In orange are the raw probabilities
-based on the synthetic trip set, here shown for the baseline scenario.
-Beta distributions are defined via a pointiness parameter, set to 200,
-such that $\alpha+\beta=200$.
-](propensity_to_travel_200.pdf){#propensity_to_travel width="70%"}
+Each individual is assigned their own propensity to travel per mode, $\delta_{i,m}$, which is a uniformly distributed random variable that is fixed and not resampled.
+Whether or not individual $i$ travels by mode $m$ depends on $\tilde{B}_{a(i),g(i),m}$, which does vary: specifically, individual $i$ travels by mode $m$ if $\delta_{i,m}<\tilde{B}_{a(i),g(i),m}$.
+(To mitigate the effects of these fixed random variables, a large population is required, e.g.
+more than 5000 individuals.)
+
+\[Distributions for propensities to travel.
+The propensity is the probability that a member of a demographic group will travel by a particular mode on a particular day.
+In orange are the raw probabilities based on the synthetic trip set, here shown for the baseline scenario.
+Beta distributions are defined via a pointiness parameter, set to 200, such that $\alpha+\beta=200$.\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("propensity_to_travel_200.pdf")
+```
+
+(){#propensity_to_travel width="70%"}
 
@@
 
-In order to extrapolate to one week of travel, we consider "trips" only
-at the stage of scenario generation, and proceed with individual-level
-statistics. Then, in our synthetic population, we have "duration
-travelled per mode" columns. None of the core ITHIM-R functions will use
-the "synthetic trip set".
-
-To get duration per mode per person, we use, for each
-demographic-group--mode combination, the probability to travel, and the
-density of the total duration travelled given that travel occurred. To
-get from one day to one week, consider two cases: one, each individual
-follows the same pattern of travel every day. That is, we sample one
-day's travel and apply it to seven days. Two, each individual follows
-the travel pattern of their demographic group on each day. That is, we
-sample seven times from the group's density to make up seven days of
-travel.
-
-In implementation, we blend individual and group-level travel through a
-"repetitiveness" parameter, $\pi$. Each individual has two travel sets:
-one from their simulated daily travel pattern, and one from the shared,
-convolved distribution. These travel sets are weighted by $\pi$ and
-$1-\pi$, respectively. $\delta_{i,m}$ defines the quantile taken for
-both the individual density, which is scaled up to one week, and the
-shared distribution, which is the convolution of the raw density seven
-times.[^6] The densities are shown in Figure
-[17](#raw_densities){reference-type="ref" reference="raw_densities"}.
-
-![Raw distributions for duration travelled per day (left) from the
-travel survey. Convolution to one week is shown on the right. We use
-""walking"" mode for demographic group "Male",
-"15--49".](raw_densities.pdf){#raw_densities width="80%"}
+In order to extrapolate to one week of travel, we consider "trips" only at the stage of scenario generation, and proceed with individual-level statistics.
+Then, in our synthetic population, we have "duration travelled per mode" columns.
+None of the core ITHIM-R functions will use the "synthetic trip set".
+
+To get duration per mode per person, we use, for each demographic-group--mode combination, the probability to travel, and the density of the total duration travelled given that travel occurred.
+To get from one day to one week, consider two cases: one, each individual follows the same pattern of travel every day.
+That is, we sample one day's travel and apply it to seven days.
+Two, each individual follows the travel pattern of their demographic group on each day.
+That is, we sample seven times from the group's density to make up seven days of travel.
+
+In implementation, we blend individual and group-level travel through a "repetitiveness" parameter, $\pi$.
+Each individual has two travel sets: one from their simulated daily travel pattern, and one from the shared, convolved distribution.
+These travel sets are weighted by $\pi$ and $1-\pi$, respectively.
+$\delta_{i,m}$ defines the quantile taken for both the individual density, which is scaled up to one week, and the shared distribution, which is the convolution of the raw density seven times.[^6] The densities are shown in Figure [17](#raw_densities){reference-type="ref" reference="raw_densities"}.
+
+\[Raw distributions for duration travelled per day (left) from the travel survey. Convolution to one week is shown on the right. We use ""walking"" mode for demographic group "Male", "15--49".\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("raw_densities.pdf")
+```
+
+(){#raw_densities width="80%"}
 
@@
 
-1.  Choose a population size of individuals ${N}$. Label each individual
-    $i=1,...,N$.
-
-2.  Allocate each individual to a demographic group according to the
-    proportions in the Accra population (Table
-    [5](#population){reference-type="ref" reference="population"}). Then
-    each demographic group has number $\tilde{N}_{a,g}$ and
-    $\sum_{a,g}\tilde{N}_{a,g}=N$.
-
-3.  Generate one standard uniform variable per person per mode,
-    $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual
-    $i$'s propensity to travel by mode $m$.
-
-4.  $\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel
-    by mode $m$ in scenario $s$. It is the proportion of raw
-    person--duration values that are zero. It is parametrised by
-    $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation
-    [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).
-
-5.  Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw
-    person--duration values for demographic group {$a,g$} and mode $m$
-    in scenario $s$:
-    $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$
-    i.e. we isolate all individuals $i$ in group $\{a,g\}$ and sum their
-    trip durations by mode $m$ in scenario $s$.
-
-6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}$ be that density corrected to
-    have proportion $\tilde{B}_{a,g,m,s}$ zeros.
-
-7.  Define each individual's travel duration as
-    $D_{i,m,s}=7\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})$
-    (see Figure [18](#distances_beta){reference-type="ref"
-    reference="distances_beta"} for resulting distributions).
-
-For sampling, we loop over steps 6--7, where each sample has unique
-$\tilde{B}_{a,g,m,s}$ values. In total, the number of uncertain
-variables is one per mode.
+1.  Choose a population size of individuals ${N}$. Label each individual $i=1,...,N$.
+
+2.
+Allocate each individual to a demographic group according to the proportions in the Accra population (Table [5](#population){reference-type="ref" reference="population"}).
+Then each demographic group has number $\tilde{N}_{a,g}$ and $\sum_{a,g}\tilde{N}_{a,g}=N$.
+
+3.  Generate one standard uniform variable per person per mode, $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual $i$'s propensity to travel by mode $m$.
+
+4.
+$\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel by mode $m$ in scenario $s$.
+It is the proportion of raw person--duration values that are zero.
+It is parametrised by $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).
+
+5.
+Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw person--duration values for demographic group {$a,g$} and mode $m$ in scenario $s$: $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$ i.e.
+we isolate all individuals $i$ in group $\{a,g\}$ and sum their trip durations by mode $m$ in scenario $s$.
+
+6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}$ be that density corrected to have proportion $\tilde{B}_{a,g,m,s}$ zeros.
+
+7.
+Define each individual's travel duration as $D_{i,m,s}=7\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})$ (see Figure [18](#distances_beta){reference-type="ref" reference="distances_beta"} for resulting distributions).
+
+For sampling, we loop over steps 6--7, where each sample has unique $\tilde{B}_{a,g,m,s}$ values. In total, the number of uncertain variables is one per mode.
 
@@
 
-1.  Choose a population size of individuals ${N}$. Label each individual
-    $i=1,...,N$.
-
-2.  Allocate each individual to a demographic group according to the
-    proportions in the Accra population (Table
-    [5](#population){reference-type="ref" reference="population"}). Then
-    each demographic group has number $\tilde{N}_{a,g}$ and
-    $\sum_{a,g}\tilde{N}_{a,g}=N$.
-
-3.  Generate one standard uniform variable per person per mode,
-    $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual
-    $i$'s propensity to travel by mode $m$.
-
-4.  $\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel
-    by mode $m$ in scenario $s$. It is the proportion of raw
-    person--duration values that are zero. It is parametrised by
-    $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation
-    [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).
-
-5.  Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw
-    person--duration values for demographic group {$a,g$} and mode $m$
-    in scenario $s$:
-    $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$
-    i.e. we isolate all individuals $i$ in group $\{a,g\}$ and sum their
-    trip durations by mode $m$ in scenario $s$.
-
-6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}^7$ be the week convolution of
-    $\tilde{\mathcal{F}}_{a,g,m,s}$ (see Figure
-    [17](#raw_densities){reference-type="ref"
-    reference="raw_densities"}, right).
-
-7.  Define each individual's travel duration as
-    $D_{i,m,s}=7\pi\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})+(1-\pi)\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}^7}^{-1}(\delta_{i,m})$
-    (see Figure [18](#distances_beta){reference-type="ref"
-    reference="distances_beta"} for resulting distributions).
-
-For sampling, we loop over steps 6--8, where each sample has unique
-$\tilde{B}_{a,g,m,s}$ and $\pi$ values. In total, the number of
-uncertain variables is one per mode plus one.
-
-![Samples of total distances travelled by mode and
-scenario.](distances_beta.pdf){#distances_beta width="80%"}
+1.  Choose a population size of individuals ${N}$. Label each individual $i=1,...,N$.
+
+2.
+Allocate each individual to a demographic group according to the proportions in the Accra population (Table [5](#population){reference-type="ref" reference="population"}).
+Then each demographic group has number $\tilde{N}_{a,g}$ and $\sum_{a,g}\tilde{N}_{a,g}=N$.
+
+3.  Generate one standard uniform variable per person per mode, $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual $i$'s propensity to travel by mode $m$.
+
+4.
+$\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel by mode $m$ in scenario $s$.
+It is the proportion of raw person--duration values that are zero.
+It is parametrised by $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).
+
+5.
+Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw person--duration values for demographic group {$a,g$} and mode $m$ in scenario $s$: $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$ i.e.
+we isolate all individuals $i$ in group $\{a,g\}$ and sum their trip durations by mode $m$ in scenario $s$.
+
+6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}^7$ be the week convolution of $\tilde{\mathcal{F}}_{a,g,m,s}$ (see Figure [17](#raw_densities){reference-type="ref" reference="raw_densities"}, right).
+
+7.
+Define each individual's travel duration as $D_{i,m,s}=7\pi\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})+(1-\pi)\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}^7}^{-1}(\delta_{i,m})$ (see Figure [18](#distances_beta){reference-type="ref" reference="distances_beta"} for resulting distributions).
+
+For sampling, we loop over steps 6--8, where each sample has unique $\tilde{B}_{a,g,m,s}$ and $\pi$ values. In total, the number of uncertain variables is one per mode plus one.
+
+\[Samples of total distances travelled by mode and scenario.\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("distances_beta.pdf")
+```
+
+(){#distances_beta width="80%"}
 
@@
 
-![YLLs in the scenario minus YLLs in the baseline for Accra, with
-motorcycle probability scalar 2 in
-Accra.](accra_walk_yll_uncertain_travel.pdf){#results width="80%"}
-
-![EVPPI for Accra's "walking" scenario for all causes of YLLs excluding
-"all cause" and "neoplasm", with propensity to motorcycle =
-2.](accra_walk_evppi_uncertain_travel.pdf){#evppi_mode_propensities_200
-width="70%"}
+\[YLLs in the scenario minus YLLs in the baseline for Accra, with motorcycle probability scalar 2 in Accra.\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("accra_walk_yll_uncertain_travel.pdf")
+```
+
+(){width="80%"}
+
+\[EVPPI for Accra's "walking" scenario for all causes of YLLs excluding "all cause" and "neoplasm", with propensity to motorcycle = 2.\]
+
+```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
+knitr::include_graphics("accra_walk_evppi_uncertain_travel.pdf")
+```
+
+(){#evppi_mode_propensities_200 width="70%"}
 
@@
 
-    $\text{Letter}$ Meaning                           $\text{Letter}$ Meaning               $\text{Letter}$ Meaning
-  ----------------- ------------------------------- ----------------- ------------------- ----------------- --------------------------
-           $\alpha$                                               $a$ age                               $A$ 
-            $\beta$                                               $b$                                   $B$ raw travel probabilities
-           $\gamma$ emission uncertainty                          $c$                                   $C$ ventilation constants
-           $\delta$ propensity to travel                          $d$ demographic group                 $D$ 
-         $\epsilon$ walk-to-bus time                              $e$                                   $E$ injury distance exponent
-            $\zeta$ traffic PM2.5 fraction                        $f$                                   $F$ PA DR function
-             $\eta$ PM2.5 concentration                           $g$ gender                            $G$ PA DR curves
-           $\theta$ PA uncertainty                                $h$ disease                           $H$ AP DR curves
-            $\iota$                                               $i$ individual                        $I$ injuries
-                 $$                                               $j$ trip                              $J$ 
-           $\kappa$ injury regression parameter                   $k$                                   $K$ AP exposure
-          $\lambda$ travel MMETs                                  $l$                                   $L$ 
-              $\mu$ mode distances                                $m$ mode                              $M$ MMETs per person
-              $\nu$                                               $n$                                   $N$ population
-              $\xi$ AP DR quantile                                $o$ outcome                           $O$ 
-              $\pi$ travel repetitiveness                         $p$                                   $P$ vehicle emissions
-                 $$                                               $q$                                   $Q$ person weights
-             $\rho$ chronic disease scalar                        $r$                                   $R$ death-to-injury ratio
-           $\sigma$ injury reporting rate                         $s$ scenario                          $S$ 
-             $\tau$                                               $t$                                   $T$ trips
-         $\upsilon$                                               $u$                                   $U$ burden of disease
-             $\phi$ PA DR quantile                                $v$                                   $V$ AP per person
-                 $$                                               $w$ AP DR parameter                   $W$ relative risks
-             $\chi$ day-to-week travel scalar                     $x$ PA DR parameter                   $X$ injury model matrix
-             $\psi$ casualty fraction of $\omega$                 $y$ year                              $Y$ individual MMETs
-           $\omega$ sum of distance exponents                     $z$ PA dose                           $Z$ speeds
-:::
-
-[^1]: Ainsworth compendium 2011 sites.google.com/site/
-    compendiumofphysicalactivities for walking and cycling
-
-[^2]: For any person of age lower than 25, we set the relative risk to
-    1.
-
-[^3]: The four parameters refer, in numerical order, to
-    $\alpha, \beta, \gamma$ and tmrel in @Burnett2014.
-
-[^4]: This is a problem for the examples of Delhi and Bangalore, whose
-    trip sets have car travellers under the age at which driving is
-    permitted and who therefore should not contribute to striking
-    distance.
-
-[^5]: Note that the assymmetry in the offset term for the whw expression
-    arises due to the absence of demographic information for striking
-    vehicles. Were we to have that, the offset would read
-    $\log(\bar{T}_{a,g,m_{\text{cas}},s})+(E_{m_{\text{cas}}}-1)\log(\hat{T}_{m_{\text{cas}},s})+\log(\bar{T}_{a,g,m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})$.
-    For our case, the latter two terms simplify to
-    $\log(\hat{T}_{m_{\text{str}},s})+(E_{m_{\text{str}}}-1)\log(\hat{T}_{m_{\text{str}},s})=E_{m_{\text{str}}}\log(\hat{T}_{m_{\text{str}},s})$.
-
-[^6]: Using the same random variable for both travel sets increases the
-    variability of the resulting dataset. Conversely, using different
-    random variables would result in more homogenous durations, and
-    fewer zeros.
+| $\text{Letter}$ | Meaning                       | $\text{Letter}$ | Meaning           | $\text{Letter}$ | Meaning                  |
+|----------------:|:------------------------------|----------------:|:------------------|----------------:|:-------------------------|
+|        $\alpha$ |                               |             $a$ | age               |             $A$ |                          |
+|         $\beta$ |                               |             $b$ |                   |             $B$ | raw travel probabilities |
+|        $\gamma$ | emission uncertainty          |             $c$ |                   |             $C$ | ventilation constants    |
+|        $\delta$ | propensity to travel          |             $d$ | demographic group |             $D$ |                          |
+|      $\epsilon$ | walk-to-bus time              |             $e$ |                   |             $E$ | injury distance exponent |
+|         $\zeta$ | traffic PM2.5 fraction        |             $f$ |                   |             $F$ | PA DR function           |
+|          $\eta$ | PM2.5 concentration           |             $g$ | gender            |             $G$ | PA DR curves             |
+|        $\theta$ | PA uncertainty                |             $h$ | disease           |             $H$ | AP DR curves             |
+|         $\iota$ |                               |             $i$ | individual        |             $I$ | injuries                 |
+|            \$\$ |                               |             $j$ | trip              |             $J$ |                          |
+|        $\kappa$ | injury regression parameter   |             $k$ |                   |             $K$ | AP exposure              |
+|       $\lambda$ | travel MMETs                  |             $l$ |                   |             $L$ |                          |
+|           $\mu$ | mode distances                |             $m$ | mode              |             $M$ | MMETs per person         |
+|           $\nu$ |                               |             $n$ |                   |             $N$ | population               |
+|           $\xi$ | AP DR quantile                |             $o$ | outcome           |             $O$ |                          |
+|           $\pi$ | travel repetitiveness         |             $p$ |                   |             $P$ | vehicle emissions        |
+|            \$\$ |                               |             $q$ |                   |             $Q$ | person weights           |
+|          $\rho$ | chronic disease scalar        |             $r$ |                   |             $R$ | death-to-injury ratio    |
+|        $\sigma$ | injury reporting rate         |             $s$ | scenario          |             $S$ |                          |
+|          $\tau$ |                               |             $t$ |                   |             $T$ | trips                    |
+|      $\upsilon$ |                               |             $u$ |                   |             $U$ | burden of disease        |
+|          $\phi$ | PA DR quantile                |             $v$ |                   |             $V$ | AP per person            |
+|            \$\$ |                               |             $w$ | AP DR parameter   |             $W$ | relative risks           |
+|          $\chi$ | day-to-week travel scalar     |             $x$ | PA DR parameter   |             $X$ | injury model matrix      |
+|          $\psi$ | casualty fraction of $\omega$ |             $y$ | year              |             $Y$ | individual MMETs         |
+|        $\omega$ | sum of distance exponents     |             $z$ | PA dose           |             $Z$ | speeds                   |
+:::
+
+[^6]: Using the same random variable for both travel sets increases the variability of the resulting dataset.
+Conversely, using different random variables would result in more homogenous durations, and fewer zeros.

### 2021-10-20 latex/output-md.Rmd
Add references

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+
+bibliography: ithim.bib
 
@@
 ::: {#inputs}
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \[\[inputs\]\]{label= "inputs"}           |                                                                 | *ITHIM-R reference*                |                               |                    |
-+:==========================================+=================================================================+:===================================+================================+================+
-| Name                                      | Description                                                     | Value in Accra model               | Label                          | Model/ Setting |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++:==========================================+=================================================================+:===================================+===============================+====================+
+| **Name**                                  | **Description**                                                 | **Value in Accra model**           | **Label**                     | **Model/ Setting** |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Global values*                           |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $U_ { a,g,h,o}$                           | Background burden of disease                                    | Constant                           | GBD_DATA                      | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $N_{a,g}$                                 | Population by age and gender in GBD_DATA                        | Constant                           | GBD_DATA                      | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\bar { N}_{a,g}$                         | Population by age and gender                                    | Constant                           | DEMOGRAPHIC                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\rho_h$                                  | Chronic disease scalar                                          | 1 or Lnorm(0,0.18)                 | CHRONIC_DISEASE_SCALAR        | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Travel values*                           |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=walk}$                            | Walk speed                                                      | 4.8                                | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=cyc.}$                            | Cycle speed                                                     | 14.5                               | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=bus}$                             | Bus speed                                                       | 15                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=car}$                             | Car speed                                                       | 21                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=mot.}$                            | Motorbike speed                                                 | 25                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| \$\\epsilon \$                            | Walk time to bus                                                | 5 or Lnorm(log (5),0.18)           | BUS\_ WALK_TIME                | Setting        |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| $\mu_{m=HGV}$ \|                          | Truck distance relative to car \|                               | 0.21 or Beta(3,10)                 | TRUCK_TO\_ CAR_RATIO \| S      | etting         |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| $\mu_{m=bus}$ \|                          | Bus driver distance relative to bus passenger distance \|       | 1/45 or Bet a(20,600) \|           | BUS\_ TO_PASSEN GER_RATIO      | Setting        |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| \$\\epsilon \$                            | Walk time to bus                                                | 5 or Lnorm(log (5),0.18)           | BUS_WALK_TIME                 | Setting            |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\mu_{m=HGV}$                             | Truck distance relative to car                                  | 0.21 or Beta(3,10)                 | TRUCK_TO\_ CAR_RATIO          | Setting            |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\mu_{m=bus}$                             | Bus driver distance relative to bus passenger distance          | 1/45 or Bet a(20,600) \|           | BUS_TO_PASSENGER_RATIO        | Setting            |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\mu_ {m=mot.}$                           | Motorcycle fleet distance relative to total motorcycle distance | 37/68 or Beta(37/3+ 1,31/3+1) \|   | FLEET_TO_MOTORCYCLE_RATIO     | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $T_j$                                     | Trip-level data                                                 | Constant                           | SYNTHETIC_TRIPS               | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Injury model values*                     |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $I _ {a,g,m_{\                            | Fatalities table                                                | Constant                           | INJURY_TABLE                  | Setting            |
 | text{vi c} },m_{\text{str}}}$             |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\sigma$                                  | Injury reporting rate                                           | 1 or Beta(8,3)                     | INJURY_REPORTING_RATE         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\omega$                                  | Sum of distance exponents                                       | 1.9 or Lnorm(l og(1.9),l og(1.03)) | SIN_EXPONENT_SUM              | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\psi$                                    | Casualty fraction of $\omega$                                   | 0.5 or Beta(20,20) \|              | CASUALTY_EXPONENT_FRACTION    | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | -   Pollution model                       |                                                                 |                                    |                               |                    |
@@
 |     values\*                              |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| $\eta$                                    | Background PM2.5 concentration                                  | 50 or Lnorm(log( 50),0.18) \|      | PM_CONC_BASE \|                | Setting        |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| $\zeta$                                   | Fraction of measured PM2.5 concentration due to road transport  | 0.225 or Beta(5,20) \|             | PM_TRANS_SHARE \|              | Setting        |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\eta$                                    | Background PM2.5 concentration                                  | 50 or Lnorm(log( 50),0.18) \|      | PM_CONC_BASE                  | Setting            |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\zeta$                                   | Fraction of measured PM2.5 concentration due to road transport  | 0.225 or Beta(5,20) \|             | PM_TRANS_SHARE                | Setting            |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\                                        | Vehicle emission inventory                                      | Constant or Dirichlet              | EMISSION_INVENTORY            | Setting            |
 | g amma_{m}$                               |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| $P$                                       | Vehicle emission confidence                                     | 1                                  | EM ISSION_INVENTORY_CONFIDENCE | Setting        |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $P$                                       | Vehicle emission confidence                                     | 1                                  | EMISSION_INVENTORY_CONFIDENCE | Setting            |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | -   Pollution & health model              |                                                                 |                                    |                               |                    |
@@
 |     values\*                              |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| $C_1$                                     | Base-level inhalation rate                                      | 1                                  | BAS E_LEVEL_I NHALATION \_RATE | Model          |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| $C_2$                                     | Exposure ratio window closed                                    | 0.5                                | CLOSED_WI NDOW_PM\_ RATIO      | Model          |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $C_1$                                     | Base-level inhalation rate                                      | 1                                  | BASE_LEVEL_INHALATION_RATE    | Model              |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $C_2$                                     | Exposure ratio window closed                                    | 0.5                                | CLOSED_WINDOW_PM\_ RATIO      | Model              |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_3$                                     | Proportion of travel with closed windows                        | 0.5                                | C LOSED_WIN DOW_RATIO         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_4$                                     | Parameter for on-road PM2.5 level                               | 3.216                              | ROAD\_ RATIO_MAX              | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_5$                                     | Parameter for on-road PM2.5 level                               | 0.379                              | ROAD_RA TIO_SLOPE             | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_6$                                     | Exposure ratio for underground                                  | 0.8                                | SUBWAY \_PM_RATIO             | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$ H \_{a,h,w}\$                          | AP dose- -res ponse--curve parameters                           | Constant                           | DR_AP                         | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\                                        | Quantiles for AP RR functions                                   | 0.5 or Unif(0,1)                   | AP_DOSE_R ESPONSE\_ QUANTILE  | Model              |
 | xi_{h,w}$                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Physical activity model values*          |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
-| $\chi$                                    | Day-to-week travel scalar                                       | 7 or 7 $\times$Beta(20,3)          | DAY_TO_WEEK \_TRAVEL_SCALAR    | Model          |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\chi$                                    | Day-to-week travel scalar                                       | 7 or 7 $\times$Beta(20,3)          | DAY_TO_WEEK_TRAVEL_SCALAR     | Model              |
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Y_i$                                     | Individual n on-transport MMETs                                 | Constant                           | PA_SET\$ work_ltpa \_marg_met | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\theta$                                  | Background PA scalar                                            | 1 or Lnorm(0,0.18)                 | BACKGROUND\_ PA_SCALAR        | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\                                        | Confidence in PA survey                                         | 1                                  | BACKGROUND_PA_C ONFIDENCE     | Setting            |
 | tilde{Y}$                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\tilde {\theta}$                         | Background PA zeros quantile                                    | 0.5 or Unif(0,1)                   | BACKGROUND \_PA_ZEROS         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\lambda_{m=\text{cyc.}}$                 | Cycling MET surplus                                             | 4.63 or Lnorm(log (4.63),1)        | MMET_CYCLING                  | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\lambda_{m=\text{walk}}$                 | Walking MET surplus                                             | 2.53 or Lnorm(log (2.53),1)        | MMET_WALKING                  | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\lambda_{m\not\in\text{wal               | In-vehicle MET surplus                                          | 0                                  |                               | Model              |
 | k,cyc.}}$                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Physical activity & health model values* |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$\\G\_{h,x,z}\$                          | Look-up table for relative risk truncated normal distribution   | Constant                           |                               | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$\phi\_{h}\$                             | Quantiles for PA RR functions                                   | 0.5 or Unif(0,1)                   | PA_DOSE_RESPONSE_QUANTILE     | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+--------------------------------+----------------+
++-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 
@@
 \[Example of the dose--response workflow for PA and total cancer.
-Top left: results of the metaanalysis: mean, lower bound and upper bound of the relationship between PA dose and relative risk of disease.
+Top left: results of the meta-analysis: mean, lower bound and upper bound of the relationship between PA dose and relative risk of disease.
 Top right: examples of 500 samples from the normal distribution defined by the mean of the top-left panel and the standard deviation defined by (upper bound minus lower bound) divided by 1.96 and truncated at 0.
@@
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| \$ \che ck{T}\_{a,g,m,s}=\frac{\bar{T}_{a,                                                                                                                   | Relative travel per mode and scenario by age and gender                                        | inj_distan ces\[\[1\]\]                      |
+| \$ \che ck{T}\_{a,g,m,s}=\frac{\bar{T}_{a,                                                                                                                   | Relative travel per mode and scenario by age and gender                                        | inj_distances\[\[1\]\]                       |
 | g, m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}\$                                                                                                                      |                                                                                                |                                              |
@@
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{T}_{m,s}=\frac{\hat{T}_{ m, s}}{\hat{T}_{m,s=\text{baseline}}}$                                                                                      | Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1 | inj_distan ces\[\[2\]\]                      |
+| $\tilde{T}_{m,s}=\frac{\hat{T}_{ m, s}}{\hat{T}_{m,s=\text{baseline}}}$                                                                                      | Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1 | inj_distances\[\[2\]\]                       |
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
@@
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\dot{P}_{m}=\left\{\begin{array}{lr} \gam ma_m & P=1 \\ \tilde{\gamma}_m, \  | tilde{\gamma}_m \sim \text{Dir}(f(P,\  | gamma_m)) & P<1 \end{array}\right.$ | Emissions by mode \| \|                                                                        | tra ns_emissions                             |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{P}_{m,s}=\dot{P}_m\hat{T} _{ m,s}/\hat{T}_{m,s=\text{baseline}}$                                                                                     | Emissions by mode in scenario relative to baseline                                             | tra ns_emissions                             |
+| $\dot{P}_{m}=\left\{\begin{array}{lr} \gam ma_m & P=1 \\ \tilde{\gamma}_m, \  | tilde{\gamma}_m \sim \text{Dir}(f(P,\  | gamma_m)) & P<1 \end{array}\right.$ | Emissions by mode \| \|                                                                        | trans_emissions                              |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{P}_{m,s}=\dot{P}_m\hat{T} _{ m,s}/\hat{T}_{m,s=\text{baseline}}$                                                                                     | Emissions by mode in scenario relative to baseline                                             | trans_emissions                              |
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
@@
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $K_s=C_4-C_5\log(\bar{P_s})$                                                                                                                                 | On-road PM2.5 exposure ratio                                                                   | on_road_of f_road_ratio                      |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{K}_{m\                                                                                                                                               | On-road PM2.5 exposure ratio for open vehicle                                                  | on_road_of f_road_ratio                      |
+| $K_s=C_4-C_5\log(\bar{P_s})$                                                                                                                                 | On-road PM2.5 exposure ratio                                                                   | on_road_off_road_ratio                       |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{K}_{m\                                                                                                                                               | On-road PM2.5 exposure ratio for open vehicle                                                  | on_road_off_road_ratio                       |
 | in\{\text{walk,cyc.,mot.}\},s}=K_s$                                                                                                                          |                                                                                                |                                              |
@@
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\                                                                                                                                                           | In-vehicle PM2.5 exposure ratio                                                                | in_v ehicle_ratio                            |
+| $\                                                                                                                                                           | In-vehicle PM2.5 exposure ratio                                                                | in_vehicle_ratio                             |
 | tilde{K}_{m\not\in\{\text{walk,cyc.
@@
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\bar{U}_{a,g,h,o}=U_{a ,g ,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$                                                                                               | GBD scaled to local population                                                                 | DI SEASE_BURDEN                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{U}_{a, g, h,o}=\rho_h\cdot \bar{U}_{a,g,h,o}$                                                                                                        | Scaled background disease for Neoplasms, IHD, Lung cancer                                      | gbd \_data_scaled                            |
+| $\bar{U}_{a,g,h,o}=U_{a ,g ,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$                                                                                               | GBD scaled to local population                                                                 | DISEASE_BURDEN                               |
++--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
+| $\tilde{U}_{a, g, h,o}=\rho_h\cdot \bar{U}_{a,g,h,o}$                                                                                                        | Scaled background disease for Neoplasms, IHD, Lung cancer                                      | gbd_data_scaled                              |
 +--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+

### 2021-10-21 latex/output-md.Rmd
Add note for strike mode regularization

@@
 
+[Note on recategorization]{.ul}: for some of the modes (like for car, bus, motorcycle and cycle), if both strike and victim modes are same, we explicitly make the strike mode as 'noov'
+
 We model the number of injuries as a Poisson-distributed variable with an offset depending on the distance(s) and the reporting rate. We write the model as follows: $$\begin{aligned}
@@
 For each person's dose, we get the response by mapping the uniform random variable onto the truncated normal defined by the mean and bounds for that dosage.
+We use an R package called {drpa}, whose description can be seen here: <https://meta-analyses.github.io/drpa/drpa.html>
+
+The relationships comes out of our work @Leandro2018, where we look at the association of `physical activity` with various diseases.
 

### 2021-10-26 code/producing_trips_rd/trip-summary.Rmd
Summary statistics of trips datasets

--- 2021-04-27 code/producing_trips_rd/trip-summary.Rmd
@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    toc_depth: 5
+    toc_float: true
+    number_sections: true
 ---
 
+Note: in order to create the markdown properly from this file, Rstudio options need to be changed.
+Tools > Global Options > R Markdown > Evaluate chunks in directory > Current.
+By default this option is set to *Document*, but here you need to change it to *Current* so everything works as expected.
+
 ```{r, include = FALSE}
-source("used_functions.R")
+source("code/producing_trips_rd/used_functions.R")
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
@@
 ```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
-trip <- read_csv("../../data/local/buenos_aires/buenos_aires_trip.csv",
+trip <- read_csv("data/local/buenos_aires/buenos_aires_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/sao_paulo/sao_paulo_trip.csv",
+trip <- read_csv("data/local/sao_paulo/sao_paulo_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/belo_horizonte/belo_horizonte_trip.csv",
+trip <- read_csv("data/local/belo_horizonte/belo_horizonte_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/mexico_city/mexico_city_trip.csv",
+trip <- read_csv("data/local/mexico_city/mexico_city_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/bogota/bogota_trips.csv",
+trip <- read_csv("data/local/bogota/bogota_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/santiago/santiago_trip.csv",
+trip <- read_csv("data/local/santiago/santiago_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/vizag/vizag_trip.csv",
+trip <- read_csv("data/local/vizag/vizag_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/delhi/delhi_trip.csv",
+trip <- read_csv("data/local/delhi/delhi_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/bangalore/bangalore_trip.csv",
+trip <- read_csv("data/local/bangalore/bangalore_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/accra/accra_trip.csv",
+trip <- read_csv("data/local/accra/accra_trip.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("../../data/local/cape_town/cape_town_trip.csv",
+trip <- read_csv("data/local/cape_town/cape_town_trip.csv",
                  guess_max = 100000)
@@
 trip_summary$Cape_Town <- value
+
+###
+trip <- read_csv("data/local/antofagasta/antofagasta_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Antofagasta <- value
+
+trip <- read_csv("data/local/arica/arica_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Arica <- value
+
+trip <- read_csv("data/local/cali/cali_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Cali <- value
+
+trip <- read_csv("data/local/copiapo/copiapo_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Copiapo <- value
+
+trip <- read_csv("data/local/coquimbo_laserena/coquimbo_laserena_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Coquimbo <- value
+
+trip <- read_csv("data/local/gran_valparaiso/gran_valparaiso_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Gran_Valparaiso <- value
+
+trip <- read_csv("data/local/iquique_altohospicio/iquique_altohospicio_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Iquique <- value
+
+trip <- read_csv("data/local/medellin/medellin_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Medellin <- value
+
+trip <- read_csv("data/local/montevideo/montevideo_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Montevideo <- value
+
+trip <- read_csv("data/local/osorno/osorno_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Osorno <- value
+
+trip <- read_csv("data/local/puerto_montt/puerto_montt_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Puerto_montt <- value
+
+trip <- read_csv("data/local/san_antonio/san_antonio_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$San_antonio <- value
+
+trip <- read_csv("data/local/temuco_padrelascasas/temuco_padrelascasas_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Temuco <- value
+
+trip <- read_csv("data/local/valdivia/valdivia_trips.csv",
+                 guess_max = 100000)
+quality_check(trip)
+trip_summary$Valdivia <- value
+
 ```
 
-## summary table
+```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
+trip <- read_csv("data/local/buenos_aires/buenos_aires_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Buenos_Aires <- value
+
+trip <- read_csv("data/local/sao_paulo/sao_paulo_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Sao_Paulo <- value
+
+trip <- read_csv("data/local/belo_horizonte/belo_horizonte_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Belo_Horizonte <- value
+
+trip <- read_csv("data/local/mexico_city/mexico_city_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Mexico <- value
+
+trip <- read_csv("data/local/bogota/bogota_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Bogota <- value
+
+trip <- read_csv("data/local/santiago/santiago_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Santiago <- value
+
+trip <- read_csv("data/local/vizag/vizag_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Visakhapatnam <- value
+
+trip <- read_csv("data/local/delhi/delhi_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Delhi <- value
+
+trip <- read_csv("data/local/bangalore/bangalore_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Bangalore <- value
+
+trip <- read_csv("data/local/accra/accra_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Accra <- value
+
+trip <- read_csv("data/local/cape_town/cape_town_trip.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Cape_Town <- value
+
+###
+trip <- read_csv("data/local/antofagasta/antofagasta_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Antofagasta <- value
+
+trip <- read_csv("data/local/arica/arica_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Arica <- value
+
+trip <- read_csv("data/local/cali/cali_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Cali <- value
+
+trip <- read_csv("data/local/copiapo/copiapo_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Copiapo <- value
+
+trip <- read_csv("data/local/coquimbo_laserena/coquimbo_laserena_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Coquimbo <- value
+
+trip <- read_csv("data/local/gran_valparaiso/gran_valparaiso_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Gran_Valparaiso <- value
+
+trip <- read_csv("data/local/iquique_altohospicio/iquique_altohospicio_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Iquique <- value
+
+trip <- read_csv("data/local/medellin/medellin_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Medellin <- value
+
+trip <- read_csv("data/local/montevideo/montevideo_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Montevideo <- value
+
+trip <- read_csv("data/local/osorno/osorno_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Osorno <- value
+
+trip <- read_csv("data/local/puerto_montt/puerto_montt_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Puerto_montt <- value
+
+trip <- read_csv("data/local/san_antonio/san_antonio_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$San_antonio <- value
+
+trip <- read_csv("data/local/temuco_padrelascasas/temuco_padrelascasas_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Temuco <- value
+
+trip <- read_csv("data/local/valdivia/valdivia_trips.csv",
+                 guess_max = 100000)
+quality_check(trip, 15, 69)
+trip_summary2$Valdivia <- value
+```
+
+## summary table (all rows)
 ```{r, include = TRUE,  echo= FALSE, results='asis'}

### 2021-10-26 summary_tables.Rmd
Regen all outputs for latam cities

@@
 cities <- names(io)[!names(io) %in% 'scen_prop']
-
+LATAM <- TRUE
 #cities <- c('accra','sao_paulo','buenos_aires')

### 2021-10-26 code/WHW/WHW_cities.R

@@ -1,0 +1,251 @@
+#' ---
+#' title: "Script to get the WHW matrices for each injury dataset"
+#' author: "Daniel"
+#' output:
+#'   html_document:
+#'     toc: true
+#'     toc_float: true
+#' ---
+#' 
+
+#' *Loading libraries*
+library(dplyr)
+library(readr)
+
+#+ warning=FALSE, message=FALSE, echo=FALSE
+# Cleaning workspace
+rm(list = ls());gc()
+
+# Printing options
+options(scipen = 50)
+
+
+#' *Definition of functions*
+#' Sourcing package functions
+#source("R/assign_age_groups.R")
+#source("R/set_injury_contingency.R")
+
+get_whw <- function(injuries, population = NULL, ithim = F,
+                    min_age = 15, max_age = 69) {
+  # Same preprocessing as in the package (ithim_load_data)
+  #filename <- paste0(local_path,"/injuries_",CITY,".csv")
+  #injuries <- read_csv(filename,col_types = cols())
+  
+  # Transform to lowercase all strings
+  injuries$cas_mode <- tolower(injuries$cas_mode)
+  injuries$strike_mode <- tolower(injuries$strike_mode)
+  if('cas_gender'%in%colnames(injuries)) injuries$cas_gender <- tolower(injuries$cas_gender)
+  #injuries$strike_mode[is.na(injuries$strike_mode)] <- 'listed_na'
+  nov_words <- c('no.other.fixed.or.stationary.object','no other vehicle','none')
+  injuries$strike_mode[injuries$strike_mode%in%nov_words] <- 'nov'
+  ## add weight column if missing
+  if(!'weight'%in%colnames(injuries)) injuries$weight <- 1
+  
+  # Similar preprocessing as in the package (ithim_load_data), i.e., filter age
+  # age range and label incidents as "nov" when strike and cas mode are the same
+  if(ithim){
+    # Filter only injuries in the age range
+    if('cas_age'%in%colnames(injuries)) {
+      injuries <- injuries %>% filter(cas_age >= min_age & cas_age <= max_age)
+      if (!is.null(population)) {
+        pop <- population %>% 
+          rowwise() %>% 
+          mutate(age_lb = unlist(strsplit(age, "-"))[1], # lower bound
+                 age_ub = unlist(strsplit(age, "-"))[2] # upper bound
[…]

### 2021-11-10 summary_tables.Rmd
Add code to generate scen prop, and save it as a csv file

--- 2021-09-29 summary_tables.Rmd
@@
 kable(scen_prop, headings = "Scenario Proportions")
+
+```
+
+```{r}
+
+dat <- list()
+
+for (city in cities){
+
+dat[[city]] <- io[[city]]$trip_scen_sets %>% mutate(scenario = case_when(
+  scenario == "Scenario 1" ~ "PED_SC", 
+  scenario == "Scenario 2" ~ "CYC_SC",
+  scenario == "Scenario 3" ~ "CAR_SC",
+  scenario == "Scenario 4" ~ "MOT_SC",
+  scenario == "Scenario 5" ~ "BUS_SC",
+  scenario == "Baseline" ~ "Baseline")) %>% group_by(scenario, trip_mode, trip_distance_cat) %>% summarise(n = dplyr::n()) %>% mutate(city = city) %>% filter(!(trip_mode %in% "truck"))
+
+td <- dat[[city]]
+
+for (i in 1:nrow(td)){
+    val <- td %>% filter(scenario == "Baseline", trip_mode == td$trip_mode[i] & 
+                           trip_distance_cat == td$trip_distance_cat[i] &
+                           city == td$city[i])
+
+    if (nrow(val) > 0){
+      td$diff[i] <- val$n - td$n[i]
+      td$prop[i] <- round(td$n[i] / val$n, 2)
+
+    }
+}
+
+dat[[city]] <- td
+
+}
+
+write_csv(bind_rows(dat), "results/multi_city/scen_prop.csv")
 

### 2021-11-18 summary_tables.Rmd
Regen latam
Rename variables and add trip proportion for scenarios as well

--- 2021-11-10 summary_tables.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')
+knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis', warning = F)
 
@@
 
-```{r}
+```{r warning=FALSE, echo=F, message=FALSE}
 
@@
   scenario == "Scenario 5" ~ "BUS_SC",
-  scenario == "Baseline" ~ "Baseline")) %>% group_by(scenario, trip_mode, trip_distance_cat) %>% summarise(n = dplyr::n()) %>% mutate(city = city) %>% filter(!(trip_mode %in% "truck"))
+  scenario == "Baseline" ~ "Baseline")) %>% group_by(scenario, trip_mode, trip_distance_cat) %>% summarise(n = dplyr::n(), dist = sum(trip_distance, na.rm = T)) %>% mutate(city = city)
 
@@
       td$diff[i] <- val$n - td$n[i]
-      td$prop[i] <- round(td$n[i] / val$n, 2)
+      td$trip_prop[i] <- round(td$n[i] / val$n, 2)
+      td$dist_prop[i] <- round(td$dist[i] / val$dist, 2)
 
@@
 for(city in cities){
-  print(city)
   df <- io[[city]]$trip_scen_sets
@@
 
-  print(city)
   count_people <- round(sum(io[[city]]$demographic$population))

### 2021-11-24 code/producing_trips_rd/trip-summary.Rmd
Changing names of files and population in some cases

@@
 ```{r,  comment=NA, message=FALSE, warning=FALSE, prompt=FALSE, cache=FALSE, echo=F}
-trip <- read_csv("data/local/buenos_aires/buenos_aires_trip.csv",
+trip <- read_csv("data/local/buenos_aires/buenos_aires_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/sao_paulo/sao_paulo_trip.csv",
+trip <- read_csv("data/local/sao_paulo/sao_paulo_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/belo_horizonte/belo_horizonte_trip.csv",
+trip <- read_csv("data/local/belo_horizonte/belo_horizonte_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/mexico_city/mexico_city_trip.csv",
+trip <- read_csv("data/local/mexico_city/mexico_city_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/santiago/santiago_trip.csv",
+trip <- read_csv("data/local/santiago/santiago_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/vizag/vizag_trip.csv",
+trip <- read_csv("data/local/vizag/vizag_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/delhi/delhi_trip.csv",
+trip <- read_csv("data/local/delhi/delhi_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/bangalore/bangalore_trip.csv",
+trip <- read_csv("data/local/bangalore/bangalore_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/accra/accra_trip.csv",
+trip <- read_csv("data/local/accra/accra_trips.csv",
                  guess_max = 100000)
@@
 
-trip <- read_csv("data/local/cape_town/cape_town_trip.csv",
+trip <- read_csv("data/local/cape_town/cape_town_trips.csv",
                  guess_max = 100000)

### 2021-11-24 code/producing_trips_rd/Preprocessing_cities/belo_horizonte/belo_horizonte.R

@@ -168,7 +168,7 @@ trip <- person %>%
   rename(age= IDADE, trip_id = Viagem)
 
 trip$meta_data <- NA
-trip$meta_data[1] <- 5595800
+trip$meta_data[1] <- '5,314,932 (2010)'
 trip$meta_data[2] <- 15134 
 trip$meta_data[3] <- "Travel Survey"
 trip$meta_data[4] <- 2012

### 2021-11-24 code/producing_trips_rd/Preprocessing_cities/buenos_aires/buenos_aires.R

@@ -218,7 +218,7 @@ trip <- person %>%
          stage_mode, stage_duration)
 
 trip$meta_data <- NA
-trip$meta_data[1] <- 13381800
+trip$meta_data[1] <- "12,806,866 (2010)"
 trip$meta_data[2] <- 23606 
 trip$meta_data[3] <- "Travel Survey"
 trip$meta_data[4] <- 2012

### 2021-11-24 code/producing_trips_rd/Preprocessing_cities/mexico_city/mexico_city.R

@@ -286,7 +286,7 @@ mexico_city_v2 <- mexico_city %>%
 
 trip <- mexico_city_v2
 trip$meta_data <- NA
-trip$meta_data[1] <- 20976700
+trip$meta_data[1] <- "9,200,318 (2020)"
 trip$meta_data[2] <- 19239
 trip$meta_data[3] <- "Travel Survey"
 trip$meta_data[4] <- 2017

### 2021-11-24 code/producing_trips_rd/Preprocessing_cities/sao_paulo/sao_paulo.R

@@ -316,7 +316,7 @@ report <- trips_saopaulo_ready %>%
   #arrange(cluster_id, household_id, participant_id, trip_id)
 
 report$meta_data <- NA
-report$meta_data[1] <- 20847500
+report$meta_data[1] <- '11,253,503 (2010)'
 report$meta_data[2] <- 20650 
 report$meta_data[3] <- "Travel Survey"
 report$meta_data[4] <- 2012

### 2021-11-24 summary_tables.Rmd
Suppress messages

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis', warning = F)
+knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis', warning = F, message = F)
 

### 2021-11-29 injury_tables.Rmd
Regen

@@
 
-collapse_ages <- function(data,ages=c('15-49','50-69'),age_lab='age_cat'){
+collapse_ages <- function(data, ages = c('15-49', '50-69'), age_lab = 'age_cat'){
   target_min_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][1]))
@@
   genders <- unique(data$sex)
-  if(ncol(data)>3) {
+  if (ncol(data) > 3) {
     reformatted <- do.call(rbind,lapply(1:length(ages),
@@
 io <- io[-2]
-for(i in 1:length(io)) {
-  io[[i]]$demographic$age[io[[i]]$demographic$age=='5-9'] <- '05-9'
+for (i in 1:length(io)) {
+  io[[i]]$demographic$age[io[[i]]$demographic$age == '5-9'] <- '05-9'
   io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
@@
 }
-for(city in cities)
-  for(type in c('hb','pathway_hb'))
-    for(out in c('deaths','ylls'))
+for (city in cities)
+  for (type in c('hb','pathway_hb'))
+    for (out in c('deaths','ylls'))
       io[[city]]$outcomes[[type]][[out]] <- collapse_ages(io[[city]]$outcomes[[type]][[out]])
@@
 
-scen_names_w_baseline <- c('bl', 'w_sc', 'bi_sc', 'car_sc', 'mc_sc', 'bus_sc')
+#scen_names_w_baseline <- c('bl', 'w_sc', 'bi_sc', 'car_sc', 'mc_sc', 'bus_sc')
+scen_names_w_baseline <- c('bl', 'bi_sc', 'car_sc', 'bus_sc')
 scen_lt <- data.frame("scen_name" = names(io[[1]]$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)
@@
 
-for(city in cities){
-  {
+for (city in cities) {
     cat('\n')
@@
       all_whw[[var]][[city]] <- td
-    }
-
     }
@@
 injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Driving'
-injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
+#injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+#injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
 
@@
 injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Driving'
-injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
+#injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+#injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
 

### 2021-11-29 summary_tables.Rmd
Regen

@@
 LATAM <- TRUE
+NSCEN <- 3
 #cities <- c('accra','sao_paulo','buenos_aires')
@@
 
-for (city in cities){
+for (city in cities) {
 
-dat[[city]] <- io[[city]]$trip_scen_sets %>% mutate(scenario = case_when(
-  scenario == "Scenario 1" ~ "PED_SC", 
-  scenario == "Scenario 2" ~ "CYC_SC",
-  scenario == "Scenario 3" ~ "CAR_SC",
-  scenario == "Scenario 4" ~ "MOT_SC",
-  scenario == "Scenario 5" ~ "BUS_SC",
-  scenario == "Baseline" ~ "Baseline")) %>% group_by(scenario, trip_mode, trip_distance_cat) %>% summarise(n = dplyr::n(), dist = sum(trip_distance, na.rm = T)) %>% mutate(city = city)
+dat[[city]] <- io[[city]]$trip_scen_sets %>% 
+  mutate(scenario = case_when(
+    # scenario == "Scenario 1" ~ "PED_SC", 
+    # scenario == "Scenario 2" ~ "CYC_SC",
+    # scenario == "Scenario 3" ~ "CAR_SC",
+    # scenario == "Scenario 4" ~ "MOT_SC",
+    # scenario == "Scenario 5" ~ "BUS_SC",
+    scenario == "Scenario 1" ~ "CYC_SC", 
+    scenario == "Scenario 2" ~ "CAR_SC",
+    scenario == "Scenario 3" ~ "BUS_SC",
+    scenario == "Baseline" ~ "Baseline")) %>% 
+  group_by(scenario, trip_mode, trip_distance_cat) %>% 
+  summarise(n = dplyr::n(), dist = sum(trip_distance, na.rm = T)) %>% 
+  mutate(city = city)
 
@@
 
-for (i in 1:nrow(td)){
+for (i in 1:nrow(td)) {
     val <- td %>% filter(scenario == "Baseline", trip_mode == td$trip_mode[i] & 
@@
 
-    if (nrow(val) > 0){
+    if (nrow(val) > 0) {
       td$diff[i] <- val$n - td$n[i]
@@
 
-for(city in cities){
+for (city in cities) { # Loop for each city
   df <- io[[city]]$trip_scen_sets
-  u_trips <- df %>% dplyr::filter(scenario == "Baseline") %>% summarise(uid <- n_distinct(trip_id)) %>% as.numeric()
-  td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% group_by(trip_mode, scenario) %>% summarise(p = round(dplyr::n() / u_trips * 100, 1)) %>% spread(key = trip_mode, value = p) %>% mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
+  u_trips <- df %>% dplyr::filter(scenario == "Baseline") %>% 
+    summarise(uid = n_distinct(trip_id)) %>% as.numeric()
+  td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% 
+    group_by(trip_mode, scenario) %>% 
+    summarise(p = round(dplyr::n() / u_trips * 100, 1)) %>% 
+    spread(key = trip_mode, value = p) %>% 
+    mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
   td <- as.data.frame(t(td))
@@
   cat("\n")
-}
+} # End loop city
 
@@
 
-trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>% mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>% janitor::adorn_totals('row')
+trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
+  mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>%
+  janitor::adorn_totals('row')
 
@@
 
-for(city in cities){
+for (city in cities) {
 
   count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
+    mutate_if(is.numeric, round, digits = round_to) %>% 
+    mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
   #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:7] <- c("BASELINE", rownames(scen_prop))
+  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
 
@@
   count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
+    mutate_if(is.numeric, round, digits = round_to) %>% 
+    mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
   #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:7] <- c("BASELINE", rownames(scen_prop))
+  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
   x <- match(td$stage_mode, trip_order$mode)
@@
 
-for(city in cities){
+for (city in cities) {
 
   count_people <- round(sum(io[[city]]$demographic$population))
-  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, list(~round((.) * 365, 3)))
+  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
+    mutate_if(is.numeric, list(~round((.) * 365, 3)))
   #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:7] <- c("BASELINE", rownames(scen_prop))
+  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
 
@@
   #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:7] <- c("BASELINE", rownames(scen_prop))
+  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
   x <- match(td$stage_mode, trip_order$mode)
@@
 yll_totals <- lapply(io[-2],function(x){
-  temp <- sapply(1:5,function(y){
+  temp <- sapply(1:NSCEN,function(y){
     xx <- x$outcomes$hb$ylls
-    xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = 5)])
+    xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
     sapply(sort(unique(xx$age_cat)),function(z)
@@
 yll_rates <- lapply(cities, function(x)
-  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],5,1))*100000,
+  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000,
         Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), 
-                                             length = 5)*100000
+                                             length = NSCEN)*100000
         )
@@
 yll_totals <- lapply(io[-2], function(x) {
-  temp <- sapply(1:5,function(y){
+  temp <- sapply(1:NSCEN,function(y){
     xx <- x$outcomes$hb$ylls
-    rowSums(xx[,seq(2 + y, ncol(xx), by = 5)])
+    rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
   })
@@
                               rownames(yll_totals[[x]])),] /
-          t(repmat(io[[x]]$demographic$population,5,1)) * 100000,
+          t(repmat(io[[x]]$demographic$population,NSCEN,1)) * 100000,
         Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),
-                                             length = 5) * 100000))
+                                             length = NSCEN) * 100000))
 names(yll_rates) <- cities
@@
   injury_col <- grep("inj", names(xx))[1] # Dan: This line is mine (not sure is ok)
-  xxx <- sapply(1:5, function(y)
+  xxx <- sapply(1:NSCEN, function(y)
     sapply(sort(unique(xx$age_cat)), function(z)
@@
   xx <- x$outcomes$pathway_hb$ylls
-  xxx <- sapply(1:5, function(y){
-    xxx <- rowSums(xx[, pa_cols[seq(y, length(pa_cols), by = 5)]])
+  xxx <- sapply(1:NSCEN, function(y){
+    xxx <- rowSums(xx[, pa_cols[seq(y, length(pa_cols), by = NSCEN)]])
     sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
@@
   xx <- x$outcomes$pathway_hb$ylls
-  xxx <- sapply(1:5,function(y) {
-    xxx <- rowSums(xx[, ap_cols[seq(y, length(ap_cols), by = 5)]])
+  xxx <- sapply(1:NSCEN,function(y) {
+    xxx <- rowSums(xx[, ap_cols[seq(y, length(ap_cols), by = NSCEN)]])
     sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
@@
   rbind(
-    injury_totals[[x]]/t(repmat(pop_by_age[[x]],5,1)) * 100000,
+    injury_totals[[x]]/t(repmat(pop_by_age[[x]],NSCEN,1)) * 100000,
     Total = colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),
-                                            length = 5) * 100000)
+                                            length = NSCEN) * 100000)
   )
@@
   rbind(
-    pa_totals[[x]] / t(repmat(pop_by_age[[x]], 5, 1)) * 100000,
+    pa_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN, 1)) * 100000,
     Total = colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),
-                                        length = 5) * 100000))
+                                        length = NSCEN) * 100000))
 
@@
   rbind(
-    ap_totals[[x]] / t(repmat(pop_by_age[[x]], 5 ,1)) * 100000,
+    ap_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN ,1)) * 100000,
     Total = colSums(ap_totals[[x]]) / rep(sum(pop_by_age[[x]]),
-                                          length = 5) * 100000))
+                                          length = NSCEN) * 100000))
 names(ap_rates) <- cities
@@
   xx <- x$outcomes$hb$deaths
-  xxx <- sapply(1:5,function(y){
-    colSums(xx[, seq(y + 2, ncol(xx), by = 5)])
+  xxx <- sapply(1:NSCEN,function(y){
+    colSums(xx[, seq(y + 2, ncol(xx), by = NSCEN)])
   })
@@
     Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]), 
-                                               length = 5)*100000))
+                                               length = NSCEN)*100000))
 
@@
   xx <- x$outcomes$hb$ylls
-  xxx <- sapply(1:5,function(y) {
-    colSums(xx[,seq(y + 2, ncol(xx), by = 5)])
+  xxx <- sapply(1:NSCEN,function(y) {
+    colSums(xx[,seq(y + 2, ncol(xx), by = NSCEN)])
   })
@@
     Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]),
-                                               length = 5) * 100000))
+                                               length = NSCEN) * 100000))
 names(disease_rates) <- cities

### 2021-11-29 summary_tables_PA_AP.Rmd
Regen

@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE}
 io <- readRDS("results/multi_city/io.rds")
@@
   # print(cities[x])
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
-                                                        "Bicycling", "Driving",
-                                                        "Motorcycling", 
-                                                        "Public Transport")
+  # names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
+  #                                                       "Bicycling", "Driving",
+  #                                                       "Motorcycling", 
+  #                                                       "Public Transport")
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:9] <- c("Baseline", "Bicycling",
+                                                        "Driving", "Public Transport")
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
-
-for (x in 1: length(cities))
-{
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc, Baseline:`Public Transport`, factor_key=TRUE)
-  summary<- as.data.frame(data_long %>% group_by(scenario) %>% summarise('mean'=mean(pm_conc),'5th'=quantile(pm_conc, 0.05),'50th'=quantile(pm_conc, 0.5),'95th'=quantile(pm_conc, 0.9)))
-  summary<- cbind(summary$scenario ,round(summary[,2:5], digits=1))
-  summary$change_PM<- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits=2)
-  names(summary)[1]<-"Scenario"
-  print(kable(summary, caption= cities[x]))
+data_long <- NA
+summary <- NA
+for (x in 1:length(cities)) {
+  #names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:9] <- c("Baseline", "Bicycling", 
+                                                     "Driving", "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
+                      Baseline:`Public Transport`, factor_key = TRUE)
+  summary <- as.data.frame(data_long %>% group_by(scenario) %>%
+                             summarise('mean' = mean(pm_conc),
+                                       '5th' = quantile(pm_conc, 0.05),
+                                       '50th' = quantile(pm_conc, 0.5),
+                                       '95th' = quantile(pm_conc, 0.9)))
+  summary <- cbind(summary$scenario ,round(summary[,2:5], digits = 1))
+  summary$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits=2)
+  names(summary)[1] <- "Scenario"
+  print(kable(summary, caption = cities[x]))
 
@@
 #pm_conc<- c(50, 18, 122, 47 )
-city<-cities
+city <- cities
 #shares<- as.data.frame(cbind(city,as.numeric(io), pm_conc))
@@
 {
-  modes<-names(unlist(io[[cities[x]]]$PM_emission_inventory))
-  emissions<-as.data.frame(unlist(io[[cities[x]]]$PM_emission_inventory))
-  city_emissions<-cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))
-  names(city_emissions)[2]<- "emissions"
-  select<- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
-  city_emissions$modes<- as.character(city_emissions$modes)
-  city_emissions$modes[!(city_emissions$modes %in% select)]<- "other"
-  summary<-city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
-  names(summary)[2]<- cities[x]
-  summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits=1)
-  summary<-as.data.frame(summary)
-  summary[nrow(summary)+1,1] <- "Transport share"
-  summary[nrow(summary),2]<- as.character(round(pm_trans_share[[cities[x]]] * 100))
-  summary[nrow(summary)+1,1] <- "PM2.5 Conc"
+  modes <- names(unlist(io[[cities[x]]]$PM_emission_inventory))
+  emissions <- as.data.frame(unlist(io[[cities[x]]]$PM_emission_inventory))
+  city_emissions <- cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))
+  names(city_emissions)[2] <- "emissions"
+  select <- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
+  city_emissions$modes <- as.character(city_emissions$modes)
+  city_emissions$modes[!(city_emissions$modes %in% select)] <- "other"
+  summary <- city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
+  names(summary)[2] <- cities[x]
+  summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits = 1)
+  summary <- as.data.frame(summary)
+  summary[nrow(summary) + 1,1] <- "Transport share"
+  summary[nrow(summary),2] <- as.character(round(pm_trans_share[[cities[x]]] * 100))
+  summary[nrow(summary) + 1,1] <- "PM2.5 Conc"
   summary[nrow(summary),2] <- as.character(pm_conc_base[[cities[x]]])
-  io[[cities[x]]]$summary_emission<- summary
+  io[[cities[x]]]$summary_emission <- summary
 
@@
 cl <- list()
-for (city in cities)
-{
+td <- NA
+for (city in cities) {
 
-  td <- round(colSums(io[[city]]$outcomes$co2_conc, na.rm = T), 1) %>% as.data.frame() %>% tibble::rownames_to_column()
+  td <- round(colSums(io[[city]]$outcomes$co2_conc, na.rm = T), 1) %>% 
+    as.data.frame() %>% tibble::rownames_to_column()
   names(td) <- c('Scenario', city)
@@
 
-cl$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+#cl$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+cl$Scenario <- c("Baseline", "Bicycling", "Driving", "Public Transport")
 
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
-limit=100
-for (x in 1: length(cities))
-{
-  names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Public Transport`, factor_key=TRUE)
-  y<-ggplot(data_long, aes(x=scenario, y=mmet, fill=scenario)) + geom_boxplot(outlier.shape = NA)+ ggtitle(cities[x])+ labs(y="Weekly MMET", x = "Scenarios")+  ylim(0, limit)
+limit = 100
+for (x in 1:length(cities)) {
+  #names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
+  names(io[[cities[x]]]$outcomes$mmets)[5:8] <- c("Baseline", "Bicycling", 
+                                                  "Driving", "Public Transport")
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Public Transport`, factor_key = TRUE)
+  y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) +
+    geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 
+    labs(y = "Weekly MMET", x = "Scenarios")+  ylim(0, limit)
   print(y)
@@
 for (x in 1:length(cities)) {
-  names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
+  #names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
+  names(io[[cities[x]]]$outcomes$mmets)[5:8] <- c("baseline","bike_scen", 
+                                                  "car_scen", "bus_scen")
   data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key = TRUE)
@@
 {
-
-  td <- round(colMeans(io[[city]]$outcomes$mmets[,5:10], na.rm = T), 1) %>% as.data.frame() %>% tibble::rownames_to_column()
+  #td <- round(colMeans(io[[city]]$outcomes$mmets[,5:10], na.rm = T), 1) %>%
+  #  as.data.frame() %>% tibble::rownames_to_column()
+  td <- round(colMeans(io[[city]]$outcomes$mmets[,5:8], na.rm = T), 1) %>%
+    as.data.frame() %>% tibble::rownames_to_column()
   names(td) <- c('Scenario', city)
@@
   # PM 2.5 table and plot
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
+  # names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
+  #                                                       "Bicycling", "Driving",
+  #                                                       "Motorcycling", 
+  #                                                       "Public Transport")
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:9] <- c("Baseline",
                                                         "Bicycling", "Driving",
-                                                        "Motorcycling", 
                                                         "Public Transport")
+
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
@@
   names(td) <- c('Scenario', 'CO2')
-  td$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling",
-                   "Public Transport")
+  # td$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling",
+  #                  "Public Transport")
+  td$Scenario <- c("Baseline", "Bicycling", "Driving", "Public Transport")
 
@@
 for (x in 1:length(cities)) {
-  names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("Baseline","Walking",
+  # names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("Baseline","Walking",
+  #                                                  "Bicycling", "Driving",
+  #                                                  "Motorcycling", 
+  #                                                  "Public Transport")
+  names(io[[cities[x]]]$outcomes$mmets)[5:8] <- c("Baseline",
                                                    "Bicycling", "Driving",
-                                                   "Motorcycling", 
                                                    "Public Transport")

### 2021-11-29 injury_tables.Rmd
Regen injuries to get distances

@@
 ```
+
+### True distances in a single file
+```{r }
+true_distances <- list()
+for (city in cities) { # Loop for each city
+  true_distances[[city]] <- io[[city]]$true_dist %>% 
+    pivot_longer(!stage_mode, names_to = "scenario", values_to = "distance") %>% 
+    mutate(city = city)
+}
+true_distances_df <- bind_rows(true_distances)
+readr::write_csv(true_distances_df, 'results/multi_city/whw_matrices/distances.csv')
+```

### 2021-12-08 injury_tables.Rmd
Remove rounding of numbers for WHW and NOV outcomes. It affects the injury risk numbers. Fix #81

@@
       # td1: Number of fatalities in nov
-      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td1 <- io[[city]]$outcomes$whw[[cs]]$nov %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
       # td2: Number of fatalities in whw
-      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums(io[[city]]$outcomes$whw[[cs]]$whw) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
       # td3: Number of fatalities: sum of nov and whw
@@
     }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
-      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- colSums(io[[city]]$outcomes$whw[[cs]]$whw) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
       #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
@@
 
-      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- io[[city]]$outcomes$whw[[cs]]$nov %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
       #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
@@
     if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
-      td1 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td2 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td1 <- (io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td2 <- colSums((io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
       td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
@@
     }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
-      td3 <- colSums(round(io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- colSums((io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
     }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
 
-      td3 <- round(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      td3 <- (io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
 

### 2022-01-19 summary_tables_PA_AP.Rmd
Save AP summary of PM2.5 and CO2 in a csv file. Ref #86

@@
 ```{r message=FALSE, warning=FALSE, echo=FALSE}
+
+l <- list()
 # I have to keep the first chunk of code to run this part
@@
   cat( paste('#### Descriptive table of Individual-level PM2.5 concentrations and CO2 emissions\n' ))
+  l[[cities[x]]] <- left_join(summary2, td, by = "Scenario")
+  l[[cities[x]]]$city <- cities[x]
   print(kable(cbind(summary2, td), caption = cities[x]))
@@
 }
+
+write_csv(bind_rows(l), "results/multi_city/AP/desc_stats.csv")
+
 ```

### 2022-01-19 injury_tables.Rmd
Save injury outputs in excel sheet. Ref #86

@@
 }
-true_distances_df <- bind_rows(true_distances)
+true_distances_df <- bind_rows(true_distances) %>% 
+  mutate(scenario = case_when(
+    scenario == "Scenario 1" ~ "Bicycling", 
+    scenario == "Scenario 2" ~ "Driving",
+    scenario == "Scenario 3" ~ "Public Transport",
+    scenario == "Baseline" ~ "Baseline"))
 readr::write_csv(true_distances_df, 'results/multi_city/whw_matrices/distances.csv')
 ```
+
+### Write all injury tables in a single excel sheet
+```{r}
+
+# Export injury datasets in a single excel sheet
+path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"
+writexl::write_xlsx(list(whw = whw_list %>% filter(str_mode != 'Total'),
+                         injury_risks_per_billion_kms = rd,
+                         injury_risks_per_100k  = injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'),
+                         true_distances = true_distances_df),
+                    path = path_inj_wb)
+
+```

### 2022-01-19 code/producing_trips_rd/trip_pivot_table_san_antonio.R

@@ -1,0 +1,24 @@
+library(rpivotTable)
+library(htmlwidgets)
+library(tidyverse)
+
+io <- readRDS("results/multi_city/io.rds")
+
+tp <- io$san_antonio$trip_scen_sets %>% mutate(scenario = case_when(
+  scenario == "Scenario 1" ~ "CYC_SC",
+  scenario == "Scenario 2" ~ "CAR_SC",
+  scenario == "Scenario 3" ~ "BUS_SC",
+  scenario == "Baseline" ~ "Baseline")) 
+
+write_csv(tp, "results/multi_city/scen_distr/san_antonio_trips.csv")
+
+tp %>% rpivotTable(
+    io$san_antonio$trip_scen_sets,
+    rows = c("sex", "age_cat"),
+    cols = c("trip_mode","stage_mode", "scenario"),
+    aggregatorName = "Sum",
+    inclusions = list( Survived = list("Yes")),
+    exclusions= list(trip_mode = list("truck"), scenario = list("CYC_SC", "CAR_SC")),
+    vals = "stage_distance",
+    rendererName = "Table Barchart", subtotals = FALSE
+  ) %>% saveWidget(file = "results/multi_city/scen_distr/san_antonio.html", selfcontained = T)

### 2022-01-19 summary_tables.Rmd
Save both as CSV and excel sheet - health outcomes. Ref #86

@@
 library(data.table)
+library(writexl)
 })
@@
 
+# # Export deaths HB
+# path_deaths_wb <- "results/multi_city/health_impacts/Deaths.xlsx"
+# deaths_wb <- readWorkbook(path_deaths_wb)
+# wb <-  loadWorkbook(path_deaths_wb)
+# removeTable(wb = wb, sheet = "Deaths", table = "Tabla1")
+# writeDataTable(wb, sheet = "Deaths", x = rbindlist(health$deaths$hb, use.names = T), tableName = "Tabla1", startCol = 1)
+# removeTable(wb = wb, sheet = "Deaths_Pathway", table = "Tabla2")
+# writeDataTable(wb, sheet = "Deaths_Pathway", x = rbindlist(health$deaths$pathway_hb, use.names = T), tableName = "Tabla2", startCol = 1)
+# saveWorkbook(wb, path_deaths_wb, overwrite =TRUE)
+# 
+# # Export ylls HB
+# path_ylls_wb <- "results/multi_city/health_impacts/YLLs.xlsx"
+# ylls_wb <- readWorkbook(path_ylls_wb)
+# wb <-  loadWorkbook(path_ylls_wb)
+# removeTable(wb = wb, sheet = "YLLs", table = "Tabla1")
+# writeDataTable(wb, sheet = "YLLs", x = rbindlist(health$ylls$hb, use.names = T), tableName = "Tabla1", startCol = 1)
+# removeTable(wb = wb, sheet = "YLLs_Pathway", table = "Tabla2")
+# writeDataTable(wb, sheet = "YLLs_Pathway", x = rbindlist(health$ylls$pathway_hb, use.names = T), tableName = "Tabla2", startCol = 1)
+# saveWorkbook(wb, path_ylls_wb, overwrite =TRUE)
+
+# # Export ylls HB
+output_file <- "results/multi_city/health_impacts/data.xlsx"
+writexl::write_xlsx(list(ylls_hb_lng = rbindlist(health$ylls$hb, use.names = T),
+                         ylls_pathway_hb_lng = rbindlist(health$ylls$pathway_hb, use.names = T),
+                         Deaths = rbindlist(health$deaths$hb, use.names = T),
+                         Deaths_Pathway = rbindlist(health$deaths$pathway_hb, use.names = T)), path = output_file)
+
 # Export ylls HB

### 2022-01-19 injury_tables.Rmd
Add missing library

@@
 library(tidyverse)
+library(writexl)
+
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')

### 2022-01-19 latex/output-md.Rmd
Update text

@@
 
-The relationships comes out of our work @Leandro2018, where we look at the association of `physical activity` with various diseases.
+The relationships comes out of our work on meta-analyses \@Leandro2018, where we look at the association of `physical activity` with various diseases.
 
@@
 
-We resample the adjusted probability parameters of Table [6](#calculations){reference-type="ref" reference="calculations"}.
-We describe each parameter with a Beta distribution as follows: $$\begin{aligned}
+We resample the adjusted probability parameters of Table [6](#calculations){reference-type="ref" reference="calculations"}. We describe each parameter with a Beta distribution as follows:
+
+\$$\hat{B}_{g,m} \tilda L_m{B}_{a,g,m}.$$
+
+Ba,g,m∼Beta(αa,g,m,βa,g,m);
+
 \label{pitilde}
-\tilde{B}_{a,g,m} & \sim \text{Beta}(\alpha_{a,g,m},\beta_{a,g,m}); \\
+
+\$$\hat{B}_{a,g,m} & \sim \text{Beta}(\alpha_{a,g,m},\beta_{a,g,m}); \\$$
+
+${\hat{B}}_{a,g,m} \sim Beta({\alpha}_{a,g,m}, {\beta}_{a,g,m})$
+${\alpha}_{a,g,m} + {\beta}_{a,g,m} = 200$
+$\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}$
+
 \alpha_{a,g,m}+\beta_{a,g,m} & = 200; \\

### 2022-01-19 summary_tables_PA_AP.Rmd
Create excel sheet. Ref #86

@@
 library(knitr)     
-library(tidyr)
-library(dplyr)
+library(tidyverse)
 #library(plotly)
@@
 
+output_file <- "results/multi_city/AP/desc_stats.xlsx"
+writexl::write_xlsx(list(PM_2.5_emission_inventory = sl,
+                         CO2_emissions = cl,
+                         summary_stats_PM2.5_CO2 = bind_rows(l)), path = output_file)
+
 write_csv(bind_rows(l), "results/multi_city/AP/desc_stats.csv")

### 2022-04-22 summary_tables.Rmd
Calculate pa distribution

@@
 ```
+
+## PA dist
+```{r}
+
+### pa distr
+cities <- names(io)[!names(io) %in% 'scen_prop']
+
+l <- list()
+
+for (city in cities){
+
+  # city <- "antofagasta"
+
+  td <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
+    rename("Baseline" = base_mmet, "CYC_SC" = scen1_mmet,
+           "CAR_SC" = scen2_mmet,  "BUS_SC" = scen3_mmet) %>% 
+    summarise(across(, ~ quantile(.x, c(0.05,0.10,0.25, 0.375, 0.50,0.75,0.90), na.rm = T)), 
+              prob = c(0.05,0.10,0.25, 0.375, 0.50,0.75,0.90)) %>%
+    pivot_longer(cols = -c(prob)) %>% mutate(city = city)
+
+  td1 <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
+    rename("Baseline" = base_mmet, "CYC_SC" = scen1_mmet,
+           "CAR_SC" = scen2_mmet,  "BUS_SC" = scen3_mmet) %>% 
+    summarise(across(, ~ mean(.x))) %>%  t() %>% as.data.frame() %>% 
+    tibble::rownames_to_column("name") %>% 
+    rename(value = V1) %>% 
+    mutate(prob = "mean") %>% mutate(city = city)
+
+  l[[city]] <- plyr::rbind.fill(td, td1)
+
+}
+
+write_csv(data.table::rbindlist(l), "results/multi_city/PA/pa_distr.csv")
+```

### 2022-05-16 summary_tables_PA_AP.Rmd
Fix overwritten object and regen

@@
 
-cl <- list()
+co2_conc <- list()
 td <- NA
@@
 
-  if (length(cl) == 0)
-    cl <- td
+  if (length(co2_conc) == 0)
+    co2_conc <- td
   else
-    cl <- left_join(cl, td)
+    co2_conc <- left_join(co2_conc, td)
 
@@
 #cl$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-cl$Scenario <- c("Baseline", "Bicycling", "Driving", "Public Transport")
+co2_conc$Scenario <- c("Baseline", "Bicycling", "Driving", "Public Transport")
 
-print(kable(cl))
+print(kable(co2_conc))
 
@@
 writexl::write_xlsx(list(PM_2.5_emission_inventory = sl,
-                         CO2_emissions = cl,
+                         CO2_conc = co2_conc,
                          summary_stats_PM2.5_CO2 = bind_rows(l)), path = output_file)

### 2022-05-26 injury_tables.Rmd
Update injury outputs catering for confidence interval. Re #100

@@
 cities <- names(io)[!names(io) %in% 'scen_prop']
+
+for (city in cities){
+  io[[city]][["trip_scen_sets"]] <- NULL
+  io[[city]][["inj_distances"]] <- NULL
+  io[[city]][["synth_pop"]] <- NULL
+
+  io[[city]]$outcomes$mmets <- io[[city]]$outcomes$pm_conc_pp <- NULL
+}
 
@@
 
+city_df <- list()
+
 for (city in cities) {
+
+  # city <- cities[1]
+
   cat('\n')
@@
 
-    for (cs in names(io[[1]]$outcomes$whw)){
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$Baseline), value = T), 2, sep = "_"))){
+    # pref <- "ub"
+    pref_name <- ""
+    if (!is.na(pref))
+      pref_name <- paste0("_", pref)
 
-      if (is.null(io[[city]]$outcomes$whw[[cs]]$whw)){
-        if (!is.null(io[[city]]$outcomes$whw[[cs]]$nov)){
-          td2 <- t(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame()
+    print(pref)
+
+    whw_qn <- paste0("whw", pref_name)
+    nov_qn <- paste0("nov", pref_name)
+
+    for (cs in names(io[[city]]$outcomes$whw)){
+      # cs <- 'Baseline'
+
+      if (is.null(io[[city]]$outcomes$whw[[cs]][[whw_qn]])){
+        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){
+          td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()
           td2$mode <- 'NOV'
@@
       }else{
-        td1 <- (io[[city]]$outcomes$whw[[cs]]$whw) %>% as.data.frame() %>% tibble::rownames_to_column("mode")
+        td1 <- (io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% tibble::rownames_to_column("mode")
         td3 <- td1
-        if (!is.null(io[[city]]$outcomes$whw[[cs]]$nov)){
-          td2 <- t(io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame()
+        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){
+          td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()
           td2$mode <- 'NOV'
@@
       var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
-
-      # scen_names <- c('Walking', 'Bicycling', 'Driving', 'Public Transport', 'Motorcycling')
 
@@
 
-      cat('####', qualified_scen_name, ' WHW', '\n')
+      cat('####', qualified_scen_name, ' ', ifelse(is.na(pref), "mean", pref), ' WHW', '\n')
 
@@
       names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+      td$measure <- ifelse(is.na(pref), "mean", pref)
 
-      all_whw[[var]][[city]] <- td
+      city_df[[var]][[city]][[unique(td$measure)]] <- td
+
+      if (length(city_df[[var]][[city]]) == 3)
+        all_whw[[var]][[city]] <- data.table::rbindlist(city_df[[var]][[city]])
+    }
+
   }
@@
 ```{r whw_for_all_cs, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
-
 st <- list()
-
 file_list <- list()
-
 td <- NULL
-
 whw_list <- list()
-
 for (cs in names(all_whw)){
@@
 
-  td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = "mode") %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))
+  td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = c("mode", "measure")) %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))
 
@@
   whw_lng <- cbind(whw_lng, col_split)
-  names(whw_lng)[4] <- 'strike_mode'
-  names(whw_lng)[5] <- 'city'
+  names(whw_lng)[5] <- 'strike_mode'
+  names(whw_lng)[6] <- 'city'
   whw_lng <- whw_lng %>% dplyr::select(-variable)
@@
 
-  f <- ggplot(data = whw_lng) +
-    aes(x = str_mode, fill = cas_mode, weight = value) +
-    geom_bar(position = "stack") +
-    scale_fill_brewer(palette = "Dark2") +
-    labs(x = "Strike Mode", y = "# of incidents", title = paste("WHW - ", qualified_scen_name, scen)) +
-    coord_flip() +
-    facet_wrap(vars(city), scales = "free_y")
-
-  file_list[[cs]] <- f
+  # f <- ggplot(data = whw_lng) +
+  #   aes(x = str_mode, fill = cas_mode, weight = value) +
+  #   geom_bar(position = "stack") +
+  #   scale_fill_brewer(palette = "Dark2") +
+  #   labs(x = "Strike Mode", y = "# of incidents", title = paste("WHW - ", qualified_scen_name, scen)) +
+  #   coord_flip() +
+  #   facet_wrap(vars(city), scales = "free_y")
+  # 
+  # file_list[[cs]] <- f
 
@@
 }
-
 # Remove total from whw long form, and write as csv
 readr::write_csv(whw_list %>% filter(str_mode != 'Total'), paste0('results/multi_city/whw_matrices/whw_lng.csv'))
-
 ```
@@
 ```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
-
 list_temp <- list()
 index <- 1
-
 scen_names <- unique(whw_list$scenario)
@@
   for (j in 1:length(cas_mode)){
-
     cm <- cas_mode[j]
@@
 }
-
 ```
-
-## Who hit whom matrix for all case cities as CSV {#whw_all_csv}
-
-[Download from](https://raw.githubusercontent.com/ITHIM/ITHIM-R/master/results/multi_city/whw_matrices/whw_lng.csv)
 
@@
 overall_dd <- list() # To save all cities results
+
+city_df <- list()
 for (city in cities){
 
-  #city <- 'delhi'
-
-  # print(city)
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$Baseline), value = T), 2, sep = "_"))){
 
     el <- list() # To save city specific results
+    # pref <- "ub"
+    pref_name <- ""
+    if (!is.na(pref))
+      pref_name <- paste0("_", pref)
 
-  #if (!city %in% c('buenos_aires', 'mexico_city', 'bogota', 'vizag')){
+    print(pref)
+
+    whw_qn <- paste0("whw", pref_name)
+    nov_qn <- paste0("nov", pref_name)
 
     for (cs in names(io[[city]]$outcomes$whw)){
-    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
+      # cs <- 'Baseline'
+      if (length(word(names(io[[city]]$outcomes$whw$Baseline), 1, sep = "_") %>% unique()) == 2){
         # td1: Number of fatalities in nov
-      td1 <- io[[city]]$outcomes$whw[[cs]]$nov %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td1 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
         # td2: Number of fatalities in whw
-      td2 <- colSums(io[[city]]$outcomes$whw[[cs]]$whw) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td2 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
         # td3: Number of fatalities: sum of nov and whw
@@
       }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
-      td3 <- colSums(io[[city]]$outcomes$whw[[cs]]$whw) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
         #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
@@
 
-      td3 <- io[[city]]$outcomes$whw[[cs]]$nov %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
         #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
@@
 
-    el <- full_join(el, td4, by = 'stage_mode')
+      el <- full_join(el, td4, by = c('stage_mode'))
 
@@
     }
-  print(kable(el, caption = city))
+
+    print(kable(el, caption = paste(city, ifelse(is.na(pref), "mean", pref))))
 
@@
     names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+    td$measure <- ifelse(is.na(pref), "mean", pref)
+    dd$measure <- ifelse(is.na(pref), "mean", pref)
 
-  overall_el[[city]] <- td
-  overall_dd[[city]] <- dd
+    city_df[["td"]][[city]][[unique(td$measure)]] <- td
+    city_df[["dd"]][[city]][[unique(td$measure)]] <- dd
+
+    if (length(city_df[["td"]][[city]]) == 3){
+      overall_el[[city]] <- data.table::rbindlist(city_df[["td"]][[city]])
+      overall_dd[[city]] <- data.table::rbindlist(city_df[["dd"]][[city]])
+    }
+
     cat("\n")
+  }
 }
@@
 require(tibble)
-
-td <- overall_el %>% purrr::reduce(full_join, by = "stage_mode") %>% as.data.frame()
+td <- overall_el %>% purrr::reduce(full_join, by = c("stage_mode", "measure")) %>% as.data.frame()
 td[is.na(td)] <- 0
@@
 colnames(td) = gsub("w_sc", "wsc", colnames(td))
-
 injury_risks_b <- td
@@
 # injury_risks_b[, 2:ncol(injury_risks_b)] <- injury_risks_b[, 2:ncol(injury_risks_b)] * 10000
-
 #readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')
@@
 injury_risks_lng <- cbind(injury_risks_lng, col_split)
-names(injury_risks_lng)[4] <- 'scenario'
-names(injury_risks_lng)[5] <- 'city'
+names(injury_risks_lng)[5] <- 'scenario'
+names(injury_risks_lng)[6] <- 'city'
 injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
@@
 injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Driving'
-#injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-#injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
-
 rd <- rename(injury_risks_lng, mode = stage_mode)
 rd <- rd %>% filter(mode != 'Total')
-
 readr::write_csv(rd, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms_lng.csv')
-
-temp <- injury_risks_lng %>% group_by(stage_mode, scenario) %>% summarise(mv = mean(value), sd = sd(value))
-# injury_risks_lng <- left_join(injury_risks_lng, td, by = c('stage_mode', 'scenario'))
-
-cols_cbrewer <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#c51b7d','#ff7f00','#cab2d6','#6a3d9a','#543005','#b15928')
-
-d <- ggplot(injury_risks_lng) +
-  aes(x = stage_mode, fill = scenario, weight = value) +
-  geom_bar(position = "dodge", colour = 'black') +
-  scale_color_manual(values = cols) +
-  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
-  coord_flip() +
-  theme_minimal() +
-  facet_wrap(vars(city)) +
-  labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) 
-df <- injury_risks_lng %>% spread(value = value, key = city) 
-df[df == 0] <- NA
-df <- df %>% mutate(mv = rowMeans(dplyr::select(., -c(stage_mode, scenario)))) %>% reshape2::melt()
-df1 <- df %>% dplyr::filter(variable == 'mv')
-df <- df %>% dplyr::filter(variable != 'mv')
-d1 <- ggplot(df, aes(x = stage_mode, fill = variable, weight = value)) +
-   geom_bar(position = "dodge", colour = 'black') +
-   scale_fill_manual(values = cols_cbrewer) +
-   theme_minimal() + 
-   geom_point(data = df1, aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
-   facet_wrap(vars(scenario)) + 
-   #scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 25)) +
-   labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
-   theme(axis.text.x = element_text(face="bold", color="#993333", size=8, angle=90))
-d2 <- ggplot(df %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')), aes(x = stage_mode, fill = variable, weight = value)) +
-    geom_bar(position = "dodge", colour = 'black') +
-    scale_fill_manual(values = cols_cbrewer) +
-    theme_minimal() + 
-    geom_point(data = df1 %>% filter(!stage_mode %in% c('auto_rickshaw', 'bus', 'truck')) , aes(x = stage_mode, fill = variable, y = value), colour = "black") + 
-    facet_grid(vars(scenario), scales='free_x') + 
-    # scale_y_continuous(breaks = seq(0, max(df$value, na.rm = T), 50)) +
-    labs(x = "Strike Mode", y = "# of incidents", title = paste("Injury risk per billion km")) +
-    theme(axis.text.x = element_text(face="bold", color="#993333", size=14))
-# ggsave(d, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km.png"), limitsize = F, dpi = 300, scale = 1.5)
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km.html")
-# htmlwidgets::saveWidget(plotly::ggplotly(d), file.path(normalizePath(dirname(fp)), basename(fp)))
-# ggsave(d1, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_mode.png"),  width = 8, height = 5, dpi = 300, units = "in", scale = 2, device='png')
- fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_mode.html")
- # htmlwidgets::saveWidget(plotly::ggplotly(d1), file.path(normalizePath(dirname(fp)), basename(fp)))
-# print(kable(format(injury_risks_b, scientific = F), caption = paste('Injury per billion km across all cities')))
-# ggsave(d2, file = paste0("results/multi_city/whw_matrices/plots/injury_risk_billion_km_by_specific_mode.png"), limitsize = F, dpi = 300, scale = 1.5)
-fp <- paste0("results/multi_city/whw_matrices/interactive_plots/injury_risk_billion_km_by_specific_mode.html")
-# htmlwidgets::saveWidget(plotly::ggplotly(d2), file.path(normalizePath(dirname(fp)), basename(fp)))
 ```
@@
 overall_el_normalized <- list()
+city_df <- list()
+
 for (city in cities){
@@
 
-  el <- list()
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$Baseline), value = T), 2, sep = "_"))){
+
+    el <- list() # To save city specific results
+    pref_name <- ""
+    if (!is.na(pref))
+      pref_name <- paste0("_", pref)
+
+    print(pref)
+
+    whw_qn <- paste0("whw", pref_name)
+    nov_qn <- paste0("nov", pref_name)
 
     for (cs in names(io[[city]]$outcomes$whw)){
-    if (length(names(io[[city]]$outcomes$whw$Baseline)) == 2){
-      td1 <- (io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      td2 <- colSums((io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+      if (length(word(names(io[[city]]$outcomes$whw$Baseline), 1, sep = "_") %>% unique()) == 2){
+        td1 <- (io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td2 <- colSums((io[[city]]$outcomes$whw[[cs]][[whw_qn]])) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
         td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
@@
       }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
-      td3 <- colSums((io[[city]]$outcomes$whw[[cs]]$whw)) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3 <- colSums((io[[city]]$outcomes$whw[[cs]][[whw_qn]])) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
       }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
 
-      td3 <- (io[[city]]$outcomes$whw[[cs]]$nov) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        td3 <- (io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
 
@@
     names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+    td$measure <- ifelse(is.na(pref), "mean", pref)
 
-  overall_el_normalized[[city]] <- td
+    city_df[["td"]][[city]][[unique(td$measure)]] <- td
+
+    if (length(city_df[["td"]][[city]]) == 3){
+      overall_el_normalized[[city]] <- data.table::rbindlist(city_df[["td"]][[city]])
+    }
 
     cat("\n")
+  }
 
@@
 
-td <- overall_el_normalized %>% purrr::reduce(full_join, by = "stage_mode") %>% as.data.frame()
+td <- overall_el_normalized %>% purrr::reduce(full_join, by = c("stage_mode", "measure")) %>% as.data.frame()
 td[is.na(td)] <- 0
 td <- td %>% dplyr::select(stage_mode, sort(names(.)))
-
 td <- rename(td, mode = stage_mode)
@@
 injury_risks_per_100k <- cbind(injury_risks_per_100k, col_split)
-names(injury_risks_per_100k)[4] <- 'scenario'
-names(injury_risks_per_100k)[5] <- 'city'
+names(injury_risks_per_100k)[5] <- 'scenario'
+names(injury_risks_per_100k)[6] <- 'city'
 injury_risks_per_100k <- injury_risks_per_100k %>% dplyr::select(-variable)
@@
 #injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
-
 readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/whw_matrices/injury_risks_per_100k_pop.csv')
-
 ```
@@
 ```{r}
-
 # Export injury datasets in a single excel sheet
@@
                     path = path_inj_wb)
-
 ```

### 2022-06-22 quarto-doc/pkg_doc.qmd
Add quarto doc and its associated files. Re #97

--- 2022-01-19 latex/output-md.Rmd
@@
 ---
-title: "ITHIM"
-author: "Rob Johnson, Ali Abbas"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+title: "Package Documentation"
+format:
+  html:
+    light: flatly
+    dark: darkly
+    toc: true
 
-bibliography: ithim.bib
+reference-location: margin
+citation-location: margin
 
-output:
-  rmdformats::downcute:
-    self_contained: true
-    thumbnails: true
-    lightbox: true
-    gallery: false
-    highlight: tango
+number-sections: true
+crossref:
+  chapters: true
+
 ---
 
+```{r, echo=FALSE}
+library(knitr)
+
+io <- readRDS("../results/multi_city/io.rds")
+cities <- names(io)[!names(io) %in% 'scen_prop']
+scen_def <- io$scen_prop
+```
+
 # Introduction
-
-## Summary of literature
-
-## Summary of aims
 
@@
 The ITHIM as described here is implemented in the ITHIM-R R package (<https://github.com/ITHIM/ITHIM-R>).
-Other cities are also included in that package (São Paulo, Delhi, Bangalore), and reference will be made to those where relevant.
+In the Latin America settings, we have a list of 20 cities, which are (Antofagasta, Arica, Belo Horizonte, Bogota, Buenos Aires, Cali, Copiapo, Coquimbo Laserena, Gran Valparaiso, Iquique Altohospicio, Medellin, Mexico City, Montevideo, Osorno, Puerto Montt, San Antonio, Santiago, Sao Paulo, Temuco Padrelascasas, Valdivia), and reference will be made to those where relevant.
 
@@
 
-!\[ITHIM workflow.
+![ITHIM workflow.
 Inputs to the model have a bold, black outline.
@@
 GBD: global burden of disease.
-BOD:
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("flowchart.pdf")
-```
+BOD: Burden of Disease](figures/flowchart.png "Flowchart: Model descrption"){fig-alt="A flowchart of model."}
 
@@
 
-::: center
-::: {#indices}
++---------------+----------------------+----------------------+
 | Label         | Name                 | Levels               |
-|:------|:---------------------|:--------------------|
-| $a$   | Age group            | 15--49              |
++===============+======================+======================+
+| *a*           | Age group            | 15--49\              |
 |               |                      | 50--69               |
-| $g$   | Gender               | Male                |
++---------------+----------------------+----------------------+
+| *g*           |                      | Male\                |
 |               |                      | Female               |
-| $h$   | Disease              |                     |
-| $i$   | Individual index     |                     |
-| $j$   | Trip index           |                     |
-| $m$   | Transport mode       | walk                |
-|       |                      | cycling (cyc.)      |
-|       |                      | bus                 |
-|       |                      | car                 |
-|       |                      | motorbike (mot.)    |
-|       |                      | goods vehicles (GV) |
++---------------+----------------------+----------------------+
+| *h*           | Disease              |                      |
++---------------+----------------------+----------------------+
+| *i*           | Individual index     |                      |
++---------------+----------------------+----------------------+
+| *j*           | Trip index           |                      |
++---------------+----------------------+----------------------+
+| *m*           | Transport mode       | walk\                |
+|               |                      | cycling (cyc.)\      |
+|               |                      | bus\                 |
+|               |                      | car\                 |
+|               |                      | motorbike (mot.)\    |
+|               |                      | goods vehicles (GV)\ |
 |               |                      | subway (sub.)        |
-| $o$   | Outcome              | Death               |
++---------------+----------------------+----------------------+
+| *o*           | Outcome              | Death\               |
 |               |                      | YLL                  |
-| $s$   | Scenario             |                     |
-| $w$   | AP RR hyperparameter | $1,2,3,4$           |
-| $x$   | PA RR hyperparameter | $1,2,3$             |
-| $z$   | PA dose              | $0,...,300$         |
++---------------+----------------------+----------------------+
+| *s*           | Scenario             | 3                    |
++---------------+----------------------+----------------------+
+| *w*           | AP RR hyperparameter | 1, 2, 3, 4           |
++---------------+----------------------+----------------------+
+| *x*           | PA RR hyperparameter | 1, 2, 3              |
++---------------+----------------------+----------------------+
+| *z*           | PA dose              |                      |
++---------------+----------------------+----------------------+
 
-: Indices used in the ITHIM variables. The "Label" is the subscript used to refer to the index in this document. The "Name" is descriptive. The "Levels" are the levels used in the Accra model.
-:::
-:::
-
-::: center
-::: {#abbrev}
 | Abbreviation | Meaning                                       |
@@
 
-: Common abbreviations.
-:::
-:::
+: Indices used in the ITHIM variables. The "Label" is the subscript used to refer to the index in this document. The "Name" is descriptive. The "Levels" are the levels used in the Accra model.
 
-::: landscape
-::: center
-::: {#inputs}
+```{r}
+
+```
+
+{#inputs}
+
 +-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
@@
 +-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| -   Pollution model                       |                                                                 |                                    |                               |                    |
-|                                           |                                                                 |                                    |                               |                    |
-|     values\*                              |                                                                 |                                    |                               |                    |
+| *Pollution model values*                  |                                                                 |                                    |                               |                    |
 +-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
@@
 +-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| -   Pollution & health model              |                                                                 |                                    |                               |                    |
-|                                           |                                                                 |                                    |                               |                    |
-|     values\*                              |                                                                 |                                    |                               |                    |
+| *Pollution & health model values*         |                                                                 |                                    |                               |                    |
 +-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
@@
 
-: All values input to the ITHIM model.
-:::
-:::
-:::
-
 ## Data module {#data}
@@
 swapping the mode of travel for particular trips and/or particular people).
-For the purposes of illustration, we choose a simple scenario for Accra in which every person gains one one-kilometre walk.
+For the purposes of illustration, we have created three scenarios mentioned in the table @tbl-scen-specs, where each mode such as `cycle` has propensities in three different distance categories: `0-2 km`, `2-6 km` and `6+ km`.
+What the scenario does that in each of those distance categories, it increases the current mode share by the propensities.
+So for instance, if in a city the baseline mode share of `cycle` is `10` percent for short trips (`0-2 km`), then in the `cycle` scenario, it increases upto 15.6 (10 + 5.6).
+
+```{r}
+#| label: tbl-scen-specs
+#| tbl-cap: Scenario propensities.
+#| warning: false
+
+kable(scen_def |> tibble::rownames_to_column("Scenarios"))
+```
 
@@
 
-```{r out.height = "800px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("cancer_PA_dose_response.pdf")
-```
-
 ## Injury module
@@
 Non-travel PA, injury reporting rate and NCD burden all act as scalars for the corresponding datasets.
-Note that the non-travel PA scalar does not affect the $\approx$<!-- -->40% of the population whose non-travel PA is 0.
-
-```{r out.height = "460px", out.width='800px', fig.align='center', fg.cap = 'Distributions of univariate uncertain parameters (1024 samples).', echo=F}
-knitr::include_graphics("uncertain_parameters.pdf")
-```
+Note that the non-travel PA scalar does not affect the $\approx$\<40% of the population whose non-travel PA is 0.
 
@@
 
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("prob_zero_PA.pdf")
-```
-
-(){#prob_zero_PA width="50%"}
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-"Distributions of Dirichlet-distributed emission inventories for four modes and varying confidence values. Raw input values were 4, 4, 20, and 60."
-
-knitr::include_graphics("emission_dist.pdf")
-```
-
 (){#emission_dist width="65%"}
@@
 See Figure [6](#cdf){reference-type="ref" reference="cdf"} for an illustration.
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F, fig.show='hold', fig.cap="We use the same uniform samples to generate random variables from different distributions.
-It means that the samples are perfectly correlated."}
-knitr::include_graphics("DR_transform.pdf")
-```
 
@@
 
-```{r out.height = "600px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("accra_walk_yll.pdf")
-```
-
 ## Value of information
@@
 \[EVPPI for Accra's "walking" scenario for all causes of YLLs excluding "all cause" and "neoplasm".\]
-
-```{r out.height = "460px", out.width='800px', fig.align='center', fig.cap="EVPPI for Accra",fig.path="accra_walk_evppi.pdf", echo=F}
-#knitr::include_graphics("accra_walk_evppi.pdf")
-```
 
@@
 
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("ithim_with_parameters.pdf")
-```
-
 (){#ithim_with_parameters width="\\textwidth"}
@@
 \[1000 samples from PA dose--response curves\]
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("PA_dose_response.pdf")
-```
 
@@
 (I'd like to find an improved notation for this.)
-:::
-
-::: {#calculationstab}
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| Name                                                                                                                                                         | Description                                                                                    | Label                                        |
-+:=============================================================================================================================================================+:===============================================================================================+:=============================================+
-| *Scenario generation*                                                                                                                                        |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| ...
-|                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| *Travel calculations*                                                                                                                                        |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\{T_{j^+}\leftarrow \epsilon; m(j^+)\leftarrow\text{walk} \}\, \forall\, T_j:m(j)=\text{bus}$                                                               | New walk trips to bus                                                                          | walk *to_bus_and* combine_scen               |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\bar{T}_{a,g,m,s}=\sum_{j:a(j )= a,g(j)=g,m(j)=m,s(j)=s}Q_{i(j)}T_j$                                                                                        | (Weighted) total travel per age, gender, mode and scenario                                     |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| \$ \che ck{T}\_{a,g,m,s}=\frac{\bar{T}_{a,                                                                                                                   | Relative travel per mode and scenario by age and gender                                        | inj_distances\[\[1\]\]                       |
-| g, m,s}}{\sum_{a,g}\bar{T}_{a,g,m,s}}\$                                                                                                                      |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\hat{ T} _{m,s}=\sum_{a,g}\bar{T}_{a,g,m,s}$                                                                                                                | Total travel per mode and scenario                                                             | dist                                         |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{T}_{m,s}=\frac{\hat{T}_{ m, s}}{\hat{T}_{m,s=\text{baseline}}}$                                                                                      | Total travel per mode and scenario relative to baseline, with car modes summed, and Tuktuk = 1 | inj_distances\[\[2\]\]                       |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| *Injury calculations*                                                                                                                                        |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $E_{m_{\text{cas}}}=\psi\omega$                                                                                                                              | Casualty exponent                                                                              | CAS_EXPONENT                                 |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $E_{m_{\t ex t{str}}}=\omega-E_{m_{\text{cas}}}$                                                                                                             | Striker exponent                                                                               | STR_EXPONENT                                 |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{I}_{a,g,m_{\text{vic}},m_{\text{str}},s}=\mathcal{S}\left(\sigma\c dot {I}_{a,g,m_{\text{vic}},m_{\text{| Fataliti                                   | es in scenarios by strike mode \| injuries                                                     |                                              |
-| str}}},\bar{T}_{a,g,m,s},E_{m_{\t ex t{cas}}},E_{m_{\text{str}}}\right)$                                                                                     |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\check{I}_{a,g,m,o=\text{death                                                                                                                              | Fatalities in scenarios                                                                        | v ictim_deaths                               |
-| },s} =\sum_{m_{\text{str}}}\tilde{I}_{ a, g,m_{\text{vic}},m_{\text{str}},s}$                                                                                |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $R_{a,g}=U_{a,g,h=\text{road                                                                                                                                 | Road YLL-to-death ratio                                                                        | GBD_INJ_YLL                                  |
-| injuries},o=\text{YLL}}/U_{a,g,h= \t ext{road injuries},o=\text{death}}$                                                                                     |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\check{ I}_{ a,g,m,o=\text{YLL},s}=\check{I}_{ a, g,m,o=\text{death},s}\cdot R_{a,g}$                                                                       | YLLs scaled from fatalities                                                                    | deaths\_ yll_injuries                        |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| *Pollution calculations*                                                                                                                                     |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\dot{P}_{m}=\left\{\begin{array}{lr} \gam ma_m & P=1 \\ \tilde{\gamma}_m, \  | tilde{\gamma}_m \sim \text{Dir}(f(P,\  | gamma_m)) & P<1 \end{array}\right.$ | Emissions by mode \| \|                                                                        | trans_emissions                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{P}_{m,s}=\dot{P}_m\hat{T} _{ m,s}/\hat{T}_{m,s=\text{baseline}}$                                                                                     | Emissions by mode in scenario relative to baseline                                             | trans_emissions                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| \$ \h at{P}*{s}=*\sum{m} \tilde{P}\_{m,s}\$                                                                                                                  | Emissions scalar for scenarios                                                                 | baseline_sum                                 |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\bar {P}_s=\eta(\zeta\hat{P}_s + 1-\zeta)$ \|                                                                                                               | Background PM2.5 concentration in scenarios \|                                                 | scenario_pm                                  |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $K_s=C_4-C_5\log(\bar{P_s})$                                                                                                                                 | On-road PM2.5 exposure ratio                                                                   | on_road_off_road_ratio                       |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{K}_{m\                                                                                                                                               | On-road PM2.5 exposure ratio for open vehicle                                                  | on_road_off_road_ratio                       |
-| in\{\text{walk,cyc.,mot.}\},s}=K_s$                                                                                                                          |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\                                                                                                                                                           | Subway PM2.5 exposure ratio \|                                                                 | subway_ratio                                 |
-| tilde{K}_{m\in\{\text{sub.}\},s}=C_6$ \|                                                                                                                     |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\                                                                                                                                                           | In-vehicle PM2.5 exposure ratio                                                                | in_vehicle_ratio                             |
-| tilde{K}_{m\not\in\{\text{walk,cyc.
-|                                                                                                |                                              |
-| ,mot.,sub.}\},s}=C_2C_3+K_s(1-C_3)$                                                                                                                          |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $V_m=C_1+\frac{1}{2}\lambda_m$                                                                                                                               | Ventilation rates                                                                              | vent_rates                                   |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde {V}_{i,m,s} = \sum_{j:i(j)=i,m(j) =m ,s(j)=s}T_j\cdot V_{m(j)}/Z_{m(j)}$ \|                                                                          | In-travel air inhaled \|                                                                       |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\bar{V}_{i,s} = \sum_{m}\t il de{V}_{i,m,s}\cdot \tilde{K}_{m,s}$                                                                                           | In-travel PM2.5 inhalation rate                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| \$ \hat{V}*{i,s} = C_1*\left(24-\sum{j                                                                                                                       | Non-travel air inhaled                                                                         |                                              |
-| :i(j)=i,s(j)=s}T_j/Z\_{m(j)}\right)\$                                                                                                                        |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| \$\check{W}*{i,s} =* \bar{ P}*s(*\bar{V}{i,s}+\hat{V}{i,s})/24                                                                                               | Total PM2.5 inhaled per hour                                                                   |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| %/(\sum\*m\                                                                                                                                                  |                                                                                                |                                              |
-| \*\tilde{V}{i,m,s}+\hat{V}\_{i,s})\$                                                                                                                         |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{H}_{a,h,w=3}=\text{CDF}_{H_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ \|                                                                              | AP dose-- response--curve parameter \|                                                         | DR_AP_LIST                                   |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{H}_{a,h,w=3} =\te xt{CDF}_{H_{a,h,w=3}|\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                           | AP dose-- response--curve parameter                                                            | DR_AP_LIST                                   |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{H }_{a,h,w=3}=\text{CDF}_{H_{a,h, w=3} |\tilde{H}_{a,h,w=3},\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$ \|                                   | AP dose-- response--curve parameter \|                                                         | DR_AP_LIST                                   |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{H}_{a,h,w=3}=\text{CD                                                                                                                                | AP dose-- response--curve parameter                                                            | DR_AP_LIST                                   |
-| F}_{H_{a,h,w=3}|\tilde{H}_{a,h, w=3} ,\tilde{H}_{a,h,w=3},\tilde{H}_{a ,h ,w=3}}^{-1}(\xi_{h,w=3})^{\dagger}$                                                |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\check{H}_{h,i,s} = 1 + \tilde{H}_{a=a(i),h,w=3}\times$                                                                                                     |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\left\{1 - \exp\left(-\tilde{H }_ {a=a(i),h,w=3} (\check{W}_{i,s} - \                                                                                       | RR of disease given pollution                                                                  | RR_AP\_ calculations                         |
-| tilde{H}_{a=a(i),h,w=3}) ^ {\tild e{ H}_{a=a(i),h,w=3} }\right)\right\}$                                                                                     |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| *Physical activity calculations*                                                                                                                             |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\dot{Y}_{a,g}=\                                                                                                                                             | Raw probability non-travel MMETs = 0 for demographic group $a,g$                               | raw_zero                                     |
-| text{Prob}(Y_i=0)|_{a(i)=a,g(i)=g}$                                                                                                                          |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\b ar{Y}_{a,g}=\left\{\begin{array }{lr } \dot{Y}_{a,g} & \tilde{Y}=1 \\ \t ext{CDF}_{\text{Beta}(f(\tilde{Y},\                                             | Probability non-travel MMETs = 0 for demographic group $a,g$                                   | zeros                                        |
-| dot{Y}_{a,g}))}^{-1}(\tilde{\thet a} ) & \tilde{Y}<1 \end{array}\right.$                                                                                     |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\check{Y}_i = \left\{\begin {a rray}{lr} 0 & \bar{Y}_{a(i),g(i)} \                                                                                          | Non-travel MMETs                                                                               | SYNTHETIC\_ POPULATION\$ work_l tpa_marg_met |
-| \ \text{Sample}(\{Y_{i'}\}|_{a( i')= a(i),g(i')=g(i),Y_{i'}>0}) & 1-\b ar {Y}_{a(i),g(i)} \end{array}\right.$                                                |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $M_{i,s}=\sum_{j :i(j )=i,s(j)=s}(\chi T_j\cdot \lambda _{ m(j)}/Z_{m(j)})+\theta \check{Y}_i$                                                               | Total MMETs per person                                                                         | mmets                                        |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{G}_{h,x}(m) = {G}_{h, x,z= \lfloor{m}\rfloor}+({G}_{h,x,z=\l ce il{m}\rceil}-{G}_{h,x,z=\lfloor{m}\                                                  | Interpolated dose--response parameters                                                         | rr, lb, ub                                   |
-| rfloor})\frac{m-\lfloor{m}\rfloor }{ \lceil{m}\rceil-\lfloor{m}\rfloor}$                                                                                     |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $F_S(s_x) = \mathcal{N}(s_{x=2},( s_ {x=3}-s_{x=1})/1.96)|_{0<F_S(s_x)}$                                                                                     | Truncated normal function                                                                      |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $W_{h,i,s}=\text{CDF}_{F_S\                                                                                                                                  | Relative risk of disease given PA                                                              | RR_PA\_ calculations                         |
-| left (\tilde{G}_{h,x}(M_{i,s})\right)} ^{ -1}\left(\phi_{h}\right)^{\dagger}$                                                                                |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| *Burden-of-disease calculations*                                                                                                                             |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{W}_{h, i, s}=W_{h,i,s}\cdot\check{H}_{h,i,s}$                                                                                                        | Combined PA and AP relative risks                                                              | RR_PA_AP\_ calculations                      |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\hat{W}_{a,g,h,s}= \su m_ {i:a(i)=a,g(i)=g}\tilde{W}_{h,i,s}$                                                                                               | Populati on-attributable fractions                                                             | pif_temp                                     |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\bar{W}_{a,g,h,s}= (\hat{W} _{a, g,h,s=\text{scen1}}-\hat{W}_{a,g, h, s})/\hat{W}_{a,g,h,s=\text{scen1}}$                                                   | Populati on-attributable fractions relative to scenario 1                                      | pif_scen                                     |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\bar{U}_{a,g,h,o}=U_{a ,g ,h,o}\frac{\bar{N}_{a,g}}{N_{a,g}}$                                                                                               | GBD scaled to local population                                                                 | DISEASE_BURDEN                               |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\tilde{U}_{a, g, h,o}=\rho_h\cdot \bar{U}_{a,g,h,o}$                                                                                                        | Scaled background disease for Neoplasms, IHD, Lung cancer                                      | gbd_data_scaled                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\hat{U}_{a,g,h,o,s}=\bar{ W} _{a,g,h,s}\cdot\tilde{U}_{a,g,h,o}$                                                                                            | Combined health and PIF                                                                        | yll_dfs, death_dfs                           |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-| $\hat{U}_{a,g,h=\text{road injury                                                                                                                            | Injury health burden                                                                           | inj                                          |
-| },o,s}=\sum_m\check{I}_{a,g,m,o,s}$                                                                                                                          |                                                                                                |                                              |
-+--------------------------------------------------------------------------------------------------------------------------------------------------------------+------------------------------------------------------------------------------------------------+----------------------------------------------+
-
-: All calculations in the ITHIM.
-:::
-:::
-:::
-:::
-
-# Uncertainty in travel data
-
-We model uncertainty in travel via the definition of "travel attributes" (in the terminology of @Mohammadian2010).
-Specifically, we define the propensity to travel in a day, and the distance travelled in a day, given that travelled occurred.
-Each travel attribute is defined per demographic group, per mode.
-
-In contrast to other travel-related synthetic populations, we have four demographic groups (rather than clusters) which consist of individuals (rather than households).
-A crucial difference between our objectives and the existing work on simulating populations in travel-demand models (that I've found so far) is that we don't need trip-level data for our calculations.
-Simulating trip-level data is computationally expensive and high-dimensional in terms of uncertain parameters.
-For our downstream computations, we need only summary data: travel per person per mode.
-Beyond scenario generation, there is no need for us to consider trips.
-
-E.g.
-@Saadi2016 focuses on (a) synthetic population via matching covariates and (b) the sequence of trips taken by each person.
-(a) isn't a big thing for Accra; we're only using two covariates.
-And for (b), I think we can focus on people as it's a lot simpler.
-For our purposes, the things in this work that might be useful to us include matching principles for (a), when we revisit it, and distributions (in particular joint distributions) for travel.
-More useful to us would be distributions for total travel, but I haven't found any in this literature (e.g.
-zero-inflated compound Poisson distribution).
-
-In contrast, my main questions are:
-
--   How to account for multi-mode trips.
-
--   What to do about zero travel for e.g. older cyclists. One possibility is to smooth over all modes and/or demographic groups.
-
--   In the ITHIM-R programme, are bus and truck drivers participants?
-
--   How should we include correlations between modes? This could also be aided by a smooth model across modes.
-
--   Whether there is a better method to get distance travelled per person than by resampling raw data.
-
-## Method
-
-### Summarise the travel survey
-
-Our first step is to summarise the demographic information and the travel survey.
-In Table [5](#population){reference-type="ref" reference="population"} are the demographic data: the populations of the city, and the number of people in the survey.
-In Table [6](#calculations){reference-type="ref" reference="calculations"} are the summary statistics from the travel survey which we will use to generate synthetic travel.
-From the raw travel data (actually the adjusted trip set, which is the raw travel plus motorcycle, bus and truck trips), we calculate the probability to travel, $B_{a,g,m}$, as the number of people who travelled by that mode divided by the number of people in the demographic group.
-It is to these values that we assign uncertainty through the variable PROPENSITY_TO_TRAVEL.
-
-::: center
-::: {#population}
-| Age    | Gender | Surveyed | Accra population |
-|:-------|:-------|---------:|-----------------:|
-| 15--49 | Male   |      279 |        5,847,716 |
-| 50--69 | Male   |       56 |        1,016,476 |
-| 15--49 | Female |      328 |        6,360,535 |
-| 50--69 | Female |       69 |        1,110,037 |
-
-: Population by demographic group.
-:::
-:::
-
-::: center
-::: ThreePartTable
-::: TableNotes
-:::
-
-::: {#calculations}
-| Age   | Gender | Mode          | Probability in raw trip set | Probability in baseline trip set |
-|:------|:-------|:--------------|----------------------------:|---------------------------------:|
-| 15-49 | Male   | Bus           |                      0.4014 |                           0.3862 |
-| 50-69 | Male   | Bus           |                      0.4286 |                           0.3636 |
-| 15-49 | Female | Bus           |                      0.3506 |                           0.3495 |
-| 50-69 | Female | Bus           |                      0.4203 |                           0.4203 |
-| 15-49 | Male   | Taxi          |                      0.0681 |                           0.0655 |
-| 50-69 | Male   | Taxi          |                      0.0357 |                           0.0303 |
-| 15-49 | Female | Taxi          |                      0.0671 |                           0.0669 |
-| 50-69 | Female | Taxi          |                      0.1159 |                           0.1159 |
-| 15-49 | Male   | Walking       |                      0.6559 |                            0.631 |
-| 50-69 | Male   | Walking       |                      0.4643 |                           0.3939 |
-| 15-49 | Female | Walking       |                      0.5427 |                            0.541 |
-| 50-69 | Female | Walking       |                      0.4203 |                           0.4203 |
-| 15-49 | Male   | Private Car   |                      0.1254 |                           0.1207 |
-| 50-69 | Male   | Private Car   |                      0.1964 |                           0.1667 |
-| 15-49 | Female | Private Car   |                      0.0549 |                           0.0547 |
-| 50-69 | Female | Private Car   |                       0.058 |                            0.058 |
-| 15-49 | Male   | Bicycle       |                      0.0179 |                           0.0172 |
-| 50-69 | Male   | Bicycle       |                           0 |                                0 |
-| 15-49 | Female | Bicycle       |                       0.003 |                            0.003 |
-| 50-69 | Female | Bicycle       |                           0 |                                0 |
-| 15-49 | Male   | Short Walking |                           0 |                           0.3862 |
-| 50-69 | Male   | Short Walking |                           0 |                           0.3636 |
-| 15-49 | Female | Short Walking |                           0 |                           0.3495 |
-| 50-69 | Female | Short Walking |                           0 |                           0.4203 |
-| 15-49 | Male   | Motorcycle    |                           0 |                           0.0621 |
-| 50-69 | Male   | Motorcycle    |                           0 |                           0.1515 |
-| 15-49 | Female | Motorcycle    |                           0 |                           0.0061 |
-| 50-69 | Female | Motorcycle    |                           0 |                           0.0145 |
-
-: Travel summary table.
-:::
-:::
-:::
-
-### Adjust for trip weight
-
-We define a trip-mode weight per mode, $L_m$, so that a single trip by mode $m$ in the baseline scenario represents $L_m$ trips in reality.
-If in a scenario the mode of the trip is changed ($m'\rightarrow m$), it retains the weight of its original mode $m'$.
-We use the trip-mode weight to (a) sample trips for scenario generation and (b) scale the probabilities to travel.
-Note that we don't use it as trip weights are ordinarily used: to scale trip distance/duration.
-The result is a scaled set of probability parameters: $$\hat{B}_{a,g,m} = L_m{B}_{a,g,m}.$$ ($L_m$ should actually be the average trip-mode weight for all trips in the scenario.
-For baseline, it will be the same.
-It can/should differ for scenarios.
-Need to work out notation.)
-
-### Smooth probabilities
-
-We estimate smoothed probabilities for any mode that has zero or one probability, using the rest of the probabilities for reference, replacing $\hat{B}_{a,g,m}$.
-The reference to other modes is slightly problematic if e.g.
-in the scenario all travel increases.
-It means a 0 in the baseline is not the same as a 0 in the scenario.
-This could be fixed with more rigid coding.
-
-### Resample probabilities
-
-We resample the adjusted probability parameters of Table [6](#calculations){reference-type="ref" reference="calculations"}. We describe each parameter with a Beta distribution as follows:
-
-\$$\hat{B}_{g,m} \tilda L_m{B}_{a,g,m}.$$
-
-Ba,g,m∼Beta(αa,g,m,βa,g,m);
-
-\label{pitilde}
-
-\$$\hat{B}_{a,g,m} & \sim \text{Beta}(\alpha_{a,g,m},\beta_{a,g,m}); \\$$
-
-${\hat{B}}_{a,g,m} \sim Beta({\alpha}_{a,g,m}, {\beta}_{a,g,m})$
-${\alpha}_{a,g,m} + {\beta}_{a,g,m} = 200$
-$\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}$
-
-\alpha_{a,g,m}+\beta_{a,g,m} & = 200; \\
-\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}.\end{aligned}$$ Here, $\tilde{B}_{a,g,m}$ is the new, variable parameter that we resample for each simulation.
-We sample via a quantile, $\nu_m\sim\mathcal{U}(0,1)$, which we use to calculate $\tilde{B}_{a,g,m}$.
-We set $\alpha_{a,g,m}$ and $\beta_{a,g,m}$ to sum to 200, making the distributions concentrated, and we constrain their relative values to enforce the distribution's mean to equal $\hat{B}_{a,g,m}$ (the right-most column in Table [6](#calculations){reference-type="ref" reference="calculations"}).
-
-### Inform individual travel
-
-Each individual is assigned their own propensity to travel per mode, $\delta_{i,m}$, which is a uniformly distributed random variable that is fixed and not resampled.
-Whether or not individual $i$ travels by mode $m$ depends on $\tilde{B}_{a(i),g(i),m}$, which does vary: specifically, individual $i$ travels by mode $m$ if $\delta_{i,m}<\tilde{B}_{a(i),g(i),m}$.
-(To mitigate the effects of these fixed random variables, a large population is required, e.g.
-more than 5000 individuals.)
-
-\[Distributions for propensities to travel.
-The propensity is the probability that a member of a demographic group will travel by a particular mode on a particular day.
-In orange are the raw probabilities based on the synthetic trip set, here shown for the baseline scenario.
-Beta distributions are defined via a pointiness parameter, set to 200, such that $\alpha+\beta=200$.\]
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("propensity_to_travel_200.pdf")
-```
-
-(){#propensity_to_travel width="70%"}
-
-## Extrapolating to one week of travel {#Extrapolation}
-
-In order to extrapolate to one week of travel, we consider "trips" only at the stage of scenario generation, and proceed with individual-level statistics.
-Then, in our synthetic population, we have "duration travelled per mode" columns.
-None of the core ITHIM-R functions will use the "synthetic trip set".
-
-To get duration per mode per person, we use, for each demographic-group--mode combination, the probability to travel, and the density of the total duration travelled given that travel occurred.
-To get from one day to one week, consider two cases: one, each individual follows the same pattern of travel every day.
-That is, we sample one day's travel and apply it to seven days.
-Two, each individual follows the travel pattern of their demographic group on each day.
-That is, we sample seven times from the group's density to make up seven days of travel.
-
-In implementation, we blend individual and group-level travel through a "repetitiveness" parameter, $\pi$.
-Each individual has two travel sets: one from their simulated daily travel pattern, and one from the shared, convolved distribution.
-These travel sets are weighted by $\pi$ and $1-\pi$, respectively.
-$\delta_{i,m}$ defines the quantile taken for both the individual density, which is scaled up to one week, and the shared distribution, which is the convolution of the raw density seven times.[^6] The densities are shown in Figure [17](#raw_densities){reference-type="ref" reference="raw_densities"}.
-
-\[Raw distributions for duration travelled per day (left) from the travel survey. Convolution to one week is shown on the right. We use ""walking"" mode for demographic group "Male", "15--49".\]
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("raw_densities.pdf")
-```
-
-(){#raw_densities width="80%"}
-
-## Simulating a synthetic population {#synthpop}
-
-1.  Choose a population size of individuals ${N}$. Label each individual $i=1,...,N$.
-
-2.
-Allocate each individual to a demographic group according to the proportions in the Accra population (Table [5](#population){reference-type="ref" reference="population"}).
-Then each demographic group has number $\tilde{N}_{a,g}$ and $\sum_{a,g}\tilde{N}_{a,g}=N$.
-
-3.  Generate one standard uniform variable per person per mode, $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual $i$'s propensity to travel by mode $m$.
-
-4.
-$\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel by mode $m$ in scenario $s$.
-It is the proportion of raw person--duration values that are zero.
-It is parametrised by $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).
-
-5.
-Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw person--duration values for demographic group {$a,g$} and mode $m$ in scenario $s$: $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$ i.e.
-we isolate all individuals $i$ in group $\{a,g\}$ and sum their trip durations by mode $m$ in scenario $s$.
-
-6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}$ be that density corrected to have proportion $\tilde{B}_{a,g,m,s}$ zeros.
-
-7.
-Define each individual's travel duration as $D_{i,m,s}=7\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})$ (see Figure [18](#distances_beta){reference-type="ref" reference="distances_beta"} for resulting distributions).
-
-For sampling, we loop over steps 6--7, where each sample has unique $\tilde{B}_{a,g,m,s}$ values. In total, the number of uncertain variables is one per mode.
-
-## Simulating a synthetic population with repetitiveness parameter & uncertain extrapolation {#synthpop2}
-
-1.  Choose a population size of individuals ${N}$. Label each individual $i=1,...,N$.
-
-2.
-Allocate each individual to a demographic group according to the proportions in the Accra population (Table [5](#population){reference-type="ref" reference="population"}).
-Then each demographic group has number $\tilde{N}_{a,g}$ and $\sum_{a,g}\tilde{N}_{a,g}=N$.
-
-3.  Generate one standard uniform variable per person per mode, $\delta_{i,m}\sim\mathcal{U}(0,1)$. $\delta_{i,m}$ is individual $i$'s propensity to travel by mode $m$.
-
-4.
-$\tilde{B}_{a,g,m,s}$ is the propensity of group {$a,g$} to travel by mode $m$ in scenario $s$.
-It is the proportion of raw person--duration values that are zero.
-It is parametrised by $\alpha_{a,g,m,s}$ and $\beta_{a,g,m,s}$ (Equation [\[pitilde\]](#pitilde){reference-type="ref" reference="pitilde"}).
-
-5.
-Let $\mathcal{F}_{a,g,m,s}$ be the raw density for raw person--duration values for demographic group {$a,g$} and mode $m$ in scenario $s$: $\mathcal{F}_{a,g,m,s}=\left\{\sum_{j:i(j)=i,m(j)=m,s(j)=s} T_j\right\}_{i:a(i)=a,g(i)=g}$ i.e.
-we isolate all individuals $i$ in group $\{a,g\}$ and sum their trip durations by mode $m$ in scenario $s$.
-
-6.  Let $\tilde{\mathcal{F}}_{a,g,m,s}^7$ be the week convolution of $\tilde{\mathcal{F}}_{a,g,m,s}$ (see Figure [17](#raw_densities){reference-type="ref" reference="raw_densities"}, right).
-
-7.
-Define each individual's travel duration as $D_{i,m,s}=7\pi\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}}^{-1}(\delta_{i,m})+(1-\pi)\cdot\text{CDF}_{\tilde{\mathcal{F}}_{a(i),g(i),m,s}^7}^{-1}(\delta_{i,m})$ (see Figure [18](#distances_beta){reference-type="ref" reference="distances_beta"} for resulting distributions).
-
-For sampling, we loop over steps 6--8, where each sample has unique $\tilde{B}_{a,g,m,s}$ and $\pi$ values. In total, the number of uncertain variables is one per mode plus one.
-
-\[Samples of total distances travelled by mode and scenario.\]
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("distances_beta.pdf")
-```
-
-(){#distances_beta width="80%"}
-
-## Results
-
-\[YLLs in the scenario minus YLLs in the baseline for Accra, with motorcycle probability scalar 2 in Accra.\]
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("accra_walk_yll_uncertain_travel.pdf")
-```
-
-(){width="80%"}
-
-\[EVPPI for Accra's "walking" scenario for all causes of YLLs excluding "all cause" and "neoplasm", with propensity to motorcycle = 2.\]
-
-```{r out.height = "460px", out.width='800px', fig.align='center', echo=F}
-knitr::include_graphics("accra_walk_evppi_uncertain_travel.pdf")
-```
-
-(){#evppi_mode_propensities_200 width="70%"}
-
-# Glossary by letter
-
-| $\text{Letter}$ | Meaning                       | $\text{Letter}$ | Meaning           | $\text{Letter}$ | Meaning                  |
-|----------------:|:------------------------------|----------------:|:------------------|----------------:|:-------------------------|
-|        $\alpha$ |                               |             $a$ | age               |             $A$ |                          |
-|         $\beta$ |                               |             $b$ |                   |             $B$ | raw travel probabilities |
-|        $\gamma$ | emission uncertainty          |             $c$ |                   |             $C$ | ventilation constants    |
-|        $\delta$ | propensity to travel          |             $d$ | demographic group |             $D$ |                          |
-|      $\epsilon$ | walk-to-bus time              |             $e$ |                   |             $E$ | injury distance exponent |
-|         $\zeta$ | traffic PM2.5 fraction        |             $f$ |                   |             $F$ | PA DR function           |
-|          $\eta$ | PM2.5 concentration           |             $g$ | gender            |             $G$ | PA DR curves             |
-|        $\theta$ | PA uncertainty                |             $h$ | disease           |             $H$ | AP DR curves             |
-|         $\iota$ |                               |             $i$ | individual        |             $I$ | injuries                 |
-|            \$\$ |                               |             $j$ | trip              |             $J$ |                          |
-|        $\kappa$ | injury regression parameter   |             $k$ |                   |             $K$ | AP exposure              |
-|       $\lambda$ | travel MMETs                  |             $l$ |                   |             $L$ |                          |
-|           $\mu$ | mode distances                |             $m$ | mode              |             $M$ | MMETs per person         |
-|           $\nu$ |                               |             $n$ |                   |             $N$ | population               |
-|           $\xi$ | AP DR quantile                |             $o$ | outcome           |             $O$ |                          |
-|           $\pi$ | travel repetitiveness         |             $p$ |                   |             $P$ | vehicle emissions        |
-|            \$\$ |                               |             $q$ |                   |             $Q$ | person weights           |
-|          $\rho$ | chronic disease scalar        |             $r$ |                   |             $R$ | death-to-injury ratio    |
-|        $\sigma$ | injury reporting rate         |             $s$ | scenario          |             $S$ |                          |
-|          $\tau$ |                               |             $t$ |                   |             $T$ | trips                    |
-|      $\upsilon$ |                               |             $u$ |                   |             $U$ | burden of disease        |
-|          $\phi$ | PA DR quantile                |             $v$ |                   |             $V$ | AP per person            |
-|            \$\$ |                               |             $w$ | AP DR parameter   |             $W$ | relative risks           |
-|          $\chi$ | day-to-week travel scalar     |             $x$ | PA DR parameter   |             $X$ | injury model matrix      |
-|          $\psi$ | casualty fraction of $\omega$ |             $y$ | year              |             $Y$ | individual MMETs         |
-|        $\omega$ | sum of distance exponents     |             $z$ | PA dose           |             $Z$ | speeds                   |
-:::
-
-[^6]: Using the same random variable for both travel sets increases the variability of the resulting dataset.
-Conversely, using different random variables would result in more homogenous durations, and fewer zeros.

### 2022-07-04 quarto-doc/pkg_doc.qmd
Minor update. Re #97

@@
   chapters: true
-
 ---
@@
 
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | Label | Name                 | Levels               |
-+===============+======================+======================+
-| *a*           | Age group            | 15--49\              |
-|               |                      | 50--69               |
-+---------------+----------------------+----------------------+
++=======+======================+======================+
+| *a*   | Age group            | 15--69\              |
+|       |                      | (five years age      |
+|                              |             group)   |
+|                              |                      |       
++-------+----------------------+----------------------+
 | *g*   |                      | Male\                |
 |       |                      | Female               |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *h*   | Disease              |                      |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *i*   | Individual index     |                      |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *j*   | Trip index           |                      |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *m*   | Transport mode       | walk\                |
@@
 |       |                      | subway (sub.)        |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *o*   | Outcome              | Death\               |
 |       |                      | YLL                  |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *s*   | Scenario             | 3                    |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *w*   | AP RR hyperparameter | 1, 2, 3, 4           |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *x*   | PA RR hyperparameter | 1, 2, 3              |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 | *z*   | PA dose              |                      |
-+---------------+----------------------+----------------------+
++-------+----------------------+----------------------+
 
@@
 
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \[\[inputs\]\]{label= "inputs"}                |                                                                 | *ITHIM-R reference*                |                               |                    |
-+:==========================================+=================================================================+:===================================+===============================+====================+
+|:-----------------------------------------------|-----------------------------------------------------------------|:-----------------------------------|-------------------------------|--------------------|
 | **Name**                                       | **Description**                                                 | **Value in Accra model**           | **Label**                     | **Model/ Setting** |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Global values*                                |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $U_ { a,g,h,o}$                                | Background burden of disease                                    | Constant                           | GBD_DATA                      | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $N_{a,g}$                                      | Population by age and gender in GBD_DATA                        | Constant                           | GBD_DATA                      | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\bar { N}_{a,g}$                              | Population by age and gender                                    | Constant                           | DEMOGRAPHIC                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\rho_h$                                       | Chronic disease scalar                                          | 1 or Lnorm(0,0.18)                 | CHRONIC_DISEASE_SCALAR        | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Travel values*                                |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=walk}$                                 | Walk speed                                                      | 4.8                                | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=cyc.}$                                 | Cycle speed                                                     | 14.5                               | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=bus}$                                  | Bus speed                                                       | 15                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=car}$                                  | Car speed                                                       | 21                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=mot.}$                                 | Motorbike speed                                                 | 25                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$\\epsilon \$                                 | Walk time to bus                                                | 5 or Lnorm(log (5),0.18)           | BUS_WALK_TIME                 | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\mu_{m=HGV}$                                  | Truck distance relative to car                                  | 0.21 or Beta(3,10)                 | TRUCK_TO\_ CAR_RATIO          | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\mu_{m=bus}$                                  | Bus driver distance relative to bus passenger distance          | 1/45 or Bet a(20,600) \|           | BUS_TO_PASSENGER_RATIO        | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\mu_ {m=mot.}$                                | Motorcycle fleet distance relative to total motorcycle distance | 37/68 or Beta(37/3+ 1,31/3+1) \|   | FLEET_TO_MOTORCYCLE_RATIO     | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $T_j$                                          | Trip-level data                                                 | Constant                           | SYNTHETIC_TRIPS               | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Injury model values*                          |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $I _ {a,g,m_{\                            | Fatalities table                                                | Constant                           | INJURY_TABLE                  | Setting            |
-| text{vi c} },m_{\text{str}}}$             |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $I _ {a,g,m_{\   text{vi c} },m_{\text{str}}}$ | Fatalities table                                                | Constant                           | INJURY_TABLE                  | Setting            |
 | $\sigma$                                       | Injury reporting rate                                           | 1 or Beta(8,3)                     | INJURY_REPORTING_RATE         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\omega$                                       | Sum of distance exponents                                       | 1.9 or Lnorm(l og(1.9),l og(1.03)) | SIN_EXPONENT_SUM              | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\psi$                                         | Casualty fraction of $\omega$                                   | 0.5 or Beta(20,20) \|              | CASUALTY_EXPONENT_FRACTION    | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Pollution model values*                       |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\eta$                                         | Background PM2.5 concentration                                  | 50 or Lnorm(log( 50),0.18) \|      | PM_CONC_BASE                  | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\zeta$                                        | Fraction of measured PM2.5 concentration due to road transport  | 0.225 or Beta(5,20) \|             | PM_TRANS_SHARE                | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $\                                        | Vehicle emission inventory                                      | Constant or Dirichlet              | EMISSION_INVENTORY            | Setting            |
-| g amma_{m}$                               |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\   g amma_{m}$                               | Vehicle emission inventory                                      | Constant or Dirichlet              | EMISSION_INVENTORY            | Setting            |
 | $P$                                            | Vehicle emission confidence                                     | 1                                  | EMISSION_INVENTORY_CONFIDENCE | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Pollution & health model values*              |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_1$                                          | Base-level inhalation rate                                      | 1                                  | BASE_LEVEL_INHALATION_RATE    | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_2$                                          | Exposure ratio window closed                                    | 0.5                                | CLOSED_WINDOW_PM\_ RATIO      | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_3$                                          | Proportion of travel with closed windows                        | 0.5                                | C LOSED_WIN DOW_RATIO         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_4$                                          | Parameter for on-road PM2.5 level                               | 3.216                              | ROAD\_ RATIO_MAX              | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_5$                                          | Parameter for on-road PM2.5 level                               | 0.379                              | ROAD_RA TIO_SLOPE             | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_6$                                          | Exposure ratio for underground                                  | 0.8                                | SUBWAY \_PM_RATIO             | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$ H \_{a,h,w}\$                               | AP dose- -res ponse--curve parameters                           | Constant                           | DR_AP                         | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $\                                        | Quantiles for AP RR functions                                   | 0.5 or Unif(0,1)                   | AP_DOSE_R ESPONSE\_ QUANTILE  | Model              |
-| xi_{h,w}$                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\   xi_{h,w}$                                 | Quantiles for AP RR functions                                   | 0.5 or Unif(0,1)                   | AP_DOSE_R ESPONSE\_ QUANTILE  | Model              |
 | *Physical activity model values*               |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\chi$                                         | Day-to-week travel scalar                                       | 7 or 7 $\times$Beta(20,3)          | DAY_TO_WEEK_TRAVEL_SCALAR     | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Y_i$                                          | Individual n on-transport MMETs                                 | Constant                           | PA_SET\$ work_ltpa \_marg_met | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\theta$                                       | Background PA scalar                                            | 1 or Lnorm(0,0.18)                 | BACKGROUND\_ PA_SCALAR        | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $\                                        | Confidence in PA survey                                         | 1                                  | BACKGROUND_PA_C ONFIDENCE     | Setting            |
-| tilde{Y}$                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\   tilde{Y}$                                 | Confidence in PA survey                                         | 1                                  | BACKGROUND_PA_C ONFIDENCE     | Setting            |
 | $\tilde {\theta}$                              | Background PA zeros quantile                                    | 0.5 or Unif(0,1)                   | BACKGROUND \_PA_ZEROS         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\lambda_{m=\text{cyc.}}$                      | Cycling MET surplus                                             | 4.63 or Lnorm(log (4.63),1)        | MMET_CYCLING                  | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\lambda_{m=\text{walk}}$                      | Walking MET surplus                                             | 2.53 or Lnorm(log (2.53),1)        | MMET_WALKING                  | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $\lambda_{m\not\in\text{wal               | In-vehicle MET surplus                                          | 0                                  |                               | Model              |
-| k,cyc.}}$                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\lambda_{m\not\in\text{wal                    
+  k,cyc.}}$                                      | In-vehicle MET surplus                                          | 0                                  |                               | Model              |
 | *Physical activity & health model values*      |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$\\G\_{h,x,z}\$                               | Look-up table for relative risk truncated normal distribution   | Constant                           |                               | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$\phi\_{h}\$                                  | Quantiles for PA RR functions                                   | 0.5 or Unif(0,1)                   | PA_DOSE_RESPONSE_QUANTILE     | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 
@@
 
-[Note on recategorization]{.ul}: for some of the modes (like for car, bus, motorcycle and cycle), if both strike and victim modes are same, we explicitly make the strike mode as 'noov'
+[Note on recategorization]{.underline}: for some of the modes (like for car, bus, motorcycle and cycle), if both strike and victim modes are same, we explicitly make the strike mode as 'noov'
 
@@
 
-::: appendix
-\[ITHIM-R workflow, with variable parameters in purple.
+::: appendix \[ITHIM-R workflow, with variable parameters in purple.
 Fixed inputs to the model have a bold, black outline, which is solid for inputs we consider "global" and dashed for inputs we consider "local".
@@
 
-+:----------------------------------------------------------------------------------:+:---------------------------------------------------------------------------------:+
+|                                                                                    |                                                                                   |
+|:----------------------------------------------------------------------------------:|:---------------------------------------------------------------------------------:|
 | ![100 samples from AP dose--response curves](cvd_ihdDR-sim.pdf){#APDR width="40%"} | \[100 samples from AP dose--response curves\] (cvd_strokeDR-sim.pdf){width="40%"} |
-+------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------+
 
-+:-----------------------------------------------------------------:+:------------------------------------------------------------:+:------------------------------------------------------------------:+
+|                                                                   |                                                              |                                                                    |
+|:-----------------------------------------------------------------:|:------------------------------------------------------------:|:------------------------------------------------------------------:|
 | ![100 samples from AP dose--response curv es](neo_lungDR-sim.pdf) | ![100 samples from AP dose--response c urves](lriDR-sim.pdf) | ![100 samples from AP dose--response curve s](resp_copdDR-sim.pdf) |
-+-------------------------------------------------------------------+--------------------------------------------------------------+--------------------------------------------------------------------+
 
-::: landscape
-# Tabulated ITHIM equations {#tableequations}
+::: landscape \# Tabulated ITHIM equations {#tableequations}
 
-::: center
-::: ThreePartTable
-::: TableNotes
-CDF: cumulative distribution function.
+::: center ::: ThreePartTable ::: TableNotes CDF: cumulative distribution function.
 We use generative distributions to generate random numbers so that samples are correlated.

### 2022-07-14 quarto-doc/pkg_doc.qmd
Update quarto doc. Re #97

--- 2022-06-22 quarto-doc/pkg_doc.qmd
@@
 
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | Label       | Name                 | Levels                        |
-+===============+======================+======================+
-| *a*           | Age group            | 15--49\              |
-|               |                      | 50--69               |
-+---------------+----------------------+----------------------+
++=============+======================+===============================+
+| *a*         | Age group            | 15--69 (five years age group) |
++-------------+----------------------+-------------------------------+
 | *g*         |                      | Male\                         |
 |             |                      | Female                        |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *h*         | Disease              |                               |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *i*         | Individual index     |                               |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *j*         | Trip index           |                               |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *m*         | Transport mode       | walk\                         |
@@
 |             |                      | subway (sub.)                 |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *o*         | Outcome              | Death\                        |
 |             |                      | YLL                           |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *s*         | Scenario             | 3                             |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *w*         | AP RR hyperparameter | 1, 2, 3, 4                    |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *x*         | PA RR hyperparameter | 1, 2, 3                       |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 | *z*         | PA dose              |                               |
-+---------------+----------------------+----------------------+
++-------------+----------------------+-------------------------------+
 
@@
 
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \[\[inputs\]\]{label= "inputs"}                 |                                                                 | *ITHIM-R reference*                |                               |                    |
-+:==========================================+=================================================================+:===================================+===============================+====================+
++:================================================+=================================================================+:===================================+===============================+====================+
 | **Name**                                        | **Description**                                                 | **Value in Accra model**           | **Label**                     | **Model/ Setting** |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Global values*                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $U_ { a,g,h,o}$                                 | Background burden of disease                                    | Constant                           | GBD_DATA                      | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $N_{a,g}$                                       | Population by age and gender in GBD_DATA                        | Constant                           | GBD_DATA                      | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\bar { N}_{a,g}$                               | Population by age and gender                                    | Constant                           | DEMOGRAPHIC                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\rho_h$                                        | Chronic disease scalar                                          | 1 or Lnorm(0,0.18)                 | CHRONIC_DISEASE_SCALAR        | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Travel values*                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=walk}$                                  | Walk speed                                                      | 4.8                                | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=cyc.}$                                  | Cycle speed                                                     | 14.5                               | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=bus}$                                   | Bus speed                                                       | 15                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=car}$                                   | Car speed                                                       | 21                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Z _ {m=mot.}$                                  | Motorbike speed                                                 | 25                                 | MODE_SPEEDS                   | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$\\epsilon \$                                  | Walk time to bus                                                | 5 or Lnorm(log (5),0.18)           | BUS_WALK_TIME                 | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\mu_{m=HGV}$                                   | Truck distance relative to car                                  | 0.21 or Beta(3,10)                 | TRUCK_TO\_ CAR_RATIO          | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\mu_{m=bus}$                                   | Bus driver distance relative to bus passenger distance          | 1/45 or Bet a(20,600) \|           | BUS_TO_PASSENGER_RATIO        | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\mu_ {m=mot.}$                                 | Motorcycle fleet distance relative to total motorcycle distance | 37/68 or Beta(37/3+ 1,31/3+1) \|   | FLEET_TO_MOTORCYCLE_RATIO     | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $T_j$                                           | Trip-level data                                                 | Constant                           | SYNTHETIC_TRIPS               | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Injury model values*                           |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $I _ {a,g,m_{\                            | Fatalities table                                                | Constant                           | INJURY_TABLE                  | Setting            |
-| text{vi c} },m_{\text{str}}}$             |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $I _ {a,g,m_{\  text{vi c} },m_{\text{str}}}$   | Fatalities table                                                | Constant                           | INJURY_TABLE                  | Setting            |
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\sigma$                                        | Injury reporting rate                                           | 1 or Beta(8,3)                     | INJURY_REPORTING_RATE         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\omega$                                        | Sum of distance exponents                                       | 1.9 or Lnorm(l og(1.9),l og(1.03)) | SIN_EXPONENT_SUM              | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\psi$                                          | Casualty fraction of $\omega$                                   | 0.5 or Beta(20,20) \|              | CASUALTY_EXPONENT_FRACTION    | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Pollution model values*                        |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\eta$                                          | Background PM2.5 concentration                                  | 50 or Lnorm(log( 50),0.18) \|      | PM_CONC_BASE                  | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\zeta$                                         | Fraction of measured PM2.5 concentration due to road transport  | 0.225 or Beta(5,20) \|             | PM_TRANS_SHARE                | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $\                                        | Vehicle emission inventory                                      | Constant or Dirichlet              | EMISSION_INVENTORY            | Setting            |
-| g amma_{m}$                               |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\  g amma_{m}$                                 | Vehicle emission inventory                                      | Constant or Dirichlet              | EMISSION_INVENTORY            | Setting            |
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $P$                                             | Vehicle emission confidence                                     | 1                                  | EMISSION_INVENTORY_CONFIDENCE | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Pollution & health model values*               |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_1$                                           | Base-level inhalation rate                                      | 1                                  | BASE_LEVEL_INHALATION_RATE    | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_2$                                           | Exposure ratio window closed                                    | 0.5                                | CLOSED_WINDOW_PM\_ RATIO      | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_3$                                           | Proportion of travel with closed windows                        | 0.5                                | C LOSED_WIN DOW_RATIO         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_4$                                           | Parameter for on-road PM2.5 level                               | 3.216                              | ROAD\_ RATIO_MAX              | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_5$                                           | Parameter for on-road PM2.5 level                               | 0.379                              | ROAD_RA TIO_SLOPE             | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $C_6$                                           | Exposure ratio for underground                                  | 0.8                                | SUBWAY \_PM_RATIO             | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$ H \_{a,h,w}\$                                | AP dose- -res ponse--curve parameters                           | Constant                           | DR_AP                         | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $\                                        | Quantiles for AP RR functions                                   | 0.5 or Unif(0,1)                   | AP_DOSE_R ESPONSE\_ QUANTILE  | Model              |
-| xi_{h,w}$                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\  xi_{h,w}$                                   | Quantiles for AP RR functions                                   | 0.5 or Unif(0,1)                   | AP_DOSE_R ESPONSE\_ QUANTILE  | Model              |
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Physical activity model values*                |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\chi$                                          | Day-to-week travel scalar                                       | 7 or 7 $\times$Beta(20,3)          | DAY_TO_WEEK_TRAVEL_SCALAR     | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $Y_i$                                           | Individual n on-transport MMETs                                 | Constant                           | PA_SET\$ work_ltpa \_marg_met | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\theta$                                        | Background PA scalar                                            | 1 or Lnorm(0,0.18)                 | BACKGROUND\_ PA_SCALAR        | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
-| $\                                        | Confidence in PA survey                                         | 1                                  | BACKGROUND_PA_C ONFIDENCE     | Setting            |
-| tilde{Y}$                                 |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
+| $\  tilde{Y}$                                   | Confidence in PA survey                                         | 1                                  | BACKGROUND_PA_C ONFIDENCE     | Setting            |
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\tilde {\theta}$                               | Background PA zeros quantile                                    | 0.5 or Unif(0,1)                   | BACKGROUND \_PA_ZEROS         | Setting            |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\lambda_{m=\text{cyc.}}$                       | Cycling MET surplus                                             | 4.63 or Lnorm(log (4.63),1)        | MMET_CYCLING                  | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\lambda_{m=\text{walk}}$                       | Walking MET surplus                                             | 2.53 or Lnorm(log (2.53),1)        | MMET_WALKING                  | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | $\lambda_{m\not\in\text{wal                     | In-vehicle MET surplus                                          | 0                                  |                               | Model              |
 |   k,cyc.}}$                                     |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | *Physical activity & health model values*       |                                                                 |                                    |                               |                    |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$\\G\_{h,x,z}\$                                | Look-up table for relative risk truncated normal distribution   | Constant                           |                               | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 | \$\phi\_{h}\$                                   | Quantiles for PA RR functions                                   | 0.5 or Unif(0,1)                   | PA_DOSE_RESPONSE_QUANTILE     | Model              |
-+-------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
++-------------------------------------------------+-----------------------------------------------------------------+------------------------------------+-------------------------------+--------------------+
 
@@
 
-### Scenarios
+### Scenarios {#scenarios}
 
@@
 
-We use each person's exposure to PM2.5 ($\check{W}_{i,s}$) to calculate their relative risk of five diseases (IHD, lung cancer, COPD, stroke, LRI), using curves parametrised by four disease-specific variables [@Burnett2014].
+We use each person's exposure to PM2.5 ($\check{W}_{i,s}$) to calculate their relative risk of seven causes/diseases (all-cause mortality, IHD, stroke, COPD, LRI, Lung cancer and Type 2 Diabetes (T2D)).
+These diseases operate in three levels.
+The Level 1 is for all-cause mortality.
+The Level 2 combines IHD and stroke to CHD, and COPD, LRI, Lung cancer and T2D as respiratory disease mortality, and then in level 3, we have all diseases separately.
 Of the five diseases, two (IHD and stroke) have parameters specific to age groups starting at age 25.[^2] The other three (lung cancer, LRI and COPD) have one set of parameters for all ages.
@@
 
-The curves are in the form of samples of the set of four parameters.
+![AP disease levels](images/ap-levels-min.png)The curves are in the form of samples of the set of four parameters.
 We model the densities of these samples (using a quantile for parameter 3, kernel density estimation for parameter 2, and GAMs for parameters 1 and 4)[^3] in order to draw either the median or random samples via their quantiles ($\xi_{h,w}$) as follows:
@@
 
-::: landscape
-# Tabulated ITHIM equations {#tableequations}
+::: landscape \# Tabulated ITHIM equations {#tableequations}
 
-::: center
-::: ThreePartTable
-::: TableNotes
-CDF: cumulative distribution function.
+::: center ::: ThreePartTable ::: TableNotes CDF: cumulative distribution function.
 We use generative distributions to generate random numbers so that samples are correlated.

### 2022-09-30 injury_tables.Rmd
Indent

@@
 #scen_names_w_baseline <- c('bl', 'w_sc', 'bi_sc', 'car_sc', 'mc_sc', 'bus_sc')
-scen_names_w_baseline <- c('bl', 'bi_sc', 'car_sc', 'bus_sc')
+scen_names_w_baseline <- c('bl', 'bi_sc', 'car_sc', 'bus_sc', 'mc_sc')
 scen_lt <- data.frame("scen_name" = names(io[[1]]$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)
@@
 injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Driving'
-#injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
 #injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
@@
     scenario == "Scenario 3" ~ "Public Transport",
+    scenario == "Scenario 4" ~ "Motorcycling",
     scenario == "Baseline" ~ "Baseline"))

### 2022-10-03 injury_tables.Rmd
Added fatalities per 100million hours travelled

@@
 io <- io[-2]
+# re-format some age and injury names
 for (i in 1:length(io)) {
@@
     for (out in c('deaths','ylls'))
-      io[[city]]$outcomes[[type]][[out]] <- collapse_ages(io[[city]]$outcomes[[type]][[out]])
-for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')
-pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population)))
+      io[[city]]$outcomes[[type]][[out]] <- collapse_ages(io[[city]]$outcomes[[type]][[out]]) # only keep two age categories
+
+for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')  # only keep two age categories
+
+pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population))) # find number of people within two age categories
 pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
-injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths)=='scen1_deaths_inj')
+injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths)=='scen1_deaths_inj') # find col number of injury data
 ap_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
@@
 
-  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$Baseline), value = T), 2, sep = "_"))){
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$Baseline), value = T), 2, sep = "_"))){ # if upper and lower bounds are given, this is NA, ub, lb
     # pref <- "ub"
@@
     if (!is.na(pref))
-      pref_name <- paste0("_", pref)
+      pref_name <- paste0("_", pref) # if pref is lb or ub e.g.
 
@@
 
-    whw_qn <- paste0("whw", pref_name)
+    whw_qn <- paste0("whw", pref_name)   # if upper and lower bounds are given, this takes one of whw, whw_ub, whw_lb
     nov_qn <- paste0("nov", pref_name)
 
-    for (cs in names(io[[city]]$outcomes$whw)){
+    for (cs in names(io[[city]]$outcomes$whw)){ # Baseline, Scenario 1, Scenario 2, etc.
       # cs <- 'Baseline'
@@
       if (is.null(io[[city]]$outcomes$whw[[cs]][[whw_qn]])){
-        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){
+        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){ # if only data for the given name exists for nov but not for whw, only create data frame from nov part
           td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()
@@
 
-      }else{
-        td1 <- (io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% tibble::rownames_to_column("mode")
+      }else{ # if whw exists
+        td1 <- (io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% tibble::rownames_to_column("mode") # whw fatalities table
         td3 <- td1
-        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){
-          td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()
+        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){ # if nov exists as well
+          td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()  # nov fatalities
           td2$mode <- 'NOV'
 
-          td3 <- plyr::rbind.fill(td1, td2)
+          td3 <- plyr::rbind.fill(td1, td2) # nov and whw fatalities
         }
       }
-      td3 <- td3 %>% mutate(rowSum = rowSums(.[2:ncol(td3)], na.rm = T))
-      td3 <- td3 %>% janitor::adorn_totals("row")
+      td3 <- td3 %>% mutate(rowSum = rowSums(.[2:ncol(td3)], na.rm = T)) # add row sums
+      td3 <- td3 %>% janitor::adorn_totals("row") # add column sums
       td3[, 2:ncol(td3)] <- round(td3[, 2:ncol(td3)], 2)
@@
 
-      td3 <- td3[,c(1, na.omit(match(smodes$exhaustive_list, colnames(td3))), ncol(td3))]
+      td3 <- td3[,c(1, na.omit(match(smodes$exhaustive_list, colnames(td3))), ncol(td3))] # remove column modes that do not appear in list of standardized modes
 
-      td3 <- td3[c(na.omit(match(smodes$exhaustive_list, td3$mode)), nrow(td3)),]
+      td3 <- td3[c(na.omit(match(smodes$exhaustive_list, td3$mode)), nrow(td3)),] # remove rows with mode names that do not appear in list of standardized modes
 
-      temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))
+      temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( .
+< 1 ~ round(., 2) ,  .
+>= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))  # convert numbers to characters
 
-      ft <- flextable(temp)
+      ft <- flextable(temp)   # convert table in correct format for printing to html file
       ft <- add_header_row(ft, values = c('Strike Mode', 'Casualty Mode'), colwidths = c(1, ncol(temp) - 1))
@@
       td <- td3
-      names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+      names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_") # add 'city_' at beginning of column names
       td$measure <- ifelse(is.na(pref), "mean", pref)
 
-      city_df[[var]][[city]][[unique(td$measure)]] <- td
+      city_df[[var]][[city]][[unique(td$measure)]] <- td   # save all data in one list called city_df
 
-      if (length(city_df[[var]][[city]]) == 3)
-        all_whw[[var]][[city]] <- data.table::rbindlist(city_df[[var]][[city]])
+      if (length(city_df[[var]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
+        all_whw[[var]][[city]] <- data.table::rbindlist(city_df[[var]][[city]]) # combine mean, lb and ub counts into one table for each scenario
     }
@@
 whw_list <- list()
-for (cs in names(all_whw)){
+for (cs in names(all_whw)){ # loop through scenarios
   # cs <- names(all_whw)[1]
@@
 
-  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total'))
+  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total')) # re-arrange order of rows
 
@@
 
-  whw_lng <- reshape2::melt(td)
+  whw_lng <- reshape2::melt(td) # create df with 1 column containing all the fatality counts
 
-  col_split <- stringr::str_split(whw_lng$variable, "_", simplify = TRUE, n = 2)
+  col_split <- stringr::str_split(whw_lng$variable, "_", simplify = TRUE, n = 2) # create one city and one strike mode column
 
+  # tidy and rename some of the columns
   whw_lng <- cbind(whw_lng, col_split)
@@
 
-  whw_lng$str_mode <- factor(whw_lng$str_mode, levels = unique(whw_lng$str_mode))
+  whw_lng$str_mode <- factor(whw_lng$str_mode, levels = unique(whw_lng$str_mode)) # assign factors
 
@@
 # Remove total from whw long form, and write as csv
-readr::write_csv(whw_list %>% filter(str_mode != 'Total'), paste0('results/multi_city/whw_matrices/whw_lng.csv'))
+readr::write_csv(whw_list %>% filter(str_mode != 'Total'), paste0('results/multi_city/whw_matrices/whw_lng.csv')) # table containing all the fatality counts
 ```
@@
 ```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
+### by casualty mode # This does contain NOV fatalities!
+
 list_temp <- list()
@@
 
+### Change in deaths due to injury per 100 million hours
+
+```{r inj_100k = "asis"}
+require(tibble)
+overall_el <- list() # To save all cities results
+overall_dd <- list() # To save all cities results
+
+city_df <- list()
+for (city in cities){
+
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$Baseline), value = T), 2, sep = "_"))){
+
+    el <- list() # To save city specific results
+    # pref <- "ub"
+    pref_name <- ""
+    if (!is.na(pref))
+      pref_name <- paste0("_", pref)
+
+    print(pref)
+
+    whw_qn <- paste0("whw", pref_name)
+    nov_qn <- paste0("nov", pref_name)
+
+    for (cs in names(io[[city]]$outcomes$whw)){ # loop through scenarios
+      # cs <- 'Baseline'
+      if (length(word(names(io[[city]]$outcomes$whw$Baseline), 1, sep = "_") %>% unique()) == 2){ # if there exist nov and whw matrices
+        # td1: Number of fatalities in nov
+        td1 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        # td2: Number of fatalities in whw
+        td2 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        # td3: Number of fatalities: sum of nov and whw
+        td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
+        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+        td3 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+
+        td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
+
+      }
+      # td4: duration travelled per mode
+      td4 <- io[[city]]$dur %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs) %>% as.data.frame()
+
+      if (length(el) == 0){
+        el <- td4 %>% dplyr::select(stage_mode)
+        dd <- el
+      }
+      # td4: merge total fatalities to distance
+      td4 <- full_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
+
+      var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+
+      names(td4)[2] <- var
+
+      td5 <- td4
+
+      td4[, 2] <- as.numeric(td4[,2])
+      # Compute risk per 100 million hours travelled
+      td4[, 2] <- round((td4[,3] / ( td4[,2] * 365)) * 
+                          100000000, 4)
+
+      names(td4)[3] <- paste(names(td4)[2], names(td4)[3], sep = "_")
+
+      el <- full_join(el, td4, by = c('stage_mode'))
+
+      # City's population
+      pop <- sum(io[[city]]$demographic$population)
+      # Compute travel time travelled per capita
+      td6 <- td5
+      td6[[paste0(var,"_risk")]] <- unlist(td4[2])
+      td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop
+      names(td6)[3] <- paste(names(td6)[2], names(td6)[3], sep = "_")
+      dd <- full_join(dd, td6, by = 'stage_mode')
+    }
+
+    print(kable(el, caption = paste(city, ifelse(is.na(pref), "mean", pref))))
+
+    td <- el %>% dplyr::select(-contains('count'))
+    names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+    td$measure <- ifelse(is.na(pref), "mean", pref)
+    dd$measure <- ifelse(is.na(pref), "mean", pref)
+
+    city_df[["td"]][[city]][[unique(td$measure)]] <- td
+    city_df[["dd"]][[city]][[unique(td$measure)]] <- dd
+
+    if (length(city_df[["td"]][[city]]) == 3){
+      overall_el[[city]] <- data.table::rbindlist(city_df[["td"]][[city]])
+      overall_dd[[city]] <- data.table::rbindlist(city_df[["dd"]][[city]])
+    }
+
+    cat("\n")
+  }
+}
+```
+
+### Change in deaths due to injury per 100 million hours across cities
+
+```{r inj_100k_all_cities = "asis"}
+#require(tibble)
+td <- overall_el %>% purrr::reduce(full_join, by = c("stage_mode", "measure")) %>% as.data.frame()
+td[is.na(td)] <- 0
+td <- td %>% dplyr::select(stage_mode, sort(names(.)))
+
+colnames(td) = gsub("bi_sc", "bisc", colnames(td))
+colnames(td) = gsub("bus_sc", "bssc", colnames(td))
+colnames(td) = gsub("car_sc", "csc", colnames(td))
+colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
+colnames(td) = gsub("w_sc", "wsc", colnames(td))
+injury_risks_100mil_hours <- td
+
+colnames(injury_risks_100mil_hours) = gsub("bi_sc", "bisc", colnames(injury_risks_100mil_hours))
+colnames(injury_risks_100mil_hours) = gsub("bus_sc", "bssc", colnames(injury_risks_100mil_hours))
+colnames(injury_risks_100mil_hours) = gsub("car_sc", "csc", colnames(injury_risks_100mil_hours))
+colnames(injury_risks_100mil_hours) = gsub("mc_sc", "mcsc", colnames(injury_risks_100mil_hours))
+colnames(injury_risks_100mil_hours) = gsub("w_sc", "wsc", colnames(injury_risks_100mil_hours))
+injury_risks_lng_hours <- reshape2::melt(injury_risks_100mil_hours)
+col_split <- stringr::str_split(injury_risks_lng_hours$variable, "_", simplify = TRUE, n = 2)
+injury_risks_lng_hours <- cbind(injury_risks_lng_hours, col_split)
+names(injury_risks_lng_hours)[5] <- 'scenario'
+names(injury_risks_lng_hours)[6] <- 'city'
+injury_risks_lng_hours <- injury_risks_lng_hours %>% dplyr::select(-variable)
+cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
+injury_risks_lng_hours$scenario <- as.character(injury_risks_lng_hours$scenario)
+injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bisc',]$scenario <- 'Bicycling'
+injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bssc',]$scenario <- 'Public Transport'
+injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'csc',]$scenario <- 'Driving'
+rd_hours <- rename(injury_risks_lng_hours, mode = stage_mode)
+rd_hours <- rd_hours %>% filter(mode != 'Total')
+readr::write_csv(rd_hours, 'results/multi_city/whw_matrices/injury_risks_per_100million_h_lng.csv')
+```
+
 ### Write all injury tables in a single excel sheet
@@
                           injury_risks_per_100k  = injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'),
+                          injury_risks_per_100million_h = rd_hours,
                           true_distances = true_distances_df),

### 2022-10-10 summary_tables_PA_AP.Rmd
Include fourth motorcycling scen

@@
 cities <- names(io)[!names(io) %in% 'scen_prop']
+
+scen_names <- c("Baseline", "Bicycling", "Driving", "Public Transport", "Motorcycling")
 
@@
   #                                                       "Public Transport")
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:9] <- c("Baseline", "Bicycling",
-                                                        "Driving", "Public Transport")
+
+  n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
+
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Public Transport`, factor_key = TRUE)
+                      Baseline:`Motorcycling`, factor_key = TRUE)
   y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
@@
   #names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:9] <- c("Baseline", "Bicycling", 
-                                                     "Driving", "Public Transport")
+  #
+  n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Public Transport`, factor_key = TRUE)
+                      Baseline:`Motorcycling`, factor_key = TRUE)
   summary <- as.data.frame(data_long %>% group_by(scenario) %>%
@@
 #cl$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-co2_conc$Scenario <- c("Baseline", "Bicycling", "Driving", "Public Transport")
+co2_conc$Scenario <- scen_names
 
@@
   #names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-  names(io[[cities[x]]]$outcomes$mmets)[5:8] <- c("Baseline", "Bicycling", 
-                                                  "Driving", "Public Transport")
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Public Transport`, factor_key = TRUE)
+  #
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
+
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Motorcycling`, factor_key = TRUE)
   y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) +
@@
   #names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  names(io[[cities[x]]]$outcomes$mmets)[5:8] <- c("baseline","bike_scen", 
-                                                  "car_scen", "bus_scen")
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, baseline:bus_scen, factor_key = TRUE)
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
+
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:Motorcycling, factor_key = TRUE)
   summary <- as.data.frame(data_long %>% group_by(scenario) %>% 
@@
   #  as.data.frame() %>% tibble::rownames_to_column()
-  td <- round(colMeans(io[[city]]$outcomes$mmets[,5:8], na.rm = T), 1) %>%
+  #  
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
+
+  td <- round(colMeans(io[[city]]$outcomes$mmets[,(n_col - 4):n_col], na.rm = T), 1) %>%
     as.data.frame() %>% tibble::rownames_to_column()
@@
 
-#cl$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-cl$Scenario <- c("Baseline", "Bicycling", "Driving", "Public Transport")
+cl$Scenario <- scen_names
 
@@
   # PM 2.5 table and plot
-  # names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
-  #                                                       "Bicycling", "Driving",
-  #                                                       "Motorcycling", 
-  #                                                       "Public Transport")
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:9] <- c("Baseline",
-                                                        "Bicycling", "Driving",
-                                                        "Public Transport")
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
+
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
 
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Public Transport`, factor_key = TRUE)
+                      Baseline:`Motorcycling`, factor_key = TRUE)
 
@@
     as.data.frame() %>% tibble::rownames_to_column()
-  #names(td) <- c('Scenario', cities[x])
   names(td) <- c('Scenario', 'CO2')
-  # td$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling",
-  #                  "Public Transport")
-  td$Scenario <- c("Baseline", "Bicycling", "Driving", "Public Transport")
+  td$Scenario <- scen_names
 
@@
   #                                                  "Public Transport")
-  names(io[[cities[x]]]$outcomes$mmets)[5:8] <- c("Baseline",
-                                                   "Bicycling", "Driving",
-                                                   "Public Transport")
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Public Transport`, factor_key = TRUE)
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Motorcycling`, factor_key = TRUE)
   y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) + geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 

### 2022-10-11 summary_tables.Rmd
Adjust for four scenarios. Ref #

--- 2021-11-29 summary_tables.Rmd
@@
 library(data.table)
+library(writexl)
 })
+
+options(dplyr.summarise.inform = FALSE)
 
@@
 LATAM <- TRUE
-NSCEN <- 3
+NSCEN <- 4
 #cities <- c('accra','sao_paulo','buenos_aires')
@@
     scenario == "Scenario 3" ~ "BUS_SC",
+    scenario == "Scenario 4" ~ "MOT_SC",
     scenario == "Baseline" ~ "Baseline")) %>% 
@@
 injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths) == 'scen1_deaths_inj')
-ap_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
-pa_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
-
+ap_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))
+ap_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))
+pa_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))
+pa_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))
 sum_and_round_and_print <- function(data, text = '') {
@@
 }
-
 scen_names <- rownames(scen_prop)
-
 #yll_totals <- lapply(io,function(x){
@@
   temp <- sapply(1:NSCEN,function(y){
-    xx <- x$outcomes$hb$ylls
+    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
     xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
@@
   )
-
 yll_rates <- lapply(cities, function(x)
@@
   )
-
 names(yll_rates) <- cities
@@
   temp <- sapply(1:NSCEN,function(y){
-    xx <- x$outcomes$hb$ylls
+    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
     rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
@@
 })
-
 yll_rates <- lapply(cities, function(x)
@@
 injury_totals <- lapply(io[-2], function(x) {
-  xx <- x$outcomes$hb$ylls
+  xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
   injury_col <- grep("inj", names(xx))[1] # Dan: This line is mine (not sure is ok)
@@
 pa_totals <- lapply(io[-2],function(x){
-  xx <- x$outcomes$pathway_hb$ylls
+  xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
   xxx <- sapply(1:NSCEN, function(y){
-    xxx <- rowSums(xx[, pa_cols[seq(y, length(pa_cols), by = NSCEN)]])
+    xxx <- rowSums(xx[, pa_yll_cols[seq(y, length(pa_yll_cols), by = NSCEN)]])
     sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
@@
 })
-
 sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
@@
 ap_totals <- lapply(io[-2],function(x){
-  xx <- x$outcomes$pathway_hb$ylls
+  xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
   xxx <- sapply(1:NSCEN,function(y) {
-    xxx <- rowSums(xx[, ap_cols[seq(y, length(ap_cols), by = NSCEN)]])
+    xxx <- rowSums(xx[, ap_yll_cols[seq(y, length(ap_yll_cols), by = NSCEN)]])
     sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
@@
   )
-
 names(injury_rates) <- cities
@@
                                         length = NSCEN) * 100000))
-
 names(pa_rates) <- cities
@@
 disease_totals <- lapply(io[-2],function(x){
-  xx <- x$outcomes$hb$deaths
+  xx <- x$outcomes$hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
   xxx <- sapply(1:NSCEN,function(y){
@@
                                                length = NSCEN)*100000))
-
 names(disease_rates) <- cities
@@
 disease_totals <- lapply(io[-2],function(x){
-  xx <- x$outcomes$hb$ylls
+  xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
   xxx <- sapply(1:NSCEN,function(y) {
@@
 } # Loop for each measure
+# # Export deaths HB
+# path_deaths_wb <- "results/multi_city/health_impacts/Deaths.xlsx"
+# deaths_wb <- readWorkbook(path_deaths_wb)
+# wb <-  loadWorkbook(path_deaths_wb)
+# removeTable(wb = wb, sheet = "Deaths", table = "Tabla1")
+# writeDataTable(wb, sheet = "Deaths", x = rbindlist(health$deaths$hb, use.names = T), tableName = "Tabla1", startCol = 1)
+# removeTable(wb = wb, sheet = "Deaths_Pathway", table = "Tabla2")
+# writeDataTable(wb, sheet = "Deaths_Pathway", x = rbindlist(health$deaths$pathway_hb, use.names = T), tableName = "Tabla2", startCol = 1)
+# saveWorkbook(wb, path_deaths_wb, overwrite =TRUE)
+# 
+# # Export ylls HB
+# path_ylls_wb <- "results/multi_city/health_impacts/YLLs.xlsx"
+# ylls_wb <- readWorkbook(path_ylls_wb)
+# wb <-  loadWorkbook(path_ylls_wb)
+# removeTable(wb = wb, sheet = "YLLs", table = "Tabla1")
+# writeDataTable(wb, sheet = "YLLs", x = rbindlist(health$ylls$hb, use.names = T), tableName = "Tabla1", startCol = 1)
+# removeTable(wb = wb, sheet = "YLLs_Pathway", table = "Tabla2")
+# writeDataTable(wb, sheet = "YLLs_Pathway", x = rbindlist(health$ylls$pathway_hb, use.names = T), tableName = "Tabla2", startCol = 1)
+# saveWorkbook(wb, path_ylls_wb, overwrite =TRUE)
+# # Export ylls HB
+output_file <- "results/multi_city/health_impacts/data.xlsx"
+writexl::write_xlsx(list(ylls_hb_lng = rbindlist(health$ylls$hb, use.names = T),
+                         ylls_pathway_hb_lng = rbindlist(health$ylls$pathway_hb, use.names = T),
+                         Deaths = rbindlist(health$deaths$hb, use.names = T),
+                         Deaths_Pathway = rbindlist(health$deaths$pathway_hb, use.names = T)), path = output_file)
 
@@
             row.names = F)
+```
 
+## PA dist
+```{r}
+### pa distr
+cities <- names(io)[!names(io) %in% 'scen_prop']
+l <- list()
+for (city in cities){
+
+  # city <- "antofagasta"
+
+  td <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
+    rename("Baseline" = base_mmet, "CYC_SC" = scen1_mmet,
+           "CAR_SC" = scen2_mmet,  "BUS_SC" = scen3_mmet) %>% 
+    summarise(across(, ~ quantile(.x, c(0.05,0.10,0.25, 0.375, 0.50,0.75,0.90), na.rm = T)), 
+              prob = c(0.05,0.10,0.25, 0.375, 0.50,0.75,0.90)) %>%
+    pivot_longer(cols = -c(prob)) %>% mutate(city = city)
+
+  td1 <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
+    rename("Baseline" = base_mmet, "CYC_SC" = scen1_mmet,
+           "CAR_SC" = scen2_mmet,  "BUS_SC" = scen3_mmet) %>% 
+    summarise(across(, ~ mean(.x))) %>%  t() %>% as.data.frame() %>% 
+    tibble::rownames_to_column("name") %>% 
+    rename(value = V1) %>% 
+    mutate(prob = "mean") %>% mutate(city = city)
+
+  l[[city]] <- plyr::rbind.fill(td, td1)
+
+}
+write_csv(data.table::rbindlist(l), "results/multi_city/PA/pa_distr.csv")
 ```

### 2022-10-11 summary_tables.Rmd
initial outputs for Africa/India

@@
             row.names = F)
+
 ```

### 2022-10-12 summary_tables.Rmd
adding mock scenarios for africa and india; updating input parameters fro Africa

@@
 cities <- names(io)[!names(io) %in% 'scen_prop']
-LATAM <- TRUE
+LATAM <- F
+AFRICA_INDIA = T
 NSCEN <- 4

### 2022-10-12 R/create_africa_india_scenarios.R

@@ -1,0 +1,129 @@
+#' Create scenarios defined for latam paper
+#' 
+#' Creates three scenarios where, in each one, the mode share is elevated to the
+#' certain values in each distance band. The scenario-modes are cycle, car, and bus.
+#' 
+#' Motorcycle trips were added as well
+#' 
+#' @param trip_set data frame, baseline scenario
+#' 
+#' @return list of baseline scenario and three mode scenarios
+#' 
+#' @export
+create_africa_india_scenarios <- function(trip_set){
+  # I created this function based on "create_max_mode_share_scenarios" function.
+  rdr <- trip_set
+  trip_set <- NULL
+  
+  rd_list <- list()
+  # In this function SCENARIO_PROPORTIONS has the increase in % percentage points
+  # for each mode. These numbers came from computing the difference between
+  # median propensity in each distance band for cycling and the top value.
+  # We computed this numbers by hand, using the SCENARIO_PROPORTIONs that results
+  # in the function "create_max_mode_share_scenarios" and the median was computed
+  # by hand across cities. All of this is computed in the file "TablesAndFigures.xlsx"
+  # in dropbox.
+  scenario_proportions <- data.frame(c(5.6, 5.6, 0, 5.6),
+                                     c(10.2, 10.2, 10.2, 10.2),
+                                     c(2,2,2, 2))
+  colnames(scenario_proportions) <- target_distances <- DIST_CAT
+  rownames(scenario_proportions) <- modes <- c("cycle", "car", "bus", 'motorcycle')
+  SCENARIO_PROPORTIONS <<- scenario_proportions
+  # Baseline scenario
+  rd_list[[1]] <- rdr
+  modes_not_changeable <- c('bus_driver', 'truck', 'car_driver')
+  rdr_not_changeable <-  rdr %>% filter(trip_mode %in% modes_not_changeable)
+  rdr_changeable <-  rdr %>% filter(!trip_mode %in% modes_not_changeable) # Trips that can be reassigned to another mode
+  rdr <- NULL
+  
+  # Split trips by distance band in a new list
+  rdr_changeable_by_distance <- list()
+  for (j in 1:ncol(SCENARIO_PROPORTIONS)) {
+    target_distance <- target_distances[j]
+    rdr_changeable_by_distance[[j]] <- rdr_changeable %>% 
+      filter(trip_distance_cat == target_distance)
[…]

### 2022-10-12 R/get_scenario_settings.R

@@ -46,6 +46,9 @@ get_scenario_settings <- function(cities = c('accra', 'bangalore', 'belo_horizon
                 'osorno', 'puerto_montt', 'san_antonio',
                 'santiago', 'sao_paulo', 'temuco_padrelascasas', 'valdivia')
   
+  if (AFRICA_INDIA)
+    cities <- c('accra', 'cape_town','nairobi','kisumu','port_louis','bangalore', 'delhi', 'vizag')
+  
   min_distances <- as.numeric(sapply(distances,function(x)strsplit(x, "[^0-9.]+")[[1]][1]))
   mode_proportions <- mode_proportions_by_distance <- list()
   

### 2022-10-12 R/get_synthetic_from_trips.R

@@ -88,6 +88,8 @@ get_synthetic_from_trips <- function(){
     SYNTHETIC_TRIPS <- create_latam_scenarios(trip_set)
   }else if (GLOBAL) {
     SYNTHETIC_TRIPS <- create_global_scenarios(trip_set)
+  }else if (AFRICA_INDIA) {
+    SYNTHETIC_TRIPS <- create_africa_india_scenarios(trip_set)
   }else{
     SYNTHETIC_TRIPS <- create_all_scenarios(trip_set)
   }

### 2022-10-12 R/run_ithim_setup.R

@@ -124,7 +124,8 @@ run_ithim_setup <- function(seed = 1,
                             DISTANCE_SCALAR_CYCLING = 1,
                             DISTANCE_SCALAR_MOTORCYCLE = 1,
                             LATAM = FALSE,
-                            GLOBAL = FALSE){
+                            GLOBAL = FALSE,
+                            AFRICA_INDIA = FALSE){
   
   ## SUMMARY OF INPUTS
   # seed = double. sets seed to allow some reproducibility.
@@ -202,6 +203,7 @@ run_ithim_setup <- function(seed = 1,
   ## SET GLOBAL VALUES
   LATAM <<- LATAM
   GLOBAL <<- GLOBAL
+  AFRICA_INDIA <<- AFRICA_INDIA
   
   ## PROGRAMMING VARIABLES
   NSAMPLES <<- NSAMPLES

### 2022-10-12 multi_city_script.R

@@ -39,8 +39,8 @@ comment <- "Added CO2 emission sampling"
 # scenario definition
 max_mode_share_scenario <- F
 latam <- F
-global <- T
-africa_india <- F
+global <- F
+africa_india <- T
 test_walk_scenario <- F
 test_cycle_scenario <- F
 reference_scenario <- 'Baseline'
@@ -184,6 +184,7 @@ print(system.time(for(city in cities){
   ithim_objects[[city]] <- run_ithim_setup(
     LATAM = latam,
     GLOBAL = global,
+    AFRICA_INDIA = africa_india,
     DIST_CAT = as.character(dist_cat),
     ADD_WALK_TO_PT_TRIPS = as.logical(add_walk_to_pt_trips[[city]]),
     CITY = city,

### 2022-10-17 summary_tables.Rmd
Get NSCEN from io object

--- 2022-10-11 summary_tables.Rmd
@@
 cities <- names(io)[!names(io) %in% 'scen_prop']
-LATAM <- TRUE
-NSCEN <- 4
+LATAM <- F
+AFRICA_INDIA = T
+NSCEN <- nrow(io$scen_prop)
 #cities <- c('accra','sao_paulo','buenos_aires')
@@
   mutate(scenario = case_when(
-    # scenario == "Scenario 1" ~ "PED_SC", 
-    # scenario == "Scenario 2" ~ "CYC_SC",
-    # scenario == "Scenario 3" ~ "CAR_SC",
-    # scenario == "Scenario 4" ~ "MOT_SC",
-    # scenario == "Scenario 5" ~ "BUS_SC",
     scenario == "Scenario 1" ~ "CYC_SC", 

### 2022-10-17 summary_tables_PA_AP.Rmd
Read params from excel sheet

@@
 library(tidyverse)
+library(readxl)
 #library(plotly)
@@
 
-all_inputs <- read.csv('all_city_parameter_inputs.csv',stringsAsFactors = F)
+input_parameter_file <- "InputParameters_v15.0.xlsx"
+# scenario definition
+scenario_name <- "GLOBAL"
+reference_scenario <- 'Baseline'
 
-#all_inputs$cape_town <- all_inputs$accra
-#all_inputs$vizag <- all_inputs$sao_paulo
+compute_mode <- 'constant' # constant parameters from the given parameters
 
+all_inputs <- read_excel(input_parameter_file, sheet = "all_city_parameter_inputs")
+all_inputs[is.na(all_inputs)] <- ""
+all_inputs <- as.data.frame(all_inputs)
+
+# get input parameters into correct format
 parameter_names <- all_inputs$parameter
-parameter_starts <- which(parameter_names != '')
+parameter_starts <- which(parameter_names!='')
 parameter_stops <- c(parameter_starts[-1] - 1, nrow(all_inputs)) 
-parameter_names <- parameter_names[parameter_names != '']
+parameter_names <- parameter_names[parameter_names!='']
 parameter_list <- list()
-compute_mode <- 'constant'
-for (i in 1:length(parameter_names)) {
+
+for(i in 1:length(parameter_names)){
   parameter_list[[parameter_names[i]]] <- list()
-  parameter_index <- which(all_inputs$parameter == parameter_names[i])
-  if (all_inputs[parameter_index,2] == '') {
+  parameter_index <- which(all_inputs$parameter==parameter_names[i]) 
+  if(all_inputs[parameter_index,2]=='')  { 
     parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
-      city_index <- which(colnames(all_inputs) == x)
+      city_index <- which(colnames(all_inputs)==x)
       val <- all_inputs[parameter_index,city_index]
-      ifelse(val %in% c('T','F'),val,as.numeric(val))
+      ifelse(val%in%c('T','F'),val,ifelse(is.numeric(val), as.numeric(val), as.character(val)))
     })
     names(parameter_list[[parameter_names[i]]]) <- cities
-  }else if (all_inputs[parameter_index,2] == 'constant') {
+  }else if(all_inputs[parameter_index,2]=='constant'){
+    if (compute_mode != 'sample'){
       indices <- 0
-    if (compute_mode == 'sample') indices <- 1:2
       parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
-      city_index <- which(colnames(all_inputs) == x)
-      val <- all_inputs[parameter_index + indices,city_index]
-      ifelse(val == '',0,as.numeric(val))
+        city_index <- which(colnames(all_inputs)==x)
+        val <- all_inputs[parameter_index+indices,city_index]
+        ifelse(val=='',0,as.numeric(val))
       })
+    }
+    if(compute_mode=='sample'){ # if sampling from distribution, check that distribution parameters exist
+      parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
+        indices <- 1:2
+        city_index <- which(colnames(all_inputs)==x)  
+        val <- all_inputs[parameter_index+indices,city_index] 
+        if (val[1] == '' & val[2]==''){  # if no distribution parameters given in input file, read in constant value instead
+          indices <-0
+          city_index <- which(colnames(all_inputs)==x) 
+          val <- all_inputs[parameter_index+indices,city_index]} 
+        val <- as.numeric(val)
+      })
+    }
     names(parameter_list[[parameter_names[i]]]) <- cities
@@
     parameter_list[[parameter_names[i]]] <- lapply(cities,function(x) {
-      city_index <- which(colnames(all_inputs) == x)
-      if (any(all_inputs[parameter_starts[i]:parameter_stops[i],
-                         city_index] != '')) {
-        sublist_indices <- which(all_inputs[parameter_starts[i]:parameter_stops[i],city_index] != '')
+      city_index <- which(colnames(all_inputs)==x)
+      if(any(all_inputs[parameter_starts[i]:parameter_stops[i],city_index]!='')){
+        sublist_indices <- which(all_inputs[parameter_starts[i]:parameter_stops[i],city_index]!='')
         thing <- as.list(as.numeric(c(all_inputs[parameter_starts[i]:parameter_stops[i],city_index])[sublist_indices]))
@@
 
-for (i in 1:length(parameter_list))  assign(names(parameter_list)[i],parameter_list[[i]])
+list2env(parameter_list, environment())
+
+# read in global parameters
+
+all_global_inputs <- read_excel(input_parameter_file, sheet = "all_global_parameter_inputs")
+all_global_inputs[is.na(all_global_inputs)] <- ""
+all_global_inputs <- as.data.frame(all_global_inputs)
+
+# get input parameters into correct format
+global_parameter_names <- all_global_inputs$parameter
+global_parameter_starts <- which(global_parameter_names!='')
+global_parameter_stops <- c(global_parameter_starts[-1] - 1, nrow(all_global_inputs)) 
+global_parameter_names <- global_parameter_names[global_parameter_names!='']
+global_parameter_list <- list()
+
+for(i in 1:length(global_parameter_names)){
+  global_parameter_list[[global_parameter_names[i]]] <- list()
+  global_parameter_index <- which(all_global_inputs$parameter==global_parameter_names[i]) 
+  if(all_global_inputs[global_parameter_index,2]=='')  {
+
+    global_parameter_list[[global_parameter_names[i]]] <- all_global_inputs[global_parameter_index,'global']
+
+  }else if(all_global_inputs[global_parameter_index,2]=='constant'){
+    if (compute_mode != 'sample'){
+      global_parameter_list[[global_parameter_names[i]]] <- ifelse(all_global_inputs[global_parameter_index,'global']=='',
+                                                                   0,as.numeric(all_global_inputs[global_parameter_index,'global']))
+    }
+    else if(compute_mode=='sample'){ # if sampling from distribution, check that distribution parameters exist
+      indices <- 1:2
+      val <- all_global_inputs[global_parameter_index+indices,'global'] 
+      if (val[1] == '' & val[2]==''){  # if no distribution parameters given in input file, read in constant value instead
+        val <- all_global_inputs[global_parameter_index,'global']} 
+      val <- as.numeric(val)
+      global_parameter_list[[global_parameter_names[i]]] <- val
+    }
+  }
+}
+
+list2env(global_parameter_list, environment())
+
+dist_cat <- unlist(strsplit(gsub(" ", "", dist_cat, fixed = TRUE), "\\,"))
+
+outcome_age_min <- as.numeric(unlist(strsplit(gsub(" ", "", outcome_age_min, fixed = TRUE), "\\,")))
+outcome_age_max <- as.numeric(unlist(strsplit(gsub(" ", "", outcome_age_max, fixed = TRUE), "\\,")))
+outcome_age_groups <- unlist(strsplit(gsub(" ", "", outcome_age_groups, fixed = TRUE), "\\,"))
+
+min_age <- as.numeric(min_age)
+max_age <- as.numeric(max_age)
+
+day_to_week_scalar <- as.numeric(day_to_week_scalar)
+
 ```
@@
 {
+  print(cities[x])
   modes <- names(unlist(io[[cities[x]]]$PM_emission_inventory))
@@
   summary2 <- cbind(summary2$scenario, round(summary2[,2:5], digits = 1))
+  summary2$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, 2)
   summary2$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits = 2)

### 2022-10-17 summary_tables_PA_AP.Rmd
Add scenario_PM

@@
   summary <- cbind(summary$scenario ,round(summary[,2:5], digits = 1))
+  summary$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, digits=2)
   summary$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits=2)

### 2022-10-19 latex/output-md.Rmd

@@
 Non-travel PA, injury reporting rate and NCD burden all act as scalars for the corresponding datasets.
-Note that the non-travel PA scalar does not affect the $\approx$<!-- -->40% of the population whose non-travel PA is 0.
+Note that the non-travel PA scalar does not affect the $\approx$\<40% of the population whose non-travel PA is 0.
 
@@
 
-\$$\hat{B}_{g,m} \tilda L_m{B}_{a,g,m}.$$
+\$$\hat{B}_{g,m} \tilda L_m{B}_{a,g,m}.$\$
 
@@
 
-\$$\hat{B}_{a,g,m} & \sim \text{Beta}(\alpha_{a,g,m},\beta_{a,g,m}); \\$$
-
-${\hat{B}}_{a,g,m} \sim Beta({\alpha}_{a,g,m}, {\beta}_{a,g,m})$
-${\alpha}_{a,g,m} + {\beta}_{a,g,m} = 200$
-$\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}$
-
-\alpha_{a,g,m}+\beta_{a,g,m} & = 200; \\
-\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}.\end{aligned}$$ Here, $\tilde{B}_{a,g,m}$ is the new, variable parameter that we resample for each simulation.
+\$$\hat{B}_{a,g,m} & \sim \text{Beta}(\alpha_{a,g,m},\beta_{a,g,m}); \\$\$
+
+${\hat{B}}_{a,g,m} \sim Beta({\alpha}_{a,g,m}, {\beta}_{a,g,m})$ ${\alpha}_{a,g,m} + {\beta}_{a,g,m} = 200$ $\hat{B}_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}$
+
+\alpha*{a,g,m}+\*\beta{a,g,m} & = 200; \\ \hat{B}\_{a,g,m} & = \frac{1}{1+\frac{\beta_{a,g,m}}{\alpha_{a,g,m}}}.\\end{aligned}\$\$ Here, $\tilde{B}_{a,g,m}$ is the new, variable parameter that we resample for each simulation.
 We sample via a quantile, $\nu_m\sim\mathcal{U}(0,1)$, which we use to calculate $\tilde{B}_{a,g,m}$.

### 2022-10-21 summary_tables.Rmd
Ignore CIs in health outcomes - easier for pivot tables. Fix #120

@@
 # saveWorkbook(wb, path_ylls_wb, overwrite =TRUE)
-# # Export ylls HB
-output_file <- "results/multi_city/health_impacts/data.xlsx"
-writexl::write_xlsx(list(ylls_hb_lng = rbindlist(health$ylls$hb, use.names = T),
-                         ylls_pathway_hb_lng = rbindlist(health$ylls$pathway_hb, use.names = T),
-                         Deaths = rbindlist(health$deaths$hb, use.names = T),
-                         Deaths_Pathway = rbindlist(health$deaths$pathway_hb, use.names = T)), path = output_file)
+#
+
+# Remove CIs
+ylls <-  rbindlist(health$ylls$hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+deaths <- rbindlist(health$deaths$hb, use.names = T)|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
 
 # Export ylls HB
-write.csv(rbindlist(health$ylls$hb, use.names = T),
-            "results/multi_city/health_impacts/ylls_hb_lng.csv",
+output_file <- "results/multi_city/health_impacts/data.xlsx"
+writexl::write_xlsx(list(ylls = ylls,
+                         ylls_pathway = ylls_pathway,
+                         deaths = deaths,
+                         deaths_pathway = deaths_pathway), 
+                    path = output_file)
+
+# Export ylls HB
+write.csv(ylls,
+            "results/multi_city/health_impacts/ylls.csv",
             row.names = F)
 # Export ylls PATHWAY_HB
-write.csv(rbindlist(health$ylls$pathway_hb, use.names = T),
-            "results/multi_city/health_impacts/ylls_pathway_hb_lng.csv",
+write.csv(ylls_pathway,
+            "results/multi_city/health_impacts/ylls_pathway.csv",
             row.names = F)
 # Export deaths HB
-write.csv(rbindlist(health$deaths$hb, use.names = T),
-            "results/multi_city/health_impacts/deaths_hb_lng.csv",
+write.csv(deaths,
+            "results/multi_city/health_impacts/deaths.csv",
             row.names = F)
 # Export deaths PATHWAY_HB
-write.csv(rbindlist(health$deaths$pathway_hb, use.names = T),
-            "results/multi_city/health_impacts/deaths_pathway_hb_lng.csv",
+write.csv(deaths_pathway,
+            "results/multi_city/health_impacts/deaths_pathway",
             row.names = F)

### 2022-10-21 summary_tables.Rmd
Correct number of scenarios

@@
 ## Trip proportion
-Case study specific trip proportions by mode, for baseline and five scenarios
+Case study specific trip proportions by mode, for baseline and four scenarios
 
@@
 ## Distance tables
-Case study specific distance tables for baseline and five scenarios
+Case study specific distance tables for baseline and four scenarios
 
@@
 ## Distance tables scaled by using total population (per year in km)
-Case study specific distance tables for baseline and five scenarios
+Case study specific distance tables for baseline and four scenarios
 
@@
 
-Case study specific duration tables for baseline and five scenarios (per day in minutes)
+Case study specific duration tables for baseline and four scenarios (per day in minutes)
 

### 2022-10-23 summary_tables.Rmd
updated input file and corrected ouput file path

@@
 write.csv(deaths_pathway,
-            "results/multi_city/health_impacts/deaths_pathway",
+            "results/multi_city/health_impacts/deaths_pathway.csv",
             row.names = F)

### 2022-10-25 injury_tables.Rmd
tidy up and addition of baseline predicted and updated duration to injury_tables.Rmd

@@
       }else if(var == "car_sc"){
-        qualified_scen_name <- 'Driving'
+        qualified_scen_name <- 'Car'
       }else if(var == "mc_sc"){
@@
   }
+  # create df with location information
+  cityname <- city
+  if (city == cities[[1]]){
+    location <- data.frame(city = cityname, country = io[[city]]$location$country, continent = io[[city]]$location$continent)
+  } else {
+    dummy_row <- data.frame(city = cityname, country = io[[city]]$location$country, continent = io[[city]]$location$continent)
+    location <- rbind(location, dummy_row)
   }
+
+  # create dataframe with original fatality counts adjusted by weight and injury_reporting_rate
+
+  # Baseline - whw
+  injury_data_count <- as.data.frame(io[[city]]$injury_table$whw)
+  injury_data_count <- injury_data_count %>% rename(value = count, str_mode = strike_mode) # rename columns
+
+  # recalculate value based on weight and injury reporting rate
+  injury_data_count$value <- injury_data_count$value / injury_data_count$weight / injury_data_count$injury_reporting_rate
+  injury_data_count <- injury_data_count[c('str_mode', 'value', 'cas_mode')] # re order columns and drop column weight
+
+  # aggregate as age and gender column has been removed
+  injury_data_count2 <- injury_data_count %>% group_by(cas_mode, str_mode) %>% summarise(value = sum(value))
+
+  # Baseline - nov
+  injury_data_count_nov <- as.data.frame(io[[city]]$injury_table$nov)
+
+  if( length(injury_data_count_nov) > 0){
+    injury_data_count_nov <- injury_data_count_nov %>% rename(value = count, str_mode = strike_mode) # rename columns
+
+    # recalculate value based on weight and injury reporting rate
+    injury_data_count_nov$value <- injury_data_count_nov$value / injury_data_count_nov$weight / injury_data_count_nov$injury_reporting_rate
+    injury_data_count_nov <- injury_data_count_nov[c('str_mode', 'value', 'cas_mode')] # re order columns and drop column weight
+
+    # aggregate as age and gender column has been removed
+    injury_data_count_nov2 <- injury_data_count_nov %>% group_by(cas_mode, str_mode) %>% summarise(value = sum(value))
+
+    city_baseline_count <- rbind(injury_data_count2, injury_data_count_nov2)
+  } else {
+    city_baseline_count <- injury_data_count2
+  }
+
+  city_baseline_count$city <- city
+  city_baseline_count$scenario <- 'Baseline'
+  city_baseline_count$country <- io[[city]]$location$country
+  city_baseline_count$continent <- io[[city]]$location$continent
+
+  if (city == cities[[1]]){
+    baseline_counts <- city_baseline_count
+  } else {
+    baseline_counts <- rbind(baseline_counts, city_baseline_count)
+  }
+
+  baseline_counts$measure <- 'mean'
+
+  ###### create casualty distance df
+  dummy_distances <- io[[city]]$true_dist
+
+  dummy_distances2 <- pivot_longer(dummy_distances, cols = -c(stage_mode), names_to = 'scenario', values_to = 'mode_distance')
+
+  baseline_pred <- dummy_distances2 %>% filter(scenario == 'Baseline')
+  baseline_pred$scenario <- 'Baseline_predicted'
+
+  dummy_distances3 <- rbind(dummy_distances2, baseline_pred)
+
+  dummy_distances3$city <- city
+
+  if (city == cities[1]){
+    true_distances <- dummy_distances3
+  } else {
+    true_distances <- rbind(true_distances,dummy_distances3)
+  }
+
+  ####### create population df
+  pop_df <- data.frame(model_population = sum(io[[city]]$demographic$population))
+  pop_df$city <- city
+
+  if (city == cities[1]){
+    all_pop_df <- pop_df
+  } else {
+    all_pop_df <- rbind(all_pop_df, pop_df)
+  }
+
+  ####### create duration df based on true distances and speed
+  speed_city <- io[[city]]$vehicle_inventory %>% dplyr::select(-c( "PM_emission_inventory",  "CO2_emission_inventory"))
+  speed_city$city <- cityname
+
+  if (city == cities[1]){
+    speed_df <- speed_city
+  } else {
+    speed_df <- rbind(speed_df, speed_city)
+  }
+
+  # dummy_dur <- io[[city]]$dur
+  # 
+  # dummy_dur2 <- pivot_longer(dummy_dur, cols = -c(stage_mode), names_to = 'scenario', values_to = 'mode_duration')
+  # 
+  # baseline_pred <- dummy_dur2 %>% filter(scenario == 'Baseline')
+  # baseline_pred$scenario <- 'Baseline_predicted'
+  # 
+  # dummy_dur3 <- rbind(dummy_dur2, baseline_pred)
+  # 
+  # dummy_dur3$city <- city
+  # 
+  # if (city == cities[1]){
+  #   duration <- dummy_dur3
+  # } else {
+  #   duration <- rbind(duration,dummy_dur3)
+  # }
+}
+
+# calculate duration
+duration <- left_join(true_distances, speed_df, by = c('city','stage_mode' ))
+duration$mode_duration <- duration$mode_distance / duration$speed
+duration <- duration %>% dplyr::select(-c(mode_distance, speed))
+
+true_distances <- true_distances %>% rename(cas_mode = stage_mode)
 
@@
   }else if(cs == "car_sc"){
-    qualified_scen_name <- 'Driving'
+    qualified_scen_name <- 'Car'
   }else if(cs == "mc_sc"){
@@
 }
+
+# add country and continent information
+whw_list <- left_join(whw_list, location, by = 'city')
+
+# add original counts and rename Baseline to Baseline predicted
+whw_list$scenario <- ifelse(whw_list$scenario == 'Baseline', 'Baseline_predicted',whw_list$scenario)
+
+whw_list <- rbind(whw_list, baseline_counts)
+
+# replace capital with lower letters
+whw_list$str_mode <- tolower(whw_list$str_mode)
+
+# update true distances scenarios
+true_distances[true_distances$scenario == 'Scenario 1',]$scenario <- 'Bicycling'
+true_distances[true_distances$scenario == 'Scenario 3',]$scenario <- 'Public Transport'
+true_distances[true_distances$scenario == 'Scenario 2',]$scenario <- 'Car'
+true_distances[true_distances$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
+
+# add mode distances
+whw_list <- left_join(whw_list, true_distances, by = c('city', 'cas_mode','scenario'))
+
 # Remove total from whw long form, and write as csv
@@
 injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
-injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline_predicted'
 injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycling'
 injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Public Transport'
-injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Driving'
+injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car'
+injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
 rd <- rename(injury_risks_lng, mode = stage_mode)
 rd <- rd %>% filter(mode != 'Total')
+
+# add country and continent information
+rd <- left_join(rd, location, by = 'city')
+
+# aggregate baseline counts by cas_mode
+baseline_counts_agg <- baseline_counts %>% group_by(city,cas_mode, country, continent, measure, scenario) %>% summarise(value = sum(value)) %>% rename(mode = cas_mode)
+
+# get true_distances into correct format
+true_distances <- true_distances  %>% rename(mode = cas_mode)
+
+# add distance information
+rd <- left_join(rd, true_distances, by = c('mode', 'city','scenario'))
+
+# add counts but first divide by distance
+baseline_counts_km <- left_join(baseline_counts_agg, true_distances, by = c('mode', 'city', 'scenario'))
+baseline_counts_km$value <- round((baseline_counts_km$value /(baseline_counts_km$mode_distance * 365))*1000000000,4)
+
+rd <- rbind(rd, baseline_counts_km )
+
 readr::write_csv(rd, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms_lng.csv')
@@
 injury_risks_per_100k$scenario <- as.character(injury_risks_per_100k$scenario)
-injury_risks_per_100k[injury_risks_per_100k$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'bl',]$scenario <- 'Baseline_predicted'
 injury_risks_per_100k[injury_risks_per_100k$scenario == 'bisc',]$scenario <- 'Bicycling'
 injury_risks_per_100k[injury_risks_per_100k$scenario == 'bssc',]$scenario <- 'Public Transport'
-injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Driving'
+injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Car'
 injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
 #injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
+
+# add distance, population and location information
+injury_risks_per_100k <- left_join(injury_risks_per_100k, true_distances, by = c('mode', 'city','scenario'))
+injury_risks_per_100k <- left_join(injury_risks_per_100k, all_pop_df, by = c('city'))
+injury_risks_per_100k <- left_join(injury_risks_per_100k, location, by = 'city')
+
+# add counts but first divide by population
+baseline_counts_pop <- left_join(baseline_counts_agg, pop_df, by = c('city'))
+baseline_counts_pop$value <- round(baseline_counts_pop$value /baseline_counts_pop$model_population *  100000, 2)
+baseline_counts_pop <- left_join(baseline_counts_pop, true_distances, by = c('city', 'mode','scenario'))
+
+injury_risks_per_100k <- rbind(injury_risks_per_100k, baseline_counts_pop)
+
+# add country and continent information
+injury_risks_per_100k <- left_join(injury_risks_per_100k, location, by = 'city')
+
 readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/whw_matrices/injury_risks_per_100k_pop.csv')
@@
 ```{r }
-true_distances <- list()
+true_distances_city <- list()
 for (city in cities) { # Loop for each city
-  true_distances[[city]] <- io[[city]]$true_dist %>% 
+  true_distances_city[[city]] <- io[[city]]$true_dist %>% 
     pivot_longer(!stage_mode, names_to = "scenario", values_to = "distance") %>% 
@@
 }
-true_distances_df <- bind_rows(true_distances) %>% 
+true_distances_df <- bind_rows(true_distances_city) %>% 
   mutate(scenario = case_when(
     scenario == "Scenario 1" ~ "Bicycling", 
-    scenario == "Scenario 2" ~ "Driving",
+    scenario == "Scenario 2" ~ "Car",
     scenario == "Scenario 3" ~ "Public Transport",
@@
 injury_risks_lng_hours$scenario <- as.character(injury_risks_lng_hours$scenario)
-injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bl',]$scenario <- 'Baseline'
+injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bl',]$scenario <- 'Baseline_predicted'
 injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bisc',]$scenario <- 'Bicycling'
 injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bssc',]$scenario <- 'Public Transport'
-injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'csc',]$scenario <- 'Driving'
+injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'csc',]$scenario <- 'Car'
 rd_hours <- rename(injury_risks_lng_hours, mode = stage_mode)
 rd_hours <- rd_hours %>% filter(mode != 'Total')
+
+# add country and continent information
+rd_hours <- left_join(rd_hours, location, by = 'city')
+
+# add distance information
+rd_hours <- left_join(rd_hours, true_distances, by = c('city','mode','scenario'))
+
+# get true_distances into correct format
+duration <- duration  %>% rename(mode = stage_mode)
+duration[duration$scenario == 'Scenario 1',]$scenario <- 'Bicycling'
+duration[duration$scenario == 'Scenario 3',]$scenario <- 'Public Transport'
+duration[duration$scenario == 'Scenario 2',]$scenario <- 'Car'
+duration[duration$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
+
+# add duration information
+rd_hours <- left_join(rd_hours, duration, by = c('mode', 'city','scenario'))
+
+# add counts but first divide by duration
+baseline_counts_dur <- left_join(baseline_counts_agg, duration, by = c('mode', 'city', 'scenario'))
+baseline_counts_dur$value <- round((baseline_counts_dur$value /(baseline_counts_dur$mode_duration * 365))*100000000,4)
+baseline_counts_dur <- left_join(baseline_counts_dur,true_distances, by = c('city', 'mode','scenario'))
+
+rd_hours <- rbind(rd_hours, baseline_counts_dur )
+
 readr::write_csv(rd_hours, 'results/multi_city/whw_matrices/injury_risks_per_100million_h_lng.csv')

### 2022-10-26 summary_tables.Rmd
Use cities from cities var and set the scenario name

--- 2022-10-21 summary_tables.Rmd
@@
 LATAM <- F
-AFRICA_INDIA = T
+AFRICA_INDIA = F
+GLOBAL <- T
 NSCEN <- nrow(io$scen_prop)
@@
 ```{r message=F, error=F, warning=FALSE}
-get_scen_settings <- ithimr::get_scenario_settings()
+get_scen_settings <- ithimr::get_scenario_settings(cities = cities)
 
@@
 # saveWorkbook(wb, path_ylls_wb, overwrite =TRUE)
-#
-
-# Remove CIs
-ylls <-  rbindlist(health$ylls$hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-deaths <- rbindlist(health$deaths$hb, use.names = T)|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+# # Export ylls HB
+output_file <- "results/multi_city/health_impacts/data.xlsx"
+writexl::write_xlsx(list(ylls_hb_lng = rbindlist(health$ylls$hb, use.names = T),
+                         ylls_pathway_hb_lng = rbindlist(health$ylls$pathway_hb, use.names = T),
+                         Deaths = rbindlist(health$deaths$hb, use.names = T),
+                         Deaths_Pathway = rbindlist(health$deaths$pathway_hb, use.names = T)), path = output_file)
 
 # Export ylls HB
-output_file <- "results/multi_city/health_impacts/data.xlsx"
-writexl::write_xlsx(list(ylls = ylls,
-                         ylls_pathway = ylls_pathway,
-                         deaths = deaths,
-                         deaths_pathway = deaths_pathway), 
-                    path = output_file)
-
-# Export ylls HB
-write.csv(ylls,
-            "results/multi_city/health_impacts/ylls.csv",
+write.csv(rbindlist(health$ylls$hb, use.names = T),
+            "results/multi_city/health_impacts/ylls_hb_lng.csv",
             row.names = F)
 # Export ylls PATHWAY_HB
-write.csv(ylls_pathway,
-            "results/multi_city/health_impacts/ylls_pathway.csv",
+write.csv(rbindlist(health$ylls$pathway_hb, use.names = T),
+            "results/multi_city/health_impacts/ylls_pathway_hb_lng.csv",
             row.names = F)
 # Export deaths HB
-write.csv(deaths,
-            "results/multi_city/health_impacts/deaths.csv",
+write.csv(rbindlist(health$deaths$hb, use.names = T),
+            "results/multi_city/health_impacts/deaths_hb_lng.csv",
             row.names = F)
 # Export deaths PATHWAY_HB
-write.csv(deaths_pathway,
-            "results/multi_city/health_impacts/deaths_pathway",
+write.csv(rbindlist(health$deaths$pathway_hb, use.names = T),
+            "results/multi_city/health_impacts/deaths_pathway_hb_lng.csv",
             row.names = F)

### 2022-10-30 summary_tables.Rmd
Make scenario var global and use global scenario instead

--- 2022-10-23 summary_tables.Rmd
@@
 # Get names of cities from the io object
-cities <- names(io)[!names(io) %in% 'scen_prop']
-LATAM <- F
-AFRICA_INDIA = T
+cities <<- names(io)[!names(io) %in% 'scen_prop']
+LATAM <<- F
+AFRICA_INDIA <<- F
 NSCEN <- nrow(io$scen_prop)

### 2022-11-29 results/multi_city/health_impacts/health_pathway_tbls.qmd
Add qmd to create pivot tables. Fix #123

---
title: "Health Burden (per 100k population) by pathway"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    page-layout: custom
execute:
  warning: false
  echo: false
---

```{r}
#| message: false

library(tidyverse)
library(rpivotTable)
library(htmlwidgets)

rel_path <- ""

deaths_pathway <- read_csv(paste0(rel_path, "deaths_pathway.csv"))

ylls_pathway <- read_csv(paste0(rel_path, "ylls_pathway.csv"))

plot_pivottable <- function(data, colname = "level1"){
  data |> filter(!str_detect(cause, "lb|ub")) |>   filter((!is.na(!!rlang::sym(colname)))) |>
    group_by(city, scenario, dose, !!rlang::sym(colname)) |>
    summarise(measure_100k = sum(measure_100k)) |> rpivotTable(
      rows = c("city"),
      cols = c("scenario"),
      aggregatorName = "Sum",
      vals = "measure_100k",
      rendererName = "Horizontal Stacked Bar Chart", 
      subtotals = FALSE,
      height = "100%", 
      overflow = "scroll"
    )
}

# shared_deaths_pathway <- SharedData$new(deaths_pathway |> pivot_longer(cols = -c(select(deaths_pathway, -contains("lev")) |> names())))

```

::: {.panel-tabset}

<!-- ## Reactable -->
<!-- ```{r} -->

<!-- data <- shared_deaths_pathway -->

<!-- htmltools::browsable( -->
<!--   tagList( -->
<!--     div( -->
<!--       div(tags$label("Filter Type", `for` = "cars-type-filter")), -->
<!--       tags$select( -->
<!--         id = "cars-type-filter", -->
<!--         onchange = "Reactable.setFilter('cars-filter-table', 'level1', this.value)", -->
<!--         tags$option("Level", value = "level1"), -->
<!--         lapply(data$name |> unique() |> as.character(), tags$option) -->
<!--       ) -->
<!--     ), -->

<!--     tags$hr("aria-hidden" = "true"), -->

<!--     data$data() |> filter(!str_detect(cause, "lb|ub")) |> filter(!is.na(level1)) |>  group_by(city, scenario, dose, level1) |> -->
<!--       summarise(measure_100k = sum(measure_100k)) |> rpivotTable( -->
<!--         rows = c("city"), -->
<!--         cols = c("scenario"), -->
<!--         aggregatorName = "Sum", -->
<!--         vals = "measure_100k", -->
<!--         rendererName = "Horizontal Stacked Bar Chart", -->
<!--         subtotals = FALSE, -->
<!--         height = "80%", -->
<!--         overflow = "scroll" -->
<!--       ) -->
<!--   ) -->
<!-- ) -->
<!-- ``` -->

## Deaths - Level 1
```{r}
plot_pivottable(deaths_pathway, "level1")
```

## Deaths - Level 2
```{r}
plot_pivottable(deaths_pathway, "level2")
```

## Deaths - Level 3
```{r}
plot_pivottable(deaths_pathway, "level3")
```

## YLLs - Level 1
```{r}
plot_pivottable(ylls_pathway, "level1")
```

## YLLs - Level 2
```{r}
plot_pivottable(ylls_pathway, "level2")
```

## YLLs - Level 3
```{r}
plot_pivottable(ylls_pathway, "level3")
```

:::

### 2022-11-29 summary_tables.Rmd
Comment code to create health outcomes as an excel sheet

@@
 # Remove CIs
-ylls <-  rbindlist(health$ylls$hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-deaths <- rbindlist(health$deaths$hb, use.names = T)|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) |> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+deaths <- rbindlist(health$deaths$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
 
-# Export ylls HB
-output_file <- "results/multi_city/health_impacts/data.xlsx"
-writexl::write_xlsx(list(ylls = ylls,
-                         ylls_pathway = ylls_pathway,
-                         deaths = deaths,
-                         deaths_pathway = deaths_pathway), 
-                    path = output_file)
+# # Export ylls HB
+# output_file <- "results/multi_city/health_impacts/data.xlsx"
+# writexl::write_xlsx(list(ylls = ylls,
+#                          ylls_pathway = ylls_pathway,
+#                          deaths = deaths,
+#                          deaths_pathway = deaths_pathway), 
+#                     path = output_file)
 

### 2022-11-29 injury_tables.Rmd
Remove excel sheet code ouput

@@
 ```
-
-### Write all injury tables in a single excel sheet
-```{r}
-# Export injury datasets in a single excel sheet
-path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"
- writexl::write_xlsx(list(whw = whw_list %>% filter(str_mode != 'Total'),
-                          injury_risks_per_billion_kms = rd,
-                          injury_risks_per_100k  = injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'),
-                          injury_risks_per_100million_h = rd_hours,
-                          true_distances = true_distances_df),
-                     path = path_inj_wb)
-```

### 2022-11-29 injury_tables.Rmd
Added active travel mode

--- 2022-10-25 injury_tables.Rmd
@@
 
+# combine active travel modes
+active_travel <- whw_list %>% filter(cas_mode == 'cycle' | cas_mode == 'pedestrian')
+active_travel <- active_travel %>% group_by(str_mode, measure, city, scenario, country, continent) %>% summarise(
+                                  value = sum(value), mode_distance = sum(mode_distance), cas_mode ='active_travel')
+
 # Remove total from whw long form, and write as csv
@@
 
-      td5 <- td4
-
       td4[, 2] <- as.numeric(td4[,2])
+
+      # add active travel mode
+      # filter out pedestrian and cycle trips, calculate sum of counts and distances and add to existing data
+      active_travel <- td4 %>% filter(stage_mode == 'pedestrian' | stage_mode == 'cycle')
+      active_travel <- active_travel %>% group_by() %>% summarise(stage_mode = 'active_travel', 
+                                                                  x = sum(active_travel[,2]),
+                                                                  y = sum(active_travel[,3]))
+      colnames(active_travel) <- colnames(td4)
+
+      td4 <- rbind(td4, active_travel)
+
+      td5 <- td4 # save copy for risk per population
+
       # Compute risk per bn km travelled
@@
 
-# add counts but first divide by distance
+# add baseline counts
+
+# merge baseline count with distance
 baseline_counts_km <- left_join(baseline_counts_agg, true_distances, by = c('mode', 'city', 'scenario'))
+
+# add active travel 
+active_travel <- baseline_counts_km %>% filter(mode == 'cycle' | mode == 'pedestrian')
+active_travel <- active_travel %>% group_by(city, scenario, country, continent, measure, ) %>% summarise(
+                                   value = sum(value), mode_distance = sum(mode_distance), mode ='active_travel')
+
+baseline_counts_km <- rbind(baseline_counts_km, active_travel)
+
+# calculate value per distance
 baseline_counts_km$value <- round((baseline_counts_km$value /(baseline_counts_km$mode_distance * 365))*1000000000,4)
@@
 
+       # add active travel mode
+      # filter out pedestrian and cycle trips, calculate sum of counts and distances and add to existing data
+      active_travel <- td4 %>% filter(stage_mode == 'pedestrian' | stage_mode == 'cycle')
+      active_travel <- active_travel %>% group_by() %>% summarise(stage_mode = 'active_travel', 
+                                                                  x = sum(active_travel[,2]))
+      colnames(active_travel) <- colnames(td4)
+
+      td4 <- rbind(td4, active_travel)
+
       td5 <- td4
@@
 
-      el <- inner_join(el, td4, by = 'stage_mode')
+      el <- full_join(el, td4, by = 'stage_mode')
     }
@@
 baseline_counts_pop <- left_join(baseline_counts_agg, pop_df, by = c('city'))
+
+# add active travel 
+active_travel <- baseline_counts_pop %>% filter(mode == 'cycle' | mode == 'pedestrian')
+active_travel <- active_travel %>% group_by(city, scenario, country, continent, measure ) %>% summarise(
+                                   value = sum(value), model_population = sum(model_population), 
+                                   mode ='active_travel')
+
+baseline_counts_pop <- rbind(baseline_counts_pop, active_travel)
+
 baseline_counts_pop$value <- round(baseline_counts_pop$value /baseline_counts_pop$model_population *  100000, 2)
@@
 
-      td5 <- td4
-
       td4[, 2] <- as.numeric(td4[,2])
+
+      # add active travel mode
+      # filter out pedestrian and cycle trips, calculate sum of counts and distances and add to existing data
+      active_travel <- td4 %>% filter(stage_mode == 'pedestrian' | stage_mode == 'cycle')
+      active_travel <- active_travel %>% group_by() %>% summarise(stage_mode = 'active_travel', 
+                                                                  x = sum(active_travel[,2]),
+                                                                  y = sum(active_travel[,3]))
+      colnames(active_travel) <- colnames(td4)
+
+      td4 <- rbind(td4, active_travel)
+
       # Compute risk per 100 million hours travelled
@@
       el <- full_join(el, td4, by = c('stage_mode'))
-
-      # City's population
-      pop <- sum(io[[city]]$demographic$population)
-      # Compute travel time travelled per capita
-      td6 <- td5
-      td6[[paste0(var,"_risk")]] <- unlist(td4[2])
-      td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop
-      names(td6)[3] <- paste(names(td6)[2], names(td6)[3], sep = "_")
-      dd <- full_join(dd, td6, by = 'stage_mode')
     }
@@
     td$measure <- ifelse(is.na(pref), "mean", pref)
-    dd$measure <- ifelse(is.na(pref), "mean", pref)
 
     city_df[["td"]][[city]][[unique(td$measure)]] <- td
-    city_df[["dd"]][[city]][[unique(td$measure)]] <- dd
 
@@
       overall_el[[city]] <- data.table::rbindlist(city_df[["td"]][[city]])
-      overall_dd[[city]] <- data.table::rbindlist(city_df[["dd"]][[city]])
          }
@@
 baseline_counts_dur <- left_join(baseline_counts_agg, duration, by = c('mode', 'city', 'scenario'))
+
+# add active travel 
+active_travel <- baseline_counts_dur %>% filter(mode == 'cycle' | mode == 'pedestrian')
+active_travel <- active_travel %>% group_by(city, scenario, country, continent, measure ) %>% summarise(
+                                   value = sum(value), mode_duration = sum(mode_duration), 
+                                   mode ='active_travel')
+
+baseline_counts_dur <- rbind(baseline_counts_dur, active_travel)
+
 baseline_counts_dur$value <- round((baseline_counts_dur$value /(baseline_counts_dur$mode_duration * 365))*100000000,4)

### 2022-11-29 results/multi_city/health_impacts/health_pathway_tbls.qmd
Change layout of results and remove unused code

@@
     summarise(measure_100k = sum(measure_100k)) |> rpivotTable(
-      rows = c("city"),
-      cols = c("scenario"),
+      cols = c("city"),
+      rows = c("dose","scenario"),
       aggregatorName = "Sum",
       vals = "measure_100k",
-      rendererName = "Horizontal Stacked Bar Chart", 
+      rendererName = "Stacked Bar Chart", 
+      inclusions = list(scenario = list("BUS_SC")),
+      exclusions= list(city = list("buenos_aires", "santiago")),
       subtotals = FALSE,
@@
 
-# shared_deaths_pathway <- SharedData$new(deaths_pathway |> pivot_longer(cols = -c(select(deaths_pathway, -contains("lev")) |> names())))
-
 ```
@@
 ::: {.panel-tabset}
-
-<!-- ## Reactable -->
-<!-- ```{r} -->
-
-<!-- data <- shared_deaths_pathway -->
-
-<!-- htmltools::browsable( -->
-<!--   tagList( -->
-<!--     div( -->
-<!--       div(tags$label("Filter Type", `for` = "cars-type-filter")), -->
-<!--       tags$select( -->
-<!--         id = "cars-type-filter", -->
-<!--         onchange = "Reactable.setFilter('cars-filter-table', 'level1', this.value)", -->
-<!--         tags$option("Level", value = "level1"), -->
-<!--         lapply(data$name |> unique() |> as.character(), tags$option) -->
-<!--       ) -->
-<!--     ), -->
-
-<!--     tags$hr("aria-hidden" = "true"), -->
-
-<!--     data$data() |> filter(!str_detect(cause, "lb|ub")) |> filter(!is.na(level1)) |>  group_by(city, scenario, dose, level1) |> -->
-<!--       summarise(measure_100k = sum(measure_100k)) |> rpivotTable( -->
-<!--         rows = c("city"), -->
-<!--         cols = c("scenario"), -->
-<!--         aggregatorName = "Sum", -->
-<!--         vals = "measure_100k", -->
-<!--         rendererName = "Horizontal Stacked Bar Chart", -->
-<!--         subtotals = FALSE, -->
-<!--         height = "80%", -->
-<!--         overflow = "scroll" -->
-<!--       ) -->
-<!--   ) -->
-<!-- ) -->
-<!-- ``` -->
 

### 2022-12-05 multi_city_script.R
Update result/docs for v20. Previous relevant commit: 37b8f5d43690b0e0aa1e859b73caf245d1958aef

@@ -30,7 +30,7 @@ cities <- c('antofagasta', 'arica', 'belo_horizonte', 'bogota', 'buenos_aires',
 
 # number of times input values are sampled from each input parameter distribution
 
-input_parameter_file <- "InputParameters_v19.0.xlsx"
+input_parameter_file <- "InputParameters_v20.0.xlsx"
 
 output_version <- "v0.3" # gives the version number of the output documents, independent of the input parameter file name
 author <- "AA"

### 2022-12-08 results/multi_city/health_impacts/health_pathway_tbls.qmd
Use overall ylls and deaths instead of pathway specific.

@@
 
-deaths_pathway <- read_csv(paste0(rel_path, "deaths_pathway.csv"))
+deaths_pathway <- read_csv(paste0(rel_path, "deaths.csv"))
 
-ylls_pathway <- read_csv(paste0(rel_path, "ylls_pathway.csv"))
+ylls_pathway <- read_csv(paste0(rel_path, "ylls.csv"))
 
@@
     group_by(city, scenario, dose, !!rlang::sym(colname)) |>
-    summarise(measure_100k = sum(measure_100k)) |> rpivotTable(
+    summarise(metric_100k = sum(metric_100k)) |> rpivotTable(
       cols = c("city"),
@@
       aggregatorName = "Sum",
-      vals = "measure_100k",
+      vals = "metric_100k",
       rendererName = "Stacked Bar Chart", 
       inclusions = list(scenario = list("BUS_SC")),
-      exclusions= list(city = list("buenos_aires", "santiago")),
+      # exclusions= list(city = list("buenos_aires", "santiago")),
       subtotals = FALSE,

### 2022-12-09 summary_tables.Rmd
Rename measure to metric for health outcomes

@@
         left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>% 
-        mutate(measure_100k = measure / pop_age * 100000,
-               measeure_100k_sex = measure / pop_age_sex * 100000)
+        mutate(metric_100k = measure / pop_age * 100000,
+               metric_100k_sex = measure / pop_age_sex * 100000)
 

### 2023-01-30 results_app/app.R
Commenting it out as it creates duplicated geographies

@@ -4,6 +4,7 @@ library(tidyverse)
 library(gridlayout)
 library(shinyWidgets)
 library(bslib)
+library(readxl)
 #library(shiny.semantic)
 
 options(scipen = 10000)
@@ -21,6 +22,14 @@ deaths_pathway$measures <- "Deaths"
 injury_risks_per_billion_kms_lng <- read_csv("../results/multi_city/whw_matrices/injury_risks_per_billion_kms_lng.csv")
 injury_risks_per_100k_pop <- read_csv("../results/multi_city/whw_matrices/injury_risks_per_100k_pop.csv")
 injury_risks_per_100million_h_lng <- read_csv("../results/multi_city/whw_matrices/injury_risks_per_100million_h_lng.csv")
+
+
+# Input params
+input_parameter_file_path <- "../InputParameters_v24.0.xlsx"
+city_input_params <- read_excel(input_parameter_file_path, sheet = "all_city_parameter_inputs")
+global_input_params <- read_excel(input_parameter_file_path, sheet = "all_global_parameter_inputs")
+
+# output$table <- DT::renderDataTable(DT::datatable({
 
 ren_scen <- function(df){
   df |> 
@@ -116,7 +125,8 @@ ui <- grid_page(
     tabsetPanel(
       id = "main_tab",
       tabPanel("Health Outcomes", plotlyOutput("in_pivot", width = "100%", height = "100%")),
-      tabPanel("Injury Risks", plotlyOutput("in_inj_pivot", width = "100%", height = "100%"))
+      tabPanel("Injury Risks", plotlyOutput("in_inj_pivot", width = "100%", height = "100%")),
+      tabPanel("Params", DT::dataTableOutput("input_params"))
     )
   ),
   grid_card(
@@ -170,8 +180,8 @@ ui <- grid_page(
                    label = "Risk Type: ",
                    choices = inj_risk_types,
                    selected = inj_risk_types[1])
-    )
-    
+    ),
+    downloadButton("download_top_data", "Download data", icon = shiny::icon("file-download"))
     
   )
   
@@ -234,32 +244,8 @@ server <- function(input, output, session) {
     filtered_cities <- cities |> filter(city %in% input$in_cities) |> dplyr::select(city) |> pull()
     filtered_modes <- input$in_inj_modes
     
-    local_df <- injury_risks_per_billion_kms_lng
-    ylab <- "Distance: risk per Billion kilometeres"
[…]

### 2023-02-01 injury_tables.Rmd
Commenting it out as it creates duplicated geographies
Added the option to have 3, 4 or 5 scenarios with fixed scenario order

@@
 smodes <- read_csv('data/global/modes/standardized_modes.csv')
+
+# NOTE: the following scenario order is assumed: 'bi_sc', 'car_sc', 'bus_sc','mc_sc','w_sc' and if there are less than 5 scenarios, the last scenarios can be missing but all of the scenarios in the middle need to exist.
+e.g.
+if you want the public transport and the walking scenario, you still need to create a car, motorcycling and bicycling scenario!
 
@@
 scen_names <- rownames(scen_prop)
-# scen_names <- c('Walking', 'Bicycling', 'Driving', 'Public Transport', 'Motorcycling')
 
@@
 
-#scen_names_w_baseline <- c('bl', 'w_sc', 'bi_sc', 'car_sc', 'mc_sc', 'bus_sc')
+scen_names_w_baseline <- c('bl',  'bi_sc', 'car_sc', 'bus_sc','mc_sc','w_sc')
+if (NSCEN == 4){
   scen_names_w_baseline <- c('bl', 'bi_sc', 'car_sc', 'bus_sc', 'mc_sc')
+} else if (NSCEN == 3){
+  scen_names_w_baseline <- c('bl', 'bi_sc', 'car_sc', 'bus_sc')
+}
 scen_lt <- data.frame("scen_name" = names(io[[1]]$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)
@@
 true_distances[true_distances$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
+
+true_distances[true_distances$scenario == 'Scenario 5',]$scenario <- 'Walking'
 
@@
 injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car'
+if (NSCEN == 4){
   injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+}
+if (NSCEN == 5){
+  injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+    injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
+}
 rd <- rename(injury_risks_lng, mode = stage_mode)
@@
 injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Car'
+if (NSCEN == 4){
   injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-#injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
+}
+if (NSCEN == 5){
+  injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+  injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
+}
 
@@
 # add country and continent information
-injury_risks_per_100k <- left_join(injury_risks_per_100k, location, by = 'city')
+# injury_risks_per_100k <- left_join(injury_risks_per_100k, location, by = 'city')
 
@@
     scenario == "Scenario 4" ~ "Motorcycling",
+    scenario == "Scenario 5" ~ "Walking",
     scenario == "Baseline" ~ "Baseline"))
@@
 injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'csc',]$scenario <- 'Car'
+if (NSCEN == 4){
+  injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+}
+if (NSCEN == 5){
+  injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+    injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'wsc',]$scenario <- 'Walking'
+}
 rd_hours <- rename(injury_risks_lng_hours, mode = stage_mode)
@@
 duration[duration$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
+duration[duration$scenario == 'Scenario 5',]$scenario <- 'Walking'
 

### 2023-03-17 code/synthetic_population/bogota_synth_pop.qmd
Add quarto doc reading the latest bogota travel survey from fc8bae5381440fcb96e6789d26d115305597e5f1
ref #128

---
title: "Synthetic Population for Bogota"
author: "Ali Abbas"
editor: visual
format:
  html:
    light: flatly
    dark: darkly
    toc: true
    embed-resources: true
reference-location: margin
citation-location: margin
execute:
  echo: false
  message: false
  warning: false
number-sections: true
crossref:
  chapters: true
---

## Load library
Load all the require pacakges
```{r}
library(rsample)
library(tidyverse)
library(summarytools)

# Set seed
set.seed(2023)

```

## Read latest bogoto travel survey with weights

Daniel has kindly added `participation weights` to each row in the travel dataset, which represents household weights

```{r, echo=TRUE}
bt <- read_csv("https://raw.githubusercontent.com/ITHIM/TravelSurveyPreprocessing/fc8bae5381440fcb96e6789d26d115305597e5f1/Data/ITHIM/bogota/trips_bogota.csv")

view(dfSummary(bt |> filter(age != 9999)))

# rts <- bt |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, 1))
# rss <- bt |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(raw_stage_freq = round(n / sum(n) * 100, 1))
# 
# et <- bt %>% mutate(pw = round(participant_wt - 1))
# exp1 <- et  |> filter(pw > 0) |> uncount(pw, .id = "pid")
# exp2 <- et  |> filter(pw > 0) |> filter(pw == 0) %>% mutate(pid = 1)
# 
# view(dfSummary(exp1 |> filter(age != 9999)))
# 
# ets <- exp1 |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
# ess <- exp1 |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(exp_stage_freq = round(n / sum(n) * 100, 1))
# 
# ts <- full_join(rts, ets)
# ss <- left_join(rss, ess)
# 
# st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode))
# ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode))
# 
# view(dfSummary(st))
# view(dfSummary(ss))
# 
# st |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
# 
# rd <- plyr::rbind.fill(exp1, exp2) %>% arrange(hh_id, person_id, trip_id)

```

### 2023-03-21 code/synthetic_population/bogota_synth_pop.qmd
Improve layout of summarytools. Ref #128

@@
     embed-resources: true
-reference-location: margin
-citation-location: margin
+# reference-location: margin
+# citation-location: margin
 execute:
@@
 
-## Load library
-Load all the require pacakges
+# Load library
+
+Load all the require packages
+
 ```{r}
@@
 set.seed(2023)
+st_options(plain.ascii = FALSE)
 
@@
 
-## Read latest bogoto travel survey with weights
+# Read latest bogoto travel survey with weights
 
@@
 
-```{r, echo=TRUE}
+```{r, results='asis'}
 bt <- read_csv("https://raw.githubusercontent.com/ITHIM/TravelSurveyPreprocessing/fc8bae5381440fcb96e6789d26d115305597e5f1/Data/ITHIM/bogota/trips_bogota.csv")
 
-view(dfSummary(bt |> filter(age != 9999)))
+raw_dataset_without_weights <- bt |> filter(age != 9999)
 
-# rts <- bt |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, 1))
-# rss <- bt |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(raw_stage_freq = round(n / sum(n) * 100, 1))
-# 
-# et <- bt %>% mutate(pw = round(participant_wt - 1))
-# exp1 <- et  |> filter(pw > 0) |> uncount(pw, .id = "pid")
-# exp2 <- et  |> filter(pw > 0) |> filter(pw == 0) %>% mutate(pid = 1)
-# 
-# view(dfSummary(exp1 |> filter(age != 9999)))
-# 
-# ets <- exp1 |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
-# ess <- exp1 |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(exp_stage_freq = round(n / sum(n) * 100, 1))
-# 
-# ts <- full_join(rts, ets)
-# ss <- left_join(rss, ess)
-# 
-# st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode))
-# ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode))
-# 
-# view(dfSummary(st))
-# view(dfSummary(ss))
+print(dfSummary(raw_dataset_without_weights, 
+                varnumbers   = FALSE, 
+                valid.col    = FALSE, 
+                graph.magnif = 0.76),
+      method = 'render')
+
+# dfSummary(bt |> filter(age != 9999), plain.ascii = FALSE, style = "grid")
+
+# print(dfSummary(dfSummary(bt |> filter(age != 9999))), method = "render")
+
+rts <- bt |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, 1))
+rss <- bt |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(raw_stage_freq = round(n / sum(n) * 100, 1))
+
+et <- bt %>% mutate(pw = round(participant_wt - 1))
+exp1 <- et  |> filter(pw > 0) |> uncount(pw, .id = "pid")
+exp2 <- et  |> filter(pw > 0) |> filter(pw == 0) %>% mutate(pid = 1)
+
+print(dfSummary(exp1 |> filter(age != 9999),
+                varnumbers   = FALSE, 
+                valid.col    = FALSE, 
+                graph.magnif = 0.76),
+      method = 'render')
+
+ets <- exp1 |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
+ess <- exp1 |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(exp_stage_freq = round(n / sum(n) * 100, 1))
+
+ts <- full_join(rts, ets)
+ss <- left_join(rss, ess)
+
+st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode))
+ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode))
+
+print(dfSummary(st,
+                varnumbers   = FALSE, 
+                valid.col    = FALSE, 
+                graph.magnif = 0.76),
+      method = 'render')
+
+print(dfSummary(ss,
+                varnumbers   = FALSE, 
+                valid.col    = FALSE, 
+                graph.magnif = 0.76),
+      method = 'render')
 # 

### 2023-03-22 code/synthetic_population/bogota_synth_pop.qmd
Use arrow and remove unused code. Ref #128

@@
 library(summarytools)
+library(arrow)
 
@@
 Daniel has kindly added `participation weights` to each row in the travel dataset, which represents household weights
+
+## Summary without weights
 
@@
 
-rts <- bt |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, 1))
-rss <- bt |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(raw_stage_freq = round(n / sum(n) * 100, 1))
+# rts <- bt |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, 1))
+# rss <- bt |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(raw_stage_freq = round(n / sum(n) * 100, 1))
+# 
+# ets <- exp1 |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
+# ess <- exp1 |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(exp_stage_freq = round(n / sum(n) * 100, 1))
+# 
+# ts <- full_join(rts, ets)
+# ss <- left_join(rss, ess)
 
-et <- bt %>% mutate(pw = round(participant_wt - 1))
-exp1 <- et  |> filter(pw > 0) |> uncount(pw, .id = "pid")
-exp2 <- et  |> filter(pw > 0) |> filter(pw == 0) %>% mutate(pid = 1)
+# 
+# st |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
+# 
+# rd <- plyr::rbind.fill(exp1, exp2) %>% arrange(hh_id, person_id, trip_id)
 
-print(dfSummary(exp1 |> filter(age != 9999),
+```
+
+## Summary with weights - expanded travel survey
+
+```{r, results='asis'}
+
+exp1 <- read_parquet("../../data/local/bogota/exp_trips.parquet")
+# Remove undefined ages
+exp1 <- exp1 |> filter(age != 9999)
+
+# exp2 <- et  |> filter(pw > 0) |> filter(pw == 0) %>% mutate(pid = 1)
+
+print(dfSummary(exp1,
                 varnumbers   = FALSE, 
@@
 
-ets <- exp1 |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
-ess <- exp1 |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(exp_stage_freq = round(n / sum(n) * 100, 1))
+```
 
-ts <- full_join(rts, ets)
-ss <- left_join(rss, ess)
+## Summary of sampled travel survey - with trip_mode strata
 
+```{r, results='asis'}
 st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode))
-ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode))
 
@@
 
+```
+
+## Summary of sampled travel survey - with stage_mode strata
+
+```{r, results='asis'}
+ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode))
 print(dfSummary(ss,
@@
       method = 'render')
-# 
-# st |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
-# 
-# rd <- plyr::rbind.fill(exp1, exp2) %>% arrange(hh_id, person_id, trip_id)
 

### 2023-03-23 code/synthetic_population/bogota_synth_pop.qmd
Add comaprison with census population - and remove distribution of sampled. Ref #128

--- 2023-03-17 code/synthetic_population/bogota_synth_pop.qmd
@@
 ---
-title: "Synthetic Population for Bogota"
+title: "Travel Survey and Census population for Bogota"
 author: "Ali Abbas"
@@
 
-## Load library
-Load all the require pacakges
 ```{r}
@@
 library(summarytools)
+library(arrow)
+library(knitr)
 
@@
 
-## Read latest bogoto travel survey with weights
+# Compare Travel Survey and Census Population by age group
 
-Daniel has kindly added `participation weights` to each row in the travel dataset, which represents household weights
+In this section, we compare population in the Travel Survey with Census population by sex and five year age groups.
 
-```{r, echo=TRUE}
+## Distribution of `five year age groups`
+
+The table below compares the proportion of population in Census and Travel Survey for Bogota by five year age groups.
+
+| Age Group (5-year band) | Population (N) | Percent (%) | Travel Survey Population (N) | Percent (%) |
+|:--------------|:--------------|:--------------|:--------------|:--------------|
+| 15-19                   | 562,016        | 9.8         | 3,775                        | 9.8         |
+| 20-24                   | 718,131        | 12.6        | 4,750                        | 12.3        |
+| 25-29                   | 724,454        | 12.7        | 4,296                        | 11.1        |
+| 30-34                   | 651,833        | 11.4        | 3,831                        | 9.9         |
+| 35-39                   | 599,112        | 10.5        | 3,662                        | 9.5         |
+| 40-44                   | 519,983        | 9.1         | 3,278                        | 8.5         |
+| 45-49                   | 470,334        | 8.2         | 3,100                        | 8           |
+| 50-54                   | 465,196        | 8.1         | 3,395                        | 8.8         |
+| 55-59                   | 423,099        | 7.4         | 3,253                        | 8.4         |
+| 60-64                   | 336,100        | 5.9         | 3,000                        | 7.8         |
+| 65-69                   | 246,821        | 4.3         | 2,307                        | 6           |
+
+## Distribution of `sex`
+
+The table below compares the distribution of population in Census and Travel Survey by `sex`
+
+| Sex    | Population (N) | Percent (%) | Travel Survey Population (N) | Percent (%) |
+|--------|----------------|-------------|------------------------------|-------------|
+| Female | 2,996,444      | 52.4        | 17,423                       | 52.8        |
+| Male   | 2,720,635      | 47.6        | 15,568                       | 47.2        |
+
+**Takeaway**: We show the overall numbers and percentages for both of these sources.
+Overall, percentages of both sources are relatively similar and strengthen our confidence in the representativenesss of the household travel survey as a proxy for the overall census population.
+
+# Compare Travel Survey distribution with and without household weights
+
+Daniel has kindly added `weights` to each row in the Travel Survey dataset, which represents household weights.
+In this section we compare distribution based on `trip_mode`, `stage_mode` and `sex` distribution with and without the household weights.
+In all the tables below, the column `raw_trip_freq` shows distribution [**without**]{.underline} and `weighted_trip_freq` [**with**]{.underline} `weights`.
+
+**Takeaway**: Similarly with Census, when we compare the distribution in Travel Survey with and without `weights`, they all are relatively similar across all three variables: `trip_mode`, `stage_mode` and `sex`.
+
+```{r}
+
+calc_freq <- function(x, y, mode = "trip_mode", round_to = 1){
+
+  grx <- x |> filter(!is.na(across(mode))) |> group_by(across(all_of(mode))) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, round_to))
+
+  gry <- y |> filter(!is.na(across(mode))) |> group_by(across(all_of(mode))) |> summarise(n = n()) |> mutate(weighted_trip_freq = round(n / sum(n) * 100, round_to))
+
+  return(left_join(grx, gry, by = mode) |> dplyr::select(-all_of(starts_with("n"))))
+
+}
+
 bt <- read_csv("https://raw.githubusercontent.com/ITHIM/TravelSurveyPreprocessing/fc8bae5381440fcb96e6789d26d115305597e5f1/Data/ITHIM/bogota/trips_bogota.csv")
 
-view(dfSummary(bt |> filter(age != 9999)))
+raw_dataset_without_weights <- bt |> filter(age != 9999)
 
-# rts <- bt |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(raw_trip_freq = round(n / sum(n) * 100, 1))
-# rss <- bt |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(raw_stage_freq = round(n / sum(n) * 100, 1))
-# 
-# et <- bt %>% mutate(pw = round(participant_wt - 1))
-# exp1 <- et  |> filter(pw > 0) |> uncount(pw, .id = "pid")
-# exp2 <- et  |> filter(pw > 0) |> filter(pw == 0) %>% mutate(pid = 1)
-# 
-# view(dfSummary(exp1 |> filter(age != 9999)))
-# 
-# ets <- exp1 |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
-# ess <- exp1 |> filter(!is.na(stage_mode)) |> group_by(stage_mode) |> summarise(n = n()) |> mutate(exp_stage_freq = round(n / sum(n) * 100, 1))
-# 
-# ts <- full_join(rts, ets)
-# ss <- left_join(rss, ess)
-# 
-# st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode))
-# ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode))
-# 
-# view(dfSummary(st))
-# view(dfSummary(ss))
-# 
-# st |> filter(!is.na(trip_mode)) |> group_by(trip_mode) |> summarise(n = n()) |> mutate(exp_trip_freq = round(n / sum(n) * 100, 1))
-# 
-# rd <- plyr::rbind.fill(exp1, exp2) %>% arrange(hh_id, person_id, trip_id)
+exp1 <- read_parquet("../../data/local/bogota/exp_trips.parquet")
+# Remove undefined ages
+exp1 <- exp1 |> filter(age != 9999)
 
 ```
+
+## Distribution of `trip_mode`
+
+Comparison of distribution of `trip_mode` with and without weights
+
+```{r, results='asis'}
+
+kable(calc_freq(raw_dataset_without_weights, exp1, mode = "trip_mode", round_to = 2))
+
+```
+
+## Distribution of `stage_mode`
+
+Comparison of distribution of `stage_mode` with and without weights
+
+```{r, results='asis'}
+
+kable(calc_freq(raw_dataset_without_weights, exp1, mode = "stage_mode", round_to = 2))
+```
+
+## Distribution of `sex`
+
+Comparison of distribution of `sex` with and without weights
+
+```{r, results='asis'}
+
+kable(calc_freq(raw_dataset_without_weights, exp1, mode = "sex", round_to = 2))
+```
+
+<!-- ## Summary of sampled travel survey - with trip_mode strata -->
+
+<!-- ```{r, results='asis'} -->
+
+<!-- st <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = trip_mode)) -->
+
+<!-- kable(calc_freq(raw_dataset_without_weights, st, mode = "trip_mode", round_to = 2)) -->
+
+<!-- ``` -->
+
+<!-- ## Summary of sampled travel survey - with stage_mode strata -->
+
+<!-- ```{r, results='asis'} -->
+
+<!-- ss <- rsample::training(rsample::initial_split(exp1, prop = 0.1, strata = stage_mode)) -->
+
+<!-- kable(calc_freq(raw_dataset_without_weights, ss, mode = "stage_mode", round_to = 2)) -->
+
+<!-- ``` -->

### 2023-03-29 code/synthetic_population/bogota_synth_pop.qmd
Move takeaway message to the front. Ref #128
Rename household weights as individual weights. Ref #128

@@
 
+**Takeaway**: We show the overall numbers and percentages for both of these sources.
+Overall, percentages of both sources are relatively similar and strengthen our confidence in the representativenesss of the household travel survey as a proxy for the overall census population.
+
 ## Distribution of `five year age groups`
@@
 | Sex    | Population (N) | Percent (%) | Travel Survey Population (N) | Percent (%) |
-|--------|----------------|-------------|------------------------------|-------------|
+|-------------|-------------|-------------|---------------------|-------------|
 | Female | 2,996,444      | 52.4        | 17,423                       | 52.8        |
@@
 
-**Takeaway**: We show the overall numbers and percentages for both of these sources.
-Overall, percentages of both sources are relatively similar and strengthen our confidence in the representativenesss of the household travel survey as a proxy for the overall census population.
+# Compare Travel Survey distribution with and without individual weights
 
-# Compare Travel Survey distribution with and without household weights
-
-Daniel has kindly added `weights` to each row in the Travel Survey dataset, which represents household weights.
-In this section we compare distribution based on `trip_mode`, `stage_mode` and `sex` distribution with and without the household weights.
+Daniel has kindly added `weights` to each row in the Travel Survey dataset, which represents individual weights.
+In this section we compare distribution based on `trip_mode`, `stage_mode` and `sex` distribution with and without the individual weights.
 In all the tables below, the column `raw_trip_freq` shows distribution [**without**]{.underline} and `weighted_trip_freq` [**with**]{.underline} `weights`.

### 2023-04-28 summary_tables.Rmd
Update scenario names

@@
 
-```{r warning=FALSE, echo=F, message=FALSE}
-
-dat <- list()
-
-for (city in cities) {
-
-dat[[city]] <- io[[city]]$trip_scen_sets %>% 
-  mutate(scenario = case_when(
-    scenario == "Scenario 1" ~ "CYC_SC", 
-    scenario == "Scenario 2" ~ "CAR_SC",
-    scenario == "Scenario 3" ~ "BUS_SC",
-    scenario == "Scenario 4" ~ "MOT_SC",
-    scenario == "Baseline" ~ "Baseline")) %>% 
-  group_by(scenario, trip_mode, trip_distance_cat) %>% 
-  summarise(n = dplyr::n(), dist = sum(trip_distance, na.rm = T)) %>% 
-  mutate(city = city)
-
-td <- dat[[city]]
-
-for (i in 1:nrow(td)) {
-    val <- td %>% filter(scenario == "Baseline", trip_mode == td$trip_mode[i] & 
-                           trip_distance_cat == td$trip_distance_cat[i] &
-                           city == td$city[i])
-
-    if (nrow(val) > 0) {
-      td$diff[i] <- val$n - td$n[i]
-      td$trip_prop[i] <- round(td$n[i] / val$n, 2)
-      td$dist_prop[i] <- round(td$dist[i] / val$dist, 2)
-
-    }
-}
-
-dat[[city]] <- td
-
-}
-
-write_csv(bind_rows(dat), "results/multi_city/scen_prop.csv")
-
-```
+# ```{r warning=FALSE, echo=F, message=FALSE}
+# 
+# dat <- list()
+# 
+# for (city in cities) {
+#   
+# dat[[city]] <- io[[city]]$trip_scen_sets %>% 
+#   mutate(scenario = case_when(
+#     scenario == "Scenario 1" ~ "CYC_SC", 
+#     scenario == "Scenario 2" ~ "CAR_SC",
+#     scenario == "Scenario 3" ~ "BUS_SC",
+#     scenario == "Scenario 4" ~ "MOT_SC",
+#     scenario == "Baseline" ~ "Baseline")) %>% 
+#   group_by(scenario, trip_mode, trip_distance_cat) %>% 
+#   summarise(n = dplyr::n(), dist = sum(trip_distance, na.rm = T)) %>% 
+#   mutate(city = city)
+# 
+# td <- dat[[city]]
+# 
+# for (i in 1:nrow(td)) {
+#     val <- td %>% filter(scenario == "Baseline", trip_mode == td$trip_mode[i] & 
+#                            trip_distance_cat == td$trip_distance_cat[i] &
+#                            city == td$city[i])
+#     
+#     if (nrow(val) > 0) {
+#       td$diff[i] <- val$n - td$n[i]
+#       td$trip_prop[i] <- round(td$n[i] / val$n, 2)
+#       td$dist_prop[i] <- round(td$dist[i] / val$dist, 2)
+#       
+#     }
+# }
+# 
+# dat[[city]] <- td
+# 
+# }
+# 
+# write_csv(bind_rows(dat), "results/multi_city/scen_prop.csv")
+# 
+# ```
 
@@
   td <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
-    rename("Baseline" = base_mmet, "CYC_SC" = scen1_mmet,
-           "CAR_SC" = scen2_mmet,  "BUS_SC" = scen3_mmet) %>% 
+    rename("Baseline" = base_mmet, "CYC_SC" = sc_cycle_mmet,
+           "CAR_SC" = sc_car_mmet,  "BUS_SC" = sc_bus_mmet,
+           "MOT_SC" = sc_motorcycle_mmet) %>% 
     summarise(across(, ~ quantile(.x, c(0.05,0.10,0.25, 0.375, 0.50,0.75,0.90), na.rm = T)), 
@@
   td1 <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
-    rename("Baseline" = base_mmet, "CYC_SC" = scen1_mmet,
-           "CAR_SC" = scen2_mmet,  "BUS_SC" = scen3_mmet) %>% 
+     rename("Baseline" = base_mmet, "CYC_SC" = sc_cycle_mmet,
+           "CAR_SC" = sc_car_mmet,  "BUS_SC" = sc_bus_mmet,
+           "MOT_SC" = sc_motorcycle_mmet) %>% 
     summarise(across(, ~ mean(.x))) %>%  t() %>% as.data.frame() %>% 
@@
 }
-write_csv(data.table::rbindlist(l), "results/multi_city/PA/pa_distr.csv")
+# write_csv(data.table::rbindlist(l), "results/multi_city/PA/pa_distr.csv")
 ```

### 2023-05-12 summary_tables_PA_AP.Rmd
Move back rmd docs to the root
updated input file and corrected ouput file path
Regen rmd docs
Update result/docs for v20. Previous relevant commit: 37b8f5d43690b0e0aa1e859b73caf245d1958aef
Regen using v23 InputParams
Update rmd docs for v24
Use v28 input params

@@
 
-input_parameter_file <- "InputParameters_v15.0.xlsx"
+input_parameter_file <- "InputParameters_v28.0.xlsx"
 # scenario definition
@@
 
-<!-- # Descriptive tables of emission inventory -->
-
-<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
-
-<!-- trans_share<-c(22.5, 40.0, 22.5, 28.1)  -->
-<!-- pm_conc<- c(50, 18, 122, 47 ) -->
-<!-- city<-cities -->
-<!-- shares<- as.data.frame(cbind(city,as.numeric(trans_share), pm_conc)) -->
-<!-- names(shares)[2]<- "trans_share" -->
-<!-- for (x in 1:length(cities)) -->
-<!-- { -->
-<!--   modes<-names(unlist(io[[cities[x]]]$emission_inventory)) -->
-<!--   emissions<-as.data.frame(unlist(io[[cities[x]]]$emission_inventory)) -->
-<!--   city_emissions<-cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$emission_inventory)`)) -->
-<!--   names(city_emissions)[2]<- "emissions" -->
-<!--   select<- c("car", "motorcycle", "bus_driver", "truck", "big_truck") -->
-<!--   city_emissions$modes<- as.character(city_emissions$modes) -->
-<!--   city_emissions$modes[!(city_emissions$modes %in% select)]<- "other" -->
-<!--   summary<-city_emissions %>% group_by(modes) %>% summarise(sum(emissions)) -->
-<!--   names(summary)[2]<- "emissions" -->
-<!--   summary$emissions <- round(summary$emissions*100/sum(summary$emissions), digits=1) -->
-<!--   summary<-as.data.frame(summary) -->
-<!--   summary[nrow(summary)+1,1]<-"Transport share" -->
-<!--   summary[nrow(summary),2]<-as.character(shares$trans_share[x]) -->
-<!--   summary[nrow(summary)+1,1]<-"PM2.5 Conc" -->
-<!--   summary[nrow(summary),2]<-as.character(shares$pm_conc[x]) -->
-<!--   print(kable(summary, caption= cities[x])) -->
-
-<!-- } -->
-
-<!-- ``` -->
-
 # Descriptive tables of PM 2.5 emission inventory
@@
 
-# summary_all<- cbind(io[[cities[1]]]$summary_emission,io[[cities[2]]]$summary_emission,io[[cities[3]]]$summary_emission,io[[cities[4]]]$summary_emission)
-# summary_all<- summary_all[-c(3,5,7)]
-# summary_all$modes[2]<- "bus"
-# 
-# 
 print(kable(sl))
@@
 
-#cl$Scenario <- c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
 co2_conc$Scenario <- scen_names
@@
 for (x in 1:length(cities)) {
-  #names(io[[cities[x]]]$outcomes$mmets)[5:10]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-  #
   n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
@@
 ```
-
-<!-- # Descriptive tables of injury outcomes -->
-
-<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
-<!-- for (x in 1: length(cities)) -->
-<!-- { -->
-
-<!--   summary<- as.data.frame(io[[cities[x]]]$outcomes$injuries %>% group_by(scenario) %>% summarise('Ped'=sum(pedestrian,na.rm=T),'Bike'= sum(cycle), 'Car'=sum(car), "Motorcycle"=sum(motorcycle,na.rm=T), "All modes"=sum(Deaths,na.rm=T))) -->
-<!--   summary$scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen") -->
-<!--   summary<- cbind(summary$scenario ,round(summary[,2:6], digits=0)) -->
-<!--   names(summary)[1]<-"Scenario" -->
-<!--   print(kable(summary, caption= cities[x])) -->
-
-<!-- } -->
-<!-- ``` -->
-
-<!-- # Striking vehicles in scenarios -->
-
-<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
-<!-- for (x in 1: length(cities)) -->
-<!-- { -->
-
-<!-- scenarios<- c('Baseline', 'Scenario 1', 'Scenario 2', 'Scenario 3', 'Scenario 4', 'Scenario 5') -->
-<!-- for (i in 1: length(scenarios)) -->
-<!-- { -->
-<!-- whw<-as.data.frame(rowSums(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$whw)) -->
-<!-- noov<-as.data.frame(sum(io[[cities[x]]]$outcomes$whw[[scenarios[i]]]$nov)) -->
-<!-- names(whw)[1]<- as.character(scenarios[i]) -->
-<!-- names(noov)[1]<-as.character(scenarios[i]) -->
-
-<!-- if (i ==1 ) -->
-<!-- { -->
-<!--  summary<-round(rbind(whw,nov=noov), digits = 1) -->
-<!-- }  -->
-<!-- else -->
-<!-- { -->
-<!-- summary<- cbind(summary, round(rbind(whw, nov=noov), digits=1) )  -->
-<!-- } -->
-
-<!-- } -->
-
-<!-- names(summary)<-c("base", "walk", "bike", "car", "MC", "bus") -->
-<!-- print(kable(summary, caption= cities[x])) -->
-
-<!-- } -->
-
-<!-- ``` -->
-
-<!-- # Distance share by scenarios in percentages (excludes city-specific minority modes-- auto rickshaws, subway, etc.) -->
-
-<!-- ```{r, message=FALSE, warning=FALSE, echo=FALSE} -->
-<!-- for (x in 1: length(cities)) -->
-<!-- { -->
-
-<!--   select<- c("car", "motorcycle", "walking", "cycle", "bus") -->
-
-<!--   io[[cities[x]]]$trip_scen_sets$trip_mode[!io[[cities[x]]]$trip_scen_sets$trip_mode %in% select]<- "other" -->
-
-<!--   dist_scen<- io[[cities[x]]]$trip_scen_sets %>% group_by(trip_mode, scenario) %>% summarise(sum(trip_distance)) -->
-<!--   dist_scen<- as.data.frame (dist_scen) -->
-<!--   names(dist_scen)[3]<-"distance" -->
-<!--   dist_scen$distance<- round(dist_scen$distance, digits=0) -->
-<!--     dist_scen<-spread(dist_scen, trip_mode, distance) -->
-<!--   dist_scen$sum<-rowSums (dist_scen[2:(ncol(dist_scen))], na.rm = FALSE, dims = 1) -->
-<!--   dist_scen_prop<- round(dist_scen[,2:(ncol(dist_scen))]*100/dist_scen$sum, digits=1) -->
-<!--   dist_scen_prop<- cbind(dist_scen[,1], dist_scen_prop) -->
-
-<!--   names(dist_scen_prop)[1]<-"Scenario" -->
-<!--   dist_scen_prop$Scenario<- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen") -->
-<!--   print(kable(dist_scen_prop, caption= cities[x])) -->
-<!--   print(kable(dist_scen, caption= cities[x])) -->
-
-<!-- } -->
-<!-- ``` -->
 
@@
 for (x in 1:length(cities)) {
-  # names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("Baseline","Walking",
-  #                                                  "Bicycling", "Driving",
-  #                                                  "Motorcycling", 
-  #                                                  "Public Transport")
   n_col <- ncol(io[[cities[x]]]$outcomes$mmets)

### 2023-05-12 summary_tables.Rmd
Update scenario name

@@
   df <- io[[city]]$trip_scen_sets
-  u_trips <- df %>% dplyr::filter(scenario == "Baseline") %>% 
+  u_trips <- df %>% dplyr::filter(scenario == "baseline") %>% 
     summarise(uid = n_distinct(trip_id)) %>% as.numeric()

### 2023-05-12 injury_tables.Rmd
Update scen names

@@
 
+# Get number of scenarios
+NSCEN <- length(names(io[[1]]$outcomes$whw)) - 1
+
+# Get scenario names
+scen_names <- names(io[[1]]$outcomes$whw)
+
+# Remove extra bits of info - irrelevant outputs for injury related outputs
+# This makes it the io object lighter
 for (city in cities){
@@
   }
+}
+
+get_qualified_scen_name <- function(cs){
+  qualified_scen_name <- ""
+  if (cs == 'baseline')
+    qualified_scen_name <- 'Baseline'
+  else if(cs == "sc_walk")
+    qualified_scen_name <- 'Walking'
+  else if(cs == "sc_cycle")
+    qualified_scen_name <- 'Bicycling'
+  else if(cs == "sc_car")
+    qualified_scen_name <- 'Car'
+  else if(cs == "sc_motorcycle")
+    qualified_scen_name <- 'Motorcycling'
+  else if(cs == "sc_bus")
+    qualified_scen_name <- 'Public Transport'
+
+  return(qualified_scen_name)
 }
@@
 
-scen_names_w_baseline <- c('bl',  'bi_sc', 'car_sc', 'bus_sc','mc_sc','w_sc')
-if (NSCEN == 4){
-  scen_names_w_baseline <- c('bl', 'bi_sc', 'car_sc', 'bus_sc', 'mc_sc')
-} else if (NSCEN == 3){
-  scen_names_w_baseline <- c('bl', 'bi_sc', 'car_sc', 'bus_sc')
-}
-scen_lt <- data.frame("scen_name" = names(io[[1]]$outcomes$whw), "val" = scen_names_w_baseline, stringsAsFactors = FALSE)
-
 all_whw <- list()
@@
 
-  # city <- cities[1]
-
   cat('\n')
@@
 
-  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$Baseline), value = T), 2, sep = "_"))){ # if upper and lower bounds are given, this is NA, ub, lb
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){ # if upper and lower bounds are given, this is NA, ub, lb
     # pref <- "ub"
@@
     for (cs in names(io[[city]]$outcomes$whw)){ # Baseline, Scenario 1, Scenario 2, etc.
-      # cs <- 'Baseline'
+      # cs <- 'baseline'
 
@@
 
-      var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
-
-      if (var == 'bl'){
-        qualified_scen_name <- 'Baseline'
-        scen <- ''
-      }else if(var == "w_sc"){
-        qualified_scen_name <- 'Walking'
-      }else if(var == "bi_sc"){
-        qualified_scen_name <- 'Bicycling'
-      }else if(var == "car_sc"){
-        qualified_scen_name <- 'Car'
-      }else if(var == "mc_sc"){
-        qualified_scen_name <- 'Motorcycling'
-      }else if(var == "bus_sc"){
-        qualified_scen_name <- 'Public Transport'
-      }
+      qualified_scen_name <- get_qualified_scen_name(cs)
 
@@
 
-      city_df[[var]][[city]][[unique(td$measure)]] <- td   # save all data in one list called city_df
-
-      if (length(city_df[[var]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
-        all_whw[[var]][[city]] <- data.table::rbindlist(city_df[[var]][[city]]) # combine mean, lb and ub counts into one table for each scenario
+      city_df[[cs]][[city]][[unique(td$measure)]] <- td   # save all data in one list called city_df
+
+      if (length(city_df[[cs]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
+        all_whw[[cs]][[city]] <- data.table::rbindlist(city_df[[cs]][[city]]) # combine mean, lb and ub counts into one table for each scenario
     }
@@
 
-  if (cs == 'bl'){
-    qualified_scen_name <- 'Baseline'
-    scen <- ''
-  }else if(cs == "w_sc"){
-    qualified_scen_name <- 'Walking'
-  }else if(cs == "bi_sc"){
-    qualified_scen_name <- 'Bicycling'
-  }else if(cs == "car_sc"){
-    qualified_scen_name <- 'Car'
-  }else if(cs == "mc_sc"){
-    qualified_scen_name <- 'Motorcycling'
-  }
+ qualified_scen_name <- get_qualified_scen_name(cs)
 
@@
 # update true distances scenarios
-true_distances[true_distances$scenario == 'Scenario 1',]$scenario <- 'Bicycling'
-true_distances[true_distances$scenario == 'Scenario 3',]$scenario <- 'Public Transport'
-true_distances[true_distances$scenario == 'Scenario 2',]$scenario <- 'Car'
-true_distances[true_distances$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
-
-true_distances[true_distances$scenario == 'Scenario 5',]$scenario <- 'Walking'
+true_distances <- true_distances |> rowwise() |> mutate(scenario = get_qualified_scen_name(scenario))
 
@@
 
-  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$Baseline), value = T), 2, sep = "_"))){
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){
 
@@
     for (cs in names(io[[city]]$outcomes$whw)){
-      # cs <- 'Baseline'
-      if (length(word(names(io[[city]]$outcomes$whw$Baseline), 1, sep = "_") %>% unique()) == 2){
+      # cs <- 'baseline'
+      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){
         # td1: Number of fatalities in nov
@@
 
-      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){
         td3 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
         #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){
 
@@
 
-      var <- scen_lt %>% filter(scen_name == cs) %>% dplyr::select(val) %>% as.character()
+      var <- get_qualified_scen_name(cs)
 
@@
       td6 <- td5
-      td6[[paste0(var,"_risk")]] <- unlist(td4[2])
+      td6[[paste0(var,"_risk")]] <- td4[2]
       td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop
@@
 #readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks_per_100k_kms.csv')
-colnames(td) = gsub("bi_sc", "bisc", colnames(td))
-colnames(td) = gsub("bus_sc", "bssc", colnames(td))
-colnames(td) = gsub("car_sc", "csc", colnames(td))
-colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
-colnames(td) = gsub("w_sc", "wsc", colnames(td))
+# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
+# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
+# colnames(td) = gsub("car_sc", "csc", colnames(td))
+# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
+# colnames(td) = gsub("w_sc", "wsc", colnames(td))
 injury_risks_b <- td
@@
 #readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')
-colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))
-colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))
-colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))
-colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))
-colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))
+# colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))
+# colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))
+# colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))
+# colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))
+# colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))
 injury_risks_lng <- reshape2::melt(injury_risks_b)
@@
 injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
-injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline_predicted'
-injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycling'
-injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Public Transport'
-injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car'
-if (NSCEN == 4){
-  injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-}
-if (NSCEN == 5){
-  injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-    injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
-}
+# injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline_predicted'
+# injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycling'
+# injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Public Transport'
+# injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car'
+# if (NSCEN == 4){
+#   injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+# }
+# if (NSCEN == 5){
+#   injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+#     injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
+# }
 rd <- rename(injury_risks_lng, mode = stage_mode)
@@
     for (cs in names(io[[city]]$outcomes$whw)){
-      if (length(word(names(io[[city]]$outcomes$whw$Baseline), 1, sep = "_") %>% unique()) == 2){
+      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){
         td1 <- (io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
@@
 
-      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'whw'){
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){
         td3 <- colSums((io[[city]]$outcomes$whw[[cs]][[whw_qn]])) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      }else if(length(names(io[[city]]$outcomes$whw$Baseline)) == 1 && names(io[[city]]$outcomes$whw$Baseline) == 'nov'){
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){
 
@@
 td <- rename(td, mode = stage_mode)
-colnames(td) = gsub("bi_sc", "bisc", colnames(td))
-colnames(td) = gsub("bus_sc", "bssc", colnames(td))
-colnames(td) = gsub("car_sc", "csc", colnames(td))
-colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
-colnames(td) = gsub("w_sc", "wsc", colnames(td))
+# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
+# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
+# colnames(td) = gsub("car_sc", "csc", colnames(td))
+# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
+# colnames(td) = gsub("w_sc", "wsc", colnames(td))
 injury_risks_per_100k <- reshape2::melt(td)
@@
 injury_risks_per_100k$scenario <- as.character(injury_risks_per_100k$scenario)
-injury_risks_per_100k[injury_risks_per_100k$scenario == 'bl',]$scenario <- 'Baseline_predicted'
-injury_risks_per_100k[injury_risks_per_100k$scenario == 'bisc',]$scenario <- 'Bicycling'
-injury_risks_per_100k[injury_risks_per_100k$scenario == 'bssc',]$scenario <- 'Public Transport'
-injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Car'
-if (NSCEN == 4){
-  injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-}
-if (NSCEN == 5){
-  injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-  injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
-}
+# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bl',]$scenario <- 'Baseline_predicted'
+# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bisc',]$scenario <- 'Bicycling'
+# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bssc',]$scenario <- 'Public Transport'
+# injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Car'
+# if (NSCEN == 4){
+#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+# }
+# if (NSCEN == 5){
+#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
+# }
 
@@
 }
-true_distances_df <- bind_rows(true_distances_city) %>% 
-  mutate(scenario = case_when(
-    scenario == "Scenario 1" ~ "Bicycling", 
-    scenario == "Scenario 2" ~ "Car",
-    scenario == "Scenario 3" ~ "Public Transport",
-    scenario == "Scenario 4" ~ "Motorcycling",
-    scenario == "Scenario 5" ~ "Walking",
-    scenario == "Baseline" ~ "Baseline"))
+true_distances_df <- bind_rows(true_distances_city) # %>% 
+  # mutate(scenario = case_when(
+  #   scenario == "Scenario 1" ~ "Bicycling", 
+  #   scenario == "Scenario 2" ~ "Car",
+  #   scenario == "Scenario 3" ~ "Public Transport",
+  #   scenario == "Scenario 4" ~ "Motorcycling",
+  #   scenario == "Scenario 5" ~ "Walking",
+  #   scenario == "Baseline" ~ "Baseline"))
 readr::write_csv(true_distances_df, 'results/multi_city/whw_matrices/distances.csv')
@@
       # cs <- 'Baseline'
-      if (length(word(names(io[[city]]$outcomes$whw$Baseline), 1, sep = "_") %>% unique()) == 2){ # if there exist nov and whw matrices
+      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){ # if there exist nov and whw matrices
         # td1: Number of fatalities in nov
@@
 
-colnames(td) = gsub("bi_sc", "bisc", colnames(td))
-colnames(td) = gsub("bus_sc", "bssc", colnames(td))
-colnames(td) = gsub("car_sc", "csc", colnames(td))
-colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
-colnames(td) = gsub("w_sc", "wsc", colnames(td))
+# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
+# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
+# colnames(td) = gsub("car_sc", "csc", colnames(td))
+# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
+# colnames(td) = gsub("w_sc", "wsc", colnames(td))
 injury_risks_100mil_hours <- td
 
-colnames(injury_risks_100mil_hours) = gsub("bi_sc", "bisc", colnames(injury_risks_100mil_hours))
-colnames(injury_risks_100mil_hours) = gsub("bus_sc", "bssc", colnames(injury_risks_100mil_hours))
-colnames(injury_risks_100mil_hours) = gsub("car_sc", "csc", colnames(injury_risks_100mil_hours))
-colnames(injury_risks_100mil_hours) = gsub("mc_sc", "mcsc", colnames(injury_risks_100mil_hours))
-colnames(injury_risks_100mil_hours) = gsub("w_sc", "wsc", colnames(injury_risks_100mil_hours))
+# colnames(injury_risks_100mil_hours) = gsub("bi_sc", "bisc", colnames(injury_risks_100mil_hours))
+# colnames(injury_risks_100mil_hours) = gsub("bus_sc", "bssc", colnames(injury_risks_100mil_hours))
+# colnames(injury_risks_100mil_hours) = gsub("car_sc", "csc", colnames(injury_risks_100mil_hours))
+# colnames(injury_risks_100mil_hours) = gsub("mc_sc", "mcsc", colnames(injury_risks_100mil_hours))
+# colnames(injury_risks_100mil_hours) = gsub("w_sc", "wsc", colnames(injury_risks_100mil_hours))
 injury_risks_lng_hours <- reshape2::melt(injury_risks_100mil_hours)
@@
 injury_risks_lng_hours$scenario <- as.character(injury_risks_lng_hours$scenario)
-injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bl',]$scenario <- 'Baseline_predicted'
-injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bisc',]$scenario <- 'Bicycling'
-injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bssc',]$scenario <- 'Public Transport'
-injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'csc',]$scenario <- 'Car'
-if (NSCEN == 4){
-  injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-}
-if (NSCEN == 5){
-  injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-    injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'wsc',]$scenario <- 'Walking'
-}
+# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bl',]$scenario <- 'Baseline_predicted'
+# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bisc',]$scenario <- 'Bicycling'
+# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bssc',]$scenario <- 'Public Transport'
+# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'csc',]$scenario <- 'Car'
+# if (NSCEN == 4){
+#   injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+# }
+# if (NSCEN == 5){
+#   injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
+#     injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'wsc',]$scenario <- 'Walking'
+# }
 rd_hours <- rename(injury_risks_lng_hours, mode = stage_mode)
@@
 duration <- duration  %>% rename(mode = stage_mode)
-duration[duration$scenario == 'Scenario 1',]$scenario <- 'Bicycling'
-duration[duration$scenario == 'Scenario 3',]$scenario <- 'Public Transport'
-duration[duration$scenario == 'Scenario 2',]$scenario <- 'Car'
-duration[duration$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
-duration[duration$scenario == 'Scenario 5',]$scenario <- 'Walking'
+# duration[duration$scenario == 'Scenario 1',]$scenario <- 'Bicycling'
+# duration[duration$scenario == 'Scenario 3',]$scenario <- 'Public Transport'
+# duration[duration$scenario == 'Scenario 2',]$scenario <- 'Car'
+# duration[duration$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
+# duration[duration$scenario == 'Scenario 5',]$scenario <- 'Walking'
 

### 2023-06-21 summary_tables.Rmd

@@
 }
-scen_names <- rownames(scen_prop)
+scen_names <- rownames(io$scen_prop)
 #yll_totals <- lapply(io,function(x){
@@
                scenario = case_when(
-                 grepl("scen1", cause) ~ scen_names[1],
-                 grepl("scen2", cause) ~ scen_names[2],
-                 grepl("scen3", cause) ~ scen_names[3],
-                 grepl("scen4", cause) ~ scen_names[4],
-                 grepl("scen5", cause) ~ scen_names[5],
+                 grepl("sc_cycle", cause) ~ scen_names[1],
+                 grepl("sc_car", cause) ~ scen_names[2],
+                 grepl("sc_bus", cause) ~ scen_names[3],
+                 grepl("sc_motorcycle", cause) ~ scen_names[4]
                ),
@@
                  grepl("CVD", cause) ~ "L2: CVD",
-                 grepl("COPD", cause) ~ "L2: Respiratory",
-                 grepl("LRI", cause) ~ "L2: Respiratory",
+                 grepl("respiratory", cause) ~ "L2: Respiratory",
                  grepl("T2D", cause) ~ "L2: Other",
@@
                  grepl("myeloma", cause) ~ "L3: myeloma",
+                 grepl("myeloid_leukemia", cause) ~ "L3: Myeloid Leukemia",
                  grepl("Parkinson", cause) ~ "L3: Parkinson",

### 2023-07-12 results/multi_city/health_impacts/health_pathway_tbls.qmd
Use path specific outcomes instead - ignoring the interactions between AP and PA. Rel #124
Generated results using v30.1 on 120723

@@
     page-layout: custom
+    embed-resources: true
 execute:
@@
 
-deaths_pathway <- read_csv(paste0(rel_path, "deaths.csv"))
+deaths_pathway <- read_csv(paste0(rel_path, "deaths_pathway.csv"))
 
-ylls_pathway <- read_csv(paste0(rel_path, "ylls.csv"))
+ylls_pathway <- read_csv(paste0(rel_path, "ylls_pathway.csv"))
 

### 2023-08-14 summary_tables.Rmd

@@
 
-  # city <- "antofagasta"
-
-  td <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
-    rename("Baseline" = base_mmet, "CYC_SC" = sc_cycle_mmet,
-           "CAR_SC" = sc_car_mmet,  "BUS_SC" = sc_bus_mmet,
-           "MOT_SC" = sc_motorcycle_mmet) %>% 
+  td <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) |> 
+    rename_with(recode, base_mmet = "Baseline", sc_cycle_mmet = "CYC_SC",
+           sc_car_mmet = "CAR_SC",  sc_bus_mmet = "BUS_SC",
+           sc_motorcycle_mmet = "MOT_SC") %>% 
     summarise(across(, ~ quantile(.x, c(0.05,0.10,0.25, 0.375, 0.50,0.75,0.90), na.rm = T)), 
@@
   td1 <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
-     rename("Baseline" = base_mmet, "CYC_SC" = sc_cycle_mmet,
-           "CAR_SC" = sc_car_mmet,  "BUS_SC" = sc_bus_mmet,
-           "MOT_SC" = sc_motorcycle_mmet) %>% 
+     rename_with(recode, base_mmet = "Baseline", sc_cycle_mmet = "CYC_SC",
+           sc_car_mmet = "CAR_SC",  sc_bus_mmet = "BUS_SC",
+           sc_motorcycle_mmet = "MOT_SC") |> 
     summarise(across(, ~ mean(.x))) %>%  t() %>% as.data.frame() %>% 

### 2023-08-14 summary_tables_PA_AP.Rmd

@@
 
-input_parameter_file <- "InputParameters_v28.0.xlsx"
+NUMS <- io$scen_prop |> nrow()
+
+if (NUMS != 4)
+  scen_names <- c("Baseline", "Bicycling", "Driving", "Public Transport")
+
+scen_length <- length(scen_names) - 1
+
+if (!exists("input_parameter_file"))
+  input_parameter_file <<- "InputParameters_v30.1.xlsx"
 # scenario definition
@@
   n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- scen_names
 
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Motorcycling`, factor_key = TRUE)
+                      Baseline:scen_names[length(scen_names)], factor_key = TRUE)
   y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
@@
   n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- scen_names
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Motorcycling`, factor_key = TRUE)
+                      Baseline:scen_names[length(scen_names)], factor_key = TRUE)
   summary <- as.data.frame(data_long %>% group_by(scenario) %>%
@@
   n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- scen_names
 
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Motorcycling`, factor_key = TRUE)
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:scen_names[length(scen_names)], factor_key = TRUE)
   y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) +
@@
   n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- scen_names
 
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:Motorcycling, factor_key = TRUE)
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:scen_names[length(scen_names)], factor_key = TRUE)
   summary <- as.data.frame(data_long %>% group_by(scenario) %>% 
@@
 
-  td <- round(colMeans(io[[city]]$outcomes$mmets[,(n_col - 4):n_col], na.rm = T), 1) %>%
+  td <- round(colMeans(io[[city]]$outcomes$mmets[,(n_col - scen_length):n_col], na.rm = T), 1) %>%
     as.data.frame() %>% tibble::rownames_to_column()
@@
 
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- scen_names
 
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Motorcycling`, factor_key = TRUE)
+                      Baseline:scen_names[length(scen_names)], factor_key = TRUE)
 
@@
   n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Motorcycling`, factor_key = TRUE)
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- scen_names
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:scen_names[length(scen_names)], factor_key = TRUE)
   y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) + geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 

### 2023-08-17 injury_tables.Rmd
Updated injury output calculations

@@
 ---
-title: "Injrury Outcomes - WHW matrices"
+title: "Injury Outcomes"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-# NOTE: the following scenario order is assumed: 'bi_sc', 'car_sc', 'bus_sc','mc_sc','w_sc' and if there are less than 5 scenarios, the last scenarios can be missing but all of the scenarios in the middle need to exist.
-e.g.
-if you want the public transport and the walking scenario, you still need to create a car, motorcycling and bicycling scenario!
+# Script to summarise injury outcomes
+
+# WARNING: Script assumes that the Baseline scenario corresponds to the known injury fatality counts in the city, i.e. 
+#          it assumes that the reference scenario is the Baseline scenario
+
+#### It produces the following output documents:
+
+## - html file containing the following information:
+
+#   - fatality counts by strike and casualty mode for each scenario and city (incl lower and upper confidence interval limits were applicable), one table for each city, scenario and mean, upper or lower confidence interval values
+
+#   - fatality counts giving lower, mean and upper bounds (where they exist) for each casualty mode by strike mode for each scenario and city, one table for each casualty mode, scenario and city
+
+#   - tables showing the fatalities per billion km for each city, scenario and mean, upper and lower confidence interval values (where applicable).
+
+#   - tables showing the number of deaths per 100,000 people for each city, scenario and mean, upper and lower confidence interval values (where applicable).
+
+#   - tables showing the fatalities per 100 million hours for each city, scenario and mean, upper and lower confidence interval values (where applicable).
+
+## - output .csv files containing the following information:
+
+#   - The inj_counts.csv file ('results/multi_city/inj/inj_counts.csv') contains the fatality counts (mean and upper and lower confidence interval boundaries if required) for each city, scenario and casualty and strike mode pair.
+It also includes the observed baseline fatality counts and the active travel fatality counts (the sum of cycle and walk casualty mode fatalities)
+
+# - The injury_risks_per_billion_kms.csv file ('results/multi_city/inj/injury_risks_per_billion_kms.csv') gives the fatalities per billion km travelled for each city, scenario and casualty mode.
+It also includes the upper and lower confidence interval boundary values if required and active travel counts
+
+# - The injury_risks_per_100k_pop.csv file ('results/multi_city/inj/injury_risks_per_100k_pop.csv') gives the fatalities per 100k people for each city, scenario and casualty mode.
+It also includes the upper and lower confidence interval boundary values if required and active travel counts
+
+# - The distances.csv file ('results/multi_city/inj/distances.csv') contains the population distances travelled per day for each city, scenario and mode
+
+# - The injury_risks_per_100million_h.csv file ('results/multi_city/inj/injury_risks_per_100million_h.csv') gives the fatalities per 100 million hours travelled for each city, scenario and casualty mode.
+It also includes the upper and lower confidence interval boundary values if required and active travel counts
+
+# - The inj_data.xlsx file (results/multi_city/inj/inj_data.xlsx) contains all injury output csv files
+
+## The script performs the following steps, assuming that ITHIM-GLOBAL module has 
+# been run and the ithim_objects object has been saved in "results/multi_city/io.rds":
+
+# - read in ithim_objects and extract relevant information such as city and scenario names 
+# - define various functions to summarise and print data and to re-name scenarios, also
+#   define function to collapse the age categories into two categories (15-49 and 50-69)
+# - calculate the health burden (yll and deaths) and the demographic information for those
+#   two age categories (15-49 and 50-69)
+
+# - loop through cities:
+#   - loop through the mean injury values but also the upper and lower confidence boundary values if given:
+#     - loop through scenarios incl baseline:
+#       - create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
+#         of nov or whw exists, only create dataframe with existing data), add row and column sums
+#       - delete casualty and strike modes that do not exist in list of standardized modes
+#       - print dataframe to html file
+#       - create one list containing the predicted injury counts for each city, each scenario and for each scenario
+#         both the mean and the upper and lower confidence interval limits are given if they exist
+#       - create a dataframe with the original fatality counts for each casualty and strike mode pair
+#         (adjusted by weight and injury reporting rate)
+#       - create dataframe containing the mode distances for all scenarios including the predicted baseline
+#       - create dataframes containing the population counts, mode speeds and durations travelled
+
+# - create dataframe that contains the fatality counts for each city, scenario and casualty and strike mode pair.
+#   It also includes the observed fatality counts and the active travel fatality counts (the sum of cycle and 
+#   walk casualty mode fatalities) - ('results/multi_city/inj/inj_counts.csv') 
+#   - print individual table for each city, scenario and mean, upper and lower confidence value fatality counts 
+#     (where applicable) by casualty and strike mode to html file
+
+# - For each city, scenario and casualty mode, print a table to the html file giving the fatality counts by strike mode
+#   for the mean (and upper and lower confidence value boundaries) values
+
+# - For each city and scenario give mean (and upper and lower confidence boundaries where required) fatalities per 
+#   billion km travelled for each casualty and strike mode pair. Create html tables.
+
+# - Create csv file ('injury_risks_per_billion_kms.csv') for fatalities per billion km travelled
+
+# - Calculate the number of fatalities per 100,000 people for each city, scenario and mean, upper and lower confidence
+#   interval values (where applicable). Print tables to .html file and also create a csv file ('injury_risks_per_100k_pop.csv')
+
+# - Create dataframe containing all mode distances for each city and scenario ('distances.csv')
+
+# - Calculate tables showing the fatalities per 100 million hours for each city, scenario and mean, upper
+#   and lower confidence interval values (where applicable). Print tables to .html file
+
+# - Create a csv file with the fatalities per 100 million hours (injury_risks_per_100million_h.csv)
+
+# - Export all injury datasets into a single excel sheet (inj_data.xlsx)
 
@@
 ```{js echo=FALSE}
-window.location.href='#Who_Hit_Whom_matrix_(normalized_to_annual_figures)';
+window.location.href='#Injury_outputs_(normalized_to_annual_figures)';
 ```
@@
 ```{r loadLibraries, echo = F, message = F}
-
+# set up 
 st_options(bootstrap.css     = FALSE,       # Already part of the theme so no need for it
@@
            footnote          = NA,          # Keeping the results minimalistic
-           subtitle.emphasis = FALSE)       # For the vignette theme, this gives
-# much better results. Your mileage may vary
+           subtitle.emphasis = FALSE)       # For the vignette theme, this gives much better results. Your mileage may vary
 
@@
 ```{r load_objects = "asis"}
+# Assumes that multi_city_script.R has been run - read in ithim_objects saved as io.rds
 io <- readRDS("results/multi_city/io.rds")
-# Assumes that multi_city_script.R has been run till 
+
 # Get names of cities from the io object
@@
 
+# define which decimal place to round to
 round_to <- 1
 
-# Set plot_mmets to F
-plot_mmets <- F
-
+# function to round and sum the data and then to print it 
+# the html output document
 sum_and_round_and_print <- function(data,text=''){
@@
 }
+
+# function to round and print data
 round_and_print <- function(data,text=''){
@@
 
+# function to rename scenarios
 get_qualified_scen_name <- function(cs){
@@
   if (cs == 'baseline')
-    qualified_scen_name <- 'Baseline'
+    qualified_scen_name <- 'Baseline predicted'
   else if(cs == "sc_walk")
@@
     qualified_scen_name <- 'Public Transport'
+  else if(cs == "Baseline predicted")
+    qualified_scen_name <- 'Baseline predicted'
 
@@
 ```{r preprocessing}
-
+# start the processing
+
+# function to collapse the given age categories into only two categories
+# '15-49' and '50-69'
 collapse_ages <- function(data, ages = c('15-49', '50-69'), age_lab = 'age_cat'){
@@
     reformatted <- do.call(rbind,lapply(1:length(ages),
-                                        function(x) t(sapply(genders,
-                                                             function(y) 
+                                        function(x) t(sapply(genders, function(y) 
                                           if(ncol(data)>3)     colSums(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
@@
 io <- io[-2]
+
 # re-format some age and injury names
@@
 }
-for (city in cities)
-  for (type in c('hb','pathway_hb'))
-    for (out in c('deaths','ylls'))
-      io[[city]]$outcomes[[type]][[out]] <- collapse_ages(io[[city]]$outcomes[[type]][[out]]) # only keep two age categories
-
+
+# calculate the demographics for the two age categories '15-49' and '50-69'
 for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')  # only keep two age categories
 
-pop_by_age <- lapply(io,function(x)sapply(unique(x$demographic$age),function(y)sum(subset(x$demographic,age==y)$population))) # find number of people within two age categories
-pop_by_gender <- lapply(io,function(x)sapply(unique(x$demographic$sex),function(y)sum(subset(x$demographic,sex==y)$population)))
-injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths)=='scen1_deaths_inj') # find col number of injury data
-ap_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('ap',as.character(x))))
-pa_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths),function(x)grepl('pa',as.character(x))))
-scen_names <- rownames(scen_prop)
-
-```
-
-# Who Hit Whom matrix (normalized to annual figures)
-
-Who hit whom matrix for all scenarios for each case study
-
-## Who hit whom matrix of each case study (by baseline + scenarios) {#whw}
+```
+
+# Injury outcome (normalized to annual figures)
+
+Injury outcome for all scenarios for each scenario including the predicted and observed baseline figures
+
+## Injury outcome for each scenario (incl predicted and observed baseline figures)
 
 ```{r whw, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
+# create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
+# of nov or whw exists, only create dataframe with existing data), add row and column sums
+# create distance, speed and duration data frames
 
@@
 
-for (city in cities) {
+for (city in cities) { # loop through all cities
 
@@
 
-  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){ # if upper and lower bounds are given, this is NA, ub, lb
+  # loop through either mean values only or, if upper and lower confidence intervals are given, loop
+  # through them as well
+  #  if upper and lower bounds are given, 'pref' is set to 'NA', 'ub' and 'lb'
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){ 
     # pref <- "ub"
@@
 
-    for (cs in names(io[[city]]$outcomes$whw)){ # Baseline, Scenario 1, Scenario 2, etc.
+    # loop through scenarios incl baseline
+    for (cs in names(io[[city]]$outcomes$whw)){ 
       # cs <- 'baseline'
 
-      if (is.null(io[[city]]$outcomes$whw[[cs]][[whw_qn]])){
-        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){ # if only data for the given name exists for nov but not for whw, only create data frame from nov part
+      # if data only exists for nov but not for whw, only create data frame from nov part
+      if (is.null(io[[city]]$outcomes$whw[[cs]][[whw_qn]])){ # if no 'whw' part exists
+        if (!is.null(io[[city]]$outcomes$whw[[cs]][[nov_qn]])){ # but if 'nov' part exists
+
           td2 <- t(io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame()
           td2$mode <- 'NOV'
-
           td2 <- td2 %>% dplyr::select(mode, names(.))
-
           td3 <- td2
@@
 
-      cat('####', qualified_scen_name, ' ', ifelse(is.na(pref), "mean", pref), ' WHW', '\n')
+      cat('####', qualified_scen_name, ' ', ifelse(is.na(pref), "mean", pref), ' fatalities by cas and str mode', '\n')
 
@@
 
-      td3 <- td3[,c(1, na.omit(match(smodes$exhaustive_list, colnames(td3))), ncol(td3))] # remove column modes that do not appear in list of standardized modes
-
-      td3 <- td3[c(na.omit(match(smodes$exhaustive_list, td3$mode)), nrow(td3)),] # remove rows with mode names that do not appear in list of standardized modes
-
-      temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( .
-< 1 ~ round(., 2) ,  .
->= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))  # convert numbers to characters
+      # remove column modes that do not appear in list of standardized modes
+      td3 <- td3[,c(1, na.omit(match(smodes$exhaustive_list, colnames(td3))), ncol(td3))]
+
+      # remove rows with mode names that do not appear in list of standardized modes
+      td3 <- td3[c(na.omit(match(smodes$exhaustive_list, td3$mode)), nrow(td3)),]
+
+      # convert numbers to characters
+      temp <- td3 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))
+                                ) %>%  mutate_if(is.numeric, ~as.character(.))
 
@@
 
-      cat(knit_print(ft))
+      cat(knit_print(ft)) # print table to html file
 
@@
     }
-
-  }
+  }
+
   # create df with location information
@@
   # create dataframe with original fatality counts adjusted by weight and injury_reporting_rate 
-
   # Baseline - whw
@@
 
+  # add further city, scenario and location information
   city_baseline_count$city <- city
@@
 
+  # create one dataframe for all cities
   if (city == cities[[1]]){
@@
 
-  baseline_pred <- dummy_distances2 %>% filter(scenario == 'Baseline')
-  baseline_pred$scenario <- 'Baseline_predicted'
+  baseline_pred <- dummy_distances2 %>% filter(scenario == 'baseline' | scenario == 'Baseline')
+  baseline_pred$scenario <- 'Baseline'
+
+  # update scenario names
+  dummy_distances2 <- dummy_distances2 |> rowwise() |> mutate(scenario = get_qualified_scen_name(scenario))
 
@@
 
-  ####### create duration df based on true distances and speed
+  # create speed dataframe
   speed_city <- io[[city]]$vehicle_inventory %>% dplyr::select(-c( "PM_emission_inventory",  "CO2_emission_inventory"))
@@
 
-  # dummy_dur <- io[[city]]$dur
-  # 
-  # dummy_dur2 <- pivot_longer(dummy_dur, cols = -c(stage_mode), names_to = 'scenario', values_to = 'mode_duration')
-  # 
-  # baseline_pred <- dummy_dur2 %>% filter(scenario == 'Baseline')
-  # baseline_pred$scenario <- 'Baseline_predicted'
-  # 
-  # dummy_dur3 <- rbind(dummy_dur2, baseline_pred)
-  # 
-  # dummy_dur3$city <- city
-  # 
-  # if (city == cities[1]){
-  #   duration <- dummy_dur3
-  # } else {
-  #   duration <- rbind(duration,dummy_dur3)
-  # }
-}
-
-# calculate duration
+}
+
+# calculate duration based on distance and speed
 duration <- left_join(true_distances, speed_df, by = c('city','stage_mode' ))
@@
 ```{r whw_for_all_cs, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
+
+# create one dataframe for all cities and scenarios (incl original baseline counts) containing one column with fatality counts (and columns for city, strike mode, cas mode, scenario, distances, location)
+# calculate active_travel fatality counts (where cas mode is either walk or cycle) and add to fatality dataframe
+# write csv file
+
 st <- list()
@@
 td <- NULL
-whw_list <- list()
+inj_counts_list <- list()
 for (cs in names(all_whw)){ # loop through scenarios
@@
 
+  # join list elements of strike and cas mode fatilty predictions for all cities for the given scenario
   td <- all_whw[[cs]] %>% purrr::reduce(full_join, by = c("mode", "measure")) %>% as.data.frame() %>% dplyr::select(mode, sort(names(.)))
 
+  td[is.na(td)] <- 0 # set na values to 0
+
+  # remove duplicate values
+  td <- (td[!duplicated(td), ])
+
+  # re-arrange order of rows
+  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total'))
+
   td[is.na(td)] <- 0
 
-  td <- (td[!duplicated(td), ])
-
-  td <- rbind(td %>% dplyr::filter(!mode %in% c("Total", "NOV")) %>% arrange(mode), td %>% filter(mode == 'NOV'), td %>% filter(mode == 'Total')) # re-arrange order of rows
-
-  td[is.na(td)] <- 0
-
   td <- td %>% mutate_if(is.numeric, round, 2)
@@
 
+  # re-name auto_rickshaw as ar
   colnames(td) = gsub("auto_rickshaw", "ar", colnames(td))
 
-  whw_lng <- reshape2::melt(td) # create df with 1 column containing all the fatality counts
-
-  col_split <- stringr::str_split(whw_lng$variable, "_", simplify = TRUE, n = 2) # create one city and one strike mode column
+  # create df with 1 column containing all the fatality counts
+  inj_counts <- reshape2::melt(td)
+
+  col_split <- stringr::str_split(inj_counts$variable, "_", simplify = TRUE, n = 2) # create one city and one strike mode column
 
   # tidy and rename some of the columns
-  whw_lng <- cbind(whw_lng, col_split)
-  names(whw_lng)[5] <- 'strike_mode'
-  names(whw_lng)[6] <- 'city'
-  whw_lng <- whw_lng %>% dplyr::select(-variable)
-  whw_lng <- whw_lng %>% dplyr::filter(strike_mode != "rowSum")
-  whw_lng <- whw_lng %>% rename(str_mode = mode, cas_mode = strike_mode)
-
-  if (nrow(whw_lng[whw_lng$cas_mode == 'ar',]) > 0){
-    whw_lng[whw_lng$cas_mode == 'ar',]$cas_mode <- "auto_rickshaw"
-  }
-
-  if (nrow(whw_lng[whw_lng$str_mode == 'ar',]) > 0){
-    whw_lng[whw_lng$str_mode == 'ar',]$str_mode <- "auto_rickshaw"
-  }
-
-  whw_lng$str_mode <- factor(whw_lng$str_mode, levels = unique(whw_lng$str_mode)) # assign factors
-
-  qualified_scen_name <- 'Public Transport'
+  inj_counts <- cbind(inj_counts, col_split)
+  names(inj_counts)[5] <- 'strike_mode'
+  names(inj_counts)[6] <- 'city'
+  inj_counts <- inj_counts %>% dplyr::select(-variable)
+  inj_counts <- inj_counts %>% dplyr::filter(strike_mode != "rowSum")
+  inj_counts <- inj_counts %>% rename(str_mode = mode, cas_mode = strike_mode)
+
+  # re-name 'ar' to auto_rickshaw again ('_' in 'auto_rickshaw' would have caused issues when splitting the city and strike mode column)
+  if (nrow(inj_counts[inj_counts$cas_mode == 'ar',]) > 0){
+    inj_counts[inj_counts$cas_mode == 'ar',]$cas_mode <- "auto_rickshaw"
+  }
+
+  if (nrow(inj_counts[inj_counts$str_mode == 'ar',]) > 0){
+    inj_counts[inj_counts$str_mode == 'ar',]$str_mode <- "auto_rickshaw"
+  }
+
+  inj_counts$str_mode <- factor(inj_counts$str_mode, levels = unique(inj_counts$str_mode)) # assign factors
+
+  #qualified_scen_name <- 'Public Transport'
   scen <- 'Scenario'
@@
 
-  whw_lng$scenario <- qualified_scen_name
-
-  if (length(whw_list) == 0)
-    whw_list <- whw_lng
+  inj_counts$scenario <- qualified_scen_name
+
+  if (length(inj_counts_list) == 0)
+    inj_counts_list <- inj_counts
   else
-    whw_list <- rbind(whw_list, whw_lng)
-
-  # f <- ggplot(data = whw_lng) +
-  #   aes(x = str_mode, fill = cas_mode, weight = value) +
-  #   geom_bar(position = "stack") +
-  #   scale_fill_brewer(palette = "Dark2") +
-  #   labs(x = "Strike Mode", y = "# of incidents", title = paste("WHW - ", qualified_scen_name, scen)) +
-  #   coord_flip() +
-  #   facet_wrap(vars(city), scales = "free_y")
-  # 
-  # file_list[[cs]] <- f
-
-  #ggsave(f, file = paste0("results/multi_city/whw_matrices/plots/whw_", cs,".png"), limitsize = F, dpi = 300, scale = 1.5)
-
-  d <- paste0("results/multi_city/whw_matrices/interactive_plots/whw_", cs,".html")
-
-  # htmlwidgets::saveWidget(plotly::ggplotly(f), file.path(normalizePath(dirname(d)), basename(d)))
-  #ggsave(f, file=paste0("results/multi_city/whw_matrices/whw_", cs,".png"), width = 14, height = 10, units = "cm")
+    inj_counts_list <- rbind(inj_counts_list, inj_counts)
 
@@
 # add country and continent information
-whw_list <- left_join(whw_list, location, by = 'city')
+inj_counts_list <- left_join(inj_counts_list, location, by = 'city')
 
 # add original counts and rename Baseline to Baseline predicted
-whw_list$scenario <- ifelse(whw_list$scenario == 'Baseline', 'Baseline_predicted',whw_list$scenario)
-
-whw_list <- rbind(whw_list, baseline_counts)
+#inj_counts_list$scenario <- ifelse(inj_counts_list$scenario == 'Baseline', 'Baseline_predicted',inj_counts_list$scenario)
+
+inj_counts_list <- rbind(inj_counts_list, baseline_counts)
 
 # replace capital with lower letters
-whw_list$str_mode <- tolower(whw_list$str_mode)
-
-# update true distances scenarios
-true_distances <- true_distances |> rowwise() |> mutate(scenario = get_qualified_scen_name(scenario))
+inj_counts_list$str_mode <- tolower(inj_counts_list$str_mode)
 
 # add mode distances
-whw_list <- left_join(whw_list, true_distances, by = c('city', 'cas_mode','scenario'))
+inj_counts_list <- left_join(inj_counts_list, true_distances, by = c('city', 'cas_mode','scenario'))
 
 # combine active travel modes
-active_travel <- whw_list %>% filter(cas_mode == 'cycle' | cas_mode == 'pedestrian')
+active_travel <- inj_counts_list %>% filter(cas_mode == 'cycle' | cas_mode == 'pedestrian')
 active_travel <- active_travel %>% group_by(str_mode, measure, city, scenario, country, continent) %>% summarise(
@@
 
+# add active travel modes to inj_counts_list
+inj_counts_list <- rbind(inj_counts_list, active_travel)
+
+# add totals for each cas_mode
+inj_counts_list <- inj_counts_list %>% filter(str_mode != 'Total' & str_mode != 'total')
+inj_counts_list_total <- inj_counts_list %>% group_by(measure, cas_mode, city, scenario, country, continent, mode_distance) %>%
+                                        summarise(str_mode = 'Total', value = sum(value))
+inj_counts_list <- rbind(inj_counts_list, inj_counts_list_total)
+
 # Remove total from whw long form, and write as csv
-readr::write_csv(whw_list %>% filter(str_mode != 'Total'), paste0('results/multi_city/whw_matrices/whw_lng.csv')) # table containing all the fatality counts
-```
-
-## Who hit whom matrix for all case cities {#whw_all}
+readr::write_csv(inj_counts_list, paste0('results/multi_city/inj/inj_counts.csv')) # table containing all the fatality counts
+```
+
+## Casualty mode fatality counts for all case cities
 
 ```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
-### by casualty mode # This does contain NOV fatalities!
+### Tables for each casualty mode, giving fatalities by strike mode -> print to html file
 
@@
 index <- 1
-scen_names <- unique(whw_list$scenario)
+scen_names <- unique(inj_counts_list$scenario)
+
+# loop through scenarios
 for (i in 1:length(scen_names)){
@@
 
-  temp <- whw_list %>% filter(scenario == sn) %>% dplyr::select(cas_mode)
+  temp <- inj_counts_list %>% filter(scenario == sn) %>% dplyr::select(cas_mode)
 
@@
 
-  for (j in 1:length(cas_mode)){
+  for (j in 1:length(cas_mode)){ # loop through casualty modes
     cm <- cas_mode[j]
-    temp1 <- whw_list %>% filter(scenario == sn & cas_mode == cm) %>% spread(value = value, key = str_mode)
-
-    temp1 <- temp1[,c(1, 2, 3, na.omit(match(smodes$exhaustive_list, colnames(temp1))), ncol(temp1))]
-
+
+    # filter by scenario and cas mode, add separate columns for each strike mode
+    temp1 <- inj_counts_list %>% filter(scenario == sn & cas_mode == cm) %>% spread(value = value, key = str_mode)
+
+    # rename nov column to NOV 
+    if ('nov' %in% names(temp1)){
+      temp1 <- temp1 %>% rename(NOV = nov)
+      }
+
+    # keep first three columns and strike modes in given list
+    temp1 <- temp1[,c(1, 2, 3, na.omit(match(smodes$exhaustive_list, colnames(temp1))))]
+
+    # round fatality counts
     temp1 <- temp1 %>% mutate_if(is.numeric, .funs = funs(case_when( . < 1 ~ round(., 2) ,  . >= 1 ~  as.numeric(round(.))))) %>%  mutate_if(is.numeric, ~as.character(.))
@@
 
-### Change in deaths due to injury per billion km
+### Fatalities per billion km travelled
 
 ```{r inj_100k = "asis", echo = F, message = F}
+
+# create html tables for injuries per billion km 
+# calculate fatalities per capita
+
 require(tibble)
@@
 city_df <- list()
-for (city in cities){
-
-  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){
+for (city in cities){ # loop through cities
+
+  # loop through mean (and upper and lower confidence interval boundary values where applicable)
+  for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){ # NA, 'ub', 'lb'
 
@@
 
-    for (cs in names(io[[city]]$outcomes$whw)){
+    for (cs in names(io[[city]]$outcomes$whw)){ # loop through scenarios
       # cs <- 'baseline'
-      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){
-        # td1: Number of fatalities in nov
+      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){ # if there are whw and nov matrices
+         # td1: Number of fatalities in nov for each casualty mode
         td1 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-        # td2: Number of fatalities in whw
+        # td2: Number of fatalities in whw for each casualty mode
         td2 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-        # td3: Number of fatalities: sum of nov and whw
+        # td3: Number of fatalities: sum of nov and whw for each casualty mode
         td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
-        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-
-      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){
+
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){ # if there are only whw matrices
+        # sum of fatalities for each casualty mode
         td3 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){
-
+
+        }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){ # if there are only nov matrices
+        # sum of fatalities for each casualty mode
         td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
 
@@
 
-      td5 <- td4 # save copy for risk per population
+      td5 <- td4 # save copy for later risk per population calculation
 
@@
 
-      el <- full_join(el, td4, by = c('stage_mode'))
-
-      # City's population
+      el <- full_join(el, td4, by = c('stage_mode')) # output for fatalities per bn km travelled
+
+      # add city's population
       pop <- sum(io[[city]]$demographic$population)
@@
       td6[[paste0(var,"_risk")]] <- td4[2]
-      td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop
+      td6[[paste0(var,"_percapita")]] <- td6[[var]] * 365 / pop # calculate fatalities per capita
       names(td6)[3] <- paste(names(td6)[2], names(td6)[3], sep = "_")
@@
 
+    # remove count columns
+    el <- el %>% dplyr::select(-contains('count'))
+
+    # print fatalities per bn km travelled to .html document
     print(kable(el, caption = paste(city, ifelse(is.na(pref), "mean", pref))))
 
-    td <- el %>% dplyr::select(-contains('count'))
+    td <- el #%>% dplyr::select(-contains('count'))
     names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
@@
 
-### Change in deaths due to injury per Billion km across cities
-
 ```{r inj_100k_all_cities = "asis"}
+
+# create csv file per billion km travelled
+
 require(tibble)
@@
 td <- td %>% dplyr::select(stage_mode, sort(names(.)))
-#readr::write_csv(td, 'results/multi_city/whw_matrices/injury_risks_per_100k_kms.csv')
-# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
-# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
-# colnames(td) = gsub("car_sc", "csc", colnames(td))
-# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
-# colnames(td) = gsub("w_sc", "wsc", colnames(td))
+
 injury_risks_b <- td
-# # Convert it into Billion by multiping the values by 10000
-# injury_risks_b[, 2:ncol(injury_risks_b)] <- injury_risks_b[, 2:ncol(injury_risks_b)] * 10000
-#readr::write_csv(injury_risks_b, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms.csv')
-# colnames(injury_risks_b) = gsub("bi_sc", "bisc", colnames(injury_risks_b))
-# colnames(injury_risks_b) = gsub("bus_sc", "bssc", colnames(injury_risks_b))
-# colnames(injury_risks_b) = gsub("car_sc", "csc", colnames(injury_risks_b))
-# colnames(injury_risks_b) = gsub("mc_sc", "mcsc", colnames(injury_risks_b))
-# colnames(injury_risks_b) = gsub("w_sc", "wsc", colnames(injury_risks_b))
+
+# change shape and update some column names
 injury_risks_lng <- reshape2::melt(injury_risks_b)
@@
 injury_risks_lng <- injury_risks_lng %>% dplyr::select(-variable)
-cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
 injury_risks_lng$scenario <- as.character(injury_risks_lng$scenario)
-# injury_risks_lng[injury_risks_lng$scenario == 'bl',]$scenario <- 'Baseline_predicted'
-# injury_risks_lng[injury_risks_lng$scenario == 'bisc',]$scenario <- 'Bicycling'
-# injury_risks_lng[injury_risks_lng$scenario == 'bssc',]$scenario <- 'Public Transport'
-# injury_risks_lng[injury_risks_lng$scenario == 'csc',]$scenario <- 'Car'
-# if (NSCEN == 4){
-#   injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-# }
-# if (NSCEN == 5){
-#   injury_risks_lng[injury_risks_lng$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-#     injury_risks_lng[injury_risks_lng$scenario == 'wsc',]$scenario <- 'Walking'
-# }
+
 rd <- rename(injury_risks_lng, mode = stage_mode)
@@
 
+# get true_distances into correct format
+true_distances <- true_distances  %>% rename(mode = cas_mode)
+
+# add active travel distances
+true_distances_at <- true_distances %>% filter(mode == 'cycle' | mode == 'pedestrian')
+true_distances_at <- true_distances_at %>% group_by(city,scenario) %>% summarise(
+                                              mode_distance = sum(mode_distance), mode = 'active_travel')
+true_distances <- rbind(true_distances, true_distances_at)
+
+# add distance information (note true_distances does not contain distance info for active_travel)
+rd <- left_join(rd, true_distances, by = c('mode', 'city','scenario'))
+
+## add baseline counts 
 # aggregate baseline counts by cas_mode
 baseline_counts_agg <- baseline_counts %>% group_by(city,cas_mode, country, continent, measure, scenario) %>% summarise(value = sum(value)) %>% rename(mode = cas_mode)
-
-# get true_distances into correct format
-true_distances <- true_distances  %>% rename(mode = cas_mode)
-
-# add distance information
-rd <- left_join(rd, true_distances, by = c('mode', 'city','scenario'))
-
-# add baseline counts
 
@@
 
-readr::write_csv(rd, 'results/multi_city/whw_matrices/injury_risks_per_billion_kms_lng.csv')
-```
-
-### # of deaths due to injury per 100,000 people
+# add total counts (excluding active travel)
+
+# output csv
+readr::write_csv(rd, 'results/multi_city/inj/injury_risks_per_billion_kms.csv')
+```
+
+### Fatalities per 100,000 people
 
@@
 
-for (city in cities){
+for (city in cities){ # loop through cities
 
@@
 
+  # loop through mean and upper and lower confidence interval boundaries if they exist
   for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){
@@
 
-    for (cs in names(io[[city]]$outcomes$whw)){
-      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){
+    for (cs in names(io[[city]]$outcomes$whw)){ # loop through scenarios
+
+      if (length(word(names(io[[city]]$outcomes$whw$baseline), 1, sep = "_") %>% unique()) == 2){ # if both whw and nov outputs exist
+        # extract nov part and get into correct format
         td1 <- (io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        # extract whw part and get into correct format
         td2 <- colSums((io[[city]]$outcomes$whw[[cs]][[whw_qn]])) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
+        # join nov and whw parts
         td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
 
-      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){ # if only whw outputs exist
+        # extract whw part and get into correct format
         td3 <- colSums((io[[city]]$outcomes$whw[[cs]][[whw_qn]])) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){
-
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){ # if only nov outputs exist
+        # extract nov part and get into correct format
         td3 <- (io[[city]]$outcomes$whw[[cs]][[nov_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
@@
 
+      # extract true distances
       td4 <- io[[city]]$true_dist %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode)
 
-      if (length(el) == 0){
+      if (length(el) == 0){ # list which gets intitialised with stage modes only
         el <- td4 %>% dplyr::select(stage_mode)
@@
 
+      # add nov and whw counts to distances
       td4 <- full_join(td4, td3 %>% dplyr::select(mode, count) %>% rename(stage_mode = mode), by = 'stage_mode')
@@
 
-       # add active travel mode
+      ## add active travel mode
       # filter out pedestrian and cycle trips, calculate sum of counts and distances and add to existing data
@@
 
+      # calculate fatalities per 100,000 people
       td4[, 2] <- round(td4[, 2] / sum(io[[city]]$demographic$population) * 100000, 2)
@@
 
-    el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
-
+    # calculate total per scenario (not including active travel)
+    #el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
+
+    el_no_at <- el %>% filter(!stage_mode == 'active_travel')
+    el_total <- el_no_at %>% ungroup() %>% janitor::adorn_totals(c('row'))
+
+    el <- rbind(el, el_total %>% filter(stage_mode == 'Total'))
+
+    # print fatalities per 100,000 people to html
     print(kable(el, caption = city))
+
+    # get into correct format and add to city_df list
     td <- el
@@
 
+    # create one table containing mean and upper and lower limits where applicable
     if (length(city_df[["td"]][[city]]) == 3){
@@
   }
-
-}
-
+}
+
+# get fatalities per 100,000 people into correct format
 td <- overall_el_normalized %>% purrr::reduce(full_join, by = c("stage_mode", "measure")) %>% as.data.frame()
@@
 td <- rename(td, mode = stage_mode)
-# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
-# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
-# colnames(td) = gsub("car_sc", "csc", colnames(td))
-# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
-# colnames(td) = gsub("w_sc", "wsc", colnames(td))
+td <- td %>% filter(!mode == 'Total')
+
+# get fatalities per 100,000 people into correct format where all scenarios are in the same column
 injury_risks_per_100k <- reshape2::melt(td)
@@
 injury_risks_per_100k$scenario <- as.character(injury_risks_per_100k$scenario)
-# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bl',]$scenario <- 'Baseline_predicted'
-# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bisc',]$scenario <- 'Bicycling'
-# injury_risks_per_100k[injury_risks_per_100k$scenario == 'bssc',]$scenario <- 'Public Transport'
-# injury_risks_per_100k[injury_risks_per_100k$scenario == 'csc',]$scenario <- 'Car'
-# if (NSCEN == 4){
-#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-# }
-# if (NSCEN == 5){
-#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-#   injury_risks_per_100k[injury_risks_per_100k$scenario == 'wsc',]$scenario <- 'Walking'
-# }
 
@@
 
-readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/whw_matrices/injury_risks_per_100k_pop.csv')
-```
+readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/inj/injury_risks_per_100k_pop.csv')
+```
+
+```{r }
 
 ### True distances in a single file
-```{r }
+
+# create dataframe containing all mode distances for each city and scenario
 true_distances_city <- list()
@@
 }
-true_distances_df <- bind_rows(true_distances_city) # %>% 
-  # mutate(scenario = case_when(
-  #   scenario == "Scenario 1" ~ "Bicycling", 
-  #   scenario == "Scenario 2" ~ "Car",
-  #   scenario == "Scenario 3" ~ "Public Transport",
-  #   scenario == "Scenario 4" ~ "Motorcycling",
-  #   scenario == "Scenario 5" ~ "Walking",
-  #   scenario == "Baseline" ~ "Baseline"))
-readr::write_csv(true_distances_df, 'results/multi_city/whw_matrices/distances.csv')
-```
-
-### Change in deaths due to injury per 100 million hours
+true_distances_df <- bind_rows(true_distances_city)
+
+# add active travel distances
+true_distances_df_at <- true_distances_df %>% filter(stage_mode == 'cycle' | stage_mode == 'pedestrian')
+true_distances_df_at <- true_distances_df_at %>% group_by(city, scenario) %>% summarise(stage_mode = 'active_travel', distance = sum(distance))
+true_distances_df <- rbind(true_distances_df, true_distances_df_at)
+
+# write to csv
+readr::write_csv(true_distances_df, 'results/multi_city/inj/distances.csv')
+```
+
+### Fatalities per 100 million hours travelled
 
@@
 city_df <- list()
-for (city in cities){
-
+for (city in cities){ # loop through cities
+
+  # loop through mean and upper and lower confidence interval boundary values where required
   for (pref in (word(grep("whw",names(io[[city]]$outcomes$whw$baseline), value = T), 2, sep = "_"))){
@@
     # pref <- "ub"
+
+    # set up names
     pref_name <- ""
@@
         td3 <- full_join(td2, td1, by = 'mode') %>% mutate(count = rowSums(.[2:3], na.rm = T)) %>% dplyr::select(-c('count.x', 'count.y'))
-        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-
-      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){
+
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'whw'){ # if there exist only whw matrices
+
+        # td3: number of fatalities for whw only
         td3 <- colSums(io[[city]]$outcomes$whw[[cs]][[whw_qn]]) %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
-      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){
-
+
+      }else if(length(names(io[[city]]$outcomes$whw$baseline)) == 1 && names(io[[city]]$outcomes$whw$baseline) == 'nov'){ # if there exist only nov matrices
+
+        # td3: number of fatalities for whw only
         td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
-        #td3[td3$mode == 'pedestrian', ]$mode <- 'walking'
 
@@
       # td4: duration travelled per mode
-      td4 <- io[[city]]$dur %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs) %>% as.data.frame()
+      #td4 <- io[[city]]$dur %>% filter(stage_mode %in% td3$mode) %>% dplyr::select(stage_mode, cs) %>% as.data.frame()
+      td4 <- duration %>% filter(stage_mode %in% td3$mode & scenario == get_qualified_scen_name(cs) & 
+                                   city == city) %>% dplyr::select(stage_mode, mode_duration)
 
@@
 
+      # rename mode duration column as scenario column
       names(td4)[2] <- var
-
-      td4[, 2] <- as.numeric(td4[,2])
+      td4[, 2] <- as.numeric(unlist(td4[,2]))
+      td4[, 3] <- as.numeric(unlist(td4[,3]))
 
@@
                                                                   y = sum(active_travel[,3]))
+
       colnames(active_travel) <- colnames(td4)
-
       td4 <- rbind(td4, active_travel)
@@
 
+      # remove count columns
+      td4 <- td4 %>% dplyr::select(-contains('count'))
+
+      # join mean and upper and lower confidence interval limits if given
       el <- full_join(el, td4, by = c('stage_mode'))
@@
 
+    # print fatalities per 100 mil h travelled and total counts to html document
     print(kable(el, caption = paste(city, ifelse(is.na(pref), "mean", pref))))
 
-    td <- el %>% dplyr::select(-contains('count'))
+    td <- el # %>% dplyr::select(-contains('count'))
+    # add city name to column names
     names(td)[2:ncol(td)] <- paste(names(td)[2:ncol(td)], city, sep = "_")
+    # add 'measure' columns
     td$measure <- ifelse(is.na(pref), "mean", pref)
 
+    # add to city_df output list  
     city_df[["td"]][[city]][[unique(td$measure)]] <- td
@@
 
-### Change in deaths due to injury per 100 million hours across cities
-
 ```{r inj_100k_all_cities = "asis"}
+
+# Change in deaths due to injury per 100 million hours across cities - produce csv file
+
 #require(tibble)
+
+# initialise dataframe 
 td <- overall_el %>% purrr::reduce(full_join, by = c("stage_mode", "measure")) %>% as.data.frame()
@@
 
-# colnames(td) = gsub("bi_sc", "bisc", colnames(td))
-# colnames(td) = gsub("bus_sc", "bssc", colnames(td))
-# colnames(td) = gsub("car_sc", "csc", colnames(td))
-# colnames(td) = gsub("mc_sc", "mcsc", colnames(td))
-# colnames(td) = gsub("w_sc", "wsc", colnames(td))
 injury_risks_100mil_hours <- td
 
-# colnames(injury_risks_100mil_hours) = gsub("bi_sc", "bisc", colnames(injury_risks_100mil_hours))
-# colnames(injury_risks_100mil_hours) = gsub("bus_sc", "bssc", colnames(injury_risks_100mil_hours))
-# colnames(injury_risks_100mil_hours) = gsub("car_sc", "csc", colnames(injury_risks_100mil_hours))
-# colnames(injury_risks_100mil_hours) = gsub("mc_sc", "mcsc", colnames(injury_risks_100mil_hours))
-# colnames(injury_risks_100mil_hours) = gsub("w_sc", "wsc", colnames(injury_risks_100mil_hours))
+# get data into correct format
 injury_risks_lng_hours <- reshape2::melt(injury_risks_100mil_hours)
@@
 injury_risks_lng_hours <- injury_risks_lng_hours %>% dplyr::select(-variable)
-cols <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")
 injury_risks_lng_hours$scenario <- as.character(injury_risks_lng_hours$scenario)
-# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bl',]$scenario <- 'Baseline_predicted'
-# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bisc',]$scenario <- 'Bicycling'
-# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'bssc',]$scenario <- 'Public Transport'
-# injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'csc',]$scenario <- 'Car'
-# if (NSCEN == 4){
-#   injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-# }
-# if (NSCEN == 5){
-#   injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'mcsc',]$scenario <- 'Motorcycling'
-#     injury_risks_lng_hours[injury_risks_lng_hours$scenario == 'wsc',]$scenario <- 'Walking'
-# }
+
 rd_hours <- rename(injury_risks_lng_hours, mode = stage_mode)
@@
 
-# get true_distances into correct format
+# get duration into correct format
 duration <- duration  %>% rename(mode = stage_mode)
-# duration[duration$scenario == 'Scenario 1',]$scenario <- 'Bicycling'
-# duration[duration$scenario == 'Scenario 3',]$scenario <- 'Public Transport'
-# duration[duration$scenario == 'Scenario 2',]$scenario <- 'Car'
-# duration[duration$scenario == 'Scenario 4',]$scenario <- 'Motorcycling'
-# duration[duration$scenario == 'Scenario 5',]$scenario <- 'Walking'
+
+# add active travel mode duration
+duration_at <- duration %>% filter(mode == 'pedestrian' | mode == 'cycle')
+duration_at <- duration_at %>% group_by(scenario, city) %>% summarise(mode = 'active_travel',
+                                                                      mode_duration = sum(mode_duration))
+duration <- rbind(duration, duration_at)
 
@@
 
-# add counts but first divide by duration
+# add baseline counts - join with duration information
 baseline_counts_dur <- left_join(baseline_counts_agg, duration, by = c('mode', 'city', 'scenario'))
 
-# add active travel 
+# add active travel to baseline counts
 active_travel <- baseline_counts_dur %>% filter(mode == 'cycle' | mode == 'pedestrian')
@@
 
+# calculate fatalities per 100mil h travelled for baseine counts
 baseline_counts_dur$value <- round((baseline_counts_dur$value /(baseline_counts_dur$mode_duration * 365))*100000000,4)
@@
 
+# add baseline rate to predicted rates
 rd_hours <- rbind(rd_hours, baseline_counts_dur )
 
-readr::write_csv(rd_hours, 'results/multi_city/whw_matrices/injury_risks_per_100million_h_lng.csv')
-```
-
-### Write all injury tables in a single excel sheet
+# write as csv
+readr::write_csv(rd_hours, 'results/multi_city/inj/injury_risks_per_100million_h.csv')
+```
+
 ```{r}
@@
 path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"
- writexl::write_xlsx(list(whw = whw_list %>% filter(str_mode != 'Total'),
-                          injury_risks_per_billion_kms = rd,
-                          injury_risks_per_100k  = injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'),
-                          injury_risks_per_100million_h = rd_hours,
+ writexl::write_xlsx(list(inj_counts = inj_counts_list, #%>% filter(str_mode != 'Total'),
+                          injury_risks_billion_kms = rd,
+                          injury_risks_100k_people  = injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'),
+                          injury_risks_100mil_h = rd_hours,
                           true_distances = true_distances_df),

### 2023-08-18 injury_tables.Rmd
updated injury and PA/AP output calculations

@@
 
-# WARNING: Script assumes that the Baseline scenario corresponds to the known injury fatality counts in the city, i.e. 
-#          it assumes that the reference scenario is the Baseline scenario
+# WARNING: Script assumes that the Baseline scenario corresponds to the known injury fatality counts in the city, i.e it assumes that the reference scenario is the Baseline scenario
 
@@
 ```{r}
-# Export injury datasets in a single excel sheet
+# Export injury datasets in a single excel file
 path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"

### 2023-08-18 summary_tables_PA_AP.Rmd
updated injury and PA/AP output calculations

--- 2023-05-12 summary_tables_PA_AP.Rmd
@@
 
+# Script to summarise the PM2.5 exposure and levels, PM2.5 and CO2 concentrations and a physical activity levels
+
+# This script looks at individual level PM2.5 exposure for all scenarios, it summarise the PM2.5 and CO2 concentration levels for all scenarios and summarises the individual physical activity levels for all scenarios.
+
+#### It produces the following output documents:
+
+### - html file containing the following information:
+
+# - boxplots for each city of daily individual PM2.5 exposure levels for each scenario
+
+# - one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario
+
+# - one table detailing the percentage Baseline PM2.5 emissions attributed to each transport mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities
+
+# - one table detailing the CO2 concentration levels for each scenario for all cities
+
+# - boxplots for each city of individual mMET physical activity levels for each scenario
+
+# - one table for each city giving the mean, 5th, 50th and 95th percentile levels of individual physical activity levels (mMETs) for each scenario
+
+# - one table detailing the average mMETs levels for each scenario for all cities
+
+# 2nd Option:
+
+# - boxplots for each city of daily individual PM2.5 exposure levels for each scenario
+
+# - one table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities
+
+# - one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario.
+Table also contains CO2 concentrations for each scenario
+
+# - boxplots for each city of individual mMET physical activity levels for each scenario
+
+### - output .csv/.xlsx files containing the following information:
+
+# - desc_stats.csv file (results/multi_city/AP/desc_stats.csv) containing the summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the baseline scenario and the total transport related CO2 emission concentrations for each scenario and city
+
+# - desc_stats.xlsx file (results/multi_city/AP/desc_stats.xlsx) containing the following tabs:
+
+#   - PM_2.5_emission_inventory: table detailing the percentage Baseline PM2.5 emissions attributed to each transport mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities
+
+#   - CO2_conc: table detailing the CO2 concentration levels for each scenario for all cities
+
+#   - summary_stats_PM2.5_CO2: summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the baseline scenario and the total transport related CO2 emission concentrations for each scenario and city
+
+#### The script performs the following steps, assuming that ITHIM-GLOBAL module has been run and the ithim_objects object has been saved in "results/multi_city/io.rds":
+
+# - read in input parameter values and define function that renames the existing scenarios
+
+# - create boxplots of individual PM2.5 levels for each scenario, one for each city, and write to the .html file
+
+# - create one table for each city detailing the individual PM2.5 exposure levels incl.
+the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario
+
+# - create a single table detailing the percentage Baseline PM2.5 emissions attributed to each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities
+
+# - create a single summary table showing the CO2 concentrations in each scenario for all cities
+
+# - create boxplots for each city showing the individual mMET levels for each scenario
+
+# - create one table for each city giving the mean, 5th, 50th and 95th percentile levels of individual physical activity levels (mMETs) for each scenario
+
+# - create a single table detailing the average mMETs levels for each scenario for all cities
+
+# - 2nd option:
+
+# - create boxplots for each city of daily individual PM2.5 exposure levels for each scenario
+
+# - create one table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities
+
+# - create one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario.
+Table also contains CO2 concentrations for each scenario
+
+# - create boxplots for each city of individual mMET physical activity levels for each scenario
+
 ```
@@
 
-1) Individual-level PM2.5 concentrations for baseline and scenarios
+1) Individual PM2.5 exposure levels and total PM2.5 concentrations for baseline and scenarios
 2) Overall CO2 emissions for baseline and scenarios
-3) Individual-level physical activity for baseline and scenarios
-
-```{r, echo=FALSE}
-#setwd('C:/Users/rg574/Dropbox/spatiotemporal analysis fatalities inida/Rajasthan tourism road deaths') #to create pretty tables
-```
+3) Individual physical activity levels (mMETs) for baseline and scenarios
 
@@
 
-# Boxplots of Individual-level PM2.5 Concentrations
+```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE}
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, cache=FALSE}
+###### Read in input parameters
+
+# Assumes that multi_city_script.R has been run  
+# read in input file
 io <- readRDS("results/multi_city/io.rds")
 
-# Assumes that multi_city_script.R has been run till 
 # Get names of cities from the io object
@@
 
-scen_names <- c("Baseline", "Bicycling", "Driving", "Public Transport", "Motorcycling")
+# function to convert scenario names
+get_qualified_scen_name <- function(cs){
+  qualified_scen_name <- ""
+  if (cs == 'base')
+    qualified_scen_name <- 'Baseline'
+  else if(cs == "sc_walk")
+    qualified_scen_name <- 'Walking'
+  else if(cs == "sc_cycle")
+    qualified_scen_name <- 'Bicycling'
+  else if(cs == "sc_car")
+    qualified_scen_name <- 'Car'
+  else if(cs == "sc_motorcycle")
+    qualified_scen_name <- 'Motorcycling'
+  else if(cs == "sc_bus")
+    qualified_scen_name <- 'Public Transport'
 
-input_parameter_file <- "InputParameters_v28.0.xlsx"
+  return(qualified_scen_name)
+}
+
+# initialise parameters using first city in list
+city <- cities[1]
+scen_length <- length(io[[city]]$ithim_run$scenario_names) - 1
+
+# input parameter file name
+input_parameter_file <<- io[[city]]$ithim_run$input_parameter_file
+
 # scenario definition
-scenario_name <- "GLOBAL"
-reference_scenario <- 'Baseline'
+scenario_name <- io[[city]]$ithim_run$scenarios_used
+reference_scenario <- io[[city]]$ithim_run$reference_scenario
 
-compute_mode <- 'constant' # constant parameters from the given parameters
+compute_mode <- io[[city]]$ithim_run$compute_mode # constant parameters from the given parameters
 
+# read in local input parameter inputs
 all_inputs <- read_excel(input_parameter_file, sheet = "all_city_parameter_inputs")
@@
 # read in global parameters
-
 all_global_inputs <- read_excel(input_parameter_file, sheet = "all_global_parameter_inputs")
@@
 
+# Boxplots of individual daily PM2.5 exposure levels
+
 ```{r}
-# high<- c(70, 25, 150,60)
-# low<- c(40,15,100,40)
-# l <- htmltools::tagList()
+
+#### for each city create a boxplot showing the individual daily PM2.5 exposure levels for all scenarios
+
+# loop through cities
 for (x in 1:length(cities)) {
-  # print(cities[x])
-  # names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11] <- c("Baseline","Walking",
-  #                                                       "Bicycling", "Driving",
-  #                                                       "Motorcycling", 
-  #                                                       "Public Transport")
 
   n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
 
+  # extract original scenario names and convert to updated scenario names
+  col_names <- names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col]
+  orig_scen_names <- sub("pm_conc_", "", col_names)
+  new_scen_names  <- unlist(lapply(orig_scen_names, FUN=get_qualified_scen_name))
+
+  # rename columns
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- new_scen_names
+
+  # get PM2.5 individual exposure level data into correct format
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Motorcycling`, factor_key = TRUE)
+                      Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
+
+  # create boxplot
   y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
     geom_boxplot(outlier.shape = 8) + ggtitle(cities[x]) +
-    labs(y = "Daily PM2.5 Concentration", x = "Scenarios") #+
-  # +ylim(low[x],high[x])
-  # yy <- ggplot() +
-  #   geom_density(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
-  #   ggtitle(cities[x]) +
-  #   labs(y = "Daily PM2.5 Concentration", x = "Scenarios")
+    labs(y = "Daily PM2.5 exposure levels", x = "Scenarios") #+
+
+  # print to html
   print(y)
-  #print(yy)
-  # l[[i]] <- ggplotly(ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
-  #   geom_boxplot(outlier.shape = 8) + ggtitle(cities[x]) +
-  #   labs(y = "Daily PM2.5 Concentration", x = "Scenarios"))
-  # print(ggplotly(yy))
+
 }
@@
 
-# Descriptive tables of Individual-level PM2.5 Concentrations
+# Summary statistics of individual PM2.5 exposure levels
 
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
+
+# for each city create a summary table of individual PM2.5 exposure levels giving the mean and the 5th, 50th and 95th percentile levels.
+Tables also show the PM2.5 concentration levels in the city for each scenario and the changes compared to the baseline scenario
+
 data_long <- NA
@@
 for (x in 1:length(cities)) {
-  #names(io[[cities[x]]]$outcomes$pm_conc_pp)[6:11]<-c("Baseline","Walking", "Bicycling", "Driving", "Motorcycling", "Public Transport")
-  #
+
   n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
+  # rename columns with updated scenario names
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- new_scen_names
+
+  # get data into correct format
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Motorcycling`, factor_key = TRUE)
+                      Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
+
+  # create summary stats
   summary <- as.data.frame(data_long %>% group_by(scenario) %>%
@@
   summary <- cbind(summary$scenario ,round(summary[,2:5], digits = 1))
+
+  # add scenario specific PM2.5 concentration levels and the change in PM2.5 compared to the baseline
   summary$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, digits=2)
@@
   names(summary)[1] <- "Scenario"
+
+  # print to html file
   print(kable(summary, caption = cities[x]))
@@
 
-# Descriptive tables of PM 2.5 emission inventory
+# Summary table of PM 2.5 emission inventories
 
@@
 
-#trans_share<-c(22.5, 40.0, 22.5, 28.1) 
-#pm_conc<- c(50, 18, 122, 47 )
-city <- cities
-#shares<- as.data.frame(cbind(city,as.numeric(io), pm_conc))
-#names(shares)[2]<- "trans_share"
+# Create a single table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities
+
+#city <- cities
 
 sl <- list()
-for (x in 1:length(cities))
-{
+for (x in 1:length(cities)){ # loop through cities
   print(cities[x])
+
+  # find the modes given in the PM emission inventory for that city
   modes <- names(unlist(io[[cities[x]]]$PM_emission_inventory))
+
+  # extract the PM emissions for each mode
   emissions <- as.data.frame(unlist(io[[cities[x]]]$PM_emission_inventory))
+
+  # combine modes and emissions into one dataframe
   city_emissions <- cbind(as.data.frame(modes),  as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))
   names(city_emissions)[2] <- "emissions"
+
+  # define modes of interest
   select <- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
   city_emissions$modes <- as.character(city_emissions$modes)
+
+  # call all other modes not in list of modes of interest 'other'
   city_emissions$modes[!(city_emissions$modes %in% select)] <- "other"
+
+  # calculate the sum of PM emissions by mode (primarily summing across all newly classified 'other' modes)
   summary <- city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
   names(summary)[2] <- cities[x]
+
+  # calculate the percentages of PM emissions attributed to each mode
   summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits = 1)
   summary <- as.data.frame(summary)
-  summary[nrow(summary) + 1,1] <- "Transport share"
+
+  # rename mode columns including the word percentage
+  summary$modes <- paste0(summary$modes, ' (percentage)')
+
+  # add transport share 
+  summary[nrow(summary) + 1,1] <- "Transport share (percentage)"
   summary[nrow(summary),2] <- as.character(round(pm_trans_share[[cities[x]]] * 100))
+
+  # add city wide PM2.5 concentration levels
   summary[nrow(summary) + 1,1] <- "PM2.5 Conc"
   summary[nrow(summary),2] <- as.character(pm_conc_base[[cities[x]]])
+
+  # add to io list
   io[[cities[x]]]$summary_emission <- summary
 
+  # create one dataframe containing all cities
   if (length(sl) == 0){
@@
 
+# print table to html
 print(kable(sl))
@@
 
-# Descriptive table for CO2 Emissions
+# Summary table of CO2 concentrations related to transport in each scenario
+
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
 
+# create summary table of CO2 concentrations in each scenario
 co2_conc <- list()
 td <- NA
-for (city in cities) {
 
+for (city in cities) { # loop through cities
+
+  # extract Co2 concentrations
   td <- round(colSums(io[[city]]$outcomes$co2_conc, na.rm = T), 1) %>% 
@@
 
+  # create one dataframe containing all cities
   if (length(co2_conc) == 0)
@@
 
-co2_conc$Scenario <- scen_names
+co2_conc$Scenario <- new_scen_names
 
+# print to html
 print(kable(co2_conc))
@@
 
-# Boxplots of Individual-level Physical activity (MMETs)
+# Boxplots of individual physical activity (MMETs) levels
 
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
+# create boxplots for each city showing the individual mMET levels for each scenario
+
 limit = 100
-for (x in 1:length(cities)) {
-  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
 
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Motorcycling`, factor_key = TRUE)
+for (x in 1:length(cities)) { # loop through cities
+
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # define columns of interests
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names
+
+  # extract data of interest
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
+
+  # create the boxplots
   y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) +
     geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 
-    labs(y = "Weekly MMET", x = "Scenarios")+  ylim(0, limit)
+    labs(y = "Weekly mMET", x = "Scenarios")+  ylim(0, limit)
   print(y)
@@
 ```
-# Descriptive tables of Individual-level Physical activity (MMETs)
+
+# Summary statistics of individual physical activity levels (MMETs)
 
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
-for (x in 1:length(cities)) {
-  #names(io[[cities[x]]]$outcomes$mmets)[5:10] <- c("baseline","walk_scen", "bike_scen", "car_scen", "MC_scen", "bus_scen")
-  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
 
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:Motorcycling, factor_key = TRUE)
+# create one table for each city giving the mean, 5th, 50th and 95th percentile levels of individual physical activity levels (mMETs) for each scenario
+
+for (x in 1:length(cities)) { # loop through the cities
+
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find the number of columns
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names # assign new column names
+
+  # create dataframe with individual mMET values
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
+
+  # create summary stats
   summary <- as.data.frame(data_long %>% group_by(scenario) %>% 
@@
   names(summary)[1] <- "Scenario"
+
+  # print to html
   print(kable(summary, caption = cities[x]))
@@
 
-# Descriptive table for Physical Activity (MMETs)
+# Average physical activity levels (MMETs) in each city
+
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
+# create a single table detailing the average mMETs levels for each scenario for all cities
 
 cl <- list()
-for (city in cities)
-{
-  #td <- round(colMeans(io[[city]]$outcomes$mmets[,5:10], na.rm = T), 1) %>%
-  #  as.data.frame() %>% tibble::rownames_to_column()
-  #  
-  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
+for (city in cities){ # loop through cities
 
-  td <- round(colMeans(io[[city]]$outcomes$mmets[,(n_col - 4):n_col], na.rm = T), 1) %>%
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find number of columns of interest
+
+  # extract the mean mMET values and create dataframe
+  td <- round(colMeans(io[[city]]$outcomes$mmets[,(n_col - scen_length):n_col], na.rm = T), 1) %>%
     as.data.frame() %>% tibble::rownames_to_column()
@@
 
+  # create one dataframe containing information about all cities
   if (length(cl) == 0)
@@
 
-cl$Scenario <- scen_names
+cl$Scenario <- new_scen_names
 
+# print to html 
 print(kable(cl))
@@
 
-# PM 2.5 summary and CO2
+# PM2.5 and CO2 results
+
 ```{r message=FALSE, warning=FALSE, echo=FALSE}
 
+# create PM2.5 and CO2 boxplots and tables:
+
+# - boxplots for each city of daily individual PM2.5 exposure levels for each scenario
+
+# - one table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities
+
+# - one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario.
+Table also contains CO2 concentrations for each scenario
+
 l <- list()
-# I have to keep the first chunk of code to run this part
-# high<- c(70, 25, 150,60)
-# low<- c(40,15,100,40)
-# l <- htmltools::tagList()
-for (x in 1:length(cities)) {
+# Have to keep the first chunk of code to run this part
+
+for (x in 1:length(cities)) { # loop through cities
   # print(cities[x])
@@
   cat( paste0('## **', cities[x], '** \n' ))
-  # Emission inventory
+
+  ## Emission inventory
+
+  # find the mode names in the PM2.5 emissions inventory
   modes <- names(unlist(io[[cities[x]]]$PM_emission_inventory))
+
+  # extract the mode emissions
   emissions <- as.data.frame(unlist(io[[cities[x]]]$PM_emission_inventory))
-  city_emissions <- cbind(as.data.frame(modes),                           as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))
+
+  # create one table with modes and corresponding PM2.5 emissions
+  city_emissions <- cbind(as.data.frame(modes),
+                          as.data.frame(emissions$`unlist(io[[cities[x]]]$PM_emission_inventory)`))
+
   names(city_emissions)[2] <- "emissions"
+
+  # define list with modes of interest
   select <- c("car", "motorcycle", "bus_driver", "truck", "big_truck")
   city_emissions$modes <- as.character(city_emissions$modes)
+
+  # define all other modes not in list of interest as 'other'
   city_emissions$modes[!(city_emissions$modes %in% select)] <- "other"
+
+  # sum the emissions (mainly to find the total emissions for all newly defined 'other' modes)
   summary <- city_emissions %>% group_by(modes) %>% summarise(sum(emissions))
   names(summary)[2] <- cities[x]
+
+  # calculate the percentage contribution of each mode
   summary[[cities[x]]] <- round(summary[[cities[x]]]*100/sum(summary[[cities[x]]]), digits = 1)
   summary <- as.data.frame(summary)
-  summary[nrow(summary)+1,1] <- "Transport share"
+
+  # rename mode columns including the word percentage
+  summary$modes <- paste0(summary$modes, ' (percentage)')
+
+  # add the percentage attributed to transport
+  summary[nrow(summary)+1,1] <- "Transport share (percentage)"
   summary[nrow(summary),2] <- as.character(round(pm_trans_share[[cities[x]]] * 100))
+
+  # add the total PM2.5 concentration levels
   summary[nrow(summary)+1,1] <- "PM2.5 Conc"
   summary[nrow(summary),2] <- as.character(pm_conc_base[[cities[x]]])
+
+  # add to io objects
   io[[cities[x]]]$summary_emission <- summary
 
-  # PM 2.5 table and plot
-  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
+  ## PM 2.5 table and plot
 
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - 4):n_col] <- scen_names
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find number of columns
 
+  # assign new scenario names to columns
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- new_scen_names
+
+  # extract data with PM2.5 exposure levels for each scenario
   data_long <- gather(io[[cities[x]]]$outcomes$pm_conc_pp, scenario, pm_conc,
-                      Baseline:`Motorcycling`, factor_key = TRUE)
+                      Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
 
+  # create summary statistics
   summary2 <- as.data.frame(data_long %>% group_by(scenario) %>%
@@
   summary2 <- cbind(summary2$scenario, round(summary2[,2:5], digits = 1))
+
+  # add total PM concentration levels for each scenario
   summary2$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, 2)
+
+  # calculate difference to baseline scenario
   summary2$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits = 2)
@@
 
-  # Plot
+  # Plot individual level PM2.5 exposure levels by scenario
   y <- ggplot(data_long, aes(x = scenario, y = pm_conc, fill = scenario)) +
@@
   # CO2
+  # extract CO2 transport emissions for each scenario
   td <- round(colSums(io[[cities[x]]]$outcomes$co2_conc, na.rm = T), 1) %>%
@@
   names(td) <- c('Scenario', 'CO2')
-  td$Scenario <- scen_names
+  td$Scenario <- new_scen_names
 
+  # print various boxplots and tables to html
   cat("\n \n")
-  cat( paste('#### Boxplot of Individual-level PM2.5 concentrations \n' ))
+  cat( paste('#### Boxplots of individual PM2.5 exposure levels \n' ))
   print(y)
@@
   cat("\n \n")
-  cat( paste('#### Emission inventory \n' )) 
+  cat( paste('#### PM2.5 emissions inventory \n' )) 
   print(kable(summary))
@@
   cat("\n \n")
-  cat( paste('#### Descriptive table of Individual-level PM2.5 concentrations and CO2 emissions\n' ))
+  cat( paste('#### Summary statistics of individual level PM2.5 exposure levels and total transport related CO2 emissions\n' ))
   l[[cities[x]]] <- left_join(summary2, td, by = "Scenario")
@@
 
+# create output Excel file containing the basic PM2.5 and CO2 exposure / concentration information
 output_file <- "results/multi_city/AP/desc_stats.xlsx"
@@
 
+# create csv file (summary_stats_Pm2.5_CO2 tab in Excel file above)
 write_csv(bind_rows(l), "results/multi_city/AP/desc_stats.csv")
@@
 
-# Boxplots of Individual-level Physical activity (MMETs)
+# Boxplots of individual physical activity levels (MMETs)
 
 ```{r, message=FALSE, warning=FALSE, echo=FALSE}
+# create boxplots for each city of individual mMET physical activity levels for each scenario
+
 limit = 100
-for (x in 1:length(cities)) {
-  n_col <- ncol(io[[cities[x]]]$outcomes$mmets)
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - 4):n_col] <- scen_names
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:`Motorcycling`, factor_key = TRUE)
+for (x in 1:length(cities)) { # loop through cities
+
+  n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find number of columns
+
+  # assign new scenario names
+  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names
+
+  # extract mMETs data
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
+
+  # create boxplots
   y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) + geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 
     labs(y = "Weekly MMET", x = "Scenarios") + ylim(0, limit)
+
+  # print to html
   print(y)

### 2023-08-21 injury_tables.Rmd
updated comments

@@
 
-knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')
-
-smodes <- read_csv('data/global/modes/standardized_modes.csv')
-
 # Script to summarise injury outcomes
 
-# WARNING: Script assumes that the Baseline scenario corresponds to the known injury fatality counts in the city, i.e it assumes that the reference scenario is the Baseline scenario
+# WARNING: Script assumes that the Baseline scenario corresponds to the known injury fatality counts in the city
 
@@
 
-## - html file containing the following information:
-
-#   - fatality counts by strike and casualty mode for each scenario and city (incl lower and upper confidence interval limits were applicable), one table for each city, scenario and mean, upper or lower confidence interval values
-
-#   - fatality counts giving lower, mean and upper bounds (where they exist) for each casualty mode by strike mode for each scenario and city, one table for each casualty mode, scenario and city
-
-#   - tables showing the fatalities per billion km for each city, scenario and mean, upper and lower confidence interval values (where applicable).
+### - html file containing the following information:
+
+#   - fatality counts by strike and casualty mode for each scenario and city (incl lower and upper confidence interval limits were applicable); one table for each city, scenario and mean, upper or lower confidence interval values
+
+#   - fatality counts giving lower, mean and upper bounds (where they exist) for each casualty mode by strike mode for each scenario and city; one table for each casualty mode, scenario and city
+
+#   - tables showing the fatalities per billion km travelled for each city, scenario and mean, upper and lower confidence interval values (where applicable).
 
@@
 
-#   - tables showing the fatalities per 100 million hours for each city, scenario and mean, upper and lower confidence interval values (where applicable).
-
-## - output .csv files containing the following information:
+#   - tables showing the fatalities per 100 million hours travelled for each city, scenario and mean, upper and lower confidence interval values (where applicable).
+
+### - output .csv files containing the following information:
 
@@
 
-## The script performs the following steps, assuming that ITHIM-GLOBAL module has 
-# been run and the ithim_objects object has been saved in "results/multi_city/io.rds":
+## The script performs the following steps, assuming that ITHIM-GLOBAL has been run and the ithim_objects object has been saved in "results/multi_city/io.rds":
 
 # - read in ithim_objects and extract relevant information such as city and scenario names 
-# - define various functions to summarise and print data and to re-name scenarios, also
-#   define function to collapse the age categories into two categories (15-49 and 50-69)
-# - calculate the health burden (yll and deaths) and the demographic information for those
-#   two age categories (15-49 and 50-69)
+# - define various functions to summarise and print data and to re-name scenarios
 
@@
 #   - loop through the mean injury values but also the upper and lower confidence boundary values if given:
-#     - loop through scenarios incl baseline:
+#     - loop through scenarios incl baseline (original counts) and baseline predicted:
 #       - create dataframe with nov and whw injury counts by casualty and strike mode pair (if only one
@@
 #   It also includes the observed fatality counts and the active travel fatality counts (the sum of cycle and 
-#   walk casualty mode fatalities) - ('results/multi_city/inj/inj_counts.csv') 
-#   - print individual table for each city, scenario and mean, upper and lower confidence value fatality counts 
-#     (where applicable) by casualty and strike mode to html file
+#   walk casualty mode fatalities) ('results/multi_city/inj/inj_counts.csv')
 
@@
 # - For each city and scenario give mean (and upper and lower confidence boundaries where required) fatalities per 
-#   billion km travelled for each casualty and strike mode pair. Create html tables.
+#   billion km travelled for each casualty mode. Create html tables.
 
@@
 
-# - Create dataframe containing all mode distances for each city and scenario ('distances.csv')
+# - Create dataframe containing all daily population mode distances for each city and scenario ('distances.csv')
 
@@
 
-# - Export all injury datasets into a single excel sheet (inj_data.xlsx)
+# - Export all injury datasets into a single excel file (inj_data.xlsx)
+
+knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message = F, warning = F, results='asis')
+
+# read in table detailing the standardized modes
+smodes <- read_csv('data/global/modes/standardized_modes.csv')
 
@@
 # Remove extra bits of info - irrelevant outputs for injury related outputs
-# This makes it the io object lighter
+# This makes the io object lighter
 for (city in cities){
@@
   io[[city]][["synth_pop"]] <- NULL
-
   io[[city]]$outcomes$mmets <- io[[city]]$outcomes$pm_conc_pp <- NULL
@@
 
-# function to round and sum the data and then to print it 
+# function to round and sum the data and then to print it to
 # the html output document
@@
 
-```{r preprocessing}
-# start the processing
-
-# function to collapse the given age categories into only two categories
-# '15-49' and '50-69'
-collapse_ages <- function(data, ages = c('15-49', '50-69'), age_lab = 'age_cat'){
-  target_min_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][1]))
-  target_max_age <- as.numeric(sapply(ages,function(x)strsplit(x,'-')[[1]][2]))
-  min_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][1]))
-  max_ages <- as.numeric(sapply(data[[age_lab]],function(x)strsplit(x,'-')[[1]][2]))
-  genders <- unique(data$sex)
-  if (ncol(data) > 3) {
-    reformatted <- do.call(rbind,lapply(1:length(ages),
-                                        function(x) t(sapply(genders, function(y) 
-                                          if(ncol(data)>3)     colSums(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
-                                                             else sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
-                                        ))
-    ))
-    data.frame(age_cat=rep(ages,each=2),sex=rep(genders,2),reformatted,stringsAsFactors=F)
-  }else{
-    reformatted <- do.call(c,lapply(1:length(ages),
-                                    function(x) sapply(genders,
-                                                       function(y) 
-                                                         sum(data[min_ages<=target_max_age[x]&max_ages>=target_min_age[x]&data$sex==y,-c(1,2),drop=F])
-                                    )
-    ))
-    data.frame(age=rep(ages,each=2),sex=rep(genders,2),population=as.numeric(reformatted),stringsAsFactors=F)
-  }
-}
-
-scen_prop <- io$scen_prop
-io <- io[-2]
-
-# re-format some age and injury names
-for (i in 1:length(io)) {
-  io[[i]]$demographic$age[io[[i]]$demographic$age == '5-9'] <- '05-9'
-  io[[i]]$outcomes$hb$deaths$age_cat[io[[i]]$outcomes$hb$deaths$age_cat=='5-9'] <- '05-9'
-  io[[i]]$outcomes$hb$ylls$age_cat[io[[i]]$outcomes$hb$ylls$age_cat=='5-9'] <- '05-9'
-  io[[i]]$outcomes$pathway_hb$deaths$age_cat[io[[i]]$outcomes$pathway_hb$deaths$age_cat=='5-9'] <- '05-9'
-  io[[i]]$outcomes$pathway_hb$ylls$age_cat[io[[i]]$outcomes$pathway_hb$ylls$age_cat=='5-9'] <- '05-9'
-  io[[i]]$outcomes$hb$deaths <- io[[i]]$outcomes$hb$deaths[,!sapply(names(io[[i]]$outcomes$hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
-  io[[i]]$outcomes$hb$ylls <- io[[i]]$outcomes$hb$ylls[,!sapply(names(io[[i]]$outcomes$hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
-  io[[i]]$outcomes$pathway_hb$deaths <-      io[[i]]$outcomes$pathway_hb$deaths[,!sapply(names(io[[i]]$outcomes$pathway_hb$deaths),function(x)grepl('ac|neo',as.character(x)))]
-  io[[i]]$outcomes$pathway_hb$ylls <- io[[i]]$outcomes$pathway_hb$ylls[,!sapply(names(io[[i]]$outcomes$pathway_hb$ylls),function(x)grepl('ac|neo',as.character(x)))]
-}
-
-# calculate the demographics for the two age categories '15-49' and '50-69'
-for(city in cities) io[[city]]$demographic <- collapse_ages(io[[city]]$demographic,age_lab='age')  # only keep two age categories
-
-```
-
 # Injury outcome (normalized to annual figures)
@@
 # of nov or whw exists, only create dataframe with existing data), add row and column sums
-# create distance, speed and duration data frames
+# create distance, speed and duration dataframes
 
@@
 
-    # loop through scenarios incl baseline
+    # loop through scenarios incl baseline predicted
     for (cs in names(io[[city]]$outcomes$whw)){ 
@@
       if (length(city_df[[cs]][[city]]) == 3) # if for each fatality sum we have mean, lb and ub
-        all_whw[[cs]][[city]] <- data.table::rbindlist(city_df[[cs]][[city]]) # combine mean, lb and ub counts into one table for each scenario
+        # combine mean, lb and ub counts into one table for each scenario
+        all_whw[[cs]][[city]] <- data.table::rbindlist(city_df[[cs]][[city]]) 
     }
@@
     injury_data_count_nov$value <- injury_data_count_nov$value / injury_data_count_nov$weight / injury_data_count_nov$injury_reporting_rate
-    injury_data_count_nov <- injury_data_count_nov[c('str_mode', 'value', 'cas_mode')] # re order columns and drop column weight
+    # re order columns and drop column weight
+    injury_data_count_nov <- injury_data_count_nov[c('str_mode', 'value', 'cas_mode')]
 
@@
 
+    # combine whw and nove data
     city_baseline_count <- rbind(injury_data_count2, injury_data_count_nov2)
@@
 
-  col_split <- stringr::str_split(inj_counts$variable, "_", simplify = TRUE, n = 2) # create one city and one strike mode column
+  # create one city and one strike mode column
+  col_split <- stringr::str_split(inj_counts$variable, "_", simplify = TRUE, n = 2)
 
@@
 
-  #qualified_scen_name <- 'Public Transport'
   scen <- 'Scenario'
 
+  # rename scenarios
   qualified_scen_name <- get_qualified_scen_name(cs)
@@
 inj_counts_list <- left_join(inj_counts_list, location, by = 'city')
-
-# add original counts and rename Baseline to Baseline predicted
-#inj_counts_list$scenario <- ifelse(inj_counts_list$scenario == 'Baseline', 'Baseline_predicted',inj_counts_list$scenario)
 
@@
 
-# Remove total from whw long form, and write as csv
+# write to csv
 readr::write_csv(inj_counts_list, paste0('results/multi_city/inj/inj_counts.csv')) # table containing all the fatality counts
@@
 ```{r whw_print, results = 'asis', fig.width=7, fig.height=4, echo=FALSE}
+
 ### Tables for each casualty mode, giving fatalities by strike mode -> print to html file
@@
     cat('#### Casualty mode: ', stringr::str_to_title(cm), '\n')
-    #print(kable(format(temp1, scientific = F), caption = paste('Who Hit Whom (WHW) for ', scen_names[i], 'scenario and ', cas_mode[j], ' mode')))
-
+
+    # updated formating 
     ft <- flextable(temp1)
@@
 
+      # get updated scenario names
       var <- get_qualified_scen_name(cs)
@@
       pop <- sum(io[[city]]$demographic$population)
+
        # Compute distance travelled per capita
@@
 
-# add distance information (note true_distances does not contain distance info for active_travel)
+# add distance information 
 rd <- left_join(rd, true_distances, by = c('mode', 'city','scenario'))
@@
 rd <- rbind(rd, baseline_counts_km )
-
-# add total counts (excluding active travel)
 
@@
     }
-
-    # calculate total per scenario (not including active travel)
-    #el <- el %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
 
@@
 
-# add country and continent information
-# injury_risks_per_100k <- left_join(injury_risks_per_100k, location, by = 'city')
-
+# write to csv
 readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/inj/injury_risks_per_100k_pop.csv')
@@
 
-        # td3: number of fatalities for whw only
+        # td3: number of fatalities for nov only
         td3 <- io[[city]]$outcomes$whw[[cs]][[nov_qn]] %>% as.data.frame() %>% rownames_to_column() %>% rename(mode = rowname) %>% rename_at(2, ~"count")
@@
     td <- el # %>% dplyr::select(-contains('count'))
+
     # add city name to column names
@@
 ```{r}
-# Export injury datasets in a single excel file
+# Export injury datasets as a single excel file
 path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"

### 2023-08-22 R/run_ithim.R
updated documentation

@@ -5,7 +5,7 @@
 #' This function works by creating a switch to run the computation by calling 
 #' ithim_calculation_sequence() directly, or to divert to the sampling case
 #' ithim_uncertainty.R which first extracts the sampled parameters and then calls
-#' the ithim_calculation_sequence()
+#' the ithim_calculation_sequence().
 #' 
 #' 
 #' @param ithim_object list of input data needed to calculate the health burden
@@ -37,45 +37,61 @@ run_ithim <- function(ithim_object,seed=1){
 #' 
 #' This function performs the following steps:
 #' 
-#' - extract all lists and variables from the ithim_object list
+#' \enumerate{
 #' 
-#' - (1) air pollution pathway:
-#'    - calculate the PM2.5 exposure for each person in the synthetic population 
+#' \item extract all lists and variables from the ithim_object list
+#' 
+#' \item  air pollution pathway:
+#'    \itemize{
+#'    \item calculate the PM2.5 exposure for each person in the synthetic population 
 #'      and PM2.5 emissions for each mode and scenario (scenario_pm_calculations.R)
-#'    - calculate the CO2 emissions for each mode and scenario (scenario_co2_calculations.R)
-#'    - assign relative risk to each person in the synthetic population for each disease
+#'    \item calculate the CO2 emissions for each mode and scenario (scenario_co2_calculations.R)
+#'    \item assign relative risk to each person in the synthetic population for each disease
 #'      related to PM pollution and each scenario based on the individual PM exposure 
 #'      levels (gen_ap_rr.R)
+#'      }
 #'      
-#' - (2) physical activity pathway:
-#'    - calculate total mMETs for each person in the synthetic population (total_mmet.R)
-#'    - assign relative risk to each person in the synthetic population for each disease
+#' \item  physical activity pathway:
+#'    \itemize{
+#'    \item calculate total mMETs for each person in the synthetic population (total_mmet.R)
+#'    \item assign relative risk to each person in the synthetic population for each disease
 #'      related to physical activity levels and each scenario based on the individual mMET
 #'      values (gen_pa_rr.R)
+#'    }
 #'  
-#' - (3) physical activity and air pollution combined:
-#'    - combine the PA and AP datasets by joining the two datasets. For disease affected by 
+#' \item  physical activity and air pollution combined:
+#'    \itemize{
+#'    \item combine the PA and AP datasets by joining the two datasets. For disease affected by 
[…]

### 2023-08-22 R/run_ithim_setup.R

@@ -70,7 +70,7 @@
 #'
 #' \item if none of the corresponding input parameters were sampled from a distribution,
 #'   call get_all_distances.R to calculate trip distances
-#' 
+#' }
 #' 
 #' @param seed set seed to get the same results when sampling from a distribution
 #' @param CITY name of the city, and name of the directory containing city data files

### 2023-08-22 multi_city_script.R

@@ -321,6 +321,15 @@ print(system.time(for(city in cities){
   ithim_objects[[city]]$new_walk_trips_count$bus <- count_new_walk_trips_bus
   ithim_objects[[city]]$new_walk_trips_count$rail <- count_new_walk_trips_rail
   
+  # add run relevant information to ithim_objects list
+  ithim_objects[[city]]$ithim_run <- list()
+  ithim_objects[[city]]$ithim_run$input_parameter_file <- input_parameter_file
+  ithim_objects[[city]]$ithim_run$scenarios_used <- scenario_name
+  ithim_objects[[city]]$ithim_run$reference_scenario <- reference_scenario
+  ithim_objects[[city]]$ithim_run$scenario_increase <- scenario_increase
+  ithim_objects[[city]]$ithim_run$scenario_names <- SCEN
+  ithim_objects[[city]]$ithim_run$compute_mode <- compute_mode
+  
   # store results to plot
   min_ages <- sapply(ithim_objects[[city]]$outcome$hb$ylls$age_cat,function(x)as.numeric(strsplit(x,'-')[[1]][1]))
   max_ages <- sapply(ithim_objects[[city]]$outcome$hb$ylls$age_cat,function(x)as.numeric(strsplit(x,'-')[[1]][2]))

### 2023-08-23 summary_tables.Rmd
updated use of reference scenario and added comments

--- 2023-05-12 summary_tables.Rmd
@@
 ```{r setup, include=FALSE}
+
+# Script summarises mode splits, distances and durations travelled and looks at health outcomes for the different air pollution, injury and physical activity pathways but also by disease
+
+# WARNING: This script only works if ITHIM-Global has been run in constant mode!
+
+#### It produces the following output documents:
+
+### - html file containing the following information:
+
+# ( - table displaying highest propensity (%) for each distance category, for a given mode. - CURRENTLY commented out as not relevant for BOGOTA, GLOBAL, LATAM or AFRICA_INDIA scenario definitions)
+
+## Mode split
+# - for each city a table is displayed showing the Baseline mode split for each distance category
+
+# - for each city a table is given showing the mode split for each scenario
+
+# - one table showing the baseline mode split for all cities
+
+## Distances travelled
+# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey by mode for all scenarios (incl Baseline)
+
+# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)
+
+# - one table giving the city specific baseline average daily distances (km) travelled by each person in the respective travel surveys for all cities
+
+# - city specific distance tables showing total yearly distances in km travelled by the total  population considered in the model for each mode for all scenarios (incl Baseline)
+
+# - one table giving the city specific total yearly baseline distances in km travelled by the total population for each mode for all cities
+
+# - city specific distance tables showing total daily distances in km travelled by the total population for each mode for all scenarios (incl Baseline)
+
+# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)
+
+# - one table giving the city specific baseline average daily distances (km) travelled by each person in the respective travel surveys for all cities
+
+## Travel duration
+# - city specific duration tables showing average daily duration (h) spent travelling by each person in the travel survey by mode for all scenarios (incl Baseline)
+
+# - city specific duration tables showing average daily duration (h) spent travelling by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)
+
+# - one table giving the city specific baseline average duration (h) spent travelling by each person in the respective travel surveys for all cities
+
+## Health outcomes - injury, air pollution and physical activity
+# - one table for each city showing the change in total YLLs in the scenarios compared to the Baseline for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age and sex category
+
+# - one table for each city showing the change in total YLLs due to changes in accident fatalities in the scenarios compared to the Baseline for each age group
+
+# - one table for each city showing the change in total YLLs due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group
+
+# - one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalities in the scenarios compared to the Baseline for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group
+
+## Health outcomes - disease
+# - one table for each city showing the change in deaths in the scenarios compared to the Baseline for each disease
+
+# - one table for each city showing the change in deaths per 100,000 people in the scenarios compared to the Baseline for each disease
+
+# - one table for each city showing the change in YLLs in the scenarios compared to the Baseline for each disease
+
+# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each disease
+
+### - output .csv/.xlsx files containing the following information:
+
+# - the ylls.csv file (results/multi_city/health_impacts/ylls.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people, for each city, disease, scenario, age and sex category.
+It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and, the age and age and sex specific population levels.
+
+# - the ylls_pathway.csv file (results/multi_city/health_impacts/ylls_pathway.csv)
+
+# - the deaths.csv file (results/multi_city/health_impacts/deaths.csv)
+
+# - the deaths_pathway.csv file (results/multi_city/health_impacts/deaths_pathway.csv)
+
+#### The script performs the following steps:
+
+# - set up helper function to re-format large numbers and re-name scenarios
+
+# ???????????????
+
+# call get_scen_settings.R function to find the Baseline mode splits for each distance category for each city
+
 knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis', warning = F, message = F)
@@
 
-```{r helper_functions, echo = F, message = F}
-
-# Source: https://stackoverflow.com/questions/28159936/formatting-large-currency-or-dollar-values-to-millions-billions
-comprss <- function(tx) { 
-  div <- findInterval(as.numeric(gsub("\\,", "", tx)), 
-                      c(0, 1e3, 1e6, 1e9, 1e12) )  # modify this if negative numbers are possible
-  paste(round( as.numeric(gsub("\\,","",tx))/10^(3*(div-1)), 2), 
-        c("","K","M","B","T")[div] )
-}
-
-```
-
-```{r loadLibraries, echo = F, message = F}
+```{r loadLibraries, echo=F, message=FALSE}
 suppressWarnings({
@@
 
+```{r helper_functions, echo = F, message = F}
+# Function to format numbers with million (M) and billion (B) suffixes
+# Source: https://stackoverflow.com/questions/28159936/formatting-large-currency-or-dollar-values-to-millions-billions
+comprss <- function(tx) { 
+  div <- findInterval(as.numeric(gsub("\\,", "", tx)), 
+                      c(0, 1e3, 1e6, 1e9, 1e12) )  # modify this if negative numbers are possible
+  paste(round( as.numeric(gsub("\\,","",tx))/10^(3*(div-1)), 2), 
+        c("","K","M","B","T")[div] )
+}
+
+# function to rename scenarios
+get_qualified_scen_name <- function(cs){
+  qualified_scen_name <- ""
+  if (cs == 'base' | cs == 'baseline' | cs == 'Baseline')
+    qualified_scen_name <- 'Baseline'
+  else if(cs == "sc_walk")
+    qualified_scen_name <- 'Walking'
+  else if(cs == "sc_cycle")
+    qualified_scen_name <- 'Bicycling'
+  else if(cs == "sc_car")
+    qualified_scen_name <- 'Car'
+  else if(cs == "sc_motorcycle")
+    qualified_scen_name <- 'Motorcycling'
+  else if(cs == "sc_bus")
+    qualified_scen_name <- 'Public Transport'
+
+  return(qualified_scen_name)
+}
+
+# define function to sum the columns of the data, print the data to html and add a caption
+sum_and_round_and_print <- function(data, text = '') {
+  data <- lapply(data, function(x) round(x, round_to))
+  data <- lapply(data, function(x) rbind(x, Total = colSums(x)))
+  for (city in cities) {
+    print(kable(data[[city]], caption = paste(text, city)))
+    cat("\n")
+  }
+}
+
+# define function to round the data and print it to html with a caption
+round_and_print <- function(data,text=''){
+  data <- lapply(data, function(x)round(x,round_to))
+  for(city in cities) {
+    print(kable(data[[city]], caption = paste(text, city)))
+    cat("\n")
+  }
+}
+
+```
+
 ```{r load_objects = "asis", echo = F, message = F}
+
+# read in io object which is produced at the end of an ITHIM-Global run
+# Assumes that multi_city_script.R has been run 
 io <- readRDS("results/multi_city/io.rds")
-# Assumes that multi_city_script.R has been run till 
+
 # Get names of cities from the io object
 cities <<- names(io)[!names(io) %in% 'scen_prop']
-LATAM <<- F
-AFRICA_INDIA <<- F
 NSCEN <- nrow(io$scen_prop)
-#cities <- c('accra','sao_paulo','buenos_aires')
 
-# cities <- c('sao_paulo')
+# find reference scenario
+ref_scen <- io[[cities[1]]]$ithim_run$reference_scenario
+reference_scenario <- get_qualified_scen_name(ref_scen) # update scen name
 
@@
 
-# round to
+# define which decimal place to round to
 round_to <- 1
-
-# Set plot_mmets to F
-plot_mmets <- F
 
@@
 ## Scenario definition 
-Table displays highest propensity for each distance category, for a given mode.
+<!-- Table displays highest propensity for each distance category, for a given mode. -->
 
-Table displays highest propensity (%) for each distance category, for a given mode.
+<!-- Table displays highest propensity (%) for each distance category, for a given mode. -->
 
-In scenario generation, trips are sampled without replacement to increase the share of the requested mode up to the total shown.
+<!-- In scenario generation, trips are sampled without replacement to increase the share of the requested mode up to the total shown. -->
 
-Distance categories are:
+Distance categories are defined as:
 
-0-2 km = \{trip distance < 2\}
+0-2 km = \{trip distance < 2km\}
 
-2-6 km = \{2 <= trip distance < 6\}
+2-6 km = \{2km <= trip distance < 6km\}
 
-6+ km = \{6 <= trip distance\}
+6+ km = \{6km <= trip distance\}
 
@@
 
-#rownames(scen_prop) <- paste(rownames(io$scen_prop), "scen", sep = "_")
-rownames(scen_prop) <- paste0(substr(toupper(rownames(io$scen_prop)),1,3),"_SC")
+rownames(scen_prop) <- paste("sc",rownames(io$scen_prop),  sep = "_")
 
-kable(scen_prop, headings = "Scenario Proportions")
+orig_scen_names_updated <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))
+
+# # rename scenarios
+# rownames(scen_prop) <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))
+
+# kable(scen_prop, headings = "Scenario Proportions")
 
@@
 
-# ```{r warning=FALSE, echo=F, message=FALSE}
-# 
-# dat <- list()
-# 
-# for (city in cities) {
-#   
-# dat[[city]] <- io[[city]]$trip_scen_sets %>% 
-#   mutate(scenario = case_when(
-#     scenario == "Scenario 1" ~ "CYC_SC", 
-#     scenario == "Scenario 2" ~ "CAR_SC",
-#     scenario == "Scenario 3" ~ "BUS_SC",
-#     scenario == "Scenario 4" ~ "MOT_SC",
-#     scenario == "Baseline" ~ "Baseline")) %>% 
-#   group_by(scenario, trip_mode, trip_distance_cat) %>% 
-#   summarise(n = dplyr::n(), dist = sum(trip_distance, na.rm = T)) %>% 
-#   mutate(city = city)
-# 
-# td <- dat[[city]]
-# 
-# for (i in 1:nrow(td)) {
-#     val <- td %>% filter(scenario == "Baseline", trip_mode == td$trip_mode[i] & 
-#                            trip_distance_cat == td$trip_distance_cat[i] &
-#                            city == td$city[i])
-#     
-#     if (nrow(val) > 0) {
-#       td$diff[i] <- val$n - td$n[i]
-#       td$trip_prop[i] <- round(td$n[i] / val$n, 2)
-#       td$dist_prop[i] <- round(td$dist[i] / val$dist, 2)
-#       
-#     }
-# }
-# 
-# dat[[city]] <- td
-# 
-# }
-# 
-# write_csv(bind_rows(dat), "results/multi_city/scen_prop.csv")
-# 
-# ```
+## City mode split for each distance category
 
-## Case study specific propensities
 ```{r message=F, error=F, warning=FALSE}
-get_scen_settings <- ithimr::get_scenario_settings()
+# call get_scen_settings function to find the mode split for each distance category for each city
+get_scen_settings <- ithimr::get_scenario_settings(cities)
 
@@
 
-## Trip proportion
-Case study specific trip proportions by mode, for baseline and four scenarios
+## Mode split for each scenario by city
+Case study specific trip proportions by mode, for baseline and scenarios
 
@@
 ```{r trip_mode_dist = "asis", echo = F}
+# find the modal share of for all cities and scenarios
 
@@
 for (city in cities) { # Loop for each city
+
+  # extract the trip data for all scenarios
   df <- io[[city]]$trip_scen_sets
+
+  # find the number of unique trips in the Baseline
   u_trips <- df %>% dplyr::filter(scenario == "baseline") %>% 
     summarise(uid = n_distinct(trip_id)) %>% as.numeric()
+
+  # find the proportion of trips made by each mode in each scenario
   td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% 
@@
     mutate(row_sums = rowSums(.[sapply(., is.numeric)], na.rm = TRUE))
+
+  # find and update scenario names
+  scen_names <- td$scenario
+  scen_names_updated <- unlist(lapply(scen_names, FUN=get_qualified_scen_name))
+
+  # take transpose
   td <- as.data.frame(t(td))
-  names(td) <- lapply(append('BASELINE', rownames(scen_prop)), as.character)
+
+  # rename columns by using updated scenario names
+  names(td) <- scen_names_updated
+
+  # drop first column and re-order by pre-defined mode order
   td <- td[-1, ]
@@
 
+  # convert row names into column of data and create new dataframe
   td1 <- td %>% rownames_to_column()
 
+  # re-name column
   names(td1)[1] <- 'stage_mode'
 
+  # populate trip_prop list with stage modes
   if (length(trip_prop) == 0){
@@
 
-  td1 <- td1 %>% dplyr::select(stage_mode, BASELINE)
+  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
 
@@
 
+  # add proportions for all cities but only keep stage modes that appear in the trip data of all cities
   trip_prop <- inner_join(trip_prop, td1, by = 'stage_mode')
 
+  # print mode share for individual cities
   print(kable(td, caption = paste("Trip proportion (%) by mode for ", city)))
@@
 
-### Baseline trip proportion (%)
+### Baseline mode split for all cities (%)
 ```{r}
+# create baseline mode share table containing information for all cities
 
+# prep data
 trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
@@
 
-kable(trip_prop, caption = 'Trip prop (%) for all case studies - only selected modes')
+# print table to html
+kable(trip_prop, caption = 'Baseline mode share (%) for all cities - only selected modes')
 
@@
 ## Distance tables
-Case study specific distance tables for baseline and four scenarios
+City specific distance tables showing average distance in km travelled by each person by mode for all scenarios (incl Baseline)
 
 ```{r trip_dist = "asis"}
+# find the average distance a person spends travelling on each mode per day
 
@@
 
-for (city in cities) {
+for (city in cities) { # loop through cities
 
+  # find number of rows / people in the synthetic population
   count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
+
+  # extract distances and divide by number of people
+  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
     mutate_if(is.numeric, round, digits = round_to) %>% 
     mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
 
+  # update scenario names - cycle, car, bus
+  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
+
+  # match with pre-defined mode order
   x <- match(td$stage_mode, trip_order$mode)
@@
 
+  # write to .html
   print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person")))
@@
 
+  # initialise trip_dist
   if (length(trip_dist) == 0){
@@
 
-  td1 <- td1 %>% dplyr::select(stage_mode, BASELINE)
+  # extract baseline distances
+  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
 
@@
 
+  # join baseline mode distances for all cities
   trip_dist <- inner_join(trip_dist, td1, by = 'stage_mode')
 
+  ## repeat but only for people who are travelling
+  # extract number of people travelling from the trip data rather than the synthetic population
   count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
@@
     mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
+  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
   x <- match(td$stage_mode, trip_order$mode)
@@
 
-  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips)")))
+  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips ONLY)")))
 
@@
 
-### Baseline average distance (km) per day for all case studies
+### Distance table showing average baseline distances in km travelled by each person by mode for all cities
+
 ```{r}
-
+# create table showing average baseline distances by mode for all cities
 trip_dist <- trip_dist %>% janitor::adorn_totals('row')
 
-kable(trip_dist, caption = "Avg. distance per person (km) per day across all case studies")
+kable(trip_dist, caption = "Avg. baseline distance per person (km) per day across all cities")
 
@@
 
-## Distance tables scaled by using total population (per year in km)
-Case study specific distance tables for baseline and four scenarios
+## Distance tables scaled by using total model age population (per year in km)
+
+City specific distance tables for baseline and scenarios for the entire population with age range considered in the model
 
 ```{r trip_tot_dist = "asis", message=F}
-
+# create distance tables for the yearly distances travelled by the entire population
 bl_td <- list()
 
-for (city in cities) {
+for (city in cities) { # loop through cities
 
-  count_people <- round(sum(io[[city]]$demographic$population))
-  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
+  # find the total city level population considered in the model only
+  #count_people <- round(sum(io[[city]]$demographic$population))
+
+  # extract daily population distances for ages considered in model and multiply by 365 to to get yearly distances
+  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
     mutate_if(is.numeric, list(~round((.) * 365, 3)))
-  #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
 
-  #
+  # update scenario names
+  colnames(td)[2:ncol(td)] <- scen_names_updated
 
+  # match with pre-defined mode order
   x <- match(td$stage_mode, trip_order$mode)
@@
 
+  # initialise bl_td with stage_mode
   if (length(bl_td) == 0){
@@
 
+  # calculate row and column totals
   td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
 
-  td1 <- td %>% dplyr::select(stage_mode, BASELINE)
+  # extract baseline distances
+  td1 <- td %>% dplyr::select(stage_mode, Baseline)
 
@@
 
+  # join baseline distances for all cities
   bl_td <- inner_join(bl_td, td1, by = 'stage_mode')
 
+  # format table
   td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)
 
+  # print to html
+  cat(paste("Yearly distance (km) travelled by the entire population with ages considered in the model for ", city))
   print(td1)
 
-  # print(kable(td, caption = paste("Distance table (km) for ", city, "( ", sitools::f2si(count_people), " ) per person")))
   cat("\n")
@@
 
-### Baseline total distance for all case studies (per year in km)
+### Baseline total distance for all cities (per year in km)
+
+Baseline distance travelled by entire population with ages considered in the model for all cities
 
@@
 
-# tab <- bl_td %>% gt() %>% fmt_number(columns = 2:ncol(bl_td), decimals = T, suffixing = T)
+# create baseline yearly distance table for all cities
 
@@
 
+# add row totals
 td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')
 
+# remove all spaces/blanks
 for (i in 2:ncol(td)){
   td[, i] <- comprss(td[, i])
-
 }
 
-print(kable(td, caption = paste("Distance table (km)")))
+print(kable(td, caption = paste("Yearly baseline distance table (km) for population with ages considered in the model")))
 
@@
 ```{r trip_dist_mode_figs = "asis"}
+# for each city create a table showing the total distance, the number of trips and the proportion of trips in each distance category
 
-for (city in cities){
+for (city in cities){ # loop through cities
 
-  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), prop = round(n_vals / nrow(.) * 100, 1))
-  print(kable(df, caption = paste("Distance by mode by for  ", city)))
+  # extract trip data for the baseline scenario and calculate the total distance, the number of trips and the proportion of trips in each distance category
+  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline' | scenario == 'baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% 
+    summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), proportion = round(n_vals / nrow(.) * 100, 1))
+
+ # print to html
+  print(kable(df, caption = paste("Distance in travel survey by mode for the ages considered in the model for  ", city)))
 }
@@
 
-## True Distance by total population (per day in km)
+## Daily population distances (per day in km) for ages considered in model
 ### Tables
@@
 
-for (city in cities){
+for (city in cities){ # loop through distances
 
+  # extract total daily distances travelled by the entire population with ages considered in the model
   df <- io[[city]]$true_dist
 
-  names(df)[-1] <- lapply(append('BASELINE', rownames(scen_prop)), as.character)
+  names(df)[-1] <- scen_names_updated
 
-  print(kable(df, caption = paste("True distance by mode per day for  ", city)))
+  print(kable(df, caption = paste("Model age population distance by mode per day for  ", city)))
 }
@@
 
-<!-- ### Figures -->
-
-<!-- ```{r trip_dist_mode_figs = "asis"} -->
-
-<!-- for (city in cities){ -->
-
-<!--   df <- io[[city]]$trip_scen_sets %>% filter(scenario == 'Baseline' & trip_mode %in% c('walking', 'bicycle', 'car', 'motorcycle', 'bus')) %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_mode, trip_distance_cat) %>% summarise(sum_dist = sum(trip_distance), n_vals = n()) %>%  -->
-<!--     group_by(trip_mode) %>% mutate(n_vals_1 = sum(n_vals)) %>%  -->
-<!--     mutate(prop = round(n_vals/n_vals_1 * 100, 1)) -->
-
-<!--   print(ggplot(data=df, aes(x=trip_distance_cat, y=prop)) + -->
-<!--           geom_bar(stat="identity", fill="steelblue")+ -->
-<!--           facet_wrap(vars(trip_mode)) + -->
-<!--           geom_text(aes(label=prop), vjust = -0.3, size = 3.5)+ -->
-<!--           theme_minimal() + -->
-<!--           theme(panel.spacing = unit(2, "lines")) + -->
-<!--           labs(title = paste(city, ' by dist by mode'))) -->
-<!-- } -->
-
-<!-- ``` -->
-
 ## Duration tables
 
-Case study specific duration tables for baseline and four scenarios (per day in minutes)
+City specific duration tables for baseline and scenarios (per day in minutes)
 
 ```{r trip_dur = "asis"}
+# create tables showing the average duration each person spends travelling per mode
 
 trip_dur <- list()
-
 l <- list()
 
-for(city in cities){
+for(city in cities){ # loop through cities
   count_people <- nrow(io[[city]]$synth_pop)
-  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
+
+  # extract the duration for the synthetic population
+  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+
+  # update scenario names
+  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
+
+  # order by pre-defined mode order
   x <- match(td$stage_mode, trip_order$mode)
@@
 
+  # add output to list
   l[[city]] <- td
@@
 
+  # initialise list for all cities
   if (length(trip_dur) == 0){
@@
 
-  td1 <- td1 %>% dplyr::select(stage_mode, BASELINE)
+  # extract baseline data
+  td1 <- td1 %>% dplyr::select(stage_mode, Baseline)
 
@@
 
+  # create one baseline table containing all city information
   trip_dur <- inner_join(trip_dur, td1, by = 'stage_mode')
 
-  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (everyone)")))
+  # print to html
+  print(kable(td, caption = paste("Duration table (mins) per day for model synthetic population for ", city, "( ", count_people, " ) per person (everyone)")))
 
-  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
-  #colnames(td)[3:7] <- rownames(scen_prop)
-  colnames(td)[2:ncol(td)] <- c("BASELINE", rownames(scen_prop))
+  # repeat but only counting people who are actually taking a trip
+  count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id)) # find unique trip ids rather than rows in synthetic population
+  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
   x <- match(td$stage_mode, trip_order$mode)
@@
 
-  print(kable(td, caption = paste("Duration table (mins) per day for ", city, "( ", count_people, " ) per person (people with trips)")))
+  print(kable(td, caption = paste("Duration table (mins) per day based on trips taken in travel survey for ", city, "( ", count_people, " ) per person (people with trips)")))
   cat("\n")
@@
 
-### Baseline average duration (mins) per day for all case studies
+### Baseline average duration (mins) per day for all cities
+
 ```{r}
+# Average baseline duration travelled on each mode for all cities for all people in synthetic population
 
+# calculate the row sum
 trip_dur <- trip_dur %>% janitor::adorn_totals('row')
 
-kable(trip_dur, caption = "Avg. duration per person (mins) per day across all case studies")
+kable(trip_dur, caption = "Avg. duration (min) per person in the synthetic population per day across all cities")
 
@@
 
-### Change in YLL {#change_YLL}
+### Change in YLLs compared to reference scenario {#change_YLL}
+
 ```{r scen_prop = "asis"}
-injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths) == 'scen1_deaths_inj')
-ap_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))
+
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# Calculate the changes in YLL compared to the reference scenario by age groups
+
+# find injury death column
+#injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths) == paste0(scen_names[2],'_deaths_inj'))
+
+# find ap death columns
+# ap_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))
+
+# find ap ylls columns
 ap_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))
-pa_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))
+
+# find pa death columns
+# pa_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))
+
+# find pa yll columns
 pa_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))
-sum_and_round_and_print <- function(data, text = '') {
-  data <- lapply(data, function(x) round(x, round_to))
-  data <- lapply(data, function(x) rbind(x, Total = colSums(x)))
-  for (city in cities) {
-    print(kable(data[[city]], caption = paste(text, city)))
-    cat("\n")
-  }
-}
-round_and_print <- function(data,text=''){
-  data <- lapply(data, function(x)round(x,round_to))
-  for(city in cities) {
-    print(kable(data[[city]], caption = paste(text, city)))
-    cat("\n")
-  }
-}
-scen_names <- rownames(scen_prop)
-#yll_totals <- lapply(io,function(x){
+
+# # find scenario names of health burden as the reference scenario is not necessarily the baseline scenario
+# hb_columns <- names(io[[cities[1]]]$outcomes$hb$ylls)[3: length(names(io[[cities[1]]]$outcomes$hb$ylls))]
+# 
+# # find the unique scenario names in the hb columns
+# inj_scen_cols <- grep("inj", hb_columns) # find location of injury columns
+# # extract unique scenario names
+# hb_scen_names <- unique(sapply(stringr::str_split(hb_columns[inj_scen_names], "_yll_inj", n = 3), function(x) x[1]))
+# # update scenario names
+# hb_scen_names_upd <- unlist(lapply(hb_scen_names, FUN=get_qualified_scen_name))
+
+# from the known scenarios, remove the reference scenario
+all_orig_scen_upd <- c('Baseline',orig_scen_names_updated)
+hb_scen_names_upd <- all_orig_scen_upd[all_orig_scen_upd!=reference_scenario]
+
+# calculate ylls by age category and scenario
+# i.e. sum over all disease and remove sex 
 yll_totals <- lapply(io[-2],function(x){
-  temp <- sapply(1:NSCEN,function(y){
+  temp <- sapply(1:NSCEN,function(y){# loop through scenarios
     xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
-    xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
+    xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)]) # sum over all diseases
     sapply(sort(unique(xx$age_cat)),function(z)
-      sum(xxx[xx$age_cat == z]))
+      sum(xxx[xx$age_cat == z])) # sum age categories ignoring sex
   })
-  colnames(temp) <- scen_names
+  colnames(temp) <- hb_scen_names_upd # update scenario names
   temp
 })
-sum_and_round_and_print(yll_totals,"Change in YLL total in ")
+
+# add the column total, round the results and print to html
+sum_and_round_and_print(yll_totals,"Change total in YLLs compared to the reference scenario in ")
 ```
+
+### Change in YLLs compared to reference scenario per 100,000 people {#change_YLL_age_100k}
 
@@
 
-### Change in YLLs per 100,000 {#change_YLL_age_100k}
 ```{r scen_prop = "asis"}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate change in YLLs compared to the reference scenario per 100,000 people
+
+# extract population statistics
 pop_by_age <- lapply(io[-2], function(x)
@@
   )
+
+# calculate ylls per 100,000 people
 yll_rates <- lapply(cities, function(x)
-  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000,
-        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), 
+  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000, # divide by population
+        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), # calculate rate for total YLLs
                                              length = NSCEN)*100000
@@
   )
+
 names(yll_rates) <- cities
+
+# round results and print to html
 round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
@@
 
-YLLs per 100,000 people by gender by age group scenario
+### Change in YLLs compared to reference scenario per 100,000 people by age and sex  {#change_YLL_gender_age_100k}
 
-### Change in YLLs per 100,000 {#change_YLL_gender_age_100k}
+Change in YLLs per 100,000 people by age and sex category for each scenario compared to the reference scenario
 ```{r scen_prop = "asis"}
-#yll_totals <- lapply(io,function(x){
+cat(' ')
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate changes in YLLs for each scenario compared to the Baseline per 100,000 people by age and sex category
+
+# calculate the total ylls for each scenario summing across all diseases
 yll_totals <- lapply(io[-2], function(x) {
@@
                           function(z)paste0(z[2], '_', z[1]))
-  colnames(temp) <- scen_names
+  colnames(temp) <- orig_scen_names_updated
   temp
 })
+
+# calculate the YLLs per 100,000 people
 yll_rates <- lapply(cities, function(x)
@@
                                              length = NSCEN) * 100000))
+
 names(yll_rates) <- cities
+
+# round and print ylls per 100,000 people to html
 round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
@@
 
-### Change in YLLs due to injury {#change_yll_injury}
+### Change in YLLs due to injury compared to the reference scenario  {#change_yll_injury}
+
 ```{r}
+
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate change in YLLs due to injury
+
+# calculate change in YLLs due to injury by summing across all age categories for the injury results
 injury_totals <- lapply(io[-2], function(x) {
   xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
-  injury_col <- grep("inj", names(xx))[1] # Dan: This line is mine (not sure is ok)
+  injury_col <- grep("inj", names(xx))[1] 
   xxx <- sapply(1:NSCEN, function(y)
@@
       sum(xx[xx$age_cat == z, injury_col - 1 + y])))
-  colnames(xxx) <- scen_names
+  colnames(xxx) <- hb_scen_names_upd
   xxx
 })
+
+# calculate the total, round the results and print to html
 sum_and_round_and_print(injury_totals, "Change in YLLs due to injury in ")
@@
 
-### Change in YLLs due to PA {#change_yll_PA}
+### Change in YLLs due to PA compared to the reference scenario {#change_yll_PA}
+
 ```{r}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate the changes in YLLs compared to the reference scenario caused by changes in PA for the unique age categories
+
+# sum the YLLs for each scenario by age category
 pa_totals <- lapply(io[-2],function(x){
@@
   })
-  colnames(xxx) <- scen_names
+  colnames(xxx) <- hb_scen_names_upd
   xxx
 })
+
+# calculate column sums, round the results and print to html
 sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
@@
 
-### Change in YLLs due to AP {#change_yll_AP}
+### Change in YLLs due to AP compared to the reference scenario {#change_yll_AP}
 ```{r}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate the changes in YLLS compared to the reference scenario due to changes in AP
+
+# sum the YLLs by scenario and age category
 ap_totals <- lapply(io[-2],function(x){
@@
   })
-  colnames(xxx) <- scen_names
+  colnames(xxx) <- hb_scen_names_upd
   xxx
 })
+
+# calculate column sums, round the results and print to html
 sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ")
@@
 
-YLLs per 100,000 people by age group by scenario
+### Change in YLLs due to injury compared to the reference scenario per 100,000 people {#change_yll_injury_100k}
 
-### Change in YLLs due to injury per 100,000 {#change_yll_injury_100k}
+YLLs per 100,000 people by age group by scenario - injury
+
 ```{r scen_prop = "asis"}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to injury
 injury_rates <- lapply(cities, function(x)
@@
 names(injury_rates) <- cities
-round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 in ")
+
+# round results and print to html
+round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 people in ")
 ```
 
-### Change in YLLs due to PA per 100,000 {#change_yll_PA_100k}
+### Change in YLLs due to PA compared to the reference scenario per 100,000 people {#change_yll_PA_100k}
+
 ```{r}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to PA
 pa_rates <- lapply(cities, function(x)
@@
 names(pa_rates) <- cities
-round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 in ")
+
+# round results and print to html
+round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 people in ")
 ```
 
-### Change in YLLs due to injury AP 100,000 {#change_yll_AP_100k}
+### Change in YLLs due to AP compared to the reference scenario per 100,000 people {#change_yll_AP_100k}
+
 ```{r}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to AP
 ap_rates <- lapply(cities, function(x)
@@
 names(ap_rates) <- cities
-round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 in ")
+
+# round results and print to html
+round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 people in ")
 ```
@@
 ## BY DISEASE
-Change in deaths total (for the city based on real population size) by scenario
 
-### Change in deaths due to disease {#change_death_disease}
+Change in total deaths compared to the reference scenario (for the city based on real population size) by scenario
+
+### Change in deaths by disease compared to the reference scenario {#change_death_disease}
+
 ```{r}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate the changes in deaths compared to the reference scenario by disease
+
+# sum the total deaths by scenario
 disease_totals <- lapply(io[-2],function(x){
+  # extract disease counts
   xx <- x$outcomes$hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) 
-  xxx <- sapply(1:NSCEN,function(y){
+  xxx <- sapply(1:NSCEN,function(y){ # sum over scenarios
     colSums(xx[, seq(y + 2, ncol(xx), by = NSCEN)])
   })
-  colnames(xxx) <- scen_names
-  rownames(xxx) <- gsub("scen1_deaths_", "", rownames(xxx))
-  #rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_', '', y))
-  # rownames(xxx) <- c("PA_AllCauses", "PA_AP_IschemicHeartDisease",
-  #                    "PA_Neoplasms",
-  #                    "PA_AP_LungCancer", "AP_COPD", "PA_AP_Stroke",
-  #                    "PA_AP_T2Diabetes", "AP_LowerRespiratoryInfections",
-  #                    "PA_BreastCancer", "PA_ColonCancer", "PA_UterineCancer",
-  #                    "Injuries")
+  colnames(xxx) <- hb_scen_names_upd
+
+  # find scenario name of row names
+  row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2]))
+
+  # update row names
+  rownames(xxx) <- gsub(paste0("sc_",row_scen,"_deaths_"), "", rownames(xxx))
+
   xxx
 })
-sum_and_round_and_print(disease_totals,"Change in deaths due to disease in ")
+
+# sum, round and print the results 
+sum_and_round_and_print(disease_totals,"Change in deaths by disease in ")
 ```
 
-Change in deaths per 100,000 people by scenario
+### Change in deaths by disease compared to the reference scenario per 100,000 people {#change_death_disease_100k}
 
-### Change in deaths due to disease per 100,000 {#change_death_disease_100k}
+Change in deaths per 100,000 people by scenario in comparison to the reference scenario
+
 ```{r scen_prop = "asis"}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate the change in deaths compared to the reference scenario per 100,000 people
+
+# divided total number of deaths by population
 disease_rates <- lapply(cities, function(x)
@@
 names(disease_rates) <- cities
-round_and_print(disease_rates,"Change in deaths due to disease per 100,000 in ")
+
+# round results and print to html
+round_and_print(disease_rates,"Change in deaths by disease per 100,000 people in ")
 ```
 
-Change in YLL total (for the city based on real population size) by scenario
+### Change in total YLLs by disease compared to the reference scenario {#change_yll_disease}
 
-### Change in YLL due to disease {#change_yll_disease}
+Change in total YLLs by disease compared to the reference scenario (for the city based on real population size, for ages considered within the ITHIM model)
+
 ```{r}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# sum the total ylls by scenario
 disease_totals <- lapply(io[-2],function(x){
@@
   })
-  colnames(xxx) <- scen_names
-  rownames(xxx) <- gsub("scen1_ylls_", "", rownames(xxx))
-  #rownames(xxx) <- sapply(rownames(xxx), function(y) gsub('scen1_','',y))
-  # rownames(xxx) <- c("PA_AllCauses", "PA_AP_IschemicHeartDisease",
-  #                    "PA_Neoplasms",
-  #                    "PA_AP_LungCancer", "AP_COPD", "PA_AP_Stroke",
-  #                    "PA_AP_T2Diabetes", "AP_LowerRespiratoryInfections",
-  #                    "PA_BreastCancer", "PA_ColonCancer", "PA_UterineCancer",
-  #                    "Injuries")
+  colnames(xxx) <- hb_scen_names_upd
+
+  # find scenario name of row names
+  row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2]))
+
+  # update row names
+  rownames(xxx) <- gsub(paste0("sc_",row_scen,"_ylls_"), "", rownames(xxx))
   xxx
 })
-sum_and_round_and_print(disease_totals,"Change in YLL due to disease in ")
+
+# sum and round results, print to html
+sum_and_round_and_print(disease_totals,"Change in YLL by disease in ")
 ```
 
-### Change in YLL due to disease per 100,000 {#change_yll_disease_100k}
+### Change in YLL due to disease compared to the reference scenario per 100,000 people {#change_yll_disease_100k}
+
 ```{r scen_prop = "asis"}
+cat(paste0('Reference scenario:  ', reference_scenario))
+
+# calculate the change in YLL by disease by dividing by the population
 disease_rates <- lapply(cities, function(x)
@@
 names(disease_rates) <- cities
-round_and_print(disease_rates,"Change in YLL due to disease per 100,000 in ")
+
+# round and print results to html
+round_and_print(disease_rates,"Change in YLL by disease per 100,000 in ")
 ```
@@
 ```{r, echo=FALSE, results='hide', message=FALSE}
+
 # Export all health results
+
+# hb - gives all AP (for diseases affected by AP only), PA (for diseases affected by PA only), combined AP and PA (for diseases affected by both AP and PA) and inj results
+# pathway_hb - gives all AP, PA and inj results (but does not combine the results for diseases affected by both AP and PA)
+
+# initialise variables
 health <- list()
@@
 pathway <- c("hb", "pathway_hb")
+
 for (k in measure) { # loop for ylls or deaths
@@
     for (i in names(io[-2])) { # loop for cities
-      # Temporal datset
+
+      # Temporal dataset for health outcomes
       temp <- io[[i]]$outcomes[[j]][[k]]
+
       # Population by age, and age-sex
@@
 
-      # Long format and adding new variables
+      # get the health burden into the correct format by creating a long format and adding additional columns
+      # create long format
       health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure", 
                                       -sex, -age_cat) %>% 
-        mutate(city = i,
-               scenario = case_when(
-                 grepl("scen1", cause) ~ scen_names[1],
-                 grepl("scen2", cause) ~ scen_names[2],
-                 grepl("scen3", cause) ~ scen_names[3],
-                 grepl("scen4", cause) ~ scen_names[4],
-                 grepl("scen5", cause) ~ scen_names[5],
+        mutate(city = i, # create city column
+               scenario = case_when( # create scenario column
+                 grepl("sc_cycle", cause) ~ get_qualified_scen_name('sc_cycle'),
+                 grepl("sc_car", cause) ~ get_qualified_scen_name('sc_car'),
+                 grepl("sc_bus", cause) ~ get_qualified_scen_name('sc_bus'),
+                 grepl("sc_motorcycle", cause) ~ get_qualified_scen_name('sc_motorcycle')
                ),
-               level1 = case_when(
+               level1 = case_when( # create level 1 column
                  grepl("all_cause", cause) ~ "L1: All Cause",
@@
                ),
-               level2 = case_when(
+               level2 = case_when( # create level 2 column
                  grepl("total_cancer", cause) ~ "L2: Cancer",
                  grepl("CVD", cause) ~ "L2: CVD",
-                 grepl("COPD", cause) ~ "L2: Respiratory",
-                 grepl("LRI", cause) ~ "L2: Respiratory",
+                 grepl("respiratory", cause) ~ "L2: Respiratory",
                  grepl("T2D", cause) ~ "L2: Other",
@@
                ),
-               level3 = case_when(
+               level3 = case_when( # create level 3 column
                  grepl("IHD", cause) ~ "L3: IHD",
@@
                ),
-               dose = case_when(
+               dose = case_when( # create dose column
                  grepl("_pa_ap_", cause) ~ "PA and AP",
@@
         ) %>% # End mutate 
+        # join with demographic information
         left_join(pop_by_age, by = c("age_cat" = "age")) %>%  
         left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>%
+        # create new columns showing burden per 100000 people by age or by age and sex
         mutate(metric_100k = measure / pop_age * 100000, 
@@
 } # Loop for each measure
-# # Export deaths HB
-# path_deaths_wb <- "results/multi_city/health_impacts/Deaths.xlsx"
-# deaths_wb <- readWorkbook(path_deaths_wb)
-# wb <-  loadWorkbook(path_deaths_wb)
-# removeTable(wb = wb, sheet = "Deaths", table = "Tabla1")
-# writeDataTable(wb, sheet = "Deaths", x = rbindlist(health$deaths$hb, use.names = T), tableName = "Tabla1", startCol = 1)
-# removeTable(wb = wb, sheet = "Deaths_Pathway", table = "Tabla2")
-# writeDataTable(wb, sheet = "Deaths_Pathway", x = rbindlist(health$deaths$pathway_hb, use.names = T), tableName = "Tabla2", startCol = 1)
-# saveWorkbook(wb, path_deaths_wb, overwrite =TRUE)
-# 
-# # Export ylls HB
-# path_ylls_wb <- "results/multi_city/health_impacts/YLLs.xlsx"
-# ylls_wb <- readWorkbook(path_ylls_wb)
-# wb <-  loadWorkbook(path_ylls_wb)
-# removeTable(wb = wb, sheet = "YLLs", table = "Tabla1")
-# writeDataTable(wb, sheet = "YLLs", x = rbindlist(health$ylls$hb, use.names = T), tableName = "Tabla1", startCol = 1)
-# removeTable(wb = wb, sheet = "YLLs_Pathway", table = "Tabla2")
-# writeDataTable(wb, sheet = "YLLs_Pathway", x = rbindlist(health$ylls$pathway_hb, use.names = T), tableName = "Tabla2", startCol = 1)
-# saveWorkbook(wb, path_ylls_wb, overwrite =TRUE)
-#
 
-# Remove CIs
+# for each health burden and pathway join all cities into one dataframe
 ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
@@
 ```
-
-## PA dist
-```{r}
-### pa distr
-cities <- names(io)[!names(io) %in% 'scen_prop']
-l <- list()
-for (city in cities){
-
-  # city <- "antofagasta"
-
-  td <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
-    rename("Baseline" = base_mmet, "CYC_SC" = sc_cycle_mmet,
-           "CAR_SC" = sc_car_mmet,  "BUS_SC" = sc_bus_mmet,
-           "MOT_SC" = sc_motorcycle_mmet) %>% 
-    summarise(across(, ~ quantile(.x, c(0.05,0.10,0.25, 0.375, 0.50,0.75,0.90), na.rm = T)), 
-              prob = c(0.05,0.10,0.25, 0.375, 0.50,0.75,0.90)) %>%
-    pivot_longer(cols = -c(prob)) %>% mutate(city = city)
-
-  td1 <- io[[city]]$outcomes$mmets %>% dplyr::select(ends_with("mmet")) %>%
-     rename("Baseline" = base_mmet, "CYC_SC" = sc_cycle_mmet,
-           "CAR_SC" = sc_car_mmet,  "BUS_SC" = sc_bus_mmet,
-           "MOT_SC" = sc_motorcycle_mmet) %>% 
-    summarise(across(, ~ mean(.x))) %>%  t() %>% as.data.frame() %>% 
-    tibble::rownames_to_column("name") %>% 
-    rename(value = V1) %>% 
-    mutate(prob = "mean") %>% mutate(city = city)
-
-  l[[city]] <- plyr::rbind.fill(td, td1)
-
-}
-# write_csv(data.table::rbindlist(l), "results/multi_city/PA/pa_distr.csv")
-```

### 2023-08-23 summary_tables_PA_AP.Rmd
updated documentation
updated how the reference scenario is used

@@
   qualified_scen_name <- ""
-  if (cs == 'base')
+  if (cs == 'base' | cs == 'Baseline' | cs == 'baseline')
     qualified_scen_name <- 'Baseline'
@@
 scenario_name <- io[[city]]$ithim_run$scenarios_used
-reference_scenario <- io[[city]]$ithim_run$reference_scenario
+ref_scen <- io[[cities[1]]]$ithim_run$reference_scenario
+reference_scenario <- get_qualified_scen_name(ref_scen)
 
@@
 
+cat(paste0('The change in PM2.5 levels is given as the difference to the reference scenario: ', reference_scenario))
+
 # for each city create a summary table of individual PM2.5 exposure levels giving the mean and the 5th, 50th and 95th percentile levels.
@@
   summary$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, digits=2)
-  summary$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits=2)
+
+  # find location of reference scenario in vector of all scenarios
+  loc_ref <- which(new_scen_names == reference_scenario)
+
+  # calculate the PM2.5 changes between the reference scenario and the other scenarios
+  summary$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[loc_ref], digits=2)
   names(summary)[1] <- "Scenario"
@@
 
-# Average physical activity levels (MMETs) in each city
+# Average baseline physical activity levels (MMETs) in each city
 
@@
 
-  # calculate difference to baseline scenario
-  summary2$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[1], digits = 2)
+  # calculate difference to reference scenario
+  summary2$change_PM <- round(io[[cities[x]]]$outcomes$scenario_pm - io[[cities[x]]]$outcomes$scenario_pm[loc_ref], digits = 2)
   names(summary2)[1] <- "Scenario"
@@
   cat( paste('#### Summary statistics of individual level PM2.5 exposure levels and total transport related CO2 emissions\n' ))
+  cat(paste0('The change in PM2.5 levels is given as the difference to the reference scenario: ', reference_scenario))
   l[[cities[x]]] <- left_join(summary2, td, by = "Scenario")

### 2023-08-24 summary_tables.Rmd
updated comments

@@
 
-# Script summarises mode splits, distances and durations travelled and looks at health outcomes for the different air pollution, injury and physical activity pathways but also by disease
+# Script summarises mode splits, distances and durations travelled and looks at health outcomes (YLLs and deaths) for the different air pollution, injury and physical activity pathways but also by disease
 
@@
 ## Distances travelled
-# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey by mode for all scenarios (incl Baseline)
-
-# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)
-
-# - one table giving the city specific baseline average daily distances (km) travelled by each person in the respective travel surveys for all cities
-
-# - city specific distance tables showing total yearly distances in km travelled by the total  population considered in the model for each mode for all scenarios (incl Baseline)
-
-# - one table giving the city specific total yearly baseline distances in km travelled by the total population for each mode for all cities
-
-# - city specific distance tables showing total daily distances in km travelled by the total population for each mode for all scenarios (incl Baseline)
-
-# - city specific distance tables showing average daily distances in km travelled by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)
-
-# - one table giving the city specific baseline average daily distances (km) travelled by each person in the respective travel surveys for all cities
+# - city specific distance tables showing average daily distances in km travelled by each person in the synthetic population by mode for all scenarios (incl Baseline)
+
+# - city specific distance tables showing average daily distances in km travelled by each person in the synthetic population WITH a trip by mode for all scenarios (incl Baseline)
+
+# - one table giving the city specific baseline average daily distances (km) travelled by each person in the synthetic population for all cities
+
+# - city specific distance tables showing total yearly distances in km travelled by the total  population with ages considered in the model for each mode for all scenarios (incl Baseline)
+
+# - one table giving the city specific total yearly baseline distances in km travelled by the total population with ages considered in the model for each mode for all cities
+
+# - city specific distance tables showing total distances travelled by distance category toegether with the number of trips and proportions of all trips within the distance categories.
+
+# - city specific distance tables showing total daily distances in km travelled by the total population  with ages considered in the model for each mode for all scenarios (incl Baseline)
 
 ## Travel duration
-# - city specific duration tables showing average daily duration (h) spent travelling by each person in the travel survey by mode for all scenarios (incl Baseline)
-
-# - city specific duration tables showing average daily duration (h) spent travelling by each person in the travel survey WITH a trip by mode for all scenarios (incl Baseline)
-
-# - one table giving the city specific baseline average duration (h) spent travelling by each person in the respective travel surveys for all cities
+# - city specific duration tables showing average daily duration (h) spent travelling by each person in the synthetic population by mode for all scenarios (incl Baseline)
+
+# - city specific duration tables showing average daily duration (h) spent travelling by each person in the synthetic population WITH a trip by mode for all scenarios (incl Baseline)
+
+# - one table giving the city specific baseline average duration (h) spent travelling by each person in the respective synthetic populations for all cities
 
@@
 
-# - the ylls.csv file (results/multi_city/health_impacts/ylls.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people, for each city, disease, scenario, age and sex category.
-It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and, the age and age and sex specific population levels.
-
-# - the ylls_pathway.csv file (results/multi_city/health_impacts/ylls_pathway.csv)
-
-# - the deaths.csv file (results/multi_city/health_impacts/deaths.csv)
-
-# - the deaths_pathway.csv file (results/multi_city/health_impacts/deaths_pathway.csv)
+# - the ylls.csv file (results/multi_city/health_impacts/ylls.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
+It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age and age and sex specific population levels.
+The results are for either AP and PA only if the diseases are affected by only AP or PA or the results for the combined AP and PA effects if the diseases are affected by both AP and PA.
+
+# - the ylls_pathway.csv file (results/multi_city/health_impacts/ylls_pathway.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
+It also shows the disease level, whether the disease is exasperated by changes in AP or PA, and the age and age and sex specific population levels.
+The results are given for AP and PA independently even if a disease is affected by both pathways.
+
+# - the deaths.csv file (results/multi_city/health_impacts/deaths.csv) gives the changes in deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
+It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age and age and sex specific population levels.
+The results are for either AP and PA only if the diseases are affected by only AP or PA or the results for the combined AP and PA effects if the diseases are affected by both AP and PA.
+
+# - the deaths_pathway.csv file (results/multi_city/health_impacts/deaths_pathway.csv) gives the changes in deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
+It also shows the disease level, whether the disease is exasperated by changes in AP or PA, and the age and age and sex specific population levels.
+The results are given for AP and PA independently even if a disease is affected by both pathways.
 
@@
 
-# - set up helper function to re-format large numbers and re-name scenarios
-
-# ???????????????
-
-# call get_scen_settings.R function to find the Baseline mode splits for each distance category for each city
+# - set up helper function to re-format large numbers, re-name scenarios and format and print output tables
+
+# - read in io object containing the results from an ITHIM-Global run and extract relevant variables. Read in mode_order.csv which sets up a set mode order
+
+# (extract table displaying highest propensity (%) for each distance category, for a given mode.
+- CURRENTLY commented out as not relevant for BOGOTA, GLOBAL, LATAM or AFRICA_INDIA scenario definitions)
+
+# - call get_scen_settings.R function to find the Baseline mode splits for each distance category for each city and print tables to html file
+
+# - for each city create a table showing the mode split for each scenario
+
+# - create one table showing the baseline mode split for all cities
+
+# - create city specific distance tables showing average daily distances in km travelled by each person in the synthetic population by mode for all scenarios (incl Baseline)
+
+# - create  city specific distance tables showing average daily distances in km travelled by each person in the synthetic population WITH a trip by mode for all scenarios (incl Baseline)
+
+# - create one table giving the city specific baseline average daily distances (km) travelled by each person in the synthetic population for all cities
+
+# - create city specific distance tables showing total yearly distances in km travelled by the total  population with ages considered in the model for each mode for all scenarios (incl Baseline)
+
+# - create one table giving the city specific total yearly baseline distances in km travelled by the total population with ages considered in the model for each mode for all cities
+
+# - create city specific distance tables showing total distances travelled by distance category together with the number of trips and proportions of all trips within the distance categories
+
+# - create city specific distance tables showing total daily distances in km travelled by the total population  with ages considered in the model for each mode for all scenarios (incl Baseline)
+
+# - create city specific duration tables showing average daily duration (h) spent travelling by each person in the synthetic population by mode for all scenarios (incl Baseline)
+
+# - create city specific duration tables showing average daily duration (h) spent travelling by each person in the synthetic population WITH a trip by mode for all scenarios (incl Baseline)
+
+# - create one table giving the city specific baseline average duration (h) spent travelling by each person in the respective synthetic populations for all cities
+
+# - create one table for each city showing the change in total YLLs in the scenarios compared to the Baseline for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age and sex category
+
+# - create one table for each city showing the change in total YLLs due to changes in accident fatalities in the scenarios compared to the Baseline for each age group
+
+# - create one table for each city showing the change in total YLLs due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group
+
+# - create one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalities in the scenarios compared to the Baseline for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group
+
+# - create one table for each city showing the change in deaths in the scenarios compared to the Baseline for each disease
+
+# - create one table for each city showing the change in deaths per 100,000 people in the scenarios compared to the Baseline for each disease
+
+# - create one table for each city showing the change in YLLs in the scenarios compared to the Baseline for each disease
+
+# - create one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each disease
+
+# - create 4 individual .csv files in (results/multi_city/health_impacts/) for both death and YLLs considering the effect of AP and PA separately (*_pathway.csv) or together (*.csv).
+These files give the changes in YLLs / deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
+They also show the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age and age and sex specific population levels.
 
@@
 ## Mode split for each scenario by city
-Case study specific trip proportions by mode, for baseline and scenarios
+City specific trip proportions by mode, for baseline and scenarios
 
@@
 ```{r trip_mode_dist = "asis", echo = F}
-# find the modal share of for all cities and scenarios
+# find the modal share for all cities and scenarios
 

### 2023-08-24 summary_tables.Rmd
updated comments

@@
 
-# WARNING: This script only works if ITHIM-Global has been run in constant mode!
+# WARNING: This script only works if ITHIM-Global has been run in constant mode using either the BOGOTA, LATAM, AFRICA_INDIA or GLOBAL scenario definitions
 
@@
 # call get_scen_settings function to find the mode split for each distance category for each city
-get_scen_settings <- ithimr::get_scenario_settings(cities)
+
+get_scen_settings <- ithimr::get_scenario_settings(cities = cities)
 

### 2023-08-24 R/add_ghost_trips.R

@@ -1,7 +1,7 @@
 #' Addition of ghost trips
 #'
 #' Add trips that do not get assigned any physical activity component, 
-#' often used as proxy for vehicle distances.
+#' can be used as proxy for vehicle distances.
 #' 
 #' This function creates trips based on distance relative to another (reference) mode, 
 #' without altering the synthetic population. I.e. it adds trips that do not get assigned any physical 

### 2023-08-24 R/combine_health_and_pif.R

@@ -1,7 +1,7 @@
 #' Combine health and potential impact fraction (PIF)
 #' 
 #' Applies PIF calculated from relative risks (RRs) to the current observed health burden from the 
-#' GBD to generate the scenario health burdens
+#' Global Burden of Disease data to generate the scenario health burdens
 #' 
 #' This function performs the following steps:
 #'

### 2023-08-24 R/compute_evppi.R

@@ -1,4 +1,4 @@
-#' Compute evppi, designed to be run in parallel
+#' Compute evppi, designed to be run in parallel - ONLY USED in SAMPLING MODE
 #' 
 #' Creates a list of EVPPI values one parameter (set) at a time across some pre-defined outcomes
 #' Uses Chris Jackson's VoI Github package https://github.com/chjackson/voi

### 2023-08-24 R/create_africa_india_scenarios.R

@@ -1,14 +1,12 @@
-#' Create scenarios based on difference between median and max propensities for all cities
+#' Create scenarios for African and Indian cities - NOT CURRENTLY USED
 #' 
-#' Creates four scenarios where, in each one, the mode share is elevated to the
-#' certain values in each distance band. The scenario-modes are cycle, car, and bus and motorcycle.
+#' Creates three scenarios where in each one, the mode share of a given mode is elevated by a set
+#' percentage of the total trips. The scenario-modes are cycle, car, and bus and motorcycle.
 #' 
-#' Based on create_latam_scenarios
-#' 
+#'  
 #' Add 5% of trips overall in such a way that the average mean mode share for each mode across the
 #' three distance bands is preserved.
 #' 
-#' https://www.dropbox.com/home/ITHIM%20Global/Methods%20and%20Processes/ScenarioDefn/GlobalScenario
 #' 
 #' @param trip_set data frame, baseline scenario
 #' 

### 2023-08-24 R/create_bogota_scenarios.R

@@ -1,6 +1,6 @@
 #' Create scenarios for Bogota
 #' 
-#' Creates three scenarios where, in each one, the mode share of a given mode is elevated by a set
+#' Creates three scenarios where in each one, the mode share of a given mode is elevated by a set
 #' percentage of the total trips. The scenario modes are cycle, car, and bus.
 #' 
 #' This function creates three scenarios increasing the mode shares of cycling, car
@@ -31,6 +31,7 @@
 #'   trips are still in distance band 0-2km.
 #'}   
 #'   
+#'   
 #' Example:
 #' 
 #' Assume that there are only two distance bands A and B and that 80% of all 
@@ -50,6 +51,7 @@
 #' band A and 20% in distance band B and preserving the total mode shares of 60% in
 #' distance band A and 40% in distance band B. The total number of trips is also
 #' preserved.
+#' 
 #' 
 #' The function performs the following steps:
 #' 

### 2023-08-24 R/create_cycle_scenarios.R

@@ -1,4 +1,4 @@
-#' Creates cycling scenarios
+#' Creates cycling scenarios - NOT CURRENTLY USED
 #' 
 #' Creates five scenarios with 10-50% cycling
 #' 

### 2023-08-24 R/create_global_scenarios.R

@@ -1,13 +1,12 @@
-#' Create scenarios based on difference between median and max propensities for all cities
+#' Create scenarios for all cities - NOT CURRENTLY USED
 #' 
-#' Creates four scenarios where, in each one, the mode share is elevated to the
-#' certain values in each distance band. The scenario-modes are cycle, car, and bus and motorcycle.
+#' Creates three scenarios where in each one, the mode share of a given mode is elevated by a set
+#' percentage of the total trips. The scenario-modes are cycle, car, and bus and motorcycle.
 #' 
-#' 
+#'  
 #' Add 5% of trips overall in such a way that the average mean mode share for each mode across the
 #' three distance bands is preserved.
 #' 
-#' https://www.dropbox.com/home/ITHIM%20Global/Methods%20and%20Processes/ScenarioDefn/GlobalScenario
 #' 
 #' @param trip_set data frame, baseline scenario
 #' 

### 2023-08-24 R/create_latam_scenarios.R

@@ -1,6 +1,12 @@
-#' Create scenarios defined for latam paper
+#' Create scenarios for all Latin American cities - NOT CURRENTLY USED
 #' 
+#' Creates three scenarios where in each one, the mode share of a given mode is elevated by a set
+#' percentage of the total trips. The scenario-modes are cycle, car, and bus and motorcycle.
 #' 
+#'  
+#' Add 5% of trips overall in such a way that the average mean mode share for each mode across the
+#' three distance bands is preserved.
+#'  
 #' 
 #' @param trip_set data frame, baseline scenario
 #' 

### 2023-08-24 R/create_max_mode_share_scenarios.R

@@ -1,4 +1,4 @@
-#' Create scenarios defined by maximum mode share
+#' Create scenarios defined by maximum mode share - CURRENTLY NOT USED  
 #' 
 #' Creates five scenarios where, in each one, the mode share is elevated to the maximum observed across the cities.
 #' The scenario-modes are pedestrian, cycle, car, motorcycle and bus

### 2023-08-24 R/create_synth_pop.R

@@ -1,6 +1,7 @@
 #' Creates synthetic population
 #' 
-#' Creates a synthetic population by matching individuals in the trip set to individuals in the PA set
+#' Creates a synthetic population by matching individuals in the trip set
+#' to individuals in the physical activity (PA) dataset
 #' 
 #' The function performs the following steps:
 #' 
@@ -13,23 +14,25 @@
 #'     information from the trip data (not including bus driver, truck, car driver and commercial 
 #'     motorcycle trips)
 #'     
-#'   \item to assign non - travel (i.e. work and leisure) physical activity MMET values to this 
+#'   \item to assign non-occupational physical activity MMET values to this 
 #'     synthetic population, the following steps are performed:
+#'     
 #'    \itemize{
-#'     \item for each sex and age category, find the proportion of people with zero work and leisure MMET values
+#'     \item for each sex and age category, find the proportion of people with zero non-occupational MMET values
 #'       and also find the list of people with non-zero MMET values
 #'       
 #'     \item if BACKGROUND_PA_CONFIDENCE < 1 when calling the value of information script, i.e. when 
 #'       input values are sampled from distributions, a beta distribution is built from which the 
-#'       proportion of people with zero work and leisure MMET values is sample using the  
+#'       proportion of people with zero work and leisure MMET values is sampled using the  
 #'       known proportion as mean of this distribution
 #'       
 #'     \item sample with replacement from a vector with 0 MMET values and the vector non-zero MMET values (from
 #'       the people having non-zero work and leisure MMET values) using the proportion of people with
 #'       zero work and leisure MMET values and assign those sampled MMET values to the synthetic population
+#'       
 #'      }
 #'   }    
-#' \item remove any non-travel trips from the trip data
+#' \item remove participants with trip or stage modes that are not in Vehicle inventory
 #'}
 #' 
 #' 
@@ -74,23 +77,32 @@ create_synth_pop <- function(raw_trip_set){
   # match only for "real" people (i.e. not `ghost drivers', whose participant id is 0)
   # extract unique participant ids including age, sex and age category information from the trip data
   synthetic_population <- subset(trip_set,!duplicated(participant_id)&participant_id>0)[,names(trip_set)%in%c("participant_id","age","sex","age_cat")]
[…]

### 2023-08-24 R/get_scenario_settings.R

@@ -1,16 +1,46 @@
-#' Get values for max mode share scenario
+#' Get values for mode share across distance categories and for max mode share scenario - CURRENTLY only used when calling summary_tables.Rmd 
 #' 
-#' Computes the maximum mode share for specified mode types and specified distance categories across specified (stored) cities.
-#' Used for max mode share scenario generation.
+#' Computes the mode shares for the different distance categories and then the maximum mode
+#' share for specified mode types and specified distance 
+#' categories across specified (stored) cities. Used for max mode share scenario generation.
+#' 
+#' The function performs the following steps:
+#' 
+#' \itemize{
+#' \item define the minimum distances in each distance category
+#' 
+#' \item loop through the pre-defined cities:
+#'    \itemize{
+#'    \item read in trip data and get data into correct format
+#'    
+#'    \item assign distance categories
+#'    
+#'    \item for each distance category and pre-defined mode find the proportional modal share 
+#'      (for each distance category, the proportion of the modes adds up to 100%)
+#'    
+#'    \item if rail trips exist, they are added to the proportion of the bus trips to get one value 
+#'      for public transport
+#'    }
+#'    
+#' \item find the maximum mode shares for each mode and distance category
+#'    
+#' \item for each city print one table showing the mode shares for each distance category    
+#'}    
+#'    
+#'    
+#'    
+#'    
 #' 
 #' @param cities which cities to use
 #' @param modes which modes to use
 #' @param distances which distance categories to use
 #' @param speeds named list of mode speeds (to be applied to all cities)
 #' 
-#' @return data frame of proportions by mode and distance category
+#' @return data frame of maximum proportions by mode and distance category
 #' 
 #' @export
+
+
 get_scenario_settings <- function(cities = c('accra', 'bangalore', 'belo_horizonte', 'bogota', 'buenos_aires', 'cape_town',
                                              'delhi', 'mexico_city', 'santiago', 'sao_paulo', 'vizag'),
                                   modes=c("pedestrian","cycle","car","motorcycle","bus"),     
[…]

### 2023-08-24 R/ithim_load_data.R

@@ -220,7 +220,7 @@ ithim_load_data <- function(speeds =
      (!exists('SCENARIO_PROPORTIONS')||
       exists('SCENARIO_PROPORTIONS')&&!isTRUE(base::all.equal(DIST_CAT,colnames(SCENARIO_PROPORTIONS)))
      )){
-    SCENARIO_PROPORTIONS <<- get_scenario_settings(distances=DIST_CAT,speeds=speeds)
+    SCENARIO_PROPORTIONS <<- get_scenario_settings( distances=DIST_CAT,speeds=speeds)
   }
   
   

### 2023-08-25 summary_tables.Rmd

@@
 
-# Script summarises mode splits, distances and durations travelled and looks at health outcomes (YLLs and deaths) for the different air pollution, injury and physical activity pathways but also by disease
+# Script summarises mode splits, distances and durations travelled and looks at health outcomes (YLLs and deaths) for the different air pollution (AP), injury and physical activity (PA) pathways but also by disease
 
@@
 ## Health outcomes - injury, air pollution and physical activity
-# - one table for each city showing the change in total YLLs in the scenarios compared to the Baseline for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age and sex category
-
-# - one table for each city showing the change in total YLLs due to changes in accident fatalities in the scenarios compared to the Baseline for each age group
-
-# - one table for each city showing the change in total YLLs due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group
-
-# - one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalities in the scenarios compared to the Baseline for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group
+# - one table for each city showing the change in total YLLs for each of the scenarios compared to the reference scenario for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each age and sex category
+
+# - one table for each city showing the change in total YLLs due to changes in accident fatalities for each of the scenarios compared to the reference scenario for each age group
+
+# - one table for each city showing the change in total YLLs due to changes in physical activity levels for each of the scenarios compared to the reference scenario for each age group
+
+# - one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels for each of the scenarios compared to the reference scenario for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalitiesfor each of the scenarios compared to the reference scenario for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levelsfor each of the scenarios compared to the reference scenario for each age group
+
+# - one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels for each of the scenarios compared to the reference scenario for each age group
 
 ## Health outcomes - disease
-# - one table for each city showing the change in deaths in the scenarios compared to the Baseline for each disease
-
-# - one table for each city showing the change in deaths per 100,000 people in the scenarios compared to the Baseline for each disease
-
-# - one table for each city showing the change in YLLs in the scenarios compared to the Baseline for each disease
-
-# - one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each disease
+# - one table for each city showing the change in deaths for each of the scenarios compared to the reference scenario for each disease
+
+# - one table for each city showing the change in deaths per 100,000 people for each of the scenarios compared to the reference scenario for each disease
+
+# - one table for each city showing the change in YLLs for each of the scenarios compared to the reference scenario for each disease
+
+# - one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each disease
 
@@
 
-# - the ylls.csv file (results/multi_city/health_impacts/ylls.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
-It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age and age and sex specific population levels.
-The results are for either AP and PA only if the diseases are affected by only AP or PA or the results for the combined AP and PA effects if the diseases are affected by both AP and PA.
-
-# - the ylls_pathway.csv file (results/multi_city/health_impacts/ylls_pathway.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
-It also shows the disease level, whether the disease is exasperated by changes in AP or PA, and the age and age and sex specific population levels.
+# - the ylls.csv file (results/multi_city/health_impacts/ylls.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario compared to the reference scenario.
+It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age, and age and sex specific population levels.
+The results are for either AP and PA only if the diseases are affected by only AP or PA or the results are for the combined AP and PA effects if the diseases are affected by both AP and PA.
+
+# - the ylls_pathway.csv file (results/multi_city/health_impacts/ylls_pathway.csv) gives the changes in YLLs for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario compared to the reference scenario.
+It also shows the disease level, whether the disease is exasperated by changes in AP or PA, and the age, and age and sex specific population levels.
 The results are given for AP and PA independently even if a disease is affected by both pathways.
 
-# - the deaths.csv file (results/multi_city/health_impacts/deaths.csv) gives the changes in deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
-It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age and age and sex specific population levels.
-The results are for either AP and PA only if the diseases are affected by only AP or PA or the results for the combined AP and PA effects if the diseases are affected by both AP and PA.
-
-# - the deaths_pathway.csv file (results/multi_city/health_impacts/deaths_pathway.csv) gives the changes in deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
-It also shows the disease level, whether the disease is exasperated by changes in AP or PA, and the age and age and sex specific population levels.
+# - the deaths.csv file (results/multi_city/health_impacts/deaths.csv) gives the changes in deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario compared to the reference scenario.
+It also shows the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age, and age and sex specific population levels.
+The results are for either AP and PA only if the diseases are affected by only AP or PA or the results are for the combined AP and PA effects if the diseases are affected by both AP and PA.
+
+# - the deaths_pathway.csv file (results/multi_city/health_impacts/deaths_pathway.csv) gives the changes in deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario compared to the reference scenario.
+It also shows the disease level, whether the disease is exasperated by changes in AP or PA, and the age, and age and sex specific population levels.
 The results are given for AP and PA independently even if a disease is affected by both pathways.
@@
 
-# - set up helper function to re-format large numbers, re-name scenarios and format and print output tables
+# - set up helper functions to re-format large numbers, re-name scenarios and format and print output tables
 
@@
 
-# - create one table for each city showing the change in total YLLs in the scenarios compared to the Baseline for each age group
-
-# - create one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age group
-
-# - create one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each age and sex category
-
-# - create one table for each city showing the change in total YLLs due to changes in accident fatalities in the scenarios compared to the Baseline for each age group
-
-# - create one table for each city showing the change in total YLLs due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group
-
-# - create one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group
-
-# - create one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalities in the scenarios compared to the Baseline for each age group
-
-# - create one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levels in in the scenarios compared to the Baseline for each age group
-
-# - create one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels in in the scenarios compared to the Baseline for each age group
-
-# - create one table for each city showing the change in deaths in the scenarios compared to the Baseline for each disease
-
-# - create one table for each city showing the change in deaths per 100,000 people in the scenarios compared to the Baseline for each disease
-
-# - create one table for each city showing the change in YLLs in the scenarios compared to the Baseline for each disease
-
-# - create one table for each city showing the change in YLLs per 100,000 people in the scenarios compared to the Baseline for each disease
-
-# - create 4 individual .csv files in (results/multi_city/health_impacts/) for both death and YLLs considering the effect of AP and PA separately (*_pathway.csv) or together (*.csv).
-These files give the changes in YLLs / deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario.
-They also show the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age and age and sex specific population levels.
+# - create one table for each city showing the change in total YLLsfor each of the scenarios compared to the reference scenario  for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each age and sex category
+
+# - create one table for each city showing the change in total YLLs due to changes in accident fatalities for each of the scenarios compared to the reference scenario for each age group
+
+# - create one table for each city showing the change in total YLLs due to changes in physical activity levels for each of the scenarios compared to the reference scenario for each age group
+
+# - create one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels for each of the scenarios compared to the reference scenario for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalities for each of the scenarios compared to the reference scenario for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levels for each of the scenarios compared to the reference scenario for each age group
+
+# - create one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levelsfor each of the scenarios compared to the reference scenario for each age group
+
+# - create one table for each city showing the change in deaths for each of the scenarios compared to the reference scenario for each disease
+
+# - create one table for each city showing the change in deaths per 100,000 people for each of the scenarios compared to the reference scenario for each disease
+
+# - create one table for each city showing the change in YLLs for each of the scenarios compared to the reference scenario for each disease
+
+# - create one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each disease
+
+# - create 4 individual .csv files in 'results/multi_city/health_impacts/' for both death and YLLs considering the effect of AP and PA separately (*_pathway.csv) or together (*.csv).
+These files give the changes in YLLs or deaths for each age and sex category, for each age category per 100,000 people, and for each age and sex category per 100,000 people for each city, disease and scenario compared to the reference scenario.
+They also give the disease level, whether the disease is exasperated by changes in AP, PA or both, and the age and age and sex specific population levels.
 
@@
 ```{r helper_functions, echo = F, message = F}
+# set up helper functions
+
 # Function to format numbers with million (M) and billion (B) suffixes
@@
 cities <<- names(io)[!names(io) %in% 'scen_prop']
+
+# extract number of scenarios
 NSCEN <- nrow(io$scen_prop)
@@
 ```
-# **1st option**
-## Scenario definition 
+<!-- # **1st option** -->
+
+## Distance categories
+
+<!-- ## Scenario definition  -->
 <!-- Table displays highest propensity for each distance category, for a given mode. -->
@@
 
+# extract and update scenario names
 scen_prop <- round(io$scen_prop, round_to)
@@
 
+# rename scenarios
 orig_scen_names_updated <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))
@@
 ## Mode split for each scenario by city
+
 City specific trip proportions by mode, for baseline and scenarios
@@
 ## Distance tables
-City specific distance tables showing average distance in km travelled by each person by mode for all scenarios (incl Baseline)
+City specific distance tables showing average distance in km travelled by each person in the synthetic population by mode for all scenarios (incl Baseline)
 
 ```{r trip_dist = "asis"}
-# find the average distance a person spends travelling on each mode per day
+# find the average distance a person in the synthetic population spends travelling on each mode per day
 
@@
 
-  # update scenario names - cycle, car, bus
+  # update scenario names 
   colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
@@
   # write to .html
-  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person")))
+  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person in the synthetic population")))
   cat("\n")
@@
 
-  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person (people with trips ONLY)")))
-
-}
-
-```
-
-### Distance table showing average baseline distances in km travelled by each person by mode for all cities
+  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person in the synthetic population (people with trips ONLY)")))
+
+}
+
+```
+
+### Distance table showing average baseline distances in km travelled by each person in the synthetic population by mode for all cities
 
@@
 
-kable(trip_dist, caption = "Avg. baseline distance per person (km) per day across all cities")
+kable(trip_dist, caption = "Avg. baseline distance (km) per person in the synthetic population per day across all cities")
 
@@
 
-City specific distance tables for baseline and scenarios for the entire population with age range considered in the model
+City specific distance tables for baseline and scenarios for the entire population with ages considered in the model
 
@@
 for (city in cities) { # loop through cities
-
-  # find the total city level population considered in the model only
-  #count_people <- round(sum(io[[city]]$demographic$population))
 
@@
 
-# create baseline yearly distance table for all cities
+# create baseline yearly distance table for the entire population with ages considered in the model for all cities
 
@@
 
-  print(kable(df, caption = paste("Model age population distance by mode per day for  ", city)))
+  print(kable(df, caption = paste("Total population distance with ages considered in model by mode per day for  ", city)))
 }
@@
 
-City specific duration tables for baseline and scenarios (per day in minutes)
+City specific duration tables for the synthetic population for baseline and scenarios (per day in minutes)
 
 ```{r trip_dur = "asis"}
-# create tables showing the average duration each person spends travelling per mode
+
+# create tables showing the average duration each person in the synthetic population spends travelling per mode
 
@@
   # print to html
-  print(kable(td, caption = paste("Duration table (mins) per day for model synthetic population for ", city, "( ", count_people, " ) per person (everyone)")))
+  print(kable(td, caption = paste("Duration table (mins) per day for synthetic population for ", city, "( ", count_people, " ) per person (everyone)")))
 
@@
 
-# find injury death column
-#injury_col <- which(colnames(io[[1]]$outcomes$hb$deaths) == paste0(scen_names[2],'_deaths_inj'))
-
-# find ap death columns
-# ap_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))
-
 # find ap ylls columns
 ap_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))
-
-# find pa death columns
-# pa_death_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))
 
@@
 Change in YLLs per 100,000 people by age and sex category for each scenario compared to the reference scenario
+
 ```{r scen_prop = "asis"}
-cat(' ')
+
 cat(paste0('Reference scenario:  ', reference_scenario))
@@
 
-YLLs per 100,000 people by age group by scenario - injury
+<!-- YLLs per 100,000 people by age group by scenario - injury -->
 
@@
 
-Change in total deaths compared to the reference scenario (for the city based on real population size) by scenario
+Change in total deaths compared to the reference scenario (for the city based on real population size but only for ages considered in the model) by scenario
 
@@
 
-Change in deaths per 100,000 people by scenario in comparison to the reference scenario
+Change in deaths per 100,000 people (based on real population size but only for ages considered in the model) by scenario in comparison to the reference scenario
 
@@
 
-Change in total YLLs by disease compared to the reference scenario (for the city based on real population size, for ages considered within the ITHIM model)
+Change in total YLLs by disease compared to the reference scenario (based on real population size, for ages considered within the ITHIM model)
 
@@
 ### Change in YLL due to disease compared to the reference scenario per 100,000 people {#change_yll_disease_100k}
+
+Change in total YLLs by disease compared to the reference scenario per 100,000 people (based on real population size, for ages considered within the ITHIM model)
 

### 2023-08-25 summary_tables_PA_AP.Rmd

@@
 
-# Script to summarise the PM2.5 exposure and levels, PM2.5 and CO2 concentrations and a physical activity levels
+# Script to summarise the PM2.5 personal exposure levels, PM2.5 and CO2 concentrations and physical activity levels
 
@@
 
-# - one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario
+# - one table for each city detailing the individual PM2.5 exposure levels including the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario
 
@@
 
-# - one table detailing the CO2 concentration levels for each scenario for all cities
+# - one table detailing the Baseline CO2 concentration levels for each scenario for all cities
 
@@
 
-# - one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario.
+# - one table for each city detailing the individual PM2.5 exposure levels including the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario.
 Table also contains CO2 concentrations for each scenario
@@
 
-# - desc_stats.csv file (results/multi_city/AP/desc_stats.csv) containing the summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the baseline scenario and the total transport related CO2 emission concentrations for each scenario and city
+# - desc_stats.csv file (results/multi_city/AP/desc_stats.csv) containing the summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the reference scenario and the total transport related CO2 emission concentrations for each scenario and city
 
@@
 
-#   - summary_stats_PM2.5_CO2: summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the baseline scenario and the total transport related CO2 emission concentrations for each scenario and city
+#   - summary_stats_PM2.5_CO2: summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the reference scenario and the total transport related CO2 emission concentrations for each scenario and city
 
-#### The script performs the following steps, assuming that ITHIM-GLOBAL module has been run and the ithim_objects object has been saved in "results/multi_city/io.rds":
+#### The script performs the following steps, assuming that the ITHIM-Global model has been run in CONSTANT mode and the ithim_objects object has been saved in "results/multi_city/io.rds":
 
@@
 # - create one table for each city detailing the individual PM2.5 exposure levels incl.
-the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario
+the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario
 
@@
 
-# - create one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario.
+# - create one table for each city detailing the individual PM2.5 exposure levels including the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario.
 Table also contains CO2 concentrations for each scenario
@@
 
-# scenario definition
+# scenario and reference scenario definitions
 scenario_name <- io[[city]]$ithim_run$scenarios_used
@@
 
-compute_mode <- io[[city]]$ithim_run$compute_mode # constant parameters from the given parameters
+compute_mode <- io[[city]]$ithim_run$compute_mode
 
@@
 
+# extract local parameter information
 for(i in 1:length(parameter_names)){
@@
 
+# extract global parameter information
 for(i in 1:length(global_parameter_names)){
@@
 
+# get parameters into correct format
 dist_cat <- unlist(strsplit(gsub(" ", "", dist_cat, fixed = TRUE), "\\,"))
@@
 # for each city create a summary table of individual PM2.5 exposure levels giving the mean and the 5th, 50th and 95th percentile levels.
-Tables also show the PM2.5 concentration levels in the city for each scenario and the changes compared to the baseline scenario
+Tables also show the PM2.5 concentration levels in the city for each scenario and the changes compared to the reference scenario
 
@@
   n_col <- ncol(io[[cities[x]]]$outcomes$pm_conc_pp)
+
   # rename columns with updated scenario names
@@
 
-  # add scenario specific PM2.5 concentration levels and the change in PM2.5 compared to the baseline
+  # add scenario specific PM2.5 concentration levels 
   summary$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, digits=2)
@@
 # Create a single table detailing the Baseline PM2.5 emissions for each mode, the city wide PM2.5 concentration levels and the percentage of these concentrations attributed to transport for all cities
-
-#city <- cities
 
@@
 
-  # extract Co2 concentrations
+  # extract CO2 concentrations
   td <- round(colSums(io[[city]]$outcomes$co2_conc, na.rm = T), 1) %>% 
@@
     labs(y = "Weekly mMET", x = "Scenarios")+  ylim(0, limit)
+
+  # print to html
   print(y)
@@
 
-# - one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the baseline scenario.
+# - one table for each city detailing the individual PM2.5 exposure levels incl the mean, 5th, 50th and 95th percentiles, the mean PM2.5 concentration levels for each scenario and the change in PM2.5 compared to the reference scenario.
 Table also contains CO2 concentrations for each scenario
@@
 
-  # add to io objects
+  # add to io object
   io[[cities[x]]]$summary_emission <- summary
 
-  ## PM 2.5 table and plot
+  ## PM2.5 table and plot
 

### 2023-08-26 injury_tables.Rmd
added output version control

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
+params:
+  output_version: ''
 output:
@@
 ```{r load_objects = "asis"}
+output_version <- "v0.3"
+
 # Assumes that multi_city_script.R has been run - read in ithim_objects saved as io.rds
-io <- readRDS("results/multi_city/io.rds")
+io <- readRDS(paste0("results/multi_city/io_", output_version, ".rds"))
 
 # Get names of cities from the io object
-cities <- names(io)[!names(io) %in% 'scen_prop']
+cities <- names(io)[!names(io) %in% c('scen_prop','ithim_run' )]
 
@@
 scen_names <- names(io[[1]]$outcomes$whw)
+
+# input parameter file name
+input_parameter_file <<- io$ithim_run$input_parameter_file
+
+# further model run information
+compute_mode <- io$ithim_run$compute_mode 
+timestamp_model <- io$ithim_run$timestamp
+comments_model <- io$ithim_run$comment
 
@@
 
+# print model run information to screen:
+cat("  \n")
+cat(paste0('Output version: ', output_version)) 
+cat("  \n")
+cat(paste0('Timestamp of model run: ', timestamp_model))
+cat("  \n")
+cat(paste0('Comments from model run: ', comments_model))
+cat("  \n")
+
 ```
@@
 # write to csv
-readr::write_csv(inj_counts_list, paste0('results/multi_city/inj/inj_counts.csv')) # table containing all the fatality counts
+readr::write_csv(inj_counts_list, paste0('results/multi_city/inj/inj_counts_',output_version,'.csv')) # table containing all the fatality counts
 ```
@@
 # output csv
-readr::write_csv(rd, 'results/multi_city/inj/injury_risks_per_billion_kms.csv')
+readr::write_csv(rd,paste0( 'results/multi_city/inj/injury_risks_per_billion_kms_',output_version,'.csv'))
 ```
@@
 # write to csv
-readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), 'results/multi_city/inj/injury_risks_per_100k_pop.csv')
+readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), paste0('results/multi_city/inj/injury_risks_per_100k_pop_',output_version,'.csv'))
 ```
@@
 # write to csv
-readr::write_csv(true_distances_df, 'results/multi_city/inj/distances.csv')
+readr::write_csv(true_distances_df, paste0('results/multi_city/inj/distances_',output_version,'.csv'))
 ```
@@
 # write as csv
-readr::write_csv(rd_hours, 'results/multi_city/inj/injury_risks_per_100million_h.csv')
+readr::write_csv(rd_hours, paste0('results/multi_city/inj/injury_risks_per_100million_h_',output_version,'.csv'))
 ```
@@
 # Export injury datasets as a single excel file
-path_inj_wb <- "results/multi_city/inj/inj_data.xlsx"
+path_inj_wb <- paste0('results/multi_city/inj/inj_data_',output_version,'.xlsx')
  writexl::write_xlsx(list(inj_counts = inj_counts_list, #%>% filter(str_mode != 'Total'),

### 2023-08-26 summary_tables.Rmd
added output version control

@@
 
-### - output .csv/.xlsx files containing the following information:
+### - output .csv files containing the following information:
 
@@
 ```{r load_objects = "asis", echo = F, message = F}
-
+#output_version <- "v0.3"
 # read in io object which is produced at the end of an ITHIM-Global run
 # Assumes that multi_city_script.R has been run 
-io <- readRDS("results/multi_city/io.rds")
+# read in input file
+io <- readRDS(paste0("results/multi_city/io_", output_version, ".rds"))
 
 # Get names of cities from the io object
-cities <<- names(io)[!names(io) %in% 'scen_prop']
+cities <- names(io)[!names(io) %in% c('scen_prop','ithim_run' )]
 
@@
 
-# find reference scenario
-ref_scen <- io[[cities[1]]]$ithim_run$reference_scenario
-reference_scenario <- get_qualified_scen_name(ref_scen) # update scen name
+# input parameter file name
+input_parameter_file <<- io$ithim_run$input_parameter_file
+
+# scenario and reference scenario definitions
+scenario_name <- io$ithim_run$scenarios_used
+ref_scen <- io$ithim_run$reference_scenario
+reference_scenario <- get_qualified_scen_name(ref_scen)
+
+# further model run information
+compute_mode <- io$ithim_run$compute_mode 
+timestamp_model <- io$ithim_run$timestamp
+comments_model <- io$ithim_run$comment
 
@@
 
-## Distance categories
+# Distance categories
 
@@
 # i.e. sum over all disease and remove sex 
-yll_totals <- lapply(io[-2],function(x){
+yll_totals <- lapply(io[-c(2,3)],function(x){
   temp <- sapply(1:NSCEN,function(y){# loop through scenarios
@@
 # extract population statistics
-pop_by_age <- lapply(io[-2], function(x)
+pop_by_age <- lapply(io[-c(2,3)], function(x)
   sapply(unique(x$demographic$age), 
@@
 # calculate the total ylls for each scenario summing across all diseases
-yll_totals <- lapply(io[-2], function(x) {
+yll_totals <- lapply(io[-c(2,3)], function(x) {
   temp <- sapply(1:NSCEN,function(y){
@@
 # calculate change in YLLs due to injury by summing across all age categories for the injury results
-injury_totals <- lapply(io[-2], function(x) {
+injury_totals <- lapply(io[-c(2,3)], function(x) {
   xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
@@
 # sum the YLLs for each scenario by age category
-pa_totals <- lapply(io[-2],function(x){
+pa_totals <- lapply(io[-c(2,3)],function(x){
   xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
@@
 # sum the YLLs by scenario and age category
-ap_totals <- lapply(io[-2],function(x){
+ap_totals <- lapply(io[-c(2,3)],function(x){
   xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
@@
 # sum the total deaths by scenario
-disease_totals <- lapply(io[-2],function(x){
+disease_totals <- lapply(io[-c(2,3)],function(x){
   # extract disease counts
@@
 # sum the total ylls by scenario
-disease_totals <- lapply(io[-2],function(x){
+disease_totals <- lapply(io[-c(2,3)],function(x){
   xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
@@
   for (j in pathway) { # loop for hb or pathway_hb
-    for (i in names(io[-2])) { # loop for cities
+    for (i in names(io[-c(2,3)])) { # loop for cities
 
@@
 write.csv(ylls,
-            "results/multi_city/health_impacts/ylls.csv",
+            paste0('results/multi_city/health_impacts/ylls_',output_version,'.csv'),
             row.names = F)
@@
 write.csv(ylls_pathway,
-            "results/multi_city/health_impacts/ylls_pathway.csv",
+            paste0('results/multi_city/health_impacts/ylls_pathway_',output_version,'.csv'),
             row.names = F)
@@
 write.csv(deaths,
-            "results/multi_city/health_impacts/deaths.csv",
+            paste0('results/multi_city/health_impacts/deaths_',output_version,'.csv'),
             row.names = F)
@@
 write.csv(deaths_pathway,
-            "results/multi_city/health_impacts/deaths_pathway.csv",
+            paste0('results/multi_city/health_impacts/deaths_pathway_',output_version,'.csv'),
             row.names = F)

### 2023-08-26 summary_tables_PA_AP.Rmd
added output version control

@@
 
-# Introduction
-These are the summary tables of the following items
-
-1) Individual PM2.5 exposure levels and total PM2.5 concentrations for baseline and scenarios
-2) Overall CO2 emissions for baseline and scenarios
-3) Individual physical activity levels (mMETs) for baseline and scenarios
-
 ```{r, message=FALSE, warning=FALSE}
@@
 ###### Read in input parameters
+output_version <- output_version
 
@@
 # read in input file
-io <- readRDS("results/multi_city/io.rds")
+io <- readRDS(paste0("results/multi_city/io_", output_version, ".rds"))
 
 # Get names of cities from the io object
-cities <- names(io)[!names(io) %in% 'scen_prop']
+cities <- names(io)[!names(io) %in% c('scen_prop','ithim_run' )]
 
@@
 city <- cities[1]
-scen_length <- length(io[[city]]$ithim_run$scenario_names) - 1
+scen_length <- length(io$ithim_run$scenario_names) - 1
 
 # input parameter file name
-input_parameter_file <<- io[[city]]$ithim_run$input_parameter_file
+input_parameter_file <<- io$ithim_run$input_parameter_file
 
 # scenario and reference scenario definitions
-scenario_name <- io[[city]]$ithim_run$scenarios_used
-ref_scen <- io[[cities[1]]]$ithim_run$reference_scenario
+scenario_name <- io$ithim_run$scenarios_used
+ref_scen <- io$ithim_run$reference_scenario
 reference_scenario <- get_qualified_scen_name(ref_scen)
 
-compute_mode <- io[[city]]$ithim_run$compute_mode
+# further model run information
+compute_mode <- io$ithim_run$compute_mode 
+timestamp_model <- io$ithim_run$timestamp
+comments_model <- io$ithim_run$comment
 
@@
 
+# print model run information to screen:
+cat("  \n")
+cat(paste0('Output version: ', output_version)) 
+cat("  \n")
+cat(paste0('Timestamp of model run: ', timestamp_model))
+cat("  \n")
+cat(paste0('Comments from model run: ', comments_model))
+cat("  \n")
+
 ```
+
+# Introduction
+These are the summary tables of the following items
+
+1) Individual PM2.5 exposure levels and total PM2.5 concentrations for baseline and scenarios
+2) Overall CO2 emissions for baseline and scenarios
+3) Individual physical activity levels (mMETs) for baseline and scenarios
 
@@
 # create output Excel file containing the basic PM2.5 and CO2 exposure / concentration information
-output_file <- "results/multi_city/AP/desc_stats.xlsx"
+output_file <- paste0('results/multi_city/AP/desc_stats_', output_version ,".xlsx")
 writexl::write_xlsx(list(PM_2.5_emission_inventory = sl,
@@
 # create csv file (summary_stats_Pm2.5_CO2 tab in Excel file above)
-write_csv(bind_rows(l), "results/multi_city/AP/desc_stats.csv")
+write_csv(bind_rows(l), paste0('results/multi_city/AP/desc_stats_', output_version ,".csv"))
 

### 2023-08-26 Call_Rmd_outputs.R

@@ -1,0 +1,30 @@
+# Script to knit the various .Rmd files and re-name the corresponding output .html files
+# to contain the output version of the corresponding ithim run
+
+# if output_version is not already defined then you need to define it before calling the various .Rmd files
+
+
+
+# output_version <- 
+
+# create summary tables for AP and PA
+rmarkdown::render('summary_tables_PA_AP.Rmd', params = list(output_version = output_version))
+
+# rename output .html files
+file.rename('summary_tables_PA_AP.html', paste0('summary_tables_PA_AP_',output_version,'.html'))
+
+
+
+# create summary tables
+rmarkdown::render('summary_tables.Rmd', params = list(output_version = output_version))
+
+# rename output .html files
+file.rename('summary_tables.html', paste0('summary_tables_',output_version,'.html'))
+
+
+
+# create injury summary tables
+rmarkdown::render('injury_tables.Rmd', params = list(output_version = output_version))
+
+# rename output .html files
+file.rename('injury_tables.html', paste0('injury_tables_',output_version,'.html'))

### 2023-08-26 multi_city_script.R

@@ -76,7 +76,7 @@ if (!require("drpa",character.only = TRUE)) {
 # #African & Indian cities
 #cities <- c('accra','cape_town','kisumu', 'nairobi', 'port_louis', 'bangalore', 'delhi', 'vizag')
 
-cities <- 'bogota'
+cities <- c('bogota')
 
 
 input_parameter_file <- "InputParameters_v32.0.xlsx" # file containing the local and global input parameters
@@ -242,6 +242,7 @@ ithim_objects <- outcome <- outcome_pp <- yll_per_hundred_thousand <- list()
 
 
 print(system.time(for(city in cities){
+  cat("\n")
   print(city)
   # run code to prepare the input data for the actual ITHIM Global health impact assessment
   ithim_objects[[city]] <- run_ithim_setup(
@@ -321,15 +322,6 @@ print(system.time(for(city in cities){
   ithim_objects[[city]]$new_walk_trips_count$bus <- count_new_walk_trips_bus
   ithim_objects[[city]]$new_walk_trips_count$rail <- count_new_walk_trips_rail
   
-  # add run relevant information to ithim_objects list
-  ithim_objects[[city]]$ithim_run <- list()
-  ithim_objects[[city]]$ithim_run$input_parameter_file <- input_parameter_file
-  ithim_objects[[city]]$ithim_run$scenarios_used <- scenario_name
-  ithim_objects[[city]]$ithim_run$reference_scenario <- reference_scenario
-  ithim_objects[[city]]$ithim_run$scenario_increase <- scenario_increase
-  ithim_objects[[city]]$ithim_run$scenario_names <- SCEN
-  ithim_objects[[city]]$ithim_run$compute_mode <- compute_mode
-  
   # store results to plot
   min_ages <- sapply(ithim_objects[[city]]$outcome$hb$ylls$age_cat,function(x)as.numeric(strsplit(x,'-')[[1]][1]))
   max_ages <- sapply(ithim_objects[[city]]$outcome$hb$ylls$age_cat,function(x)as.numeric(strsplit(x,'-')[[1]][2]))
@@ -359,6 +351,21 @@ print(system.time(for(city in cities){
     disease_list[[i]][,which(cities == city)] <- result_mat_plot[1:NSCEN + (i - 1) * NSCEN]/sum(subset(DEMOGRAPHIC,min_pop_ages >= min_age & max_pop_ages <= max_age)$population)
 }))
 
+
+# add run relevant information to ithim_objects list
+timestamp <- Sys.time()
[…]

### 2023-08-27 multi_city_script.R
First version of results

@@ -88,9 +88,9 @@ input_parameter_file <- "InputParameters_v35.0.xlsx" # file containing the local
 # the number of samples (which is 1 in constant mode), the path to any other input files,
 # any comments and the runtime of the code
 write_output_control = T # whether you want to save the model run specifics or not
-output_version <- "v49f1a71_test_run" # gives the version number of the output documents, independent of the input parameter file name
-author <- "AKS"
-comment <- "Test whether version control working"
+output_version <- "c28f173d_test_run" # gives the version number of the output documents, independent of the input parameter file name
+author <- "DGS"
+comment <- "First test under bogota branch"
 
 
 # scenario definition

### 2023-09-15 injury_tables.Rmd
Added output version numbering - works on Latam branch
First version of results
Added csv outputs without version number

@@
 # read in table detailing the standardized modes
-smodes <- read_csv('data/global/modes/standardized_modes.csv')
+# smodes <- read_csv('data/global/modes/standardized_modes.csv')
+smodes <- read_csv('results/multi_city/standardized_modes.csv')
 
@@
 ```{r load_objects = "asis"}
-output_version <- "v0.3"
+#output_version <- "v0.3"
 
@@
 
-# write to csv
+# write to csv with and without output version number
 readr::write_csv(inj_counts_list, paste0('results/multi_city/inj/inj_counts_',output_version,'.csv')) # table containing all the fatality counts
+readr::write_csv(inj_counts_list, paste0('results/multi_city/inj/inj_counts.csv'))
 ```
@@
 
-# output csv
+# output csv with and without output version number
 readr::write_csv(rd,paste0( 'results/multi_city/inj/injury_risks_per_billion_kms_',output_version,'.csv'))
+readr::write_csv(rd,paste0( 'results/multi_city/inj/injury_risks_per_billion_kms.csv'))
 ```
@@
 active_travel <- baseline_counts_pop %>% filter(mode == 'cycle' | mode == 'pedestrian')
-active_travel <- active_travel %>% group_by(city, scenario, country, continent, measure ) %>% summarise(
-                                   value = sum(value), model_population = sum(model_population), 
+active_travel <- active_travel %>% group_by(city, scenario, country, continent, measure, model_population 
+                                            ) %>% summarise(
+                                   value = sum(value), 
                                    mode ='active_travel')
@@
 
-# write to csv
+# write to csv with and without output version number
 readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), paste0('results/multi_city/inj/injury_risks_per_100k_pop_',output_version,'.csv'))
+
+readr::write_csv(injury_risks_per_100k %>% filter(mode != 'Total' & scenario != 'Total'), paste0('results/multi_city/inj/injury_risks_per_100k_pop.csv'))
 ```
@@
 
-# write to csv
+# write to csv with and without output version number
 readr::write_csv(true_distances_df, paste0('results/multi_city/inj/distances_',output_version,'.csv'))
+readr::write_csv(true_distances_df, paste0('results/multi_city/inj/distances.csv'))
 ```
@@
 
-# write as csv
+# write as csv with and without output version number
 readr::write_csv(rd_hours, paste0('results/multi_city/inj/injury_risks_per_100million_h_',output_version,'.csv'))
+readr::write_csv(rd_hours, paste0('results/multi_city/inj/injury_risks_per_100million_h.csv'))
 ```

### 2023-09-15 summary_tables.Rmd
First version of results
Added csv outputs without version number

@@
 # Read trip_order
-trip_order <- read_csv("data/global/trips/mode_order.csv")
+# trip_order <- read_csv("data/global/trips/mode_order.csv")
+trip_order <- read_csv("results/multi_city/mode_order.csv")
 
@@
 # Export ylls HB
+
+# write csv with output version number
 write.csv(ylls,
@@
 
-```
+# write csv without output version number
+write.csv(ylls,
+            paste0('results/multi_city/health_impacts/ylls.csv'),
+            row.names = F)
+# Export ylls PATHWAY_HB
+write.csv(ylls_pathway,
+            paste0('results/multi_city/health_impacts/ylls_pathway.csv'),
+            row.names = F)
+# Export deaths HB
+write.csv(deaths,
+            paste0('results/multi_city/health_impacts/deaths.csv'),
+            row.names = F)
+# Export deaths PATHWAY_HB
+write.csv(deaths_pathway,
+            paste0('results/multi_city/health_impacts/deaths_pathway.csv'),
+            row.names = F)
+
+```

### 2023-09-15 summary_tables_PA_AP.Rmd
Added output version numbering - works on Latam branch
Added csv outputs without version number

@@
 ###### Read in input parameters
-output_version <- output_version
+#output_version <-
 
@@
 
-# create csv file (summary_stats_Pm2.5_CO2 tab in Excel file above)
+# create csv file (summary_stats_Pm2.5_CO2 tab in Excel file above) with and without output version number
 write_csv(bind_rows(l), paste0('results/multi_city/AP/desc_stats_', output_version ,".csv"))
+write_csv(bind_rows(l), paste0('results/multi_city/AP/desc_stats.csv'))
 

### 2023-09-15 Create_output_summaries.R

@@ -28,3 +28,4 @@ rmarkdown::render('injury_tables.Rmd', params = list(output_version = output_ver
 
 # rename output .html files
 file.rename('injury_tables.html', paste0('injury_tables_',output_version,'.html'))
+

### 2023-10-04 summary_tables_PA_AP.Rmd
Index to columns in summary_tables_PA_AP. There were some errors that are now fixed

@@
   # add scenario specific PM2.5 concentration levels 
+  ## Dan - TODO: This can be problematic because we are assuming that the order from "pm_conc_pp"
+  ## is the same as in "scenario_pm". And maybe it is, but don't know if this will
+  ## apply always.
+  ## To fix it we have to export "scenario_pm" with names and here compare the order of
+  ## both objects.
   summary$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, digits=2)
@@
 
-co2_conc$Scenario <- new_scen_names
+# Dan: here the order is different (bus before car)
+# co2_conc$Scenario <- new_scen_names
+co2_conc$Scenario <- sapply(co2_conc$Scenario, FUN = get_qualified_scen_name)
 
@@
   n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # define columns of interests
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names
+  # names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names
+  change_names <- which(grepl("_mmet", names(io[[cities[x]]]$outcomes$mmets), fixed = TRUE))
+  names(io[[cities[x]]]$outcomes$mmets)[change_names] <- sapply(gsub("_mmet" , "", names(io[[cities[x]]]$outcomes$mmets)[change_names]), FUN = get_qualified_scen_name)
 
   # extract data of interest
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, 
+                      Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
 
@@
   n_col <- ncol(io[[cities[x]]]$outcomes$mmets) # find the number of columns
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names # assign new column names
+  # names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names # assign new column names
+  change_names <- which(grepl("_mmet", names(io[[cities[x]]]$outcomes$mmets), fixed = TRUE))
+  names(io[[cities[x]]]$outcomes$mmets)[change_names] <- sapply(gsub("_mmet" , "", names(io[[cities[x]]]$outcomes$mmets)[change_names]), FUN = get_qualified_scen_name)
 
   # create dataframe with individual mMET values
-  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
+  data_long <- gather(io[[cities[x]]]$outcomes$mmets, scenario, mmet, 
+                      Baseline:new_scen_names[length(new_scen_names)], factor_key = TRUE)
 
@@
   # extract the mean mMET values and create dataframe
-  td <- round(colMeans(io[[city]]$outcomes$mmets[,(n_col - scen_length):n_col], na.rm = T), 1) %>%
+  # Dan: we change mmets column names before
+  # td <- round(colMeans(io[[city]]$outcomes$mmets[,(n_col - scen_length):n_col], na.rm = T), 1) %>%
+  #   as.data.frame() %>% tibble::rownames_to_column()
+  td <- round(colMeans(io[[city]]$outcomes$mmets[,new_scen_names], na.rm = T), 1) %>%
     as.data.frame() %>% tibble::rownames_to_column()
@@
 
-cl$Scenario <- new_scen_names
+# We dont need this anymore
+# cl$Scenario <- new_scen_names
 
@@
   # assign new scenario names to columns
-  names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- new_scen_names
+  # names(io[[cities[x]]]$outcomes$pm_conc_pp)[(n_col - scen_length):n_col] <- new_scen_names
+  change_names <- which(grepl("pm_conc_", names(io[[cities[x]]]$outcomes$pm_conc_pp), fixed = TRUE))
+  names(io[[cities[x]]]$outcomes$pm_conc_pp)[change_names] <- sapply(gsub("pm_conc_" , "", names(io[[cities[x]]]$outcomes$pm_conc_pp)[change_names]), FUN = get_qualified_scen_name)
 
@@
   # add total PM concentration levels for each scenario
+  ## Dan - TODO: This can be problematic because we are assuming that the order from "pm_conc_pp"
+  ## is the same as in "scenario_pm". And maybe it is, but don't know if this will
+  ## apply always.
+  ## To fix it we have to export "scenario_pm" with names and here compare the order of
+  ## both objects.
   summary2$scenario_PM <- round(io[[cities[x]]]$outcomes$scenario_pm, 2)
@@
   names(td) <- c('Scenario', 'CO2')
-  td$Scenario <- new_scen_names
+
+  # Dan: here the order is different (bus before car)
+  # td$Scenario <- new_scen_names
+  td$Scenario <- sapply(td$Scenario, FUN = get_qualified_scen_name)
 
@@
   # assign new scenario names
-  names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names
+  # Dan: This is not needed anymore
+  # names(io[[cities[x]]]$outcomes$mmets)[(n_col - scen_length):n_col] <- new_scen_names
 
@@
   # create boxplots
-  y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) + geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 
+  y <- ggplot(data_long, aes(x = scenario, y = mmet, fill = scenario)) + 
+    geom_boxplot(outlier.shape = NA) + ggtitle(cities[x]) + 
     labs(y = "Weekly MMET", x = "Scenarios") + ylim(0, limit)

### 2023-10-16 summary_tables.Rmd
Fix metric per 100k for summary docs/shiny app

@@
   for (j in pathway) { # loop for hb or pathway_hb
-    for (i in names(io[-c(2,3)])) { # loop for cities
+    for (i in names(io)[!names(io) %in% c('scen_prop','ithim_run' )]) { # loop for cities
 
@@
 
-      # Population by age, and age-sex
-      pop_by_age <- io[[i]]$demographic %>% group_by(age) %>% 
-        summarize(pop_age = sum(population))
-      pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>% 
-        summarize(pop_age_sex = sum(population))
+      # Total population
+      overall_pop <- sum(io[[i]]$demographic$population)
 
@@
         ) %>% # End mutate 
-        # join with demographic information
-        left_join(pop_by_age, by = c("age_cat" = "age")) %>%  
-        left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>%
+        # # join with demographic information
+        # left_join(pop_by_age, by = c("age_cat" = "age")) %>%  
+        # left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>%
         # create new columns showing burden per 100000 people by age or by age and sex
-        mutate(metric_100k = measure / pop_age * 100000, 
-               metric_100k_sex = measure / pop_age_sex * 100000)
+        mutate(metric_100k = measure / overall_pop * 100000, 
+               metric_100k_sex = measure / overall_pop * 100000)
 

### 2023-11-22 summary_tables.Rmd
corrected output of summary_tables.rmd, removed metric per 100,000, only include measure and population size

@@
       # Total population
-      overall_pop <- sum(io[[i]]$demographic$population)
+      #overall_pop <- sum(io[[i]]$demographic$population)
+
+      # Population by age and sex
+      pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>% 
+        summarize(pop_age_sex = sum(population))
 
@@
         ) %>% # End mutate 
-        # # join with demographic information
-        # left_join(pop_by_age, by = c("age_cat" = "age")) %>%  
-        # left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>%
+        # join with demographic information
+        left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex"))
+
+        # %>%
         # create new columns showing burden per 100000 people by age or by age and sex
-        mutate(metric_100k = measure / overall_pop * 100000, 
-               metric_100k_sex = measure / overall_pop * 100000)
+        # mutate(metric_100k = measure / overall_pop * 100000, 
+        #        metric_100k_sex = measure / overall_pop * 100000)
 

### 2023-11-23 R/distances_for_injury_function.R
Update linking of functions and references of how to run ithim vignette

@@ -27,11 +27,11 @@
 #'    Poisson injury model.
 #'
 #'  \item A new list (injuries_list) is created containing all strike and casualty mode and age and sex combinations
-#'    together with strike and casualty mode distances (add_distance_columns.R) for the baseline and all scenarios. For
+#'    together with strike and casualty mode distances (\code{\link{add_distance_columns()}}) for the baseline and all scenarios. For
 #'    the whw model, any strike mode and casualty pairs where strike mode equals casualty mode
 #'    are removed as fatalities for these combinations have already been added to the nov matrix.
 #'    Combinations which do not have a non-zero strike or casualty mode distance
-#'    are also removed. - This list will later be used in the injuries_function_2.R function
+#'    are also removed. This list will later be used in the \code{\link{injuries_function_2()}} function
 #'    to predict fatality counts using the Poisson injury regression model.
 #'
 #'  \item The casualty and strike mode exponents used to account for the safety in number effect

### 2023-11-23 R/gen_ap_rr.R

@@ -21,7 +21,7 @@
 #'    \item loop through the scenarios:
 #'      \itemize{
 #'      \item assign the relative risk for the given disease, age group, quantile and scenario to the
-#'        relevant people in the synthetic population by calling the AP_dose_response.R function
+#'        relevant people in the synthetic population by calling the \code{\link{AP_dose_response()}} function
 #'      }
 #'    }
 #' }

### 2023-11-23 R/get_all_distances.R

@@ -7,7 +7,7 @@
 #'
 #' \itemize{
 #' \item generate distance and duration matrices by age, sex, mode and scenario from the ithim_object$trip_scen_sets
-#'   for the synthetic population by calling the dist_dur_tbls.R function
+#'   for the synthetic population by calling the \code{\link{dist_dur_tbls()}} function
 #'
 #' \item find the total mode distances for each scenario and scale this up to the distance travelled by the entire
 #'   population by using the demographic information for the city
@@ -18,7 +18,7 @@
 #'   are then multiplied by these proportions to find the total population distances travelled by each mode
 #'   and age and sex category.
 #'
-#' \item the distances_for_injury_function.R function is called which creates a list inj_distances that is added
+#' \item the \code{\link{distances_for_injury_function()}} function is called which creates a list inj_distances that is added
 #'   to ithim_object containing the following matrices:
 #'   \itemize{
 #'   \item true_distances (population mode distances by age and sex with all walking modes and all car modes combined and

### 2023-11-23 R/injuries_function_2.R

@@ -3,7 +3,7 @@
 #' Predict injuries for baseline and scenarios based on Poisson regression model fitted
 #' on baseline fatality counts and distances
 #'
-#' This function uses the Poisson regression model built in the distances_for_injury_function.R to predict fatality
+#' This function uses the Poisson regression model built in the \code{\link{distances_for_injury_function()}} to predict fatality
 #' counts for the Baseline and all the scenarios. It performs the following steps:
 #'
 #' \itemize{

### 2023-11-23 R/ithim_setup_parameters.R

@@ -35,7 +35,7 @@
 #'   and assign a quantile to them by sampling from a uniform distribution between 0 and 1
 #' }
 #'
-#' At the bottom of this function, the dirichlet_pointiness() function is defined which
+#' At the bottom of this function, the \code{\link{dirichlet_pointiness()}} function is defined which
 #' parameterises the Dirichlet distributions for the PM and CO2 emission inventories.
 #'
 #'

### 2023-11-23 R/ithim_uncertainty.R

@@ -9,16 +9,16 @@
 #' \item extract the ithim_object list entries
 #'
 #' \item extract the sampled parameters which are stored in the ithim_object
-#' \item call set_vehicle_inventory.R to update emissions if any
+#' \item call \code{\link{set_vehicle_inventory()}} to update emissions if any
 #'   emission parameters have been sampled from a distribution
 #'
-#' \item call get_synthetic_from_trips.R to update the synthetic trips and synthetic population if any
+#' \item call \code{\link{get_synthetic_from_trips()}} to update the synthetic trips and synthetic population if any
 #'   relevant input parameters have been sampled from a distribution
 #'
-#' \item call get_all_distances.R to recalculate the distances if any of the relevant input parameters have
+#' \item call \code{\link{get_all_distances()}} to recalculate the distances if any of the relevant input parameters have
 #'   been sampled from a distribution
 #'
-#' \item run ITHIM-Global model by calling ithim_calculation_sequence.R
+#' \item run ITHIM-Global model by calling \code{\link{ithim_calculation_sequence()}}
 #' }
 #'
 #' @param ithim_object list of necessary inputs, including parameters

### 2023-11-23 R/run_ithim.R

@@ -43,32 +43,32 @@ run_ithim <- function(ithim_object, seed = 1) {
 #' \item  air pollution pathway:
 #'    \itemize{
 #'    \item calculate the PM2.5 exposure for each person in the synthetic population
-#'      and PM2.5 emissions for each mode and scenario (scenario_pm_calculations.R)
+#'      and PM2.5 emissions for each mode and scenario (\code{\link{scenario_pm_calculations()}})
 #'
-#'    \item calculate the CO2 emissions for each mode and scenario (scenario_co2_calculations.R)
+#'    \item calculate the CO2 emissions for each mode and scenario (\code{\link{scenario_co2_calculations()}})
 #'
 #'    \item assign relative risk to each person in the synthetic population for each disease
 #'      related to PM pollution and each scenario based on the individual PM exposure
-#'      levels (gen_ap_rr.R)
+#'      levels (\code{\link{gen_ap_rr()}})
 #'      }
 #'
 #' \item  physical activity pathway:
 #'    \itemize{
-#'    \item calculate total mMETs for each person in the synthetic population (total_mmet.R)
+#'    \item calculate total mMETs for each person in the synthetic population (\code{\link{total_mmet()}})
 #'
 #'    \item assign relative risk to each person in the synthetic population for each disease
 #'      related to physical activity levels and each scenario based on the individual mMET
-#'      values (gen_pa_rr.R)
+#'      values (\code{\link{gen_pa_rr()}})
 #'    }
 #'
 #' \item  physical activity and air pollution combined:
 #'    \itemize{
 #'    \item combine the PA and AP datasets by joining the two datasets. For disease affected by
 #'      both PA and AP calculate the joined relative risk by multiplying the PA and AP
-#'      relative risks (combined_rr_ap_pa.R)
+#'      relative risks (\code{\link{combined_rr_ap_pa()}})
 #'
 #'    \item calculate the health burden (Yll and deaths) for each disease and age and sex
-#'      category (health_burden.R):
+#'      category (\code{\link{health_burden()}}):
 #'      \itemize{
 #'      \item calculate the health burden (Yll and deaths) for each disease and age
 #'        and sex category. Combine the AP and PA pathways for diseases affected
@@ -83,16 +83,16 @@ run_ithim <- function(ithim_object, seed = 1) {
 #'    \itemize{
 #'    \item estimate the injury deaths for the baseline and each scenario by age and sex category,
[…]

### 2023-11-23 R/run_ithim_setup.R

@@ -49,27 +49,27 @@
 #'    \item update default values if city specific values are given as input parameters
 #'    }
 #'
-#' \item load and processe data from files by calling ithim_load_data.R
+#' \item load and process data from files by calling \code{\link{ithim_load_data()}}
 #'
-#' \item call ithim_setup_parameters.R to set the given input parameters to the global
+#' \item call \code{\link{ithim_setup_parameters()}} to set the given input parameters to the global
 #'   environment if running in constant mode or to obtain NSAMPLE samples from the
 #'   given distributions for each of the input parameters if running in sample mode
 #'
 #' \item set flags which cause certain parts of the model to be called at a later stage
-#'  (ithim_uncertainty.R) IF certain input parameters were sampled from a distribution
+#'  (\code{\link{ithim_uncertainty()}}) IF certain input parameters were sampled from a distribution
 #'
-#' \item call complete_trip_distance_duration.R to add any missing stage or distance
+#' \item call \code{\link{complete_trip_distance_duration()}} to add any missing stage or distance
 #'   information to the trip data
 #'
 #' \item if none of the corresponding input parameters were sampled from a distribution,
-#'   call set_vehicle_inventory.R to create a dataframe with mode specific speed,
+#'   call \code{\link{set_vehicle_inventory()}} to create a dataframe with mode specific speed,
 #'   distance and emission information
 #'
 #' \item if none of the corresponding input parameters were sampled from a distribution,
-#'   call get_synthetic_from_trips to set synthetic trips and synthetic population
+#'   call \code{\link{get_synthetic_from_trips()}} to set synthetic trips and synthetic population
 #'
 #' \item if none of the corresponding input parameters were sampled from a distribution,
-#'   call get_all_distances.R to calculate trip distances
+#'   call \code{\link{get_all_distances()}} to calculate trip distances
 #' }
 #'
 #' @param seed set seed to get the same results when sampling from a distribution

### 2023-11-23 R/set_injury_contingency.R

@@ -8,7 +8,7 @@
 #'
 #' \itemize{
 #' \item The data is split into a WHW (who hit who) matrix where both casualty and strike mode are given
-#'   and a NOV (no other vehicle) matrix where strike mode was set to NOV in ithim_load_data.R or
+#'   and a NOV (no other vehicle) matrix where strike mode was set to NOV in \code{\link{ithim_load_data()}} or
 #'   no other vehicle was involved in the accident.
 #'
 #' \item If no age and gender information is given, then the counts are multiplied by the

### 2023-11-23 R/walk_to_pt_and_combine_scen.R

@@ -15,7 +15,7 @@
 #'   \item divide public transport trips into those with and without a 'walk to pt' stage
 #'
 #'  \item add a 'walk to pt' stage to those public transport trips without a walking stage
-#'     (add_walk_trips.R)
+#'     (\code{\link{add_walk_trips()}})
 #'     }
 #' \item combine all trips from all scenarios into one dataframe
 #'

### 2023-11-29 summary_tables.Rmd
Commented out health related outputs and kept only the necessary bits re distance, duration and definition of scenario

--- 2023-10-16 summary_tables.Rmd
@@
 
-# Distance categories
+## Distance categories
 
@@
 
-### Baseline mode split for all cities (%)
-```{r}
-# create baseline mode share table containing information for all cities
-
-# prep data
-trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>%
-  mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>%
-  janitor::adorn_totals('row')
-
-# print table to html
-kable(trip_prop, caption = 'Baseline mode share (%) for all cities - only selected modes')
-
-```
-
-```{r unload_tidyverse, echo = F, message = F}
-suppressWarnings({
-  detach("package:tidyverse", character.only = T)
-})
-
-```
+<!-- ### Baseline mode split for all cities (%) -->
+<!-- ```{r} -->
+<!-- # create baseline mode share table containing information for all cities -->
+
+<!-- # prep data -->
+<!-- trip_prop <- trip_prop %>% dplyr::filter(stage_mode != "row_sums") %>% -->
+<!--   mutate_at(2:ncol(trip_prop), ~ as.numeric(as.character(.x))) %>% -->
+<!--   janitor::adorn_totals('row') -->
+
+<!-- # print table to html -->
+<!-- kable(trip_prop, caption = 'Baseline mode share (%) for all cities - only selected modes') -->
+
+<!-- ``` -->
+
+<!-- ```{r unload_tidyverse, echo = F, message = F} -->
+<!-- suppressWarnings({ -->
+<!--   detach("package:tidyverse", character.only = T) -->
+<!-- }) -->
+
+<!-- ``` -->
 
@@
 
-### Distance table showing average baseline distances in km travelled by each person in the synthetic population by mode for all cities
-
-```{r}
-# create table showing average baseline distances by mode for all cities
-trip_dist <- trip_dist %>% janitor::adorn_totals('row')
-
-kable(trip_dist, caption = "Avg. baseline distance (km) per person in the synthetic population per day across all cities")
-
-```
-
-## Distance tables scaled by using total model age population (per year in km)
-
-City specific distance tables for baseline and scenarios for the entire population with ages considered in the model
-
-```{r trip_tot_dist = "asis", message=F}
-# create distance tables for the yearly distances travelled by the entire population
-bl_td <- list()
-
-for (city in cities) { # loop through cities
-
-  # extract daily population distances for ages considered in model and multiply by 365 to to get yearly distances
-  td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
-    mutate_if(is.numeric, list(~round((.) * 365, 3)))
-
-  # update scenario names
-  colnames(td)[2:ncol(td)] <- scen_names_updated
-
-  # match with pre-defined mode order
-  x <- match(td$stage_mode, trip_order$mode)
-  x[!is.na(x)]
-  td <- td[order(x),]
-  row.names(td) <- NULL
-
-  # initialise bl_td with stage_mode
-  if (length(bl_td) == 0){
-    bl_td <- td %>% dplyr::select(stage_mode)
-  }
-
-  # calculate row and column totals
-  td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col'))
-
-  # extract baseline distances
-  td1 <- td %>% dplyr::select(stage_mode, Baseline)
-
-  names(td1)[2] <- city
-
-  # join baseline distances for all cities
-  bl_td <- inner_join(bl_td, td1, by = 'stage_mode')
-
-  # format table
-  td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T)
-
-  # print to html
-  cat(paste("Yearly distance (km) travelled by the entire population with ages considered in the model for ", city))
-  print(td1)
-
-  cat("\n")
-
-}
-
-```
-
-### Baseline total distance for all cities (per year in km)
-
-Baseline distance travelled by entire population with ages considered in the model for all cities
-
-```{r trip_dist_all_cities = "asis", message=F}
-
-# create baseline yearly distance table for the entire population with ages considered in the model for all cities
-
-backup <- bl_td
-
-# add row totals
-td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row')
-
-# remove all spaces/blanks
-for (i in 2:ncol(td)){
-  td[, i] <- comprss(td[, i])
-}
-
-print(kable(td, caption = paste("Yearly baseline distance table (km) for population with ages considered in the model")))
-
-```
-
-## Distance by distance category
-### Tables
-
-```{r trip_dist_mode_figs = "asis"}
-# for each city create a table showing the total distance, the number of trips and the proportion of trips in each distance category
-
-for (city in cities){ # loop through cities
-
-  # extract trip data for the baseline scenario and calculate the total distance, the number of trips and the proportion of trips in each distance category
-  df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline' | scenario == 'baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>% 
-    summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), proportion = round(n_vals / nrow(.) * 100, 1))
-
- # print to html
-  print(kable(df, caption = paste("Distance in travel survey by mode for the ages considered in the model for  ", city)))
-}
-
-```
-
-## Daily population distances (per day in km) for ages considered in model
-### Tables
-
-```{r trip_dist_mode_figs = "asis"}
-
-for (city in cities){ # loop through distances
-
-  # extract total daily distances travelled by the entire population with ages considered in the model
-  df <- io[[city]]$true_dist
-
-  names(df)[-1] <- scen_names_updated
-
-  print(kable(df, caption = paste("Total population distance with ages considered in model by mode per day for  ", city)))
-}
-
-```
+<!-- ### Distance table showing average baseline distances in km travelled by each person in the synthetic population by mode for all cities -->
+
+<!-- ```{r} -->
+<!-- # create table showing average baseline distances by mode for all cities -->
+<!-- trip_dist <- trip_dist %>% janitor::adorn_totals('row') -->
+
+<!-- kable(trip_dist, caption = "Avg. baseline distance (km) per person in the synthetic population per day across all cities") -->
+
+<!-- ``` -->
+
+<!-- ## Distance tables scaled by using total model age population (per year in km) -->
+
+<!-- City specific distance tables for baseline and scenarios for the entire population with ages considered in the model -->
+
+<!-- ```{r trip_tot_dist = "asis", message=F} -->
+<!-- # create distance tables for the yearly distances travelled by the entire population -->
+<!-- bl_td <- list() -->
+
+<!-- for (city in cities) { # loop through cities -->
+
+<!--   # extract daily population distances for ages considered in model and multiply by 365 to to get yearly distances -->
+<!--   td <- io[[city]]$true_dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% -->
+<!--     mutate_if(is.numeric, list(~round((.) * 365, 3))) -->
+
+<!--   # update scenario names -->
+<!--   colnames(td)[2:ncol(td)] <- scen_names_updated -->
+
+<!--   # match with pre-defined mode order -->
+<!--   x <- match(td$stage_mode, trip_order$mode) -->
+<!--   x[!is.na(x)] -->
+<!--   td <- td[order(x),] -->
+<!--   row.names(td) <- NULL -->
+
+<!--   # initialise bl_td with stage_mode -->
+<!--   if (length(bl_td) == 0){ -->
+<!--     bl_td <- td %>% dplyr::select(stage_mode) -->
+<!--   } -->
+
+<!--   # calculate row and column totals -->
+<!--   td <- td %>% ungroup() %>% janitor::adorn_totals(c('row', 'col')) -->
+
+<!--   # extract baseline distances -->
+<!--   td1 <- td %>% dplyr::select(stage_mode, Baseline) -->
+
+<!--   names(td1)[2] <- city -->
+
+<!--   # join baseline distances for all cities -->
+<!--   bl_td <- inner_join(bl_td, td1, by = 'stage_mode') -->
+
+<!--   # format table -->
+<!--   td1 <- td %>% gt() %>% fmt_number(columns = 2:ncol(td), decimals = T, suffixing = T) -->
+
+<!--   # print to html -->
+<!--   cat(paste("Yearly distance (km) travelled by the entire population with ages considered in the model for ", city)) -->
+<!--   print(td1) -->
+
+<!--   cat("\n") -->
+
+<!-- } -->
+
+<!-- ``` -->
+
+<!-- ### Baseline total distance for all cities (per year in km) -->
+
+<!-- Baseline distance travelled by entire population with ages considered in the model for all cities -->
+
+<!-- ```{r trip_dist_all_cities = "asis", message=F} -->
+
+<!-- # create baseline yearly distance table for the entire population with ages considered in the model for all cities -->
+
+<!-- backup <- bl_td -->
+
+<!-- # add row totals -->
+<!-- td <- as.data.frame(bl_td) %>% janitor::adorn_totals('row') -->
+
+<!-- # remove all spaces/blanks -->
+<!-- for (i in 2:ncol(td)){ -->
+<!--   td[, i] <- comprss(td[, i]) -->
+<!-- } -->
+
+<!-- print(kable(td, caption = paste("Yearly baseline distance table (km) for population with ages considered in the model"))) -->
+
+<!-- ``` -->
+
+<!-- ## Distance by distance category -->
+<!-- ### Tables -->
+
+<!-- ```{r trip_dist_mode_figs = "asis"} -->
+<!-- # for each city create a table showing the total distance, the number of trips and the proportion of trips in each distance category -->
+
+<!-- for (city in cities){ # loop through cities -->
+
+<!--   # extract trip data for the baseline scenario and calculate the total distance, the number of trips and the proportion of trips in each distance category -->
+<!--   df <- io[[city]]$trip_scen_sets %>% dplyr::filter(scenario == 'Baseline' | scenario == 'baseline') %>% distinct(trip_id, .keep_all = T) %>% group_by(trip_distance_cat) %>%  -->
+<!--     summarise(sum_dist = sum(trip_distance), n_vals = dplyr::n(), proportion = round(n_vals / nrow(.) * 100, 1)) -->
+
+<!--  # print to html -->
+<!--   print(kable(df, caption = paste("Distance in travel survey by mode for the ages considered in the model for  ", city))) -->
+<!-- } -->
+
+<!-- ``` -->
+
+<!-- ## Daily population distances (per day in km) for ages considered in model -->
+<!-- ### Tables -->
+
+<!-- ```{r trip_dist_mode_figs = "asis"} -->
+
+<!-- for (city in cities){ # loop through distances -->
+
+<!--   # extract total daily distances travelled by the entire population with ages considered in the model -->
+<!--   df <- io[[city]]$true_dist -->
+
+<!--   names(df)[-1] <- scen_names_updated -->
+
+<!--   print(kable(df, caption = paste("Total population distance with ages considered in model by mode per day for  ", city))) -->
+<!-- } -->
+
+<!-- ``` -->
 
@@
 
-### Baseline average duration (mins) per day for all cities
-
-```{r}
-# Average baseline duration travelled on each mode for all cities for all people in synthetic population
-
-# calculate the row sum
-trip_dur <- trip_dur %>% janitor::adorn_totals('row')
-
-kable(trip_dur, caption = "Avg. duration (min) per person in the synthetic population per day across all cities")
-
-```
-
-## Health outcomes
-
-### Change in YLLs compared to reference scenario {#change_YLL}
-
-```{r scen_prop = "asis"}
-
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# Calculate the changes in YLL compared to the reference scenario by age groups
-
-# find ap ylls columns
-ap_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x))))
-
-# find pa yll columns
-pa_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x))))
-
-# # find scenario names of health burden as the reference scenario is not necessarily the baseline scenario
-# hb_columns <- names(io[[cities[1]]]$outcomes$hb$ylls)[3: length(names(io[[cities[1]]]$outcomes$hb$ylls))]
-# 
-# # find the unique scenario names in the hb columns
-# inj_scen_cols <- grep("inj", hb_columns) # find location of injury columns
-# # extract unique scenario names
-# hb_scen_names <- unique(sapply(stringr::str_split(hb_columns[inj_scen_names], "_yll_inj", n = 3), function(x) x[1]))
-# # update scenario names
-# hb_scen_names_upd <- unlist(lapply(hb_scen_names, FUN=get_qualified_scen_name))
-
-# from the known scenarios, remove the reference scenario
-all_orig_scen_upd <- c('Baseline',orig_scen_names_updated)
-hb_scen_names_upd <- all_orig_scen_upd[all_orig_scen_upd!=reference_scenario]
-
-# calculate ylls by age category and scenario
-# i.e. sum over all disease and remove sex 
-yll_totals <- lapply(io[-c(2,3)],function(x){
-  temp <- sapply(1:NSCEN,function(y){# loop through scenarios
-    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
-    xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)]) # sum over all diseases
-    sapply(sort(unique(xx$age_cat)),function(z)
-      sum(xxx[xx$age_cat == z])) # sum age categories ignoring sex
-  })
-  colnames(temp) <- hb_scen_names_upd # update scenario names
-  temp
-})
-
-# add the column total, round the results and print to html
-sum_and_round_and_print(yll_totals,"Change total in YLLs compared to the reference scenario in ")
-```
-
-### Change in YLLs compared to reference scenario per 100,000 people {#change_YLL_age_100k}
-
-YLLs per 100,000 people by age group by scenario
-
-```{r scen_prop = "asis"}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate change in YLLs compared to the reference scenario per 100,000 people
-
-# extract population statistics
-pop_by_age <- lapply(io[-c(2,3)], function(x)
-  sapply(unique(x$demographic$age), 
-         function(y)
-           sum(subset(x$demographic, age == y)$population)
-         )
-  )
-
-# calculate ylls per 100,000 people
-yll_rates <- lapply(cities, function(x)
-  rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000, # divide by population
-        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), # calculate rate for total YLLs
-                                             length = NSCEN)*100000
-        )
-  )
-
-names(yll_rates) <- cities
-
-# round results and print to html
-round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
-```
-
-### Change in YLLs compared to reference scenario per 100,000 people by age and sex  {#change_YLL_gender_age_100k}
-
-Change in YLLs per 100,000 people by age and sex category for each scenario compared to the reference scenario
-
-```{r scen_prop = "asis"}
-
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate changes in YLLs for each scenario compared to the Baseline per 100,000 people by age and sex category
-
-# calculate the total ylls for each scenario summing across all diseases
-yll_totals <- lapply(io[-c(2,3)], function(x) {
-  temp <- sapply(1:NSCEN,function(y){
-    xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
-    rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)])
-  })
-  rownames(temp) <- apply(x$outcomes$hb$deaths[,c('sex','age_cat')], 1,
-                          function(z)paste0(z[2], '_', z[1]))
-  colnames(temp) <- orig_scen_names_updated
-  temp
-})
-
-# calculate the YLLs per 100,000 people
-yll_rates <- lapply(cities, function(x)
-  rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')], 1,
-                                    function(z) paste0(z[2],'_',z[1])), 
-                              rownames(yll_totals[[x]])),] /
-          t(repmat(io[[x]]$demographic$population,NSCEN,1)) * 100000,
-        Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]),
-                                             length = NSCEN) * 100000))
-
-names(yll_rates) <- cities
-
-# round and print ylls per 100,000 people to html
-round_and_print(yll_rates,"Change in YLLs per 100,000 in ")
-```
-
-### Change in YLLs due to injury compared to the reference scenario  {#change_yll_injury}
-
-```{r}
-
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate change in YLLs due to injury
-
-# calculate change in YLLs due to injury by summing across all age categories for the injury results
-injury_totals <- lapply(io[-c(2,3)], function(x) {
-  xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
-  injury_col <- grep("inj", names(xx))[1] 
-  xxx <- sapply(1:NSCEN, function(y)
-    sapply(sort(unique(xx$age_cat)), function(z)
-      sum(xx[xx$age_cat == z, injury_col - 1 + y])))
-  colnames(xxx) <- hb_scen_names_upd
-  xxx
-})
-
-# calculate the total, round the results and print to html
-sum_and_round_and_print(injury_totals, "Change in YLLs due to injury in ")
-```
-
-### Change in YLLs due to PA compared to the reference scenario {#change_yll_PA}
-
-```{r}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate the changes in YLLs compared to the reference scenario caused by changes in PA for the unique age categories
-
-# sum the YLLs for each scenario by age category
-pa_totals <- lapply(io[-c(2,3)],function(x){
-  xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
-  xxx <- sapply(1:NSCEN, function(y){
-    xxx <- rowSums(xx[, pa_yll_cols[seq(y, length(pa_yll_cols), by = NSCEN)]])
-    sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
-  })
-  colnames(xxx) <- hb_scen_names_upd
-  xxx
-})
-
-# calculate column sums, round the results and print to html
-sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ")
-```
-
-### Change in YLLs due to AP compared to the reference scenario {#change_yll_AP}
-```{r}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate the changes in YLLS compared to the reference scenario due to changes in AP
-
-# sum the YLLs by scenario and age category
-ap_totals <- lapply(io[-c(2,3)],function(x){
-  xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
-  xxx <- sapply(1:NSCEN,function(y) {
-    xxx <- rowSums(xx[, ap_yll_cols[seq(y, length(ap_yll_cols), by = NSCEN)]])
-    sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z]))
-  })
-  colnames(xxx) <- hb_scen_names_upd
-  xxx
-})
-
-# calculate column sums, round the results and print to html
-sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ")
-```
-
-### Change in YLLs due to injury compared to the reference scenario per 100,000 people {#change_yll_injury_100k}
-
-<!-- YLLs per 100,000 people by age group by scenario - injury -->
-
-```{r scen_prop = "asis"}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to injury
-injury_rates <- lapply(cities, function(x)
-  rbind(
-    injury_totals[[x]]/t(repmat(pop_by_age[[x]],NSCEN,1)) * 100000,
-    Total = colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]),
-                                            length = NSCEN) * 100000)
-  )
-names(injury_rates) <- cities
-
-# round results and print to html
-round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 people in ")
-```
-
-### Change in YLLs due to PA compared to the reference scenario per 100,000 people {#change_yll_PA_100k}
-
-```{r}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to PA
-pa_rates <- lapply(cities, function(x)
-  rbind(
-    pa_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN, 1)) * 100000,
-    Total = colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]),
-                                        length = NSCEN) * 100000))
-names(pa_rates) <- cities
-
-# round results and print to html
-round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 people in ")
-```
-
-### Change in YLLs due to AP compared to the reference scenario per 100,000 people {#change_yll_AP_100k}
-
-```{r}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate the changes in YLLs compared to the reference scenario per 100,000 people due to AP
-ap_rates <- lapply(cities, function(x)
-  rbind(
-    ap_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN ,1)) * 100000,
-    Total = colSums(ap_totals[[x]]) / rep(sum(pop_by_age[[x]]),
-                                          length = NSCEN) * 100000))
-names(ap_rates) <- cities
-
-# round results and print to html
-round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 people in ")
-```
-
-## BY DISEASE
-
-Change in total deaths compared to the reference scenario (for the city based on real population size but only for ages considered in the model) by scenario
-
-### Change in deaths by disease compared to the reference scenario {#change_death_disease}
-
-```{r}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate the changes in deaths compared to the reference scenario by disease
-
-# sum the total deaths by scenario
-disease_totals <- lapply(io[-c(2,3)],function(x){
-  # extract disease counts
-  xx <- x$outcomes$hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) 
-  xxx <- sapply(1:NSCEN,function(y){ # sum over scenarios
-    colSums(xx[, seq(y + 2, ncol(xx), by = NSCEN)])
-  })
-  colnames(xxx) <- hb_scen_names_upd
-
-  # find scenario name of row names
-  row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2]))
-
-  # update row names
-  rownames(xxx) <- gsub(paste0("sc_",row_scen,"_deaths_"), "", rownames(xxx))
-
-  xxx
-})
-
-# sum, round and print the results 
-sum_and_round_and_print(disease_totals,"Change in deaths by disease in ")
-```
-
-### Change in deaths by disease compared to the reference scenario per 100,000 people  {#change_death_disease_100k}
-
-Change in deaths per 100,000 people (based on real population size but only for ages considered in the model) by scenario in comparison to the reference scenario
-
-```{r scen_prop = "asis"}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate the change in deaths compared to the reference scenario per 100,000 people
-
-# divided total number of deaths by population
-disease_rates <- lapply(cities, function(x)
-  rbind(
-    disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000,
-    Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]), 
-                                               length = NSCEN)*100000))
-names(disease_rates) <- cities
-
-# round results and print to html
-round_and_print(disease_rates,"Change in deaths by disease per 100,000 people in ")
-```
-
-### Change in total YLLs by disease compared to the reference scenario {#change_yll_disease}
-
-Change in total YLLs by disease compared to the reference scenario (based on real population size, for ages considered within the ITHIM model)
-
-```{r}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# sum the total ylls by scenario
-disease_totals <- lapply(io[-c(2,3)],function(x){
-  xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))
-  xxx <- sapply(1:NSCEN,function(y) {
-    colSums(xx[,seq(y + 2, ncol(xx), by = NSCEN)])
-  })
-  colnames(xxx) <- hb_scen_names_upd
-
-  # find scenario name of row names
-  row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2]))
-
-  # update row names
-  rownames(xxx) <- gsub(paste0("sc_",row_scen,"_ylls_"), "", rownames(xxx))
-  xxx
-})
-
-# sum and round results, print to html
-sum_and_round_and_print(disease_totals,"Change in YLL by disease in ")
-```
-
-### Change in YLL due to disease compared to the reference scenario per 100,000 people {#change_yll_disease_100k}
-
-Change in total YLLs by disease compared to the reference scenario per 100,000 people (based on real population size, for ages considered within the ITHIM model)
-
-```{r scen_prop = "asis"}
-cat(paste0('Reference scenario:  ', reference_scenario))
-
-# calculate the change in YLL by disease by dividing by the population
-disease_rates <- lapply(cities, function(x)
-  rbind(
-    disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000,
-    Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]),
-                                               length = NSCEN) * 100000))
-names(disease_rates) <- cities
-
-# round and print results to html
-round_and_print(disease_rates,"Change in YLL by disease per 100,000 in ")
-```
-
-```{r, echo=FALSE, results='hide', message=FALSE}
-
-# Export all health results
-
-# hb - gives all AP (for diseases affected by AP only), PA (for diseases affected by PA only), combined AP and PA (for diseases affected by both AP and PA) and inj results
-# pathway_hb - gives all AP, PA and inj results (but does not combine the results for diseases affected by both AP and PA)
-
-# initialise variables
-health <- list()
-measure <- c("ylls", "deaths")
-pathway <- c("hb", "pathway_hb")
-
-for (k in measure) { # loop for ylls or deaths
-  for (j in pathway) { # loop for hb or pathway_hb
-    for (i in names(io)[!names(io) %in% c('scen_prop','ithim_run' )]) { # loop for cities
-
-      # Temporal dataset for health outcomes
-      temp <- io[[i]]$outcomes[[j]][[k]]
-
-      # Total population
-      overall_pop <- sum(io[[i]]$demographic$population)
-
-      # get the health burden into the correct format by creating a long format and adding additional columns
-      # create long format
-      health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure", 
-                                      -sex, -age_cat) %>% 
-        mutate(city = i, # create city column
-               scenario = case_when( # create scenario column
-                 grepl("sc_cycle", cause) ~ get_qualified_scen_name('sc_cycle'),
-                 grepl("sc_car", cause) ~ get_qualified_scen_name('sc_car'),
-                 grepl("sc_bus", cause) ~ get_qualified_scen_name('sc_bus'),
-                 grepl("sc_motorcycle", cause) ~ get_qualified_scen_name('sc_motorcycle')
-               ),
-               level1 = case_when( # create level 1 column
-                 grepl("all_cause", cause) ~ "L1: All Cause",
-                 grepl("inj", cause) ~ "L1: RTI"
-               ),
-               level2 = case_when( # create level 2 column
-                 grepl("total_cancer", cause) ~ "L2: Cancer",
-                 grepl("CVD", cause) ~ "L2: CVD",
-                 grepl("respiratory", cause) ~ "L2: Respiratory",
-                 grepl("T2D", cause) ~ "L2: Other",
-                 grepl("total_dementia", cause) ~ "L2: Other",
-                 grepl("Parkinson", cause) ~ "L2: Other",
-                 grepl("inj", cause) ~ "L2: RTI"
-               ),
-               level3 = case_when( # create level 3 column
-                 grepl("IHD", cause) ~ "L3: IHD",
-                 grepl("lung_cancer", cause) ~ "L3: lung_cancer",
-                 grepl("COPD", cause) ~ "L3: COPD",
-                 grepl("stroke", cause) ~ "L3: stroke",
-                 grepl("T2D", cause) ~ "L3: T2D",
-                 grepl("LRI", cause) ~ "L3: LRI",
-                 grepl("breast_cancer", cause) ~ "L3: breast_cancer",
-                 grepl("colon_cancer", cause) ~ "L3: colon_cancer",
-                 grepl("endo_cancer", cause) ~ "L3: endo_cancer",
-                 grepl("liver_cancer", cause) ~ "L3: liver_cancer",
-                 grepl("total_dementia", cause) ~ "L3: total_dementia",
-                 grepl("myeloma", cause) ~ "L3: myeloma",
-                 grepl("myeloid_leukemia", cause) ~ "L3: Myeloid Leukemia",
-                 grepl("Parkinson", cause) ~ "L3: Parkinson",
-                 grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer",
-                 grepl("stomach_cancer", cause) ~ "L3: stomach_cancer",
-                 grepl("inj", cause) ~ "L3: RTI"
-               ),
-               dose = case_when( # create dose column
-                 grepl("_pa_ap_", cause) ~ "PA and AP",
-                 grepl("_pa_", cause) ~ "PA",
-                 grepl("_ap_", cause) ~ "AP",
-                 grepl("_inj", cause) ~ "RTI"
-               )
-        ) %>% # End mutate 
-        # # join with demographic information
-        # left_join(pop_by_age, by = c("age_cat" = "age")) %>%  
-        # left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex")) %>%
-        # create new columns showing burden per 100000 people by age or by age and sex
-        mutate(metric_100k = measure / overall_pop * 100000, 
-               metric_100k_sex = measure / overall_pop * 100000)
-
-    } # Loop for each city
-  } # Loop for each pathway
-} # Loop for each measure
-
-# for each health burden and pathway join all cities into one dataframe
-ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-deaths <- rbindlist(health$deaths$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
-
-# # Export ylls HB
-# output_file <- "results/multi_city/health_impacts/data.xlsx"
-# writexl::write_xlsx(list(ylls = ylls,
-#                          ylls_pathway = ylls_pathway,
-#                          deaths = deaths,
-#                          deaths_pathway = deaths_pathway), 
-#                     path = output_file)
-
-# Export ylls HB
-
-# write csv with output version number
-write.csv(ylls,
-            paste0('results/multi_city/health_impacts/ylls_',output_version,'.csv'),
-            row.names = F)
-# Export ylls PATHWAY_HB
-write.csv(ylls_pathway,
-            paste0('results/multi_city/health_impacts/ylls_pathway_',output_version,'.csv'),
-            row.names = F)
-# Export deaths HB
-write.csv(deaths,
-            paste0('results/multi_city/health_impacts/deaths_',output_version,'.csv'),
-            row.names = F)
-# Export deaths PATHWAY_HB
-write.csv(deaths_pathway,
-            paste0('results/multi_city/health_impacts/deaths_pathway_',output_version,'.csv'),
-            row.names = F)
-
-# write csv without output version number
-write.csv(ylls,
-            paste0('results/multi_city/health_impacts/ylls.csv'),
-            row.names = F)
-# Export ylls PATHWAY_HB
-write.csv(ylls_pathway,
-            paste0('results/multi_city/health_impacts/ylls_pathway.csv'),
-            row.names = F)
-# Export deaths HB
-write.csv(deaths,
-            paste0('results/multi_city/health_impacts/deaths.csv'),
-            row.names = F)
-# Export deaths PATHWAY_HB
-write.csv(deaths_pathway,
-            paste0('results/multi_city/health_impacts/deaths_pathway.csv'),
-            row.names = F)
-
-```
+<!-- ### Baseline average duration (mins) per day for all cities -->
+
+<!-- ```{r} -->
+<!-- # Average baseline duration travelled on each mode for all cities for all people in synthetic population -->
+
+<!-- # calculate the row sum -->
+<!-- trip_dur <- trip_dur %>% janitor::adorn_totals('row') -->
+
+<!-- kable(trip_dur, caption = "Avg. duration (min) per person in the synthetic population per day across all cities") -->
+
+<!-- ``` -->
+
+<!-- ## Health outcomes -->
+
+<!-- ### Change in YLLs compared to reference scenario {#change_YLL} -->
+
+<!-- ```{r scen_prop = "asis"} -->
+
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # Calculate the changes in YLL compared to the reference scenario by age groups -->
+
+<!-- # find ap ylls columns -->
+<!-- ap_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('ap',as.character(x)))) -->
+
+<!-- # find pa yll columns -->
+<!-- pa_yll_cols <- which(sapply(colnames(io[[1]]$outcomes$pathway_hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))),function(x)grepl('pa',as.character(x)))) -->
+
+<!-- # # find scenario names of health burden as the reference scenario is not necessarily the baseline scenario -->
+<!-- # hb_columns <- names(io[[cities[1]]]$outcomes$hb$ylls)[3: length(names(io[[cities[1]]]$outcomes$hb$ylls))] -->
+<!-- #  -->
+<!-- # # find the unique scenario names in the hb columns -->
+<!-- # inj_scen_cols <- grep("inj", hb_columns) # find location of injury columns -->
+<!-- # # extract unique scenario names -->
+<!-- # hb_scen_names <- unique(sapply(stringr::str_split(hb_columns[inj_scen_names], "_yll_inj", n = 3), function(x) x[1])) -->
+<!-- # # update scenario names -->
+<!-- # hb_scen_names_upd <- unlist(lapply(hb_scen_names, FUN=get_qualified_scen_name)) -->
+
+<!-- # from the known scenarios, remove the reference scenario -->
+<!-- all_orig_scen_upd <- c('Baseline',orig_scen_names_updated) -->
+<!-- hb_scen_names_upd <- all_orig_scen_upd[all_orig_scen_upd!=reference_scenario] -->
+
+<!-- # calculate ylls by age category and scenario -->
+<!-- # i.e. sum over all disease and remove sex  -->
+<!-- yll_totals <- lapply(io[-c(2,3)],function(x){ -->
+<!--   temp <- sapply(1:NSCEN,function(y){# loop through scenarios -->
+<!--     xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
+<!--     xxx <- rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)]) # sum over all diseases -->
+<!--     sapply(sort(unique(xx$age_cat)),function(z) -->
+<!--       sum(xxx[xx$age_cat == z])) # sum age categories ignoring sex -->
+<!--   }) -->
+<!--   colnames(temp) <- hb_scen_names_upd # update scenario names -->
+<!--   temp -->
+<!-- }) -->
+
+<!-- # add the column total, round the results and print to html -->
+<!-- sum_and_round_and_print(yll_totals,"Change total in YLLs compared to the reference scenario in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLLs compared to reference scenario per 100,000 people {#change_YLL_age_100k} -->
+
+<!-- YLLs per 100,000 people by age group by scenario -->
+
+<!-- ```{r scen_prop = "asis"} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate change in YLLs compared to the reference scenario per 100,000 people -->
+
+<!-- # extract population statistics -->
+<!-- pop_by_age <- lapply(io[-c(2,3)], function(x) -->
+<!--   sapply(unique(x$demographic$age),  -->
+<!--          function(y) -->
+<!--            sum(subset(x$demographic, age == y)$population) -->
+<!--          ) -->
+<!--   ) -->
+
+<!-- # calculate ylls per 100,000 people -->
+<!-- yll_rates <- lapply(cities, function(x) -->
+<!--   rbind(yll_totals[[x]] / t(repmat(pop_by_age[[x]],NSCEN,1))*100000, # divide by population -->
+<!--         Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), # calculate rate for total YLLs -->
+<!--                                              length = NSCEN)*100000 -->
+<!--         ) -->
+<!--   ) -->
+
+<!-- names(yll_rates) <- cities -->
+
+<!-- # round results and print to html -->
+<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLLs compared to reference scenario per 100,000 people by age and sex  {#change_YLL_gender_age_100k} -->
+
+<!-- Change in YLLs per 100,000 people by age and sex category for each scenario compared to the reference scenario -->
+
+<!-- ```{r scen_prop = "asis"} -->
+
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate changes in YLLs for each scenario compared to the Baseline per 100,000 people by age and sex category -->
+
+<!-- # calculate the total ylls for each scenario summing across all diseases -->
+<!-- yll_totals <- lapply(io[-c(2,3)], function(x) { -->
+<!--   temp <- sapply(1:NSCEN,function(y){ -->
+<!--     xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
+<!--     rowSums(xx[,seq(2 + y, ncol(xx), by = NSCEN)]) -->
+<!--   }) -->
+<!--   rownames(temp) <- apply(x$outcomes$hb$deaths[,c('sex','age_cat')], 1, -->
+<!--                           function(z)paste0(z[2], '_', z[1])) -->
+<!--   colnames(temp) <- orig_scen_names_updated -->
+<!--   temp -->
+<!-- }) -->
+
+<!-- # calculate the YLLs per 100,000 people -->
+<!-- yll_rates <- lapply(cities, function(x) -->
+<!--   rbind(yll_totals[[x]][match(apply(io[[x]]$demographic[,c('sex','age')], 1, -->
+<!--                                     function(z) paste0(z[2],'_',z[1])),  -->
+<!--                               rownames(yll_totals[[x]])),] / -->
+<!--           t(repmat(io[[x]]$demographic$population,NSCEN,1)) * 100000, -->
+<!--         Total = colSums(yll_totals[[x]])/rep(sum(pop_by_age[[x]]), -->
+<!--                                              length = NSCEN) * 100000)) -->
+
+<!-- names(yll_rates) <- cities -->
+
+<!-- # round and print ylls per 100,000 people to html -->
+<!-- round_and_print(yll_rates,"Change in YLLs per 100,000 in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLLs due to injury compared to the reference scenario  {#change_yll_injury} -->
+
+<!-- ```{r} -->
+
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate change in YLLs due to injury -->
+
+<!-- # calculate change in YLLs due to injury by summing across all age categories for the injury results -->
+<!-- injury_totals <- lapply(io[-c(2,3)], function(x) { -->
+<!--   xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
+<!--   injury_col <- grep("inj", names(xx))[1]  -->
+<!--   xxx <- sapply(1:NSCEN, function(y) -->
+<!--     sapply(sort(unique(xx$age_cat)), function(z) -->
+<!--       sum(xx[xx$age_cat == z, injury_col - 1 + y]))) -->
+<!--   colnames(xxx) <- hb_scen_names_upd -->
+<!--   xxx -->
+<!-- }) -->
+
+<!-- # calculate the total, round the results and print to html -->
+<!-- sum_and_round_and_print(injury_totals, "Change in YLLs due to injury in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLLs due to PA compared to the reference scenario {#change_yll_PA} -->
+
+<!-- ```{r} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate the changes in YLLs compared to the reference scenario caused by changes in PA for the unique age categories -->
+
+<!-- # sum the YLLs for each scenario by age category -->
+<!-- pa_totals <- lapply(io[-c(2,3)],function(x){ -->
+<!--   xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
+<!--   xxx <- sapply(1:NSCEN, function(y){ -->
+<!--     xxx <- rowSums(xx[, pa_yll_cols[seq(y, length(pa_yll_cols), by = NSCEN)]]) -->
+<!--     sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z])) -->
+<!--   }) -->
+<!--   colnames(xxx) <- hb_scen_names_upd -->
+<!--   xxx -->
+<!-- }) -->
+
+<!-- # calculate column sums, round the results and print to html -->
+<!-- sum_and_round_and_print(pa_totals,"Change in YLLs due to PA in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLLs due to AP compared to the reference scenario {#change_yll_AP} -->
+<!-- ```{r} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate the changes in YLLS compared to the reference scenario due to changes in AP -->
+
+<!-- # sum the YLLs by scenario and age category -->
+<!-- ap_totals <- lapply(io[-c(2,3)],function(x){ -->
+<!--   xx <- x$outcomes$pathway_hb$ylls  |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
+<!--   xxx <- sapply(1:NSCEN,function(y) { -->
+<!--     xxx <- rowSums(xx[, ap_yll_cols[seq(y, length(ap_yll_cols), by = NSCEN)]]) -->
+<!--     sapply(sort(unique(xx$age_cat)), function(z) sum(xxx[xx$age_cat == z])) -->
+<!--   }) -->
+<!--   colnames(xxx) <- hb_scen_names_upd -->
+<!--   xxx -->
+<!-- }) -->
+
+<!-- # calculate column sums, round the results and print to html -->
+<!-- sum_and_round_and_print(ap_totals,"Change in YLLs due to AP in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLLs due to injury compared to the reference scenario per 100,000 people {#change_yll_injury_100k} -->
+
+<!-- <!-- YLLs per 100,000 people by age group by scenario - injury --> -->
+
+<!-- ```{r scen_prop = "asis"} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate the changes in YLLs compared to the reference scenario per 100,000 people due to injury -->
+<!-- injury_rates <- lapply(cities, function(x) -->
+<!--   rbind( -->
+<!--     injury_totals[[x]]/t(repmat(pop_by_age[[x]],NSCEN,1)) * 100000, -->
+<!--     Total = colSums(injury_totals[[x]])/rep(sum(pop_by_age[[x]]), -->
+<!--                                             length = NSCEN) * 100000) -->
+<!--   ) -->
+<!-- names(injury_rates) <- cities -->
+
+<!-- # round results and print to html -->
+<!-- round_and_print(injury_rates,"Change in YLLs due to injury per 100,000 people in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLLs due to PA compared to the reference scenario per 100,000 people {#change_yll_PA_100k} -->
+
+<!-- ```{r} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate the changes in YLLs compared to the reference scenario per 100,000 people due to PA -->
+<!-- pa_rates <- lapply(cities, function(x) -->
+<!--   rbind( -->
+<!--     pa_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN, 1)) * 100000, -->
+<!--     Total = colSums(pa_totals[[x]])/rep(sum(pop_by_age[[x]]), -->
+<!--                                         length = NSCEN) * 100000)) -->
+<!-- names(pa_rates) <- cities -->
+
+<!-- # round results and print to html -->
+<!-- round_and_print(pa_rates,"Change in YLLs due to PA per 100,000 people in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLLs due to AP compared to the reference scenario per 100,000 people {#change_yll_AP_100k} -->
+
+<!-- ```{r} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate the changes in YLLs compared to the reference scenario per 100,000 people due to AP -->
+<!-- ap_rates <- lapply(cities, function(x) -->
+<!--   rbind( -->
+<!--     ap_totals[[x]] / t(repmat(pop_by_age[[x]], NSCEN ,1)) * 100000, -->
+<!--     Total = colSums(ap_totals[[x]]) / rep(sum(pop_by_age[[x]]), -->
+<!--                                           length = NSCEN) * 100000)) -->
+<!-- names(ap_rates) <- cities -->
+
+<!-- # round results and print to html -->
+<!-- round_and_print(ap_rates,"Change in YLLs due to AP per 100,000 people in ") -->
+<!-- ``` -->
+
+<!-- ## BY DISEASE -->
+
+<!-- Change in total deaths compared to the reference scenario (for the city based on real population size but only for ages considered in the model) by scenario -->
+
+<!-- ### Change in deaths by disease compared to the reference scenario {#change_death_disease} -->
+
+<!-- ```{r} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate the changes in deaths compared to the reference scenario by disease -->
+
+<!-- # sum the total deaths by scenario -->
+<!-- disease_totals <- lapply(io[-c(2,3)],function(x){ -->
+<!--   # extract disease counts -->
+<!--   xx <- x$outcomes$hb$deaths |> dplyr::select(-c(ends_with("lb") | ends_with("ub")))  -->
+<!--   xxx <- sapply(1:NSCEN,function(y){ # sum over scenarios -->
+<!--     colSums(xx[, seq(y + 2, ncol(xx), by = NSCEN)]) -->
+<!--   }) -->
+<!--   colnames(xxx) <- hb_scen_names_upd -->
+
+<!--   # find scenario name of row names -->
+<!--   row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2])) -->
+
+<!--   # update row names -->
+<!--   rownames(xxx) <- gsub(paste0("sc_",row_scen,"_deaths_"), "", rownames(xxx)) -->
+
+<!--   xxx -->
+<!-- }) -->
+
+<!-- # sum, round and print the results  -->
+<!-- sum_and_round_and_print(disease_totals,"Change in deaths by disease in ") -->
+<!-- ``` -->
+
+<!-- ### Change in deaths by disease compared to the reference scenario per 100,000 people  {#change_death_disease_100k} -->
+
+<!-- Change in deaths per 100,000 people (based on real population size but only for ages considered in the model) by scenario in comparison to the reference scenario -->
+
+<!-- ```{r scen_prop = "asis"} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate the change in deaths compared to the reference scenario per 100,000 people -->
+
+<!-- # divided total number of deaths by population -->
+<!-- disease_rates <- lapply(cities, function(x) -->
+<!--   rbind( -->
+<!--     disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000, -->
+<!--     Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]),  -->
+<!--                                                length = NSCEN)*100000)) -->
+<!-- names(disease_rates) <- cities -->
+
+<!-- # round results and print to html -->
+<!-- round_and_print(disease_rates,"Change in deaths by disease per 100,000 people in ") -->
+<!-- ``` -->
+
+<!-- ### Change in total YLLs by disease compared to the reference scenario {#change_yll_disease} -->
+
+<!-- Change in total YLLs by disease compared to the reference scenario (based on real population size, for ages considered within the ITHIM model)  -->
+
+<!-- ```{r} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # sum the total ylls by scenario -->
+<!-- disease_totals <- lapply(io[-c(2,3)],function(x){ -->
+<!--   xx <- x$outcomes$hb$ylls |> dplyr::select(-c(ends_with("lb") | ends_with("ub"))) -->
+<!--   xxx <- sapply(1:NSCEN,function(y) { -->
+<!--     colSums(xx[,seq(y + 2, ncol(xx), by = NSCEN)]) -->
+<!--   }) -->
+<!--   colnames(xxx) <- hb_scen_names_upd -->
+
+<!--   # find scenario name of row names -->
+<!--   row_scen <- unique(sapply(stringr::str_split(rownames(xxx), "_", n = 3), function(x) x[2])) -->
+
+<!--   # update row names -->
+<!--   rownames(xxx) <- gsub(paste0("sc_",row_scen,"_ylls_"), "", rownames(xxx)) -->
+<!--   xxx -->
+<!-- }) -->
+
+<!-- # sum and round results, print to html -->
+<!-- sum_and_round_and_print(disease_totals,"Change in YLL by disease in ") -->
+<!-- ``` -->
+
+<!-- ### Change in YLL due to disease compared to the reference scenario per 100,000 people {#change_yll_disease_100k} -->
+
+<!-- Change in total YLLs by disease compared to the reference scenario per 100,000 people (based on real population size, for ages considered within the ITHIM model)  -->
+
+<!-- ```{r scen_prop = "asis"} -->
+<!-- cat(paste0('Reference scenario:  ', reference_scenario)) -->
+
+<!-- # calculate the change in YLL by disease by dividing by the population -->
+<!-- disease_rates <- lapply(cities, function(x) -->
+<!--   rbind( -->
+<!--     disease_totals[[x]] / sum(pop_by_age[[x]]) * 100000, -->
+<!--     Total = colSums(disease_totals[[x]]) / rep(sum(pop_by_age[[x]]), -->
+<!--                                                length = NSCEN) * 100000)) -->
+<!-- names(disease_rates) <- cities -->
+
+<!-- # round and print results to html -->
+<!-- round_and_print(disease_rates,"Change in YLL by disease per 100,000 in ") -->
+<!-- ``` -->
+
+<!-- ```{r, echo=FALSE, results='hide', message=FALSE} -->
+
+<!-- # Export all health results -->
+
+<!-- # hb - gives all AP (for diseases affected by AP only), PA (for diseases affected by PA only), combined AP and PA (for diseases affected by both AP and PA) and inj results -->
+<!-- # pathway_hb - gives all AP, PA and inj results (but does not combine the results for diseases affected by both AP and PA) -->
+
+<!-- # initialise variables -->
+<!-- health <- list() -->
+<!-- measure <- c("ylls", "deaths") -->
+<!-- pathway <- c("hb", "pathway_hb") -->
+
+<!-- for (k in measure) { # loop for ylls or deaths -->
+<!--   for (j in pathway) { # loop for hb or pathway_hb -->
+<!--     for (i in names(io)[!names(io) %in% c('scen_prop','ithim_run' )]) { # loop for cities -->
+
+<!--       # Temporal dataset for health outcomes -->
+<!--       temp <- io[[i]]$outcomes[[j]][[k]] -->
+
+<!--       # Total population -->
+<!--       #overall_pop <- sum(io[[i]]$demographic$population) -->
+
+<!--       # Population by age and sex -->
+<!--       pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>%  -->
+<!--         summarize(pop_age_sex = sum(population)) -->
+
+<!--       # get the health burden into the correct format by creating a long format and adding additional columns -->
+<!--       # create long format -->
+<!--       health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure",  -->
+<!--                                       -sex, -age_cat) %>%  -->
+<!--         mutate(city = i, # create city column -->
+<!--                scenario = case_when( # create scenario column -->
+<!--                  grepl("sc_cycle", cause) ~ get_qualified_scen_name('sc_cycle'), -->
+<!--                  grepl("sc_car", cause) ~ get_qualified_scen_name('sc_car'), -->
+<!--                  grepl("sc_bus", cause) ~ get_qualified_scen_name('sc_bus'), -->
+<!--                  grepl("sc_motorcycle", cause) ~ get_qualified_scen_name('sc_motorcycle') -->
+<!--                ), -->
+<!--                level1 = case_when( # create level 1 column -->
+<!--                  grepl("all_cause", cause) ~ "L1: All Cause", -->
+<!--                  grepl("inj", cause) ~ "L1: RTI" -->
+<!--                ), -->
+<!--                level2 = case_when( # create level 2 column -->
+<!--                  grepl("total_cancer", cause) ~ "L2: Cancer", -->
+<!--                  grepl("CVD", cause) ~ "L2: CVD", -->
+<!--                  grepl("respiratory", cause) ~ "L2: Respiratory", -->
+<!--                  grepl("T2D", cause) ~ "L2: Other", -->
+<!--                  grepl("total_dementia", cause) ~ "L2: Other", -->
+<!--                  grepl("Parkinson", cause) ~ "L2: Other", -->
+<!--                  grepl("inj", cause) ~ "L2: RTI" -->
+<!--                ), -->
+<!--                level3 = case_when( # create level 3 column -->
+<!--                  grepl("IHD", cause) ~ "L3: IHD", -->
+<!--                  grepl("lung_cancer", cause) ~ "L3: lung_cancer", -->
+<!--                  grepl("COPD", cause) ~ "L3: COPD", -->
+<!--                  grepl("stroke", cause) ~ "L3: stroke", -->
+<!--                  grepl("T2D", cause) ~ "L3: T2D", -->
+<!--                  grepl("LRI", cause) ~ "L3: LRI", -->
+<!--                  grepl("breast_cancer", cause) ~ "L3: breast_cancer", -->
+<!--                  grepl("colon_cancer", cause) ~ "L3: colon_cancer", -->
+<!--                  grepl("endo_cancer", cause) ~ "L3: endo_cancer", -->
+<!--                  grepl("liver_cancer", cause) ~ "L3: liver_cancer", -->
+<!--                  grepl("total_dementia", cause) ~ "L3: total_dementia", -->
+<!--                  grepl("myeloma", cause) ~ "L3: myeloma", -->
+<!--                  grepl("myeloid_leukemia", cause) ~ "L3: Myeloid Leukemia", -->
+<!--                  grepl("Parkinson", cause) ~ "L3: Parkinson", -->
+<!--                  grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer", -->
+<!--                  grepl("stomach_cancer", cause) ~ "L3: stomach_cancer", -->
+<!--                  grepl("inj", cause) ~ "L3: RTI" -->
+<!--                ), -->
+<!--                dose = case_when( # create dose column -->
+<!--                  grepl("_pa_ap_", cause) ~ "PA and AP", -->
+<!--                  grepl("_pa_", cause) ~ "PA", -->
+<!--                  grepl("_ap_", cause) ~ "AP", -->
+<!--                  grepl("_inj", cause) ~ "RTI" -->
+<!--                ) -->
+<!--         ) %>% # End mutate  -->
+<!--         # join with demographic information -->
+<!--         left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex"))  -->
+
+<!--         # %>% -->
+<!--         # create new columns showing burden per 100000 people by age or by age and sex -->
+<!--         # mutate(metric_100k = measure / overall_pop * 100000,  -->
+<!--         #        metric_100k_sex = measure / overall_pop * 100000) -->
+
+<!--     } # Loop for each city -->
+<!--   } # Loop for each pathway -->
+<!-- } # Loop for each measure -->
+
+<!-- # for each health burden and pathway join all cities into one dataframe -->
+<!-- ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
+<!-- ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
+<!-- deaths <- rbindlist(health$deaths$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
+<!-- deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
+
+<!-- # # Export ylls HB -->
+<!-- # output_file <- "results/multi_city/health_impacts/data.xlsx" -->
+<!-- # writexl::write_xlsx(list(ylls = ylls, -->
+<!-- #                          ylls_pathway = ylls_pathway, -->
+<!-- #                          deaths = deaths, -->
+<!-- #                          deaths_pathway = deaths_pathway),  -->
+<!-- #                     path = output_file) -->
+
+<!-- # Export ylls HB -->
+
+<!-- # write csv with output version number -->
+<!-- write.csv(ylls, -->
+<!--             paste0('results/multi_city/health_impacts/ylls_',output_version,'.csv'), -->
+<!--             row.names = F) -->
+<!-- # Export ylls PATHWAY_HB -->
+<!-- write.csv(ylls_pathway, -->
+<!--             paste0('results/multi_city/health_impacts/ylls_pathway_',output_version,'.csv'), -->
+<!--             row.names = F) -->
+<!-- # Export deaths HB -->
+<!-- write.csv(deaths, -->
+<!--             paste0('results/multi_city/health_impacts/deaths_',output_version,'.csv'), -->
+<!--             row.names = F) -->
+<!-- # Export deaths PATHWAY_HB -->
+<!-- write.csv(deaths_pathway, -->
+<!--             paste0('results/multi_city/health_impacts/deaths_pathway_',output_version,'.csv'), -->
+<!--             row.names = F) -->
+
+<!-- # write csv without output version number -->
+<!-- write.csv(ylls, -->
+<!--             paste0('results/multi_city/health_impacts/ylls.csv'), -->
+<!--             row.names = F) -->
+<!-- # Export ylls PATHWAY_HB -->
+<!-- write.csv(ylls_pathway, -->
+<!--             paste0('results/multi_city/health_impacts/ylls_pathway.csv'), -->
+<!--             row.names = F) -->
+<!-- # Export deaths HB -->
+<!-- write.csv(deaths, -->
+<!--             paste0('results/multi_city/health_impacts/deaths.csv'), -->
+<!--             row.names = F) -->
+<!-- # Export deaths PATHWAY_HB -->
+<!-- write.csv(deaths_pathway, -->
+<!--             paste0('results/multi_city/health_impacts/deaths_pathway.csv'), -->
+<!--             row.names = F) -->
+
+<!-- ``` -->

### 2023-11-29 summary_tables.Rmd
Uncomment code to display scenario definition

@@
 
+# Summary tables
+List of tables for scenario, distance and duration
+
 ## Distance categories
@@
 # # rename scenarios
-# rownames(scen_prop) <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))
-
-# kable(scen_prop, headings = "Scenario Proportions")
+rownames(scen_prop) <- unlist(lapply(rownames(scen_prop), FUN=get_qualified_scen_name))
+
+kable(scen_prop, headings = "Scenario Proportions")
 

### 2023-12-06 summary_tables.Rmd
Uncomment code to generate health outcomes as CSV files

@@
     toc_depth: 5
-    toc_float: true
+    toc_float: 
+      collapsed: false
     number_sections: true
@@
 
-<!-- ```{r, echo=FALSE, results='hide', message=FALSE} -->
-
-<!-- # Export all health results -->
-
-<!-- # hb - gives all AP (for diseases affected by AP only), PA (for diseases affected by PA only), combined AP and PA (for diseases affected by both AP and PA) and inj results -->
-<!-- # pathway_hb - gives all AP, PA and inj results (but does not combine the results for diseases affected by both AP and PA) -->
-
-<!-- # initialise variables -->
-<!-- health <- list() -->
-<!-- measure <- c("ylls", "deaths") -->
-<!-- pathway <- c("hb", "pathway_hb") -->
-
-<!-- for (k in measure) { # loop for ylls or deaths -->
-<!--   for (j in pathway) { # loop for hb or pathway_hb -->
-<!--     for (i in names(io)[!names(io) %in% c('scen_prop','ithim_run' )]) { # loop for cities -->
-
-<!--       # Temporal dataset for health outcomes -->
-<!--       temp <- io[[i]]$outcomes[[j]][[k]] -->
-
-<!--       # Total population -->
-<!--       #overall_pop <- sum(io[[i]]$demographic$population) -->
-
-<!--       # Population by age and sex -->
-<!--       pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>%  -->
-<!--         summarize(pop_age_sex = sum(population)) -->
-
-<!--       # get the health burden into the correct format by creating a long format and adding additional columns -->
-<!--       # create long format -->
-<!--       health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure",  -->
-<!--                                       -sex, -age_cat) %>%  -->
-<!--         mutate(city = i, # create city column -->
-<!--                scenario = case_when( # create scenario column -->
-<!--                  grepl("sc_cycle", cause) ~ get_qualified_scen_name('sc_cycle'), -->
-<!--                  grepl("sc_car", cause) ~ get_qualified_scen_name('sc_car'), -->
-<!--                  grepl("sc_bus", cause) ~ get_qualified_scen_name('sc_bus'), -->
-<!--                  grepl("sc_motorcycle", cause) ~ get_qualified_scen_name('sc_motorcycle') -->
-<!--                ), -->
-<!--                level1 = case_when( # create level 1 column -->
-<!--                  grepl("all_cause", cause) ~ "L1: All Cause", -->
-<!--                  grepl("inj", cause) ~ "L1: RTI" -->
-<!--                ), -->
-<!--                level2 = case_when( # create level 2 column -->
-<!--                  grepl("total_cancer", cause) ~ "L2: Cancer", -->
-<!--                  grepl("CVD", cause) ~ "L2: CVD", -->
-<!--                  grepl("respiratory", cause) ~ "L2: Respiratory", -->
-<!--                  grepl("T2D", cause) ~ "L2: Other", -->
-<!--                  grepl("total_dementia", cause) ~ "L2: Other", -->
-<!--                  grepl("Parkinson", cause) ~ "L2: Other", -->
-<!--                  grepl("inj", cause) ~ "L2: RTI" -->
-<!--                ), -->
-<!--                level3 = case_when( # create level 3 column -->
-<!--                  grepl("IHD", cause) ~ "L3: IHD", -->
-<!--                  grepl("lung_cancer", cause) ~ "L3: lung_cancer", -->
-<!--                  grepl("COPD", cause) ~ "L3: COPD", -->
-<!--                  grepl("stroke", cause) ~ "L3: stroke", -->
-<!--                  grepl("T2D", cause) ~ "L3: T2D", -->
-<!--                  grepl("LRI", cause) ~ "L3: LRI", -->
-<!--                  grepl("breast_cancer", cause) ~ "L3: breast_cancer", -->
-<!--                  grepl("colon_cancer", cause) ~ "L3: colon_cancer", -->
-<!--                  grepl("endo_cancer", cause) ~ "L3: endo_cancer", -->
-<!--                  grepl("liver_cancer", cause) ~ "L3: liver_cancer", -->
-<!--                  grepl("total_dementia", cause) ~ "L3: total_dementia", -->
-<!--                  grepl("myeloma", cause) ~ "L3: myeloma", -->
-<!--                  grepl("myeloid_leukemia", cause) ~ "L3: Myeloid Leukemia", -->
-<!--                  grepl("Parkinson", cause) ~ "L3: Parkinson", -->
-<!--                  grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer", -->
-<!--                  grepl("stomach_cancer", cause) ~ "L3: stomach_cancer", -->
-<!--                  grepl("inj", cause) ~ "L3: RTI" -->
-<!--                ), -->
-<!--                dose = case_when( # create dose column -->
-<!--                  grepl("_pa_ap_", cause) ~ "PA and AP", -->
-<!--                  grepl("_pa_", cause) ~ "PA", -->
-<!--                  grepl("_ap_", cause) ~ "AP", -->
-<!--                  grepl("_inj", cause) ~ "RTI" -->
-<!--                ) -->
-<!--         ) %>% # End mutate  -->
-<!--         # join with demographic information -->
-<!--         left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex"))  -->
-
-<!--         # %>% -->
-<!--         # create new columns showing burden per 100000 people by age or by age and sex -->
-<!--         # mutate(metric_100k = measure / overall_pop * 100000,  -->
-<!--         #        metric_100k_sex = measure / overall_pop * 100000) -->
-
-<!--     } # Loop for each city -->
-<!--   } # Loop for each pathway -->
-<!-- } # Loop for each measure -->
-
-<!-- # for each health burden and pathway join all cities into one dataframe -->
-<!-- ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
-<!-- ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
-<!-- deaths <- rbindlist(health$deaths$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
-<!-- deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub")) -->
-
-<!-- # # Export ylls HB -->
-<!-- # output_file <- "results/multi_city/health_impacts/data.xlsx" -->
-<!-- # writexl::write_xlsx(list(ylls = ylls, -->
-<!-- #                          ylls_pathway = ylls_pathway, -->
-<!-- #                          deaths = deaths, -->
-<!-- #                          deaths_pathway = deaths_pathway),  -->
-<!-- #                     path = output_file) -->
-
-<!-- # Export ylls HB -->
-
-<!-- # write csv with output version number -->
-<!-- write.csv(ylls, -->
-<!--             paste0('results/multi_city/health_impacts/ylls_',output_version,'.csv'), -->
-<!--             row.names = F) -->
-<!-- # Export ylls PATHWAY_HB -->
-<!-- write.csv(ylls_pathway, -->
-<!--             paste0('results/multi_city/health_impacts/ylls_pathway_',output_version,'.csv'), -->
-<!--             row.names = F) -->
-<!-- # Export deaths HB -->
-<!-- write.csv(deaths, -->
-<!--             paste0('results/multi_city/health_impacts/deaths_',output_version,'.csv'), -->
-<!--             row.names = F) -->
-<!-- # Export deaths PATHWAY_HB -->
-<!-- write.csv(deaths_pathway, -->
-<!--             paste0('results/multi_city/health_impacts/deaths_pathway_',output_version,'.csv'), -->
-<!--             row.names = F) -->
-
-<!-- # write csv without output version number -->
-<!-- write.csv(ylls, -->
-<!--             paste0('results/multi_city/health_impacts/ylls.csv'), -->
-<!--             row.names = F) -->
-<!-- # Export ylls PATHWAY_HB -->
-<!-- write.csv(ylls_pathway, -->
-<!--             paste0('results/multi_city/health_impacts/ylls_pathway.csv'), -->
-<!--             row.names = F) -->
-<!-- # Export deaths HB -->
-<!-- write.csv(deaths, -->
-<!--             paste0('results/multi_city/health_impacts/deaths.csv'), -->
-<!--             row.names = F) -->
-<!-- # Export deaths PATHWAY_HB -->
-<!-- write.csv(deaths_pathway, -->
-<!--             paste0('results/multi_city/health_impacts/deaths_pathway.csv'), -->
-<!--             row.names = F) -->
-
-<!-- ``` -->
+```{r, echo=FALSE, results='hide', message=FALSE}
+
+# Export all health results
+
+# hb - gives all AP (for diseases affected by AP only), PA (for diseases affected by PA only), combined AP and PA (for diseases affected by both AP and PA) and inj results
+# pathway_hb - gives all AP, PA and inj results (but does not combine the results for diseases affected by both AP and PA)
+
+# initialise variables
+health <- list()
+measure <- c("ylls", "deaths")
+pathway <- c("hb", "pathway_hb")
+
+for (k in measure) { # loop for ylls or deaths
+  for (j in pathway) { # loop for hb or pathway_hb
+    for (i in names(io)[!names(io) %in% c('scen_prop','ithim_run' )]) { # loop for cities
+
+      # Temporal dataset for health outcomes
+      temp <- io[[i]]$outcomes[[j]][[k]]
+
+      # Total population
+      #overall_pop <- sum(io[[i]]$demographic$population)
+
+      # Population by age and sex
+      pop_by_age_sex <- io[[i]]$demographic %>% group_by(age, sex) %>%
+        summarize(pop_age_sex = sum(population))
+
+      # get the health burden into the correct format by creating a long format and adding additional columns
+      # create long format
+      health[[k]][[j]][[i]] <- gather(temp, key = "cause", "measure",
+                                      -sex, -age_cat) %>%
+        mutate(city = i, # create city column
+               scenario = case_when( # create scenario column
+                 grepl("sc_cycle", cause) ~ get_qualified_scen_name('sc_cycle'),
+                 grepl("sc_car", cause) ~ get_qualified_scen_name('sc_car'),
+                 grepl("sc_bus", cause) ~ get_qualified_scen_name('sc_bus'),
+                 grepl("sc_motorcycle", cause) ~ get_qualified_scen_name('sc_motorcycle')
+               ),
+               level1 = case_when( # create level 1 column
+                 grepl("all_cause", cause) ~ "L1: All Cause",
+                 grepl("inj", cause) ~ "L1: RTI"
+               ),
+               level2 = case_when( # create level 2 column
+                 grepl("total_cancer", cause) ~ "L2: Cancer",
+                 grepl("CVD", cause) ~ "L2: CVD",
+                 grepl("respiratory", cause) ~ "L2: Respiratory",
+                 grepl("T2D", cause) ~ "L2: Other",
+                 grepl("total_dementia", cause) ~ "L2: Other",
+                 grepl("Parkinson", cause) ~ "L2: Other",
+                 grepl("inj", cause) ~ "L2: RTI"
+               ),
+               level3 = case_when( # create level 3 column
+                 grepl("IHD", cause) ~ "L3: IHD",
+                 grepl("lung_cancer", cause) ~ "L3: lung_cancer",
+                 grepl("COPD", cause) ~ "L3: COPD",
+                 grepl("stroke", cause) ~ "L3: stroke",
+                 grepl("T2D", cause) ~ "L3: T2D",
+                 grepl("LRI", cause) ~ "L3: LRI",
+                 grepl("breast_cancer", cause) ~ "L3: breast_cancer",
+                 grepl("colon_cancer", cause) ~ "L3: colon_cancer",
+                 grepl("endo_cancer", cause) ~ "L3: endo_cancer",
+                 grepl("liver_cancer", cause) ~ "L3: liver_cancer",
+                 grepl("total_dementia", cause) ~ "L3: total_dementia",
+                 grepl("myeloma", cause) ~ "L3: myeloma",
+                 grepl("myeloid_leukemia", cause) ~ "L3: Myeloid Leukemia",
+                 grepl("Parkinson", cause) ~ "L3: Parkinson",
+                 grepl("head_neck_cancer", cause) ~ "L3: head_neck_cancer",
+                 grepl("stomach_cancer", cause) ~ "L3: stomach_cancer",
+                 grepl("inj", cause) ~ "L3: RTI"
+               ),
+               dose = case_when( # create dose column
+                 grepl("_pa_ap_", cause) ~ "PA and AP",
+                 grepl("_pa_", cause) ~ "PA",
+                 grepl("_ap_", cause) ~ "AP",
+                 grepl("_inj", cause) ~ "RTI"
+               )
+        ) %>% # End mutate
+        # join with demographic information
+        left_join(pop_by_age_sex, by = c("age_cat" = "age", "sex" = "sex"))
+
+        # %>%
+        # create new columns showing burden per 100000 people by age or by age and sex
+        # mutate(metric_100k = measure / overall_pop * 100000,
+        #        metric_100k_sex = measure / overall_pop * 100000)
+
+    } # Loop for each city
+  } # Loop for each pathway
+} # Loop for each measure
+
+# for each health burden and pathway join all cities into one dataframe
+ylls <-  rbindlist(health$ylls$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+ylls_pathway <- rbindlist(health$ylls$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+deaths <- rbindlist(health$deaths$hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+deaths_pathway <- rbindlist(health$deaths$pathway_hb, use.names = T) #|> filter(!str_detect(cause, "_lb") & !str_detect(cause, "_ub"))
+
+# # Export ylls HB
+# output_file <- "results/multi_city/health_impacts/data.xlsx"
+# writexl::write_xlsx(list(ylls = ylls,
+#                          ylls_pathway = ylls_pathway,
+#                          deaths = deaths,
+#                          deaths_pathway = deaths_pathway),
+#                     path = output_file)
+
+# Export ylls HB
+
+# write csv with output version number
+write.csv(ylls,
+            paste0('results/multi_city/health_impacts/ylls_',output_version,'.csv'),
+            row.names = F)
+# Export ylls PATHWAY_HB
+write.csv(ylls_pathway,
+            paste0('results/multi_city/health_impacts/ylls_pathway_',output_version,'.csv'),
+            row.names = F)
+# Export deaths HB
+write.csv(deaths,
+            paste0('results/multi_city/health_impacts/deaths_',output_version,'.csv'),
+            row.names = F)
+# Export deaths PATHWAY_HB
+write.csv(deaths_pathway,
+            paste0('results/multi_city/health_impacts/deaths_pathway_',output_version,'.csv'),
+            row.names = F)
+
+# write csv without output version number
+write.csv(ylls,
+            paste0('results/multi_city/health_impacts/ylls.csv'),
+            row.names = F)
+# Export ylls PATHWAY_HB
+write.csv(ylls_pathway,
+            paste0('results/multi_city/health_impacts/ylls_pathway.csv'),
+            row.names = F)
+# Export deaths HB
+write.csv(deaths,
+            paste0('results/multi_city/health_impacts/deaths.csv'),
+            row.names = F)
+# Export deaths PATHWAY_HB
+write.csv(deaths_pathway,
+            paste0('results/multi_city/health_impacts/deaths_pathway.csv'),
+            row.names = F)
+
+```

### 2023-12-11 summary_tables.Rmd
updated function description

@@
 
-# - one table showing the baseline mode split for all cities
+# (- one table showing the baseline mode split for all cities - CURRENTLY COMMENTED OUT)
 
@@
 
-# - city specific distance tables showing total yearly distances in km travelled by the total population with ages considered in the model for each mode for all scenarios (incl Baseline)
-
-# - one table giving the city specific total yearly baseline distances in km travelled by the total population with ages considered in the model for each mode for all cities
-
-# - city specific distance tables showing total distances travelled by distance category toegether with the number of trips and proportions of all trips within the distance categories.
-
-# - city specific distance tables showing total daily distances in km travelled by the total population  with ages considered in the model for each mode for all scenarios (incl Baseline)
+#( - city specific distance tables showing total yearly distances in km travelled by the total population with ages considered in the model for each mode for all scenarios (incl Baseline)  - CURRENTLY COMMENTED OUT)
+
+#( - one table giving the city specific total yearly baseline distances in km travelled by the total population with ages considered in the model for each mode for all cities - CURRENTLY COMMENTED OUT)
+#( - city specific distance tables showing total distances travelled by distance category toegether with the number of trips and proportions of all trips within the distance categories - CURRENTLY COMMENTED OUT)
+
+#( - city specific distance tables showing total daily distances in km travelled by the total population  with ages considered in the model for each mode for all scenarios (incl Baseline)  - CURRENTLY COMMENTED OUT)
 
@@
 
-# - one table giving the city specific baseline average duration (h) spent travelling by each person in the respective synthetic populations for all cities
-
-## Health outcomes - injury, air pollution and physical activity
-# - one table for each city showing the change in total YLLs for each of the scenarios compared to the reference scenario for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each age and sex category
-
-# - one table for each city showing the change in total YLLs due to changes in accident fatalities for each of the scenarios compared to the reference scenario for each age group
-
-# - one table for each city showing the change in total YLLs due to changes in physical activity levels for each of the scenarios compared to the reference scenario for each age group
-
-# - one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels for each of the scenarios compared to the reference scenario for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalitiesfor each of the scenarios compared to the reference scenario for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levelsfor each of the scenarios compared to the reference scenario for each age group
-
-# - one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels for each of the scenarios compared to the reference scenario for each age group
-
-## Health outcomes - disease
-# - one table for each city showing the change in deaths for each of the scenarios compared to the reference scenario for each disease
-
-# - one table for each city showing the change in deaths per 100,000 people for each of the scenarios compared to the reference scenario for each disease
-
-# - one table for each city showing the change in YLLs for each of the scenarios compared to the reference scenario for each disease
-
-# - one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each disease
+#( - one table giving the city specific baseline average duration (h) spent travelling by each person in the respective synthetic populations for all cities  - CURRENTLY COMMENTED OUT)
+
+## (Health outcomes - injury, air pollution and physical activity  - CURRENTLY COMMENTED OUT)
+# (- one table for each city showing the change in total YLLs for each of the scenarios compared to the reference scenario for each age group - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each age group - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each age and sex category - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in total YLLs due to changes in accident fatalities for each of the scenarios compared to the reference scenario for each age group - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in total YLLs due to changes in physical activity levels for each of the scenarios compared to the reference scenario for each age group - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in total YLLs due to changes in air pollution (PM2.5) levels for each of the scenarios compared to the reference scenario for each age group - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in YLLs per 100,000 people due to changes in accident fatalities for each of the scenarios compared to the reference scenario for each age group - CURRENTLY COMMENTED OUT)
+
+# - (one table for each city showing the change in YLLs per 100,000 people due to changes in physical activity levelsfor each of the scenarios compared to the reference scenario for each age group - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in YLLs per 100,000 people due to changes in air pollution (PM2.5) levels for each of the scenarios compared to the reference scenario for each age group  - CURRENTLY COMMENTED OUT)
+
+## (Health outcomes - disease  - CURRENTLY COMMENTED OUT)
+# (- one table for each city showing the change in deaths for each of the scenarios compared to the reference scenario for each disease  - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in deaths per 100,000 people for each of the scenarios compared to the reference scenario for each disease  - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in YLLs for each of the scenarios compared to the reference scenario for each disease - CURRENTLY COMMENTED OUT)
+
+# (- one table for each city showing the change in YLLs per 100,000 people for each of the scenarios compared to the reference scenario for each disease - CURRENTLY COMMENTED OUT)
 
@@
 
-kable(scen_prop, headings = "Scenario Proportions")
+#kable(scen_prop, headings = "Scenario Proportions")
 

### 2023-12-11 injury_tables.Rmd
Updated function documentation, minor changes

@@
 
-### - output .csv files containing the following information:
+### - output .csv files containing the following information. (Note this files are also saved with their output_version in the file name):
 
@@
 
-# - The inj_data.xlsx file (results/multi_city/inj/inj_data.xlsx) contains all injury output csv files
+# - The inj_data_*output_version*.xlsx file (results/multi_city/inj/inj_data_*output_version*.xlsx) contains all injury output csv files
 

### 2023-12-12 results_app/results_presentation.qmd

---
title: "ITHIM Results"
author: "Ali Abbas"
format: revealjs
editor: visual
---

## ITHIM Overview

### Structure

#### Figure

#### Pathways

## Outcomes

## Scenario Definition - only 5% change in trips

## Distance and Duration tables

## PA outcomes - average MMETs per peron per week

## AP Outcomes - PM2.5 exposure levels

## Health Outcomes

### YLLs and Deaths by total, sex and gender stratification - with and without interaction for three levels and covering all three paths

### Touch on unexpected results when median values of MMETs per week compared in bus and cycle scenarios

## Injury risks by population, distance and duration - overall and not stratified

### 2023-12-12 summary_tables_PA_AP.Rmd

@@
 
-# - desc_stats.csv file (results/multi_city/AP/desc_stats.csv) containing the summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the reference scenario and the total transport related CO2 emission concentrations for each scenario and city
+# - desc_stats.csv file (results/multi_city/AP/desc_stats.csv, also saved with output_version number in file name) containing the summary statistics of individual PM2.5 exposure levels, the total PM2.5 concentrations levels, the change in PM2.5 concentration levels compared to the reference scenario and the total transport related CO2 emission concentrations for each scenario and city
 
-# - desc_stats.xlsx file (results/multi_city/AP/desc_stats.xlsx) containing the following tabs:
+# - desc_stats_*output_version*.xlsx file (results/multi_city/AP/desc_stats_*output_version*.xlsx) containing the following tabs:
 

### 2023-12-13 results_app/results_presentation.qmd
Add presentation on results

@@
 author: "Ali Abbas"
-format: revealjs
-editor: visual
+date: today
+date-format: long
+title-slide-attributes:
+  data-background-image: "images/ithim-accra.png"
+  data-background-size: contain
+  data-background-opacity: "0.2"
+execute:
+    echo: false
+    warning: false
+    message: false
+    cache: false
+format:
+  revealjs:
+    standalone: false
+    embed-resources: true
+    width: 1600
+    height: 900
+    theme: quartomonothemer.scss
+    slide-number: true
+    fig-cap-location: top
+    pdf-separate-fragments: false
+    transition: slide
+    background-transition: fade
+
+output: 
+  revealjs::revealjs_presentation: default
+
+reveal_options:
+  minScale: 1.0
+  maxScale: 1.0
 ---
+
+```{r}
+library(tidyverse)
+library(kableExtra)
+library(modelsummary)
+library(fixest)
+library(gt)
+library(gtExtras)
+library(gghighlight)
+library(quartomonothemer)
+library(showtext)
+```
+
+```{r}
+font_title <- "Josefin Sans"
+font_text <- "Montserrat"
+font_sans <- "Noto Sans" 
+color_base <- "#009F8C"
+color_base_light <- "#95DFD6"
+color_accent <- "#B75C9D"
+color_accent_light <- "#DBA6CC"
+gray <- "#bebebe"
+darkgray <- "#6d6d6d"
+
+font_add_google(font_title)
+font_add_google(font_text)
+showtext_auto()
+
+style_mono_quarto(
+  font_title = font_title,
+  font_text = font_text,
+  font_sans = font_sans,
+  color_base = color_base,
+  color_accent = color_accent,
+  color_link = color_accent,
+  color_code = color_base,
+  size_base = 30,
+  path_scss = here::here("quartomonothemer.scss")
+)
+```
+
+```{r}
+# function to convert scenario names
+get_qualified_scen_name <- function(cs){
+  qualified_scen_name <- ""
+  if (cs == 'base' | cs == 'Baseline' | cs == 'baseline')
+    qualified_scen_name <- 'Baseline'
+  else if(cs == "sc_walk")
+    qualified_scen_name <- 'Walking'
+  else if(cs == "sc_cycle")
+    qualified_scen_name <- 'Bicycling'
+  else if(cs == "sc_car")
+    qualified_scen_name <- 'Car'
+  else if(cs == "sc_motorcycle")
+    qualified_scen_name <- 'Motorcycling'
+  else if(cs == "sc_bus")
+    qualified_scen_name <- 'Public Transport'
+
+  return(qualified_scen_name)
+}
+
+```
+
+```{r}
+
+library(tidyverse)
+library(plotly)
+
+library(here)
+library(tidyverse)
+library(gghighlight)
+library(gt)
+library(gtExtras)
+
+repo_sha <-  as.character(readLines(file.path("repo_sha")))
+io <- readRDS(paste0("../results/multi_city/io_", repo_sha, "_test_run.rds"))
+
+PIF <- read_csv("data/PIF.csv")
+
+PIF <- PIF |> rename(Scenario = name) |> mutate(Scenario = case_when(
+                 Scenario  == "sc_cycle_PIF" ~ "Bicycling",
+                 Scenario  == "sc_car_PIF" ~ "Car",
+                 Scenario  == "sc_bus_PIF" ~ "Public Transport",
+               )
+)
+
+prop <- io$bogota$trip_scen_sets |> distinct(scenario, trip_id, .keep_all = T) |> count(scenario, trip_mode) |> mutate(freq = round(prop.table(n) * 100, 1), .by = scenario) |> mutate(pd = round(freq - freq[scenario == 'baseline'], 1)) |> filter(pd != 0)
+
+prop[prop$scenario == "baseline",]$scenario <- "Baseline"
+prop[prop$scenario == "sc_bus",]$scenario <- "Public Transport"
+prop[prop$scenario == "sc_car",]$scenario <- "Car"
+prop[prop$scenario == "sc_cycle",]$scenario <- "Bicycling"
+
+scen_colours <- c("Baseline" = '#ffffbf',
+                  "Bicycling" = '#abdda4',
+                  "Car" = '#d7191c',
+                  "Public Transport" = '#2b83ba',
+                  "Motorcycle" = '#fdae61')
+
+```
+
+# Scenario Definition - (5% change)
+
+## Scenario Proportions
+
+![](images/trip-prop.png)
+
+## Cycling Scenario
+
+![](images/cycle-scenario.png)
+
+## Bus Scenario
+
+![](images/bus-transport-scenario.png)
+
+## Car Scenario
+
+![](images/car-scenario.png)
+
+## Comparison
+
+```{r}
+
+plotly::ggplotly(ggplot(prop) +
+     aes(x = trip_mode, fill = scenario, weight = pd) +
+     geom_bar() +
+     scale_fill_hue(direction = 1) +
+    scale_x_discrete(guide = guide_axis(angle = 90)) + 
+    geom_text(aes(x = trip_mode, y = round(pd, 1), label = round(pd, 1)), size = 4, 
+                                 position = position_dodge()) + scale_fill_manual(values = scen_colours) +
+     facet_wrap(vars(scenario)) + 
+     labs(title = "Scenario definition (by mode)", y = "Percentage (%)", x = "Trip Mode") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))
+
+```
+
+# Distance tables - (km)
+
+## Distance table
+
+![](images/distance-table.png)
+
+## Cycling Scenario distance table
+
+![](images/distance-table-cycle-scenario.png)
+
+## Bus Scenario distance table
+
+![](images/distance-table-bus-scenario.png)
+
+## Car Scenario distance table
+
+![](images/distance-table-car-scenario.png)
+
+# Duration tables - (minutes)
+
+## Duration table
+
+![](images/duration-table.png)
+
+## Cycling Scenario duration table
+
+![](images/duration-table-cycle-scenario.png)
+
+## Bus Scenario duration table
+
+![](images/duration-table-bus-scenario.png)
+
+## Car Scenario duration table
+
+![](images/duration-table-car-scenario.png)
 
@@
 
-## Scenario Definition - only 5% change in trips
+## Duration table
 
-## Distance and Duration tables
+![](images/duration-table.png)
 
-## PA outcomes - average MMETs per peron per week
+## PM Emission Inventory
 
-## AP Outcomes - PM2.5 exposure levels
+```{r fig.height=5, fig.width=8}
+
+gtb <- io$bogota$vehicle_inventory |> 
+  dplyr::select(-c(speed, CO2_emission_inventory)) |> 
+  filter(PM_emission_inventory !=0) |> 
+  mutate(PM_emission_inventory = round(PM_emission_inventory * 100, 1)) |> 
+  gt() |> 
+  cols_label(stage_mode = "Mode", PM_emission_inventory = "PM Emission Inventory") |>
+  gt_plt_bar_pct(PM_emission_inventory, scaled = TRUE, labels = TRUE)
+
+gtb
+```
+
+## AP Outcomes - PM2.5 exposure levels ($\mu$/$m^3$)
+
+```{r mmets}
+
+mmets <- io$bogota$outcomes$mmets #|> pivot_longer(cols = ends_with("mmet"))
+
+change_names <- which(grepl("_mmet", names(mmets), fixed = TRUE))
+  names(mmets)[change_names] <- sapply(gsub("_mmet" , "", names(mmets)[change_names]), FUN = get_qualified_scen_name)
+
+mmets <- mmets |> pivot_longer(cols = -c(1:4)) |> rename(Scenario = name)
+
+plotly::ggplotly(ggplot(mmets, aes(x=value, fill=Scenario)) + 
+                   geom_density(alpha=.3) + 
+                   scale_fill_manual(values = scen_colours) +
+                   xlim(0, 120) +
+                   labs(title = "Cumulative Distribution of MMETs per person per week (by scenario)",
+                        x = "MMETs per person per week",
+                        y = "Density")
+                   )
+
+```
+
+## PA outcomes - Marginal METs per person per week
+
+```{r}
+mmets |> 
+    group_by(Scenario) |> 
+    reframe(qs = c(quantile(value, prob = 0.25), median(value), mean(value))) |>
+    mutate(qs = round(qs, 1)) |> group_by(Scenario) |> summarise(ld = list(qs)) |> gt() |> gt_plt_bar_stack(column = ld, labels = c("Minimum", "Median", "Mean"), palette = c("#fee8c8", "#fdbb84", "#e34a33"))
+
+```
+
+## Potential Impact Fraction using different censor points
+
+```{r}
+
+ggplotly(ggplot(PIF) +
+  aes(x = group, y = value, fill = Scenario) +
+  geom_boxplot() +
+  scale_fill_hue(direction = 1) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_fill_manual(values = scen_colours) +
+  facet_wrap(vars(Scenario)) +
+  labs("Potential Impact Fraction using all-cause DRF",
+  subtitle = "We have used three different DRFs where censor point varies.
+PIF_75p where censor point is at 75th percentile, and PIF_WHO_DRL (censor point is at 35 MMETs per person per week and PIF_none (without any censor point)",
+  x = "", y = "PIF (%) (+ve benefit)"))
+
+```
 

### 2024-02-29 summary_tables.Rmd

@@
 
-### - output .csv files containing the following information:
+### - output .csv files containing the following information. (Note that all these files are also saved with their version number in the file name):
 
@@
   td <- df %>% distinct(trip_id, scenario, .keep_all = T) %>% 
+    filter(!trip_mode %in% c("bus_driver", "taxi", "rail", "auto_rickshaw", "truck", "other", "car_driver")) |> 
     group_by(trip_mode, scenario) %>% 
@@
   # extract distances and divide by number of people
-  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
+  td <- io[[city]]$dist %>% 
+    # dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>%
+    filter(!stage_mode %in% c("bus_driver", "taxi", "rail", "auto_rickshaw", "truck", "other", "car_driver")) |> 
     mutate_if(is.numeric, round, digits = round_to) %>% 
@@
   # write to .html
-  print(kable(td, caption = paste("Distance table (km) per day for ", city, "( ", count_people, " ) per person in the synthetic population")))
+  print(kable(td, caption = paste("Distance table (km) per day for ", city, "(", count_people, ") per person in the synthetic population")))
   cat("\n")
@@
   count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id))
-  td <- io[[city]]$dist %>% dplyr::filter(stage_mode != 'bus_driver') %>%
+  td <- io[[city]]$dist %>% 
+    # dplyr::filter(stage_mode != 'bus_driver') %>%
+    filter(!stage_mode %in% c("bus_driver", "taxi", "rail", "auto_rickshaw", "truck", "other", "car_driver")) |> 
     mutate_if(is.numeric, round, digits = round_to) %>% 
@@
 
-  print(kable(td, caption = paste("Distance table (km) for ", city, "( ", count_people, " ) per person in the synthetic population (people with trips ONLY)")))
+  print(kable(td, caption = paste("Distance table (km) for ", city, "(", count_people, ") per person in the synthetic population (people with trips ONLY)")))
 
@@
   # extract the duration for the synthetic population
-  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  td <- io[[city]]$dur %>% 
+    # dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% 
+    filter(!stage_mode %in% c("bus_driver", "taxi", "rail", "auto_rickshaw", "truck", "other", "car_driver")) |> 
+    mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
 
@@
   # print to html
-  print(kable(td, caption = paste("Duration table (mins) per day for synthetic population for ", city, "( ", count_people, " ) per person (everyone)")))
+  print(kable(td, caption = paste("Duration table (mins) per day for synthetic population for ", city, "(", count_people, ") per person (everyone)")))
 
@@
   count_people <- length(unique(io[[city]]$trip_scen_sets$participant_id)) # find unique trip ids rather than rows in synthetic population
-  td <- io[[city]]$dur %>% dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
+  td <- io[[city]]$dur %>% 
+    # dplyr::filter(stage_mode != 'bus_driver' & stage_mode != 'car_driver') %>% 
+    filter(!stage_mode %in% c("bus_driver", "taxi", "rail", "auto_rickshaw", "truck", "other", "car_driver")) |> 
+    mutate_if(is.numeric, round, digits = round_to) %>% mutate_if(is.numeric, list(~round((.) / count_people, round_to)))
   colnames(td)[2:ncol(td)] <- c('Baseline', orig_scen_names_updated)
@@
 
-  print(kable(td, caption = paste("Duration table (mins) per day based on trips taken in travel survey for ", city, "( ", count_people, " ) per person (people with trips)")))
+  print(kable(td, caption = paste("Duration table (mins) per day based on trips taken in travel survey for ", city, "(", count_people, ") per person (people with trips)")))
   cat("\n")

### 2024-03-13 summary_tables_PA_AP.Rmd
Use co2_emission instead of conc

@@
 ---
+
 ```{r setup, include=FALSE}
@@
 
-#   - CO2_conc: table detailing the CO2 concentration levels for each scenario for all cities
+#   - C02_emission_inventory: table detailing the CO2 emission levels for each scenario for all cities
 
@@
 # create summary table of CO2 concentrations in each scenario
-co2_conc <- list()
+co2_emission_inventory <- list()
 td <- NA
@@
   # extract CO2 concentrations
-  td <- round(colSums(io[[city]]$outcomes$co2_conc, na.rm = T), 1) %>% 
+  td <- round(colSums(io[[city]]$outcomes$co2_emission_inventory, na.rm = T), 1) %>% 
     as.data.frame() %>% tibble::rownames_to_column()
@@
   # create one dataframe containing all cities
-  if (length(co2_conc) == 0)
-    co2_conc <- td
+  if (length(co2_emission_inventory) == 0)
+    co2_emission_inventory <- td
   else
-    co2_conc <- left_join(co2_conc, td)
+    co2_emission_inventory <- left_join(co2_emission_inventory, td)
 
@@
 # Dan: here the order is different (bus before car)
-# co2_conc$Scenario <- new_scen_names
-co2_conc$Scenario <- sapply(co2_conc$Scenario, FUN = get_qualified_scen_name)
+# co2_emission_inventory$Scenario <- new_scen_names
+co2_emission_inventory$Scenario <- sapply(co2_emission_inventory$Scenario, FUN = get_qualified_scen_name)
 
 # print to html
-print(kable(co2_conc))
+print(kable(co2_emission_inventory))
 
@@
   # extract CO2 transport emissions for each scenario
-  td <- round(colSums(io[[cities[x]]]$outcomes$co2_conc, na.rm = T), 1) %>%
+  td <- round(colSums(io[[cities[x]]]$outcomes$co2_emission_inventory, na.rm = T), 1) %>%
     as.data.frame() %>% tibble::rownames_to_column()
@@
 writexl::write_xlsx(list(PM_2.5_emission_inventory = sl,
-                         CO2_conc = co2_conc,
+                         co2_emission_inventory = co2_emission_inventory,
                          summary_stats_PM2.5_CO2 = bind_rows(l)), path = output_file)

### 2024-03-19 summary_tables_PA_AP.Rmd
Add missing scenario names

@@
     qualified_scen_name <- 'Walking'
-  else if(cs == "sc_cycle")
+  else if(cs == "sc_cycle" | cs == "Bicycling")
     qualified_scen_name <- 'Bicycling'
-  else if(cs == "sc_car")
+  else if(cs == "sc_car" | cs == "Car")
     qualified_scen_name <- 'Car'
@@
     qualified_scen_name <- 'Motorcycling'
-  else if(cs == "sc_bus")
+  else if(cs == "sc_bus" | cs == "Public Transport")
     qualified_scen_name <- 'Public Transport'
